### 2023-04-11 Notes/nonmonotone_project.tex
Initial commit. Added README, TODO, nonmonotone.R and some initial write ups.

\documentclass[12pt]{article}

\usepackage{amsmath, amssymb, mathrsfs, fancyhdr}
\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
\usepackage{biblatex}
\usepackage{booktabs}
\usepackage{float}

\addbibresource{references.bib}

\hypersetup{colorlinks = true, urlcolor = black}

\headheight     15pt
\topmargin      -1.5cm   % read Lamport p.163
\oddsidemargin  -0.04cm  % read Lamport p.163
\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
\textwidth      16.59cm
\textheight     23.94cm
\parskip         7.2pt   % sets spacing between paragraphs
\parindent         0pt   % sets leading space for paragraphs
\pagestyle{empty}        % Uncomment if don't want page numbers
\pagestyle{fancyplain}

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\logistic}{{\text{logistic}}}

\begin{document}

%\lhead{Caleb Leedy}
\chead{Nonmonotone Missingness}
%\chead{STAT 615 - Advanced Bayesian Methods}
%\rhead{Page \thepage\ of \pageref{LastPage}}
\rhead{April 4, 2023}

\section*{Overview:}

The goal of this project is to outperform existing techniques in the literature
related to nonmonotone missing data.

\section*{Completed:}

\begin{itemize}
  \item \textbf{Implemented simulation of monotone MAR data}: This is 
  correspondingly easier
  than the subsequent nonmonotone MAR simulation. For this simulation we use the 
  following approach:
  \begin{enumerate}
      \item Generate $X$, $Y_1$, and $Y_2$ for elements $i = 1, \dots, n$.
      \item Using the covariate $X$, determine the probability $p_1$ of $Y_1$
      being observed for each element $i$.
      \item Based on $p_1$, determine if $R_1 = 1$.
      \item If $R_1 = 0$, then $R_2 = 0$. Otherwise, using variables $X$ and $Y_1$,
      determine the probability $p_{12}$.
      \item Based on $p_{12}$ determine if $R_2 = 1$.
  \end{enumerate}
  At the end of the algorithm, we have determined the values of binary variables 
  $R_1$ and $R_2$ for each $i$ and if either of them are equal to 1, the corresponding
  level of $Y_{k}$. As is common in this literature, the values of $R_1$ and $R_2$
  determine if the corresponding variable $Y_1$ or $Y_2$ is missing or observed
  with $R = 1$ indicating $Y$ being observed.

  \item \textbf{Implemented simulation of nonmonotone MAR data}: 
  Following the approach
    of \cite{robins1997non}, I construct a nonmonotone MAR simulation with two 
    response variables $Y_1$ and $Y_2$ and one covariate $X$. The algorithm to
    generate the data is the following:
    \begin{enumerate}
        \item Generate $X$, $Y_1$, and $Y_2$ for elements $i = 1, \dots, n$.
        \item Using the covariate $X_i$, generate probabilities for each element $i$
        $p_0$, $p_1$, and $p_2$ such that $p_0 + p_1 + p_2 = 1$. 
        \item Select one option based on the three probabilities for each element $i$.
        If 0 is selected: $R_1 = 0$ and $R_2 = 0$. If 1 is selected $R_1 = 1$. If 
        2 is selected, $R_2 = 1$.
        \item We take the next step in multiple cases. If 0 was selected, we are done.
        If 1 was selected, we generate probabilities $p_{12}$ based on $X$ and $Y_1$.
        Then based on this probability, we determine if $R_2 = 1$. In the same manner,
        if 2 was selected in the previous step, we generate probabilities $p_{21}$
        based on $X$ and $Y_2$. Then based on this probability, we determine if $R_2 = 1$.
    \end{enumerate}
    Like the monotone MAR simulation this algorithm produces similar final results
    with the determination of binary variables 
    $R_1$ and $R_2$ and variables $X$, $Y_1$, and $Y_2$. Unlike the monotone MAR case,
    the nonmonotone MAR includes observations with $Y_2$ observed and $Y_1$ missing.

    \item \textbf{Analyzed simulation from monotone MAR}:
    Following the algorithm described in the monotone MAR simulation bullet, 
    we first generate data from the following distributions:
    \begin{align*}
        X_i &\stackrel{iid}{\sim} N(0, 1) \\
        Y_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
        Y_{2i} &\stackrel{iid}{\sim} N(\theta, 1)
    \end{align*}

    Then, we create the probabilities $p_1 = \logistic(x_i)$ and 
    $p_{12} = \logistic(y_{1i})$.
    Since, both $x_i$ and $y_1$ are standard normal distributions, each of these
    probabilities is approximately $0.5$ in expectation.

    The goal of this simulation is to estimate $\theta$. Alternatively, we can
    express this as solving the estimating equation:

    \[g(\theta) \equiv Y_2 - \theta = 0.\]

    We estimate $\theta$ using the following procedures:

    \begin{itemize}
        \item Oracle: This computes $\bar Y$ using \textit{both} the observed and
        missing data.
        \item IPW-Oracle: This is an IPW estimator using only the observed values
        of $Y_2$. The weights (inverse probabilities) use the actual probabilities.
        \item IPW-Est: This is an IPW estimator using the probabilities that have
        been estimated by a logistic model.
        \item Semi: This is the monotone semiparametric efficient estimator from
        Slide 11 (Equation 2) of Dr. Kim's Nonmonotone Missingness presentation.
    \end{itemize}

    We run this simulation with different values of $\theta$, sample size of 1000, and
    1000 Monte Carlo replications. Each algorithm for each replication generates
    $\hat \theta$. In the subsequent tables, we compute the bias, standard deviation (sd),
    t-statistic (where we test for a significant difference between the Monte Carlo
    mean $\hat \theta$ and the true $\theta$) and the p-value of the t-statistic.

    \newpage

    \input{../Tables/monomar_t-5.tex}
    \input{../Tables/monomar_t0.tex}
    \input{../Tables/monomar_t5.tex}

    Overall, these results are mostly what I would have expected. All of the algorithms
    estimate the true value of $\theta$ correctly in each case, with the oracle 
    estimate having the smallest variance followed by the semiparametric algorithm.
    If there is anything surprising it is that the IPW estimator has better 
    performance with the estimated weights compared to the true weights. However,
    I think that this is a known phenomenon.

    \newpage

    \item \textbf{Analyzed simulation from nonmonotone MAR}:
    Like the monotone simulation, we generate data using:
    \begin{align*}
        X_i &\stackrel{iid}{\sim} N(0, 1) \\
        Y_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
        Y_{2i} &\stackrel{iid}{\sim} N(\theta, 1)
    \end{align*}

    However, because we have nonmonotone data, our ``Stage 1'' probabilities are
    different. We compute the true Stage 1 probabilities being proportional to the
    following values:

    \begin{align*}
        p_0 &\propto |x - 2|\\
        p_1 &\propto |x|\\
        p_2 &\propto |x + 2|.
    \end{align*}

   However, we keep the same structure for the Stage 2 probabilities with: 
   $p_{12} = \logistic(y_1)$ and $p_{21} = \logistic(y_2)$. The goal remains to estimate
   $\theta$. We continue to use the Oracle algorithm and the IPW-Oracle algorithm. Since
   we have nonmonotone MAR data, we use the ``Proposed'' algorithm that is described 
   on Slide 25 (Equation 12) of Dr. Kim's presentation. The response models and outcome
   models were estimated using logistic regression and OLS and correctly specified.
   This yields the following results:

    \input{../Tables/nonmonomar_t-5}
    \input{../Tables/nonmonomar_t0}
    \input{../Tables/nonmonomar_t5}

    \newpage
\end{itemize}

\section*{Todo:}

\begin{itemize}
    \item \textbf{Estimate IPW probabilities for nonmonotone MAR}:
\end{itemize}

\newpage

\printbibliography

\end{document}

### 2023-04-11 Simulations/nonmonotone.R

@@ -1,0 +1,515 @@
+# Title: nonmonotone.R
+# Author: Caleb Leedy
+# Date Created: March 29, 2023
+# Purpose: This document contains the simulations for computing estimating
+# equations with both monotone and nonmonotone missingness. See
+# Notes/Non-monotone.pdf for more information.
+#
+# Questions:
+# 1. What estimating functions should we use? 
+# - U(\theta, Y) = (\bar y_1 + \bar y_2) / 2 - \theta
+# - U(\theta, Y) = \bar y_1 + \bar y_2 - \theta
+# - U(\theta, Y) = \bar y_1 - \bar y_2 - \theta
+# Other ideas include moments of a distribution of quantiles.
+#
+# 2. How should nonmonotone missingness data be generated?
+# In the monotone setting we think about data being generated conditionally (or
+# at least the missingness is conditionally generated). This is because we want
+# to have data MAR instead of MCAR. The challenge in this is that we also have
+# to think about how to ensure that 
+#
+# \[p(R_1, R_2 | X) = p(R_1 | R_2, X) p(R_2 | X) = p(R_2 | R_1, X) p(R_1 | X).\]
+# 
+# A further challenge is that these conditional probabilites can depend on Y_1
+# and Y_2 if they have been observed. However, we cannot predetermine if an
+# observation should have wlog R_2 depend on Y_1 because we don't know if Y_1 is
+# observed for that element.
+#
+# Answers:
+# 1. Currently, we will use U(\theta, Y) = \bar y_2 - \theta.
+#
+# 2. While one cannot have a global model this is MAR, using the ideas of Robins
+# and Gill (1998), we can construct an MAR model where each probability is
+# conditional.
+
+# *************
+# * Libraries *
+# *************
+
+library(MASS)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(purrr)
+
+library(doParallel)
+library(doRNG)
[…]

### 2023-04-12 Notes/nonmonotone_project.tex
Updated Notes/nonmmontone_project.tex

@@
 \newcommand{\argmin}{{\text{argmin}}}
+\newcommand{\Cov}{{\text{Cov}}}
 \newcommand{\logistic}{{\text{logistic}}}
@@
       \item Based on $p_1$, determine if $R_1 = 1$.
-      \item If $R_1 = 0$, then $R_2 = 0$. Otherwise, using variables $X$ and $Y_1$,
-      determine the probability $p_{12}$.
+      \item If $R_1 = 0$, then $R_2 = 0$. Otherwise, using variables $X$ and
+        $Y_1$, determine the probability $p_{12}$.
       \item Based on $p_{12}$ determine if $R_2 = 1$.
@@
   At the end of the algorithm, we have determined the values of binary variables
-  $R_1$ and $R_2$ for each $i$ and if either of them are equal to 1, the corresponding
-  level of $Y_{k}$. As is common in this literature, the values of $R_1$ and $R_2$
-  determine if the corresponding variable $Y_1$ or $Y_2$ is missing or observed
-  with $R = 1$ indicating $Y$ being observed.
+    $R_1$ and $R_2$ for each $i$ and if either of them are equal to 1, the
+    corresponding level of $Y_{k}$. As is common in this literature, the values
+    of $R_1$ and $R_2$ determine if the corresponding variable $Y_1$ or $Y_2$ is
+    missing or observed with $R = 1$ indicating $Y$ being observed.
 
@@
         \item Generate $X$, $Y_1$, and $Y_2$ for elements $i = 1, \dots, n$.
-        \item Using the covariate $X_i$, generate probabilities for each element $i$
-        $p_0$, $p_1$, and $p_2$ such that $p_0 + p_1 + p_2 = 1$. 
-        \item Select one option based on the three probabilities for each element $i$.
-        If 0 is selected: $R_1 = 0$ and $R_2 = 0$. If 1 is selected $R_1 = 1$. If 
-        2 is selected, $R_2 = 1$.
-        \item We take the next step in multiple cases. If 0 was selected, we are done.
-        If 1 was selected, we generate probabilities $p_{12}$ based on $X$ and $Y_1$.
-        Then based on this probability, we determine if $R_2 = 1$. In the same manner,
-        if 2 was selected in the previous step, we generate probabilities $p_{21}$
-        based on $X$ and $Y_2$. Then based on this probability, we determine if $R_2 = 1$.
+        \item Using the covariate $X_i$, generate probabilities for each element
+          $i$ $p_0$, $p_1$, and $p_2$ such that $p_0 + p_1 + p_2 = 1$. 
+        \item Select one option based on the three probabilities for each
+          element $i$. If 0 is selected: $R_1 = 0$ and $R_2 = 0$. If 1 is
+          selected $R_1 = 1$. If 2 is selected, $R_2 = 1$.
+        \item We take the next step in multiple cases. If 0 was selected, we are
+          done. If 1 was selected, we generate probabilities $p_{12}$ based on
+          $X$ and $Y_1$. Then based on this probability, we determine if $R_2 =
+          1$. In the same manner, if 2 was selected in the previous step, we
+          generate probabilities $p_{21}$ based on $X$ and $Y_2$. Then based on
+          this probability, we determine if $R_2 = 1$.
     \end{enumerate}
-    Like the monotone MAR simulation this algorithm produces similar final results
-    with the determination of binary variables 
-    $R_1$ and $R_2$ and variables $X$, $Y_1$, and $Y_2$. Unlike the monotone MAR case,
-    the nonmonotone MAR includes observations with $Y_2$ observed and $Y_1$ missing.
+    Like the monotone MAR simulation this algorithm produces similar final
+    results with the determination of binary variables $R_1$ and $R_2$ and
+    variables $X$, $Y_1$, and $Y_2$. Unlike the monotone MAR case, the
+    nonmonotone MAR includes observations with $Y_2$ observed and $Y_1$ missing.
 
@@
     \begin{itemize}
-        \item Oracle: This computes $\bar Y$ using \textit{both} the observed and
-        missing data.
-        \item IPW-Oracle: This is an IPW estimator using only the observed values
-        of $Y_2$. The weights (inverse probabilities) use the actual probabilities.
-        \item IPW-Est: This is an IPW estimator using the probabilities that have
-        been estimated by a logistic model.
+        \item Oracle: This computes $\bar Y$ using \textit{both} the observed
+          and missing data.
+        \item IPW-Oracle: This is an IPW estimator using only the observed
+          values of $Y_2$. The weights (inverse probabilities) use the actual
+          probabilities.
+        \item IPW-Est: This is an IPW estimator using the probabilities that
+          have been estimated by a logistic model.
         \item Semi: This is the monotone semiparametric efficient estimator from
-        Slide 11 (Equation 2) of Dr. Kim's Nonmonotone Missingness presentation.
+          Slide 11 (Equation 2) of Dr. Kim's Nonmonotone Missingness
+          presentation.
     \end{itemize}
 
-    We run this simulation with different values of $\theta$, sample size of 1000, and
-    1000 Monte Carlo replications. Each algorithm for each replication generates
-    $\hat \theta$. In the subsequent tables, we compute the bias, standard deviation (sd),
-    t-statistic (where we test for a significant difference between the Monte Carlo
-    mean $\hat \theta$ and the true $\theta$) and the p-value of the t-statistic.
+    We run this simulation with different values of $\theta$, sample size of
+    1000, and 1000 Monte Carlo replications. Each algorithm for each replication
+    generates $\hat \theta$. In the subsequent tables, we compute the bias,
+    standard deviation (sd), t-statistic (where we test for a significant
+    difference between the Monte Carlo mean $\hat \theta$ and the true $\theta$)
+    and the p-value of the t-statistic.
 
@@
 
-    \input{../Tables/monomar_t-5.tex}
-    \input{../Tables/monomar_t0.tex}
-    \input{../Tables/monomar_t5.tex}
+    \input{Tables/monomar_t-5.tex}
+    \input{Tables/monomar_t0.tex}
+    \input{Tables/monomar_t5.tex}
 
-    Overall, these results are mostly what I would have expected. All of the algorithms
-    estimate the true value of $\theta$ correctly in each case, with the oracle 
-    estimate having the smallest variance followed by the semiparametric algorithm.
-    If there is anything surprising it is that the IPW estimator has better 
-    performance with the estimated weights compared to the true weights. However,
-    I think that this is a known phenomenon.
+    Overall, these results are mostly what I would have expected. All of the
+    algorithms estimate the true value of $\theta$ correctly in each case, with
+    the oracle estimate having the smallest variance followed by the
+    semiparametric algorithm. If there is anything surprising it is that the IPW
+    estimator has better performance with the estimated weights compared to the
+    true weights. However, I think that this is a known phenomenon.
 
@@
     However, because we have nonmonotone data, our ``Stage 1'' probabilities are
-    different. We compute the true Stage 1 probabilities being proportional to the
-    following values:
+    different. We compute the true Stage 1 probabilities being proportional to
+    the following values:
 
@@
    However, we keep the same structure for the Stage 2 probabilities with:
-   $p_{12} = \logistic(y_1)$ and $p_{21} = \logistic(y_2)$. The goal remains to estimate
-   $\theta$. We continue to use the Oracle algorithm and the IPW-Oracle algorithm. Since
-   we have nonmonotone MAR data, we use the ``Proposed'' algorithm that is described 
-   on Slide 25 (Equation 12) of Dr. Kim's presentation. The response models and outcome
-   models were estimated using logistic regression and OLS and correctly specified.
-   This yields the following results:
+   $p_{12} = \logistic(y_1)$ and $p_{21} = \logistic(y_2)$. The goal remains to
+   estimate $\theta$. We continue to use the Oracle algorithm and the IPW-Oracle
+   algorithm. Since we have nonmonotone MAR data, we use the ``Proposed''
+   algorithm that is described on Slide 25 (Equation 12) of Dr. Kim's
+   presentation. The response models and outcome models were estimated using
+   logistic regression and OLS and correctly specified. This yields the
+   following results:
 
-    \input{../Tables/nonmonomar_t-5}
-    \input{../Tables/nonmonomar_t0}
-    \input{../Tables/nonmonomar_t5}
+    \input{Tables/nonmonomar_t-5.tex}
+    \input{Tables/nonmonomar_t0.tex}
+    \input{Tables/nonmonomar_t5.tex}
 
     \newpage
+
+    While most of these numbers are to be expected, the proposed algorithm
+    exhibits considerable bias with the true value of $\theta = -5$. Why does
+    this happen? When $\theta = -5$, $p_{21} = \logistic(y_2)$ is very small
+    ($\logistic(-5) = 0.007$). This means that in many replications every single
+    observation that is initially selected to have $R_2 = 1$ chooses to have
+    $R_1 = 0$. I think that this is causing the problem, but I am not totally
+    sure how to check.
+
+    \item \textbf{Simulation 2 with Nonmonotone MAR}:
+    We also want to simulate data that is correlated. For this simulation, we
+    focus on $\Cov(Y_1, Y_2)$. While the probabilities are generated the same as
+    the previous simulation the data generating process is
+
+    \[\begin{bmatrix}
+    X_i \\ Y_{1i} \\ Y_{2i}
+    \end{bmatrix} \stackrel{iid}{\sim}
+    N\left(
+    \begin{bmatrix}
+        0 \\ 0 \\ \theta
+    \end{bmatrix},
+    \begin{bmatrix}
+        1 & 0 & 0 \\
+        0 & 1 & \sigma_{yy}\\
+        0 & \sigma_{yy} & 1
+    \end{bmatrix}
+    \right).\]
+
+    We are still interested in $\bar Y_2$ and we still run 1000 simulation with
+    1000 observations. In all of the next simulations the true value of $\theta
+    = 0$. The results are the following:
+
+    \input{Tables/nonmonomar_t0cy1y20.1.tex}
+    \input{Tables/nonmonomar_t0cy1y20.5.tex}
+    \input{Tables/nonmonomar_t0cy1y20.9.tex}
+
+    \newpage
+
+    While a small correlation (0.1) in Table 7 (also see Table 5 for no
+    correlation), shows that the proposed algorithm with an insignificant level
+    of bias, with a stronger correlation the bias increases.
 \end{itemize}
 
-\section*{Todo:}
-
+{\bf Commments from JK}
 \begin{itemize}
-    \item \textbf{Estimate IPW probabilities for nonmonotone MAR}:
+    \item Three variables $(X, Y_1, Y_2)$ should be correlated. 
+    \item I think the proposed method is doubly robust. So, you may make a
+      simulation setup testing DR property. I will explain it further. 
 \end{itemize}

### 2023-04-13 Notes/nonmonotone_project.tex
Wrote double robustness.

--- 2023-04-11 Notes/nonmonotone_project.tex
@@
         \item Generate $X$, $Y_1$, and $Y_2$ for elements $i = 1, \dots, n$.
-        \item Using the covariate $X_i$, generate probabilities for each element $i$
-        $p_0$, $p_1$, and $p_2$ such that $p_0 + p_1 + p_2 = 1$. 
-        \item Select one option based on the three probabilities for each element $i$.
-        If 0 is selected: $R_1 = 0$ and $R_2 = 0$. If 1 is selected $R_1 = 1$. If 
-        2 is selected, $R_2 = 1$.
-        \item We take the next step in multiple cases. If 0 was selected, we are done.
-        If 1 was selected, we generate probabilities $p_{12}$ based on $X$ and $Y_1$.
-        Then based on this probability, we determine if $R_2 = 1$. In the same manner,
-        if 2 was selected in the previous step, we generate probabilities $p_{21}$
-        based on $X$ and $Y_2$. Then based on this probability, we determine if $R_2 = 1$.
+        \item Using the covariate $X_i$, generate probabilities for each element
+          $i$ $p_0$, $p_1$, and $p_2$ such that $p_0 + p_1 + p_2 = 1$. 
+        \item Select one option based on the three probabilities for each
+          element $i$. If 0 is selected: $R_1 = 0$ and $R_2 = 0$. If 1 is
+          selected $R_1 = 1$. If 2 is selected, $R_2 = 1$.
+        \item We take the next step in multiple cases. If 0 was selected, we are
+          done. If 1 was selected, we generate probabilities $p_{12}$ based on
+          $X$ and $Y_1$. Then based on this probability, we determine if $R_2 =
+          1$. In the same manner, if 2 was selected in the previous step, we
+          generate probabilities $p_{21}$ based on $X$ and $Y_2$. Then based on
+          this probability, we determine if $R_2 = 1$.
     \end{enumerate}
-    Like the monotone MAR simulation this algorithm produces similar final results
-    with the determination of binary variables 
-    $R_1$ and $R_2$ and variables $X$, $Y_1$, and $Y_2$. Unlike the monotone MAR case,
-    the nonmonotone MAR includes observations with $Y_2$ observed and $Y_1$ missing.
+    Like the monotone MAR simulation this algorithm produces similar final
+    results with the determination of binary variables $R_1$ and $R_2$ and
+    variables $X$, $Y_1$, and $Y_2$. Unlike the monotone MAR case, the
+    nonmonotone MAR includes observations with $Y_2$ observed and $Y_1$ missing.
 
-    \item \textbf{Analyzed simulation from monotone MAR}:
+  \item \textbf{Simulation 1 with Monotone MAR}:
     Following the algorithm described in the monotone MAR simulation bullet, 
@@
     \begin{itemize}
-        \item Oracle: This computes $\bar Y$ using \textit{both} the observed and
-        missing data.
-        \item IPW-Oracle: This is an IPW estimator using only the observed values
-        of $Y_2$. The weights (inverse probabilities) use the actual probabilities.
-        \item IPW-Est: This is an IPW estimator using the probabilities that have
-        been estimated by a logistic model.
+        \item Oracle: This computes $\bar Y$ using \textit{both} the observed
+          and missing data.
+        \item IPW-Oracle: This is an IPW estimator using only the observed
+          values of $Y_2$. The weights (inverse probabilities) use the actual
+          probabilities.
+        \item IPW-Est: This is an IPW estimator using the probabilities that
+          have been estimated by a logistic model.
         \item Semi: This is the monotone semiparametric efficient estimator from
-        Slide 11 (Equation 2) of Dr. Kim's Nonmonotone Missingness presentation.
+          Slide 11 (Equation 2) of Dr. Kim's Nonmonotone Missingness
+          presentation.
     \end{itemize}
 
-    We run this simulation with different values of $\theta$, sample size of 1000, and
-    1000 Monte Carlo replications. Each algorithm for each replication generates
-    $\hat \theta$. In the subsequent tables, we compute the bias, standard deviation (sd),
-    t-statistic (where we test for a significant difference between the Monte Carlo
-    mean $\hat \theta$ and the true $\theta$) and the p-value of the t-statistic.
+    We run this simulation with different values of $\theta$, sample size of
+    2000, and 2000 Monte Carlo replications. Each algorithm for each replication
+    generates $\hat \theta$. In the subsequent tables, we compute the bias,
+    standard deviation (sd), t-statistic (where we test for a significant
+    difference between the Monte Carlo mean $\hat \theta$ and the true $\theta$)
+    and the p-value of the t-statistic.
 
@@
 
-    \item \textbf{Analyzed simulation from nonmonotone MAR}:
-    Like the monotone simulation, we generate data using:
-    \begin{align*}
-        X_i &\stackrel{iid}{\sim} N(0, 1) \\
-        Y_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
-        Y_{2i} &\stackrel{iid}{\sim} N(\theta, 1)
-    \end{align*}
+    \item \textbf{Simulation 1 with Nonmonotone MAR}:
 
-    However, because we have nonmonotone data, our ``Stage 1'' probabilities are
-    different. We compute the true Stage 1 probabilities being proportional to the
-    following values:
+    We generate variables $(X, Y_1, Y_2)$ using the following setup:
+
+    \[\begin{bmatrix}
+    X_i \\ \varepsilon_{1i} \\ \varepsilon_{2i}
+    \end{bmatrix} \stackrel{iid}{\sim}
+    N\left(
+    \begin{bmatrix}
+        0 \\ 0 \\ \theta
+    \end{bmatrix},
+    \begin{bmatrix}
+        1 & 0 & 0 \\
+        0 & 1 & \sigma_{yy}\\
+        0 & \sigma_{yy} & 1
+    \end{bmatrix}
+    \right).\]
+
+    Then,
+
+    \[y_{1i} = x_i + \varepsilon_{1i} \text{ and } 
+    y_{2i} = x_i + \varepsilon_{2i}.\]
+
+    Since we have nonmonotone data, our ``Stage 1'' probabilities are
+    different. We compute the true Stage 1 probabilities being proportional to
+    the following values:
 
     \begin{align*}
-        p_0 &\propto |x - 2|\\
-        p_1 &\propto |x|\\
-        p_2 &\propto |x + 2|.
+        p_0 &= 0.2 \\
+        p_1 &= 0.4 \\
+        p_2 &= 0.4 \\
     \end{align*}
@@
    However, we keep the same structure for the Stage 2 probabilities with:
-   $p_{12} = \logistic(y_1)$ and $p_{21} = \logistic(y_2)$. The goal remains to estimate
-   $\theta$. We continue to use the Oracle algorithm and the IPW-Oracle algorithm. Since
-   we have nonmonotone MAR data, we use the ``Proposed'' algorithm that is described 
-   on Slide 25 (Equation 12) of Dr. Kim's presentation. The response models and outcome
-   models were estimated using logistic regression and OLS and correctly specified.
-   This yields the following results:
+   $p_{12} = \logistic(y_1)$ and $p_{21} = \logistic(y_2)$. The goal remains to
+   estimate $\theta$. We continue to use the Oracle algorithm and the IPW-Oracle
+   algorithm. Since we have nonmonotone MAR data, we use the ``Proposed''
+   algorithm that is described on Slide 25 (Equation 12) of Dr. Kim's
+   presentation. The response models and outcome models were estimated using
+   logistic regression and OLS and correctly specified. This yields the
+   following results:
 
-    \input{../Tables/nonmonomar_t-5}
-    \input{../Tables/nonmonomar_t0}
-    \input{../Tables/nonmonomar_t5}
+    \input{Tables/nonmonocmar_t-5cy1y20.tex}
+    \input{Tables/nonmonocmar_t0cy1y20.tex}
+    \input{Tables/nonmonocmar_t5cy1y20.tex}
 
     \newpage
+
+    \item \textbf{Simulation 2 with Nonmonotone MAR}:
+    We also want to simulate data that is correlated. For this simulation, we
+    focus on $\Cov(Y_1, Y_2)$. The data generating process now has $\sigma_{yy}
+    \neq 0$. We are still interested in $\bar Y_2$ and we still run 2000
+    simulation with 2000 observations. In all of the next simulations the true
+    value of $\theta = 0$. The results are the following:
+
+    \input{Tables/nonmonocmar_t0cy1y20.1.tex}
+    \input{Tables/nonmonocmar_t0cy1y20.5.tex}
+    \input{Tables/nonmonocmar_t0cy1y20.9.tex}
+
+    \newpage
+
+  \item \textbf{Simulation 3 with Nonmonotone MAR}:
+    This simulation aims to see if the proposed algorithm is doubly robust.
+    First, we check with a misspecified outcome model. In this case the data
+    generating procedure is the following:
+
+    \[\begin{bmatrix}
+    X_i \\ \varepsilon_{1i} \\ \varepsilon_{2i}
+    \end{bmatrix} \stackrel{iid}{\sim}
+    N\left(
+    \begin{bmatrix}
+        0 \\ 0 \\ \theta
+    \end{bmatrix},
+    \begin{bmatrix}
+        1 & 0 & 0 \\
+        0 & 1 & \sigma_{yy}\\
+        0 & \sigma_{yy} & 1
+    \end{bmatrix}
+    \right).\]
+
+    Then,
+
+    \[y_{1i} = x_i + x_i^2 \varepsilon_{1i} \text{ and } 
+    y_{2i} = -x_i + x_i^3 + \varepsilon_{2i}.\]
+
+    This procedure causes $X$ to influence both $Y_1$ and $Y_2$ and we still
+    have correlation in the error terms of $Y_1$ and $Y_2$. However, since
+    neither $Y_1$ nor $Y_2$ are linear in $X$, the model will be misspecified.
+    The response mechanisms are first generated MCAR with a probability of
+    either $Y_1$ or $Y_2$ being the first variable observed to be $0.4$. (There
+    is a $0.2$ probability neither is observed.) Then the probability of the
+    other variable being observed is proportional to $\logistic(y_k)$ where
+    $y_k$ is the $y$ that has been observed. To ensure that the proposed method
+    has the correct propensity score we use the oracle probabilites instead of
+    estimating them. This yields the following:
+
+    \input{Tables/nmonocmar_t0cy1y20.tex}
+    \input{Tables/nmonocmar_t0cy1y20.1.tex}
+    \input{Tables/nmonocmar_t0cy1y20.5.tex}
+
+    \newpage
+
+    Clearly, the proposed method is unbiased with a misspecified outcome model.
+    When the response model is misspecified as in Simulation 1, the procedure
+    still works.
+
 \end{itemize}
 
-\section*{Todo:}
-
+{\bf Commments from JK}
 \begin{itemize}
-    \item \textbf{Estimate IPW probabilities for nonmonotone MAR}:
+    \item Three variables $(X, Y_1, Y_2)$ should be correlated. 
 \end{itemize}

### 2023-04-13 Simulations/nonmonotone.R

@@ -461,17 +461,17 @@ registerDoParallel(clust)
 
 B <- 2000
 true_mean <- 0
-cor_y1y2 <- 0.5
+cor_y1y2 <- 0
 
 mc_theta <- foreach(b = 1:B, 
                   #.options.RNG = 1,
                   .packages = c("dplyr")) %dorng% {
 
   # Parameters
-  n_sim <- 1000
+  n_sim <- 2000
 
   df <- nonmono_mar(n_sim, mean_y2 = true_mean, cor_y1y2 = cor_y1y2,
-                    r_ind_y = FALSE, miss_out = TRUE)
+                    r_ind_y = FALSE, miss_out = FALSE)
 
   # Estimation
   # We want to compute:
@@ -505,11 +505,9 @@ mc_theta %>%
   tidyr::pivot_longer(cols = everything(),
                names_to = c(".value", "algorithm"),
                names_pattern = "(.*)_(.*)") %>%
-  mutate(pval = pt(-abs(tstat), df = B)) 
-
-%>%
+  mutate(pval = pt(-abs(tstat), df = B)) %>%
   knitr::kable("latex", booktabs = TRUE, digits = 3,
    caption = paste0("True Value is ", true_mean, ". Cor(Y1, Y2) = ", cor_y1y2)) %>%
-  cat(., file = paste0("../Tables/nonmonomar_t" , true_mean, "cy1y2", cor_y1y2, ".tex"))
+  cat(., file = paste0("../Tables/nonmonocmar_t" , true_mean, "cy1y2", cor_y1y2, ".tex"))
   
 

### 2023-04-17 Notes/nonmonotone_project.tex
Wrote up double robustness. Modified simulations in Simulations/nonmonotone.R so that tables are reproducible.

@@
 
-\section*{Completed:}
+\section*{Initial Simulations:}
+% Tables generated from Simulations/nonmonotone.R
 
@@
    algorithm that is described on Slide 25 (Equation 12) of Dr. Kim's
-   presentation. The response models and outcome models were estimated using
-   logistic regression and OLS and correctly specified. This yields the
+   presentation. The outcome models were estimated using
+   logistic regression and OLS and correctly specified. The response model used
+   the oracle estimates of the probabilites. This yields the
    following results:
 
-    \input{Tables/nonmonocmar_t-5cy1y20.tex}
-    \input{Tables/nonmonocmar_t0cy1y20.tex}
-    \input{Tables/nonmonocmar_t5cy1y20.tex}
+    \input{Tables/nonmonosim1m-5.tex}
+    \input{Tables/nonmonosim1m0.tex}
+    \input{Tables/nonmonosim1m5.tex}
 
@@
 
-    \input{Tables/nonmonocmar_t0cy1y20.1.tex}
-    \input{Tables/nonmonocmar_t0cy1y20.5.tex}
-    \input{Tables/nonmonocmar_t0cy1y20.9.tex}
+    \input{Tables/nonmonosim2c0.1.tex}
+    \input{Tables/nonmonosim2c0.5.tex}
+    \input{Tables/nonmonosim2c0.9.tex}
 
@@
 
-    \input{Tables/nmonocmar_t0cy1y20.tex}
-    \input{Tables/nmonocmar_t0cy1y20.1.tex}
-    \input{Tables/nmonocmar_t0cy1y20.5.tex}
+    \input{Tables/nonmonosim3c0.tex}
+    \input{Tables/nonmonosim3c0.1.tex}
+    \input{Tables/nonmonosim3c0.5.tex}
 
@@
 
-    Clearly, the proposed method is unbiased with a misspecified outcome model.
-    When the response model is misspecified as in Simulation 1, the procedure
-    still works.
+    Thus, the proposed method is unbiased with a misspecified outcome model.
+    We now show a simulation where the outcome model is correctly specified but
+    the response model is not.
+
+  \item \textbf{Simulation 4 with Nonmonotone MAR}:
+    Continuing to test if the proposed algorithm is doubly robust, this
+    simulation checks a misspecified response model. Instead of using oracle
+    weights as in Simulation 3, we estimate the weights for the proposed method.
+    The algorithms to which we compare still use the oracle weights.
+
+    \input{Tables/nonmonosim4c0.tex}
+    \input{Tables/nonmonosim4c0.1.tex}
+    \input{Tables/nonmonosim4c0.5.tex}
+
+    This is clear evidence that the proposed method is doubly robust.
 
@@
 
-{\bf Commments from JK}
+\newpage
+
+\section*{Full Nonmonotone Estimator}
+
 \begin{itemize}
-    \item Three variables $(X, Y_1, Y_2)$ should be correlated. 
+  \item
+
 \end{itemize}

### 2023-04-17 Simulations/clean_tables.R

@@ -1,0 +1,29 @@
+# Title: clean_tables.R
+# Author: Caleb Leedy
+# Date Created: April 17, 2023
+# Purpose: This script edits the tables in Tables/ so that I do not have to edit
+# them by hand.
+
+# *************
+# * Libraries *
+# *************
+
+library(stringr)
+
+# Steps:
+# 1. Read .tex files from Tables/
+# 2. Check to see if "[h!]" is at the end of line 1. 
+# 3. If not add it.
+
+tables_to_check <- list.files("../Tables/", full.names = TRUE)
+
+for (tab in tables_to_check) {
+
+  tab_vec <- readLines(tab)
+
+  if (str_sub(tab_vec[1], start = -4) != "[h!]") {
+    tab_vec[1] <- str_c(tab_vec[1], "[h!]")
+    cat(tab_vec, file = tab, sep = "\n")
+  }
+
+}

### 2023-04-17 Simulations/nonmonotone.R

@@ -456,58 +456,171 @@ twophase_reg <- function(df) {
 # * Run Simulation: Nonmonotone Missingness *
 # *******************************************
 
-clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
-registerDoParallel(clust)
+run_init_sims <- function(n_sim, B, true_mean, cor_y1y2,
+                          miss_out, oracle, seed) {
 
-B <- 2000
-true_mean <- 0
-cor_y1y2 <- 0
 
-mc_theta <- foreach(b = 1:B, 
-                  #.options.RNG = 1,
-                  .packages = c("dplyr")) %dorng% {
+  clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
+  registerDoParallel(clust)
 
-  # Parameters
-  n_sim <- 2000
+  mc_theta <- 
+    foreach(b = 1:B, 
+            .options.RNG = seed,
+            .export = c("nonmono_mar", "ipw_cc_oracle", "est_nonmono", "expit"),
+            .packages = c("dplyr")) %dorng% {
 
-  df <- nonmono_mar(n_sim, mean_y2 = true_mean, cor_y1y2 = cor_y1y2,
-                    r_ind_y = FALSE, miss_out = FALSE)
+    # Parameters
+    df <- nonmono_mar(n_sim, mean_y2 = true_mean, cor_y1y2 = cor_y1y2,
+                      r_ind_y = FALSE, miss_out = miss_out)
 
-  # Estimation
-  # We want to compute:
-  # 1. Oracle estimate
-  # 2. pi_11 estimate (ipw using fully observed cases)
-  # 3. New estimate
+    # Estimation
+    # We want to compute:
+    # 1. Oracle estimate
+    # 2. pi_11 estimate (ipw using fully observed cases)
+    # 3. New estimate
 
-  tibble(oracle = mean(df$y2),
-         ipworacle = ipw_cc_oracle(df),
-         proposed = est_nonmono(df, oracle = TRUE))
+    tibble(oracle = mean(df$y2),
+           ipworacle = ipw_cc_oracle(df),
+           proposed = est_nonmono(df, oracle = oracle))
 
-} %>% bind_rows()
+  } %>% bind_rows()
 
-stopCluster(clust)
+  stopCluster(clust)
 
-# ****************
[…]

### 2023-04-18 Notes/nonmonotone_project.tex
Updated after meeting with Dr. Kim. We talked about simulation study that did not show robustness to misspecified response model. TODO: Fix estimation of (X_i, Y_{2i})$ and rerun Simulation 4.

@@
    logistic regression and OLS and correctly specified. The response model used
-   the oracle estimates of the probabilites. This yields the
+   the oracle estimates of the probabilities. This yields the
    following results:
@@
     $y_k$ is the $y$ that has been observed. To ensure that the proposed method
-    has the correct propensity score we use the oracle probabilites instead of
+    has the correct propensity score we use the oracle probabilities instead of
     estimating them. This yields the following:
@@
 
-    This is clear evidence that the proposed method is doubly robust.
+    Thus, there is strong evidence that the proposed method is \textit{not}
+    doubly robust. If we have a misspecified response model, then it seems like
+    we induce bias. I will explore this more in the next section.
 
@@
 
-\section*{Full Nonmonotone Estimator}
+\section*{Estimating the Response Model}
 
 \begin{itemize}
-  \item
+  \item In most (critically \textit{not} Simulation 4) of the simulations of the
+    previous section, we used the oracle weights when estimating our proposed
+    method. The reasoning for this was straightforward{--}we wanted to ensure
+    that the response model was correctly specified and the best case scenario
+    is to use the oracle weights which we did.
+
+  \item This section focuses more on estimating these response weights. Instead
+    of focusing on the proposed method, we will actually be working on
+    estimation of the complete case IPW estimator. This model is less complex
+    and will thus make it easier to understand where we are making mistakes.
+
+  \item We use the following simulation study:
+    \begin{align*}
+      x_i &\stackrel{iid}{\sim} N(0, 1)\\
+      \varepsilon_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
+      \varepsilon_{2i} &\stackrel{iid}{\sim} N(0, 1)\\
+      y_{1i} &= x_i + \varepsilon_{1i} \\
+      y_{2i} &= x_i + \varepsilon_{2i} \\
+    \end{align*}
+
+    To select a missingness pattern for each $i$, we have sequence: first, we
+    select the first variable to observe (or neither), then we either select the
+    second variable or we do not. In the first step, we select $R_{1i} = 1$ with
+    probability $0.4$, $R_{2i} = 1$ with probability $0.4$ and neither variable
+    with probability $0.2$. For the second step, we have the probability of
+    observing the other variable be $\logistic(x_i)$. This yields the following:
+
+    \input{Tables/ipwsim.tex}
+    % TODO: estimate IPW with \pi_{2|1} and \pi_{1|2}.
+
+    The problem with this is that our estimate is biased. For one particular
+    realization of the simulation, here is the distribution of the difference
+    between the estimated and true probabilities of $\pi_{11}$:
+
+    \includegraphics[width = 0.7\linewidth]{diffhist.png}
+
+\end{itemize}
+
+\newpage
+
+\section*{Proposal: Full Nonmonotone Estimator}
+
+\begin{itemize}
+
+  \item When constructing
+    estimates of the response model, we estimate $\pi_{11}$ using
+    $\pi_{A_2 = 1|A_1 = 1}$ and we never include information about $\pi_{A_1 =
+    1|A_2 = 1}$. I would like to create a full estimator where we are able to
+    use this information.
+
+  \item The current estimator is the following (See Slide 24 of Non-monotone
+    Presentation):
+
+    \begin{align*}
+      \hat \theta_{eff} &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \\
+      &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)} (b_2(X_i, Y_{1i}) - 
+      E[g_i \mid X_i]) \\
+      &+ n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)} (a_2(X_i, Y_{2i}) -
+      E[g_i \mid X_i]) \\
+      &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_1, Y_{1i})}(g_i 
+      - b_2(X_i, Y_{1i}) - a_2(X_i, Y_{2i}) + E[g_i \mid X_i])
+    \end{align*}
+
+  \item The problem with this is that we assume $\pi_{11}(X, Y_{1}) =
+    \pi_{2|1} (X, Y_{1}) \pi_{1+}(X)$ when we could have $\pi_{11}(X, Y_2) =
+    \pi_{1|2} (X, Y_2) \pi_{2+}(X)$. In other words, the previous result
+    implicitly assumes that $\pi_{11}$ is a function of $X$ and $Y_1$ when it
+    could be a function of $X$ and $Y_2$ as well.
+
+  \item I think that we could use the
+    following (small modification) instead,
+
+    \[\pi_{11}(X, Y_1, Y_2) = \alpha \pi_{1|2} \pi_{2+} + (1 - \alpha) \pi_{2|1}
+    \pi_{1+}\]
+
+    for some $\alpha \in [0, 1]$.
+
+  \item The reasoning behind this is two-fold. First, it allows us to think of
+    the nonmonotone missingness case as a linear combination of two monotone
+    cases, which I think is useful. Second, we can make explicit our choice of
+    $\alpha$, which in the existing estimator is simply $\alpha = 0$.
+
+  \item The problem with adding another parameter $\alpha$ is that it makes the
+    overall model unidentifiable. We cannot estimate $\alpha$ and the
+    conditional and marginal distributions that we have without additional
+    assumptions. So for now, I think that we should just assume that $\alpha$ is
+    known and then we apply this method to a data integration problem we can
+    review what $\alpha$ makes sense or figure out an additional assumption that
+    can help us estimate $\alpha$ (for example if there is a variable correlated
+    with $1|2$ versus $2|1$).
+
+  \item This is not just a pure Bayes' rule.
+    Formally, let $L = (X, Y_1, Y_2)$. By Bayes' rule we have
+
+    \begin{align*}
+      \Pr(R_1 &= 1 \mid R_2 = 1, L) \Pr(R_2 = 1 \mid L) \\
+      &= \Pr(R_1 = 1, R_2 = 1 \mid L) \\
+      & = \Pr(R_2 = 1 \mid R_1 = 1, L) \Pr(R_1 = 1 \mid L).
+    \end{align*}
+
+    Yet to estimate this model, we need to assume something like the following:
+    \begin{align*}
+      \Pr(R_1 &= 1 \mid R_2 = 1, X, Y_2) \Pr(R_2 = 1 \mid X, Y_2)\\
+      &= \Pr(R_1 = 1, R_2 = 1 \mid L) \\
+      &= \Pr(R_2 = 1 \mid R_1 = 1, X, Y_1) \Pr(R_1 = 1 \mid X, Y_1)
+    \end{align*}
+
+    in which case the two sides are clearly different.
+
+    \item Unfortunately, early simulation results are not promising because we
+      cannot distinguish points that should have $\pi_{11}$ conditional on $Y_1$
+      and points that should be conditional on $Y_2$.
+
+      \input{Tables/alphasim.tex}
+
+    \item Dr. Kim, please let me know what you think about this idea and if it
+      is worth pursuing more.
 

### 2023-04-18 Simulations/nonmonotone.R

@@ -80,7 +80,7 @@ mono_mar <- function(n, mean_y2 = 0) {
 
 nonmono_mar <- 
   function(n, mean_y2 = 0, cor_xy1 = 0, cor_xy2 = 0, cor_y1y2 = 0,
-           r_ind_y = FALSE, miss_out = FALSE, miss_resp = FALSE) {
+           r_ind_y = FALSE, miss_out = FALSE, miss_resp = FALSE, mcar = FALSE) {
 
   # Steps:
   # 1. Generate X, Y_1, Y_2.
@@ -127,8 +127,14 @@ nonmono_mar <-
     } else {
 
       p0 <- 0.2
-      p1 <- expit(1 + 2 * x_vec)
-      p2 <- expit(-1 + -2 * x_vec)
+
+      if (mcar) {
+        p1 <- 0.4 
+        p2 <- 0.4
+      } else {
+        p1 <- expit(2 * x_vec)
+        p2 <- expit(-2 * x_vec)
+      }
       tot <- p0 + p1 + p2
       p0 <- p0 / tot
       p1 <- p1 / tot
@@ -173,9 +179,9 @@ nonmono_mar <-
     } else {
 
       p0 <- 0.2
-      # p1 <- expit(1 + 2 * x_vec)
+      # p1 <- expit(2 * x_vec)
       p1 <- 0.4
-      # p2 <- expit(-1 + -2 * x_vec)
+      # p2 <- expit(-2 * x_vec)
       p2 <- 0.4
       tot <- p0 + p1 + p2
       p0 <- p0 / tot
@@ -336,7 +342,7 @@ mc_theta %>%
 # * Estimating Nonmonotone Missingness *
 # **************************************
 
-est_nonmono <- function(df, oracle = FALSE) {
+est_nonmono <- function(df, oracle = FALSE, alpha = NA) {
 
   # g(x, y1, y2) = y2
   gfun <- "y2"
@@ -359,13 +365,20 @@ est_nonmono <- function(df, oracle = FALSE) {
 
   } else {
     pi_11_mod <- glm(r2 ~ x + y1, data = df_10, family = binomial(link = logit))
-    pi_11 <- pi_11_mod$fitted.values # NOTE: This is NOT of length n.
+    pi_12 <- pi_11_mod$fitted.values # NOTE: This is NOT of length n.
 
     df$pi_11 <- 0
-    df$pi_11[df$r1 == 1] <- pi_11
[…]

### 2023-04-20 Notes/nonmonotone_project.tex
Added simulation and proof for double robustness of estimator. We still need to make progress to see what to do if the response model is sometimes a function of Y.

@@
     weights as in Simulation 3, we estimate the weights for the proposed method.
+    However, unlike the true probabilities of being proportional to
+    $\logistic(y_k)$, this simulation has the true probabilities being
+    proportional to $\logistic(x_i)$.
     The algorithms to which we compare still use the oracle weights.
@@
 
-    Thus, there is strong evidence that the proposed method is \textit{not}
-    doubly robust. If we have a misspecified response model, then it seems like
-    we induce bias. I will explore this more in the next section.
+    Thus, there is strong evidence that the proposed method is 
+    doubly robust. Previously, we did not get this result and the reason is 
+    the way that the selection probabilities work. When they were proportional
+    to $\logistic(y_k)$, the estimator is biased (see the next section);
+    however, when the probabilities are proportional to $\logistic(x_i)$ then it
+    works.
+
+\end{itemize}
+
+\newpage
+
+\section*{Missingness Mechanism}
+
+\begin{itemize}
+  \item First I am going to reproduce the proof of double robustness that we
+    talked about during our last meeting. I think it is insightful for future
+    comments:
+
+    \begin{align*}
+      E[\hat \theta_{eff} - \theta_n] 
+      &= E\left[n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] - g_i\right]\\
+      &\quad+ E\left[n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)}
+      (b_2(X_i, Y_{1i}) - E[g_i \mid X_i])\right]\\
+      &\quad+ E\left[n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)}
+      (a_2(X_i, Y_{2i}) - E[g_i \mid X_i])\right]\\
+      &\quad+ E\left[n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_i)} 
+      (g_i - a_2(X_i, Y_{2i}) - b_2(X_i, Y_{1i}) + E[g_i \mid X_i])\right]\\
+      &= n^{-1} \sum_{i = 1}^n (E[E[g_i \mid X_i]] - E[g_i])\\
+      &\quad+ n^{-1} \sum_{i = 1}^n E\left[E\left[\frac{R_{1i}}{\pi_{1+}(X_i)} 
+      (b_2(X_i, Y_{1i}) - E[g_i \mid X_i]) \mid X_i\right]\right] \\ 
+      &\quad+ n^{-1} \sum_{i = 1}^n E\left[E\left[\frac{R_{2i}}{\pi_{2+}(X_i)} 
+      (a_2(X_i, Y_{2i}) - E[g_i \mid X_i]) \mid X_i\right]\right] \\
+      &\quad+ n^{-1} \sum_{i = 1}^n E\left[E\left[\frac{R_{1i}
+      R_{2i}}{\pi_{11}(X_i)} (g_i - a_2(X_i, Y_{2i}) - b_2(X_i, Y_{1i}) + E[g_i
+      \mid X_i]) \mid X_i \right]\right]\\
+      \intertext{Since $R_{1i} \perp Y_{1i} \mid X_i$, $R_{2i} \perp Y_{2i} \mid
+      X_i$, $(R_{1i}, R_{2i}) \perp (Y_{1i}, Y_{2i}) \mid X_i$ and $\pi_{1+},
+      \pi_{2+}$ and $\pi_{11}$ are all free of $Y_{1i}$ and $Y_{2i}$.}
+      &= n^{-1} \sum_{i = 1}^n E\left[\frac{R_{1i}}{\pi_{1+}(X_i)} E\left[
+        (E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i]) \mid X_i\right]\right] \\ 
+      &\quad+ n^{-1} \sum_{i = 1}^n E\left[\frac{R_{2i}}{\pi_{2+}(X_i)} 
+      E\left[E[g_i \mid X_i, Y_{2i}] - E[g_i \mid X_i] \mid X_i\right]\right] \\
+      &\quad+ n^{-1} \sum_{i = 1}^n E\left[\frac{R_{1i}
+      R_{2i}}{\pi_{11}(X_i)} E\left[(g_i - E[g_i \mid X_i, Y_{2i}] - 
+      E[g_i \mid X_i, Y_{1i}]+ E[g_i \mid X_i]) \mid X_i \right]\right]\\
+      \intertext{Since $E[E[g_i \mid X_i, Y_{ki}] \mid X_i] = E[g_i \mid X_i] =
+      0$,}
+      &= 0.
+    \end{align*}
+
+    Thus, if the outcome models are correctly specified $\hat \theta_{eff}$ is
+    unbiased. If the response models are correctly specified it is easy to see
+    that $\hat \theta_{eff}$ is also unbiased. This means that $\hat
+    \theta_{eff}$ is doubly robust.
+
+  \item However, one of the key steps is that \textit{all} of the response
+    models are free of $Y$. In a previous iteration of Simulation 4, we had
+    adopted the framework of \cite{robins1997non} where we first to observe the
+    first variable and see if we observe the second variable. In this case, the
+    second step can depend on the result of the first step and this is what we
+    did. However, this makes it the case that $\pi_11$ is a function of $X_i$
+    and $Y_1$ and $Y_2$. In this case $\hat \theta_{eff}$ is not unbiased if the
+    reponse model is misspecified (or even just estimated).
+
+  \item If we modify Simulation 4, such that the second step of observed the
+    second varible is proportional to $\logistic(y_k)$ then we get the same
+    result as before:
+
+    \input{Tables/nonmonosim5c0.tex}
+    \input{Tables/nonmonosim5c0.1.tex}
+    \input{Tables/nonmonosim5c0.5.tex}
 

### 2023-04-20 Simulations/nonmonotone.R

@@ -494,7 +494,7 @@ twophase_reg <- function(df) {
 # instead it is a parameter for est_nonmono. This should be changed to something
 # like est_weights.
 run_init_sims <- function(n_sim, B, true_mean, cor_y1y2,
-                          miss_out, miss_resp, seed) {
+                          miss_out, miss_resp, seed, r_ind_y = FALSE) {
 
 
   clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
@@ -508,7 +508,7 @@ run_init_sims <- function(n_sim, B, true_mean, cor_y1y2,
 
     # Parameters
     df <- nonmono_mar(n_sim, mean_y2 = true_mean, cor_y1y2 = cor_y1y2,
-                      r_ind_y = FALSE, miss_out = miss_out)
+                      r_ind_y = r_ind_y, miss_out = miss_out)
 
     # Estimation
     # We want to compute:
@@ -627,24 +627,49 @@ run_init_sims(n_sim = 2000, B = 2000,
 
 run_init_sims(n_sim = 2000, B = 2000,
               true_mean = 0, cor_y1y2 = 0,
-              miss_out = FALSE, miss_resp = TRUE, seed = 4) %>%
+              miss_out = FALSE, miss_resp = TRUE, seed = 4, r_ind_y = TRUE) %>%
   knitr::kable("latex", booktabs = TRUE, digits = 3,
   caption = "True Value is 0. Cor(Y1, Y2) = 0") %>%
   cat(., file = paste0("../Tables/nonmonosim4" , "c0", ".tex"))
 
 run_init_sims(n_sim = 2000, B = 2000,
               true_mean = 0, cor_y1y2 = 0.1,
-              miss_out = FALSE, miss_resp = TRUE, seed = 4) %>%
+              miss_out = FALSE, miss_resp = TRUE, seed = 4, r_ind_y = TRUE) %>%
   knitr::kable("latex", booktabs = TRUE, digits = 3,
   caption = "True Value is 0. Cor(Y1, Y2) = 0.1") %>%
   cat(., file = paste0("../Tables/nonmonosim4" , "c0.1", ".tex"))
 
 run_init_sims(n_sim = 2000, B = 2000,
               true_mean = 0, cor_y1y2 = 0.5,
[…]

### 2023-04-20 Simulations/sim_kim2021combining.R
Added simulation from Kim et al. (2021). Added papers to Reference folder.

@@ -1,0 +1,140 @@
+# Title: sim_kim2021combining.R
+# Author: Caleb Leedy
+# Date Created: April 20, 2023
+# Purpose: This document contains the code to reproduce the simulation study of
+# Kim et al. (2021) (see ../Reference/kim2021combining.pdf).
+
+# *************
+# * Libraries *
+# *************
+
+library(dplyr)
+library(purrr)
+
+library(doParallel)
+library(parallelly)
+library(doRNG)
+
+# ***************************
+# * Data Generating Process *
+# ***************************
+
+gen_data <- function(n_B, model_num) {
+
+  # Finite population
+  N <- 100000
+  x_vec <- rnorm(N, mean = 2, sd = 1)
+  e_vec <- rnorm(N)
+
+  if (model_num == 1) {
+    y_vec <- 1 + 2 * x_vec + e_vec
+  } else if (model_num == 2) {
+    y_vec <- 3 + x_vec + 2 * e_vec
+  } else if (model_num == 3) {
+    y_vec <- 2.5 + 0.5 * x_vec^2 + e_vec
+  } else {
+    stop("model_num must be 1, 2, or 3")
+  }
+
+  pop_df <- tibble(X = x_vec, Y = y_vec)
+
+  # Sample A
+  n_A <- 500
+  A_df <- slice_sample(pop_df, n = n_A)
+
+  # Sample B
+  strat1 <- filter(pop_df, X <= 2)
+  strat2 <- filter(pop_df, X > 2)
+  B_df <- bind_rows(slice_sample(strat1, n = 0.7 * n_B),
+                    slice_sample(strat2, n = 0.3 * n_B))
+
+
+  return(list(A = A_df, B = B_df))
+
+}
+
+# *************************
+# * Estimation Algorithms *
+# *************************
+
+mean_A <- function(data) {
+  mean(data$A$Y)
+}
+
+mean_B <- function(data) {
+  mean(data$B$Y)
+}
+
+mass_imp <- function(data) {
+
+  imp_mod <- lm(Y ~ X, data = data$B)
+  mean(predict(imp_mod, data$A))
+
+}
+
+chen_ipw <- function(data) {
+
+  A_df <- data$A
+  B_df <- data$B
+
+  return(1)
+  # FIXME:
+
+}
+
+# **************************
+# * Monte Carlo Experiment *
+# **************************
+
+run_sim <- function(n_B, model_number, seed) {
+
[…]

### 2023-04-24 Notes/main.tex
Added optimal a_2 b_2 theory to notes.

--- 2023-04-20 Notes/nonmonotone_project.tex
@@
 \newcommand{\Cov}{{\text{Cov}}}
+\newcommand{\Var}{{\text{Var}}}
 \newcommand{\logistic}{{\text{logistic}}}
@@
     and $Y_1$ and $Y_2$. In this case $\hat \theta_{eff}$ is not unbiased if the
-    reponse model is misspecified (or even just estimated).
+    response model is misspecified (or even just estimated).
 
   \item If we modify Simulation 4, such that the second step of observed the
-    second varible is proportional to $\logistic(y_k)$ then we get the same
+    second variable is proportional to $\logistic(y_k)$ then we get the same
     result as before:
@@
 
-\section*{Proposal: Full Nonmonotone Estimator}
+\section*{Minimizing the Variance}
 
 \begin{itemize}
+  \item The goal of this section is to find optimal values of $b_2(X, Y_1)$ and
+    $a_2(X, Y_2)$ such that the variance of $\hat \theta_{eff}$ is minimized.
 
-  \item When constructing
-    estimates of the response model, we estimate $\pi_{11}$ using
-    $\pi_{A_2 = 1|A_1 = 1}$ and we never include information about $\pi_{A_1 =
-    1|A_2 = 1}$. I would like to create a full estimator where we are able to
-    use this information.
-
-  \item The current estimator is the following (See Slide 24 of Non-monotone
-    Presentation):
+  \item Recall:
 
     \begin{align*}
-      \hat \theta_{eff} &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \\
-      &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)} (b_2(X_i, Y_{1i}) - 
-      E[g_i \mid X_i]) \\
-      &+ n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)} (a_2(X_i, Y_{2i}) -
-      E[g_i \mid X_i]) \\
-      &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_1, Y_{1i})}(g_i 
-      - b_2(X_i, Y_{1i}) - a_2(X_i, Y_{2i}) + E[g_i \mid X_i])
+      \hat \theta_{eff} - \hat \theta_n 
+      &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \left(1 - \frac{R_{1i}}{\pi_{1+}}
+      - \frac{R_{2i}}{\pi_{2+}} + \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \\
+      &+ n^{-1} \sum_{i = 1}^n b_2(X_i, Y_{1i}) \left(\frac{R_{1i}}{\pi_{1+}} -
+      \frac{R_{1i}R_{2i}}{\pi_{11}}\right)\\
+      &+ n^{-1} \sum_{i = 1}^n a_2(X_i, Y_{2i}) \left(\frac{R_{2i}}{\pi_{2+}} -
+      \frac{R_{1i}R_{2i}}{\pi_{11}}\right)\\
+      &+ n^{-1} \sum_{i = 1}^n g_i \left(\frac{R_{1i}R_{2i}}{\pi_{11}} -
+      1\right) \\
+      &\equiv A + B + C + D.
     \end{align*}
 
-  \item The problem with this is that we assume $\pi_{11}(X, Y_{1}) =
-    \pi_{2|1} (X, Y_{1}) \pi_{1+}(X)$ when we could have $\pi_{11}(X, Y_2) =
-    \pi_{1|2} (X, Y_2) \pi_{2+}(X)$. In other words, the previous result
-    implicitly assumes that $\pi_{11}$ is a function of $X$ and $Y_1$ when it
-    could be a function of $X$ and $Y_2$ as well.
+  \item Notice that we will suppress the fact that response models are functions
+    of $X$ (i.e. we write $\pi_{11}$ instead of $\pi_{11}(X)$).
 
-  \item I think that we could use the
-    following (small modification) instead,
+  \item To compute the variance, we first solve for each covariance combination.
+    Basically, all these computations rely of the following ideas. First, we
+    assume that the response model is correctly specified. Consequently,
+    $E[A] = E[B] = E[C] = E[D] = 0$ and things work out better. This helps when
+    we take the covariance conditional on $X$ because the inner expectations are
+    zero. The second key insight it to notice that $E[R_{j}^k] = E[R_j]$ for $j
+    \in \{1, 2\}$ and $k \in \mathbb{N}$. This is because $R$ is a binary
+    variable. Third, since we assume that the response models are correctly
+    specified, we have $E[R_1 \mid X] = \pi_{1+}$, $E[R_2 \mid X] = \pi_{2+}$,
+    and $E[R_1, R_2 \mid X] = \pi_{11}$.
 
-    \[\pi_{11}(X, Y_1, Y_2) = \alpha \pi_{1|2} \pi_{2+} + (1 - \alpha) \pi_{2|1}
-    \pi_{1+}\]
-
-    for some $\alpha \in [0, 1]$.
-
-  \item The reasoning behind this is two-fold. First, it allows us to think of
-    the nonmonotone missingness case as a linear combination of two monotone
-    cases, which I think is useful. Second, we can make explicit our choice of
-    $\alpha$, which in the existing estimator is simply $\alpha = 0$.
-
-  \item The problem with adding another parameter $\alpha$ is that it makes the
-    overall model unidentifiable. We cannot estimate $\alpha$ and the
-    conditional and marginal distributions that we have without additional
-    assumptions. So for now, I think that we should just assume that $\alpha$ is
-    known and then we apply this method to a data integration problem we can
-    review what $\alpha$ makes sense or figure out an additional assumption that
-    can help us estimate $\alpha$ (for example if there is a variable correlated
-    with $1|2$ versus $2|1$).
-
-  \item This is not just a pure Bayes' rule.
-    Formally, let $L = (X, Y_1, Y_2)$. By Bayes' rule we have
+    The overall approach to each of this computations is the following:
+    (1) take conditional expectations with respect to $X$ (the $\Cov(E[\cdot])$
+    term is zero), (2) expand the covariance to $E[XY] - E[X]E[Y]$ (the second
+    term is also zero), (3) by the MAR assumption $g, a_2, b_2$ are independent
+    of $R_1$ and $R_2$ and we can take the latter out of the expectation, (4)
+    evaluate and simplify expressions involving $E[R]$.
 
     \begin{align*}
-      \Pr(R_1 &= 1 \mid R_2 = 1, L) \Pr(R_2 = 1 \mid L) \\
-      &= \Pr(R_1 = 1, R_2 = 1 \mid L) \\
-      & = \Pr(R_2 = 1 \mid R_1 = 1, L) \Pr(R_1 = 1 \mid L).
+      \Cov(A, B) &= n^{-2} \sum_{i = 1}^n E\left[\Cov\left(E[g_i \mid X] \left(1
+      - \frac{R_{1i}}{\pi_{1+}} - \frac{R_{2i}}{\pi_{2+}} +
+      \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \mid X_i,\right. \right.\\
+      &\qquad \qquad \qquad \qquad b_2(X_i, Y_{1i}) \left. \left.
+      \left(\frac{R_{1i}}{\pi_{1+}} - \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \mid
+      X_i\right)\right]\\
+      &= n^{-1} E\left[E\left[E[g \mid X] \left(1 - \frac{R_{1i}}{\pi_{1+}} -
+      \frac{R_{2i}}{\pi_{2+}} + \frac{R_{1i}R_{2i}}{\pi_{11}}\right)
+      b_2(X_i, Y_{1i}) \left(\frac{R_{1i}}{\pi_{1+}} -
+      \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \mid X \right]\right] \\
+      &= n^{-1} E\left[E[g \mid X] E[b_2(X, Y_1) \mid X]
+      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
+      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right].
     \end{align*}
 
-    Yet to estimate this model, we need to assume something like the following:
+    By symmetry,
     \begin{align*}
-      \Pr(R_1 &= 1 \mid R_2 = 1, X, Y_2) \Pr(R_2 = 1 \mid X, Y_2)\\
-      &= \Pr(R_1 = 1, R_2 = 1 \mid L) \\
-      &= \Pr(R_2 = 1 \mid R_1 = 1, X, Y_1) \Pr(R_1 = 1 \mid X, Y_1)
+      \Cov(A, C) 
+      &= n^{-1} E\left[E[g \mid X] E[a_2(X, Y_2) \mid X]
+      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
+      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right].
     \end{align*}
 
-    in which case the two sides are clearly different.
+    \begin{align*}
+      \Cov(A, D) &= n^{-1} E\left[E\left[E[g \mid X] \left(1 -
+      \frac{R_{1i}}{\pi_{1+}} - \frac{R_{2i}}{\pi_{2+}} +
+      \frac{R_{1i}R_{2i}}{\pi_{11}}\right) g \left(\frac{R_1 R_2}{\pi_{11}} -
+      1\right) \mid X\right] \right]\\
+      &= n^{-1} E\left[E[g \mid X]^2 \left(\frac{-1}{\pi_{1+}} -
+      \frac{1}{\pi_{2+}} + 2\right)\right].
+    \end{align*}
 
-    \item Unfortunately, early simulation results are not promising because we
-      cannot distinguish points that should have $\pi_{11}$ conditional on $Y_1$
-      and points that should be conditional on $Y_2$.
+    \begin{align*}
+      \Cov(B, C) &= n^{-1} E\left[E[b_2(X, Y_1) \mid X] E[a_2(X, Y_2) \mid X]
+      E\left[\left(\frac{R_1}{\pi_{1+}} - \frac{R_1 R_2}{\pi_{11}}\right)
+      \left(\frac{R_2}{\pi_{2+}} - \frac{R_1 R_2}{\pi_{11}}\right) \mid
+      X\right]\right]\\
+      &= n^{-1} E\left[E[b_2(X, Y_1) \mid X] E[a_2(X, Y_2) \mid X]
+      \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} - \frac{1}{\pi_{1+}} -
+      \frac{1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right)\right].
+    \end{align*}
 
-      \input{Tables/alphasim.tex}
+    \begin{align*}
+      \Cov(B, D) &= n^{-1} E\left[ E[b_2(X, Y_1) \mid X] E[g \mid X]
+      E\left[\left(\frac{R_1}{\pi_{1+}} - \frac{R_1 R_2}{\pi_{11}}\right)
+      \left(\frac{R_1 R_2}{\pi_{11}} - 1\right) \mid X\right] \right]\\
+      &= n^{-1} E\left[ E[b_2(X, Y_1) \mid X] E[g \mid X]
+      \left(\frac{1}{\pi_{1+}} - \frac{1}{\pi_{11}}\right)\right].
+    \end{align*}
 
-    \item Dr. Kim, please let me know what you think about this idea and if it
-      is worth pursuing more.
+    By symmetry,
+    \begin{align*}
+      \Cov(C, D) 
+      &= n^{-1} E\left[ E[a_2(X, Y_2) \mid X] E[g \mid X]
+      \left(\frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}}\right)\right].
+    \end{align*}
+
+    We also compute the variance terms,
+
+    \begin{align*}
+      \Cov(A, A) &= n^{-1} E\left[ E\left[ E[g \mid X]^2 \left(1
+      - \frac{R_{1i}}{\pi_{1+}} - \frac{R_{2i}}{\pi_{2+}} +
+      \frac{R_{1i}R_{2i}}{\pi_{11}}\right)^2 \mid X\right] \right]\\
+      &= n^{-1} E \left[ E[g \mid X]^2 \left(-1 + \frac{2 \pi_{11}}{\pi_{1+}
+      \pi_{2+}} - \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
+      \frac{1}{\pi_{11}}\right)\right].
+    \end{align*}
+
+    \begin{align*}
+      \Cov(B, B) &= n^{-1} E\left[ E\left[b(X, Y_1)^2 \left(\frac{R_1}{\pi_{1+}}
+      - \frac{R_1 R_2}{\pi_{11}}\right)^2 \mid X\right] \right]\\
+      &= n^{-1} E\left[ E[b_2(X, Y_1)^2 \mid X] \left(\frac{-1}{\pi_{1+}} +
+      \frac{1}{\pi_{11}}\right)\right].
+    \end{align*}
+
+    \begin{align*}
+      \Cov(C, C) &= n^{-1} E\left[ E[a_2(X, Y_2)^2 \mid X]
+      \left(\frac{-1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right) \right].
+    \end{align*}
+
+    \begin{align*}
+      \Cov(D, D) &= n^{-1} E\left[ E\left[g_i^2 \left(\frac{R_1 R_2}{\pi_{11}} -
+      1\right)^2 \mid X\right] \right]\\
+      &= n^{-1} E\left[ E[g^2 \mid X] \left(\frac{1}{\pi_{11}} -
+      1\right)\right].
+    \end{align*}
+
+    This means that
+    \begin{align*}
+      \Var(\hat \theta_{eff} - \hat \theta_n) 
+      &= \Cov(A, A) + 2 \Cov(A, B) + 2 \Cov(A, C) + 2 \Cov(A, D) + \Cov(B, B) \\
+      &+ 2 \Cov(B, C) + 2 \Cov(B, D) + \Cov(C, C) + 2 \Cov(C, D) + \Cov(D, D) \\
+      &= n^{-1} E \left[ E[g \mid X]^2 \left(-1 + \frac{2 \pi_{11}}{\pi_{1+}
+      \pi_{2+}} - \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
+      \frac{1}{\pi_{11}}\right)\right] \\
+      &+ 2 n^{-1} E\left[E[g \mid X] E[b_2(X, Y_1) \mid X]
+      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
+      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right]\\
+      &+ 2 n^{-1} E\left[E[g \mid X] E[a_2(X, Y_2) \mid X]
+      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
+      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right]\\
+      &+ 2 n^{-1} E\left[E[g \mid X]^2 \left(\frac{-1}{\pi_{1+}} -
+      \frac{1}{\pi_{2+}} + 2\right)\right]\\
+      &+ n^{-1} E\left[ E[b_2(X, Y_1)^2 \mid X] \left(\frac{-1}{\pi_{1+}} +
+      \frac{1}{\pi_{11}}\right)\right]\\
+      &+ 2 n^{-1} E\left[E[b_2(X, Y_1) \mid X] E[a_2(X, Y_2) \mid X]
+      \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} - \frac{1}{\pi_{1+}} -
+      \frac{1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right)\right]\\
+      &+ 2 n^{-1} E\left[ E[b_2(X, Y_1) \mid X] E[g \mid X]
+      \left(\frac{1}{\pi_{1+}} - \frac{1}{\pi_{11}}\right)\right]\\
+      &+  n^{-1} E\left[ E[a_2(X, Y_2)^2 \mid X]
+      \left(\frac{-1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right) \right]\\
+      &+ 2 n^{-1} E\left[ E[a_2(X, Y_2) \mid X] E[g \mid X]
+      \left(\frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}}\right)\right]\\
+      &+ n^{-1} E\left[ E[g^2 \mid X] \left(\frac{1}{\pi_{11}} -
+      1\right)\right].
+    \end{align*}
+
+    Differentiating yields:
+
+    \begin{align*}
+      \frac{\partial}{\partial a_2} \Var(\hat \theta_{eff} - \hat \theta_n)
+      &= E\left[E[g \mid X] \left(\frac{1}{\pi_{1+}} + \frac{2}{\pi_{2+}} -
+      \frac{2}{\pi_{11}} - \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right] \\
+      &+ E\left[E[b_2(X, Y_1) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
+      \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
+      \frac{1}{\pi_{11}}\right)\right]\\
+      &+ E\left[ E[a_2(X, Y_2) \mid X] \left(\frac{-1}{\pi_{2+}} +
+      \frac{1}{\pi_{11}}\right)\right]\\
+      &\equiv 0, \text{ and }\\
+      \frac{\partial}{\partial b_2} \Var(\hat \theta_{eff} - \hat \theta_n)
+      &= E\left[E[g \mid X] \left(\frac{2}{\pi_{1+}} + \frac{1}{\pi_{2+}} -
+      \frac{2}{\pi_{11}} - \frac{\pi_{11}}{\pi_{1+}\pi_{2+}}\right)\right]\\
+      &+ E\left[E[a_2(X, Y_2) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
+      \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
+      \frac{1}{\pi_{11}}\right)\right]\\
+      &+ E\left[ E[b_2(X, Y_2) \mid X] \left(\frac{-1}{\pi_{1+}} +
+      \frac{1}{\pi_{11}}\right)\right]\\
+      &\equiv 0. 
+    \end{align*}
+
+    Substitution shows that these constraints are equivalent to:
+
+    \begin{align*}
+      E&\left[E[g \mid X] \left(\frac{-1}{\pi_{1+}} +
+    \frac{1}{\pi_{2+}}\right)\right] + E\left[ E[b_2(X, Y_1) \mid X]
+    \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -\frac{1}{\pi_{2+}}\right)\right]\\
+      &- E\left[E[a_2(X, Y_2) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} - 
+    \frac{1}{\pi_{1+}}\right)\right] \equiv 0
+    \end{align*}
+
+    where is the same as,
+
+    \begin{align*}
+      E&\left[ E[b_2(X, Y_1) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
+    \frac{1}{\pi_{2+}}\right) \right] + E\left[E[g \mid X]
+      \left(\frac{1}{\pi_{2+}}\right)\right] \\
+      &=
+    E\left[ E[a_2(X, Y_2) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
+    \frac{1}{\pi_{1+}}\right) \right] + E\left[E[g \mid X]
+    \left(\frac{1}{\pi_{1+}}\right)\right].
+    \end{align*}
+
+  \item These constraints can be satisfied (this is sufficient but maybe not
+    necessary) if
+
+    \begin{align*}
+      E&\left[(E[b_2(X, Y_1) \mid X] - E[a_2(X, Y_2) \mid X])
+      \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right) \right] = 0\\
+      E&\left[\left(\frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}}\right)(E[a_2(X, Y_2)
+      \mid X] + E[b_2(X, Y_1) \mid X] - 2E[g \mid X])\right] = 0.
+    \end{align*}
 
 \end{itemize}
+
+\textbf{Questions for Dr. Kim}
+\begin{itemize}
+  \item Where do we go from here?
+  \item This is a functional expectation that I need to calibrate. Should I use
+    some sort of basis spline to try to estimate these expectations and then set
+    them equal to zero? I don't think that I have worked with this kind of
+    constraint before.
+\end{itemize}
+%\input{prop_full_nonmonoest.tex}
 

### 2023-04-24 Notes/prop_full_nonmonoest.tex
Added optimal a_2 b_2 theory to notes.

--- 2023-04-18 Notes/nonmonotone_project.tex
@@
-\documentclass[12pt]{article}
-
-\usepackage{amsmath, amssymb, mathrsfs, fancyhdr}
-\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
-\usepackage{biblatex}
-\usepackage{booktabs}
-\usepackage{float}
-
-\addbibresource{references.bib}
-
-\hypersetup{colorlinks = true, urlcolor = black}
-
-\headheight     15pt
-\topmargin      -1.5cm   % read Lamport p.163
-\oddsidemargin  -0.04cm  % read Lamport p.163
-\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
-\textwidth      16.59cm
-\textheight     23.94cm
-\parskip         7.2pt   % sets spacing between paragraphs
-\parindent         0pt   % sets leading space for paragraphs
-\pagestyle{empty}        % Uncomment if don't want page numbers
-\pagestyle{fancyplain}
-
-\newcommand{\MAP}{{\text{MAP}}}
-\newcommand{\argmax}{{\text{argmax}}}
-\newcommand{\argmin}{{\text{argmin}}}
-\newcommand{\Cov}{{\text{Cov}}}
-\newcommand{\logistic}{{\text{logistic}}}
-
-\begin{document}
-
-%\lhead{Caleb Leedy}
-\chead{Nonmonotone Missingness}
-%\chead{STAT 615 - Advanced Bayesian Methods}
-%\rhead{Page \thepage\ of \pageref{LastPage}}
-\rhead{April 4, 2023}
-
-\section*{Overview:}
-
-The goal of this project is to outperform existing techniques in the literature
-related to nonmonotone missing data.
-
-\section*{Initial Simulations:}
-% Tables generated from Simulations/nonmonotone.R
-
-\begin{itemize}
-  \item \textbf{Implemented simulation of monotone MAR data}: This is 
-  correspondingly easier
-  than the subsequent nonmonotone MAR simulation. For this simulation we use the 
-  following approach:
-  \begin{enumerate}
-      \item Generate $X$, $Y_1$, and $Y_2$ for elements $i = 1, \dots, n$.
-      \item Using the covariate $X$, determine the probability $p_1$ of $Y_1$
-      being observed for each element $i$.
-      \item Based on $p_1$, determine if $R_1 = 1$.
-      \item If $R_1 = 0$, then $R_2 = 0$. Otherwise, using variables $X$ and
-        $Y_1$, determine the probability $p_{12}$.
-      \item Based on $p_{12}$ determine if $R_2 = 1$.
-  \end{enumerate}
-  At the end of the algorithm, we have determined the values of binary variables
-    $R_1$ and $R_2$ for each $i$ and if either of them are equal to 1, the
-    corresponding level of $Y_{k}$. As is common in this literature, the values
-    of $R_1$ and $R_2$ determine if the corresponding variable $Y_1$ or $Y_2$ is
-    missing or observed with $R = 1$ indicating $Y$ being observed.
-
-  \item \textbf{Implemented simulation of nonmonotone MAR data}: 
-  Following the approach
-    of \cite{robins1997non}, I construct a nonmonotone MAR simulation with two 
-    response variables $Y_1$ and $Y_2$ and one covariate $X$. The algorithm to
-    generate the data is the following:
-    \begin{enumerate}
-        \item Generate $X$, $Y_1$, and $Y_2$ for elements $i = 1, \dots, n$.
-        \item Using the covariate $X_i$, generate probabilities for each element
-          $i$ $p_0$, $p_1$, and $p_2$ such that $p_0 + p_1 + p_2 = 1$. 
-        \item Select one option based on the three probabilities for each
-          element $i$. If 0 is selected: $R_1 = 0$ and $R_2 = 0$. If 1 is
-          selected $R_1 = 1$. If 2 is selected, $R_2 = 1$.
-        \item We take the next step in multiple cases. If 0 was selected, we are
-          done. If 1 was selected, we generate probabilities $p_{12}$ based on
-          $X$ and $Y_1$. Then based on this probability, we determine if $R_2 =
-          1$. In the same manner, if 2 was selected in the previous step, we
-          generate probabilities $p_{21}$ based on $X$ and $Y_2$. Then based on
-          this probability, we determine if $R_2 = 1$.
-    \end{enumerate}
-    Like the monotone MAR simulation this algorithm produces similar final
-    results with the determination of binary variables $R_1$ and $R_2$ and
-    variables $X$, $Y_1$, and $Y_2$. Unlike the monotone MAR case, the
-    nonmonotone MAR includes observations with $Y_2$ observed and $Y_1$ missing.
-
-  \item \textbf{Simulation 1 with Monotone MAR}:
-    Following the algorithm described in the monotone MAR simulation bullet, 
-    we first generate data from the following distributions:
-    \begin{align*}
-        X_i &\stackrel{iid}{\sim} N(0, 1) \\
-        Y_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
-        Y_{2i} &\stackrel{iid}{\sim} N(\theta, 1)
-    \end{align*}
-
-    Then, we create the probabilities $p_1 = \logistic(x_i)$ and 
-    $p_{12} = \logistic(y_{1i})$.
-    Since, both $x_i$ and $y_1$ are standard normal distributions, each of these
-    probabilities is approximately $0.5$ in expectation.
-
-    The goal of this simulation is to estimate $\theta$. Alternatively, we can
-    express this as solving the estimating equation:
-
-    \[g(\theta) \equiv Y_2 - \theta = 0.\]
-
-    We estimate $\theta$ using the following procedures:
-
-    \begin{itemize}
-        \item Oracle: This computes $\bar Y$ using \textit{both} the observed
-          and missing data.
-        \item IPW-Oracle: This is an IPW estimator using only the observed
-          values of $Y_2$. The weights (inverse probabilities) use the actual
-          probabilities.
-        \item IPW-Est: This is an IPW estimator using the probabilities that
-          have been estimated by a logistic model.
-        \item Semi: This is the monotone semiparametric efficient estimator from
-          Slide 11 (Equation 2) of Dr. Kim's Nonmonotone Missingness
-          presentation.
-    \end{itemize}
-
-    We run this simulation with different values of $\theta$, sample size of
-    2000, and 2000 Monte Carlo replications. Each algorithm for each replication
-    generates $\hat \theta$. In the subsequent tables, we compute the bias,
-    standard deviation (sd), t-statistic (where we test for a significant
-    difference between the Monte Carlo mean $\hat \theta$ and the true $\theta$)
-    and the p-value of the t-statistic.
-
-    \newpage
-
-    \input{Tables/monomar_t-5.tex}
-    \input{Tables/monomar_t0.tex}
-    \input{Tables/monomar_t5.tex}
-
-    Overall, these results are mostly what I would have expected. All of the
-    algorithms estimate the true value of $\theta$ correctly in each case, with
-    the oracle estimate having the smallest variance followed by the
-    semiparametric algorithm. If there is anything surprising it is that the IPW
-    estimator has better performance with the estimated weights compared to the
-    true weights. However, I think that this is a known phenomenon.
-
-    \newpage
-
-    \item \textbf{Simulation 1 with Nonmonotone MAR}:
-
-    We generate variables $(X, Y_1, Y_2)$ using the following setup:
-
-    \[\begin{bmatrix}
-    X_i \\ \varepsilon_{1i} \\ \varepsilon_{2i}
-    \end{bmatrix} \stackrel{iid}{\sim}
-    N\left(
-    \begin{bmatrix}
-        0 \\ 0 \\ \theta
-    \end{bmatrix},
-    \begin{bmatrix}
-        1 & 0 & 0 \\
-        0 & 1 & \sigma_{yy}\\
-        0 & \sigma_{yy} & 1
-    \end{bmatrix}
-    \right).\]
-
-    Then,
-
-    \[y_{1i} = x_i + \varepsilon_{1i} \text{ and } 
-    y_{2i} = x_i + \varepsilon_{2i}.\]
-
-    Since we have nonmonotone data, our ``Stage 1'' probabilities are
-    different. We compute the true Stage 1 probabilities being proportional to
-    the following values:
-
-    \begin{align*}
-        p_0 &= 0.2 \\
-        p_1 &= 0.4 \\
-        p_2 &= 0.4 \\
-    \end{align*}
-
-   However, we keep the same structure for the Stage 2 probabilities with:
-   $p_{12} = \logistic(y_1)$ and $p_{21} = \logistic(y_2)$. The goal remains to
-   estimate $\theta$. We continue to use the Oracle algorithm and the IPW-Oracle
-   algorithm. Since we have nonmonotone MAR data, we use the ``Proposed''
-   algorithm that is described on Slide 25 (Equation 12) of Dr. Kim's
-   presentation. The outcome models were estimated using
-   logistic regression and OLS and correctly specified. The response model used
-   the oracle estimates of the probabilities. This yields the
-   following results:
-
-    \input{Tables/nonmonosim1m-5.tex}
-    \input{Tables/nonmonosim1m0.tex}
-    \input{Tables/nonmonosim1m5.tex}
-
-    \newpage
-
-    \item \textbf{Simulation 2 with Nonmonotone MAR}:
-    We also want to simulate data that is correlated. For this simulation, we
-    focus on $\Cov(Y_1, Y_2)$. The data generating process now has $\sigma_{yy}
-    \neq 0$. We are still interested in $\bar Y_2$ and we still run 2000
-    simulation with 2000 observations. In all of the next simulations the true
-    value of $\theta = 0$. The results are the following:
-
-    \input{Tables/nonmonosim2c0.1.tex}
-    \input{Tables/nonmonosim2c0.5.tex}
-    \input{Tables/nonmonosim2c0.9.tex}
-
-    \newpage
-
-  \item \textbf{Simulation 3 with Nonmonotone MAR}:
-    This simulation aims to see if the proposed algorithm is doubly robust.
-    First, we check with a misspecified outcome model. In this case the data
-    generating procedure is the following:
-
-    \[\begin{bmatrix}
-    X_i \\ \varepsilon_{1i} \\ \varepsilon_{2i}
-    \end{bmatrix} \stackrel{iid}{\sim}
-    N\left(
-    \begin{bmatrix}
-        0 \\ 0 \\ \theta
-    \end{bmatrix},
-    \begin{bmatrix}
-        1 & 0 & 0 \\
-        0 & 1 & \sigma_{yy}\\
-        0 & \sigma_{yy} & 1
-    \end{bmatrix}
-    \right).\]
-
-    Then,
-
-    \[y_{1i} = x_i + x_i^2 \varepsilon_{1i} \text{ and } 
-    y_{2i} = -x_i + x_i^3 + \varepsilon_{2i}.\]
-
-    This procedure causes $X$ to influence both $Y_1$ and $Y_2$ and we still
-    have correlation in the error terms of $Y_1$ and $Y_2$. However, since
-    neither $Y_1$ nor $Y_2$ are linear in $X$, the model will be misspecified.
-    The response mechanisms are first generated MCAR with a probability of
-    either $Y_1$ or $Y_2$ being the first variable observed to be $0.4$. (There
-    is a $0.2$ probability neither is observed.) Then the probability of the
-    other variable being observed is proportional to $\logistic(y_k)$ where
-    $y_k$ is the $y$ that has been observed. To ensure that the proposed method
-    has the correct propensity score we use the oracle probabilities instead of
-    estimating them. This yields the following:
-
-    \input{Tables/nonmonosim3c0.tex}
-    \input{Tables/nonmonosim3c0.1.tex}
-    \input{Tables/nonmonosim3c0.5.tex}
-
-    \newpage
-
-    Thus, the proposed method is unbiased with a misspecified outcome model.
-    We now show a simulation where the outcome model is correctly specified but
-    the response model is not.
-
-  \item \textbf{Simulation 4 with Nonmonotone MAR}:
-    Continuing to test if the proposed algorithm is doubly robust, this
-    simulation checks a misspecified response model. Instead of using oracle
-    weights as in Simulation 3, we estimate the weights for the proposed method.
-    The algorithms to which we compare still use the oracle weights.
-
-    \input{Tables/nonmonosim4c0.tex}
-    \input{Tables/nonmonosim4c0.1.tex}
-    \input{Tables/nonmonosim4c0.5.tex}
-
-    Thus, there is strong evidence that the proposed method is \textit{not}
-    doubly robust. If we have a misspecified response model, then it seems like
-    we induce bias. I will explore this more in the next section.
-
-\end{itemize}
-
-\newpage
-
-\section*{Estimating the Response Model}
-
-\begin{itemize}
-  \item In most (critically \textit{not} Simulation 4) of the simulations of the
-    previous section, we used the oracle weights when estimating our proposed
-    method. The reasoning for this was straightforward{--}we wanted to ensure
-    that the response model was correctly specified and the best case scenario
-    is to use the oracle weights which we did.
-
-  \item This section focuses more on estimating these response weights. Instead
-    of focusing on the proposed method, we will actually be working on
-    estimation of the complete case IPW estimator. This model is less complex
-    and will thus make it easier to understand where we are making mistakes.
-
-  \item We use the following simulation study:
-    \begin{align*}
-      x_i &\stackrel{iid}{\sim} N(0, 1)\\
-      \varepsilon_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
-      \varepsilon_{2i} &\stackrel{iid}{\sim} N(0, 1)\\
-      y_{1i} &= x_i + \varepsilon_{1i} \\
-      y_{2i} &= x_i + \varepsilon_{2i} \\
-    \end{align*}
-
-    To select a missingness pattern for each $i$, we have sequence: first, we
-    select the first variable to observe (or neither), then we either select the
-    second variable or we do not. In the first step, we select $R_{1i} = 1$ with
-    probability $0.4$, $R_{2i} = 1$ with probability $0.4$ and neither variable
-    with probability $0.2$. For the second step, we have the probability of
-    observing the other variable be $\logistic(x_i)$. This yields the following:
-
-    \input{Tables/ipwsim.tex}
-    % TODO: estimate IPW with \pi_{2|1} and \pi_{1|2}.
-
-    The problem with this is that our estimate is biased. For one particular
-    realization of the simulation, here is the distribution of the difference
-    between the estimated and true probabilities of $\pi_{11}$:
-
-    \includegraphics[width = 0.7\linewidth]{diffhist.png}
-
-\end{itemize}
-
-\newpage
-
 \section*{Proposal: Full Nonmonotone Estimator}
@@
 \end{itemize}
-
-\newpage
-
-\printbibliography
-
-\end{document}

### 2023-04-25 Notes/estimating_response_mod.tex
Fixed incorrect testing of MAR model in paper.

--- 2023-04-18 Notes/nonmonotone_project.tex
@@
-\documentclass[12pt]{article}
-
-\usepackage{amsmath, amssymb, mathrsfs, fancyhdr}
-\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
-\usepackage{biblatex}
-\usepackage{booktabs}
-\usepackage{float}
-
-\addbibresource{references.bib}
-
-\hypersetup{colorlinks = true, urlcolor = black}
-
-\headheight     15pt
-\topmargin      -1.5cm   % read Lamport p.163
-\oddsidemargin  -0.04cm  % read Lamport p.163
-\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
-\textwidth      16.59cm
-\textheight     23.94cm
-\parskip         7.2pt   % sets spacing between paragraphs
-\parindent         0pt   % sets leading space for paragraphs
-\pagestyle{empty}        % Uncomment if don't want page numbers
-\pagestyle{fancyplain}
-
-\newcommand{\MAP}{{\text{MAP}}}
-\newcommand{\argmax}{{\text{argmax}}}
-\newcommand{\argmin}{{\text{argmin}}}
-\newcommand{\Cov}{{\text{Cov}}}
-\newcommand{\logistic}{{\text{logistic}}}
-
-\begin{document}
-
-%\lhead{Caleb Leedy}
-\chead{Nonmonotone Missingness}
-%\chead{STAT 615 - Advanced Bayesian Methods}
-%\rhead{Page \thepage\ of \pageref{LastPage}}
-\rhead{April 4, 2023}
-
-\section*{Overview:}
-
-The goal of this project is to outperform existing techniques in the literature
-related to nonmonotone missing data.
-
-\section*{Initial Simulations:}
-% Tables generated from Simulations/nonmonotone.R
-
-\begin{itemize}
-  \item \textbf{Implemented simulation of monotone MAR data}: This is 
-  correspondingly easier
-  than the subsequent nonmonotone MAR simulation. For this simulation we use the 
-  following approach:
-  \begin{enumerate}
-      \item Generate $X$, $Y_1$, and $Y_2$ for elements $i = 1, \dots, n$.
-      \item Using the covariate $X$, determine the probability $p_1$ of $Y_1$
-      being observed for each element $i$.
-      \item Based on $p_1$, determine if $R_1 = 1$.
-      \item If $R_1 = 0$, then $R_2 = 0$. Otherwise, using variables $X$ and
-        $Y_1$, determine the probability $p_{12}$.
-      \item Based on $p_{12}$ determine if $R_2 = 1$.
-  \end{enumerate}
-  At the end of the algorithm, we have determined the values of binary variables
-    $R_1$ and $R_2$ for each $i$ and if either of them are equal to 1, the
-    corresponding level of $Y_{k}$. As is common in this literature, the values
-    of $R_1$ and $R_2$ determine if the corresponding variable $Y_1$ or $Y_2$ is
-    missing or observed with $R = 1$ indicating $Y$ being observed.
-
-  \item \textbf{Implemented simulation of nonmonotone MAR data}: 
-  Following the approach
-    of \cite{robins1997non}, I construct a nonmonotone MAR simulation with two 
-    response variables $Y_1$ and $Y_2$ and one covariate $X$. The algorithm to
-    generate the data is the following:
-    \begin{enumerate}
-        \item Generate $X$, $Y_1$, and $Y_2$ for elements $i = 1, \dots, n$.
-        \item Using the covariate $X_i$, generate probabilities for each element
-          $i$ $p_0$, $p_1$, and $p_2$ such that $p_0 + p_1 + p_2 = 1$. 
-        \item Select one option based on the three probabilities for each
-          element $i$. If 0 is selected: $R_1 = 0$ and $R_2 = 0$. If 1 is
-          selected $R_1 = 1$. If 2 is selected, $R_2 = 1$.
-        \item We take the next step in multiple cases. If 0 was selected, we are
-          done. If 1 was selected, we generate probabilities $p_{12}$ based on
-          $X$ and $Y_1$. Then based on this probability, we determine if $R_2 =
-          1$. In the same manner, if 2 was selected in the previous step, we
-          generate probabilities $p_{21}$ based on $X$ and $Y_2$. Then based on
-          this probability, we determine if $R_2 = 1$.
-    \end{enumerate}
-    Like the monotone MAR simulation this algorithm produces similar final
-    results with the determination of binary variables $R_1$ and $R_2$ and
-    variables $X$, $Y_1$, and $Y_2$. Unlike the monotone MAR case, the
-    nonmonotone MAR includes observations with $Y_2$ observed and $Y_1$ missing.
-
-  \item \textbf{Simulation 1 with Monotone MAR}:
-    Following the algorithm described in the monotone MAR simulation bullet, 
-    we first generate data from the following distributions:
-    \begin{align*}
-        X_i &\stackrel{iid}{\sim} N(0, 1) \\
-        Y_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
-        Y_{2i} &\stackrel{iid}{\sim} N(\theta, 1)
-    \end{align*}
-
-    Then, we create the probabilities $p_1 = \logistic(x_i)$ and 
-    $p_{12} = \logistic(y_{1i})$.
-    Since, both $x_i$ and $y_1$ are standard normal distributions, each of these
-    probabilities is approximately $0.5$ in expectation.
-
-    The goal of this simulation is to estimate $\theta$. Alternatively, we can
-    express this as solving the estimating equation:
-
-    \[g(\theta) \equiv Y_2 - \theta = 0.\]
-
-    We estimate $\theta$ using the following procedures:
-
-    \begin{itemize}
-        \item Oracle: This computes $\bar Y$ using \textit{both} the observed
-          and missing data.
-        \item IPW-Oracle: This is an IPW estimator using only the observed
-          values of $Y_2$. The weights (inverse probabilities) use the actual
-          probabilities.
-        \item IPW-Est: This is an IPW estimator using the probabilities that
-          have been estimated by a logistic model.
-        \item Semi: This is the monotone semiparametric efficient estimator from
-          Slide 11 (Equation 2) of Dr. Kim's Nonmonotone Missingness
-          presentation.
-    \end{itemize}
-
-    We run this simulation with different values of $\theta$, sample size of
-    2000, and 2000 Monte Carlo replications. Each algorithm for each replication
-    generates $\hat \theta$. In the subsequent tables, we compute the bias,
-    standard deviation (sd), t-statistic (where we test for a significant
-    difference between the Monte Carlo mean $\hat \theta$ and the true $\theta$)
-    and the p-value of the t-statistic.
-
-    \newpage
-
-    \input{Tables/monomar_t-5.tex}
-    \input{Tables/monomar_t0.tex}
-    \input{Tables/monomar_t5.tex}
-
-    Overall, these results are mostly what I would have expected. All of the
-    algorithms estimate the true value of $\theta$ correctly in each case, with
-    the oracle estimate having the smallest variance followed by the
-    semiparametric algorithm. If there is anything surprising it is that the IPW
-    estimator has better performance with the estimated weights compared to the
-    true weights. However, I think that this is a known phenomenon.
-
-    \newpage
-
-    \item \textbf{Simulation 1 with Nonmonotone MAR}:
-
-    We generate variables $(X, Y_1, Y_2)$ using the following setup:
-
-    \[\begin{bmatrix}
-    X_i \\ \varepsilon_{1i} \\ \varepsilon_{2i}
-    \end{bmatrix} \stackrel{iid}{\sim}
-    N\left(
-    \begin{bmatrix}
-        0 \\ 0 \\ \theta
-    \end{bmatrix},
-    \begin{bmatrix}
-        1 & 0 & 0 \\
-        0 & 1 & \sigma_{yy}\\
-        0 & \sigma_{yy} & 1
-    \end{bmatrix}
-    \right).\]
-
-    Then,
-
-    \[y_{1i} = x_i + \varepsilon_{1i} \text{ and } 
-    y_{2i} = x_i + \varepsilon_{2i}.\]
-
-    Since we have nonmonotone data, our ``Stage 1'' probabilities are
-    different. We compute the true Stage 1 probabilities being proportional to
-    the following values:
-
-    \begin{align*}
-        p_0 &= 0.2 \\
-        p_1 &= 0.4 \\
-        p_2 &= 0.4 \\
-    \end{align*}
-
-   However, we keep the same structure for the Stage 2 probabilities with:
-   $p_{12} = \logistic(y_1)$ and $p_{21} = \logistic(y_2)$. The goal remains to
-   estimate $\theta$. We continue to use the Oracle algorithm and the IPW-Oracle
-   algorithm. Since we have nonmonotone MAR data, we use the ``Proposed''
-   algorithm that is described on Slide 25 (Equation 12) of Dr. Kim's
-   presentation. The outcome models were estimated using
-   logistic regression and OLS and correctly specified. The response model used
-   the oracle estimates of the probabilities. This yields the
-   following results:
-
-    \input{Tables/nonmonosim1m-5.tex}
-    \input{Tables/nonmonosim1m0.tex}
-    \input{Tables/nonmonosim1m5.tex}
-
-    \newpage
-
-    \item \textbf{Simulation 2 with Nonmonotone MAR}:
-    We also want to simulate data that is correlated. For this simulation, we
-    focus on $\Cov(Y_1, Y_2)$. The data generating process now has $\sigma_{yy}
-    \neq 0$. We are still interested in $\bar Y_2$ and we still run 2000
-    simulation with 2000 observations. In all of the next simulations the true
-    value of $\theta = 0$. The results are the following:
-
-    \input{Tables/nonmonosim2c0.1.tex}
-    \input{Tables/nonmonosim2c0.5.tex}
-    \input{Tables/nonmonosim2c0.9.tex}
-
-    \newpage
-
-  \item \textbf{Simulation 3 with Nonmonotone MAR}:
-    This simulation aims to see if the proposed algorithm is doubly robust.
-    First, we check with a misspecified outcome model. In this case the data
-    generating procedure is the following:
-
-    \[\begin{bmatrix}
-    X_i \\ \varepsilon_{1i} \\ \varepsilon_{2i}
-    \end{bmatrix} \stackrel{iid}{\sim}
-    N\left(
-    \begin{bmatrix}
-        0 \\ 0 \\ \theta
-    \end{bmatrix},
-    \begin{bmatrix}
-        1 & 0 & 0 \\
-        0 & 1 & \sigma_{yy}\\
-        0 & \sigma_{yy} & 1
-    \end{bmatrix}
-    \right).\]
-
-    Then,
-
-    \[y_{1i} = x_i + x_i^2 \varepsilon_{1i} \text{ and } 
-    y_{2i} = -x_i + x_i^3 + \varepsilon_{2i}.\]
-
-    This procedure causes $X$ to influence both $Y_1$ and $Y_2$ and we still
-    have correlation in the error terms of $Y_1$ and $Y_2$. However, since
-    neither $Y_1$ nor $Y_2$ are linear in $X$, the model will be misspecified.
-    The response mechanisms are first generated MCAR with a probability of
-    either $Y_1$ or $Y_2$ being the first variable observed to be $0.4$. (There
-    is a $0.2$ probability neither is observed.) Then the probability of the
-    other variable being observed is proportional to $\logistic(y_k)$ where
-    $y_k$ is the $y$ that has been observed. To ensure that the proposed method
-    has the correct propensity score we use the oracle probabilities instead of
-    estimating them. This yields the following:
-
-    \input{Tables/nonmonosim3c0.tex}
-    \input{Tables/nonmonosim3c0.1.tex}
-    \input{Tables/nonmonosim3c0.5.tex}
-
-    \newpage
-
-    Thus, the proposed method is unbiased with a misspecified outcome model.
-    We now show a simulation where the outcome model is correctly specified but
-    the response model is not.
-
-  \item \textbf{Simulation 4 with Nonmonotone MAR}:
-    Continuing to test if the proposed algorithm is doubly robust, this
-    simulation checks a misspecified response model. Instead of using oracle
-    weights as in Simulation 3, we estimate the weights for the proposed method.
-    The algorithms to which we compare still use the oracle weights.
-
-    \input{Tables/nonmonosim4c0.tex}
-    \input{Tables/nonmonosim4c0.1.tex}
-    \input{Tables/nonmonosim4c0.5.tex}
-
-    Thus, there is strong evidence that the proposed method is \textit{not}
-    doubly robust. If we have a misspecified response model, then it seems like
-    we induce bias. I will explore this more in the next section.
-
-\end{itemize}
-
-\newpage
-
 \section*{Estimating the Response Model}
@@
 \newpage
-
-\section*{Proposal: Full Nonmonotone Estimator}
-
-\begin{itemize}
-
-  \item When constructing
-    estimates of the response model, we estimate $\pi_{11}$ using
-    $\pi_{A_2 = 1|A_1 = 1}$ and we never include information about $\pi_{A_1 =
-    1|A_2 = 1}$. I would like to create a full estimator where we are able to
-    use this information.
-
-  \item The current estimator is the following (See Slide 24 of Non-monotone
-    Presentation):
-
-    \begin{align*}
-      \hat \theta_{eff} &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \\
-      &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)} (b_2(X_i, Y_{1i}) - 
-      E[g_i \mid X_i]) \\
-      &+ n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)} (a_2(X_i, Y_{2i}) -
-      E[g_i \mid X_i]) \\
-      &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_1, Y_{1i})}(g_i 
-      - b_2(X_i, Y_{1i}) - a_2(X_i, Y_{2i}) + E[g_i \mid X_i])
-    \end{align*}
-
-  \item The problem with this is that we assume $\pi_{11}(X, Y_{1}) =
-    \pi_{2|1} (X, Y_{1}) \pi_{1+}(X)$ when we could have $\pi_{11}(X, Y_2) =
-    \pi_{1|2} (X, Y_2) \pi_{2+}(X)$. In other words, the previous result
-    implicitly assumes that $\pi_{11}$ is a function of $X$ and $Y_1$ when it
-    could be a function of $X$ and $Y_2$ as well.
-
-  \item I think that we could use the
-    following (small modification) instead,
-
-    \[\pi_{11}(X, Y_1, Y_2) = \alpha \pi_{1|2} \pi_{2+} + (1 - \alpha) \pi_{2|1}
-    \pi_{1+}\]
-
-    for some $\alpha \in [0, 1]$.
-
-  \item The reasoning behind this is two-fold. First, it allows us to think of
-    the nonmonotone missingness case as a linear combination of two monotone
-    cases, which I think is useful. Second, we can make explicit our choice of
-    $\alpha$, which in the existing estimator is simply $\alpha = 0$.
-
-  \item The problem with adding another parameter $\alpha$ is that it makes the
-    overall model unidentifiable. We cannot estimate $\alpha$ and the
-    conditional and marginal distributions that we have without additional
-    assumptions. So for now, I think that we should just assume that $\alpha$ is
-    known and then we apply this method to a data integration problem we can
-    review what $\alpha$ makes sense or figure out an additional assumption that
-    can help us estimate $\alpha$ (for example if there is a variable correlated
-    with $1|2$ versus $2|1$).
-
-  \item This is not just a pure Bayes' rule.
-    Formally, let $L = (X, Y_1, Y_2)$. By Bayes' rule we have
-
-    \begin{align*}
-      \Pr(R_1 &= 1 \mid R_2 = 1, L) \Pr(R_2 = 1 \mid L) \\
-      &= \Pr(R_1 = 1, R_2 = 1 \mid L) \\
-      & = \Pr(R_2 = 1 \mid R_1 = 1, L) \Pr(R_1 = 1 \mid L).
-    \end{align*}
-
-    Yet to estimate this model, we need to assume something like the following:
-    \begin{align*}
-      \Pr(R_1 &= 1 \mid R_2 = 1, X, Y_2) \Pr(R_2 = 1 \mid X, Y_2)\\
-      &= \Pr(R_1 = 1, R_2 = 1 \mid L) \\
-      &= \Pr(R_2 = 1 \mid R_1 = 1, X, Y_1) \Pr(R_1 = 1 \mid X, Y_1)
-    \end{align*}
-
-    in which case the two sides are clearly different.
-
-    \item Unfortunately, early simulation results are not promising because we
-      cannot distinguish points that should have $\pi_{11}$ conditional on $Y_1$
-      and points that should be conditional on $Y_2$.
-
-      \input{Tables/alphasim.tex}
-
-    \item Dr. Kim, please let me know what you think about this idea and if it
-      is worth pursuing more.
-
-\end{itemize}
-
-\newpage
-
-\printbibliography
-
-\end{document}

### 2023-04-25 Notes/main.tex
Fixed incorrect testing of MAR model in paper.

@@
 \begin{itemize}
-  \item \textbf{Implemented simulation of monotone MAR data}: This is 
-  correspondingly easier
-  than the subsequent nonmonotone MAR simulation. For this simulation we use the 
-  following approach:
+  \item \textbf{Implemented simulation of monotone MAR data}: 
+    This is correspondingly easier than the subsequent nonmonotone MAR
+    simulation. For this simulation we use the following approach:
   \begin{enumerate}
@@
   \end{enumerate}
-  At the end of the algorithm, we have determined the values of binary variables
-    $R_1$ and $R_2$ for each $i$ and if either of them are equal to 1, the
-    corresponding level of $Y_{k}$. As is common in this literature, the values
-    of $R_1$ and $R_2$ determine if the corresponding variable $Y_1$ or $Y_2$ is
-    missing or observed with $R = 1$ indicating $Y$ being observed.
+    At the end of the algorithm, we have determined the values of binary
+    variables $R_1$ and $R_2$ for each $i$ and if either of them are equal to 1,
+    the corresponding level of $Y_{k}$. As is common in this literature, the
+    values of $R_1$ and $R_2$ determine if the corresponding variable $Y_1$ or
+    $Y_2$ is missing or observed with $R = 1$ indicating $Y$ being observed.
 
   \item \textbf{Implemented simulation of nonmonotone MAR data}: 
-  Following the approach
-    of \cite{robins1997non}, I construct a nonmonotone MAR simulation with two 
-    response variables $Y_1$ and $Y_2$ and one covariate $X$. The algorithm to
-    generate the data is the following:
+    Following the approach of \cite{robins1997non}, I construct a nonmonotone
+    MAR simulation with two response variables $Y_1$ and $Y_2$ and one covariate
+    $X$. The algorithm to generate the data is the following:
     \begin{enumerate}
@@
 
-    Then,
+    Then, the true outcome model is,
 
@@
     $\logistic(y_k)$, this simulation has the true probabilities being
-    proportional to $\logistic(x_i)$.
+    proportional to $\logistic(x_i)$. Thus, the true response model is the
+    following:
+
+    \begin{enumerate}
+      \item (Stage 1) Choose a variable observe. We choose $Y_1$ with
+        probability $0.4$, $Y_2$ with probability $0.4$ and neither with
+        probability $0.2$. If neither, $R_1 = 0$ and $R_2 = 0$. Otherwise,
+        continue to Step 2.
+      \item (Stage 2) With probability $p_i \propto \logistic(x_i)$, choose to
+        observed the other $Y$ variable. 
+    \end{enumerate}
+
+    This sequence generates the missingness indicators $R_1$ and $R_2$. Since,
+    the Stage 1 probabilities are fixed and known and the Stage 2 probabilities
+    only depend on $x_i$, the missingness is MAR and only a function of $x_i$.
     The algorithms to which we compare still use the oracle weights.
@@
 
+    The previous version of this simulation used the Stage 2 probability $p_i
+    \propto \logistic(y_i)$ where $y_i$ was the observed $Y$ value in Stage 1.
+    However, under this setup, our method is biased. This is because
+
+    \[ E\left[E\left[\frac{R_1 R_2}{\pi_{11}(X, Y_1)} Y_1 \mid X\right]\right] 
+    \neq
+    E\left[\frac{E[Y_1 \mid X]}{\pi_{11}(X, Y_1)} E[R_1 R_2 \mid X]\right] \]
+
+    but
+
+    \[ E\left[E\left[\frac{R_1 R_2}{\pi_{11}(X)} Y_1 \mid X\right]\right] 
+    =
+    E\left[\frac{E[Y_1 \mid X]}{\pi_{11}(X)} E[R_1 R_2 \mid X]\right].\]
+
     Thus, there is strong evidence that the proposed method is 
-    doubly robust. Previously, we did not get this result and the reason is 
-    the way that the selection probabilities work. When they were proportional
-    to $\logistic(y_k)$, the estimator is biased (see the next section);
-    however, when the probabilities are proportional to $\logistic(x_i)$ then it
-    works.
-
+    doubly robust because it is robust to both misspecification in the outcome
+    and response model.
 \end{itemize}
@@
 
-\section*{Estimating the Response Model}
-
-\begin{itemize}
-  \item In most (critically \textit{not} Simulation 4) of the simulations of the
-    previous section, we used the oracle weights when estimating our proposed
-    method. The reasoning for this was straightforward{--}we wanted to ensure
-    that the response model was correctly specified and the best case scenario
-    is to use the oracle weights which we did.
-
-  \item This section focuses more on estimating these response weights. Instead
-    of focusing on the proposed method, we will actually be working on
-    estimation of the complete case IPW estimator. This model is less complex
-    and will thus make it easier to understand where we are making mistakes.
-
-  \item We use the following simulation study:
-    \begin{align*}
-      x_i &\stackrel{iid}{\sim} N(0, 1)\\
-      \varepsilon_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
-      \varepsilon_{2i} &\stackrel{iid}{\sim} N(0, 1)\\
-      y_{1i} &= x_i + \varepsilon_{1i} \\
-      y_{2i} &= x_i + \varepsilon_{2i} \\
-    \end{align*}
-
-    To select a missingness pattern for each $i$, we have sequence: first, we
-    select the first variable to observe (or neither), then we either select the
-    second variable or we do not. In the first step, we select $R_{1i} = 1$ with
-    probability $0.4$, $R_{2i} = 1$ with probability $0.4$ and neither variable
-    with probability $0.2$. For the second step, we have the probability of
-    observing the other variable be $\logistic(x_i)$. This yields the following:
-
-    \input{Tables/ipwsim.tex}
-    % TODO: estimate IPW with \pi_{2|1} and \pi_{1|2}.
-
-    The problem with this is that our estimate is biased. For one particular
-    realization of the simulation, here is the distribution of the difference
-    between the estimated and true probabilities of $\pi_{11}$:
-
-    \includegraphics[width = 0.7\linewidth]{diffhist.png}
-
-\end{itemize}
-
-\newpage
+% \input{estimating_response_mod.tex}
 

### 2023-04-25 Simulations/nonmonotone.R

@@ -469,15 +469,13 @@ ipw_cc_est <- function(df, oracle = FALSE, alpha = NA) {
 
 }
 
-png("diffhist.png")
-filter(df, r1 == 1, r2 == 1) %>%
-  dplyr::select(-p0, -p1, -p2, -p12, -p21) %>%
-  mutate(diff = pi_11 - true_pi11) %>%
-  pull(diff) %>% 
-  hist(main = "Histogram of Estimated - True") 
-dev.off()
-
-  
+# png("diffhist.png")
+# filter(df, r1 == 1, r2 == 1) %>%
+#   dplyr::select(-p0, -p1, -p2, -p12, -p21) %>%
+#   mutate(diff = pi_11 - true_pi11) %>%
+#   pull(diff) %>% 
+#   hist(main = "Histogram of Estimated - True") 
+# dev.off()
 
 twophase_reg <- function(df) {
 

### 2023-04-26 Notes/calibration_comparison.tex
Refactored Simulations/nonmonotone.R into Simulations/generate_data.R and Simulations/nmono_est_funs.R.

\documentclass[12pt]{article}

\usepackage{amsmath, amssymb, mathrsfs, fancyhdr}
\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
\usepackage{biblatex}
\usepackage{booktabs}
\usepackage{float}

\addbibresource{references.bib}

\hypersetup{colorlinks = true, urlcolor = black}

\headheight     15pt
\topmargin      -1.5cm   % read Lamport p.163
\oddsidemargin  -0.04cm  % read Lamport p.163
\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
\textwidth      16.59cm
\textheight     23.94cm
\parskip         7.2pt   % sets spacing between paragraphs
\parindent         0pt   % sets leading space for paragraphs
\pagestyle{empty}        % Uncomment if don't want page numbers
\pagestyle{fancyplain}

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}

\begin{document}

%\lhead{Caleb Leedy}
\chead{Nonmonotone Missingness}
%\chead{STAT 615 - Advanced Bayesian Methods}
%\rhead{Page \thepage\ of \pageref{LastPage}}
\rhead{April 4, 2023}

\section*{Comparison with Calibration Estimator}

% Outline
% * [ ] Regression estimator comparison with calibration in monotone case
% * [ ] Equivalent calibration in nonmonotone case

\subsection*{Monotone Case}

\begin{itemize}
  \item In the monotone case the efficient estimator is

    \begin{align*}
      \hat \theta_{eff} &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \\
      &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)}(
        E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i]) \\
      &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_i)} (
      E[g_i \mid X_i, Y_{1i}, Y_{2i}] - E[g_i \mid X_i, Y_{1i}]). 
    \end{align*}

  \item This should be very similar to the following calibration estimator, for
    $\sum_{i = 1}^n w_i y_{2i}$

    \begin{align*}
      \argmin_w \sum_{i = 1}^n w_i^2& \text{ such that }\\
      \sum_{i = 1}^n x_i &= \sum_{i = 1}^n R_{1i} w_{1i} x_i \\
      \sum_{i = 1}^n w_{1i} (x_i, y_{1i}) &= \sum_{i = 1}^n R_{1i} R_{2i} w_{2i}
      (x_i, y_{1i}) \\
    \end{align*}

  \item The reason that these should be the same is because they are similar in
    relationship to a calibration and regression estimator which are exactly the
    same.

  \item % TODO: Add simulation study results
\end{itemize}

\subsection*{Nonmonotone Case}

\begin{itemize}
  \item Similar to the monotone case, we have an idea of the efficient
    estimator. Now we want to show that it is similar to a calibration equation.
    Unlike the monotone case where $R_{1i} = 0$ implies $R_{2i} = 0$, the
    nonmonotone case does not have this relationship. Instead we believe that we
    have the following calibration equations:

    \begin{align*}
      \sum_{i = 1}^n E[g_i \mid X_i] &= \sum_{i = 1}^n R_{1i} w_{1i} E[g_i \mid
      X_i]\\
      \sum_{i = 1}^n E[g_i \mid X_i] &= \sum_{i = 1}^n R_{2i} w_{2i} E[g_i \mid
      X_i]\\
      \sum_{i = 1}^n R_{1i} w_{1i} E[g_i \mid X_i, Y_{1i}] &= \sum_{i = 1}^n
      R_{1i} R_{2i} w_{ci} E[g_i \mid X_i, Y_{1i}]\\
      \sum_{i = 1}^n R_{2i} w_{2i} E[g_i \mid X_i, Y_{1i}] &= \sum_{i = 1}^n
      R_{1i} R_{2i} w_{ci} E[g_i \mid X_i, Y_{2i}]\\
      \sum_{i = 1}^n E[g_i \mid X_i] &= \sum_{i = 1}^n R_{1i} R_{2i} w_{ci}
      E[g_i \mid X_i].
    \end{align*}

  \item % TODO: Add simulation results
\end{itemize}

\newpage
\end{document}

### 2023-04-26 Simulations/calibration_comp.R

@@ -1,0 +1,38 @@
+# Title: Simulations/calibration_comp.R
+# Author: Caleb Leedy
+# Date Created: April 25, 2023
+# Purpose: This script contains the code to compare the calibration estimators
+# with the regression estimators. The regression estimators can be originally
+# found in Simulations/nonmonotone.R but they have been imported here.
+
+# ******************
+# * Script Outline *
+# ******************
+# A. Monotone Case
+#
+# B. Nonmonotone Case
+
+# *************
+# * Libraries *
+# *************
+
+library(dplyr)
+library(CVXR)
+
+library(doParallel)
+library(doRNG)
+library(parallelly)
+
+# *****************************
+# * Data Generation Functions *
+# *****************************
+source("generate_data.R")
+
+# *************************************
+# * Non/monotone Estimation Functions *
+# *************************************
+source("nmono_est_funs.R")
+
+# ************************
+# * Monotone Calibration *
+# ************************

### 2023-04-26 Simulations/generate_data.R

@@ -1,0 +1,197 @@
+# Title: Simulations/generate_data.R
+# Author: Caleb Leedy
+# Date Created: April 25, 2023
+# Purpose: This file contains the code to generate data for different simulation
+# studies. Instead of writing the functions in Simulations/nonmonotone.R and
+# then copying them to other files, I thought that it would be a good idea to
+# put them all into one script and then source it.
+
+# ------------------------------------------------------------------------------
+
+# ********************
+# * Script Functions *
+# ********************
+# 1. mono_mar -- This generates monotone missingness.
+# 2. nonmono_mar -- This generates nonmonotone missing data.
+# 3. expit -- A helper function for generating probabilites. 
+
+# ------------------------------------------------------------------------------
+
+# **********************
+# * Required Libraries *
+# **********************
+
+require(MASS)
+require(dplyr)
+
+# ------------------------------------------------------------------------------
+
+# ************************
+# * Additional Functions *
+# ************************
+
+expit <- function(x) {1 / (1 + exp(-x))}
+
+# *****************
+# * Generate Data *
+# *****************
+
+mono_mar <- function(n, mean_y2 = 0) {
+
+  # Steps:
+  # 1. Generate X, Y_1, Y_2.
+  # 2. Get probabilities conditional on past.
+  # 3. Reveal hidden variables.
+
+  # 1. Generate X, Y_1, Y_2.
+  x_vec <- rnorm(n)
+  y1 <- rnorm(n)
+  y2 <- rnorm(n, mean = mean_y2)
+
+  # 2. Get probabilities conditional on past.
+  r1 <- rbinom(n, 1, expit(x_vec))
+  r2 <- rbinom(n, 1, expit(y1))
+
+  # 3. Reveal hidden variables.
+  tibble(x = x_vec, y1 = y1, y2 = y2, r1 = r1, r2 = r2) %>%
[…]

### 2023-04-26 Simulations/nmono_est_funs.R

@@ -1,0 +1,247 @@
+# Title: ../Simulations/nmono_est_funs.R
+# Author: Caleb Leedy
+# Date Created: April 25, 2023
+# Purpose: This script contains the functions used to estimate monotone and
+# nonmonotone missingness. Like Simulations/generate_data.R, I wanted to
+# separate the function estimation from the simulation study itself in order to
+# gain more modular code.
+
+# ------------------------------------------------------------------------------
+
+# ********************
+# * Script Functions *
+# ********************
+# A. Monotone estimators
+#   1. est_mono -- Classical monotone estimator
+#   2. ipw_mono_est -- IPW estimator with estimated weights
+# B. Nonmonotone estimators
+#   1. est_nonmono -- The proposed estimator in the nonmonotone setting
+#   2. ipw_cc_oracle -- IPW estimator with oracle weights
+#   3. ipw_cc_est -- IPW estimator with weights estimated
+#   4. twophase_reg -- Unfinished
+
+# ------------------------------------------------------------------------------
+
+# **********************
+# * Required Libraries *
+# **********************
+
+require(dplyr)
+
+# ------------------------------------------------------------------------------
+
+# ***********************************
+# * Estimating Monotone Missingness *
+# ***********************************
+
+est_mono <- function(df) {
+
+  # g(x, y1, y2) = y2
+  gfun <- "y2"
+
+  # Algorithm Steps:
+  # 1. Estimate pi_11, pi_1+
+  # 2. Estimate b1 and b2
+  # 3. Estimate \hat \theta
+
+  df_1 <- filter(df, r1 == 1)  # Contains data from first stage
+  df_11 <- filter(df, r2 == 1) # Contains data from first and second stage
+
+  # 1. Estimate pi_11, pi_1+
[…]

### 2023-04-26 Simulations/nonmonotone.R

@@ -46,243 +46,18 @@ library(doParallel)
 library(doRNG)
 library(parallelly)
 
-# ************************
-# * Additional Functions *
-# ************************
-
-expit <- function(x) {1 / (1 + exp(-x))}
-
 # *****************
 # * Generate Data *
 # *****************
 
-mono_mar <- function(n, mean_y2 = 0) {
+# Run from Simulations/
+source("generate_data.R")
 
-  # Steps:
-  # 1. Generate X, Y_1, Y_2.
-  # 2. Get probabilities conditional on past.
-  # 3. Reveal hidden variables.
+# ************************
+# * Estimating Functions *
+# ************************
 
-  # 1. Generate X, Y_1, Y_2.
-  x_vec <- rnorm(n)
-  y1 <- rnorm(n)
-  y2 <- rnorm(n, mean = mean_y2)
-
-  # 2. Get probabilities conditional on past.
-  r1 <- rbinom(n, 1, expit(x_vec))
-  r2 <- rbinom(n, 1, expit(y1))
-
-  # 3. Reveal hidden variables.
-  tibble(x = x_vec, y1 = y1, y2 = y2, r1 = r1, r2 = r2) %>%
-    mutate(r2 = ifelse(r1 == 0, 0, r2))
-
-}
-
-nonmono_mar <- 
-  function(n, mean_y2 = 0, cor_xy1 = 0, cor_xy2 = 0, cor_y1y2 = 0,
-           r_ind_y = FALSE, miss_out = FALSE, miss_resp = FALSE, mcar = FALSE) {
-
-  # Steps:
-  # 1. Generate X, Y_1, Y_2.
-  # 2. Get probabilities conditional on past.
-  # 3. Reveal hidden variables.
-
-  # 1. Generate X, Y_1, Y_2.
-  mean_vec <- c(0, 0, mean_y2)
-  Sigma_mat <- 
-    matrix(c(1, cor_xy1, cor_xy2, cor_xy1, 1, cor_y1y2, cor_xy2, cor_y1y2, 1),
-           nrow = 3)
-  data <- MASS::mvrnorm(n, mean_vec, Sigma_mat)
-  x_vec <- data[, 1]
-  y1 <- x_vec + data[, 2]
-  y2 <- x_vec + data[, 3]
-
-  # Misspecified outcome model
-  if (miss_out) {
[…]

### 2023-04-26 Simulations/calibration_comp.R
Added monotone calibration simulations. They seem to work well. I might want to set the seed in the future.

@@ -36,3 +36,111 @@ source("nmono_est_funs.R")
 # ************************
 # * Monotone Calibration *
 # ************************
+
+mono_est_weights <- function(df) {
+
+  # Steps:
+  # 1. Calibrate with w1 weights on Y_1
+  # 2. Calibrate with w2 weights on Y_2 using w1 weights on Y_1.
+
+  # 1. Calibrate with w1 weights on Y_1
+  n1 <- sum(df$r1)
+  w1 <- Variable(n1)
+  objective <- Minimize(sum_squares(w1))
+  Tx <- mean(df$x)
+  xr1_vec <-
+    filter(df, r1 == 1) %>%
+    pull(x)
+  constraints <- 
+    list(t(w1) %*% matrix(xr1_vec, ncol = 1) == Tx,
+         w1 >= 0,
+         t(w1) %*% matrix(rep(1, n1), ncol = 1) == 1)
+  p1 <- Problem(objective, constraints)
+  res1 <- solve(p1)
+
+  df$w1 <- 0
+  df$w1[df$r1 == 1] <- res1$getValue(w1)
+
+  # 2. Calibrate with w2 weights on Y_2 using w1 weights on Y_1.
+  n2 <- sum(df$r2) # Only works because of monotone.
+  w2 <- Variable(n2)
+  objective <- Minimize(sum_squares(w2))
+  Exw <- sum(df$w1 * df$x)
+  Ey1w <- sum(df$w1 * df$y1)
+  xr2_vec <- filter(df, r2 == 1) %>% pull(x)
+  y1r2_vec <- filter(df, r2 == 1) %>% pull(y1)
+  constraints <-
+    list(w2 >= 0,
+         t(w2) %*% matrix(rep(1, n2), ncol = 1) == 1,
+         t(w2) %*% matrix(xr2_vec, ncol = 1) == Exw,
+         t(w2) %*% matrix(y1r2_vec, ncol = 1) == Ey1w)
+  p2 <- Problem(objective, constraints)
+  res2 <- solve(p2)
+
+  df$w2 <- 0
+  df$w2[df$r2 == 1] <- res2$getValue(w2)
+
+  # Objective is E[g] = E[y_2]
+  sum(df$w2 * df$y2)
+
+}
+
+# *****************************
+# * Monotone Simulation Study *
+# *****************************
+
+clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
[…]

### 2023-04-27 Notes/calibration_comparison.tex
Added nonmonotone calibration simulations. The results look pretty good.

@@
-\documentclass[12pt]{article}
-
-\usepackage{amsmath, amssymb, mathrsfs, fancyhdr}
-\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
-\usepackage{biblatex}
-\usepackage{booktabs}
-\usepackage{float}
-
-\addbibresource{references.bib}
-
-\hypersetup{colorlinks = true, urlcolor = black}
-
-\headheight     15pt
-\topmargin      -1.5cm   % read Lamport p.163
-\oddsidemargin  -0.04cm  % read Lamport p.163
-\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
-\textwidth      16.59cm
-\textheight     23.94cm
-\parskip         7.2pt   % sets spacing between paragraphs
-\parindent         0pt   % sets leading space for paragraphs
-\pagestyle{empty}        % Uncomment if don't want page numbers
-\pagestyle{fancyplain}
-
-\newcommand{\MAP}{{\text{MAP}}}
-\newcommand{\argmax}{{\text{argmax}}}
-\newcommand{\argmin}{{\text{argmin}}}
-\newcommand{\Cov}{{\text{Cov}}}
-\newcommand{\Var}{{\text{Var}}}
-\newcommand{\logistic}{{\text{logistic}}}
-
-\begin{document}
-
-%\lhead{Caleb Leedy}
-\chead{Nonmonotone Missingness}
-%\chead{STAT 615 - Advanced Bayesian Methods}
-%\rhead{Page \thepage\ of \pageref{LastPage}}
-\rhead{April 4, 2023}
-
 \section*{Comparison with Calibration Estimator}
@@
 % Outline
-% * [ ] Regression estimator comparison with calibration in monotone case
-% * [ ] Equivalent calibration in nonmonotone case
+% * [X] Regression estimator comparison with calibration in monotone case
+% * [X] Equivalent calibration in nonmonotone case
 
@@
 
-  \item % TODO: Add simulation study results
+  \item To test the idea that the monotone regression estimator is similar to
+    the calibration estimator we run several simulation studies. In the monotone
+    case data is generating in the following steps:
+
+    \begin{enumerate}
+      \item The variables $X$, $Y_1$, and $Y_2$ are simulated from the following
+        distributions:
+        \begin{align*}
+          X_i &\stackrel{iid}{\sim} N(0, 1) \\
+          Y_{1i} &\stackrel{iid}{\sim} N(0, 1) \\
+          Y_{2i} &\stackrel{iid}{\sim} N(\theta, 1).
+        \end{align*}
+
+      \item After the variables have been simulated, we see which variables are
+        observed. We always observe $X_i$. We observed $Y_1$ with
+        probability $p_{1i} \propto \logistic(x_i)$. If $Y_{1i}$ is observed,
+        then we observe $Y_{2i}$ with probability $p_{2i} \propto
+        \logistic(y_{1i})$. If $Y_{1i}$ is not observed, we do not observe
+        $Y_{2i}$.
+    \end{enumerate}
+
+  \item The goal of this simulation study is the estimate $\theta = E[Y_2]$. We
+    use the previous monotone data generating process with different true values
+    of $\theta$ and compute the bias, standard deviation, T-statistic and
+    p-value. (The T-statistic and p-value test if the estimated value of $\hat
+    \theta$ is significantly different from the true value of $\theta$.)
+
+    \input{Tables/calimono_t-5.tex}
+    \input{Tables/calimono_t0.tex}
+    \input{Tables/calimono_t5.tex}
+
 \end{itemize}
+
+\newpage
 
@@
 
-  \item % TODO: Add simulation results
+  \item We still have the same goal of the simulation study: estimate $\theta =
+    E[Y_2]$. We use the previous nonmonotone data generating process with
+    different true values of $\theta$ to estimate the bias, standard deviation,
+    T-statistic, and p-value. For clarity here is a reminder of the simulation
+    setup.
+
+    \begin{enumerate}
+      \item Generate $X_i$, $\varepsilon_{1i}$, and $\varepsilon_{2i}$ from the
+        following distributions:
+
+        \begin{align*}
+          x_i &\stackrel{iid}{\sim} N(0, 1)\\
+          \varepsilon_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
+          \varepsilon_{2i} &\stackrel{iid}{\sim} N(\theta, 1)\\
+        \end{align*}
+
+        Then we have
+        \[y_{1i} = x_i + \varepsilon_{1i} \text{ and } y_{2i} = x_i +
+        \varepsilon_{2i}.\]
+
+      \item Then we have to select the variables to observe. We always observe
+        $X_i$. Then we choose to either observe $Y_1$ with probability $0.4$,
+        $Y_2$ with probability $0.4$ or neither with probability $0.2$.
+
+      \item If neither then $R_{1i} = 0$ and $R_{2i} = 0$. If we observe $Y_1$
+        then $R_1 = 1$ and if we observe $Y_2$ then $R_2 = 1$.
+
+      \item If we observe either $Y_1$ or $Y_2$ then with probability $p \propto
+        \logistic(Y_k)$ where $Y_k$ is the observed $Y$ variable we choose to
+        observe the other $Y$ variable.
+
+      \item If the other $Y$ variable is observed then the corresponding $R_k =
+        1$. Otherwise, $R_k = 0$.
+    \end{enumerate}
+
+  \item For this simulation setup, we estimate $\theta = E[Y_2]$. Like the
+    previous nonmonotone simulations, we compare this calibration estimator to
+    the oracle estimator which uses the average value of $Y_2$ if $R_2 = 0$ or
+    $R_2 = 1$, an IPW estimator with the correct weights, and the proposed
+    regression estimator. These are currently run with a sample size of $n =
+    1000$ with the number of Monte Carlo simulations of $B = 1000$.
+
+    \input{Tables/calinonmono1m-5.tex}
+    \input{Tables/calinonmono1m0.tex}
+    \input{Tables/calinonmono1m5.tex}
+
 \end{itemize}
@@
 \newpage
-\end{document}

### 2023-04-27 Simulations/calibration_comp.R

@@ -144,3 +144,193 @@ mc_theta %>%
   
 # Overall, pretty good. It looks like this is working as expected.
 
+# ***************************
+# * Nonmonotone Calibration *
+# ***************************
+
+nonmono_est_weights <- function(df) {
+
+  gfun <- "y2"
+
+  # Steps:
+  # 1. Compute w1--the weights between U and A1.
+  # 2. Compute w2--the weights between U and A2.
+  # 3. Compute wc--the weights between A1, A2, U, and the core points.
+
+  # 1. Compute w1--the weights between U and A1.
+  #   a. Estimate E[g_i \mid X_i].
+  #   b. Compute weights on A1.
+  Eg_mod <- lm(as.formula(paste0(gfun, " ~ x")), data = df)
+  df$Egx <- predict(Eg_mod)
+
+  Mgx <- mean(df$Egx)
+  n1 <- sum(df$r1)
+  w1 <- Variable(n1)
+  Egx_r1 <- filter(df, r1 == 1) %>% pull(Egx)
+  const <- list(w1 >= 0,
+                t(w1) %*% matrix(rep(1, n1), ncol = 1) == 1,
+                t(w1) %*% matrix(Egx_r1, ncol = 1) == Mgx)
+
+  p1 <- Problem(Minimize(sum_squares(w1)), const)
+  res1 <- solve(p1)
+
+  df$w1 <- 0
+  df$w1[df$r1 == 1] <- res1$getValue(w1)
+
+  # 2. Compute w2--the weights between U and A2.
+  #   a. Compute weights on A2.
+  n2 <- sum(df$r2)
+  w2 <- Variable(n2)
+  Egx_r2 <- filter(df, r2 == 1) %>% pull(Egx)
+  const <- list(w2 >= 0,
+                t(w2) %*% matrix(rep(1, n2), ncol = 1) == 1,
+                t(w2) %*% matrix(Egx_r2, ncol = 1) == Mgx)
+
+  p2 <- Problem(Minimize(sum_squares(w2)), const)
+  res2 <- solve(p2)
+
+  df$w2 <- 0
+  df$w2[df$r2 == 1] <- res2$getValue(w2)
+
+  # 3. Compute wc--the weights between A1, A2, U, and the core points.
+  #   a. Estimate E[g_i \mid X_i, Y_{1i}].
+  #   b. Estimate E[g_i \mid X_i, Y_{2i}].
+  #   c. Compute core weights on each part simultaneously.
+  df_1 <- filter(df, r1 == 1)
[…]

### 2023-04-27 Simulations/nonmonotone.R

@@ -117,11 +117,8 @@ mc_theta %>%
 # * Run Simulation: Nonmonotone Missingness *
 # *******************************************
 
-# FIXME: miss_resp is misleading. This should be a parameter of nonmono_mar
-# instead it is a parameter for est_nonmono. This should be changed to something
-# like est_weights.
 run_init_sims <- function(n_sim, B, true_mean, cor_y1y2,
-                          miss_out, miss_resp, seed, r_ind_y = FALSE) {
+                          miss_out, est_weights, seed, r_ind_y = FALSE) {
 
 
   clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
@@ -145,7 +142,7 @@ run_init_sims <- function(n_sim, B, true_mean, cor_y1y2,
 
     tibble(oracle = mean(df$y2),
            ipworacle = ipw_cc_oracle(df),
-           proposed = est_nonmono(df, oracle = !miss_resp))
+           proposed = est_nonmono(df, oracle = !est_weights))
 
   } %>% bind_rows()
 
@@ -179,21 +176,21 @@ run_init_sims <- function(n_sim, B, true_mean, cor_y1y2,
 
 run_init_sims(n_sim = 2000, B = 2000,
               true_mean = -5, cor_y1y2 = 0,
-              miss_out = FALSE, miss_resp = FALSE, seed = 1) %>%
+              miss_out = FALSE, est_weights = FALSE, seed = 1) %>%
   knitr::kable("latex", booktabs = TRUE, digits = 3,
   caption = "True Value is -5. Cor(Y1, Y2) = 0") %>%
   cat(., file = paste0("../Tables/nonmonosim1" , "m-5", ".tex"))
 
 run_init_sims(n_sim = 2000, B = 2000,
               true_mean = 0, cor_y1y2 = 0,
-              miss_out = FALSE, miss_resp = FALSE, seed = 1) %>%
+              miss_out = FALSE, est_weights = FALSE, seed = 1) %>%
   knitr::kable("latex", booktabs = TRUE, digits = 3,
   caption = "True Value is 0. Cor(Y1, Y2) = 0") %>%
   cat(., file = paste0("../Tables/nonmonosim1" , "m0", ".tex"))
 
 run_init_sims(n_sim = 2000, B = 2000,
[…]

### 2023-04-28 Notes/annote_bib.tex
Added annotated bibliography.

\documentclass[12pt]{article}

\usepackage{amsmath, amssymb, mathrsfs, fancyhdr}
\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
\usepackage{biblatex}
\usepackage{booktabs}
\usepackage{float}

\addbibresource{references.bib}

\hypersetup{colorlinks = true, urlcolor = black}

\headheight     15pt
\topmargin      -1.5cm   % read Lamport p.163
\oddsidemargin  -0.04cm  % read Lamport p.163
\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
\textwidth      16.59cm
\textheight     23.94cm
\parskip         7.2pt   % sets spacing between paragraphs
\parindent         0pt   % sets leading space for paragraphs
\pagestyle{empty}        % Uncomment if don't want page numbers
\pagestyle{fancyplain}

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}

\begin{document}

%\lhead{Caleb Leedy}
\chead{Nonmonotone Missingness}
%\chead{STAT 615 - Advanced Bayesian Methods}
%\rhead{Page \thepage\ of \pageref{LastPage}}
\rhead{April 28, 2023}

\section*{Annotated Bibliography}

\subsection*{Merkouris et al. (2023)}

\begin{description}
  \item[Title:] Combining National Surveys with Composite Calibration to Improve
    the Precision of Estimates from the United Kingdom's Living Costs and Food
    Supply

  \item[Authors:] Takis Merkouris, Paul A. Smith, Andy Fallows

  \item[Journal (Year):] JSSAM (2023)

  \item[Summary:] In some sense this is the application of ideas from Merkouris
    (2004) and Merkouris (2010) to combine information from the Living Costs and
    Food Survey (LCF) and the Labour Force Survey (LFS) in the United Kingdom.
    The overall approach is to use a linear combination of regression estimators
    from each survey:

    \[\hat Z^{CR} = B \hat Z_1^R + (I - B) Z_2^R\]

    where $B$ is a matrix, $Z_1^R$ is the regression estimator from the first
    survey and $Z_2^R$ is the regression estimator from the second survey. Like
    Merkouris (2004), they choose the optimal $B$ that minimizes the variance of
    the estimator while still having the estimation of $\hat Z$ be the same for
    each survey. To incorporate results from Merkouris (2010), they also control
    totals at the regional level (a small-area estimation problem). (More
    details regarding the derivation of the optimal $B$ is left to the summary
    of Merkouris (2004) and Merkouris (2010).)

    Methodologically, the biggest innovation is the jackknife variance
    estimation that they use. However, even this seems relatively
    straightforward. I think that the main contribution of this paper is to show
    a dramatic reduction in the variance of regression estimates due to data
    integration.

  \item[Limitations:]\;
    \begin{itemize}
      \item They only integrate two surveys.
      \item They do not consider informative sampling. % FIXME
    \end{itemize}
  \item[Extensions:]\;
    \begin{itemize}
      \item I think that it would be interesting if the supports of several
        control variables were not the same. This would be evidence of selection
        bias and we would need to account for it.
    \end{itemize}
\end{description}

\subsection*{Merkouris (2004)}

\begin{description}
  \item[Title:] Combining Independent Regression Estimators from Multiple
    Surveys

  \item[Author:] Takis Merkouris

  \item[Journal (Year):] JASA (2004)
  \item[Summary:] This paper develops the methodology to combine regression
    estimators from multiple surveys. Given two surveys the paper suggests
    combining the regression estimators as,

    \[\hat Z^{CR} = B\hat Z_1^R + (I - B) \hat Z_2^R\]

    where $B = \phi Z_2' L_2 Z_2 ((1 - \phi) Z_1'L_1 Z_1 + \phi Z_2'L_2
    Z_2)^{-1}$, $\phi = \frac{n_1 / d_1}{n_1 / d_1 + n_2 / d_2}$, $n_i$ is the
    sample size from survey $i$, $d_i$ is the design effects from survey $i$ of
    $z$, $Z_i$ is a matrix of common columns between the two surveys, 
    $L_i = I - X_i (X_i' X_i)^{-1} X_i'$, and $X_i$ are additional covariates
    that we have information to calibrate. In this way, the combined regression
    estimator can have minimal variance while having the estimates of $Z$ from
    each survey match. The weights also adjust for sample size and the strength
    of the correlation between $Z$ and $X$. This is the main method used in
    Merkouris (2023) except that Merkouris (2023) also adjusts for regional
    totals as developed in Merkouris (2010).

\end{description}

\subsection*{Merkouris (2010)}

\begin{description}
  \item[Title:] Combining information from multiple surveys by using regression
    for efficient small domain estimation

  \item[Author:] Takis Merkouris

  \item[Journal (Year):] JRSSB (2010)
  \item[Summary:] The bulk of this paper compares three estimators: the
    population control $\hat Y_d^R \equiv \hat Y_d + \hat \beta_d' (t_x - \hat
    X)$ where $\hat \beta = (X_s' \Lambda_s X_s)^{-1} X_s'\Lambda_s Y_{s_d}$,
    the domain control $\check Y_d^R \equiv \hat Y_d + \check \beta_d' (t_{x_d}
    - \hat X_d)$ where $\check \beta_d = (X_{s_d}' \Lambda_{s_d} X_{s_d})^{-1}
    X_{s_d}' \Lambda_{s_d} Y_{s_d}$, and the domain size control $\tilde Y_d^R
    \equiv \hat Y_d^R + \tilde \beta_d' (N_d - \hat N_d^R)$ where $\tilde
    \beta_d = (1_{s_d}' L_s 1_{s_d})^{-1} 1_{s_d}' L_s Y_{s_d}$. Notice that the
    last estimator is based off of the population control $\hat Y_d^R$ instead
    of the HT-estimator $\hat Y_d$. Since this model incorporates both the
    population variables and the domain indicators, the corresponding regression
    estimator can be thought of as an application of the augmented regression in 
    Merkouris (2004).

\end{description}

% \begin{description}
%   \item[Title:]
%   \item[Author:]
%   \item[Journal (Year):]
%   \item[Summary:]
%   \item[Limitations:]\;
%     \begin{itemize}
%       \item
%     \end{itemize}
%   \item[Extensions:]\;
%     \begin{itemize}
%       \item
%     \end{itemize}
% \end{description}

\end{document}

### 2023-05-02 Notes/efficiency_twophase.tex
Wrote up efficiency comparison with regression estimators. Our methods seem to outperform them.

\section*{Efficiency of Proposed Estimator}

One of our main goals is to show that we have an efficient estimator. We
want to ensure that our estimator is more efficient (lower MSE or really zero
bias and a smaller variance) than other competing estimators. We have already
demonstrated superior performance compared to the IPW estimator with known
weights. Now, we want to compare our estimator with two-phase and three-phase 
regression estimators.

\subsection*{Two and Three Phase Regression Estimators}

\begin{itemize}
  \item The simulation setup that we have been running generates monotone and
    nonmonotone missing patterns because we always observe $X$, yet we observe
    $Y_1$ and $Y_2$ with missingness. This creates the need to extend the
    traditional two-phase estimator to become a three-phase estimator. The
    two-phase regression estimator is really just a regression estimator with
    the ``finite population'' being Phase 1 of the sample. Thus, a valid
    two-phase estimator for $\theta = E[Y_2]$ is

    \[\hat \theta = \bar y_2 + (\bar x_1 - \bar x_2) \hat \beta\]

    where $\bar y_2 = n_2^{-1} \sum_{i \in U} I(i \in A_2) y_2$ and $\bar x_k =
    n_k^{-1} \sum_{i \in U} I(i \in A_k) x_i$ where $n_k$ is the number of
    elements in $A_k$ where $A_k$ is the Phase $k$ sample. In this case, $\hat
    \beta$ solves the following equation for $\beta_0$ and $\beta_1$,

    \[\sum_{i \in A_2} (y_{2i} - \beta_0 - \beta_1 x_i)^2 = 0.\]

  \item Notice that the previous construction of the two-phase estimator ignored
    the variable $Y_1$. We can incorporate this into the model using a 
    three-phase estimator. From \cite{fuller2009sampling}, the three-phase
    estimator is

    \begin{align*}
      \bar y_{2, 2p} &= \bar y_2 + (\bar x_0 - \bar x_{2}) \hat \beta_1 + 
      (\bar y_{1, reg} - \bar y_{1, 2p}) \hat \beta_2
    \end{align*}

    where 
    \begin{align*}
      \bar x_0 &= n^{-1} \sum_{i \in U} x_i \\
      \bar x_1 &= \left(\sum_{i \in A_1} \pi_{1i}^{-1}\right)^{-1} 
                  \sum_{i \in A_1} \pi_{1i}^{-1} x_i\\
      \bar x_2 &= \left(\sum_{i \in A_2} \pi_{2i}^{-1}\right)^{-1} 
                  \sum_{i \in A_2} \pi_{2i}^{-1} x_i\\
      \bar y_{1, reg} &= \bar y_{1, 1p} + (\bar x_0 - \bar x_1)\hat \beta_{1p}\\
      \bar y_{1, 1p} &= \left(\sum_{i \in A_1} \pi_{1i}^{-1}\right)^{-1} 
                        \sum_{i \in A_1} \pi_{1i}^{-1} y_{1i}\\
      \bar y_{1, 2p} &= \left(\sum_{i \in A_2} \pi_{2i}^{-1}\right)^{-1} 
                        \sum_{i \in A_2} \pi_{2i}^{-1} y_{1i}\\
      \hat \beta_{1p} &= \left(\sum_{i \in A_1} (x_i - \bar x_1)^2
      \pi_{1i}^{-1}\right)^{-1} \sum_{i \in A_1} (x_i - \bar x_1)\pi_{1i}^{-1}
      (y_{1i} - \bar y_{1, 1p})\\
      \hat \beta_1 &= \left(\sum_{i \in A_2} (x_i - \bar x_2, y_{1i} - \bar
      y_{1, 2p})' \pi_{2i}^{-1} (x_i - \bar x_2, y_{1i} - \bar y_{1,
      2p})\right)^{-1} \sum_{i \in A_2} (x_i - \bar x_2) \pi_{2i}^{-1} (y_{2i} -
      \bar y_2)\\
      \hat \beta_2 &= \left(\sum_{i \in A_2} (x_i - \bar x_2, y_{1i} - \bar
      y_{1, 2p})' \pi_{2i}^{-1} (x_i - \bar x_2, y_{1i} - \bar y_{1,
      2p})\right)^{-1} \sum_{i \in A_2} (y_{1i} - \bar y_{1, 2p}) \pi_{2i}^{-1} 
      (y_{2i} - \bar y_2)\\
    \end{align*}

  \item Notice that the three-phase estimator implicitly assumes a monotone 
    missingness model. It ignores the values of $y_2$ that have an unobserved
    $y_1$.

\end{itemize}

\subsection*{Monotone Results}

\begin{itemize}
  \item First, we test the two-phase and three-phase regression estimators with
    a monotone simulation. Like the previous monotone simulations,
    we first generate data from the following distributions:
    \begin{align*}
        X_i &\stackrel{iid}{\sim} N(0, 1) \\
        Y_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
        Y_{2i} &\stackrel{iid}{\sim} N(\theta, 1)
    \end{align*}

    Then, we create the probabilities $p_1 = \logistic(x_i)$ and 
    $p_{12} = \logistic(y_{1i})$. Including the calibration estimator from the
    previous section (note, these are the same as the previous monotone tables)
    this yields,

    \input{Tables/calimono_t-5.tex}
    \input{Tables/calimono_t0.tex}
    \input{Tables/calimono_t5.tex}

  \item The proposed semiparametric estimator and the calibration estimator
    seem to both outperform the regression estimators. With smaller bias and
    smaller variance our estimators do better.

  \item It may be puzzling to see that the three-phase estimator does worse than
    the two-phase; however, I think that there is a good reason for this. In the
    simulation, $y_2 = x + \varepsilon$. This is the regression from the
    two-phase estimator. However, the three-phase estimator measures $y_2 \sim x
    + y_1 + \varepsilon$, which just adds noise. This is why I think the
    standard deviation of the three-phase regression estimator is larger.

\end{itemize}

\subsection*{Nonmonotone Results}

\begin{itemize}
  \item Similar to the monotone results, we also use the same simulation as the
    nonmonotone calibration estimators. Repeating the simulation outline, we are
    trying to estimate $\theta = E[y_2]$ and we have

    \begin{enumerate}
      \item Generate $X_i$, $\varepsilon_{1i}$, and $\varepsilon_{2i}$ from the
        following distributions:

        \begin{align*}
          x_i &\stackrel{iid}{\sim} N(0, 1)\\
          \varepsilon_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
          \varepsilon_{2i} &\stackrel{iid}{\sim} N(\theta, 1)\\
        \end{align*}

        Then we have
        \[y_{1i} = x_i + \varepsilon_{1i} \text{ and } y_{2i} = x_i +
        \varepsilon_{2i}.\]

      \item Then we have to select the variables to observe. We always observe
        $X_i$. Then we choose to either observe $Y_1$ with probability $0.4$,
        $Y_2$ with probability $0.4$ or neither with probability $0.2$.

      \item If neither then $R_{1i} = 0$ and $R_{2i} = 0$. Otherwise, if we 
        observe $Y_1$ then $R_1 = 1$ and if we observe $Y_2$ then $R_2 = 1$.

      \item If we observe either $Y_1$ or $Y_2$ then with probability $p \propto
        \logistic(Y_k)$ where $Y_k$ is the observed $Y$ variable we choose to
        observe the other $Y$ variable.

      \item If the other $Y$ variable is observed then the corresponding $R_k =
        1$. Otherwise, $R_k = 0$.
    \end{enumerate}

    \input{Tables/calinonmono1m-5.tex}
    \input{Tables/calinonmono1m0.tex}
    \input{Tables/calinonmono1m5.tex}

  \item Overall, it seems that the proposed estimator and calibration estimator
    outperform the two regression estimators. The regression estimators display
    slight bias (perhaps because of the nonmonotonicity).

\end{itemize}

\newpage

### 2023-05-02 Notes/main.tex
Refactored Simulations/nonmonotone.R into Simulations/generate_data.R and Simulations/nmono_est_funs.R.
Wrote up efficiency comparison with regression estimators. Our methods seem to outperform them.

@@
         \item Select one option based on the three probabilities for each
-          element $i$. If 0 is selected: $R_1 = 0$ and $R_2 = 0$. If 1 is
-          selected $R_1 = 1$. If 2 is selected, $R_2 = 1$.
+          element $i$. If 0 is selected: $R_1 = 0$ and $R_2 = 0$; if 1 is
+          selected $R_1 = 1$; if 2 is selected, $R_2 = 1$.
         \item We take the next step in multiple cases. If 0 was selected, we are
@@
 
-    Overall, these results are mostly what I would have expected. All of the
+    Overall, these results are mostly what I would have expected. All the
     algorithms estimate the true value of $\theta$ correctly in each case, with
@@
     \neq 0$. We are still interested in $\bar Y_2$ and we still run 2000
-    simulation with 2000 observations. In all of the next simulations the true
+    simulation with 2000 observations. In all the next simulations the true
     value of $\theta = 0$. The results are the following:
@@
     Thus, the proposed method is unbiased with a misspecified outcome model.
-    We now show a simulation where the outcome model is correctly specified but
+    We now show a simulation where the outcome model is correctly specified, but
     the response model is not.
@@
       \item (Stage 2) With probability $p_i \propto \logistic(x_i)$, choose to
-        observed the other $Y$ variable. 
+        observe the other $Y$ variable. 
     \end{enumerate}
@@
 
-  \item However, one of the key steps is that \textit{all} of the response
+  \item However, one of the key steps is that \textit{all} the response
     models are free of $Y$. In a previous iteration of Simulation 4, we had
@@
   \item To compute the variance, we first solve for each covariance combination.
-    Basically, all these computations rely of the following ideas. First, we
+    Basically, all these computations rely on the following ideas. First, we
     assume that the response model is correctly specified. Consequently,
@@
 
-    The overall approach to each of this computations is the following:
+    The overall approach to each of these computations is the following:
     (1) take conditional expectations with respect to $X$ (the $\Cov(E[\cdot])$
@@
 
-\textbf{Questions for Dr. Kim}
-\begin{itemize}
-  \item Where do we go from here?
-  \item This is a functional expectation that I need to calibrate. Should I use
-    some sort of basis spline to try to estimate these expectations and then set
-    them equal to zero? I don't think that I have worked with this kind of
-    constraint before.
-\end{itemize}
+%\textbf{Questions for Dr. Kim}
+%\begin{itemize}
+%  \item Where do we go from here?
+%  \item This is a functional expectation that I need to calibrate. Should I use
+%    some sort of basis spline to try to estimate these expectations and then set
+%    them equal to zero? I don't think that I have worked with this kind of
+%    constraint before.
+%\end{itemize}
+
 %\input{prop_full_nonmonoest.tex}
@@
 
+\input{calibration_comparison.tex}
+
+\input{efficiency_twophase.tex}
+
 \printbibliography

### 2023-05-02 Simulations/calibration_comp.R

@@ -33,57 +33,9 @@ source("generate_data.R")
 # *************************************
 source("nmono_est_funs.R")
 
-# ************************
-# * Monotone Calibration *
-# ************************
-
-mono_est_weights <- function(df) {
-
-  # Steps:
-  # 1. Calibrate with w1 weights on Y_1
-  # 2. Calibrate with w2 weights on Y_2 using w1 weights on Y_1.
-
-  # 1. Calibrate with w1 weights on Y_1
-  n1 <- sum(df$r1)
-  w1 <- Variable(n1)
-  objective <- Minimize(sum_squares(w1))
-  Tx <- mean(df$x)
-  xr1_vec <-
-    filter(df, r1 == 1) %>%
-    pull(x)
-  constraints <- 
-    list(t(w1) %*% matrix(xr1_vec, ncol = 1) == Tx,
-         w1 >= 0,
-         t(w1) %*% matrix(rep(1, n1), ncol = 1) == 1)
-  p1 <- Problem(objective, constraints)
-  res1 <- solve(p1)
-
-  df$w1 <- 0
-  df$w1[df$r1 == 1] <- res1$getValue(w1)
-
-  # 2. Calibrate with w2 weights on Y_2 using w1 weights on Y_1.
-  n2 <- sum(df$r2) # Only works because of monotone.
-  w2 <- Variable(n2)
-  objective <- Minimize(sum_squares(w2))
-  Exw <- sum(df$w1 * df$x)
-  Ey1w <- sum(df$w1 * df$y1)
-  xr2_vec <- filter(df, r2 == 1) %>% pull(x)
-  y1r2_vec <- filter(df, r2 == 1) %>% pull(y1)
-  constraints <-
-    list(w2 >= 0,
-         t(w2) %*% matrix(rep(1, n2), ncol = 1) == 1,
-         t(w2) %*% matrix(xr2_vec, ncol = 1) == Exw,
-         t(w2) %*% matrix(y1r2_vec, ncol = 1) == Ey1w)
-  p2 <- Problem(objective, constraints)
-  res2 <- solve(p2)
-
-  df$w2 <- 0
-  df$w2[df$r2 == 1] <- res2$getValue(w2)
-
-  # Objective is E[g] = E[y_2]
-  sum(df$w2 * df$y2)
-
-}
+# *******************************
+# * Monotone: Current Functions *
+# *******************************
 
 # *****************************
 # * Monotone Simulation Study *
[…]

### 2023-05-02 Simulations/generate_data.R

@@ -54,6 +54,11 @@ mono_mar <- function(n, mean_y2 = 0) {
 
   # 3. Reveal hidden variables.
   tibble(x = x_vec, y1 = y1, y2 = y2, r1 = r1, r2 = r2) %>%
+    mutate(p1 = expit(x_vec)) %>%
+    mutate(p12 = expit(y1)) %>%
+    mutate(p21 = 0) %>%
+    mutate(p2 = 0) %>%
+    mutate(pi_11 = p1 * p12) %>%
     mutate(r2 = ifelse(r1 == 0, 0, r2))
 
 }

### 2023-05-02 Simulations/nmono_est_funs.R

@@ -14,11 +14,13 @@
 # A. Monotone estimators
 #   1. est_mono -- Classical monotone estimator
 #   2. ipw_mono_est -- IPW estimator with estimated weights
+#   3. mono_est_weights -- Calibration estimator
 # B. Nonmonotone estimators
 #   1. est_nonmono -- The proposed estimator in the nonmonotone setting
 #   2. ipw_cc_oracle -- IPW estimator with oracle weights
 #   3. ipw_cc_est -- IPW estimator with weights estimated
 #   4. twophase_reg -- Unfinished
+#   5. nonmono_est_weights -- Calibration estimator 
 
 # ------------------------------------------------------------------------------
 
@@ -99,6 +101,53 @@ ipw_mono_est <- function(df, est = FALSE) {
 
 }
 
+mono_est_weights <- function(df) {
+
+  # Steps:
+  # 1. Calibrate with w1 weights on Y_1
+  # 2. Calibrate with w2 weights on Y_2 using w1 weights on Y_1.
+
+  # 1. Calibrate with w1 weights on Y_1
+  n1 <- sum(df$r1)
+  w1 <- Variable(n1)
+  objective <- Minimize(sum_squares(w1))
+  Tx <- mean(df$x)
+  xr1_vec <-
+    filter(df, r1 == 1) %>%
+    pull(x)
+  constraints <- 
+    list(t(w1) %*% matrix(xr1_vec, ncol = 1) == Tx,
+         w1 >= 0,
+         t(w1) %*% matrix(rep(1, n1), ncol = 1) == 1)
+  p1 <- Problem(objective, constraints)
+  res1 <- solve(p1)
+
+  df$w1 <- 0
+  df$w1[df$r1 == 1] <- res1$getValue(w1)
+
+  # 2. Calibrate with w2 weights on Y_2 using w1 weights on Y_1.
+  n2 <- sum(df$r2) # Only works because of monotone.
+  w2 <- Variable(n2)
+  objective <- Minimize(sum_squares(w2))
+  Exw <- sum(df$w1 * df$x)
+  Ey1w <- sum(df$w1 * df$y1)
+  xr2_vec <- filter(df, r2 == 1) %>% pull(x)
[…]

### 2023-05-03 Notes/main.tex
Recapped lohr2017combining.

@@
 
-% \input{estimating_response_mod.tex}
-
-\section*{Minimizing the Variance}
-
-\begin{itemize}
-  \item The goal of this section is to find optimal values of $b_2(X, Y_1)$ and
-    $a_2(X, Y_2)$ such that the variance of $\hat \theta_{eff}$ is minimized.
-
-  \item Recall:
-
-    \begin{align*}
-      \hat \theta_{eff} - \hat \theta_n 
-      &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \left(1 - \frac{R_{1i}}{\pi_{1+}}
-      - \frac{R_{2i}}{\pi_{2+}} + \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \\
-      &+ n^{-1} \sum_{i = 1}^n b_2(X_i, Y_{1i}) \left(\frac{R_{1i}}{\pi_{1+}} -
-      \frac{R_{1i}R_{2i}}{\pi_{11}}\right)\\
-      &+ n^{-1} \sum_{i = 1}^n a_2(X_i, Y_{2i}) \left(\frac{R_{2i}}{\pi_{2+}} -
-      \frac{R_{1i}R_{2i}}{\pi_{11}}\right)\\
-      &+ n^{-1} \sum_{i = 1}^n g_i \left(\frac{R_{1i}R_{2i}}{\pi_{11}} -
-      1\right) \\
-      &\equiv A + B + C + D.
-    \end{align*}
-
-  \item Notice that we will suppress the fact that response models are functions
-    of $X$ (i.e. we write $\pi_{11}$ instead of $\pi_{11}(X)$).
-
-  \item To compute the variance, we first solve for each covariance combination.
-    Basically, all these computations rely on the following ideas. First, we
-    assume that the response model is correctly specified. Consequently,
-    $E[A] = E[B] = E[C] = E[D] = 0$ and things work out better. This helps when
-    we take the covariance conditional on $X$ because the inner expectations are
-    zero. The second key insight it to notice that $E[R_{j}^k] = E[R_j]$ for $j
-    \in \{1, 2\}$ and $k \in \mathbb{N}$. This is because $R$ is a binary
-    variable. Third, since we assume that the response models are correctly
-    specified, we have $E[R_1 \mid X] = \pi_{1+}$, $E[R_2 \mid X] = \pi_{2+}$,
-    and $E[R_1, R_2 \mid X] = \pi_{11}$.
-
-    The overall approach to each of these computations is the following:
-    (1) take conditional expectations with respect to $X$ (the $\Cov(E[\cdot])$
-    term is zero), (2) expand the covariance to $E[XY] - E[X]E[Y]$ (the second
-    term is also zero), (3) by the MAR assumption $g, a_2, b_2$ are independent
-    of $R_1$ and $R_2$ and we can take the latter out of the expectation, (4)
-    evaluate and simplify expressions involving $E[R]$.
-
-    \begin{align*}
-      \Cov(A, B) &= n^{-2} \sum_{i = 1}^n E\left[\Cov\left(E[g_i \mid X] \left(1
-      - \frac{R_{1i}}{\pi_{1+}} - \frac{R_{2i}}{\pi_{2+}} +
-      \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \mid X_i,\right. \right.\\
-      &\qquad \qquad \qquad \qquad b_2(X_i, Y_{1i}) \left. \left.
-      \left(\frac{R_{1i}}{\pi_{1+}} - \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \mid
-      X_i\right)\right]\\
-      &= n^{-1} E\left[E\left[E[g \mid X] \left(1 - \frac{R_{1i}}{\pi_{1+}} -
-      \frac{R_{2i}}{\pi_{2+}} + \frac{R_{1i}R_{2i}}{\pi_{11}}\right)
-      b_2(X_i, Y_{1i}) \left(\frac{R_{1i}}{\pi_{1+}} -
-      \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \mid X \right]\right] \\
-      &= n^{-1} E\left[E[g \mid X] E[b_2(X, Y_1) \mid X]
-      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
-      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right].
-    \end{align*}
-
-    By symmetry,
-    \begin{align*}
-      \Cov(A, C) 
-      &= n^{-1} E\left[E[g \mid X] E[a_2(X, Y_2) \mid X]
-      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
-      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right].
-    \end{align*}
-
-    \begin{align*}
-      \Cov(A, D) &= n^{-1} E\left[E\left[E[g \mid X] \left(1 -
-      \frac{R_{1i}}{\pi_{1+}} - \frac{R_{2i}}{\pi_{2+}} +
-      \frac{R_{1i}R_{2i}}{\pi_{11}}\right) g \left(\frac{R_1 R_2}{\pi_{11}} -
-      1\right) \mid X\right] \right]\\
-      &= n^{-1} E\left[E[g \mid X]^2 \left(\frac{-1}{\pi_{1+}} -
-      \frac{1}{\pi_{2+}} + 2\right)\right].
-    \end{align*}
-
-    \begin{align*}
-      \Cov(B, C) &= n^{-1} E\left[E[b_2(X, Y_1) \mid X] E[a_2(X, Y_2) \mid X]
-      E\left[\left(\frac{R_1}{\pi_{1+}} - \frac{R_1 R_2}{\pi_{11}}\right)
-      \left(\frac{R_2}{\pi_{2+}} - \frac{R_1 R_2}{\pi_{11}}\right) \mid
-      X\right]\right]\\
-      &= n^{-1} E\left[E[b_2(X, Y_1) \mid X] E[a_2(X, Y_2) \mid X]
-      \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} - \frac{1}{\pi_{1+}} -
-      \frac{1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right)\right].
-    \end{align*}
-
-    \begin{align*}
-      \Cov(B, D) &= n^{-1} E\left[ E[b_2(X, Y_1) \mid X] E[g \mid X]
-      E\left[\left(\frac{R_1}{\pi_{1+}} - \frac{R_1 R_2}{\pi_{11}}\right)
-      \left(\frac{R_1 R_2}{\pi_{11}} - 1\right) \mid X\right] \right]\\
-      &= n^{-1} E\left[ E[b_2(X, Y_1) \mid X] E[g \mid X]
-      \left(\frac{1}{\pi_{1+}} - \frac{1}{\pi_{11}}\right)\right].
-    \end{align*}
-
-    By symmetry,
-    \begin{align*}
-      \Cov(C, D) 
-      &= n^{-1} E\left[ E[a_2(X, Y_2) \mid X] E[g \mid X]
-      \left(\frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}}\right)\right].
-    \end{align*}
-
-    We also compute the variance terms,
-
-    \begin{align*}
-      \Cov(A, A) &= n^{-1} E\left[ E\left[ E[g \mid X]^2 \left(1
-      - \frac{R_{1i}}{\pi_{1+}} - \frac{R_{2i}}{\pi_{2+}} +
-      \frac{R_{1i}R_{2i}}{\pi_{11}}\right)^2 \mid X\right] \right]\\
-      &= n^{-1} E \left[ E[g \mid X]^2 \left(-1 + \frac{2 \pi_{11}}{\pi_{1+}
-      \pi_{2+}} - \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
-      \frac{1}{\pi_{11}}\right)\right].
-    \end{align*}
-
-    \begin{align*}
-      \Cov(B, B) &= n^{-1} E\left[ E\left[b(X, Y_1)^2 \left(\frac{R_1}{\pi_{1+}}
-      - \frac{R_1 R_2}{\pi_{11}}\right)^2 \mid X\right] \right]\\
-      &= n^{-1} E\left[ E[b_2(X, Y_1)^2 \mid X] \left(\frac{-1}{\pi_{1+}} +
-      \frac{1}{\pi_{11}}\right)\right].
-    \end{align*}
-
-    \begin{align*}
-      \Cov(C, C) &= n^{-1} E\left[ E[a_2(X, Y_2)^2 \mid X]
-      \left(\frac{-1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right) \right].
-    \end{align*}
-
-    \begin{align*}
-      \Cov(D, D) &= n^{-1} E\left[ E\left[g_i^2 \left(\frac{R_1 R_2}{\pi_{11}} -
-      1\right)^2 \mid X\right] \right]\\
-      &= n^{-1} E\left[ E[g^2 \mid X] \left(\frac{1}{\pi_{11}} -
-      1\right)\right].
-    \end{align*}
-
-    This means that
-    \begin{align*}
-      \Var(\hat \theta_{eff} - \hat \theta_n) 
-      &= \Cov(A, A) + 2 \Cov(A, B) + 2 \Cov(A, C) + 2 \Cov(A, D) + \Cov(B, B) \\
-      &+ 2 \Cov(B, C) + 2 \Cov(B, D) + \Cov(C, C) + 2 \Cov(C, D) + \Cov(D, D) \\
-      &= n^{-1} E \left[ E[g \mid X]^2 \left(-1 + \frac{2 \pi_{11}}{\pi_{1+}
-      \pi_{2+}} - \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
-      \frac{1}{\pi_{11}}\right)\right] \\
-      &+ 2 n^{-1} E\left[E[g \mid X] E[b_2(X, Y_1) \mid X]
-      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
-      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right]\\
-      &+ 2 n^{-1} E\left[E[g \mid X] E[a_2(X, Y_2) \mid X]
-      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
-      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right]\\
-      &+ 2 n^{-1} E\left[E[g \mid X]^2 \left(\frac{-1}{\pi_{1+}} -
-      \frac{1}{\pi_{2+}} + 2\right)\right]\\
-      &+ n^{-1} E\left[ E[b_2(X, Y_1)^2 \mid X] \left(\frac{-1}{\pi_{1+}} +
-      \frac{1}{\pi_{11}}\right)\right]\\
-      &+ 2 n^{-1} E\left[E[b_2(X, Y_1) \mid X] E[a_2(X, Y_2) \mid X]
-      \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} - \frac{1}{\pi_{1+}} -
-      \frac{1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right)\right]\\
-      &+ 2 n^{-1} E\left[ E[b_2(X, Y_1) \mid X] E[g \mid X]
-      \left(\frac{1}{\pi_{1+}} - \frac{1}{\pi_{11}}\right)\right]\\
-      &+  n^{-1} E\left[ E[a_2(X, Y_2)^2 \mid X]
-      \left(\frac{-1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right) \right]\\
-      &+ 2 n^{-1} E\left[ E[a_2(X, Y_2) \mid X] E[g \mid X]
-      \left(\frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}}\right)\right]\\
-      &+ n^{-1} E\left[ E[g^2 \mid X] \left(\frac{1}{\pi_{11}} -
-      1\right)\right].
-    \end{align*}
-
-    Differentiating yields:
-
-    \begin{align*}
-      \frac{\partial}{\partial a_2} \Var(\hat \theta_{eff} - \hat \theta_n)
-      &= E\left[E[g \mid X] \left(\frac{1}{\pi_{1+}} + \frac{2}{\pi_{2+}} -
-      \frac{2}{\pi_{11}} - \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right] \\
-      &+ E\left[E[b_2(X, Y_1) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
-      \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
-      \frac{1}{\pi_{11}}\right)\right]\\
-      &+ E\left[ E[a_2(X, Y_2) \mid X] \left(\frac{-1}{\pi_{2+}} +
-      \frac{1}{\pi_{11}}\right)\right]\\
-      &\equiv 0, \text{ and }\\
-      \frac{\partial}{\partial b_2} \Var(\hat \theta_{eff} - \hat \theta_n)
-      &= E\left[E[g \mid X] \left(\frac{2}{\pi_{1+}} + \frac{1}{\pi_{2+}} -
-      \frac{2}{\pi_{11}} - \frac{\pi_{11}}{\pi_{1+}\pi_{2+}}\right)\right]\\
-      &+ E\left[E[a_2(X, Y_2) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
-      \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
-      \frac{1}{\pi_{11}}\right)\right]\\
-      &+ E\left[ E[b_2(X, Y_2) \mid X] \left(\frac{-1}{\pi_{1+}} +
-      \frac{1}{\pi_{11}}\right)\right]\\
-      &\equiv 0. 
-    \end{align*}
-
-    Substitution shows that these constraints are equivalent to:
-
-    \begin{align*}
-      E&\left[E[g \mid X] \left(\frac{-1}{\pi_{1+}} +
-    \frac{1}{\pi_{2+}}\right)\right] + E\left[ E[b_2(X, Y_1) \mid X]
-    \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -\frac{1}{\pi_{2+}}\right)\right]\\
-      &- E\left[E[a_2(X, Y_2) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} - 
-    \frac{1}{\pi_{1+}}\right)\right] \equiv 0
-    \end{align*}
-
-    where is the same as,
-
-    \begin{align*}
-      E&\left[ E[b_2(X, Y_1) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
-    \frac{1}{\pi_{2+}}\right) \right] + E\left[E[g \mid X]
-      \left(\frac{1}{\pi_{2+}}\right)\right] \\
-      &=
-    E\left[ E[a_2(X, Y_2) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
-    \frac{1}{\pi_{1+}}\right) \right] + E\left[E[g \mid X]
-    \left(\frac{1}{\pi_{1+}}\right)\right].
-    \end{align*}
-
-  \item These constraints can be satisfied (this is sufficient but maybe not
-    necessary) if
-
-    \begin{align*}
-      E&\left[(E[b_2(X, Y_1) \mid X] - E[a_2(X, Y_2) \mid X])
-      \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right) \right] = 0\\
-      E&\left[\left(\frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}}\right)(E[a_2(X, Y_2)
-      \mid X] + E[b_2(X, Y_1) \mid X] - 2E[g \mid X])\right] = 0.
-    \end{align*}
-
-\end{itemize}
-
-%\textbf{Questions for Dr. Kim}
-%\begin{itemize}
-%  \item Where do we go from here?
-%  \item This is a functional expectation that I need to calibrate. Should I use
-%    some sort of basis spline to try to estimate these expectations and then set
-%    them equal to zero? I don't think that I have worked with this kind of
-%    constraint before.
-%\end{itemize}
+%\input{estimating_response_mod.tex}
 
@@
 
-\newpage
+\input{minimizing_variance.tex}
 
@@
 
+\input{multiphase_estimators.tex}
+
 \printbibliography

### 2023-08-02 Notes/init_sims.tex
Commit at end of summer. About to start again.

--- 2023-05-03 Notes/main.tex
@@
-\documentclass[12pt]{article}
-
-\usepackage{amsmath, amssymb, mathrsfs, fancyhdr}
-\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
-\usepackage{biblatex}
-\usepackage{booktabs}
-\usepackage{float}
-
-\addbibresource{references.bib}
-
-\hypersetup{colorlinks = true, urlcolor = black}
-
-\headheight     15pt
-\topmargin      -1.5cm   % read Lamport p.163
-\oddsidemargin  -0.04cm  % read Lamport p.163
-\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
-\textwidth      16.59cm
-\textheight     23.94cm
-\parskip         7.2pt   % sets spacing between paragraphs
-\parindent         0pt   % sets leading space for paragraphs
-\pagestyle{empty}        % Uncomment if don't want page numbers
-\pagestyle{fancyplain}
-
-\newcommand{\MAP}{{\text{MAP}}}
-\newcommand{\argmax}{{\text{argmax}}}
-\newcommand{\argmin}{{\text{argmin}}}
-\newcommand{\Cov}{{\text{Cov}}}
-\newcommand{\Var}{{\text{Var}}}
-\newcommand{\logistic}{{\text{logistic}}}
-
-\begin{document}
-
-%\lhead{Caleb Leedy}
-\chead{Nonmonotone Missingness}
-%\chead{STAT 615 - Advanced Bayesian Methods}
-%\rhead{Page \thepage\ of \pageref{LastPage}}
-\rhead{April 4, 2023}
-
-\section*{Overview:}
-
-The goal of this project is to outperform existing techniques in the literature
-related to nonmonotone missing data.
-
 \section*{Initial Simulations:}
@@
 \newpage
-
-\section*{Missingness Mechanism}
-
-\begin{itemize}
-  \item First I am going to reproduce the proof of double robustness that we
-    talked about during our last meeting. I think it is insightful for future
-    comments:
-
-    \begin{align*}
-      E[\hat \theta_{eff} - \theta_n] 
-      &= E\left[n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] - g_i\right]\\
-      &\quad+ E\left[n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)}
-      (b_2(X_i, Y_{1i}) - E[g_i \mid X_i])\right]\\
-      &\quad+ E\left[n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)}
-      (a_2(X_i, Y_{2i}) - E[g_i \mid X_i])\right]\\
-      &\quad+ E\left[n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_i)} 
-      (g_i - a_2(X_i, Y_{2i}) - b_2(X_i, Y_{1i}) + E[g_i \mid X_i])\right]\\
-      &= n^{-1} \sum_{i = 1}^n (E[E[g_i \mid X_i]] - E[g_i])\\
-      &\quad+ n^{-1} \sum_{i = 1}^n E\left[E\left[\frac{R_{1i}}{\pi_{1+}(X_i)} 
-      (b_2(X_i, Y_{1i}) - E[g_i \mid X_i]) \mid X_i\right]\right] \\ 
-      &\quad+ n^{-1} \sum_{i = 1}^n E\left[E\left[\frac{R_{2i}}{\pi_{2+}(X_i)} 
-      (a_2(X_i, Y_{2i}) - E[g_i \mid X_i]) \mid X_i\right]\right] \\
-      &\quad+ n^{-1} \sum_{i = 1}^n E\left[E\left[\frac{R_{1i}
-      R_{2i}}{\pi_{11}(X_i)} (g_i - a_2(X_i, Y_{2i}) - b_2(X_i, Y_{1i}) + E[g_i
-      \mid X_i]) \mid X_i \right]\right]\\
-      \intertext{Since $R_{1i} \perp Y_{1i} \mid X_i$, $R_{2i} \perp Y_{2i} \mid
-      X_i$, $(R_{1i}, R_{2i}) \perp (Y_{1i}, Y_{2i}) \mid X_i$ and $\pi_{1+},
-      \pi_{2+}$ and $\pi_{11}$ are all free of $Y_{1i}$ and $Y_{2i}$.}
-      &= n^{-1} \sum_{i = 1}^n E\left[\frac{R_{1i}}{\pi_{1+}(X_i)} E\left[
-        (E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i]) \mid X_i\right]\right] \\ 
-      &\quad+ n^{-1} \sum_{i = 1}^n E\left[\frac{R_{2i}}{\pi_{2+}(X_i)} 
-      E\left[E[g_i \mid X_i, Y_{2i}] - E[g_i \mid X_i] \mid X_i\right]\right] \\
-      &\quad+ n^{-1} \sum_{i = 1}^n E\left[\frac{R_{1i}
-      R_{2i}}{\pi_{11}(X_i)} E\left[(g_i - E[g_i \mid X_i, Y_{2i}] - 
-      E[g_i \mid X_i, Y_{1i}]+ E[g_i \mid X_i]) \mid X_i \right]\right]\\
-      \intertext{Since $E[E[g_i \mid X_i, Y_{ki}] \mid X_i] = E[g_i \mid X_i] =
-      0$,}
-      &= 0.
-    \end{align*}
-
-    Thus, if the outcome models are correctly specified $\hat \theta_{eff}$ is
-    unbiased. If the response models are correctly specified it is easy to see
-    that $\hat \theta_{eff}$ is also unbiased. This means that $\hat
-    \theta_{eff}$ is doubly robust.
-
-  \item However, one of the key steps is that \textit{all} the response
-    models are free of $Y$. In a previous iteration of Simulation 4, we had
-    adopted the framework of \cite{robins1997non} where we first to observe the
-    first variable and see if we observe the second variable. In this case, the
-    second step can depend on the result of the first step and this is what we
-    did. However, this makes it the case that $\pi_11$ is a function of $X_i$
-    and $Y_1$ and $Y_2$. In this case $\hat \theta_{eff}$ is not unbiased if the
-    response model is misspecified (or even just estimated).
-
-  \item If we modify Simulation 4, such that the second step of observed the
-    second variable is proportional to $\logistic(y_k)$ then we get the same
-    result as before:
-
-    \input{Tables/nonmonosim5c0.tex}
-    \input{Tables/nonmonosim5c0.1.tex}
-    \input{Tables/nonmonosim5c0.5.tex}
-
-\end{itemize}
-
-\newpage
-
-%\input{estimating_response_mod.tex}
-
-%\input{prop_full_nonmonoest.tex}
-
-\input{minimizing_variance.tex}
-
-\input{calibration_comparison.tex}
-
-\input{efficiency_twophase.tex}
-
-\input{multiphase_estimators.tex}
-
-\printbibliography
-
-\end{document}

### 2023-08-02 Notes/main.tex
Commit at end of summer. About to start again.

@@
 
-\section*{Initial Simulations:}
-% Tables generated from Simulations/nonmonotone.R
+%\input{init_sims.tex}
 
-\begin{itemize}
-  \item \textbf{Implemented simulation of monotone MAR data}: 
-    This is correspondingly easier than the subsequent nonmonotone MAR
-    simulation. For this simulation we use the following approach:
-  \begin{enumerate}
-      \item Generate $X$, $Y_1$, and $Y_2$ for elements $i = 1, \dots, n$.
-      \item Using the covariate $X$, determine the probability $p_1$ of $Y_1$
-      being observed for each element $i$.
-      \item Based on $p_1$, determine if $R_1 = 1$.
-      \item If $R_1 = 0$, then $R_2 = 0$. Otherwise, using variables $X$ and
-        $Y_1$, determine the probability $p_{12}$.
-      \item Based on $p_{12}$ determine if $R_2 = 1$.
-  \end{enumerate}
-    At the end of the algorithm, we have determined the values of binary
-    variables $R_1$ and $R_2$ for each $i$ and if either of them are equal to 1,
-    the corresponding level of $Y_{k}$. As is common in this literature, the
-    values of $R_1$ and $R_2$ determine if the corresponding variable $Y_1$ or
-    $Y_2$ is missing or observed with $R = 1$ indicating $Y$ being observed.
+%\input{missing_mechanism.tex}
 
-  \item \textbf{Implemented simulation of nonmonotone MAR data}: 
-    Following the approach of \cite{robins1997non}, I construct a nonmonotone
-    MAR simulation with two response variables $Y_1$ and $Y_2$ and one covariate
-    $X$. The algorithm to generate the data is the following:
-    \begin{enumerate}
-        \item Generate $X$, $Y_1$, and $Y_2$ for elements $i = 1, \dots, n$.
-        \item Using the covariate $X_i$, generate probabilities for each element
-          $i$ $p_0$, $p_1$, and $p_2$ such that $p_0 + p_1 + p_2 = 1$. 
-        \item Select one option based on the three probabilities for each
-          element $i$. If 0 is selected: $R_1 = 0$ and $R_2 = 0$; if 1 is
-          selected $R_1 = 1$; if 2 is selected, $R_2 = 1$.
-        \item We take the next step in multiple cases. If 0 was selected, we are
-          done. If 1 was selected, we generate probabilities $p_{12}$ based on
-          $X$ and $Y_1$. Then based on this probability, we determine if $R_2 =
-          1$. In the same manner, if 2 was selected in the previous step, we
-          generate probabilities $p_{21}$ based on $X$ and $Y_2$. Then based on
-          this probability, we determine if $R_2 = 1$.
-    \end{enumerate}
-    Like the monotone MAR simulation this algorithm produces similar final
-    results with the determination of binary variables $R_1$ and $R_2$ and
-    variables $X$, $Y_1$, and $Y_2$. Unlike the monotone MAR case, the
-    nonmonotone MAR includes observations with $Y_2$ observed and $Y_1$ missing.
+%%%\input{estimating_response_mod.tex}
 
-  \item \textbf{Simulation 1 with Monotone MAR}:
-    Following the algorithm described in the monotone MAR simulation bullet, 
-    we first generate data from the following distributions:
-    \begin{align*}
-        X_i &\stackrel{iid}{\sim} N(0, 1) \\
-        Y_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
-        Y_{2i} &\stackrel{iid}{\sim} N(\theta, 1)
-    \end{align*}
+%%%\input{prop_full_nonmonoest.tex}
 
-    Then, we create the probabilities $p_1 = \logistic(x_i)$ and 
-    $p_{12} = \logistic(y_{1i})$.
-    Since, both $x_i$ and $y_1$ are standard normal distributions, each of these
-    probabilities is approximately $0.5$ in expectation.
+%\input{minimizing_variance.tex}
 
-    The goal of this simulation is to estimate $\theta$. Alternatively, we can
-    express this as solving the estimating equation:
+%\input{calibration_comparison.tex}
 
-    \[g(\theta) \equiv Y_2 - \theta = 0.\]
-
-    We estimate $\theta$ using the following procedures:
-
-    \begin{itemize}
-        \item Oracle: This computes $\bar Y$ using \textit{both} the observed
-          and missing data.
-        \item IPW-Oracle: This is an IPW estimator using only the observed
-          values of $Y_2$. The weights (inverse probabilities) use the actual
-          probabilities.
-        \item IPW-Est: This is an IPW estimator using the probabilities that
-          have been estimated by a logistic model.
-        \item Semi: This is the monotone semiparametric efficient estimator from
-          Slide 11 (Equation 2) of Dr. Kim's Nonmonotone Missingness
-          presentation.
-    \end{itemize}
-
-    We run this simulation with different values of $\theta$, sample size of
-    2000, and 2000 Monte Carlo replications. Each algorithm for each replication
-    generates $\hat \theta$. In the subsequent tables, we compute the bias,
-    standard deviation (sd), t-statistic (where we test for a significant
-    difference between the Monte Carlo mean $\hat \theta$ and the true $\theta$)
-    and the p-value of the t-statistic.
-
-    \newpage
-
-    \input{Tables/monomar_t-5.tex}
-    \input{Tables/monomar_t0.tex}
-    \input{Tables/monomar_t5.tex}
-
-    Overall, these results are mostly what I would have expected. All the
-    algorithms estimate the true value of $\theta$ correctly in each case, with
-    the oracle estimate having the smallest variance followed by the
-    semiparametric algorithm. If there is anything surprising it is that the IPW
-    estimator has better performance with the estimated weights compared to the
-    true weights. However, I think that this is a known phenomenon.
-
-    \newpage
-
-    \item \textbf{Simulation 1 with Nonmonotone MAR}:
-
-    We generate variables $(X, Y_1, Y_2)$ using the following setup:
-
-    \[\begin{bmatrix}
-    X_i \\ \varepsilon_{1i} \\ \varepsilon_{2i}
-    \end{bmatrix} \stackrel{iid}{\sim}
-    N\left(
-    \begin{bmatrix}
-        0 \\ 0 \\ \theta
-    \end{bmatrix},
-    \begin{bmatrix}
-        1 & 0 & 0 \\
-        0 & 1 & \sigma_{yy}\\
-        0 & \sigma_{yy} & 1
-    \end{bmatrix}
-    \right).\]
-
-    Then,
-
-    \[y_{1i} = x_i + \varepsilon_{1i} \text{ and } 
-    y_{2i} = x_i + \varepsilon_{2i}.\]
-
-    Since we have nonmonotone data, our ``Stage 1'' probabilities are
-    different. We compute the true Stage 1 probabilities being proportional to
-    the following values:
-
-    \begin{align*}
-        p_0 &= 0.2 \\
-        p_1 &= 0.4 \\
-        p_2 &= 0.4 \\
-    \end{align*}
-
-   However, we keep the same structure for the Stage 2 probabilities with:
-   $p_{12} = \logistic(y_1)$ and $p_{21} = \logistic(y_2)$. The goal remains to
-   estimate $\theta$. We continue to use the Oracle algorithm and the IPW-Oracle
-   algorithm. Since we have nonmonotone MAR data, we use the ``Proposed''
-   algorithm that is described on Slide 25 (Equation 12) of Dr. Kim's
-   presentation. The outcome models were estimated using
-   logistic regression and OLS and correctly specified. The response model used
-   the oracle estimates of the probabilities. This yields the
-   following results:
-
-    \input{Tables/nonmonosim1m-5.tex}
-    \input{Tables/nonmonosim1m0.tex}
-    \input{Tables/nonmonosim1m5.tex}
-
-    \newpage
-
-    \item \textbf{Simulation 2 with Nonmonotone MAR}:
-    We also want to simulate data that is correlated. For this simulation, we
-    focus on $\Cov(Y_1, Y_2)$. The data generating process now has $\sigma_{yy}
-    \neq 0$. We are still interested in $\bar Y_2$ and we still run 2000
-    simulation with 2000 observations. In all the next simulations the true
-    value of $\theta = 0$. The results are the following:
-
-    \input{Tables/nonmonosim2c0.1.tex}
-    \input{Tables/nonmonosim2c0.5.tex}
-    \input{Tables/nonmonosim2c0.9.tex}
-
-    \newpage
-
-  \item \textbf{Simulation 3 with Nonmonotone MAR}:
-    This simulation aims to see if the proposed algorithm is doubly robust.
-    First, we check with a misspecified outcome model. In this case the data
-    generating procedure is the following:
-
-    \[\begin{bmatrix}
-    X_i \\ \varepsilon_{1i} \\ \varepsilon_{2i}
-    \end{bmatrix} \stackrel{iid}{\sim}
-    N\left(
-    \begin{bmatrix}
-        0 \\ 0 \\ \theta
-    \end{bmatrix},
-    \begin{bmatrix}
-        1 & 0 & 0 \\
-        0 & 1 & \sigma_{yy}\\
-        0 & \sigma_{yy} & 1
-    \end{bmatrix}
-    \right).\]
-
-    Then, the true outcome model is,
-
-    \[y_{1i} = x_i + x_i^2 \varepsilon_{1i} \text{ and } 
-    y_{2i} = -x_i + x_i^3 + \varepsilon_{2i}.\]
-
-    This procedure causes $X$ to influence both $Y_1$ and $Y_2$ and we still
-    have correlation in the error terms of $Y_1$ and $Y_2$. However, since
-    neither $Y_1$ nor $Y_2$ are linear in $X$, the model will be misspecified.
-    The response mechanisms are first generated MCAR with a probability of
-    either $Y_1$ or $Y_2$ being the first variable observed to be $0.4$. (There
-    is a $0.2$ probability neither is observed.) Then the probability of the
-    other variable being observed is proportional to $\logistic(y_k)$ where
-    $y_k$ is the $y$ that has been observed. To ensure that the proposed method
-    has the correct propensity score we use the oracle probabilities instead of
-    estimating them. This yields the following:
-
-    \input{Tables/nonmonosim3c0.tex}
-    \input{Tables/nonmonosim3c0.1.tex}
-    \input{Tables/nonmonosim3c0.5.tex}
-
-    \newpage
-
-    Thus, the proposed method is unbiased with a misspecified outcome model.
-    We now show a simulation where the outcome model is correctly specified, but
-    the response model is not.
-
-  \item \textbf{Simulation 4 with Nonmonotone MAR}:
-    Continuing to test if the proposed algorithm is doubly robust, this
-    simulation checks a misspecified response model. Instead of using oracle
-    weights as in Simulation 3, we estimate the weights for the proposed method.
-    However, unlike the true probabilities of being proportional to
-    $\logistic(y_k)$, this simulation has the true probabilities being
-    proportional to $\logistic(x_i)$. Thus, the true response model is the
-    following:
-
-    \begin{enumerate}
-      \item (Stage 1) Choose a variable observe. We choose $Y_1$ with
-        probability $0.4$, $Y_2$ with probability $0.4$ and neither with
-        probability $0.2$. If neither, $R_1 = 0$ and $R_2 = 0$. Otherwise,
-        continue to Step 2.
-      \item (Stage 2) With probability $p_i \propto \logistic(x_i)$, choose to
-        observe the other $Y$ variable. 
-    \end{enumerate}
-
-    This sequence generates the missingness indicators $R_1$ and $R_2$. Since,
-    the Stage 1 probabilities are fixed and known and the Stage 2 probabilities
-    only depend on $x_i$, the missingness is MAR and only a function of $x_i$.
-    The algorithms to which we compare still use the oracle weights.
-
-    \input{Tables/nonmonosim4c0.tex}
-    \input{Tables/nonmonosim4c0.1.tex}
-    \input{Tables/nonmonosim4c0.5.tex}
-
-    The previous version of this simulation used the Stage 2 probability $p_i
-    \propto \logistic(y_i)$ where $y_i$ was the observed $Y$ value in Stage 1.
-    However, under this setup, our method is biased. This is because
-
-    \[ E\left[E\left[\frac{R_1 R_2}{\pi_{11}(X, Y_1)} Y_1 \mid X\right]\right] 
-    \neq
-    E\left[\frac{E[Y_1 \mid X]}{\pi_{11}(X, Y_1)} E[R_1 R_2 \mid X]\right] \]
-
-    but
-
-    \[ E\left[E\left[\frac{R_1 R_2}{\pi_{11}(X)} Y_1 \mid X\right]\right] 
-    =
-    E\left[\frac{E[Y_1 \mid X]}{\pi_{11}(X)} E[R_1 R_2 \mid X]\right].\]
-
-    Thus, there is strong evidence that the proposed method is 
-    doubly robust because it is robust to both misspecification in the outcome
-    and response model.
-\end{itemize}
-
-\newpage
-
-\section*{Missingness Mechanism}
-
-\begin{itemize}
-  \item First I am going to reproduce the proof of double robustness that we
-    talked about during our last meeting. I think it is insightful for future
-    comments:
-
-    \begin{align*}
-      E[\hat \theta_{eff} - \theta_n] 
-      &= E\left[n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] - g_i\right]\\
-      &\quad+ E\left[n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)}
-      (b_2(X_i, Y_{1i}) - E[g_i \mid X_i])\right]\\
-      &\quad+ E\left[n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)}
-      (a_2(X_i, Y_{2i}) - E[g_i \mid X_i])\right]\\
-      &\quad+ E\left[n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_i)} 
-      (g_i - a_2(X_i, Y_{2i}) - b_2(X_i, Y_{1i}) + E[g_i \mid X_i])\right]\\
-      &= n^{-1} \sum_{i = 1}^n (E[E[g_i \mid X_i]] - E[g_i])\\
-      &\quad+ n^{-1} \sum_{i = 1}^n E\left[E\left[\frac{R_{1i}}{\pi_{1+}(X_i)} 
-      (b_2(X_i, Y_{1i}) - E[g_i \mid X_i]) \mid X_i\right]\right] \\ 
-      &\quad+ n^{-1} \sum_{i = 1}^n E\left[E\left[\frac{R_{2i}}{\pi_{2+}(X_i)} 
-      (a_2(X_i, Y_{2i}) - E[g_i \mid X_i]) \mid X_i\right]\right] \\
-      &\quad+ n^{-1} \sum_{i = 1}^n E\left[E\left[\frac{R_{1i}
-      R_{2i}}{\pi_{11}(X_i)} (g_i - a_2(X_i, Y_{2i}) - b_2(X_i, Y_{1i}) + E[g_i
-      \mid X_i]) \mid X_i \right]\right]\\
-      \intertext{Since $R_{1i} \perp Y_{1i} \mid X_i$, $R_{2i} \perp Y_{2i} \mid
-      X_i$, $(R_{1i}, R_{2i}) \perp (Y_{1i}, Y_{2i}) \mid X_i$ and $\pi_{1+},
-      \pi_{2+}$ and $\pi_{11}$ are all free of $Y_{1i}$ and $Y_{2i}$.}
-      &= n^{-1} \sum_{i = 1}^n E\left[\frac{R_{1i}}{\pi_{1+}(X_i)} E\left[
-        (E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i]) \mid X_i\right]\right] \\ 
-      &\quad+ n^{-1} \sum_{i = 1}^n E\left[\frac{R_{2i}}{\pi_{2+}(X_i)} 
-      E\left[E[g_i \mid X_i, Y_{2i}] - E[g_i \mid X_i] \mid X_i\right]\right] \\
-      &\quad+ n^{-1} \sum_{i = 1}^n E\left[\frac{R_{1i}
-      R_{2i}}{\pi_{11}(X_i)} E\left[(g_i - E[g_i \mid X_i, Y_{2i}] - 
-      E[g_i \mid X_i, Y_{1i}]+ E[g_i \mid X_i]) \mid X_i \right]\right]\\
-      \intertext{Since $E[E[g_i \mid X_i, Y_{ki}] \mid X_i] = E[g_i \mid X_i] =
-      0$,}
-      &= 0.
-    \end{align*}
-
-    Thus, if the outcome models are correctly specified $\hat \theta_{eff}$ is
-    unbiased. If the response models are correctly specified it is easy to see
-    that $\hat \theta_{eff}$ is also unbiased. This means that $\hat
-    \theta_{eff}$ is doubly robust.
-
-  \item However, one of the key steps is that \textit{all} the response
-    models are free of $Y$. In a previous iteration of Simulation 4, we had
-    adopted the framework of \cite{robins1997non} where we first to observe the
-    first variable and see if we observe the second variable. In this case, the
-    second step can depend on the result of the first step and this is what we
-    did. However, this makes it the case that $\pi_11$ is a function of $X_i$
-    and $Y_1$ and $Y_2$. In this case $\hat \theta_{eff}$ is not unbiased if the
-    response model is misspecified (or even just estimated).
-
-  \item If we modify Simulation 4, such that the second step of observed the
-    second variable is proportional to $\logistic(y_k)$ then we get the same
-    result as before:
-
-    \input{Tables/nonmonosim5c0.tex}
-    \input{Tables/nonmonosim5c0.1.tex}
-    \input{Tables/nonmonosim5c0.5.tex}
-
-\end{itemize}
-
-\newpage
-
-%\input{estimating_response_mod.tex}
-
-%\input{prop_full_nonmonoest.tex}
-
-\input{minimizing_variance.tex}
-
-\input{calibration_comparison.tex}
-
-\input{efficiency_twophase.tex}
+%\input{efficiency_twophase.tex}
 

### 2023-08-02 Simulations/generate_data.R

@@ -21,8 +21,8 @@
 # * Required Libraries *
 # **********************
 
-require(MASS)
-require(dplyr)
+library(MASS)
+library(dplyr)
 
 # ------------------------------------------------------------------------------
 

### 2023-08-02 Simulations/multiphase_est.R

@@ -1,0 +1,269 @@
+# Title: Simulations/multiphase_est.R
+# Author: Caleb Leedy
+# Date Created: May  3, 2023
+# Purpose: This document first compares the two three-phase sampling estimators:
+# F3P and K3P (see Notes/multiphase_estimators.tex or page 33 in the research
+# journal).
+
+# ******************
+# * Script Outline *
+# ******************
+# 1. Setup
+# 2. Compare three-phase estimators
+
+# *************
+# * Libraries *
+# *************
+
+library(dplyr)
+
+library(doParallel)
+library(doRNG)
+library(parallelly)
+
+# Run from Simulations/
+source("generate_data.R")
+source("nmono_est_funs.R")
+
+# ********************
+# * Estimator 1: F3P *
+# ********************
+f3p <- threephase_reg
+
+# ********************
+# * Estimator 2: K3P *
+# ********************
+k3p <- function(df) {
+
+  gfun <- "y2"
+
+  # Steps:
+  # 1. Get weights
+  # 2. Regression estimator Phase 2
+  # 3. Regression estimator Phase 1
+
+  # 1. Get weights
+  pi_11 <- df$p12 * df$p1 + df$p21 * df$p2
+  df$w <- 1 / pi_11
+
+  df$w_y1 <- 1 / df$p1
+
+  # 2. Regression estimator Phase 2
+  df_11 <- filter(df, r1 == 1, r2 == 1)
+  mean_y2 <- sum(df_11$w * df_11$y2) / sum(df_11$w)
+  mean_x2 <- sum(df_11$w * df_11$x) / sum(df_11$w)
+  mean_y1 <- sum(df_11$w * df_11$y1) / sum(df_11$w)
+
+  df_11[[gfun]] <- df_11[[gfun]] - mean_y2
+  df_11$x <- df_11$x - mean_x2
+  df_11$y1 <- df_11$y1 - mean_y1
+
+  mod <- lm(as.formula(paste0(gfun, " ~ x + y1")), weights = w, data = df_11)
+
+  # 3. Regression estimator Phase 1
+  mod_1 <- lm(as.formula(paste0(gfun, " ~ x")), weights = w, data = df_11)
+
+  df_1 <- 
+    filter(df, r1 == 1) %>%
+    mutate(w_y1 = 1 / p1) 
+
+  mean_x <- mean(df$x)
[…]

### 2023-08-02 Simulations/nmono_est_funs.R

@@ -28,7 +28,25 @@
 # * Required Libraries *
 # **********************
 
-require(dplyr)
+library(dplyr)
+library(CVXR)
+
+# ************************
+# * Additional Functions *
+# ************************
+
+#' Thanks Hengfang Wang for this matrix inversion function!
+hw_inv = function(X, eps = 1e-12) {
+  eig.X = eigen(X, symmetric = TRUE)
+  P = eig.X[[2]] 
+  lambda = eig.X[[1]] 
+  ind = lambda > eps
+  lambda[ind] = 1 / lambda[ind] 
+  lambda[!ind] = 0
+  ans = P %*% diag(lambda) %*% t(P)
+  return(ans)
+}
+
 
 # ------------------------------------------------------------------------------
 
@@ -148,6 +166,7 @@ mono_est_weights <- function(df) {
   sum(df$w2 * df$y2)
 
 }
+
 # **************************************
 # * Estimating Nonmonotone Missingness *
 # **************************************
@@ -444,4 +463,45 @@ nonmono_est_weights <- function(df) {
 
 }
 
+f3p_cali <- function(df) {
 
+  # Steps
+  # 1. Construct vectors / matrices.
+  # 2. Run calibration
+
+  # Do I need different steps if the data is monotone vs nonmonotone? No.
+  pi_11 <- df$p12 * df$p1 + df$p21 * df$p2
+  df$w <- 1 / pi_11
+  df_11 <- filter(df, r1 == 1, r2 == 1)
+  df_1 <- filter(df, r1 == 1)
+  df_1$w_1 <- 1 / df_1$p1
+
+  # tot_y2 <- sum(df_11$w * df_11$y2)
+  tot_x <- sum(df$x)
+  tot_z <- sum(df_1$w_1 * df_1$y1) 
+
+  n2 <- nrow(df_11)
+
+  # 1. Construct vectors / matrices.
+  w_vec <- matrix(df_11$w, ncol = 1)
+  L_mat <- base::diag(df_11$w)
+  L_inv <- base::diag(1 / df_11$w)
+  x_vec <- df_11$x
+  z_vec <- df_11$y1
+  c_vec <- Variable(n2)
+
+  # 2. Run calibration
+  const <- 
+    list(c_vec >= 0,
+         t(c_vec) %*% matrix(rep(1, n2), ncol = 1) == n2, # We estimate totals
[…]

### 2023-08-02 Simulations/regression_est.R

@@ -1,0 +1,68 @@
+# Title: regression_est.R
+# Author: Caleb Leedy
+# Date Created: May  5, 2023
+# Purpose: This file runs some code that I use for better understanding of the
+# regression estimator.
+
+# ***********
+# * Outline *
+# ***********
+# 1. Setup
+#   a. Libraries
+#   b. Additional Code
+# 2. Least Squares Estimator
+# 3. Least Squares Estimator without Intercept
+# 4. Regression Estimator
+
+# ------------------------------------------------------------------------------
+
+# *********
+# * Setup *
+# *********
+
+# *************
+# * Libraries *
+# *************
+
+library(dplyr)
+library(CVXR)
+
+# *******************
+# * Additional Code *
+# *******************
+
+source("generate_data.R")
+source("nmono_est_funs.R")
+
+df <- mono_mar(n = 1000)
+df_2 <- filter(df, r2 == 1)
+
+# ***************************
+# * Least Squares Estimator *
+# ***************************
+ls_est <- lm(y2 ~ x, data = df_2)
+
+# *********************************************
+# * Least Squares Estimator without Intercept *
+# *********************************************
+ls0_est <- lm(y2 ~ 0 + x, data = df_2)
+
+# ************************
+# * Regression Estimator *
+# ************************
+mean(df_2$y2) + (mean(df$x) - mean(df_2$x)) * ls_est$coefficients[2]
+
+# ***********
+# * Scratch *
+# ***********
+
+lm(y2 ~ x, data = df_2)
+lm(y2 ~ I(x - mean(df$x)), data = df_2)
+lm(y2 ~ 0 + I(x - mean(df$x)), data = df_2)
+lm(y2 ~ 0 + x, data = df_2)
+
+mean(df_2$y2) + (mean(df$x) - mean(df_2$x)) * ls0_est$coefficients[1]
+
+
+
+

### 2023-08-02 rsync_with_millipede.sh

@@ -1,0 +1,16 @@
+#!/bin/bash
+
+# File: rsync_with_millipede.sh
+# Created by: Caleb Leedy
+# Created on: April 12, 2021
+# Updated on: May 4, 2023
+# 
+# Purpose:
+# This file syncs my local repository with my repository on millipede.
+
+echo "Current Directory is `pwd`"
+
+# rsync -rtvz ./ cleedy@prontodtn.las.iastate.edu:~/Data_Integration/
+# rsync -rtvz ./ cleedy@millipede.cssm.iastate.edu:~/Data_Integration/
+rsync -rtvz Simulations/ cleedy@millipede.cssm.iastate.edu:~/Data_Integration/Simulations
+

### 2023-08-21 Notes/temp.tex
Updated notation.

--- 2023-04-17 Notes/nonmonotone_project.tex
@@
     weights as in Simulation 3, we estimate the weights for the proposed method.
+    However, unlike the true probabilities of being proportional to
+    $\logistic(y_k)$, this simulation has the true probabilities being
+    proportional to $\logistic(x_i)$. Thus, the true response model is the
+    following:
+
+    \begin{enumerate}
+      \item (Stage 1) Choose a variable observe. We choose $Y_1$ with
+        probability $0.4$, $Y_2$ with probability $0.4$ and neither with
+        probability $0.2$. If neither, $R_1 = 0$ and $R_2 = 0$. Otherwise,
+        continue to Step 2.
+      \item (Stage 2) With probability $p_i \propto \logistic(x_i)$, choose to
+        observe the other $Y$ variable. 
+    \end{enumerate}
+
+    This sequence generates the missingness indicators $R_1$ and $R_2$. Since,
+    the Stage 1 probabilities are fixed and known and the Stage 2 probabilities
+    only depend on $x_i$, the missingness is MAR and only a function of $x_i$.
     The algorithms to which we compare still use the oracle weights.
@@
 
-    This is clear evidence that the proposed method is doubly robust.
+    The previous version of this simulation used the Stage 2 probability $p_i
+    \propto \logistic(y_i)$ where $y_i$ was the observed $Y$ value in Stage 1.
+    However, under this setup, our method is biased. This is because
+
+    \[ E\left[E\left[\frac{R_1 R_2}{\pi_{11}(X, Y_1)} Y_1 \mid X\right]\right] 
+    \neq
+    E\left[\frac{E[Y_1 \mid X]}{\pi_{11}(X, Y_1)} E[R_1 R_2 \mid X]\right] \]
+
+    but
+
+    \[ E\left[E\left[\frac{R_1 R_2}{\pi_{11}(X)} Y_1 \mid X\right]\right] 
+    =
+    E\left[\frac{E[Y_1 \mid X]}{\pi_{11}(X)} E[R_1 R_2 \mid X]\right].\]
+
+    Thus, there is strong evidence that the proposed method is 
+    doubly robust because it is robust to both misspecification in the outcome
+    and response model.
+\end{itemize}
+
+\newpage
+
+\section*{Missingness Mechanism}
+
+\begin{itemize}
+  \item First I am going to reproduce the proof of double robustness that we
+    talked about during our last meeting. I think it is insightful for future
+    comments:
+
+    \begin{align*}
+      E[\hat \theta_{eff} - \theta_n] 
+      &= E\left[n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] - g_i\right]\\
+      &\quad+ E\left[n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)}
+      (b_2(X_i, Y_{1i}) - E[g_i \mid X_i])\right]\\
+      &\quad+ E\left[n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)}
+      (a_2(X_i, Y_{2i}) - E[g_i \mid X_i])\right]\\
+      &\quad+ E\left[n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_i)} 
+      (g_i - a_2(X_i, Y_{2i}) - b_2(X_i, Y_{1i}) + E[g_i \mid X_i])\right]\\
+      &= n^{-1} \sum_{i = 1}^n (E[E[g_i \mid X_i]] - E[g_i])\\
+      &\quad+ n^{-1} \sum_{i = 1}^n E\left[E\left[\frac{R_{1i}}{\pi_{1+}(X_i)} 
+      (b_2(X_i, Y_{1i}) - E[g_i \mid X_i]) \mid X_i\right]\right] \\ 
+      &\quad+ n^{-1} \sum_{i = 1}^n E\left[E\left[\frac{R_{2i}}{\pi_{2+}(X_i)} 
+      (a_2(X_i, Y_{2i}) - E[g_i \mid X_i]) \mid X_i\right]\right] \\
+      &\quad+ n^{-1} \sum_{i = 1}^n E\left[E\left[\frac{R_{1i}
+      R_{2i}}{\pi_{11}(X_i)} (g_i - a_2(X_i, Y_{2i}) - b_2(X_i, Y_{1i}) + E[g_i
+      \mid X_i]) \mid X_i \right]\right]\\
+      \intertext{Since $R_{1i} \perp Y_{1i} \mid X_i$, $R_{2i} \perp Y_{2i} \mid
+      X_i$, $(R_{1i}, R_{2i}) \perp (Y_{1i}, Y_{2i}) \mid X_i$ and $\pi_{1+},
+      \pi_{2+}$ and $\pi_{11}$ are all free of $Y_{1i}$ and $Y_{2i}$.}
+      &= n^{-1} \sum_{i = 1}^n E\left[\frac{R_{1i}}{\pi_{1+}(X_i)} E\left[
+        (E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i]) \mid X_i\right]\right] \\ 
+      &\quad+ n^{-1} \sum_{i = 1}^n E\left[\frac{R_{2i}}{\pi_{2+}(X_i)} 
+      E\left[E[g_i \mid X_i, Y_{2i}] - E[g_i \mid X_i] \mid X_i\right]\right] \\
+      &\quad+ n^{-1} \sum_{i = 1}^n E\left[\frac{R_{1i}
+      R_{2i}}{\pi_{11}(X_i)} E\left[(g_i - E[g_i \mid X_i, Y_{2i}] - 
+      E[g_i \mid X_i, Y_{1i}]+ E[g_i \mid X_i]) \mid X_i \right]\right]\\
+      \intertext{Since $E[E[g_i \mid X_i, Y_{ki}] \mid X_i] = E[g_i \mid X_i] =
+      0$,}
+      &= 0.
+    \end{align*}
+
+    Thus, if the outcome models are correctly specified $\hat \theta_{eff}$ is
+    unbiased. If the response models are correctly specified it is easy to see
+    that $\hat \theta_{eff}$ is also unbiased. This means that $\hat
+    \theta_{eff}$ is doubly robust.
+
+  \item However, one of the key steps is that \textit{all} the response
+    models are free of $Y$. In a previous iteration of Simulation 4, we had
+    adopted the framework of \cite{robins1997non} where we first to observe the
+    first variable and see if we observe the second variable. In this case, the
+    second step can depend on the result of the first step and this is what we
+    did. However, this makes it the case that $\pi_11$ is a function of $X_i$
+    and $Y_1$ and $Y_2$. In this case $\hat \theta_{eff}$ is not unbiased if the
+    response model is misspecified (or even just estimated).
+
+  \item If we modify Simulation 4, such that the second step of observed the
+    second variable is proportional to $\logistic(y_k)$ then we get the same
+    result as before:
+
+    \input{Tables/nonmonosim5c0.tex}
+    \input{Tables/nonmonosim5c0.1.tex}
+    \input{Tables/nonmonosim5c0.5.tex}
 
@@
 
-\section*{Full Nonmonotone Estimator}
+% \input{estimating_response_mod.tex}
+
+\section*{Minimizing the Variance}
 
 \begin{itemize}
-  \item
+  \item The goal of this section is to find optimal values of $b_2(X, Y_1)$ and
+    $a_2(X, Y_2)$ such that the variance of $\hat \theta_{eff}$ is minimized.
+
+  \item Recall:
+
+    \begin{align*}
+      \hat \theta_{eff} - \hat \theta_n 
+      &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \left(1 - \frac{R_{1i}}{\pi_{1+}}
+      - \frac{R_{2i}}{\pi_{2+}} + \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \\
+      &+ n^{-1} \sum_{i = 1}^n b_2(X_i, Y_{1i}) \left(\frac{R_{1i}}{\pi_{1+}} -
+      \frac{R_{1i}R_{2i}}{\pi_{11}}\right)\\
+      &+ n^{-1} \sum_{i = 1}^n a_2(X_i, Y_{2i}) \left(\frac{R_{2i}}{\pi_{2+}} -
+      \frac{R_{1i}R_{2i}}{\pi_{11}}\right)\\
+      &+ n^{-1} \sum_{i = 1}^n g_i \left(\frac{R_{1i}R_{2i}}{\pi_{11}} -
+      1\right) \\
+      &\equiv A + B + C + D.
+    \end{align*}
+
+  \item Notice that we will suppress the fact that response models are functions
+    of $X$ (i.e. we write $\pi_{11}$ instead of $\pi_{11}(X)$).
+
+  \item To compute the variance, we first solve for each covariance combination.
+    Basically, all these computations rely on the following ideas. First, we
+    assume that the response model is correctly specified. Consequently,
+    $E[A] = E[B] = E[C] = E[D] = 0$ and things work out better. This helps when
+    we take the covariance conditional on $X$ because the inner expectations are
+    zero. The second key insight it to notice that $E[R_{j}^k] = E[R_j]$ for $j
+    \in \{1, 2\}$ and $k \in \mathbb{N}$. This is because $R$ is a binary
+    variable. Third, since we assume that the response models are correctly
+    specified, we have $E[R_1 \mid X] = \pi_{1+}$, $E[R_2 \mid X] = \pi_{2+}$,
+    and $E[R_1, R_2 \mid X] = \pi_{11}$.
+
+    The overall approach to each of these computations is the following:
+    (1) take conditional expectations with respect to $X$ (the $\Cov(E[\cdot])$
+    term is zero), (2) expand the covariance to $E[XY] - E[X]E[Y]$ (the second
+    term is also zero), (3) by the MAR assumption $g, a_2, b_2$ are independent
+    of $R_1$ and $R_2$ and we can take the latter out of the expectation, (4)
+    evaluate and simplify expressions involving $E[R]$.
+
+    \begin{align*}
+      \Cov(A, B) &= n^{-2} \sum_{i = 1}^n E\left[\Cov\left(E[g_i \mid X] \left(1
+      - \frac{R_{1i}}{\pi_{1+}} - \frac{R_{2i}}{\pi_{2+}} +
+      \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \mid X_i,\right. \right.\\
+      &\qquad \qquad \qquad \qquad b_2(X_i, Y_{1i}) \left. \left.
+      \left(\frac{R_{1i}}{\pi_{1+}} - \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \mid
+      X_i\right)\right]\\
+      &= n^{-1} E\left[E\left[E[g \mid X] \left(1 - \frac{R_{1i}}{\pi_{1+}} -
+      \frac{R_{2i}}{\pi_{2+}} + \frac{R_{1i}R_{2i}}{\pi_{11}}\right)
+      b_2(X_i, Y_{1i}) \left(\frac{R_{1i}}{\pi_{1+}} -
+      \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \mid X \right]\right] \\
+      &= n^{-1} E\left[E[g \mid X] E[b_2(X, Y_1) \mid X]
+      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
+      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right].
+    \end{align*}
+
+    By symmetry,
+    \begin{align*}
+      \Cov(A, C) 
+      &= n^{-1} E\left[E[g \mid X] E[a_2(X, Y_2) \mid X]
+      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
+      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right].
+    \end{align*}
+
+    \begin{align*}
+      \Cov(A, D) &= n^{-1} E\left[E\left[E[g \mid X] \left(1 -
+      \frac{R_{1i}}{\pi_{1+}} - \frac{R_{2i}}{\pi_{2+}} +
+      \frac{R_{1i}R_{2i}}{\pi_{11}}\right) g \left(\frac{R_1 R_2}{\pi_{11}} -
+      1\right) \mid X\right] \right]\\
+      &= n^{-1} E\left[E[g \mid X]^2 \left(\frac{-1}{\pi_{1+}} -
+      \frac{1}{\pi_{2+}} + 2\right)\right].
+    \end{align*}
+
+    \begin{align*}
+      \Cov(B, C) &= n^{-1} E\left[E[b_2(X, Y_1) \mid X] E[a_2(X, Y_2) \mid X]
+      E\left[\left(\frac{R_1}{\pi_{1+}} - \frac{R_1 R_2}{\pi_{11}}\right)
+      \left(\frac{R_2}{\pi_{2+}} - \frac{R_1 R_2}{\pi_{11}}\right) \mid
+      X\right]\right]\\
+      &= n^{-1} E\left[E[b_2(X, Y_1) \mid X] E[a_2(X, Y_2) \mid X]
+      \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} - \frac{1}{\pi_{1+}} -
+      \frac{1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right)\right].
+    \end{align*}
+
+    \begin{align*}
+      \Cov(B, D) &= n^{-1} E\left[ E[b_2(X, Y_1) \mid X] E[g \mid X]
+      E\left[\left(\frac{R_1}{\pi_{1+}} - \frac{R_1 R_2}{\pi_{11}}\right)
+      \left(\frac{R_1 R_2}{\pi_{11}} - 1\right) \mid X\right] \right]\\
+      &= n^{-1} E\left[ E[b_2(X, Y_1) \mid X] E[g \mid X]
+      \left(\frac{1}{\pi_{1+}} - \frac{1}{\pi_{11}}\right)\right].
+    \end{align*}
+
+    By symmetry,
+    \begin{align*}
+      \Cov(C, D) 
+      &= n^{-1} E\left[ E[a_2(X, Y_2) \mid X] E[g \mid X]
+      \left(\frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}}\right)\right].
+    \end{align*}
+
+    We also compute the variance terms,
+
+    \begin{align*}
+      \Cov(A, A) &= n^{-1} E\left[ E\left[ E[g \mid X]^2 \left(1
+      - \frac{R_{1i}}{\pi_{1+}} - \frac{R_{2i}}{\pi_{2+}} +
+      \frac{R_{1i}R_{2i}}{\pi_{11}}\right)^2 \mid X\right] \right]\\
+      &= n^{-1} E \left[ E[g \mid X]^2 \left(-1 + \frac{2 \pi_{11}}{\pi_{1+}
+      \pi_{2+}} - \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
+      \frac{1}{\pi_{11}}\right)\right].
+    \end{align*}
+
+    \begin{align*}
+      \Cov(B, B) &= n^{-1} E\left[ E\left[b(X, Y_1)^2 \left(\frac{R_1}{\pi_{1+}}
+      - \frac{R_1 R_2}{\pi_{11}}\right)^2 \mid X\right] \right]\\
+      &= n^{-1} E\left[ E[b_2(X, Y_1)^2 \mid X] \left(\frac{-1}{\pi_{1+}} +
+      \frac{1}{\pi_{11}}\right)\right].
+    \end{align*}
+
+    \begin{align*}
+      \Cov(C, C) &= n^{-1} E\left[ E[a_2(X, Y_2)^2 \mid X]
+      \left(\frac{-1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right) \right].
+    \end{align*}
+
+    \begin{align*}
+      \Cov(D, D) &= n^{-1} E\left[ E\left[g_i^2 \left(\frac{R_1 R_2}{\pi_{11}} -
+      1\right)^2 \mid X\right] \right]\\
+      &= n^{-1} E\left[ E[g^2 \mid X] \left(\frac{1}{\pi_{11}} -
+      1\right)\right].
+    \end{align*}
+
+    This means that
+    \begin{align*}
+      \Var(\hat \theta_{eff} - \hat \theta_n) 
+      &= \Cov(A, A) + 2 \Cov(A, B) + 2 \Cov(A, C) + 2 \Cov(A, D) + \Cov(B, B) \\
+      &+ 2 \Cov(B, C) + 2 \Cov(B, D) + \Cov(C, C) + 2 \Cov(C, D) + \Cov(D, D) \\
+      &= n^{-1} E \left[ E[g \mid X]^2 \left(-1 + \frac{2 \pi_{11}}{\pi_{1+}
+      \pi_{2+}} - \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
+      \frac{1}{\pi_{11}}\right)\right] \\
+      &+ 2 n^{-1} E\left[E[g \mid X] E[b_2(X, Y_1) \mid X]
+      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
+      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right]\\
+      &+ 2 n^{-1} E\left[E[g \mid X] E[a_2(X, Y_2) \mid X]
+      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
+      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right]\\
+      &+ 2 n^{-1} E\left[E[g \mid X]^2 \left(\frac{-1}{\pi_{1+}} -
+      \frac{1}{\pi_{2+}} + 2\right)\right]\\
+      &+ n^{-1} E\left[ E[b_2(X, Y_1)^2 \mid X] \left(\frac{-1}{\pi_{1+}} +
+      \frac{1}{\pi_{11}}\right)\right]\\
+      &+ 2 n^{-1} E\left[E[b_2(X, Y_1) \mid X] E[a_2(X, Y_2) \mid X]
+      \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} - \frac{1}{\pi_{1+}} -
+      \frac{1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right)\right]\\
+      &+ 2 n^{-1} E\left[ E[b_2(X, Y_1) \mid X] E[g \mid X]
+      \left(\frac{1}{\pi_{1+}} - \frac{1}{\pi_{11}}\right)\right]\\
+      &+  n^{-1} E\left[ E[a_2(X, Y_2)^2 \mid X]
+      \left(\frac{-1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right) \right]\\
+      &+ 2 n^{-1} E\left[ E[a_2(X, Y_2) \mid X] E[g \mid X]
+      \left(\frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}}\right)\right]\\
+      &+ n^{-1} E\left[ E[g^2 \mid X] \left(\frac{1}{\pi_{11}} -
+      1\right)\right].
+    \end{align*}
+
+    Differentiating yields:
+
+    \begin{align*}
+      \frac{\partial}{\partial a_2} \Var(\hat \theta_{eff} - \hat \theta_n)
+      &= E\left[E[g \mid X] \left(\frac{1}{\pi_{1+}} + \frac{2}{\pi_{2+}} -
+      \frac{2}{\pi_{11}} - \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right] \\
+      &+ E\left[E[b_2(X, Y_1) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
+      \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
+      \frac{1}{\pi_{11}}\right)\right]\\
+      &+ E\left[ E[a_2(X, Y_2) \mid X] \left(\frac{-1}{\pi_{2+}} +
+      \frac{1}{\pi_{11}}\right)\right]\\
+      &\equiv 0, \text{ and }\\
+      \frac{\partial}{\partial b_2} \Var(\hat \theta_{eff} - \hat \theta_n)
+      &= E\left[E[g \mid X] \left(\frac{2}{\pi_{1+}} + \frac{1}{\pi_{2+}} -
+      \frac{2}{\pi_{11}} - \frac{\pi_{11}}{\pi_{1+}\pi_{2+}}\right)\right]\\
+      &+ E\left[E[a_2(X, Y_2) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
+      \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
+      \frac{1}{\pi_{11}}\right)\right]\\
+      &+ E\left[ E[b_2(X, Y_2) \mid X] \left(\frac{-1}{\pi_{1+}} +
+      \frac{1}{\pi_{11}}\right)\right]\\
+      &\equiv 0. 
+    \end{align*}
+
+    Substitution shows that these constraints are equivalent to:
+
+    \begin{align*}
+      E&\left[E[g \mid X] \left(\frac{-1}{\pi_{1+}} +
+    \frac{1}{\pi_{2+}}\right)\right] + E\left[ E[b_2(X, Y_1) \mid X]
+    \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -\frac{1}{\pi_{2+}}\right)\right]\\
+      &- E\left[E[a_2(X, Y_2) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} - 
+    \frac{1}{\pi_{1+}}\right)\right] \equiv 0
+    \end{align*}
+
+    where is the same as,
+
+    \begin{align*}
+      E&\left[ E[b_2(X, Y_1) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
+    \frac{1}{\pi_{2+}}\right) \right] + E\left[E[g \mid X]
+      \left(\frac{1}{\pi_{2+}}\right)\right] \\
+      &=
+    E\left[ E[a_2(X, Y_2) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
+    \frac{1}{\pi_{1+}}\right) \right] + E\left[E[g \mid X]
+    \left(\frac{1}{\pi_{1+}}\right)\right].
+    \end{align*}
+
+  \item These constraints can be satisfied (this is sufficient but maybe not
+    necessary) if
+
+    \begin{align*}
+      E&\left[(E[b_2(X, Y_1) \mid X] - E[a_2(X, Y_2) \mid X])
+      \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right) \right] = 0\\
+      E&\left[\left(\frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}}\right)(E[a_2(X, Y_2)
+      \mid X] + E[b_2(X, Y_1) \mid X] - 2E[g \mid X])\right] = 0.
+    \end{align*}
 
@@
 
+%\textbf{Questions for Dr. Kim}
+%\begin{itemize}
+%  \item Where do we go from here?
+%  \item This is a functional expectation that I need to calibrate. Should I use
+%    some sort of basis spline to try to estimate these expectations and then set
+%    them equal to zero? I don't think that I have worked with this kind of
+%    constraint before.
+%\end{itemize}
+
+%\input{prop_full_nonmonoest.tex}
+
 \newpage
+
+\input{calibration_comparison.tex}
+
+\input{efficiency_twophase.tex}
 

### 2023-08-21 R/opt_est.R

@@ -1,0 +1,15 @@
+# Title: opt_est.R
+# Created by: Caleb Leedy
+# Created on: August 19, 2023
+# Purpose: This file contains the functions needed for computation of the 
+# optimal estimator in the non-monotone case.
+
+# ************
+# * Overview *
+# ************
+# The goal of this file is to create code to estimate the following:
+# \hat \theta = n^{-1} \sum_{i = 1}^n E[g_i | X_i] +
+#   n^{-1} \sum_{i = 1}^n (R_{1i} / \pi_{1+}) (E[g_i | X_i, Y_{1i}] - E[g_i | X_i]) +
+#   n^{-1} \sum_{i = 1}^n (R_{2i} / \pi_{2+}) (E[g_i | X_i, Y_{2i}] - E[g_i | X_i]) +
+#   n^{-1} \sum_{i = 1}^n (R_{1i}R_{2i} / \pi_{11}) (E[g_i | X_i, Y_{1i}, Y_{2i}] -
+#     E[g_i | X_i, Y_{1i}] - E[g_i | X_i, Y_{2i}] + E[g_i | X_i])

### 2023-08-21 Simulations/optimal_simulation.R

@@ -1,0 +1,14 @@
+# Title: optimal_simulation.R
+# Created by: Caleb Leedy
+# Created on: August 19, 2023
+# Purpose: This file contains the code to test the proposed estimator with a
+# optimal estimator in the case where everything is normal.
+
+# ***********
+# * Outline *
+# ***********
+# 1. Generate data.
+
+# *************
+# * Libraries *
+# *************

### 2023-08-25 Notes/annote_bib.tex
Prepared report for Dr. Fuller about linear estimator being better than proposed estimator.

@@
 
+% \subsection*{Chen and Kim (2014)}
+% \begin{description}
+%   \item[Title:] Population Empirical Likelihodd for Nonparametric Inference in
+%     Survey Sampling
+%   \item[Author:] Sixia Chen and Jae Kwang Kim
+%   \item[Journal (Year):] Statistica Sinica (2014)
+%   \item[Summary:]
+%   \item[Limitations:]\;
+%     \begin{itemize}
+%       \item
+%     \end{itemize}
+%   \item[Extensions:]\;
+%     \begin{itemize}
+%       \item
+%     \end{itemize}
+% \end{description}
+
+\subsection*{Chen et al. (2022)}
+\begin{description}
+  \item[Title:] Nonparametric Mass Imputation for Data Integration
+  \item[Author:] Sixia Chen, Shu Yang, Jae Kwang Kim
+  \item[Journal (Year):] JSSM (2022)
+  \item[Summary:] This paper aims to combine data from a probability sample
+    $A$ and a non-probability sample $B$ using non-parametric methods. We 
+    assume the $X$ is observed in $A$ and $B$ but $Y$ is only available in $B$. 
+    It achieves this goal by using a kernel estimator for the outcome model
+    $m(x)$ on sample $B$ and then using this model to predict missing values of
+    $Y$ for $A$. Since the sample weights in $A$ are valid, these are used to 
+    adjust the predicted values $\hat m(x)$. Variance estimates are also given.
+  \item[Limitations:]\;
+    \begin{itemize}
+      \item The model require MAR for the response model in $B$.
+    \end{itemize}
+  \item[Extensions:]\;
+    \begin{itemize}
+      \item Can we extend this to multiple non-probability samples?
+    \end{itemize}
+\end{description}
+
+\subsection*{Hidiroglou (2001)}
+\begin{description}
+  \item[Title:] Double Sampling
+  \item[Author:] M. A. Hidiroglou
+  \item[Journal (Year):] Survey Methodology (2001)
+  \item[Summary:] This paper produces the optimal regression estimator for 
+    non-nested probability samples. This occurs when sample $A$ observes $X$ 
+    while $B$ observes $X$ and $Y$ but unlike two-phase sampling there is no
+    overlap between the observed values for $X$ in $A$ and $B$.
+
+    This paper also explores the connection between non-nested samples and 
+    calibration.
+  \item[Limitations:]\;
+    \begin{itemize}
+      \item Unclear to me which class of estimators we are optimizing within.
+    \end{itemize}
+  \item[Extensions:]\;
+    \begin{itemize}
+      \item How to extend to a third category?
+    \end{itemize}
+\end{description}
+
+\subsection*{Park and Kim (2019)}
+\begin{description}
+  \item[Title:] Mass Imputation for Two-Phase Sampling
+  \item[Author:] Seho Park and Jae Kwang Kim
+  \item[Journal (Year):] arXiv (2019)
+  \item[Summary:] This paper proposes a mass imputation estimator for 
+    estimating the total from a two-phase sample. Instead of using an
+    estimate from phase 1 data and combining it with phase 2 data, this
+    approach uses a regression model to predict the missing data in phase 
+    1. The paper shows how this result is consistent with previous 
+    estimates and they provide a variance estimator.
+  \item[Limitations:]\;
+    \begin{itemize}
+      \item What if the regression model is incorrect?
+      \item Does this account for enough uncertainty if just the predicted
+        values are used?
+    \end{itemize}
+  \item[Extensions:]\;
+    \begin{itemize}
+      \item How can this be extended to another phase.
+    \end{itemize}
+\end{description}
+
 % \begin{description}

### 2023-08-25 Notes/model_assumptions.tex
Prepared report for Dr. Fuller about linear estimator being better than proposed estimator.

\documentclass[12pt]{article}

\usepackage{amsmath, amssymb, mathrsfs, fancyhdr}
\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
\usepackage{biblatex}
\usepackage{booktabs}
\usepackage{float}

\addbibresource{references.bib}

\hypersetup{colorlinks = true, urlcolor = black}

\headheight     15pt
\topmargin      -1.5cm   % read Lamport p.163
\oddsidemargin  -0.04cm  % read Lamport p.163
\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
\textwidth      16.59cm
\textheight     23.94cm
\parskip         7.2pt   % sets spacing between paragraphs
\parindent         0pt   % sets leading space for paragraphs
\pagestyle{empty}        % Uncomment if don't want page numbers
\pagestyle{fancyplain}

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}

\renewcommand{\arraystretch}{1.2}

\begin{document}

\lhead{Caleb Leedy}
\chead{Non-monotone Missingness}
%\chead{STAT 615 - Advanced Bayesian Methods}
%\rhead{Page \thepage\ of \pageref{LastPage}}
\rhead{\today}

\section*{Introduction}

This document reports the optimality of the proposed non-monotone estimator.
It turns out that the proposed estimator is not optimal. We show a better
estimator and present a simulation study.

\section*{Setup}

\begin{itemize}
  \item Let $U = \{1, \dots N\}$ be a finite population.
  \item Consider the random variables $(X_i, Y_{1i}, Y_{2i}, \pi_i) 
    \stackrel{\text{ind}}{\sim} F$ for some unknown $F$ for all $i \in U$.
  \item Let $I_i \stackrel{\text{ind}}{\sim} \pi_i$ for $i \in U$ be the 
    sample inclusion indicator.
  \item We assume that $x_i$ is observed throughout the finite population.
  \item However, $(y_{1i}, y_{2i}, \pi_i)$ is only observed if $I_i = 1$.
  \item The goal is to estimate $\theta = E[g(X, Y_1, Y_2)]$ in the population.
  \item The proposed estimator is

  \begin{align}\label{eq:prop}
    \hat \theta_{\text{eff}} 
    &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \nonumber \\ 
    &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)}
      (E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i])\nonumber  \\ 
    &+ n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)}
      (E[g_i \mid X_i, Y_{2i}] - E[g_i \mid X_i])\nonumber  \\ 
    &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_i)}
      (g_i - E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i, Y_{2i}] + E[g_i \mid X_i])
  \end{align}

  \item The goal is to show that this estimator is optimal within the class
    of estimators:

  \begin{align*}
    \hat \theta_{\text{eff}} 
    &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \\ 
    &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)}
      (b(X_i, Y_{1i}) - E[g_i \mid X_i]) \\ 
    &+ n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)}
      (a(X_i, Y_{2i}) - E[g_i \mid X_i]) \\ 
    &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_i)}
      (g_i - b_i - a_i + E[g_i \mid X_i]).
  \end{align*}

\end{itemize}

\section*{Results}

I have done a lot to explore this area but I have several points of confusion.
The first part recaps what Dr. Fuller and I discussed this past week.

\begin{itemize}
  \item Let $g = E[Y_2]$, and suppose that we know the distribution of $X$ and 
    the covariance structure of $[X, Y_1, Y_2, \pi]$. Then instead of modeling
    the relationship between $X$ and $Y_k$, we can use difference estimators:
    $W_1 \equiv Y_1 - b_1 \tilde X$ and $W_2 = \equiv Y_2 - b_2 \tilde X$, where
    $\tilde X = X - E[X]$.
    To minimize the variance of $W_k$ we can choose the optimal value of $b_k$,
     which is

     \[b_k = \frac{\Cov(Y_k, X)}{\Var(X)}.\]

    Since these values are known, $b_k$ is known. This means that we now have
    the following table:

    \begin{figure}[!ht]
      \centering
      \begin{tabular}{lrr}
        \toprule
         & $W_1$ & $W_2$ \\
         \midrule
        $A_{11}$ & \checkmark & \checkmark \\
        $A_{10}$ & \checkmark & \\
        $A_{01}$ & & \checkmark \\
        $A_{00}$ & & \\
        \bottomrule
      \end{tabular}
    \end{figure}

    Because we assumed that the distribution of $X$ is known, the section
    $A_{00}$ contains no additional information about $Y_1$ or $Y_2$. Let
    $\mu_k = E[Y_k]$.

  \item We now consider a normal model:

    \[ 
      \begin{pmatrix}
        x_i \\ e_{1i} \\ e_{2i}
      \end{pmatrix} \stackrel{ind}{\sim}
      N \left(
      \begin{bmatrix}
        0 \\ 0 \\ 0
      \end{bmatrix},
      \begin{bmatrix}
        1 & 0 & 0 \\
        0 & \sigma_{11} & 0 \\ 
        0 & 0 & \sigma_{22}
      \end{bmatrix},
      \right)
    \]

    and define $y_{1i} = x_i + e_{1i}$ and $y_{2i} = x_i + e_{2i}$. Then $b_1 =
    b_2 = 1$. We define $\bar w_k^{(ij)}$ as the mean of $y_k$ in segment
    $A_{ij}$. This means that we have means $\bar w_1^{(11)}$, $\bar
    w_2^{(11)}$, $\bar w_1^{(10)}$, and $\bar w_2^{(01)}$. Let 
    $W = [\bar w_1^{(11)}, \bar w_2^{(11)}, \bar w_1^{(10)}, \bar w_2^{(01)}]'$,
    then for $n_{ij} = |A_{ij}|$, we have

    \[W - M \mu \sim N(\vec 0, V)\]

    where

    \[M = 
      \begin{bmatrix}
        1 & 0 \\
        0 & 1 \\
        1 & 0 \\
        0 & 1 \\
      \end{bmatrix}
      \text{ and }
      V = 
      \begin{bmatrix}
        \frac{\sigma_{11}}{n_{11}} & 0 & 0 & 0 \\
        0 & \frac{\sigma_{22}}{n_{11}} & 0 & 0 \\
        0 & 0 & \frac{\sigma_{11}}{n_{10}} & 0 \\
        0 & 0 & 0 & \frac{\sigma_{22}}{n_{01}} \\
      \end{bmatrix}.
    \]

    Thus, the BLUE for $\mu = [\mu_1, \mu_2]'$ is

    \[ \hat \mu = (M' V^{-1} M)^{-1} M' V^{-1} W. \]

  \item If we do the matrix algebra, $V^{-1} = 
    \text{diag}\left(\frac{n_{11}}{\sigma_{11}}, \frac{n_{11}}{\sigma_{22}},
    \frac{n_{10}}{\sigma_{11}}, \frac{n_{01}}{\sigma_{22}}\right)$ and hence,

    {\renewcommand{\arraystretch}{2}
    \begin{equation}\label{eq:wls}
    \hat \mu = 
      \begin{bmatrix}
        \frac{n_{11}\bar w_1^{(11)} + n_{10}\bar w_1^{(10)}}{n_{11} + n_{10}} \\
        \frac{n_{11}\bar w_2^{(11)} + n_{01}\bar w_2^{(01)}}{n_{11} + n_{01}} \\
      \end{bmatrix}
    \end{equation}
    }

  \item Since this is the BLUE, we would expect it to be at least as good as
    the proposed estimator. And if the proposed estimator is optimal it should
    be equivalent to the BLUE in the case that $X$, $Y_1$ and $Y_2$ are normal.
    So I ran a simulation study to test this. For $i = \{1, \dots, n = 1000\}$,

    \[
      \begin{pmatrix}
        x_i \\ e_{1i} \\ e_{2i}
      \end{pmatrix} \stackrel{ind}{\sim}
      N \left(
      \begin{bmatrix}
        0 \\ 0 \\ 0
      \end{bmatrix},
      \begin{bmatrix}
        1 & 0 & 0 \\
        0 & 1 & 0 \\ 
        0 & 0 & 1 
      \end{bmatrix},
      \right) \text{ and }
      y_{1i} = x_i + e_{1i}, y_{2i} = x_i + e_{2i}.
    \]

    Each observation $i$ was then assigned a segment $A_{11}$, $A_{10}$,
    $A_{01}$ or $A_{00}$ independently with each draw having probability
    $p_{11} = 0.4$, $p_{10} = 0.2$, $p_{01}= 0.2$ and $p_{00} = 0.2$
    respectively. This means that $\pi_{11} = 0.4$ and 
    $\pi_{1+} = 0.6 = \pi_{2+}$. We let $g = E[Y_2]$ and we test the following
    estimators:

    \begin{itemize}
      \item Oracle: This computes the average value of $Y_2$ in all 
        observations (even when $Y_2$ is not supposed to be observed).
        \[\hat \theta = n^{-1} \sum_{i = 1}^n y_{2i}.\]

      \item CC: This computes the average value of $Y_2$ in all observations 
        in which $Y_2$ is observed.
        \[\hat \theta = \frac{\sum_{i = 1}^n \delta_{2i} y_{2i}}
          {\sum_{i = 1}^n \delta_{2i}}.\]

      \item IPW: This computes the survey weighted average of $Y_2$ when both
        $Y_1$ and $Y_2$ are observed.
        \[\hat \theta = \frac{\sum_{i = 1}^n \delta_{1i}\delta_{2i} y_{2i}}
        {\sum_{i = 1}^n \delta_{1i}\delta_{2i}}.\]
      \item Proposed: This is the proposed estimator from Equation~\ref{eq:prop}.
      \item WLS: This is the weighted linear estimator from Equation~\ref{eq:wls}.
        (Note, since $g = E[Y_2]$ this only contains the second element from
        Equation~\ref{eq:wls}.)
    \end{itemize}

    The results are shown in the table below. This simulation was run with the 
    number of observations $n = 1000$ and the Monte Carlo sample size of 
    $B = 1000$.

    \begin{table}[ht!]
      \centering
      \caption{This table shows the estimators of $\theta = E[Y_2]$. The true 
      value of $\theta$ is $0$. The bias column shows the average bias of the 
      estimator and the actual value of $\theta = 0$ across the $B = 1000$ 
      simulations. The SD column shows the average standard deviation across the 
      $B = 1000$ simulations for each algorithm. The Tstat column displays the 
      t-statistic of a t-test comparing the estimator to the actual value. 
      The value of this column is computed via $\frac{\bar{\hat{\theta}} - 
      \theta}{\sqrt{\Var{\hat \theta}/B}}$. The Pval column displays the 
      p-value of the t-statistic.}
      \begin{tabular}[t]{lrrrr}
        \toprule
        Algorithm & Bias & SD & Tstat & Pval\\
        \midrule
        Oracle & -0.001 & 0.045 & -0.953 & 0.170\\
        CC & 0.000 & 0.056 & -0.059 & 0.476\\
        IPW & 0.000 & 0.071 & -0.183 & 0.428\\
        Prop & 0.000 & 0.051 & -0.131 & 0.448\\
        WLS & -0.001 & 0.042 & -0.470 & 0.319\\
        \bottomrule
      \end{tabular}
      \label{tab:res}
    \end{table}

  \item There are a couple things of note from Table~\ref{tab:res}. First, the 
    WLS estimator has a smaller standard error than the Oracle estimator. This 
    seems incorrect. (Please see the next section for questions.) Second, the 
    proposed estimator did not have a similar standard error as the WLS estimator.
    This suggests that it is not optimal.
\end{itemize}

\section*{Questions for Dr. Fuller}

\begin{enumerate}
  \item For the linear estimator I don't think that I am incorporating sampling
    weights. I am a bit confused about how to do this. Should I compute the 
    means $\bar w_{k}^{(ij)}$ using weights according to the sample weights? 
  \item Is it ok for the linear estimator to outperform the oracle estimator?
    This seems odd to me because the oracle estimator seems to contain more 
    information.
  \item On Tuesday, you asked me to compute the linear estimator in terms of 
    means of $Y$ and estimators of $0$. I think that the WLS estimator is the 
    estimator in terms of estimators of $0$ because $y_k - b_k x$ is an 
    estimator of $0$. To transform the estimator into an estimator in terms of
    $Y$, should I write the form in terms of $Y$ instead of $W$?
\end{enumerate}

\end{document}

### 2023-08-25 Presentations/kimfuller_intro.tex
Prepared report for Dr. Fuller about linear estimator being better than proposed estimator.

\documentclass{beamer} %
\usetheme{CambridgeUS}
\usepackage[latin1]{inputenc}
\usefonttheme{professionalfonts}
\usepackage{times}
\usepackage{graphics}
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{biblatex}

\usetikzlibrary{shapes, positioning, arrows}
\addbibresource{references.bib}

\renewcommand{\L}{\ensuremath{\mathcal{L}}}
\providecommand{\bY}{\ensuremath{\mathbf{Y}}}
\providecommand{\bX}{\ensuremath{\mathbf{X}}}
\providecommand{\bV}{\ensuremath{\mathbf{V}}}
\providecommand{\bK}{\ensuremath{\mathbf{K}}}
\providecommand{\bmu}{\ensuremath{\mathbf{\mu}}}
\providecommand{\bSigma}{\ensuremath{\mathbf{\Sigma}}}
\providecommand{\bbeta}{\ensuremath{\boldsymbol{\beta}}}
\providecommand{\beps}{\ensuremath{\boldsymbol{\varepsilon}}}

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}
%\setbeamersize{text margin left=0.3mm,text margin right=0.1mm}

\author{Caleb Leedy}
\title[Data Integration]{Previous Work on Non-montone Missingness}

\begin{document}

\everymath{\displaystyle}
\setbeamertemplate{title page}[default][colsep=-4bp,rounded=true]
\setbeamertemplate{itemize items}[circle]
\setbeamercolor{block title}{bg=red!50,fg=black}
\frame{\titlepage}

\begin{frame}{Outline}

  \begin{itemize}
    \item Notation
    \item Simulation Studies
  \end{itemize}
\end{frame}

\begin{frame}{Notation}

  \begin{tabular}{cc}
    Case 1: Monotone missingness &
  Case 2: Non-monotone missingness \\
  \begin{tabular}{lrrrrr}
    \toprule
    & $X$ & $Y_1$ & $Y_2$ & $R_1$ & $R_2$ \\
    \midrule
    $A_{11}$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & 1 & 1 \\
    $A_{10}$ & $\checkmark$ & $\checkmark$ &          & 1 & 0 \\
    $A_{00}$ & $\checkmark$ &          &          & 0 & 0 \\
    \bottomrule
  \end{tabular} &
  \begin{tabular}{lrrrrr}
    \toprule
    & $X$ & $Y_1$ & $Y_2$ & $R_1$ & $R_2$ \\
    \midrule
    $A_{11}$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & 1 & 1 \\
    $A_{10}$ & $\checkmark$ & $\checkmark$ &          & 1 & 0 \\
    $A_{01}$ & $\checkmark$ &          & $\checkmark$ & 0 & 1 \\
    $A_{00}$ & $\checkmark$ &          &          & 0 & 0 \\
    \bottomrule
  \end{tabular}
  \end{tabular}

  The goal is to estimate $\theta = E[g(X, Y_1, Y_2)]$.

\end{frame}

\begin{frame}{Proposed Estimator: Monotone}
  For, $b(X_i, Y_{1i}) = E[g_i \mid X_i, Y_{1i}]$ and $i = \{1, \dots, n\}$,

  \begin{align*}
\hat \theta_{\text{eff}} &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \\ 
&+ n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)}(b(X_i, Y_{1i}) - E[g_i \mid X_i]) \\ 
&+ n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_i)}(g_i - b_i)
  \end{align*}

\end{frame}

\begin{frame}{Proposed Estimator: Non-monotone}

  Define,
\begin{align*}
b(X_i, Y_{1i}) &= E[g_i \mid X_i, Y_{1i}], \\ 
a(X_i, Y_{2i}) &= E[g_i \mid X_i, Y_{2i}], \text{ and } \\
i &= \{1, \dots, n\},
\end{align*}

\end{frame}

\begin{frame}{Proposed Estimator: Non-monotone}

\begin{align*}
\hat \theta_{\text{eff}} &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \\ 
&+ n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)}(b(X_i, Y_{1i}) - E[g_i \mid X_i]) \\ 
&+ n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)}(a(X_i, Y_{2i}) - E[g_i \mid X_i]) \\ 
&+ n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_i)}(g_i - b_i - a_i + E[g_i \mid X_i])
\end{align*}

\end{frame}

\begin{frame}{Simulation Outline: Monotone MAR}

  For this simulation we use the following approach:
  \begin{enumerate}
      \item Generate $X$, $Y_1$, and $Y_2$ for elements $i = 1, \dots, n$.
      \item Using the covariate $X$, determine the probability $p_1$ of $Y_1$
      being observed for each element $i$.
      \item Based on $p_1$, determine if $R_1 = 1$.
      \item If $R_1 = 0$, then $R_2 = 0$. Otherwise, using variables $X$ and
        $Y_1$, determine the probability $p_{12}$.
      \item Based on $p_{12}$ determine if $R_2 = 1$.
  \end{enumerate}
\end{frame}

\begin{frame}{Simulation Outline: Non-monotone MAR}

    Following the approach of \cite{robins1997non},
    the algorithm to generate the data is the following:
    \begin{enumerate}
        \item Generate $X$, $Y_1$, and $Y_2$ for elements $i = 1, \dots, n$.
        \item Using the covariate $X_i$, generate probabilities for each element
          $i$ $p_0$, $p_1$, and $p_2$ such that $p_0 + p_1 + p_2 = 1$. 
        \item Select one option based on the three probabilities for each
          element $i$. If 0 is selected: $R_1 = 0$ and $R_2 = 0$; if 1 is
          selected $R_1 = 1$; if 2 is selected, $R_2 = 1$.
        \item We take the next step in multiple cases. If 0 was selected, we are
          done. If 1 was selected, we generate probabilities $p_{12}$ based on
          $X$ and $Y_1$. Then based on this probability, we determine if $R_2 =
          1$. In the same manner, if 2 was selected in the previous step, we
          generate probabilities $p_{21}$ based on $X$ and $Y_2$. Then based on
          this probability, we determine if $R_2 = 1$.
    \end{enumerate}
\end{frame}

\begin{frame}{Simulation 1: Monotone MAR}

    We generate data from the following distributions:
    \begin{align*}
        X_i &\stackrel{iid}{\sim} N(0, 1) \\
        Y_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
        Y_{2i} &\stackrel{iid}{\sim} N(\theta, 1)
    \end{align*}

    Then, we create the probabilities $p_1 = \logistic(x_i)$ and 
    $p_{12} = \logistic(y_{1i})$.
    Since, both $x_i$ and $y_1$ are standard normal distributions, each of these
    probabilities is approximately $0.5$ in expectation.

    The goal of this simulation is to estimate $\theta$. Alternatively, we can
    express this as solving the estimating equation:

    \[g(\theta) \equiv Y_2 - \theta = 0.\]

\end{frame}

\begin{frame}{Simulation 1: Monotone MAR}

    We estimate $\theta$ using the following procedures:

    \begin{itemize}
        \item Oracle: This computes $\bar Y_2$ using \textit{both} the observed
          and missing data.
        \item IPW-Oracle: This is an IPW estimator using only the observed
          values of $Y_2$. The weights (inverse probabilities) use the actual
          probabilities.
        \item IPW-Est: This is an IPW estimator using the probabilities that
          have been estimated by a logistic model.
        \item Semi: The monotone efficient estimator.
    \end{itemize}

    \begin{itemize}
      \item Sample size ($n$): 2000
      \item Monte Carlo replications: 2000
    \end{itemize}

\end{frame}

\begin{frame}

    \input{Tables/monomar_t-5.tex}

\end{frame}

\begin{frame}

    \input{Tables/monomar_t0.tex}

\end{frame}

\begin{frame}

    \input{Tables/monomar_t5.tex}

\end{frame}

\begin{frame}{Simulation 1: Non-monotone MAR}

    We generate variables $(X, Y_1, Y_2)$ using the following setup:

    \[\begin{bmatrix}
    X_i \\ \varepsilon_{1i} \\ \varepsilon_{2i}
    \end{bmatrix} \stackrel{iid}{\sim}
    N\left(
    \begin{bmatrix}
        0 \\ 0 \\ \theta
    \end{bmatrix},
    \begin{bmatrix}
        1 & 0 & 0 \\
        0 & 1 & \sigma_{yy}\\
        0 & \sigma_{yy} & 1
    \end{bmatrix}
    \right).\]

    Then,

    \[y_{1i} = x_i + \varepsilon_{1i} \text{ and } 
    y_{2i} = x_i + \varepsilon_{2i}.\]

\end{frame}

\begin{frame}

    Since we have nonmonotone data, our ``Stage 1'' probabilities are
    different. We compute the true Stage 1 probabilities being proportional to
    the following values:

    \begin{align*}
        p_0 &= 0.2 \\
        p_1 &= 0.4 \\
        p_2 &= 0.4 \\
    \end{align*}

   However, we keep the same structure for the Stage 2 probabilities with:
   $p_{12} = \logistic(y_1)$ and $p_{21} = \logistic(y_2)$.

\end{frame}

\begin{frame}

  \input{Tables/nonmonosim1m-5.tex}

\end{frame}

\begin{frame}

  \input{Tables/nonmonosim1m0.tex}

\end{frame}

\begin{frame}

  \input{Tables/nonmonosim1m5.tex}

\end{frame}

\begin{frame}{Simulation 2: Non-monotone MAR}

    For this simulation, we
    focus on $\Cov(Y_1, Y_2)$. The data generating process now has $\sigma_{yy}
    \neq 0$. We are still interested in $\bar Y_2$ and we still run 2000
    simulation with 2000 observations. In all the next simulations the true
    value of $\theta = 0$. The results are the following:

\end{frame}

\begin{frame}

    \input{Tables/nonmonosim2c0.1.tex}

\end{frame}

\begin{frame}

    \input{Tables/nonmonosim2c0.5.tex}

\end{frame}

\begin{frame}

    \input{Tables/nonmonosim2c0.9.tex}

\end{frame}

\begin{frame}{Comparing with a Calibration Estimator}

  The efficient monotone estimator should be very similar to the
    following calibration estimator, for $\sum_{i = 1}^n w_i y_{2i}$,

    \begin{align*}
      \argmin_w \sum_{i = 1}^n w_i^2& \text{ such that }\\
      \sum_{i = 1}^n x_i &= \sum_{i = 1}^n R_{1i} w_{1i} x_i \\
      \sum_{i = 1}^n w_{1i} (x_i, y_{1i}) &= \sum_{i = 1}^n R_{1i} R_{2i} w_{2i}
      (x_i, y_{1i}) \\
    \end{align*}

  The reason that these should be the same is because they are similar in
    relationship to a calibration and regression estimator which are equivalent.

\end{frame}

\begin{frame}{Calibration Comparison: Monotone}
  To test the idea that the monotone regression estimator is similar to
    the calibration estimator we run several simulation studies. In the monotone
    case data is generating in the following steps:

    \begin{enumerate}
      \item The variables $X$, $Y_1$, and $Y_2$ are simulated from the following
        distributions:
        \begin{align*}
          X_i &\stackrel{iid}{\sim} N(0, 1) \\
          Y_{1i} &\stackrel{iid}{\sim} N(0, 1) \\
          Y_{2i} &\stackrel{iid}{\sim} N(\theta, 1).
        \end{align*}

      \item After the variables have been simulated, we see which variables are
        observed. We always observe $X_i$. We observed $Y_1$ with
        probability $p_{1i} \propto \logistic(x_i)$. If $Y_{1i}$ is observed,
        then we observe $Y_{2i}$ with probability $p_{2i} \propto
        \logistic(y_{1i})$. If $Y_{1i}$ is not observed, we do not observe
        $Y_{2i}$.
    \end{enumerate}

\end{frame}

\begin{frame}{Additional Estimators}
  \begin{itemize}
\item HT estimator of $\theta=E(Y_2)$: 
$$ \hat{\theta}_{\rm HT} = \frac{1}{n} \sum_{i=1}^n \frac{R_{1i} R_{2i}}{ \pi_{11}(X_i) } y_{2i}  $$
\item The three-phase regression estimator of $\theta$:
\begin{eqnarray*} 
\hat{\theta}_{\rm reg} &=& \frac{1}{n} \sum_{i \in A_2}\frac{1}{ \pi_{2i}  }\left\{ y_{i}  - \hat{E} ( Y \mid x_i, z_{i} ) \right\} \\
&+& \frac{1}{n} \sum_{i \in A_1}\frac{1}{ \pi_{1i}  }\left\{ \hat{E} ( Y \mid x_i, z_{i} ) - \hat{E} ( Y \mid x_i ) \right\} + \frac{1}{n} \sum_{i \in U} \hat{E} ( Y \mid x_i ) \\
&=& \bar{x}_0' \hat{\beta} + \left( \bar{x}_1' \hat{\gamma}_x + \bar{z}_1' \hat{\gamma}_z - \bar{x}_1' \hat{\beta} \right) + \{ \bar{y}_2- (\bar{x}_2' \hat{\gamma}_x + \bar{z}_2' \hat{\gamma}_z ) \} \\
&=& \bar{y}_2 + \{ \bar{x}_1' \hat{\gamma}_x + \bar{z}_1' \hat{\gamma}_z  -  (\bar{x}_2' \hat{\gamma}_x + \bar{z}_2' \hat{\gamma}_z ) \} +\left(  \bar{x}_0' \hat{\beta} - \bar{x}_1' \hat{\beta} \right) \end{eqnarray*} 
\item We can view the above three-phase regression estimator as a projection estimator of \cite{kim2012combining}.
\end{itemize}
\end{frame}

\begin{frame}

    \input{Tables/calimono_t-5.tex}

\end{frame}

\begin{frame}

    \input{Tables/calimono_t0.tex}

\end{frame}

\begin{frame}

    \input{Tables/calimono_t5.tex}

\end{frame}

\begin{frame}{Non-monotone Calibration}

    For the non-monotone case,
    we believe that we have the following calibration equations:

    \begin{align*}
      \sum_{i = 1}^n E[g_i \mid X_i] &= \sum_{i = 1}^n R_{1i} w_{1i} E[g_i \mid
      X_i]\\
      \sum_{i = 1}^n E[g_i \mid X_i] &= \sum_{i = 1}^n R_{2i} w_{2i} E[g_i \mid
      X_i]\\
      \sum_{i = 1}^n R_{1i} w_{1i} E[g_i \mid X_i, Y_{1i}] &= \sum_{i = 1}^n
      R_{1i} R_{2i} w_{ci} E[g_i \mid X_i, Y_{1i}]\\
      \sum_{i = 1}^n R_{2i} w_{2i} E[g_i \mid X_i, Y_{1i}] &= \sum_{i = 1}^n
      R_{1i} R_{2i} w_{ci} E[g_i \mid X_i, Y_{2i}]\\
      \sum_{i = 1}^n E[g_i \mid X_i] &= \sum_{i = 1}^n R_{1i} R_{2i} w_{ci}
      E[g_i \mid X_i].
    \end{align*}

\end{frame}

\begin{frame}{Non-monotone Calibration}
  We still have the same goal of the simulation study: estimate $\theta =
    E[Y_2]$.

    \begin{enumerate}
      \item[1.] Generate $X_i$, $\varepsilon_{1i}$, and $\varepsilon_{2i}$ from the
        following distributions:

        \begin{align*}
          x_i &\stackrel{iid}{\sim} N(0, 1)\\
          \varepsilon_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
          \varepsilon_{2i} &\stackrel{iid}{\sim} N(\theta, 1)\\
        \end{align*}

        Then we have
        \[y_{1i} = x_i + \varepsilon_{1i} \text{ and } y_{2i} = x_i +
        \varepsilon_{2i}.\]

    \end{enumerate}

\end{frame}

\begin{frame}
  \begin{enumerate}
    \item[2.] Then we have to select the variables to observe. We always observe
        $X_i$. Then we choose to either observe $Y_1$ with probability $0.4$,
        $Y_2$ with probability $0.4$ or neither with probability $0.2$.

      \item[3.] If neither then $R_{1i} = 0$ and $R_{2i} = 0$. If we observe $Y_1$
        then $R_1 = 1$ and if we observe $Y_2$ then $R_2 = 1$.

      \item[4.] If we observe either $Y_1$ or $Y_2$ then with probability $p \propto
        \logistic(Y_k)$ where $Y_k$ is the observed $Y$ variable we choose to
        observe the other $Y$ variable.

      \item[5.] If the other $Y$ variable is observed then the corresponding $R_k =
        1$. Otherwise, $R_k = 0$.
      \end{enumerate}
    \end{frame}

\begin{frame}

    \input{Tables/calinonmono1m-5.tex}

\end{frame}

\begin{frame}

    \input{Tables/calinonmono1m0.tex}

\end{frame}

\begin{frame}

    \input{Tables/calinonmono1m5.tex}

\end{frame}

\begin{frame}[allowframebreaks]{References}
  \printbibliography
\end{frame}

\end{document}

### 2023-08-25 R/opt_est.R

@@ -1,15 +1,418 @@
 # Title: opt_est.R
 # Created by: Caleb Leedy
 # Created on: August 19, 2023
-# Purpose: This file contains the functions needed for computation of the 
+# Updated on: August 22, 2023
+# Purpose: This file contains the functions needed for computation of the
 # optimal estimator in the non-monotone case.
 
-# ************
-# * Overview *
-# ************
-# The goal of this file is to create code to estimate the following:
-# \hat \theta = n^{-1} \sum_{i = 1}^n E[g_i | X_i] +
-#   n^{-1} \sum_{i = 1}^n (R_{1i} / \pi_{1+}) (E[g_i | X_i, Y_{1i}] - E[g_i | X_i]) +
-#   n^{-1} \sum_{i = 1}^n (R_{2i} / \pi_{2+}) (E[g_i | X_i, Y_{2i}] - E[g_i | X_i]) +
-#   n^{-1} \sum_{i = 1}^n (R_{1i}R_{2i} / \pi_{11}) (E[g_i | X_i, Y_{1i}, Y_{2i}] -
-#     E[g_i | X_i, Y_{1i}] - E[g_i | X_i, Y_{2i}] + E[g_i | X_i])
+# ***********
+# * Outline *
+# ***********
+# 1. Generate data.
+# 2. EM Algorithm for Optimal Estimate
+# 3. Proposed Estimator
+
+# *************
+# * Libraries *
+# *************
+
+library(dplyr)
+library(stringr)
+
+# ************************
+# * Additional Functions *
+# ************************
+
+expit <- function(x) {1 / (1 + exp(-x))}
+
+# *****************
+# * Generate Data *
+# *****************
+
+#' This function generates a multivariate normal distribution with missingness.
+#'
+#' @param n - An integer for the number of observations
+#'
+#' @return df - A data frame.
+gen_optsim_data <- function(n, theta = 0, cor_xe1 = 0, cor_xe2 = 0, cor_e1e2 = 0) {
+
+  # 1. Generate X, Y1, Y2
+  mean_vec <- c(0, 0, 0)
+  sigma_mat <-
+    matrix(c(1, cor_xe1, cor_xe2, cor_xe1, 1, cor_e1e2, cor_xe2, cor_e1e2, 1),
+           nrow = 3)
+
[…]

### 2023-08-25 Simulations/generate_data.R

@@ -4,140 +4,6 @@
 # Purpose: This file contains the code to generate data for different simulation
 # studies. Instead of writing the functions in Simulations/nonmonotone.R and
 # then copying them to other files, I thought that it would be a good idea to
-# put them all into one script and then source it.
-
-# ------------------------------------------------------------------------------
-
-# ********************
-# * Script Functions *
-# ********************
-# 1. mono_mar -- This generates monotone missingness.
-# 2. nonmono_mar -- This generates nonmonotone missing data.
-# 3. expit -- A helper function for generating probabilites. 
-
-# ------------------------------------------------------------------------------
-
-# **********************
-# * Required Libraries *
-# **********************
-
-library(MASS)
-library(dplyr)
-
-# ------------------------------------------------------------------------------
-
-# ************************
-# * Additional Functions *
-# ************************
-
-expit <- function(x) {1 / (1 + exp(-x))}
-
-# *****************
-# * Generate Data *
-# *****************
-
-mono_mar <- function(n, mean_y2 = 0) {
-
-  # Steps:
-  # 1. Generate X, Y_1, Y_2.
-  # 2. Get probabilities conditional on past.
-  # 3. Reveal hidden variables.
-
-  # 1. Generate X, Y_1, Y_2.
-  x_vec <- rnorm(n)
-  y1 <- rnorm(n)
-  y2 <- rnorm(n, mean = mean_y2)
-
-  # 2. Get probabilities conditional on past.
-  r1 <- rbinom(n, 1, expit(x_vec))
-  r2 <- rbinom(n, 1, expit(y1))
-
-  # 3. Reveal hidden variables.
-  tibble(x = x_vec, y1 = y1, y2 = y2, r1 = r1, r2 = r2) %>%
-    mutate(p1 = expit(x_vec)) %>%
-    mutate(p12 = expit(y1)) %>%
-    mutate(p21 = 0) %>%
-    mutate(p2 = 0) %>%
-    mutate(pi_11 = p1 * p12) %>%
[…]

### 2023-08-25 Simulations/nmono_est_funs.R

@@ -20,7 +20,7 @@
 #   2. ipw_cc_oracle -- IPW estimator with oracle weights
 #   3. ipw_cc_est -- IPW estimator with weights estimated
 #   4. twophase_reg -- Unfinished
-#   5. nonmono_est_weights -- Calibration estimator 
+#   5. nonmono_est_weights -- Calibration estimator
 
 # ------------------------------------------------------------------------------
 

### 2023-08-25 Simulations/optimal_simulation.R

@@ -7,8 +7,94 @@
 # ***********
 # * Outline *
 # ***********
-# 1. Generate data.
+# 1. Load Libraries and code.
+# 2. Run Monte Carlo simulation.
 
 # *************
 # * Libraries *
 # *************
+
+library(dplyr)
+library(stringr)
+
+library(doParallel)
+library(doRNG)
+library(parallelly)
+
+source("R/opt_est.R")
+
+# ***************************
+# * Monte Carlo Simulations *
+# ***************************
+
+clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
+registerDoParallel(clust)
+
+B <- 1000
+n_obs <- 1000
+true_theta <- 0
+
+mc_theta <-
+  foreach(iter = 1:B, .options.RNG = 1, .packages = c("dplyr", "stringr")) %dorng% {
+
+    # Generate Data
+    df <- gen_optsim_data(n = n_obs, theta = true_theta)
+
+    # Get Estimates
+    oracle_est <- mean(df$Y2)
+    ipw_est <- 
+      filter(df, delta_1 == 1, delta_2 == 1) |>
+      mutate(w_est = Y2 / prob_11) |>
+      pull(w_est) |>
+      sum() / nrow(df)
+    cc_est <- 
+      filter(df, delta_2 == 1) |>
+      pull(Y2) |>
+      mean()
+    # em_est <- em_optsim(df)
+    wls_est <- opt_lin_est(df)
+    prop_est <- prop_nmono_est(df)
+    prop2_est <- prop2_nmono_est(df)
+
+    return(tibble(oracle = oracle_est,
+                  cc = cc_est,
+                  ipw = ipw_est,
+                  # em = em_est,
+                  wls = wls_est,
+                  prop = prop_est,
+                  prop2 = prop2_est))
+  } |> bind_rows()
+
+stopCluster(clust)
+
+# *******************
+# * Analyze Results *
+# *******************
+
+mc_theta |>
+  summarize(
+    bias_oracle = mean(oracle) - true_theta,
+    bias_cc = mean(cc) - true_theta,
+    bias_ipw = mean(ipw) - true_theta,
+    bias_wls = mean(wls) - true_theta,
+    bias_prop = mean(prop) - true_theta,
+    bias_prop2 = mean(prop2) - true_theta,
+    sd_oracle = sd(oracle),
[…]

### 2023-08-25 tests/testthat/test-opt_est.R

@@ -1,0 +1,97 @@
+# Title: test-opt_est.R
+# Created by: Caleb Leedy
+# Created on: August 21, 2023
+# Purpose: This file contains the code to check to see if all of the functions
+# in R/optimal_simulation.R work.
+
+# *********
+# * Tests *
+# *********
+
+# Test expit
+test_that("expit works", {
+  expect_equal(expit(0), 0.5)
+  expect_equal(expit(Inf), 1)
+  expect_equal(expit(-Inf), 0)
+})
+
+# Test expect_g
+df <- 
+  tibble(X = rep(c(1, 2), 5)) |>
+  mutate(Y1 = 1 + 2 * X) |>
+  mutate(Y2 = 2 + 3 * X) |>
+  mutate(delta_1 = c(1, 1, 1, 0, 0, 0, 1, 1, 1, 0)) |>
+  mutate(delta_2 = c(1, 1, 1, 0, 1, 1, 0, 0, 1, 0)) |>
+  mutate(prob_11 = 0.4, prob_1 = 0.6, prob_2 = 0.6)
+
+df_1 <- filter(df, delta_1 == 1)
+df_2 <- filter(df, delta_2 == 1)
+tmp1 <- rep(0, nrow(df))
+tmp1[df$delta_1 == 1] <- df$Y2[df$delta_1 == 1]
+tmp2 <- rep(0, nrow(df))
+tmp2[df$delta_2 == 1] <- df$Y2[df$delta_2 == 1]
+
+test_that("expect_g works", {
+  expect_equal(expect_g(df, "Y2", c("X"), df, rep(1, nrow(df))), df$Y2)
+  expect_equal(expect_g(df, "Y2", c("X", "Y1"), df_1, df$delta_1), tmp1)
+  expect_equal(expect_g(df, "Y2", c("X", "Y2"), df_2, df$delta_2), tmp2)
+})
+
+# Test get_cond_exp
+test_that("get_cond_exp works with cols = 1", {
+  df <- gen_optsim_data(n = 1000, theta = 0)
+  obs_y2 <- matrix(df$Y2, ncol = 1)
+  mean_y1 <- filter(df, delta_1 == 1) |> pull(Y1) |> mean()
+  mean_y2 <- filter(df, delta_2 == 1) |> pull(Y2) |> mean()
+  sigma_mat <- list(list(2, 1), list(1, 2))
+  cond_exp <- get_cond_exp(obs_y2, mean_y1, mean_y2, sigma_mat)
[…]

### 2023-09-07 Notes/model_assumptions.tex
Reorganized Simulations/ directory. Updated README.

@@
     \stackrel{\text{ind}}{\sim} F$ for some unknown $F$ for all $i \in U$.
+  \item We assume that $x_i$ is observed throughout the finite population.
   \item Let $I_i \stackrel{\text{ind}}{\sim} \pi_i$ for $i \in U$ be the 
-    sample inclusion indicator.
-  \item We assume that $x_i$ is observed throughout the finite population.
-  \item However, $(y_{1i}, y_{2i}, \pi_i)$ is only observed if $I_i = 1$.
+    first phase sample inclusion indicator.
+  \item If $i \in \{1, \dots, N\}$ is in the first phase sample ($I_i = 1$)
+    then $\pi_i$ is observed. 
+  \item The second phase sampling indicators are $\delta_{00}$, $\delta_{01}$,
+    $\delta_{10}$, and $\delta_{11}$. These mutual exclusive variables encode 
+    the following: if both $Y_1$ and $Y_2$ are observed then $\delta_{11} = 1$;
+    if $Y_1$ is observed and $Y_2$ is missing then $\delta_{10} = 1$; if $Y_1$
+    is missing and $Y_2$ is observed then $\delta_{01} = 1$; and if both $Y_1$
+    and $Y_2$ are missing then $\delta_{00} = 1$. (For notational purposes, it
+    is sometimes convienent to write $\delta_{1} = \delta_{11} + \delta_{10}$ 
+    and $\delta_2 = \delta_{11} + \delta_{01}$. A similar notation will be 
+    used for $\pi$. We have $\pi_{1+} = \pi_{11} + \pi_{10}$ and 
+    $\pi_{2+} = \pi_{11} + \pi_{01}$.)
+  \item All of the sampling indicators were drawn via \textit{Poisson}
+    sampling. So the sample size of each category $A_{ij}$ is random. However,
+    this does ensure independence between observations.
   \item The goal is to estimate $\theta = E[g(X, Y_1, Y_2)]$ in the population.
@@
     &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \nonumber \\ 
-    &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)}
+    &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{1i}}{\pi_{1+}(X_i)}
       (E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i])\nonumber  \\ 
-    &+ n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)}
+    &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{2i}}{\pi_{2+}(X_i)}
       (E[g_i \mid X_i, Y_{2i}] - E[g_i \mid X_i])\nonumber  \\ 
-    &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_i)}
+    &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{1i} \delta_{2i}}{\pi_{11}(X_i)}
       (g_i - E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i, Y_{2i}] + E[g_i \mid X_i])
@@
 
-  \begin{align*}
+    \begin{align}\label{eq:propclass}
     \hat \theta_{\text{eff}} 
-    &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \\ 
-    &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)}
-      (b(X_i, Y_{1i}) - E[g_i \mid X_i]) \\ 
-    &+ n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)}
-      (a(X_i, Y_{2i}) - E[g_i \mid X_i]) \\ 
-    &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_i)}
+    &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i]\nonumber \\ 
+    &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{1i}}{\pi_{1+}(X_i)}
+      (b(X_i, Y_{1i}) - E[g_i \mid X_i])\nonumber \\ 
+    &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{2i}}{\pi_{2+}(X_i)}
+      (a(X_i, Y_{2i}) - E[g_i \mid X_i])\nonumber \\ 
+    &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{1i} \delta_{2i}}{\pi_{11}(X_i)}
       (g_i - b_i - a_i + E[g_i \mid X_i]).
-  \end{align*}
+    \end{align}
 
@@
 
-I have done a lot to explore this area but I have several points of confusion.
-The first part recaps what Dr. Fuller and I discussed this past week.
+The first part recaps what Dr. Fuller and I discussed previously.
 
@@
     the relationship between $X$ and $Y_k$, we can use difference estimators:
-    $W_1 \equiv Y_1 - b_1 \tilde X$ and $W_2 = \equiv Y_2 - b_2 \tilde X$, where
+    $Z_1 \equiv Y_1 - b_1 \tilde X$ and $Z_2 = \equiv Y_2 - b_2 \tilde X$, where
     $\tilde X = X - E[X]$.
-    To minimize the variance of $W_k$ we can choose the optimal value of $b_k$,
+    To minimize the variance of $Z_k$ we can choose the optimal value of $b_k$,
      which is
@@
 
-    Since these values are known, $b_k$ is known. This means that we now have
-    the following table:
+    Since these covariance values are known, $b_k$ is known. This means that we
+    now have the following table:
 
@@
         \toprule
-         & $W_1$ & $W_2$ \\
+         & $Z_1$ & $Z_2$ \\
          \midrule
@@
         1 & 0 & 0 \\
-        0 & \sigma_{11} & 0 \\ 
-        0 & 0 & \sigma_{22}
+        0 & \sigma_{11} & \sigma_{12} \\ 
+        0 & \sigma_{12} & \sigma_{22}
       \end{bmatrix},
@@
 
-    and define $y_{1i} = x_i + e_{1i}$ and $y_{2i} = x_i + e_{2i}$. Then $b_1 =
-    b_2 = 1$. We define $\bar w_k^{(ij)}$ as the mean of $y_k$ in segment
-    $A_{ij}$. This means that we have means $\bar w_1^{(11)}$, $\bar
-    w_2^{(11)}$, $\bar w_1^{(10)}$, and $\bar w_2^{(01)}$. Let 
-    $W = [\bar w_1^{(11)}, \bar w_2^{(11)}, \bar w_1^{(10)}, \bar w_2^{(01)}]'$,
-    then for $n_{ij} = |A_{ij}|$, we have
+    and define $y_{1i} = \theta_1 + x_i + e_{1i}$ and $y_{2i} = \theta_2 + x_i
+    + e_{2i}$. Then $b_1 = b_2 = 1$. We define $\bar z_k^{(ij)}$ as the mean of
+    $y_k$ in segment $A_{ij}$. This means that we have means $\bar z_1^{(11)}$,
+    $\bar z_2^{(11)}$, $\bar z_1^{(10)}$, and $\bar z_2^{(01)}$. Let $W = [\bar
+    z_1^{(11)}, \bar z_2^{(11)}, \bar z_1^{(10)}, \bar z_2^{(01)}]'$, then for
+    $n_{ij} = |A_{ij}|$, we have
 
-    \[W - M \mu \sim N(\vec 0, V)\]
+    \[Z - M \mu \sim N(\vec 0, V)\]
 
@@
       \begin{bmatrix}
-        \frac{\sigma_{11}}{n_{11}} & 0 & 0 & 0 \\
-        0 & \frac{\sigma_{22}}{n_{11}} & 0 & 0 \\
+        \frac{\sigma_{11}}{n_{11}} & \frac{\sigma_{12}}{n_{11}} & 0 & 0 \\
+        \frac{\sigma_{12}}{n_{11}} & \frac{\sigma_{22}}{n_{11}} & 0 & 0 \\
         0 & 0 & \frac{\sigma_{11}}{n_{10}} & 0 \\
@@
 
-    \[ \hat \mu = (M' V^{-1} M)^{-1} M' V^{-1} W. \]
-
-  \item If we do the matrix algebra, $V^{-1} = 
-    \text{diag}\left(\frac{n_{11}}{\sigma_{11}}, \frac{n_{11}}{\sigma_{22}},
-    \frac{n_{10}}{\sigma_{11}}, \frac{n_{01}}{\sigma_{22}}\right)$ and hence,
-
-    {\renewcommand{\arraystretch}{2}
     \begin{equation}\label{eq:wls}
-    \hat \mu = 
-      \begin{bmatrix}
-        \frac{n_{11}\bar w_1^{(11)} + n_{10}\bar w_1^{(10)}}{n_{11} + n_{10}} \\
-        \frac{n_{11}\bar w_2^{(11)} + n_{01}\bar w_2^{(01)}}{n_{11} + n_{01}} \\
-      \end{bmatrix}
+    \hat \mu = (M' V^{-1} M)^{-1} M' V^{-1} W.
     \end{equation}
-    }
 
@@
         1 & 0 & 0 \\
-        0 & 1 & 0 \\ 
-        0 & 0 & 1 
+        0 & 1 & \sigma_{12} \\ 
+        0 & \sigma_{12} & 1 
       \end{bmatrix},
       \right) \text{ and }
-      y_{1i} = x_i + e_{1i}, y_{2i} = x_i + e_{2i}.
+      y_{1i} = \theta_1 + x_i + e_{1i}, y_{2i} = \theta_2 + x_i + e_{2i}.
     \]
@@
 
+      \item OracleX: This computes the average value of $Y_2 - X$ in 
+        all observations (even when $Y_2$ is not supposed to be observed.)
+        \[\hat \theta = n^{-1} \sum_{i = 1}^n (y_{2i} - x_i).\]
+
       \item CC: This computes the average value of $Y_2$ in all observations 
@@
 
-      \item IPW: This computes the survey weighted average of $Y_2$ when both
-        $Y_1$ and $Y_2$ are observed.
-        \[\hat \theta = \frac{\sum_{i = 1}^n \delta_{1i}\delta_{2i} y_{2i}}
-        {\sum_{i = 1}^n \delta_{1i}\delta_{2i}}.\]
       \item Proposed: This is the proposed estimator from Equation~\ref{eq:prop}.
       \item WLS: This is the weighted linear estimator from Equation~\ref{eq:wls}.
-        (Note, since $g = E[Y_2]$ this only contains the second element from
-        Equation~\ref{eq:wls}.)
+        Note, since $g = E[Y_2]$ this only contains the second element from
+        Equation~\ref{eq:wls}.
     \end{itemize}
@@
     number of observations $n = 1000$ and the Monte Carlo sample size of 
-    $B = 1000$.
+    $B = 3000$.
 
     \begin{table}[ht!]
-      \centering
+      \label{tab:res}
       \caption{This table shows the estimators of $\theta = E[Y_2]$. The true 
-      value of $\theta$ is $0$. The bias column shows the average bias of the 
-      estimator and the actual value of $\theta = 0$ across the $B = 1000$ 
+        value of $\theta$ is $5$ and the true value of $\sigma_{12} = 0.5$.
+        The bias column shows the average bias of the 
+      estimator and the actual value of $\theta = 5$ across the $B = 3000$ 
       simulations. The SD column shows the average standard deviation across the 
-      $B = 1000$ simulations for each algorithm. The Tstat column displays the 
+      $B = 3000$ simulations for each algorithm. The Tstat column displays the 
       t-statistic of a t-test comparing the estimator to the actual value. 
@@
       p-value of the t-statistic.}
+      \centering
       \begin{tabular}[t]{lrrrr}
@@
         \midrule
-        Oracle & -0.001 & 0.045 & -0.953 & 0.170\\
-        CC & 0.000 & 0.056 & -0.059 & 0.476\\
-        IPW & 0.000 & 0.071 & -0.183 & 0.428\\
-        Prop & 0.000 & 0.051 & -0.131 & 0.448\\
-        WLS & -0.001 & 0.042 & -0.470 & 0.319\\
+        Oracle & 0.001 & 0.044 & 1.546 & 0.061\\
+        OracleX & 0.000 & 0.032 & 0.479 & 0.316\\
+        CC & 0.002 & 0.058 & 1.549 & 0.061\\
+        WLS & 0.001 & 0.040 & 1.207 & 0.114\\
+        Prop & 0.002 & 0.051 & 1.918 & 0.028\\
         \bottomrule
       \end{tabular}
-      \label{tab:res}
     \end{table}
 
-  \item There are a couple things of note from Table~\ref{tab:res}. First, the 
-    WLS estimator has a smaller standard error than the Oracle estimator. This 
-    seems incorrect. (Please see the next section for questions.) Second, the 
-    proposed estimator did not have a similar standard error as the WLS estimator.
-    This suggests that it is not optimal.
 \end{itemize}
 
-\section*{Questions for Dr. Fuller}
+\section*{Computing the Variance by Hand}
 
-\begin{enumerate}
-  \item For the linear estimator I don't think that I am incorporating sampling
-    weights. I am a bit confused about how to do this. Should I compute the 
-    means $\bar w_{k}^{(ij)}$ using weights according to the sample weights? 
-  \item Is it ok for the linear estimator to outperform the oracle estimator?
-    This seems odd to me because the oracle estimator seems to contain more 
-    information.
-  \item On Tuesday, you asked me to compute the linear estimator in terms of 
-    means of $Y$ and estimators of $0$. I think that the WLS estimator is the 
-    estimator in terms of estimators of $0$ because $y_k - b_k x$ is an 
-    estimator of $0$. To transform the estimator into an estimator in terms of
-    $Y$, should I write the form in terms of $Y$ instead of $W$?
-\end{enumerate}
+Table~\ref{tab:res} demonstrates that the WLS estimator outperforms the 
+Oracle estiamator. I was initially confused about why this is the case, but
+realized that I could approximate the standard error of both estimators
+reasonably well. It turns out that I was able to get reasonable estimates 
+for the variance of all of the proposed estimators includeing $\hat
+\theta_{prop}$. In this part of the report, I detail my computations and show
+how the simulated results are accurate.
+
+\begin{align}\label{eq:varor}
+  \Var(\hat \theta_{Oracle}) 
+  &= \Var\left(n^{-1} \sum_{i = 1}^n y_{2i}\right) \nonumber \\
+  &= n^{-1} \Var(y_{2}) \nonumber \\
+  &= n^{-1} (1 + \sigma_{22}) & \text{because $\Var(x) = 1$.}
+\end{align}
+
+\begin{align}\label{eq:varorx}
+  \Var(\hat \theta_{OracleX}) 
+  &= \Var\left(n^{-1} \sum_{i = 1}^n y_{2i} - x_i\right) \nonumber \\
+  &= n^{-1} \Var(e_{2}) \nonumber \\
+  &= n^{-1} \sigma_{22}.
+\end{align}
+
+Using a Taylor expansion,
+
+\begin{align}\label{eq:varcc}
+  \Var(\hat \theta_{cc})
+  &= \Var\left(\theta_2 + \pi_{2+}^{-1} \left(n^{-1}\sum_{i = 1}^n \delta_{2i} y_{2i}
+  - \pi_{2+}\theta_2\right) - \frac{\theta_2 \pi_{2+}}{\pi_{2+}^2} 
+  \left(n^{-1} \sum_{i = 1}^n \delta_{2i} - \pi_{2+}\right) + o_p(1)\right)\nonumber  \\
+  &\quad (\text{I can add more steps later if necessary.})\\
+  &= \frac{1 + \sigma_{22}}{n \pi_{2+}} + o_p(1).
+\end{align}
+
+To compute the variance of $\hat \theta_{WLS}$, we have,
+
+\begin{align}\label{eq:varwls}
+  \Var(\hat \theta_{WLS})
+  &= \Var((M'V^{-1}V)^{-1}M'V^{-1}Z) \nonumber \\ 
+  &= E[\Var((M'V^{-1}V)^{-1}M'V^{-1}Z \mid n_1)]\nonumber  \\
+  &= E[(M'V^{-1}M)]\\
+\end{align}
+
+By Slutky's Theorem, $\Var(\hat \theta_{WLS}) \stackrel{P}{\to} (M'E[V]^{-1}M)^{-1}$.
+
+Finally, we compute the variance for the proposed estimator. First, we
+estimate the variance for a similar estimator $\tilde \theta_{prop}$. This
+estimator is the same as the proposed estimator except that we assume that 
+the estimates from $E[Y_{2i} \mid X, Y_1]$ and $E[Y_{2i} \mid X]$ are 
+population coefficients and independent of the rest of the sample. This
+could occur if there is a separate sample in which we can estimate these 
+expectations. Since the population coefficients of the regression 
+$Y = \beta_0 + \beta_1 X$ are
+
+\[\beta_0 = E[Y] - E[X] \Var(X)^{-1} Cov(X'Y) \text{ and } 
+\beta_1 = \Var(X)^{-1}\Cov(X, Y),\]
+
+we have that,
+
+\begin{align*}
+  E[Y_2 \mid X] &= E[Y_2] - \beta_1 E[X] + \beta_1 x = \theta_2 - x
+\end{align*}
+
+and
+
+\begin{align*}
+  E[Y_2 \mid X, Y_1] 
+  &= \beta_0 + [X, Y_1][\beta_1 \beta_2]' \\ 
+  &= E[Y_2] - E[[X, Y_1]] \Var([X, Y_1])^{-1}\Cov([X, Y_1], Y_2) + 
+  [X, Y_1]\Var([X, Y_1])^{-1} \Cov([X, Y_1], Y_2) \\ 
+  &= \theta_2 + [X, Y_1] \sigma_{11}^{-1} 
+  \begin{bmatrix}
+    1 + \sigma_{11} & -1 \\ -1 & 1
+  \end{bmatrix}
+  \begin{bmatrix}
+    1 & 1 + \sigma_{12}
+  \end{bmatrix} \\
+  &= \theta_2 + x \left(\frac{\sigma_{11} - \sigma_{12}}{\sigma_{11}}\right) + 
+  y_1 \left(\frac{\sigma_{12}}{\sigma_{11}}\right)\\
+  &= \theta_2 + x(1 - \sigma_{12}) + y_1(\sigma_{12}). 
+\end{align*}
+
+This means that
+
+\[E[Y_2 \mid X, Y_1] - E[Y_2 \mid X] = \sigma_{12}(y_1 - x) = \sigma_{12} e_1.\]
+
+Hence, whereas for $\theta = E[Y_2]$,
+
+\[\hat \theta_{prop} = n^{-1} \sum_{i = 1}^n \left\{E[Y_2 \mid X] +
+    \left(\frac{\delta_{1i}}{\pi_{1+}} - \frac{\delta_{11i}}{\pi_{11}}\right)
+    (E[Y_2 \mid X, Y_1] - E[Y_2 \mid X]) + 
+    \frac{\delta_{2i}}{\pi_{2+}}(y_{2i} - E[Y_2 \mid X])
+\right\}\]
+
+for the similar estimator we have
+
+\[\tilde \theta_{prop} = n^{-1} \sum_{i = 1}^n \left( \theta_2 + x_i + 
+    \left(\frac{\delta_{1i}}{\pi_{1+}} - \frac{\delta_{11i}}{\pi_{11}}\right)
+    \sigma_{12} e_{1i} + \frac{\delta_{2i}}{\pi_{2+}} (e_{2i}) \right)
+\]
+
+This means that
+
+\begin{align*}
+  \Var(\tilde \theta_{prop})
+  &= n^{-1}\Var\left(x + 
+    \left(\frac{\delta_{1}}{\pi_{1+}} - \frac{\delta_{11}}{\pi_{11}}\right)
+    \sigma_{12} e_{1} + \frac{\delta_{2}}{\pi_{2+}} (e_{2}) \right)\\
+  &= n^{-1} \left(\Var(x) + \Var\left(
+    \left(\frac{\delta_{1}}{\pi_{1+}} - \frac{\delta_{11}}{\pi_{11}}\right)
+    \sigma_{12} e_{1}
+    \right) + \Var\left( \frac{\delta_{2}}{\pi_{2+}} (e_{2}) \right)\right. \\ 
+  &\qquad+ \left.2\Cov\left(
+    \left(\frac{\delta_{1}}{\pi_{1+}} - \frac{\delta_{11}}{\pi_{11}}\right)
+    \sigma_{12} e_{1},
+     \frac{\delta_{2}}{\pi_{2+}} (e_{2}) 
+    \right) \right)\\
+  &\equiv A + B + C + D.
+\end{align*}
+
+\begin{align*}
+  A &= \Var(x) = 1.
+\end{align*}
+
+\begin{align*}
+  B &= \sigma_{12}^2 \sigma_{11} \Var\left(
+    \left(\frac{\delta_{1}}{\pi_{1+}} - \frac{\delta_{11}}{\pi_{11}}\right)
+  \right)\\
+    &=  \sigma_{12}^2 \sigma_{11} \left(\Var(\delta_1 / \pi_{1+}) + 
+      \Var(\delta_{11}/\pi_{11}) - 
+      2\Cov(\delta_1 / \pi_{1+}, \delta_{11} / \pi_{11}) \right)\\
+    &= \sigma_{12}^2 \sigma_{11}(\pi_{1+}^{-1} + \pi_{11}^{-1} - 2 - 2(\pi_{11}
+    - \pi_{1+} \pi_{11}) / (\pi_{1+}\pi_{11}))\\
+    &= \sigma_{12}^2 \sigma_{11}(\pi_{11}^{-1} - \pi_{1+}^{-1}).
+\end{align*}
+
+\begin{align*}
+  C &= \Var(\delta_2 e_2 / \pi_{2+})\\
+    &= \pi_{2+}^{-2}(\Var(\delta_2)\Var(e_2) + \Var(\delta_2)E[e_2]^2 +
+    \Var(e_2)E[\delta_2]^2)\\
+    &= \pi_{2+}^{-2} (\pi_{2+}(1 - \pi_{2+}) \sigma_{22} + \sigma_{22}\pi_{2+}^2)\\
+    &= \sigma_{22} / \pi_{2+}.
+\end{align*}
+
+\begin{align*}
+  D &= 2\Cov\left(
+    \left(\frac{\delta_{1}}{\pi_{1+}} - \frac{\delta_{11}}{\pi_{11}}\right)
+    \sigma_{12} e_{1}, \frac{\delta_{2}}{\pi_{2+}} (e_{2}) \right)\\
+    &= 2E\left[
+    \left(\frac{\delta_{1}}{\pi_{1+}} - \frac{\delta_{11}}{\pi_{11}}\right)
+    \sigma_{12} e_{1} \frac{\delta_{2}}{\pi_{2+}} (e_{2}) 
+    \right] \\ 
+    &= 2E\left[
+    \left(\frac{\delta_{1}}{\pi_{1+}} - \frac{\delta_{11}}{\pi_{11}}\right)
+    \frac{\delta_{2}}{\pi_{2+}}
+    \right] \sigma_{12}^2\\
+    &= 2E\left[\frac{\delta_{11}}{\pi_{1+}\pi_{2+}} - \frac{\delta_{11}}{\pi_{2+}}
+    \right] \sigma_{12}^2\\
+    &= 2 \frac{\pi_{11}}{\pi_{2+}} (\pi_{1+}^{-1} - 1) \sigma_{12}^2.
+\end{align*}
+
+Hence,
+
+\begin{equation}\label{eq:varprop}
+\Var(\tilde \theta_{prop}) = n^{-1}(1 +
+  \sigma_{12}^2\sigma_{11}(\pi_{11}^{-1} - \pi_{1+}^{-1}) + \sigma_{22}
+  \pi_{2+}^{-1} 2\sigma_{12}^2 (\pi_{1+}^{-1} - 1) \pi_{11}/ \pi_{2+}).
+\end{equation}
+
+Since $\hat \beta - \beta = O(n)$ from Theorem 2.2.1 of \cite{fuller2009sampling}, 
+$\Var(\hat \theta_{prop}) \to \Var(\tilde \theta_{prop})$. (I think that 
+I also need to show that the dependency between the estimated 
+coefficients and the observed data is bounded and that this bound goes 
+to zero as the sample size increases.)
+
+These estimated variances are almost exactly in line with the estimated
+standard deviations from the simulation study. With values of $n = 1000$,
+$\sigma_{11} = 1$, $\sigma_{22} = 1$, $\sigma_{12} = 0.5$, $\theta_2 = 5$,
+$\pi_{11} = 0.4$, $\pi_{10} = 0.2$, $\pi_{01} = 0.2$, and $\pi_{00} = 0.2$
+the standard deviations from the previous calculations and from the Monte Carlo
+simulation are listed in the Table~\ref{tab:calmc}:
+
+\begin{table}[!ht]
+  \centering
+  \label{tab:calmc}
+  \caption{This table compares the calculated and estimated standard 
+  errors for each of the algorithms listed. The calculated values are 
+  computed as the square root of the variance derived in the previous 
+  section. We use Equations~\ref{eq:varor}, \ref{eq:varorx}, \ref{eq:varcc}
+  \ref{eq:varwls}, and \ref{eq:varprop} to calculate the variance of 
+  Oracle, OracleX, CC, WLS, and Prop respectively and then take the square
+  root of the answer. For the estimated standard error, we take the 
+  standard deviation of the $B$ estimators for each algorithm.}
+  \begin{tabular}{lrr}
+    \toprule
+    Algorithm & Calculated Standard Error & Estimated Standard Error \\
+    \midrule
+    Oracle  & 0.045 & 0.044 \\
+    OracleX & 0.032 & 0.032 \\
+    CC      & 0.058 & 0.058 \\
+    WLS     & 0.040 & 0.040 \\
+    Prop    & 0.056 & 0.051 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\section*{Next Steps}
+
+\begin{itemize}
+  \item[1.] Since we have a way to estimate the variance, the goal is still 
+    to find the optimal $a$ and $b$ for Equation~\ref{eq:propclass}.
+\end{itemize}
+
+\printbibliography
 

### 2023-09-07 R/opt_est.R

@@ -295,11 +295,9 @@ expect_g <- function(df, g, cond, pred_df, d_vec) {
 #'  - df has the following columns: X, Y1, Y2, delta_y1, delta_y2, prob_vec
 #'  - If delta_yi = 1 then we assume that Y_i is observed if delta_yi = 0 then
 #'  - we assume that Y_i is missing.
-prop_nmono_est <- function(df, oracle = TRUE) {
+prop_nmono_est <- function(df, gfun = "Y2", oracle = TRUE) {
 
-  # HACK: We only test the case where we have g(X, Y1, Y2) = Y2.
-  gfun <- "Y2"
-
+  df <- mutate(df, g_i = eval(rlang::parse_expr(gfun)))
   df_11 <- filter(df, delta_1 == 1, delta_2 == 1)
   df_1 <- filter(df, delta_1 == 1)
   df_2 <- filter(df, delta_2 == 1)
@@ -321,11 +319,12 @@ prop_nmono_est <- function(df, oracle = TRUE) {
   # E[g | X]
   # E[g | X, Y1]
   # E[g | X, Y2] = Y2
-  eg_x <- expect_g(df, g = gfun, cond = c("X"), pred_df = df, d_vec = rep(1, nrow(df)))
+  eg_x <- 
+    expect_g(df, g = "g_i", cond = c("X"), pred_df = df, d_vec = rep(1, nrow(df)))
   eg_xy1 <-
-    expect_g(df, g = gfun, cond = c("X", "Y1"), pred_df = df_1, d_vec = df$delta_1)
+    expect_g(df, g = "g_i", cond = c("X", "Y1"), pred_df = df_1, d_vec = df$delta_1)
   eg_xy2 <-
-    expect_g(df, g = gfun, cond = c("X", "Y2"), pred_df = df_2, d_vec = df$delta_2)
+    expect_g(df, g = "g_i", cond = c("X", "Y2"), pred_df = df_2, d_vec = df$delta_2)
 
   # 3. Get estimate.
   # This is correct because eg_xyi both estimate E[g | X, Yi] using all of the 
@@ -333,86 +332,47 @@ prop_nmono_est <- function(df, oracle = TRUE) {
   mean(eg_x) + 
     mean(df$delta_1 / pi1_w * (eg_xy1 - eg_x)) +
     mean(df$delta_2 / pi2_w * (eg_xy2 - eg_x)) +
[…]

### 2023-09-07 R/utilities.R

@@ -1,0 +1,21 @@
+# Title: utilities.R
+# Created by: Caleb Leedy
+# Created on: September 07, 2023
+# Purpose: This file contains code that is common and used in 
+# multiple scripts but is still very general.
+
+# TODO: Test me!
+#' Thanks Hengfang Wang for this matrix inversion function!
+hw_inv = function(X, eps = 1e-12) {
+  eig.X = eigen(X, symmetric = TRUE)
+  P = eig.X[[2]] 
+  lambda = eig.X[[1]] 
+  ind = lambda > eps
+  lambda[ind] = 1 / lambda[ind] 
+  lambda[!ind] = 0
+  ans = P %*% diag(lambda) %*% t(P)
+  return(ans)
+}
+
+# TODO: Test me!
+expit <- function(x) {1 / (1 + exp(-x))}

### 2023-09-07 Simulations/Comparing_Estimator_Types/calibration_comp.R

@@ -1,4 +1,4 @@
-# Title: Simulations/calibration_comp.R
+# Title: Comparing_Estimator_Types/calibration_comp.R
 # Author: Caleb Leedy
 # Date Created: April 25, 2023
 # Purpose: This script contains the code to compare the calibration estimators

### 2023-09-07 Simulations/Comparing_Estimator_Types/generate_data.R

@@ -1,0 +1,202 @@
+# Title: Comparing_Estimator_Types/generate_data.R
+# Author: Caleb Leedy
+# Date Created: April 25, 2023
+# Purpose: This file contains the code to generate data for different simulation
+# studies. Instead of writing the functions in Simulations/nonmonotone.R and
+# then copying them to other files, I thought that it would be a good idea to
+# put them all into one script and then source it.
+
+# ------------------------------------------------------------------------------
+
+# ********************
+# * Script Functions *
+# ********************
+# 1. mono_mar -- This generates monotone missingness.
+# 2. nonmono_mar -- This generates nonmonotone missing data.
+# 3. expit -- A helper function for generating probabilites. 
+
+# ------------------------------------------------------------------------------
+
+# **********************
+# * Required Libraries *
+# **********************
+
+library(MASS)
+library(dplyr)
+
+# ------------------------------------------------------------------------------
+
+# ************************
+# * Additional Functions *
+# ************************
+
+expit <- function(x) {1 / (1 + exp(-x))}
+
+# *****************
+# * Generate Data *
+# *****************
+
+mono_mar <- function(n, mean_y2 = 0) {
+
+  # Steps:
+  # 1. Generate X, Y_1, Y_2.
+  # 2. Get probabilities conditional on past.
+  # 3. Reveal hidden variables.
+
+  # 1. Generate X, Y_1, Y_2.
+  x_vec <- rnorm(n)
+  y1 <- rnorm(n)
+  y2 <- rnorm(n, mean = mean_y2)
+
+  # 2. Get probabilities conditional on past.
+  r1 <- rbinom(n, 1, expit(x_vec))
+  r2 <- rbinom(n, 1, expit(y1))
+
+  # 3. Reveal hidden variables.
+  tibble(x = x_vec, y1 = y1, y2 = y2, r1 = r1, r2 = r2) %>%
[…]

### 2023-09-07 Simulations/Comparing_Estimator_Types/multiphase_est.R

@@ -1,6 +1,6 @@
-# Title: Simulations/multiphase_est.R
+# Title: Comparing_Estimator_Types/multiphase_est.R
 # Author: Caleb Leedy
-# Date Created: May  3, 2023
+# Date Created: May 3, 2023
 # Purpose: This document first compares the two three-phase sampling estimators:
 # F3P and K3P (see Notes/multiphase_estimators.tex or page 33 in the research
 # journal).

### 2023-09-07 Simulations/Comparing_Estimator_Types/nmono_est_funs.R

@@ -1,4 +1,4 @@
-# Title: ../Simulations/nmono_est_funs.R
+# Title: Comparing_Estimator_Types/nmono_est_funs.R
 # Author: Caleb Leedy
 # Date Created: April 25, 2023
 # Purpose: This script contains the functions used to estimate monotone and

### 2023-09-07 Simulations/Comparing_Estimator_Types/nonmonotone.R

@@ -1,4 +1,4 @@
-# Title: nonmonotone.R
+# Title: Comparing_Estimator_Types/nonmonotone.R
 # Author: Caleb Leedy
 # Date Created: March 29, 2023
 # Purpose: This document contains the simulations for computing estimating

### 2023-09-07 Simulations/Comparing_Estimator_Types/regression_est.R

@@ -1,4 +1,4 @@
-# Title: regression_est.R
+# Title: Comparing_Estimator_Types/regression_est.R
 # Author: Caleb Leedy
 # Date Created: May  5, 2023
 # Purpose: This file runs some code that I use for better understanding of the

### 2023-09-07 Simulations/Optimal_Nonmonotone_Est/comp_ests_g.R

@@ -1,0 +1,139 @@
+# Title: Optimal_Nonmonotone_Est/comp_ests_g.R
+# Created by: Caleb Leedy
+# Created on: September 06, 2023
+# Purpose: This file contains the code to test the proposed estimator with a 
+# g function of g = "Y1^2Y2".
+
+# ***********
+# * Outline *
+# ***********
+# 1. Load Libraries and code.
+# 2. Run Monte Carlo simulation.
+
+# *************
+# * Libraries *
+# *************
+
+library(dplyr)
+library(stringr)
+
+library(doParallel)
+library(doRNG)
+library(parallelly)
+
+source("R/opt_est.R")
+
+# ***************************
+# * Monte Carlo Simulations *
+# ***************************
+
+clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
+registerDoParallel(clust)
+
+B <- 3000
+n_obs <- 1000
+true_theta <- 5
+cor_e1e2 <- 0.5
+
+mc_theta <-
+  foreach(iter = 1:B,
+          .options.RNG = 1,
+          .packages = c("dplyr", "stringr")) %dorng% {
+
+    # g_fun <- "Y1^2Y2"
+    
+    # Generate Data
+    df <- gen_optsim_data(n = n_obs, theta = true_theta, cor_e1e2 = cor_e1e2)
+
+    # Get Estimates
+    oracle_est <- mean(df$Y1^2 * df$Y2)
+    oraclex_est <- mean((df$Y1 - df$X)^2 * (df$Y2 - df$X))
+    ipw_est <- 
+      filter(df, delta_1 == 1, delta_2 == 1) |>
+      mutate(g_f = Y1^2 * Y2) |>
+      mutate(g_est = g_f / prob_11) |>
+      pull(g_est) |>
+      sum() / nrow(df)
+    cc_est <- 
+      filter(df, delta_2 == 1) |>
+      mutate(g_f = Y1^2 * Y2) |>
+      pull(g_f) |>
+      mean()
+    # FIXME: Change g_fun
+    wls_est <- opt_lin_est(df, cov_y1y2 = cor_e1e2)
+    # FIXME: Change g_fun
+    prop_est <- prop_nmono_est(df)
+
+    return(tibble(oracle = oracle_est,
+                  oraclex = oraclex_est,
[…]

### 2023-09-07 Simulations/Optimal_Nonmonotone_Est/optimal_simulation.R

@@ -1,4 +1,4 @@
-# Title: optimal_simulation.R
+# Title: Optimal_Nonmonotone_Est/optimal_simulation.R
 # Created by: Caleb Leedy
 # Created on: August 19, 2023
 # Purpose: This file contains the code to test the proposed estimator with a
@@ -30,33 +30,38 @@ source("R/opt_est.R")
 clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
 registerDoParallel(clust)
 
-B <- 1000
+B <- 3000
 n_obs <- 1000
-true_theta <- 0
+true_theta <- 5
+cor_e1e2 <- 0.5
 
 mc_theta <-
-  foreach(iter = 1:B, .options.RNG = 1, .packages = c("dplyr", "stringr")) %dorng% {
+  foreach(iter = 1:B,
+          .options.RNG = 1,
+          .packages = c("dplyr", "stringr")) %dorng% {
 
     # Generate Data
-    df <- gen_optsim_data(n = n_obs, theta = true_theta)
+    df <- gen_optsim_data(n = n_obs, theta = true_theta, cor_e1e2 = cor_e1e2)
 
     # Get Estimates
     oracle_est <- mean(df$Y2)
+    oraclex_est <- mean(df$Y2 - df$X)
     ipw_est <- 
       filter(df, delta_1 == 1, delta_2 == 1) |>
-      mutate(w_est = Y2 / prob_11) |>
-      pull(w_est) |>
+      mutate(g_est = Y2 / prob_11) |>
+      pull(g_est) |>
       sum() / nrow(df)
     cc_est <- 
       filter(df, delta_2 == 1) |>
       pull(Y2) |>
       mean()
     # em_est <- em_optsim(df)
-    wls_est <- opt_lin_est(df)
+    wls_est <- opt_lin_est(df, cov_y1y2 = cor_e1e2)
     prop_est <- prop_nmono_est(df)
     prop2_est <- prop2_nmono_est(df)
 
     return(tibble(oracle = oracle_est,
+                  oraclex = oraclex_est,
                   cc = cc_est,
                   ipw = ipw_est,
                   # em = em_est,
@@ -74,18 +79,21 @@ stopCluster(clust)
 mc_theta |>
   summarize(
[…]

### 2023-09-07 Simulations/Paper_Replications/sim_kim2021combining.R

@@ -1,4 +1,4 @@
-# Title: sim_kim2021combining.R
+# Title: Paper_Replications/sim_kim2021combining.R
 # Author: Caleb Leedy
 # Date Created: April 20, 2023
 # Purpose: This document contains the code to reproduce the simulation study of

### 2023-09-07 Simulations/generate_data.R

@@ -1,68 +1,0 @@
-# Title: Simulations/generate_data.R
-# Author: Caleb Leedy
-# Date Created: April 25, 2023
-# Purpose: This file contains the code to generate data for different simulation
-# studies. Instead of writing the functions in Simulations/nonmonotone.R and
-# then copying them to other files, I thought that it would be a good idea to
-    # p2 = Pr(R2 = 1) # NOTE: This is NOT true because R2 = 1 if p12 works.
-    # p12 = Pr(R2 = 1 | R1 == 1)
-    # p21 = Pr(R1 = 1 | R2 == 1)
-    # Instead, p0, p1, and p2 are Pr(i in Group(r)) which determines the order
-    # which we fill missing values. Thus,
-    # Pr(R1 = 1) = p1 + p21 * p2
-    # Pr(R2 = 1) = p2 + p12 * p1
-
-    if (miss_resp) {
-
-      p0 <- abs(x_vec^2 - 1)
-      p1 <- abs(x_vec^2 - 2)
-      p2 <- abs(x_vec^2)
-
-      tot <- p0 + p1 + p2
-      p0 <- p0 / tot
-      p1 <- p1 / tot
-      p2 <- p2 / tot
-
-      p12 <- expit(y1^2)
-      p21 <- expit(y2^2)
-
-
-    } else {
-
-      p0 <- 0.2
-      # p1 <- expit(2 * x_vec)
-      p1 <- 0.4
-      # p2 <- expit(-2 * x_vec)
-      p2 <- 0.4
-      tot <- p0 + p1 + p2
-      p0 <- p0 / tot
-      p1 <- p1 / tot
-      p2 <- p2 / tot
-
-      p12 <- expit(y1)
-      p21 <- expit(y2)
-
-    }
-
-    rand1 <- runif(n)
-    rand2 <- runif(n)
-
-  }
-
-  # 3. Reveal hidden variables.
-  tibble(x = x_vec, y1 = y1, y2 = y2,
-         p0 = p0, p1 = p1, p2 = p2,
-         p12 = p12, p21 = p21,
-         rand1 = rand1, rand2 = rand2) %>%
-  mutate(out = case_when(
-                rand1 < p0 ~ "00",
-                rand1 < p0 + p1 & rand2 < 1 - p12 ~ "10",
-                rand1 < p0 + p1 & rand2 > 1 - p12 ~ "11a",
[…]

### 2023-09-11 Notes/model_assumptions.tex
Added non-linear simulation for estimation comparison.

@@
 the standard deviations from the previous calculations and from the Monte Carlo
-simulation are listed in the Table~\ref{tab:calmc}:
+simulation are listed in Table~2:
+%Table~\ref{tab:calmc}:
 
@@
 
+\newpage
+  \quad
+\newpage
+
+\section*{Non-linear Estimation}
+
+\begin{itemize}
+  \item While we have previously shown that in the case of the linear 
+    estimator, that the weighted least squares (WLS) estimator is superior 
+    to the proposed estimator. However, we also want to see if this holds 
+    for a non-linear function. Consider the estimating equation $g(X, Y_1, Y_2)
+     = Y_1^2 Y_2$. Using the same setup as the previous simulation, we have the
+     same definitions of $X$, $Y$, $n$, $\pi$, and $B$.
+
+     For the definition of each estimator we have the following algorithms:
+
+     \begin{itemize}
+      \item Oracle: This computes the average value of $Y_1^2Y_2$ in all 
+        observations (even when $Y_1$ or $Y_2$ is not supposed to be observed).
+        \[\hat \theta = n^{-1} \sum_{i = 1}^n y_{1i}^2y_{2i}.\]
+
+      \item OracleX: This computes the average value of $Y_1^2 (Y_2 - X)$ in 
+        all observations (even when $Y_1$ or $Y_2$ is not supposed to be observed.)
+        \[\hat \theta = n^{-1} \sum_{i = 1}^n y_{1i}^2(y_{2i} - x_i).\]
+
+      \item CC: This computes the average value of $Y_1^2Y_2$ in all
+        observations in which $Y_1$ and $Y_2$ are observed.
+        \[\hat \theta = \frac{\sum_{i = 1}^n \delta_{1i}\delta_{2i} y_{1i}^2y_{2i}}
+        {\sum_{i = 1}^n \delta_{1i}\delta_{2i}}.\]
+
+      \item Proposed: This is the proposed estimator from Equation~\ref{eq:prop}.
+        Note that since,
+
+        \begin{align*}
+          Y_1^2 Y_2 &= (x + e_1)^2 (\theta_2 + x + e_2) \\ 
+                    &= x^3 + x^2 (\theta_2 + 2 e_1 + e_2) 
+                    + x(2\theta_2 e_1 + 2 e_1 e_2 + e_1^2) + e_1^2 \theta_2 
+                    + e_1^2 e_2
+        \end{align*}
+
+        for the conditional expectations, we have
+        \begin{align*}
+          E[Y_1^2 Y_2 \mid X] &= x^3 + x^2 \theta_2 + x(2\sigma_{12} + 1) + \theta_2\\
+          E[Y_1^2 Y_2 \mid X, Y_1] &= Y_1^2 (x + \theta_2) \text{ and }\\
+          E[Y_1^2 Y_2 \mid X, Y_2] &= (x^2 + 1) Y_2.
+        \end{align*}
+
+      \item WLS: To derive the WLS estimator we can view the sample as 
+        consisting of four independent parts: $A_{11}, A_{10}, A_{01}$, and 
+        $A_{00}$, where $A_{ij}$ indicates the missingness levels of $Y_1$ (if 
+        i = 0) or $Y_2$ (if $j = 0$). We know that the correctly specified
+        regression model for $g$ in each of these segments is the following:
+
+        \begin{align*}
+          A_{11}&: Y_1^2 Y_2 \\
+          A_{10}&: Y_1^2  (x + \theta_2) \\
+          A_{01}&: (x^2 + 1) Y_2\\
+          A_{00}&: x^3 + x^2 \theta_2 + x(2\sigma_{12} + 1) + \theta_2
+        \end{align*}
+
+        The estimator is then to estimate $\theta_2$ using Equation~\ref{eq:wls}.
+        Then, we can plug in the estimated value of $\theta_2$ into the previous 
+        equations to get estimates of $g$. The overall estimate is the estimated 
+        variance weighted average of the mean estimate from each segment.
+
+      \item WLSTT: Since the previous estimator is biased, we use the same
+        estimation procedure with the true value of $\theta_2$ to the estimator 
+        $\hat \theta_{WLSTT}$.
+     \end{itemize}
+
+     \begin{table}[!ht]
+       \caption{True g is 10. $\Cov(e_1, e_2) = 0.5$}
+       \centering
+       \begin{tabular}[t]{lrrrr}
+         \toprule
+         Algorithm & Bias & SD & Tstat & Pval\\
+         \midrule
+         Oracle & 0.007 & 0.529 & 0.741 & 0.229\\
+         Oraclex & 0.004 & 0.475 & 0.510 & 0.305\\
+         CC & 0.023 & 0.824 & 1.518 & 0.065\\
+         WLS & -0.131 & 0.387 & -18.504 & 0.000\\
+         WLSTT & -0.132 & 0.380 & -19.007 & 0.000\\
+         Prop & 0.011 & 0.625 & 0.952 & 0.171\\
+         \bottomrule
+       \end{tabular}
+     \end{table}
+
+     In this scenario, the WLS estimator is biased. Other than the oracle 
+     estimators, the proposed estimator is the best unbiased estimator.
+
+\end{itemize}
+
 \section*{Next Steps}
@@
 \begin{itemize}
-  \item[1.] Since we have a way to estimate the variance, the goal is still 
-    to find the optimal $a$ and $b$ for Equation~\ref{eq:propclass}.
+  \item I am slightly unclear about how to proceed. The proposed estimator 
+    seemed to work quite well for the non-linear experiment. However, it was 
+    easily out-performed by the WLS estimator in the linear case. The WLS 
+    estimator is a relatively intuitive estimator that is optimal in that 
+    situation. However, it seems to have worse performance in the non-linear 
+    case.
 \end{itemize}

### 2023-09-11 R/opt_est.R

@@ -19,12 +19,6 @@
 library(dplyr)
 library(stringr)
 
-# ************************
-# * Additional Functions *
-# ************************
-
-expit <- function(x) {1 / (1 + exp(-x))}
-
 # *****************
 # * Generate Data *
 # *****************
@@ -282,6 +276,40 @@ expect_g <- function(df, g, cond, pred_df, d_vec) {
   }
 }
 
+#' This gets all of the formula terms for higher order formulas.
+#'
+#' @param str_vec - A vector of strings
+#' @param pow - An integer greater than or equal to 1
+#'
+#' @returns ret - A vector of strings
+#'
+#' @details
+#' Preconditions:
+#'  - !("" %in% str_vec)
+get_v_powers <- function(str_vec, pow = 1) {
+  v <- c(str_vec, "")
+  lst <- list(v)
+
+  if (pow == 1) {
+    res <- v
+  } else {
+
+    for (iter in 2:pow) {
+      lst[[iter]] <- v
+    }
+
+    df <- expand.grid(lst, stringsAsFactors = FALSE)
+    res <- c()
+    for (i in 1:nrow(df)) {
+      res[i] <- str_flatten(sort(as.character(df[i, ])), collapse = "*")
+    }
+  }
+
+  res <- unique(str_replace(res, "^\\*+", ""))
+  res <- res[purrr::map_lgl(res, function(s) {str_length(s) != 0})]
+  return(paste0("I(", res, ")"))
+}
+
 #' The proposed non-monotone estimator. This function gives the result based
 #' on an input data frame.
 #'
@@ -295,7 +323,7 @@ expect_g <- function(df, g, cond, pred_df, d_vec) {
 #'  - df has the following columns: X, Y1, Y2, delta_y1, delta_y2, prob_vec
 #'  - If delta_yi = 1 then we assume that Y_i is observed if delta_yi = 0 then
 #'  - we assume that Y_i is missing.
-prop_nmono_est <- function(df, gfun = "Y2", oracle = TRUE) {
+prop_nmono_est <- function(df, gfun = "Y2", oracle = TRUE, pow = 1) {
 
[…]

### 2023-09-11 Simulations/T2Optimal_Nonmonotone_Est/comp_ests_g.R

@@ -2,7 +2,9 @@
 # Created by: Caleb Leedy
 # Created on: September 06, 2023
 # Purpose: This file contains the code to test the proposed estimator with a 
-# g function of g = "Y1^2Y2".
+# g function of g = "Y1^2Y2". This code is very similar to 
+# ./optimal_simulation.R except that the g function considered is Y1^2 Y2 
+# instead of Y2.
 
 # ***********
 # * Outline *
@@ -47,7 +49,7 @@ mc_theta <-
 
     # Get Estimates
     oracle_est <- mean(df$Y1^2 * df$Y2)
-    oraclex_est <- mean((df$Y1 - df$X)^2 * (df$Y2 - df$X))
+    oraclex_est <- mean((df$Y1)^2 * (df$Y2 - df$X))
     ipw_est <- 
       filter(df, delta_1 == 1, delta_2 == 1) |>
       mutate(g_f = Y1^2 * Y2) |>
@@ -55,21 +57,21 @@ mc_theta <-
       pull(g_est) |>
       sum() / nrow(df)
     cc_est <- 
-      filter(df, delta_2 == 1) |>
+      filter(df, delta_1 == 1, delta_2 == 1) |>
       mutate(g_f = Y1^2 * Y2) |>
       pull(g_f) |>
       mean()
-    # FIXME: Change g_fun
-    wls_est <- opt_lin_est(df, cov_y1y2 = cor_e1e2)
-    # FIXME: Change g_fun
-    prop_est <- prop_nmono_est(df)
+    wls_est <- comb_lin_est(df, gfun = "Y1^2 * Y2", cov_e1e2 = cor_e1e2)
+    wlstt_est <- 
+      comb_lin_est(df, gfun = "Y1^2 * Y2", cov_e1e2 = cor_e1e2, theta2 = true_theta)
+    prop_est <- prop_nmono_est(df, gfun = "Y1^2 * Y2", pow = 3)
 
     return(tibble(oracle = oracle_est,
                   oraclex = oraclex_est,
                   cc = cc_est,
                   ipw = ipw_est,
-                  # em = em_est,
                   wls = wls_est,
+                  wlstt = wlstt_est,
                   prop = prop_est))
   } |> 
   bind_rows()
@@ -79,30 +81,31 @@ stopCluster(clust)
[…]

### 2023-09-11 tests/testthat/test-opt_est.R

@@ -9,6 +9,7 @@
 # *********
 
 # Test expit
+# expit has been moved to R/utilities.R
 test_that("expit works", {
   expect_equal(expit(0), 0.5)
   expect_equal(expit(Inf), 1)
@@ -35,6 +36,12 @@ test_that("expect_g works", {
   expect_equal(expect_g(df, "Y2", c("X"), df, rep(1, nrow(df))), df$Y2)
   expect_equal(expect_g(df, "Y2", c("X", "Y1"), df_1, df$delta_1), tmp1)
   expect_equal(expect_g(df, "Y2", c("X", "Y2"), df_2, df$delta_2), tmp2)
+})
+
+# Test get_v_powers
+test_that("get_v_powers works", {
+  expect_equal(get_v_powers(c("X"), 1), c("I(X)"))
+  expect_equal(get_v_powers(c("X", "Y1"), 1), c("I(X)", "I(Y1)"))
 })
 
 # Test get_cond_exp

### 2023-10-23 Notes/model_assumptions.tex
Added October 23 2023

--- 2023-09-07 Notes/model_assumptions.tex
@@
 
-\section*{Next Steps}
-
-\begin{itemize}
-  \item[1.] Since we have a way to estimate the variance, the goal is still 
-    to find the optimal $a$ and $b$ for Equation~\ref{eq:propclass}.
-\end{itemize}
+%\newpage
+%  \quad
+%\newpage
+%
+%\section*{Non-linear Estimation}
+%
+%\begin{itemize}
+%  \item While we have previously shown that in the case of the linear 
+%    estimator, that the weighted least squares (WLS) estimator is superior 
+%    to the proposed estimator. However, we also want to see if this holds 
+%    for a non-linear function. Consider the estimating equation $g(X, Y_1, Y_2)
+%     = Y_1^2 Y_2$. Using the same setup as the previous simulation, we have the
+%     same definitions of $X$, $Y$, $n$, $\pi$, and $B$.
+%
+%     For the definition of each estimator we have the following algorithms:
+%
+%     \begin{itemize}
+%      \item Oracle: This computes the average value of $Y_1^2Y_2$ in all 
+%        observations (even when $Y_1$ or $Y_2$ is not supposed to be observed).
+%        \[\hat \theta = n^{-1} \sum_{i = 1}^n y_{1i}^2y_{2i}.\]
+%
+%      \item OracleX: This computes the average value of $Y_1^2 (Y_2 - X)$ in 
+%        all observations (even when $Y_1$ or $Y_2$ is not supposed to be observed.)
+%        \[\hat \theta = n^{-1} \sum_{i = 1}^n y_{1i}^2(y_{2i} - x_i).\]
+%
+%      \item CC: This computes the average value of $Y_1^2Y_2$ in all
+%        observations in which $Y_1$ and $Y_2$ are observed.
+%        \[\hat \theta = \frac{\sum_{i = 1}^n \delta_{1i}\delta_{2i} y_{1i}^2y_{2i}}
+%        {\sum_{i = 1}^n \delta_{1i}\delta_{2i}}.\]
+%
+%      \item Proposed: This is the proposed estimator from Equation~\ref{eq:prop}.
+%        Note that since,
+%
+%        \begin{align*}
+%          Y_1^2 Y_2 &= (x + e_1)^2 (\theta_2 + x + e_2) \\ 
+%                    &= x^3 + x^2 (\theta_2 + 2 e_1 + e_2) 
+%                    + x(2\theta_2 e_1 + 2 e_1 e_2 + e_1^2) + e_1^2 \theta_2 
+%                    + e_1^2 e_2
+%        \end{align*}
+%
+%        for the conditional expectations, we have
+%        \begin{align*}
+%          E[Y_1^2 Y_2 \mid X] &= x^3 + x^2 \theta_2 + x(2\sigma_{12} + 1) + \theta_2\\
+%          E[Y_1^2 Y_2 \mid X, Y_1] &= Y_1^2 (x + \theta_2) \text{ and }\\
+%          E[Y_1^2 Y_2 \mid X, Y_2] &= (x^2 + 1) Y_2.
+%        \end{align*}
+%
+%      \item WLS: To derive the WLS estimator we can view the sample as 
+%        consisting of four independent parts: $A_{11}, A_{10}, A_{01}$, and 
+%        $A_{00}$, where $A_{ij}$ indicates the missingness levels of $Y_1$ (if 
+%        i = 0) or $Y_2$ (if $j = 0$). We know that the correctly specified
+%        regression model for $g$ in each of these segments is the following:
+%
+%        \begin{align*}
+%          A_{11}&: Y_1^2 Y_2 \\
+%          A_{10}&: Y_1^2  (x + \theta_2) \\
+%          A_{01}&: (x^2 + 1) Y_2\\
+%          A_{00}&: x^3 + x^2 \theta_2 + x(2\sigma_{12} + 1) + \theta_2
+%        \end{align*}
+%
+%        The estimator is then to estimate $\theta_2$ using Equation~\ref{eq:wls}.
+%        Then, we can plug in the estimated value of $\theta_2$ into the previous 
+%        equations to get estimates of $g$. The overall estimate is the estimated 
+%        variance weighted average of the mean estimate from each segment.
+%
+%      \item WLSTT: Since the previous estimator is biased, we use the same
+%        estimation procedure with the true value of $\theta_2$ to the estimator 
+%        $\hat \theta_{WLSTT}$.
+%     \end{itemize}
+%
+%     \begin{table}[!ht]
+%       \caption{True g is 10. $\Cov(e_1, e_2) = 0.5$}
+%       \centering
+%       \begin{tabular}[t]{lrrrr}
+%         \toprule
+%         Algorithm & Bias & SD & Tstat & Pval\\
+%         \midrule
+%         Oracle & 0.007 & 0.529 & 0.741 & 0.229\\
+%         Oraclex & 0.004 & 0.475 & 0.510 & 0.305\\
+%         CC & 0.023 & 0.824 & 1.518 & 0.065\\
+%         WLS & -0.131 & 0.387 & -18.504 & 0.000\\
+%         WLSTT & -0.132 & 0.380 & -19.007 & 0.000\\
+%         Prop & 0.011 & 0.625 & 0.952 & 0.171\\
+%         \bottomrule
+%       \end{tabular}
+%     \end{table}
+%
+%     In this scenario, the WLS estimator is biased. Other than the oracle 
+%     estimators, the proposed estimator is the best unbiased estimator.
+%
+%\end{itemize}
+%
+%\section*{Next Steps}
+%
+%\begin{itemize}
+%  \item I am slightly unclear about how to proceed. The proposed estimator 
+%    seemed to work quite well for the non-linear experiment. However, it was 
+%    easily out-performed by the WLS estimator in the linear case. The WLS 
+%    estimator is a relatively intuitive estimator that is optimal in that 
+%    situation. However, it seems to have worse performance in the non-linear 
+%    case.
+%\end{itemize}
 

### 2023-10-24 Notes/optest_update.tex
Added working notes for Dr. Fuller.

\documentclass[12pt]{article}

\usepackage{amsmath, amssymb, mathrsfs, fancyhdr}
\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
\usepackage{biblatex}
\usepackage{booktabs}
\usepackage{float}

\addbibresource{references.bib}

\hypersetup{colorlinks = true, urlcolor = black}

\headheight     15pt
\topmargin      -1.5cm   % read Lamport p.163
\oddsidemargin  -0.04cm  % read Lamport p.163
\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
\textwidth      16.59cm
\textheight     23.94cm
\parskip         7.2pt   % sets spacing between paragraphs
\parindent         0pt   % sets leading space for paragraphs
\pagestyle{empty}        % Uncomment if don't want page numbers
\pagestyle{fancyplain}

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}

\renewcommand{\arraystretch}{1.2}

\begin{document}

\lhead{Caleb Leedy}
\chead{Non-monotone Missingness}
%\chead{STAT 615 - Advanced Bayesian Methods}
%\rhead{Page \thepage\ of \pageref{LastPage}}
\rhead{\today}

\section*{Setup}

Consider the following setup. Let $(X, Y_1, Y_2, \delta)  \stackrel{ind}{\sim}
F$ for some distribution $F$ that is unknown. We define $Z = (X, Y_1, Y_2$ and
we observe the following:

\begin{table}[!ht]
  \centering
  \caption{This table shows some of our notation and some of the corresponding
  notation from \cite{tsiatis2006semiparametric}.}
\begin{tabular}{cccccccc}
  \toprule
  Segments & $X$ & $Y_1$ & $Y_2$ & Prob. Element in Segment & $\delta$ & $C$ & $G_C(Z)$ \\
  \midrule
  $A_{00}$ & $\checkmark$ &               &         & $\pi_{00}$ & $\delta_{00}$ & 1 & $\{X\}$ \\
  $A_{10}$ & $\checkmark$ & $\checkmark$  &         & $\pi_{10}$ & $\delta_{10}$ & 2 & $\{X, Y_1\}$ \\
  $A_{01}$ & $\checkmark$ &               & $\checkmark$ & $\pi_{01}$ & $\delta_{01}$ & 3 & $\{X, Y_2\}$ \\
  $A_{11}$ & $\checkmark$ & $\checkmark$  & $\checkmark$ & $\pi_{11}$ & $\delta_{11}$ & $\infty$ & $\{X, Y_1, Y_2\}$ \\
  \bottomrule
\end{tabular}
\end{table}

Define $\varpi(r, Z) = \Pr(C = r \mid Z)$. For now, we assume that $\varpi(r,
Z)$ is known. Notice that $\varpi(\infty, Z) = \pi_11$, $\varpi(3, Z) = \pi_{01}$,
$\varpi(2, Z) = \pi_{10}$, and $\varpi(1, Z) = \pi_{00}$. Since $\pi_{00} +
\pi_{10} + \pi_{01} + \pi_{11} = 1$, we only need to define three inclusion 
probabilities.\\

Suppose that we want to estimate $\theta = E[g(X, Y_1, Y_2)]$ for a known
function $g$. The proposed estimator is

\begin{align}\label{eq:prop}
  &\hat \theta_{prop} =\\ \nonumber
  &n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] + n^{-1} \sum_{i = 1}^n
  \frac{\delta_{10}}{\pi_{10}} (E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i]) 
  + n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} (E[g_i \mid X_i, Y_{2i}] - E[g_i \mid X_i]) \\ \nonumber
  &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} (g_i - E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i, Y_{2i}] + E[g_i \mid X_i]).
\end{align}

The goal is the show that over all functions $b_1(X, Y_1)$ and $b_2(X, Y_2)$,
$\hat \theta_{prop}$ is the optimal estimator in the form:

\begin{align}
  &\hat \theta =\\ \nonumber
  &n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] + n^{-1} \sum_{i = 1}^n
  \frac{\delta_{10}}{\pi_{10}} (b_1(X_i, Y_{1i}) - E[g_i \mid X_i]) 
  + n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} (b_2(X_i, Y_{2i}) - E[g_i \mid X_i]) \\ \nonumber
  &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} (g_i - b_1(X_i, Y_{1i}) - b_2(X_i, Y_{2i}) + E[g_i \mid X_i])
\end{align}

\section*{Semiparametric Inference}

We know from Theorem 7.2 of \cite{tsiatis2006semiparametric} that if $\Pr(C =
\infty \mid Z) = \pi_{11} > 0$ 
then the semiparametric influence function has the form (see page 20 of notes):

\[ \frac{I(C = \infty) g(Z)}{\varpi(\infty, Z)} + \frac{I(C =
  \infty)}{\varpi(\infty, Z)} \left(\sum_{r \neq \infty} \varpi(r, G_r(Z))
  L_{2r}(G_r(Z))\right) - \sum_{r \neq \infty} I(C = \infty)L_{2r}(G_r(Z))\]

where $L_{2r}(G_r(Z))$ is an arbitrary function of $G_r(Z)$. Notice, that this
form does not identify \textit{the} optimal estimator but a class of 
semiparametric functions. A reasonable choice of an estimator for $L_{2r}(G_r(Z))$ is

\[L_{2r}(G_r(Z)) = E[g(Z) \mid G_r(Z)].\]

It turns out that $\hat \theta_{prop}$ from Equation~\ref{eq:prop} is contained within 
the previous class of estimators with

\[L_{2r}(G_r(Z)) = \varpi(r, G_r(Z)) E[g(Z) \mid G_r(Z)].\]

(See pages 81-82 of notes for proof.) \\

\subsection*{Linear Expectations}

To simplify this problem, we consider the following estimator\footnote{This
estimator has slightly different coefficients compared to the initial
estimator.}:

\begin{align}
  &\hat \theta_c \\ \nonumber
  &= \frac{\delta_{11}}{\pi_{11}} g(Z) + \left(1 -
    \left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}}\right) -
    \left(\frac{\delta_{01} + \delta_{11}}{\pi_{01} + \pi_{11}}\right) + 
  \frac{\delta_{11}}{\pi_{11}}\right) c_0 E[g \mid X] \\ \nonumber
  &+
  \left(\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}}\right) - 
  \frac{\delta_{11}}{\pi_{11}}\right) c_1 E[g \mid X, Y_1] +
  \left(\left(\frac{\delta_{01} + \delta_{11}}{\pi_{01} + \pi_{11}}\right) - 
  \frac{\delta_{11}}{\pi_{11}}\right) c_2 E[g \mid X, Y_2]
\end{align}

\subsection*{Projection onto Nuisance Tangent Space}

The goal is now to find the coefficients $c_0, c_1$, and $c_2$ such that 
$\langle \hat \theta_c, L_2\rangle \equiv E[\hat \theta_c L_2] = 0$ for 
all $L_2 \in \Lambda_2$ (see \cite{tsiatis2006semiparametric} for definition 
of $\Lambda_2$). If we can find such coefficients that the estimator $\hat
\theta_c$ will be orthogonal to $\Lambda_2$ and hence by Theorem 10.1 of 
\cite{tsiatis2006semiparametric} semiparametrically optimal. The good news 
is that we know (CITEME) that any element $L_2 \in \Lambda_2$ has a form:

\begin{align}
  L_2 &= 
  \left(\frac{\delta_{11}}{\pi_{11}}\pi_{00} - \delta_{00}\right) L_{20}(X) +
  \left(\frac{\delta_{11}}{\pi_{11}}\pi_{10} - \delta_{10}\right) L_{21}(X, Y_1) +
  \left(\frac{\delta_{11}}{\pi_{11}}\pi_{01} - \delta_{01}\right) L_{22}(X, Y_2).
\end{align}

Then expanding and solving $E[\hat \theta_c L_2] = 0$ yields:

{\tiny
\begin{align*}
  0 &= E[\hat \theta_c L_2] \\ 
    &E\left[\left(\frac{\pi_{00}}{\pi_{11}} + \left(\frac{\pi_{10}}{\pi_{10} + \pi_{11}}\right)\frac{\pi_{00}c_1}{\pi_{11}} +
      \left(\frac{\pi_{01}}{\pi_{01} + \pi_{11}}\right) \frac{\pi_{00}c_2}{\pi_{11}} + \frac{\pi_{00}(\pi_{10}\pi_{01} -
  \pi_{11}^2)c_0}{(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})\pi_{11}}\right) E[g \mid X] L_{20}(X)\right. \\ 
    &+ \frac{\pi_{10}}{\pi_{11}} \left(E[g(Z) L_{21}(X, Y_1) \mid X] - c_1 E[E[g(Z) \mid X, Y_1] L_{21}(X, Y_1) \mid X] +
      \frac{\pi_{01}c_2}{\pi_{10} + \pi_{11}} E[E[g \mid X, Y_2] L_{21}(X, Y_1) \mid X] +
      \frac{\pi_{10} \pi_{01}}{\pi_{11}(\pi_{01} + \pi_{11})} E[g \mid X]
      E[L_{21}(X, Y_1) \mid X]c_0\right)\\
    &+ \left.\frac{\pi_{01}}{\pi_{11}} \left(E[g(Z) L_{22}(X, Y_2) \mid X] + \frac{\pi_{10} c_1}{\pi_{10} + \pi_{11}} E[E[g(Z) \mid X, Y_1] L_{22}(X, Y_2) \mid X] -
      c_2 E[E[g \mid X, Y_2] L_{22}(X, Y_2) \mid X] +
      \frac{\pi_{10}}{(\pi_{01} + \pi_{11})} E[g \mid X]
      E[L_{22}(X, Y_2) \mid X]c_0\right)\right]
\end{align*}
}

To solve for $c_0, c_1$, and $c_2$ we need the following to hold for any
$L_{21}(X, Y_1)$ and $L_{22}(X, Y_2)$:

{\small
\begin{align*}
  &1 + c_0 \frac{\pi_{01} \pi_{10} - \pi_{11}^2}{(\pi_{10} + \pi_{11})(\pi_{01}
  + \pi_{11})} + c_1 \frac{\pi_{10}}{\pi_{01} + \pi_{11}} + c_2
  \frac{\pi_{01}}{\pi_{01} + \pi_{11}} = 0\\ 
  &E\left[\left(g(Z) + c_0 \frac{\pi_{01}}{\pi_{01} + \pi_{11}}E[g(Z) \mid X] - c_1 E[g(Z) \mid X, Y_1] + c_2 \frac{\pi_{01}}{\pi_{10} + \pi_{11}}E[g(Z) \mid X, Y_2]\right)L_{21}(X, Y_1) \mid X\right] = 0 \\ 
  &E\left[\left(g(Z) + c_0 \frac{\pi_{10}}{\pi_{10} + \pi_{11}}E[g(Z) \mid X] + c_1 \frac{\pi_{10}}{\pi_{10} + \pi_{11}} E[g(Z) \mid X, Y_1] - c_2 E[g(Z) \mid X, Y_2]\right)L_{22}(X, Y_2) \mid X\right] = 0
\end{align*}
}

\subsection*{Optimal Normal Model}

An alternative optimization problem is to find coefficients $c_0$, $c_1$, and
$c_2$, that minimize the variance of $\hat \theta_c$. After expanding the 
variance and differentiating by each $c_i$, the optimal parameters are

\begin{align*}
  \begin{bmatrix} \hat c_0 \\ \hat c_1 \\ \hat c_2 \end{bmatrix} 
  &= -
  \begin{bmatrix} 
    M_{11} & M_{12} & M_{13} \\
    M_{21} & M_{22} & M_{23} \\
    M_{31} & M_{32} & M_{33} \\
  \end{bmatrix}^{-1}
  \times 
  \begin{bmatrix}
    E[E[g \mid X]^2] \left(1 + \frac{\pi_{10}\pi_{01} - \pi_{11}^2}{\pi_{11}(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})}\right) \\
    E[E[g \mid X, Y_1]^2] \left(\frac{-\pi_{10}}{\pi_{11}(\pi_{10} + \pi_{11})}\right) \\
    E[E[g \mid X, Y_2]^2] \left(\frac{-\pi_{01}}{\pi_{11}(\pi_{01} + \pi_{11})}\right) \\
  \end{bmatrix}
\end{align*}

where

\begin{align*}
  M_{11} &= E[E[g \mid X]^2] \left(\frac{\pi_{11}^2 + \pi_{10}\pi_{01}}{\pi_{11}(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})} - 1\right) \\
  M_{12} &= E[E[g \mid X]^2] \left(\frac{-\pi_{10}\pi_{01}}{\pi_{11}(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})}\right) \\
  M_{13} &= E[E[g \mid X]^2] \left(\frac{-\pi_{10}\pi_{01}}{\pi_{11}(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})}\right) \\
  M_{22} &= E[V(E[g \mid X, Y_1] \mid X)] \left(\frac{\pi_{10}}{\pi_{11}(\pi_{10} + \pi_{11})}\right) \\
  M_{23} &= E[E[g \mid X, Y_1]E[g \mid X, Y_2]] \left(\frac{\pi_{10}\pi_{01}}{\pi_{11}(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})}\right) \\
  M_{33} &= E[V(E[g \mid X, Y_2] \mid X)] \left(\frac{\pi_{01}}{\pi_{11}(\pi_{01} + \pi_{11})}\right)
\end{align*}

\newpage

\printbibliography

\end{document}

### 2023-11-07 Notes/optest_update.tex
Added simulations with optimal delta estimator.

@@
 
-\[ \frac{I(C = \infty) g(Z)}{\varpi(\infty, Z)} + \frac{I(C =
+\begin{equation}\label{lambda2}
+ \frac{I(C = \infty) g(Z)}{\varpi(\infty, Z)} + \frac{I(C =
   \infty)}{\varpi(\infty, Z)} \left(\sum_{r \neq \infty} \varpi(r, G_r(Z))
-  L_{2r}(G_r(Z))\right) - \sum_{r \neq \infty} I(C = \infty)L_{2r}(G_r(Z))\]
+  L_{2r}(G_r(Z))\right) - \sum_{r \neq \infty} I(C = r)L_{2r}(G_r(Z))
+\end{equation}
 
@@
 
-It turns out that $\hat \theta_{prop}$ from Equation~\ref{eq:prop} is contained within 
-the previous class of estimators with
-
-\[L_{2r}(G_r(Z)) = \varpi(r, G_r(Z)) E[g(Z) \mid G_r(Z)].\]
-
-(See pages 81-82 of notes for proof.) \\
+%It turns out that $\hat \theta_{prop}$ from Equation~\ref{eq:prop} is contained within 
+%the previous class of estimators with 
+%
+%\[L_{2r}(G_r(Z)) = \varpi(r, G_r(Z)) E[g(Z) \mid G_r(Z)].\]
+%
+%(See pages 81-82 of notes for proof.) \\
 
@@
 \cite{tsiatis2006semiparametric} semiparametrically optimal. The good news 
-is that we know (CITEME) that any element $L_2 \in \Lambda_2$ has a form:
+is that we know (from Theorem 7.2) that any element $L_2 \in \Lambda_2$ has a
+form:
 
@@
 
-\subsection*{Optimal Normal Model}
+\subsection*{Optimal Model}
 
-An alternative optimization problem is to find coefficients $c_0$, $c_1$, and
-$c_2$, that minimize the variance of $\hat \theta_c$. After expanding the 
-variance and differentiating by each $c_i$, the optimal parameters are
+Due to the fact that I do not know how to solve the previous equations, there 
+is another way to construct an optimal model. We can find the values of 
+$c_0$, $c_1$, and $c_2$ in $\hat \theta_c$ that minimize the variance the 
+estimator. We can find these values by differentiating by $c_i$ and solving 
+for $c_i$:
 
@@
 
+This estimator is similar to several other estimators:
+
+\begin{align}\label{thetaind}
+  \hat \theta^{ind}_c 
+  &= \frac{\delta_{11}}{\pi_{11}} g(Z) + \left(1 -
+    \left(\frac{\delta_{10}}{\pi_{10}}\right) -
+    \left(\frac{\delta_{01}}{\pi_{01}}\right) + 
+  \frac{\delta_{11}}{\pi_{11}}\right) c_0 E[g \mid X] \\ \nonumber
+  &+
+  \left(\frac{\delta_{10}}{\pi_{10}} - 
+  \frac{\delta_{11}}{\pi_{11}}\right) c_1 E[g \mid X, Y_1] +
+  \left(\frac{\delta_{01}}{\pi_{01}} - 
+  \frac{\delta_{11}}{\pi_{11}}\right) c_2 E[g \mid X, Y_2]
+\end{align}
+
+\begin{align}\label{thetadelta}
+  \hat \theta^{\delta}_c 
+  &= \frac{\delta_{11}}{\pi_{11}} g(Z) 
+  + \left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{00}}{\pi_{00}}\right) 
+   c_0 E[g \mid X] \\ \nonumber
+  &+
+  \left(\frac{\delta_{11}}{\pi_{11}} - 
+  \frac{\delta_{10}}{\pi_{10}}\right) c_1 E[g \mid X, Y_1] +
+  \left(\frac{\delta_{11}}{\pi_{11}} - 
+  \frac{\delta_{01}}{\pi_{01}}\right) c_2 E[g \mid X, Y_2]
+\end{align}
+
+The first expression (Equation~\ref{thetaind}) is the proposed estimator 
+with independent differences in each segment, while the second expression 
+(Equation~\ref{thetadelta}) is the optimal estimator with values of $\delta$
+such that $\hat \theta^{\delta}_c \in \Lambda_2$, which means that it has 
+the form of the semiparametric in Equation~\ref{lambda2}.
+
 \newpage

### 2023-11-07 R/opt_est.R

@@ -73,7 +73,11 @@ gen_optsim_data <- function(n, theta = 0, cor_xe1 = 0, cor_xe2 = 0, cor_e1e2 = 0
 
   tibble(X = x_vec, Y1 = y1, Y2 = y2,
          delta_1 = delta_y1, delta_2 = delta_y2,
-         prob_11 = prob_11, prob_1 = prob_1, prob_2 = prob_2)
+         prob_11 = prob_11, prob_1 = prob_1, prob_2 = prob_2,
+         prob_10 = 0.2, prob_01 = 0.2, prob_00 = 0.2,
+         delta_11 = delta_y1 * delta_y2, delta_10 = delta_y1 * (1 - delta_y2),
+         delta_01 = (1 - delta_y1) * delta_y2,
+         delta_00 = (1 - delta_y1) * (1 - delta_y2))
 
 }
 
@@ -121,7 +125,6 @@ get_cond_exp <- function(obs_y2, mean_y1, mean_y2, sigma_mat) {
   return(t(out))
 
 }
-
 
 # The EM algorithm consists of two steps: expectation and maximization.
 # The E-Step consists of:
@@ -323,7 +326,7 @@ get_v_powers <- function(str_vec, pow = 1) {
 #'  - df has the following columns: X, Y1, Y2, delta_y1, delta_y2, prob_vec
 #'  - If delta_yi = 1 then we assume that Y_i is observed if delta_yi = 0 then
 #'  - we assume that Y_i is missing.
-prop_nmono_est <- function(df, gfun = "Y2", oracle = TRUE, pow = 1) {
+prop_nmono_est <- function(df, gfun = "Y2", oracle = TRUE, pow = 1, prop_ind = FALSE) {
 
   df <- mutate(df, g_i = eval(rlang::parse_expr(gfun)))
   df_11 <- filter(df, delta_1 == 1, delta_2 == 1)
@@ -343,24 +346,40 @@ prop_nmono_est <- function(df, gfun = "Y2", oracle = TRUE, pow = 1) {
   pi2_w <- df$prob_2
   pi11_w <- df$prob_11
 
+  pi_11 <- unique(df$prob_11)
+  pi_10 <- unique(df$prob_10)
+  pi_01 <- unique(df$prob_01)
+  pi_00 <- unique(df$prob_00)
+
   # 2. Compute expectations.
[…]

### 2023-11-07 Simulations/T2Optimal_Nonmonotone_Est/comp_ests_g.R

@@ -65,6 +65,7 @@ mc_theta <-
     wlstt_est <- 
       comb_lin_est(df, gfun = "Y1^2 * Y2", cov_e1e2 = cor_e1e2, theta2 = true_theta)
     prop_est <- prop_nmono_est(df, gfun = "Y1^2 * Y2", pow = 3)
+    semiopt_est <- opt_semi_est(df, gfun = "Y1^2 * Y2")
 
     return(tibble(oracle = oracle_est,
                   oraclex = oraclex_est,
@@ -72,7 +73,8 @@ mc_theta <-
                   ipw = ipw_est,
                   wls = wls_est,
                   wlstt = wlstt_est,
-                  prop = prop_est))
+                  prop = prop_est,
+                  semiopt = semiopt_est))
   } |> 
   bind_rows()
 
@@ -92,6 +94,7 @@ mc_theta |>
     bias_wls = mean(wls) - true_g,
     bias_wlstt = mean(wlstt) - true_g,
     bias_prop = mean(prop) - true_g,
+    bias_semiopt = mean(semiopt) - true_g,
     sd_oracle = sd(oracle),
     sd_oraclex = sd(oraclex),
     sd_cc = sd(cc),
@@ -99,19 +102,21 @@ mc_theta |>
     sd_wls = sd(wls),
     sd_wlstt = sd(wlstt),
     sd_prop = sd(prop),
+    sd_semiopt = sd(semiopt),
     tstat_oracle = (mean(oracle) - true_g) / sqrt(var(oracle) / B),
     tstat_oraclex = (mean(oraclex) - true_g) / sqrt(var(oraclex) / B),
     tstat_cc = (mean(cc) - true_g) / sqrt(var(cc) / B),
     tstat_ipw = (mean(ipw) - true_g) / sqrt(var(ipw) / B),
     tstat_wls = (mean(wls) - true_g) / sqrt(var(wls) / B),
     tstat_wlstt = (mean(wlstt) - true_g) / sqrt(var(wlstt) / B),
-    tstat_prop = (mean(prop) - true_g) / sqrt(var(prop) / B)
+    tstat_prop = (mean(prop) - true_g) / sqrt(var(prop) / B),
+    tstat_semiopt = (mean(semiopt) - true_g) / sqrt(var(semiopt) / B)
   ) |>
   tidyr::pivot_longer(cols = everything(),
                names_to = c(".value", "algorithm"),
                names_pattern = "(.*)_(.*)") |>
[…]

### 2023-11-07 Simulations/T2Optimal_Nonmonotone_Est/optimal_simulation.R

@@ -32,7 +32,7 @@ registerDoParallel(clust)
 
 B <- 3000
 n_obs <- 1000
-true_theta <- 5
+true_theta <- 5 
 cor_e1e2 <- 0.5
 
 mc_theta <-
@@ -58,7 +58,11 @@ mc_theta <-
     # em_est <- em_optsim(df)
     wls_est <- opt_lin_est(df, cov_y1y2 = cor_e1e2)
     prop_est <- prop_nmono_est(df)
-    prop2_est <- prop2_nmono_est(df)
+    propind_est <- prop_nmono_est(df, prop_ind = TRUE)
+    semiopt_est <- opt_semi_est(df)
+    semidef_est <- opt_semi_est(df, est = "default")
+    semidel_est <- opt_delta_c(df)
+    semideldef_est <- opt_delta_c(df, est = "default")
 
     return(tibble(oracle = oracle_est,
                   oraclex = oraclex_est,
@@ -67,8 +71,14 @@ mc_theta <-
                   # em = em_est,
                   wls = wls_est,
                   prop = prop_est,
-                  prop2 = prop2_est))
-  } |> bind_rows()
+                  propind = propind_est,
+                  semiopt = semiopt_est,
+                  semidef = semidef_est,
+                  semidel = semidel_est,
+                  semideldef = semideldef_est
+    ))
+  } |>
+  bind_rows()
 
 stopCluster(clust)
 
@@ -84,27 +94,39 @@ mc_theta |>
     bias_ipw = mean(ipw) - true_theta,
     bias_wls = mean(wls) - true_theta,
     bias_prop = mean(prop) - true_theta,
-    bias_prop2 = mean(prop2) - true_theta,
+    bias_propind = mean(propind) - true_theta,
+    bias_semiopt = mean(semiopt) - true_theta,
+    bias_semidef = mean(semidef) - true_theta,
+    bias_semidel = mean(semidel) - true_theta,
+    bias_semideldef = mean(semideldef) - true_theta,
     sd_oracle = sd(oracle),
     sd_oraclex = sd(oraclex),
     sd_cc = sd(cc),
     sd_ipw = sd(ipw),
     sd_wls = sd(wls),
     sd_prop = sd(prop),
-    sd_prop2 = sd(prop2),
[…]

### 2023-11-07 Simulations/T2Optimal_Nonmonotone_Est/sim_testing.R

@@ -1,0 +1,202 @@
+# Title: sim_testing.R
+# Created by: Caleb Leedy
+# Created on: November 01, 2023
+# Purpose: This script tests the different components of the variance estimate 
+# of \hat \theta_c. We want to ensure that each component is estimated 
+# correctly.
+
+# ***********
+# * Outline *
+# ***********
+# 1. Load Libraries and code.
+# 2. Run Monte Carlo simulation for each variance component
+
+# *************
+# * Libraries *
+# *************
+
+library(dplyr)
+library(stringr)
+
+library(doParallel)
+library(doRNG)
+library(parallelly)
+
+source("R/opt_est.R")
+
+# ***************************
+# * Monte Carlo Simulations *
+# ***************************
+
+clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
+registerDoParallel(clust)
+
+B <- 5000
+n_obs <- 1000
+true_theta <- 5
+cor_e1e2 <- 0.5
+
+mc_ests <-
+  foreach(iter = 1:B,
+          .options.RNG = 1,
+          .packages = c("dplyr", "stringr")) %dorng% {
+
+    # Generate Data
+    df <- gen_optsim_data(n = n_obs, theta = true_theta, cor_e1e2 = cor_e1e2)
+
+    # Estimate different parts of theta_hat_c
+    # semidef <- opt_semi_est(df, est = "opt", test = TRUE)
+    semidef <- opt_delta_c(df, est = "opt", test = TRUE)
+
+    return(tibble(est = semidef$est,
+                  ipw = semidef$ipw,
+                  A0 = semidef$A0,
+                  A1 = semidef$A1,
+                  A2 = semidef$A2,
+                  ipwA0 = semidef$ipwA0,
+                  ipwA1 = semidef$ipwA1,
+                  ipwA2 = semidef$ipwA2,
+                  A0A1 = semidef$A0A1,
[…]

### 2023-11-07 rsync_with_nova.sh

@@ -1,0 +1,11 @@
+#!/bin/bash
+
+# File: rsync_with_nova.sh
+# Created by: Caleb Leedy
+# Created on: October 25, 2023
+# Updated on: October 25, 2023
+# 
+# Purpose:
+# This file syncs my local repository with my repository on millipede.
+rsync -rtvz R/ cleedy@nova.its.iastate.edu:~/Data_Integration/R
+

### 2023-11-07 Notes/optest_update.tex
Added simulation 2.

@@
 Suppose that we want to estimate $\theta = E[g(X, Y_1, Y_2)]$ for a known
-function $g$. The proposed estimator is
-
-\begin{align}\label{eq:prop}
-  &\hat \theta_{prop} =\\ \nonumber
-  &n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] + n^{-1} \sum_{i = 1}^n
-  \frac{\delta_{10}}{\pi_{10}} (E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i]) 
-  + n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} (E[g_i \mid X_i, Y_{2i}] - E[g_i \mid X_i]) \\ \nonumber
-  &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} (g_i - E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i, Y_{2i}] + E[g_i \mid X_i]).
-\end{align}
-
+function $g$. 
 The goal is the show that over all functions $b_1(X, Y_1)$ and $b_2(X, Y_2)$,
-$\hat \theta_{prop}$ is the optimal estimator in the form:
+we have the optimal estimator in the form:
 
@@
 \end{align}
+
+%The proposed estimator is 
+%
+%\begin{align}\label{eq:propind}
+%  &\hat \theta^{ind}_{prop} =\\ \nonumber
+%  &n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] + n^{-1} \sum_{i = 1}^n
+%  \frac{\delta_{10}}{\pi_{10}} (E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i]) 
+%  + n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} (E[g_i \mid X_i, Y_{2i}] - E[g_i \mid X_i]) \\ \nonumber
+%  &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} (g_i - E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i, Y_{2i}] + E[g_i \mid X_i]).
+%\end{align}
 
@@
 
-\[L_{2r}(G_r(Z)) = E[g(Z) \mid G_r(Z)].\]
+\[L_{2r}(G_r(Z)) = c_r E[g(Z) \mid G_r(Z)]\]
+
+where each $c_r$ is a constant. This actually suggests a variety of estimators 
+that each have different values of coefficients of $\delta$. Initially, we tried 
+a direct projection onto the nuisance tangent space, but I got stuck. See Appendix A 
+for this work.
+
+\subsection*{Different Classes of Estimators}
+
+All of the considered estimators have a similar form:
+
+\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
+\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
+\mid X, Y_2].\]
+
+In this case, $c$ is a vector of constants $c = (c_0, c_1, c_2)$ that are 
+chosen to minimize the variance of $\hat \theta$. See Appendix B for details on 
+an example about choosing the values of $\hat c$.
+
+\begin{tabular}{lrrr}
+  \toprule
+  Estimator & $\beta_0(\delta, c_0)$ & $\beta_1(\delta, c_1)$ & Implemented \\
+  \midrule
+  $\hat \theta_{prop}$ & $\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - 
+  \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
+  $\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)$ & \checkmark \\
+  $\hat \theta^{ind}_{prop}$ & $\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - 
+  \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
+  $\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)$ & \checkmark \\
+  $\hat \theta_{c}$ & $c_0\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - 
+  \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
+  $c_1\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)$ & \\
+  $\hat \theta_{c, ind}$ & $c_0\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - 
+  \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
+  $c_1\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)$ & \checkmark \\
+  $\hat \theta_{\delta}$ & $c_0\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{00}}{\pi_{00}}\right)$ &
+  $c_1\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{10}}{\pi_{10}}\right)$ & \checkmark \\
+  \bottomrule
+\end{tabular}
 
@@
 
-\subsection*{Linear Expectations}
+\section*{Simulation Studies}
+
+\subsection*{Simulation 1}
+
+To see which of these estimators is the best, we run a simulation study. In this 
+study, we have the following variables:
+
+\begin{align*}
+  \begin{bmatrix} x \\ e_1 \\ e_2 \end{bmatrix} 
+  &\stackrel{ind}{\sim} N\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
+  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}\right) \\
+  y_1 &= x + e_1 \\
+  y_2 &= \theta + x + e_2 \\
+\end{align*}
+
+Furthermore, $\pi_{11} = 0.4$ and $\pi_{00} = \pi_{10} = \pi_{01} = 0.2$.
+The goal of this simulation study is to find $\theta = E[Y_2]$. In other words,
+$g(Z) = Y_2$. There are several algorithms for comparison which are defined 
+as the following:
+
+\begin{align*}
+  Oracle &= n^{-1} \sum_{i = 1}^n g(Z_i)\\
+  CC &= \frac{\sum_{i = 1}^n \delta_{11} g(Z_i)}{\sum_{i = 1}^n \delta_{11}} \\
+  IPW &= \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g(Z_i)\\
+\end{align*}
+
+\begin{table}[ht!]
+\caption{The true value of $\theta$ is 5. $\rho = 0.5$. The test statistic and 
+p-value are from a one-sample t-test to see if the estimator is biased.}
+\centering
+\begin{tabular}[t]{lrrrr}
+\toprule
+Algorithm & Bias & SD & Tstat & Pval\\
+\midrule
+Oracle & 0.001 & 0.044 & 1.546 & 0.061\\
+CC & 0.002 & 0.058 & 1.549 & 0.061\\
+IPW & 0.006 & 0.210 & 1.504 & 0.066\\
+$\hat \theta_{prop}$ & 0.002 & 0.051 & 1.911 & 0.028\\
+$\hat \theta_{prop}^{ind}$ & 0.002 & 0.091 & 0.983 & 0.163\\
+$\hat \theta_{c, ind}$ & 0.001 & 0.074 & 0.784 & 0.216\\
+$\hat \theta_{\delta}$ & 0.002 & 0.051 & 1.964 & 0.025\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+The main takeaway from this simulation study is that $\hat \theta_\delta$
+and $\hat\theta_{prop}$ have the smallest standard deviation among the 
+non-oracle estimators.
+
+\subsection*{Simulation 2}
+We now test a second simulation study. The setup of this simulation study 
+is the same as the previous simulation. The only difference is that we 
+are interested in estimating $\theta = E[Y_1^2 Y_2]$. In other words, now 
+we have $g(Z) = Y_1^2 Y_2$.
+
+\begin{table}[ht!]
+  \caption{The true value of $E[g(Z)]$ is 10. $\rho = 0.5$. The test statistic and 
+p-value are from a one-sample t-test to see if the estimator is biased.}
+\centering
+\begin{tabular}[t]{lrrrr}
+\toprule
+Algorithm & Bias & SD & Tstat & Pval\\
+\midrule
+Oracle & 0.007 & 0.529 & 0.741 & 0.229\\
+CC & 0.023 & 0.824 & 1.518 & 0.065\\
+IPW & 0.031 & 0.915 & 1.846 & 0.032\\
+$\hat \theta_{prop}$ & 0.011 & 0.635 & 0.912 & 0.181\\
+$\hat \theta_{c, ind}$ & 0.013 & 0.662 & 1.060 & 0.145\\
+$\hat \theta_{\delta}$ & 0.011 & 0.636 & 0.983 & 0.163\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Like the first simulation, $\hat \theta_{prop}$ and $\hat \theta_{\delta}$ 
+performed the best.
+
+\newpage
+
+\section*{Next Steps}
+
+\begin{itemize}
+  \item[1.] It turns out that \cite{tsiatis2006semiparametric} contains 
+    a general technique for solving non-monotone missing data. They show 
+    that the optimal estimator is found via the following double projection:
+
+    \[L_2 = \Pi\left(\Pi\left(\frac{\delta_{11}}{\pi_{11}}g(Z) \mid C,
+    G_C(Z)\right) \mid Z\right)\]
+
+    I can show that $\hat \theta_\delta$ is NOT such a double projection. 
+    However, I think it is worth exploring what such an estimator (especially 
+    if we consider a linear estimator) looks like.
+  \item[2.] Similar to the previous point, \cite{tsiatis2006semiparametric} 
+    discuss what the optimal estimator of a non-monotone problem looks like 
+    and they determine that it solves the estimating equation
+
+    \[\sum_{i = 1}^n \mathcal{L}(\mathcal{M}^{-1}(g(Z)))\]
+
+    where $\mathcal{M}(g(Z)) = E[\mathcal{L}(g(Z)) \mid Z]$ and 
+    $\mathcal{L}(g(Z)) = E[g(Z) \mid C, G_C(Z)]$. The difficult part is 
+    understanding $\mathcal{M}^{-1}$, but there is conveniently the fact that
+
+    \[\mathcal{M}^{-1}(g(Z)) = \lim_{n \to \infty} \phi_{n + 1}(Z) \text{ where } 
+      \phi_{n + 1}(Z) = (I - \mathcal{M}) \phi_n(Z) + g(Z) \text{ and }
+    \phi_0(Z) = g(Z).\]
+
+    I have been investigating what this estimating equation yields if we 
+    approximate $\mathcal{M}^{-1}$ with $\phi_1$. However, I have not 
+    finished this yet.
+  \item[3.] A different method to find the projection onto $\Lambda_2$ 
+    could be to try to minimize the (KL) divergence between 
+    $L_2$ and $\frac{\delta_{11}}{\pi_{11}}g(Z)$.
+\end{itemize}
+
+\newpage
+
+\printbibliography
+
+\newpage
+
+\appendix
+
+\section{Linear Expectations}
 
@@
 
-\subsection*{Projection onto Nuisance Tangent Space}
+\subsection{Projection onto Nuisance Tangent Space}
 
@@
 
-\subsection*{Optimal Model}
+Unfortunately, I am now stuck because it is unclear to me how to use any actual 
+data to solve this problem.
 
-Due to the fact that I do not know how to solve the previous equations, there 
-is another way to construct an optimal model. We can find the values of 
+\section{Solving for $\hat c$}
+
+We can find the values of 
 $c_0$, $c_1$, and $c_2$ in $\hat \theta_c$ that minimize the variance the 
@@
 
-\newpage
-
-\printbibliography
-
 \end{document}

### 2023-11-07 R/opt_est.R

@@ -474,7 +474,7 @@ comb_lin_est <- function(df, gfun = "Y1^2 * Y2",
 # * Optimal Semiparametric Estimator *
 # ************************************
 
-opt_semi_est <- function(df, gfun = "Y2", est = "opt", test = FALSE) {
+opt_semi_est <- function(df, gfun = "Y2", est = "opt", test = FALSE, pow = 1) {
 
   # Steps: 
   # 1. Compute E[g_i | X_i]
@@ -489,17 +489,37 @@ opt_semi_est <- function(df, gfun = "Y2", est = "opt", test = FALSE) {
   df_01 <- filter(df, delta_1 == 0, delta_2 == 1)
   df_00 <- filter(df, delta_1 == 0, delta_2 == 0)
 
-  # 1. Compute E[g_i | X_i]
-  mod_gx <- lm(g_i ~ X, data = df_11)
+  if (pow == 1) {
+    # 1. Compute E[g_i | X_i]
+    mod_gx <- lm(g_i ~ X, data = df_11)
 
-  # 2. Compute E[E[g | X]^2], E[E[g | X, Y1]^2], E[E[g | X, Y2]^2], 
-  #   E[E[g | X, Y1]E[g | X, Y2]]
-  exp_gx2 <- mean(predict(mod_gx)^2)
-  mod_gy1 <- lm(g_i ~ X + Y1, data = df_11)
-  exp_gy12 <- mean(predict(mod_gy1)^2)
-  mod_gy2 <- lm(g_i ~ X + Y2, data = df_11)
-  exp_gy22 <- mean(predict(mod_gy2)^2)
-  exp_gy1y2 <- mean(predict(mod_gy1) * predict(mod_gy2))
+    # 2. Compute E[E[g | X]^2], E[E[g | X, Y1]^2], E[E[g | X, Y2]^2], 
+    #   E[E[g | X, Y1]E[g | X, Y2]]
+    exp_gx2 <- mean(predict(mod_gx)^2)
+    mod_gy1 <- lm(g_i ~ X + Y1, data = df_11)
+    exp_gy12 <- mean(predict(mod_gy1)^2)
+    mod_gy2 <- lm(g_i ~ X + Y2, data = df_11)
+    exp_gy22 <- mean(predict(mod_gy2)^2)
+    exp_gy1y2 <- mean(predict(mod_gy1) * predict(mod_gy2))
[…]

### 2023-11-07 Simulations/T2Optimal_Nonmonotone_Est/comp_ests_g.R

@@ -65,7 +65,10 @@ mc_theta <-
     wlstt_est <- 
       comb_lin_est(df, gfun = "Y1^2 * Y2", cov_e1e2 = cor_e1e2, theta2 = true_theta)
     prop_est <- prop_nmono_est(df, gfun = "Y1^2 * Y2", pow = 3)
-    semiopt_est <- opt_semi_est(df, gfun = "Y1^2 * Y2")
+    semiopt_est <- opt_semi_est(df, gfun = "Y1^2 * Y2", pow = 3)
+    semidef_est <- opt_semi_est(df, gfun = "Y1^2 * Y2", est = "default", pow = 3)
+    semidel_est <- opt_delta_c(df, gfun = "Y1^2 * Y2", pow = 3)
+    semideldef_est <- opt_delta_c(df, gfun = "Y1^2 * Y2", est = "default", pow = 3)
 
     return(tibble(oracle = oracle_est,
                   oraclex = oraclex_est,
@@ -74,7 +77,10 @@ mc_theta <-
                   wls = wls_est,
                   wlstt = wlstt_est,
                   prop = prop_est,
-                  semiopt = semiopt_est))
+                  semiopt = semiopt_est,
+                  semidef = semidef_est,
+                  semidel = semidel_est,
+                  semideldef = semideldef_est))
   } |> 
   bind_rows()
 
@@ -95,6 +101,9 @@ mc_theta |>
     bias_wlstt = mean(wlstt) - true_g,
     bias_prop = mean(prop) - true_g,
     bias_semiopt = mean(semiopt) - true_g,
+    bias_semidef = mean(semidef) - true_g,
+    bias_semidel = mean(semidel) - true_g,
+    bias_semideldef = mean(semideldef) - true_g,
     sd_oracle = sd(oracle),
     sd_oraclex = sd(oraclex),
     sd_cc = sd(cc),
@@ -103,6 +112,9 @@ mc_theta |>
     sd_wlstt = sd(wlstt),
     sd_prop = sd(prop),
     sd_semiopt = sd(semiopt),
+    sd_semidef = sd(semidef),
+    sd_semidel = sd(semidel),
+    sd_semideldef = sd(semideldef),
     tstat_oracle = (mean(oracle) - true_g) / sqrt(var(oracle) / B),
     tstat_oraclex = (mean(oraclex) - true_g) / sqrt(var(oraclex) / B),
[…]

### 2023-11-07 Notes/optest_update.tex
Updated description of estimators in update.

@@
 
+I wanted to give a brief comment on the origin of each of these estimators. The 
+first estimator $\hat \theta_{prop}$ is the estimator proposed by Dr. Kim. The 
+original goal was to show that this is optimal. The second estimator, 
+$\hat \theta^{ind}_{prop}$ is the first estimator with different functions of 
+$\delta$ to try to make each component more independent from each other. The 
+problem with this is that the components are still dependent and it has worse 
+properties in simulation studies (next section). The estimator $\hat \theta_c$ 
+is the originial estimator $\hat \theta_{prop}$ with different coefficients that
+hopefully make it slightly more efficient. The estimator $\hat \theta_{c, ind}$ 
+does the same thing with $\hat \theta_{prop}^{ind}$. Finally, 
+$\hat \theta_{\delta}$ uses the functional form of Theorem 7.2 from 
+\cite{tsiatis2006semiparametric} and optimized coefficients.
+
 %It turns out that $\hat \theta_{prop}$ from Equation~\ref{eq:prop} is contained within 
@@
 performed the best.
-
-\newpage
 

### 2023-12-11 Notes/optest_update.tex
I understand why the linear model is not optimal in the non-linear case. The weights are higher for smaller estimates with a smaller variance.

@@
   \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
-  $c_1\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)$ & \\
+  $c_1\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)$ & \checkmark \\
   $\hat \theta_{c, ind}$ & $c_0\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - 
@@
 $\hat \theta_{prop}$ & 0.002 & 0.051 & 1.911 & 0.028\\
+$\hat \theta_{c}$ & 0.002 & 0.051 & 1.920 & 0.027\\
 $\hat \theta_{prop}^{ind}$ & 0.002 & 0.091 & 0.983 & 0.163\\
@@
 
-The main takeaway from this simulation study is that $\hat \theta_\delta$
+The main takeaway from this simulation study is that $\hat \theta_\delta$,
+$\hat \theta_c$
 and $\hat\theta_{prop}$ have the smallest standard deviation among the 
-non-oracle estimators.
+non-oracle estimators. In fact $\hat \theta_c$ and $\hat \theta_{prop}$ are 
+virtually identical because the optimal values of $c$ in $\hat \theta_c$ 
+are $(c_0, c_1, c_2)' \approx (1, 1, 1)'$.
 
 \subsection*{Simulation 2}
+
 We now test a second simulation study. The setup of this simulation study 
@@
 $\hat \theta_{prop}$ & 0.011 & 0.635 & 0.912 & 0.181\\
+$\hat \theta_{c}$ & 0.012 & 0.632 & 1.010 & 0.156\\
 $\hat \theta_{c, ind}$ & 0.013 & 0.662 & 1.060 & 0.145\\
@@
 
-Like the first simulation, $\hat \theta_{prop}$ and $\hat \theta_{\delta}$ 
-performed the best.
+Like the first simulation, $\hat \theta_{prop}$, $\hat \theta_c$
+and $\hat \theta_{\delta}$ performed the best.
 

### 2023-12-11 Notes/prob_wlinmod.tex
I understand why the linear model is not optimal in the non-linear case. The weights are higher for smaller estimates with a smaller variance.

\documentclass[12pt]{article}

\usepackage{amsmath, amssymb, mathrsfs, fancyhdr}
\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
\usepackage{biblatex}
\usepackage{booktabs}
\usepackage{float}

\addbibresource{references.bib}

\hypersetup{colorlinks = true, urlcolor = black}

\headheight     15pt
\topmargin      -1.5cm   % read Lamport p.163
\oddsidemargin  -0.04cm  % read Lamport p.163
\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
\textwidth      16.59cm
\textheight     23.94cm
\parskip         7.2pt   % sets spacing between paragraphs
\parindent         0pt   % sets leading space for paragraphs
\pagestyle{empty}        % Uncomment if don't want page numbers
\pagestyle{fancyplain}

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}

\renewcommand{\arraystretch}{1.2}

\begin{document}

\lhead{Caleb Leedy}
\chead{Non-monotone Missingness}
%\chead{STAT 615 - Advanced Bayesian Methods}
%\rhead{Page \thepage\ of \pageref{LastPage}}
\rhead{\today}

\section*{The Problem with the Weighted Linear Model}

Besides the semiparametric model, we have been comparing our results to an 
optimal weighted linear model. This model can be summarized as the following:

\begin{itemize}
  \item[1.] For each segment $A_k$ with a distinct missing pattern, let 
    $G_k(Z)$ be the observed data in segment $A_k$.
  \item[2.] For each $A_k$, estimate $\theta = E[g(X, Y_1, Y_2)]$ as 
    $\hat \theta_k = E[g \mid G_k(Z)]$.
  \item[3.] Combine these estimates using the inverse variance weighted average
    of each $\hat \theta_k$:

    \[\hat \theta = \sum_{k} w_k \hat \theta_k \text{ where } 
      w_k = \frac{V(\hat \theta_k)}{\sum_k V(\hat \theta_k)}.\]
\end{itemize}

By linear model theory this is the BLUE in the linear model case, and
this works really well in Simulation 1 in which

\begin{align*}
  \begin{bmatrix} x_i \\ e_{1i} \\ e_{1i} \end{bmatrix} 
  &\stackrel{ind}{\sim} N\left(
  \begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix},
  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}
  \right)\\
  y_{1i} &= x_i + e_{1i} \\
  y_{2i} &= \theta + x_i + e_{2i} \\
  g(X, Y_1, Y_2) &= Y_2
\end{align*}

As demonstrated in Table~\ref{tab:sim1}, the WLS estimator outperforms all of 
the other estimators except the Oraclex estimator which uses all of the data 
and even some elements that are missing!

\begin{table}[ht!]
  \caption{True Theta is 5. $\Cov(e_1, e_2) = 0.5$}
  \label{tab:sim1}
  \centering
  \begin{tabular}[t]{lrrrr}
    \toprule
    Algorithm & Bias & SD & T-stat & P-val\\
    \midrule
    Oracle & 0.001 & 0.044 & 1.546 & 0.061\\
    Oraclex & 0.000 & 0.032 & 0.479 & 0.316\\
    CC & 0.002 & 0.058 & 1.549 & 0.061\\
    IPW & 0.006 & 0.210 & 1.504 & 0.066\\
    WLS & 0.001 & 0.040 & 1.207 & 0.114\\
    Prop & 0.002 & 0.051 & 1.911 & 0.028\\
    OptSemi & 0.002 & 0.051 & 1.964 & 0.025\\
    \bottomrule
  \end{tabular}
\end{table}

However, in Simulation 2, in which we keep the same setup as Simulation 1,
but change $g(X, Y_1, Y_2) = Y_1^2 Y_2$. In this case, we have

\begin{table}[ht!]
  \caption{True g is 10. $\Cov(e_1, e_2) = 0.5$}
  \label{tab:sim2}
  \centering
  \begin{tabular}[t]{lrrrr}
    \toprule
    Algorithm & Bias & SD & T-stat & P-val\\
    \midrule
    Oracle & 0.007 & 0.529 & 0.741 & 0.229\\
    Oraclex & 0.004 & 0.475 & 0.510 & 0.305\\
    CC & 0.023 & 0.824 & 1.518 & 0.065\\
    IPW & 0.031 & 0.915 & 1.846 & 0.032\\
    WLS & -0.131 & 0.387 & -18.504 & 0.000\\
    Prop & 0.011 & 0.635 & 0.912 & 0.181\\
    SemiOpt & 0.011 & 0.636 & 0.983 & 0.163\\
    \bottomrule
  \end{tabular}
\end{table}

Table~\ref{tab:sim2} clearly shows that the WLS estimator is biased, but 
since this is a weighted average of unbiased linear estimators, how 
can this be biased?

The answer comes from the fact that we weight based on the variance of 
each unbiased estimator. A closer examination of how the bias of 
one of the estimators from $A_k$ affects variance can be seen in
Figure~\ref{fig:biasvar}.

\begin{figure}[ht!]
  \caption{A scatterplot of the Bias-Variance relationship for Section $A_{11}$,
  in which all of the data is observed. A similar relationship is found in each 
  of the other sections too.}
  \label{fig:biasvar}
  \centering
  \includegraphics[width=7cm]{Images/biasvar_11.png}
\end{figure}

Figure~\ref{fig:biasvar} shows that the variance increases as the bias 
becomes more positive. This means that higher estimates will have 
a higher variance. Since we take the inverse variance weighted average 
of the four estimators, we are putting more weight on smaller estimators 
which will lead to a negative bias of our linear estimator which we 
observe in Table~\ref{tab:sim2}.

\end{document}

### 2023-12-11 R/opt_est.R

@@ -81,6 +81,65 @@ gen_optsim_data <- function(n, theta = 0, cor_xe1 = 0, cor_xe2 = 0, cor_e1e2 = 0
 
 }
 
+#' This function generates a multivariate normal distribution with missingness.
+#' The main difference is that the category sizes are unbalanced.
+#'
+#' @param n - An integer for the number of observations
+#'
+#' @return df - A data frame.
+gen_unbal_data <- function(n, theta = 0, cor_xe1 = 0, cor_xe2 = 0, cor_e1e2 = 0) {
+
+  # 1. Generate X, Y1, Y2
+  mean_vec <- c(0, 0, 0)
+  sigma_mat <-
+    matrix(c(1, cor_xe1, cor_xe2, cor_xe1, 1, cor_e1e2, cor_xe2, cor_e1e2, 1),
+           nrow = 3)
+
+  data <- MASS::mvrnorm(n, mean_vec, sigma_mat)
+
+  x_vec <- data[, 1]
+  y1 <- x_vec + data[, 2]
+  y2 <- theta + x_vec + data[, 3]
+
+  # 2. Generate the missing patterns:
+  # | Pattern | Prob |
+  # |---------|------|
+  # | 11      | 0.2  |
+  # | 10      | 0.4  |
+  # | 01      | 0.1  |
+  # | 00      | 0.3  |
+  u_vec <- runif(n)
+  patt_vec <-
+    case_when(
+      u_vec < 0.2 ~ "11",
+      u_vec < 0.6 ~ "10",
+      u_vec < 0.7 ~ "01",
+      u_vec < 1.0 ~ "00"
+    )
+  delta_y1 <-
+    case_when(
+      patt_vec == "11" | patt_vec == "10" ~ 1,
+      patt_vec == "01" | patt_vec == "00" ~ 0
+    )
+  delta_y2 <-
+    case_when(
+      patt_vec == "11" | patt_vec == "01" ~ 1,
+      patt_vec == "10" | patt_vec == "00" ~ 0
+    )
+  prob_1 <- 0.6
+  prob_2 <- 0.3
+  prob_11 <- 0.2
+
+  tibble(X = x_vec, Y1 = y1, Y2 = y2,
+         delta_1 = delta_y1, delta_2 = delta_y2,
+         prob_11 = prob_11, prob_1 = prob_1, prob_2 = prob_2,
+         prob_10 = 0.4, prob_01 = 0.1, prob_00 = 0.3,
[…]

### 2023-12-11 Simulations/T2Optimal_Nonmonotone_Est/fix_linmod.R

@@ -1,0 +1,148 @@
+# Title: fix_linmod.R
+# Created by: Caleb Leedy
+# Created on: December 8, 2023
+# Purpose: This file contains the code to test if the optimal linear model 
+# contains bias for the non-linear case of g = Y1^2 * Y2.
+
+# *************
+# * Libraries *
+# *************
+
+library(dplyr)
+library(stringr)
+
+library(doParallel)
+library(doRNG)
+library(parallelly)
+
+source("R/opt_est.R")
+
+# ***************************
+# * Monte Carlo Simulations *
+# ***************************
+
+clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
+registerDoParallel(clust)
+
+B <- 3000
+n_obs <- 1000
+true_theta <- 5
+cor_e1e2 <- 0.5
+
+mc_theta <-
+  foreach(iter = 1:B,
+          .options.RNG = 1,
+          .packages = c("dplyr", "stringr")) %dorng% {
+
+    # Generate Data
+    df <- gen_optsim_data(n = n_obs, theta = true_theta, cor_e1e2 = cor_e1e2)
+
+    # Get Estimates
+    wls_est <- comb_lin_est(df, gfun = "Y1^2 * Y2", cov_e1e2 = cor_e1e2)
+    wls_est_test <- 
+      comb_lin_est(df, gfun = "Y1^2 * Y2", cov_e1e2 = cor_e1e2, test = TRUE)
+    wlstt_est <- 
+      comb_lin_est(df, gfun = "Y1^2 * Y2", cov_e1e2 = cor_e1e2, theta2 = true_theta)
+    wlstt_est_test <- 
+      comb_lin_est(df, gfun = "Y1^2 * Y2", cov_e1e2 = cor_e1e2,
+                   theta2 = true_theta, test = TRUE)
+
+    return(tibble(wls = wls_est,
+                  wls11 = wls_est_test$est_vec[1],
+                  wls10 = wls_est_test$est_vec[2],
+                  wls01 = wls_est_test$est_vec[3],
+                  wls00 = wls_est_test$est_vec[4],
+                  w_11 = wls_est_test$w_vec[1],
+                  w_10 = wls_est_test$w_vec[2],
+                  w_01 = wls_est_test$w_vec[3],
+                  w_00 = wls_est_test$w_vec[4],
+                  v_11 = wls_est_test$v_vec[1],
[…]

### 2023-12-11 Simulations/T2Optimal_Nonmonotone_Est/optimal_simulation.R

@@ -57,6 +57,7 @@ mc_theta <-
       mean()
     # em_est <- em_optsim(df)
     wls_est <- opt_lin_est(df, cov_y1y2 = cor_e1e2)
+    wlsalt_est <- comb_lin_est_lin(df, theta2 = true_theta, cov_e1e2 = cor_e1e2)
     prop_est <- prop_nmono_est(df)
     propind_est <- prop_nmono_est(df, prop_ind = TRUE)
     propopt_est <- opt_theta_c(df)
@@ -72,6 +73,7 @@ mc_theta <-
                   ipw = ipw_est,
                   # em = em_est,
                   wls = wls_est,
+                  wlsalt = wlsalt_est,
                   prop = prop_est,
                   propind = propind_est,
                   propopt = propopt_est,
@@ -97,6 +99,7 @@ mc_theta |>
     bias_cc = mean(cc) - true_theta,
     bias_ipw = mean(ipw) - true_theta,
     bias_wls = mean(wls) - true_theta,
+    bias_wlsalt = mean(wlsalt) - true_theta,
     bias_prop = mean(prop) - true_theta,
     bias_propind = mean(propind) - true_theta,
     bias_propopt = mean(propopt) - true_theta,
@@ -110,6 +113,7 @@ mc_theta |>
     sd_cc = sd(cc),
     sd_ipw = sd(ipw),
     sd_wls = sd(wls),
+    sd_wlsalt = sd(wlsalt),
     sd_prop = sd(prop),
     sd_propind = sd(propind),
     sd_propopt = sd(propopt),
@@ -123,6 +127,7 @@ mc_theta |>
     tstat_cc = (mean(cc) - true_theta) / sqrt(var(cc) / B),
     tstat_ipw = (mean(ipw) - true_theta) / sqrt(var(ipw) / B),
     tstat_wls = (mean(wls) - true_theta) / sqrt(var(wls) / B),
+    tstat_wlsalt = (mean(wlsalt) - true_theta) / sqrt(var(wlsalt) / B),
     tstat_prop = (mean(prop) - true_theta) / sqrt(var(prop) / B),
     tstat_propind = (mean(propind) - true_theta) / sqrt(var(propind) / B),
     tstat_propopt = (mean(propopt) - true_theta) / sqrt(var(propopt) / B),

### 2023-12-11 Simulations/T2Optimal_Nonmonotone_Est/sim_testing.R

@@ -46,7 +46,7 @@ mc_ests <-
 
     # Estimate different parts of theta_hat_c
     # semidef <- opt_semi_est(df, est = "opt", test = TRUE)
-    semidef <- opt_theta_c(df, est = "opt", test = TRUE)
+    semidef <- opt_theta_c(df, gfun ="Y1^2 * Y2", pow = 3, est = "default", test = TRUE)
 
     return(tibble(est = semidef$est,
                   ipw = semidef$ipw,
@@ -75,6 +75,7 @@ pi_11 <- 0.4
 pi_10 <- 0.2
 pi_01 <- 0.2
 pi_00 <- 0.2
+pi_1112 <- pi_11 * (pi_10 + pi_11) * (pi_01 + pi_11)
 
 # For theta_c
 c_mat <- 
@@ -225,7 +226,6 @@ c0 <- c_mat[1]
 c1 <- c_mat[2]
 c2 <- c_mat[3]
 
-pi_1112 <- pi_11 * (pi_10 + pi_11) * (pi_01 + pi_11)
 var_ipw <- (2 + (1 - pi_11) / pi_11 * (2 + true_theta^2)) / n_obs
 var_A0 <- (c0^2 * (true_theta^2 + 1) * 
               ((pi_10 * pi_01 + pi_11^2) / pi_1112 - 1)) / n_obs

### 2023-12-11 Simulations/T2Optimal_Nonmonotone_Est/unbalanced_sim.R

@@ -1,0 +1,254 @@
+# Title: unbalanced_sim.R
+# Created by: Caleb Leedy
+# Created on: November 29, 2023
+# Purpose: This file contains the code to run an unbalanced simulation.
+# Instead of having A_00, A_10, and A_01 have the same probability in 
+# expectation (like optimal_simulation.R and comp_ests_g.R), this changes the 
+# weights.
+
+# *************
+# * Libraries *
+# *************
+
+library(dplyr)
+library(stringr)
+
+library(doParallel)
+library(doRNG)
+library(parallelly)
+
+source("R/opt_est.R")
+
+# ***************************
+# * Monte Carlo Simulations *
+# ***************************
+
+clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
+registerDoParallel(clust)
+
+B <- 3000
+n_obs <- 1000
+true_theta <- 5 
+cor_e1e2 <- 0.5
+
+mc_theta <-
+  foreach(iter = 1:B,
+          .options.RNG = 1,
+          .packages = c("dplyr", "stringr")) %dorng% {
+
+    # Generate Data
+    df <- gen_unbal_data(n = n_obs, theta = true_theta, cor_e1e2 = cor_e1e2)
+
+    # Get Estimates
+    oracle_est <- mean(df$Y2)
+    oraclex_est <- mean(df$Y2 - df$X)
+    ipw_est <- 
+      filter(df, delta_1 == 1, delta_2 == 1) |>
+      mutate(g_est = Y2 / prob_11) |>
+      pull(g_est) |>
+      sum() / nrow(df)
+    cc_est <- 
+      filter(df, delta_2 == 1) |>
+      pull(Y2) |>
+      mean()
+    # em_est <- em_optsim(df)
+    wls_est <- opt_lin_est(df, cov_y1y2 = cor_e1e2)
+    prop_est <- prop_nmono_est(df)
+    propind_est <- prop_nmono_est(df, prop_ind = TRUE)
+    propopt_est <- opt_theta_c(df)
+    propoptdef_est <- opt_theta_c(df, est = "default")
+    semiopt_est <- opt_semi_est(df)
+    semidef_est <- opt_semi_est(df, est = "default")
+    semidel_est <- opt_delta_c(df)
+    semideldef_est <- opt_delta_c(df, est = "default")
+
+    return(tibble(oracle = oracle_est,
[…]

### 2024-01-16 Notes/efficiencyloss_semi.tex
Updated my response to Dr. Fuller in which I compared parametric and semiparametric estimators to understand their efficiency.

\documentclass[12pt]{article}

\usepackage{amsmath, amssymb, mathrsfs, fancyhdr}
\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
\usepackage{biblatex}
\usepackage{booktabs}
\usepackage{float}

\addbibresource{references.bib}

\hypersetup{colorlinks = true, urlcolor = black}

\headheight     15pt
\topmargin      -1.5cm   % read Lamport p.163
\oddsidemargin  -0.04cm  % read Lamport p.163
\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
\textwidth      16.59cm
\textheight     23.94cm
\parskip         7.2pt   % sets spacing between paragraphs
\parindent         0pt   % sets leading space for paragraphs
\pagestyle{empty}        % Uncomment if don't want page numbers
\pagestyle{fancyplain}

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}

\renewcommand{\arraystretch}{1.2}

\begin{document}

\lhead{Caleb Leedy}
\chead{Non-monotone Missingness}
%\chead{STAT 615 - Advanced Bayesian Methods}
%\rhead{Page \thepage\ of \pageref{LastPage}}
\rhead{\today}

\section*{The Efficiency Loss of Semiparametric Inference}

Inspired by Dr. Fuller's note, I set out to see if we could assess the amount 
of efficiency loss due to using semiparametric inference. First, let's define 
some notation.

\begin{table}[ht!]
\centering
\begin{tabular}{lrrr}
  \toprule
  Segment & Quantity of Interest & Estimator & Coefficient \\
  \midrule
  $A_{00}$ & $E[g \mid X]$ & $\hat \gamma_{11}$ & $c_{11}$ \\
  $A_{10}$ & $E[g \mid X]$ & $\hat \gamma_{21}$ & $c_{21}$ \\
  $A_{10}$ & $E[g \mid X, Y_1]$ & $\hat \gamma_{22}$ & $c_{22}$ \\
  $A_{01}$ & $E[g \mid X]$ & $\hat \gamma_{31}$ & $c_{31}$ \\
  $A_{01}$ & $E[g \mid X, Y_2]$ & $\hat \gamma_{33}$ & $c_{33}$ \\
  $A_{11}$ & $E[g \mid X]$ & $\hat \gamma_{41}$ & $c_{41}$ \\
  $A_{11}$ & $E[g \mid X, Y_1]$ & $\hat \gamma_{42}$ & $c_{42}$ \\
  $A_{11}$ & $E[g \mid X, Y_2]$ & $\hat \gamma_{43}$ & $c_{43}$ \\
  $A_{11}$ & $E[g \mid X, Y_1, Y_2]$ & $\hat \gamma_{44}$ & $c_{44}$ \\
  \bottomrule
\end{tabular}
\caption{This table shows the relationship between different quantities of
interest in each segment, their estimators, and the corresponding coefficients
of their estimators.}
\label{tab:reftab}
\end{table}

This is a similar framework to understanding the class of estimators 
that we discussed in \verb|optest_update.pdf|, where we write the estimator 
as a weighted average of different expectation with functional coefficients.

\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
\mid X, Y_2].\]

However, now we write the estimator as

\[\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}\]

where $\hat \gamma_{kt} = \frac{\delta_{ij}}{\pi_{ij}}E[g \mid G_{ij}(X, Y_1,
Y_2)]$ and $ij$ corresponds to the segment $A_{ij}$ associated with $\hat
\gamma_{kt}$ in Table~\ref{tab:reftab}.

\newpage

\subsection*{Different Classes of Models}

As noted by Dr. Fuller, we can understand parametric estimation as solving 
the following constrained optimization problems:

\begin{align*}
  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
  \sum_{k,t} c_{kt} = 1 \\
       &\text{or} \\
  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
  \sum_{(k,t): (k,t) \neq (4, 4)} c_{kt} = 0 \text{ and } c_{44} = 1. \\
\end{align*}

To be robust to the outcome model, we can construct a similar optimization
problem with different constraints:

\begin{align*}
  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
  c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0,\\
       &\text{and } c_{44} = 1.
\end{align*}

Likewise to be robust to the response model, we can have the problem:

\begin{align*}
  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
  c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \\
       &\text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}.
\end{align*}

To be double robust (robust to the outcome and response model) we can combine 
the last two constraints. This is summarized in Table~\ref{tab:mods}.

\begin{table}[ht!]
  \centering
  \caption{This table identifies the different constraints for each model type.}
  \label{tab:mods}
  \begin{tabular}{lp{12cm}}
    \toprule
    Type & Constraints \\
    \midrule
    Parametric 1 & $\sum_{k, t} c_{kt} = 1$ \\
    Parametric 2 &  $\sum_{k, t: (k, t) \neq (4, 4)} c_{kt} = 0, c_{44} = 1$ \\
    Outcome Robust & $c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} =
    0, c_{33} + c_{43} = 0, \text{ and } c_{44} = 1.$ \\
    Response Robust & $c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} +
    c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} =
    \pi_{11}$ \\
    Double Robust & $c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} =
    0, c_{33} + c_{43} = 0, c_{44} = 1,$ \hspace{3cm} $
    c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} +
    c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}$ \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection*{Simulation Study}

\subsubsection*{Simulation 1}

We use the following simulation setup

\begin{align*}
  \begin{bmatrix} x \\ e_1 \\ e_2 \end{bmatrix} 
  &\stackrel{ind}{\sim} N\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}\right) \\
  y_1 &= x + e_1 \\
  y_2 &= \theta + x + e_2 \\
\end{align*}

Furthermore, $\pi_{11} = 0.2$ and $\pi_{00} = 0.3, \pi_{10} = 0.4$, and 
$\pi_{01} = 0.1$.
The goal of this simulation study is to find $\theta = E[Y_2]$. In other words,
$g(Z) = Y_2$. There are several algorithms for comparison which are defined 
as the following:

\begin{align*}
  Oracle &= n^{-1} \sum_{i = 1}^n g(Z_i)\\
  CC &= \frac{\sum_{i = 1}^n \delta_{11} g(Z_i)}{\sum_{i = 1}^n \delta_{11}} \\
  IPW &= \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g(Z_i)\\
\end{align*}

Furthermore, we include four existing estimators that we have already 
proposed in the past: WLS, Prop, PropInd, and SemiDelta.

The estimator WLS is a a weight least square estimator derived in the 
following manner. We now consider a normal model:

    \[ 
      \begin{pmatrix}
        x_i \\ e_{1i} \\ e_{2i}
      \end{pmatrix} \stackrel{ind}{\sim}
      N \left(
      \begin{bmatrix}
        0 \\ 0 \\ 0
      \end{bmatrix},
      \begin{bmatrix}
        1 & 0 & 0 \\
        0 & \sigma_{11} & \sigma_{12} \\ 
        0 & \sigma_{12} & \sigma_{22}
      \end{bmatrix},
      \right)
    \]

    and define $y_{1i} = \theta_1 + x_i + e_{1i}$ and $y_{2i} = \theta_2 + x_i
    + e_{2i}$. Then $b_1 = b_2 = 1$. We define $\bar z_k^{(ij)}$ as the mean of
    $y_k$ in segment $A_{ij}$. This means that we have means $\bar z_1^{(11)}$,
    $\bar z_2^{(11)}$, $\bar z_1^{(10)}$, and $\bar z_2^{(01)}$. Let $W = [\bar
    z_1^{(11)}, \bar z_2^{(11)}, \bar z_1^{(10)}, \bar z_2^{(01)}]'$, then for
    $n_{ij} = |A_{ij}|$, we have

    \[Z - M \mu \sim N(\vec 0, V)\]

    where

    \[M = 
      \begin{bmatrix}
        1 & 0 \\
        0 & 1 \\
        1 & 0 \\
        0 & 1 \\
      \end{bmatrix}
      \text{ and }
      V = 
      \begin{bmatrix}
        \frac{\sigma_{11}}{n_{11}} & \frac{\sigma_{12}}{n_{11}} & 0 & 0 \\
        \frac{\sigma_{12}}{n_{11}} & \frac{\sigma_{22}}{n_{11}} & 0 & 0 \\
        0 & 0 & \frac{\sigma_{11}}{n_{10}} & 0 \\
        0 & 0 & 0 & \frac{\sigma_{22}}{n_{01}} \\
      \end{bmatrix}.
    \]

    Thus, the BLUE for $\mu = [\mu_1, \mu_2]'$ is

    \[\hat \mu = (M' V^{-1} M)^{-1} M' V^{-1} W.\]

    Hence, WLS is $\mu_2$ as $g(X, Y_1, Y_2) = Y_2$.

The remaining three estimators are drived from the following expression
where the values for $\beta$ are provided in Table~\ref{tab:beta}. These 
are good estimators to compare to because Prop is the original proposed 
estimator. PropInd has the same form as Prop except the values for $\beta$ 
is different. PropInd shares the same values of $\beta$ as all of the new 
models with constraints. SemiDelta is useful because it is the best 
estimator in general so far.

\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
\mid X, Y_2].\]

% TODO: We use the unequal weights
\begin{table}[ht!]
  \centering
  \caption{This table displays the values of $\beta$ for different estimator 
  types.}
  \label{tab:beta}
\begin{tabular}{lrr}
  \toprule
  Estimator & $\beta_0(\delta, c_0)$ & $\beta_1(\delta, c_1)$ \\
  \midrule
  Prop & $\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - 
  \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
  $\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)$ \\
  PropInd & $\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - 
  \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
  $\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)$ \\
%  $\hat \theta_{c}$ & $c_0\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - 
%  \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
%  $c_1\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)$ \\
%  $\hat \theta_{c, ind}$ & $c_0\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - 
%  \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
%  $c_1\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)$ \\
  SemiDelta & $c_0\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{00}}{\pi_{00}}\right)$ &
  $c_1\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{10}}{\pi_{10}}\right)$ \\
  \bottomrule
\end{tabular}
\end{table}

In the simulation results in Table~\ref{tab:sim1}, the new results have the 
same label as the value from the ``Type'' column in Table~\ref{tab:mods}.

\begin{table}[ht!]
  \caption{Results from Simulation 1. True values: $\theta = 5, \rho = 0.5$.
  The test conducted for the T-statistic and P-value is a two sample test to 
  see if the estimator is unbiased.}
  \label{tab:sim1}
  \centering
  \begin{tabular}[t]{lrrrr}
    \toprule
    Algorithm & Bias & SD & Tstat & P-value\\
    \midrule
    Oracle & -0.002 & 0.045 & -1.669 & 0.048\\
    CC & 0.001 & 0.083 & 0.263 & 0.396\\
    IPW & 0.010 & 0.357 & 0.872 & 0.192\\
    WLS & 0.000 & 0.054 & -0.065 & 0.474\\
    %WLS (with true $\theta$) & -0.001 & 0.035 & -1.158 & 0.124\\
    Prop & -0.002 & 0.063 & -0.768 & 0.221\\
    PropInd & -0.003 & 0.120 & -0.900 & 0.184\\
    SemiDelta & -0.001 & 0.064 & -0.342 & 0.366\\
    Parametric 1 & -0.001 & 0.061 & -0.509 & 0.305\\
    Parametric 2 & -0.001 & 0.061 & -0.509 & 0.305\\
    Outcome Robust & -0.001 & 0.061 & -0.543 & 0.294\\
    Response Robust & -0.001 & 0.061 & -0.509 & 0.305\\
    Double Robust & -0.001 & 0.061 & -0.543 & 0.294\\
    \bottomrule
  \end{tabular}
\end{table}

\newpage

\subsubsection*{Simulation 2}

Next, we used the same simulation setup except that we are interested in 
estimating $\theta = E[g] = E[Y_1^2 Y_2]$. We do not use the WLS estimator 
because it does not work.

\begin{table}[ht!]
  \caption{Results from Simulation 2. True values: $\theta = 10, \rho = 0.5$.
  The test conducted for the T-statistic and P-value is a two sample test to 
  see if the estimator is unbiased.}
  \label{tab:sim2}
  \centering
  \begin{tabular}[t]{lrrrr}
    \toprule
    Algorithm & Bias & SD & Tstat & P-value\\
    \midrule
    Oracle & -0.037 & 0.528 & -2.214 & 0.014\\
    CC & 0.007 & 1.196 & 0.179 & 0.429\\
    IPW & 0.023 & 1.405 & 0.520 & 0.302\\
    Prop & -0.044 & 0.691 & -2.027 & 0.021\\
    SemiDelta & -0.032 & 0.674 & -1.521 & 0.064\\
    Parametric 1 & -0.004 & 0.301 & -0.460 & 0.323\\
    Parametric 2 & -0.004 & 0.344 & -0.341 & 0.367\\
    Outcome Robust & -0.040 & 0.654 & -1.950 & 0.026\\
    Response Robust & -0.004 & 0.301 & -0.461 & 0.322\\
    Double Robust & -0.040 & 0.654 & -1.954 & 0.025\\
    \bottomrule
  \end{tabular}
\end{table}

\newpage

\subsection*{Discussion}

\begin{itemize}
  \item In the first simulation there does not seem to be much of a cost 
    to using a semiparametric estimator with the Double Robust model 
    performing as well as the Outcome Robust, Response Robust and 
    Parametric models. However, in the second simulation the Parametric 
    models and Response Robust estimators perform the best.
  \item Strangly, the Double Robust estimator in the second simulation 
    has picked up the bad characteristics of the Outcome Model. This 
    can make sense if the optimal Outcome Model is contained in the 
    space of the Double Robust model but the optimal Response Model 
    is not.
\end{itemize}

\end{document}

### 2024-01-16 Notes/fall_recap.tex
Updated my response to Dr. Fuller in which I compared parametric and semiparametric estimators to understand their efficiency.

--- 2023-11-07 Notes/optest_update.tex
@@
 
-\section*{Setup}
+\section*{Recap of Fall 2023}
 
-Consider the following setup. Let $(X, Y_1, Y_2, \delta)  \stackrel{ind}{\sim}
-F$ for some distribution $F$ that is unknown. We define $Z = (X, Y_1, Y_2$ and
-we observe the following:
-
-\begin{table}[!ht]
-  \centering
-  \caption{This table shows some of our notation and some of the corresponding
-  notation from \cite{tsiatis2006semiparametric}.}
-\begin{tabular}{cccccccc}
-  \toprule
-  Segments & $X$ & $Y_1$ & $Y_2$ & Prob. Element in Segment & $\delta$ & $C$ & $G_C(Z)$ \\
-  \midrule
-  $A_{00}$ & $\checkmark$ &               &         & $\pi_{00}$ & $\delta_{00}$ & 1 & $\{X\}$ \\
-  $A_{10}$ & $\checkmark$ & $\checkmark$  &         & $\pi_{10}$ & $\delta_{10}$ & 2 & $\{X, Y_1\}$ \\
-  $A_{01}$ & $\checkmark$ &               & $\checkmark$ & $\pi_{01}$ & $\delta_{01}$ & 3 & $\{X, Y_2\}$ \\
-  $A_{11}$ & $\checkmark$ & $\checkmark$  & $\checkmark$ & $\pi_{11}$ & $\delta_{11}$ & $\infty$ & $\{X, Y_1, Y_2\}$ \\
-  \bottomrule
-\end{tabular}
-\end{table}
-
-Define $\varpi(r, Z) = \Pr(C = r \mid Z)$. For now, we assume that $\varpi(r,
-Z)$ is known. Notice that $\varpi(\infty, Z) = \pi_11$, $\varpi(3, Z) = \pi_{01}$,
-$\varpi(2, Z) = \pi_{10}$, and $\varpi(1, Z) = \pi_{00}$. Since $\pi_{00} +
-\pi_{10} + \pi_{01} + \pi_{11} = 1$, we only need to define three inclusion 
-probabilities.\\
-
-Suppose that we want to estimate $\theta = E[g(X, Y_1, Y_2)]$ for a known
-function $g$. 
-The goal is the show that over all functions $b_1(X, Y_1)$ and $b_2(X, Y_2)$,
-we have the optimal estimator in the form:
+Fall 2023 started with the goal of showing that $\hat \theta_{prop}$ in
+Equation~\ref{eq:prop} is (or is not) an optimal estimator.
 
 \begin{align}
-  &\hat \theta =\\ \nonumber
+  \hat \theta_{prop} &= \label{eq:prop}\\ \nonumber
   &n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] + n^{-1} \sum_{i = 1}^n
-  \frac{\delta_{10}}{\pi_{10}} (b_1(X_i, Y_{1i}) - E[g_i \mid X_i]) 
-  + n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} (b_2(X_i, Y_{2i}) - E[g_i \mid X_i]) \\ \nonumber
-  &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} (g_i - b_1(X_i, Y_{1i}) - b_2(X_i, Y_{2i}) + E[g_i \mid X_i])
+  \frac{\delta_{1+}}{\pi_{1+}} (E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i]) \\
+  \nonumber
+  &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{2+}}{\pi_{2+}} (E[g_i \mid X_i,
+  Y_{2i}] - E[g_i \mid X_i]) \\ \nonumber
+  &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} (g_i - E[g_i \mid X_i,
+  Y_{1i}] - E[g_i \mid X_i, Y_{2i}] + E[g_i \mid X_i]). \nonumber
 \end{align}
 
-%The proposed estimator is 
-%
-%\begin{align}\label{eq:propind}
-%  &\hat \theta^{ind}_{prop} =\\ \nonumber
-%  &n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] + n^{-1} \sum_{i = 1}^n
-%  \frac{\delta_{10}}{\pi_{10}} (E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i]) 
-%  + n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} (E[g_i \mid X_i, Y_{2i}] - E[g_i \mid X_i]) \\ \nonumber
-%  &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} (g_i - E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i, Y_{2i}] + E[g_i \mid X_i]).
-%\end{align}
+While this estimator performs quite well overall, it is outperformed by 
+$\hat \theta_\delta$ (see Table~\ref{tab:ests}) in a simulation where the 
+number of observations in segments $A_{10}$ differed from $A_{01}$. Studying 
+this estimator led to the creation of a class of estimators in
+Equation~\ref{eq:class} with specific examples in Table~\ref{tab:ests}.
 
-\section*{Semiparametric Inference}
-
-We know from Theorem 7.2 of \cite{tsiatis2006semiparametric} that if $\Pr(C =
-\infty \mid Z) = \pi_{11} > 0$ 
-then the semiparametric influence function has the form (see page 20 of notes):
-
-\begin{equation}\label{lambda2}
- \frac{I(C = \infty) g(Z)}{\varpi(\infty, Z)} + \frac{I(C =
-  \infty)}{\varpi(\infty, Z)} \left(\sum_{r \neq \infty} \varpi(r, G_r(Z))
-  L_{2r}(G_r(Z))\right) - \sum_{r \neq \infty} I(C = r)L_{2r}(G_r(Z))
+\begin{equation}\label{eq:class}
+\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
+\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
+\mid X, Y_2].
 \end{equation}
 
-where $L_{2r}(G_r(Z))$ is an arbitrary function of $G_r(Z)$. Notice, that this
-form does not identify \textit{the} optimal estimator but a class of 
-semiparametric functions. A reasonable choice of an estimator for $L_{2r}(G_r(Z))$ is
-
-\[L_{2r}(G_r(Z)) = c_r E[g(Z) \mid G_r(Z)]\]
-
-where each $c_r$ is a constant. This actually suggests a variety of estimators 
-that each have different values of coefficients of $\delta$. Initially, we tried 
-a direct projection onto the nuisance tangent space, but I got stuck. See Appendix A 
-for this work.
-
-\subsection*{Different Classes of Estimators}
-
-All of the considered estimators have a similar form:
-
-\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
-\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
-\mid X, Y_2].\]
-
-In this case, $c$ is a vector of constants $c = (c_0, c_1, c_2)$ that are 
-chosen to minimize the variance of $\hat \theta$. See Appendix B for details on 
-an example about choosing the values of $\hat c$.
-
+\begin{table}[ht!]
+  \centering
+  \caption{Specific examples of estimators from the larger class in
+  Equation~\ref{eq:class}.}
+  \label{tab:ests}
 \begin{tabular}{lrrr}
@@
 \end{tabular}
-
-%It turns out that $\hat \theta_{prop}$ from Equation~\ref{eq:prop} is contained within 
-%the previous class of estimators with 
-%
-%\[L_{2r}(G_r(Z)) = \varpi(r, G_r(Z)) E[g(Z) \mid G_r(Z)].\]
-%
-%(See pages 81-82 of notes for proof.) \\
-
-\section*{Simulation Studies}
-
-\subsection*{Simulation 1}
-
-To see which of these estimators is the best, we run a simulation study. In this 
-study, we have the following variables:
-
-\begin{align*}
-  \begin{bmatrix} x \\ e_1 \\ e_2 \end{bmatrix} 
-  &\stackrel{ind}{\sim} N\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
-  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}\right) \\
-  y_1 &= x + e_1 \\
-  y_2 &= \theta + x + e_2 \\
-\end{align*}
-
-Furthermore, $\pi_{11} = 0.4$ and $\pi_{00} = \pi_{10} = \pi_{01} = 0.2$.
-The goal of this simulation study is to find $\theta = E[Y_2]$. In other words,
-$g(Z) = Y_2$. There are several algorithms for comparison which are defined 
-as the following:
-
-\begin{align*}
-  Oracle &= n^{-1} \sum_{i = 1}^n g(Z_i)\\
-  CC &= \frac{\sum_{i = 1}^n \delta_{11} g(Z_i)}{\sum_{i = 1}^n \delta_{11}} \\
-  IPW &= \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g(Z_i)\\
-\end{align*}
-
-\begin{table}[ht!]
-\caption{The true value of $\theta$ is 5. $\rho = 0.5$. The test statistic and 
-p-value are from a one-sample t-test to see if the estimator is biased.}
-\centering
-\begin{tabular}[t]{lrrrr}
-\toprule
-Algorithm & Bias & SD & Tstat & Pval\\
-\midrule
-Oracle & 0.001 & 0.044 & 1.546 & 0.061\\
-CC & 0.002 & 0.058 & 1.549 & 0.061\\
-IPW & 0.006 & 0.210 & 1.504 & 0.066\\
-$\hat \theta_{prop}$ & 0.002 & 0.051 & 1.911 & 0.028\\
-$\hat \theta_{prop}^{ind}$ & 0.002 & 0.091 & 0.983 & 0.163\\
-$\hat \theta_{c, ind}$ & 0.001 & 0.074 & 0.784 & 0.216\\
-$\hat \theta_{\delta}$ & 0.002 & 0.051 & 1.964 & 0.025\\
-\bottomrule
-\end{tabular}
 \end{table}
 
-The main takeaway from this simulation study is that $\hat \theta_\delta$
-and $\hat\theta_{prop}$ have the smallest standard deviation among the 
-non-oracle estimators.
+The challenge with working with this class is the $\beta$ coefficients. These
+are functional coefficients and so it is difficult for me to understand how to 
+show that a particular class is optimal.
 
-\subsection*{Simulation 2}
-We now test a second simulation study. The setup of this simulation study 
-is the same as the previous simulation. The only difference is that we 
-are interested in estimating $\theta = E[Y_1^2 Y_2]$. In other words, now 
-we have $g(Z) = Y_1^2 Y_2$.
+\section*{Current Work}
 
-\begin{table}[ht!]
-  \caption{The true value of $E[g(Z)]$ is 10. $\rho = 0.5$. The test statistic and 
-p-value are from a one-sample t-test to see if the estimator is biased.}
-\centering
-\begin{tabular}[t]{lrrrr}
-\toprule
-Algorithm & Bias & SD & Tstat & Pval\\
-\midrule
-Oracle & 0.007 & 0.529 & 0.741 & 0.229\\
-CC & 0.023 & 0.824 & 1.518 & 0.065\\
-IPW & 0.031 & 0.915 & 1.846 & 0.032\\
-$\hat \theta_{prop}$ & 0.011 & 0.635 & 0.912 & 0.181\\
-$\hat \theta_{c, ind}$ & 0.013 & 0.662 & 1.060 & 0.145\\
-$\hat \theta_{\delta}$ & 0.011 & 0.636 & 0.983 & 0.163\\
-\bottomrule
-\end{tabular}
-\end{table}
+As discussed in a related note, we are currently trying to understand the 
+loss of efficiency of using a semiparametric estimator with the model is 
+correctly specified. I am having difficulties with the current simulation 
+setup but this work is contained in \verb|efficiencyloss_semi.tex|.
 
-Like the first simulation, $\hat \theta_{prop}$ and $\hat \theta_{\delta}$ 
-performed the best.
-
-\newpage
-
-\section*{Next Steps}
+\section*{Potential Ideas to Pursue}
 
 \begin{itemize}
-  \item[1.] It turns out that \cite{tsiatis2006semiparametric} contains 
-    a general technique for solving non-monotone missing data. They show 
-    that the optimal estimator is found via the following double projection:
-
-    \[L_2 = \Pi\left(\Pi\left(\frac{\delta_{11}}{\pi_{11}}g(Z) \mid C,
-    G_C(Z)\right) \mid Z\right)\]
-
-    I can show that $\hat \theta_\delta$ is NOT such a double projection. 
-    However, I think it is worth exploring what such an estimator (especially 
-    if we consider a linear estimator) looks like.
-  \item[2.] Similar to the previous point, \cite{tsiatis2006semiparametric} 
-    discuss what the optimal estimator of a non-monotone problem looks like 
-    and they determine that it solves the estimating equation
-
-    \[\sum_{i = 1}^n \mathcal{L}(\mathcal{M}^{-1}(g(Z)))\]
-
-    where $\mathcal{M}(g(Z)) = E[\mathcal{L}(g(Z)) \mid Z]$ and 
-    $\mathcal{L}(g(Z)) = E[g(Z) \mid C, G_C(Z)]$. The difficult part is 
-    understanding $\mathcal{M}^{-1}$, but there is conveniently the fact that
-
-    \[\mathcal{M}^{-1}(g(Z)) = \lim_{n \to \infty} \phi_{n + 1}(Z) \text{ where } 
-      \phi_{n + 1}(Z) = (I - \mathcal{M}) \phi_n(Z) + g(Z) \text{ and }
-    \phi_0(Z) = g(Z).\]
-
-    I have been investigating what this estimating equation yields if we 
-    approximate $\mathcal{M}^{-1}$ with $\phi_1$. However, I have not 
-    finished this yet.
-  \item[3.] A different method to find the projection onto $\Lambda_2$ 
-    could be to try to minimize the (KL) divergence between 
-    $L_2$ and $\frac{\delta_{11}}{\pi_{11}}g(Z)$.
+  \item Tsiatis (2006) describes a recursive method to get the optimal 
+    semiparametric estimator for non-monotone data. This is difficult and 
+    confusing. However, we can pose the estimation as the solution to an 
+    integral equation. While people know that one can do this, it has not 
+    been done. The real work would be estimating the variance.
+  \item Often in simulations, the optimal value for $c_i$ would be $1$ or 
+    $-1$. Since the optimal semiparametric estimator is a double projection
+    (see Chapter 10 of Tsiatis (2006)), I hypothesis that it may be possible to
+    construct a test to see how close to the optimal we are by projecting 
+    the the $\beta$ coefficients onto some contants $c$.
 \end{itemize}
 
-\newpage
-
-\printbibliography
-
-\newpage
-
-\appendix
-
-\section{Linear Expectations}
-
-To simplify this problem, we consider the following estimator\footnote{This
-estimator has slightly different coefficients compared to the initial
-estimator.}:
-
-\begin{align}
-  &\hat \theta_c \\ \nonumber
-  &= \frac{\delta_{11}}{\pi_{11}} g(Z) + \left(1 -
-    \left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}}\right) -
-    \left(\frac{\delta_{01} + \delta_{11}}{\pi_{01} + \pi_{11}}\right) + 
-  \frac{\delta_{11}}{\pi_{11}}\right) c_0 E[g \mid X] \\ \nonumber
-  &+
-  \left(\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}}\right) - 
-  \frac{\delta_{11}}{\pi_{11}}\right) c_1 E[g \mid X, Y_1] +
-  \left(\left(\frac{\delta_{01} + \delta_{11}}{\pi_{01} + \pi_{11}}\right) - 
-  \frac{\delta_{11}}{\pi_{11}}\right) c_2 E[g \mid X, Y_2]
-\end{align}
-
-\subsection{Projection onto Nuisance Tangent Space}
-
-The goal is now to find the coefficients $c_0, c_1$, and $c_2$ such that 
-$\langle \hat \theta_c, L_2\rangle \equiv E[\hat \theta_c L_2] = 0$ for 
-all $L_2 \in \Lambda_2$ (see \cite{tsiatis2006semiparametric} for definition 
-of $\Lambda_2$). If we can find such coefficients that the estimator $\hat
-\theta_c$ will be orthogonal to $\Lambda_2$ and hence by Theorem 10.1 of 
-\cite{tsiatis2006semiparametric} semiparametrically optimal. The good news 
-is that we know (from Theorem 7.2) that any element $L_2 \in \Lambda_2$ has a
-form:
-
-\begin{align}
-  L_2 &= 
-  \left(\frac{\delta_{11}}{\pi_{11}}\pi_{00} - \delta_{00}\right) L_{20}(X) +
-  \left(\frac{\delta_{11}}{\pi_{11}}\pi_{10} - \delta_{10}\right) L_{21}(X, Y_1) +
-  \left(\frac{\delta_{11}}{\pi_{11}}\pi_{01} - \delta_{01}\right) L_{22}(X, Y_2).
-\end{align}
-
-Then expanding and solving $E[\hat \theta_c L_2] = 0$ yields:
-
-{\tiny
-\begin{align*}
-  0 &= E[\hat \theta_c L_2] \\ 
-    &E\left[\left(\frac{\pi_{00}}{\pi_{11}} + \left(\frac{\pi_{10}}{\pi_{10} + \pi_{11}}\right)\frac{\pi_{00}c_1}{\pi_{11}} +
-      \left(\frac{\pi_{01}}{\pi_{01} + \pi_{11}}\right) \frac{\pi_{00}c_2}{\pi_{11}} + \frac{\pi_{00}(\pi_{10}\pi_{01} -
-  \pi_{11}^2)c_0}{(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})\pi_{11}}\right) E[g \mid X] L_{20}(X)\right. \\ 
-    &+ \frac{\pi_{10}}{\pi_{11}} \left(E[g(Z) L_{21}(X, Y_1) \mid X] - c_1 E[E[g(Z) \mid X, Y_1] L_{21}(X, Y_1) \mid X] +
-      \frac{\pi_{01}c_2}{\pi_{10} + \pi_{11}} E[E[g \mid X, Y_2] L_{21}(X, Y_1) \mid X] +
-      \frac{\pi_{10} \pi_{01}}{\pi_{11}(\pi_{01} + \pi_{11})} E[g \mid X]
-      E[L_{21}(X, Y_1) \mid X]c_0\right)\\
-    &+ \left.\frac{\pi_{01}}{\pi_{11}} \left(E[g(Z) L_{22}(X, Y_2) \mid X] + \frac{\pi_{10} c_1}{\pi_{10} + \pi_{11}} E[E[g(Z) \mid X, Y_1] L_{22}(X, Y_2) \mid X] -
-      c_2 E[E[g \mid X, Y_2] L_{22}(X, Y_2) \mid X] +
-      \frac{\pi_{10}}{(\pi_{01} + \pi_{11})} E[g \mid X]
-      E[L_{22}(X, Y_2) \mid X]c_0\right)\right]
-\end{align*}
-}
-
-To solve for $c_0, c_1$, and $c_2$ we need the following to hold for any
-$L_{21}(X, Y_1)$ and $L_{22}(X, Y_2)$:
-
-{\small
-\begin{align*}
-  &1 + c_0 \frac{\pi_{01} \pi_{10} - \pi_{11}^2}{(\pi_{10} + \pi_{11})(\pi_{01}
-  + \pi_{11})} + c_1 \frac{\pi_{10}}{\pi_{01} + \pi_{11}} + c_2
-  \frac{\pi_{01}}{\pi_{01} + \pi_{11}} = 0\\ 
-  &E\left[\left(g(Z) + c_0 \frac{\pi_{01}}{\pi_{01} + \pi_{11}}E[g(Z) \mid X] - c_1 E[g(Z) \mid X, Y_1] + c_2 \frac{\pi_{01}}{\pi_{10} + \pi_{11}}E[g(Z) \mid X, Y_2]\right)L_{21}(X, Y_1) \mid X\right] = 0 \\
-  &E\left[\left(g(Z) + c_0 \frac{\pi_{10}}{\pi_{10} + \pi_{11}}E[g(Z) \mid X] + c_1 \frac{\pi_{10}}{\pi_{10} + \pi_{11}} E[g(Z) \mid X, Y_1] - c_2 E[g(Z) \mid X, Y_2]\right)L_{22}(X, Y_2) \mid X\right] = 0
-\end{align*}
-}
-
-Unfortunately, I am now stuck because it is unclear to me how to use any actual 
-data to solve this problem.
-
-\section{Solving for $\hat c$}
-
-We can find the values of 
-$c_0$, $c_1$, and $c_2$ in $\hat \theta_c$ that minimize the variance the 
-estimator. We can find these values by differentiating by $c_i$ and solving 
-for $c_i$:
-
-\begin{align*}
-  \begin{bmatrix} \hat c_0 \\ \hat c_1 \\ \hat c_2 \end{bmatrix} 
-  &= -
-  \begin{bmatrix} 
-    M_{11} & M_{12} & M_{13} \\
-    M_{21} & M_{22} & M_{23} \\
-    M_{31} & M_{32} & M_{33} \\
-  \end{bmatrix}^{-1}
-  \times 
-  \begin{bmatrix}
-    E[E[g \mid X]^2] \left(1 + \frac{\pi_{10}\pi_{01} - \pi_{11}^2}{\pi_{11}(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})}\right) \\
-    E[E[g \mid X, Y_1]^2] \left(\frac{-\pi_{10}}{\pi_{11}(\pi_{10} + \pi_{11})}\right) \\
-    E[E[g \mid X, Y_2]^2] \left(\frac{-\pi_{01}}{\pi_{11}(\pi_{01} + \pi_{11})}\right) \\
-  \end{bmatrix}
-\end{align*}
-
-where
-
-\begin{align*}
-  M_{11} &= E[E[g \mid X]^2] \left(\frac{\pi_{11}^2 + \pi_{10}\pi_{01}}{\pi_{11}(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})} - 1\right) \\
-  M_{12} &= E[E[g \mid X]^2] \left(\frac{-\pi_{10}\pi_{01}}{\pi_{11}(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})}\right) \\
-  M_{13} &= E[E[g \mid X]^2] \left(\frac{-\pi_{10}\pi_{01}}{\pi_{11}(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})}\right) \\
-  M_{22} &= E[V(E[g \mid X, Y_1] \mid X)] \left(\frac{\pi_{10}}{\pi_{11}(\pi_{10} + \pi_{11})}\right) \\
-  M_{23} &= E[E[g \mid X, Y_1]E[g \mid X, Y_2]] \left(\frac{\pi_{10}\pi_{01}}{\pi_{11}(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})}\right) \\
-  M_{33} &= E[V(E[g \mid X, Y_2] \mid X)] \left(\frac{\pi_{01}}{\pi_{11}(\pi_{01} + \pi_{11})}\right)
-\end{align*}
-
-This estimator is similar to several other estimators:
-
-\begin{align}\label{thetaind}
-  \hat \theta^{ind}_c 
-  &= \frac{\delta_{11}}{\pi_{11}} g(Z) + \left(1 -
-    \left(\frac{\delta_{10}}{\pi_{10}}\right) -
-    \left(\frac{\delta_{01}}{\pi_{01}}\right) + 
-  \frac{\delta_{11}}{\pi_{11}}\right) c_0 E[g \mid X] \\ \nonumber
-  &+
-  \left(\frac{\delta_{10}}{\pi_{10}} - 
-  \frac{\delta_{11}}{\pi_{11}}\right) c_1 E[g \mid X, Y_1] +
-  \left(\frac{\delta_{01}}{\pi_{01}} - 
-  \frac{\delta_{11}}{\pi_{11}}\right) c_2 E[g \mid X, Y_2]
-\end{align}
-
-\begin{align}\label{thetadelta}
-  \hat \theta^{\delta}_c 
-  &= \frac{\delta_{11}}{\pi_{11}} g(Z) 
-  + \left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{00}}{\pi_{00}}\right) 
-   c_0 E[g \mid X] \\ \nonumber
-  &+
-  \left(\frac{\delta_{11}}{\pi_{11}} - 
-  \frac{\delta_{10}}{\pi_{10}}\right) c_1 E[g \mid X, Y_1] +
-  \left(\frac{\delta_{11}}{\pi_{11}} - 
-  \frac{\delta_{01}}{\pi_{01}}\right) c_2 E[g \mid X, Y_2]
-\end{align}
-
-The first expression (Equation~\ref{thetaind}) is the proposed estimator 
-with independent differences in each segment, while the second expression 
-(Equation~\ref{thetadelta}) is the optimal estimator with values of $\delta$
-such that $\hat \theta^{\delta}_c \in \Lambda_2$, which means that it has 
-the form of the semiparametric in Equation~\ref{lambda2}.
-
 \end{document}

### 2024-01-16 Simulations/T3Assess_Semiparametric_Cost/compare_algs.R

@@ -44,53 +44,88 @@ min_var_alg <-
   }
 
   # We use the population functional forms for gamma_hat and v_gamma
-  g_11 <- mean(df$delta_00 / df$prob_00 * (theta + df$X))
-  g_21 <- mean(df$delta_10 / df$prob_10 * (theta + df$X))
-  g_22 <- mean(df$delta_10 / df$prob_10 * (theta + df$X + cov_e1e2 * (df$Y1 - df$X)))
-  g_31 <- mean(df$delta_01 / df$prob_01 * (theta + df$X))
-  g_33 <- mean(df$delta_01 / df$prob_01 * (df$Y2))
-  g_41 <- mean(df$delta_11 / df$prob_11 * (theta + df$X))
-  g_42 <- mean(df$delta_11 / df$prob_11 * (theta + df$X + cov_e1e2 * (df$Y1 - df$X)))
-  g_43 <- mean(df$delta_11 / df$prob_11 * (df$Y2))
-  g_44 <- mean(df$delta_11 / df$prob_11 * (df$Y2))
+  if (gfun == "Y2") {
+
+    Eg <- theta
+    Egx <- theta + df$X 
+    Egxy1 <- theta + df$X + cov_e1e2 * (df$Y1 - df$X)
+    Egxy2 <- df$Y2
+    EEg2 <- theta^2 + 2
+    EEgx2 <- theta^2 + 1
+    EEgxy12 <- theta^2 + 1 + cov_e1e2^2
+    EEgxy22 <- theta^2 + 2
+    EEgxy1Egxy2 <- theta^2 + 1 + cov_e1e2^2
+
+  } else if (gfun == "Y1^2 * Y2") {
+
+    Eg <- 2 * theta
+    Egx <- df$X^3 + theta * df$X^2 + theta + 2 * df$X * cov_e1e2 + df$X
+    Egxy1 <- df$Y1^2 * (theta + df$X + cov_e1e2 * (df$Y1 - df$X))
+    Egxy2 <- df$Y2 * (df$X^2 + 2 * cov_e1e2 * df$X * (df$Y2 - df$X - theta) + 
+                      1 + cov_e1e2^2 * (df$Y2 - df$X - theta)^2 - cov_e1e2^2)
[…]

### 2024-01-16 Simulations/T3Assess_Semiparametric_Cost/semi_cost.R

@@ -30,6 +30,10 @@ source("Simulations/T3Assess_Semiparametric_Cost/compare_algs.R")
 # * Monte Carlo Simulations *
 # ***************************
 
+# ************************
+# * theta = E[g] = E[Y2] *
+# ************************
+
 clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
 registerDoParallel(clust)
 
@@ -44,7 +48,8 @@ mc_theta <-
           .packages = c("dplyr", "stringr", "CVXR")) %dorng% {
 
     # Generate Data
-    df <- gen_optsim_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)
+    # df <- gen_optsim_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)
+    df <- gen_unbal_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)
 
     # Get Estimates
     oracle_est <- mean(df$Y2)
@@ -175,31 +180,147 @@ mc_theta |>
                caption = paste0("True Theta is ", true_theta,
                                 ". Cov_e1e2 = ", cov_e1e2 ))
 
-# * Testing the standard deviations
-# When sigma_11 = 1 and sigma_22 = 1, we have:
-# Expected sd: 
-# Oracle:
-sqrt(2 / n_obs)
-# OracleX: 
-sqrt(1 / n_obs)
-# CC:
-sqrt(2 / (n_obs * 0.6))
-# IPW:
-sqrt((2 + (0.6 * true_theta^2)) / (n_obs * 0.4))
-# WLS:
-m_mat <- matrix(c(1, 0, 1, 0, 0, 1, 0, 1), ncol = 2)
-v_mat <- matrix(c(1 / (n_obs * 0.4), cov_e1e2 / (n_obs * 0.4), 0, 0,
-                  cov_e1e2 / (n_obs * 0.4), 1 / (n_obs * 0.4), 0, 0,
-                  0, 0, 1 / (n_obs * 0.2), 0,
-                  0, 0, 0, 1 / (n_obs * 0.2)), nrow = 4)
-exp_inv_cov <- solve(t(m_mat) %*% solve(v_mat) %*% m_mat)
-sqrt(exp_inv_cov[2, 2])
-# Prop
-exp_var <- 
[…]

### 2024-01-22 Notes/fall_recap.tex
Added Quarto document with semiparametric constraint simulation.

@@
 are functional coefficients and so it is difficult for me to understand how to 
-show that a particular class is optimal.
+show that a particular class is optimal. Dispite this difficulty, we were able 
+to show that the proposed estimator is not optimal via simulation. When this 
+estimator was compared to other estimators in Table~\ref{eq:class}, it did not 
+perform the best when the segments were unbalanced and the estimating function 
+was nonlinear. This is shown in Table~\ref{tab:unbalanced}
+
+\begin{table}[ht!]
+  \caption{True Values: $\theta = 10$, $\rho = 0.5$. This simulation assesses 
+  the bias and standard deviation (SD) of different estimators of the function 
+$\theta = E[Y_1^2 Y_2]$ where $Y_1 = x + \varepsilon_1$, $Y_2 = \beta + x + 
+\varepsilon_2$, $x \perp (\varepsilon_1, \varepsilon_2)$, and 
+$(\varepsilon_1, \varepsilon_2)$ come from a mean zero bivariate normal 
+distribution with unit variance and covariance of $\rho$. This simulation uses 
+a sample size of $N = 1000$ and $B = 3000$ Monte Carlo simulations. Each 
+observation has the following independent probabilities of landing into a 
+respective segment: $\pi_{11} = 0.2$, $\pi_{10} = 0.4$, $\pi_{01} = 0.1$, and
+$\pi_{00} = 0.3$. The columns of Tstat and P-value give the test statistic and
+p-value from a two sample test for unbiasedness.}
+  \label{tab:unbalanced}
+  \centering
+  \begin{tabular}[t]{lrrrr}
+    \toprule
+    Algorithm & Bias & SD & Tstat & P-value\\
+    \midrule
+    Oracle & 0.007 & 0.529 & 0.741 & 0.229\\
+    CC & 0.036 & 1.190 & 1.667 & 0.048\\
+    IPW & 0.029 & 1.372 & 1.170 & 0.121\\
+    $\hat \theta_{prop}$ & 0.004 & 0.694 & 0.292 & 0.385\\
+    $\hat \theta_{c}$ & 0.007 & 0.670 & 0.582 & 0.280\\
+    $\hat \theta_{\delta}$ & 0.018 & 0.675 & 1.478 & 0.070\\
+    \bottomrule
+  \end{tabular}
+\end{table}
 

### 2024-01-22 Notes/fstat_question.qmd
Added Quarto document with semiparametric constraint simulation.

---
title: "The F Test"
author:
  # Add additional authors here if needed, by replicating the above entries.
  - name: Caleb Leedy
# The short* versions of title and author are used for constructing page headers
# and are optional
date: today
date-format: "D MMMM YYYY"
#abstract: |
format:
  simple-article-pdf:
    include-in-header: "latex_header.tex"
    keep-tex: true
    indent: false
    classoption: 
     - 12pt
# reference-section-title: References
# bibliography: ../references.bib
---

# Computing the F-Test

To compute an F test we can use @eq-fstat.

$$ F = \frac{(SSE_{Reduced} - SSE_{Full}) / (DFE_{Reduced} - DFE_{Full})}{
SSE_{Full} / DFE_{Full}}. $$ {#eq-fstat}

Let's assess what this would mean to compare models `Parametric 1` and 
`Outcome Robust` from @tbl-mods.

| Type | Constraints |
| -----| ------------|
| Parametric 1 | $\sum_{k, t} c_{kt} = 1$ |
| Parametric 2 |  $\sum_{k, t: (k, t) \neq (4, 4)} c_{kt} = 0, c_{44} = 1$ |
| Outcome Robust | $c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, \text{ and } c_{44} = 1.$ |
| Response Robust | $c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}$ |
| Double Robust | $c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, c_{44} = 1, c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}$ |
: This table identifies the different constraints for each model type. {#tbl-mods}

Consider the case where we run estimate each model on a data set with $n = 1000$
observations, and define the following notation: let $\hat \theta^{(P)}$ and 
$\hat \theta^{(OR)}$ be the estimated values of $\theta = E[Y_2]$ for the
`Parametric 1` model and the `Outcome Robust` model respectively. Define 
the estimated coefficients to be $\hat c_j^{(P)}$ and $\hat c_j^{(OR)}$ where 
$j = 1, \dots 9$ for the `Parametric 1` and `Outcome Robust` models
respectively. The one can compute the SSE with the following:

$$ SSE = n^{-1} \sum_{i = 1}^n (\hat y_{2i} - \theta)^2 \text{ where }
\hat y_{2i} = \sum_{j = 1}^9 \hat c_j \hat \gamma_j $$

and $\hat \gamma_0 := \hat \gamma_{00} = \frac{\delta_{00i}}{\pi_{00}} E[Y_2
\mid x_i]$, $\hat \gamma_1 := \hat \gamma_{11} = \frac{\delta_{10i}}{\pi_{10}} E[Y_2
\mid x_i, y_{1i}]$, etc. Note that this is different from the previous estimate
of the variance which used the Monte Carlo variance (standard deviation) defined
by $$ \frac{1}{n - 1} \sum_{b = 1}^B (\hat \theta_b - \bar \theta_B)^2 $$ where
$B$ is the number of Monte Carlo estimates and $\bar \theta_B = \frac{1}{B}
\sum_{b = 1}^B \hat \theta_b$.

Likewise, one can compute the degrees of freedom by noticing that each model has
a degrees of freedom equal to nine minus the number of constraints. This means 
that we can compute the model degrees of freedom and error degrees of freedom
for each model type, which we do in @tbl-df

| Model | Model Degrees of Freedom | Error Degrees of Freedom |
| -----| ------------------------ | ------------------------ |
| Parametric 1    | $9 - 1 = 8$ | $n - 1 - 8 = n - 9$ |
| Parametric 2    | $9 - 1 = 8$ | $n - 1 - 8 = n - 9$ |
| Outcome Robust  | $9 - 4 = 5$ | $n - 1 - 5 = n - 6$ |
| Response Robust | $9 - 4 = 5$ | $n - 1 - 5 = n - 6$ |
| Double Robust   | $9 - 8 = 1$ | $n - 1 - 1 = n - 2$ |
: This table displays the degrees of freedom for each model {#tbl-df}

So continuing our first example, if $SSE^{(P)} = 993$ and $SSE^{(OR)} = 3428$ 
with $n = 1000$ then the F statistic is

$$F = \frac{(3428 - 993) / (3)}{993 / (1000 - 9)} = 810.$$

The critical value we want to compare this with is the $0.95$ quantile of 
$F_{3, 993}$ which is $2.61$. So there *is* a significant difference 
between the fit of these two model. (The p-value is basically zero.)

### 2024-01-22 Notes/note.tex
Added Quarto document with semiparametric constraint simulation.

\documentclass[12pt]{article}

\usepackage{amsmath, amssymb, mathrsfs, fancyhdr}
\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
\usepackage{biblatex}
\usepackage{booktabs}
\usepackage{float}

\addbibresource{references.bib}

\hypersetup{colorlinks = true, urlcolor = black}

\headheight     15pt
\topmargin      -1.5cm   % read Lamport p.163
\oddsidemargin  -0.04cm  % read Lamport p.163
\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
\textwidth      16.59cm
\textheight     23.94cm
\parskip         7.2pt   % sets spacing between paragraphs
\parindent         0pt   % sets leading space for paragraphs
\pagestyle{empty}        % Uncomment if don't want page numbers
\pagestyle{fancyplain}

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}

\begin{document}

%\lhead{Caleb Leedy}
\title{A note on optimal estimation under non-monotone missingness by design }
\author{Jae-Kwang Kim}
%\chead{STAT 615 - Advanced Bayesian Methods}
%\rhead{Page \thepage\ of \pageref{LastPage}}
%\rhead{April 4, 2023}
\maketitle

\begin{itemize}
%\item The sample $A$ is partitioned as $A= A_{11} \cup A_{10} \cup A_{01} \cup A_{00}$.

\item Define 
$$ \delta_{i1} =
\left\{ \begin{array}{ll}
 1 & \mbox{ if } Y_{1i} \mbox{ is observed} \\
 0 & \mbox{ otherwise}
\end{array}
\right. 
$$
and, similarly, we can define 
$\delta_{2i} $.

\item We  define $A_{01}= \{ i ; \delta_{1i} =0 \mbox{ and } \delta_{2i}=1\}$.

\item We are interested in estimating $\theta=E(Y_2)$. Note that, if we ignore $A_2$, 
then the missingness pattern is monotone and the following three-phase regression estimator can be used. 
$$
\hat{\theta}_1 = \frac{1}{n} \sum_{i=1}^n g_1^* (x_i) + \frac{1}{n} \sum_{i=1}^n \frac{\delta_{1i}}{ \pi_{1i}} \left\{ g_2^* (x_i, y_{1i}) - g_1^*(x_i) \right\} + \frac{1}{n} \sum_{i=1}^n \frac{\delta_{1i} \delta_{2i} }{ \pi_{12i}} \left\{ y_{2i} - g_2^*(x_i, y_i) \right\} $$
where 
$ g_1^*(x)= E(Y_2 \mid x)$ and $g_2^*(x,y_1) = E(Y_2 \mid x, y_1)$.

\item Now, to incorporate the information in $A_{01}$, we consider the following two-phase regression estimator 
$$
\hat{\theta}_2 = \frac{1}{n} \sum_{i=1}^n g_1^* (x_i) + \frac{1}{n} \sum_{i=1}^n \frac{(1-\delta_{1i} )\delta_{2i} }{ \pi_{2i} -\pi_{12i}} \left\{ y_{2i} - g_1^*(x_i) \right\} $$

\item The two estimators are both unbiased. We can consider 
$$ \hat{\theta}_\alpha = \alpha \hat{\theta}_1 + (1- \alpha) \hat{\theta}_2 $$
and obtain the optimal choice of $\alpha$ that minimizes the variance.
\item Please check Section 11.4 of the sampling book that I wrote. 
\end{itemize}

\end{document}

### 2024-01-22 Notes/semiconstraints.qmd
Added Quarto document with semiparametric constraint simulation.

---
title: "Understanding Semiparametric Constraints"
author:
  - name: Caleb Leedy
date: today
date-format: "D MMMM YYYY"
format:
  pdf:
    include-in-header: "latex_header.tex"
    keep-tex: true
---

```{r}
#| label: global variables
#| echo: false
B <- 1000

```

# Introduction {#sec-intro}

The goal of this report is to understand how additional constraints effect
semiparametric models. 
Inspired by Dr. Fuller's note, I set out to see if we could assess the amount 
of efficiency loss due to using semiparametric inference. First, let's define 
some notation.

| Segment | Quantity of Interest | Estimator | Coefficient |
| --------|----------------------| --------- | ------------ |
| $A_{00}$ | $E[g \mid X]$           | $\hat \gamma_{11}$ | $c_{11}$ |
| $A_{10}$ | $E[g \mid X]$           | $\hat \gamma_{21}$ | $c_{21}$ |
| $A_{10}$ | $E[g \mid X, Y_1]$      | $\hat \gamma_{22}$ | $c_{22}$ |
| $A_{01}$ | $E[g \mid X]$           | $\hat \gamma_{31}$ | $c_{31}$ |
| $A_{01}$ | $E[g \mid X, Y_2]$      | $\hat \gamma_{33}$ | $c_{33}$ |
| $A_{11}$ | $E[g \mid X]$           | $\hat \gamma_{41}$ | $c_{41}$ |
| $A_{11}$ | $E[g \mid X, Y_1]$      | $\hat \gamma_{42}$ | $c_{42}$ |
| $A_{11}$ | $E[g \mid X, Y_2]$      | $\hat \gamma_{43}$ | $c_{43}$ |
| $A_{11}$ | $E[g \mid X, Y_1, Y_2]$ | $\hat \gamma_{44}$ | $c_{44}$ |
: This table shows the relationship between different quantities of
interest in each segment, their estimators, and the corresponding coefficients
of their estimators. {#tbl-reftab}

This is a similar framework to understanding the class of estimators 
that we discussed in \verb|optest_update.pdf|, where we write the estimator 
as a weighted average of different expectation with functional coefficients.

$$\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
\mid X, Y_2].$$

However, now we write the estimator as

$$\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}$$

where $\hat \gamma_{kt} = \frac{\delta_{ij}}{\pi_{ij}}E[g \mid G_{ij}(X, Y_1,
Y_2)]$ and $ij$ corresponds to the segment $A_{ij}$ associated with $\hat
\gamma_{kt}$ in @tbl-reftab.

# Methods {#sec-methods}

As noted by Dr. Fuller, we can understand parametric estimation as solving 
the following constrained optimization problems:

\begin{align*}
  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
  \sum_{k,t} c_{kt} = 1 \\
       &\text{or} \\
  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
  \sum_{(k,t): (k,t) \neq (4, 4)} c_{kt} = 0 \text{ and } c_{44} = 1. \\
\end{align*}

To be robust to the outcome model, we can construct a similar optimization
problem with different constraints:

\begin{align*}
  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
  c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0,\\
       &\text{and } c_{44} = 1.
\end{align*}

Likewise to be robust to the response model, we can have the problem:

\begin{align*}
  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
  c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \\
       &\text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}.
\end{align*}

To be double robust (robust to the outcome and response model) we can combine 
the last two constraints. This is summarized in @tbl-mods.

| Type | Constraints |
| -----| ------------|
| Parametric 1 | $\sum_{k, t} c_{kt} = 1$ |
| Parametric 2 |  $\sum_{k, t: (k, t) \neq (4, 4)} c_{kt} = 0, c_{44} = 1$ |
| Outcome Robust | $c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, \text{ and } c_{44} = 1.$ |
| Response Robust | $c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}$ |
| Double Robust | $c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, c_{44} = 1, c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}$ |
: This table identifies the different constraints for each model type. {#tbl-mods}

# Simulation Study {#sec-simulations}

## Simulation 1 {.unnumbered}

We use the following simulation setup

\begin{align*}
  \begin{bmatrix} x \\ e_1 \\ e_2 \end{bmatrix} 
  &\stackrel{ind}{\sim} N\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}\right) \\
  y_1 &= x + e_1 \\
  y_2 &= \theta + x + e_2 \\
\end{align*}

Furthermore, $\pi_{11} = 0.2$ and $\pi_{00} = 0.3, \pi_{10} = 0.4$, and 
$\pi_{01} = 0.1$.
The goal of this simulation study is to find $\theta = E[Y_2]$. In other words,
$g(Z) = Y_2$. There are several algorithms for comparison which are defined 
as the following:

\begin{align*}
  Oracle &= n^{-1} \sum_{i = 1}^n g(Z_i)\\
  CC &= \frac{\sum_{i = 1}^n \delta_{11} g(Z_i)}{\sum_{i = 1}^n \delta_{11}} \\
  IPW &= \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g(Z_i)\\
\end{align*}

Furthermore, we include four existing estimators that we have already 
proposed in the past: WLS, Prop, PropInd, and SemiDelta.

The estimator WLS is a a weight least square estimator derived in the 
following manner. We now consider a normal model:

$$ 
  \begin{pmatrix}
    x_i \\ e_{1i} \\ e_{2i}
  \end{pmatrix} \stackrel{ind}{\sim}
  N \left(
  \begin{bmatrix}
    0 \\ 0 \\ 0
  \end{bmatrix},
  \begin{bmatrix}
    1 & 0 & 0 \\
    0 & \sigma_{11} & \sigma_{12} \\ 
    0 & \sigma_{12} & \sigma_{22}
  \end{bmatrix},
  \right)
$$

and define $y_{1i} = \theta_1 + x_i + e_{1i}$ and $y_{2i} = \theta_2 + x_i
+ e_{2i}$. Then $b_1 = b_2 = 1$. We define $\bar z_k^{(ij)}$ as the mean of
$y_k$ in segment $A_{ij}$. This means that we have means $\bar z_1^{(11)}$,
$\bar z_2^{(11)}$, $\bar z_1^{(10)}$, and $\bar z_2^{(01)}$. Let $W = [\bar
z_1^{(11)}, \bar z_2^{(11)}, \bar z_1^{(10)}, \bar z_2^{(01)}]'$, then for
$n_{ij} = |A_{ij}|$, we have

$$Z - M \mu \sim N(\vec 0, V)$$

where

$$M = 
  \begin{bmatrix}
    1 & 0 \\
    0 & 1 \\
    1 & 0 \\
    0 & 1 \\
  \end{bmatrix}
  \text{ and }
  V = 
  \begin{bmatrix}
    \frac{\sigma_{11}}{n_{11}} & \frac{\sigma_{12}}{n_{11}} & 0 & 0 \\
    \frac{\sigma_{12}}{n_{11}} & \frac{\sigma_{22}}{n_{11}} & 0 & 0 \\
    0 & 0 & \frac{\sigma_{11}}{n_{10}} & 0 \\
    0 & 0 & 0 & \frac{\sigma_{22}}{n_{01}} \\
  \end{bmatrix}.
$$

Thus, the BLUE for $\mu = [\mu_1, \mu_2]'$ is

$$\hat \mu = (M' V^{-1} M)^{-1} M' V^{-1} W.$$

Hence, WLS is $\mu_2$ as $g(X, Y_1, Y_2) = Y_2$.

The remaining three estimators are drived from the following expression
where the values for $\beta$ are provided in @tbl-beta. These 
are good estimators to compare to because Prop is the original proposed 
estimator. PropInd has the same form as Prop except the values for $\beta$ 
is different. PropInd shares the same values of $\beta$ as all of the new 
models with constraints. SemiDelta is useful because it is the best 
estimator in general so far.

$$
\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + 
\beta_0(\delta, c_0)E[g(Z) \mid X] + 
\beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + 
\beta_2(\delta, c_2) E[g(Z)
\mid X, Y_2].
$$

| Estimator | $\beta_0(\delta, c_0)$ | $\beta_1(\delta, c_1)$ |
| --------- | ---------------------- | ---------------------- |
  Prop | $\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)$ | $\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)$ |
  PropInd | $\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)$ | $\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)$ |
  SemiDelta | $c_0\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{00}}{\pi_{00}}\right)$ | $c_1\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{10}}{\pi_{10}}\right)$ |

: This table displays the values of $\beta$ for different estimator types. {#tbl-beta}

In the simulation results in @tbl-sim1, the new results have the 
same label as the value from the `Type` column in @tbl-mods.

```{r}
#| label: libraries
#| echo: false
#| warning: false
#| message: false

library(dplyr)
library(tidyr)
library(stringr)
library(CVXR)
library(ggplot2)

library(doParallel)
library(doRNG)
library(parallelly)

source("R/opt_est.R")
source("Simulations/T3Assess_Semiparametric_Cost/compare_algs.R")

```

```{r}
#| label: mcmc E[g] = Y2
#| echo: false

clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
registerDoParallel(clust)

n_obs <- 1000
true_theta <- 5 
cov_e1e2 <- 0.5

mc_theta <-
  foreach(iter = 1:B,
          .options.RNG = 1,
          .packages = c("dplyr", "stringr", "CVXR")) %dorng% {

    # Generate Data
    # df <- gen_optsim_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)
    df <- gen_unbal_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)

    # Get Estimates
    oracle_est <- mean(df$Y2)
    oraclex_est <- mean(df$Y2 - df$X)
    ipw_est <- 
      filter(df, delta_1 == 1, delta_2 == 1) |>
      mutate(g_est = Y2 / prob_11) |>
      pull(g_est) |>
      sum() / nrow(df)
    cc_est <- 
      filter(df, delta_2 == 1) |>
      pull(Y2) |>
      mean()
    # em_est <- em_optsim(df)
    wls_est <- opt_lin_est(df, cov_y1y2 = cov_e1e2)
    wlsalt_est <- comb_lin_est_lin(df, theta2 = true_theta, cov_e1e2 = cov_e1e2)
    prop_est <- prop_nmono_est(df)
    propind_est <- prop_nmono_est(df, prop_ind = TRUE)
    propopt_est <- opt_theta_c(df)
    propoptdef_est <- opt_theta_c(df, est = "default")
    semiopt_est <- opt_semi_est(df)
    semidef_est <- opt_semi_est(df, est = "default")
    semidel_est <- opt_delta_c(df)
    semideldef_est <- opt_delta_c(df, est = "default")

    conalg_para <- min_var_alg(df, mod = "para", cov_e1e2 = cov_e1e2)
    conalg_sumzero <- min_var_alg(df, mod = "sumzero", cov_e1e2 = cov_e1e2)
    conalg_outrob <- min_var_alg(df, mod = "out_rob", cov_e1e2 = cov_e1e2)
    conalg_resprob <- min_var_alg(df, mod = "resp_rob", cov_e1e2 = cov_e1e2)
    conalg_doubrob <- min_var_alg(df, mod = "double_rob", cov_e1e2 = cov_e1e2)

    return(tibble(oracle = oracle_est,
                  oraclex = oraclex_est,
                  cc = cc_est,
                  ipw = ipw_est,
                  # em = em_est,
                  wls = wls_est,
                  wlsalt = wlsalt_est,
                  prop = prop_est,
                  propind = propind_est,
                  propopt = propopt_est,
                  propoptdef = propoptdef_est,
                  semiopt = semiopt_est,
                  semidef = semidef_est,
                  semidel = semidel_est,
                  semideldef = semideldef_est,
                  conpara = conalg_para$theta_est,
                  conzero = conalg_sumzero$theta_est,
                  conout = conalg_outrob$theta_est,
                  conresp = conalg_resprob$theta_est,
                  condoub = conalg_doubrob$theta_est
    ))
  } |>
  bind_rows()

stopCluster(clust)

```

```{r}
#| label: tbl-sim1
#| echo: false
#| tbl-cap: |
#|    Results from Simulation 1. True values: $\theta = 5, \rho = 0.5$. 
#|    The test conducted for the T-statistic and P-value is a two sample test 
#|    to see if the estimator is unbiased. This simulation uses 
#|    $Y_1 = x + \varepsilon_1$, $Y_2 = \theta + x + \varepsilon_2$ where 
#|    $X \sim N(0, 1)$ and $(\varepsilon_1, \varepsilon_2)$ come from a mean 
#|    zero bivariate normal distribution with unit variance and covariance 
#|    $\rho$. The segments are unbalanced with: $\pi_{11} = 0.2$, 
#|    $\pi_{10} = 0.4$, $\pi_{01} = 0.1$, and $\pi_{00} = 0.3$.

mc_theta |>
  summarize(
    bias_oracle = mean(oracle) - true_theta,
    bias_cc = mean(cc) - true_theta,
    bias_ipw = mean(ipw) - true_theta,
    bias_wls = mean(wls) - true_theta,
    bias_prop = mean(prop) - true_theta,
    bias_semidel = mean(semidel) - true_theta,
    bias_conpara = mean(conpara) - true_theta,
    bias_conzero = mean(conzero) - true_theta,
    bias_conout = mean(conout) - true_theta,
    bias_conresp = mean(conresp) - true_theta,
    bias_condoub = mean(condoub) - true_theta,
    sd_oracle = sd(oracle),
    sd_cc = sd(cc),
    sd_ipw = sd(ipw),
    sd_wls = sd(wls),
    sd_prop = sd(prop),
    sd_semidel = sd(semidel),
    sd_conpara = sd(conpara),
    sd_conzero = sd(conzero),
    sd_conout = sd(conout),
    sd_conresp = sd(conresp),
    sd_condoub = sd(condoub),
    tstat_oracle = (mean(oracle) - true_theta) / sqrt(var(oracle) / B),
    tstat_cc = (mean(cc) - true_theta) / sqrt(var(cc) / B),
    tstat_ipw = (mean(ipw) - true_theta) / sqrt(var(ipw) / B),
    tstat_wls = (mean(wls) - true_theta) / sqrt(var(wls) / B),
    tstat_prop = (mean(prop) - true_theta) / sqrt(var(prop) / B),
    tstat_semidel = (mean(semidel) - true_theta) / sqrt(var(semidel) / B),
    tstat_conpara = (mean(conpara) - true_theta) / sqrt(var(conpara) / B),
    tstat_conzero = (mean(conzero) - true_theta) / sqrt(var(conzero) / B),
    tstat_conout = (mean(conout) - true_theta) / sqrt(var(conout) / B),
    tstat_conresp = (mean(conresp) - true_theta) / sqrt(var(conresp) / B),
    tstat_condoub = (mean(condoub) - true_theta) / sqrt(var(condoub) / B)
  ) |>
  tidyr::pivot_longer(cols = everything(),
                      names_to = c(".value", "Algorithm"),
                      names_pattern = "(.*)_(.*)") |>
  mutate(pval = pt(-abs(tstat), df = B)) |>
  mutate(Algorithm = case_when(
    Algorithm == "oracle" ~ "Oracle",
    Algorithm == "cc" ~ "CC",
    Algorithm == "ipw" ~ "IPW",
    Algorithm == "wls" ~ "WLS",
    Algorithm == "prop" ~ "$\\hat \\theta_{prop}$",
    Algorithm == "semidel" ~ "$\\hat \\theta_{\\delta}$",
    Algorithm == "conpara" ~ "Parametric 1",
    Algorithm == "conzero" ~ "Parametric 2",
    Algorithm == "conout" ~ "Outcome Robust",
    Algorithm == "conresp" ~ "Response Robust",
    Algorithm == "condoub" ~ "Double Robust",
    TRUE ~ "ERROR"
  )) |>
  knitr::kable(digits = 3)

```

## Simulation 2 {.unnumbered}

Next, we used the same simulation setup except that we are interested in 
estimating $\theta = E[g] = E[Y_1^2 Y_2]$. We do not use the WLS estimator 
because it does not work.

```{r}
#| label: mcmc E[g] = Y1^2 Y2
#| echo: false

clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
registerDoParallel(clust)

mc_theta <-
  foreach(iter = 1:B,
          .options.RNG = 1,
          .packages = c("dplyr", "stringr", "CVXR")) %dorng% {

    # g_fun <- "Y1^2Y2"

    # Generate Data
    # df <- gen_optsim_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)
    df <- gen_unbal_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)

    # Get Estimates
    oracle_est <- mean(df$Y1^2 * df$Y2)
    oraclex_est <- mean((df$Y1)^2 * (df$Y2 - df$X))
    ipw_est <- 
      filter(df, delta_1 == 1, delta_2 == 1) |>
      mutate(g_f = Y1^2 * Y2) |>
      mutate(g_est = g_f / prob_11) |>
      pull(g_est) |>
      sum() / nrow(df)
    cc_est <- 
      filter(df, delta_1 == 1, delta_2 == 1) |>
      mutate(g_f = Y1^2 * Y2) |>
      pull(g_f) |>
      mean()
    wls_est <- comb_lin_est(df, gfun = "Y1^2 * Y2", cov_e1e2 = cov_e1e2)
    wlstt_est <- 
      comb_lin_est(df, gfun = "Y1^2 * Y2", cov_e1e2 = cov_e1e2, theta2 = true_theta)
    prop_est <- prop_nmono_est(df, gfun = "Y1^2 * Y2", pow = 3)
    propopt_est <- opt_theta_c(df, gfun = "Y1^2 * Y2", pow = 3)
    semiopt_est <- opt_semi_est(df, gfun = "Y1^2 * Y2", pow = 3)
    semidef_est <- opt_semi_est(df, gfun = "Y1^2 * Y2", est = "default", pow = 3)
    semidel_est <- opt_delta_c(df, gfun = "Y1^2 * Y2", pow = 3)
    semideldef_est <- opt_delta_c(df, gfun = "Y1^2 * Y2", est = "default", pow = 3)

    conalg_para <- 
      min_var_alg(df, gfun = "Y1^2 * Y2", mod = "para", cov_e1e2 = cov_e1e2)
    conalg_sumzero <- 
      min_var_alg(df, gfun = "Y1^2 * Y2", mod = "sumzero", cov_e1e2 = cov_e1e2)
    conalg_outrob <- 
      min_var_alg(df, gfun = "Y1^2 * Y2", mod = "out_rob", cov_e1e2 = cov_e1e2)
    conalg_resprob <- 
      min_var_alg(df, gfun = "Y1^2 * Y2", mod = "resp_rob", cov_e1e2 = cov_e1e2)
    conalg_doubrob <- 
      min_var_alg(df, gfun = "Y1^2 * Y2", mod = "double_rob", cov_e1e2 = cov_e1e2)

    return(tibble(oracle = oracle_est,
                  oraclex = oraclex_est,
                  cc = cc_est,
                  ipw = ipw_est,
                  wls = wls_est,
                  wlstt = wlstt_est,
                  prop = prop_est,
                  propopt = propopt_est,
                  semiopt = semiopt_est,
                  semidef = semidef_est,
                  semidel = semidel_est,
                  semideldef = semideldef_est,
                  conpara = conalg_para$theta_est,
                  conzero = conalg_sumzero$theta_est,
                  conout = conalg_outrob$theta_est,
                  conresp = conalg_resprob$theta_est,
                  condoub = conalg_doubrob$theta_est
    ))
  } |> 
  bind_rows()

stopCluster(clust)

```

```{r}
#| label: tbl-sim2
#| echo: false
#| tbl-cap: |
#|    Results from Simulation 2. True values: $\theta = 10, \rho = 0.5$. 
#|    The test conducted for the T-statistic and P-value is a two sample test 
#|    to see if the estimator is unbiased. This simulation uses the same
#|    setup as Simulation 1 with
#|    $Y_1 = x + \varepsilon_1$, $Y_2 = \theta + x + \varepsilon_2$ where 
#|    $X \sim N(0, 1)$ and $(\varepsilon_1, \varepsilon_2)$ come from a mean 
#|    zero bivariate normal distribution with unit variance and covariance 
#|    $\rho$. The segments are unbalanced with: $\pi_{11} = 0.2$, 
#|    $\pi_{10} = 0.4$, $\pi_{01} = 0.1$, and $\pi_{00} = 0.3$.

true_g <- 2 * true_theta

mc_theta |>
  summarize(
    bias_oracle = mean(oracle) - true_g,
    bias_oraclex = mean(oraclex) - true_g,
    bias_cc = mean(cc) - true_g,
    bias_ipw = mean(ipw) - true_g,
    bias_wls = mean(wls) - true_g,
    bias_wlstt = mean(wlstt) - true_g,
    bias_prop = mean(prop) - true_g,
    bias_propopt = mean(propopt) - true_g,
    bias_semiopt = mean(semiopt) - true_g,
    bias_semidef = mean(semidef) - true_g,
    bias_semidel = mean(semidel) - true_g,
    bias_semideldef = mean(semideldef) - true_g,
    bias_conpara = mean(conpara) - true_g,
    bias_conzero = mean(conzero) - true_g,
    bias_conout = mean(conout) - true_g,
    bias_conresp = mean(conresp) - true_g,
    bias_condoub = mean(condoub) - true_g,
    sd_oracle = sd(oracle),
    sd_oraclex = sd(oraclex),
    sd_cc = sd(cc),
    sd_ipw = sd(ipw),
    sd_wls = sd(wls),
    sd_wlstt = sd(wlstt),
    sd_prop = sd(prop),
    sd_propopt = sd(propopt),
    sd_semiopt = sd(semiopt),
    sd_semidef = sd(semidef),
    sd_semidel = sd(semidel),
    sd_semideldef = sd(semideldef),
    sd_conpara = sd(conpara),
    sd_conzero = sd(conzero),
    sd_conout = sd(conout),
    sd_conresp = sd(conresp),
    sd_condoub = sd(condoub),
    tstat_oracle = (mean(oracle) - true_g) / sqrt(var(oracle) / B),
    tstat_oraclex = (mean(oraclex) - true_g) / sqrt(var(oraclex) / B),
    tstat_cc = (mean(cc) - true_g) / sqrt(var(cc) / B),
    tstat_ipw = (mean(ipw) - true_g) / sqrt(var(ipw) / B),
    tstat_wls = (mean(wls) - true_g) / sqrt(var(wls) / B),
    tstat_wlstt = (mean(wlstt) - true_g) / sqrt(var(wlstt) / B),
    tstat_prop = (mean(prop) - true_g) / sqrt(var(prop) / B),
    tstat_propopt = (mean(propopt) - true_g) / sqrt(var(propopt) / B),
    tstat_semiopt = (mean(semiopt) - true_g) / sqrt(var(semiopt) / B),
    tstat_semidef = (mean(semidef) - true_g) / sqrt(var(semidef) / B),
    tstat_semidel = (mean(semidel) - true_g) / sqrt(var(semidel) / B),
    tstat_semideldef = (mean(semideldef) - true_g) / sqrt(var(semideldef) / B),
    tstat_conpara = (mean(conpara) - true_g) / sqrt(var(conpara) / B),
    tstat_conzero = (mean(conzero) - true_g) / sqrt(var(conzero) / B),
    tstat_conout = (mean(conout) - true_g) / sqrt(var(conout) / B),
    tstat_conresp = (mean(conresp) - true_g) / sqrt(var(conresp) / B),
    tstat_condoub = (mean(condoub) - true_g) / sqrt(var(condoub) / B)
  ) |>
  tidyr::pivot_longer(cols = everything(),
               names_to = c(".value", "Algorithm"),
               names_pattern = "(.*)_(.*)") |>
  mutate(pval = pt(-abs(tstat), df = B)) |>
  filter(Algorithm %in% c("oracle", "cc", "ipw", "prop", "semidel", "conpara",
           "conzero", "conout", "conresp", "condoub")) |>
  mutate(Algorithm = case_when(
    Algorithm == "oracle" ~ "Oracle",
    Algorithm == "cc" ~ "CC",
    Algorithm == "ipw" ~ "IPW",
    Algorithm == "prop" ~ "$\\hat \\theta_{prop}$",
    Algorithm == "semidel" ~ "$\\hat \\theta_{\\delta}$",
    Algorithm == "conpara" ~ "Parametric 1",
    Algorithm == "conzero" ~ "Parametric 2",
    Algorithm == "conout" ~ "Outcome Robust",
    Algorithm == "conresp" ~ "Response Robust",
    Algorithm == "condoub" ~ "Double Robust",
    TRUE ~ "ERROR"
  )) |>
  knitr::kable(digits = 3)

```

## Discussion {.unnumbered}

There has been concern about why the constrained results are the same in
@tbl-sim1. This section aims to address these concerns.

```{r}
#| label: MCMC extended Sim 1
#| echo: false

clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
registerDoParallel(clust)

n_obs <- 1000
true_theta <- 5 
cov_e1e2 <- 0.5

mc_theta <-
  foreach(iter = 1:B,
          .options.RNG = 1,
          .packages = c("dplyr", "stringr", "CVXR", "tidyr")) %dorng% {

    # Generate Data
    # df <- gen_optsim_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)
    df <- gen_unbal_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)

    # Get Estimates
    conalg_para <- min_var_alg(df, mod = "para", cov_e1e2 = cov_e1e2)
    conalg_sumzero <- min_var_alg(df, mod = "sumzero", cov_e1e2 = cov_e1e2)
    conalg_outrob <- min_var_alg(df, mod = "out_rob", cov_e1e2 = cov_e1e2)
    conalg_resprob <- min_var_alg(df, mod = "resp_rob", cov_e1e2 = cov_e1e2)
    conalg_doubrob <- min_var_alg(df, mod = "double_rob", cov_e1e2 = cov_e1e2)

    return(
      tibble(conpara = conalg_para$theta_est,
             conzero = conalg_sumzero$theta_est,
             conout = conalg_outrob$theta_est,
             conresp = conalg_resprob$theta_est,
             condoub = conalg_doubrob$theta_est,
             parac = conalg_para$c_hat,
             zeroc = conalg_sumzero$c_hat,
             outc = conalg_outrob$c_hat,
             respc = conalg_resprob$c_hat,
             doubc = conalg_doubrob$c_hat,
             iter = iter,
             c_i = 1:9
      ) |> 
      group_by(conpara, conzero, conout, conresp, condoub) |> 
      nest()
    )
  } |>
  bind_rows()

stopCluster(clust)

```

Interestingly enough, the optimization algorithm does not have a lot of
variability in the output. The variance for each coefficient is found 
in @tbl-estvar. Since these values are all basically zero, we can see that
for each algorithm we are getting basically the same coefficients each
iteration. The coefficients that we see are displayed in @tbl-estc.

```{r}
#| label: tbl-estvar
#| echo: false
#| tbl-cap: |
#|    This table displays the variance of each coefficient for each estimator.
#|    A number close to zero means that the optimization algorithm chose 
#|    very similar values for each Monte Carlo simulation. A value of exactly 
#|    zero indicated that the optimization algorithm chose the exact same
#|    number.

c_tab <-
  bind_rows(mc_theta$data) |>
  pivot_longer(cols = parac:doubc, names_to = "estimator", values_to = "c_val") |>
  mutate(c_i = paste0("c_", c_i)) |>
  pivot_wider(names_from = "c_i", values_from = "c_val")

# Variance of each coefficient per estimator
alg_vec <- c("Para. 1", "Para. 2", "Outcome",
             "Response", "Double")
names(alg_vec) <- c("parac", "zeroc", "outc", "respc", "doubc")

ord_vec <- 1:5
names(ord_vec) <- c("parac", "zeroc", "outc", "respc", "doubc")

c_tab |>
  group_by(estimator) |>
  summarize(across(starts_with("c_"), var), .groups = "drop") |>
  rename_with(~paste0("$", .x, "$"), starts_with("c_")) |>
  mutate(order = ord_vec[estimator]) |>
  mutate(estimator = alg_vec[estimator]) |>
  rename(Estimator = estimator) |>
  arrange(order) |>
  select(-order) |>
  knitr::kable(digits = 3)

```

```{r}
#| label: tbl-estc
#| echo: false
#| tbl-cap: This table displays the estimated coefficients for each algorithm on
#|  iteration 1. Since the variance between iterations is small, these are
#|  reasonable for subsequent iterations.

c_tab |>
  filter(iter == 1) |>
  rename_with(~paste0("$", .x, "$"), starts_with("c_")) |>
  mutate(estimator = alg_vec[estimator]) |>
  rename(Estimator = estimator) |>
  select(-iter) |>
  knitr::kable(digits = 3)

```

```{r}
#| echo: false

#' get_f_statistic is a function to compute the F-statistic between two of 
#' the constrained models.
#'
#' @param c_tab - A table with values for the estimated c coefficients
#' @param reduced_mod - A string indicating which model is the reduced model
#' @param full_mod - A string indicating which model is the full model
#' @param df - A data set simulated from the same data generating process that 
#'    the c_tab coefficients were estimated from.
#'
#' @note We have previously shown that the variance of the estimated c_tab 
#' variables is close to or exactly zero. Hence, conducting a new simulation 
#' should not effect the results of the model much.
get_f_statistic <- function(c_tab, reduced_mod, full_mod,
                            df, theta = true_theta, cov_e1e2 = cov_e1e2) {

  vars_lst <- list(reduced_mod, full_mod)
  # SSE reduced 
  red_cvec <-
    c_tab |>
    filter(iter == 1) |>
    filter(.data[["estimator"]] == vars_lst[[1]]) |>
    select(starts_with("c_")) |>
    as.numeric()

  # SSE full
  full_cvec <-
    c_tab |>
    filter(iter == 1) |>
    filter(.data[["estimator"]] == vars_lst[[2]]) |>
    select(starts_with("c_")) |>
    as.numeric()

  # Constructing gam_vec
  Eg <- theta
  Egx <- theta + df$X 
  Egxy1 <- theta + df$X + cov_e1e2 * (df$Y1 - df$X)
  Egxy2 <- df$Y2
  EEg2 <- theta^2 + 2
  EEgx2 <- theta^2 + 1
  EEgxy12 <- theta^2 + 1 + cov_e1e2^2
  EEgxy22 <- theta^2 + 2
  EEgxy1Egxy2 <- theta^2 + 1 + cov_e1e2^2

  g_11 <- df$delta_00 / df$prob_00 * (Egx)
  g_21 <- df$delta_10 / df$prob_10 * (Egx)
  g_22 <- df$delta_10 / df$prob_10 * (Egxy1)
  g_31 <- df$delta_01 / df$prob_01 * (Egx)
  g_33 <- df$delta_01 / df$prob_01 * (Egxy2)
  g_41 <- df$delta_11 / df$prob_11 * (Egx)
  g_42 <- df$delta_11 / df$prob_11 * (Egxy1)
  g_43 <- df$delta_11 / df$prob_11 * (Egxy2)
  g_44 <- df$delta_11 / df$prob_11 * (df$Y2)

  gam_mat <- base::cbind(g_11, g_21, g_22, g_31, g_33, g_41, g_42, g_43, g_44)

  # SSE
  sse_red <- sum((as.numeric(gam_mat %*% t(t(red_cvec))) - theta)^2)
  sse_full <- sum((as.numeric(gam_mat %*% t(t(full_cvec))) - theta)^2)

  # Error Degrees of Freedom
  df_vec <- nrow(df) - 1 - c(9 - 1, 9 - 2, 9 - 4, 9 - 4, 9 - 8)
  names(df_vec) <- c("parac", "zeroc", "outc", "respc", "doubc")

  df_red <- df_vec[reduced_mod]
  df_full <- df_vec[full_mod]

  fstat <- ((sse_red - sse_full) / (df_red - df_full)) / (sse_full / df_full)

  pval <- 1 - pf(fstat, df_red - df_full, df_full)
  return(list(f = fstat, p = pval))

}

```

```{r}
#| label: tbl-fstat
#| echo: false
#| tbl-cap: This table computes the F-statistic and P-value for an F-test
#|    comparing the `Full Model` with a nested model in `Reduced Model`. The
#|    computation of the F-statistic is found in the `get_f_statistic` function.

df <- gen_unbal_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)
rf_tab <- tibble(rmod = c("zeroc", "outc",  "respc", "doubc",
                 "outc",  "respc", "doubc", "doubc", "doubc"),
                 fmod = c("parac", "parac", "parac", "parac",
                 "zeroc", "zeroc", "zeroc", "outc",  "respc"))

fstat_tab <-
  purrr::pmap(rf_tab, function(rmod, fmod) {
    get_f_statistic(c_tab, rmod, fmod, df = df, cov_e1e2 = cov_e1e2)
  }) |>
  bind_rows()

change_tab_names <- c("Parametric 1", "Parametric 2", "Outcome Robust",
                      "Response Robust", "Double Robust")
names(change_tab_names) <- c("parac", "zeroc", "outc", "respc", "doubc")

rf_tab |>
  mutate(Fstat = fstat_tab$f) |>
  mutate(fmod = change_tab_names[fmod]) |>
  mutate(rmod = change_tab_names[rmod]) |>
  select(fmod, rmod, Fstat) |>
  rename(`Reduced Model` = rmod) |>
  rename(`Full Model` = fmod) |>
  mutate(`P-Value` = fstat_tab$p) |>
  knitr::kable(digits = 2)

```

From the analysis in @tbl-fstat, we can see that there are significant
differences between several of the models. Why is this different from the 
previous analysis in @tbl-sim1? The answer comes from how we compute the
variance versus the sum of squares of error. The variance is computed from each
estimator created in a Monte Carlo trial; whereas the sum of squares of error is
summed from each individuals estimate for the mean of $Y_2$. For more details
about how the F-test is computed see Appendix A and Appendix B.

{{< pagebreak >}}

# Appendix A: Computing the F-Test {.appendix}

To compute an F test we can use @eq-fstat.

$$ F = \frac{(SSE_{Reduced} - SSE_{Full}) / (DFE_{Reduced} - DFE_{Full})}{
SSE_{Full} / DFE_{Full}}. $$ {#eq-fstat}

Let's assess what this would mean to compare models `Parametric 1` and 
`Outcome Robust` from @tbl-mods.

| Type | Constraints |
| -----| ------------|
| Parametric 1 | $\sum_{k, t} c_{kt} = 1$ |
| Parametric 2 |  $\sum_{k, t: (k, t) \neq (4, 4)} c_{kt} = 0, c_{44} = 1$ |
| Outcome Robust | $c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, \text{ and } c_{44} = 1.$ |
| Response Robust | $c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}$ |
| Double Robust | $c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, c_{44} = 1, c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}$ |
: This table identifies the different constraints for each model type. {#tbl-mods}

Consider the case where we run estimate each model on a data set with $n = 1000$
observations, and define the following notation: let $\hat \theta^{(P)}$ and 
$\hat \theta^{(OR)}$ be the estimated values of $\theta = E[Y_2]$ for the
`Parametric 1` model and the `Outcome Robust` model respectively. Define 
the estimated coefficients to be $\hat c_j^{(P)}$ and $\hat c_j^{(OR)}$ where 
$j = 1, \dots 9$ for the `Parametric 1` and `Outcome Robust` models
respectively. The one can compute the SSE with the following:

$$ SSE = n^{-1} \sum_{i = 1}^n (\hat y_{2i} - \theta)^2 \text{ where }
\hat y_{2i} = \sum_{j = 1}^9 \hat c_j \hat \gamma_j $$

and $\hat \gamma_0 := \hat \gamma_{00} = \frac{\delta_{00i}}{\pi_{00}} E[Y_2
\mid x_i]$, $\hat \gamma_1 := \hat \gamma_{11} = \frac{\delta_{10i}}{\pi_{10}} E[Y_2
\mid x_i, y_{1i}]$, etc. Note that this is different from the previous estimate
of the variance which used the Monte Carlo variance (standard deviation) defined
by $$ \frac{1}{n - 1} \sum_{b = 1}^B (\hat \theta_b - \bar \theta_B)^2 $$ where
$B$ is the number of Monte Carlo estimates and $\bar \theta_B = \frac{1}{B}
\sum_{b = 1}^B \hat \theta_b$.

Likewise, one can compute the degrees of freedom by noticing that each model has
a degrees of freedom equal to nine minus the number of constraints. This means 
that we can compute the model degrees of freedom and error degrees of freedom
for each model type, which we do in @tbl-df

| Model | Model Degrees of Freedom | Error Degrees of Freedom |
| -----| ------------------------ | ------------------------ |
| Parametric 1    | $9 - 1 = 8$ | $n - 1 - 8 = n - 9$ |
| Parametric 2    | $9 - 1 = 8$ | $n - 1 - 8 = n - 9$ |
| Outcome Robust  | $9 - 4 = 5$ | $n - 1 - 5 = n - 6$ |
| Response Robust | $9 - 4 = 5$ | $n - 1 - 5 = n - 6$ |
| Double Robust   | $9 - 8 = 1$ | $n - 1 - 1 = n - 2$ |
: This table displays the degrees of freedom for each model {#tbl-df}

So continuing our first example^[SSE values might differ in @tbl-fstat slightly
depending on the actual data values.], if $SSE^{(P)} = 993$ and $SSE^{(OR)} =
3428$ with $n = 1000$ then the F statistic is

$$F = \frac{(3428 - 993) / (3)}{993 / (1000 - 9)} = 810.$$

The critical value we want to compare this with is the $0.95$ quantile of 
$F_{3, 993}$ which is $2.61$. So there *is* a significant difference 
between the fit of these two model. (The p-value is basically zero.) The 
code to do this automatically is in Appendix B.

{{< pagebreak >}}

# Appendix B: R Functions {.appendix}

```{r}
#| lst-label: lst-getfstat
#| lst-cap: The R code for estimating the F statistic.
#| echo: true

#' get_f_statistic is a function to compute the F-statistic between two of 
#' the constrained models.
#'
#' @param c_tab - A table with values for the estimated c coefficients
#' @param reduced_mod - A string indicating which model is the reduced model
#' @param full_mod - A string indicating which model is the full model
#' @param df - A data set simulated from the same data generating process that 
#'    the c_tab coefficients were estimated from.
#'
#' @note We have previously shown that the variance of the estimated c_tab 
#' variables is close to or exactly zero. Hence, conducting a new simulation 
#' should not effect the results of the model much.
get_f_statistic <- 
  function(c_tab, reduced_mod, full_mod, df, theta = true_theta, cov_e1e2) {

  vars_lst <- list(reduced_mod, full_mod)

  # SSE reduced 
  red_cvec <-
    c_tab |>
    filter(iter == 1) |>
    filter(.data[["estimator"]] == vars_lst[[1]]) |>
    select(starts_with("c_")) |>
    as.numeric()

  # SSE full
  full_cvec <-
    c_tab |>
    filter(iter == 1) |>
    filter(.data[["estimator"]] == vars_lst[[2]]) |>
    select(starts_with("c_")) |>
    as.numeric()

  # Constructing gam_vec
  Eg <- theta
  Egx <- theta + df$X 
  Egxy1 <- theta + df$X + cov_e1e2 * (df$Y1 - df$X)
  Egxy2 <- df$Y2
  EEg2 <- theta^2 + 2
  EEgx2 <- theta^2 + 1
  EEgxy12 <- theta^2 + 1 + cov_e1e2^2
  EEgxy22 <- theta^2 + 2
  EEgxy1Egxy2 <- theta^2 + 1 + cov_e1e2^2

  g_11 <- df$delta_00 / df$prob_00 * (Egx)
  g_21 <- df$delta_10 / df$prob_10 * (Egx)
  g_22 <- df$delta_10 / df$prob_10 * (Egxy1)
  g_31 <- df$delta_01 / df$prob_01 * (Egx)
  g_33 <- df$delta_01 / df$prob_01 * (Egxy2)
  g_41 <- df$delta_11 / df$prob_11 * (Egx)
  g_42 <- df$delta_11 / df$prob_11 * (Egxy1)
  g_43 <- df$delta_11 / df$prob_11 * (Egxy2)
  g_44 <- df$delta_11 / df$prob_11 * (df$Y2)

  gam_mat <- base::cbind(g_11, g_21, g_22, g_31, g_33, g_41, g_42, g_43, g_44)

  # SSE
  sse_red <- sum((as.numeric(gam_mat %*% t(t(red_cvec))) - theta)^2)
  sse_full <- sum((as.numeric(gam_mat %*% t(t(full_cvec))) - theta)^2)

  # Error Degrees of Freedom
  df_vec <- nrow(df) - 1 - c(9 - 1, 9 - 2, 9 - 4, 9 - 4, 9 - 8)
  names(df_vec) <- c("parac", "zeroc", "outc", "respc", "doubc")

  df_red <- df_vec[reduced_mod]
  df_full <- df_vec[full_mod]

  fstat <- ((sse_red - sse_full) / (df_red - df_full)) / (sse_full / df_full)

  pval <- 1 - pf(fstat, df_red - df_full, df_full)
  return(list(f = fstat, p = pval))

}

```

### 2024-01-22 Notes/semiconstraints.tex
Added Quarto document with semiparametric constraint simulation.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Understanding Semiparametric Constraints},
  pdfauthor={Caleb Leedy},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Understanding Semiparametric Constraints}
\author{Caleb Leedy}
\date{19 January 2024}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, frame hidden]}{\end{tcolorbox}}\fi

\hypertarget{sec-intro}{%
\section{Introduction}\label{sec-intro}}

The goal of this report is to understand how additional constraints
effect semiparametric models. Inspired by Dr.~Fuller's note, I set out
to see if we could assess the amount of efficiency loss due to using
semiparametric inference. First, let's define some notation.

\hypertarget{tbl-reftab}{}
\begin{longtable}[]{@{}llll@{}}
\caption{\label{tbl-reftab}This table shows the relationship between
different quantities of interest in each segment, their estimators, and
the corresponding coefficients of their estimators.}\tabularnewline
\toprule\noalign{}
Segment & Quantity of Interest & Estimator & Coefficient \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Segment & Quantity of Interest & Estimator & Coefficient \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(A_{00}\) & \(E[g \mid X]\) & \(\hat \gamma_{11}\) & \(c_{11}\) \\
\(A_{10}\) & \(E[g \mid X]\) & \(\hat \gamma_{21}\) & \(c_{21}\) \\
\(A_{10}\) & \(E[g \mid X, Y_1]\) & \(\hat \gamma_{22}\) & \(c_{22}\) \\
\(A_{01}\) & \(E[g \mid X]\) & \(\hat \gamma_{31}\) & \(c_{31}\) \\
\(A_{01}\) & \(E[g \mid X, Y_2]\) & \(\hat \gamma_{33}\) & \(c_{33}\) \\
\(A_{11}\) & \(E[g \mid X]\) & \(\hat \gamma_{41}\) & \(c_{41}\) \\
\(A_{11}\) & \(E[g \mid X, Y_1]\) & \(\hat \gamma_{42}\) & \(c_{42}\) \\
\(A_{11}\) & \(E[g \mid X, Y_2]\) & \(\hat \gamma_{43}\) & \(c_{43}\) \\
\(A_{11}\) & \(E[g \mid X, Y_1, Y_2]\) & \(\hat \gamma_{44}\) &
\(c_{44}\) \\
\end{longtable}

This is a similar framework to understanding the class of estimators
that we discussed in \verb|optest_update.pdf|, where we write the
estimator as a weighted average of different expectation with functional
coefficients.

\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
\mid X, Y_2].\]

However, now we write the estimator as

\[\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}\]

where
\(\hat \gamma_{kt} = \frac{\delta_{ij}}{\pi_{ij}}E[g \mid G_{ij}(X, Y_1, Y_2)]\)
and \(ij\) corresponds to the segment \(A_{ij}\) associated with
\(\hat \gamma_{kt}\) in Table~\ref{tbl-reftab}.

\hypertarget{sec-methods}{%
\section{Methods}\label{sec-methods}}

As noted by Dr.~Fuller, we can understand parametric estimation as
solving the following constrained optimization problems:

\begin{align*}
  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
  \sum_{k,t} c_{kt} = 1 \\
       &\text{or} \\
  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
  \sum_{(k,t): (k,t) \neq (4, 4)} c_{kt} = 0 \text{ and } c_{44} = 1. \\
\end{align*}

To be robust to the outcome model, we can construct a similar
optimization problem with different constraints:

\begin{align*}
  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
  c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0,\\
       &\text{and } c_{44} = 1.
\end{align*}

Likewise to be robust to the response model, we can have the problem:

\begin{align*}
  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
  c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \\
       &\text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}.
\end{align*}

To be double robust (robust to the outcome and response model) we can
combine the last two constraints. This is summarized in
Table~\ref{tbl-mods}.

\hypertarget{tbl-mods}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2941}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7059}}@{}}
\caption{\label{tbl-mods}This table identifies the different constraints
for each model type.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Constraints
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Constraints
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Parametric 1 & \(\sum_{k, t} c_{kt} = 1\) \\
Parametric 2 &
\(\sum_{k, t: (k, t) \neq (4, 4)} c_{kt} = 0, c_{44} = 1\) \\
Outcome Robust &
\(c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, \text{ and } c_{44} = 1.\) \\
Response Robust &
\(c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}\) \\
Double Robust &
\(c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, c_{44} = 1, c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}\) \\
\end{longtable}

\hypertarget{sec-simulations}{%
\section{Simulation Study}\label{sec-simulations}}

\hypertarget{simulation-1}{%
\subsection*{Simulation 1}\label{simulation-1}}
\addcontentsline{toc}{subsection}{Simulation 1}

We use the following simulation setup

\begin{align*}
  \begin{bmatrix} x \\ e_1 \\ e_2 \end{bmatrix} 
  &\stackrel{ind}{\sim} N\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}\right) \\
  y_1 &= x + e_1 \\
  y_2 &= \theta + x + e_2 \\
\end{align*}

Furthermore, \(\pi_{11} = 0.2\) and \(\pi_{00} = 0.3, \pi_{10} = 0.4\),
and \(\pi_{01} = 0.1\). The goal of this simulation study is to find
\(\theta = E[Y_2]\). In other words, \(g(Z) = Y_2\). There are several
algorithms for comparison which are defined as the following:

\begin{align*}
  Oracle &= n^{-1} \sum_{i = 1}^n g(Z_i)\\
  CC &= \frac{\sum_{i = 1}^n \delta_{11} g(Z_i)}{\sum_{i = 1}^n \delta_{11}} \\
  IPW &= \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g(Z_i)\\
\end{align*}

Furthermore, we include four existing estimators that we have already
proposed in the past: WLS, Prop, PropInd, and SemiDelta.

The estimator WLS is a a weight least square estimator derived in the
following manner. We now consider a normal model:

\[ 
  \begin{pmatrix}
    x_i \\ e_{1i} \\ e_{2i}
  \end{pmatrix} \stackrel{ind}{\sim}
  N \left(
  \begin{bmatrix}
    0 \\ 0 \\ 0
  \end{bmatrix},
  \begin{bmatrix}
    1 & 0 & 0 \\
    0 & \sigma_{11} & \sigma_{12} \\ 
    0 & \sigma_{12} & \sigma_{22}
  \end{bmatrix},
  \right)
\]

and define \(y_{1i} = \theta_1 + x_i + e_{1i}\) and
\(y_{2i} = \theta_2 + x_i + e_{2i}\). Then \(b_1 = b_2 = 1\). We define
\(\bar z_k^{(ij)}\) as the mean of \(y_k\) in segment \(A_{ij}\). This
means that we have means \(\bar z_1^{(11)}\), \(\bar z_2^{(11)}\),
\(\bar z_1^{(10)}\), and \(\bar z_2^{(01)}\). Let
\(W = [\bar z_1^{(11)}, \bar z_2^{(11)}, \bar z_1^{(10)}, \bar z_2^{(01)}]'\),
then for \(n_{ij} = |A_{ij}|\), we have

\[Z - M \mu \sim N(\vec 0, V)\]

where

\[M = 
  \begin{bmatrix}
    1 & 0 \\
    0 & 1 \\
    1 & 0 \\
    0 & 1 \\
  \end{bmatrix}
  \text{ and }
  V = 
  \begin{bmatrix}
    \frac{\sigma_{11}}{n_{11}} & \frac{\sigma_{12}}{n_{11}} & 0 & 0 \\
    \frac{\sigma_{12}}{n_{11}} & \frac{\sigma_{22}}{n_{11}} & 0 & 0 \\
    0 & 0 & \frac{\sigma_{11}}{n_{10}} & 0 \\
    0 & 0 & 0 & \frac{\sigma_{22}}{n_{01}} \\
  \end{bmatrix}.
\]

Thus, the BLUE for \(\mu = [\mu_1, \mu_2]'\) is

\[\hat \mu = (M' V^{-1} M)^{-1} M' V^{-1} W.\]

Hence, WLS is \(\mu_2\) as \(g(X, Y_1, Y_2) = Y_2\).

The remaining three estimators are drived from the following expression
where the values for \(\beta\) are provided in Table~\ref{tbl-beta}.
These are good estimators to compare to because Prop is the original
proposed estimator. PropInd has the same form as Prop except the values
for \(\beta\) is different. PropInd shares the same values of \(\beta\)
as all of the new models with constraints. SemiDelta is useful because
it is the best estimator in general so far.

\[
\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + 
\beta_0(\delta, c_0)E[g(Z) \mid X] + 
\beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + 
\beta_2(\delta, c_2) E[g(Z)
\mid X, Y_2].
\]

\hypertarget{tbl-beta}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1698}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4151}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4151}}@{}}
\caption{\label{tbl-beta}This table displays the values of \(\beta\) for
different estimator types.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Estimator
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(\beta_0(\delta, c_0)\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(\beta_1(\delta, c_1)\)
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Estimator
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(\beta_0(\delta, c_0)\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(\beta_1(\delta, c_1)\)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Prop &
\(\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)\)
&
\(\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)\) \\
PropInd &
\(\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)\)
&
\(\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)\) \\
SemiDelta &
\(c_0\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{00}}{\pi_{00}}\right)\)
&
\(c_1\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{10}}{\pi_{10}}\right)\) \\
\end{longtable}

In the simulation results in Table~\ref{tbl-sim1}, the new results have
the same label as the value from the \texttt{Type} column in
Table~\ref{tbl-mods}.

\hypertarget{tbl-sim1}{}
\begin{longtable}[]{@{}lrrrr@{}}
\caption{\label{tbl-sim1}Results from Simulation 1. True values:
\(\theta = 5, \rho = 0.5\). The test conducted for the T-statistic and
P-value is a two sample test to see if the estimator is unbiased. This
simulation uses \(Y_1 = x + \varepsilon_1\),
\(Y_2 = \theta + x + \varepsilon_2\) where \(X \sim N(0, 1)\) and
\((\varepsilon_1, \varepsilon_2)\) come from a mean zero bivariate
normal distribution with unit variance and covariance \(\rho\). The
segments are unbalanced with: \(\pi_{11} = 0.2\), \(\pi_{10} = 0.4\),
\(\pi_{01} = 0.1\), and \(\pi_{00} = 0.3\).}\tabularnewline
\toprule\noalign{}
Algorithm & bias & sd & tstat & pval \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Algorithm & bias & sd & tstat & pval \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Oracle & -0.002 & 0.045 & -1.669 & 0.048 \\
CC & 0.001 & 0.083 & 0.263 & 0.396 \\
IPW & 0.010 & 0.357 & 0.872 & 0.192 \\
WLS & 0.000 & 0.054 & -0.065 & 0.474 \\
\(\hat \theta_{prop}\) & -0.002 & 0.063 & -0.768 & 0.221 \\
\(\hat \theta_{\delta}\) & -0.001 & 0.064 & -0.342 & 0.366 \\
Parametric 1 & -0.001 & 0.061 & -0.509 & 0.305 \\
Parametric 2 & -0.001 & 0.061 & -0.509 & 0.305 \\
Outcome Robust & -0.001 & 0.061 & -0.543 & 0.294 \\
Response Robust & -0.001 & 0.061 & -0.509 & 0.305 \\
Double Robust & -0.001 & 0.061 & -0.543 & 0.294 \\
\end{longtable}

\hypertarget{simulation-2}{%
\subsection*{Simulation 2}\label{simulation-2}}
\addcontentsline{toc}{subsection}{Simulation 2}

Next, we used the same simulation setup except that we are interested in
estimating \(\theta = E[g] = E[Y_1^2 Y_2]\). We do not use the WLS
estimator because it does not work.

\hypertarget{tbl-sim2}{}
\begin{longtable}[]{@{}lrrrr@{}}
\caption{\label{tbl-sim2}Results from Simulation 2. True values:
\(\theta = 10, \rho = 0.5\). The test conducted for the T-statistic and
P-value is a two sample test to see if the estimator is unbiased. This
simulation uses the same setup as Simulation 1 with
\(Y_1 = x + \varepsilon_1\), \(Y_2 = \theta + x + \varepsilon_2\) where
\(X \sim N(0, 1)\) and \((\varepsilon_1, \varepsilon_2)\) come from a
mean zero bivariate normal distribution with unit variance and
covariance \(\rho\). The segments are unbalanced with:
\(\pi_{11} = 0.2\), \(\pi_{10} = 0.4\), \(\pi_{01} = 0.1\), and
\(\pi_{00} = 0.3\).}\tabularnewline
\toprule\noalign{}
Algorithm & bias & sd & tstat & pval \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Algorithm & bias & sd & tstat & pval \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Oracle & -0.037 & 0.528 & -2.214 & 0.014 \\
CC & 0.007 & 1.196 & 0.179 & 0.429 \\
IPW & 0.023 & 1.405 & 0.520 & 0.302 \\
\(\hat \theta_{prop}\) & -0.044 & 0.691 & -2.027 & 0.021 \\
\(\hat \theta_{\delta}\) & -0.032 & 0.674 & -1.521 & 0.064 \\
Parametric 1 & -0.004 & 0.301 & -0.461 & 0.322 \\
Parametric 2 & -0.004 & 0.344 & -0.344 & 0.366 \\
Outcome Robust & -0.040 & 0.654 & -1.950 & 0.026 \\
Response Robust & -0.004 & 0.301 & -0.461 & 0.322 \\
Double Robust & -0.040 & 0.654 & -1.950 & 0.026 \\
\end{longtable}

\hypertarget{discussion}{%
\subsection*{Discussion}\label{discussion}}
\addcontentsline{toc}{subsection}{Discussion}

There has been concern about why the constrained results are the same in
Table~\ref{tbl-sim1}. This section aims to address these concerns.

Interestingly enough, the optimization algorithm does not have a lot of
variability in the output. The variance for each coefficient is found in
Table~\ref{tbl-estvar}. Since these values are all basically zero, we
can see that for each algorithm we are getting basically the same
coefficients each iteration. The coefficients that we see are displayed
in Table~\ref{tbl-estc}.

\hypertarget{tbl-estvar}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1562}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}@{}}
\caption{\label{tbl-estvar}This table displays the variance of each
coefficient for each estimator. A number close to zero means that the
optimization algorithm chose very similar values for each Monte Carlo
simulation. A value of exactly zero indicated that the optimization
algorithm chose the exact same number.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Estimator
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_1\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_2\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_3\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_4\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_5\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_6\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_7\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_8\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_9\)
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Estimator
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_1\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_2\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_3\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_4\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_5\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_6\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_7\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_8\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_9\)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Para. 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
Para. 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
Outcome & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
Response & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
Double & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
\end{longtable}

\hypertarget{tbl-estc}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1449}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0870}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1014}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0870}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1014}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0870}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1014}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1014}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1014}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0870}}@{}}
\caption{\label{tbl-estc}This table displays the estimated coefficients
for each algorithm on iteration 1. Since the variance between iterations
is small, these are reasonable for subsequent
iterations.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Estimator
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_1\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_2\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_3\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_4\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_5\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_6\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_7\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_8\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_9\)
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Estimator
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_1\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_2\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_3\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_4\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_5\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_6\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_7\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_8\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(c_9\)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Para. 1 & 0.3 & 0.400 & 0.000 & 0.100 & 0.000 & 0.200 & 0.000 & 0.000 &
0 \\
Para. 2 & 0.3 & 0.400 & 0.000 & 0.100 & 0.000 & 0.200 & 0.000 & -1.000 &
1 \\
Outcome & 0.3 & -0.071 & 0.471 & -0.194 & 0.294 & -0.035 & -0.471 &
-0.294 & 1 \\
Response & 0.3 & 0.400 & 0.000 & 0.100 & 0.000 & 0.200 & 0.000 & 0.000 &
0 \\
Double & 0.3 & -0.071 & 0.471 & -0.194 & 0.294 & -0.035 & -0.471 &
-0.294 & 1 \\
\end{longtable}

\hypertarget{tbl-fstat}{}
\begin{longtable}[]{@{}llrr@{}}
\caption{\label{tbl-fstat}This table computes the F-statistic and
P-value for an F-test comparing the \texttt{Full\ Model} with a nested
model in \texttt{Reduced\ Model}. The computation of the F-statistic is
found in the \texttt{get\_f\_statistic} function.}\tabularnewline
\toprule\noalign{}
Full Model & Reduced Model & Fstat & P-Value \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Full Model & Reduced Model & Fstat & P-Value \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Parametric 1 & Parametric 2 & 0.00 & 1 \\
Parametric 1 & Outcome Robust & 1190.30 & 0 \\
Parametric 1 & Response Robust & 0.00 & 1 \\
Parametric 1 & Double Robust & 510.13 & 0 \\
Parametric 2 & Outcome Robust & 1787.26 & 0 \\
Parametric 2 & Response Robust & 0.00 & 1 \\
Parametric 2 & Double Robust & 595.75 & 0 \\
Outcome Robust & Double Robust & 0.00 & 1 \\
Response Robust & Double Robust & 895.43 & 0 \\
\end{longtable}

From the analysis in Table~\ref{tbl-fstat}, we can see that there are
significant differences between several of the models. Why is this
different from the previous analysis in Table~\ref{tbl-sim1}? The answer
comes from how we compute the variance versus the sum of squares of
error. The variance is computed from each estimator created in a Monte
Carlo trial; whereas the sum of squares of error is summed from each
individuals estimate for the mean of \(Y_2\). For more details about how
the F-test is computed see Appendix A and Appendix B.

\newpage{}

\hypertarget{appendix-a-computing-the-f-test}{%
\section{Appendix A: Computing the
F-Test}\label{appendix-a-computing-the-f-test}}

To compute an F test we can use Equation~\ref{eq-fstat}.

\begin{equation}\protect\hypertarget{eq-fstat}{}{ F = \frac{(SSE_{Reduced} - SSE_{Full}) / (DFE_{Reduced} - DFE_{Full})}{
SSE_{Full} / DFE_{Full}}. }\label{eq-fstat}\end{equation}

Let's assess what this would mean to compare models
\texttt{Parametric\ 1} and \texttt{Outcome\ Robust} from
Table~\ref{tbl-mods}.

\hypertarget{tbl-mods}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2941}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7059}}@{}}
\caption{\label{tbl-mods}This table identifies the different constraints
for each model type.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Constraints
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Constraints
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Parametric 1 & \(\sum_{k, t} c_{kt} = 1\) \\
Parametric 2 &
\(\sum_{k, t: (k, t) \neq (4, 4)} c_{kt} = 0, c_{44} = 1\) \\
Outcome Robust &
\(c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, \text{ and } c_{44} = 1.\) \\
Response Robust &
\(c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}\) \\
Double Robust &
\(c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, c_{44} = 1, c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}\) \\
\end{longtable}

Consider the case where we run estimate each model on a data set with
\(n = 1000\) observations, and define the following notation: let
\(\hat \theta^{(P)}\) and \(\hat \theta^{(OR)}\) be the estimated values
of \(\theta = E[Y_2]\) for the \texttt{Parametric\ 1} model and the
\texttt{Outcome\ Robust} model respectively. Define the estimated
coefficients to be \(\hat c_j^{(P)}\) and \(\hat c_j^{(OR)}\) where
\(j = 1, \dots 9\) for the \texttt{Parametric\ 1} and
\texttt{Outcome\ Robust} models respectively. The one can compute the
SSE with the following:

\[ SSE = n^{-1} \sum_{i = 1}^n (\hat y_{2i} - \theta)^2 \text{ where }
\hat y_{2i} = \sum_{j = 1}^9 \hat c_j \hat \gamma_j \]

and
\(\hat \gamma_0 := \hat \gamma_{00} = \frac{\delta_{00i}}{\pi_{00}} E[Y_2 \mid x_i]\),
\(\hat \gamma_1 := \hat \gamma_{11} = \frac{\delta_{10i}}{\pi_{10}} E[Y_2 \mid x_i, y_{1i}]\),
etc. Note that this is different from the previous estimate of the
variance which used the Monte Carlo variance (standard deviation)
defined by
\[ \frac{1}{n - 1} \sum_{b = 1}^B (\hat \theta_b - \bar \theta_B)^2 \]
where \(B\) is the number of Monte Carlo estimates and
\(\bar \theta_B = \frac{1}{B} \sum_{b = 1}^B \hat \theta_b\).

Likewise, one can compute the degrees of freedom by noticing that each
model has a degrees of freedom equal to nine minus the number of
constraints. This means that we can compute the model degrees of freedom
and error degrees of freedom for each model type, which we do in
Table~\ref{tbl-df}

\hypertarget{tbl-df}{}
\begin{longtable}[]{@{}lll@{}}
\caption{\label{tbl-df}This table displays the degrees of freedom for
each model}\tabularnewline
\toprule\noalign{}
Model & Model Degrees of Freedom & Error Degrees of Freedom \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Model & Model Degrees of Freedom & Error Degrees of Freedom \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Parametric 1 & \(9 - 1 = 8\) & \(n - 1 - 8 = n - 9\) \\
Parametric 2 & \(9 - 1 = 8\) & \(n - 1 - 8 = n - 9\) \\
Outcome Robust & \(9 - 4 = 5\) & \(n - 1 - 5 = n - 6\) \\
Response Robust & \(9 - 4 = 5\) & \(n - 1 - 5 = n - 6\) \\
Double Robust & \(9 - 8 = 1\) & \(n - 1 - 1 = n - 2\) \\
\end{longtable}

So continuing our first example\footnote{SSE values might differ in
  Table~\ref{tbl-fstat} slightly depending on the actual data values.},
if \(SSE^{(P)} = 993\) and \(SSE^{(OR)} = 3428\) with \(n = 1000\) then
the F statistic is

\[F = \frac{(3428 - 993) / (3)}{993 / (1000 - 9)} = 810.\]

The critical value we want to compare this with is the \(0.95\) quantile
of \(F_{3, 993}\) which is \(2.61\). So there \emph{is} a significant
difference between the fit of these two model. (The p-value is basically
zero.) The code to do this automatically is in Appendix B.

\newpage{}

\hypertarget{appendix-b-r-functions}{%
\section{Appendix B: R Functions}\label{appendix-b-r-functions}}

\hypertarget{lst-getfstat}{%
\label{lst-getfstat}}%
\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\textquotesingle{} get\_f\_statistic is a function to compute the F{-}statistic between two of }
\CommentTok{\#\textquotesingle{} the constrained models.}
\CommentTok{\#\textquotesingle{}}
\CommentTok{\#\textquotesingle{} @param c\_tab {-} A table with values for the estimated c coefficients}
\CommentTok{\#\textquotesingle{} @param reduced\_mod {-} A string indicating which model is the reduced model}
\CommentTok{\#\textquotesingle{} @param full\_mod {-} A string indicating which model is the full model}
\CommentTok{\#\textquotesingle{} @param df {-} A data set simulated from the same data generating process that }
\CommentTok{\#\textquotesingle{}    the c\_tab coefficients were estimated from.}
\CommentTok{\#\textquotesingle{}}
\CommentTok{\#\textquotesingle{} @note We have previously shown that the variance of the estimated c\_tab }
\CommentTok{\#\textquotesingle{} variables is close to or exactly zero. Hence, conducting a new simulation }
\CommentTok{\#\textquotesingle{} should not effect the results of the model much.}
\NormalTok{get\_f\_statistic }\OtherTok{\textless{}{-}} 
  \ControlFlowTok{function}\NormalTok{(c\_tab, reduced\_mod, full\_mod, df, }\AttributeTok{theta =}\NormalTok{ true\_theta, cov\_e1e2) \{}

\NormalTok{  vars\_lst }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{(reduced\_mod, full\_mod)}

  \CommentTok{\# SSE reduced }
\NormalTok{  red\_cvec }\OtherTok{\textless{}{-}}
\NormalTok{    c\_tab }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{filter}\NormalTok{(iter }\SpecialCharTok{==} \DecValTok{1}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{filter}\NormalTok{(.data[[}\StringTok{"estimator"}\NormalTok{]] }\SpecialCharTok{==}\NormalTok{ vars\_lst[[}\DecValTok{1}\NormalTok{]]) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{select}\NormalTok{(}\FunctionTok{starts\_with}\NormalTok{(}\StringTok{"c\_"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{as.numeric}\NormalTok{()}

  \CommentTok{\# SSE full}
\NormalTok{  full\_cvec }\OtherTok{\textless{}{-}}
\NormalTok{    c\_tab }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{filter}\NormalTok{(iter }\SpecialCharTok{==} \DecValTok{1}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{filter}\NormalTok{(.data[[}\StringTok{"estimator"}\NormalTok{]] }\SpecialCharTok{==}\NormalTok{ vars\_lst[[}\DecValTok{2}\NormalTok{]]) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{select}\NormalTok{(}\FunctionTok{starts\_with}\NormalTok{(}\StringTok{"c\_"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{as.numeric}\NormalTok{()}

  \CommentTok{\# Constructing gam\_vec}
\NormalTok{  Eg }\OtherTok{\textless{}{-}}\NormalTok{ theta}
\NormalTok{  Egx }\OtherTok{\textless{}{-}}\NormalTok{ theta }\SpecialCharTok{+}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{X }
\NormalTok{  Egxy1 }\OtherTok{\textless{}{-}}\NormalTok{ theta }\SpecialCharTok{+}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{X }\SpecialCharTok{+}\NormalTok{ cov\_e1e2 }\SpecialCharTok{*}\NormalTok{ (df}\SpecialCharTok{$}\NormalTok{Y1 }\SpecialCharTok{{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{X)}
\NormalTok{  Egxy2 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{Y2}
\NormalTok{  EEg2 }\OtherTok{\textless{}{-}}\NormalTok{ theta}\SpecialCharTok{\^{}}\DecValTok{2} \SpecialCharTok{+} \DecValTok{2}
\NormalTok{  EEgx2 }\OtherTok{\textless{}{-}}\NormalTok{ theta}\SpecialCharTok{\^{}}\DecValTok{2} \SpecialCharTok{+} \DecValTok{1}
\NormalTok{  EEgxy12 }\OtherTok{\textless{}{-}}\NormalTok{ theta}\SpecialCharTok{\^{}}\DecValTok{2} \SpecialCharTok{+} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ cov\_e1e2}\SpecialCharTok{\^{}}\DecValTok{2}
\NormalTok{  EEgxy22 }\OtherTok{\textless{}{-}}\NormalTok{ theta}\SpecialCharTok{\^{}}\DecValTok{2} \SpecialCharTok{+} \DecValTok{2}
\NormalTok{  EEgxy1Egxy2 }\OtherTok{\textless{}{-}}\NormalTok{ theta}\SpecialCharTok{\^{}}\DecValTok{2} \SpecialCharTok{+} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ cov\_e1e2}\SpecialCharTok{\^{}}\DecValTok{2}

\NormalTok{  g\_11 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_00 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_00 }\SpecialCharTok{*}\NormalTok{ (Egx)}
\NormalTok{  g\_21 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_10 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_10 }\SpecialCharTok{*}\NormalTok{ (Egx)}
\NormalTok{  g\_22 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_10 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_10 }\SpecialCharTok{*}\NormalTok{ (Egxy1)}
\NormalTok{  g\_31 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_01 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_01 }\SpecialCharTok{*}\NormalTok{ (Egx)}
\NormalTok{  g\_33 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_01 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_01 }\SpecialCharTok{*}\NormalTok{ (Egxy2)}
\NormalTok{  g\_41 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_11 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_11 }\SpecialCharTok{*}\NormalTok{ (Egx)}
\NormalTok{  g\_42 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_11 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_11 }\SpecialCharTok{*}\NormalTok{ (Egxy1)}
\NormalTok{  g\_43 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_11 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_11 }\SpecialCharTok{*}\NormalTok{ (Egxy2)}
\NormalTok{  g\_44 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_11 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_11 }\SpecialCharTok{*}\NormalTok{ (df}\SpecialCharTok{$}\NormalTok{Y2)}

\NormalTok{  gam\_mat }\OtherTok{\textless{}{-}}\NormalTok{ base}\SpecialCharTok{::}\FunctionTok{cbind}\NormalTok{(g\_11, g\_21, g\_22, g\_31, g\_33, g\_41, g\_42, g\_43, g\_44)}

  \CommentTok{\# SSE}
\NormalTok{  sse\_red }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{((}\FunctionTok{as.numeric}\NormalTok{(gam\_mat }\SpecialCharTok{\%*\%} \FunctionTok{t}\NormalTok{(}\FunctionTok{t}\NormalTok{(red\_cvec))) }\SpecialCharTok{{-}}\NormalTok{ theta)}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}
\NormalTok{  sse\_full }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{((}\FunctionTok{as.numeric}\NormalTok{(gam\_mat }\SpecialCharTok{\%*\%} \FunctionTok{t}\NormalTok{(}\FunctionTok{t}\NormalTok{(full\_cvec))) }\SpecialCharTok{{-}}\NormalTok{ theta)}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}

  \CommentTok{\# Error Degrees of Freedom}
\NormalTok{  df\_vec }\OtherTok{\textless{}{-}} \FunctionTok{nrow}\NormalTok{(df) }\SpecialCharTok{{-}} \DecValTok{1} \SpecialCharTok{{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{9} \SpecialCharTok{{-}} \DecValTok{1}\NormalTok{, }\DecValTok{9} \SpecialCharTok{{-}} \DecValTok{2}\NormalTok{, }\DecValTok{9} \SpecialCharTok{{-}} \DecValTok{4}\NormalTok{, }\DecValTok{9} \SpecialCharTok{{-}} \DecValTok{4}\NormalTok{, }\DecValTok{9} \SpecialCharTok{{-}} \DecValTok{8}\NormalTok{)}
  \FunctionTok{names}\NormalTok{(df\_vec) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"parac"}\NormalTok{, }\StringTok{"zeroc"}\NormalTok{, }\StringTok{"outc"}\NormalTok{, }\StringTok{"respc"}\NormalTok{, }\StringTok{"doubc"}\NormalTok{)}

\NormalTok{  df\_red }\OtherTok{\textless{}{-}}\NormalTok{ df\_vec[reduced\_mod]}
\NormalTok{  df\_full }\OtherTok{\textless{}{-}}\NormalTok{ df\_vec[full\_mod]}

\NormalTok{  fstat }\OtherTok{\textless{}{-}}\NormalTok{ ((sse\_red }\SpecialCharTok{{-}}\NormalTok{ sse\_full) }\SpecialCharTok{/}\NormalTok{ (df\_red }\SpecialCharTok{{-}}\NormalTok{ df\_full)) }\SpecialCharTok{/}\NormalTok{ (sse\_full }\SpecialCharTok{/}\NormalTok{ df\_full)}

\NormalTok{  pval }\OtherTok{\textless{}{-}} \DecValTok{1} \SpecialCharTok{{-}} \FunctionTok{pf}\NormalTok{(fstat, df\_red }\SpecialCharTok{{-}}\NormalTok{ df\_full, df\_full)}
  \FunctionTok{return}\NormalTok{(}\FunctionTok{list}\NormalTok{(}\AttributeTok{f =}\NormalTok{ fstat, }\AttributeTok{p =}\NormalTok{ pval))}

\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\end{document}

### 2024-01-22 Notes/template.qmd
Added Quarto document with semiparametric constraint simulation.

---
title: "Title"
author:
  # Add additional authors here if needed, by replicating the above entries.
  - name: Caleb Leedy
    affiliation: Iowa State University
    email: cleedy@iastate.edu
# The short* versions of title and author are used for constructing page headers
# and are optional
shorttitle: "A Short Title"
shortauthor: "C. Leedy"
date: today
date-format: "D MMMM YYYY"
abstract: |
  Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Etiam lobortis 
  facilisis sem. Nullam nec mi et neque pharetra sollicitudin. Praesent 
  imperdiet mi nec ante. Donec ullamcorper, felis non sodales com- modo, 
  lectus velit ultrices augue, a dignissim nibh lectus placerat pede. 
  Vivamus nunc nunc, molestie ut, ultricies vel, semper in, velit. Ut 
  porttitor. Praesent in sapien. Lorem ipsum dolor sit amet, consectetuer 
  adipiscing elit. This is the abstract.
format:
  simple-article-pdf:
    keep-tex: true
    indent: false
    classoption: 
     - 9pt
     - twoside
reference-section-title: References
bibliography: ../references.bib
---

# Introduction {#sec-intro}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. Etiam lobortis facilisis sem. Nullam nec mi
et neque pharetra sollicitudin. Praesent imperdiet mi nec ante. Donec
ullamcorper, felis non sodales commodo, lectus velit ultrices augue, a
dignissim nibh lectus placerat pede. Vivamus nunc nunc, molestie ut, ultricies
vel, semper in, velit. Ut porttitor. Praesent in sapien. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. Duis fringilla tristique neque. Sed
interdum libero ut metus. Pellentesque placerat. Nam rutrum augue a leo. Morbi
sed elit sit amet ante lobortis sollicitudin. Praesent blandit blandit mauris.
Praesent lectus tellus, aliquet aliquam, luctus a, egestas a, turpis. Mauris
lacinia lorem sit amet ipsum. Nunc quis urna dictum turpis accumsan semper.

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Etiam lobortis
facilisis sem. Nullam nec mi et neque pharetra sollicitudin. Praesent imperdiet
mi nec ante. Donec ullamcorper, felis non sodales commodo, lectus velit
ultrices augue, a dignissim nibh lectus placerat pede. Vivamus nunc nunc, mo-
lestie ut, ultricies vel, semper in, velit. Ut porttitor. Praesent in sapien.
Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Duis fringilla
tristique neque. Sed interdum libero ut metus. Pellentesque placerat. Nam
rutrum augue a leo. Morbi sed elit sit amet ante lobortis sollicitudin.
Praesent blandit blandit mauris. Praesent lectus tellus, aliquet aliquam,
luctus a, egestas a, turpis. Mauris lacinia lorem sit amet ipsum. Nunc quis
urna dictum turpis accumsan semper.

# Methods {#sec-methods}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Etiam lobortis
facilisis sem. Nullam nec mi et neque pharetra sollicitudin. Praesent imperdiet
mi nec ante. Donec ullamcorper, felis non sodales commodo, lectus velit
ultrices augue, a dignissim nibh lectus placerat pede. Vivamus nunc nunc, mo-
lestie ut, ultricies vel, semper in, velit. Ut porttitor. Praesent in sapien.
Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Duis fringilla
tristique neque. Sed interdum libero ut metus. Pellentesque placerat. Nam
rutrum augue a leo. Morbi sed elit sit amet ante lobortis sollicitudin.
Praesent blandit blandit mauris. Praesent lectus tellus, aliquet aliquam,
luctus a, egestas a, turpis. Mauris lacinia lorem sit amet ipsum. Nunc quis
urna dictum turpis accumsan semper.

Maecenas sed ultricies felis. Sed imperdiet dictum arcu a egestas.

-   Donec dolor arcu, rutrum id molestie in, viverra sed diam
-   Curabitur feugiat
-   Turpis sed auctor facilisis
-   Arcu eros accumsan lorem, at posuere mi diam sit amet tortor
-   Fusce fermentum, mi sit amet euismod rutrum
-   Sem lorem molestie diam, iaculis aliquet sapien tortor non nisi
-   Pellentesque bibendum pretium aliquet

```{r}
kableExtra::kbl(head(mtcars), caption = "Cars", label = "tbl-cars", booktabs = TRUE, table.env = 'table*')
```

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Etiam lobortis
facilisis sem. Nullam nec mi et neque pharetra sollicitudin. Praesent imperdiet
mi nec ante. Donec ullamcorper, felis non sodales com- modo, lectus velit
ultrices augue, a dignissim nibh lectus placerat pede. Vivamus nunc nunc,
molestie ut, ultricies vel, semper in, velit. Ut porttitor. Praesent in sapien.
Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Duis fringilla
tristique neque. Sed interdum libero ut metus. Pellentesque placerat. Nam
rutrum augue a leo. Morbi sed elit sit amet ante lobortis sollicitudin.
Praesent blandit blandit mauris. Praesent lectus tellus, aliquet aliquam,
luctus a, egestas a, turpis. Mauris lacinia lorem sit amet ipsum. Nunc quis
urna dictum turpis accumsan semper.

# Results {#sec-results}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Etiam lobortis
facilisis sem. Nullam nec mi et neque pharetra sollicitudin. Praesent imperdiet
mi nec ante. Donec ullamcorper, felis non sodales com- modo, lectus velit
ultrices augue, a dignissim nibh lectus placerat pede. Vivamus nunc nunc,
molestie ut, ultricies vel, semper in, velit. Ut porttitor. Praesent in sapien.
Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Duis fringilla
tristique neque. Sed interdum libero ut metus. Pellentesque placerat. Nam
rutrum augue a leo. Morbi sed elit sit amet ante lobortis sollicitudin.
Praesent blandit blandit mauris. Praesent lectus tellus, aliquet aliquam,
luctus a, egestas a, turpis. Mauris lacinia lorem sit amet ipsum. Nunc quis
urna dictum turpis accumsan semper.

$$E = m c^2$$ {#eq-einstein}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Etiam lobortis
facilisis sem. Nullam nec mi et neque pharetra sollicitudin. Praesent imperdiet
mi nec ante. Donec ullamcorper, felis non sodales commodo, lectus velit
ultrices augue, a dignissim nibh lectus placerat pede. Vivamus nunc nunc, mo-
lestie ut, ultricies vel, semper in, velit. Ut porttitor. Praesent in sapien.
Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Duis fringilla
tristique neque. Sed interdum libero ut metus. Pellentesque placerat. Nam
rutrum augue a leo. Morbi sed elit sit amet ante lobortis sollicitudin.
Praesent blandit blandit mauris. Praesent lectus tellus, aliquet aliquam,
luctus a, egestas a, turpis. Mauris lacinia lorem sit amet ipsum. Nunc quis
urna dictum turpis accumsan semper.

# Discussion {#sec-discussion}

## First Subsection

Lorem ipsum dolor sit amet [@kim2021combining], consectetuer adipiscing elit.
Etiam lobortis facilisis sem. Nullam nec mi et neque pharetra sollicitudin.
Praesent imperdiet mi nec ante. Donec ullamcorper, felis non sodales commodo,
lectus velit ultrices augue, a dignissim nibh lectus placerat pede. Vivamus
nunc nunc, molestie ut, ultricies vel, semper in, velit. Ut porttitor. Praesent
in sapien. Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Duis
fringilla tristique neque. Sed interdum libero ut metus. Pellentesque placerat.
Nam rutrum augue a leo. Morbi sed elit sit amet ante lobortis sollicitudin.
Praesent blandit blandit mauris. Praesent lectus tellus, aliquet aliquam,
luctus a, egestas a, turpis. Mauris lacinia lorem sit amet ipsum. Nunc quis
urna dictum turpis accumsan semper.

## Second Subsection

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Etiam lobortis
facilisis sem. Nullam nec mi et neque pharetra sollicitudin. Praesent imperdiet
mi nec ante. Donec ullamcorper, felis non sodales com- modo, lectus velit
ultrices augue, a dignissim nibh lectus placerat pede. Vivamus nunc nunc,
molestie ut, ultricies vel, semper in, velit. Ut porttitor. Praesent in sapien.
Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Duis fringilla
tristique neque. Sed interdum libero ut metus. Pellentesque placerat. Nam
rutrum augue a leo. Morbi sed elit sit amet ante lobortis sollicitudin.
Praesent blandit blandit mauris. Praesent lectus tellus, aliquet aliquam,
luctus a, egestas a, turpis. Mauris lacinia lorem sit amet ipsum. Nunc quis
urna dictum turpis accumsan semper.

### 2024-01-22 Simulations/T3Assess_Semiparametric_Cost/analyze_semiconstraints.R

@@ -1,0 +1,97 @@
+# Title: analyze_semiconstraints.R
+# Created by: Caleb Leedy
+# Created on: January 17, 2024
+# Purpose: This file aims to explore the results of `semi_cost.R` in more
+# detail. I do not understand why adding a constraint does not change the 
+# estimation in the first simulation.
+
+# *************
+# * Libraries *
+# *************
+
+library(dplyr)
+library(tidyr)
+library(stringr)
+library(CVXR)
+library(ggplot2)
+
+library(doParallel)
+library(doRNG)
+library(parallelly)
+
+source("R/opt_est.R")
+source("Simulations/T3Assess_Semiparametric_Cost/compare_algs.R")
+
+# ***************************
+# * Monte Carlo Simulations *
+# ***************************
+
+# ************************
+# * theta = E[g] = E[Y2] *
+# ************************
+
+clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
+registerDoParallel(clust)
+
+B <- 100
+n_obs <- 1000
+true_theta <- 5 
+cov_e1e2 <- 0.5
+
+mc_theta <-
+  foreach(iter = 1:B,
+          .options.RNG = 1,
+          .packages = c("dplyr", "stringr", "CVXR", "tidyr")) %dorng% {
+
+    # Generate Data
+    # df <- gen_optsim_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)
+    df <- gen_unbal_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)
+
+    # Get Estimates
+    conalg_para <- min_var_alg(df, mod = "para", cov_e1e2 = cov_e1e2)
+    conalg_sumzero <- min_var_alg(df, mod = "sumzero", cov_e1e2 = cov_e1e2)
+    conalg_outrob <- min_var_alg(df, mod = "out_rob", cov_e1e2 = cov_e1e2)
+    conalg_resprob <- min_var_alg(df, mod = "resp_rob", cov_e1e2 = cov_e1e2)
+    conalg_doubrob <- min_var_alg(df, mod = "double_rob", cov_e1e2 = cov_e1e2)
+
+    return(
+      tibble(conpara = conalg_para$theta_est,
+             conzero = conalg_sumzero$theta_est,
+             conout = conalg_outrob$theta_est,
+             conresp = conalg_resprob$theta_est,
+             condoub = conalg_doubrob$theta_est,
[…]

### 2024-01-22 Simulations/T3Assess_Semiparametric_Cost/compare_algs.R

@@ -26,12 +26,7 @@ source("R/opt_est.R")
 #' "parametric", "sumzero", "outcome_robust", "response_robust", or
 #' "double_robust".
 min_var_alg <- 
-  function(df,
-           gfun = "Y2",
-           mod = "para",
-           theta = NA,
-           mean_x = 0,
-           cov_e1e2 = 0) {
+  function(df, gfun = "Y2", mod = "para", theta = NA, mean_x = 0, cov_e1e2 = 0) {
 
   df <- mutate(df, g_i = eval(rlang::parse_expr(gfun)))
   df_11 <- filter(df, delta_1 == 1, delta_2 == 1)
@@ -125,58 +120,35 @@ min_var_alg <-
   # CVXR
   c_hat <- Variable(length(gam_vec))
 
-  # TODO: We can change the inequality constraints to equality constraints.
   if (mod == "para") {
-    con1 <- matrix(rep(1, length(gam_vec)), nrow = 1) %*% c_hat <= 1
-    con2 <- matrix(rep(1, length(gam_vec)), nrow = 1) %*% c_hat >= 1
-    con_list <- list(con1, con2)
+    con1 <- matrix(rep(1, length(gam_vec)), nrow = 1) %*% c_hat == 1
+    con_list <- list(con1)
   } else if (mod == "sumzero") {
-    con1 <- matrix(c(rep(0, length(gam_vec) - 1), 1), nrow = 1) %*% c_hat <= 1
-    con2 <- matrix(c(rep(0, length(gam_vec) - 1), 1), nrow = 1) %*% c_hat >= 1
-    con3 <- matrix(c(rep(1, length(gam_vec) - 1), 0), nrow = 1) %*% c_hat <= 0
-    con4 <- matrix(c(rep(1, length(gam_vec) - 1), 0), nrow = 1) %*% c_hat >= 0
-    con_list <- list(con1, con2, con3, con4)
+    con1 <- matrix(c(rep(0, length(gam_vec) - 1), 1), nrow = 1) %*% c_hat == 1
[…]

### 2024-01-23 Notes/semiconstraints.qmd
Automatically updated on Tue Jan 23 10:28:25 AM CST 2024

@@
 that we discussed in \verb|optest_update.pdf|, where we write the estimator 
-as a weighted average of different expectation with functional coefficients.
+as a weighted average of different expectations with functional coefficients.
 
-$$\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
-\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
+$$\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta)E[g(Z)
+\mid X] + \beta_1(\delta)E[g(Z) \mid X, Y_1] + \beta_2(\delta) E[g(Z)
 \mid X, Y_2].$$
+
+In this case, we assume that $E[\beta_j(\delta)] = 0$ for $j = 0, 1, 2$.
 
@@
 
-$$\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}$$
+$$\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}$$ {#eq-linest}
 
@@
 Y_2)]$ and $ij$ corresponds to the segment $A_{ij}$ associated with $\hat
-\gamma_{kt}$ in @tbl-reftab.
+\gamma_{kt}$ in @tbl-reftab. In general the estimator has the following form,
+
+\begin{align*}
+\hat \theta = n^{-1} \sum_{i = 1}^n 
+  &\left\{c_{11}\frac{\delta_{00}}{\pi_{00}} E[g \mid X]+ \right.\\
+  &c_{21}\frac{\delta_{10}}{\pi_{10}} E[g \mid X]+ 
+  c_{22}\frac{\delta_{10}}{\pi_{10}} E[g \mid X, Y_1]+ \\
+  &c_{31}\frac{\delta_{01}}{\pi_{01}} E[g \mid X]+ 
+  c_{33}\frac{\delta_{01}}{\pi_{01}} E[g \mid X, Y_2]+ \\
+  &c_{41}\frac{\delta_{11}}{\pi_{11}} E[g \mid X]+ 
+  c_{42}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_1]+
+  c_{43}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_2]+
+  \left.c_{44}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_1, Y_2]\right\}
+\end{align*}
+
+which we have condensed to  @eq-linest.
 
@@
 : This table identifies the different constraints for each model type. {#tbl-mods}
+
+We can justify the constraints for the Outcome Robust model by rewriting the 
+estimator into the following form:
+
+\begin{align*}
+E_\delta[\hat \theta] &= n^{-1} \sum_{i = 1}^n E_\delta\left[
+E[g \mid X] \left(\frac{\delta_{00}}{\pi_{00}} c_{11} +
+\frac{\delta_{10}}{\pi_{10}}c_{21} \frac{\delta_{01}}{\pi_{01}}c_{31} +
+\frac{\delta_{11}}{\pi_{11}}c_{41}\right)\right. + \\ 
+&\qquad\qquad\qquad\qquad E[g \mid X, Y_1] \left(\frac{\delta_{10}}{\pi_{10}}c_{22} + 
+\frac{\delta_{11}}{\pi_{11}}c_{42}\right) +\\ 
+&\qquad\qquad\qquad\qquad E[g \mid X, Y_2] \left(\frac{\delta_{01}}{\pi_{01}}c_{33} + 
+\frac{\delta_{11}}{\pi_{11}}c_{43}\right) +\\ 
+&\qquad\qquad\qquad\qquad \left.g \frac{\delta_{11}}{\pi_{11}}c_{44}\right] \\
+&= n^{-1} \sum_{i = 1}^n \left\{E[g \mid X](c_{11} + c_{21} + c_{31} + c_{41}) +
+E[g \mid X, Y_1] (c_{22} + c_{42}) +
+E[g \mid X, Y_2] (c_{33} + c_{43}) + g c_{44}\right\}
+\end{align*}
+
+Thus, if the Outcome Robust constraints are satisfied, 
+$E_\delta[\hat \theta] = g$ 
+when we take the expectation with respect to the response variable $\delta$.
+These constraints make sense because if they are satisfied $E[\hat \theta]$ 
+does not depend on the correct specification of the models
+for $E[g \mid G_r(X, Y_1, Y_2)]$.
+
+Likewise, we can justify the constraints for the Response Robust model 
+by rewriting @eq-linest into the form
+
+\begin{align*}
+E_Z [\hat \theta] &= n^{-1} \sum_{i = 1}^n E_Z\left[
+\frac{\delta_{00}}{\pi_{00}}c_{11} E[g \mid X]\right. + \\
+&\qquad\qquad\qquad\qquad\frac{\delta_{10}}{\pi_{10}}\left(c_{21}E[g \mid X] + 
+  c_{22}E[g \mid X, Y_1]\right) +\\
+&\qquad\qquad\qquad\qquad\frac{\delta_{01}}{\pi_{01}}\left(c_{31}E[g \mid X] + 
+  c_{33}E[g \mid X, Y_2]\right) +\\
+&\qquad\qquad\qquad\qquad\left.\frac{\delta_{11}}{\pi_{11}}\left(c_{41}E[g \mid X] + 
+  c_{42}E[g \mid X, Y_1] + c_{43}E[g \mid X, Y_2] + c_{44}g\right)\right] \\
+& n^{-1} \sum_{i = 1}^n \left(\frac{\delta_{00}}{\pi_{00}}c_{11} + 
+\frac{\delta_{10}}{\pi_{10}}(c_{21} + c_{22}) + 
+\frac{\delta_{01}}{\pi_{01}}(c_{31} + c_{33}) + 
+\frac{\delta_{11}}{\pi_{11}}(c_{41} + c_{42} + c_{43} + c_{44})\right)g
+\end{align*}
+
+Thus, if the Response Robust constraints are satisfied, $E_X[\hat \theta] = g$
+when we take the expectation with respect to the outcome variable 
+$Z = (X, Y_1, Y_2)'$. These constraints make sense because when they are 
+satisfied $E[\hat \theta]$ does not depend on the correct response 
+probabilities $\pi = E[\delta]$.
 

### 2024-01-23 Notes/semiconstraints.tex
Automatically updated on Tue Jan 23 10:28:25 AM CST 2024

@@
 \author{Caleb Leedy}
-\date{19 January 2024}
+\date{22 January 2024}
 
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 that we discussed in \verb|optest_update.pdf|, where we write the
-estimator as a weighted average of different expectation with functional
-coefficients.
+estimator as a weighted average of different expectations with
+functional coefficients.
 
-\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
-\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
+\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta)E[g(Z)
+\mid X] + \beta_1(\delta)E[g(Z) \mid X, Y_1] + \beta_2(\delta) E[g(Z)
 \mid X, Y_2].\]
+
+In this case, we assume that \(E[\beta_j(\delta)] = 0\) for
+\(j = 0, 1, 2\).
 
@@
 
-\[\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}\]
+\begin{equation}\protect\hypertarget{eq-linest}{}{\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}}\label{eq-linest}\end{equation}
 
@@
 and \(ij\) corresponds to the segment \(A_{ij}\) associated with
-\(\hat \gamma_{kt}\) in Table~\ref{tbl-reftab}.
+\(\hat \gamma_{kt}\) in Table~\ref{tbl-reftab}. In general the estimator
+has the following form,
+
+\begin{align*}
+\hat \theta = n^{-1} \sum_{i = 1}^n 
+  &\left\{c_{11}\frac{\delta_{00}}{\pi_{00}} E[g \mid X]+ \right.\\
+  &c_{21}\frac{\delta_{10}}{\pi_{10}} E[g \mid X]+ 
+  c_{22}\frac{\delta_{10}}{\pi_{10}} E[g \mid X, Y_1]+ \\
+  &c_{31}\frac{\delta_{01}}{\pi_{01}} E[g \mid X]+ 
+  c_{33}\frac{\delta_{01}}{\pi_{01}} E[g \mid X, Y_2]+ \\
+  &c_{41}\frac{\delta_{11}}{\pi_{11}} E[g \mid X]+ 
+  c_{42}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_1]+
+  c_{43}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_2]+
+  \left.c_{44}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_1, Y_2]\right\}
+\end{align*}
+
+which we have condensed to Equation~\ref{eq-linest}.
 
@@
 \end{longtable}
+
+We can justify the constraints for the Outcome Robust model by rewriting
+the estimator into the following form:
+
+\begin{align*}
+E_\delta[\hat \theta] &= n^{-1} \sum_{i = 1}^n E_\delta\left[
+E[g \mid X] \left(\frac{\delta_{00}}{\pi_{00}} c_{11} +
+\frac{\delta_{10}}{\pi_{10}}c_{21} \frac{\delta_{01}}{\pi_{01}}c_{31} +
+\frac{\delta_{11}}{\pi_{11}}c_{41}\right)\right. + \\ 
+&\qquad\qquad\qquad\qquad E[g \mid X, Y_1] \left(\frac{\delta_{10}}{\pi_{10}}c_{22} + 
+\frac{\delta_{11}}{\pi_{11}}c_{42}\right) +\\ 
+&\qquad\qquad\qquad\qquad E[g \mid X, Y_2] \left(\frac{\delta_{01}}{\pi_{01}}c_{33} + 
+\frac{\delta_{11}}{\pi_{11}}c_{43}\right) +\\ 
+&\qquad\qquad\qquad\qquad \left.g \frac{\delta_{11}}{\pi_{11}}c_{44}\right] \\
+&= n^{-1} \sum_{i = 1}^n \left\{E[g \mid X](c_{11} + c_{21} + c_{31} + c_{41}) +
+E[g \mid X, Y_1] (c_{22} + c_{42}) +
+E[g \mid X, Y_2] (c_{33} + c_{43}) + g c_{44}\right\}
+\end{align*}
+
+Thus, if the Outcome Robust constraints are satisfied,
+\(E_\delta[\hat \theta] = g\) when we take the expectation with respect
+to the response variable \(\delta\). These constraints make sense
+because if they are satisfied \(E[\hat \theta]\) does not depend on the
+correct specification of the models for \(E[g \mid G_r(X, Y_1, Y_2)]\).
+
+Likewise, we can justify the constraints for the Response Robust model
+by rewriting Equation~\ref{eq-linest} into the form
+
+\begin{align*}
+E_Z [\hat \theta] &= n^{-1} \sum_{i = 1}^n E_Z\left[
+\frac{\delta_{00}}{\pi_{00}}c_{11} E[g \mid X]\right. + \\
+&\qquad\qquad\qquad\qquad\frac{\delta_{10}}{\pi_{10}}\left(c_{21}E[g \mid X] + 
+  c_{22}E[g \mid X, Y_1]\right) +\\
+&\qquad\qquad\qquad\qquad\frac{\delta_{01}}{\pi_{01}}\left(c_{31}E[g \mid X] + 
+  c_{33}E[g \mid X, Y_2]\right) +\\
+&\qquad\qquad\qquad\qquad\left.\frac{\delta_{11}}{\pi_{11}}\left(c_{41}E[g \mid X] + 
+  c_{42}E[g \mid X, Y_1] + c_{43}E[g \mid X, Y_2] + c_{44}g\right)\right] \\
+& n^{-1} \sum_{i = 1}^n \left(\frac{\delta_{00}}{\pi_{00}}c_{11} + 
+\frac{\delta_{10}}{\pi_{10}}(c_{21} + c_{22}) + 
+\frac{\delta_{01}}{\pi_{01}}(c_{31} + c_{33}) + 
+\frac{\delta_{11}}{\pi_{11}}(c_{41} + c_{42} + c_{43} + c_{44})\right)g
+\end{align*}
+
+Thus, if the Response Robust constraints are satisfied,
+\(E_X[\hat \theta] = g\) when we take the expectation with respect to
+the outcome variable \(Z = (X, Y_1, Y_2)'\). These constraints make
+sense because when they are satisfied \(E[\hat \theta]\) does not depend
+on the correct response probabilities \(\pi = E[\delta]\).
 
@@
 Parametric 1 & Parametric 2 & 0.00 & 1 \\
-Parametric 1 & Outcome Robust & 1190.30 & 0 \\
+Parametric 1 & Outcome Robust & 799.85 & 0 \\
 Parametric 1 & Response Robust & 0.00 & 1 \\
-Parametric 1 & Double Robust & 510.13 & 0 \\
-Parametric 2 & Outcome Robust & 1787.26 & 0 \\
+Parametric 1 & Double Robust & 342.79 & 0 \\
+Parametric 2 & Outcome Robust & 1200.99 & 0 \\
 Parametric 2 & Response Robust & 0.00 & 1 \\
-Parametric 2 & Double Robust & 595.75 & 0 \\
+Parametric 2 & Double Robust & 400.33 & 0 \\
 Outcome Robust & Double Robust & 0.00 & 1 \\
-Response Robust & Double Robust & 895.43 & 0 \\
+Response Robust & Double Robust & 601.70 & 0 \\
 \end{longtable}

### 2024-01-30 Notes/simulation_setup.tex
Finished GLS simulation.

--- 2024-01-16 Notes/efficiencyloss_semi.tex
@@
 
-\renewcommand{\arraystretch}{1.2}
-
 \begin{document}
 
-\lhead{Caleb Leedy}
-\chead{Non-monotone Missingness}
+%\lhead{Caleb Leedy}
+\chead{Nonmonotone Missingness}
 %\chead{STAT 615 - Advanced Bayesian Methods}
 %\rhead{Page \thepage\ of \pageref{LastPage}}
-\rhead{\today}
+\rhead{January 25, 2024}
 
-\section*{The Efficiency Loss of Semiparametric Inference}
-
-Inspired by Dr. Fuller's note, I set out to see if we could assess the amount 
-of efficiency loss due to using semiparametric inference. First, let's define 
-some notation.
-
-\begin{table}[ht!]
-\centering
-\begin{tabular}{lrrr}
-  \toprule
-  Segment & Quantity of Interest & Estimator & Coefficient \\
-  \midrule
-  $A_{00}$ & $E[g \mid X]$ & $\hat \gamma_{11}$ & $c_{11}$ \\
-  $A_{10}$ & $E[g \mid X]$ & $\hat \gamma_{21}$ & $c_{21}$ \\
-  $A_{10}$ & $E[g \mid X, Y_1]$ & $\hat \gamma_{22}$ & $c_{22}$ \\
-  $A_{01}$ & $E[g \mid X]$ & $\hat \gamma_{31}$ & $c_{31}$ \\
-  $A_{01}$ & $E[g \mid X, Y_2]$ & $\hat \gamma_{33}$ & $c_{33}$ \\
-  $A_{11}$ & $E[g \mid X]$ & $\hat \gamma_{41}$ & $c_{41}$ \\
-  $A_{11}$ & $E[g \mid X, Y_1]$ & $\hat \gamma_{42}$ & $c_{42}$ \\
-  $A_{11}$ & $E[g \mid X, Y_2]$ & $\hat \gamma_{43}$ & $c_{43}$ \\
-  $A_{11}$ & $E[g \mid X, Y_1, Y_2]$ & $\hat \gamma_{44}$ & $c_{44}$ \\
-  \bottomrule
-\end{tabular}
-\caption{This table shows the relationship between different quantities of
-interest in each segment, their estimators, and the corresponding coefficients
-of their estimators.}
-\label{tab:reftab}
-\end{table}
-
-This is a similar framework to understanding the class of estimators 
-that we discussed in \verb|optest_update.pdf|, where we write the estimator 
-as a weighted average of different expectation with functional coefficients.
-
-\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
-\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
-\mid X, Y_2].\]
-
-However, now we write the estimator as
-
-\[\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}\]
-
-where $\hat \gamma_{kt} = \frac{\delta_{ij}}{\pi_{ij}}E[g \mid G_{ij}(X, Y_1,
-Y_2)]$ and $ij$ corresponds to the segment $A_{ij}$ associated with $\hat
-\gamma_{kt}$ in Table~\ref{tab:reftab}.
-
-\newpage
-
-\subsection*{Different Classes of Models}
-
-As noted by Dr. Fuller, we can understand parametric estimation as solving 
-the following constrained optimization problems:
-
-\begin{align*}
-  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
-  \sum_{k,t} c_{kt} = 1 \\
-       &\text{or} \\
-  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
-  \sum_{(k,t): (k,t) \neq (4, 4)} c_{kt} = 0 \text{ and } c_{44} = 1. \\
-\end{align*}
-
-To be robust to the outcome model, we can construct a similar optimization
-problem with different constraints:
-
-\begin{align*}
-  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
-  c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0,\\
-       &\text{and } c_{44} = 1.
-\end{align*}
-
-Likewise to be robust to the response model, we can have the problem:
-
-\begin{align*}
-  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
-  c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \\
-       &\text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}.
-\end{align*}
-
-To be double robust (robust to the outcome and response model) we can combine 
-the last two constraints. This is summarized in Table~\ref{tab:mods}.
-
-\begin{table}[ht!]
-  \centering
-  \caption{This table identifies the different constraints for each model type.}
-  \label{tab:mods}
-  \begin{tabular}{lp{12cm}}
-    \toprule
-    Type & Constraints \\
-    \midrule
-    Parametric 1 & $\sum_{k, t} c_{kt} = 1$ \\
-    Parametric 2 &  $\sum_{k, t: (k, t) \neq (4, 4)} c_{kt} = 0, c_{44} = 1$ \\
-    Outcome Robust & $c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} =
-    0, c_{33} + c_{43} = 0, \text{ and } c_{44} = 1.$ \\
-    Response Robust & $c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} +
-    c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} =
-    \pi_{11}$ \\
-    Double Robust & $c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} =
-    0, c_{33} + c_{43} = 0, c_{44} = 1,$ \hspace{3cm} $
-    c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} +
-    c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}$ \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\subsection*{Simulation Study}
-
-\subsubsection*{Simulation 1}
+\section*{Simulation Study Setup}
 
@@
   y_1 &= x + e_1 \\
-  y_2 &= \theta + x + e_2 \\
+  y_2 &= \mu + x + e_2 \\
 \end{align*}
 
-Furthermore, $\pi_{11} = 0.2$ and $\pi_{00} = 0.3, \pi_{10} = 0.4$, and 
-$\pi_{01} = 0.1$.
-The goal of this simulation study is to find $\theta = E[Y_2]$. In other words,
-$g(Z) = Y_2$. There are several algorithms for comparison which are defined 
-as the following:
+This yields outcome variables $Y_1$ and $Y_2$ that are correlated both with $X$
+and additionally with each other. To generate the missingness pattern, we draw 
+from a categorical distribution with selection probabilities $(\pi_{00}, \pi_{10},
+\pi_{01}, \pi_{11})$. Each selection probability, $\pi_{j_1, j_2}$ indicates 
+the probability that an observation is selected into a segment $A_{j_1, j_2}$ 
+where $A_{00}$ indicates that only $X$ is observed, $A_{10}$ means that $X$ and
+$Y_1$ are observed, $A_{01}$ has only $X$ and $Y_2$ observed, and $A_{11}$
+observes $X$, $Y_1$, and $Y_2$.
+
+% NOTE: I change theta to mu.
+In addition to varying the values of the parameters $\mu$ and $\rho$, there
+are two main factors of the simulation that change: the distribution of the
+categories and the parameter of interest, $\theta$. In a ``balanced''
+distribution, we have the following selection probabilities: $\pi_{00} = 0.2$,
+$\pi_{10} = 0.2$, $\pi_{01} = 0.2$, and $\pi_{11} = 0.4$. In the ``unbalanced''
+distribution we have $\pi_{00} = 0.3$, $\pi_{10} = 0.4$, $\pi_{01} = 0.1$, and
+$\pi_{11} = 0.2$. The two types of $\theta$ values that we consider are a
+linear estimate, $\theta = E[g(Z)] = E[Y_2]$ and a non-linear estimate, 
+$\theta = E[g(Z)] = E[Y_1^2 Y_2]$. For notational purposes, we define $Z = (X,
+Y, Z)$ and occasionally we will use the notation $G_r(Z)$ to indicate the
+general form of the observed data in rth segment. We also use $n$ to denote to
+total number of observations in the sample. While $n$ is known, due to our
+setup, the number of elements in each segment $A_{j_1, j_2}$ is random.
+
+There are several algorithms for comparison which are defined as the following:
 
@@
 
-Furthermore, we include four existing estimators that we have already 
-proposed in the past: WLS, Prop, PropInd, and SemiDelta.
-
-The estimator WLS is a a weight least square estimator derived in the 
-following manner. We now consider a normal model:
-
-    \[ 
-      \begin{pmatrix}
-        x_i \\ e_{1i} \\ e_{2i}
-      \end{pmatrix} \stackrel{ind}{\sim}
-      N \left(
-      \begin{bmatrix}
-        0 \\ 0 \\ 0
-      \end{bmatrix},
-      \begin{bmatrix}
-        1 & 0 & 0 \\
-        0 & \sigma_{11} & \sigma_{12} \\ 
-        0 & \sigma_{12} & \sigma_{22}
-      \end{bmatrix},
-      \right)
-    \]
-
-    and define $y_{1i} = \theta_1 + x_i + e_{1i}$ and $y_{2i} = \theta_2 + x_i
-    + e_{2i}$. Then $b_1 = b_2 = 1$. We define $\bar z_k^{(ij)}$ as the mean of
-    $y_k$ in segment $A_{ij}$. This means that we have means $\bar z_1^{(11)}$,
-    $\bar z_2^{(11)}$, $\bar z_1^{(10)}$, and $\bar z_2^{(01)}$. Let $W = [\bar
-    z_1^{(11)}, \bar z_2^{(11)}, \bar z_1^{(10)}, \bar z_2^{(01)}]'$, then for
-    $n_{ij} = |A_{ij}|$, we have
-
-    \[Z - M \mu \sim N(\vec 0, V)\]
-
-    where
-
-    \[M = 
-      \begin{bmatrix}
-        1 & 0 \\
-        0 & 1 \\
-        1 & 0 \\
-        0 & 1 \\
-      \end{bmatrix}
-      \text{ and }
-      V = 
-      \begin{bmatrix}
-        \frac{\sigma_{11}}{n_{11}} & \frac{\sigma_{12}}{n_{11}} & 0 & 0 \\
-        \frac{\sigma_{12}}{n_{11}} & \frac{\sigma_{22}}{n_{11}} & 0 & 0 \\
-        0 & 0 & \frac{\sigma_{11}}{n_{10}} & 0 \\
-        0 & 0 & 0 & \frac{\sigma_{22}}{n_{01}} \\
-      \end{bmatrix}.
-    \]
-
-    Thus, the BLUE for $\mu = [\mu_1, \mu_2]'$ is
-
-    \[\hat \mu = (M' V^{-1} M)^{-1} M' V^{-1} W.\]
-
-    Hence, WLS is $\mu_2$ as $g(X, Y_1, Y_2) = Y_2$.
-
-The remaining three estimators are drived from the following expression
-where the values for $\beta$ are provided in Table~\ref{tab:beta}. These 
-are good estimators to compare to because Prop is the original proposed 
-estimator. PropInd has the same form as Prop except the values for $\beta$ 
-is different. PropInd shares the same values of $\beta$ as all of the new 
-models with constraints. SemiDelta is useful because it is the best 
-estimator in general so far.
-
-\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
-\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
-\mid X, Y_2].\]
-
-% TODO: We use the unequal weights
-\begin{table}[ht!]
-  \centering
-  \caption{This table displays the values of $\beta$ for different estimator 
-  types.}
-  \label{tab:beta}
-\begin{tabular}{lrr}
-  \toprule
-  Estimator & $\beta_0(\delta, c_0)$ & $\beta_1(\delta, c_1)$ \\
-  \midrule
-  Prop & $\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - 
-  \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
-  $\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)$ \\
-  PropInd & $\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - 
-  \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
-  $\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)$ \\
-%  $\hat \theta_{c}$ & $c_0\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - 
-%  \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
-%  $c_1\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)$ \\
-%  $\hat \theta_{c, ind}$ & $c_0\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - 
-%  \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
-%  $c_1\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)$ \\
-  SemiDelta & $c_0\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{00}}{\pi_{00}}\right)$ &
-  $c_1\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{10}}{\pi_{10}}\right)$ \\
-  \bottomrule
-\end{tabular}
-\end{table}
-
-In the simulation results in Table~\ref{tab:sim1}, the new results have the 
-same label as the value from the ``Type'' column in Table~\ref{tab:mods}.
-
-\begin{table}[ht!]
-  \caption{Results from Simulation 1. True values: $\theta = 5, \rho = 0.5$.
-  The test conducted for the T-statistic and P-value is a two sample test to 
-  see if the estimator is unbiased.}
-  \label{tab:sim1}
-  \centering
-  \begin{tabular}[t]{lrrrr}
-    \toprule
-    Algorithm & Bias & SD & Tstat & P-value\\
-    \midrule
-    Oracle & -0.002 & 0.045 & -1.669 & 0.048\\
-    CC & 0.001 & 0.083 & 0.263 & 0.396\\
-    IPW & 0.010 & 0.357 & 0.872 & 0.192\\
-    WLS & 0.000 & 0.054 & -0.065 & 0.474\\
-    %WLS (with true $\theta$) & -0.001 & 0.035 & -1.158 & 0.124\\
-    Prop & -0.002 & 0.063 & -0.768 & 0.221\\
-    PropInd & -0.003 & 0.120 & -0.900 & 0.184\\
-    SemiDelta & -0.001 & 0.064 & -0.342 & 0.366\\
-    Parametric 1 & -0.001 & 0.061 & -0.509 & 0.305\\
-    Parametric 2 & -0.001 & 0.061 & -0.509 & 0.305\\
-    Outcome Robust & -0.001 & 0.061 & -0.543 & 0.294\\
-    Response Robust & -0.001 & 0.061 & -0.509 & 0.305\\
-    Double Robust & -0.001 & 0.061 & -0.543 & 0.294\\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\newpage
-
-\subsubsection*{Simulation 2}
-
-Next, we used the same simulation setup except that we are interested in 
-estimating $\theta = E[g] = E[Y_1^2 Y_2]$. We do not use the WLS estimator 
-because it does not work.
-
-\begin{table}[ht!]
-  \caption{Results from Simulation 2. True values: $\theta = 10, \rho = 0.5$.
-  The test conducted for the T-statistic and P-value is a two sample test to 
-  see if the estimator is unbiased.}
-  \label{tab:sim2}
-  \centering
-  \begin{tabular}[t]{lrrrr}
-    \toprule
-    Algorithm & Bias & SD & Tstat & P-value\\
-    \midrule
-    Oracle & -0.037 & 0.528 & -2.214 & 0.014\\
-    CC & 0.007 & 1.196 & 0.179 & 0.429\\
-    IPW & 0.023 & 1.405 & 0.520 & 0.302\\
-    Prop & -0.044 & 0.691 & -2.027 & 0.021\\
-    SemiDelta & -0.032 & 0.674 & -1.521 & 0.064\\
-    Parametric 1 & -0.004 & 0.301 & -0.460 & 0.323\\
-    Parametric 2 & -0.004 & 0.344 & -0.341 & 0.367\\
-    Outcome Robust & -0.040 & 0.654 & -1.950 & 0.026\\
-    Response Robust & -0.004 & 0.301 & -0.461 & 0.322\\
-    Double Robust & -0.040 & 0.654 & -1.954 & 0.025\\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\newpage
-
-\subsection*{Discussion}
-
-\begin{itemize}
-  \item In the first simulation there does not seem to be much of a cost 
-    to using a semiparametric estimator with the Double Robust model 
-    performing as well as the Outcome Robust, Response Robust and 
-    Parametric models. However, in the second simulation the Parametric 
-    models and Response Robust estimators perform the best.
-  \item Strangly, the Double Robust estimator in the second simulation 
-    has picked up the bad characteristics of the Outcome Model. This 
-    can make sense if the optimal Outcome Model is contained in the 
-    space of the Double Robust model but the optimal Response Model 
-    is not.
-\end{itemize}
-
 \end{document}

### 2024-01-30 Simulations/T4GetVarianceMin/variance_of_minimum.qmd
Finished GLS simulation.

---
title: "Non-Monotone Missingness: The Minimal Variance of the Linear Estimate"
author: "Caleb Leedy"
date: today
date-format: "D MMMM YYYY"
format: 
  pdf:
    keep-tex: true
    include-in-header: latex_header.tex
---

# Introduction

The goal of this project is to get an optimal (or near optimal) estimator of a
general estimator when the data has a non-monotone missing structure. So far we
have reduced the problem to estimating the following:

$$
\hat \theta = \frac{\delta_{11}}{\pi_{11}} g(Z) + 
\lambda_0(\delta) \alpha_0(X) +
\lambda_1(\delta) \alpha_1(X, Y_1) +
\lambda_2(\delta) \alpha_2(X, Y_2).
$$ {#eq-estclass}

For notation we have variables $Z = (X, Y_1, Y_2)$ that are observed in the
segment noted in @tbl-segments.

| Segment | Variables | 
| ------- | ----------|
| $A_{00}$| $X$       |
| $A_{10}$| $X, Y_1$  |
| $A_{01}$| $X, Y_2$  |
| $A_{11}$| $X, Y_1, Y_2$ |
: This table matches the segment with its associated observed variables. {#tbl-segments}

For each segment $A_{d_1 d_2}$ the probability of an observation being in a specific
segment is known to be $\pi_{d_1 d_2}$, and the associated random variable that
indicates whether observation $i$ is in $A_{d_1 d_2}$ is $\delta_{d_1 d_2 i}$.
For brevity, it will sometimes be convienent to drop the subscript $i$. To get
an optimal estimator, we need to choose values of $\lambda$ and $\alpha$ to
minimize the variance of $\hat \theta$.

# Simplifications

To simplify this model consider the case in which each $\lambda$ *and* $\alpha$
are linear functions. This means that we have the following: 
$\lambda_0 = \lambda_0^{(0)} + \lambda_1^{(0)}\delta_{00} + \lambda_2^{(0)}
\delta_{10} + \lambda_3^{(0)} \delta_{01} + \lambda_4^{(0)}\delta_{11}$,
$\lambda_1 = \lambda_0^{(1)} + \lambda_2^{(1)} \delta_{10} +
\lambda_4^{(1)}\delta_{11}$, and 
$\lambda_2 = \lambda_0^{(2)} + \lambda_3^{(2)} \delta_{01} +
\lambda_4^{(2)}\delta_{11}$. Also, 
$\alpha_0 = \alpha_0^{(0)} + \alpha_1^{(0)} x$, 
$\alpha_1 = \alpha_0^{(1)} + \alpha_1^{(1)} x + \alpha_2^{(1)} y_1$, and
$\alpha_2 = \alpha_0^{(2)} + \alpha_1^{(1)} x + \alpha_3^{(1)} y_2$. This means
that we can express the estimator in @eq-estclass as the following matrix
equation:

$$
\hat \theta = n^{-1} \bf{1}_n' \left(\frac{\delta_{11}}{\pi_{11}} g(Z) + 
(\pmb{\alpha} \pmb{\lambda}' \pmb{\delta})' \bf Z\right).
$$

where

$$
\pmb{\alpha} = 
\begin{bmatrix}
\alpha_0^{(0)} & \alpha_0^{(1)} & \alpha_0^{(2)} \\
\alpha_1^{(0)} & \alpha_1^{(1)} & \alpha_1^{(2)} \\
0 & \alpha_2^{(1)} & 0 \\
0 & 0 & \alpha_3^{(2)} \\
\end{bmatrix},
\pmb{\lambda} = 
\begin{bmatrix}
\lambda_0^{(0)} & \lambda_0^{(1)} & \lambda_0^{(2)} \\
\lambda_1^{(0)} & 0 & 0 \\
\lambda_2^{(0)} & \lambda_2^{(1)} & 0 \\
\lambda_3^{(0)} & 0 & \lambda_3^{(2)} \\
\lambda_4^{(0)} & \lambda_4^{(1)} & \lambda_4^{(2)} \\
\end{bmatrix},
\pmb{\delta}= 
\begin{bmatrix}
1 \\ \delta_{00} \\ \delta_{10} \\ \delta_{01} \\ \delta_{11}
\end{bmatrix}
\bf{Z}_i = 
\begin{bmatrix}
1 \\ X_i \\ Y_{1i} \\ Y_{2i}
\end{bmatrix}
$$

This is a linear estimator and the goal is to find the optimal $\pmb{\alpha}$
and $\pmb{\lambda}$ that,

$$ \text{minimize } \Var(\hat \theta) \text{ such that } 
\pmb{\lambda}' E[\pmb{\delta}] = 0.$$

# Simpliying the Variance

We know that

\begin{align*}
\Var\left(\frac{\delta_{11}}{\pi_{11}}g(Z)) + 
  (\pmb \alpha \pmb \lambda' \pmb \delta)' \bf Z\right) 
&= \Var\left(\frac{\delta_{11}}{\pi_{11}}g(Z)\right) +
  \Var((\pmb \alpha \pmb \lambda' \pmb \delta)'\bf Z) +
  2 \Cov\left(\frac{\delta_{11}}{\pi_{11}}g(Z), 
  (\pmb \alpha \pmb \lambda' \pmb \delta)' \bf Z\right)\\
&:= V_1 + V_2 + V_3
\end{align*}

Hence, we have for $\Pi := E[\delta]$,

\begin{align*}
V_1 &= \Var\left(\frac{\delta_{11}}{\pi_{11}}g(Z)\right) \\ 
&= \Var\left(E\left[\frac{\delta_{11}}{\pi_{11}}g(Z) \mid Z\right]\right)
+ E\left[\Var\left(\frac{\delta_{11}}{\pi_{11}}g(Z) \mid Z\right)\right]\\
  &= E[g^2(Z)] - E[g(Z)]^2 + \frac{1}{\pi_{11}}E[g^2(Z)] - E[g^2(Z)] \\ 
  &= \frac{1}{\pi_{11}} E[g^2(Z)] - E[g(Z)]^2
\end{align*}

\begin{align*}
V_2 &= \Var((\alpha \lambda' \delta)'Z) \\ 
&= \Var(E[(\alpha \lambda' \delta)'Z \mid Z]) + 
E[\Var((\alpha \lambda' \delta)'Z \mid Z)] \\ 
&= \Pi' \lambda \alpha'  \Var(Z) \alpha \lambda' \Pi + 
E[Z'\alpha \lambda' \Var(\delta)\lambda \alpha' Z]\\
&= \Pi' \lambda \alpha'  (E[ZZ'] - E[Z]E[Z']) \alpha \lambda' \Pi + 
E[Z'\alpha \lambda' (E[\delta \delta'] - \Pi\Pi')\lambda \alpha' Z]\\
&= \Pi' \lambda \alpha'  E[ZZ'] \alpha \lambda' \Pi -
\Pi' \lambda \alpha' E[Z]E[Z'] \alpha \lambda' \Pi + 
E[Z'\alpha \lambda' E[\delta \delta']\lambda \alpha' Z] - 
E[Z'\alpha \lambda' \Pi\Pi'\lambda \alpha' Z]\\
&= \Pi' \lambda \alpha'  E[ZZ'] \alpha \lambda' \Pi -
\Pi' \lambda \alpha' E[Z]E[Z'] \alpha \lambda' \Pi + 
E[Z'\alpha \lambda' E[\delta \delta']\lambda \alpha' Z] - 
\Pi' \lambda \alpha' E[ZZ'] \alpha \lambda' \Pi \\
&= \Pi' \lambda \alpha' E[Z]E[Z'] \alpha \lambda' \Pi + 
E[Z'\alpha \lambda' E[\delta \delta']\lambda \alpha' Z] \\
&= E[Z'\alpha \lambda' E[\delta \delta']\lambda \alpha' Z] \\
\end{align*}

The third to last equality holds because a $1 \times 1$ matrix is always
symmetric and the last equality holds because $\Pi' \lambda = (\lambda' \Pi)' =
0'$. We can expand this expression but we get the horribly ugly result:

\begin{RaggedRight}

```{python}
#| echo: false
#| output: asis

import sympy

l00, l01, l02 = sympy.symbols('\lambda_0^{(0)} \lambda_0^{(1)} \lambda_0^{(2)}')
l10, l20, l21 = sympy.symbols('\lambda_1^{(0)} \lambda_2^{(0)} \lambda_0^{(1)}')
l30, l32 = sympy.symbols('\lambda_3^{(0)} \lambda_3^{(2)}')
l40, l41, l42 = sympy.symbols('\lambda_4^{(0)} \lambda_4^{(1)} \lambda_4^{(2)}')
lmat = sympy.Matrix([
  [l00, l01, l02],
  [l10, 0,   0 ],
  [l20, l21, 0],
  [l30, 0,   l32],
  [l40, l41, l42]
])

a00, a01, a02 = sympy.symbols('\\alpha_0^{(0)} \\alpha_0^{(1)} \\alpha_0^{(2)}')
a10, a11, a12 = sympy.symbols('\\alpha_1^{(0)} \\alpha_1^{(1)} \\alpha_1^{(2)}')
a21, a32 = sympy.symbols('\\alpha_2^{(1)} \\alpha_3^{(2)}')
amat = sympy.Matrix([
  [a00, a01, a02],
  [a10, a11, a12],
  [0,   a21, 0],
  [0,   0,   a32]
])

p00, p10, p01, p11 = sympy.symbols('\pi_{00} \pi_{10} \pi_{01} \pi_{11}')
ddmat = sympy.Matrix([
  [1,   p00, p10, p01, p11],
  [p00, p00, 0,   0,   0],
  [p10, 0,   p10, 0,   0],
  [p01, 0,   0,   p01, 0],
  [p11, 0,   0,   0,   p11]
])

x, y1, y2 = sympy.symbols('X Y_1 Y_2')
zmat = sympy.Matrix([[1, x, y1, y2]]).T

ans = zmat.T * amat * lmat.T * ddmat * lmat * amat.T * zmat
print(sympy.latex(sympy.expand(ans[0, 0]), itex = True, mode = "inline"))

```
\end{RaggedRight}

Finally, to understand $V_3$ we can solve for the third covariance term.

\begin{align*}
\Cov&\left(\frac{\delta_{11}}{\pi_{11}}g(Z), (\alpha \lambda' \delta)' Z\right)\\
&= E\left[\frac{\delta_{11}}{\pi_{11}}g(Z)(\alpha \lambda' \delta)' Z\right] - 
E\left[\frac{\delta_{11}}{\pi_{11}} g(Z)\right] E[(\alpha \lambda' \delta)'Z]\\
&= E\left[g(Z)\frac{\delta_{11}}{\pi_{11}}\delta' \lambda \alpha' Z\right] - 
E[g(Z)] (\alpha \lambda' \Pi)' E[Z]\\
&= E[g(Z) \begin{bmatrix}1 & 0 & 0 & 0 & 1\end{bmatrix} \lambda \alpha' Z] \\
&= (\lambda_0^{(0)} + \lambda_4^{(0)})(\alpha_0^{(0)} + \alpha_1^{(0)}E[g(Z)x])
+ (\lambda_0^{(1)} + \lambda_4^{(1)})(\alpha_0^{(1)} + \alpha_1^{(1)}E[g(Z)x] +
\alpha_2^{(1)}E[g(Z)y_1]) \\
&\qquad \qquad+ (\lambda_0^{(2)} + \lambda_4^{(2)})(\alpha_0^{(2)} +
\alpha_1^{(2)}E[g(Z)x] + \alpha_3^{(2)}E[g(Z)y_2])
\end{align*}

Since we have an understanding of $\Var(\hat \theta)$, we can find the minimum 
by differentiating with respect to each coefficient in $\lambda$ and $\alpha$.

```{python}
#| echo: false

egx, egy1, egy2 = sympy.symbols('E[gX] E[gY_1] E[gY_2]')
cov = (l00 + l40) * (a00 + a10 * egx) + (l01 + l41) * (a01 + a11 * egx + a21 *
  egy1) + (l02 + l42) * (a02 + a12 * egx + a32 * egy2)

var = ans[0, 0] + 2 * cov

# sympy.diff(var, l00)
# sympy.diff(var, l01)
# sympy.diff(var, l02)
# sympy.diff(var, l10)
# sympy.diff(var, l20)
# sympy.diff(var, l21)
# sympy.diff(var, l30)
# sympy.diff(var, l32)
# sympy.diff(var, l40)
# sympy.diff(var, l41)
# sympy.diff(var, l42)
# 
# sympy.diff(var, a00)
# sympy.diff(var, a01)
# sympy.diff(var, a02)
# sympy.diff(var, a10)
# sympy.diff(var, a11)
# sympy.diff(var, a12)
# sympy.diff(var, a21)
# sympy.diff(var, a32)

```

This is still a work in progress. I need to run a simulation and test it out.
I also believe that the terms $\alpha_0^{(0)}$, $\alpha_0^{(1)}$, and 
$\alpha_0^{(2)}$ hinder identifiability. But I will work on it.

```{r}
#| echo: false

library(dplyr)
library(CVXR)

source("~/Research/Data_Integration/R/opt_est.R")
source("~/Research/Data_Integration/Simulations/T3Assess_Semiparametric_Cost/compare_algs.R")

```

```{r}

linear_const_min <- function(df, gfun = "Y2") {

  df <- mutate(df, g_i = eval(rlang::parse_expr(gfun)))
  df_11 <- filter(df, delta_1 == 1, delta_2 == 1)
  df_10 <- filter(df, delta_1 == 1, delta_2 == 0)
  df_01 <- filter(df, delta_1 == 0, delta_2 == 1)
  df_00 <- filter(df, delta_1 == 0, delta_2 == 0)

  # Set up objective
  lambda <- Variable(rows = 5, cols = 3)
  alpha <- Variable(rows = 3, cols = 3)
  Z <- cbind(df$X, df$Y1, df$Y2)
  dd_mat <- matrix(c(1, pi_00, pi_10, pi_01, pi_11,
                     pi_00, pi_00, 0, 0, 0,
                     pi_10, 0, pi_10, 0, 0,
                     pi_01, 0, 0, pi_01, 0,
                     pi_11, 0, 0, 0, pi_11), nrow = 5, byrow = TRUE)

  mat <- quad_form(quad_form(t(Z %*% alpha %*% t(lambda)), dd_mat), diag(3))
  gz_mat <- matrix(df$g_i, ncol = 1)
  s_mat <- matrix(c(1, 0, 0, 0, 1), ncol = 1)
  covariance <- 
    matrix_trace(quad_form(Z %*% alpha %*% t(lambda) %*% s_mat %*% gz_mat),
                           diag(nrow(df)))
  obj <- Minimize(multiply(matrix_trace(mat), 1 / nrow(df)) + covariance))

}

```

### 2024-01-31 Notes/gls_explanation.tex
Updated simulation to incorporate single selection of theta vector.

--- 2024-01-22 Notes/semiconstraints.tex
@@
 
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{241,243,245}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
-\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
-
 \providecommand{\tightlist}{%
@@
 \newcommand{\logistic}{{\text{logistic}}}
+\renewcommand{\arraystretch}{1.4}
 \KOMAoption{captions}{tableheading}
@@
 \hypersetup{
-  pdftitle={Understanding Semiparametric Constraints},
+  pdftitle={GLS Explanation},
   pdfauthor={Caleb Leedy},
@@
 
-\title{Understanding Semiparametric Constraints}
+\title{GLS Explanation}
 \author{Caleb Leedy}
-\date{19 January 2024}
+\date{31 January 2024}
 
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, breakable]}{\end{tcolorbox}}\fi
 
-\hypertarget{sec-intro}{%
-\section{Introduction}\label{sec-intro}}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-The goal of this report is to understand how additional constraints
-effect semiparametric models. Inspired by Dr.~Fuller's note, I set out
-to see if we could assess the amount of efficiency loss due to using
-semiparametric inference. First, let's define some notation.
+This is a write up explaning the dual form of generalized least squares
+(GLS). We think of GLS as an extension of ordinary least squares (OLS),
+so I will discuss this first and then comment on how these are connected
+and the dual formulation for both.
 
-\hypertarget{tbl-reftab}{}
-\begin{longtable}[]{@{}llll@{}}
-\caption{\label{tbl-reftab}This table shows the relationship between
-different quantities of interest in each segment, their estimators, and
-the corresponding coefficients of their estimators.}\tabularnewline
-\toprule\noalign{}
-Segment & Quantity of Interest & Estimator & Coefficient \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-Segment & Quantity of Interest & Estimator & Coefficient \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-\(A_{00}\) & \(E[g \mid X]\) & \(\hat \gamma_{11}\) & \(c_{11}\) \\
-\(A_{10}\) & \(E[g \mid X]\) & \(\hat \gamma_{21}\) & \(c_{21}\) \\
-\(A_{10}\) & \(E[g \mid X, Y_1]\) & \(\hat \gamma_{22}\) & \(c_{22}\) \\
-\(A_{01}\) & \(E[g \mid X]\) & \(\hat \gamma_{31}\) & \(c_{31}\) \\
-\(A_{01}\) & \(E[g \mid X, Y_2]\) & \(\hat \gamma_{33}\) & \(c_{33}\) \\
-\(A_{11}\) & \(E[g \mid X]\) & \(\hat \gamma_{41}\) & \(c_{41}\) \\
-\(A_{11}\) & \(E[g \mid X, Y_1]\) & \(\hat \gamma_{42}\) & \(c_{42}\) \\
-\(A_{11}\) & \(E[g \mid X, Y_2]\) & \(\hat \gamma_{43}\) & \(c_{43}\) \\
-\(A_{11}\) & \(E[g \mid X, Y_1, Y_2]\) & \(\hat \gamma_{44}\) &
-\(c_{44}\) \\
-\end{longtable}
+\hypertarget{the-gauss-markov-theorem}{%
+\section{The Gauss-Markov Theorem}\label{the-gauss-markov-theorem}}
 
-This is a similar framework to understanding the class of estimators
-that we discussed in \verb|optest_update.pdf|, where we write the
-estimator as a weighted average of different expectation with functional
-coefficients.
+It all comes back to the Gauss-Markov Theorem, which states that for a
+linear model in the form,
 
-\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
-\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
-\mid X, Y_2].\]
+\[ Y = X\beta + \varepsilon\]
 
-However, now we write the estimator as
+with \(E[\varepsilon] = 0\) and \(\Var(\varepsilon) = \sigma^2 I_n\),
+the BLUE (best linear unbiased estimator) of \(C\beta\) for \(C = AX\)
+for some matrix \(A\) is (assuming \(X'X\) has an inverse),
 
-\[\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}\]
+\[ C\hat \beta = C(X'X)^{-1} X'Y = AX(X'X)^{-1}X'Y.\]
 
-where
-\(\hat \gamma_{kt} = \frac{\delta_{ij}}{\pi_{ij}}E[g \mid G_{ij}(X, Y_1, Y_2)]\)
-and \(ij\) corresponds to the segment \(A_{ij}\) associated with
-\(\hat \gamma_{kt}\) in Table~\ref{tbl-reftab}.
+\hypertarget{gls-extension}{%
+\subsection{GLS Extension}\label{gls-extension}}
 
-\hypertarget{sec-methods}{%
-\section{Methods}\label{sec-methods}}
+To extend this to the GLS setting consider the case where
+\(\Var(\varepsilon) = \sigma^2 V\) for some known positive-definite
+matrix \(V\). In this case, we can rewrite \(Y = X\beta + \varepsilon\)
+as
 
-As noted by Dr.~Fuller, we can understand parametric estimation as
-solving the following constrained optimization problems:
+\[ Z = W\beta + \tilde \varepsilon\]
 
-\begin{align*}
-  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
-  \sum_{k,t} c_{kt} = 1 \\
-       &\text{or} \\
-  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
-  \sum_{(k,t): (k,t) \neq (4, 4)} c_{kt} = 0 \text{ and } c_{44} = 1. \\
-\end{align*}
+where \(Z = V^{-1/2} Y\), \(W = V^{-1/2}X\) and
+\(\tilde \varepsilon = V^{-1/2} \varepsilon\). But this formulation
+implies that \(E[\tilde \varepsilon] = 0\) and
+\(\Var(\tilde \varepsilon) = \sigma^2 I_n\) and so hence by the
+Gauss-Markov Theorem, the BLUE is
 
-To be robust to the outcome model, we can construct a similar
-optimization problem with different constraints:
+\[ (W'W)^{-1} W'Z = (X'V^{-1}V)^{-1} X'V^{-1}Y.\]
 
-\begin{align*}
-  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
-  c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0,\\
-       &\text{and } c_{44} = 1.
-\end{align*}
+\hypertarget{dual-form}{%
+\section{Dual Form}\label{dual-form}}
 
-Likewise to be robust to the response model, we can have the problem:
+All of this comes back to the fact that we want to estimate the BLUE.
+However, we could do this directly. By definition, the BLUE is the
+minimum variance linear unbiased estimator. This means that we consider
+a class of estimators \(CW'Y\) such that \(E[CW'Y] = C\beta\) that
+minimizes the variance \(CW'\Var(Y) WC' \propto CW'WC'\). Since we
+assume the model form of \(Y = X\beta + \varepsilon\), the unbiasedness
+condition is equivalent to,
 
-\begin{align*}
-  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
-  c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \\
-       &\text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}.
-\end{align*}
+\[ E[CW'Y] = CW'E[Y] = CW'X\beta = C\beta. \]
 
-To be double robust (robust to the outcome and response model) we can
-combine the last two constraints. This is summarized in
-Table~\ref{tbl-mods}.
+Since this is true for all \(\beta\), it must be the case that
 
-\hypertarget{tbl-mods}{}
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2941}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7059}}@{}}
-\caption{\label{tbl-mods}This table identifies the different constraints
-for each model type.}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Type
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Constraints
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Type
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Constraints
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Parametric 1 & \(\sum_{k, t} c_{kt} = 1\) \\
-Parametric 2 &
-\(\sum_{k, t: (k, t) \neq (4, 4)} c_{kt} = 0, c_{44} = 1\) \\
-Outcome Robust &
-\(c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, \text{ and } c_{44} = 1.\) \\
-Response Robust &
-\(c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}\) \\
-Double Robust &
-\(c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, c_{44} = 1, c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}\) \\
-\end{longtable}
+\[ W'X = I_p \text{ for } p = \dim(\beta).\]
 
-\hypertarget{sec-simulations}{%
-\section{Simulation Study}\label{sec-simulations}}
+Hence, we have a Lagrangian
 
-\hypertarget{simulation-1}{%
-\subsection*{Simulation 1}\label{simulation-1}}
-\addcontentsline{toc}{subsection}{Simulation 1}
+\[L(W) \propto CW'WC' - (W'X - I_p)\lambda.\]
 
-We use the following simulation setup
+Thus, the first-order conditions are \(W - X\lambda = 0\) and
+\(W'X - I_p = 0\). This yields a solution of \(W = X(X'X)^{-1}\) which
+is exactly what we get from OLS because the final result is
+\(CW'Y = C(X'X)^{-1} X'Y\). To get the GLS form, we can keep the same
+constraint except minimize the variance \(W'VW\).
 
-\begin{align*}
-  \begin{bmatrix} x \\ e_1 \\ e_2 \end{bmatrix} 
-  &\stackrel{ind}{\sim} N\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
-  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}\right) \\
-  y_1 &= x + e_1 \\
-  y_2 &= \theta + x + e_2 \\
-\end{align*}
+\hypertarget{estimating-the-variance}{%
+\section{Estimating the Variance}\label{estimating-the-variance}}
 
-Furthermore, \(\pi_{11} = 0.2\) and \(\pi_{00} = 0.3, \pi_{10} = 0.4\),
-and \(\pi_{01} = 0.1\). The goal of this simulation study is to find
-\(\theta = E[Y_2]\). In other words, \(g(Z) = Y_2\). There are several
-algorithms for comparison which are defined as the following:
-
-\begin{align*}
-  Oracle &= n^{-1} \sum_{i = 1}^n g(Z_i)\\
-  CC &= \frac{\sum_{i = 1}^n \delta_{11} g(Z_i)}{\sum_{i = 1}^n \delta_{11}} \\
-  IPW &= \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g(Z_i)\\
-\end{align*}
-
-Furthermore, we include four existing estimators that we have already
-proposed in the past: WLS, Prop, PropInd, and SemiDelta.
-
-The estimator WLS is a a weight least square estimator derived in the
-following manner. We now consider a normal model:
-
-\[ 
-  \begin{pmatrix}
-    x_i \\ e_{1i} \\ e_{2i}
-  \end{pmatrix} \stackrel{ind}{\sim}
-  N \left(
-  \begin{bmatrix}
-    0 \\ 0 \\ 0
-  \end{bmatrix},
-  \begin{bmatrix}
-    1 & 0 & 0 \\
-    0 & \sigma_{11} & \sigma_{12} \\ 
-    0 & \sigma_{12} & \sigma_{22}
-  \end{bmatrix},
-  \right)
-\]
-
-and define \(y_{1i} = \theta_1 + x_i + e_{1i}\) and
-\(y_{2i} = \theta_2 + x_i + e_{2i}\). Then \(b_1 = b_2 = 1\). We define
-\(\bar z_k^{(ij)}\) as the mean of \(y_k\) in segment \(A_{ij}\). This
-means that we have means \(\bar z_1^{(11)}\), \(\bar z_2^{(11)}\),
-\(\bar z_1^{(10)}\), and \(\bar z_2^{(01)}\). Let
-\(W = [\bar z_1^{(11)}, \bar z_2^{(11)}, \bar z_1^{(10)}, \bar z_2^{(01)}]'\),
-then for \(n_{ij} = |A_{ij}|\), we have
-
-\[Z - M \mu \sim N(\vec 0, V)\]
-
-where
-
-\[M = 
-  \begin{bmatrix}
-    1 & 0 \\
-    0 & 1 \\
-    1 & 0 \\
-    0 & 1 \\
-  \end{bmatrix}
-  \text{ and }
-  V = 
-  \begin{bmatrix}
-    \frac{\sigma_{11}}{n_{11}} & \frac{\sigma_{12}}{n_{11}} & 0 & 0 \\
-    \frac{\sigma_{12}}{n_{11}} & \frac{\sigma_{22}}{n_{11}} & 0 & 0 \\
-    0 & 0 & \frac{\sigma_{11}}{n_{10}} & 0 \\
-    0 & 0 & 0 & \frac{\sigma_{22}}{n_{01}} \\
-  \end{bmatrix}.
-\]
-
-Thus, the BLUE for \(\mu = [\mu_1, \mu_2]'\) is
-
-\[\hat \mu = (M' V^{-1} M)^{-1} M' V^{-1} W.\]
-
-Hence, WLS is \(\mu_2\) as \(g(X, Y_1, Y_2) = Y_2\).
-
-The remaining three estimators are drived from the following expression
-where the values for \(\beta\) are provided in Table~\ref{tbl-beta}.
-These are good estimators to compare to because Prop is the original
-proposed estimator. PropInd has the same form as Prop except the values
-for \(\beta\) is different. PropInd shares the same values of \(\beta\)
-as all of the new models with constraints. SemiDelta is useful because
-it is the best estimator in general so far.
-
-\[
-\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + 
-\beta_0(\delta, c_0)E[g(Z) \mid X] + 
-\beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + 
-\beta_2(\delta, c_2) E[g(Z)
-\mid X, Y_2].
-\]
-
-\hypertarget{tbl-beta}{}
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1698}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4151}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4151}}@{}}
-\caption{\label{tbl-beta}This table displays the values of \(\beta\) for
-different estimator types.}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Estimator
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\(\beta_0(\delta, c_0)\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\(\beta_1(\delta, c_1)\)
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Estimator
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\(\beta_0(\delta, c_0)\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\(\beta_1(\delta, c_1)\)
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Prop &
-\(\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)\)
-&
-\(\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)\) \\
-PropInd &
-\(\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)\)
-&
-\(\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)\) \\
-SemiDelta &
-\(c_0\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{00}}{\pi_{00}}\right)\)
-&
-\(c_1\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{10}}{\pi_{10}}\right)\) \\
-\end{longtable}
-
-In the simulation results in Table~\ref{tbl-sim1}, the new results have
-the same label as the value from the \texttt{Type} column in
-Table~\ref{tbl-mods}.
-
-\hypertarget{tbl-sim1}{}
-\begin{longtable}[]{@{}lrrrr@{}}
-\caption{\label{tbl-sim1}Results from Simulation 1. True values:
-\(\theta = 5, \rho = 0.5\). The test conducted for the T-statistic and
-P-value is a two sample test to see if the estimator is unbiased. This
-simulation uses \(Y_1 = x + \varepsilon_1\),
-\(Y_2 = \theta + x + \varepsilon_2\) where \(X \sim N(0, 1)\) and
-\((\varepsilon_1, \varepsilon_2)\) come from a mean zero bivariate
-normal distribution with unit variance and covariance \(\rho\). The
-segments are unbalanced with: \(\pi_{11} = 0.2\), \(\pi_{10} = 0.4\),
-\(\pi_{01} = 0.1\), and \(\pi_{00} = 0.3\).}\tabularnewline
-\toprule\noalign{}
-Algorithm & bias & sd & tstat & pval \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-Algorithm & bias & sd & tstat & pval \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Oracle & -0.002 & 0.045 & -1.669 & 0.048 \\
-CC & 0.001 & 0.083 & 0.263 & 0.396 \\
-IPW & 0.010 & 0.357 & 0.872 & 0.192 \\
-WLS & 0.000 & 0.054 & -0.065 & 0.474 \\
-\(\hat \theta_{prop}\) & -0.002 & 0.063 & -0.768 & 0.221 \\
-\(\hat \theta_{\delta}\) & -0.001 & 0.064 & -0.342 & 0.366 \\
-Parametric 1 & -0.001 & 0.061 & -0.509 & 0.305 \\
-Parametric 2 & -0.001 & 0.061 & -0.509 & 0.305 \\
-Outcome Robust & -0.001 & 0.061 & -0.543 & 0.294 \\
-Response Robust & -0.001 & 0.061 & -0.509 & 0.305 \\
-Double Robust & -0.001 & 0.061 & -0.543 & 0.294 \\
-\end{longtable}
-
-\hypertarget{simulation-2}{%
-\subsection*{Simulation 2}\label{simulation-2}}
-\addcontentsline{toc}{subsection}{Simulation 2}
-
-Next, we used the same simulation setup except that we are interested in
-estimating \(\theta = E[g] = E[Y_1^2 Y_2]\). We do not use the WLS
-estimator because it does not work.
-
-\hypertarget{tbl-sim2}{}
-\begin{longtable}[]{@{}lrrrr@{}}
-\caption{\label{tbl-sim2}Results from Simulation 2. True values:
-\(\theta = 10, \rho = 0.5\). The test conducted for the T-statistic and
-P-value is a two sample test to see if the estimator is unbiased. This
-simulation uses the same setup as Simulation 1 with
-\(Y_1 = x + \varepsilon_1\), \(Y_2 = \theta + x + \varepsilon_2\) where
-\(X \sim N(0, 1)\) and \((\varepsilon_1, \varepsilon_2)\) come from a
-mean zero bivariate normal distribution with unit variance and
-covariance \(\rho\). The segments are unbalanced with:
-\(\pi_{11} = 0.2\), \(\pi_{10} = 0.4\), \(\pi_{01} = 0.1\), and
-\(\pi_{00} = 0.3\).}\tabularnewline
-\toprule\noalign{}
-Algorithm & bias & sd & tstat & pval \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-Algorithm & bias & sd & tstat & pval \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Oracle & -0.037 & 0.528 & -2.214 & 0.014 \\
-CC & 0.007 & 1.196 & 0.179 & 0.429 \\
-IPW & 0.023 & 1.405 & 0.520 & 0.302 \\
-\(\hat \theta_{prop}\) & -0.044 & 0.691 & -2.027 & 0.021 \\
-\(\hat \theta_{\delta}\) & -0.032 & 0.674 & -1.521 & 0.064 \\
-Parametric 1 & -0.004 & 0.301 & -0.461 & 0.322 \\
-Parametric 2 & -0.004 & 0.344 & -0.344 & 0.366 \\
-Outcome Robust & -0.040 & 0.654 & -1.950 & 0.026 \\
-Response Robust & -0.004 & 0.301 & -0.461 & 0.322 \\
-Double Robust & -0.040 & 0.654 & -1.950 & 0.026 \\
-\end{longtable}
-
-\hypertarget{discussion}{%
-\subsection*{Discussion}\label{discussion}}
-\addcontentsline{toc}{subsection}{Discussion}
-
-There has been concern about why the constrained results are the same in
-Table~\ref{tbl-sim1}. This section aims to address these concerns.
-
-Interestingly enough, the optimization algorithm does not have a lot of
-variability in the output. The variance for each coefficient is found in
-Table~\ref{tbl-estvar}. Since these values are all basically zero, we
-can see that for each algorithm we are getting basically the same
-coefficients each iteration. The coefficients that we see are displayed
-in Table~\ref{tbl-estc}.
-
-\hypertarget{tbl-estvar}{}
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1562}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}@{}}
-\caption{\label{tbl-estvar}This table displays the variance of each
-coefficient for each estimator. A number close to zero means that the
-optimization algorithm chose very similar values for each Monte Carlo
-simulation. A value of exactly zero indicated that the optimization
-algorithm chose the exact same number.}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Estimator
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_1\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_2\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_3\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_4\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_5\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_6\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_7\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_8\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_9\)
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Estimator
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_1\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_2\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_3\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_4\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_5\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_6\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_7\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_8\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_9\)
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Para. 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-Para. 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-Outcome & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-Response & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-Double & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-\end{longtable}
-
-\hypertarget{tbl-estc}{}
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1449}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0870}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1014}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0870}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1014}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0870}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1014}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1014}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1014}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0870}}@{}}
-\caption{\label{tbl-estc}This table displays the estimated coefficients
-for each algorithm on iteration 1. Since the variance between iterations
-is small, these are reasonable for subsequent
-iterations.}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Estimator
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_1\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_2\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_3\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_4\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_5\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_6\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_7\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_8\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_9\)
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Estimator
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_1\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_2\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_3\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_4\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_5\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_6\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_7\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_8\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_9\)
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Para. 1 & 0.3 & 0.400 & 0.000 & 0.100 & 0.000 & 0.200 & 0.000 & 0.000 &
-0 \\
-Para. 2 & 0.3 & 0.400 & 0.000 & 0.100 & 0.000 & 0.200 & 0.000 & -1.000 &
-1 \\
-Outcome & 0.3 & -0.071 & 0.471 & -0.194 & 0.294 & -0.035 & -0.471 &
--0.294 & 1 \\
-Response & 0.3 & 0.400 & 0.000 & 0.100 & 0.000 & 0.200 & 0.000 & 0.000 &
-0 \\
-Double & 0.3 & -0.071 & 0.471 & -0.194 & 0.294 & -0.035 & -0.471 &
--0.294 & 1 \\
-\end{longtable}
-
-\hypertarget{tbl-fstat}{}
-\begin{longtable}[]{@{}llrr@{}}
-\caption{\label{tbl-fstat}This table computes the F-statistic and
-P-value for an F-test comparing the \texttt{Full\ Model} with a nested
-model in \texttt{Reduced\ Model}. The computation of the F-statistic is
-found in the \texttt{get\_f\_statistic} function.}\tabularnewline
-\toprule\noalign{}
-Full Model & Reduced Model & Fstat & P-Value \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-Full Model & Reduced Model & Fstat & P-Value \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Parametric 1 & Parametric 2 & 0.00 & 1 \\
-Parametric 1 & Outcome Robust & 1190.30 & 0 \\
-Parametric 1 & Response Robust & 0.00 & 1 \\
-Parametric 1 & Double Robust & 510.13 & 0 \\
-Parametric 2 & Outcome Robust & 1787.26 & 0 \\
-Parametric 2 & Response Robust & 0.00 & 1 \\
-Parametric 2 & Double Robust & 595.75 & 0 \\
-Outcome Robust & Double Robust & 0.00 & 1 \\
-Response Robust & Double Robust & 895.43 & 0 \\
-\end{longtable}
-
-From the analysis in Table~\ref{tbl-fstat}, we can see that there are
-significant differences between several of the models. Why is this
-different from the previous analysis in Table~\ref{tbl-sim1}? The answer
-comes from how we compute the variance versus the sum of squares of
-error. The variance is computed from each estimator created in a Monte
-Carlo trial; whereas the sum of squares of error is summed from each
-individuals estimate for the mean of \(Y_2\). For more details about how
-the F-test is computed see Appendix A and Appendix B.
-
-\newpage{}
-
-\hypertarget{appendix-a-computing-the-f-test}{%
-\section{Appendix A: Computing the
-F-Test}\label{appendix-a-computing-the-f-test}}
-
-To compute an F test we can use Equation~\ref{eq-fstat}.
-
-\begin{equation}\protect\hypertarget{eq-fstat}{}{ F = \frac{(SSE_{Reduced} - SSE_{Full}) / (DFE_{Reduced} - DFE_{Full})}{
-SSE_{Full} / DFE_{Full}}. }\label{eq-fstat}\end{equation}
-
-Let's assess what this would mean to compare models
-\texttt{Parametric\ 1} and \texttt{Outcome\ Robust} from
-Table~\ref{tbl-mods}.
-
-\hypertarget{tbl-mods}{}
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2941}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7059}}@{}}
-\caption{\label{tbl-mods}This table identifies the different constraints
-for each model type.}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Type
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Constraints
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Type
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Constraints
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Parametric 1 & \(\sum_{k, t} c_{kt} = 1\) \\
-Parametric 2 &
-\(\sum_{k, t: (k, t) \neq (4, 4)} c_{kt} = 0, c_{44} = 1\) \\
-Outcome Robust &
-\(c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, \text{ and } c_{44} = 1.\) \\
-Response Robust &
-\(c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}\) \\
-Double Robust &
-\(c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, c_{44} = 1, c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}\) \\
-\end{longtable}
-
-Consider the case where we run estimate each model on a data set with
-\(n = 1000\) observations, and define the following notation: let
-\(\hat \theta^{(P)}\) and \(\hat \theta^{(OR)}\) be the estimated values
-of \(\theta = E[Y_2]\) for the \texttt{Parametric\ 1} model and the
-\texttt{Outcome\ Robust} model respectively. Define the estimated
-coefficients to be \(\hat c_j^{(P)}\) and \(\hat c_j^{(OR)}\) where
-\(j = 1, \dots 9\) for the \texttt{Parametric\ 1} and
-\texttt{Outcome\ Robust} models respectively. The one can compute the
-SSE with the following:
-
-\[ SSE = n^{-1} \sum_{i = 1}^n (\hat y_{2i} - \theta)^2 \text{ where }
-\hat y_{2i} = \sum_{j = 1}^9 \hat c_j \hat \gamma_j \]
-
-and
-\(\hat \gamma_0 := \hat \gamma_{00} = \frac{\delta_{00i}}{\pi_{00}} E[Y_2 \mid x_i]\),
-\(\hat \gamma_1 := \hat \gamma_{11} = \frac{\delta_{10i}}{\pi_{10}} E[Y_2 \mid x_i, y_{1i}]\),
-etc. Note that this is different from the previous estimate of the
-variance which used the Monte Carlo variance (standard deviation)
-defined by
-\[ \frac{1}{n - 1} \sum_{b = 1}^B (\hat \theta_b - \bar \theta_B)^2 \]
-where \(B\) is the number of Monte Carlo estimates and
-\(\bar \theta_B = \frac{1}{B} \sum_{b = 1}^B \hat \theta_b\).
-
-Likewise, one can compute the degrees of freedom by noticing that each
-model has a degrees of freedom equal to nine minus the number of
-constraints. This means that we can compute the model degrees of freedom
-and error degrees of freedom for each model type, which we do in
-Table~\ref{tbl-df}
-
-\hypertarget{tbl-df}{}
-\begin{longtable}[]{@{}lll@{}}
-\caption{\label{tbl-df}This table displays the degrees of freedom for
-each model}\tabularnewline
-\toprule\noalign{}
-Model & Model Degrees of Freedom & Error Degrees of Freedom \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-Model & Model Degrees of Freedom & Error Degrees of Freedom \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Parametric 1 & \(9 - 1 = 8\) & \(n - 1 - 8 = n - 9\) \\
-Parametric 2 & \(9 - 1 = 8\) & \(n - 1 - 8 = n - 9\) \\
-Outcome Robust & \(9 - 4 = 5\) & \(n - 1 - 5 = n - 6\) \\
-Response Robust & \(9 - 4 = 5\) & \(n - 1 - 5 = n - 6\) \\
-Double Robust & \(9 - 8 = 1\) & \(n - 1 - 1 = n - 2\) \\
-\end{longtable}
-
-So continuing our first example\footnote{SSE values might differ in
-  Table~\ref{tbl-fstat} slightly depending on the actual data values.},
-if \(SSE^{(P)} = 993\) and \(SSE^{(OR)} = 3428\) with \(n = 1000\) then
-the F statistic is
-
-\[F = \frac{(3428 - 993) / (3)}{993 / (1000 - 9)} = 810.\]
-
-The critical value we want to compare this with is the \(0.95\) quantile
-of \(F_{3, 993}\) which is \(2.61\). So there \emph{is} a significant
-difference between the fit of these two model. (The p-value is basically
-zero.) The code to do this automatically is in Appendix B.
-
-\newpage{}
-
-\hypertarget{appendix-b-r-functions}{%
-\section{Appendix B: R Functions}\label{appendix-b-r-functions}}
-
-\hypertarget{lst-getfstat}{%
-\label{lst-getfstat}}%
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\#\textquotesingle{} get\_f\_statistic is a function to compute the F{-}statistic between two of }
-\CommentTok{\#\textquotesingle{} the constrained models.}
-\CommentTok{\#\textquotesingle{}}
-\CommentTok{\#\textquotesingle{} @param c\_tab {-} A table with values for the estimated c coefficients}
-\CommentTok{\#\textquotesingle{} @param reduced\_mod {-} A string indicating which model is the reduced model}
-\CommentTok{\#\textquotesingle{} @param full\_mod {-} A string indicating which model is the full model}
-\CommentTok{\#\textquotesingle{} @param df {-} A data set simulated from the same data generating process that }
-\CommentTok{\#\textquotesingle{}    the c\_tab coefficients were estimated from.}
-\CommentTok{\#\textquotesingle{}}
-\CommentTok{\#\textquotesingle{} @note We have previously shown that the variance of the estimated c\_tab }
-\CommentTok{\#\textquotesingle{} variables is close to or exactly zero. Hence, conducting a new simulation }
-\CommentTok{\#\textquotesingle{} should not effect the results of the model much.}
-\NormalTok{get\_f\_statistic }\OtherTok{\textless{}{-}} 
-  \ControlFlowTok{function}\NormalTok{(c\_tab, reduced\_mod, full\_mod, df, }\AttributeTok{theta =}\NormalTok{ true\_theta, cov\_e1e2) \{}
-
-\NormalTok{  vars\_lst }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{(reduced\_mod, full\_mod)}
-
-  \CommentTok{\# SSE reduced }
-\NormalTok{  red\_cvec }\OtherTok{\textless{}{-}}
-\NormalTok{    c\_tab }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{filter}\NormalTok{(iter }\SpecialCharTok{==} \DecValTok{1}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{filter}\NormalTok{(.data[[}\StringTok{"estimator"}\NormalTok{]] }\SpecialCharTok{==}\NormalTok{ vars\_lst[[}\DecValTok{1}\NormalTok{]]) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{select}\NormalTok{(}\FunctionTok{starts\_with}\NormalTok{(}\StringTok{"c\_"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{as.numeric}\NormalTok{()}
-
-  \CommentTok{\# SSE full}
-\NormalTok{  full\_cvec }\OtherTok{\textless{}{-}}
-\NormalTok{    c\_tab }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{filter}\NormalTok{(iter }\SpecialCharTok{==} \DecValTok{1}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{filter}\NormalTok{(.data[[}\StringTok{"estimator"}\NormalTok{]] }\SpecialCharTok{==}\NormalTok{ vars\_lst[[}\DecValTok{2}\NormalTok{]]) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{select}\NormalTok{(}\FunctionTok{starts\_with}\NormalTok{(}\StringTok{"c\_"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{as.numeric}\NormalTok{()}
-
-  \CommentTok{\# Constructing gam\_vec}
-\NormalTok{  Eg }\OtherTok{\textless{}{-}}\NormalTok{ theta}
-\NormalTok{  Egx }\OtherTok{\textless{}{-}}\NormalTok{ theta }\SpecialCharTok{+}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{X }
-\NormalTok{  Egxy1 }\OtherTok{\textless{}{-}}\NormalTok{ theta }\SpecialCharTok{+}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{X }\SpecialCharTok{+}\NormalTok{ cov\_e1e2 }\SpecialCharTok{*}\NormalTok{ (df}\SpecialCharTok{$}\NormalTok{Y1 }\SpecialCharTok{{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{X)}
-\NormalTok{  Egxy2 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{Y2}
-\NormalTok{  EEg2 }\OtherTok{\textless{}{-}}\NormalTok{ theta}\SpecialCharTok{\^{}}\DecValTok{2} \SpecialCharTok{+} \DecValTok{2}
-\NormalTok{  EEgx2 }\OtherTok{\textless{}{-}}\NormalTok{ theta}\SpecialCharTok{\^{}}\DecValTok{2} \SpecialCharTok{+} \DecValTok{1}
-\NormalTok{  EEgxy12 }\OtherTok{\textless{}{-}}\NormalTok{ theta}\SpecialCharTok{\^{}}\DecValTok{2} \SpecialCharTok{+} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ cov\_e1e2}\SpecialCharTok{\^{}}\DecValTok{2}
-\NormalTok{  EEgxy22 }\OtherTok{\textless{}{-}}\NormalTok{ theta}\SpecialCharTok{\^{}}\DecValTok{2} \SpecialCharTok{+} \DecValTok{2}
-\NormalTok{  EEgxy1Egxy2 }\OtherTok{\textless{}{-}}\NormalTok{ theta}\SpecialCharTok{\^{}}\DecValTok{2} \SpecialCharTok{+} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ cov\_e1e2}\SpecialCharTok{\^{}}\DecValTok{2}
-
-\NormalTok{  g\_11 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_00 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_00 }\SpecialCharTok{*}\NormalTok{ (Egx)}
-\NormalTok{  g\_21 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_10 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_10 }\SpecialCharTok{*}\NormalTok{ (Egx)}
-\NormalTok{  g\_22 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_10 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_10 }\SpecialCharTok{*}\NormalTok{ (Egxy1)}
-\NormalTok{  g\_31 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_01 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_01 }\SpecialCharTok{*}\NormalTok{ (Egx)}
-\NormalTok{  g\_33 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_01 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_01 }\SpecialCharTok{*}\NormalTok{ (Egxy2)}
-\NormalTok{  g\_41 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_11 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_11 }\SpecialCharTok{*}\NormalTok{ (Egx)}
-\NormalTok{  g\_42 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_11 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_11 }\SpecialCharTok{*}\NormalTok{ (Egxy1)}
-\NormalTok{  g\_43 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_11 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_11 }\SpecialCharTok{*}\NormalTok{ (Egxy2)}
-\NormalTok{  g\_44 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_11 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_11 }\SpecialCharTok{*}\NormalTok{ (df}\SpecialCharTok{$}\NormalTok{Y2)}
-
-\NormalTok{  gam\_mat }\OtherTok{\textless{}{-}}\NormalTok{ base}\SpecialCharTok{::}\FunctionTok{cbind}\NormalTok{(g\_11, g\_21, g\_22, g\_31, g\_33, g\_41, g\_42, g\_43, g\_44)}
-
-  \CommentTok{\# SSE}
-\NormalTok{  sse\_red }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{((}\FunctionTok{as.numeric}\NormalTok{(gam\_mat }\SpecialCharTok{\%*\%} \FunctionTok{t}\NormalTok{(}\FunctionTok{t}\NormalTok{(red\_cvec))) }\SpecialCharTok{{-}}\NormalTok{ theta)}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}
-\NormalTok{  sse\_full }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{((}\FunctionTok{as.numeric}\NormalTok{(gam\_mat }\SpecialCharTok{\%*\%} \FunctionTok{t}\NormalTok{(}\FunctionTok{t}\NormalTok{(full\_cvec))) }\SpecialCharTok{{-}}\NormalTok{ theta)}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}
-
-  \CommentTok{\# Error Degrees of Freedom}
-\NormalTok{  df\_vec }\OtherTok{\textless{}{-}} \FunctionTok{nrow}\NormalTok{(df) }\SpecialCharTok{{-}} \DecValTok{1} \SpecialCharTok{{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{9} \SpecialCharTok{{-}} \DecValTok{1}\NormalTok{, }\DecValTok{9} \SpecialCharTok{{-}} \DecValTok{2}\NormalTok{, }\DecValTok{9} \SpecialCharTok{{-}} \DecValTok{4}\NormalTok{, }\DecValTok{9} \SpecialCharTok{{-}} \DecValTok{4}\NormalTok{, }\DecValTok{9} \SpecialCharTok{{-}} \DecValTok{8}\NormalTok{)}
-  \FunctionTok{names}\NormalTok{(df\_vec) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"parac"}\NormalTok{, }\StringTok{"zeroc"}\NormalTok{, }\StringTok{"outc"}\NormalTok{, }\StringTok{"respc"}\NormalTok{, }\StringTok{"doubc"}\NormalTok{)}
-
-\NormalTok{  df\_red }\OtherTok{\textless{}{-}}\NormalTok{ df\_vec[reduced\_mod]}
-\NormalTok{  df\_full }\OtherTok{\textless{}{-}}\NormalTok{ df\_vec[full\_mod]}
-
-\NormalTok{  fstat }\OtherTok{\textless{}{-}}\NormalTok{ ((sse\_red }\SpecialCharTok{{-}}\NormalTok{ sse\_full) }\SpecialCharTok{/}\NormalTok{ (df\_red }\SpecialCharTok{{-}}\NormalTok{ df\_full)) }\SpecialCharTok{/}\NormalTok{ (sse\_full }\SpecialCharTok{/}\NormalTok{ df\_full)}
-
-\NormalTok{  pval }\OtherTok{\textless{}{-}} \DecValTok{1} \SpecialCharTok{{-}} \FunctionTok{pf}\NormalTok{(fstat, df\_red }\SpecialCharTok{{-}}\NormalTok{ df\_full, df\_full)}
-  \FunctionTok{return}\NormalTok{(}\FunctionTok{list}\NormalTok{(}\AttributeTok{f =}\NormalTok{ fstat, }\AttributeTok{p =}\NormalTok{ pval))}
-
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
+When estimating GLS, we assume that the covariance matrix \(V\) is fully
+known. What do we do if we have to estimate \(V\)? In this case, we need
+to use (finite-dimensional) semiparametric theory to estimate the
+optimal \(W\) such that \(W\) belongs to the orthogonal nuisance tangent
+space of \(\hat V\). This is something that we could explore further if
+we want to.
 

### 2024-02-02 Notes/sim_no00.qmd
Fixed gls_sim so that the covariance matrix reflects the SRS design. Tested to see if removing A_00 made estimator worse in sim_no00. It did.

---
title: "Non-Monotone GLS Simulation: No $A_{00}$"
author:
  - name: Caleb Leedy
date: today
date-format: "D MMMM YYYY"
format:
  pdf:
    include-in-header: 
      - "latex_header.tex"
    keep-tex: true
---

```{r}
#| label: global variables
#| echo: false
B <- 1000

obs_seg <- 250
true_mu <- 5 
cov_e1e2 <- 0.5

```

# Introduction

The goal of this report is to conduct a simulation study to show the validity of
using a GLS estimator when intermediate models do not also estimate the
parameter in question. This setup is different from the previous setup is a
couple of ways:

1. We do not use $E[g \mid G_r(Z)]$ as the $g$-functions. Instead, we have $g = (g_1,
   g_2, g_3)' = (X, Y_1, Y_2)'$.
2. We have fewer comparison estimators. Since we changed the $g$-functions it
   makes less sense to compare then to other estimators that are using different
   intermediate estimators.
3. We use a simple random sample (SRS) instead of a Poisson sample. For this
   setup, we have each segment being totally independent of each other. Each
   segment also has a fixed sample size of `r obs_seg` instead of having
   segments with random sample sizes with a total observation count of 1000.
4. The GLS estimator is now only estimating $\theta = E[Y_2] = E[g_3]$.

# Notation and Setup

Let $Z = (X, Y_1, Y_2)'$. We want to estimate the parameter $\theta = E[Y_2]$
where we may not always observe $Y_1$ and
$Y_2$. Define segments that contain observations in which the same variables are
observed as in @tbl-vars1.

| Segment  | Variables Observed |
| -------  | ------------------ |
| $A_{00}$ | $X$                |
| $A_{10}$ | $X, Y_1$           |
| $A_{01}$ | $X, Y_2$           |
| $A_{11}$ | $X, Y_1, Y_2$      | 
: This table identifies which variables are observed in each segment. Since $X$
is always observed, the subscript for each segment identifies which of variables
$Y_1$ and $Y_2$ are in the segment based on the position of a 1. {#tbl-vars1}

Let $\delta_{i, j_1, j_2}$ be the sample inclusion indicator for observation $i$
in segment $A_{j_1, j_2}$, and let $\pi_{j_1, j_2}$ be the probability of
selecting an element into $A_{j_1, j_2}$.

We consider the vector $g(Z) = (g_1(X), g_2(Y_1), g_3(Y_2))'$ and 
*for this simulation setup*, let $g_1$, $g_2$, and $g_3$ all be the identity
function $I()$. This means that $g(Z) = (X, Y_1, Y_2)'$. Notice, that we
have $\theta = E[Y_2] = E[g_3]$. In each segment $A_{j_1, j_2}$ we can obtain
estimators of some of the $g$ elements. We have the following:

\begin{align*}
g_1^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_1(x_i) \\
g_2^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_2(y_{1i}) \\
g_3^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_3(y_{2i}) \\
g_1^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} g_1(x_i) \\
g_2^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} g_2(y_{1i}) \\
g_1^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} g_1(x_i) \\
g_3^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} g_2(y_{2i}) \\
g_1^{(00)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{00}}{\pi_{00}} g_1(x_i) \\
\end{align*}

This yields the following linear estimator,

$$ \hat g = Zg + e $$

where

$$\hat g = 
\begin{bmatrix}
g_1^{(11)} \\
g_2^{(11)} \\
g_3^{(11)} \\
g_1^{(10)} \\
g_2^{(10)} \\
g_1^{(01)} \\
g_3^{(01)} \\
g_1^{(00)} \\
\end{bmatrix},
Z = 
\begin{bmatrix}
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 0 & 1 \\
1 & 0 & 0 \\
0 & 1 & 0 \\
1 & 0 & 0 \\
0 & 0 & 1 \\
1 & 0 & 0 \\
\end{bmatrix},
E[e] = 0,
\text{ and } 
\Var(e) = n^{-1}
\begin{bmatrix}
V_{11} & 0 & 0 & 0 \\
0 & V_{10} & 0 & 0 \\
0 & 0 & V_{01} & 0 \\
0 & 0 & 0 & V_{00} \\
\end{bmatrix}.
$$

Here, we also have

$$
V_{11} = 
\begin{bmatrix}
\frac{1}{\pi_{11}} E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{11}} E[g_1g_2] -
E[g_1]E[g_2] & \frac{1}{\pi_{11}} E[g_1g_3] - E[g_1]E[g_3] \\
\frac{1}{\pi_{11}} E[g_1g_2] - E[g_1]E[g_2] & \frac{1}{\pi_{11}} E[g_2^2] -
E[g_2]^2 & \frac{1}{\pi_{11}} E[g_2g_3] - E[g_2]E[g_3] \\
\frac{1}{\pi_{11}} E[g_1g_3] - E[g_1]E[g_3] & \frac{1}{\pi_{11}} E[g_2g_3] -
E[g_2]E[g_3] & \frac{1}{\pi_{11}} E[g_3^2] - E[g_3]^2 \\
\end{bmatrix},
$$
$$
V_{10} = 
\begin{bmatrix}
\frac{1}{\pi_{10}}E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{10}}E[g_1g_2] - E[g_1]E[g_2]\\
\frac{1}{\pi_{10}}E[g_1g_2] - E[g_1]E[g_2] & \frac{1}{\pi_{10}}E[g_2^2] - E[g_2]^2 \\
\end{bmatrix},
$$
$$
V_{01} = 
\begin{bmatrix}
\frac{1}{\pi_{01}}E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{01}}E[g_1g_3] - E[g_1]E[g_3]\\
\frac{1}{\pi_{01}}E[g_1g_3] - E[g_1]E[g_3] & \frac{1}{\pi_{01}}E[g_3^2] - E[g_3]^2 \\
\end{bmatrix}, \text{ and }
V_{00} = 
\begin{bmatrix}
\frac{1}{\pi_{00}}E[g_1^2] - E[g_1]^2
\end{bmatrix}.
$$

{{< pagebreak >}}

# Simulation

We use the following simulation setup

\begin{align*}
  \begin{bmatrix} x \\ e_1 \\ e_2 \end{bmatrix} 
  &\stackrel{ind}{\sim} N\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}\right) \\
  y_1 &= x + e_1 \\
  y_2 &= \mu + x + e_2 \\
\end{align*}

This yields outcome variables $Y_1$ and $Y_2$ that are correlated both with $X$
and additionally with each other. To generate the missingness pattern, we select
`r obs_seg` observations into the four segments independently.

```{r}
#| label: setup
#| echo: false
#| warning: false
#| message: false

library(dplyr)

library(doParallel)
library(doRNG)
library(parallelly)

```

```{r}
#| label: simulation function
#| echo: false

gen_simple_data <- function(n_obs_sec, mu, rho) {

  # A_11
  x <- rnorm(n_obs_sec)
  e_mat <- MASS::mvrnorm(n_obs_sec, c(0, 0), matrix(c(1, rho, rho, 1), nrow = 2))
  e1 <- e_mat[, 1]
  e2 <- e_mat[, 2]
  y1 <- x + e1
  y2 <- mu + x + e1

  df_11 <- 
    tibble(X = x, Y1 = y1, Y2 = y2,
           delta_00 = 0, delta_10 = 0, delta_01 = 0, delta_11 = 1,
           pi_00 = 1/3, pi_10 = 1/3, pi_01 = 1/3, pi_11 = 1/3)
  # A_10
  x <- rnorm(n_obs_sec)
  e_mat <- MASS::mvrnorm(n_obs_sec, c(0, 0), matrix(c(1, rho, rho, 1), nrow = 2))
  e1 <- e_mat[, 1]
  e2 <- e_mat[, 2]
  y1 <- x + e1
  y2 <- mu + x + e1

  df_10 <- 
    tibble(X = x, Y1 = y1, Y2 = y2,
           delta_00 = 0, delta_10 = 1, delta_01 = 0, delta_11 = 0,
           pi_00 = 1/3, pi_10 = 1/3, pi_01 = 1/3, pi_11 = 1/3)
  # A_01
  x <- rnorm(n_obs_sec)
  e_mat <- MASS::mvrnorm(n_obs_sec, c(0, 0), matrix(c(1, rho, rho, 1), nrow = 2))
  e1 <- e_mat[, 1]
  e2 <- e_mat[, 2]
  y1 <- x + e1
  y2 <- mu + x + e1

  df_01 <- 
    tibble(X = x, Y1 = y1, Y2 = y2,
           delta_00 = 0, delta_10 = 0, delta_01 = 1, delta_11 = 0,
           pi_00 = 1/3, pi_10 = 1/3, pi_01 = 1/3, pi_11 = 1/3)
  # A_00
  x <- rnorm(n_obs_sec)
  e_mat <- MASS::mvrnorm(n_obs_sec, c(0, 0), matrix(c(1, rho, rho, 1), nrow = 2))
  e1 <- e_mat[, 1]
  e2 <- e_mat[, 2]
  y1 <- x + e1
  y2 <- mu + x + e1

  df_00 <- 
    tibble(X = x, Y1 = y1, Y2 = y2,
           delta_00 = 1, delta_10 = 0, delta_01 = 0, delta_11 = 0,
           pi_00 = 1/3, pi_10 = 1/3, pi_01 = 1/3, pi_11 = 1/3)

  df <- bind_rows(df_11, df_10, df_01)

  # Return
  # X, Y_1, Y_2,
  # delta_00, delta_10, delta_01, delta_11,
  # pi_00, pi_10, pi_01, pi_11
  return(df)

}

```

```{r}
#| label: gls functions
#| echo: false

gls_est <- function(df, gfun, class = "basic", mu, rho) {

  if (class == "basic") {
    z_mat <- matrix(c(1, 0, 0,
                      0, 1, 0,
                      0, 0, 1,
                      1, 0, 0,
                      0, 1, 0,
                      1, 0, 0,
                      0, 0, 1), ncol = 3, byrow = TRUE)

    g_hat <- 
      c(mutate(df, tmp = delta_11 / pi_11 * X) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * Y1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * Y2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * Y1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * X) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * Y2) |> pull(tmp) |> mean())

    v_11 <- 1 / df$pi_11[1] *
      matrix(c(1, 1, 1,
               1, 2, 1 + rho,
               1, 1 + rho, mu^2 + 2), nrow = 3, byrow = TRUE) -
      matrix(c(0, 0, 0,
               0, 0, 0,
               0, 0, mu^2), nrow = 3, byrow = TRUE)
    v_10 <- 1 / df$pi_10[1] * matrix(c(1, 1, 1, 2), nrow = 2, byrow = TRUE)
    v_01 <- 
      1 / df$pi_01[1] * matrix(c(1, 1, 1, mu^2 + 2), nrow = 2) - 
      matrix(c(0, 0, 0, mu^2), nrow = 2)

    v_mat <- matrix(rep(0, 49), ncol = 7)
    v_mat[1:3, 1:3] <- v_11
    v_mat[4:5, 4:5] <- v_10
    v_mat[6:7, 6:7] <- v_01

    v_mat <- v_mat / nrow(df)

    ghat <- MASS::ginv(t(z_mat) %*% MASS::ginv(v_mat) %*% z_mat) %*% 
      t(z_mat) %*% MASS::ginv(v_mat) %*% g_hat

    return(as.numeric(t(c(0, 0, 1) %*% ghat)))

  } else {
    stop("We have only 1 class: 'basic'.")
  }
}

```

```{r}
#| label: MCMC
#| echo: false
#| tbl-cap: |
#|   Results from simulations study with independent equally sized segments
#|   $A_{11}$, $A_{10}$, and $A_{01}$ all of size $n = 250$. In this
#|   simulation we have the true mean of $Y_2$ equal to $\mu = 5$ and the covariance
#|   between $e_1$ and $e_2$ is $\rho = 0.5$. The goal is to estimate 
#|   $E[Y_2] = \mu$. For the GLS estimation, we use the
#|   true known covariance matrix (using the true values of $\mu$ and $\rho$),
#|   and we use g-functions $g_1 = X$, $g_2 = Y_1$ and $g_3 = Y_2$.

clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
registerDoParallel(clust)

mc_theta <-
  foreach(iter = 1:B,
          .options.RNG = 1,
          .packages = c("dplyr", "stringr")) %dorng% {

    # Generate Data
    df <- gen_simple_data(obs_seg, mu = true_mu, rho = cov_e1e2)

    # Estimators
    oracle <- mean(df$Y2)
    cc <- df |>
      filter(delta_11 == 1) |>
      pull(Y2) |>
      mean()
    ipw <- df |>
      mutate(tmp = delta_11 / pi_11 * Y2) |>
      pull(tmp) |>
      mean()

    gls <- gls_est(df, gfun = "Y2", mu = true_mu, rho = cov_e1e2)

    return(tibble(Oracle = oracle,
                  CC = cc,
                  IPW = ipw,
                  GLS = gls
    ))

  } |>
  bind_rows()

stopCluster(clust)

# Analyze Results
mc_theta |>
  summarize(
    across(everything(),
    list(Bias = ~mean(.) - true_mu,
         SD = sd,
         Tstat = ~(mean(.) - true_mu) / sqrt(var(.) / B)),
    .names = "{.fn}_{.col}")) |> 
  tidyr::pivot_longer(cols = everything(),
                      names_to = c(".value", "Algorithm"),
                      names_pattern = "(.*)_(.*)") |>
  mutate(Pval = pt(-abs(Tstat), df = B)) |>
  knitr::kable(digits = 3)

```

A quick comparison between the standard deviation of the GLS estimator and the
standard deviation of the GLS estimator in `gls_sim.qmd` shows that this
standard deviation is slightly larger, which means that we do lose efficiency by
not including $A_{00}$.

### 2024-02-02 Notes/sim_no00.tex
Fixed gls_sim so that the covariance matrix reflects the SRS design. Tested to see if removing A_00 made estimator worse in sim_no00. It did.

--- 2024-01-22 Notes/semiconstraints.tex
@@
 \newcommand{\logistic}{{\text{logistic}}}
+\renewcommand{\arraystretch}{1.3}
 \KOMAoption{captions}{tableheading}
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
 \makeatletter
@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
 \@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
 \makeatother
@@
 \fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{bookmark}
+
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \hypersetup{
-  pdftitle={Understanding Semiparametric Constraints},
+  pdftitle={Non-Monotone GLS Simulation: No A\_\{00\}},
   pdfauthor={Caleb Leedy},
@@
 
-\title{Understanding Semiparametric Constraints}
+\title{Non-Monotone GLS Simulation: No \(A_{00}\)}
 \author{Caleb Leedy}
-\date{19 January 2024}
+\date{1 February 2024}
 
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, frame hidden]}{\end{tcolorbox}}\fi
 
-\hypertarget{sec-intro}{%
-\section{Introduction}\label{sec-intro}}
+\section{Introduction}\label{introduction}
 
-The goal of this report is to understand how additional constraints
-effect semiparametric models. Inspired by Dr.~Fuller's note, I set out
-to see if we could assess the amount of efficiency loss due to using
-semiparametric inference. First, let's define some notation.
+The goal of this report is to conduct a simulation study to show the
+validity of using a GLS estimator when intermediate models do not also
+estimate the parameter in question. This setup is different from the
+previous setup is a couple of ways:
 
-\hypertarget{tbl-reftab}{}
-\begin{longtable}[]{@{}llll@{}}
-\caption{\label{tbl-reftab}This table shows the relationship between
-different quantities of interest in each segment, their estimators, and
-the corresponding coefficients of their estimators.}\tabularnewline
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  We do not use \(E[g \mid G_r(Z)]\) as the \(g\)-functions. Instead, we
+  have \(g = (g_1,
+  g_2, g_3)' = (X, Y_1, Y_2)'\).
+\item
+  We have fewer comparison estimators. Since we changed the
+  \(g\)-functions it makes less sense to compare then to other
+  estimators that are using different intermediate estimators.
+\item
+  We use a simple random sample (SRS) instead of a Poisson sample. For
+  this setup, we have each segment being totally independent of each
+  other. Each segment also has a fixed sample size of 250 instead of
+  having segments with random sample sizes with a total observation
+  count of 1000.
+\item
+  The GLS estimator is now only estimating \(\theta = E[Y_2] = E[g_3]\).
+\end{enumerate}
+
+\section{Notation and Setup}\label{notation-and-setup}
+
+Let \(Z = (X, Y_1, Y_2)'\). We want to estimate the parameter
+\(\theta = E[Y_2]\) where we may not always observe \(Y_1\) and \(Y_2\).
+Define segments that contain observations in which the same variables
+are observed as in Table~\ref{tbl-vars1}.
+
+\begin{longtable}[]{@{}ll@{}}
+\caption{This table identifies which variables are observed in each
+segment. Since \(X\) is always observed, the subscript for each segment
+identifies which of variables \(Y_1\) and \(Y_2\) are in the segment
+based on the position of a 1.}\label{tbl-vars1}\tabularnewline
 \toprule\noalign{}
-Segment & Quantity of Interest & Estimator & Coefficient \\
+Segment & Variables Observed \\
 \midrule\noalign{}
@@
 \toprule\noalign{}
-Segment & Quantity of Interest & Estimator & Coefficient \\
+Segment & Variables Observed \\
 \midrule\noalign{}
@@
 \endlastfoot
-\(A_{00}\) & \(E[g \mid X]\) & \(\hat \gamma_{11}\) & \(c_{11}\) \\
-\(A_{10}\) & \(E[g \mid X]\) & \(\hat \gamma_{21}\) & \(c_{21}\) \\
-\(A_{10}\) & \(E[g \mid X, Y_1]\) & \(\hat \gamma_{22}\) & \(c_{22}\) \\
-\(A_{01}\) & \(E[g \mid X]\) & \(\hat \gamma_{31}\) & \(c_{31}\) \\
-\(A_{01}\) & \(E[g \mid X, Y_2]\) & \(\hat \gamma_{33}\) & \(c_{33}\) \\
-\(A_{11}\) & \(E[g \mid X]\) & \(\hat \gamma_{41}\) & \(c_{41}\) \\
-\(A_{11}\) & \(E[g \mid X, Y_1]\) & \(\hat \gamma_{42}\) & \(c_{42}\) \\
-\(A_{11}\) & \(E[g \mid X, Y_2]\) & \(\hat \gamma_{43}\) & \(c_{43}\) \\
-\(A_{11}\) & \(E[g \mid X, Y_1, Y_2]\) & \(\hat \gamma_{44}\) &
-\(c_{44}\) \\
+\(A_{00}\) & \(X\) \\
+\(A_{10}\) & \(X, Y_1\) \\
+\(A_{01}\) & \(X, Y_2\) \\
+\(A_{11}\) & \(X, Y_1, Y_2\) \\
 \end{longtable}
 
-This is a similar framework to understanding the class of estimators
-that we discussed in \verb|optest_update.pdf|, where we write the
-estimator as a weighted average of different expectation with functional
-coefficients.
+Let \(\delta_{i, j_1, j_2}\) be the sample inclusion indicator for
+observation \(i\) in segment \(A_{j_1, j_2}\), and let
+\(\pi_{j_1, j_2}\) be the probability of selecting an element into
+\(A_{j_1, j_2}\).
 
-\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
-\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
-\mid X, Y_2].\]
+We consider the vector \(g(Z) = (g_1(X), g_2(Y_1), g_3(Y_2))'\) and
+\emph{for this simulation setup}, let \(g_1\), \(g_2\), and \(g_3\) all
+be the identity function \(I()\). This means that
+\(g(Z) = (X, Y_1, Y_2)'\). Notice, that we have
+\(\theta = E[Y_2] = E[g_3]\). In each segment \(A_{j_1, j_2}\) we can
+obtain estimators of some of the \(g\) elements. We have the following:
 
-However, now we write the estimator as
+\begin{align*}
+g_1^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_1(x_i) \\
+g_2^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_2(y_{1i}) \\
+g_3^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_3(y_{2i}) \\
+g_1^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} g_1(x_i) \\
+g_2^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} g_2(y_{1i}) \\
+g_1^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} g_1(x_i) \\
+g_3^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} g_2(y_{2i}) \\
+g_1^{(00)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{00}}{\pi_{00}} g_1(x_i) \\
+\end{align*}
 
-\[\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}\]
+This yields the following linear estimator,
+
+\[ \hat g = Zg + e \]
 
 where
-\(\hat \gamma_{kt} = \frac{\delta_{ij}}{\pi_{ij}}E[g \mid G_{ij}(X, Y_1, Y_2)]\)
-and \(ij\) corresponds to the segment \(A_{ij}\) associated with
-\(\hat \gamma_{kt}\) in Table~\ref{tbl-reftab}.
 
-\hypertarget{sec-methods}{%
-\section{Methods}\label{sec-methods}}
+\[\hat g = 
+\begin{bmatrix}
+g_1^{(11)} \\
+g_2^{(11)} \\
+g_3^{(11)} \\
+g_1^{(10)} \\
+g_2^{(10)} \\
+g_1^{(01)} \\
+g_3^{(01)} \\
+g_1^{(00)} \\
+\end{bmatrix},
+Z = 
+\begin{bmatrix}
+1 & 0 & 0 \\
+0 & 1 & 0 \\
+0 & 0 & 1 \\
+1 & 0 & 0 \\
+0 & 1 & 0 \\
+1 & 0 & 0 \\
+0 & 0 & 1 \\
+1 & 0 & 0 \\
+\end{bmatrix},
+E[e] = 0,
+\text{ and } 
+\Var(e) = n^{-1}
+\begin{bmatrix}
+V_{11} & 0 & 0 & 0 \\
+0 & V_{10} & 0 & 0 \\
+0 & 0 & V_{01} & 0 \\
+0 & 0 & 0 & V_{00} \\
+\end{bmatrix}.
+\]
 
-As noted by Dr.~Fuller, we can understand parametric estimation as
-solving the following constrained optimization problems:
+Here, we also have
 
-\begin{align*}
-  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
-  \sum_{k,t} c_{kt} = 1 \\
-       &\text{or} \\
-  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
-  \sum_{(k,t): (k,t) \neq (4, 4)} c_{kt} = 0 \text{ and } c_{44} = 1. \\
-\end{align*}
+\[
+V_{11} = 
+\begin{bmatrix}
+\frac{1}{\pi_{11}} E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{11}} E[g_1g_2] -
+E[g_1]E[g_2] & \frac{1}{\pi_{11}} E[g_1g_3] - E[g_1]E[g_3] \\
+\frac{1}{\pi_{11}} E[g_1g_2] - E[g_1]E[g_2] & \frac{1}{\pi_{11}} E[g_2^2] -
+E[g_2]^2 & \frac{1}{\pi_{11}} E[g_2g_3] - E[g_2]E[g_3] \\
+\frac{1}{\pi_{11}} E[g_1g_3] - E[g_1]E[g_3] & \frac{1}{\pi_{11}} E[g_2g_3] -
+E[g_2]E[g_3] & \frac{1}{\pi_{11}} E[g_3^2] - E[g_3]^2 \\
+\end{bmatrix},
+\] \[
+V_{10} = 
+\begin{bmatrix}
+\frac{1}{\pi_{10}}E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{10}}E[g_1g_2] - E[g_1]E[g_2]\\
+\frac{1}{\pi_{10}}E[g_1g_2] - E[g_1]E[g_2] & \frac{1}{\pi_{10}}E[g_2^2] - E[g_2]^2 \\
+\end{bmatrix},
+\] \[
+V_{01} = 
+\begin{bmatrix}
+\frac{1}{\pi_{01}}E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{01}}E[g_1g_3] - E[g_1]E[g_3]\\
+\frac{1}{\pi_{01}}E[g_1g_3] - E[g_1]E[g_3] & \frac{1}{\pi_{01}}E[g_3^2] - E[g_3]^2 \\
+\end{bmatrix}, \text{ and }
+V_{00} = 
+\begin{bmatrix}
+\frac{1}{\pi_{00}}E[g_1^2] - E[g_1]^2
+\end{bmatrix}.
+\]
 
-To be robust to the outcome model, we can construct a similar
-optimization problem with different constraints:
+\newpage{}
 
-\begin{align*}
-  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
-  c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0,\\
-       &\text{and } c_{44} = 1.
-\end{align*}
-
-Likewise to be robust to the response model, we can have the problem:
-
-\begin{align*}
-  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
-  c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \\
-       &\text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}.
-\end{align*}
-
-To be double robust (robust to the outcome and response model) we can
-combine the last two constraints. This is summarized in
-Table~\ref{tbl-mods}.
-
-\hypertarget{tbl-mods}{}
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2941}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7059}}@{}}
-\caption{\label{tbl-mods}This table identifies the different constraints
-for each model type.}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Type
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Constraints
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Type
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Constraints
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Parametric 1 & \(\sum_{k, t} c_{kt} = 1\) \\
-Parametric 2 &
-\(\sum_{k, t: (k, t) \neq (4, 4)} c_{kt} = 0, c_{44} = 1\) \\
-Outcome Robust &
-\(c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, \text{ and } c_{44} = 1.\) \\
-Response Robust &
-\(c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}\) \\
-Double Robust &
-\(c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, c_{44} = 1, c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}\) \\
-\end{longtable}
-
-\hypertarget{sec-simulations}{%
-\section{Simulation Study}\label{sec-simulations}}
-
-\hypertarget{simulation-1}{%
-\subsection*{Simulation 1}\label{simulation-1}}
-\addcontentsline{toc}{subsection}{Simulation 1}
+\section{Simulation}\label{simulation}
 
@@
   y_1 &= x + e_1 \\
-  y_2 &= \theta + x + e_2 \\
+  y_2 &= \mu + x + e_2 \\
 \end{align*}
 
-Furthermore, \(\pi_{11} = 0.2\) and \(\pi_{00} = 0.3, \pi_{10} = 0.4\),
-and \(\pi_{01} = 0.1\). The goal of this simulation study is to find
-\(\theta = E[Y_2]\). In other words, \(g(Z) = Y_2\). There are several
-algorithms for comparison which are defined as the following:
+This yields outcome variables \(Y_1\) and \(Y_2\) that are correlated
+both with \(X\) and additionally with each other. To generate the
+missingness pattern, we select 250 observations into the four segments
+independently.
 
-\begin{align*}
-  Oracle &= n^{-1} \sum_{i = 1}^n g(Z_i)\\
-  CC &= \frac{\sum_{i = 1}^n \delta_{11} g(Z_i)}{\sum_{i = 1}^n \delta_{11}} \\
-  IPW &= \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g(Z_i)\\
-\end{align*}
-
-Furthermore, we include four existing estimators that we have already
-proposed in the past: WLS, Prop, PropInd, and SemiDelta.
-
-The estimator WLS is a a weight least square estimator derived in the
-following manner. We now consider a normal model:
-
-\[ 
-  \begin{pmatrix}
-    x_i \\ e_{1i} \\ e_{2i}
-  \end{pmatrix} \stackrel{ind}{\sim}
-  N \left(
-  \begin{bmatrix}
-    0 \\ 0 \\ 0
-  \end{bmatrix},
-  \begin{bmatrix}
-    1 & 0 & 0 \\
-    0 & \sigma_{11} & \sigma_{12} \\ 
-    0 & \sigma_{12} & \sigma_{22}
-  \end{bmatrix},
-  \right)
-\]
-
-and define \(y_{1i} = \theta_1 + x_i + e_{1i}\) and
-\(y_{2i} = \theta_2 + x_i + e_{2i}\). Then \(b_1 = b_2 = 1\). We define
-\(\bar z_k^{(ij)}\) as the mean of \(y_k\) in segment \(A_{ij}\). This
-means that we have means \(\bar z_1^{(11)}\), \(\bar z_2^{(11)}\),
-\(\bar z_1^{(10)}\), and \(\bar z_2^{(01)}\). Let
-\(W = [\bar z_1^{(11)}, \bar z_2^{(11)}, \bar z_1^{(10)}, \bar z_2^{(01)}]'\),
-then for \(n_{ij} = |A_{ij}|\), we have
-
-\[Z - M \mu \sim N(\vec 0, V)\]
-
-where
-
-\[M = 
-  \begin{bmatrix}
-    1 & 0 \\
-    0 & 1 \\
-    1 & 0 \\
-    0 & 1 \\
-  \end{bmatrix}
-  \text{ and }
-  V = 
-  \begin{bmatrix}
-    \frac{\sigma_{11}}{n_{11}} & \frac{\sigma_{12}}{n_{11}} & 0 & 0 \\
-    \frac{\sigma_{12}}{n_{11}} & \frac{\sigma_{22}}{n_{11}} & 0 & 0 \\
-    0 & 0 & \frac{\sigma_{11}}{n_{10}} & 0 \\
-    0 & 0 & 0 & \frac{\sigma_{22}}{n_{01}} \\
-  \end{bmatrix}.
-\]
-
-Thus, the BLUE for \(\mu = [\mu_1, \mu_2]'\) is
-
-\[\hat \mu = (M' V^{-1} M)^{-1} M' V^{-1} W.\]
-
-Hence, WLS is \(\mu_2\) as \(g(X, Y_1, Y_2) = Y_2\).
-
-The remaining three estimators are drived from the following expression
-where the values for \(\beta\) are provided in Table~\ref{tbl-beta}.
-These are good estimators to compare to because Prop is the original
-proposed estimator. PropInd has the same form as Prop except the values
-for \(\beta\) is different. PropInd shares the same values of \(\beta\)
-as all of the new models with constraints. SemiDelta is useful because
-it is the best estimator in general so far.
-
-\[
-\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + 
-\beta_0(\delta, c_0)E[g(Z) \mid X] + 
-\beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + 
-\beta_2(\delta, c_2) E[g(Z)
-\mid X, Y_2].
-\]
-
-\hypertarget{tbl-beta}{}
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1698}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4151}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4151}}@{}}
-\caption{\label{tbl-beta}This table displays the values of \(\beta\) for
-different estimator types.}\tabularnewline
+\begin{longtable}[]{@{}lrrrr@{}}
+\caption{Results from simulations study with independent equally sized
+segments \(A_{11}\), \(A_{10}\), and \(A_{01}\) all of size \(n = 250\).
+In this simulation we have the true mean of \(Y_2\) equal to \(\mu = 5\)
+and the covariance between \(e_1\) and \(e_2\) is \(\rho = 0.5\). The
+goal is to estimate \(E[Y_2] = \mu\). For the GLS estimation, we use the
+true known covariance matrix (using the true values of \(\mu\) and
+\(\rho\)), and we use g-functions \(g_1 = X\), \(g_2 = Y_1\) and
+\(g_3 = Y_2\).}\tabularnewline
 \toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Estimator
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\(\beta_0(\delta, c_0)\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\(\beta_1(\delta, c_1)\)
-\end{minipage} \\
+Algorithm & Bias & SD & Tstat & Pval \\
 \midrule\noalign{}
@@
 \toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Estimator
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\(\beta_0(\delta, c_0)\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\(\beta_1(\delta, c_1)\)
-\end{minipage} \\
+Algorithm & Bias & SD & Tstat & Pval \\
 \midrule\noalign{}
@@
 \endlastfoot
-Prop &
-\(\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)\)
-&
-\(\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)\) \\
-PropInd &
-\(\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)\)
-&
-\(\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)\) \\
-SemiDelta &
-\(c_0\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{00}}{\pi_{00}}\right)\)
-&
-\(c_1\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{10}}{\pi_{10}}\right)\) \\
+Oracle & -0.001 & 0.049 & -0.818 & 0.207 \\
+CC & -0.004 & 0.089 & -1.317 & 0.094 \\
+IPW & -0.004 & 0.089 & -1.317 & 0.094 \\
+GLS & -0.002 & 0.053 & -1.490 & 0.068 \\
 \end{longtable}
 
-In the simulation results in Table~\ref{tbl-sim1}, the new results have
-the same label as the value from the \texttt{Type} column in
-Table~\ref{tbl-mods}.
-
-\hypertarget{tbl-sim1}{}
-\begin{longtable}[]{@{}lrrrr@{}}
-\caption{\label{tbl-sim1}Results from Simulation 1. True values:
-\(\theta = 5, \rho = 0.5\). The test conducted for the T-statistic and
-P-value is a two sample test to see if the estimator is unbiased. This
-simulation uses \(Y_1 = x + \varepsilon_1\),
-\(Y_2 = \theta + x + \varepsilon_2\) where \(X \sim N(0, 1)\) and
-\((\varepsilon_1, \varepsilon_2)\) come from a mean zero bivariate
-normal distribution with unit variance and covariance \(\rho\). The
-segments are unbalanced with: \(\pi_{11} = 0.2\), \(\pi_{10} = 0.4\),
-\(\pi_{01} = 0.1\), and \(\pi_{00} = 0.3\).}\tabularnewline
-\toprule\noalign{}
-Algorithm & bias & sd & tstat & pval \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-Algorithm & bias & sd & tstat & pval \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Oracle & -0.002 & 0.045 & -1.669 & 0.048 \\
-CC & 0.001 & 0.083 & 0.263 & 0.396 \\
-IPW & 0.010 & 0.357 & 0.872 & 0.192 \\
-WLS & 0.000 & 0.054 & -0.065 & 0.474 \\
-\(\hat \theta_{prop}\) & -0.002 & 0.063 & -0.768 & 0.221 \\
-\(\hat \theta_{\delta}\) & -0.001 & 0.064 & -0.342 & 0.366 \\
-Parametric 1 & -0.001 & 0.061 & -0.509 & 0.305 \\
-Parametric 2 & -0.001 & 0.061 & -0.509 & 0.305 \\
-Outcome Robust & -0.001 & 0.061 & -0.543 & 0.294 \\
-Response Robust & -0.001 & 0.061 & -0.509 & 0.305 \\
-Double Robust & -0.001 & 0.061 & -0.543 & 0.294 \\
-\end{longtable}
-
-\hypertarget{simulation-2}{%
-\subsection*{Simulation 2}\label{simulation-2}}
-\addcontentsline{toc}{subsection}{Simulation 2}
-
-Next, we used the same simulation setup except that we are interested in
-estimating \(\theta = E[g] = E[Y_1^2 Y_2]\). We do not use the WLS
-estimator because it does not work.
-
-\hypertarget{tbl-sim2}{}
-\begin{longtable}[]{@{}lrrrr@{}}
-\caption{\label{tbl-sim2}Results from Simulation 2. True values:
-\(\theta = 10, \rho = 0.5\). The test conducted for the T-statistic and
-P-value is a two sample test to see if the estimator is unbiased. This
-simulation uses the same setup as Simulation 1 with
-\(Y_1 = x + \varepsilon_1\), \(Y_2 = \theta + x + \varepsilon_2\) where
-\(X \sim N(0, 1)\) and \((\varepsilon_1, \varepsilon_2)\) come from a
-mean zero bivariate normal distribution with unit variance and
-covariance \(\rho\). The segments are unbalanced with:
-\(\pi_{11} = 0.2\), \(\pi_{10} = 0.4\), \(\pi_{01} = 0.1\), and
-\(\pi_{00} = 0.3\).}\tabularnewline
-\toprule\noalign{}
-Algorithm & bias & sd & tstat & pval \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-Algorithm & bias & sd & tstat & pval \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Oracle & -0.037 & 0.528 & -2.214 & 0.014 \\
-CC & 0.007 & 1.196 & 0.179 & 0.429 \\
-IPW & 0.023 & 1.405 & 0.520 & 0.302 \\
-\(\hat \theta_{prop}\) & -0.044 & 0.691 & -2.027 & 0.021 \\
-\(\hat \theta_{\delta}\) & -0.032 & 0.674 & -1.521 & 0.064 \\
-Parametric 1 & -0.004 & 0.301 & -0.461 & 0.322 \\
-Parametric 2 & -0.004 & 0.344 & -0.344 & 0.366 \\
-Outcome Robust & -0.040 & 0.654 & -1.950 & 0.026 \\
-Response Robust & -0.004 & 0.301 & -0.461 & 0.322 \\
-Double Robust & -0.040 & 0.654 & -1.950 & 0.026 \\
-\end{longtable}
-
-\hypertarget{discussion}{%
-\subsection*{Discussion}\label{discussion}}
-\addcontentsline{toc}{subsection}{Discussion}
-
-There has been concern about why the constrained results are the same in
-Table~\ref{tbl-sim1}. This section aims to address these concerns.
-
-Interestingly enough, the optimization algorithm does not have a lot of
-variability in the output. The variance for each coefficient is found in
-Table~\ref{tbl-estvar}. Since these values are all basically zero, we
-can see that for each algorithm we are getting basically the same
-coefficients each iteration. The coefficients that we see are displayed
-in Table~\ref{tbl-estc}.
-
-\hypertarget{tbl-estvar}{}
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1562}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0938}}@{}}
-\caption{\label{tbl-estvar}This table displays the variance of each
-coefficient for each estimator. A number close to zero means that the
-optimization algorithm chose very similar values for each Monte Carlo
-simulation. A value of exactly zero indicated that the optimization
-algorithm chose the exact same number.}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Estimator
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_1\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_2\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_3\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_4\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_5\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_6\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_7\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_8\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_9\)
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Estimator
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_1\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_2\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_3\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_4\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_5\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_6\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_7\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_8\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_9\)
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Para. 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-Para. 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-Outcome & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-Response & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-Double & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-\end{longtable}
-
-\hypertarget{tbl-estc}{}
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1449}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0870}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1014}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0870}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1014}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0870}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1014}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1014}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1014}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0870}}@{}}
-\caption{\label{tbl-estc}This table displays the estimated coefficients
-for each algorithm on iteration 1. Since the variance between iterations
-is small, these are reasonable for subsequent
-iterations.}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Estimator
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_1\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_2\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_3\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_4\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_5\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_6\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_7\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_8\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_9\)
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Estimator
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_1\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_2\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_3\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_4\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_5\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_6\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_7\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_8\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\(c_9\)
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Para. 1 & 0.3 & 0.400 & 0.000 & 0.100 & 0.000 & 0.200 & 0.000 & 0.000 &
-0 \\
-Para. 2 & 0.3 & 0.400 & 0.000 & 0.100 & 0.000 & 0.200 & 0.000 & -1.000 &
-1 \\
-Outcome & 0.3 & -0.071 & 0.471 & -0.194 & 0.294 & -0.035 & -0.471 &
--0.294 & 1 \\
-Response & 0.3 & 0.400 & 0.000 & 0.100 & 0.000 & 0.200 & 0.000 & 0.000 &
-0 \\
-Double & 0.3 & -0.071 & 0.471 & -0.194 & 0.294 & -0.035 & -0.471 &
--0.294 & 1 \\
-\end{longtable}
-
-\hypertarget{tbl-fstat}{}
-\begin{longtable}[]{@{}llrr@{}}
-\caption{\label{tbl-fstat}This table computes the F-statistic and
-P-value for an F-test comparing the \texttt{Full\ Model} with a nested
-model in \texttt{Reduced\ Model}. The computation of the F-statistic is
-found in the \texttt{get\_f\_statistic} function.}\tabularnewline
-\toprule\noalign{}
-Full Model & Reduced Model & Fstat & P-Value \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-Full Model & Reduced Model & Fstat & P-Value \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Parametric 1 & Parametric 2 & 0.00 & 1 \\
-Parametric 1 & Outcome Robust & 1190.30 & 0 \\
-Parametric 1 & Response Robust & 0.00 & 1 \\
-Parametric 1 & Double Robust & 510.13 & 0 \\
-Parametric 2 & Outcome Robust & 1787.26 & 0 \\
-Parametric 2 & Response Robust & 0.00 & 1 \\
-Parametric 2 & Double Robust & 595.75 & 0 \\
-Outcome Robust & Double Robust & 0.00 & 1 \\
-Response Robust & Double Robust & 895.43 & 0 \\
-\end{longtable}
-
-From the analysis in Table~\ref{tbl-fstat}, we can see that there are
-significant differences between several of the models. Why is this
-different from the previous analysis in Table~\ref{tbl-sim1}? The answer
-comes from how we compute the variance versus the sum of squares of
-error. The variance is computed from each estimator created in a Monte
-Carlo trial; whereas the sum of squares of error is summed from each
-individuals estimate for the mean of \(Y_2\). For more details about how
-the F-test is computed see Appendix A and Appendix B.
-
-\newpage{}
-
-\hypertarget{appendix-a-computing-the-f-test}{%
-\section{Appendix A: Computing the
-F-Test}\label{appendix-a-computing-the-f-test}}
-
-To compute an F test we can use Equation~\ref{eq-fstat}.
-
-\begin{equation}\protect\hypertarget{eq-fstat}{}{ F = \frac{(SSE_{Reduced} - SSE_{Full}) / (DFE_{Reduced} - DFE_{Full})}{
-SSE_{Full} / DFE_{Full}}. }\label{eq-fstat}\end{equation}
-
-Let's assess what this would mean to compare models
-\texttt{Parametric\ 1} and \texttt{Outcome\ Robust} from
-Table~\ref{tbl-mods}.
-
-\hypertarget{tbl-mods}{}
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2941}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7059}}@{}}
-\caption{\label{tbl-mods}This table identifies the different constraints
-for each model type.}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Type
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Constraints
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Type
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Constraints
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Parametric 1 & \(\sum_{k, t} c_{kt} = 1\) \\
-Parametric 2 &
-\(\sum_{k, t: (k, t) \neq (4, 4)} c_{kt} = 0, c_{44} = 1\) \\
-Outcome Robust &
-\(c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, \text{ and } c_{44} = 1.\) \\
-Response Robust &
-\(c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}\) \\
-Double Robust &
-\(c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, c_{44} = 1, c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}\) \\
-\end{longtable}
-
-Consider the case where we run estimate each model on a data set with
-\(n = 1000\) observations, and define the following notation: let
-\(\hat \theta^{(P)}\) and \(\hat \theta^{(OR)}\) be the estimated values
-of \(\theta = E[Y_2]\) for the \texttt{Parametric\ 1} model and the
-\texttt{Outcome\ Robust} model respectively. Define the estimated
-coefficients to be \(\hat c_j^{(P)}\) and \(\hat c_j^{(OR)}\) where
-\(j = 1, \dots 9\) for the \texttt{Parametric\ 1} and
-\texttt{Outcome\ Robust} models respectively. The one can compute the
-SSE with the following:
-
-\[ SSE = n^{-1} \sum_{i = 1}^n (\hat y_{2i} - \theta)^2 \text{ where }
-\hat y_{2i} = \sum_{j = 1}^9 \hat c_j \hat \gamma_j \]
-
-and
-\(\hat \gamma_0 := \hat \gamma_{00} = \frac{\delta_{00i}}{\pi_{00}} E[Y_2 \mid x_i]\),
-\(\hat \gamma_1 := \hat \gamma_{11} = \frac{\delta_{10i}}{\pi_{10}} E[Y_2 \mid x_i, y_{1i}]\),
-etc. Note that this is different from the previous estimate of the
-variance which used the Monte Carlo variance (standard deviation)
-defined by
-\[ \frac{1}{n - 1} \sum_{b = 1}^B (\hat \theta_b - \bar \theta_B)^2 \]
-where \(B\) is the number of Monte Carlo estimates and
-\(\bar \theta_B = \frac{1}{B} \sum_{b = 1}^B \hat \theta_b\).
-
-Likewise, one can compute the degrees of freedom by noticing that each
-model has a degrees of freedom equal to nine minus the number of
-constraints. This means that we can compute the model degrees of freedom
-and error degrees of freedom for each model type, which we do in
-Table~\ref{tbl-df}
-
-\hypertarget{tbl-df}{}
-\begin{longtable}[]{@{}lll@{}}
-\caption{\label{tbl-df}This table displays the degrees of freedom for
-each model}\tabularnewline
-\toprule\noalign{}
-Model & Model Degrees of Freedom & Error Degrees of Freedom \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-Model & Model Degrees of Freedom & Error Degrees of Freedom \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Parametric 1 & \(9 - 1 = 8\) & \(n - 1 - 8 = n - 9\) \\
-Parametric 2 & \(9 - 1 = 8\) & \(n - 1 - 8 = n - 9\) \\
-Outcome Robust & \(9 - 4 = 5\) & \(n - 1 - 5 = n - 6\) \\
-Response Robust & \(9 - 4 = 5\) & \(n - 1 - 5 = n - 6\) \\
-Double Robust & \(9 - 8 = 1\) & \(n - 1 - 1 = n - 2\) \\
-\end{longtable}
-
-So continuing our first example\footnote{SSE values might differ in
-  Table~\ref{tbl-fstat} slightly depending on the actual data values.},
-if \(SSE^{(P)} = 993\) and \(SSE^{(OR)} = 3428\) with \(n = 1000\) then
-the F statistic is
-
-\[F = \frac{(3428 - 993) / (3)}{993 / (1000 - 9)} = 810.\]
-
-The critical value we want to compare this with is the \(0.95\) quantile
-of \(F_{3, 993}\) which is \(2.61\). So there \emph{is} a significant
-difference between the fit of these two model. (The p-value is basically
-zero.) The code to do this automatically is in Appendix B.
-
-\newpage{}
-
-\hypertarget{appendix-b-r-functions}{%
-\section{Appendix B: R Functions}\label{appendix-b-r-functions}}
-
-\hypertarget{lst-getfstat}{%
-\label{lst-getfstat}}%
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\#\textquotesingle{} get\_f\_statistic is a function to compute the F{-}statistic between two of }
-\CommentTok{\#\textquotesingle{} the constrained models.}
-\CommentTok{\#\textquotesingle{}}
-\CommentTok{\#\textquotesingle{} @param c\_tab {-} A table with values for the estimated c coefficients}
-\CommentTok{\#\textquotesingle{} @param reduced\_mod {-} A string indicating which model is the reduced model}
-\CommentTok{\#\textquotesingle{} @param full\_mod {-} A string indicating which model is the full model}
-\CommentTok{\#\textquotesingle{} @param df {-} A data set simulated from the same data generating process that }
-\CommentTok{\#\textquotesingle{}    the c\_tab coefficients were estimated from.}
-\CommentTok{\#\textquotesingle{}}
-\CommentTok{\#\textquotesingle{} @note We have previously shown that the variance of the estimated c\_tab }
-\CommentTok{\#\textquotesingle{} variables is close to or exactly zero. Hence, conducting a new simulation }
-\CommentTok{\#\textquotesingle{} should not effect the results of the model much.}
-\NormalTok{get\_f\_statistic }\OtherTok{\textless{}{-}} 
-  \ControlFlowTok{function}\NormalTok{(c\_tab, reduced\_mod, full\_mod, df, }\AttributeTok{theta =}\NormalTok{ true\_theta, cov\_e1e2) \{}
-
-\NormalTok{  vars\_lst }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{(reduced\_mod, full\_mod)}
-
-  \CommentTok{\# SSE reduced }
-\NormalTok{  red\_cvec }\OtherTok{\textless{}{-}}
-\NormalTok{    c\_tab }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{filter}\NormalTok{(iter }\SpecialCharTok{==} \DecValTok{1}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{filter}\NormalTok{(.data[[}\StringTok{"estimator"}\NormalTok{]] }\SpecialCharTok{==}\NormalTok{ vars\_lst[[}\DecValTok{1}\NormalTok{]]) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{select}\NormalTok{(}\FunctionTok{starts\_with}\NormalTok{(}\StringTok{"c\_"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{as.numeric}\NormalTok{()}
-
-  \CommentTok{\# SSE full}
-\NormalTok{  full\_cvec }\OtherTok{\textless{}{-}}
-\NormalTok{    c\_tab }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{filter}\NormalTok{(iter }\SpecialCharTok{==} \DecValTok{1}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{filter}\NormalTok{(.data[[}\StringTok{"estimator"}\NormalTok{]] }\SpecialCharTok{==}\NormalTok{ vars\_lst[[}\DecValTok{2}\NormalTok{]]) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{select}\NormalTok{(}\FunctionTok{starts\_with}\NormalTok{(}\StringTok{"c\_"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}}
-    \FunctionTok{as.numeric}\NormalTok{()}
-
-  \CommentTok{\# Constructing gam\_vec}
-\NormalTok{  Eg }\OtherTok{\textless{}{-}}\NormalTok{ theta}
-\NormalTok{  Egx }\OtherTok{\textless{}{-}}\NormalTok{ theta }\SpecialCharTok{+}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{X }
-\NormalTok{  Egxy1 }\OtherTok{\textless{}{-}}\NormalTok{ theta }\SpecialCharTok{+}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{X }\SpecialCharTok{+}\NormalTok{ cov\_e1e2 }\SpecialCharTok{*}\NormalTok{ (df}\SpecialCharTok{$}\NormalTok{Y1 }\SpecialCharTok{{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{X)}
-\NormalTok{  Egxy2 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{Y2}
-\NormalTok{  EEg2 }\OtherTok{\textless{}{-}}\NormalTok{ theta}\SpecialCharTok{\^{}}\DecValTok{2} \SpecialCharTok{+} \DecValTok{2}
-\NormalTok{  EEgx2 }\OtherTok{\textless{}{-}}\NormalTok{ theta}\SpecialCharTok{\^{}}\DecValTok{2} \SpecialCharTok{+} \DecValTok{1}
-\NormalTok{  EEgxy12 }\OtherTok{\textless{}{-}}\NormalTok{ theta}\SpecialCharTok{\^{}}\DecValTok{2} \SpecialCharTok{+} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ cov\_e1e2}\SpecialCharTok{\^{}}\DecValTok{2}
-\NormalTok{  EEgxy22 }\OtherTok{\textless{}{-}}\NormalTok{ theta}\SpecialCharTok{\^{}}\DecValTok{2} \SpecialCharTok{+} \DecValTok{2}
-\NormalTok{  EEgxy1Egxy2 }\OtherTok{\textless{}{-}}\NormalTok{ theta}\SpecialCharTok{\^{}}\DecValTok{2} \SpecialCharTok{+} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ cov\_e1e2}\SpecialCharTok{\^{}}\DecValTok{2}
-
-\NormalTok{  g\_11 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_00 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_00 }\SpecialCharTok{*}\NormalTok{ (Egx)}
-\NormalTok{  g\_21 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_10 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_10 }\SpecialCharTok{*}\NormalTok{ (Egx)}
-\NormalTok{  g\_22 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_10 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_10 }\SpecialCharTok{*}\NormalTok{ (Egxy1)}
-\NormalTok{  g\_31 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_01 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_01 }\SpecialCharTok{*}\NormalTok{ (Egx)}
-\NormalTok{  g\_33 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_01 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_01 }\SpecialCharTok{*}\NormalTok{ (Egxy2)}
-\NormalTok{  g\_41 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_11 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_11 }\SpecialCharTok{*}\NormalTok{ (Egx)}
-\NormalTok{  g\_42 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_11 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_11 }\SpecialCharTok{*}\NormalTok{ (Egxy1)}
-\NormalTok{  g\_43 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_11 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_11 }\SpecialCharTok{*}\NormalTok{ (Egxy2)}
-\NormalTok{  g\_44 }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{delta\_11 }\SpecialCharTok{/}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{prob\_11 }\SpecialCharTok{*}\NormalTok{ (df}\SpecialCharTok{$}\NormalTok{Y2)}
-
-\NormalTok{  gam\_mat }\OtherTok{\textless{}{-}}\NormalTok{ base}\SpecialCharTok{::}\FunctionTok{cbind}\NormalTok{(g\_11, g\_21, g\_22, g\_31, g\_33, g\_41, g\_42, g\_43, g\_44)}
-
-  \CommentTok{\# SSE}
-\NormalTok{  sse\_red }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{((}\FunctionTok{as.numeric}\NormalTok{(gam\_mat }\SpecialCharTok{\%*\%} \FunctionTok{t}\NormalTok{(}\FunctionTok{t}\NormalTok{(red\_cvec))) }\SpecialCharTok{{-}}\NormalTok{ theta)}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}
-\NormalTok{  sse\_full }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{((}\FunctionTok{as.numeric}\NormalTok{(gam\_mat }\SpecialCharTok{\%*\%} \FunctionTok{t}\NormalTok{(}\FunctionTok{t}\NormalTok{(full\_cvec))) }\SpecialCharTok{{-}}\NormalTok{ theta)}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}
-
-  \CommentTok{\# Error Degrees of Freedom}
-\NormalTok{  df\_vec }\OtherTok{\textless{}{-}} \FunctionTok{nrow}\NormalTok{(df) }\SpecialCharTok{{-}} \DecValTok{1} \SpecialCharTok{{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{9} \SpecialCharTok{{-}} \DecValTok{1}\NormalTok{, }\DecValTok{9} \SpecialCharTok{{-}} \DecValTok{2}\NormalTok{, }\DecValTok{9} \SpecialCharTok{{-}} \DecValTok{4}\NormalTok{, }\DecValTok{9} \SpecialCharTok{{-}} \DecValTok{4}\NormalTok{, }\DecValTok{9} \SpecialCharTok{{-}} \DecValTok{8}\NormalTok{)}
-  \FunctionTok{names}\NormalTok{(df\_vec) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"parac"}\NormalTok{, }\StringTok{"zeroc"}\NormalTok{, }\StringTok{"outc"}\NormalTok{, }\StringTok{"respc"}\NormalTok{, }\StringTok{"doubc"}\NormalTok{)}
-
-\NormalTok{  df\_red }\OtherTok{\textless{}{-}}\NormalTok{ df\_vec[reduced\_mod]}
-\NormalTok{  df\_full }\OtherTok{\textless{}{-}}\NormalTok{ df\_vec[full\_mod]}
-
-\NormalTok{  fstat }\OtherTok{\textless{}{-}}\NormalTok{ ((sse\_red }\SpecialCharTok{{-}}\NormalTok{ sse\_full) }\SpecialCharTok{/}\NormalTok{ (df\_red }\SpecialCharTok{{-}}\NormalTok{ df\_full)) }\SpecialCharTok{/}\NormalTok{ (sse\_full }\SpecialCharTok{/}\NormalTok{ df\_full)}
-
-\NormalTok{  pval }\OtherTok{\textless{}{-}} \DecValTok{1} \SpecialCharTok{{-}} \FunctionTok{pf}\NormalTok{(fstat, df\_red }\SpecialCharTok{{-}}\NormalTok{ df\_full, df\_full)}
-  \FunctionTok{return}\NormalTok{(}\FunctionTok{list}\NormalTok{(}\AttributeTok{f =}\NormalTok{ fstat, }\AttributeTok{p =}\NormalTok{ pval))}
-
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
+A quick comparison between the standard deviation of the GLS estimator
+and the standard deviation of the GLS estimator in \texttt{gls\_sim.qmd}
+shows that this standard deviation is slightly larger, which means that
+we do lose efficiency by not including \(A_{00}\).
 

### 2024-02-02 final/paper/old_paper/init_sims.tex
Changed project structure.

--- 2023-04-18 Notes/nonmonotone_project.tex
@@
-\documentclass[12pt]{article}
-
-\usepackage{amsmath, amssymb, mathrsfs, fancyhdr}
-\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
-\usepackage{biblatex}
-\usepackage{booktabs}
-\usepackage{float}
-
-\addbibresource{references.bib}
-
-\hypersetup{colorlinks = true, urlcolor = black}
-
-\headheight     15pt
-\topmargin      -1.5cm   % read Lamport p.163
-\oddsidemargin  -0.04cm  % read Lamport p.163
-\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
-\textwidth      16.59cm
-\textheight     23.94cm
-\parskip         7.2pt   % sets spacing between paragraphs
-\parindent         0pt   % sets leading space for paragraphs
-\pagestyle{empty}        % Uncomment if don't want page numbers
-\pagestyle{fancyplain}
-
-\newcommand{\MAP}{{\text{MAP}}}
-\newcommand{\argmax}{{\text{argmax}}}
-\newcommand{\argmin}{{\text{argmin}}}
-\newcommand{\Cov}{{\text{Cov}}}
-\newcommand{\logistic}{{\text{logistic}}}
-
-\begin{document}
-
-%\lhead{Caleb Leedy}
-\chead{Nonmonotone Missingness}
-%\chead{STAT 615 - Advanced Bayesian Methods}
-%\rhead{Page \thepage\ of \pageref{LastPage}}
-\rhead{April 4, 2023}
-
-\section*{Overview:}
-
-The goal of this project is to outperform existing techniques in the literature
-related to nonmonotone missing data.
-
 \section*{Initial Simulations:}
@@
         Y_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
-        Y_{2i} &\stackrel{iid}{\sim} N(\theta, 1)
+        Y_{2i} &\stackrel{iid}{\sim} \theta + N(0, 1)
     \end{align*}
@@
     \begin{bmatrix}
-        0 \\ 0 \\ \theta
+        0 \\ 0 \\ 0
     \end{bmatrix},
@@
         1 & 0 & 0 \\
-        0 & 1 & \sigma_{yy}\\
-        0 & \sigma_{yy} & 1
+        0 & 1 & \rho_{yy}\\
+        0 & \rho_{yy} & 1
     \end{bmatrix}
@@
     \[y_{1i} = x_i + \varepsilon_{1i} \text{ and } 
-    y_{2i} = x_i + \varepsilon_{2i}.\]
+    y_{2i} = \theta + x_i + \varepsilon_{2i}.\]
 
@@
     We also want to simulate data that is correlated. For this simulation, we
-    focus on $\Cov(Y_1, Y_2)$. The data generating process now has $\sigma_{yy}
+    focus on $\Cov(\varepsilon_1, \varepsilon_2)$. The data generating process now has $\rho_{yy}
     \neq 0$. We are still interested in $\bar Y_2$ and we still run 2000
-    simulation with 2000 observations. In all of the next simulations the true
+    simulation with 2000 observations. In all the next simulations the true
     value of $\theta = 0$. The results are the following:
@@
 
-    Then,
+    Then, the true outcome model is,
 
     \[y_{1i} = x_i + x_i^2 \varepsilon_{1i} \text{ and } 
-    y_{2i} = -x_i + x_i^3 + \varepsilon_{2i}.\]
+    y_{2i} = \theta -x_i + x_i^3 + \varepsilon_{2i}.\]
 
@@
 
-    Thus, there is strong evidence that the proposed method is \textit{not}
-    doubly robust. If we have a misspecified response model, then it seems like
-    we induce bias. I will explore this more in the next section.
+    The previous version of this simulation used the Stage 2 probability $p_i
+    \propto \logistic(y_i)$ where $y_i$ was the observed $Y$ value in Stage 1.
+    However, under this setup, our method is biased. This is because
 
+    \[ E\left[E\left[\frac{R_1 R_2}{\pi_{11}(X, Y_1)} Y_1 \mid X\right]\right] 
+    \neq
+    E\left[\frac{E[Y_1 \mid X]}{\pi_{11}(X, Y_1)} E[R_1 R_2 \mid X]\right] \]
+
+    but
+
+    \[ E\left[E\left[\frac{R_1 R_2}{\pi_{11}(X)} Y_1 \mid X\right]\right] 
+    =
+    E\left[\frac{E[Y_1 \mid X]}{\pi_{11}(X)} E[R_1 R_2 \mid X]\right].\]
+
+    Thus, there is strong evidence that the proposed method is 
+    doubly robust because it is robust to both misspecification in the outcome
+    and response model.
 \end{itemize}
@@
 \newpage
-
-\section*{Estimating the Response Model}
-
-\begin{itemize}
-  \item In most (critically \textit{not} Simulation 4) of the simulations of the
-    previous section, we used the oracle weights when estimating our proposed
-    method. The reasoning for this was straightforward{--}we wanted to ensure
-    that the response model was correctly specified and the best case scenario
-    is to use the oracle weights which we did.
-
-  \item This section focuses more on estimating these response weights. Instead
-    of focusing on the proposed method, we will actually be working on
-    estimation of the complete case IPW estimator. This model is less complex
-    and will thus make it easier to understand where we are making mistakes.
-
-  \item We use the following simulation study:
-    \begin{align*}
-      x_i &\stackrel{iid}{\sim} N(0, 1)\\
-      \varepsilon_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
-      \varepsilon_{2i} &\stackrel{iid}{\sim} N(0, 1)\\
-      y_{1i} &= x_i + \varepsilon_{1i} \\
-      y_{2i} &= x_i + \varepsilon_{2i} \\
-    \end{align*}
-
-    To select a missingness pattern for each $i$, we have sequence: first, we
-    select the first variable to observe (or neither), then we either select the
-    second variable or we do not. In the first step, we select $R_{1i} = 1$ with
-    probability $0.4$, $R_{2i} = 1$ with probability $0.4$ and neither variable
-    with probability $0.2$. For the second step, we have the probability of
-    observing the other variable be $\logistic(x_i)$. This yields the following:
-
-    \input{Tables/ipwsim.tex}
-    % TODO: estimate IPW with \pi_{2|1} and \pi_{1|2}.
-
-    The problem with this is that our estimate is biased. For one particular
-    realization of the simulation, here is the distribution of the difference
-    between the estimated and true probabilities of $\pi_{11}$:
-
-    \includegraphics[width = 0.7\linewidth]{diffhist.png}
-
-\end{itemize}
-
-\newpage
-
-\section*{Proposal: Full Nonmonotone Estimator}
-
-\begin{itemize}
-
-  \item When constructing
-    estimates of the response model, we estimate $\pi_{11}$ using
-    $\pi_{A_2 = 1|A_1 = 1}$ and we never include information about $\pi_{A_1 =
-    1|A_2 = 1}$. I would like to create a full estimator where we are able to
-    use this information.
-
-  \item The current estimator is the following (See Slide 24 of Non-monotone
-    Presentation):
-
-    \begin{align*}
-      \hat \theta_{eff} &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \\
-      &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)} (b_2(X_i, Y_{1i}) - 
-      E[g_i \mid X_i]) \\
-      &+ n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)} (a_2(X_i, Y_{2i}) -
-      E[g_i \mid X_i]) \\
-      &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_1, Y_{1i})}(g_i 
-      - b_2(X_i, Y_{1i}) - a_2(X_i, Y_{2i}) + E[g_i \mid X_i])
-    \end{align*}
-
-  \item The problem with this is that we assume $\pi_{11}(X, Y_{1}) =
-    \pi_{2|1} (X, Y_{1}) \pi_{1+}(X)$ when we could have $\pi_{11}(X, Y_2) =
-    \pi_{1|2} (X, Y_2) \pi_{2+}(X)$. In other words, the previous result
-    implicitly assumes that $\pi_{11}$ is a function of $X$ and $Y_1$ when it
-    could be a function of $X$ and $Y_2$ as well.
-
-  \item I think that we could use the
-    following (small modification) instead,
-
-    \[\pi_{11}(X, Y_1, Y_2) = \alpha \pi_{1|2} \pi_{2+} + (1 - \alpha) \pi_{2|1}
-    \pi_{1+}\]
-
-    for some $\alpha \in [0, 1]$.
-
-  \item The reasoning behind this is two-fold. First, it allows us to think of
-    the nonmonotone missingness case as a linear combination of two monotone
-    cases, which I think is useful. Second, we can make explicit our choice of
-    $\alpha$, which in the existing estimator is simply $\alpha = 0$.
-
-  \item The problem with adding another parameter $\alpha$ is that it makes the
-    overall model unidentifiable. We cannot estimate $\alpha$ and the
-    conditional and marginal distributions that we have without additional
-    assumptions. So for now, I think that we should just assume that $\alpha$ is
-    known and then we apply this method to a data integration problem we can
-    review what $\alpha$ makes sense or figure out an additional assumption that
-    can help us estimate $\alpha$ (for example if there is a variable correlated
-    with $1|2$ versus $2|1$).
-
-  \item This is not just a pure Bayes' rule.
-    Formally, let $L = (X, Y_1, Y_2)$. By Bayes' rule we have
-
-    \begin{align*}
-      \Pr(R_1 &= 1 \mid R_2 = 1, L) \Pr(R_2 = 1 \mid L) \\
-      &= \Pr(R_1 = 1, R_2 = 1 \mid L) \\
-      & = \Pr(R_2 = 1 \mid R_1 = 1, L) \Pr(R_1 = 1 \mid L).
-    \end{align*}
-
-    Yet to estimate this model, we need to assume something like the following:
-    \begin{align*}
-      \Pr(R_1 &= 1 \mid R_2 = 1, X, Y_2) \Pr(R_2 = 1 \mid X, Y_2)\\
-      &= \Pr(R_1 = 1, R_2 = 1 \mid L) \\
-      &= \Pr(R_2 = 1 \mid R_1 = 1, X, Y_1) \Pr(R_1 = 1 \mid X, Y_1)
-    \end{align*}
-
-    in which case the two sides are clearly different.
-
-    \item Unfortunately, early simulation results are not promising because we
-      cannot distinguish points that should have $\pi_{11}$ conditional on $Y_1$
-      and points that should be conditional on $Y_2$.
-
-      \input{Tables/alphasim.tex}
-
-    \item Dr. Kim, please let me know what you think about this idea and if it
-      is worth pursuing more.
-
-\end{itemize}
-
-\newpage
-
-\printbibliography
-
-\end{document}

### 2024-02-02 final/paper/old_paper/missing_mechanism.tex
Changed project structure.

--- 2023-04-20 Notes/nonmonotone_project.tex
@@
-\documentclass[12pt]{article}
-
-\usepackage{amsmath, amssymb, mathrsfs, fancyhdr}
-\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
-\usepackage{biblatex}
-\usepackage{booktabs}
-\usepackage{float}
-
-\addbibresource{references.bib}
-
-\hypersetup{colorlinks = true, urlcolor = black}
-
-\headheight     15pt
-\topmargin      -1.5cm   % read Lamport p.163
-\oddsidemargin  -0.04cm  % read Lamport p.163
-\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
-\textwidth      16.59cm
-\textheight     23.94cm
-\parskip         7.2pt   % sets spacing between paragraphs
-\parindent         0pt   % sets leading space for paragraphs
-\pagestyle{empty}        % Uncomment if don't want page numbers
-\pagestyle{fancyplain}
-
-\newcommand{\MAP}{{\text{MAP}}}
-\newcommand{\argmax}{{\text{argmax}}}
-\newcommand{\argmin}{{\text{argmin}}}
-\newcommand{\Cov}{{\text{Cov}}}
-\newcommand{\logistic}{{\text{logistic}}}
-
-\begin{document}
-
-%\lhead{Caleb Leedy}
-\chead{Nonmonotone Missingness}
-%\chead{STAT 615 - Advanced Bayesian Methods}
-%\rhead{Page \thepage\ of \pageref{LastPage}}
-\rhead{April 4, 2023}
-
-\section*{Overview:}
-
-The goal of this project is to outperform existing techniques in the literature
-related to nonmonotone missing data.
-
-\section*{Initial Simulations:}
-% Tables generated from Simulations/nonmonotone.R
-
-\begin{itemize}
-  \item \textbf{Implemented simulation of monotone MAR data}: This is 
-  correspondingly easier
-  than the subsequent nonmonotone MAR simulation. For this simulation we use the 
-  following approach:
-  \begin{enumerate}
-      \item Generate $X$, $Y_1$, and $Y_2$ for elements $i = 1, \dots, n$.
-      \item Using the covariate $X$, determine the probability $p_1$ of $Y_1$
-      being observed for each element $i$.
-      \item Based on $p_1$, determine if $R_1 = 1$.
-      \item If $R_1 = 0$, then $R_2 = 0$. Otherwise, using variables $X$ and
-        $Y_1$, determine the probability $p_{12}$.
-      \item Based on $p_{12}$ determine if $R_2 = 1$.
-  \end{enumerate}
-  At the end of the algorithm, we have determined the values of binary variables
-    $R_1$ and $R_2$ for each $i$ and if either of them are equal to 1, the
-    corresponding level of $Y_{k}$. As is common in this literature, the values
-    of $R_1$ and $R_2$ determine if the corresponding variable $Y_1$ or $Y_2$ is
-    missing or observed with $R = 1$ indicating $Y$ being observed.
-
-  \item \textbf{Implemented simulation of nonmonotone MAR data}: 
-  Following the approach
-    of \cite{robins1997non}, I construct a nonmonotone MAR simulation with two 
-    response variables $Y_1$ and $Y_2$ and one covariate $X$. The algorithm to
-    generate the data is the following:
-    \begin{enumerate}
-        \item Generate $X$, $Y_1$, and $Y_2$ for elements $i = 1, \dots, n$.
-        \item Using the covariate $X_i$, generate probabilities for each element
-          $i$ $p_0$, $p_1$, and $p_2$ such that $p_0 + p_1 + p_2 = 1$. 
-        \item Select one option based on the three probabilities for each
-          element $i$. If 0 is selected: $R_1 = 0$ and $R_2 = 0$. If 1 is
-          selected $R_1 = 1$. If 2 is selected, $R_2 = 1$.
-        \item We take the next step in multiple cases. If 0 was selected, we are
-          done. If 1 was selected, we generate probabilities $p_{12}$ based on
-          $X$ and $Y_1$. Then based on this probability, we determine if $R_2 =
-          1$. In the same manner, if 2 was selected in the previous step, we
-          generate probabilities $p_{21}$ based on $X$ and $Y_2$. Then based on
-          this probability, we determine if $R_2 = 1$.
-    \end{enumerate}
-    Like the monotone MAR simulation this algorithm produces similar final
-    results with the determination of binary variables $R_1$ and $R_2$ and
-    variables $X$, $Y_1$, and $Y_2$. Unlike the monotone MAR case, the
-    nonmonotone MAR includes observations with $Y_2$ observed and $Y_1$ missing.
-
-  \item \textbf{Simulation 1 with Monotone MAR}:
-    Following the algorithm described in the monotone MAR simulation bullet, 
-    we first generate data from the following distributions:
-    \begin{align*}
-        X_i &\stackrel{iid}{\sim} N(0, 1) \\
-        Y_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
-        Y_{2i} &\stackrel{iid}{\sim} N(\theta, 1)
-    \end{align*}
-
-    Then, we create the probabilities $p_1 = \logistic(x_i)$ and 
-    $p_{12} = \logistic(y_{1i})$.
-    Since, both $x_i$ and $y_1$ are standard normal distributions, each of these
-    probabilities is approximately $0.5$ in expectation.
-
-    The goal of this simulation is to estimate $\theta$. Alternatively, we can
-    express this as solving the estimating equation:
-
-    \[g(\theta) \equiv Y_2 - \theta = 0.\]
-
-    We estimate $\theta$ using the following procedures:
-
-    \begin{itemize}
-        \item Oracle: This computes $\bar Y$ using \textit{both} the observed
-          and missing data.
-        \item IPW-Oracle: This is an IPW estimator using only the observed
-          values of $Y_2$. The weights (inverse probabilities) use the actual
-          probabilities.
-        \item IPW-Est: This is an IPW estimator using the probabilities that
-          have been estimated by a logistic model.
-        \item Semi: This is the monotone semiparametric efficient estimator from
-          Slide 11 (Equation 2) of Dr. Kim's Nonmonotone Missingness
-          presentation.
-    \end{itemize}
-
-    We run this simulation with different values of $\theta$, sample size of
-    2000, and 2000 Monte Carlo replications. Each algorithm for each replication
-    generates $\hat \theta$. In the subsequent tables, we compute the bias,
-    standard deviation (sd), t-statistic (where we test for a significant
-    difference between the Monte Carlo mean $\hat \theta$ and the true $\theta$)
-    and the p-value of the t-statistic.
-
-    \newpage
-
-    \input{Tables/monomar_t-5.tex}
-    \input{Tables/monomar_t0.tex}
-    \input{Tables/monomar_t5.tex}
-
-    Overall, these results are mostly what I would have expected. All of the
-    algorithms estimate the true value of $\theta$ correctly in each case, with
-    the oracle estimate having the smallest variance followed by the
-    semiparametric algorithm. If there is anything surprising it is that the IPW
-    estimator has better performance with the estimated weights compared to the
-    true weights. However, I think that this is a known phenomenon.
-
-    \newpage
-
-    \item \textbf{Simulation 1 with Nonmonotone MAR}:
-
-    We generate variables $(X, Y_1, Y_2)$ using the following setup:
-
-    \[\begin{bmatrix}
-    X_i \\ \varepsilon_{1i} \\ \varepsilon_{2i}
-    \end{bmatrix} \stackrel{iid}{\sim}
-    N\left(
-    \begin{bmatrix}
-        0 \\ 0 \\ \theta
-    \end{bmatrix},
-    \begin{bmatrix}
-        1 & 0 & 0 \\
-        0 & 1 & \sigma_{yy}\\
-        0 & \sigma_{yy} & 1
-    \end{bmatrix}
-    \right).\]
-
-    Then,
-
-    \[y_{1i} = x_i + \varepsilon_{1i} \text{ and } 
-    y_{2i} = x_i + \varepsilon_{2i}.\]
-
-    Since we have nonmonotone data, our ``Stage 1'' probabilities are
-    different. We compute the true Stage 1 probabilities being proportional to
-    the following values:
-
-    \begin{align*}
-        p_0 &= 0.2 \\
-        p_1 &= 0.4 \\
-        p_2 &= 0.4 \\
-    \end{align*}
-
-   However, we keep the same structure for the Stage 2 probabilities with:
-   $p_{12} = \logistic(y_1)$ and $p_{21} = \logistic(y_2)$. The goal remains to
-   estimate $\theta$. We continue to use the Oracle algorithm and the IPW-Oracle
-   algorithm. Since we have nonmonotone MAR data, we use the ``Proposed''
-   algorithm that is described on Slide 25 (Equation 12) of Dr. Kim's
-   presentation. The outcome models were estimated using
-   logistic regression and OLS and correctly specified. The response model used
-   the oracle estimates of the probabilities. This yields the
-   following results:
-
-    \input{Tables/nonmonosim1m-5.tex}
-    \input{Tables/nonmonosim1m0.tex}
-    \input{Tables/nonmonosim1m5.tex}
-
-    \newpage
-
-    \item \textbf{Simulation 2 with Nonmonotone MAR}:
-    We also want to simulate data that is correlated. For this simulation, we
-    focus on $\Cov(Y_1, Y_2)$. The data generating process now has $\sigma_{yy}
-    \neq 0$. We are still interested in $\bar Y_2$ and we still run 2000
-    simulation with 2000 observations. In all of the next simulations the true
-    value of $\theta = 0$. The results are the following:
-
-    \input{Tables/nonmonosim2c0.1.tex}
-    \input{Tables/nonmonosim2c0.5.tex}
-    \input{Tables/nonmonosim2c0.9.tex}
-
-    \newpage
-
-  \item \textbf{Simulation 3 with Nonmonotone MAR}:
-    This simulation aims to see if the proposed algorithm is doubly robust.
-    First, we check with a misspecified outcome model. In this case the data
-    generating procedure is the following:
-
-    \[\begin{bmatrix}
-    X_i \\ \varepsilon_{1i} \\ \varepsilon_{2i}
-    \end{bmatrix} \stackrel{iid}{\sim}
-    N\left(
-    \begin{bmatrix}
-        0 \\ 0 \\ \theta
-    \end{bmatrix},
-    \begin{bmatrix}
-        1 & 0 & 0 \\
-        0 & 1 & \sigma_{yy}\\
-        0 & \sigma_{yy} & 1
-    \end{bmatrix}
-    \right).\]
-
-    Then,
-
-    \[y_{1i} = x_i + x_i^2 \varepsilon_{1i} \text{ and } 
-    y_{2i} = -x_i + x_i^3 + \varepsilon_{2i}.\]
-
-    This procedure causes $X$ to influence both $Y_1$ and $Y_2$ and we still
-    have correlation in the error terms of $Y_1$ and $Y_2$. However, since
-    neither $Y_1$ nor $Y_2$ are linear in $X$, the model will be misspecified.
-    The response mechanisms are first generated MCAR with a probability of
-    either $Y_1$ or $Y_2$ being the first variable observed to be $0.4$. (There
-    is a $0.2$ probability neither is observed.) Then the probability of the
-    other variable being observed is proportional to $\logistic(y_k)$ where
-    $y_k$ is the $y$ that has been observed. To ensure that the proposed method
-    has the correct propensity score we use the oracle probabilities instead of
-    estimating them. This yields the following:
-
-    \input{Tables/nonmonosim3c0.tex}
-    \input{Tables/nonmonosim3c0.1.tex}
-    \input{Tables/nonmonosim3c0.5.tex}
-
-    \newpage
-
-    Thus, the proposed method is unbiased with a misspecified outcome model.
-    We now show a simulation where the outcome model is correctly specified but
-    the response model is not.
-
-  \item \textbf{Simulation 4 with Nonmonotone MAR}:
-    Continuing to test if the proposed algorithm is doubly robust, this
-    simulation checks a misspecified response model. Instead of using oracle
-    weights as in Simulation 3, we estimate the weights for the proposed method.
-    However, unlike the true probabilities of being proportional to
-    $\logistic(y_k)$, this simulation has the true probabilities being
-    proportional to $\logistic(x_i)$.
-    The algorithms to which we compare still use the oracle weights.
-
-    \input{Tables/nonmonosim4c0.tex}
-    \input{Tables/nonmonosim4c0.1.tex}
-    \input{Tables/nonmonosim4c0.5.tex}
-
-    Thus, there is strong evidence that the proposed method is 
-    doubly robust. Previously, we did not get this result and the reason is 
-    the way that the selection probabilities work. When they were proportional
-    to $\logistic(y_k)$, the estimator is biased (see the next section);
-    however, when the probabilities are proportional to $\logistic(x_i)$ then it
-    works.
-
-\end{itemize}
-
-\newpage
-
 \section*{Missingness Mechanism}

### 2024-02-02 final/presentations/old_kimfuller_intro.qmd
Changed project structure.

---
title: "Combining Information from Multiple Surveys Using Probability Sampling"
author: "Caleb Leedy"
date: "August 15, 2023"
date-format: long
format: 
  revealjs:
    slide-number: c/t
    theme: [default, pres.scss]
bibliography: references.bib
---
<!---
Outline:
* Introduce notation

* Introduce problem:
  * semiparametrically efficient estimator to combine non-nested samples

* Proposed approach:
  * Use model-assisted semiparametric inference to get better / more robust results
  * i.e. use both a regression estimator and a model for the response probabilities

* Alternative Approaches:
  * Modified regression estimation of Merkouris (2004), Merkouris et al. (2023)
  * Multiphase regression estimator of Fuller (2009)

* Goals:
  * TODO:

* Next steps:
  * TODO:

Notation:
* Hidiroglou: Population universe (U), Sample i (s_i)
* Merkouris: same as above

* Inclusion probabilities: $\pi_k$
* Inclusion indicator R_i
* Variable names: x, y, z

--->

## Overall Goal

* We want to combine information from separate surveys and use 
semiparametric theory to get an optimal estimator.

# Existing Work

## Surveys with Missingness {.smaller}

Non-monotone Missingness

| Sample | $X$ | $Y_1$ | $Y_2$ | $R_1$ | $R_2$ |
|:------:|:---:|:-----:|:-----:|:-----:|:-----:|
|$A_{11}$| ✔️   | ✔️     | ✔️     | 1     | 1     |
|$A_{10}$| ✔️   | ✔️     |       | 1     | 0     |
|$A_{01}$| ✔️   |       | ✔️     | 0     | 1     |
|$A_{00}$| ✔️   |       |       | 0     | 0     |

* The goal is the estimate $\theta = E[g(X, Y_1, Y_2)]$.

## Estimating $\theta$ {.smaller}

Then for

\begin{align*}
b(X_i, Y_{1i}) &= E[g_i \mid X_i, Y_{1i}], \\ 
a(X_i, Y_{2i}) &= E[g_i \mid X_i, Y_{2i}], \text{ and } \\
i &= \{1, \dots, n\},
\end{align*}

\begin{align*}
\hat \theta_{\text{eff}} &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \\ 
&+ n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)}(b(X_i, Y_{1i}) - E[g_i \mid X_i]) \\ 
&+ n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)}(a(X_i, Y_{2i}) - E[g_i \mid X_i]) \\ 
&+ n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_i)}(g_i - b_i - a_i + E[g_i \mid X_i])
\end{align*}

# Simulation Studies

## Simulation 1

We use the following simulation setup

$$
\begin{bmatrix}
X_i \\ 
\varepsilon_{1i} \\
\varepsilon_{2i} \\
\end{bmatrix} \stackrel{iid}{\sim}
N\left(
\begin{bmatrix}
0 \\ 0 \\ \theta
\end{bmatrix},
\begin{bmatrix}
1 & 0 & 0 \\ 
0 & 1 & \sigma_{yy} \\
0 & \sigma_{yy} & 1
\end{bmatrix}
\right)
$$

* $y_{1i} = x_i + \varepsilon_{1i}$
* $y_{2i} = x_i + \varepsilon_{2i}$

## Simulation 1

* Following the graphical MAR of [@robins1997non], for Stage 1, we have,

$$ p_0 = 0.2, p_1 = 0.4, \text{ and } p_2 = 0.4. $$

* For Stage 2, we have,

$$ p_{12} = \text{logistic}(y_1) \text{ and } p_{21} = \text{logistic}(y_2). $$

* The goal is to estimate $g = Y_2 - \theta$.

## Simulation 1

Then we compare three estimators:

* Oracle: we use both the observed and missing values.
* IPW-Oracle: we use 
$$ \hat \theta_{IPW} = n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_i}y_{2i}.$$
* The proposed estimator $\hat \theta_{eff}$.

## Results

```{r}

library(MASS)
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)

library(doParallel)
library(doRNG)
library(parallelly)

# *****************
# * Generate Data *
# *****************

source("../Simulations/generate_data.R")

# ************************
# * Estimating Functions *
# ************************

source("../Simulations/nmono_est_funs.R")

run_init_sims <- function(n_sim, B, true_mean, cor_y1y2,
                          miss_out, est_weights, seed, r_ind_y = FALSE) {

  clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
  registerDoParallel(clust)

  mc_theta <- 
    foreach(b = 1:B, 
            .options.RNG = seed,
            .export = c("nonmono_mar", "ipw_cc_oracle", "est_nonmono", "expit"),
            .packages = c("dplyr")) %dorng% {

    # Parameters
    df <- nonmono_mar(n_sim, mean_y2 = true_mean, cor_y1y2 = cor_y1y2,
                      r_ind_y = r_ind_y, miss_out = miss_out)

    # Estimation
    # We want to compute:
    # 1. Oracle estimate
    # 2. pi_11 estimate (ipw using fully observed cases)
    # 3. New estimate

    tibble(oracle = mean(df$y2),
           ipworacle = ipw_cc_oracle(df),
           proposed = est_nonmono(df, oracle = !est_weights))

  } %>% bind_rows()

  stopCluster(clust)

  # ****************
  # * Analyze Data *
  # ****************

  mc_theta %>%
   summarize(
     bias_oracle = mean(oracle) - true_mean,
     bias_ipworacle = mean(ipworacle) - true_mean,
     bias_proposed = mean(proposed) - true_mean,
     sd_oracle = sd(oracle),
     sd_ipworacle = sd(ipworacle),
     sd_proposed = sd(proposed),
     tstat_oracle = (mean(oracle) - true_mean) / sqrt(var(oracle) / B),
     tstat_ipworacle = (mean(ipworacle) - true_mean) / sqrt(var(ipworacle) / B),
     tstat_proposed = (mean(proposed) - true_mean) / sqrt(var(proposed) / B))%>%
    tidyr::pivot_longer(cols = everything(),
                 names_to = c(".value", "algorithm"),
                 names_pattern = "(.*)_(.*)") %>%
    mutate(pval = pt(-abs(tstat), df = B))

}

# **********************************
# * Simulations: Sim 1 Nonmonotone *
# **********************************

run_init_sims(n_sim = 2000, B = 2000,
              true_mean = -5, cor_y1y2 = 0,
              miss_out = FALSE, est_weights = FALSE, seed = 1) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = "True Value is -5. Cor(eps1, eps2) = 0")

```

## Results

```{r}
run_init_sims(n_sim = 2000, B = 2000,
              true_mean = 0, cor_y1y2 = 0,
              miss_out = FALSE, est_weights = FALSE, seed = 1) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = "True Value is 0. Cor(eps1, eps2) = 0")

```

## Results

```{r}
run_init_sims(n_sim = 2000, B = 2000,
              true_mean = 5, cor_y1y2 = 0,
              miss_out = FALSE, est_weights = FALSE, seed = 1) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = "True Value is 5. Cor(eps1, eps2) = 0")

```

## Simulation 2

Simulation 2 follows the same structure are Simulation 1 except that 
instead of testing variation win the mean, we test differences in the covariance
of $\varepsilon_1$ and $\varepsilon_2$.

## Results

```{r}

run_init_sims(n_sim = 2000, B = 2000,
              true_mean = 0, cor_y1y2 = 0.1,
              miss_out = FALSE, est_weights = FALSE, seed = 2) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = "True Value is 0. Cor(Y1, Y2) = 0.1")

```

## Results

```{r}

run_init_sims(n_sim = 2000, B = 2000,
              true_mean = 0, cor_y1y2 = 0.5,
              miss_out = FALSE, est_weights = FALSE, seed = 2) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = "True Value is 0. Cor(Y1, Y2) = 0.5")

```

## Results

```{r}

run_init_sims(n_sim = 2000, B = 2000,
              true_mean = 0, cor_y1y2 = 0.9,
              miss_out = FALSE, est_weights = FALSE, seed = 2) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = "True Value is 0. Cor(Y1, Y2) = 0.9")

```

## Simulation 3

<!--- FIXME: -->

## Results

```{r}

library(dplyr)
library(CVXR)

library(doParallel)
library(doRNG)
library(parallelly)

# *****************************
# * Data Generation Functions *
# *****************************
source("generate_data.R")

# *************************************
# * Non/monotone Estimation Functions *
# *************************************
source("nmono_est_funs.R")

run_cali_sims <- function(n_sim, B, true_mean, cor_y1y2,
                          miss_out, est_weights, seed, r_ind_y = FALSE) {

  clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
  registerDoParallel(clust)

  mc_theta <- 
    foreach(b = 1:B, 
           .options.RNG = seed,
           .export = c("nonmono_mar", "ipw_cc_oracle", "est_nonmono", "expit",
                       "nonmono_est_weights", "twophase_reg", "threephase_reg"),
           .packages = c("dplyr", "CVXR")) %dorng% {

    # Parameters
    df <- nonmono_mar(n_sim, mean_y2 = true_mean, cor_y1y2 = cor_y1y2,
                      r_ind_y = r_ind_y, miss_out = miss_out)

    # Estimation
    # We want to compute:
    # 1. Oracle estimate
    # 2. pi_11 estimate (ipw using fully observed cases)
    # 3. New estimate

    tibble(oracle = mean(df$y2),
           ipworacle = ipw_cc_oracle(df),
           reg2p = twophase_reg(df, reg_on_y1 = FALSE),
           reg3p = threephase_reg(df),
           proposed = est_nonmono(df, oracle = !est_weights),
           calib = nonmono_est_weights(df))

  } %>% bind_rows()

  stopCluster(clust)

  # ****************
  # * Analyze Data *
  # ****************

  mc_theta %>%
   summarize(
     bias_oracle = mean(oracle) - true_mean,
     bias_ipworacle = mean(ipworacle) - true_mean,
     bias_proposed = mean(proposed) - true_mean,
     bias_reg2p = mean(reg2p) - true_mean,
     bias_reg3p = mean(reg3p) - true_mean,
     bias_calib = mean(calib) - true_mean,
     sd_oracle = sd(oracle),
     sd_ipworacle = sd(ipworacle),
     sd_proposed = sd(proposed),
     sd_reg2p = sd(reg2p),
     sd_reg3p = sd(reg3p),
     sd_calib = sd(calib),
     tstat_oracle = (mean(oracle) - true_mean) / sqrt(var(oracle) / B),
     tstat_ipworacle = (mean(ipworacle) - true_mean) / sqrt(var(ipworacle) / B),
     tstat_proposed = (mean(proposed) - true_mean) / sqrt(var(proposed) / B),
     tstat_reg2p = (mean(reg2p) - true_mean) / sqrt(var(reg2p) / B),
     tstat_reg3p = (mean(reg3p) - true_mean) / sqrt(var(reg3p) / B),
     tstat_calib = (mean(calib) - true_mean) / sqrt(var(calib) / B)) %>%
    tidyr::pivot_longer(cols = everything(),
                 names_to = c(".value", "algorithm"),
                 names_pattern = "(.*)_(.*)") %>%
    mutate(pval = pt(-abs(tstat), df = B))

}

tr_m <- -5
cy1y2 <- 0
run_cali_sims(n_sim = 1000, B = 1000,
              true_mean = tr_m, cor_y1y2 = cy1y2,
              miss_out = FALSE, est_weights = FALSE, seed = 1) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = paste0("True Value is ", tr_m, ". Cor(Y1, Y2) = ", cy1y2)) %>%

```

## Results

```{r}

tr_m <- 0
cy1y2 <- 0
run_cali_sims(n_sim = 1000, B = 1000,
              true_mean = tr_m, cor_y1y2 = cy1y2,
              miss_out = FALSE, est_weights = FALSE, seed = 2) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = paste0("True Value is ", tr_m, ". Cor(Y1, Y2) = ", cy1y2))

```

## Results

```{r}

tr_m <- 5
cy1y2 <- 0
run_cali_sims(n_sim = 1000, B = 1000,
              true_mean = tr_m, cor_y1y2 = cy1y2,
              miss_out = FALSE, est_weights = FALSE, seed = 4) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = paste0("True Value is ", tr_m, ". Cor(Y1, Y2) = ", cy1y2))

```

## Problem Setup

::: {.incremental}
* Let $U$ be a finite population.
* Let $s_1$ and $s_2$ be samples of $U$
  - It may be the case that $s_1 \cap s_2 = \emptyset$ or $s_1 \cap s_2 \neq \emptyset$
  but $s_1 \subsetneq s_2$ and $s_2 \subsetneq s_1$.
  - This is an example of *non-nested* sampling [@hidiroglou2001double]
:::

## {}

::: {.incremental}
* Let $X$ be observed in both $s_1$ and $s_2$.
* Let $Y_1$ be only observed in $s_1$.
* Let $Y_2$ be only observed in $s_2$.
* We denote the first-order inclusion probabilities as $\pi^{(i)}_{j}$ of 
element $Y_j$ in survey $s_i$.
* If element $Y_j$ is in $s_i$, then let $R_j^{(i)} = 1$. Otherwise, $R_j^{(i)} = 0$.
:::

## {}

* This setup is very similar to the NRI project with $X$ consisting of core data
points and $Y_1$ and $Y_2$ being part of the supplemental samples.

## Research Problem

* How to estimate $\bar Y_1$ efficiently from information in
samples $s_1$ and $s_2$?

## Similar Methods

* Weight adjustment [@hidiroglou2001double]
* GREG Estimation [@merkouris2004combining, @merkouris2023combining]
* Projection estimator [@kim2012combining]
* Mass imputation [@chen2022nonparametric]

## Proposed Approach

* Use semiparametric theory to construct an efficient estimator.
* We fist consider the case of non-informative sampling.

# Comparison with Existing Methods

## Weight adjustment [@hidiroglou2001double]

:::: {.columns}

::: {.column width="50%"}
Similarities with Proposed Method

* Estimates $\bar Y_1$. 
* Considers the non-nested sampling model.
* Incorporates a regression model.

:::

::: {.column width="50%"}
Differences with Proposed Method

* No $Y_2$ in $s_2$.
* Does not have semiparametric theory.
* Does not use a response model.

:::

::::

## GREG estimator [@merkouris2004combining]

:::: {.columns}

::: {.column width="50%"}
Similarities with Proposed Method

* Estimates $\bar Y_1$. 
* Considers the non-nested sampling model.
* Constructs optimal regression model (conditions out calibrated variables).

:::

::: {.column width="50%"}
Differences with Proposed Method

* No $Y_2$ in $s_2$.
* Does not use a response model.

:::

::::

## Projection estimator [@kim2012combining]

:::: {.columns}

::: {.column width="50%"}
Similarities with Proposed Method

* Estimates $\bar Y_1$.
* Considers $Y_2 \in s_2$.
* Considers a non-nested sampling model.

:::

::: {.column width="50%"}
Differences with Proposed Method

* Imputes missing values of $Y_2$ with a mass imputation estimator.

:::

::::

## Mass imputation [@chen2022nonparametric]

:::: {.columns}

::: {.column width="50%"}
Similarities with Proposed Method

* Can be used to estimate $\bar Y_1$.
* Considers $Y_2 \in s_2$.
* Non-nested model.

:::

::: {.column width="50%"}
Differences with Proposed Method

* Nonparametric estimator.
* No response model.
* Considers $s_2$ to be non-probability sample.

:::

::::

## Proposed Next Steps

* Propose semiparametric estimator.

## References

## Simulation 2

## Simulation 3

## Problem Setup

::: {.incremental}
* Let $U$ be a finite population.
* Let $s_1$ and $s_2$ be samples of $U$
  - It may be the case that $s_1 \cap s_2 = \emptyset$ or $s_1 \cap s_2 \neq \emptyset$
  but $s_1 \subsetneq s_2$ and $s_2 \subsetneq s_1$.
  - This is an example of *non-nested* sampling [@hidiroglou2001double]
:::

## {}
::: {.incremental}
* Let $X$ be observed in both $s_1$ and $s_2$.
* Let $Y_1$ be only observed in $s_1$.
* Let $Y_2$ be only observed in $s_2$.
* We denote the first-order inclusion probabilities as $\pi^{(i)}_{j}$ of 
element $Y_j$ in survey $s_i$.
* If element $Y_j$ is in $s_i$, then let $R_j^{(i)} = 1$. Otherwise, $R_j^{(i)} = 0$.
:::

## {}

* This setup is very similar to the NRI project with $X$ consisting of core data
points and $Y_1$ and $Y_2$ being part of the supplemental samples.

## Research Problem

* How to estimate $\bar Y_1$ efficiently from information in
samples $s_1$ and $s_2$?

## Similar Methods

* Weight adjustment [@hidiroglou2001double]
* GREG Estimation [@merkouris2004combining, @merkouris2023combining]
* Projection estimator [@kim2012combining]
* Mass imputation [@chen2022nonparametric]

## Proposed Approach

* Use semiparametric theory to construct an efficient estimator.
* We fist consider the case of non-informative sampling.

# Comparison with Existing Methods

## Weight adjustment [@hidiroglou2001double]

:::: {.columns}

::: {.column width="50%"}
Similarities with Proposed Method

* Estimates $\bar Y_1$. 
* Considers the non-nested sampling model.
* Incorporates a regression model.

:::

::: {.column width="50%"}
Differences with Proposed Method

* No $Y_2$ in $s_2$.
* Does not have semiparametric theory.
* Does not use a response model.

:::

::::

## GREG estimator [@merkouris2004combining]

:::: {.columns}

::: {.column width="50%"}
Similarities with Proposed Method

* Estimates $\bar Y_1$. 
* Considers the non-nested sampling model.
* Constructs optimal regression model (conditions out calibrated variables).

:::

::: {.column width="50%"}
Differences with Proposed Method

* No $Y_2$ in $s_2$.
* Does not use a response model.

:::

::::

## Projection estimator [@kim2012combining]

:::: {.columns}

::: {.column width="50%"}
Similarities with Proposed Method

* Estimates $\bar Y_1$.
* Considers $Y_2 \in s_2$.
* Considers a non-nested sampling model.

:::

::: {.column width="50%"}
Differences with Proposed Method

* Imputes missing values of $Y_2$ with a mass imputation estimator.

:::

::::

## Mass imputation [@chen2022nonparametric]

:::: {.columns}

::: {.column width="50%"}
Similarities with Proposed Method

* Can be used to estimate $\bar Y_1$.
* Considers $Y_2 \in s_2$.
* Non-nested model.

:::

::: {.column width="50%"}
Differences with Proposed Method

* Nonparametric estimator.
* No response model.
* Considers $s_2$ to be non-probability sample.

:::

::::

## Proposed Next Steps

* Propose semiparametric estimator.

## References

### 2024-02-02 src/explore/20240131-gls_explanation.qmd
Changed project structure.

---
title: "GLS Explanation"
author:
  - name: Caleb Leedy
date: today
date-format: "D MMMM YYYY"
format:
  pdf:
    include-in-header: 
      - "latex_header.tex"
      - text: \renewcommand{\arraystretch}{1.4}
    keep-tex: true
---

# Introduction

This is a write up explaning the dual form of generalized least squares (GLS).
We think of GLS as an extension of ordinary least squares (OLS), so I will
discuss this first and then comment on how these are connected and the dual
formulation for both.

# The Gauss-Markov Theorem

It all comes back to the Gauss-Markov Theorem, which states that for a linear
model in the form,

$$ Y = X\beta + \varepsilon$$

with $E[\varepsilon] = 0$ and $\Var(\varepsilon) = \sigma^2 I_n$, the BLUE (best
linear unbiased estimator) of $C\beta$ for $C = AX$ for some matrix $A$ is
(assuming $X'X$ has an inverse),

$$ C\hat \beta = C(X'X)^{-1} X'Y = AX(X'X)^{-1}X'Y.$$

## GLS Extension

To extend this to the GLS setting consider the case where $\Var(\varepsilon) =
\sigma^2 V$ for some known positive-definite matrix $V$. In this case, we can
rewrite $Y = X\beta + \varepsilon$ as

$$ Z = W\beta + \tilde \varepsilon$$

where $Z = V^{-1/2} Y$, $W = V^{-1/2}X$ and 
$\tilde \varepsilon = V^{-1/2} \varepsilon$. But this formulation implies that
$E[\tilde \varepsilon] = 0$ and 
$\Var(\tilde \varepsilon) = \sigma^2 I_n$ and so hence by the Gauss-Markov
Theorem, the BLUE is

$$ (W'W)^{-1} W'Z = (X'V^{-1}V)^{-1} X'V^{-1}Y.$$

# Dual Form

All of this comes back to the fact that we want to estimate the BLUE. However,
we could do this directly. By definition, the BLUE is the minimum variance
linear unbiased estimator. This means that we consider a class of estimators
$CW'Y$ such that $E[CW'Y] = C\beta$ that minimizes the variance $CW'\Var(Y) WC'
\propto CW'WC'$. Since we assume the model form of $Y = X\beta + \varepsilon$,
the unbiasedness condition is equivalent to,

$$ E[CW'Y] = CW'E[Y] = CW'X\beta = C\beta. $$

Since this is true for all $\beta$, it must be the case that

$$ W'X = I_p \text{ for } p = \dim(\beta).$$

Hence, we have a Lagrangian

$$L(W) \propto CW'WC' - (W'X - I_p)\lambda.$$

Thus, the first-order conditions are $W - X\lambda = 0$ and $W'X - I_p = 0$.
This yields a solution of $W = X(X'X)^{-1}$ which is exactly what we get from
OLS because the final result is $CW'Y = C(X'X)^{-1} X'Y$. To get the GLS form,
we can keep the same constraint except minimize the variance $W'VW$.

<!---
# Additional Constraints

In the previous section, we assumed that we only had a single constraint: $W'X =
I_p$. Now, consider the case in which we add linear equality constraints that
are free of additional parameters. We can denote these additional constraints
by,

$$ W'\tilde X = \alpha.$$

Putting these constraints together, yields,

$$W' [X, \tilde X] - [I_p, \alpha] := W'Z - M = 0.$$

Solving this constrained optimization problem yields a linear estimator

$$W'Y = M(Z'V^{-1}Z)^{-1}Z'V^{-1}Y.$$

This could be useful for our simulation approach as we compare estimators with
different constraints.
--->

# Estimating the Variance

When estimating GLS, we assume that the covariance matrix $V$ is fully known.
What do we do if we have to estimate $V$? In this case, we need to use
(finite-dimensional) semiparametric theory to estimate the optimal $W$ such that
$W$ belongs to the orthogonal nuisance tangent space of $\hat V$. This is
something that we could explore further if we want to.

### 2024-02-02 src/explore/20240202-gls_sim.qmd
Changed project structure.

--- 2024-02-02 Notes/sim_no00.qmd
@@
 ---
-title: "Non-Monotone GLS Simulation: No $A_{00}$"
+title: "Non-Monotone GLS Simulation: Simple"
 author:
@@
   y1 <- x + e1
-  y2 <- mu + x + e1
+  y2 <- mu + x + e2
 
@@
            delta_00 = 0, delta_10 = 0, delta_01 = 0, delta_11 = 1,
-           pi_00 = 1/3, pi_10 = 1/3, pi_01 = 1/3, pi_11 = 1/3)
+           pi_00 = 0.25, pi_10 = 0.25, pi_01 = 0.25, pi_11 = 0.25)
   # A_10
@@
            delta_00 = 0, delta_10 = 1, delta_01 = 0, delta_11 = 0,
-           pi_00 = 1/3, pi_10 = 1/3, pi_01 = 1/3, pi_11 = 1/3)
+           pi_00 = 0.25, pi_10 = 0.25, pi_01 = 0.25, pi_11 = 0.25)
   # A_01
@@
            delta_00 = 0, delta_10 = 0, delta_01 = 1, delta_11 = 0,
-           pi_00 = 1/3, pi_10 = 1/3, pi_01 = 1/3, pi_11 = 1/3)
+           pi_00 = 0.25, pi_10 = 0.25, pi_01 = 0.25, pi_11 = 0.25)
   # A_00
@@
            delta_00 = 1, delta_10 = 0, delta_01 = 0, delta_11 = 0,
-           pi_00 = 1/3, pi_10 = 1/3, pi_01 = 1/3, pi_11 = 1/3)
+           pi_00 = 0.25, pi_10 = 0.25, pi_01 = 0.25, pi_11 = 0.25)
 
-  df <- bind_rows(df_11, df_10, df_01)
+  df <- bind_rows(df_11, df_10, df_01, df_00)
 
@@
                       1, 0, 0,
-                      0, 0, 1), ncol = 3, byrow = TRUE)
+                      0, 0, 1,
+                      1, 0, 0), ncol = 3, byrow = TRUE)
 
@@
         mutate(df, tmp = delta_01 / pi_01 * X) |> pull(tmp) |> mean(),
-        mutate(df, tmp = delta_01 / pi_01 * Y2) |> pull(tmp) |> mean())
+        mutate(df, tmp = delta_01 / pi_01 * Y2) |> pull(tmp) |> mean(),
+        mutate(df, tmp = delta_00 / pi_00 * X) |> pull(tmp) |> mean())
 
@@
                1, 2, 1 + rho,
-               1, 1 + rho, mu^2 + 2), nrow = 3, byrow = TRUE) -
-      matrix(c(0, 0, 0,
-               0, 0, 0,
-               0, 0, mu^2), nrow = 3, byrow = TRUE)
-    v_10 <- 1 / df$pi_10[1] * matrix(c(1, 1, 1, 2), nrow = 2, byrow = TRUE)
-    v_01 <- 
-      1 / df$pi_01[1] * matrix(c(1, 1, 1, mu^2 + 2), nrow = 2) - 
-      matrix(c(0, 0, 0, mu^2), nrow = 2)
+               1, 1 + rho, 2), nrow = 3)
+    v_10 <- 1 / df$pi_10[1] * matrix(c(1, 1, 1, 2), nrow = 2)
+    v_01 <- 1 / df$pi_01[1] * matrix(c(1, 1, 1, 2), nrow = 2)
+    v_00 <- 1 / df$pi_00[1]
 
-    v_mat <- matrix(rep(0, 49), ncol = 7)
+    v_mat <- matrix(rep(0, 64), ncol = 8)
     v_mat[1:3, 1:3] <- v_11
@@
     v_mat[6:7, 6:7] <- v_01
+    v_mat[8, 8] <- v_00
 
@@
 }
-
 ```
@@
 #|   Results from simulations study with independent equally sized segments
-#|   $A_{11}$, $A_{10}$, and $A_{01}$ all of size $n = 250$. In this
+#|   $A_{11}$, $A_{10}$, $A_{01}$, and $A_{00}$ all of size $n = 250$. In this
 #|   simulation we have the true mean of $Y_2$ equal to $\mu = 5$ and the covariance
@@
     list(Bias = ~mean(.) - true_mu,
-         SD = sd,
+         SD = ~sd(.),
          Tstat = ~(mean(.) - true_mu) / sqrt(var(.) / B)),
@@
 
-A quick comparison between the standard deviation of the GLS estimator and the
-standard deviation of the GLS estimator in `gls_sim.qmd` shows that this
-standard deviation is slightly larger, which means that we do lose efficiency by
-not including $A_{00}$.
+The theoretical variance of $\hat g_{GLS}$ is
+
+```{r}
+#| echo: false
+
+z_mat <- matrix(c(1, 0, 0,
+                  0, 1, 0,
+                  0, 0, 1,
+                  1, 0, 0,
+                  0, 1, 0,
+                  1, 0, 0,
+                  0, 0, 1,
+                  1, 0, 0), ncol = 3, byrow = TRUE)
+
+v_11 <- 1 / 0.25 *
+  matrix(c(1, 1, 1,
+           1, 2, 1 + cov_e1e2,
+           1, 1 + cov_e1e2, 2), nrow = 3)
+v_10 <- 1 / 0.25 * matrix(c(1, 1, 1, 2), nrow = 2)
+v_01 <- 1 / 0.25 * matrix(c(1, 1, 1, 2), nrow = 2)
+v_00 <- 1 / 0.25
+
+v_mat <- matrix(rep(0, 64), ncol = 8)
+v_mat[1:3, 1:3] <- v_11
+v_mat[4:5, 4:5] <- v_10
+v_mat[6:7, 6:7] <- v_01
+v_mat[8, 8] <- v_00
+
+v_mat <- v_mat / (4 * obs_seg)
+ones_v <- t(c(0, 0, 1))
+
+theo_var <- 
+  ones_v %*% MASS::ginv(t(z_mat) %*% MASS::ginv(v_mat) %*% z_mat) %*% t(ones_v)
+
+```
+
+$$ V(\hat g_{GLS}) = [0, 0, 1] (Z'V^{-1}Z)^{-1} [0, 0, 1]'.$$
+
+This is equal to `{r} theo_var` which means that the standard deviation 
+is `{r} sqrt(theo_var)`.

### 2024-02-05 src/explore/20240202-estcovmat.qmd
I realized that the Gram-Schmidt process was not going to work. We resort to estimating the covariance and variance directly.

---
title: "Estimating the Covariance Matrix"
author:
  - name: Caleb Leedy
date: today
date-format: "D MMMM YYYY"
format:
  pdf:
    include-in-header: 
      - "latex_header.tex"
---

```{r}
#| label: global variables
#| echo: false
B <- 1000

obs_seg <- 250
true_mu <- 5
cov_e1e2 <- 0.5

```

# Summary

In this document, we

1. Propose a new technique for estimating a covariance matrix, and 
2. Show via simulation studies that it works.

# Problem and Proposal

## Problem

Previously, have a model $\hat g = Z g + e$ where $\hat g$ is defined by

\begin{align*}
g_1^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_1(x_i) \\
g_2^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_2(y_{1i}) \\
g_3^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_3(y_{2i}) \\
g_1^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} g_1(x_i) \\
g_2^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} g_2(y_{1i}) \\
g_1^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} g_1(x_i) \\
g_3^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} g_2(y_{2i}) \\
g_1^{(00)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{00}}{\pi_{00}} g_1(x_i) \\
\end{align*}

and

$$\hat g = 
\begin{bmatrix}
g_1^{(11)} \\
g_2^{(11)} \\
g_3^{(11)} \\
g_1^{(10)} \\
g_2^{(10)} \\
g_1^{(01)} \\
g_3^{(01)} \\
g_1^{(00)} \\
\end{bmatrix},
Z = 
\begin{bmatrix}
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 0 & 1 \\
1 & 0 & 0 \\
0 & 1 & 0 \\
1 & 0 & 0 \\
0 & 0 & 1 \\
1 & 0 & 0 \\
\end{bmatrix},
E[e] = 0,
\text{ and } 
\Var(e) = n^{-1}
\begin{bmatrix}
V_{11} & 0 & 0 & 0 \\
0 & V_{10} & 0 & 0 \\
0 & 0 & V_{01} & 0 \\
0 & 0 & 0 & V_{00} \\
\end{bmatrix}.
$$

We also have

$$
V_{11} = 
\begin{bmatrix}
\frac{1}{\pi_{11}} E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{11}} E[g_1g_2] -
E[g_1]E[g_2] & \frac{1}{\pi_{11}} E[g_1g_3] - E[g_1]E[g_3] \\
\frac{1}{\pi_{11}} E[g_1g_2] - E[g_1]E[g_2] & \frac{1}{\pi_{11}} E[g_2^2] -
E[g_2]^2 & \frac{1}{\pi_{11}} E[g_2g_3] - E[g_2]E[g_3] \\
\frac{1}{\pi_{11}} E[g_1g_3] - E[g_1]E[g_3] & \frac{1}{\pi_{11}} E[g_2g_3] -
E[g_2]E[g_3] & \frac{1}{\pi_{11}} E[g_3^2] - E[g_3]^2 \\
\end{bmatrix},
$$
$$
V_{10} = 
\begin{bmatrix}
\frac{1}{\pi_{10}}E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{10}}E[g_1g_2] - E[g_1]E[g_2]\\
\frac{1}{\pi_{10}}E[g_1g_2] - E[g_1]E[g_2] & \frac{1}{\pi_{10}}E[g_2^2] - E[g_2]^2 \\
\end{bmatrix},
$$
$$
V_{01} = 
\begin{bmatrix}
\frac{1}{\pi_{01}}E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{01}}E[g_1g_3] - E[g_1]E[g_3]\\
\frac{1}{\pi_{01}}E[g_1g_3] - E[g_1]E[g_3] & \frac{1}{\pi_{01}}E[g_3^2] - E[g_3]^2 \\
\end{bmatrix}, \text{ and }
V_{00} = 
\begin{bmatrix}
\frac{1}{\pi_{00}}E[g_1^2] - E[g_1]^2
\end{bmatrix}.
$$

However, this challenge with actually using this model is that we have to use
the matrix $V$, which we assume to be *known*. In this write up, we will not
make this assumption.

## Proposals

### Proposal 1: REML

One potential solution would be to directly estimate $V$ and use some sort of
$\hat V$. We can do this with a REML estimate.

### Proposal 2: Orthogonal g-functions

For this idea we choose functions that after estimation are approximately
independent from each
other with a standard variance. If this is true, the covariance matrix $V$ is
approximately the identity matrix. Not only do we not have to estimate $V$, but
we do not even need to invert it! I now give a more detailed explanation of this
method.

Suppose that we observe variables $Z = (X_{m_1}, Y_{m_2})$ with an objective of
estimating $\theta = E[g(Y_{m_2})$ for some function $g$ such that the
variables are subject to missingness. We assume that there are $R$ unique
combinations of observed variables including a fully observed case. We can index
the combinations of observed variables by $r$ and assume that the fully observed
case occurs at $r = 1$. Let $G_r(Z)$ be the variables that are observed at a
particular value of $r$. We can choose a sequence of functions $f_1, \dots, f_K$
that we want to estimate. Each $f_k$ is assumed to be a function of a subset of
$Z$ and it makes sense to assume (since these are chosen by the analyst) that
there is at least one $f_k$ for each combination of observed variables $G_r(Z)$.
Let $A_k$ be the sets of observed variable combinations that can be evaluate by
the function $f_k$. If $f_k$ can be evaluated by the observed variables
$G_r(Z)$, this will consist of all of the combinations of variables
$r'$ such that $G_r(Z) \subseteq G_r'(Z)$. Because we assume that $G_1(Z) = Z$,
$A_k$ is always non-empty.

Once we have the original $f_1, \dots, f_K$ we orthogonalize them using a
Gram-Schmidt process. Let $g_1 = f_1 / \hat \Var(f_1)$. Then for $k > 1$
we can continue using the sequence,

$$\tilde g_k = f_k - \sum_{i = 1}^k \frac{\hat \Cov(f_k, g_i)}{\hat \Var(f_k)}
g_i \text{ followed by } g_k = \tilde g_k / \hat \Var(\tilde g_k).$$

However, the parameters for $\tilde g_k$ and $k > 1$ are all regression
coefficients and the multiplier for $g_k$ is simply computing a variance.
Then to
achieve efficiency, we propose estimating the regression coefficients by running
the coresponding regression on all of the data points in $A_k$ as well as
computing the variance of $\tilde g_k$ within $A_k$.

# Simulation

<!---# Compare with: Known V in non-monotone case --->
## Non-Monotone Case

First, we consider the following non-monotone simulation setup.

\begin{align*}
  \begin{bmatrix} x_1 \\ e_1 \\ e_2 \end{bmatrix} 
  &\stackrel{ind}{\sim} N\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}\right) \\
  x_2 &= x + e_1 \\
  y &= \mu + x + e_2 \\
\end{align*}

This yields outcome variables $X_1$ and $X_2$ that are correlated both with $Y$
and additionally with each other. To generate the missingness pattern, we select
`r obs_seg` observations into the four segments independently using simple
random sampling. The goal of this estimation problem is to estimate $\theta = E[Y]$.

```{r}
#| label: setup
#| echo: false
#| warning: false
#| message: false

library(dplyr)

library(doParallel)
library(doRNG)
library(parallelly)

# Run from src/
source("R/data_generation.R")

```

```{r}
#| label: gls functions
#| echo: false

#' This function is also used in 20240202-gls_sim.qmd. I have modified the code
#' to account for the change from X, Y1, Y2 to X1, X2, Y.
gls_est <- function(df, class = "basic", rho) {

  if (class == "basic") {
    z_mat <- matrix(c(1, 0, 0,
                      0, 1, 0,
                      0, 0, 1,
                      1, 0, 0,
                      0, 1, 0,
                      1, 0, 0,
                      0, 0, 1,
                      1, 0, 0), ncol = 3, byrow = TRUE)

    g_hat <- 
      c(mutate(df, tmp = delta_11 / pi_11 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * X2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_00 / pi_00 * X1) |> pull(tmp) |> mean())

    v_11 <- 1 / df$pi_11[1] *
      matrix(c(1, 1, 1,
               1, 2, 1 + rho,
               1, 1 + rho, 2), nrow = 3)
    v_10 <- 1 / df$pi_10[1] * matrix(c(1, 1, 1, 2), nrow = 2)
    v_01 <- 1 / df$pi_01[1] * matrix(c(1, 1, 1, 2), nrow = 2)
    v_00 <- 1 / df$pi_00[1]

    v_mat <- matrix(rep(0, 64), ncol = 8)
    v_mat[1:3, 1:3] <- v_11
    v_mat[4:5, 4:5] <- v_10
    v_mat[6:7, 6:7] <- v_01
    v_mat[8, 8] <- v_00

    v_mat <- v_mat / nrow(df)

    ghat <- MASS::ginv(t(z_mat) %*% MASS::ginv(v_mat) %*% z_mat) %*% 
      t(z_mat) %*% MASS::ginv(v_mat) %*% g_hat

    return(as.numeric(t(c(0, 0, 1) %*% ghat)))

  } else {
    stop("We have only 1 class: 'basic'.")
  }
}

```

```{r}
#| label: proposed estimation
#| echo: false

ortho_gls <- function(df, class = "basic") {

  # For class == 'basic', we consider the functions f_k to be originally,
  # f1 = X1, f2 = X2, f3 = Y.

  if (class == "basic") {

    # Get functions g_k
    g1 <- df$X1 / var(df$X1)
    df <- mutate(df, g1 = g1)
    df_1 <- filter(df, delta_11 == 1 | delta_10 == 1)
    t2 <- lm(X2 ~ g1, data = df_1)
    g2 <- df_1$X2 - t2$coefficients[2] * df_1$g1
    g2 <- g2 / var(g2)
    df_1 <- mutate(df_1, g2 = g2)
    df_11 <- filter(df_1, delta_11 == 1)
    t3 <- lm(Y ~ X1 + X2, data = df_11)
    g3 <- 
      df_11$Y - t3$coefficients[2] * df_11$g1 - t3$coefficients[3] * df_11$g2
    g3 <- g3 / var(g3)

    g_hat <- 
      c(mutate(df, tmp = delta_11 / pi_11 * g1) |> pull(tmp) |> sum(),
        mutate(df_1, tmp = delta_11 / pi_11 * g2) |> pull(tmp) |> sum(),
        mutate(df_11, tmp = delta_11 / pi_11 * g3) |> pull(tmp) |> sum(),
        mutate(df, tmp = delta_10 / pi_10 * g1) |> pull(tmp) |> sum(),
        mutate(df_1, tmp = delta_10 / pi_10 * g2) |> pull(tmp) |> sum(),
        mutate(df, tmp = delta_01 / pi_01 * g1) |> pull(tmp) |> sum(),
        # ERROR: This does not work!
        mutate(df_11, tmp = delta_01 / pi_01 * g3) |> pull(tmp) |> sum(),
        mutate(df, tmp = delta_00 / pi_00 * g1) |> pull(tmp) |> sum())

    g_hat <- g_hat / nrow(df)

  } else {
    stop("Only class == 'basic' has been implemented.")
  }
}

```

<!---# Compare with: Best case in monotone data --->
## Monotone Case

# Variance Estimation

# Conclusion

Did it work?

### 2024-02-05 src/R/data_generation.R

@@ -1,0 +1,80 @@
+# Title: data_generation.R
+# Created by: Caleb Leedy
+# Created on: February 03, 2024
+# Purpose: This file contains the code for all of the functions used to generate
+# data within simulation studies. I needed to store them here so that they could
+# be reused.
+
+#' A data generation function to create a simple random sample of equally sized
+#' segments of X_1, X_2, and Y. There variables are correlated but all normal.
+#'
+#' @param n_obs_seg - An integer greater than zero. This is the number of
+#' observations in each segment
+#' @param mu - The true mean of Y
+#' @param rho - The correlation between (X_1 - X_2) and (Y - X_1).
+#'
+#' @details We always observe X1 but X2 and Y are only observed depending on the
+#' value of \delta_{ij}. If i = 1 then X2 is observed. If j = 1 then Y is
+#' observed.
+gen_simple_data <- function(n_obs_seg, mu, rho) {
+
+  # This code is modified from the case where we had X, Y_1, and Y_2. It only
+  # changes at the end to X_1, X_2, and Y.
+  # A_11
+  x <- rnorm(n_obs_seg)
+  e_mat <- MASS::mvrnorm(n_obs_seg, c(0, 0), matrix(c(1, rho, rho, 1), nrow = 2))
+  e1 <- e_mat[, 1]
+  e2 <- e_mat[, 2]
+  y1 <- x + e1
+  y2 <- mu + x + e2
+
+  df_11 <- 
+    tibble(X1 = x, X2 = y1, Y = y2,
+           delta_00 = 0, delta_10 = 0, delta_01 = 0, delta_11 = 1,
+           pi_00 = 0.25, pi_10 = 0.25, pi_01 = 0.25, pi_11 = 0.25)
+  # A_10
+  x <- rnorm(n_obs_seg)
+  e_mat <- MASS::mvrnorm(n_obs_seg, c(0, 0), matrix(c(1, rho, rho, 1), nrow = 2))
+  e1 <- e_mat[, 1]
+  e2 <- e_mat[, 2]
+  y1 <- x + e1
+  y2 <- mu + x + e2
+
+  df_10 <- 
[…]

### 2024-02-05 src/explore/20240202-estcovmat.qmd
Finished initial simulation for estimating the covariance matrix.

@@
 
-## Proposals
+## Proposal
 
-### Proposal 1: REML
+Instead of using the known covariance matrix $V$, we estimate it instead.
+Suppose that we observe variables $Z = (X_{m_1}, Y_{m_2})$ with an objective of
+estimating $\theta = E[g(Y_{m_2})$ for some known function $g$ such that the
+variables $Z$ are subject to missingness. We assume that there are $R$ unique
+combinations of observed variables including a fully observed case. We can index
+the combinations of observed variables by $r$ and assume that the fully observed
+case occurs at $r = 1$. Let $G_r(Z)$ be the variables that are observed at a
+particular value of $r$. We can choose a sequence of functions $f_1, \dots, f_K$
+that we want to estimate. Each $f_k$ is assumed to be a function of a subset of
+$Z$ and it makes sense to assume (since these are chosen by the analyst) that
+there is at least one $f_k$ for each combination of observed variables $G_r(Z)$.
+Let $A_k$ be the sets of observed variable combinations that can be evaluate by
+the function $f_k$. If $f_k$ can be evaluated by the observed variables
+$G_r(Z)$, this will consist of all of the combinations of variables
+$r'$ such that $G_r(Z) \subseteq G_r'(Z)$. Because we assume that $G_1(Z) = Z$,
+$A_k$ is always non-empty.
 
-One potential solution would be to directly estimate $V$ and use some sort of
-$\hat V$. We can do this with a REML estimate.
+To estimate the covariance matrix $V$, we estimate the covariance between 
+$f_{k_1}$ and $f_{k_2}$ directly by computing the estimated covariance on 
+$A_{k_1} \cap A_{k_2}$.
 
+<!---# This did not work because the orthogonalization created variables that we
+needed to impute
 ### Proposal 2: Orthogonal g-functions
@@
 computing the variance of $\tilde g_k$ within $A_k$.
+--->
 
@@
 
-# Run from src/
+# Run from src/ or src/explore/
 source("R/data_generation.R")
@@
 
-ortho_gls <- function(df, class = "basic") {
-
-  # For class == 'basic', we consider the functions f_k to be originally,
-  # f1 = X1, f2 = X2, f3 = Y.
+est_var_gls <- function(df, class = "basic") {
 
@@
 
-    # Get functions g_k
-    g1 <- df$X1 / var(df$X1)
-    df <- mutate(df, g1 = g1)
-    df_1 <- filter(df, delta_11 == 1 | delta_10 == 1)
-    t2 <- lm(X2 ~ g1, data = df_1)
-    g2 <- df_1$X2 - t2$coefficients[2] * df_1$g1
-    g2 <- g2 / var(g2)
-    df_1 <- mutate(df_1, g2 = g2)
-    df_11 <- filter(df_1, delta_11 == 1)
-    t3 <- lm(Y ~ X1 + X2, data = df_11)
-    g3 <- 
-      df_11$Y - t3$coefficients[2] * df_11$g1 - t3$coefficients[3] * df_11$g2
-    g3 <- g3 / var(g3)
+    z_mat <- matrix(c(1, 0, 0,
+                      0, 1, 0,
+                      0, 0, 1,
+                      1, 0, 0,
+                      0, 1, 0,
+                      1, 0, 0,
+                      0, 0, 1,
+                      1, 0, 0), ncol = 3, byrow = TRUE)
 
     g_hat <- 
-      c(mutate(df, tmp = delta_11 / pi_11 * g1) |> pull(tmp) |> sum(),
-        mutate(df_1, tmp = delta_11 / pi_11 * g2) |> pull(tmp) |> sum(),
-        mutate(df_11, tmp = delta_11 / pi_11 * g3) |> pull(tmp) |> sum(),
-        mutate(df, tmp = delta_10 / pi_10 * g1) |> pull(tmp) |> sum(),
-        mutate(df_1, tmp = delta_10 / pi_10 * g2) |> pull(tmp) |> sum(),
-        mutate(df, tmp = delta_01 / pi_01 * g1) |> pull(tmp) |> sum(),
-        # ERROR: This does not work!
-        mutate(df_11, tmp = delta_01 / pi_01 * g3) |> pull(tmp) |> sum(),
-        mutate(df, tmp = delta_00 / pi_00 * g1) |> pull(tmp) |> sum())
+      c(mutate(df, tmp = delta_11 / pi_11 * X1) |> pull(tmp) |> mean(),
+        mutate(df, tmp = delta_11 / pi_11 * X2) |> pull(tmp) |> mean(),
+        mutate(df, tmp = delta_11 / pi_11 * Y) |> pull(tmp) |> mean(),
+        mutate(df, tmp = delta_10 / pi_10 * X1) |> pull(tmp) |> mean(),
+        mutate(df, tmp = delta_10 / pi_10 * X2) |> pull(tmp) |> mean(),
+        mutate(df, tmp = delta_01 / pi_01 * X1) |> pull(tmp) |> mean(),
+        mutate(df, tmp = delta_01 / pi_01 * Y) |> pull(tmp) |> mean(),
+        mutate(df, tmp = delta_00 / pi_00 * X1) |> pull(tmp) |> mean())
 
-    g_hat <- g_hat / nrow(df)
+    # Estimate Covariance Matrix
+    df_1p <- filter(df, delta_11 == 1 | delta_10 == 1)
+    df_p1 <- filter(df, delta_11 == 1 | delta_01 == 1)
+    df_11 <- filter(df, delta_11 == 1)
+    v_x1 <- var(df$X1)
+    v_x2 <- var(df_1p$X2)
+    v_y <- var(df_p1$Y)
+    cov_x1x2 <- cov(df_1p$X1, df_1p$X2)
+    cov_x1y <- cov(df_p1$X1, df_p1$Y)
+    cov_x2y <- cov(df_11$X2, df_11$Y)
 
+    v_11 <- 1 / df$pi_11[1] *
+      matrix(c(v_x1, cov_x1x2, cov_x1y,
+               cov_x1x2, v_x2, cov_x2y,
+               cov_x1y, cov_x2y, v_y), nrow = 3)
+
+    v_10 <- 1 / df$pi_10[1] * matrix(c(v_x1, cov_x1x2, cov_x1x2, v_x2), nrow = 2)
+    v_01 <- 1 / df$pi_01[1] * matrix(c(v_x1, cov_x1y, cov_x1y, v_y), nrow = 2)
+    v_00 <- 1 / df$pi_00[1] * v_x1
+
+    v_mat <- matrix(rep(0, 64), ncol = 8)
+    v_mat[1:3, 1:3] <- v_11
+    v_mat[4:5, 4:5] <- v_10
+    v_mat[6:7, 6:7] <- v_01
+    v_mat[8, 8] <- v_00
+
+    v_mat <- v_mat / nrow(df)
+
+    ghat <- MASS::ginv(t(z_mat) %*% MASS::ginv(v_mat) %*% z_mat) %*% 
+      t(z_mat) %*% MASS::ginv(v_mat) %*% g_hat
+
+    return(as.numeric(t(c(0, 0, 1) %*% ghat)))
   } else {
@@
 
-<!---# Compare with: Best case in monotone data --->
-## Monotone Case
+```{r}
+#| label: MCMC
+#| echo: false
+#| tbl-cap: |
+#|   Results from simulations study with independent equally sized segments
+#|   $A_{11}$, $A_{10}$, $A_{01}$, and $A_{00}$ all of size $n = 250$. In this
+#|   simulation we have the true mean of $Y_2$ equal to $\mu = 5$ and the covariance
+#|   between $e_1$ and $e_2$ is $\rho = 0.5$. The goal is to estimate 
+#|   $E[Y_2] = \mu$. For the GLS estimation, we use the
+#|   estimated covariance matrix $\hat V$ with 
+#|   f-functions $f_1 = X_1$, $f_2 = X_2$ and $f_3 = Y$.
 
-# Variance Estimation
+clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
+registerDoParallel(clust)
+
+mc_theta <-
+  foreach(iter = 1:B,
+          .options.RNG = 1,
+          .packages = c("dplyr", "stringr")) %dorng% {
+
+    # Generate Data
+    df <- gen_simple_data(obs_seg, mu = true_mu, rho = cov_e1e2)
+
+    # Estimators
+    oracle <- mean(df$Y)
+    cc <- df |>
+      filter(delta_11 == 1) |>
+      pull(Y) |>
+      mean()
+    ipw <- df |>
+      mutate(tmp = delta_11 / pi_11 * Y) |>
+      pull(tmp) |>
+      mean()
+
+    gls <- gls_est(df, rho = cov_e1e2)
+    gls_ev <- est_var_gls(df)
+
+    return(tibble(Oracle = oracle,
+                  CC = cc,
+                  IPW = ipw,
+                  GLS = gls,
+                  GLSEstVar = gls_ev
+    ))
+
+  } |>
+  bind_rows()
+
+stopCluster(clust)
+
+# Analyze Results
+mc_theta |>
+  summarize(
+    across(everything(),
+    list(Bias = ~mean(.) - true_mu,
+         SD = ~sd(.),
+         Tstat = ~(mean(.) - true_mu) / sqrt(var(.) / B)),
+    .names = "{.fn}_{.col}")) |> 
+  tidyr::pivot_longer(cols = everything(),
+                      names_to = c(".value", "Algorithm"),
+                      names_pattern = "(.*)_(.*)") |>
+  mutate(Pval = pt(-abs(Tstat), df = B)) |>
+  knitr::kable(digits = 3)
+
+```
+
+<!---# Compare with: Best case in monotone data 
+## Monotone Case--->
+
+# Optimal choice of g-functions
+
+TODO for the next set of simulations.
 
@@
 
-Did it work?
+Did it work? Yes, estimating the covariance matrix did not produce a loss in
+efficiency.

### 2024-02-07 src/explore/20240202-estcovmat.qmd
Finished CSSM presentation.

@@
 ---
-title: "Estimating the Covariance Matrix"
+title: "Using GLS and Estimating the Covariance Matrix"
 author:

### 2024-02-07 src/explore/20240206-covmat.qmd
Finished CSSM presentation.

---
title: "Estimating the Covariance Matrix"
author:
  - name: Caleb Leedy
date: today
date-format: "D MMMM YYYY"
format:
  pdf:
    include-in-header: 
      - "latex_header.tex"
---

```{r}
#| label: global variables
#| echo: false

B <- 2000
obs_seg <- 250
true_mu <- 5
cov_e1e2 <- 0.5

```

# Summary

In this document, we do the following:

1. Verify numerically the true covariance matrix in the simple
normal model.

# Problem

In `20240202-estcovmat.pdf` and `20240202-gls_sim.pdf`, we ran simulations in
which we used the *true* covariance matrix to get the best estimator.
Unfortunately, further analysis suggests that the covariance matrix used
previously was slightly incorrect. This report tests to see if this new result
is correct.

The previously accepted true variance estimate^[See Page 219 of notes] was the
following:

$$\Cov(\hat f_1, \hat f_2) = N^{-1} \pi^{-1} I(\delta_1 = \delta_2) \Cov(f_1,
f_2) $$

and the new variance estimate^[See Page 221 of notes] is

$$\Cov(\hat f_1, \hat f_2) = N^{-1}(I(\delta_1 \neq \delta_2) \Cov(f_1, f_2) +
\pi^{-1} I(\delta_1 = \delta_2) \Cov(f_1, f_2)).$$

The only difference between the estimates is the covariance terms between
segments. Are these independent or not?

# Simulation

```{r}
#| label: setup
#| echo: false
#| warning: false
#| message: false

library(dplyr)

library(doParallel)
library(doRNG)
library(parallelly)

# Run from src/ or src/explore/
source("R/data_generation.R")

```

```{r}
#| label: MCMC
#| echo: false

clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
registerDoParallel(clust)

mc_g <-
  foreach(iter = 1:B,
          .options.RNG = 1,
          .packages = c("dplyr", "stringr")) %dorng% {

# Generate Data
  df <- gen_simple_data(obs_seg, mu = true_mu, rho = cov_e1e2)
  g_hat <- 
    c(mutate(df, tmp = delta_11 / pi_11 * X1) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_11 / pi_11 * X2) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_11 / pi_11 * Y) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_10 / pi_10 * X1) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_10 / pi_10 * X2) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_01 / pi_01 * X1) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_01 / pi_01 * Y) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_00 / pi_00 * X1) |> pull(tmp) |> mean())

  return(tibble(g1 = g_hat[1],
                g2 = g_hat[2],
                g3 = g_hat[3],
                g4 = g_hat[4],
                g5 = g_hat[5],
                g6 = g_hat[6],
                g7 = g_hat[7],
                g8 = g_hat[8]))
  } |>
  bind_rows()

stopCluster(clust)
cov_vec <- 
  c(cov(mc_g$g1, mc_g$g1),
    cov(mc_g$g1, mc_g$g2),
    cov(mc_g$g1, mc_g$g3),
    cov(mc_g$g1, mc_g$g4),
    cov(mc_g$g1, mc_g$g5),
    cov(mc_g$g1, mc_g$g6),
    cov(mc_g$g1, mc_g$g7),
    cov(mc_g$g1, mc_g$g8),
    cov(mc_g$g2, mc_g$g2),
    cov(mc_g$g2, mc_g$g3),
    cov(mc_g$g2, mc_g$g4),
    cov(mc_g$g2, mc_g$g5),
    cov(mc_g$g2, mc_g$g6),
    cov(mc_g$g2, mc_g$g7),
    cov(mc_g$g2, mc_g$g8),
    cov(mc_g$g3, mc_g$g3),
    cov(mc_g$g3, mc_g$g4),
    cov(mc_g$g3, mc_g$g5),
    cov(mc_g$g3, mc_g$g6),
    cov(mc_g$g3, mc_g$g7),
    cov(mc_g$g3, mc_g$g8),
    cov(mc_g$g4, mc_g$g4),
    cov(mc_g$g4, mc_g$g5),
    cov(mc_g$g4, mc_g$g6),
    cov(mc_g$g4, mc_g$g7),
    cov(mc_g$g4, mc_g$g8),
    cov(mc_g$g5, mc_g$g5),
    cov(mc_g$g5, mc_g$g6),
    cov(mc_g$g5, mc_g$g7),
    cov(mc_g$g5, mc_g$g8),
    cov(mc_g$g6, mc_g$g6),
    cov(mc_g$g6, mc_g$g7),
    cov(mc_g$g6, mc_g$g8),
    cov(mc_g$g7, mc_g$g7),
    cov(mc_g$g7, mc_g$g8),
    cov(mc_g$g8, mc_g$g8))

cov_f1f2 <- 
  c(1, 1, 1, 1, 1, 1, 1, 1,
    2, 1 + cov_e1e2, 1, 2, 1, 1 + cov_e1e2, 1,
    2, 1, 1 + cov_e1e2, 1, 2, 1,
    1, 1, 1, 1, 1,
    2, 1, 1 + cov_e1e2, 1,
    1, 1, 1,
    2, 1,
    1)
exp1_cov <- 
  c(4, 4, 4, 0, 0, 0, 0, 0,
    4, 4, 0, 0, 0, 0, 0,
    4, 0, 0, 0, 0, 0,
    4, 4, 0, 0, 0,
    4, 0, 0, 0,
    4, 4, 0,
    4, 0,
    4) / (4 * obs_seg)
exp2_cov <- 
  c(4, 4, 4, 1, 1, 1, 1, 1,
    4, 4, 1, 1, 1, 1, 1,
    4, 1, 1, 1, 1, 1,
    4, 4, 1, 1, 1,
    4, 1, 1, 1,
    4, 4, 1,
    4, 1,
    4) / (4 * obs_seg)

exp1 <- exp1_cov * cov_f1f2
exp2 <- exp2_cov * cov_f1f2

tibble(Actual = cov_vec, Exp1 = exp1, Exp2 = exp2) |>
  knitr::kable(digits = 5)

```

# Conclusion

Which is correct? It looks like the first derivation is correct. This does make
sense as I view the different segments as independent. The question is then what
mistake did I make in my own analysis?

### 2024-02-07 src/paper_sims/cheng2022handling.qmd
Finished CSSM presentation.

---
title: "Simulation Study from Cheng et al. (2022)"
author: "Caleb Leedy"
date: "February 6, 2024"
date-format: "D MMMM YYYY"
format:
  pdf:
    include-in-header: 
      - "latex_header.tex"
bibliography: references.bib
---

```{r}
#| label: parameters

B <- 10
n_obs <- 2000

```

# Summary

This document does the following:

1. Recreates the simulation study for the Single Primary Variable simulation 
   from @cheng2022handling,
2. Recreates the simulation study for Multiple Variables from
   @cheng2022handling.

# Simulation

```{r}
#| label: libraries
#| echo: false

library(dplyr)
library(purrr)
library(knitr)
library(doParallel)
library(doRNG)
library(parallelly)

```

```{r}
#| label: generate data
#| echo: false

gen_data <- function(n_obs) {
  # Steps:
  # 1. Identify missingness pattern
  # 2. Generate data

  # I am going to use SRS of Equal sizes 
  seg_size <- n_obs / 8

  mu_1 <- 1
  mu_2 <- c(1, -1)
  mu_3 <- c(0, -1, -1)

  sig_1 <- 1
  sig_2 <- matrix(c(1, 0.5, 0.5, 1), nrow = 2)
  sig_3 <- matrix(c(1, 0.5, 0.5, 0.5, 1, 0.5, 0.5, 0.5, 1), nrow = 3)
  # In terms of (Y1, Y2, Y3) we have the following segments:
  # 111, 011, 101, 001, 110, 010, 100, 000

  # 111
  y3y1y2 <- MASS::mvrnorm(seg_size, mu = mu_3, Sigma = sig_3)

  # 011
  y1y2 <- MASS::mvrnorm(seg_size, mu = mu_2, Sigma = sig_2)

  # 101
  y3y2 <- MASS::mvrnorm(seg_size, mu = mu_2, Sigma = sig_2)

  # 001
  y2 <- rnorm(seg_size, mean = mu_1, sd = sig_1)

  # 110
  y3y1 <- MASS::mvrnorm(seg_size, mu = mu_2, Sigma = sig_2)
  # 010
  y1 <- rnorm(seg_size, mean = mu_1, sd = sig_1)

  # 100
  y3 <- rnorm(seg_size, mean = mu_1, sd = sig_1)

  # 000
  # nothing

  t111 <- tibble(Y3 = y3y1y2[, 1], Y1 = y3y1y2[, 2], Y2 = y3y1y2[, 3])
  t011 <- tibble(Y3 = NA, Y1 = y1y2[, 1], Y2 = y1y2[, 2])
  t101 <- tibble(Y3 = y3y2[, 1], Y1 = NA, Y2 = y3y2[, 2])
  t001 <- tibble(Y3 = NA, Y1 = NA, Y2 = y2)
  t110 <- tibble(Y3 = y3y1[, 1], Y1 = y3y1[, 2], Y2 = NA)
  t010 <- tibble(Y3 = NA, Y1 = y1, Y2 = NA)
  t100 <- tibble(Y3 = y3, Y1 = NA, Y2 = NA)

  return(bind_rows(t111, t011, t101, t001, t110, t010, t100))
}

```

```{r}
#| label: implementing estimators

ipw_est <- function(df, n_tot) {
  # Steps:
  # 1. Estimate O00, O10, O01, O11
  # 2. Combine

  # 1. Estimate O00, O10, O01, O11
  df_00 <- mutate(df, A = !is.na(Y3))
  mod00 <- glm(A ~ 1, data = df_00, family = "binomial")
  O00_00 <- predict(mod00, filter(df, !is.na(Y3), is.na(Y1), is.na(Y2)))
  O00_10 <- predict(mod00, filter(df, !is.na(Y3), !is.na(Y1), is.na(Y2)))
  O00_01 <- predict(mod00, filter(df, !is.na(Y3), is.na(Y1), !is.na(Y2)))
  O00_11 <- predict(mod00, filter(df, !is.na(Y3), !is.na(Y1), !is.na(Y2)))

  df_10 <- 
    filter(df, !is.na(Y1)) |>
    mutate(A = !is.na(Y3))
  mod10 <- glm(A ~ Y1, data = df_10, family = "binomial")
  O10_10 <- predict(mod10, filter(df, !is.na(Y3), !is.na(Y1), is.na(Y2)))
  O10_11 <- predict(mod10, filter(df, !is.na(Y3), !is.na(Y1), !is.na(Y2)))

  df_01 <- 
    filter(df, !is.na(Y2)) |>
    mutate(A = !is.na(Y3))
  mod01 <- glm(A ~ Y2, data = df_01, family = "binomial")
  O01_01 <- predict(mod01, filter(df, !is.na(Y3), is.na(Y1), !is.na(Y2)))
  O01_11 <- predict(mod01, filter(df, !is.na(Y3), !is.na(Y1), !is.na(Y2)))

  df_11 <- 
    filter(df, !is.na(Y1), !is.na(Y2)) |>
    mutate(A = !is.na(Y3))
  mod11 <- glm(A ~ Y1 + Y2, data = df_11, family = "binomial")
  O11 <- predict(mod11, filter(df, !is.na(Y3), !is.na(Y1), !is.na(Y2)))

  # 2. Combine
  y_111 <- filter(df, !is.na(Y3), !is.na(Y1), !is.na(Y2)) |> pull(Y3)
  y_101 <- filter(df, !is.na(Y3), is.na(Y1), !is.na(Y2)) |> pull(Y3)
  y_110 <- filter(df, !is.na(Y3), !is.na(Y1), is.na(Y2)) |> pull(Y3)
  y_100 <- filter(df, !is.na(Y3), is.na(Y1), is.na(Y2)) |> pull(Y3)

  res <- 
    y_111 * (1 + O11 + O01_11 + O10_11 + O00_11) +
    y_101 * (1 + O01_01 + O00_01) +
    y_110 * (1 + O10_10 + O00_10) +
    y_100 * (1 + O00_00)

  return(sum(res) / n_tot)
}

reg_est <- function(df, n_tot) {
  # Steps:
  # 1. Estimate m00, m10, m01, m11
  # 2. Combine

  # 1. Estimate m00, m10, m01, m11
  df_00 <- filter(df, !is.na(Y3))
  m00 <- mean(df_00$Y3)

  df_10 <- filter(df, !is.na(Y3), !is.na(Y1))
  mod10 <- lm(Y3 ~ Y1, data = df_10)
  m10 <- predict(mod10, data = filter(df, is.na(Y3), !is.na(Y1)))

  df_01 <- filter(df, !is.na(Y3), !is.na(Y2))
  mod01 <- lm(Y3 ~ Y2, data = df_01)
  m01 <- predict(mod01, data = filter(df, is.na(Y3), !is.na(Y2)))

  df_11 <- filter(df, !is.na(Y3), !is.na(Y1), !is.na(Y2))
  mod01 <- lm(Y3 ~ Y1 + Y2, data = df_11)
  m11 <- predict(mod11, data = filter(df, is.na(Y3), !is.na(Y1), !is.na(Y2)))

  # 2. Combine
  obs_y3 <- filter(df, !is.na(Y3)) |> pull(Y3)

  return(sum(obs_y3 + m00 + m10 + m01 + m11) / n_tot)
}

```

```{r}
#| label: MCMC
#| echo: false

clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
registerDoParallel(clust)

mc_theta <-
  foreach(iter = 1:B,
          .options.RNG = 1,
          .packages = c("dplyr")) %dorng% {

    # Generate Data
    df <- gen_data(n_obs)
    ipw <- ipw_est(df, n_obs)
    reg <- reg_est(df, n_obs)

    cc <- filter(df, !is.na(Y3)) |> pull(Y3) |> mean()

    return(tibble(IPW = ipw,
                  REG = reg,
                  CC = cc))

  } |>
  bind_rows()

stopCluster(clust)

# Analyze Results
true_mu <- 89 / 96
mc_theta |>
  summarize(
    across(everything(),
    list(Bias = ~mean(.) - true_mu,
         SD = ~sd(.),
         Tstat = ~(mean(.) - true_mu) / sqrt(var(.) / B)),
    .names = "{.fn}_{.col}")) |> 
  tidyr::pivot_longer(cols = everything(),
                      names_to = c(".value", "Algorithm"),
                      names_pattern = "(.*)_(.*)") |>
  mutate(Pval = pt(-abs(Tstat), df = B)) |>
  knitr::kable(digits = 3)

```

### 2024-02-13 references/literature/papers.qmd
It looks like we can estimate the variance. This holds if we just use X1 and X2 or if we use f(X1) and f(X2).

---
title: "Papers Read"
author: "Caleb Leedy"
date: February 12, 2024
format: pdf
bibliography: references.bib
---

# @zhou2012efficient

Title: An efficient method of estimation for longitudinal surveys with monotone
missing data

Author(s): Ming Zhou and Jae-Kwang Kim

Journal (Year): Biometrika (2012)

Importance: I think that this paper does something very similar to our project
except with monotone data rather than nonmonotone data. The newness in the paper
is that the MAR assumption depends on variables oberseved in the previous time
period.

Innovation: They use a likelihood method for estimating the propensity score. I
am also curious about how they got Equation 9, which they claim is an optimal
estimator.

{{< pagebreak >}}

# References

### 2024-02-13 src/explore/20240208-optimalf.qmd
It looks like we can estimate the variance. This holds if we just use X1 and X2 or if we use f(X1) and f(X2).

---
title: "Finding Optimal f-Functions"
author:
  - name: Caleb Leedy
date: today
date-format: "D MMMM YYYY"
format:
  pdf:
    include-in-header: 
      - "latex_header.tex"
---

# Summary

This document takes the case of observing three variables $(X_1, X_2, Y)$ under
nonmonotone missingness and discusses the optimal function $f_1$, $f_2$, and
$f_3$ under a simple random sampling design.

# Notation and Problem Setup

Consider the case where $Z = (X_1, X_2, Y)$ and we want to estimate the
parameter $\theta = E[Y]$. Suppose that we have a finite population of size $N$.
Instead of observing the entire data set we observe 
the segments in @tbl-vars1. Each of the segments is a simple random sample of
size $n$ and independent from the other segments. Hence this means that the
first order selection probability $\pi_i = \pi = n / N$ for every $i$.

| Segment  | Variables Observed |
| -------  | ------------------ |
| $A_{00}$ | $X_1$              |
| $A_{10}$ | $X_1, X_2$         |
| $A_{01}$ | $X_1, Y$           |
| $A_{11}$ | $X_1, X_2, Y$      | 
: This table identifies which variables are observed in each segment. Since $X_1$
is always observed, the subscript for each segment identifies which of variables
$X_2$ and $Y$ are in the segment based on the position of a 1. {#tbl-vars1}

As an analyst, we can choose functions $f_1, f_2, f_3$ such that $g = Zf + e$
where

\begin{align*}
g_1^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} f_1(x_{1i}) \\
g_2^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} f_2(x_{2i}) \\
g_3^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} f_3(y_{i}) \\
g_1^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} f_1(x_{1i}) \\
g_2^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} f_2(x_{2i}) \\
g_1^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} f_1(x_{1i}) \\
g_3^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} f_2(y_{i}) \\
g_1^{(00)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{00}}{\pi_{00}} f_1(x_{1i}) \\
\end{align*}

where

$$\hat g = 
\begin{bmatrix}
g_1^{(11)} \\
g_2^{(11)} \\
g_3^{(11)} \\
g_1^{(10)} \\
g_2^{(10)} \\
g_1^{(01)} \\
g_3^{(01)} \\
g_1^{(00)} \\
\end{bmatrix},
Z = 
\begin{bmatrix}
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 0 & 1 \\
1 & 0 & 0 \\
0 & 1 & 0 \\
1 & 0 & 0 \\
0 & 0 & 1 \\
1 & 0 & 0 \\
\end{bmatrix},
E[e] = 0,
\text{ and } 
\Var(e) = V.$$

For now assume that $V$ is known. Due to the construction of $g$, we know that

$$V = 
\begin{bmatrix}
V_{11} & 0 & 0 & 0 \\
0 & V_{10} & 0 & 0 \\
0 & 0 & V_{01} & 0 \\
0 & 0 & 0 & V_{00} \\
\end{bmatrix}
$$

where

$$
V_{11} = N^{-1} \pi^{-1} \Cov(f, f), 
V_{10} = N^{-1} \pi^{-1} 
\begin{bmatrix}
\Cov(f_1, f_1) & \Cov(f_1, f_2) \\ \Cov(f_1, f_2) & \Cov(f_2, f_2) 
\end{bmatrix},
$$

$$
V_{01} = N^{-1} \pi^{-1} 
\begin{bmatrix}
\Cov(f_1, f_1) & \Cov(f_1, f_3) \\ \Cov(f_1, f_3) & \Cov(f_3, f_3) 
\end{bmatrix}, \text{ and }
V_{00} = N^{-1} \pi^{-1} \Cov(f_3, f_3).
$$

Previously, we have shown that the optimal estimator is the GLS estimator $\hat
f = (Z'V^{-1}Z)^{-1} Z'V^{-1}g$ and by linear model theory 
$\Var(\hat f) = (Z'V^{-1}Z)^{-1}$ since $\Var(g) = V$ by construction.

```{python}
#| echo: false

import sympy

s1, s2, s3 = sympy.symbols('\sigma_1 \sigma_2 \sigma_3')
p12, p13, p23 = sympy.symbols('\\rho_{12} \\rho_{13} \\rho_{23}')

v_11 = sympy.Matrix([
  [s1 * s1, p12 * s1 * s2, p13 * s1 * s3],
  [p12 * s1 * s2, s2 * s2, p23 * s2 * s3],
  [p13 * s1 * s3, p23 * s2 * s3, s3 * s3]
])
v_10 = sympy.Matrix([
  [s1 * s1, p12 * s1 * s2],
  [p12 * s1 * s2, s2 * s2]
])
v_01 = sympy.Matrix([
  [s1 * s1, p13 * s1 * s3],
  [p13 * s1 * s3, s3 * s3]
])

v_00 = sympy.Matrix([s1 * s1])

v_mat = diag(v_11, v_10, v_01, v_00)

z_mat = sympy.Matrix([
  [1, 0, 0],
  [0, 1, 0],
  [0, 0, 1],
  [1, 0, 0],
  [0, 1, 0],
  [1, 0, 0],
  [0, 0, 1],
  [1, 0, 0]
])
```

### 2024-02-13 src/explore/20240213-optcoefs.qmd
It looks like we can estimate the variance. This holds if we just use X1 and X2 or if we use f(X1) and f(X2).

--- 2024-02-05 src/explore/20240202-estcovmat.qmd
@@
 ---
-title: "Estimating the Covariance Matrix"
+title: "Finding the Optimal Coefficient Forms"
 author:
@@
 #| echo: false
-B <- 1000
 
+B <- 2000
 obs_seg <- 250
@@
 
-In this document, we
+In this document, we do the following:
 
-1. Propose a new technique for estimating a covariance matrix, and 
-2. Show via simulation studies that it works.
+1. Vary the functional forms of the coefficients, 
+2. Check to ensure that our estimated variance is still accurate.
 
-# Problem and Proposal
+# Idea
 
-## Problem
+In `20240202-estcovmat.qmd`, we tested an estimated covariance matrix. The results
+suggest that estimating the covariance matrix suffers little penalty. In this
+report, we want to see if choosing different functional forms of $f_1$ and $f_2$ 
+(that depend on the data) cause the covariance estimation to be biased.
 
-Previously, have a model $\hat g = Z g + e$ where $\hat g$ is defined by
-
-\begin{align*}
-g_1^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_1(x_i) \\
-g_2^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_2(y_{1i}) \\
-g_3^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_3(y_{2i}) \\
-g_1^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} g_1(x_i) \\
-g_2^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} g_2(y_{1i}) \\
-g_1^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} g_1(x_i) \\
-g_3^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} g_2(y_{2i}) \\
-g_1^{(00)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{00}}{\pi_{00}} g_1(x_i) \\
-\end{align*}
-
-and
-
-$$\hat g = 
-\begin{bmatrix}
-g_1^{(11)} \\
-g_2^{(11)} \\
-g_3^{(11)} \\
-g_1^{(10)} \\
-g_2^{(10)} \\
-g_1^{(01)} \\
-g_3^{(01)} \\
-g_1^{(00)} \\
-\end{bmatrix},
-Z = 
-\begin{bmatrix}
-1 & 0 & 0 \\
-0 & 1 & 0 \\
-0 & 0 & 1 \\
-1 & 0 & 0 \\
-0 & 1 & 0 \\
-1 & 0 & 0 \\
-0 & 0 & 1 \\
-1 & 0 & 0 \\
-\end{bmatrix},
-E[e] = 0,
-\text{ and } 
-\Var(e) = n^{-1}
-\begin{bmatrix}
-V_{11} & 0 & 0 & 0 \\
-0 & V_{10} & 0 & 0 \\
-0 & 0 & V_{01} & 0 \\
-0 & 0 & 0 & V_{00} \\
-\end{bmatrix}.
-$$
-
-We also have
-
-$$
-V_{11} = 
-\begin{bmatrix}
-\frac{1}{\pi_{11}} E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{11}} E[g_1g_2] -
-E[g_1]E[g_2] & \frac{1}{\pi_{11}} E[g_1g_3] - E[g_1]E[g_3] \\
-\frac{1}{\pi_{11}} E[g_1g_2] - E[g_1]E[g_2] & \frac{1}{\pi_{11}} E[g_2^2] -
-E[g_2]^2 & \frac{1}{\pi_{11}} E[g_2g_3] - E[g_2]E[g_3] \\
-\frac{1}{\pi_{11}} E[g_1g_3] - E[g_1]E[g_3] & \frac{1}{\pi_{11}} E[g_2g_3] -
-E[g_2]E[g_3] & \frac{1}{\pi_{11}} E[g_3^2] - E[g_3]^2 \\
-\end{bmatrix},
-$$
-$$
-V_{10} = 
-\begin{bmatrix}
-\frac{1}{\pi_{10}}E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{10}}E[g_1g_2] - E[g_1]E[g_2]\\
-\frac{1}{\pi_{10}}E[g_1g_2] - E[g_1]E[g_2] & \frac{1}{\pi_{10}}E[g_2^2] - E[g_2]^2 \\
-\end{bmatrix},
-$$
-$$
-V_{01} = 
-\begin{bmatrix}
-\frac{1}{\pi_{01}}E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{01}}E[g_1g_3] - E[g_1]E[g_3]\\
-\frac{1}{\pi_{01}}E[g_1g_3] - E[g_1]E[g_3] & \frac{1}{\pi_{01}}E[g_3^2] - E[g_3]^2 \\
-\end{bmatrix}, \text{ and }
-V_{00} = 
-\begin{bmatrix}
-\frac{1}{\pi_{00}}E[g_1^2] - E[g_1]^2
-\end{bmatrix}.
-$$
-
-However, this challenge with actually using this model is that we have to use
-the matrix $V$, which we assume to be *known*. In this write up, we will not
-make this assumption.
-
-## Proposal
-
-Instead of using the known covariance matrix $V$, we estimate it instead.
-Suppose that we observe variables $Z = (X_{m_1}, Y_{m_2})$ with an objective of
-estimating $\theta = E[g(Y_{m_2})$ for some known function $g$ such that the
-variables $Z$ are subject to missingness. We assume that there are $R$ unique
-combinations of observed variables including a fully observed case. We can index
-the combinations of observed variables by $r$ and assume that the fully observed
-case occurs at $r = 1$. Let $G_r(Z)$ be the variables that are observed at a
-particular value of $r$. We can choose a sequence of functions $f_1, \dots, f_K$
-that we want to estimate. Each $f_k$ is assumed to be a function of a subset of
-$Z$ and it makes sense to assume (since these are chosen by the analyst) that
-there is at least one $f_k$ for each combination of observed variables $G_r(Z)$.
-Let $A_k$ be the sets of observed variable combinations that can be evaluate by
-the function $f_k$. If $f_k$ can be evaluated by the observed variables
-$G_r(Z)$, this will consist of all of the combinations of variables
-$r'$ such that $G_r(Z) \subseteq G_r'(Z)$. Because we assume that $G_1(Z) = Z$,
-$A_k$ is always non-empty.
-
-To estimate the covariance matrix $V$, we estimate the covariance between 
-$f_{k_1}$ and $f_{k_2}$ directly by computing the estimated covariance on 
-$A_{k_1} \cap A_{k_2}$.
-
-<!---# This did not work because the orthogonalization created variables that we
-needed to impute
-### Proposal 2: Orthogonal g-functions
-
-For this idea we choose functions that after estimation are approximately
-independent from each
-other with a standard variance. If this is true, the covariance matrix $V$ is
-approximately the identity matrix. Not only do we not have to estimate $V$, but
-we do not even need to invert it! I now give a more detailed explanation of this
-method.
-
-Suppose that we observe variables $Z = (X_{m_1}, Y_{m_2})$ with an objective of
-estimating $\theta = E[g(Y_{m_2})$ for some function $g$ such that the
-variables are subject to missingness. We assume that there are $R$ unique
-combinations of observed variables including a fully observed case. We can index
-the combinations of observed variables by $r$ and assume that the fully observed
-case occurs at $r = 1$. Let $G_r(Z)$ be the variables that are observed at a
-particular value of $r$. We can choose a sequence of functions $f_1, \dots, f_K$
-that we want to estimate. Each $f_k$ is assumed to be a function of a subset of
-$Z$ and it makes sense to assume (since these are chosen by the analyst) that
-there is at least one $f_k$ for each combination of observed variables $G_r(Z)$.
-Let $A_k$ be the sets of observed variable combinations that can be evaluate by
-the function $f_k$. If $f_k$ can be evaluated by the observed variables
-$G_r(Z)$, this will consist of all of the combinations of variables
-$r'$ such that $G_r(Z) \subseteq G_r'(Z)$. Because we assume that $G_1(Z) = Z$,
-$A_k$ is always non-empty.
-
-Once we have the original $f_1, \dots, f_K$ we orthogonalize them using a
-Gram-Schmidt process. Let $g_1 = f_1 / \hat \Var(f_1)$. Then for $k > 1$
-we can continue using the sequence,
-
-$$\tilde g_k = f_k - \sum_{i = 1}^k \frac{\hat \Cov(f_k, g_i)}{\hat \Var(f_k)}
-g_i \text{ followed by } g_k = \tilde g_k / \hat \Var(\tilde g_k).$$
-
-However, the parameters for $\tilde g_k$ and $k > 1$ are all regression
-coefficients and the multiplier for $g_k$ is simply computing a variance.
-Then to
-achieve efficiency, we propose estimating the regression coefficients by running
-the coresponding regression on all of the data points in $A_k$ as well as
-computing the variance of $\tilde g_k$ within $A_k$.
---->
+Instead of using $f_1(X_1) = X_1$ and $f_2(X_2) = X_2$ as we did in 
+`20240202-estcovmat.qmd` we want to use $f_1(X_1) = E[Y \mid X_1]$ and 
+$f_2(X_2) = E[Y \mid X_2]$, *and* we need to also test them with estimated
+versions of these functions.
 
 # Simulation
-
-<!---# Compare with: Known V in non-monotone case --->
-## Non-Monotone Case
-
-First, we consider the following non-monotone simulation setup.
-
-\begin{align*}
-  \begin{bmatrix} x_1 \\ e_1 \\ e_2 \end{bmatrix} 
-  &\stackrel{ind}{\sim} N\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
-  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}\right) \\
-  x_2 &= x + e_1 \\
-  y &= \mu + x + e_2 \\
-\end{align*}
-
-This yields outcome variables $X_1$ and $X_2$ that are correlated both with $Y$
-and additionally with each other. To generate the missingness pattern, we select
-`r obs_seg` observations into the four segments independently using simple
-random sampling. The goal of this estimation problem is to estimate $\theta = E[Y]$.
 
@@
 ```{r}
-#| label: proposed estimation
+#| label: estimated variance basic
 #| echo: false
@@
     return(as.numeric(t(c(0, 0, 1) %*% ghat)))
+
     } else {
     stop("Only class == 'basic' has been implemented.")
+  }
+}
+
+```
+
+```{r}
+#| label: estimated variance opt
+#| echo: false
+
+estvar_glsopt <- function(df, class = "opt") {
+
+  if (class == "opt") {
+
+    z_mat <- matrix(c(1, 0, 0,
+                      0, 1, 0,
+                      0, 0, 1,
+                      1, 0, 0,
+                      0, 1, 0,
+                      1, 0, 0,
+                      0, 0, 1,
+                      1, 0, 0), ncol = 3, byrow = TRUE)
+
+    z_mat <- matrix(rep(1, 8), ncol = 1)
+
+    # Get Regression Estimates
+    mx1 <- lm(Y ~ X1, data = filter(df, delta_11 == 1 | delta_01 == 1))
+    mx2 <- lm(Y ~ X2, data = filter(df, delta_11 == 1))
+    df <- df |>
+      mutate(mx1 = predict(mx1, newdata = df)) |>
+      mutate(mx2 = predict(mx2, newdata = df))
+
+    # Put Estimates into g_hat
+    g_hat <- 
+      c(mutate(df, tmp = delta_11 / pi_11 * mx1) |> pull(tmp) |> mean(),
+        mutate(df, tmp = delta_11 / pi_11 * mx2) |> pull(tmp) |> mean(),
+        mutate(df, tmp = delta_11 / pi_11 * Y) |> pull(tmp) |> mean(),
+        mutate(df, tmp = delta_10 / pi_10 * mx1) |> pull(tmp) |> mean(),
+        mutate(df, tmp = delta_10 / pi_10 * mx2) |> pull(tmp) |> mean(),
+        mutate(df, tmp = delta_01 / pi_01 * mx1) |> pull(tmp) |> mean(),
+        mutate(df, tmp = delta_01 / pi_01 * Y) |> pull(tmp) |> mean(),
+        mutate(df, tmp = delta_00 / pi_00 * mx1) |> pull(tmp) |> mean())
+
+    # Estimate Covariance Matrix
+    df_1p <- filter(df, delta_11 == 1 | delta_10 == 1)
+    df_p1 <- filter(df, delta_11 == 1 | delta_01 == 1)
+    df_11 <- filter(df, delta_11 == 1)
+    v_x1 <- var(df$mx1)
+    v_x2 <- var(df_1p$mx2)
+    v_y <- var(df_p1$Y)
+    cov_x1x2 <- cov(df_1p$mx1, df_1p$mx2)
+    cov_x1y <- cov(df_p1$mx1, df_p1$Y)
+    cov_x2y <- cov(df_11$mx2, df_11$Y)
+
+    v_11 <- 1 / df$pi_11[1] *
+      matrix(c(v_x1, cov_x1x2, cov_x1y,
+               cov_x1x2, v_x2, cov_x2y,
+               cov_x1y, cov_x2y, v_y), nrow = 3)
+
+    v_10 <- 1 / df$pi_10[1] * matrix(c(v_x1, cov_x1x2, cov_x1x2, v_x2), nrow = 2)
+    v_01 <- 1 / df$pi_01[1] * matrix(c(v_x1, cov_x1y, cov_x1y, v_y), nrow = 2)
+    v_00 <- 1 / df$pi_00[1] * v_x1
+
+    v_mat <- matrix(rep(0, 64), ncol = 8)
+    v_mat[1:3, 1:3] <- v_11
+    v_mat[4:5, 4:5] <- v_10
+    v_mat[6:7, 6:7] <- v_01
+    v_mat[8, 8] <- v_00
+
+    v_mat <- v_mat / nrow(df)
+
+    ghat <- MASS::ginv(t(z_mat) %*% MASS::ginv(v_mat) %*% z_mat) %*% 
+      t(z_mat) %*% MASS::ginv(v_mat) %*% g_hat
+
+    return(as.numeric(ghat))
+  } else {
+    stop("Only class == 'opt' has been implemented.")
   }
@@
 #|   estimated covariance matrix $\hat V$ with 
-#|   f-functions $f_1 = X_1$, $f_2 = X_2$ and $f_3 = Y$.
+#|   f-functions $f_1 = X_1$, $f_2 = X_2$ and $f_3 = Y$. For the GLS optimal
+#|   estimation we are trying $f_1 = E[Y \mid X_1]$, $f_2 = E[Y \mid X_2]$ and 
+#|   $f_3 = Y$. Each of these expectations are estimated with using linear
+#|   regression. So $E[Y \mid X_i]$ is implemented as a regression estimator
+#|
 
@@
     gls_ev <- est_var_gls(df)
+    gls_opt <- estvar_glsopt(df)
 
@@
                   GLS = gls,
-                  GLSEstVar = gls_ev
+                  GLSEstVar = gls_ev,
+                  GLSOpt = gls_opt
     ))
@@
 ```
-
-<!---# Compare with: Best case in monotone data 
-## Monotone Case--->
-
-# Optimal choice of g-functions
-
-TODO for the next set of simulations.
-
 # Conclusion
 
-Did it work? Yes, estimating the covariance matrix did not produce a loss in
-efficiency.
+Is the covariance estimation still working? (By working we mean causing the
+estimator to be unbiased.)
+
+* It looks like estimating the variance is not causing additional bias in the
+  estimator. This is great. 
+* We may, however, need to check more complicated simulation setups because the
+  simpicity of the current setup might make it look better than it actually is.
+* One suprising thing is that the "optimal" estimator is not better than the
+  other estimators. This could be because of the simplicity of the setup, but we
+  may need to check it out.

### 2024-02-14 src/paper_sims/cheng2022handling.qmd
Updated simulation to correct problems with implementation. I think that the basic IPW and regression estimators are correct.

@@
 
-B <- 10
+B <- 1000
 n_obs <- 2000
@@
   # 1. Estimate O00, O10, O01, O11
-  df_00 <- mutate(df, A = !is.na(Y3))
+  df_00 <- df |>
+    filter(!is.na(Y3)) |>
+    add_row(Y3 = rep(NA, 250), Y1 = rep(NA, 250), Y2 = rep(NA, 250)) |>
+    mutate(A = !is.na(Y3)) 
   mod00 <- glm(A ~ 1, data = df_00, family = "binomial")
-  O00_00 <- predict(mod00, filter(df, !is.na(Y3), is.na(Y1), is.na(Y2)))
-  O00_10 <- predict(mod00, filter(df, !is.na(Y3), !is.na(Y1), is.na(Y2)))
-  O00_01 <- predict(mod00, filter(df, !is.na(Y3), is.na(Y1), !is.na(Y2)))
-  O00_11 <- predict(mod00, filter(df, !is.na(Y3), !is.na(Y1), !is.na(Y2)))
+  O00_00 <- exp(-predict(mod00, filter(df, !is.na(Y3), is.na(Y1), is.na(Y2))))
+  O00_10 <- exp(-predict(mod00, filter(df, !is.na(Y3), !is.na(Y1), is.na(Y2))))
+  O00_01 <- exp(-predict(mod00, filter(df, !is.na(Y3), is.na(Y1), !is.na(Y2))))
+  O00_11 <- exp(-predict(mod00, filter(df, !is.na(Y3), !is.na(Y1), !is.na(Y2))))
 
   df_10 <- 
-    filter(df, !is.na(Y1)) |>
+    filter(df, (!is.na(Y1) & !is.na(Y3)) | (is.na(Y3) & !is.na(Y1) & is.na(Y2))) |>
     mutate(A = !is.na(Y3))
   mod10 <- glm(A ~ Y1, data = df_10, family = "binomial")
-  O10_10 <- predict(mod10, filter(df, !is.na(Y3), !is.na(Y1), is.na(Y2)))
-  O10_11 <- predict(mod10, filter(df, !is.na(Y3), !is.na(Y1), !is.na(Y2)))
+  O10_10 <- exp(-predict(mod10, filter(df, !is.na(Y3), !is.na(Y1), is.na(Y2))))
+  O10_11 <- exp(-predict(mod10, filter(df, !is.na(Y3), !is.na(Y1), !is.na(Y2))))
 
   df_01 <- 
-    filter(df, !is.na(Y2)) |>
+    filter(df, (!is.na(Y2) & !is.na(Y3)) | (is.na(Y3) & is.na(Y1) & !is.na(Y2))) |>
     mutate(A = !is.na(Y3))
   mod01 <- glm(A ~ Y2, data = df_01, family = "binomial")
-  O01_01 <- predict(mod01, filter(df, !is.na(Y3), is.na(Y1), !is.na(Y2)))
-  O01_11 <- predict(mod01, filter(df, !is.na(Y3), !is.na(Y1), !is.na(Y2)))
+  O01_01 <- exp(-predict(mod01, filter(df, !is.na(Y3), is.na(Y1), !is.na(Y2))))
+  O01_11 <- exp(-predict(mod01, filter(df, !is.na(Y3), !is.na(Y1), !is.na(Y2))))
 
@@
   mod11 <- glm(A ~ Y1 + Y2, data = df_11, family = "binomial")
-  O11 <- predict(mod11, filter(df, !is.na(Y3), !is.na(Y1), !is.na(Y2)))
+  O11 <- exp(-predict(mod11, filter(df, !is.na(Y3), !is.na(Y1), !is.na(Y2))))
 
@@
   df_00 <- filter(df, !is.na(Y3))
-  m00 <- mean(df_00$Y3)
+  mod00 <- mean(df_00$Y3)
+  m00 <- mod00 * rep(1, n_tot / 8)
 
@@
   mod10 <- lm(Y3 ~ Y1, data = df_10)
-  m10 <- predict(mod10, data = filter(df, is.na(Y3), !is.na(Y1)))
+  m10 <- predict(mod10, newdata = filter(df, is.na(Y3), !is.na(Y1), is.na(Y2)))
 
@@
   mod01 <- lm(Y3 ~ Y2, data = df_01)
-  m01 <- predict(mod01, data = filter(df, is.na(Y3), !is.na(Y2)))
+  m01 <- predict(mod01, newdata = filter(df, is.na(Y3), is.na(Y1), !is.na(Y2)))
 
   df_11 <- filter(df, !is.na(Y3), !is.na(Y1), !is.na(Y2))
-  mod01 <- lm(Y3 ~ Y1 + Y2, data = df_11)
-  m11 <- predict(mod11, data = filter(df, is.na(Y3), !is.na(Y1), !is.na(Y2)))
+  mod11 <- lm(Y3 ~ Y1 + Y2, data = df_11)
+  m11 <- predict(mod11, newdata = filter(df, is.na(Y3), !is.na(Y1), !is.na(Y2)))
 
@@
 
-  return(sum(obs_y3 + m00 + m10 + m01 + m11) / n_tot)
+  return((sum(obs_y3) + sum(m00) + sum(m10) + sum(m01) + sum(m11)) / n_tot)
 }
@@
 ```
+
+# Discussion
+
+* While I think that I implemented this correctly, the biggest problem with this
+  paper is the fact that their ACCMV assumption is almost equivalent to a MCAR
+  assumption.

### 2024-02-17 src/explore/20240215-sympy.qmd
Updated simulation with true and estimated marginal and conditional GLS. See 20240216-optest.qmd.

---
title: "SymPy Calculations"
author:
  - name: Caleb Leedy
date: today
date-format: "D MMMM YYYY"
format:
  pdf:
    classoption: [landscape]
    include-in-header: 
      - "latex_header.tex"
---

# Summary

This document records computations that I think would be easier to make in SymPy
than on paper. For now it does the following:

1. Compute the optimal values of $\theta$, $\eta_1$ and $\eta_2$ for the simple
nonmonotone GLS estimator (Notes: Pages 237-243+).

\tiny
```{python}
#| label: sympyres1
#| echo: false
#| output: asis

import sympy

# Steps:
# 1. Solve for eta_1
# 2. Solve for eta_2
# 3. Solve for theta

# We need variables for what we observe:
x1_1, x2_1, x3_1, x4_1 = sympy.symbols('X^1_1 X^2_1 X^3_1 X^4_1')
x1_2, x2_2, y1, y2 = sympy.symbols('X^1_2 X^2_2 Y^1 Y^2')

# Variances 
vx1_1, vx2_1 = sympy.symbols('V(X^1_{1}) V(X^2_{1})')
vx3_1, vx4_1 = sympy.symbols('V(X^3_{1}) V(X^4_{1})')
vx1_2, vx2_2, vy1, vy2 = sympy.symbols('V(X^1_{2}) V(X^2_{2}) V(Y^{1}) V(Y^{2})')

# Covariances
cx11x12, cx21x22 = sympy.symbols('C(X^1_{1}X^1_{2}) C(X^2_{1}X^2_{2})')
cx11y1 = sympy.symbols('C(X^1_{1}Y^{1})')
cx12y1, cx31y2 = sympy.symbols('C(X^1_{2}Y^{1}) C(X^3_{1}Y^{2})')

# 1. Solve for eta_1
x_1_mat = sympy.Matrix([[x1_1], [x2_1], [x3_1], [x4_1]])
vq1 = sympy.diag(vx1_1, vx2_1, vx3_1, vx4_1)
denom_q1 = (sympy.ones(1, 4) * (vq1**(-1)) * sympy.ones(4, 1))**(-1)
numer_q1 = sympy.ones(1, 4) * (vq1**(-1)) * x_1_mat
eta1 = (denom_q1 * numer_q1)[0, 0]

eta1 = sympy.symbols('\eta_1')

# 2. Solve for eta_2
x_2_mat = sympy.Matrix([[x1_2], [x2_2]])
q2adj_mat = sympy.Matrix([
  [cx11x12 / vx1_1 * (x1_1 - eta1)],
  [cx21x22 / vx2_1 * (x2_1 - eta1)]
])
vq2 = sympy.diag(vx1_2 - cx11x12**2 / vx1_1, vx2_2 - cx21x22**2 / vx2_1)
denom_q2 = (sympy.ones(1, 2) * (vq2**(-1)) * sympy.ones(2, 1))**(-1)
numer_q2 = sympy.ones(1, 2) * (vq2**(-1)) * (x_2_mat - q2adj_mat)
eta2 = (denom_q2 * numer_q2)[0, 0]

# 3. Solve for theta
qv11 = sympy.diag(vx1_1, vx2_1, vx3_1, vx4_1, vx1_2, vx2_2)
qv11[0, 4] = cx11x12
qv11[4, 0] = cx11x12
qv11[1, 5] = cx21x22
qv11[5, 1] = cx21x22

qv11_inv = qv11**(-1)
qv21 = sympy.Matrix([
  [cx11y1, 0, 0, 0, cx12y1, 0],
  [0, 0, cx31y2, 0, 0, 0]
])
qv22 = sympy.diag(vy1, vy2)
qv = qv22 - qv21 * qv11_inv * qv21.T

denom_q = (sympy.ones(1, 2) * qv**(-1) * sympy.ones(2, 1))**(-1)
diff_mat = sympy.Matrix([
  [x1_1 - eta1],
  [x2_1 - eta1],
  [x3_1 - eta1],
  [x4_1 - eta1],
  [x1_2 - eta2],
  [x2_2 - eta2]
])
y_mat = sympy.Matrix([[y1], [y2]])
numer_q = sympy.ones(1, 2) * qv**(-1) * (y_mat - qv21 * qv11_inv * diff_mat)
theta = (denom_q * numer_q)[0, 0]
# sympy.simplify(theta)

print(sympy.latex(sympy.simplify(theta), itex = True, mode = "inline"))

```

### 2024-02-17 src/explore/20240216-optest.qmd
Updated simulation with true and estimated marginal and conditional GLS. See 20240216-optest.qmd.

---
title: "Optimal Estimation with Nuisance Parameters"
author: "Caleb Leedy"
date: "February 16, 2024"
format: 
  pdf:
    include-in-header: 
      - "latex_header.tex"
bibliography: references.bib
---

```{r}
#| label: global variables
#| echo: false
B <- 1000

obs_seg <- 250
true_mu <- 5
cov_e1e2 <- 0.5

```

# Summary

In this document, we do the following:

1. Estimate the optimal $\theta$ with a known covariance function,
2. Estimate the optimal $\theta$ with an estimated covariance function, and
3. Compare these results to other estimates.

# Simulation Setup

We consider a normal model for $(X_1, X_2, Y)$ where $X_1 \sim N(0, 1)$, and 
$e_1, e_2$ are from a bivariate normal distribution each with mean zero,
variance one, and covariance $\rho = $ `r cov_e1e2`. Then $X_2 = X_1 + e_1$ and
$Y = \mu + X_1 + e_1$ where $\mu = $ `{r} true_mu`. Each observation is
independent of the others. Instead of always observing all three variables, we
observe them with missingness. The observed segments can be viewed in @tbl-segs.
Each segment is independent of each other and of the same size. We can view each
segment as an independent simple random sample of the superpopulation of size
$n = $ `{r} obs_seg`.

| Segment | Observed Variables |
| --------| -------------------|
| $A_{11}$| $X_1, X_2, Y$      |
| $A_{10}$| $X_1, X_2$         |
| $A_{01}$| $X_1, Y$           |
| $A_{00}$| $X_1$              |
: This table shows the segment names with each variable that they contain. {#tbl-segs}

# Estimation

The goal of the simulation is to estimate $\theta = E[Y]$. We also estimate
several nuisance parameters $\eta_1 = E[X_1]$ and $\eta_2 = E[X_2]$. All of
these parameters can be estimated using GLS estimation. The model that we
propose is the following. Consider the model, for 
$g = (\eta_1, \eta_2, \theta)'$

$$ \hat f = Zg + e \text{ where } $$

\begin{align*}
f_1^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_1(x_{1i}) \\
f_2^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_2(x_{2i}) \\
f_3^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_3(y_{i}) \\
f_1^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} g_1(x_{1i}) \\
f_2^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} g_2(x_{2i}) \\
f_1^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} g_1(x_{1i}) \\
f_3^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} g_2(y_{i}) \\
f_1^{(00)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{00}}{\pi_{00}} g_1(x_{1i}) \\
\end{align*}

and

$$\hat f = 
\begin{bmatrix}
f_1^{(11)} \\
f_2^{(11)} \\
f_3^{(11)} \\
f_1^{(10)} \\
f_2^{(10)} \\
f_1^{(01)} \\
f_3^{(01)} \\
f_1^{(00)} \\
\end{bmatrix},
Z = 
\begin{bmatrix}
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 0 & 1 \\
1 & 0 & 0 \\
0 & 1 & 0 \\
1 & 0 & 0 \\
0 & 0 & 1 \\
1 & 0 & 0 \\
\end{bmatrix},
E[e] = 0,
\text{ and } 
\Var(e) = n^{-1}
\begin{bmatrix}
V_{11} & 0 & 0 & 0 \\
0 & V_{10} & 0 & 0 \\
0 & 0 & V_{01} & 0 \\
0 & 0 & 0 & V_{00} \\
\end{bmatrix}.
$$

We also have

$$
V_{11} = 
\begin{bmatrix}
\frac{1}{\pi_{11}} E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{11}} E[g_1g_2] -
E[g_1]E[g_2] & \frac{1}{\pi_{11}} E[g_1g_3] - E[g_1]E[g_3] \\
\frac{1}{\pi_{11}} E[g_1g_2] - E[g_1]E[g_2] & \frac{1}{\pi_{11}} E[g_2^2] -
E[g_2]^2 & \frac{1}{\pi_{11}} E[g_2g_3] - E[g_2]E[g_3] \\
\frac{1}{\pi_{11}} E[g_1g_3] - E[g_1]E[g_3] & \frac{1}{\pi_{11}} E[g_2g_3] -
E[g_2]E[g_3] & \frac{1}{\pi_{11}} E[g_3^2] - E[g_3]^2 \\
\end{bmatrix},
$$
$$
V_{10} = 
\begin{bmatrix}
\frac{1}{\pi_{10}}E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{10}}E[g_1g_2] - E[g_1]E[g_2]\\
\frac{1}{\pi_{10}}E[g_1g_2] - E[g_1]E[g_2] & \frac{1}{\pi_{10}}E[g_2^2] - E[g_2]^2 \\
\end{bmatrix},
$$
$$
V_{01} = 
\begin{bmatrix}
\frac{1}{\pi_{01}}E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{01}}E[g_1g_3] - E[g_1]E[g_3]\\
\frac{1}{\pi_{01}}E[g_1g_3] - E[g_1]E[g_3] & \frac{1}{\pi_{01}}E[g_3^2] - E[g_3]^2 \\
\end{bmatrix}, \text{ and }
V_{00} = 
\begin{bmatrix}
\frac{1}{\pi_{00}}E[g_1^2] - E[g_1]^2
\end{bmatrix}.
$$

Using GLS directly, yields an estimator
$\hat g = (\hat \eta_1, \hat \eta_2, \hat \theta)$. Previously, we have used 
the estimator^[The "M" is short for marginal.]

$$ \hat \theta_{M} := [0, 0, 1] \cdot \hat g. $$

However, we can do better if we assume that the distribution of $\hat g$ is a
joint multivariate normal model and compute $\hat \theta \mid \hat \eta_1, \hat
\eta_2$. We can represent this model^[The "C" is for conditional.] as 
$\hat \theta_C := E[\hat \theta \mid \hat \eta]$. For each of these models we
use the true covariance matrix and the estimated covariance matrix. For a full
discussion of the derivation of $\hat \theta_C$ see Appendix A.

# Simulation Study

```{r}
#| label: setup
#| echo: false
#| warning: false
#| message: false

library(dplyr)

library(doParallel)
library(doRNG)
library(parallelly)

# Run from src/ or src/explore/
source("R/data_generation.R")

```

```{r}
#| label: gls M, true
#| echo: false

#' This function is also used in 20240202-gls_sim.qmd. I have modified the code
#' to account for the change from X, Y1, Y2 to X1, X2, Y.
gls_m_true <- function(df, class = "basic", rho) {

  if (class == "basic") {
    z_mat <- matrix(c(1, 0, 0,
                      0, 1, 0,
                      0, 0, 1,
                      1, 0, 0,
                      0, 1, 0,
                      1, 0, 0,
                      0, 0, 1,
                      1, 0, 0), ncol = 3, byrow = TRUE)

    g_hat <- 
      c(mutate(df, tmp = delta_11 / pi_11 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * X2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_00 / pi_00 * X1) |> pull(tmp) |> mean())

    v_11 <- 1 / df$pi_11[1] *
      matrix(c(1, 1, 1,
               1, 2, 1 + rho,
               1, 1 + rho, 2), nrow = 3)
    v_10 <- 1 / df$pi_10[1] * matrix(c(1, 1, 1, 2), nrow = 2)
    v_01 <- 1 / df$pi_01[1] * matrix(c(1, 1, 1, 2), nrow = 2)
    v_00 <- 1 / df$pi_00[1]

    v_mat <- matrix(rep(0, 64), ncol = 8)
    v_mat[1:3, 1:3] <- v_11
    v_mat[4:5, 4:5] <- v_10
    v_mat[6:7, 6:7] <- v_01
    v_mat[8, 8] <- v_00

    v_mat <- v_mat / nrow(df)

    ghat <- MASS::ginv(t(z_mat) %*% MASS::ginv(v_mat) %*% z_mat) %*% 
      t(z_mat) %*% MASS::ginv(v_mat) %*% g_hat

    return(as.numeric(t(c(0, 0, 1) %*% ghat)))

  } else {
    stop("We have only 1 class: 'basic'.")
  }
}

```

```{r}
#| label: GLS M est
#| echo: false

gls_m_est <- function(df, class = "basic") {

  if (class == "basic") {

    z_mat <- matrix(c(1, 0, 0,
                      0, 1, 0,
                      0, 0, 1,
                      1, 0, 0,
                      0, 1, 0,
                      1, 0, 0,
                      0, 0, 1,
                      1, 0, 0), ncol = 3, byrow = TRUE)

    g_hat <- 
      c(mutate(df, tmp = delta_11 / pi_11 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * X2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_00 / pi_00 * X1) |> pull(tmp) |> mean())

    # Estimate Covariance Matrix
    df_1p <- filter(df, delta_11 == 1 | delta_10 == 1)
    df_p1 <- filter(df, delta_11 == 1 | delta_01 == 1)
    df_11 <- filter(df, delta_11 == 1)
    v_x1 <- var(df$X1)
    v_x2 <- var(df_1p$X2)
    v_y <- var(df_p1$Y)
    cov_x1x2 <- cov(df_1p$X1, df_1p$X2)
    cov_x1y <- cov(df_p1$X1, df_p1$Y)
    cov_x2y <- cov(df_11$X2, df_11$Y)

    v_11 <- 1 / df$pi_11[1] *
      matrix(c(v_x1, cov_x1x2, cov_x1y,
               cov_x1x2, v_x2, cov_x2y,
               cov_x1y, cov_x2y, v_y), nrow = 3)

    v_10 <- 1 / df$pi_10[1] * matrix(c(v_x1, cov_x1x2, cov_x1x2, v_x2), nrow = 2)
    v_01 <- 1 / df$pi_01[1] * matrix(c(v_x1, cov_x1y, cov_x1y, v_y), nrow = 2)
    v_00 <- 1 / df$pi_00[1] * v_x1

    v_mat <- matrix(rep(0, 64), ncol = 8)
    v_mat[1:3, 1:3] <- v_11
    v_mat[4:5, 4:5] <- v_10
    v_mat[6:7, 6:7] <- v_01
    v_mat[8, 8] <- v_00

    v_mat <- v_mat / nrow(df)

    ghat <- MASS::ginv(t(z_mat) %*% MASS::ginv(v_mat) %*% z_mat) %*% 
      t(z_mat) %*% MASS::ginv(v_mat) %*% g_hat

    return(as.numeric(t(c(0, 0, 1) %*% ghat)))
  } else {
    stop("Only class == 'basic' has been implemented.")
  }
}

```

```{r}
#| label: GLS C true
#| echo: false

gls_c <- function(df, case = "true") {

  # Steps:
  # 1. Get fhat
  # 2. Get covariance matrix
  # 3. Solve for eta1
  # 4. Solve for eta2
  # 5. Solve for theta

  # 1. Get fhat
  fhat <- 
    c(mutate(df, tmp = delta_11 / pi_11 * X1) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_10 / pi_10 * X1) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_01 / pi_01 * X1) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_00 / pi_00 * X1) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_11 / pi_11 * X2) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_10 / pi_10 * X2) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_11 / pi_11 * Y) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_01 / pi_01 * Y) |> pull(tmp) |> mean()
    )

  # 2. Get covariance matrix

  if (case == "true") {

    Ex1 <- 0
    Ex2 <- 0
    Ey <- true_mu
    Ex12 <- 1
    Ex22 <- 2
    Ey2 <- true_mu^2 + 2
    Ex1x2 <- 1
    Ex1y <- 1
    Ex2y <- 1 + cov_e1e2

  } else if (case == "est") {

    Ex1 <- mean(df$X1)
    Ex2 <- 
      mutate(df, tmp = (delta_10 + delta_11) * X2) %>%
      filter(tmp != 0) %>% 
      pull(tmp) %>% 
      mean()
    Ey <-
      mutate(df, tmp = (delta_01 + delta_11) * Y) %>%
      filter(tmp != 0) %>% 
      pull(tmp) %>% 
      mean()
    Ex12 <- mean(df$X1^2)
    Ex22 <-
      mutate(df, tmp = (delta_10 + delta_11) * X2^2) %>%
      filter(tmp != 0) %>% 
      pull(tmp) %>% 
      mean()
    Ey2 <-
      mutate(df, tmp = (delta_01 + delta_11) * Y^2) %>%
      filter(tmp != 0) %>% 
      pull(tmp) %>% 
      mean()
    Ex1x2 <-
      mutate(df, tmp = (delta_10 + delta_11) * X1 * X2) %>%
      filter(tmp != 0) %>% 
      pull(tmp) %>% 
      mean()
    Ex1y <-
      mutate(df, tmp = (delta_01 + delta_11) * X1 * Y) %>%
      filter(tmp != 0) %>% 
      pull(tmp) %>% 
      mean()
    Ex2y <-
      mutate(df, tmp = (delta_11) * X2 * Y) %>%
      filter(tmp != 0) %>% 
      pull(tmp) %>% 
      mean()

  } else {
    stop("We have only implemented case == 'true' and 'est'.")
  }

  big_v <- matrix(0, nrow = 8, ncol = 8)
  big_v[1:4, 1:4] <- 
    diag(c(1 / df$pi_11[1] * Ex12 - Ex1^2,
           1 / df$pi_10[1] * Ex12 - Ex1^2,
           1 / df$pi_01[1] * Ex12 - Ex1^2,
           1 / df$pi_00[1] * Ex12 - Ex1^2))
  big_v[5:6, 5:6] <- 
  diag(c(1 / df$pi_11[1] * Ex22 - Ex2^2,
         1 / df$pi_10[1] * Ex22 - Ex2^2))
  big_v[7:8, 7:8] <- 
    diag(c(1 / df$pi_11[1] * Ey2 - Ey^2,
           1 / df$pi_01[1] * Ey2 - Ey^2))

  big_v[1, 5] <- 1 / df$pi_11[1] * Ex1x2 - Ex1 * Ex2
  big_v[1, 7] <- 1 / df$pi_11[1] * Ex1y - Ex1 * Ey
  big_v[5, 7] <- 1 / df$pi_11[1] * Ex2y - Ex2 * Ey
  big_v[2, 6] <- 1 / df$pi_10[1] * Ex1x2 - Ex1 * Ex2
  big_v[3, 8] <- 1 / df$pi_01[1] *  Ex1y - Ex1 * Ey

  big_v[5, 1] <- big_v[1, 5]
  big_v[7, 1] <- big_v[1, 7]
  big_v[7, 5] <- big_v[5, 7]
  big_v[6, 2] <- big_v[2, 6]
  big_v[8, 3] <- big_v[3, 8]

  # 3. Solve for eta1
  x <- t(t(c(1, 1, 1, 1)))
  v <- big_v[1:4, 1:4]
  y <- fhat[1:4]

  eta1 <- 
    MASS::ginv(t(x) %*% MASS::ginv(v) %*% x) %*% t(x) %*% MASS::ginv(v) %*% y

  # 4. Solve for eta2
  x <- t(t(c(1, 1)))
  v <- 
  big_v[5:6, 5:6] - big_v[5:6, 1:4] %*% MASS::ginv(big_v[1:4, 1:4]) %*% big_v[1:4, 5:6]
  y <- 
    fhat[5:6] - big_v[5:6, 1:4] %*% MASS::ginv(big_v[1:4, 1:4]) %*% 
    (fhat[1:4] - c(eta1, eta1, eta1, eta1))

  eta2 <- 
    MASS::ginv(t(x) %*% MASS::ginv(v) %*% x) %*% t(x) %*% MASS::ginv(v) %*% y

  # 5. Solve for theta
  x <- t(t(c(1, 1)))
  v <-
  big_v[7:8, 7:8] - big_v[7:8, 1:6] %*% MASS::ginv(big_v[1:6, 1:6]) %*% big_v[1:6, 7:8]
  y <-
    fhat[7:8] - big_v[7:8, 1:6] %*% MASS::ginv(big_v[1:6, 1:6]) %*% 
    (fhat[1:6] - c(eta1, eta1, eta1, eta1, eta2, eta2))

  theta <- 
    MASS::ginv(t(x) %*% MASS::ginv(v) %*% x) %*% t(x) %*% MASS::ginv(v) %*% y

  return(as.numeric(theta))
}

```

```{r}
#| label: MCMC
#| echo: false
#| tbl-cap: |
#|   Results from simulations study with independent equally sized segments
#|   $A_{11}$, $A_{10}$, $A_{01}$, and $A_{00}$ all of size $n = 250$. In this
#|   simulation we have the true mean of $Y_2$ equal to $\mu = 5$ and the covariance
#|   between $e_1$ and $e_2$ is $\rho = 0.5$. The goal is to estimate 
#|   $E[Y_2] = \mu$. For the estimators with 'est', we use the
#|   estimated covariance matrix $\hat V$ with 
#|   f-functions $f_1 = X_1$, $f_2 = X_2$ and $f_3 = Y$. For the estimators
#|   without we use the true covariance matrix. GLSM is the marginal model while
#|   GLSC is the conditional model.
clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
registerDoParallel(clust)

mc_theta <-
  foreach(iter = 1:B,
          .options.RNG = 1,
          .packages = c("dplyr", "stringr")) %dorng% {

    # Generate Data
    df <- gen_simple_data(obs_seg, mu = true_mu, rho = cov_e1e2)

    # Estimators
    oracle <- mean(df$Y)
    cc <- df |>
      filter(delta_11 == 1) |>
      pull(Y) |>
      mean()
    ipw <- df |>
      mutate(tmp = delta_11 / pi_11 * Y) |>
      pull(tmp) |>
      mean()

    est_m_true <- gls_m_true(df, rho = cov_e1e2)
    est_m_est <- gls_m_est(df)
    est_c_true <- gls_c(df, case = "true")
    est_c_est <- gls_c(df, case = "est")

    return(tibble(Oracle = oracle,
                  CC = cc,
                  IPW = ipw,
                  GLSM = est_m_true,
                  GLSMEst = est_m_est,
                  GLSC = est_c_true,
                  GLSCEst = est_c_est
    ))

  } |>
  bind_rows()

stopCluster(clust)

# Analyze Results
mc_theta |>
  summarize(
    across(everything(),
    list(Bias = ~mean(.) - true_mu,
         SD = ~sd(.),
         Tstat = ~(mean(.) - true_mu) / sqrt(var(.) / B)),
    .names = "{.fn}_{.col}")) |> 
  tidyr::pivot_longer(cols = everything(),
                      names_to = c(".value", "Algorithm"),
                      names_pattern = "(.*)_(.*)") |>
  mutate(Pval = pt(-abs(Tstat), df = B)) |>
  knitr::kable(digits = 3)

```

# Results and Discussion

1. Estimating the covariance in our conditional model cause dependence issues.
2. This particular simulation does not differentiate between the GLS models. We
   might need a different setup to distinguish between the models (such as
   unequal segment size).

# Next Steps

1. Find the minimum variance estimator GLSC as a function of $g$.
2. Estimate the covariance matrix so that it does not interfere with estimation
   of $\hat f$.

{{< pagebreak >}}

# Appendix A: Derivation of $\hat \theta_C$

The derivation in this appendix follows the idea in @zhou2012efficient. The
initial goal is to minimize $Q(\eta_1, \eta_2, \theta)$ where

$$
Q(\eta_1, \eta_2, \theta) = z'V^{-1}z \text{ for }
z = 
\begin{bmatrix}
f_1^{(11)} - \eta_1 \\
f_1^{(10)} - \eta_1 \\
f_1^{(01)} - \eta_1 \\
f_1^{(00)} - \eta_1 \\
f_2^{(11)} - \eta_2 \\
f_2^{(10)} - \eta_2 \\
f_3^{(11)} - \theta \\
f_3^{(01)} - \theta \\
\end{bmatrix} \text{ and }
V = V(z).
$$

We can notice that $Q$ is the kernel of a multivariate normal distribution,
which means that we can decompose it into a marginal and conditional part,

$$ 
\min_{\eta, \theta} Q(\eta_1, \eta_2, \theta) = 
\min_{\eta, \theta} (Q(\eta) + Q(\theta \mid \eta)).
$$

Since $Q(\eta)$ is the kernal of a normal distribution we can apply the same
technique to get

$$ 
\min_{\eta, \theta} Q(\eta_1, \eta_2, \theta) = 
\min_{\eta, \theta} \{Q(\eta_1) + Q(\eta_2 \mid \eta_1) + Q(\theta \mid \eta_1,
\eta_2)\}.
$$

Then we can solve these sequentially.

1. We find $\hat \eta_1 = \argmin Q(\eta_1)$.
2. We can substitute this back into $Q(\eta_2 \mid \eta_1)$ to get the 
  optimal $\hat \eta_2$ of 
  $$ \hat \eta_2 = \argmin Q(\eta_2 \mid \hat \eta_1). $$
3. We can use the results of the previous two steps to get,
  $$ \hat \theta = \argmin Q(\theta \mid \hat \eta_1, \hat \eta_2).$$

Unfortunately, the nonmonotone case does not have any nice simplifications like
how the monotone case reduces to a regression estimator. This is due to the fact
that the marginal and conditional distributions are generally multivariate
normal distributions instead of a one-dimensional distribution. This causes the
result to be weighted averages instead of difference estimators.

## Finding $\hat \eta_1$

We start with the fact that

$$
\hat \eta_1 = \argmin Q(\eta_1) \text{ where } 
Q(\eta_1) = z_{1:4}' V^{-1}_{1:4, 1:4} z_{1:4}.
$$

This means that for the simulation study in this report, since each segment is
independent,

$$
Q(\eta_1) = 
\begin{bmatrix}
\hat f_1^{(11)} - \eta_1 \\
\hat f_1^{(10)} - \eta_1 \\
\hat f_1^{(01)} - \eta_1 \\
\hat f_1^{(00)} - \eta_1 \\
\end{bmatrix}'
\begin{bmatrix}
(\Var(\hat f_1^{(11)}))^{-1} & 0 & 0 & 0 \\
0 & (\Var(\hat f_1^{(10)}))^{-1} & 0 & 0 \\
0 & 0 & (\Var(\hat f_1^{(01)}))^{-1} & 0 \\
0 & 0 & 0 & (\Var(\hat f_1^{(00)}))^{-1} \\
\end{bmatrix}
\begin{bmatrix}
\hat f_1^{(11)} - \eta_1 \\
\hat f_1^{(10)} - \eta_1 \\
\hat f_1^{(01)} - \eta_1 \\
\hat f_1^{(00)} - \eta_1 \\
\end{bmatrix}.
$$

Since the dual of this problem is a GLS estimator with a diagonal variance
matrix, $\hat \eta_1$ is the inverse-variance weighted average of all of the
individual estimators,

$$
\hat \eta_1 = 
\frac{\frac{\hat f_1^{(11)}}{\Var(\hat f_1^{(11)})} + 
\frac{\hat f_1^{(10)}}{\Var(\hat f_1^{(10)})} + 
\frac{\hat f_1^{(01)}}{\Var(\hat f_1^{(01)})} + 
\frac{\hat f_1^{(00)}}{\Var(\hat f_1^{(00)})}}{
\frac{1}{\Var(\hat f_1^{(11)})} + 
\frac{1}{\Var(\hat f_1^{(10)})} + 
\frac{1}{\Var(\hat f_1^{(01)})} + 
\frac{1}{\Var(\hat f_1^{(00)})}
}.
$$

## Finding $\hat \eta_2$

Similar to the previous step, we can notice that $Q(\eta_2 \mid \eta_1)$ is the
kernel of the conditional distribution

$$
\begin{bmatrix}
\hat f_2^{11} \\
\hat f_2^{10}
\end{bmatrix} 
\sim N \left(
\begin{bmatrix}
\eta_2 \\ \eta_2
\end{bmatrix}  + 
\Sigma_{\eta_2, \eta_1} \Sigma_{\eta_1}^{-1}
\begin{bmatrix}
\hat f_1^{(11)} - \eta_1 \\
\hat f_1^{(10)} - \eta_1 \\
\hat f_1^{(01)} - \eta_1 \\
\hat f_1^{(00)} - \eta_1 \\
\end{bmatrix},
\Sigma_{\eta_2} - \Sigma_{\eta_2, \eta_1} \Sigma_{\eta_1}^{-1} \Sigma_{\eta_1, \eta_2}
\right).
$$

Since we have $\Sigma_{\eta_1} = V_{1:4, 1:4}$ and

$$
\Sigma_{\eta_2, \eta_1} = 
\begin{bmatrix}
\Cov(\hat f_1^{(11)}, \hat f_2^{(11)}) & 0 & 0 & 0 \\
0 & \Cov(\hat f_1^{(10)}, \hat f_2^{(10)}) & 0 & 0 \\
\end{bmatrix},
$$

we have

$$
\Sigma_{\eta_2} - \Sigma_{\eta_2, \eta_1} \Sigma_{\eta_1}^{-1} 
\Sigma_{\eta_2, \eta_1} = 
\begin{bmatrix}
\Var(\hat f_2^{(11)}) - \frac{\Cov(\hat f_1^{(11)}, \hat{f}_2^{(11)})^2}{\Var(\hat f_1^{(11)})} & 0 \\
0 & \Var(\hat f_2^{(10)}) - \frac{\Cov(\hat f_1^{(10)}, \hat{f}_2^{(10)})^2}{\Var(\hat f_1^{(10)})} \\
\end{bmatrix}.
$$

This means that

$$
\footnotesize
\hat \eta_2 =
\left(\begin{bmatrix} 1 & 1 \end{bmatrix} 
(\Sigma_{\eta_2} - \Sigma_{\eta_2, \eta_1} \Sigma_{\eta_1}^{-1}\Sigma_{\eta_1, \eta_2})^{-1} 
\begin{bmatrix} 1 \\ 1 \end{bmatrix} 
\right)^{-1}
\begin{bmatrix} 1 & 1 \end{bmatrix} 
(\Sigma_{\eta_2} - \Sigma_{\eta_2, \eta_1} \Sigma_{\eta_1}^{-1}\Sigma_{\eta_1, \eta_2})^{-1} 
\begin{bmatrix} 
\hat f_2^{(11)} - \frac{\Cov(\hat f_2^{(11)}, \hat f_1^{(11)})}{\Var(\hat f_1^{(11)})} (\hat f_1^{(11)} -
\hat \eta_1) \\
\hat f_2^{(10)} - \frac{\Cov(\hat f_2^{(10)}, \hat f_1^{(10)})}{\Var(\hat f_1^{(10)})} (\hat f_1^{(10)} -
\hat \eta_1) \\
\end{bmatrix}.
$$

## Finding $\hat \theta$

Finally, to find $\hat \theta$ we use a similar process as shown in the previous
step. This time it is used to compute $Q(\theta \mid \hat \eta_1, \hat \eta_2)$.
We can notice that $Q(\theta \mid \hat \eta)$ is the kernel of a conditional
multivariate normal distribution,

$$
\begin{bmatrix}
\hat f_3^{(11)} \\
\hat f_3^{(01)}
\end{bmatrix} 
\sim N \left(
\begin{bmatrix}
\theta \\ \theta
\end{bmatrix}  + 
\Sigma_{\theta, \eta} \Sigma_{\eta}^{-1}
\begin{bmatrix}
\hat f_1^{(11)} - \eta_1 \\
\hat f_1^{(10)} - \eta_1 \\
\hat f_1^{(01)} - \eta_1 \\
\hat f_1^{(00)} - \eta_1 \\
\hat f_2^{(11)} - \eta_2 \\
\hat f_2^{(10)} - \eta_2 \\
\end{bmatrix},
\Sigma_{\theta} - \Sigma_{\theta, \eta} \Sigma_{\eta}^{-1} \Sigma_{\eta, \theta}
\right).
$$

In this case we have,

$$
\Sigma_{\theta} = 
\begin{bmatrix}
V(\hat f_3^{(11)}) & 0 \\
0 & V(\hat f_3^{(10)})
\end{bmatrix},
\Sigma_{\theta, \eta} = 
\begin{bmatrix}
\Cov(\hat f_3^{(11)}, \hat f_1^{(11)}) & 0 & 0 & 0 & \Cov(\hat f_3^{(11)}, \hat
f_2^{(10)}) & 0 \\
0 & 0 & \Cov(\hat f_3^{(01)}, \hat f_1^{(01)}) & 0 & 0 & 0
\end{bmatrix},
$$

and,

$$
\Sigma_{\eta} = 
\begin{bmatrix}
V(\hat f_1^{(11)}) & 0 & 0 & 0 & \Cov(\hat f_1^{(11)}, \hat f_2^{(11)}) & 0 \\
0 & V(\hat f_1^{(10)}) & 0 & 0 & 0 & \Cov(\hat f_1^{(10)}, \hat f_2^{(10)}) \\
0 & 0 & V(\hat f_1^{(01)}) & 0 & 0 & 0 \\
0 & 0 & 0 & V(\hat f_1^{(00)}) & 0 & 0 \\
\Cov(\hat f_1^{(11)}, \hat f_2^{(11)}) & 0 & 0 & 0 & V(\hat f_2^{(11)}) & 0 \\
0 & \Cov(\hat f_1^{(10)}, \hat f_2^{(10)}) & 0 & 0 & 0 & V(\hat f_2^{(10)}) \\
\end{bmatrix}.
$$

Hence,

$$
\footnotesize
\hat \theta =
\left(\begin{bmatrix} 1 & 1 \end{bmatrix} 
(\Sigma_{\theta} - \Sigma_{\theta, \eta} \Sigma_{\eta}^{-1}\Sigma_{\eta, \theta})^{-1} 
\begin{bmatrix} 1 \\ 1 \end{bmatrix} 
\right)^{-1}
\begin{bmatrix} 1 & 1 \end{bmatrix} 
(\Sigma_{\theta} - \Sigma_{\theta, \eta} \Sigma_{\eta}^{-1}\Sigma_{\eta, \theta})^{-1} 
\left(
\begin{bmatrix}
\hat f_3^{(11)} \\
\hat f_3^{(01)}
\end{bmatrix}
-\Sigma_{\theta, \eta} \Sigma_{\eta}^{-1}
\begin{bmatrix}
\hat f_1^{(11)} - \hat \eta_1 \\
\hat f_1^{(10)} - \hat \eta_1 \\
\hat f_1^{(01)} - \hat \eta_1 \\
\hat f_1^{(00)} - \hat \eta_1 \\
\hat f_2^{(11)} - \hat \eta_2 \\
\hat f_2^{(10)} - \hat \eta_2 \\
\end{bmatrix}
\right).
$$

While this might still be slightly unclear, I have been able to write down the
closed form solution to this estimator, which is given below in @fig-sympy as a
function of $\eta_1$. (Including $\hat \eta_1$ made it too big for the page.)

![A closed form solution to $\hat \theta$.](sympy_opttheta.png){#fig-sympy}

{{< pagebreak >}}

# References

### 2024-02-19 references/literature/papers.qmd
Added REML papers.

@@
 
+# @patterson1971recovery
+
+Title: Recovery of inter-block information when block sizes are unequal
+
+Author(s): H.D. Patterson and R. Thompson
+
+Journal (Year): Biometrika (1971)
+
+Importance: This describes REML for unbalanced designs.
+
+Innovation: They decompose the likelihood into error constrasts and other. (These
+are the $S$ and $Q$ matrices.)
+
+# @harville1977maximum
+
+Title: Maximum Likelihood Approaches to Variance Component Estimation and to
+Related Problems
+
+Author(s): David A. Harville
+
+Journal (Year): JASA (1977)
+
+Importance: This paper summarizes the use of ML and REML methods and focuses on
+their use and computation.
+
+Innovation: This is a good summary of the problems that REML solves.
+
+Title:
+
+Author(s):
+
+Journal (Year):
+
+Importance:
+
+Innovation:
+
+Title:
+
+Author(s):
+
+Journal (Year):
+
+Importance:
+
+Innovation:
+
+Title:
+
+Author(s):
+
+Journal (Year):
+
+Importance:
+
+Innovation:
+
 {{< pagebreak >}}

### 2024-02-19 src/explore/20240216-optest.qmd
Added REML papers.

@@
 several nuisance parameters $\eta_1 = E[X_1]$ and $\eta_2 = E[X_2]$. All of
-these parameters can be estimated using GLS estimation. The model that we
-propose is the following. Consider the model, for 
+these parameters can be estimated using GLS estimation.
+
+## The Marginal GLS Model
+
+The model that we propose is the following. Consider the model, for 
 $g = (\eta_1, \eta_2, \theta)'$
@@
 
+## The Conditional GLS Model
+
 However, we can do better if we assume that the distribution of $\hat g$ is a
@@
 \eta_2$. We can represent this model^[The "C" is for conditional.] as 
-$\hat \theta_C := E[\hat \theta \mid \hat \eta]$. For each of these models we
-use the true covariance matrix and the estimated covariance matrix. For a full
-discussion of the derivation of $\hat \theta_C$ see Appendix A.
+$\hat \theta_C := E[\hat \theta \mid \hat \eta]$. In this representation, 
+the initial goal is to minimize $Q(\eta_1, \eta_2, \theta)$ where
+
+$$
+Q(\eta_1, \eta_2, \theta) = z'V^{-1}z \text{ for }
+z = 
+\begin{bmatrix}
+f_1^{(11)} - \eta_1 \\
+f_1^{(10)} - \eta_1 \\
+f_1^{(01)} - \eta_1 \\
+f_1^{(00)} - \eta_1 \\
+f_2^{(11)} - \eta_2 \\
+f_2^{(10)} - \eta_2 \\
+f_3^{(11)} - \theta \\
+f_3^{(01)} - \theta \\
+\end{bmatrix} \text{ and }
+V = V(z).
+$$
+
+More details about how we solve this model in in Appendix A.
+For each of these models we use the true covariance matrix and the estimated
+covariance matrix.
 
@@
 
+Comment on the second bullet point. Currently, I estimate the covariance matrix
+by estimating the corresponding expected values separately using all of the
+available data. See Appendix B for more details. For example, to estimate 
+$E[X_1 X_2]$ we use
+
+$$\hat E[X_1 X_2] = 
+\frac{\sum_{i = 1}^n (\delta_{10} + \delta_{11})(x_{1i}x_{2i})}{\sum_{i = 1}^n
+\delta_{10} + \delta_{11}}.
+$$
+
+I think that this may be overcome using REML. I will be working on this.
+
 {{< pagebreak >}}
@@
 
+# Appendix B: Computing the Conditional Covariance Matrix
+
+For the simulation study in this report the covariance matrix for the
+conditional model $\hat \theta_C$. Since each segment is considered to be
+independent, using the current estimator $\hat \theta_C$ we only need to compute
+the variance of each $\hat f_j$ and the following covariances: 
+$\Cov(\hat f_1^{(11)}, \hat f_2^{(11)})$,
+$\Cov(\hat f_1^{(11)}, \hat f_3^{(11)})$,
+$\Cov(\hat f_2^{(11)}, \hat f_3^{(11)})$,
+$\Cov(\hat f_1^{(10)}, \hat f_2^{(10)})$, and 
+$\Cov(\hat f_1^{(01)}, \hat f_3^{(01)})$.
+
+The variances of each $\hat f_j$ are relatively straightforward. As an example,
+I will derive $\Var(\hat f_1^{(11)})$.
+
+$$
+\Var(\hat f_1^{(11)}) = n^{-1}(\pi_{11}^{-1} E[X_1^2] - E[X_1]^2).
+$$
+
+For other values of $\hat f_j$ we simply use the correct $\pi_{ij}$ and variable
+from $Z = (X_1, X_2, Y)$. In a similar manner, the covariance 
+$\Cov(\hat f_1^{(11)}, \hat f_2^{(11)})$, can be computed as
+
+$$
+\Cov(\hat f_1^{(11)}, \hat f_2^{(11)}) =
+n^{-1}(\pi_{11}^{-1} E[X_1 X_2] - E[X_1]E[X_2]).
+$$
+
+For the other covariances, we can substitute the corresponding $\pi_{ij}$ and
+other variables from $Z$.
+
+We using the true covariance, I use the population covariance that I know from
+the construction of the problem. To estimate the covariance matrix, I estimate
+each required expectation (such as $E[X_1 X_2]$, $E[X_1]$, and $E[X_2]$ for the
+previous covariance) using all of the available data. This means that
+
+$$
+\hat E[X_1 X_2] = \frac{\sum_{i = 1}^n (\delta_{11} + \delta_{10})x_{1i}x_{2i}}{
+\sum_{i = 1}^n (\delta_{11} + \delta_{10})},
+\hat E[X_1] = n^{-1} \sum_{i = 1}^n x_{1i} \text{ and } 
+\hat E[X_2] = \frac{\sum_{i = 1}^n (\delta_{01} + \delta_{11})x_{2i}}{\sum_{i =
+1}^n \delta_{01} + \delta_{11}}.
+$$
+
+{{< pagebreak >}}
+
 # References

### 2024-03-06 src/explore/20240216-optest.qmd
I showed that the conditional expectation is the optimal function.

@@
   big_v[1:4, 1:4] <- 
-    diag(c(1 / df$pi_11[1] * Ex12 - Ex1^2,
-           1 / df$pi_10[1] * Ex12 - Ex1^2,
-           1 / df$pi_01[1] * Ex12 - Ex1^2,
-           1 / df$pi_00[1] * Ex12 - Ex1^2))
+    diag(c(1 / df$pi_11[1] * (Ex12 - Ex1^2),
+           1 / df$pi_10[1] * (Ex12 - Ex1^2),
+           1 / df$pi_01[1] * (Ex12 - Ex1^2),
+           1 / df$pi_00[1] * (Ex12 - Ex1^2)))
   big_v[5:6, 5:6] <- 
-  diag(c(1 / df$pi_11[1] * Ex22 - Ex2^2,
-         1 / df$pi_10[1] * Ex22 - Ex2^2))
+  diag(c(1 / df$pi_11[1] * (Ex22 - Ex2^2),
+         1 / df$pi_10[1] * (Ex22 - Ex2^2)))
   big_v[7:8, 7:8] <- 
-    diag(c(1 / df$pi_11[1] * Ey2 - Ey^2,
-           1 / df$pi_01[1] * Ey2 - Ey^2))
+    diag(c(1 / df$pi_11[1] * (Ey2 - Ey^2),
+           1 / df$pi_01[1] * (Ey2 - Ey^2)))
 
-  big_v[1, 5] <- 1 / df$pi_11[1] * Ex1x2 - Ex1 * Ex2
-  big_v[1, 7] <- 1 / df$pi_11[1] * Ex1y - Ex1 * Ey
-  big_v[5, 7] <- 1 / df$pi_11[1] * Ex2y - Ex2 * Ey
-  big_v[2, 6] <- 1 / df$pi_10[1] * Ex1x2 - Ex1 * Ex2
-  big_v[3, 8] <- 1 / df$pi_01[1] *  Ex1y - Ex1 * Ey
+  big_v[1, 5] <- 1 / df$pi_11[1] * (Ex1x2 - Ex1 * Ex2)
+  big_v[1, 7] <- 1 / df$pi_11[1] * (Ex1y - Ex1 * Ey)
+  big_v[5, 7] <- 1 / df$pi_11[1] * (Ex2y - Ex2 * Ey)
+  big_v[2, 6] <- 1 / df$pi_10[1] * (Ex1x2 - Ex1 * Ex2)
+  big_v[3, 8] <- 1 / df$pi_01[1] *  (Ex1y - Ex1 * Ey)
 
@@
 $$
-\Var(\hat f_1^{(11)}) = n^{-1}(\pi_{11}^{-1} E[X_1^2] - E[X_1]^2).
+\Var(\hat f_1^{(11)}) = n^{-1}\pi_{11}^{-1} (E[X_1^2] - E[X_1]^2).
 $$
@@
 \Cov(\hat f_1^{(11)}, \hat f_2^{(11)}) =
-n^{-1}(\pi_{11}^{-1} E[X_1 X_2] - E[X_1]E[X_2]).
+n^{-1}\pi_{11}^{-1} (E[X_1 X_2] - E[X_1]E[X_2]).
 $$

### 2024-03-06 src/explore/20240222-sympyvarest.qmd
I showed that the conditional expectation is the optimal function.

---
title: "SymPy Representation for Optimal Regression Estimation"
author: "Caleb Leedy"
date: "February 22, 2024"
format: html
---

# Summary: Part 1

This part prints out the optimal estimator for a general $f(x)$ and for when
$f(x)  = E[Y \mid X]$.

```{python}
#| label: sympyres1
#| echo: false
#| output: asis

import sympy

# We need variables for what we observe:
f11, f10, y = sympy.symbols('f_11 f_10 y')
s11, s10, sy, sfy = sympy.symbols('s11 s10 sy sfy')

v_all = sympy.Matrix([
  [s11, 0, sfy],
  [0, s10, 0],
  [sfy, 0, sy]
])

one3 = sympy.ones(3, 1)
y_vec = sympy.Matrix([
  [f11],
  [f10],
  [y]
])

out = (one3.T * v_all**(-1) * one3)**(-1) * one3.T * v_all**(-1) * y_vec
sympy.simplify(out[0, 0])

sympy.factor(4 - 4 * y - 3 * y**2 + 2 * y**3 + y**4)

```

# Summary: Part 2

We consider the model where $f(X) = E[Y \mid X] + \ell(X)$. We want to get the
variance of $\hat \mu_y$. Then we want to show

1. $V(\hat \mu_y) = a + b \mu_{\ell}$
2. $0 = \argmin_{\sigma_\ell^2} V(\hat \mu_y)$
3. $0 = \argmin_{\sigma_{\ell, y}} V(\hat \mu_y)$

```{python}
#| label: sympyres2
#| echo: false
#| output: asis

import sympy

# We need variables for what we observe:
f11, f10, y = sympy.symbols('f_11 f_10 y')
sl, sx, sy = sympy.symbols('sl sx sy')
pcy, ply = sympy.symbols('pcy ply')

v_all = sympy.Matrix([
  [sx**2 + sl**2 + 2 * ply * sl * sy, 0, sx**2 + ply * sl * sy],
  [0, sx**2 + sl**2 + 2 * ply * sl * sy, 0],
  [sx**2 + ply * sl * sy, 0, sy**2]
])

mod_mat = sympy.Matrix([
  [1, 1],
  [1, 1],
  [1, 0]
])
y_vec = sympy.Matrix([
  [f11],
  [f10],
  [y]
])

out = (mod_mat.T * v_all**(-1) * mod_mat)**(-1) * mod_mat.T * v_all**(-1) * y_vec
sympy.simplify(out[0, 0])
sympy.simplify((mod_mat.T * v_all**(-1) * mod_mat)**(-1))[0, 0]

```

# Summary: Part 3

We consider the model where $f(X) = E[Y \mid X] + \ell(X)$. We want to get the
variance of $\hat \mu_y$. Then we want to show

1. $V(\hat \mu_y) = a + b \mu_{\ell}$
2. $0 = \argmin_{\sigma_\ell^2} V(\hat \mu_y)$
3. $0 = \argmin_{\sigma_{\ell, y}} V(\hat \mu_y)$

Unlike Part 2, in this model we have a true nonmonotone missing problem with the
following parts observed:

| Segment | $X$ | $Y$ |
| ------- | --- | --- |
|$A_{11}$ | ✔️   | ✔️   |
|$A_{10}$ | ✔️   |     |
|$A_{01}$ |     | ✔️   |

We assume that $|A_{11}| = |A_{10}| = |A_{01}|$.

```{python}
#| label: sympyres3
#| echo: false
#| output: asis

import sympy

# We need variables for what we observe:
f11, f10, y11, y01 = sympy.symbols('f_11 f_10 y11 y01')
sl, sy = sympy.symbols('sl sy')
pcy, ply = sympy.symbols('pcy ply')

v_all = sympy.Matrix([
  [pcy**2 * sy**2 + sl**2 + 2 * ply * sl * sy, 0, pcy**2 * sy**2 + ply * sl * sy, 0],
  [0, pcy**2 * sy**2 + sl**2 + 2 * ply * sl * sy, 0, 0],
  [pcy**2 * sy**2 + ply * sl * sy, 0, sy**2, 0],
  [0, 0, 0, sy**2]
])

mod_mat = sympy.Matrix([
  [1, 1],
  [1, 1],
  [1, 0],
  [1, 0]
])
y_vec = sympy.Matrix([
  [f11],
  [f10],
  [y11],
  [y01]
])

xvinvx = mod_mat.T * v_all**(-1) * mod_mat 
xvinvx_inv = sympy.Matrix([
  [xvinvx[1, 1], -xvinvx[0, 1]],
  [-xvinvx[1, 0], xvinvx[0, 0]]
]) * 1 / (xvinvx[0, 0] * xvinvx[1, 1] - xvinvx[1, 0] * xvinvx[0, 1])

out = xvinvx_inv * mod_mat.T * v_all**(-1) * y_vec
out = xvinvx**(-1) * mod_mat.T * v_all**(-1) * y_vec
sympy.simplify(out[0, 0])
sympy.simplify(out[1, 0])
vhatmuy = sympy.simplify(xvinvx_inv[0, 0])
sympy.simplify(sympy.diff(vhatmuy, sl))

alt_vhatmuy = sy**2 * (1 + (2 * pcy**2 * sy**2 + 4 * ply * sl * sy + 2 * sl**2) * (pcy**4 * sy**2 + 2 * pcy**2 * ply * sl * sy - 4 * pcy**2 * sy**2 + ply**2 * sl**2 - 8 * ply * sl * sy - 4 * sl**2)**(-1))

# These are the same
sympy.simplify(vhatmuy - alt_vhatmuy)

sympy.simplify(sympy.diff(alt_vhatmuy, sl))

```

# Summary: Part 4

We consider the model where $f(X_1) = E[Y \mid X_1] + \ell(X_1)$, $f(X_2) = E[Y
\mid X_2]$ and $f(X_1, X_2) = E[Y \mid X_1, X_2] + \ell(X_1, X_2)$. We want to
get the variance of $\hat \mu_y$. Then we want to show

1. $0 = \argmin_{\sigma_\ell^2} V(\hat \mu_y)$ for each $\sigma_\ell$

This model we have a true nonmonotone missing problem with the
following parts observed:

| Segment | $X_1$ | $X_2$ | $Y$ |
| ------- | ----- | --- | --- |
|$A_{11}$ | ✔️     | ✔️   |     |
|$A_{10}$ | ✔️     |     | ✔️   |

We assume that $|A_{11}| = |A_{10}|$.

```{python}
#| label: sympyres3
#| echo: false
#| output: asis

import sympy

def matinv_2x2(mat):
  assert mat.shape == (2, 2)

  det_mat = mat[0, 0] * mat[1, 1] - mat[0, 1] * mat[1, 0]
  ret = sympy.Matrix([
    [mat[1, 1], -mat[0, 1]],
    [-mat[1, 0], mat[0, 0]]
  ])

  return(ret / det_mat)

def matinv_3x3(mat):
  assert mat.shape == (3, 3)

  mat = mat.T

  adj_mat = sympy.Matrix([
    [mat[[1, 2], [1, 2]].det(), -mat[[1, 2], [0, 2]].det(),  mat[[1, 2], [0, 1]].det()],
    [-mat[[0, 2], [1, 2]].det(), mat[[0, 2], [0, 2]].det(), -mat[[0, 2], [0, 1]].det()],
    [mat[[0, 1], [1, 2]].det(), -mat[[0, 1], [0, 2]].det(),  mat[[0, 1], [0, 1]].det()],
  ])

  return(adj_mat / mat.det())

def matdet_2x2(mat):
  assert mat.shape == (2, 2)
  return(mat[0, 0] * mat[1, 1] - mat[0, 1] * mat[1, 0])

def matdet_3x3(mat):
  assert mat.shape == (3, 3)
  return(mat[0, 0] * matdet_2x2(mat[[1, 2], [1, 2]]) \
    - mat[0, 1] * matdet_2x2(mat[[1, 2], [0, 2]]) \
    + mat[0, 2] * matdet_2x2(mat[[1, 2], [0, 1]]))

def matdet_4x4(mat):
  assert mat.shape == (4, 4)
  return(mat[0, 0] * matdet_3x3(mat[[1, 2, 3], [1, 2, 3]]) \
    - mat[0, 1] * matdet_3x3(mat[[1, 2, 3], [0, 2, 3]]) \
    + mat[0, 2] * matdet_3x3(mat[[1, 2, 3], [0, 1, 3]]) \
    - mat[0, 3] * matdet_3x3(mat[[1, 2, 3], [0, 1, 2]]))

# We need variables for what we observe:
f11x1, f10x1, f11x2, f11x12, y01 = sympy.symbols('f11x1 f10x1 f11x2 f11x12 y01')
sx1, sx2, sx12, sl1, sl2, sl12, sy = sympy.symbols('sx1 sx2 sx12 sl1 sl2 sl12 sy')
pl1y, px1x2, px1l2, px2l1 = sympy.symbols('pl1y px1x2 px1l2 px2l1')
pl1l2, pl2y, pl12y = sympy.symbols('pl1l2 pl2y pl12y')
px1x12, px1l12, px12l1, pl1l12 = sympy.symbols('px1x12 px1l12 px12l1 pl1l12')
px2x12, px2l12, px12l2, pl2l12 = sympy.symbols('px2x12 px2l12 px12l2 pl2l12')

c01 = px1x2 * sx1 * sx2 + px2l1 * sx1 * sl2 + px2l1 * sx2 * sl1 + pl1l2 * sl1 * sl2
c02 = px1x12 * sx1 * sx12 + px1l12 * sx1 * sl12 + px12l1 * sx12 * sl1 + pl1l12 * sl1 * sl12
c12 = px2x12 * sx2 * sx12 + px2l12 * sx2 * sl12 + px12l2 * sx12 * sl2 + pl2l12 * sl2 * sl12

v_11 = sympy.Matrix([
  [sx1**2 + sl1**2 + 2 * pl1y * sl1 * sy, c01, c02],
  [c01, sx2**2 + sl2**2 + 2 * pl2y * sl2 * sy, c12],
  [c02, c12, sx12**2 + sl12**2 + 2 * pl12y * sl12 * sy]
])
v_10 = sympy.Matrix([
  [sx1**2 + sl1**2 + 2 * pl1y * sl1 * sy, sx1**2 + pl1y * sl1 * sy],
  [sx1**2 + pl1y * sl1 * sy, sy**2]
])

v_all = sympy.diag(v_11, v_10)

mod_mat = sympy.Matrix([
  [1, 1, 0, 0],
  [1, 0, 1, 0],
  [1, 0, 0, 1],
  [1, 1, 0, 0],
  [1, 0, 0, 0]
])
y_vec = sympy.Matrix([
  [f11x1],
  [f11x2],
  [f11x12],
  [f10x1],
  [y01]
])

v_inv = sympy.diag(matinv_3x3(v_11), matinv_2x2(v_10))
xvinvx = mod_mat.T * v_inv * mod_mat

# xvinvx_inv = matinv_4x4(xvinvx)

# out = xvinvx_inv * mod_mat.T * v_inv * y_vec
# sympy.simplify(out[0, 0])
# vhatmuy = sympy.simplify(xvinvx_inv[0, 0])
vhatmuy = matdet_3x3(xvinvx[[1, 2, 3], [1, 2, 3]]) / matdet_4x4(xvinvx)
sympy.simplify(vhatmuy)
sympy.simplify(sympy.diff(vhatmuy, sl))

```

### 2024-03-06 src/explore/20240223-optvar.qmd
I showed that the conditional expectation is the optimal function.

---
title: "Proof of Optimal Variance Estimators"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    grid:
      margin-width: 450px
bibliography: references.bib
comments:
  hypothesis: true
---

\newcommand{\Var}{\text{Var}}
\newcommand{\Cov}{\text{Cov}}
\newcommand{\argmin}{\text{argmin}}

# Summary

This report contains my current progress in establishing the optimal variance
estimators for the nonmonotone GLS estimation.

# General Notation

For now, we assume that we have a superpopulation model and a sampling
model. I would first like to establish optimality results using the simpliest
models, but it is not always clear what is simple in the nonmonotone case.
If we also consider a monotone model, we can have simple models based on if we
want to have complete cases (observations in which every variable is observed),
which I denote `CC`,
or if we want to have complete variables (variables which are never missing),
which I denote `CV`.
This leads to the following set of simple cases:

::: {#tbl-panel layout-ncol=4}
| $X$ | $Y$ |
|---|---|
| ✔️ | ✔️ |
| :heavy_check_mark: |  |
: Monotone {#tbl-one}

| $X$ | $Y$ |
|---|---|
| ✔️ | ✔️ |
| :heavy_check_mark: |  |
|  | ✔️ |
: Nonmonotone: CC {#tbl-two}

| $X_1$ | $X_2$ | $Y$ |
|-------|-------|---|
| ✔️     | ✔️     |   |
| ✔️     |       | :heavy_check_mark: | 
: Nonmonotone: CV {#tbl-three}

| $X_1$ | $X_2$ | $Y$ |
|-------|-------|---|
| ✔️     | ✔️     | :heavy_check_mark: |
| ✔️     | ✔️     |   |
| ✔️     |       | :heavy_check_mark: | 
: Nonmonotone: CC + CV {#tbl-four}

Set of Simple Monotone and Nonmonotone Cases
:::

For each of these cases, we can construct a GLS estimator from the corresponding
missingness model (I am giving the
monotone model in @tbl-one) like the following:

$$
\begin{bmatrix}
\hat f_{11}(X) \\
\hat f_{10}(X) \\
\hat Y_{11}
\end{bmatrix} \sim
\left(
\begin{bmatrix}
\mu_X \\
\mu_X \\
\mu_Y \\
\end{bmatrix},
\begin{bmatrix}
\Sigma_{11} & 0 & \Sigma_{fy} \\ 
0 & \Sigma_{10} & 0 \\ 
\Sigma_{yf} & 0 & \Sigma_y \\
\end{bmatrix}
\right).
$$

# Goals

I would like to establish the following:

1. For the corresponding missingness model, I would like to find all of the 
   $f(X)$ with $E[f(X)] = \mu_f$ such that the final estimator of $\hat \mu_y$
   has minimal variance.
2. Find the conditions under which the GLS estimator based on a superpopulation
   model is design consistent.

The second task is proving a result that is very similar to 
[@fuller2009sampling, Theorem 2.3.1].

# Results

::: {#thm-mono}
Assume that we have the simple monotone case in @tbl-one with independent
segments $A_{11}$ and $A_{10}$ having a sample size of $n_{11}$ and $n_{10}$
respectively. We assume that $n_{11} = n_{10}$. Then for GLS estimators in the
form,

$$
\begin{bmatrix}
\hat f_{11}(X) \\
\hat f_{10}(X) \\
\hat Y_{11}
\end{bmatrix} \sim
\left(
\begin{bmatrix}
\mu_X \\
\mu_X \\
\mu_Y \\
\end{bmatrix},
V
\right)
$$

where 
$\hat f_{11}(X) = n_{11}^{-1} \sum_{i = 1}^n \frac{\delta_{11i}f(x_i)}{\pi_{11i}}$,
$\hat f_{10}(X) = n_{10}^{-1} \sum_{i = 1}^n \frac{\delta_{10i}f(x_i)}{\pi_{10i}}$,
$\hat Y_{11} = n_{11}^{-1} \sum_{i = 1}^n \frac{\delta_{11i}y_i}{\pi_{11i}}$,
and the function $f(X) = E[Y \mid X]$ minimizes the variance of $\hat \mu_Y$.

:::

::: {.proof}
The goal is to show that

$$ E[Y \mid X] = \argmin_{f} \hat \mu_Y(f). $$

Our approach is the following. Consider $f(X) = E[Y \mid X] + \ell(X)$ where
$E[\ell(X)] = 0$. With this formulation of $f(X)$ we can compute the variance
matrix $V$. Define the following notation:

| Label | Symbol |
| :----- | ------: |
| $\Var(E[Y \mid X])$ | $\sigma_x^2$ |
| $\Var(Y)$ | $\sigma_y^2$ |
| $\Var(\ell(X))$ | $\sigma_\ell^2$ |
| $\Cov(E[Y \mid X], Y)$ | $\rho_{xy} \sigma_x \sigma_y$ |
| $\Cov(E[Y \mid X], \ell(X))$ | $\rho_{x\ell} \sigma_x \sigma_\ell$ |
| $\Cov(Y, \ell(X))$ | $\rho_{y\ell} \sigma_y \sigma_\ell$ |
: A table to define the notation used in the rest of the proof. {#tbl-lab1}

Notice that,

$$ \Cov(E[Y \mid X], Y) = E[E[Y \mid X]^2] - E[Y]^2 = \Var(E[Y \mid X]) $$

and

$$ \Cov(E[Y \mid X], \ell(X)) = E[E[Y \mid X]\ell(X)] = \Cov(Y, \ell(X)). $$

This means that $\sigma_x = \rho_{xy} \sigma_y$ and 
$\rho_{x\ell} \sigma_x = \rho_{y\ell} \sigma_y$. Hence, we have the following
covariance matrix $V$,

[The constant in front of the matrix is equal to $n^{-1}$ because $n_{11} = n_{10}$. Since each segment is chosen via simple random sampling, the overall variance of 
$\hat f_{11}(X)$ is $(n_{11}^{-1} - n^{-1})S_{f_{11}}^2 = (2/n - 1/n)S_{f_{11}}^2
= S^2_{f_{11}}/n$. Since $n_{11} = n_{10}$ the variance of $\hat f_{10}(X)$ is
also equal to $S_{f_{10}}^2 / n$. This holds for all of the variance and
covariance terms which allows us to factor it out of the matrix.]{.aside}

$$V =  \frac{1}{n}
\begin{bmatrix}
\rho_{xy}^2\sigma_y^2 + \sigma_\ell^2 + 2 \rho_{y\ell} \sigma_y \sigma_\ell & 0
& \rho_{xy}^2\sigma_y^2 + \rho_{y\ell} \sigma_y \sigma_\ell \\
0 & \rho_{xy}^2\sigma_y^2 + \sigma_\ell^2 + 2 \rho_{y\ell} \sigma_y \sigma_\ell & 0 \\
\rho_{xy}^2\sigma_y^2 + \rho_{y\ell} \sigma_y \sigma_\ell & 0 & \sigma_y^2
\end{bmatrix}.
$$

The goal is now to show that minimizing the variance of $\hat \mu_Y$ yields
$\sigma_\ell^2 = 0$. Using GLS results in a variance estimate of $\hat \mu$ of

$$ V(\hat \mu) = (X'V^{-1}X)^{-1}$$

where

$$ X = 
\begin{bmatrix}
1 & 1 \\
1 & 1 \\
1 & 0
\end{bmatrix}
\text{ for }
\mu = 
\begin{bmatrix}
\mu_Y \\ \mu_\ell
\end{bmatrix}.
$$

Evaluating this matrix expression and then viewing the first entry yields an
expression of the variance of $\mu_Y$,

$$\Var(\mu_Y) = n^{-1} \sigma_y^2 \left(1 + 
\frac{\rho_{xy}^4 \sigma_y^2 + 2\rho_{y\ell}\sigma_\ell \sigma_y \rho_{xy}^2 +
\rho_{y\ell}^2\sigma_{\ell}^2}{2(\rho_{xy}^2\sigma_y^2 + 2\rho_{y\ell}\sigma_\ell
\sigma_y + \sigma_\ell^2)}\right).$$

Then we can solve the equation,

$$ \frac{\partial V(\hat \mu_Y)}{\partial \sigma_\ell} = 0. $$

This yields (after some cancelations), the equation

$$ \sigma_\ell \rho_{y\ell} \sigma_y (\rho_{xy}^2 (\sigma_y - \sigma_\ell) +
\rho_{y\ell}^2 (1 - \rho_{xy}\sigma_y)) = 0.$$

Hence, the candidates for the minimum value of $V(\hat \mu_Y)$ are ($\rho_{y\ell}
= 0$, $\sigma_\ell = \sigma_y$), $\rho_{y\ell} = 0$, and $\sigma_\ell = 0$.
Plugging each of these back into $V(\hat \mu_Y)$ shows that $\sigma_\ell = 0$ is
actually the minimum. Hence, $\sigma_\ell^2 = 0$ yields the minimum variance of
$V(\hat \mu_Y)$ which means that since $E[\ell(X)] = 0$ that $\ell(X) = 0$. This
means that the optimal function of $f(X)$ is $E[Y \mid X]$. :black_medium_square:

:::

We can use the same technique to get a result for the nonmonotone case where we
have complete cases.

::: {#thm-nmcc}
Assume that we have the simple monotone case in @tbl-two with independent
segments $A_{11}$, $A_{10}$, and $A_{01}$ having sample sizes of $n_{11}$,
$n_{10}$, and $n_{01}$ respectively. We assume that $n_{11} = n_{10} = n_{01}$.
Then for GLS estimators in the form,

$$
\begin{bmatrix}
\hat f_{11}(X) \\
\hat f_{10}(X) \\
\hat Y_{11} \\
\hat Y_{10}
\end{bmatrix} \sim
\left(
\begin{bmatrix}
\mu_X \\
\mu_X \\
\mu_Y \\
\mu_Y \\
\end{bmatrix},
V
\right)
$$

where 
$\hat f_{11}(X) = n_{11}^{-1} \sum_{i = 1}^n \frac{\delta_{11i}f(x_i)}{\pi_{11i}}$,
$\hat f_{10}(X) = n_{10}^{-1} \sum_{i = 1}^n \frac{\delta_{10i}f(x_i)}{\pi_{10i}}$,
$\hat Y_{11} = n_{11}^{-1} \sum_{i = 1}^n \frac{\delta_{11i}y_i}{\pi_{11i}}$, 
$\hat Y_{10} = n_{10}^{-1} \sum_{i = 1}^n \frac{\delta_{10i}y_i}{\pi_{10i}}$,
and the function $f(X) = E[Y \mid X]$ minimizes the variance of $\hat \mu_Y$.

:::

::: {.proof}
This proof follows the same approach as @thm-mono. Using the same notation, we
have the following GLS model,

$$
\begin{bmatrix}
\hat f_{11}(X) \\
\hat f_{10}(X) \\
\hat Y_{11} \\
\hat Y_{10} \\
\end{bmatrix} \sim
\left(
\begin{bmatrix}
1 & 1 \\
1 & 1 \\
1 & 0 \\
1 & 0 \\
\end{bmatrix},
\begin{bmatrix}
\mu_Y \\
\mu_\ell
\end{bmatrix}, \frac{2}{n}
\begin{bmatrix}
\rho_{xy}^2\sigma_y^2 + \sigma_\ell^2 + 2 \rho_{y\ell} \sigma_y \sigma_\ell & 0
& \rho_{xy}^2\sigma_y^2 + \rho_{y\ell} \sigma_y \sigma_\ell & 0\\
0 & \rho_{xy}^2\sigma_y^2 + \sigma_\ell^2 + 2 \rho_{y\ell} \sigma_y \sigma_\ell & 0 & 0 \\
\rho_{xy}^2\sigma_y^2 + \rho_{y\ell} \sigma_y \sigma_\ell & 0 & \sigma_y^2 & 0 \\
0 & 0 & 0 & \sigma_y^2 \\
\end{bmatrix}.
\right)
$$

[The constant in front of the matrix is equal to $2/n$ because $n_{11} = n_{10}
= n_{01}$. Since each segment is chosen via simple random sampling, the overall
variance of 
$\hat f_{11}(X)$ is $(n_{11}^{-1} - n^{-1})S_{f_{11}}^2 = (3/n - 1/n)S_{f_{11}}^2
= 2S^2_{f_{11}}/n$. This holds for all of the variance and
covariance terms which allows us to factor it out of the matrix.]{.aside}

The variance of $\hat \mu_Y$ is

$$V(\hat \mu_Y) = 2n^{-1} \sigma_y^2 \left(1 + 
\frac{2 \rho_{xy}^2 \sigma_y^2 + 4 \rho_{y\ell} \sigma_\ell \sigma_y +
2\sigma_\ell^2}{\rho_{xy}^4 \sigma_y^2 + 2 \rho_{xy}^2 \rho_{y\ell} \sigma_\ell
\sigma_y - 4\rho_{xy}^2 \sigma_y^2 + \rho_{y\ell}^2 \sigma_\ell^2 - 8
\rho_{y\ell}\sigma_{\ell} \sigma_{y} - 4 \sigma_{\ell}^2}
\right).$$

Solving $\frac{\partial V(\hat \mu_Y)}{\partial \sigma_y} = 0$ yields (after
simplification),

$$ \sigma_{\ell} \sigma_y (\rho_{xy}^4 \sigma_y + \rho_{xy}^2 \rho_{y\ell}(\sigma_\ell
- \rho_{y\ell}\sigma_y) - \rho_{y\ell}^3 \sigma_\ell) = 0. $$

This implies that the minimum variance occurs at $\sigma_{\ell} = 0$ which means
that $\ell(X) = 0$. Hence, the optimal value of $f(X) = E[Y \mid X]$.
:black_medium_square:

:::

# Discussion

This current approach seems to be working. However, I am using SymPy [@meurer2017sympy]
to assist with and check the computational algebra. I have started to find the
optimal functions of $f(X_1), f(X_2)$, and $f(X_1, X_2)$ from @tbl-three but
inverting a $4 \times 4$ matrix is difficult to approach computationally. Since
this is very similar to the semiparametric literature, I am going to try to see
if I can use some of their techniques to simplify my work.

<!---
# Current Progress

The first thing that I have tried to consider is the optimal function $f(X)$ in
the monotone case (@tbl-one). For this missingness set, we have the GLS model:

$$
\begin{bmatrix}
\hat f_{11}(X) \\
\hat f_{10}(X) \\
\hat Y_{11}
\end{bmatrix} \sim
\left(
\begin{bmatrix}
\mu_f \\
\mu_f \\
\mu_y \\
\end{bmatrix},
\begin{bmatrix}
\Sigma_{11} & 0 & \Sigma_{fy} \\ 
0 & \Sigma_{10} & 0 \\ 
\Sigma_{yf} & 0 & \Sigma_y \\
\end{bmatrix}
\right).
$$

::: {.column-margin}

I also compared this result to the case in which we have $E[f(X)] = \mu_y$ which
occurs if $f(X) = E[Y \mid X]$. In this case, we have the model

$$
\begin{bmatrix}
\hat f_{11}(X) \\
\hat f_{10}(X) \\
\hat Y_{11}
\end{bmatrix} \sim
\left(
\begin{bmatrix}
\mu_y \\
\mu_y \\
\mu_y \\
\end{bmatrix},
\begin{bmatrix}
\Sigma_{11} & 0 & \Sigma_{fy} \\ 
0 & \Sigma_{10} & 0 \\ 
\Sigma_{yf} & 0 & \Sigma_y \\
\end{bmatrix}
\right).
$$

Then using GLS (it is not too hard to invert the covariance matrix because it is
a block diagonal), we get

$$
\hat \mu_y =
\frac{\hat f_{10}(X)(\Sigma_{11}\Sigma_y - \Sigma_{fy}^2) - \hat f_{11}(X)
\Sigma_{10}(\Sigma_{fy} - \Sigma_y) + Y_{11} \Sigma_{10}(\Sigma_{11} - \Sigma_{fy})}{
\Sigma_{10}(\Sigma_{11} - 2\Sigma_{fy} + \Sigma_y) + \Sigma_{11}\Sigma_y - \Sigma_{fy}^2
}.
$$

The variance of this is

$$
\Var(\mu_y) = 
\left(\frac{\Sigma_{11} - 2\Sigma_{fy} + \Sigma_y}{\Sigma_{11}\Sigma_y -
\Sigma_{fy}^2} + \frac{1}{\Sigma_{10}}\right)^{-1}.
$$

We can notice that if we let all of the variances 
($\Sigma_{11}, \Sigma_{10}, \Sigma_y = 1$) then we can compare this variance
result to @eq-condvarymono. If this case this marginal variance is always lower
than @eq-condvarymono unless $\Sigma_{fy} = 1$. In this case they are the same.

However, I do not think that this is a problem because in this marginal model we
are assuming that $f(X)$ is correctly specified for $E[Y \mid X]$. 
**Is this correct Dr. Kim?**

:::

Using the technique of @zhou2012efficient, we can first solve for the optimal
estimator of $\mu_f$, which is

$$
\hat \mu_f = 
\frac{\frac{\hat f_11(X)}{\Sigma_{11}} + \frac{\hat f_10(X)}{\Sigma_{10}}}{
\Sigma_{11}^{-1} + \Sigma_{10}^{-1}}.
$$

Then the conditional distribution of $Y_{11}$ given $f(X)$ is

$$
Y_{11} \mid \hat f(X) \sim \left(
\mu_y + \Sigma_{yf}\Sigma_{11}^{-1}(\hat f_11(X) - \hat \mu_f),
\Sigma_y - \Sigma_{yf} \Sigma_{11}^{-1} \Sigma_{fy}\right).
$$

This means that

$$
\hat \mu_y = Y_{11} - 
\left(\frac{\Sigma_{yf}}{\Sigma_{11} + \Sigma_{10}}\right)
(\hat f_{11}(X) - \hat f_{10}(X))
$$

and this has a variance of

$$ \Sigma_{y} - \frac{\Sigma_{yf}^2}{\Sigma_{11} + \Sigma_{10}}. $$ {#eq-condvarymono}

From semiparametric theory, I think that this result should be expected because
$\Sigma_{11} + \Sigma_{10} = \Var(\hat f_{11}(X) + \hat f_{10}(X))$. This does
generate follow up questions for me:

1. It seems like we want to find the function $f(X)$ such that $|\Sigma_{fy}|$ 
   is maximized. Does this mean that $f(X)$ should be $E[Y \mid X]$?
2. In this analysis, I am not using the superpopulation model directly. Is the
   only purpose of a superpopulation model to assert that the GLS estimator is
   optimal for a proposed model?

--->

# References

### 2024-03-06 src/explore/20240301-sagescore.qmd
I showed that the conditional expectation is the optimal function.

---
title: "SymPy Representation for Efficient Score Functions"
author: "Caleb Leedy"
date: "March 1, 2024"
format: html
jupyter:
  kernel: sagemath
---

# Summary: Part 1

This part determines the efficient score function for the simple monotone model.

```{python}
#| label: sage score 1

from sage.symbolic.integration.integral import definite_integral

f11, f10, y = var('f11 f10 y')
muy, muf, sy, sf, pfy = var('muy muf sy sf pfy')

loglikefun = -3/2 * log(2 * pi) - 1/2 * log(sy^2 * (1 - pfy^2)) - 1 / (2 * sy^2
* (1 - pfy^2)) * (y - muy - (pfy * sy / sf) * (f11 - muf))^2 - log(sf^2) - 1/2 * (
  (f11 - muf)^2 + (f10 - muf)^2) / sf^2

S_muy = loglikefun.derivative(muy)
S_muf = loglikefun.derivative(muf)
S_sy = loglikefun.derivative(sy)
S_sf = loglikefun.derivative(sf)
S_pfy = loglikefun.derivative(pfy)

v_mat = matrix([
  [sf^2, 0 , pfy * sf * sy],
  [0, sf^2, 0],
  [pfy * sf * sy, 0, sy^2]
])

mean_diff = matrix([
  [f11 - muf],
  [f10 - muf],
  [y - muy]
])

dens_fy = (2 * pi)^(3/2) * (v_mat.det())^(-1/2) * exp(-1 / 2 * mean_diff.transpose() * v_mat.inverse() * mean_diff)[0, 0]

dens_y = (2 * pi * sy^2)^(-1/2) * exp(-1/(2 * sy^2) * (y - muy)^2)

res = definite_integral(dens_y, y, -infinity, infinity)
res.full_simplify()
```

### 2024-03-06 src/explore/20240301-sagetest.qmd
I showed that the conditional expectation is the optimal function.

---
title: "Sage Math Representation for Optimal Regression Estimation"
author: "Caleb Leedy"
date: "March 1, 2024"
format: html
jupyter:
  kernel: sagemath
---

# Summary: Part 1

```{python}

f11, f10, y = var('f11 f10 y')
sl, sx, sy = var('sl sx sy')
pcy, ply = var('pcy ply')

v_all = matrix([
  [pcy**2 * sy**2 + sl**2 + 2 * ply * sl * sy, 0, pcy**2 * sy**2 + ply * sl * sy],
  [0, pcy**2 * sy**2 + sl**2 + 2 * ply * sl * sy, 0],
  [pcy**2 * sy**2 + ply * sl * sy, 0, sy**2]
])

mod_mat = matrix([
  [1, 1],
  [1, 1],
  [1, 0]
])
y_vec = matrix([
  [f11],
  [f10],
  [y]
])

out = (mod_mat.transpose() * v_all.inverse() * mod_mat).inverse() * mod_mat.transpose() * v_all.inverse() * y_vec
(out[0, 0]).full_simplify()

estvar = (mod_mat.transpose() * v_all.inverse() * mod_mat).inverse()
(estvar[0, 0]).full_simplify()

```

# Summary: Part 4

We consider the model where $f(X_1) = E[Y \mid X_1] + \ell(X_1)$, $f(X_2) = E[Y
\mid X_2]$ and $f(X_1, X_2) = E[Y \mid X_1, X_2] + \ell(X_1, X_2)$. We want to
get the variance of $\hat \mu_y$. Then we want to show

1. $0 = \argmin_{\sigma_\ell^2} V(\hat \mu_y)$ for each $\sigma_\ell$

This model we have a true nonmonotone missing problem with the
following parts observed:

| Segment | $X_1$ | $X_2$ | $Y$ |
| ------- | ----- | --- | --- |
|$A_{11}$ | ✔️     | ✔️   |     |
|$A_{10}$ | ✔️     |     | ✔️   |

We assume that $|A_{11}| = |A_{10}|$.

```{sage}
#| label: sympyres3
#| echo: false
#| output: asis

#import sympy
#
#def matinv_2x2(mat):
#  assert mat.shape == (2, 2)
#
#  det_mat = mat[0, 0] * mat[1, 1] - mat[0, 1] * mat[1, 0]
#  ret = sympy.Matrix([
#    [mat[1, 1], -mat[0, 1]],
#    [-mat[1, 0], mat[0, 0]]
#  ])
#
#  return(ret / det_mat)
#
#def matinv_3x3(mat):
#  assert mat.shape == (3, 3)
#
#  mat = mat.T
#
#  adj_mat = sympy.Matrix([
#    [mat[[1, 2], [1, 2]].det(), -mat[[1, 2], [0, 2]].det(),  mat[[1, 2], [0, 1]].det()],
#    [-mat[[0, 2], [1, 2]].det(), mat[[0, 2], [0, 2]].det(), -mat[[0, 2], [0, 1]].det()],
#    [mat[[0, 1], [1, 2]].det(), -mat[[0, 1], [0, 2]].det(),  mat[[0, 1], [0, 1]].det()],
#  ])
#
#  return(adj_mat / mat.det())
#
#def matdet_2x2(mat):
#  assert mat.shape == (2, 2)
#  return(mat[0, 0] * mat[1, 1] - mat[0, 1] * mat[1, 0])
#
#def matdet_3x3(mat):
#  assert mat.shape == (3, 3)
#  return(mat[0, 0] * matdet_2x2(mat[[1, 2], [1, 2]]) \
#    - mat[0, 1] * matdet_2x2(mat[[1, 2], [0, 2]]) \
#    + mat[0, 2] * matdet_2x2(mat[[1, 2], [0, 1]]))
#
#def matdet_4x4(mat):
#  assert mat.shape == (4, 4)
#  return(mat[0, 0] * matdet_3x3(mat[[1, 2, 3], [1, 2, 3]]) \
#    - mat[0, 1] * matdet_3x3(mat[[1, 2, 3], [0, 2, 3]]) \
#    + mat[0, 2] * matdet_3x3(mat[[1, 2, 3], [0, 1, 3]]) \
#    - mat[0, 3] * matdet_3x3(mat[[1, 2, 3], [0, 1, 2]]))

# We need variables for what we observe:
f11x1, f10x1, f11x2, f11x12, y01 = var('f11x1 f10x1 f11x2 f11x12 y01')
sx1, sx2, sx12, sl1, sl2, sl12, sy = var('sx1 sx2 sx12 sl1 sl2 sl12 sy')
pl1y, px1x2, px1l2, px2l1 = var('pl1y px1x2 px1l2 px2l1')
pl1l2, pl2y, pl12y = var('pl1l2 pl2y pl12y')
px1x12, px1l12, px12l1, pl1l12 = var('px1x12 px1l12 px12l1 pl1l12')
px2x12, px2l12, px12l2, pl2l12 = var('px2x12 px2l12 px12l2 pl2l12')

c01 = px1x2 * sx1 * sx2 + px2l1 * sx1 * sl2 + px2l1 * sx2 * sl1 + pl1l2 * sl1 * sl2
c02 = px1x12 * sx1 * sx12 + px1l12 * sx1 * sl12 + px12l1 * sx12 * sl1 + pl1l12 * sl1 * sl12
c12 = px2x12 * sx2 * sx12 + px2l12 * sx2 * sl12 + px12l2 * sx12 * sl2 + pl2l12 * sl2 * sl12

v_11 = matrix([
  [sx1**2 + sl1**2 + 2 * pl1y * sl1 * sy, c01, c02],
  [c01, sx2**2 + sl2**2 + 2 * pl2y * sl2 * sy, c12],
  [c02, c12, sx12**2 + sl12**2 + 2 * pl12y * sl12 * sy]
])
v_10 = matrix([
  [sx1**2 + sl1**2 + 2 * pl1y * sl1 * sy, sx1**2 + pl1y * sl1 * sy],
  [sx1**2 + pl1y * sl1 * sy, sy**2]
])

v_all = block_matrix([[v_11, 0], [0, v_10]])

mod_mat = matrix([
  [1, 1, 0, 0],
  [1, 0, 1, 0],
  [1, 0, 0, 1],
  [1, 1, 0, 0],
  [1, 0, 0, 0]
])
y_vec = matrix([
  [f11x1],
  [f11x2],
  [f11x12],
  [f10x1],
  [y01]
])

v_inv = sympy.diag(matinv_3x3(v_11), matinv_2x2(v_10))
xvinvx = mod_mat.transpose() * v_all.inverse() * mod_mat

# xvinvx_inv = matinv_4x4(xvinvx)

out = xvinvx.inverse() * mod_mat.transpose() * v_all.inverse() * y_vec
res = out[0, 0]
res.full_simplify()

# vhatmuy = sympy.simplify(xvinvx_inv[0, 0])
vhatmuy = matdet_3x3(xvinvx[[1, 2, 3], [1, 2, 3]]) / matdet_4x4(xvinvx)
sympy.simplify(vhatmuy)
sympy.simplify(sympy.diff(vhatmuy, sl))

```

### 2024-03-06 src/explore/20240304-altoptvar.qmd
I showed that the conditional expectation is the optimal function.

---
title: "Alternate: Proof of Optimal Variance Estimators"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

\newcommand{\Var}{\text{Var}}
\newcommand{\argmin}{\text{argmin}}

# Summary

This report contains my current progress in establishing the optimal variance
estimators for the nonmonotone GLS estimation. I think that I got it.

# Report (Monday, March 3)

Previously^[See 20240223-optvar.html.] we argued that the conditional
expectation of $Y$ on the observed variables is the optimal form of $f(X)$. We
showed this by allowing $f(X) = E[Y \mid X] + \ell(X)$ where $E[\ell(X)] = 0$.
Then the minimal variance estimator of $E[Y]$ yielded $\Var(\ell(X)) = 0$, which
indicated that the optimal function of $f(X) = E[Y \mid X]$. Unfortunately, the
details in this proof made it difficult to scale beyond the case of having two
variables, and so we sought an alternative method which could scale better.

One approach is to use semiparametric theory for parametric models. In the
monotone case, the problem that we want to solve is to find the functional form
of $f(X)$ such that we have the minimum variance of $\hat \beta = E[Y]$ for the
following model, which holds if $n_{11} = n_{10}$,

$$
\begin{bmatrix}
\hat f_{11}(X) \\
\hat f_{10}(X) \\
\hat Y_{11}
\end{bmatrix} \sim
\left(
\begin{bmatrix}
\mu_f \\
\mu_f \\
\mu_y \\
\end{bmatrix},
n^{-1}
\begin{bmatrix}
\sigma_f^2 & 0 & \rho_{fy} \sigma_f \sigma_y \\ 
0 & \sigma_f^2 & 0 \\ 
\rho_{fy} \sigma_f \sigma_y & 0 & \sigma_y^2 \\
\end{bmatrix}
\right).
$$

We now have a parametric model with $\beta = \mu_y$ and $\eta = (\mu_f,
\sigma_y, \sigma_f, \rho_{fy})^T$. Previously, we were thinking about how the
choice of $f(X)$ determines $\mu_f$, $\sigma_f$, and $\rho_{fy}$, which is true.
However, since these parameters are the only things that we need for the model,
we can find the efficient influence function by letting $\theta = (\beta,
\eta^T)^T$ and then we have,

$$ \varphi_{eff} = \left(\frac{\partial \theta}{\partial \beta}\right)^{-1}
E[S_\theta S_\theta^T]^{-1} S_\theta.$$

This looks promising. However, the inverse information matrix, $E[S_\theta
S_\theta^T]^{-1}$ is very difficult to compute. Even for the monotone case that
we currently consider it is a 5x5 matrix.^[Technically, it is a block diagonal
matrix with blocks of size 1 and size 4. The 4x4 block is very challenging
because it requires the actual expectations of difference squared terms.] This
means that this technique also does not generalize well.

We have learned that we can work with the model directly, but this will require
slightly more advanced techniques like something similar to the partial linear
model (PLM).

# Update (Tuesday, March 4)

I think that I have a good method of proof. However, I want to check to make
sure that the details are correct. I argue that the optimal function for $f(X)$
in the GLS estimator is $f(X) = E[Y \mid X]$. It takes three steps:

1. We argue that $\alpha(X) = E[Y \mid X]$ is optimal in the PLM,
2. Then we show that the PLM holds for each independent segment, and
3. Finally, we combine the segment estimators using the variance weighted
   average.

::: {#lem-plm}
Let $Y$ and $X$ be random variables that are possibly vector valued with finite
variance, and denote $\mu_Y = E[Y]$. Then, for

$$ \alpha^*(X) = \argmin_{\alpha(X)} E[(Y - \mu_Y - \alpha(X))^2] \text{ such
that } E[\alpha(X)] = 0$$

we have $\alpha^*(X) = E[Y \mid X] - \mu_Y$.
:::

::: {.proof}
Let $\hat \alpha(X) = E[Y \mid X] - \mu_Y$. Notice, that

$$ E[\hat \alpha(X)] = E[E[Y \mid X]] - \mu_Y = 0. $$

Furthermore, let $\ell(X)$ satisfy $\alpha(X) = \hat \alpha(X) + \ell(X)$. Then,

\begin{align*}
E[(Y - \mu_y - \alpha(X))^2] 
&= E[(Y - \mu_y - \hat \alpha(X) - \ell(X))^2] \\
&= E[(Y - \mu_y - \hat \alpha(X))^2] - 2 E[(Y - \mu_Y - \hat \alpha(X))\ell(X)]
+ E[\ell(X)^2] \\
&= E[(Y - \mu_y - \hat \alpha(X))^2] - 2 E[E[(Y - E[Y \mid X])\ell(X)
\mid X]] + E[\ell(X)^2] \\
&= E[(Y - \mu_y - \hat \alpha(X))^2] - 2 E[(E[Y \mid X] - E[Y \mid X])\ell(X)] +
E[\ell(X)^2] \\
&= E[(Y - \mu_y - \hat \alpha(X))^2] + E[\ell(X)^2].
\end{align*}

Hence, $\alpha^*(X) = \hat \alpha(X) = E[Y \mid X] - \mu_Y$.
:::

## Step 1

Now consider the following partial linear model

$$Y = \mu_Y + \alpha(X) + \varepsilon$$

where $E[\alpha(X)] = 0$.^[Note that this partial linear model is different than
the one used by @robinson1988root.] By @lem-plm, the function of $\alpha$ that
minimizes the variance of $\varepsilon$ is $\alpha(X) = E[Y \mid X] - \mu_Y$.

## Step 2

Consider a single segment where we observe $Y$ and some variables $X$ where $X$
could be a vector. We consider the sample superpopulation model,

$$
\begin{bmatrix} X \\ Y \end{bmatrix} \sim \left(
\begin{bmatrix} \mu_X \\ \mu_Y \end{bmatrix},
\begin{bmatrix} 
\Sigma_{XX} & \Sigma_{XY} \\ \Sigma_{YX} & \sigma_Y^2
\end{bmatrix}
\right).
$$

The goal is to find the function of $X$, $f(X)$, such that we have the minimum
variance. Mathematically, this is

$$f^*(X) = \argmin_{f(X)} \Var(Y - \mu_Y - f(X)).$$

For identification purposes we also require that $E[f(X)] = 0$. This makes the
problem the same as Step 1, in which case the optimal form of $f(X)$ is $E[Y
\mid X] - \mu_Y$.^[If our model is assumed to be normal, then we can see what
the form of $f(X)$ is from the conditional distribution: $f(X) =
\Sigma_{YX}\Sigma_{XX}^{-1}(X - \mu_X)$.] Note that we define the
optimal form of $f(X)$ in the sample *superpopulation* model. We will worry
about estimating $f(X)$ later.

## Step 3

The overall goal is to find the optimal functional form of $f(G_r(X))$ where
$G_r(X)$ is the set of $X$ variables observed in the $r$-th segment. However,
since we assume that each segment is independent of the rest, and that $X$ and
$Y$ follow a normal model, by the Gauss-Markov theorem, the inverse-variance
weighted average of each segment contains best linear unbiased estimator of
$\theta = E[Y]$. This means that if we have the problem,

$$
\begin{bmatrix} f(G_0(X)) \\ f(G_1(X)) \\ \vdots \\ Y_0 \\ \vdots \end{bmatrix} \sim \left(
\begin{bmatrix} \mu_{G_0(X)} \\ \mu_{G_1(X)} \\ \vdots \\ \mu_Y \\ \vdots \end{bmatrix},
V
\right).
$$

The optimal form of $f(G_r(X))$ is $E[Y \mid G_r(X)] - \mu_Y$.

# Questions

1. I am not totally sure how to cover segments that do not contain $Y$. Should I
   say that the best estimator for $Y - \mu_Y$ is $E[Y \mid X] - \mu_Y$?

# Next Steps

* Try estimating the variance with REML so that we have unbiased estimation.
* Extend proof to semiparametric case where we do not necessarily assume a
normal model.
* Find the optimal estimator under MAR instead of missing by design.

### 2024-04-11 final/reports/prelim_paper.tex
Writing prelim_paper added some of Dr. Kim's comments.

\documentclass[12pt]{article}

\usepackage{amsmath, amssymb, amsthm, mathrsfs, fancyhdr}
\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
\usepackage[style=authoryear]{biblatex}
\usepackage{booktabs}
\usepackage{float}

\addbibresource{references.bib}

\hypersetup{colorlinks = true, urlcolor = black}

\headheight     15pt
\topmargin      -1.5cm   % read Lamport p.163
\oddsidemargin  -0.04cm  % read Lamport p.163
\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
\textwidth      16.59cm
\textheight     23.94cm
\parskip         7.2pt   % sets spacing between paragraphs
\parindent         0pt   % sets leading space for paragraphs
\pagestyle{empty}        % Uncomment if don't want page numbers
\pagestyle{fancyplain}

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}

\newcommand{\bx}{\mathbf{x}}
\newcommand{\R}{\mathbb{R}}
\renewcommand{\bf}[1]{\mathbf{#1}}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

\begin{document}

\title{Debiased calibration for generalized two-Phase sampling}
\author{Caleb Leedy}
\maketitle

\section{Introduction}

\begin{itemize}
\item Combining information from several sources is an important practical problem. 
\item In many cases, we do not have direct access to the other sources. We can only obtain summary statistics (and their standard errors) for the external data sources. We wish to incorporate the information from external  sources in our in-house data effectively using calibration weighting.  
\item We formulate the problem as a generalized two-phase sampling where the first phase sample can be obtained from multiple sources. The second-phase sample is our in-house data in which we want to construct the calibration weights. 
\item To achieve the goal, we first consider  the  classical two-phase sampling setup where the second-phase sample is a subset of the first-phase sample. After that, we extend the setup to more general cases such as non-nested two-phase sampling or multiple independent surveys with some common measurements. 
\item (To simplify the presentation, I wonder whether we can just use SRS in the first-phase sampling. ) 
\end{itemize}
\section{Topic 1: Classical Two-Phase Sampling}

\subsection{Background and Introduction}

Consider a finite population of size $N$ containing elements $(X_i, Y_i)$ where
an initial (Phase 1) sample of size $n$ is selected and $X_i$ is observed. Then
from the Phase 1 sample of elements, a (Phase 2) sample of size $r < n$ is
selected and $Y_i$ is observed. This is two-phase sampling (See 
\cite{fuller2009sampling}, \cite{kim2024statistics} for general references.) The
goal of two-phase sampling is to construct an estimator of $Y$ not
only using the observed information from the Phase 2 sample but also
incorporating the extra auxilary information of $X$ from the Phase 1 sample, and
the challenge is doing this efficiently.

An easy-to-implement unbiased estimator in the spirit of a Horvitz-Thompson
estimator (\cite{horvitz1952generalization}, \cite{narain1951sampling}) is the
$\pi^*$-estimator. Let $\pi_i^{(2)}$ be the response probability of element $i$
being observed in the Phase 2 sample. Then, allowing the elements in the Phase 1
sample to be represented by $A_1$ and the elements in the Phase 2 sample to be
denoted as $A_2$,

\begin{align*}
  \pi_i &= \sum_{A_2: i \in A_2} \Pr(A_2) \\ 
        &= \sum_{A_1: A_2 \subseteq A_1} \sum_{A_2: i \in A_2} \Pr(A_2 \mid
        A_1) \Pr(A_1) \\
        &= \sum_{A_1: i \in A_1} \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1) \Pr(A_1).
\end{align*}

If we define $\pi_{2i | 1}^{(2)} = \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1)$ and
$\pi_i^{(1)} = \sum_{A_1: i \in A_1} \Pr(A_1)$ then,

$$ \pi_i = \pi_{2i | 1}^{(2)} \pi_i^{(1)}.$$
\textcolor{red}{(note: there is an abuse of notation. You do not need to introduce superscript $(2)$ on $\pi_{2i \mid 1}$.) }
This means that we can define the $\pi^*$-estimator as the following design
unbiased estimator:

$$ \hat Y_{\pi^*} = \sum_{i \in A_2} \frac{y_i}{\pi_{2i | 1}^{(2)} \pi_i^{(1)}}.$$

\textcolor{red}{Dr. Kim: Should I provide the proof for design unbiasedness of
$\hat Y_{\pi^*}$? } \textcolor{blue}{Not necessary. You may cite a book or a paper  for this claim. }

While unbiased, the $\pi^*$-estimator unfortunately does not account for the
additional information contained in the auxiliary Phase 1 variable $X$. The
two-phase regression estimator $\hat Y_{reg, tp}$ does incorporate \textcolor{blue}{$\hat{X}_1$ obtained from the phase one sample. That is, we can leverage the external information $\hat{X}_1$ to improve the $\pi^*$-estimator in the second-phase sample. } 
The two-phase 
regression estimator has the form,

$$ \hat Y_{reg, tp} 
= \sum_{i \in A_1} \frac{1}{\pi_i^{(1)}} x_i \hat \beta_2+ \sum_{i \in A_2}
\frac{1}{\pi_i^{(1)}\pi_{2i|1}^{(2)}} (y_i - x_i \hat \beta_2)$$
where $$\hat \beta_2 = \left(\sum_{i \in A_2} 
  \frac{x_i x_i'}{\pi_{2i|1}^{(2)}}\right)^{-1} 
  \sum_{i \in A_2} \frac{x_i y_i}{\pi_i^{(1)}\pi_{2i|1}^{(2)}}. 
  \footnote{Caleb, we do not have to use $\pi_{2i \mid 1}^{(2)}$ in computing $\hat{\beta}_2$. Please check my book. }
  $$

\textcolor{red}{I suggest that we use 
$$\hat \beta_2 = \left(\sum_{i \in A_2} 
  \frac{x_i x_i'}{\pi_{i}^{(1)} q_i }\right)^{-1} 
  \sum_{i \in A_2} \frac{x_i y_i}{\pi_i^{(1)} q_i }
$$ 
where $q_i=q(\bx_i)$ is a function of $\bx_i$. This will have some connection with model-optimal regression estimator (under superpopulation model).  
}

  The regression
estimator is the minumum variance design consistent linear estimator which is
easily shown to be the case because $\hat Y_{reg, tp} = \sum_{i \in A_2} \hat
w_{2i} y_i / \pi_i^{(1)}$ where

$$\hat w_i = \argmin_{w} \sum_{i \in A_2} (w_i - \pi_{2i|1}^{-1})^2 \text{ such
that } \sum_{i \in A_2} w_i x_i / \pi_i^{(1)} = \sum_{i \in A_1} x_i /
\pi_i^{(1)}.$$

\textcolor{red}{Using $q_i$, we can construct 
$$\hat w_{2i} = \argmin_{w} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2  q_i \text{ such
that } \sum_{i \in A_2} w_{2i}  x_i / \pi_i^{(1)} = \sum_{i \in A_1} x_i /
\pi_i^{(1)}$$
as a way to implement the two-phase regression estimator indirectly using calibration weighting. 
}

This means that $\hat Y_{reg, tp}$ is also a calibration estimator. The idea
that regression estimation is a form of calibration was extended by 
\cite{deville1992calibration} to consider loss functions other than just squared
loss. They generalized the loss function to minimize $\sum_i G(w_i, d_i)q_i$ for
weights $w_i$ and design-weights $d_i$ where $G()$ is non-negative, strictly
convex function with respect to $w$, defined on an interval containing $d_i$,
with $g(w_i, d_i) = \partial G / \partial w$ continuous.
\footnote{The \cite{deville1992calibration} paper considers regression estimators
for a single phase setup, which we apply to our two-phase example. } This
generalization includes empirical likelihood estimation, and maximum entropy
estimation amoung others. The variance estimation is based on a linearization
that shows that minimizing the generalized loss function subject to the
calibration constraints is asymptotically equivalent to a regression estimator.

While this generalization is useful to an analyst who may want different
properties of their estimator from maximum entropy estimation rather than the
minimal squared loss, unless $\pi_{2i|1}^{-1} \in C(X)$, the estimator is not
design consistent. \footnote{ \textcolor{red}{I do not understand this part.} } 
Furthermore, at a conceptual level, the regression estimator
has a nice feature that its two terms can be thought about as minimizing
variance and bias correction,

$$ \hat Y_{reg, tp} 
= \underbrace{\sum_{i \in A_1} \frac{x_i \hat \beta_2}{\pi_i^{(1)}}}_{Minimizing
the variance} + \underbrace{\sum_{i \in A_2}
\frac{1}{\pi_i^{(1)}\pi_{2i|1}^{(2)}} (y_i - x_i \hat \beta_2)}_{Bias
correction}.$$

The \cite{deville1992calibration} method incorporates the design weights into
the loss function, which is the part minimizing the variance. Instead, there is
a desire to get design consistency from the calibration term. In 
\cite{kwon2024debiased}, the authors show that for a generalized entropy
function $G(w)$, including a term of $g(\pi_{2i|1}^{-1})$ into the calibration
for $g = \partial G / \partial w$ not only creates a design consistent
estimator, but it also has better efficiency than the generalized regression
estimators of \cite{deville1992calibration}.

Unfortunately, the method of \cite{kwon2024debiased} requires known calibration
levels (no uncertainty) for the finite population. It does not handle the
two-phase setup where we need to estimate the finite population total of $x$
from the Phase 1 sample. Hence, we extend this method to have a valid variance 
estimator when including estimated Phase 1 weights.

\subsection*{Methodology}

We follow the approach of \cite{kwon2024debiased} for the debiased calibration
method. We consider maximizing the generalized entropy \cite{gneiting2007strictly},

\begin{equation}\label{eq:primalloss}
H(w) = - \sum_{i \in A_2} G(w_i)
\end{equation}

where $G: \mathcal{V} \to \R$ is strictly convex, differentiable function
subject to the constraints:

\begin{equation}\label{eq:calconst1}
\sum_{i \in A_2} \frac{x_i w_i}{\pi_i^{(1)}} = 
\sum_{i \in A_1} \frac{x_i}{\pi_i^{(1)}}
\end{equation}

and

\begin{equation}\label{eq:calconst2}
\sum_{i \in A_2} \frac{g(\pi_{2i|1}^{-1})w_i}{\pi_i^{(1)}} = 
\sum_{i \in A_1} \frac{g(\pi_{2i|1}^{-1})}{\pi_i^{(1)}}.
\end{equation}

The first constraint is the existing calibration constraint and the second
ensures that design consistency is achieved. Here, $g(w) = \partial G / \partial
w$. The original method of \cite{kwon2024debiased} only considered having finite
population quantities on the right hand side of \ref{eq:calconst1}. Let $z_i =
(x_i, g(\pi_{2i|1}^{-1})$. To solve this minimization problem, we use the convex
conjugate function $F(w) = -G(g^{-1}) + g^{-1}(w)w$, and instead find

$$ \hat \lambda = \argmin_{\lambda \in \Lambda_A} \sum_{i \in A_2} F(w_i) +
\lambda_1^T \sum_{i \in A_1} \frac{z_i}{\pi_{2i|1}}$$

for $\Lambda_A = \{\lambda: \lambda^T z_i \in g(\mathcal{V}) \text{ for all } i
\in A_2\}$.

\subsection*{Theoretical Results}

We prove two results in this section is the same spirit as \cite{kwon2024debiased}.
First, we state the assumptions of our analysis then we give the results.

\begin{itemize}
  \item[[A1]] $G(w)$ is strictly convex and twice differentiable in an interval
    $\mathcal{V} \in \R$,
  \item[[A2]] There exists $c_1, c_2 \in \mathcal{V}$ with $c_1, c_2 > 0$ such
    that $c_1 < \pi_{2i|1} < c_2$ for $i = 1, \dots, n$,
  \item[[A3]] If $\pi_{2ij|1}$ is joint inclusion probability of elements $i$
    and $j$ and $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$ then,
    $$\limsup_{n \to \infty} \max_{i,j \in A_1: i \neq j} |\Delta_{2ij|1}| <
    \infty,$$
  % TODO: Check if A4 should be stated for the finite population (as written) or
    % if this should be stated for the Phase 1 sample.
  \item[[A4]] Assume that $\Sigma_z = \lim_{N \to \infty} \sum_{i \in U} z_i
    z_i^T/N$ exists and is positive definite, the average fourth moment of $(y_i,
    x_i^T)$ is finite ($\limsup_{N \to \infty} \sum_{i \in U} ||(y_i, x_i^T)||^4
    / N < \infty$), and $\Gamma(\lambda) = \lim_{N \to \infty}  \sum_{i \in U}
    f'(\lambda^T z_i) z_i z_i^T / N$ exists in a neighborhood around $\lambda_0
    = (\bf 0, 1)$.
  \item[[A5]] FIXME: We need to make an assumption about how Phase 2 relates to
    Phase 1 which relates to the finite population.
\end{itemize}

\begin{theorem}{Design Consistency}
  Suppose that Conditions [A1] - [A5] hold. Then the solution $\hat w$ to 
  \ref{eq:primalloss} subject to \ref{eq:calconst1} and \ref{eq:calconst2} 
  exists and is unique with probability approaching 1. Furthermore, the
  estimator $\hat \theta_{DCE} = \sum_{i \in A_2} \frac{\hat w y_i}{\pi_i^{(1)}}$
  satisfies

  $$\hat \theta_{DCE} = \hat \theta_{DC} + o_p(FIXME)$$

  where $\hat \theta_{DC}$ is the debiased calibration estimator of 
  \cite{kwon2024debiased}.

\end{theorem}

\begin{theorem}(Asymptotic Normality)

\end{theorem}

% TODO: Provide asymptotic results showing the variance estimation of the
% estimated debiased calibration method
% DUE: Thursday April 11 (I have already showed the proof but I might finish the
% simulation study before writing this up.)

\subsection*{Simulation Studies}

% TODO: Conduct a simulation study highlighting the proposed method
% DUE: Friday April 12

\section*{Topic 2: Non-nested Two-Phase Sampling}

\section*{Topic 3: Multi-source Two-Phase Sampling}

\printbibliography

\end{document}

### 2024-04-11 final/reports/prelim_paper_ov.tex
Writing prelim_paper added some of Dr. Kim's comments.

--- 2024-04-11 final/reports/prelim_paper.tex
@@
 
-We follow the approach of \cite{kwon2024debiased} for the debiased calibration
-method. We consider maximizing the generalized entropy \cite{gneiting2007strictly},
-
-\begin{equation}\label{eq:primalloss}
-H(w) = - \sum_{i \in A_2} G(w_i)
-\end{equation}
-
-where $G: \mathcal{V} \to \R$ is strictly convex, differentiable function
-subject to the constraints:
-
-\begin{equation}\label{eq:calconst1}
-\sum_{i \in A_2} \frac{x_i w_i}{\pi_i^{(1)}} = 
-\sum_{i \in A_1} \frac{x_i}{\pi_i^{(1)}}
-\end{equation}
-
-and
-
-\begin{equation}\label{eq:calconst2}
-\sum_{i \in A_2} \frac{g(\pi_{2i|1}^{-1})w_i}{\pi_i^{(1)}} = 
-\sum_{i \in A_1} \frac{g(\pi_{2i|1}^{-1})}{\pi_i^{(1)}}.
-\end{equation}
-
-The first constraint is the existing calibration constraint and the second
-ensures that design consistency is achieved. Here, $g(w) = \partial G / \partial
-w$. The original method of \cite{kwon2024debiased} only considered having finite
-population quantities on the right hand side of \ref{eq:calconst1}. Let $z_i =
-(x_i, g(\pi_{2i|1}^{-1})$. To solve this minimization problem, we use the convex
-conjugate function $F(w) = -G(g^{-1}) + g^{-1}(w)w$, and instead find
-
-$$ \hat \lambda = \argmin_{\lambda \in \Lambda_A} \sum_{i \in A_2} F(w_i) +
-\lambda_1^T \sum_{i \in A_1} \frac{z_i}{\pi_{2i|1}}$$
-
-for $\Lambda_A = \{\lambda: \lambda^T z_i \in g(\mathcal{V}) \text{ for all } i
-\in A_2\}$.
+% TODO: Add detailed method for the setup of the debiased calibration method and
+% explain how we estimate the weights in two-phase sampling.
+% DUE: Thursday April 11
 
 \subsection*{Theoretical Results}
-
-We prove two results in this section is the same spirit as \cite{kwon2024debiased}.
-First, we state the assumptions of our analysis then we give the results.
-
-\begin{itemize}
-  \item[[A1]] $G(w)$ is strictly convex and twice differentiable in an interval
-    $\mathcal{V} \in \R$,
-  \item[[A2]] There exists $c_1, c_2 \in \mathcal{V}$ with $c_1, c_2 > 0$ such
-    that $c_1 < \pi_{2i|1} < c_2$ for $i = 1, \dots, n$,
-  \item[[A3]] If $\pi_{2ij|1}$ is joint inclusion probability of elements $i$
-    and $j$ and $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$ then,
-    $$\limsup_{n \to \infty} \max_{i,j \in A_1: i \neq j} |\Delta_{2ij|1}| <
-    \infty,$$
-  % TODO: Check if A4 should be stated for the finite population (as written) or
-    % if this should be stated for the Phase 1 sample.
-  \item[[A4]] Assume that $\Sigma_z = \lim_{N \to \infty} \sum_{i \in U} z_i
-    z_i^T/N$ exists and is positive definite, the average fourth moment of $(y_i,
-    x_i^T)$ is finite ($\limsup_{N \to \infty} \sum_{i \in U} ||(y_i, x_i^T)||^4
-    / N < \infty$), and $\Gamma(\lambda) = \lim_{N \to \infty}  \sum_{i \in U}
-    f'(\lambda^T z_i) z_i z_i^T / N$ exists in a neighborhood around $\lambda_0
-    = (\bf 0, 1)$.
-  \item[[A5]] FIXME: We need to make an assumption about how Phase 2 relates to
-    Phase 1 which relates to the finite population.
-\end{itemize}
-
-\begin{theorem}{Design Consistency}
-  Suppose that Conditions [A1] - [A5] hold. Then the solution $\hat w$ to 
-  \ref{eq:primalloss} subject to \ref{eq:calconst1} and \ref{eq:calconst2} 
-  exists and is unique with probability approaching 1. Furthermore, the
-  estimator $\hat \theta_{DCE} = \sum_{i \in A_2} \frac{\hat w y_i}{\pi_i^{(1)}}$
-  satisfies
-
-  $$\hat \theta_{DCE} = \hat \theta_{DC} + o_p(FIXME)$$
-
-  where $\hat \theta_{DC}$ is the debiased calibration estimator of 
-  \cite{kwon2024debiased}.
-
-\end{theorem}
-
-\begin{theorem}(Asymptotic Normality)
-
-\end{theorem}
 

### 2024-04-11 src/explore/20240313-eststeps.qmd
Writing prelim_paper added some of Dr. Kim's comments.

---
title: "Steps in Estimation Process of Optimal GLS Estimator"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    code-fold: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

```{r}
#| label: global variables
B <- 1000

obs_seg <- 250
true_mu <- 5
rho <- 0.5

```

```{r}
#| label: setup

library(dplyr)

```

\newcommand{\Var}{\text{Var}}
\newcommand{\Cov}{\text{Cov}}

# Summary

For the nonmonotone GLS estimation we need to estimate the following:

1. The functional form of $f(G_r(X)) = E[Y \mid G_r(X)]$,
2. The variance, $\hat{V}(E[Y \mid G_r(X)])$,
3. The function, $\hat E[Y \mid G_r(X)]$, and 
4. The goal, $\hat \theta = E[Y]$.

In [Optimal Estimation with Nuisance Parameters](20240216-optest.qmd), we tested
to see if we could
get an unbiased estimate of $\hat \theta$ when we estimated the covariance
matrix. This report is similar. We compare the following:

* The Oracle estimator of $\hat \theta$ $ = E[Y] = n^{-1} \sum_{i = 1}^n y_i$,
* The Complete case estimator of $\hat \theta$,
* The IPW estimator of $\hat \theta$,
* The GLS estimate with correctly specified $E[Y \mid G_r(X)]$, and correctly
specified $V(E[Y \mid G_r(X)])$,
* The GLS estimate with estimated $\hat E[Y \mid G_r(X)]$, and correctly
specified $V(E[Y \mid G_r(X)])$, and 
* The GLS estimate with estimated $\hat E[Y \mid G_r(X)]$, and estimated
$\hat V(E[Y \mid G_r(X)])$.

These estimators are compared to each other under the nonmonotone complete case
and variable setting.

# Simulation Setup

## Nonmonotone Complete Case and Variable

For this nonmonotone case, we have three segments $A_1$, $A_2$, and $A_3$ with
missing patterns shown in @tbl-four.

| Segment | $X_1$ | $X_2$ | $Y$ |
|----|-------|-------|---|
| $A_1$ | ✔️     | ✔️     | :heavy_check_mark: |
| $A_2$ | ✔️     | ✔️     |   |
| $A_3$ | ✔️     |       | :heavy_check_mark: | 
: Nonmonotone: Complete Case + Complete Variable {#tbl-four}

The estimating vector is the following

$$
\begin{bmatrix} 
\hat f_1(X_1) \\ 
\hat f_1(X_2) \\ 
\hat f_1(X_1, X_2) \\ 
\hat Y_1 \\ 
\hat f_2(X_1) \\ 
\hat f_2(X_2) \\ 
\hat f_2(X_1, X_2) \\ 
\hat f_3(X_1) \\ 
\hat Y_3 
\end{bmatrix} =
\begin{bmatrix}
n_1^{-1} \sum_{i = 1}^n \frac{\delta_{1i}E[Y \mid x_{1i}]}{\pi_{1i}} \\
n_1^{-1} \sum_{i = 1}^n \frac{\delta_{1i}E[Y \mid x_{2i}]}{\pi_{1i}} \\
n_1^{-1} \sum_{i = 1}^n \frac{\delta_{1i}E[Y \mid x_{1i}, x_{2i}]}{\pi_{1i}} \\
n_1^{-1} \sum_{i = 1}^n \frac{\delta_{1i}y_i}{\pi_{1i}} \\
n_2^{-1} \sum_{i = 1}^n \frac{\delta_{2i}E[Y \mid x_{1i}]}{\pi_{2i}} \\
n_2^{-1} \sum_{i = 1}^n \frac{\delta_{2i}E[Y \mid x_{2i}]}{\pi_{2i}} \\
n_2^{-1} \sum_{i = 1}^n \frac{\delta_{2i}E[Y \mid x_{1i}, x_{2i}]}{\pi_{2i}} \\
n_3^{-1} \sum_{i = 1}^n \frac{\delta_{3i}E[Y \mid x_{1i}]}{\pi_{3i}} \\
n_3^{-1} \sum_{i = 1}^n \frac{\delta_{3i}y_i}{\pi_{3i}} \\
\end{bmatrix}.
$$

## Implementation Details

For this setup, we consider segments of equal size 
$n_1 = n_2 = n_3 =$ `{r} obs_seg` and a
data generating process of

$$
\begin{bmatrix} X_1 \\ \varepsilon_1 \\ \varepsilon_2 \end{bmatrix} \sim 
\left(
\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix},
\begin{bmatrix} 
1 & 0 & 0 \\ 
0 & 1 & \rho \\ 
0 & \rho & 1 \\
\end{bmatrix}
\right)
$$

in which $X_2 = X_1 + \varepsilon_1$ and $Y = \mu_Y + X_1 + \varepsilon_2$.
While we can vary the parameters of $\mu_Y$ and $\rho$, for this study we have,
$\mu_Y =$ `{r} true_mu`, and $\rho =$ `{r} rho`.

```{r}
#| label: data generating process

#' A data generation function to create a simple random sample of equally sized
#' segments of X_1, X_2, and Y. There variables are correlated but all normal.
#'
#' @param n_obs_seg - An integer greater than zero. This is the number of
#' observations in each segment
#' @param mu - The true mean of Y
#' @param rho - The correlation between (X_1 - X_2) and (Y - X_1).
#'
#' @details We always observe X1 but X2 and Y are only observed depending on the
#' value of \delta_{ij}. If i = 1 then X2 is observed. If j = 1 then Y is
#' observed.
gen_simple_data <- function(n_obs_seg, mu, rho) {

  # This code is modified from the case where we had X, Y_1, and Y_2. It only
  # changes at the end to X_1, X_2, and Y.
  # A_11
  x <- rnorm(n_obs_seg)
  e_mat <- MASS::mvrnorm(n_obs_seg, c(0, 0), matrix(c(1, rho, rho, 1), nrow = 2))
  e1 <- e_mat[, 1]
  e2 <- e_mat[, 2]
  y1 <- x + e1
  y2 <- mu + x + e2

  df_11 <- 
    tibble(X1 = x, X2 = y1, Y = y2,
           delta_00 = 0, delta_10 = 0, delta_01 = 0, delta_11 = 1,
           pi_00 = 0.25, pi_10 = 0.25, pi_01 = 0.25, pi_11 = 0.25)
  # A_10
  x <- rnorm(n_obs_seg)
  e_mat <- MASS::mvrnorm(n_obs_seg, c(0, 0), matrix(c(1, rho, rho, 1), nrow = 2))
  e1 <- e_mat[, 1]
  e2 <- e_mat[, 2]
  y1 <- x + e1
  y2 <- mu + x + e2

  df_10 <- 
    tibble(X1 = x, X2 = y1, Y = y2,
           delta_00 = 0, delta_10 = 1, delta_01 = 0, delta_11 = 0,
           pi_00 = 0.25, pi_10 = 0.25, pi_01 = 0.25, pi_11 = 0.25)
  # A_01
  x <- rnorm(n_obs_seg)
  e_mat <- MASS::mvrnorm(n_obs_seg, c(0, 0), matrix(c(1, rho, rho, 1), nrow = 2))
  e1 <- e_mat[, 1]
  e2 <- e_mat[, 2]
  y1 <- x + e1
  y2 <- mu + x + e2

  df_01 <- 
    tibble(X1 = x, X2 = y1, Y = y2,
           delta_00 = 0, delta_10 = 0, delta_01 = 1, delta_11 = 0,
           pi_00 = 0.25, pi_10 = 0.25, pi_01 = 0.25, pi_11 = 0.25)
  # A_00
  x <- rnorm(n_obs_seg)
  e_mat <- MASS::mvrnorm(n_obs_seg, c(0, 0), matrix(c(1, rho, rho, 1), nrow = 2))
  e1 <- e_mat[, 1]
  e2 <- e_mat[, 2]
  y1 <- x + e1
  y2 <- mu + x + e2

  # No empty segment
  # df_00 <- 
  #   tibble(X1 = x, X2 = y1, Y = y2,
  #          delta_00 = 1, delta_10 = 0, delta_01 = 0, delta_11 = 0,
  #          pi_00 = 0.25, pi_10 = 0.25, pi_01 = 0.25, pi_11 = 0.25)

  df <- bind_rows(df_11, df_10, df_01)

  # Return
  # X, Y_1, Y_2,
  # delta_00, delta_10, delta_01, delta_11,
  # pi_00, pi_10, pi_01, pi_11
  return(df)

}

```

## Estimation Details

We assume that we always use the correct functional form of $E[Y \mid G_r(X)]$.
The correctly specified and estimated forms of $f(G_r(X))$ are given in
@tbl-meanform

| Function | Correctly Specified Form | Estimated Form |
| -------- | :----------------------: | :------------: |
| $E[Y \mid X_1]$ | $\hat \mu + x_1$  | $\hat \beta_0 + \hat \beta_1 x_1$ |
| $E[Y \mid X_2]$ | $\hat \mu + \left(\frac{1 + \rho}{2}\right)x_2$ | $\hat \beta_0 + \hat \beta_2 x_2$ |
| $E[Y \mid X_1, X_2]$ | $\hat \mu + x_1 + \rho (x_2 - x_1)$ | $\hat \beta_0 +
\hat \beta_1 x_1 + \hat \beta_2 x_2$ |
: {#tbl-meanform}

In @tbl-meanform, the estimation of $\hat \mu$ for the correctly specified form
is always carried out as the mean of a difference. For example, to estimate 
the $\hat \mu$ in $E[Y \mid X_1]$, we use

$$ \hat \mu = \frac{\sum_{i = 1}^n \frac{\delta_{2i}}{\pi_{2i}}(y_i -
x_{1i})}{\sum_{i = 1}^n \frac{\delta_{2i}}{\pi_{2i}}}. $$

The estimation of $\beta_i$ in the `Estimated Form` column of @tbl-meanform, is
computed using least squares on all of the observations containing all of the
variables $G_r(X)$ and $Y$.

Likewise, we have a similar result for the covariance matrix in @tbl-covform.

| Covariance | Correctly Specified Form | Estimated Form |
| ---------- | :----------------------: | :------------: |
| $\Var(E[Y \mid X_1])$ | $1$ | $\hat \sigma_{X_1}^2$ |
| $\Var(E[Y \mid X_2])$ | $\frac{(1 + \rho)^2}{2}$ | $\left(\frac{1 + \rho}{2}\right)^2
\hat \sigma_{X_2}^2$ |
| $\Var(E[Y \mid X_1, X_2])$ | $1 + \rho^2$ | $\sigma_{X_1}^2 + \rho^2$ |
| $\Var(Y)$ | $2$ | $\hat \sigma_y^2$ |
| $\Cov(E[Y \mid X_1], E[Y \mid X_2])$ | $\frac{1 + \rho}{2}$ | $\left(\frac{1 +
\rho}{2}\right)\hat \sigma_{X_1}^2$ |
| $\Cov(E[Y \mid X_1], E[Y \mid X_1, X_2])$ | $1$ | $\hat \sigma_{X_1}^2$ |
| $\Cov(E[Y \mid X_2], E[Y \mid X_1, X_2])$ | $\frac{(1 + \rho)^2}{2}$ | $\frac{(1 + \rho)^2}{2}$ |
| $\Cov(E[Y \mid X_1], Y)$ | $1$ | $\hat \sigma_{X_1}^2$ |
| $\Cov(E[Y \mid X_2], Y)$ | $\frac{(1 + \rho)^2}{2}$ | $\frac{(1 + \rho)^2}{2}$ |
| $\Cov(E[Y \mid X_1, X_2], Y)$ | $1 + \rho^2$ | $\hat \sigma_{X_1}^2 + \rho^2$ |
: {#tbl-covform}

The estimated variances $\hat \sigma_{X_1}^2$, $\hat \sigma_{X_2}^2$, and $\hat
\sigma_Y^2$ are estimated from all of the observations in which that value
occurs.

In @tbl-covform, I don't want to have to estimate $\rho$ because I am not sure
what the REML estimator of $\rho$ is at the moment. If the REML estimated form
for the variances works, then I will consider how to estimate the covariance.

```{r}
#| label: cor mean, cor var

cormeancorvar <- function(df, rho) {

  # Steps:
  # 1. Separate into segments
  # 2. Construct estimates for each segment
  # 3. Combine estimates

  # 1. Separate into segments
  df_11 <- filter(df, delta_11 == 1)
  df_10 <- filter(df, delta_10 == 1)
  df_01 <- filter(df, delta_01 == 1)

  # 2. Construct estimates for each segment
  # 3. Combine estimates

}

```

```{r}
#| label: est mean, cor var

estmeancorvar <- function(df, rho) {

}

```

```{r}
#| label: est mean, est var

estmeanestvar <- function(df, rho) {

}

```

```{r}
#| label: MCMC

df <- gen_simple_data(obs_seg, true_mu, rho)

```

### 2024-04-11 src/explore/20240405-glsdcsim.qmd
Writing prelim_paper added some of Dr. Kim's comments.

---
title: "GLS-Debiased Calibration with Estimated Population Totals Simulation"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    code-fold: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

# Summary

This report simulates the GLS-DC setup. We check for the following:

1. Bias of $\hat Y_1$ and $\hat Y_2$, and 
2. Bias of $\hat V(\hat Y_1)$ and $\hat V(\hat Y_2)$.

For each estimator $\hat Y_1$ and $\hat Y_1$, we construct

1. The two-phase regression estimator,
2. The Deville and Sarndal (1992) estimator,
3. The debiased calibration estimator, and
4. The debiased calibration estimator with estimated population totals.

# Simulation Setup

```{r}
#| label: setup

N_obs <- 10000
n1_obs <- 500
n2_obs <- 500

library(dplyr)
library(GECal)
library(doParallel)

```

We have the following superpopulation model:

$$
\begin{aligned}
X_{1i} &= 1 + e_{1i} \\
X_{2i} &= 2 + e_{2i} \\
X_{3i} &= 3 + e_{3i} \\
Y_{1i} &= 2 + 3x_{1i} + x_{2i} + e_{y_1, i}} \\
Y_{2i} &= 1 + 2x_{1i} + 2 x_{2i} + x_{3i} + e_{y_2, i}\\
\end{aligned}
$$

where $e_{m, i} \stackrel{ind}{\sim} N(0, 1)$. Then we generate a finite
population of size $N = $`{r} N_obs`. Sample $A_1$ has size $n_1 =$`{r} n1_obs`
and sample $A_2$ has size $n_2 =$`{r} n2_obs`. Each sample is an independent
Poisson from the finite population.

```{r}
#| label: data generating process

gen_data <- function(obs) {
  tibble(X1 = 1 + rnorm(obs),
         X2 = 2 + rnorm(obs),
         X3 = 3 + rnorm(obs)) %>%
    mutate(Y1 = 2 + 3 * X1 + X2 + rnorm(obs)) %>%
    mutate(Y2 = 1 + 2 * X1 + 2 * X2 + X3 + rnorm(obs))

}

gen_samples <- function(pop_df, n1, n2) {

  a1_ind <- sample(1:nrow(pop_df), n1)
  a2_ind <- sample(1:nrow(pop_df), n2)

  a1_df <- filter(pop_df, row_number() %in% a1_ind)
  a2_df <- filter(pop_df, row_number() %in% a2_ind)

  a1_df <- select(a1_df, X1, X2, Y1)
  a2_df <- select(a2_df, X1, X3, Y2)

  return(list(A1 = a1_df, A2 = a2_df))

}

```

# Estimation

```{r}
#| label: MCMC

set.seed(1)
pop_df <- gen_data(N_obs)
samps <- gen_samples(pop_df, n1_obs, n2_obs)
s1_df <- samps[[1]]
s2_df <- samps[[2]]

```

### 2024-04-11 src/explore/20240408-redokwongsim.qmd
Writing prelim_paper added some of Dr. Kim's comments.

---
title: "Debiased Calibration Simulation"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    code-fold: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

# Summary

This report simulates part of the debiased calibration setup. I need to do this
to ensure that I understand how to use their package. I will compare their
method with the @deville1992calibration as done in the simulation section of 
@kwon2024debiased.

# Simulation Setup

```{r}
#| label: setup

N_obs <- 10000
sample_draws <- 1000

library(dplyr)
library(sampling)
library(GECal)
library(doParallel)
library(doRNG)

```

```{r}
#| label: data generation

gen_pop <- function(obs, method = 1) {
  x1 <- rnorm(obs, 2, 1)
  x2 <- runif(obs, 0, 4)
  mu_x <- x1 + x2
  z <- rnorm(obs, 0, 1)
  if (method == 1) {
    y <- mu_x + z + rnorm(obs, 0, 1)
  } else if (method == 2) {
    y <- mu_x + z^2 + rnorm(obs, 0, 1)
  } else {
    stop("method must be 1 or 2")
  }

  pi_vec <- ifelse(pt(-z - 2, 3) > 0.7, 0.7, pt(-z - 2, 3))

  return(tibble(X1 = x1, X2 = x2, Y = y, Prob_i = pi_vec))
}

gen_samp <- function(pop_df) {
  delta <- sampling::UPpoisson(pop_df$Prob_i)
  return(filter(pop_df, row_number() %in% which(delta == 1)))
}

```

```{r}
#| label: temp

f = function(lambda, d, Xs, total, entropy, del, ..., returnw = F){
  # w = d * ginv(drop(Xs %*% lambda), entropy = entropy)
  if(entropy == "HD" & any(Xs %*% lambda >= 0)) return(rep(Inf, length(lambda)))
  if(entropy == "PH" & any(abs(Xs %*% lambda) >= del)) return(rep(Inf, length(lambda)))

  if(entropy == "SL"){
    w = d * drop(Xs %*% lambda)
  }else if(entropy == "EL"){
    w = -d / drop(Xs %*% lambda)
  }else if(entropy == "ET"){
    w = d * exp(drop(Xs %*% lambda))
  }else if(entropy == "CE"){
    w = d / (1 - exp(drop(Xs %*% lambda)))
  }else if(entropy == "HD"){
    w = d / drop(Xs %*% lambda)^2
  }else if(entropy == "PH"){
    w = d / sqrt(1 / drop(Xs %*% lambda)^2 - 1 / del^2)
  }
  if(entropy != "SL" & any(w <= 0)) return(rep(Inf, length(lambda)))
  if(entropy == "CE" & any(w <= 1)) return(rep(Inf, length(lambda)))
  if(returnw == T){
    return(w)
  }else{
    return(colSums(Xs * w) - total)
  }
}

h = function(lambda, d, Xs, total, entropy, del){
  # return(t(Xs) %*% (Xs * d * ginvprime(drop(Xs %*% lambda), entropy = entropy)))
  if(entropy == "SL"){
    return(t(Xs) %*% (Xs * d))
  }else if(entropy == "EL"){
    w = -d / drop(Xs %*% lambda)
    return(t(Xs) %*% (Xs * (w^2 / d)))
  }else if(entropy == "ET"){
    w = d * exp(drop(Xs %*% lambda))
    return(t(Xs) %*% (Xs * w))
  }else if(entropy == "CE"){
    p_Stmp = 1 / (1 - exp(drop(Xs %*% lambda)))
    return(t(Xs) %*% (Xs * d * p_Stmp * (p_Stmp - 1)))
  }else if(entropy == "HD"){
    return(t(Xs) %*% (Xs * (-2 * d / drop(Xs %*% lambda)^3)))
  }else if(entropy == "PH"){
    return(t(Xs) %*% (Xs * (d * (1 - (drop(Xs %*% lambda) / del)^2)^(-1.5))))
  }
}
#' @export
GEcalib = function(Xs, d, total, entropy = c("SL", "EL", "ET", "CE", "HD", "PH"), ...,
       DS = T, method = "Newton", control = list(maxit = 1e5, allowSingular = T)){

  del = quantile(d, 0.80)

  init = rep(0, length(total))
  if(!DS){
    init[length(init)] = 1
    # init = c(0, 0, 1)
  }else{
    if(entropy == "EL"){
      init[1] = -1
    }else if(entropy == "ET"){
      # init[1] = 0
    }else if(entropy == "SL"){
      init[1] = 1
    }else if(entropy == "CE"){
      init[1] = -1
    }else if(entropy == "HD"){
      init[1] = -1
    }else if(entropy == "PH"){
      init[1] = 1 / sqrt(1 + 1 / del^2)
    }
  }

  nleqslv_res = nleqslv(init, f, jac = h, d = d, Xs = Xs, 
                        total = total, entropy = entropy, del = del,
                        method = method, control = control)
  if(nleqslv_res$termcd != 1){
    if(max(abs(f(nleqslv_res$x, d = d, Xs = Xs, total = total,
                 entropy = entropy, del = del))) > 1e-5)
      print(nleqslv_res)
      w = NA
  }else{
    w = f(nleqslv_res$x, d = d, Xs = Xs, total = total,
          entropy = entropy, del = del, returnw = T)             
  }
  return(w)
}

```

```{r}
#| label: estimation

set.seed(1)
pop_df <- gen_pop(N_obs)
pop_Xtot <- c(nrow(pop_df), sum(pop_df$X1), sum(pop_df$X2))
pop_d <- 1 / pop_df$Prob_i

clust <- makeCluster(min(detectCores() - 2, 100))
registerDoParallel(clust)

mc_res <- foreach(ind = 1:sample_draws, .packages = c("dplyr", "GECal")) %dorng% {

# Sample
samp_df <- gen_samp(pop_df)
samp_X <- samp_df %>% select(X1, X2) %>% as.matrix()
samp_X <- cbind(rep(1, nrow(samp_X)), samp_X)
d_weights <- 1 / samp_df$Prob_i

## Method: EL
method = "EL"
if (method == "EL") {
  dtot <- sum(-1 / pop_d)
  dsamp <- -1 / d_weights
  samp_df <- samp_df %>%
    mutate(gdi = dsamp) %>%
    mutate(gpinv = 1 / Prob_i^2)
  ds_w <- GEcalib(cbind(samp_X, dsamp), d = d_weights, c(pop_Xtot, dtot),
    entropy = "EL", DS = TRUE)
  dc_w <- GEcalib(cbind(samp_X, dsamp), d = rep(1, nrow(samp_df)), c(pop_Xtot, dtot),
   entropy = "EL", DS = FALSE)
} else if (method == "ET") {
  dtot <- sum(log(d_weights))
  dsamp <- log(d_weights)
  samp_df <- samp_df %>%
    mutate(gdi = dsamp) %>%
    mutate(gpinv = 1 / Prob_i)
  ds_w <- GEcalib(cbind(samp_X, dsamp), d = d_weights, c(pop_Xtot, dtot),
    entropy = "ET", DS = TRUE)
  dc_w <- GEcalib(cbind(samp_X, dsamp), d = d_weights, c(pop_Xtot, dtot),
    entropy = "ET", DS = FALSE)
}

# Mean Estimation
y_ds <- sum(samp_df$Y * ds_w) / N_obs
y_dc <- sum(samp_df$Y * dc_w) / N_obs
y_hajek <- sum(samp_df$Y / samp_df$Prob_i) / sum(1 / samp_df$Prob_i)

## Variance Estimation
#ds_mod <- lm(Y ~ X1 + X2, weights = gpinv, data = samp_df)
#dc_mod <- lm(Y ~ X1 + X2 + gdi, weights = gpinv, data = samp_df)
#
#v_ds <- sum((1 - samp_df$Prob_i) * (ds_w / N_obs)^2 *
#    (samp_df$Y - predict(ds_mod, samp_df))^2)
#v_dc <- sum((1 - samp_df$Prob_i) * (dc_w / N_obs)^2 *
#    (samp_df$Y - predict(dc_mod, samp_df))^2)
v_hajek <- 1 / N_obs^2 *
sum((samp_df$Y - y_hajek)^2 * (1 / samp_df$Prob_i^2 - 1 / samp_df$Prob_i))

Pi_ds <- diag(1 - samp_df$Prob_i) * (ds_w / N_obs)^2
z_ds <- model.matrix(~X1 + X2 + gdi, data = samp_df)
gam_ds <- solve(t(z_ds) %*% diag(1 / samp_df$Prob_i) %*% z_ds,
                t(z_ds) %*% diag(1 / samp_df$Prob_i) %*% samp_df$Y)
v_ds <- (samp_df$Y - t(gam_ds) %*% t(z_ds)) %*% 
             Pi_ds %*% t(samp_df$Y - t(gam_ds) %*% t(z_ds))

Pi_dc <- diag(1 - samp_df$Prob_i) * (dc_w / N_obs)^2
z_dc <- model.matrix(~X1 + X2 + gdi, data = samp_df)
gam_dc <- solve(t(z_dc) %*% diag(samp_df$gpinv) %*% z_dc,
                t(z_dc) %*% diag(samp_df$gpinv) %*% samp_df$Y)
v_dc <- 
  (samp_df$Y - t(gam_dc) %*% t(z_dc)) %*% Pi_dc %*% t(samp_df$Y - t(gam_dc) %*% t(z_dc))

return(tibble(y_ds = y_ds, y_dc = y_dc, y_h = y_hajek,
              v_ds = as.numeric(v_ds), v_dc = as.numeric(v_dc), v_h = v_hajek))

} %>% bind_rows()

stopCluster(clust)

## Assessment
bias_ds <- mean(mc_res$y_ds - mean(pop_df$Y))
bias_dc <- mean(mc_res$y_dc - mean(pop_df$Y))
bias_h <- mean(mc_res$y_h - mean(pop_df$Y))

rmse_ds <- sqrt(mean((mc_res$y_ds - mean(pop_df$Y))^2))
rmse_dc <- sqrt(mean((mc_res$y_dc - mean(pop_df$Y))^2))
rmse_h <- sqrt(mean((mc_res$y_h - mean(pop_df$Y))^2))

cr_ds <- mean(abs(mc_res$y_ds - mean(pop_df$Y)) < qnorm(0.975) * sqrt(mc_res$v_ds))
cr_dc <- mean(abs(mc_res$y_dc - mean(pop_df$Y)) < qnorm(0.975) * sqrt(mc_res$v_dc))
cr_h <- mean(abs(mc_res$y_h - mean(pop_df$Y)) < qnorm(0.975) * sqrt(mc_res$v_h))

```

### 2024-04-11 src/explore/Sampling.R

@@ -1,0 +1,717 @@
+# Simulation setup
+
+if(!interactive()){
+  args <- as.numeric(commandArgs(trailingOnly = TRUE))
+}else{
+  args <- c(5)
+}
+
+timenow1 = Sys.time()
+timenow0 = gsub(' ', '_', gsub('[-:]', '', timenow1))
+timenow = paste(timenow0, ".txt", sep = "")
+
+# install.packages( "MatrixModels", type="win.binary" )  
+
+library(np)
+# library(ggplot2)
+library(nleqslv)
+library(knitr)
+# suppressMessages(library(CVXR))
+# suppressMessages(library(kableExtra))
+suppressMessages(library(dplyr))
+suppressMessages(library(foreach))
+suppressMessages(library(doParallel))
+
+set.seed(11)
+SIMNUM = args[1]
+
+if(!interactive()){
+  dir.create(timenow0)
+  setwd(timenow0)
+  
+  sink(timenow, append=TRUE)
+}
+# setup parallel backend to use many processors
+cores = min(detectCores() - 3, 101)
+print(paste("cores =", cores))
+# cl <- makeCluster(cores, outfile = timenow) #not to overload your computer
+cl <- makeCluster(cores)
+registerDoParallel(cl)
+
+# N = 100000
+# x = matrix(rnorm(N, 1, 1), nc = 1)
+# e = rnorm(N, 0, 1)
+# beta = c(1, 1)
+# y = beta[1] + x[,1] * beta[2] + e
+# # y = beta[1] + x[,1]^2 * beta[2] + e
+# # y = (x[,1] - 1)^2 + e
+# # y = beta[1] + 1 / (x[,1]^2 + 1)* beta[2] + e
+# 
+# En = 200
+# pi = 1 / ((x + y + 10) / 100)
+# pi = pi / sum(pi) * En
+
+# pi = 1 / (1 + exp(-(-6 + .2 * x[,1]^2)))
+# pi = 1 / (1 + exp(-(-4 + .2 * x[,1]^2)))
+# pi = 1 / (1 + exp(-(-6 -  0.2 * x[,1]^2)))
+# pi = 1 / (1 + exp(-(-6 -  0.2 * x[,1]^2)))
+# pi = 1 / (1 + exp(-(rep(-6, N))))
+# pi = 1 / (1 + exp(-(-7 + .2 * x[,1] + 0.2 * y)))
+# pi = 1 / (1 + exp(-(-6 + 0.2 * e)))
[…]

### 2024-04-11 final/reports/prelim_paper.tex
Working on TP DC simulation. About to switch from SRS to Poisson sampling because then I can actually use the debiased calibration. Hopefully, the GEcalib function will then work.

@@
 
-\title{Debiased calibration for generalized two-Phase sampling}
+\title{Debiased Calibration for Generalized Two-Phase Sampling}
 \author{Caleb Leedy}
@@
 \begin{itemize}
-\item Combining information from several sources is an important practical problem. 
-\item In many cases, we do not have direct access to the other sources.
-We can only obtain summary statistics (and their standard errors) for the external data sources.
-We wish to incorporate the information from external  sources in our in-house data effectively using calibration weighting.
-\item We formulate the problem as a generalized two-phase sampling where the first phase sample can be obtained from multiple sources.
-The second-phase sample is our in-house data in which we want to construct the calibration weights.
-\item To achieve the goal, we first consider  the  classical two-phase sampling setup where the second-phase sample is a subset of the first-phase sample.
-After that, we extend the setup to more general cases such as non-nested two-phase sampling or multiple independent surveys with some common measurements.
-\item (To simplify the presentation, I wonder whether we can just use SRS in the first-phase sampling. ) 
+\item Combining information from several sources is an important practical
+  problem. 
+\item In many cases, we do not have direct access to the other sources. We can
+  only obtain summary statistics (and their standard errors) for the external
+  data sources. We wish to incorporate the information from external  sources in
+  our in-house data effectively using calibration weighting.  
+\item We formulate the problem as a generalized two-phase sampling where the
+  first phase sample can be obtained from multiple sources. The second-phase
+  sample is our in-house data in which we want to construct the calibration
+  weights. 
+\item To achieve the goal, we first consider  the  classical two-phase sampling
+  setup where the second-phase sample is a subset of the first-phase sample.
+  After that, we extend the setup to more general cases such as non-nested
+  two-phase sampling or multiple independent surveys with some common
+  measurements. 
+\item (To simplify the presentation, I wonder whether we can just use SRS in the
+  first-phase sampling. ) 
 \end{itemize}
@@
 
-\textcolor{red}{Dr. Kim: Should I provide the proof for design unbiasedness of
-$\hat Y_{\pi^*}$? } \textcolor{blue}{Not necessary. You may cite a book or a paper  for this claim. }
-
-While unbiased, the $\pi^*$-estimator unfortunately does not account for the
-additional information contained in the auxiliary Phase 1 variable $X$. The
-two-phase regression estimator $\hat Y_{reg, tp}$ does incorporate \textcolor{blue}{$\hat{X}_1$ obtained from the phase one sample.
-That is, we can leverage the external information $\hat{X}_1$ to improve the $\pi^*$-estimator in the second-phase sample.
-}
-The two-phase 
-regression estimator has the form,
+While unbiased \cite{kim2024statistics}, the $\pi^*$-estimator unfortunately
+does not account for the additional information contained in the auxiliary Phase
+1 variable $X$. The two-phase regression estimator $\hat Y_{reg, tp}$ does incorporate 
+\textcolor{blue}{$\hat{X}_1$ obtained from the phase one sample. That is, we can
+leverage the external information $\hat{X}_1$ to improve the $\pi^*$-estimator
+in the second-phase sample. } 
+The two-phase regression estimator has the form,
 
@@
 loss. They generalized the loss function to minimize $\sum_i G(w_i, d_i)q_i$ for
-weights $w_i$ and design-weights $d_i$ where $G()$ is non-negative, strictly
+weights $w_i$ and design-weights $d_i$ where $G(\cdot)$ is non-negative, strictly
 convex function with respect to $w$, defined on an interval containing $d_i$,
@@
 properties of their estimator from maximum entropy estimation rather than the
-minimal squared loss, unless $\pi_{2i|1}^{-1} \in C(X)$, the estimator is not
-design consistent. \footnote{ \textcolor{red}{I do not understand this part.} } 
+minimal squared loss, 
+% TODO: I am not sure if this is correct -----
+unless $\pi_{2i|1}^{-1} = x_i'a$ for some $a$, the
+estimator is not design consistent. 
+\footnote{ \textcolor{red}{I do not understand this part.} I changed the
+notation so that it would not be the unused column space notation. This is from
+Equation (11.17) of \cite{kim2024statistics}.} 
+% --------------------------------------------
+
 Furthermore, at a conceptual level, the regression estimator
@@
 $$ \hat Y_{reg, tp} 
-= \underbrace{\sum_{i \in A_1} \frac{x_i \hat \beta_2}{\pi_i^{(1)}}}_{Minimizing
-the variance} + \underbrace{\sum_{i \in A_2}
-\frac{1}{\pi_i^{(1)}\pi_{2i|1}^{(2)}} (y_i - x_i \hat \beta_2)}_{Bias
-correction}.$$
+= \underbrace{\sum_{i \in A_1} \frac{x_i \hat \beta_2}{\pi_i^{(1)}}}_{
+  \text{ Minimizing the variance}} + \underbrace{\sum_{i \in A_2}
+\frac{1}{\pi_i^{(1)}\pi_{2i|1}^{(2)}} (y_i - x_i \hat \beta_2)}_{
+\text{Bias correction}}.$$
 
@@
 \begin{theorem}(Asymptotic Normality)
-
+% FIXME: Add result.
 \end{theorem}

### 2024-04-11 src/explore/20240411-tpdcsim.qmd
Working on TP DC simulation. About to switch from SRS to Poisson sampling because then I can actually use the debiased calibration. Hopefully, the GEcalib function will then work.

---
title: "Two Phase Debiased Calibration Simulation"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    code-fold: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

# Summary

This report simulates two-phase sampling with the debiased calibration of
@kwon2024debiased. We check for the following:

1. Bias of $\hat Y$ and,
2. Bias of $\hat V(\hat Y)$.

For the estimator $\hat Y$ we construct

1. The $\pi^*$-estimator,
2. The two-phase regression estimator,
3. The debiased calibration estimator (with known finite population total), and
4. The debiased calibration estimator with estimated population totals from the
   Phase 1 sample.

# Simulation Setup

```{r}
#| label: setup

B_sims <- 1000

N_obs <- 10000
n1_obs <- 1000
n2_obs <- 500

library(dplyr)
library(GECal)
library(doParallel)
library(doRNG)

```

We have the following superpopulation model:

$$
\begin{aligned}
X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
X_{3i} &\stackrel{ind}{\sim} N(-1, 1) \\
\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i.
\end{aligned}
$$

and the two-phase sampling scheme is to select a SRS of size $n = $`{r} n1_obs`
from a finite population of size $N = $`{r} N_obs` for the Phase 1 sample and a
SRS of size $r =$`{r} n2_obs` from the Phase 1 sample of size 
$n = $`{r} n1_obs`.

```{r}
#| label: gen data

gen_pop <- function(obs) {
  x1 <- rnorm(obs, 2, 1)
  x2 <- runif(obs, 0, 4)
  x3 <- rnorm(obs, -1, 1)
  eps <- rnorm(obs)

  y <- 3 * x1 + 2 * x2 + eps

  return(tibble(X1 = x1, X2 = x2, X3 = x3, Y = y))
}

gen_twophase <- function(pop_df, p1_obs, p2_obs) {

  # Phase 1 
  p1_ind <- sample(1:nrow(pop_df), size = p1_obs)
  p1_df <- filter(pop_df, row_number() %in% p1_ind)

  # Phase 2
  p2_ind <- sample(1:nrow(p1_df), size = p2_obs)
  p2_df <- filter(p1_df, row_number() %in% p2_ind)

  return(list(P1 = p1_df, P2 = p2_df))

}

```

```{r}
#| label: estimators

# Each estimator should return two outcomes:
# 1. Point estimate, and
# 2. Variance estimate.
pi_star <- function(p1_df, p2_df, pop_obs, ...) {
  extra <- list(...)
  n1 <- nrow(p1_df)
  n2 <- nrow(p2_df)

  theta <- sum(p2_df$Y / n2)
  vhat <- (1 / n2 - 1 / pop_obs) * var(p2_df$Y)

  return(list(theta = theta, Var = vhat))
}

tp_reg <- function(p1_df, p2_df, pop_obs, ...) {
  extra <- list(...)
  n1 <- nrow(p1_df)
  n2 <- nrow(p2_df)

  # Step 1: Get beta
  mod <- lm(Y ~ X1 + X2, data = p2_df)

  # Step 2: Predict
  theta <- mean(predict(mod, p1_df))

  # Step 3: Get variance estimator
  vhat <-
    (1 / n1 - 1 / pop_obs) * (mod$coefficients[2]^2 * var(p1_df$X1) + 
    mod$coefficients[3]^2 * var(p1_df$X2)) +
    (1 / n2 - 1 / n1) * var(p2_df$Y - predict(mod, p2_df))

  return(list(theta = theta, Var = vhat))

}

dc_poptots <- function(p1_df, p2_df, pop_obs, ...) {
  extra <- list(...)
  n1 <- nrow(p1_df)
  n2 <- nrow(p2_df)

  if ("entropy" %in% names(extra)) {
    entropy <- extra$entropy
  } else {
    entropy <- "EL"
  }

  samp_X <- select(p2_df, X1, X2) %>% as.matrix()
  samp_X <- cbind(rep(1, nrow(p2_df)), samp_X)
  pop_tot <- extra$pop_tot
  # FIXME: This does not work
  # TODO: Change everything to Poisson sample.
  dc_w <-  
    GEcalib(samp_X,
            d = rep(1, nrow(p2_df)),
            total = pop_tot,
            entropy = entropy,
            DS = TRUE)

  # Mean Estimation
  theta <- sum(p2_df$Y * dc_w) / pop_obs

  # Variance Estimation

  Pi_dc <- diag(1 - n2 / pop_obs) * (dc_w / N_obs)^2
  z_dc <- model.matrix(~X1 + X2, data = p2_df)
  gam_dc <- solve(t(z_dc) %*% z_dc, t(z_dc) %*% samp_df$Y)
  vhat <- 
  (p2_df$Y - t(gam_dc) %*% t(z_dc)) %*% Pi_dc %*% t(p2_df$Y - t(gam_dc) %*% t(z_dc))

  return(list(theta = theta, Var = vhat))
}

dc_esttots <- function(p1_df, p2_df, pop_obs, ...) {
  extra <- list(...)
  n1 <- nrow(p1_df)
  n2 <- nrow(p2_df)

}

```

```{r}
#| label: MCMC

set.seed(1)
pop_df <- gen_pop(N_obs)
true_theta <- mean(pop_df$Y)

clust <- makeCluster(min(detectCores() - 2, 100))
registerDoParallel(clust)

mc_res <- foreach(iter=1:B_sims, .packages = c("dplyr", "GECal")) %dorng% {

  samps <- gen_twophase(pop_df, n1_obs, n2_obs)
  p1_df <- samps[[1]]
  p2_df <- samps[[2]]

  pistar_est <- pi_star(p1_df, p2_df, nrow(pop_df))
  tpreg_est <- tp_reg(p1_df, p2_df, nrow(pop_df))
  dcpop_est <- 
    dc_poptots(p1_df, p2_df, nrow(pop_df),
               pop_tot = c(nrow(pop_df), sum(pop_df$X1), sum(pop_df$X2)))

  return(
    tibble(Est = c("pistar", "tpreg"),
           Theta = c(pistar_est[[1]], tpreg_est[[1]]),
           Var = c(pistar_est[[2]], tpreg_est[[2]]))
  )

} %>% bind_rows()

stopCluster(clust)

```

```{r}
#| label: Analysis

# Analyze Mean
mc_res %>%
  mutate(err = Theta - true_theta) %>%
  mutate(CI = abs(err) < qnorm(0.975) * sqrt(Var)) %>%
  group_by(Est) %>%
  summarize(bias = mean(err),
            rmse = sqrt(mean(err^2)),
            EmpCI = mean(CI),
            sdest = sd(Theta)) %>%
  ungroup() %>%
  mutate(Ttest = abs(bias) / sqrt(sdest^2 / B_sims)) %>%
  select(-sdest)

# Analyze Variance
mc_res %>%
  group_by(Est) %>%
  summarize(mcvar = var(Theta),
            estvar = mean(Var),
            varvar = var(Var)) %>%
  ungroup() %>%
  mutate(Ttest = abs(mcvar - estvar) / sqrt(varvar / B_sims))

```

### 2024-04-16 final/reports/prelim_paper.tex
Updated prelim_paper.tex. Mostly finished with section 1.

@@
 \usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
-\usepackage[style=authoryear]{biblatex}
+%\usepackage[style=authoryear]{biblatex}
 \usepackage{booktabs}
@@
 
-\addbibresource{references.bib}
+\usepackage{amsmath, amsthm, mathtools,commath}
+\usepackage{graphics, color}
+\usepackage{latexsym}
+\usepackage{amssymb, amsfonts, bm}
+\usepackage{mathrsfs}
 
-\hypersetup{colorlinks = true, urlcolor = black}
+\usepackage{graphicx}
+\usepackage{epsfig}
+\usepackage{makeidx}
+\usepackage{fullpage}
+\usepackage{booktabs, arydshln}
+\usepackage{comment} 
+%\makeindex
 
-\headheight     15pt
-\topmargin      -1.5cm   % read Lamport p.163
-\oddsidemargin  -0.04cm  % read Lamport p.163
-\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
-\textwidth      16.59cm
-\textheight     23.94cm
-\parskip         7.2pt   % sets spacing between paragraphs
-\parindent         0pt   % sets leading space for paragraphs
-\pagestyle{empty}        % Uncomment if don't want page numbers
-\pagestyle{fancyplain}
+\hbadness=10000 \tolerance=10000 \hyphenation{en-vi-ron-ment
+in-ven-tory e-num-er-ate char-ac-ter-is-tic}
+
+\usepackage[round]{natbib}
+%\bibliographystyle{apalike2}
+% \bibliographystyle{jmr}
+
+\newcommand{\biblist}{\begin{list}{}
+{\listparindent 0.0cm \leftmargin 0.50cm \itemindent -0.50 cm
+\labelwidth 0 cm \labelsep 0.50 cm
+\usecounter{list}}\clubpenalty4000\widowpenalty4000}
+\newcommand{\ebiblist}{\end{list}}
+
+\newcounter{list}
+
+%\usepackage{setspace}
+
+%\usepackage{hangpar}
+\newcommand{\lbl}[1]{\label{#1}{\ensuremath{^{\fbox{\tiny\upshape#1}}}}}
+% remove % from next line for final copy
+\renewcommand{\lbl}[1]{\label{#1}}
+
+\newtheorem{theorem}{Theorem}
+\newtheorem{corollary}{Corollary}
+\newtheorem{definition}{Definition}
+\newtheorem{example}{Example}
+\newtheorem{remark}{Remark}
+\newtheorem{result}{Result}
+
+\newtheorem{lemma}{Lemma}
 
@@
 
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-
 \begin{document}
@@
 
+\baselineskip .3in
+
 \section{Introduction}
 
-\begin{itemize}
-\item Combining information from several sources is an important practical
-  problem. 
-\item In many cases, we do not have direct access to the other sources. We can
-  only obtain summary statistics (and their standard errors) for the external
-  data sources. We wish to incorporate the information from external  sources in
-  our in-house data effectively using calibration weighting.  
-\item We formulate the problem as a generalized two-phase sampling where the
-  first phase sample can be obtained from multiple sources. The second-phase
-  sample is our in-house data in which we want to construct the calibration
-  weights. 
-\item To achieve the goal, we first consider  the  classical two-phase sampling
-  setup where the second-phase sample is a subset of the first-phase sample.
-  After that, we extend the setup to more general cases such as non-nested
-  two-phase sampling or multiple independent surveys with some common
-  measurements. 
-\item (To simplify the presentation, I wonder whether we can just use SRS in the
-  first-phase sampling. ) 
-\end{itemize}
+Combining information from several sources is an important practical problem. (CITEME)
+We want to incorporate information from external data sources to reduce the bias
+in our estimates or improve the estimator's efficiency. For many problems, the
+additional information consist of summary statistics with standard errors. The
+goal of this project is to incorporate external information with existing data 
+to create more efficient estimators using calibration weighting.
+
+To model this scenaro, we formulate the problem as a generalized two-phase
+sample where the first phase sample consists of data from multiple sources. The
+second phase sample contains our existing data. To motivate this setup, we
+consider the following approach: first, we consider the classical two-phase
+sampling setup where the second phase sample is a subset of the first phase
+sample; then, we extend this setup to consider non-nested two-phase samples;
+and finally, we consider the more general approach of having multiple sources as
+the first phase sample.
+
+%\begin{itemize}
+%\item To achieve the goal, we first consider  the  classical two-phase sampling
+%setup where the second-phase sample is a subset of the first-phase sample.
+%  After that, we extend the setup to more general cases such as non-nested
+%  two-phase sampling or multiple independent surveys with some common
+%  measurements. 
+%\item The proposed method can be called two-step calibration. In the
+%first-step, the best linear unbiased estimators of the auxiliary variable
+%totals are computed. In the second-step, the final calibration weights are
+%constructed to match (benchmark) with the best estimators computed from Step 1. 
+%\end{itemize}
+
 \section{Topic 1: Classical Two-Phase Sampling}
 
-\subsection{Background and Introduction}
+\subsection{Background}
 
@@
 only using the observed information from the Phase 2 sample but also
-incorporating the extra auxilary information of $X$ from the Phase 1 sample, and
+incorporating the extra auxiliary information of $X$ from the Phase 1 sample, and
 the challenge is doing this efficiently.
 
-An easy-to-implement unbiased estimator in the spirit of a Horvitz-Thompson
+An easy-to-implement unbiased estimator in the spirit of a Horvitz-Thompson (HT)
 estimator (\cite{horvitz1952generalization}, \cite{narain1951sampling}) is the
@@
 denoted as $A_2$,
+%\begin{align*}
+%  \pi_i &= \sum_{A_2: i \in A_2} \Pr(A_2) \\ 
+%        &= \sum_{A_1: A_2 \subseteq A_1} \sum_{A_2: i \in A_2} \Pr(A_2 \mid
+%        A_1) \Pr(A_1) \\
+%        &= \sum_{A_1: i \in A_1} \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1) \Pr(A_1).
+%\end{align*}
+%\textcolor{red}{(I am not sure whether this part is necessary.) }
+if we define $\pi_{2i | 1} = \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1)$ and
+$\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ then,
 
-\begin{align*}
-  \pi_i &= \sum_{A_2: i \in A_2} \Pr(A_2) \\ 
-        &= \sum_{A_1: A_2 \subseteq A_1} \sum_{A_2: i \in A_2} \Pr(A_2 \mid
-        A_1) \Pr(A_1) \\
-        &= \sum_{A_1: i \in A_1} \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1) \Pr(A_1).
-\end{align*}
+$$ \pi_i(A_1) = \pi_{2i | 1} \pi_{1i}.$$
 
-If we define $\pi_{2i | 1}^{(2)} = \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1)$ and
-$\pi_i^{(1)} = \sum_{A_1: i \in A_1} \Pr(A_1)$ then,
+%\textcolor{red}{Also, the above equality is not true under two-phase sampling!)}
+\textcolor{red}{I agree that having the superscript $(1)$ and $(2)$ is
+  unnecessary. I have removed them. I have also modified the previous equation.
+  I agree that the invariance condition does not hold for two-phase sampling,
+  but if we consider $\pi_i$ as a function of $A_1$ then I believe this holds.
+  Is this correct? Should we not have $\pi_{i} = \pi_{1i} \pi_{2i|1}(A_1)$
+  because this is how conditional distributions work? (Also, I thought that this
+  is equivalent to the reverse sampling idea of Fey (1992), Shao and Steel
+(1999), and Kim et al. (2006).) I would love to discuss this with you.}
 
-$$ \pi_i = \pi_{2i | 1}^{(2)} \pi_i^{(1)}.$$
-\textcolor{red}{(note: there is an abuse of notation. You do not need to introduce superscript $(2)$ on $\pi_{2i \mid 1}$.) }
 This means that we can define the $\pi^*$-estimator as the following design
@@
 
-$$ \hat Y_{\pi^*} = \sum_{i \in A_2} \frac{y_i}{\pi_{2i | 1}^{(2)} \pi_i^{(1)}}.$$
+$$ \hat Y_{\pi^*} = \sum_{i \in A_2} \frac{y_i}{\pi_{2i | 1} \pi_{1i}}.$$
 
-While unbiased \cite{kim2024statistics}, the $\pi^*$-estimator unfortunately
+While unbiased (see \cite{kim2024statistics}), the $\pi^*$-estimator
 does not account for the additional information contained in the auxiliary Phase
 1 variable $X$. The two-phase regression estimator $\hat Y_{reg, tp}$ does incorporate 
-\textcolor{blue}{$\hat{X}_1$ obtained from the phase one sample. That is, we can
-leverage the external information $\hat{X}_1$ to improve the $\pi^*$-estimator
-in the second-phase sample. } 
+information for $X$ by using the estimate $\hat X_1$ from the Phase 1 sample.
+This is how we can leverage the external information $\hat X_1$ to improve the
+initial $\pi^*$-estimator in the second phase sample.
 The two-phase regression estimator has the form,
@@
 $$ \hat Y_{reg, tp} 
-= \sum_{i \in A_1} \frac{1}{\pi_i^{(1)}} x_i \hat \beta_2+ \sum_{i \in A_2}
-\frac{1}{\pi_i^{(1)}\pi_{2i|1}^{(2)}} (y_i - x_i \hat \beta_2)$$
-where $$\hat \beta_2 = \left(\sum_{i \in A_2} 
-  \frac{x_i x_i'}{\pi_{2i|1}^{(2)}}\right)^{-1} 
-  \sum_{i \in A_2} \frac{x_i y_i}{\pi_i^{(1)}\pi_{2i|1}^{(2)}}. 
-  \footnote{Caleb, we do not have to use $\pi_{2i \mid 1}^{(2)}$ in computing $\hat{\beta}_2$. Please check my book. }
+= \sum_{i \in A_1} \frac{1}{\pi_{1i}} x_i \hat \beta_2+ \sum_{i \in A_2}
+\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_2)$$
+
+where for $q_i = q(x_i)$ and is a function of $x_i$,
+$$
+\hat \beta_2 = \left(\sum_{i \in A_2} 
+  \frac{x_i x_i'}{\pi_{1i} q_i}\right)^{-1} 
+\sum_{i \in A_2} \frac{x_i y_i}{\pi_{1i} q_i}. 
+\footnote{
+  \textcolor{blue}{Caleb, we do not have to use $\pi_{2i \mid 1}^{(2)}$ in computing
+  $\hat{\beta}_2$. Please check my book.}
+  \textcolor{red}{Thanks for the check. I
+checked it and I must have gotten $\pi_{1i}$ and $\pi_{2i}$ confused.}}
+\footnote{\textcolor{red}{I was initially thinking of using $\hat \beta_2$ without $q_i$ to be
+  more consistent with how I defined the two-phase regression estimator;
+  however, I see your point about using $q_i$ to connect this estimator with the
+superpopulation model.}}
   $$
 
-\textcolor{red}{I suggest that we use 
-$$\hat \beta_2 = \left(\sum_{i \in A_2} 
-  \frac{x_i x_i'}{\pi_{i}^{(1)} q_i }\right)^{-1} 
-  \sum_{i \in A_2} \frac{x_i y_i}{\pi_i^{(1)} q_i }
-$$ 
-where $q_i=q(\bx_i)$ is a function of $\bx_i$. This will have some connection with model-optimal regression estimator (under superpopulation model).  
-}
+The regression estimator is the minimum variance design consistent linear
+estimator which is easily shown to be the case because $\hat Y_{reg, tp} =
+\sum_{i \in A_2} \hat w_{2i} y_i / \pi_{1i}$ where
 
-The regression
-estimator is the minumum variance design consistent linear estimator which is
-easily shown to be the case because $\hat Y_{reg, tp} = \sum_{i \in A_2} \hat
-w_{2i} y_i / \pi_i^{(1)}$ where
-
-$$\hat w_i = \argmin_{w} \sum_{i \in A_2} (w_i - \pi_{2i|1}^{-1})^2 \text{ such
-that } \sum_{i \in A_2} w_i x_i / \pi_i^{(1)} = \sum_{i \in A_1} x_i /
-\pi_i^{(1)}.$$
-
-\textcolor{red}{Using $q_i$, we can construct 
 $$\hat w_{2i} = \argmin_{w} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i \text{ such
-that } \sum_{i \in A_2} w_{2i}  x_i / \pi_i^{(1)} = \sum_{i \in A_1} x_i /
-\pi_i^{(1)}$$
-as a way to implement the two-phase regression estimator indirectly using calibration weighting. 
-}
+that } \sum_{i \in A_2} w_{2i} x_i / \pi_{1i} = \sum_{i \in A_1} x_i / \pi_{1i}.$$
 
@@
 generalization includes empirical likelihood estimation, and maximum entropy
-estimation amoung others. The variance estimation is based on a linearization
+estimation among others. The variance estimation is based on a linearization
 that shows that minimizing the generalized loss function subject to the
@@
 estimator is not design consistent. 
-\footnote{ \textcolor{red}{I do not understand this part.} I changed the
+\footnote{ \textcolor{blue}{I do not understand this part.} 
+  \textcolor{red}{I changed the
 notation so that it would not be the unused column space notation. This is from
-Equation (11.17) of \cite{kim2024statistics}.} 
+Equation (11.17) of \cite{kim2024statistics}.}} 
+\footnote{\textcolor{red}{I think that I am confused about the purpose of having
+the assumption from Equation (11.17) in your sampling book. Could we discuss
+this?}}
 % --------------------------------------------
@@
 $$ \hat Y_{reg, tp} 
-= \underbrace{\sum_{i \in A_1} \frac{x_i \hat \beta_2}{\pi_i^{(1)}}}_{
+= \underbrace{\sum_{i \in A_1} \frac{x_i \hat \beta_2}{\pi_{1i}}}_{
   \text{ Minimizing the variance}} + \underbrace{\sum_{i \in A_2}
-\frac{1}{\pi_i^{(1)}\pi_{2i|1}^{(2)}} (y_i - x_i \hat \beta_2)}_{
+\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_2)}_{
 \text{Bias correction}}.$$
@@
 
-Unfortunately, the method of \cite{kwon2024debiased} requires known calibration
+The method of \cite{kwon2024debiased} requires known calibration
 levels (no uncertainty) for the finite population. It does not handle the
@@
 
+%The proposed method will be something like this: 
+%\textcolor{red}{Minimize 
+%  $$ \sum_{i \in A_2} \frac{1}{\pi_i^{(1)}}  G( w_{2i} ) q_i $$ 
+%  subject to 
+%  $$ \sum_{i \in A_2} w_{2i}  x_i / \pi_i^{(1)} = \sum_{i \in A_1} x_i / \pi_i^{(1)}$$ 
+%  and 
+%  $$ \sum_{i \in A_2}  \frac{1}{\pi_i^{(1)}} w_{2i} g( \pi_{2i \mid 1}^{-1} ) q_i
+%  = \sum_{i \in A_1}  \frac{1}{\pi_i^{(1)}}  g(\pi_{2i \mid 1}^{-1} ) q_i $$ 
+%  where $g(\omega) = \partial G( \omega)/ \partial \omega$. }
+
+% You need to show that the resulting calibration estimator $\hat{Y}_{\rm cal}=
+% \sum_{i \in A_2} w_{1i} \hat{w}_{2i} y_i$ is asymptotically equivalent to the
+% two-phase regression estimator in (\ref{reg}), where $w_{1i} = 1/ \pi_i^{(1)}$.
+
+% Also, there should be a discussion about variance estimation. Once the
+% linearization form is obtained, the formula for linearized variance estimation
+% should be obtained easily.
+
 We follow the approach of \cite{kwon2024debiased} for the debiased calibration
@@
 \begin{equation}\label{eq:primalloss}
-H(w) = - \sum_{i \in A_2} G(w_i)
+  H(w) = - \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i
 \end{equation}
@@
 \begin{equation}\label{eq:calconst1}
-\sum_{i \in A_2} \frac{x_i w_i}{\pi_i^{(1)}} = 
-\sum_{i \in A_1} \frac{x_i}{\pi_i^{(1)}}
+  \sum_{i \in A_2} \frac{x_i w_{2i}}{\pi_{1i}} = 
+\sum_{i \in A_1} \frac{x_i}{\pi_{1i}}
 \end{equation}
@@
 \begin{equation}\label{eq:calconst2}
-\sum_{i \in A_2} \frac{g(\pi_{2i|1}^{-1})w_i}{\pi_i^{(1)}} = 
-\sum_{i \in A_1} \frac{g(\pi_{2i|1}^{-1})}{\pi_i^{(1)}}.
+  \sum_{i \in A_2} \frac{g(\pi_{2i|1}^{-1})w_{2i}q_i}{\pi_{1i}} = 
+  \sum_{i \in A_1} \frac{g(\pi_{2i|1}^{-1})q_i}{\pi_{1i}}.
 \end{equation}
@@
 The first constraint is the existing calibration constraint and the second
-ensures that design consistency is achieved. Here, $g(w) = \partial G / \partial
-w$. The original method of \cite{kwon2024debiased} only considered having finite
-population quantities on the right hand side of \ref{eq:calconst1}. Let $z_i =
+ensures that design consistency is achieved. Here, 
+$g(w) = \partial G / \partial w$. 
+% TODO: Elaborate on this more.
+The original method of \cite{kwon2024debiased} only considered having finite
+population quantities on the right hand side of \ref{eq:calconst1}.
+Let $z_i =
 (x_i, g(\pi_{2i|1}^{-1})$. To solve this minimization problem, we use the convex
@@
 \begin{itemize}
-  \item[[A1]] $G(w)$ is strictly convex and twice differentiable in an interval
+  \item[(A1)] $G(w)$ is strictly convex and twice differentiable in an interval
     $\mathcal{V} \in \R$,
-  \item[[A2]] There exists $c_1, c_2 \in \mathcal{V}$ with $c_1, c_2 > 0$ such
-    that $c_1 < \pi_{2i|1} < c_2$ for $i = 1, \dots, n$,
-  \item[[A3]] If $\pi_{2ij|1}$ is joint inclusion probability of elements $i$
-    and $j$ and $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$ then,
-    $$\limsup_{n \to \infty} \max_{i,j \in A_1: i \neq j} |\Delta_{2ij|1}| <
+  \item[(A2)] There exists $c_1, c_2 \in \mathcal{V}$ with $c_1, c_2 > 0$ such
+    that $c_1 < \pi_{2i|1} < c_2$ for $i = 1, \dots, n_2$, and 
+    that $c_1 < \pi_{1i} < c_2$ for $i = 1, \dots, n_1$,
+  \item[(A3)] If $\pi_{2ij|1}$ is joint inclusion probability of elements $i$
+    and $j$ in $A_2$, the Phase 2 sample and 
+    $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$ then,
+    $$\limsup_{n \to \infty} \max_{i,j \in U: i \neq j} |\Delta_{2ij|1}| <
     \infty,$$
-  % TODO: Check if A4 should be stated for the finite population (as written) or
-    % if this should be stated for the Phase 1 sample.
-  \item[[A4]] Assume that $\Sigma_z = \lim_{N \to \infty} \sum_{i \in U} z_i
+    and if $\pi_{1ij}$ is the joint inclusion probability of elements $i$ and
+    $j$ in $A_1$, the Phase 1 sample and 
+    $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ then,
+    $$\limsup_{n \to \infty} \max_{i,j \in U: i \neq j} |\Delta_{1ij}| <
+    \infty,$$
+  \item[(A4)] Assume that $\Sigma_z = \lim_{N \to \infty} \sum_{i \in U} z_i
     z_i^T/N$ exists and is positive definite, the average fourth moment of $(y_i,
@@
     = (\bf 0, 1)$.
-  \item[[A5]] FIXME: We need to make an assumption about how Phase 2 relates to
-    Phase 1 which relates to the finite population.
 \end{itemize}
 
-\begin{theorem}{Design Consistency}
-  Suppose that Conditions [A1] - [A5] hold. Then the solution $\hat w$ to 
-  \ref{eq:primalloss} subject to \ref{eq:calconst1} and \ref{eq:calconst2} 
+\begin{theorem}[Design Consistency]
+  Suppose that Conditions (A1) - (A4) hold. Then the solution $\hat w$ to 
+  Equation~\ref{eq:primalloss} subject to Equation~\ref{eq:calconst1} and 
+  Equation~\ref{eq:calconst2} 
   exists and is unique with probability approaching 1. Furthermore, the
-  estimator $\hat \theta_{DCE} = \sum_{i \in A_2} \frac{\hat w y_i}{\pi_i^{(1)}}$
+  estimator $\hat \theta_{DCE} = \sum_{i \in A_2} \frac{\hat w_i y_i}{\pi_{1i}}$
   satisfies
 
-  $$\hat \theta_{DCE} = \hat \theta_{DC} + o_p(FIXME)$$
+  $$\hat \theta_{DCE} = \hat \theta_{DC} + o_p(n_2^{-1})$$
 
@@
 
-\begin{theorem}(Asymptotic Normality)
+\begin{proof}
+  This proof follows the proof of \cite{kwon2024debiased} in a straightforward
+  manner. There are two main modifications that need to be made. The first is
+  that the proof needs to be written for two-phase sampling with $n_1$ and $n_2$
+  instead of a single $n$. The second is that we need to ensure that the result
+  still holds with an estimated population constraint that is $O_p(n_1^{-1/2})$
+  consistent.
+
+  With the Conditions (A1) - (A4) written with respect the the two-phase sample,
+  the first challenge is done. The second step largely holds from the existing
+  proof of \cite{kwon2024debiased} with modifications such as
+
+  $$\hat Q(\lambda) = n_2^{-1} \sum_{i \in A_2} F(\lambda^T z_i^*) - n_1^{-1}
+  \lambda^T \sum_{i \in A_1} z_i^*$$
+
+  ensuring that our extension holds.
+
+  \textcolor{red}{Dr. Kim, I know that I need to write up this proof in more
+  detail, but I am unsure about the level of granularity that is required now. I
+  can basically rewrite everything from the Appendix of \cite{kwon2024debiased}
+  with the two-phase modification to show this result, but I don't know if I
+  should yet.}
+\end{proof}
+
+We need the following additional assumptions to prove asymptotic normality.
+
+% FIXME: Add assumptions
+\textcolor{red}{I need to think about the assumptions that we need in two-phase
+sampling.}
+\begin{itemize}
+  \item[(B1)] 
+  \item[(B2)] The HT estimator is asymptotically normal under the sampling
+    design in the sense that
+
+\end{itemize}
+
+\begin{theorem}[Asymptotic Normality]
 % FIXME: Add result.
+\textcolor{red}{I need to write out more but we should have something like this:
+  $$V(\hat \theta_{DCE} = \Var(\hat \theta_{DC}) + \hat \gamma_{[1:3]}^T V(\bf
+x)) \hat \gamma_{[1:3]} + \Cov(\hat \theta_{DC}, \bf x \hat \gamma_{[1:3}]).$$}
 \end{theorem}
 
-% TODO: Provide asymptotic results showing the variance estimation of the
-% estimated debiased calibration method
-% DUE: Thursday April 11 (I have already showed the proof but I might finish the
-% simulation study before writing this up.)
+\textcolor{red}{Dr. Kim, I am not convinced that $\Cov(\hat \theta_{DC}, \bf x
+\hat \gamma_{[1:3]}) = 0$. The reason that this covariance holds in
+\cite{fuller2009sampling} Section 3.3.3 is that the estimating equation requires
+$\sum_i x_i \hat e_i = 0$, which means that $\Cov(\hat e, x_i) = 0$. However, we
+do not have this requirement. Instead, I see this more like a nested model where
+we have a group level error that aggregates the individual level errors. The
+problem with the current simulation that I cannot tell if this covariance should
+be zero because the variance term $\hat \gamma_{[1:3]}'V(\bf x) \hat \gamma_{[1:3]}$
+dominates the variance estimation. I will change the simulation parameters to
+try to see if I can get a better test of this.}
 
@@
 
-% TODO: Conduct a simulation study highlighting the proposed method
-% DUE: Friday April 12
+We run a simulation testing the proposed method. In this approach we have the
+following simulation setup:
+
+$$
+\begin{aligned}
+X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
+X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
+X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
+X_{4i} &\stackrel{ind}{\sim} Unif(0.1, 0.9) \\
+\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
+Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
+\pi_{1i} &= \Phi_3(-x_{3i} - 2) \\
+\pi_{2i|1} &= x_{4i}.
+\end{aligned}
+$$
+
+where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
+This is a two-phase extension of the setup in \cite{kwon2024debiased}. We
+consider a finite population of size $N = 10,000$ with both the Phase 1 and
+Phase 2 sampling occuring under Poisson (Bernoulli) sampling. This yields a
+Phase 1 sample
+size of $E[n_1] \approx 1100$ and a Phase 2 sample size of
+$E[n_2] \approx 550$. In the Phase 1 sample, we observe 
+$(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. This
+simulation does not deal with model misspecification, and we compare the
+proposed method for the parameter $\bar Y_N$ with four approaches:
+
+\begin{itemize}
+  \item[1.] $\pi^*$-estimator: $\hat Y_{\pi^*} = N^{-1} \sum_{i \in A_2}
+    \frac{y_i}{\pi_{1i} \pi_{2i|1}},$
+  \item[2.] Two Phase Regression estimator (TP-Reg): 
+    $\hat Y_{reg} = \sum_{i \in A_1} \frac{\bf x_i' \hat \beta}{\pi_{1i}} + 
+    \sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}}(y_i - \bf x_i' \hat \beta)$ 
+    where $\hat \beta = 
+    \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$
+    and $\bf x_i = (x_{1i}, x_{2i})^T$,
+    \textcolor{red}{Dr. Kim, should I modify the simulation so that the
+    regression estimator also includes $g(\pi_{2i|1}^{-1})$ as a covariate?}
+  \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This is
+    the method from \cite{kwon2024debiased} with the true population level
+    constraints, and 
+  \item[4.] Debiased Calibration with Estimated Population Constraints (DC-Est):
+    This is the proposed method with the Phase 1 sample being used to estimate
+    the population level constraints.
+\end{itemize}
+
+In addition to estimating the mean parameter $\bar Y_N$, we also construct
+variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$. For each
+approach we have the following variance estimate\footnote{These variance
+estimates use the fact that we have Poisson sampling for both phases in the
+simulation.}:
+
+\begin{figure}[ht!]
+  \centering
+  \begin{tabular}{lcc}
+    \toprule
+    Estimator & Variance & Notes \\
+    \midrule
+    $\pi^*$ & $N^{-2} \sum_{i \in A_2} \left(\pi_{2i|1}^{-2} - \pi_{2i|1}^{-1}\right)
+    y_i^2 $ & \\
+    TP-Reg  & 
+      {\scriptsize $N^{-2}\left(\sum\limits_{i \in A_1} \left(\pi_{1i}^{-2} -
+          \pi_{1i}^{-1}\right) 
+      \eta_i^2 
+      + \sum\limits_{i \in A_2} \frac{1}{\pi_{1i} \pi_{2i|1}}(\pi_{2i|1}^{-1}
+  - 1) (y_i - \bf x_i' \hat \beta)^2 \right)$}
+    & 
+      {\scriptsize$\eta_i = \bf x_i \hat \beta + \frac{\delta_{2i}}{\pi_{2i|1}}(y_i - 
+      \bf x_i \hat \beta)$} \\
+      %$\hat \beta = \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$ \\
+      DC-Pop  & $(Y - \bf z^T \hat \gamma)^T \Pi (Y - \bf z^T \hat \gamma)$ & 
+      {\scriptsize$\Pi
+      = \text{diag}(1 - (\pi_{1}\pi_{2|1})^{-1}) \cdot \frac{\hat w^2}{N^2}$} \\
+      DC-Est  & $(Y - \bf z^T \hat \gamma)^T \Pi (Y - \bf z^T \hat \gamma) +
+      \hat \gamma_{[1:3]}^T V(\bf x) \hat \gamma_{[1:3]}/N^2$ & \\
+    \bottomrule
+  \end{tabular}
+  \caption{This table gives the formulas for each variance estimator used in
+  this simulation.}
+  \label{tab:varforms}
+\end{figure}
+
+We run this simulation $1000$ times for each of these methods and compute the
+Bias ($E[\hat Y] - \bar Y_N$), the RMSE ($\sqrt{\Var(\hat Y - \bar Y_N)}$), a 95\%
+empirical confidence interval ($\sum_{b = 1}^{1000} |\hat Y^{(b)} - \bar Y_N| \leq 
+\Phi(0.975)\sqrt{\hat V(\hat Y^{(b)})^{(b)}}$), and a T-test that assesses the
+unbiasedness of each estimator. The results are in Figure~\ref{fig:tpdc-mean}.
+
+\begin{figure}[ht!]
+  \centering
+\input{tables/tpdcsim_mean.tex}
+% This table was generated from /src/explore/20240411-tpdcsim.qmd
+\caption{This table shows the results of the simulation study. It displays the
+Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
+unbiasedness of each estimator for the estimators: $\pi^*$, TP-Reg, DC-Pop, and
+DC-Est.}
+\label{fig:tpdc-mean}
+\end{figure}
 
@@
 
+\textcolor{blue}{
+( Materials in Section 11.4 can be used here. I have copy-and-pasted the textbook materials below. Please modify them.  )
+}
+\textcolor{red}{I will do this shortly.}
+
+In contrast to the classical two-phase sampling framework, non-nested two-phase sampling involves two independent surveys conducted on the same target population.
+The key distinction is that the two samples, denoted as $A_1$ and $A_2$, are drawn independently rather than sequentially.
+Table \ref{table:11-1} presents the data structure for non-nested two-phase sampling.
+
+In the non-nested two-phase sampling, 
+a large probability sample $A_1$ is drawn from a finite population,  collecting  only the $\bx$ variable, and  a  smaller sample $A_2$ is drawn from the same population, providing information  on both the $y$ and $\bx$ variables.
+It is assumed that the observed variable $x$ is comparable in the two surveys. \cite{renssen1997} formally addressed this non-nested two-phase sampling problem  and 
+\cite{merkouris2004} extended the idea further to develop regression estimation combining information from multiple surveys. 
+\cite{kimrao12} considered the non-nested two-phase sampling in the context of mass imputation combining two independent surveys at the population and domain levels.
+
+\begin{table}[htb]
+\caption{Data Structure for non-nested two-phase sampling}
+\label{table:1}\par
+\vskip .2cm
+\centerline{\tabcolsep=3truept\begin{tabular}{|c|cc|}
+\hline
+Sample  & $X$ & $Y$ \\
+\hline
+$A_1$ & \checkmark &  \\
+$A_2$ & \checkmark & \checkmark  \\
+\hline
+\end{tabular}}
+\label{table:11-1}
+\end{table}
+
+To illustrate the non-nested two-phase sampling approach, let's consider the data structure shown in Table \ref{table:11-1}.
+This setup involves two independent samples, $A_1$ and $A_2$, drawn from the same target population.
+
+From these two samples, we can compute two unbiased estimators of the population total $\mathbf{X} = \sum_{i=1}^N \bx_i$ for the auxiliary variable x: $\hat{\mathbf{X}}_1 = \sum_{i \in A_1} \pi_{1i}^{-1} \bx_i$
+and 
+$\hat{\mathbf{X}}_2 = \sum_{i \in A_2} \pi_{2i}^{-1} \bx_i$. 
+Here, $\pi_{1i}$ and $\pi_{2i}$ represent the inclusion probabilities for samples $A_1$ and $A_2$, respectively.
+
+Both $\hat{\mathbf{X}}_1$ and $\hat{\mathbf{X}}_2$ are unbiased estimators of the population total $\mathbf{X}$ under the respective sampling designs. 
+The availability of these two unbiased estimators is a key feature of the non-nested two-phase sampling design, as it provides opportunities for developing enhanced estimation procedures combining information from different sources.
+
+We can  construct a combined estimator of X, denoted as $\widehat{\mathbf{X}}_c$, as follows:
+\begin{equation}
+\widehat{\mathbf{X}}_c = W \hat{\mathbf{X}}_1 + (I - W) \hat{\mathbf{X}}_2, 
+\label{xgls}
+\end{equation}
+where $W$ is a $p \times p$ symmetric matrix of constants, and $p = \text{dim}(\bx)$ is the dimension of the auxiliary variable x.
+The optimal choice of the matrix W can be determined using the Generalized Least Squares (GLS) method.
+However, other choices of W can also be used.
+The key idea is to leverage the information from these two independent surveys to obtain a more accurate and efficient estimator of the population total X for the auxiliary variable x, compared to using only one of the surveys alone.
+
+Using the combined estimator  $\widehat{\mathbf{X}}_{\rm c} $ in (\ref{xgls}), we can construct the following projection estimator:  
+\begin{equation}
+ \widehat{Y}_p = \widehat{\mathbf{X}}_{\rm c}' \hat{\bm \beta}_q  
+ \label{11-proj}
+ \end{equation}
+where the regression coefficient estimator $\hat{\bm \beta}_q$ is defined as 
+$$ 
+\hat{\bm \beta}_q = \left( \sum_{i \in A_2} \bx_i \bx_i' / q_i \right)^{-1} 
+\sum_{i \in A_2} \bx_i y_i / q_i. 
+$$
+The choice of $q_i$ in the regression coefficient estimator is somewhat arbitrary. Two possible choices are:
+\begin{enumerate}
+\item Using the model variance under a regression superpopulation model.
+\item Using $q_i = \pi_{2i}^{-2} - \pi_{2i}^{-1}$ to compute the design-optimal regression estimator under Poisson sampling.
+\end{enumerate}
+The key idea is that by using the combined estimator $\widehat{\mathbf{X}}_c$ in the projection estimator $\widehat{Y}_p$, we can leverage the information from both the A1 and A2 samples to obtain a more accurate prediction of the variable of interest Y.
+The choice of $q_i$ allows for some flexibility in how the regression coefficient is estimated.
+
+To ensure the design-consistency of the projection estimator in (\ref{11-proj}), we can use the following regression estimator under non-nested two-phase sampling:
+\begin{equation}
+\widehat{Y}_{\rm tp, reg} = \widehat{Y}_2 + \left( \widehat{\mathbf{X}}_{\rm c} - \widehat{\mathbf{X}}_2 \right)' \hat{\bm \beta}_q
+\label{eq:11-28}
+\end{equation} 
+By the definition of $\widehat{\mathbf{X}}_{\rm c}$, we can also express this as: 
+\begin{equation}
+\widehat{Y}_{\rm tp, reg} = \widehat{Y}_2 + \left( \widehat{\mathbf{X}}_{1} - \widehat{\mathbf{X}}_2 \right)' \hat{\bm \alpha}_q,
+\label{eq:11-28}
+\end{equation} 
+where $\hat{\bm \alpha}_q = W \hat{\bm \beta}_q$.
+The key points are:
+\begin{enumerate}
+\item The design-consistent regression estimator $\widehat{Y}_{\rm tp, reg}$ is constructed by adding a correction term to the projection estimator $\widehat{Y}_p$ from the second sample.
+\item The regression estimator improves the efficiency of the  design unbiased estimator  $\hat{Y}_2$ by substracting the projection of $\hat{Y}_2$ onto the augmentation space \citep{tsiatis2006}, 
+the linear space generated by 
+ the difference between the combined estimator $\widehat{\mathbf{X}}_{\rm c}$ and the estimator $\widehat{\mathbf{X}}_2$ from the second sample.
+\item 
+Alternatively, the augmentation space  can be expressed using the difference between the estimators $\widehat{\mathbf{X}}_{1}$ and $\widehat{\mathbf{X}}_2$, weighted by $\hat{\bm \alpha}_q$.
+\end{enumerate} 
+The goal is to leverage the information from both samples to obtain a design-consistent regression estimator for the variable of interest Y.
+
+Using the standard argument, we can obtain 
+\begin{eqnarray}
+\widehat{Y}_{\rm tp, reg} 
+&=& \widehat{Y}_2 +  \left( \widehat{\mathbf{X}}_{1} - \widehat{\mathbf{X}}_2 \right)' {\bm \alpha}_q^* + O_p(n^{-1} N) \label{eq:11-29}
+\end{eqnarray}
+where $\bm \alpha_q^*$ is the probability limit of $\hat{\bm \alpha}_q = W \hat{\bm \beta}_q$. By (\ref{eq:11-29}), 
+ we can obtain 
+\begin{equation}
+V \left( \widehat{Y}_{\rm tp, reg}  \right) 
+= ({\bm \alpha}_q^* )' V \left( \widehat{\mathbf{X}}_{1}  \right) {\bm \alpha}_q^* + V\left( \hat{u}_2 \right) 
+\label{eq:11-31}
+\end{equation}
+where $\hat{u}_2 = \sum_{i \in A_2} \pi_{2i}^{-1} \left( y_i -  \bx_i'\bm \alpha_q^* \right) $. From the formula in (\ref{eq:11-31}), we can construct a linearized variance estimator.
+
+Now, we can use the calibration weighting to construct the regression estimator under non-nested two-phase sampling. For  given the design weights $d_{2i} = \pi_{2i}^{-1}$, we find the minimizer of 
+    $$Q \left( {\bm \omega } \right) = \sum_{i \in A_2} \left(  {\omega_i} - d_{2i}  \right)^2 q_i $$
+    subject to 
+    $$ \sum_{i \in A_2} {\omega_i} \bx_i = \widehat{\mathbf{X}}_{\rm c} .$$
+    The solution is 
+    $$ \hat{\omega}_i = d_{2i} + \left( \widehat{\mathbf{X}}_{\rm c} - \widehat{\mathbf{X}}_{\rm 2} \right)^{-1} \left( \sum_{i \in A_2} q_i^{-1} \bx_i \bx_i' \right)^{-1} \bx_i q_i^{-1} . $$
+     Note that 
+    $$ \sum_{i \in A_2} \hat{\omega}_i y_i = \widehat{Y}_{\rm tp, reg} , $$
+    where  $\widehat{Y}_{\rm tp, reg}$ is defined in (\ref{eq:11-28}). 
+    Thus, the algebraic equivalence between the  regression estimator and the calibration weighting estimator is established under non-nested two-phase sampling.
+
 \section*{Topic 3: Multi-source Two-Phase Sampling}
 
-\printbibliography
+\bibliographystyle{chicago}
+\bibliography{references}
 

### 2024-04-16 final/reports/prelim_paper_ov.tex
Updated prelim_paper.tex. Mostly finished with section 1.

@@
 
-\usepackage{amsmath, amssymb, amsthm, mathrsfs, fancyhdr}
-\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
-\usepackage[style=authoryear]{biblatex}
-\usepackage{booktabs}
-\usepackage{float}
+\usepackage{amssymb}
+\usepackage{amsmath,amsthm, mathtools,commath}
+\usepackage{graphics, color}
 
-\addbibresource{references.bib}
+\usepackage{graphicx}
+\usepackage{epsfig}
+\usepackage{makeidx}
+\usepackage{kotex}
+\usepackage{fullpage}
+\usepackage{booktabs, arydshln}
+%\makeindex
 
-\hypersetup{colorlinks = true, urlcolor = black}
+\hbadness=10000 \tolerance=10000 \hyphenation{en-vi-ron-ment
+in-ven-tory e-num-er-ate char-ac-ter-is-tic}
 
-\headheight     15pt
-\topmargin      -1.5cm   % read Lamport p.163
-\oddsidemargin  -0.04cm  % read Lamport p.163
-\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
-\textwidth      16.59cm
-\textheight     23.94cm
-\parskip         7.2pt   % sets spacing between paragraphs
-\parindent         0pt   % sets leading space for paragraphs
-\pagestyle{empty}        % Uncomment if don't want page numbers
-\pagestyle{fancyplain}
+\usepackage[round]{natbib}
+%\bibliographystyle{apalike2}
+% \bibliographystyle{jmr}
+
+\newcommand{\biblist}{\begin{list}{}
+{\listparindent 0.0cm \leftmargin 0.50cm \itemindent -0.50 cm
+\labelwidth 0 cm \labelsep 0.50 cm
+\usecounter{list}}\clubpenalty4000\widowpenalty4000}
+\newcommand{\ebiblist}{\end{list}}
+
+\newcounter{list}
+\usepackage{comment}
+
+%\usepackage{setspace}
+\usepackage{latexsym}
+\usepackage{amsmath, amssymb, amsfonts, amsthm, bm}
+\usepackage{graphicx}
+\usepackage{mathrsfs}
+
+%\usepackage{hangpar}
+\newcommand{\lbl}[1]{\label{#1}{\ensuremath{^{\fbox{\tiny\upshape#1}}}}}
+% remove % from next line for final copy
+\renewcommand{\lbl}[1]{\label{#1}}
+
+\newtheorem{theorem}{Theorem}
+\newtheorem{corollary}{Corollary}
+\newtheorem{definition}{Definition}
+\newtheorem{example}{Example}
+\newtheorem{remark}{Remark}
+\newtheorem{result}{Result}
+
+\newtheorem{lemma}{Lemma}
 
@@
 \maketitle
+
+\baselineskip .3in
 
@@
 After that, we extend the setup to more general cases such as non-nested two-phase sampling or multiple independent surveys with some common measurements.
-\item (To simplify the presentation, I wonder whether we can just use SRS in the first-phase sampling. ) 
+\item The proposed method can be called two-step calibration.
+In the first-step, the best linear unbiased estimators of the auxiliary variable totals are computed.
+In the second-step, the final calibration weights are constructed to match (benchmark) with the best estimators computed from Step 1.
+%\item (To simplify the presentation, I wonder whether we can just use SRS in the first-phase sampling. ) 
 \end{itemize}
@@
 only using the observed information from the Phase 2 sample but also
-incorporating the extra auxilary information of $X$ from the Phase 1 sample, and
+incorporating the extra auxiliary information of $X$ from the Phase 1 sample, and
 the challenge is doing this efficiently.
@@
 \end{align*}
+\textcolor{red}{(I am not sure whether this part is necessary.) }
 
@@
 $$ \pi_i = \pi_{2i | 1}^{(2)} \pi_i^{(1)}.$$
-\textcolor{red}{(note: there is an abuse of notation. You do not need to introduce superscript $(2)$ on $\pi_{2i \mid 1}$.) }
+\textcolor{red}{(note: there is an abuse of notation. You do not need to introduce superscript $(2)$ on $\pi_{2i \mid 1}$. 
+Also, the above equality is not true under two-phase sampling!)  }
 This means that we can define the $\pi^*$-estimator as the following design
@@
 regression estimator has the form,
-
-$$ \hat Y_{reg, tp} 
+\begin{equation}
+\hat Y_{reg, tp} 
 = \sum_{i \in A_1} \frac{1}{\pi_i^{(1)}} x_i \hat \beta_2+ \sum_{i \in A_2}
-\frac{1}{\pi_i^{(1)}\pi_{2i|1}^{(2)}} (y_i - x_i \hat \beta_2)$$
+\frac{1}{\pi_i^{(1)}\pi_{2i|1}^{(2)}} (y_i - x_i \hat \beta_2)
+\label{reg}
+\end{equation}
 where $$\hat \beta_2 = \left(\sum_{i \in A_2} 
@@
 $$ 
-where $q_i=q(\bx_i)$ is a function of $\bx_i$. This will have some connection with model-optimal regression estimator (under superpopulation model).  
+where $q_i=q(\bx_i)$ is a function of $\bx_i$. This will have some connection with model-optimal regression estimator later (under superpopulation model).  
 }
@@
 
-\subsection*{Methodology}
+\subsection{Methodology}
+
+The proposed method will be something like this: 
+\textcolor{red}{Minimize 
+$$
+\sum_{i \in A_2} \frac{1}{\pi_i^{(1)}}  G( w_{2i} )  q_i 
+$$
+subject to 
+$$
+\sum_{i \in A_2} w_{2i}  x_i / \pi_i^{(1)} = \sum_{i \in A_1} x_i /
+\pi_i^{(1)}$$
+and 
+$$
+\sum_{i \in A_2}  \frac{1}{\pi_i^{(1)}} w_{2i} g( \pi_{2i \mid 1}^{-1} ) q_i  = \sum_{i \in A_1}  \frac{1}{\pi_i^{(1)}}  g( \pi_{2i \mid 1}^{-1} ) q_i $$
+where $g(\omega) = \partial G( \omega)/ \partial \omega$. 
+}
+
+You need to show that the resulting calibration estimator $\hat{Y}_{\rm cal}= \sum_{i \in A_2} w_{1i} \hat{w}_{2i} y_i$ is asymptotically equivalent to the two-phase regression estimator in (\ref{reg}), where $w_{1i} = 1/ \pi_i^{(1)}$.
+
+Also, there should be a discussion about variance estimation. Once the linearization form is obtained, the formula for linearized variance estimation should be obtained easily.
 
@@
 
+\textcolor{red}{
+( Materials in Section 11.4 can be used here. I have copy-and-pasted the textbook materials below. Please modify them.  )
+}
+
+In contrast to the classical two-phase sampling framework, non-nested two-phase sampling involves two independent surveys conducted on the same target population.
+The key distinction is that the two samples, denoted as $A_1$ and $A_2$, are drawn independently rather than sequentially.
+Table \ref{table:11-1} presents the data structure for non-nested two-phase sampling.
+
+In the non-nested two-phase sampling, 
+a large probability sample $A_1$ is drawn from a finite population,  collecting  only the $\bx$ variable, and  a  smaller sample $A_2$ is drawn from the same population, providing information  on both the $y$ and $\bx$ variables.
+It is assumed that the observed variable $x$ is comparable in the two surveys. \cite{renssen1997} formally addressed this non-nested two-phase sampling problem  and 
+\cite{merkouris2004} extended the idea further to develop regression estimation combining information from multiple surveys. 
+\cite{kimrao12} considered the non-nested two-phase sampling in the context of mass imputation combining two independent surveys at the population and domain levels.
+
+\begin{table}[htb]
+\caption{Data Structure for non-nested two-phase sampling}
+\label{table:1}\par
+\vskip .2cm
+\centerline{\tabcolsep=3truept\begin{tabular}{|c|cc|}
+\hline
+Sample  & $X$ & $Y$ \\
+\hline
+$A_1$ & \checkmark &  \\
+$A_2$ & \checkmark & \checkmark  \\
+\hline
+\end{tabular}}
+\label{table:11-1}
+\end{table}
+
+To illustrate the non-nested two-phase sampling approach, let's consider the data structure shown in Table \ref{table:11-1}.
+This setup involves two independent samples, $A_1$ and $A_2$, drawn from the same target population.
+
+From these two samples, we can compute two unbiased estimators of the population total $\mathbf{X} = \sum_{i=1}^N \bx_i$ for the auxiliary variable x: $\hat{\mathbf{X}}_1 = \sum_{i \in A_1} \pi_{1i}^{-1} \bx_i$
+and 
+$\hat{\mathbf{X}}_2 = \sum_{i \in A_2} \pi_{2i}^{-1} \bx_i$. 
+Here, $\pi_{1i}$ and $\pi_{2i}$ represent the inclusion probabilities for samples $A_1$ and $A_2$, respectively.
+
+Both $\hat{\mathbf{X}}_1$ and $\hat{\mathbf{X}}_2$ are unbiased estimators of the population total $\mathbf{X}$ under the respective sampling designs. 
+The availability of these two unbiased estimators is a key feature of the non-nested two-phase sampling design, as it provides opportunities for developing enhanced estimation procedures combining information from different sources.
+
+We can  construct a combined estimator of X, denoted as $\widehat{\mathbf{X}}_c$, as follows:
+\begin{equation}
+\widehat{\mathbf{X}}_c = W \hat{\mathbf{X}}_1 + (I - W) \hat{\mathbf{X}}_2, 
+\label{xgls}
+\end{equation}
+where $W$ is a $p \times p$ symmetric matrix of constants, and $p = \text{dim}(\bx)$ is the dimension of the auxiliary variable x.
+The optimal choice of the matrix W can be determined using the Generalized Least Squares (GLS) method.
+However, other choices of W can also be used.
+The key idea is to leverage the information from these two independent surveys to obtain a more accurate and efficient estimator of the population total X for the auxiliary variable x, compared to using only one of the surveys alone.
+
+Using the combined estimator  $\widehat{\mathbf{X}}_{\rm c} $ in (\ref{xgls}), we can construct the following projection estimator:  
+\begin{equation}
+ \widehat{Y}_p = \widehat{\mathbf{X}}_{\rm c}' \hat{\bm \beta}_q  
+ \label{11-proj}
+ \end{equation}
+where the regression coefficient estimator $\hat{\bm \beta}_q$ is defined as 
+$$ 
+\hat{\bm \beta}_q = \left( \sum_{i \in A_2} \bx_i \bx_i' / q_i \right)^{-1} 
+\sum_{i \in A_2} \bx_i y_i / q_i. 
+$$
+The choice of $q_i$ in the regression coefficient estimator is somewhat arbitrary. Two possible choices are:
+\begin{enumerate}
+\item Using the model variance under a regression superpopulation model.
+\item Using $q_i = \pi_{2i}^{-2} - \pi_{2i}^{-1}$ to compute the design-optimal regression estimator under Poisson sampling.
+\end{enumerate}
+The key idea is that by using the combined estimator $\widehat{\mathbf{X}}_c$ in the projection estimator $\widehat{Y}_p$, we can leverage the information from both the A1 and A2 samples to obtain a more accurate prediction of the variable of interest Y.
+The choice of $q_i$ allows for some flexibility in how the regression coefficient is estimated.
+
+To ensure the design-consistency of the projection estimator in (\ref{11-proj}), we can use the following regression estimator under non-nested two-phase sampling:
+\begin{equation}
+\widehat{Y}_{\rm tp, reg} = \widehat{Y}_2 + \left( \widehat{\mathbf{X}}_{\rm c} - \widehat{\mathbf{X}}_2 \right)' \hat{\bm \beta}_q
+\label{eq:11-28}
+\end{equation} 
+By the definition of $\widehat{\mathbf{X}}_{\rm c}$, we can also express this as: 
+\begin{equation}
+\widehat{Y}_{\rm tp, reg} = \widehat{Y}_2 + \left( \widehat{\mathbf{X}}_{1} - \widehat{\mathbf{X}}_2 \right)' \hat{\bm \alpha}_q,
+\label{eq:11-28}
+\end{equation} 
+where $\hat{\bm \alpha}_q = W \hat{\bm \beta}_q$.
+The key points are:
+\begin{enumerate}
+\item The design-consistent regression estimator $\widehat{Y}_{\rm tp, reg}$ is constructed by adding a correction term to the projection estimator $\widehat{Y}_p$ from the second sample.
+\item The regression estimator improves the efficiency of the  design unbiased estimator  $\hat{Y}_2$ by substracting the projection of $\hat{Y}_2$ onto the augmentation space \citep{tsiatis2006}, 
+the linear space generated by 
+ the difference between the combined estimator $\widehat{\mathbf{X}}_{\rm c}$ and the estimator $\widehat{\mathbf{X}}_2$ from the second sample.
+\item 
+Alternatively, the augmentation space  can be expressed using the difference between the estimators $\widehat{\mathbf{X}}_{1}$ and $\widehat{\mathbf{X}}_2$, weighted by $\hat{\bm \alpha}_q$.
+\end{enumerate} 
+The goal is to leverage the information from both samples to obtain a design-consistent regression estimator for the variable of interest Y.
+
+Using the standard argument, we can obtain 
+\begin{eqnarray}
+\widehat{Y}_{\rm tp, reg} 
+&=& \widehat{Y}_2 +  \left( \widehat{\mathbf{X}}_{1} - \widehat{\mathbf{X}}_2 \right)' {\bm \alpha}_q^* + O_p(n^{-1} N) \label{eq:11-29}
+\end{eqnarray}
+where $\bm \alpha_q^*$ is the probability limit of $\hat{\bm \alpha}_q = W \hat{\bm \beta}_q$. By (\ref{eq:11-29}), 
+ we can obtain 
+\begin{equation}
+V \left( \widehat{Y}_{\rm tp, reg}  \right) 
+= ({\bm \alpha}_q^* )' V \left( \widehat{\mathbf{X}}_{1}  \right) {\bm \alpha}_q^* + V\left( \hat{u}_2 \right) 
+\label{eq:11-31}
+\end{equation}
+where $\hat{u}_2 = \sum_{i \in A_2} \pi_{2i}^{-1} \left( y_i -  \bx_i'\bm \alpha_q^* \right) $. From the formula in (\ref{eq:11-31}), we can construct a linearized variance estimator.
+
+Now, we can use the calibration weighting to construct the regression estimator under non-nested two-phase sampling. For  given the design weights $d_{2i} = \pi_{2i}^{-1}$, we find the minimizer of 
+    $$Q \left( {\bm \omega } \right) = \sum_{i \in A_2} \left(  {\omega_i} - d_{2i}  \right)^2 q_i $$
+    subject to 
+    $$ \sum_{i \in A_2} {\omega_i} \bx_i = \widehat{\mathbf{X}}_{\rm c} .$$
+    The solution is 
+    $$ \hat{\omega}_i = d_{2i} + \left( \widehat{\mathbf{X}}_{\rm c} - \widehat{\mathbf{X}}_{\rm 2} \right)^{-1} \left( \sum_{i \in A_2} q_i^{-1} \bx_i \bx_i' \right)^{-1} \bx_i q_i^{-1} . $$
+     Note that 
+    $$ \sum_{i \in A_2} \hat{\omega}_i y_i = \widehat{Y}_{\rm tp, reg} , $$
+    where  $\widehat{Y}_{\rm tp, reg}$ is defined in (\ref{eq:11-28}). 
+    Thus, the algebraic equivalence between the  regression estimator and the calibration weighting estimator is established under non-nested two-phase sampling.
+
 \section*{Topic 3: Multi-source Two-Phase Sampling}
 
-\printbibliography
+\bibliographystyle{chicago}
+\bibliography{references}
 

### 2024-04-16 src/explore/20240408-redokwongsim.qmd
Updated prelim_paper.tex. Mostly finished with section 1.

@@
 
-  return(tibble(X1 = x1, X2 = x2, Y = y, Prob_i = pi_vec))
+  return(tibble(X1 = x1, X2 = x2, Z = z, Y = y, Prob_i = pi_vec))
 }
@@
 gen_samp <- function(pop_df) {
-  delta <- sampling::UPpoisson(pop_df$Prob_i)
+  # delta <- sampling::UPpoisson(pop_df$Prob_i)
+  delta <- rbinom(nrow(pop_df), 1, pop_df$Prob_i)
   return(filter(pop_df, row_number() %in% which(delta == 1)))
@@
 
-f = function(lambda, d, Xs, total, entropy, del, ..., returnw = F){
-  # w = d * ginv(drop(Xs %*% lambda), entropy = entropy)
-  if(entropy == "HD" & any(Xs %*% lambda >= 0)) return(rep(Inf, length(lambda)))
-  if(entropy == "PH" & any(abs(Xs %*% lambda) >= del)) return(rep(Inf, length(lambda)))
-
-  if(entropy == "SL"){
-    w = d * drop(Xs %*% lambda)
-  }else if(entropy == "EL"){
-    w = -d / drop(Xs %*% lambda)
-  }else if(entropy == "ET"){
-    w = d * exp(drop(Xs %*% lambda))
-  }else if(entropy == "CE"){
-    w = d / (1 - exp(drop(Xs %*% lambda)))
-  }else if(entropy == "HD"){
-    w = d / drop(Xs %*% lambda)^2
-  }else if(entropy == "PH"){
-    w = d / sqrt(1 / drop(Xs %*% lambda)^2 - 1 / del^2)
-  }
-  if(entropy != "SL" & any(w <= 0)) return(rep(Inf, length(lambda)))
-  if(entropy == "CE" & any(w <= 1)) return(rep(Inf, length(lambda)))
-  if(returnw == T){
-    return(w)
-  }else{
-    return(colSums(Xs * w) - total)
-  }
-}
-
-h = function(lambda, d, Xs, total, entropy, del){
-  # return(t(Xs) %*% (Xs * d * ginvprime(drop(Xs %*% lambda), entropy = entropy)))
-  if(entropy == "SL"){
-    return(t(Xs) %*% (Xs * d))
-  }else if(entropy == "EL"){
-    w = -d / drop(Xs %*% lambda)
-    return(t(Xs) %*% (Xs * (w^2 / d)))
-  }else if(entropy == "ET"){
-    w = d * exp(drop(Xs %*% lambda))
-    return(t(Xs) %*% (Xs * w))
-  }else if(entropy == "CE"){
-    p_Stmp = 1 / (1 - exp(drop(Xs %*% lambda)))
-    return(t(Xs) %*% (Xs * d * p_Stmp * (p_Stmp - 1)))
-  }else if(entropy == "HD"){
-    return(t(Xs) %*% (Xs * (-2 * d / drop(Xs %*% lambda)^3)))
-  }else if(entropy == "PH"){
-    return(t(Xs) %*% (Xs * (d * (1 - (drop(Xs %*% lambda) / del)^2)^(-1.5))))
-  }
-}
-#' @export
-GEcalib = function(Xs, d, total, entropy = c("SL", "EL", "ET", "CE", "HD", "PH"), ...,
-       DS = T, method = "Newton", control = list(maxit = 1e5, allowSingular = T)){
-
-  del = quantile(d, 0.80)
-
-  init = rep(0, length(total))
-  if(!DS){
-    init[length(init)] = 1
-    # init = c(0, 0, 1)
-  }else{
-    if(entropy == "EL"){
-      init[1] = -1
-    }else if(entropy == "ET"){
-      # init[1] = 0
-    }else if(entropy == "SL"){
-      init[1] = 1
-    }else if(entropy == "CE"){
-      init[1] = -1
-    }else if(entropy == "HD"){
-      init[1] = -1
-    }else if(entropy == "PH"){
-      init[1] = 1 / sqrt(1 + 1 / del^2)
-    }
-  }
-
-  nleqslv_res = nleqslv(init, f, jac = h, d = d, Xs = Xs, 
-                        total = total, entropy = entropy, del = del,
-                        method = method, control = control)
-  if(nleqslv_res$termcd != 1){
-    if(max(abs(f(nleqslv_res$x, d = d, Xs = Xs, total = total,
-                 entropy = entropy, del = del))) > 1e-5)
-      print(nleqslv_res)
-      w = NA
-  }else{
-    w = f(nleqslv_res$x, d = d, Xs = Xs, total = total,
-          entropy = entropy, del = del, returnw = T)             
-  }
-  return(w)
-}
+#f = function(lambda, d, Xs, total, entropy, del, ..., returnw = F){
+#  # w = d * ginv(drop(Xs %*% lambda), entropy = entropy)
+#  if(entropy == "HD" & any(Xs %*% lambda >= 0)) return(rep(Inf, length(lambda)))
+#  if(entropy == "PH" & any(abs(Xs %*% lambda) >= del)) return(rep(Inf, length(lambda)))
+#  
+#  if(entropy == "SL"){
+#    w = d * drop(Xs %*% lambda)
+#  }else if(entropy == "EL"){
+#    w = -d / drop(Xs %*% lambda)
+#  }else if(entropy == "ET"){
+#    w = d * exp(drop(Xs %*% lambda))
+#  }else if(entropy == "CE"){
+#    w = d / (1 - exp(drop(Xs %*% lambda)))
+#  }else if(entropy == "HD"){
+#    w = d / drop(Xs %*% lambda)^2
+#  }else if(entropy == "PH"){
+#    w = d / sqrt(1 / drop(Xs %*% lambda)^2 - 1 / del^2)
+#  }
+#  if(entropy != "SL" & any(w <= 0)) return(rep(Inf, length(lambda)))
+#  if(entropy == "CE" & any(w <= 1)) return(rep(Inf, length(lambda)))
+#  if(returnw == T){
+#    return(w)
+#  }else{
+#    return(colSums(Xs * w) - total)
+#  }
+#}
+#
+#h = function(lambda, d, Xs, total, entropy, del){
+#  # return(t(Xs) %*% (Xs * d * ginvprime(drop(Xs %*% lambda), entropy = entropy)))
+#  if(entropy == "SL"){
+#    return(t(Xs) %*% (Xs * d))
+#  }else if(entropy == "EL"){
+#    w = -d / drop(Xs %*% lambda)
+#    return(t(Xs) %*% (Xs * (w^2 / d)))
+#  }else if(entropy == "ET"){
+#    w = d * exp(drop(Xs %*% lambda))
+#    return(t(Xs) %*% (Xs * w))
+#  }else if(entropy == "CE"){
+#    p_Stmp = 1 / (1 - exp(drop(Xs %*% lambda)))
+#    return(t(Xs) %*% (Xs * d * p_Stmp * (p_Stmp - 1)))
+#  }else if(entropy == "HD"){
+#    return(t(Xs) %*% (Xs * (-2 * d / drop(Xs %*% lambda)^3)))
+#  }else if(entropy == "PH"){
+#    return(t(Xs) %*% (Xs * (d * (1 - (drop(Xs %*% lambda) / del)^2)^(-1.5))))
+#  }
+#}
+##' @export
+#GEcalib = function(Xs, d, total, entropy = c("SL", "EL", "ET", "CE", "HD", "PH"), ...,
+#       DS = T, method = "Newton", control = list(maxit = 1e5, allowSingular = T)){
+#  
+#  del = quantile(d, 0.80) 
+#  
+#  init = rep(0, length(total))
+#  if(!DS){
+#    init[length(init)] = 1
+#    # init = c(0, 0, 1)
+#  }else{
+#    if(entropy == "EL"){
+#      init[1] = -1
+#    }else if(entropy == "ET"){
+#      # init[1] = 0
+#    }else if(entropy == "SL"){
+#      init[1] = 1
+#    }else if(entropy == "CE"){
+#      init[1] = -1
+#    }else if(entropy == "HD"){
+#      init[1] = -1
+#    }else if(entropy == "PH"){
+#      init[1] = 1 / sqrt(1 + 1 / del^2)
+#    }
+#  }
+#  
+#  nleqslv_res = nleqslv(init, f, jac = h, d = d, Xs = Xs, 
+#                        total = total, entropy = entropy, del = del,
+#                        method = method, control = control)
+#  if(nleqslv_res$termcd != 1){
+#    if(max(abs(f(nleqslv_res$x, d = d, Xs = Xs, total = total,
+#                 entropy = entropy, del = del))) > 1e-5)
+#      print(nleqslv_res)
+#      w = NA
+#  }else{
+#    w = f(nleqslv_res$x, d = d, Xs = Xs, total = total,
+#          entropy = entropy, del = del, returnw = T)             
+#  }
+#  return(w)
+#}
 
@@
 # Sample
+set.seed(2)
 samp_df <- gen_samp(pop_df)

### 2024-04-16 src/explore/20240411-tpdcsim.qmd
Updated prelim_paper.tex. Mostly finished with section 1.

@@
 X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
-X_{3i} &\stackrel{ind}{\sim} N(-1, 1) \\
+X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
+X_{4i} &\stackrel{ind}{\sim} Unif(0.1, 0.9) \\
 \varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
-Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i.
+Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
+\pi_{1i} &= \Phi(-x_{3i} - 2)\\
+\pi_{2i|1} &= x_{4i}.
 \end{aligned}
@@
   x2 <- runif(obs, 0, 4)
-  x3 <- rnorm(obs, -1, 1)
+  x3 <- rnorm(obs, 0, 1)
+  x4 <- runif(obs, 0.1, 0.9)
   eps <- rnorm(obs)
@@
 
-  return(tibble(X1 = x1, X2 = x2, X3 = x3, Y = y))
+  return(tibble(X1 = x1, X2 = x2, X3 = x3, X4 = x4, Y = y))
 }
@@
   # Phase 1 
-  p1_ind <- sample(1:nrow(pop_df), size = p1_obs)
-  p1_df <- filter(pop_df, row_number() %in% p1_ind)
+  # Poisson sampling
+  # mean(pi_i) = 0.1
+  # ptemp <- pt(-pop_df$X3 - 2, 3)
+  # pi_i <- ifelse(ptemp > 0.7, 0.7, ptemp)
+  ptemp <- pt(-pop_df$X3, 3)
+  pi_i <- ptemp
+  del1 <- rbinom(nrow(pop_df), 1, pi_i)
+  p1_df <- pop_df %>%
+    mutate(pi1 = pi_i) %>%
+    rename(pi2 = X4) %>%
+    mutate(del1 = del1) %>%
+    filter(del1 == 1)
+
+  # SRS
+  # p1_ind <- sample(1:nrow(pop_df), size = p1_obs)
+  # p1_df <- filter(pop_df, row_number() %in% p1_ind)
 
   # Phase 2
-  p2_ind <- sample(1:nrow(p1_df), size = p2_obs)
-  p2_df <- filter(p1_df, row_number() %in% p2_ind)
+  # del2 <- rbinom(nrow(p1_df), 1, p1_df$pi2)
+  del2 <- rbinom(nrow(p1_df), 1, p1_df$pi2 / 2)
+
+  p2_df <- p1_df %>%
+    mutate(del2 = del2) %>%
+    filter(del2 == 1)
+
+  # SRS
+  # p2_ind <- sample(1:nrow(p1_df), size = p2_obs)
+  # p2_df <- filter(p1_df, row_number() %in% p2_ind)
 
@@
   extra <- list(...)
-  n1 <- nrow(p1_df)
-  n2 <- nrow(p2_df)
 
-  theta <- sum(p2_df$Y / n2)
-  vhat <- (1 / n2 - 1 / pop_obs) * var(p2_df$Y)
+  theta <- sum(p2_df$Y / (p2_df$pi1 * p2_df$pi2)) / pop_obs
+
+  # SRS Variance
+  # vhat <- (1 / n2 - 1 / pop_obs) * var(p2_df$Y)
+
+  # Poisson Variance
+  pis <- p2_df$pi1 * p2_df$pi2
+  vhat <- sum((1 / pis^2 - 1 / pis) * p2_df$Y^2) / pop_obs^2
 
@@
   extra <- list(...)
-  n1 <- nrow(p1_df)
-  n2 <- nrow(p2_df)
 
@@
   # Step 2: Predict
-  theta <- mean(predict(mod, p1_df))
+  # SRS
+  # theta <- mean(predict(mod, p1_df))
+
+  # Poisson
+  theta <- (sum(predict(mod, p1_df) / p1_df$pi1) + 
+  sum(1 / (p2_df$pi1 * p2_df$pi2) * (p2_df$Y - predict(mod, p2_df)))) / pop_obs
 
   # Step 3: Get variance estimator
-  vhat <-
-    (1 / n1 - 1 / pop_obs) * (mod$coefficients[2]^2 * var(p1_df$X1) + 
-    mod$coefficients[3]^2 * var(p1_df$X2)) +
-    (1 / n2 - 1 / n1) * var(p2_df$Y - predict(mod, p2_df))
+  # SRS
+  # vhat <-
+  #   (1 / n1 - 1 / pop_obs) * (mod$coefficients[2]^2 * var(p1_df$X1) + 
+  #   mod$coefficients[3]^2 * var(p1_df$X2)) +
+  #   (1 / n2 - 1 / n1) * var(p2_df$Y - predict(mod, p2_df))
+
+  # Poisson
+  p12_df <- 
+    left_join(p1_df, p2_df, by = join_by(X1, X2, X3, Y, pi1, pi2, del1)) %>%
+    mutate(del2 = ifelse(is.na(del2), 0, 1))
+  pred_12 <- predict(mod, p12_df)
+  eta <- pred_12 + p12_df$del2 / p12_df$pi2 * (p12_df$Y - pred_12)
+
+  v1 <- sum((1 - p12_df$pi1) / p12_df$pi1^2 * eta^2)
+  v2 <- sum(1 / (p2_df$pi1 * p2_df$pi2) * 
+           (1 / p2_df$pi2 - 1) * (p2_df$Y - predict(mod, p2_df))^2)
+
+  vhat <- (v1 + v2) / pop_obs^2
 
@@
 
-  samp_X <- select(p2_df, X1, X2) %>% as.matrix()
-  samp_X <- cbind(rep(1, nrow(p2_df)), samp_X)
-  pop_tot <- extra$pop_tot
-  # FIXME: This does not work
-  # TODO: Change everything to Poisson sample.
+  d_vec <- 1 / (p2_df$pi2 * p2_df$pi1)
+  if (entropy == "EL") {
+    gdi <- -1 / d_vec
+    gpinv <- d_vec^2
+
+  } else if (entropy == "ET") {
+    gdi <- log(d_vec)
+    gpinv <- d_vec
+
+  } else {
+    stop("This type of entropy has not yet been implemented.")
+  }
+
+  samp_X <- cbind(rep(1, nrow(p2_df)), as.matrix(select(p2_df, X1, X2)))
   dc_w <-  
-    GEcalib(samp_X,
+    GEcalib(cbind(samp_X, gdi),
             d = rep(1, nrow(p2_df)),
-            total = pop_tot,
+            extra$pop_tot,
             entropy = entropy,
-            DS = TRUE)
+            DS = FALSE)
 
@@
 
-  Pi_dc <- diag(1 - n2 / pop_obs) * (dc_w / N_obs)^2
-  z_dc <- model.matrix(~X1 + X2, data = p2_df)
-  gam_dc <- solve(t(z_dc) %*% z_dc, t(z_dc) %*% samp_df$Y)
+  Pi_dc <- diag(1 - 1 / d_vec) * (dc_w / pop_obs)^2
+  z_dc <- model.matrix(~1 + p2_df$X1 + p2_df$X2 + gdi)
+  gam_dc <- 
+    solve(t(z_dc) %*% diag(gpinv) %*% z_dc, t(z_dc) %*% diag(gpinv) %*% p2_df$Y)
   vhat <- 
@@
 
-  return(list(theta = theta, Var = vhat))
+  return(list(theta = theta, Var = as.numeric(vhat)))
 }
@@
   extra <- list(...)
-  n1 <- nrow(p1_df)
-  n2 <- nrow(p2_df)
 
+  # Get estimates of population totals
+  pop_est <- c(pop_obs,
+               sum(p1_df$X1 / p1_df$pi1),
+               sum(p1_df$X2 / p1_df$pi1),
+               extra$pop_gdi)
+
+  if ("entropy" %in% names(extra)) {
+    entropy <- extra$entropy
+  } else {
+    entropy <- "EL"
+  }
+
+  d_vec <- 1 / (p2_df$pi2 * p2_df$pi1)
+  if (entropy == "EL") {
+    gdi <- -1 / d_vec
+    gpinv <- d_vec^2
+
+  } else if (entropy == "ET") {
+    gdi <- log(d_vec)
+    gpinv <- d_vec
+
+  } else {
+    stop("This type of entropy has not yet been implemented.")
+  }
+
+  samp_X <- cbind(rep(1, nrow(p2_df)), as.matrix(select(p2_df, X1, X2)))
+  dc_w <-  
+    GEcalib(cbind(samp_X, gdi),
+            d = rep(1, nrow(p2_df)),
+            pop_est,
+            entropy = entropy,
+            DS = FALSE)
+
+  # Mean Estimation
+  theta <- sum(p2_df$Y * dc_w) / pop_obs
+
+  # Variance Estimation
+  ex_var_mat <- matrix(nrow = 2, ncol = 2)
+  ex_var_mat[1, 1] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X1^2)
+  ex_var_mat[1, 2] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X1 * p1_df$X2)
+  ex_var_mat[2, 1] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X1 * p1_df$X2)
+  ex_var_mat[2, 2] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X2^2)
+
+  Pi_dc <- diag(1 - 1 / d_vec) * (dc_w / pop_obs)^2
+  z_dc <- model.matrix(~1 + p2_df$X1 + p2_df$X2 + gdi)
+  gam_dc <- 
+    solve(t(z_dc) %*% diag(gpinv) %*% z_dc, t(z_dc) %*% diag(gpinv) %*% p2_df$Y)
+  vhat <- 
+  (p2_df$Y - t(gam_dc) %*% t(z_dc)) %*% Pi_dc %*% t(p2_df$Y - t(gam_dc) %*% t(z_dc)) +
+  (gam_dc[2:3] %*% ex_var_mat %*% gam_dc[2:3] / pop_obs^2)
+
+  return(list(theta = theta, Var = as.numeric(vhat)))
 }
@@
 
-  pistar_est <- pi_star(p1_df, p2_df, nrow(pop_df))
-  tpreg_est <- tp_reg(p1_df, p2_df, nrow(pop_df))
-  dcpop_est <- 
-    dc_poptots(p1_df, p2_df, nrow(pop_df),
-               pop_tot = c(nrow(pop_df), sum(pop_df$X1), sum(pop_df$X2)))
+  # ptemp <- pt(-pop_df$X3 - 2, 3)
+  # pop_pi1 <- ifelse(ptemp > 0.7, 0.7, ptemp)
+  ptemp <- pt(-pop_df$X3, 3)
+  pop_pi1 <- ptemp
+
+  pistar <- pi_star(p1_df, p2_df, nrow(pop_df), pop_df = pop_df)
+  tpreg <- tp_reg(p1_df, p2_df, nrow(pop_df))
+  dcpop <- 
+    dc_poptots(
+      p1_df,
+      p2_df,
+      nrow(pop_df),
+      pop_tot = c(nrow(pop_df), sum(pop_df$X1), sum(pop_df$X2), sum(-pop_pi1))
+    )
+  dcest <-
+    dc_esttots(
+      p1_df,
+      p2_df,
+      nrow(pop_df),
+      pop_gdi = sum(-pop_pi1)
+    )
 
   return(
-    tibble(Est = c("pistar", "tpreg"),
-           Theta = c(pistar_est[[1]], tpreg_est[[1]]),
-           Var = c(pistar_est[[2]], tpreg_est[[2]]))
+    tibble(Est = c("pistar", "tpreg", "dcpop", "dcest"),
+           Theta = c(pistar[[1]], tpreg[[1]], dcpop[[1]], dcest[[1]]),
+           Var = c(pistar[[2]], tpreg[[2]], dcpop[[2]], dcest[[2]]))
   )
@@
 
+# Find any failed estimates
+which(is.na(mc_res$Theta))
+which(is.na(mc_res$Var))
+
 # Analyze Mean
@@
   group_by(Est) %>%
-  summarize(bias = mean(err),
-            rmse = sqrt(mean(err^2)),
-            EmpCI = mean(CI),
-            sdest = sd(Theta)) %>%
+  summarize(Bias = mean(err, na.rm = TRUE),
+            RMSE = sqrt(mean(err^2, na.rm = TRUE)),
+            EmpCI = mean(CI, na.rm = TRUE),
+            sdest = sd(Theta, na.rm = TRUE)) %>%
   ungroup() %>%
-  mutate(Ttest = abs(bias) / sqrt(sdest^2 / B_sims)) %>%
-  select(-sdest)
+  mutate(Ttest = abs(Bias) / sqrt(sdest^2 / B_sims)) %>%
+  select(-sdest) %>%
+  mutate(Ind = case_when(
+    Est == "dcest" ~ 4,
+    Est == "dcpop" ~ 3,
+    Est == "tpreg" ~ 2,
+    Est == "pistar" ~ 1)
+  ) %>%
+  mutate(Est = case_when(
+    Est == "dcest" ~ "DC-Est",
+    Est == "dcpop" ~ "DC-Pop",
+    Est == "tpreg" ~ "TP-Reg",
+    Est == "pistar" ~ "$\\pi^*$")
+  ) %>%
+  arrange(Ind) %>%
+  select(Est, Bias, RMSE, EmpCI, Ttest) %>%
+  knitr::kable("latex", booktabs = TRUE, escape = FALSE, digits = 3) %>%
+  cat(file = "tables/tpdcsim_mean.tex")
 
@@
   group_by(Est) %>%
-  summarize(mcvar = var(Theta),
-            estvar = mean(Var),
-            varvar = var(Var)) %>%
+  summarize(mcvar = var(Theta, na.rm = TRUE),
+            estvar = mean(Var, na.rm = TRUE),
+            varvar = var(Var, na.rm = TRUE)) %>%
   ungroup() %>%

### 2024-04-18 final/reports/prelim_paper.tex
Updated prelim_paper but need to fix part 1 theory.

@@
 
+\DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\argmax}{arg\,max}
 \newcommand{\MAP}{{\text{MAP}}}
-\newcommand{\argmax}{{\text{argmax}}}
-\newcommand{\argmin}{{\text{argmin}}}
 \newcommand{\Cov}{{\text{Cov}}}
@@
 in our estimates or improve the estimator's efficiency. For many problems, the
-additional information consist of summary statistics with standard errors. The
+additional information consists of summary statistics with standard errors. The
 goal of this project is to incorporate external information with existing data 
@@
 sample; then, we extend this setup to consider non-nested two-phase samples;
-and finally, we consider the more general approach of having multiple sources as
-the first phase sample.
+and finally, we consider the more general approach of having multiple sources.
 
@@
 Consider a finite population of size $N$ containing elements $(X_i, Y_i)$ where
-an initial (Phase 1) sample of size $n$ is selected and $X_i$ is observed. Then
-from the Phase 1 sample of elements, a (Phase 2) sample of size $r < n$ is
+an initial (Phase 1) sample of size $n_1$ is selected and $X_i$ is observed. Then
+from the Phase 1 sample of elements, a (Phase 2) sample of size $n_2 < n_1$ is
 selected and $Y_i$ is observed. This is two-phase sampling (See 
 \cite{fuller2009sampling}, \cite{kim2024statistics} for general references.) The
-goal of two-phase sampling is to construct an estimator of $Y$ not
-only using the observed information from the Phase 2 sample but also
-incorporating the extra auxiliary information of $X$ from the Phase 1 sample, and
-the challenge is doing this efficiently.
+goal of two-phase sampling is to construct an estimator of $\bar Y_N$ 
+that uses both the observed information in the Phase 2 sample and also the extra
+auxiliary information from $X$ in the Phase 1 sample.
+The challenge is doing this efficiently.
 
@@
 
-$$ \pi_i(A_1) = \pi_{2i | 1} \pi_{1i}.$$
+$$ \pi_i^{(2)}(A_1) = \pi_{2i | 1} \pi_{1i}.$$
 
-%\textcolor{red}{Also, the above equality is not true under two-phase sampling!)}
-\textcolor{red}{I agree that having the superscript $(1)$ and $(2)$ is
-  unnecessary. I have removed them. I have also modified the previous equation.
-  I agree that the invariance condition does not hold for two-phase sampling,
-  but if we consider $\pi_i$ as a function of $A_1$ then I believe this holds.
-  Is this correct? Should we not have $\pi_{i} = \pi_{1i} \pi_{2i|1}(A_1)$
-  because this is how conditional distributions work? (Also, I thought that this
-  is equivalent to the reverse sampling idea of Fey (1992), Shao and Steel
-(1999), and Kim et al. (2006).) I would love to discuss this with you.}
+%\textcolor{blue}{Also, the above equality is not true under two-phase sampling!)}
+%\textcolor{red}{I agree that having the superscript $(1)$ and $(2)$ is
+%  unnecessary. I have removed them. I have also modified the previous equation.
+%  I agree that the invariance condition does not hold for two-phase sampling,
+%  but if we consider $\pi_i$ as a function of $A_1$ then I believe this holds.
+%  Is this correct? Should we not have $\pi_{i} = \pi_{1i} \pi_{2i|1}(A_1)$
+%  because this is how conditional distributions work? (Also, I thought that this
+%  is equivalent to the reverse sampling idea of Fey (1992), Shao and Steel
+%(1999), and Kim et al. (2006).) I would love to discuss this with you.}
 
@@
 $$ \hat Y_{reg, tp} 
-= \sum_{i \in A_1} \frac{1}{\pi_{1i}} x_i \hat \beta_2+ \sum_{i \in A_2}
-\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_2)$$
+= \sum_{i \in A_1} \frac{1}{\pi_{1i}} x_i \hat \beta_q+ \sum_{i \in A_2}
+\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_q)$$
 
@@
 $$
-\hat \beta_2 = \left(\sum_{i \in A_2} 
+\hat \beta_q = \left(\sum_{i \in A_2} 
   \frac{x_i x_i'}{\pi_{1i} q_i}\right)^{-1} 
-\sum_{i \in A_2} \frac{x_i y_i}{\pi_{1i} q_i}. 
-\footnote{
-  \textcolor{blue}{Caleb, we do not have to use $\pi_{2i \mid 1}^{(2)}$ in computing
-  $\hat{\beta}_2$. Please check my book.}
-  \textcolor{red}{Thanks for the check. I
-checked it and I must have gotten $\pi_{1i}$ and $\pi_{2i}$ confused.}}
-\footnote{\textcolor{red}{I was initially thinking of using $\hat \beta_2$ without $q_i$ to be
-  more consistent with how I defined the two-phase regression estimator;
-  however, I see your point about using $q_i$ to connect this estimator with the
-superpopulation model.}}
-  $$
+\sum_{i \in A_2} \frac{x_i y_i}{\pi_{1i} q_i}.$$ 
+%\footnote{
+%  \textcolor{blue}{Caleb, we do not have to use $\pi_{2i \mid 1}^{(2)}$ in computing
+%  $\hat{\beta}_2$. Please check my book.}
+%  \textcolor{red}{Thanks for the check. I
+%checked it and I must have gotten $\pi_{1i}$ and $\pi_{2i}$ confused.}}
+%\footnote{\textcolor{red}{I was initially thinking of using $\hat \beta_2$ without $q_i$ to be
+%  more consistent with how I defined the two-phase regression estimator;
+%  however, I see your point about using $q_i$ to connect this estimator with the
+%superpopulation model.}}
 
@@
 This means that $\hat Y_{reg, tp}$ is also a calibration estimator. The idea
-that regression estimation is a form of calibration was extended by 
-\cite{deville1992calibration} to consider loss functions other than just squared
-loss. They generalized the loss function to minimize $\sum_i G(w_i, d_i)q_i$ for
-weights $w_i$ and design-weights $d_i$ where $G(\cdot)$ is non-negative, strictly
-convex function with respect to $w$, defined on an interval containing $d_i$,
-with $g(w_i, d_i) = \partial G / \partial w$ continuous.
-\footnote{The \cite{deville1992calibration} paper considers regression estimators
-for a single phase setup, which we apply to our two-phase example. } This
+that regression estimation is a form of calibration was noted by
+\cite{deville1992calibration} and extended by them to consider loss functions
+other than just squared loss. Their generalized loss function minimizes
+$\sum_i G(w_i, d_i)q_i$ for weights $w_i$ and design-weights $d_i$ where
+$G(\cdot)$ is a non-negative, strictly convex function with respect to $w$,
+defined on an interval containing $d_i$, with $g(w_i, d_i) = \partial G /
+\partial w$ continuous.\footnote{The \cite{deville1992calibration} paper
+considers regression estimators for a single phase setup, which we apply to our
+two-phase example.} This
 generalization includes empirical likelihood estimation, and maximum entropy
@@
 
-While this generalization is useful to an analyst who may want different
-properties of their estimator from maximum entropy estimation rather than the
-minimal squared loss, 
-% TODO: I am not sure if this is correct -----
-unless $\pi_{2i|1}^{-1} = x_i'a$ for some $a$, the
-estimator is not design consistent. 
-\footnote{ \textcolor{blue}{I do not understand this part.} 
-  \textcolor{red}{I changed the
-notation so that it would not be the unused column space notation. This is from
-Equation (11.17) of \cite{kim2024statistics}.}} 
-\footnote{\textcolor{red}{I think that I am confused about the purpose of having
-the assumption from Equation (11.17) in your sampling book. Could we discuss
-this?}}
-% --------------------------------------------
+%While this generalization is useful to an analyst who may want different
+%properties of their estimator from maximum entropy estimation rather than the
+%minimal squared loss, 
+%% I am not sure if this is correct -----
+%unless $\pi_{2i|1}^{-1} = x_i'a$ for some $a$, the
+%estimator is not design consistent. 
+%\footnote{ \textcolor{blue}{I do not understand this part.} 
+%  \textcolor{red}{I changed the
+%notation so that it would not be the unused column space notation. This is from
+%Equation (11.17) of \cite{kim2024statistics}.}} 
+%\footnote{\textcolor{red}{I think that I am confused about the purpose of having
+%the assumption from Equation (11.17) in your sampling book. Could we discuss
+%this?}}
+%% --------------------------------------------
 
-Furthermore, at a conceptual level, the regression estimator
-has a nice feature that its two terms can be thought about as minimizing
+Furthermore, the regression estimator
+has a nice feature that its two terms can be thought about as minimizing the
 variance and bias correction,
@@
 $$ \hat Y_{reg, tp} 
-= \underbrace{\sum_{i \in A_1} \frac{x_i \hat \beta_2}{\pi_{1i}}}_{
+= \underbrace{\sum_{i \in A_1} \frac{x_i \hat \beta_q}{\pi_{1i}}}_{
   \text{ Minimizing the variance}} + \underbrace{\sum_{i \in A_2}
-\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_2)}_{
+\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_q)}_{
 \text{Bias correction}}.$$
@@
 The \cite{deville1992calibration} method incorporates the design weights into
-the loss function, which is the part minimizing the variance. Instead, there is
-a desire to get design consistency from the calibration term. In 
+the loss function, which is the part minimizing the variance. We would rather
+separate have bias calibration separate from the minimizing the variance so that
+we can control each in isolation. In
 \cite{kwon2024debiased}, the authors show that for a generalized entropy
@@
 
-The method of \cite{kwon2024debiased} requires known calibration
-levels (no uncertainty) for the finite population. It does not handle the
+The method of \cite{kwon2024debiased} requires known finite population 
+calibration levels. It does not handle the
 two-phase setup where we need to estimate the finite population total of $x$
-from the Phase 1 sample. Hence, we extend this method to have a valid variance 
-estimator when including estimated Phase 1 weights.
+from the Phase 1 sample. In the rest of the section, we extend this method to 
+two phase sampling so that we have a valid 
+estimator when including estimated Phase 1 weights with appropriate variance
+estimation.
 
-\subsection*{Methodology}
+\subsection{Methodology}
 
@@
 \begin{equation}\label{eq:calconst1}
-  \sum_{i \in A_2} \frac{x_i w_{2i}}{\pi_{1i}} = 
-\sum_{i \in A_1} \frac{x_i}{\pi_{1i}}
+  \sum_{i \in A_2} \frac{x_i w_{2i}q_i}{\pi_{1i}} = 
+\sum_{i \in A_1} \frac{x_iq_i}{\pi_{1i}}
 \end{equation}
@@
 $g(w) = \partial G / \partial w$. 
-% TODO: Elaborate on this more.
 The original method of \cite{kwon2024debiased} only considered having finite
 population quantities on the right hand side of \ref{eq:calconst1}.
-Let $z_i =
-(x_i, g(\pi_{2i|1}^{-1})$. To solve this minimization problem, we use the convex
-conjugate function $F(w) = -G(g^{-1}) + g^{-1}(w)w$, and instead find
+Let $z_i = (x_i, g(\pi_{2i|1}^{-1}))$. Letting $w_{1i} = \pi_{1i}^{-1}$,
+the the goal is to solve,
 
-$$ \hat \lambda = \argmin_{\lambda \in \Lambda_A} \sum_{i \in A_2} F(w_i) +
-\lambda_1^T \sum_{i \in A_1} \frac{z_i}{\pi_{2i|1}}$$
+\begin{equation}\label{eq:primal}
+  \argmin_{w_{2|1}} \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i 
+  \text{ such that}
+  \sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i = \sum_{i \in A_1} w_{1i} z_i
+  q_i.
+\end{equation}
 
-for $\Lambda_A = \{\lambda: \lambda^T z_i \in g(\mathcal{V}) \text{ for all } i
-\in A_2\}$.
+Let $\hat w_{2i|1}$ be the solution to Equation~\ref{eq:primal}, then the
+estimate of $Y_N$ is $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i|1} y_i$.
 
-\subsection*{Theoretical Results}
+\subsection{Theoretical Results}
 
-We prove two results in this section is the same spirit as \cite{kwon2024debiased}.
-First, we state the assumptions of our analysis then we give the results.
+%\begin{itemize}
+%  \item[(A1)] $G(w)$ is strictly convex and twice differentiable in an interval
+%    $\mathcal{V} \in \R$,
+%  \item[(A2)] There exists $c_1, c_2 \in \mathcal{V}$ with $c_1, c_2 > 0$ such
+%    that $c_1 < \pi_{2i|1} < c_2$ for $i = 1, \dots, n_2$, and 
+%    that $c_1 < \pi_{1i} < c_2$ for $i = 1, \dots, n_1$,
+%  \item[(A3)] If $\pi_{2ij|1}$ is joint inclusion probability of elements $i$
+%    and $j$ in $A_2$, the Phase 2 sample and 
+%    $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$ then,
+%    $$\limsup_{n \to \infty} \max_{i,j \in U: i \neq j} |\Delta_{2ij|1}| <
+%    \infty,$$
+%    and if $\pi_{1ij}$ is the joint inclusion probability of elements $i$ and
+%    $j$ in $A_1$, the Phase 1 sample and 
+%    $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ then,
+%    $$\limsup_{n \to \infty} \max_{i,j \in U: i \neq j} |\Delta_{1ij}| <
+%    \infty,$$
+%  \item[(A4)] Assume that $\Sigma_z = \lim_{N \to \infty} \sum_{i \in U} z_i
+%    z_i^T/N$ exists and is positive definite, the average fourth moment of $(y_i,
+%    x_i^T)$ is finite ($\limsup_{N \to \infty} \sum_{i \in U} ||(y_i, x_i^T)||^4
+%    / N < \infty$), and $\Gamma(\lambda) = \lim_{N \to \infty}  \sum_{i \in U}
+%    f'(\lambda^T z_i) z_i z_i^T / N$ exists in a neighborhood around $\lambda_0
+%    = (\bf 0, 1)$.
+%\end{itemize}
 
-\begin{itemize}
-  \item[(A1)] $G(w)$ is strictly convex and twice differentiable in an interval
-    $\mathcal{V} \in \R$,
-  \item[(A2)] There exists $c_1, c_2 \in \mathcal{V}$ with $c_1, c_2 > 0$ such
-    that $c_1 < \pi_{2i|1} < c_2$ for $i = 1, \dots, n_2$, and 
-    that $c_1 < \pi_{1i} < c_2$ for $i = 1, \dots, n_1$,
-  \item[(A3)] If $\pi_{2ij|1}$ is joint inclusion probability of elements $i$
-    and $j$ in $A_2$, the Phase 2 sample and 
-    $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$ then,
-    $$\limsup_{n \to \infty} \max_{i,j \in U: i \neq j} |\Delta_{2ij|1}| <
-    \infty,$$
-    and if $\pi_{1ij}$ is the joint inclusion probability of elements $i$ and
-    $j$ in $A_1$, the Phase 1 sample and 
-    $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ then,
-    $$\limsup_{n \to \infty} \max_{i,j \in U: i \neq j} |\Delta_{1ij}| <
-    \infty,$$
-  \item[(A4)] Assume that $\Sigma_z = \lim_{N \to \infty} \sum_{i \in U} z_i
-    z_i^T/N$ exists and is positive definite, the average fourth moment of $(y_i,
-    x_i^T)$ is finite ($\limsup_{N \to \infty} \sum_{i \in U} ||(y_i, x_i^T)||^4
-    / N < \infty$), and $\Gamma(\lambda) = \lim_{N \to \infty}  \sum_{i \in U}
-    f'(\lambda^T z_i) z_i z_i^T / N$ exists in a neighborhood around $\lambda_0
-    = (\bf 0, 1)$.
-\end{itemize}
+%\begin{theorem}[Design Consistency]
+%  Suppose that Conditions (A1) - (A4) hold. Then the solution $\hat w$ to 
+%  Equation~\ref{eq:primalloss} subject to Equation~\ref{eq:calconst1} and 
+%  Equation~\ref{eq:calconst2} 
+%  exists and is unique with probability approaching 1. Furthermore, the
+%  estimator $\hat \theta_{DCE} = \sum_{i \in A_2} \frac{\hat w_i y_i}{\pi_{1i}}$
+%  satisfies
+%
+%  $$\hat \theta_{DCE} = \hat \theta_{DC} + o_p(n_2^{-1})$$
+%
+%  where $\hat \theta_{DC}$ is the debiased calibration estimator of 
+%  \cite{kwon2024debiased}.
+%
+%\end{theorem}
+%
+%\begin{proof}
+%  This proof follows the proof of \cite{kwon2024debiased} in a straightforward
+%  manner. There are two main modifications that need to be made. The first is
+%  that the proof needs to be written for two-phase sampling with $n_1$ and $n_2$
+%  instead of a single $n$. The second is that we need to ensure that the result
+%  still holds with an estimated population constraint that is $O_p(n_1^{-1/2})$
+%  consistent.
+%
+%  With the Conditions (A1) - (A4) written with respect the the two-phase sample,
+%  the first challenge is done. The second step largely holds from the existing
+%  proof of \cite{kwon2024debiased} with modifications such as 
+%
+%  $$\hat Q(\lambda) = n_2^{-1} \sum_{i \in A_2} F(\lambda^T z_i^*) - n_1^{-1}
+%  \lambda^T \sum_{i \in A_1} z_i^*$$ 
+%
+%  ensuring that our extension holds.
+%
+%%  \textcolor{red}{Dr. Kim, I know that I need to write up this proof in more
+%%  detail, but I am unsure about the level of granularity that is required now. I
+%%  can basically rewrite everything from the Appendix of \cite{kwon2024debiased}
+%%  with the two-phase modification to show this result, but I don't know if I
+%%  should yet.}
+%\end{proof}
 
-\begin{theorem}[Design Consistency]
-  Suppose that Conditions (A1) - (A4) hold. Then the solution $\hat w$ to 
-  Equation~\ref{eq:primalloss} subject to Equation~\ref{eq:calconst1} and 
-  Equation~\ref{eq:calconst2} 
-  exists and is unique with probability approaching 1. Furthermore, the
-  estimator $\hat \theta_{DCE} = \sum_{i \in A_2} \frac{\hat w_i y_i}{\pi_{1i}}$
-  satisfies
+%We need the following additional assumptions to prove asymptotic normality.
+%
+%\begin{itemize}
+%  \item[(B1)] 
+%  \item[(B2)] The HT estimator is asymptotically normal under the sampling
+%    design in the sense that 
+%
+%\end{itemize}
 
-  $$\hat \theta_{DCE} = \hat \theta_{DC} + o_p(n_2^{-1})$$
+\begin{theorem}[Variance Estimation]
+  Let $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1} \pi_{2j|1}$ and 
+  $\hat \eta_i = z_i \hat \phi / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
+  - z_i \hat \phi / q_i\right)$ where
 
-  where $\hat \theta_{DC}$ is the debiased calibration estimator of 
-  \cite{kwon2024debiased}.
+  $$\hat \phi = \left[\sum_{i \in A_2} w_{1i}\left(\frac{z_i z_i^T}{
+  g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in A_2} w_{1i}
+    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
 
+  An unbiased estimator of the variance of $\hat Y_{DCE}$ is
+
+  \begin{align*}
+    \hat V 
+    &= \sum_{i \in A_1} \sum_{j \in A_1} \left(\frac{\pi_{1ij} -
+    \pi_{1i}\pi_{1j}}{\pi_{1ij}}
+    w_{1i} w_{1j} \hat \eta_i \hat \eta_j \right) \\ 
+    &+ \sum_{i \in A_2} \sum_{j \in A_2} w_{1ij} w_{2ij|1} w_{2i|1} w_{2j|1} 
+    \Delta_{2ij|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).
+  \end{align*}
 \end{theorem}
@@
 \begin{proof}
-  This proof follows the proof of \cite{kwon2024debiased} in a straightforward
-  manner. There are two main modifications that need to be made. The first is
-  that the proof needs to be written for two-phase sampling with $n_1$ and $n_2$
-  instead of a single $n$. The second is that we need to ensure that the result
-  still holds with an estimated population constraint that is $O_p(n_1^{-1/2})$
-  consistent.
+  First, we derive the solution for $\hat Y_{DCE}$, and then derive the
+  asymptotic variance of this estimator using the linearization technique of
+  \cite{randles1982asymptotic}.
 
-  With the Conditions (A1) - (A4) written with respect the the two-phase sample,
-  the first challenge is done. The second step largely holds from the existing
-  proof of \cite{kwon2024debiased} with modifications such as
+  We want to solve Equation~\ref{eq:primalloss} subject to the constraints in
+  Equations~\ref{eq:calconst1} and \ref{eq:calconst2}. Using the method of
+  Lagrange multipliers we have a Lagrangian of
 
-  $$\hat Q(\lambda) = n_2^{-1} \sum_{i \in A_2} F(\lambda^T z_i^*) - n_1^{-1}
-  \lambda^T \sum_{i \in A_1} z_i^*$$
+  $$\mathcal{L}(w_2) = -\sum_{i \in A_2} w_{1i} G(w_{2i})q_i + \lambda 
+  \left(\sum_{i \in A_2} w_{1i} w_{2i} z_i - \sum_{i \in A_1} w_{1i} z_i\right).$$
 
-  ensuring that our extension holds.
+  Our first order conditions yield,
 
-  \textcolor{red}{Dr. Kim, I know that I need to write up this proof in more
-  detail, but I am unsure about the level of granularity that is required now. I
-  can basically rewrite everything from the Appendix of \cite{kwon2024debiased}
-  with the two-phase modification to show this result, but I don't know if I
-  should yet.}
+  \begin{align*}
+    0 &= \frac{\partial}{\partial w_{2i}} \mathcal{L}(w_{2}) \\ 
+      &= -w_{1i} g(w_{2i}) q_i + \lambda w_{1i} z_i.
+  \end{align*}
+
+  This means that $\hat w_{2i} = g^{-1}(\hat \lambda z_i / q_i)$ and since $\hat
+  \lambda$ is determined from the solution to Equations~\ref{eq:calconst1} and
+  \ref{eq:calconst2}. Hence, $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} 
+  \hat w_{2i}(\hat \lambda) y_i$.
+
+  To get the variance of $\hat Y_{DCE}$ we use the linearization technique of
+  \cite{randles1982asymptotic}. First, notice that
+
+  $$\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i =
+  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i + \left(
+  \sum_{i \in A_1} w_{1i} z_i -
+  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) z_i\right)\phi
+  := \hat Y_\ell(\hat \lambda, \phi).$$
+
+  We can find the value of $\phi$ such that $E[\frac{\partial}{\partial
+  \lambda}\hat Y_\ell(\lambda, \phi)]_{\lambda = \lambda^*}$ where $\lambda^*$
+  is the probability limit of $\hat \lambda$. Notice, that 
+  $\lambda^* = (\bf 0, 1)$ because $\hat w_{2i} \to \pi_{2i|1}^{-1}$. This means
+  that the optimal value of $\phi$ solves,
+
+  $$E\left[\frac{\partial}{\partial \lambda} \hat Y_\ell(\lambda, \phi)\right] =
+  E\left[\sum_{i \in A_2} w_{1i} \frac{\partial}{\partial \lambda} w_{2i|1}(\lambda)
+    y_i - \phi \sum_{i \in A_2} w_{1i} z_i \frac{\partial}{\partial \lambda}
+    w_{2i|1}(\lambda) / q_i \right]$$
+
+  which means that
+
+  $$\phi^* = \left[\sum_{i \in U} \pi_{2i|1} \left(\frac{z_i z_i^T}{
+    g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in U} \pi_{2i|1}
+    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
+
+  Hence,
+
+  \begin{align*}
+    \hat Y_\ell(\lambda^*, \phi^*) 
+    &= \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1}
+    y_i + \left(\sum_{i \in A_1} w_{1i} z_i / q_i -
+      \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} z_i / q_i\right)\phi^* \\ 
+    &= \sum_{i \in A_1} w_{1i} \eta_i
+  \end{align*}
+
+  where
+
+  $$\eta_i = z_i \phi^* / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
+  - z_i \phi^* / q_i\right).$$
+
+  To analyze the variance, consider the reverse framework of
+  \cite{fay1992inferences}, \cite{kim2006replication}, and \cite{shao1999variance}.
+  In this setup, we assume that $\mathcal{R} = \{\delta_{2i|1}\}_{i \in U}$ is
+  known at the
+  finite population level. Then with this we have,
+
+  \begin{align*}
+    V(\hat Y_{\ell}(\lambda^*, \phi^*)) 
+    &= V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\right) \\
+    &= V\left(E\left[\sum_{i \in A_1} w_{1i}\eta_i(\phi^*)\mid\mathcal{R}\right]\right) 
+    +E\left[V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\mid\mathcal{R}\right)\right]\\
+    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right) + E\left[\sum_{i \in U} 
+    \sum_{j \in U} \Cov(\delta_{1i}, \delta_{1j}) 
+    w_{1i} w_{1j} \eta_i(\phi^*) \eta_j(\phi^*)\right]\\
+  \end{align*}
+
+  We can estimate this variance using
+
+  $$\hat V_{DCE} = \sum_{i \in A_1} \sum_{j \in A_1} \left(\frac{\pi_{1ij} -
+  \pi_{1i}\pi_{1j}}{\pi_{1ij}}
+    w_{1i} w_{1j} \hat \eta_i \hat \eta_j \right)
+  $$
+
+  for $\hat \eta_i = z_i \hat \phi / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
+  - z_i \hat \phi / q_i\right)$ where
+
+  $$\hat \phi = \left[\sum_{i \in A_2} w_{1i}\left(\frac{z_i z_i^T}{
+  g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in A_2} w_{1i}
+    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
+
+  The problem with this estimator is that it is biased as $E[\hat V_{DCE}]$ is
+  asymptotically equivalent to
+  $V\left(\sum_{i \in A_1} w_{1i} \eta(\phi^*)\right)$. This yields a bias of
+
+  \begin{align*}
+    \text{Bias}(\hat V_{DCE}) 
+    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right)\\ 
+    &= \sum_{i \in U} \sum_{j \in U} \Delta_{2ij|1} w_{2i|1}
+    w_{2j|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).
+  \end{align*}
+
+  where $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1} \pi_{2j|1}$. 
+  However, we can estimate this bias with
+
+  $$\widehat{\text{Bias}}(\hat V_{DCE}) = 
+  \sum_{i \in A_2} \sum_{j \in A_2} w_{1ij} w_{2ij|1} w_{2i|1} w_{2j|1} 
+  \Delta_{2ij|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).$$
+
 \end{proof}
 
-We need the following additional assumptions to prove asymptotic normality.
-
-% FIXME: Add assumptions
-\textcolor{red}{I need to think about the assumptions that we need in two-phase
-sampling.}
-\begin{itemize}
-  \item[(B1)] 
-  \item[(B2)] The HT estimator is asymptotically normal under the sampling
-    design in the sense that
-
-\end{itemize}
-
-\begin{theorem}[Asymptotic Normality]
-% FIXME: Add result.
-\textcolor{red}{I need to write out more but we should have something like this:
-  $$V(\hat \theta_{DCE} = \Var(\hat \theta_{DC}) + \hat \gamma_{[1:3]}^T V(\bf
-x)) \hat \gamma_{[1:3]} + \Cov(\hat \theta_{DC}, \bf x \hat \gamma_{[1:3}]).$$}
-\end{theorem}
-
-\textcolor{red}{Dr. Kim, I am not convinced that $\Cov(\hat \theta_{DC}, \bf x
-\hat \gamma_{[1:3]}) = 0$. The reason that this covariance holds in
-\cite{fuller2009sampling} Section 3.3.3 is that the estimating equation requires
-$\sum_i x_i \hat e_i = 0$, which means that $\Cov(\hat e, x_i) = 0$. However, we
-do not have this requirement. Instead, I see this more like a nested model where
-we have a group level error that aggregates the individual level errors. The
-problem with the current simulation that I cannot tell if this covariance should
-be zero because the variance term $\hat \gamma_{[1:3]}'V(\bf x) \hat \gamma_{[1:3]}$
-dominates the variance estimation. I will change the simulation parameters to
-try to see if I can get a better test of this.}
-
-\subsection*{Simulation Studies}
+\subsection{Simulation Studies}
 
@@
 
-\section*{Topic 2: Non-nested Two-Phase Sampling}
+\section{Topic 2: Non-nested Two-Phase Sampling}
 
 \textcolor{blue}{
-( Materials in Section 11.4 can be used here. I have copy-and-pasted the textbook materials below. Please modify them.  )
+( Materials in Section 11.4 can be used here. I have copy-and-pasted the
+textbook materials below. Please modify them.  )
 }
@@
 
-In contrast to the classical two-phase sampling framework, non-nested two-phase sampling involves two independent surveys conducted on the same target population.
-The key distinction is that the two samples, denoted as $A_1$ and $A_2$, are drawn independently rather than sequentially.
-Table \ref{table:11-1} presents the data structure for non-nested two-phase sampling.
+\subsection{Background}
 
-In the non-nested two-phase sampling, 
-a large probability sample $A_1$ is drawn from a finite population,  collecting  only the $\bx$ variable, and  a  smaller sample $A_2$ is drawn from the same population, providing information  on both the $y$ and $\bx$ variables.
-It is assumed that the observed variable $x$ is comparable in the two surveys. \cite{renssen1997} formally addressed this non-nested two-phase sampling problem  and 
-\cite{merkouris2004} extended the idea further to develop regression estimation combining information from multiple surveys. 
-\cite{kimrao12} considered the non-nested two-phase sampling in the context of mass imputation combining two independent surveys at the population and domain levels.
+Now we consider the sampling mechanism known as non-nested two phase sampling 
+(\cite{kim2024statistics}). In the last section, we considered two phase sampling
+in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
+two phase sampling the Phase 2 sample is independent of the Phase 1 sample. It
+is a separate independent sample of the same population. Like traditional two
+phase sampling, we consider the Phase 1 sample, $A_1$, to consist of
+observations of $(X_i)_{i = 1}^{n_1}$ and the Phase 2 sample, $A_2$, to consist
+of observations of $(X_i, Y_i)_{i = 1}^{n_2}$.
 
-\begin{table}[htb]
-\caption{Data Structure for non-nested two-phase sampling}
-\label{table:1}\par
-\vskip .2cm
-\centerline{\tabcolsep=3truept\begin{tabular}{|c|cc|}
-\hline
-Sample  & $X$ & $Y$ \\
-\hline
-$A_1$ & \checkmark &  \\
-$A_2$ & \checkmark & \checkmark  \\
-\hline
-\end{tabular}}
-\label{table:11-1}
-\end{table}
+Whereas the classical two phase estimator uses a single Horvitz-Thompson
+estimator of the Phase 1 sample to construct estimates for calibration totals,
+in the non-nested two phase sample we have two independent Horvitz-Thompson
+estimators of the total of $X$,
 
-To illustrate the non-nested two-phase sampling approach, let's consider the data structure shown in Table \ref{table:11-1}.
-This setup involves two independent samples, $A_1$ and $A_2$, drawn from the same target population.
+$$\hat X_1 = \sum_{i \in A_1}^{n_1} \frac{x_i}{\pi_{1i}} \text{ and } 
+\hat X_2 = \sum_{i \in A_2}^{n_2} \frac{x_i}{\pi_{2i}}$$
 
-From these two samples, we can compute two unbiased estimators of the population total $\mathbf{X} = \sum_{i=1}^N \bx_i$ for the auxiliary variable x: $\hat{\mathbf{X}}_1 = \sum_{i \in A_1} \pi_{1i}^{-1} \bx_i$
-and 
-$\hat{\mathbf{X}}_2 = \sum_{i \in A_2} \pi_{2i}^{-1} \bx_i$. 
-Here, $\pi_{1i}$ and $\pi_{2i}$ represent the inclusion probabilities for samples $A_1$ and $A_2$, respectively.
+where $\pi_{1i}$ is the probability of $i \in A_1$ and $\pi_{2i} = \Pr(i \in
+A_2)$. \textcolor{red}{I don't know if it makes sense to change the Phase 2
+selection probability as $\pi_{2i|1}$. Should I keep it at $\pi_{2i}$? But if
+I do, how is the two phase setup useful?} We can combine these estimates to get
+$\hat X_c = W \hat X_1 + (1 - W)\hat X_2$ for some matrix $W$ (see
+\cite{merkouris2004combining} for the optimal choice of $W$). We can then define
+a regression estimator as
 
-Both $\hat{\mathbf{X}}_1$ and $\hat{\mathbf{X}}_2$ are unbiased estimators of the population total $\mathbf{X}$ under the respective sampling designs. 
-The availability of these two unbiased estimators is a key feature of the non-nested two-phase sampling design, as it provides opportunities for developing enhanced estimation procedures combining information from different sources.
+$$
+\hat Y_{NN, reg} = \hat Y_2 + (\hat X_c - \hat X_2)^T \hat \beta_q = 
+\hat Y_2 + (\hat X_1 - \hat X_2)^T W\hat \beta_q 
+$$
 
-We can  construct a combined estimator of X, denoted as $\widehat{\mathbf{X}}_c$, as follows:
-\begin{equation}
-\widehat{\mathbf{X}}_c = W \hat{\mathbf{X}}_1 + (I - W) \hat{\mathbf{X}}_2, 
-\label{xgls}
+where
+
+$$\hat \beta_q = \left(\sum_{i \in A_2} \frac{x_i x_i^T}{q_i}\right)^{-1}
+\sum_{i \in A_2} \frac{x_i y_i}{q_i} \text{ and }\hat Y_2 = \sum_{i \in A_2}
+\frac{y_i}{\pi_{2i}}.$$
+
+Since the samples $A_1$ and $A_2$ are independent,
+
+$$V(\hat Y_{NN, reg}) = V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i -
+x_i^TW\beta^*_q)\right) + (\beta^*_q)^T W^T V(\hat X_1) W \beta_q^*$$
+
+where $\beta_q^*$ is the probability limit of $\hat \beta_q$. Like the two phase
+sample this regression estimator can be viewed as the solution to the following
+calibration equation where $d_{2i} = \pi_{2i}^{-1}$,
+
+\begin{equation}\label{eq:nncal}
+\hat w = \argmin_w Q(w) = \sum_{i \in A_2} (w_{2i} - d_{2i})^2 q_i \text{ such
+that } \sum_{i \in A_2} w_{2i} x_i = \hat X_c
 \end{equation}
-where $W$ is a $p \times p$ symmetric matrix of constants, and $p = \text{dim}(\bx)$ is the dimension of the auxiliary variable x.
-The optimal choice of the matrix W can be determined using the Generalized Least Squares (GLS) method.
-However, other choices of W can also be used.
-The key idea is to leverage the information from these two independent surveys to obtain a more accurate and efficient estimator of the population total X for the auxiliary variable x, compared to using only one of the surveys alone.
 
-Using the combined estimator  $\widehat{\mathbf{X}}_{\rm c} $ in (\ref{xgls}), we can construct the following projection estimator:  
-\begin{equation}
- \widehat{Y}_p = \widehat{\mathbf{X}}_{\rm c}' \hat{\bm \beta}_q  
- \label{11-proj}
- \end{equation}
-where the regression coefficient estimator $\hat{\bm \beta}_q$ is defined as 
-$$ 
-\hat{\bm \beta}_q = \left( \sum_{i \in A_2} \bx_i \bx_i' / q_i \right)^{-1} 
-\sum_{i \in A_2} \bx_i y_i / q_i. 
-$$
-The choice of $q_i$ in the regression coefficient estimator is somewhat arbitrary. Two possible choices are:
-\begin{enumerate}
-\item Using the model variance under a regression superpopulation model.
-\item Using $q_i = \pi_{2i}^{-2} - \pi_{2i}^{-1}$ to compute the design-optimal regression estimator under Poisson sampling.
-\end{enumerate}
-The key idea is that by using the combined estimator $\widehat{\mathbf{X}}_c$ in the projection estimator $\widehat{Y}_p$, we can leverage the information from both the A1 and A2 samples to obtain a more accurate prediction of the variable of interest Y.
-The choice of $q_i$ allows for some flexibility in how the regression coefficient is estimated.
+and $\hat Y_{NN, reg} = \sum_{i \in A_2} \hat w_{2i} y_i$ where $\hat
+w_{2i}$ is the solution to Equation~\ref{eq:nncal}.
 
-To ensure the design-consistency of the projection estimator in (\ref{11-proj}), we can use the following regression estimator under non-nested two-phase sampling:
-\begin{equation}
-\widehat{Y}_{\rm tp, reg} = \widehat{Y}_2 + \left( \widehat{\mathbf{X}}_{\rm c} - \widehat{\mathbf{X}}_2 \right)' \hat{\bm \beta}_q
-\label{eq:11-28}
-\end{equation} 
-By the definition of $\widehat{\mathbf{X}}_{\rm c}$, we can also express this as: 
-\begin{equation}
-\widehat{Y}_{\rm tp, reg} = \widehat{Y}_2 + \left( \widehat{\mathbf{X}}_{1} - \widehat{\mathbf{X}}_2 \right)' \hat{\bm \alpha}_q,
-\label{eq:11-28}
-\end{equation} 
-where $\hat{\bm \alpha}_q = W \hat{\bm \beta}_q$.
-The key points are:
-\begin{enumerate}
-\item The design-consistent regression estimator $\widehat{Y}_{\rm tp, reg}$ is constructed by adding a correction term to the projection estimator $\widehat{Y}_p$ from the second sample.
-\item The regression estimator improves the efficiency of the  design unbiased estimator  $\hat{Y}_2$ by substracting the projection of $\hat{Y}_2$ onto the augmentation space \citep{tsiatis2006}, 
-the linear space generated by 
- the difference between the combined estimator $\widehat{\mathbf{X}}_{\rm c}$ and the estimator $\widehat{\mathbf{X}}_2$ from the second sample.
-\item 
-Alternatively, the augmentation space  can be expressed using the difference between the estimators $\widehat{\mathbf{X}}_{1}$ and $\widehat{\mathbf{X}}_2$, weighted by $\hat{\bm \alpha}_q$.
-\end{enumerate} 
-The goal is to leverage the information from both samples to obtain a design-consistent regression estimator for the variable of interest Y.
+We extend the debiased calibration estimator of \cite{kwon2024debiased} to the
+non-nested two phase sampling case where we use a combined estimate $\hat X_c$
+as the calibration totals instead of using the true totals from the finite
+population.
 
-Using the standard argument, we can obtain 
-\begin{eqnarray}
-\widehat{Y}_{\rm tp, reg} 
-&=& \widehat{Y}_2 +  \left( \widehat{\mathbf{X}}_{1} - \widehat{\mathbf{X}}_2 \right)' {\bm \alpha}_q^* + O_p(n^{-1} N) \label{eq:11-29}
-\end{eqnarray}
-where $\bm \alpha_q^*$ is the probability limit of $\hat{\bm \alpha}_q = W \hat{\bm \beta}_q$. By (\ref{eq:11-29}), 
- we can obtain 
-\begin{equation}
-V \left( \widehat{Y}_{\rm tp, reg}  \right) 
-= ({\bm \alpha}_q^* )' V \left( \widehat{\mathbf{X}}_{1}  \right) {\bm \alpha}_q^* + V\left( \hat{u}_2 \right) 
-\label{eq:11-31}
-\end{equation}
-where $\hat{u}_2 = \sum_{i \in A_2} \pi_{2i}^{-1} \left( y_i -  \bx_i'\bm \alpha_q^* \right) $. From the formula in (\ref{eq:11-31}), we can construct a linearized variance estimator.
+\subsection{Methodology}
 
-Now, we can use the calibration weighting to construct the regression estimator under non-nested two-phase sampling. For  given the design weights $d_{2i} = \pi_{2i}^{-1}$, we find the minimizer of 
-    $$Q \left( {\bm \omega } \right) = \sum_{i \in A_2} \left(  {\omega_i} - d_{2i}  \right)^2 q_i $$
-    subject to 
-    $$ \sum_{i \in A_2} {\omega_i} \bx_i = \widehat{\mathbf{X}}_{\rm c} .$$
-    The solution is 
-    $$ \hat{\omega}_i = d_{2i} + \left( \widehat{\mathbf{X}}_{\rm c} - \widehat{\mathbf{X}}_{\rm 2} \right)^{-1} \left( \sum_{i \in A_2} q_i^{-1} \bx_i \bx_i' \right)^{-1} \bx_i q_i^{-1} . $$
-     Note that 
-    $$ \sum_{i \in A_2} \hat{\omega}_i y_i = \widehat{Y}_{\rm tp, reg} , $$
-    where  $\widehat{Y}_{\rm tp, reg}$ is defined in (\ref{eq:11-28}). 
-    Thus, the algebraic equivalence between the  regression estimator and the calibration weighting estimator is established under non-nested two-phase sampling.
+\subsection{Theoretical Results}
 
-\section*{Topic 3: Multi-source Two-Phase Sampling}
+\subsection{Simulation Studies}
+
+%In contrast to the classical two-phase sampling framework, non-nested two-phase
+%sampling involves two independent surveys conducted on the same target
+%population. The key distinction is that the two samples, denoted as $A_1$ and
+%$A_2$, are drawn independently rather than sequentially.  Table \ref{table:11-1}
+%presents the data structure for non-nested two-phase sampling. 
+%
+%In the non-nested two-phase sampling, 
+%a large probability sample $A_1$ is drawn from a finite population,  collecting
+%only the $\bx$ variable, and  a  smaller sample $A_2$ is drawn from the same
+%population, providing information  on both the $y$ and $\bx$ variables.
+%It is assumed that the observed variable $x$ is comparable in the two surveys.
+%\cite{renssen1997} formally addressed this non-nested two-phase sampling problem
+%and 
+%\cite{merkouris2004} extended the idea further to develop regression estimation
+%combining information from multiple surveys. 
+%\cite{kimrao12} considered the non-nested two-phase sampling in the context of
+%mass imputation combining two independent surveys at the population and domain
+%levels.
+%
+%
+%
+%\begin{table}[htb]
+%\caption{Data Structure for non-nested two-phase sampling}
+%\label{table:1}\par
+%\vskip .2cm
+%\centerline{\tabcolsep=3truept\begin{tabular}{|c|cc|}
+%\hline
+%Sample  & $X$ & $Y$ \\
+%\hline
+%$A_1$ & \checkmark &  \\
+%$A_2$ & \checkmark & \checkmark  \\
+%\hline
+%\end{tabular}}
+%\label{table:11-1}
+%\end{table}
+%
+%To illustrate the non-nested two-phase sampling approach, let's consider the
+%data structure shown in Table \ref{table:11-1}. This setup involves two
+%independent samples, $A_1$ and $A_2$, drawn from the same target population.
+%
+%From these two samples, we can compute two unbiased estimators of the population
+%total $\mathbf{X} = \sum_{i=1}^N \bx_i$ for the auxiliary variable x:
+%$\hat{\mathbf{X}}_1 = \sum_{i \in A_1} \pi_{1i}^{-1} \bx_i$
+%and 
+%$\hat{\mathbf{X}}_2 = \sum_{i \in A_2} \pi_{2i}^{-1} \bx_i$. 
+%Here, $\pi_{1i}$ and $\pi_{2i}$ represent the inclusion probabilities for
+%samples $A_1$ and $A_2$, respectively.
+%
+%Both $\hat{\mathbf{X}}_1$ and $\hat{\mathbf{X}}_2$ are unbiased estimators of
+%the population total $\mathbf{X}$ under the respective sampling designs. 
+%The availability of these two unbiased estimators is a key feature of the
+%non-nested two-phase sampling design, as it provides opportunities for
+%developing enhanced estimation procedures combining information from different
+%sources.  
+%
+%We can  construct a combined estimator of X, denoted as
+%$\widehat{\mathbf{X}}_c$, as follows:
+%\begin{equation}
+%\widehat{\mathbf{X}}_c = W \hat{\mathbf{X}}_1 + (I - W) \hat{\mathbf{X}}_2, 
+%\label{xgls}
+%\end{equation}
+%where $W$ is a $p \times p$ symmetric matrix of constants, and $p =
+%\text{dim}(\bx)$ is the dimension of the auxiliary variable x. The optimal
+%choice of the matrix W can be determined using the Generalized Least Squares
+%(GLS) method. However, other choices of W can also be used. The key idea is to
+%leverage the information from these two independent surveys to obtain a more
+%accurate and efficient estimator of the population total X for the auxiliary
+%variable x, compared to using only one of the surveys alone. 
+%
+%
+%Using the combined estimator  $\widehat{\mathbf{X}}_{\rm c} $ in (\ref{xgls}),
+%we can construct the following projection estimator:  
+%\begin{equation}
+% \widehat{Y}_p = \widehat{\mathbf{X}}_{\rm c}' \hat{\bm \beta}_q  
+% \label{11-proj}
+% \end{equation}
+%where the regression coefficient estimator $\hat{\bm \beta}_q$ is defined as 
+%$$ 
+%\hat{\bm \beta}_q = \left( \sum_{i \in A_2} \bx_i \bx_i' / q_i \right)^{-1} 
+%\sum_{i \in A_2} \bx_i y_i / q_i. 
+%$$
+%The choice of $q_i$ in the regression coefficient estimator is somewhat
+%arbitrary. Two possible choices are:
+%\begin{enumerate}
+%\item Using the model variance under a regression superpopulation model.
+%\item Using $q_i = \pi_{2i}^{-2} - \pi_{2i}^{-1}$ to compute the design-optimal regression estimator under Poisson sampling.
+%\end{enumerate}
+%The key idea is that by using the combined estimator $\widehat{\mathbf{X}}_c$ in
+%the projection estimator $\widehat{Y}_p$, we can leverage the information from
+%both the A1 and A2 samples to obtain a more accurate prediction of the variable
+%of interest Y. The choice of $q_i$ allows for some flexibility in how the
+%regression coefficient is estimated.
+%
+%
+%To ensure the design-consistency of the projection estimator in (\ref{11-proj}),
+%we can use the following regression estimator under non-nested two-phase
+%sampling:
+%\begin{equation}
+%\widehat{Y}_{\rm tp, reg} = \widehat{Y}_2 + \left( \widehat{\mathbf{X}}_{\rm c} - \widehat{\mathbf{X}}_2 \right)' \hat{\bm \beta}_q
+%\label{eq:11-28}
+%\end{equation} 
+%By the definition of $\widehat{\mathbf{X}}_{\rm c}$, we can also express this as: 
+%\begin{equation}
+%\widehat{Y}_{\rm tp, reg} = \widehat{Y}_2 + \left( \widehat{\mathbf{X}}_{1} - \widehat{\mathbf{X}}_2 \right)' \hat{\bm \alpha}_q,
+%\label{eq:11-28}
+%\end{equation} 
+%where $\hat{\bm \alpha}_q = W \hat{\bm \beta}_q$.
+%The key points are:
+%\begin{enumerate}
+%\item The design-consistent regression estimator $\widehat{Y}_{\rm tp, reg}$ is
+%  constructed by adding a correction term to the projection estimator
+%  $\widehat{Y}_p$ from the second sample.
+%\item The regression estimator improves the efficiency of the  design unbiased
+%  estimator  $\hat{Y}_2$ by substracting the projection of $\hat{Y}_2$ onto the
+%  augmentation space \citep{tsiatis2006}, the linear space generated by 
+% the difference between the combined estimator $\widehat{\mathbf{X}}_{\rm c}$
+% and the estimator $\widehat{\mathbf{X}}_2$ from the second sample.
+%\item 
+%Alternatively, the augmentation space  can be expressed using the difference
+%between the estimators $\widehat{\mathbf{X}}_{1}$ and $\widehat{\mathbf{X}}_2$,
+%weighted by $\hat{\bm \alpha}_q$.
+%\end{enumerate} 
+%The goal is to leverage the information from both samples to obtain a
+%design-consistent regression estimator for the variable of interest Y.
+%
+%Using the standard argument, we can obtain 
+%\begin{eqnarray}
+%\widehat{Y}_{\rm tp, reg} 
+%&=& \widehat{Y}_2 +  \left( \widehat{\mathbf{X}}_{1} - \widehat{\mathbf{X}}_2 \right)' {\bm \alpha}_q^* + O_p(n^{-1} N) \label{eq:11-29}
+%\end{eqnarray}
+%where $\bm \alpha_q^*$ is the probability limit of $\hat{\bm \alpha}_q = W \hat{\bm \beta}_q$. By (\ref{eq:11-29}), 
+% we can obtain 
+%\begin{equation}
+%V \left( \widehat{Y}_{\rm tp, reg}  \right) 
+%= ({\bm \alpha}_q^* )' V \left( \widehat{\mathbf{X}}_{1}  \right) {\bm \alpha}_q^* + V\left( \hat{u}_2 \right) 
+%\label{eq:11-31}
+%\end{equation}
+%where $\hat{u}_2 = \sum_{i \in A_2} \pi_{2i}^{-1} \left( y_i -  \bx_i'\bm
+%\alpha_q^* \right) $. From the formula in (\ref{eq:11-31}), we can construct a
+%linearized variance estimator. 
+%
+%Now, we can use the calibration weighting to construct the regression estimator
+%under non-nested two-phase sampling. For  given the design weights $d_{2i} =
+%\pi_{2i}^{-1}$, we find the minimizer of 
+%    $$Q \left( {\bm \omega } \right) = \sum_{i \in A_2} \left(  {\omega_i} - d_{2i}  \right)^2 q_i $$
+%    subject to 
+%    $$ \sum_{i \in A_2} {\omega_i} \bx_i = \widehat{\mathbf{X}}_{\rm c} .$$
+%    The solution is 
+%    $$ \hat{\omega}_i = d_{2i} + \left( \widehat{\mathbf{X}}_{\rm c} - \widehat{\mathbf{X}}_{\rm 2} \right)^{-1} \left( \sum_{i \in A_2} q_i^{-1} \bx_i \bx_i' \right)^{-1} \bx_i q_i^{-1} . $$
+%     Note that 
+%    $$ \sum_{i \in A_2} \hat{\omega}_i y_i = \widehat{Y}_{\rm tp, reg} , $$
+%    where  $\widehat{Y}_{\rm tp, reg}$ is defined in (\ref{eq:11-28}). 
+%    Thus, the algebraic equivalence between the  regression estimator and the calibration weighting estimator is established under non-nested two-phase sampling.
+
+\section{Topic 3: Multi-source Two-Phase Sampling}
 

### 2024-04-18 src/explore/20240411-tpdcsim.qmd
Updated prelim_paper but need to fix part 1 theory.

@@
   arrange(Ind) %>%
-  select(Est, Bias, RMSE, EmpCI, Ttest) %>%
+  select(Est, Bias, RMSE, EmpCI, Ttest)
+
+%>%
   knitr::kable("latex", booktabs = TRUE, escape = FALSE, digits = 3) %>%

### 2024-04-18 final/reports/prelim_paper.tex
Updated prelim paper. Hopefully done with topic 1.

@@
 
-%\textcolor{blue}{Also, the above equality is not true under two-phase sampling!)}
-%\textcolor{red}{I agree that having the superscript $(1)$ and $(2)$ is
-%  unnecessary. I have removed them. I have also modified the previous equation.
-%  I agree that the invariance condition does not hold for two-phase sampling,
-%  but if we consider $\pi_i$ as a function of $A_1$ then I believe this holds.
-%  Is this correct? Should we not have $\pi_{i} = \pi_{1i} \pi_{2i|1}(A_1)$
-%  because this is how conditional distributions work? (Also, I thought that this
-%  is equivalent to the reverse sampling idea of Fey (1992), Shao and Steel
-%(1999), and Kim et al. (2006).) I would love to discuss this with you.}
-
 This means that we can define the $\pi^*$-estimator as the following design
@@
 
-%While this generalization is useful to an analyst who may want different
-%properties of their estimator from maximum entropy estimation rather than the
-%minimal squared loss, 
-%% I am not sure if this is correct -----
-%unless $\pi_{2i|1}^{-1} = x_i'a$ for some $a$, the
-%estimator is not design consistent. 
-%\footnote{ \textcolor{blue}{I do not understand this part.} 
-%  \textcolor{red}{I changed the
-%notation so that it would not be the unused column space notation. This is from
-%Equation (11.17) of \cite{kim2024statistics}.}} 
-%\footnote{\textcolor{red}{I think that I am confused about the purpose of having
-%the assumption from Equation (11.17) in your sampling book. Could we discuss
-%this?}}
-%% --------------------------------------------
-
 Furthermore, the regression estimator
@@
 
-%The proposed method will be something like this: 
-%\textcolor{red}{Minimize 
-%  $$ \sum_{i \in A_2} \frac{1}{\pi_i^{(1)}}  G( w_{2i} ) q_i $$ 
-%  subject to 
-%  $$ \sum_{i \in A_2} w_{2i}  x_i / \pi_i^{(1)} = \sum_{i \in A_1} x_i / \pi_i^{(1)}$$ 
-%  and 
-%  $$ \sum_{i \in A_2}  \frac{1}{\pi_i^{(1)}} w_{2i} g( \pi_{2i \mid 1}^{-1} ) q_i
-%  = \sum_{i \in A_1}  \frac{1}{\pi_i^{(1)}}  g(\pi_{2i \mid 1}^{-1} ) q_i $$ 
-%  where $g(\omega) = \partial G( \omega)/ \partial \omega$. }
-
-% You need to show that the resulting calibration estimator $\hat{Y}_{\rm cal}=
-% \sum_{i \in A_2} w_{1i} \hat{w}_{2i} y_i$ is asymptotically equivalent to the
-% two-phase regression estimator in (\ref{reg}), where $w_{1i} = 1/ \pi_i^{(1)}$.
-
-% Also, there should be a discussion about variance estimation. Once the
-% linearization form is obtained, the formula for linearized variance estimation
-% should be obtained easily.
-
 We follow the approach of \cite{kwon2024debiased} for the debiased calibration
@@
 population quantities on the right hand side of \ref{eq:calconst1}.
-Let $z_i = (x_i, g(\pi_{2i|1}^{-1}))$. Letting $w_{1i} = \pi_{1i}^{-1}$,
+Let $z_i = (x_i / q_i, g(\pi_{2i|1}^{-1}))$. Letting $w_{1i} = \pi_{1i}^{-1}$,
 the the goal is to solve,
@@
   \text{ such that}
-  \sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i = \sum_{i \in A_1} w_{1i} z_i
-  q_i.
+  \sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i = \sum_{i \in A_1} w_{1i} z_i q_i.
 \end{equation}
@@
 
-%\begin{itemize}
-%  \item[(A1)] $G(w)$ is strictly convex and twice differentiable in an interval
-%    $\mathcal{V} \in \R$,
-%  \item[(A2)] There exists $c_1, c_2 \in \mathcal{V}$ with $c_1, c_2 > 0$ such
-%    that $c_1 < \pi_{2i|1} < c_2$ for $i = 1, \dots, n_2$, and 
-%    that $c_1 < \pi_{1i} < c_2$ for $i = 1, \dots, n_1$,
-%  \item[(A3)] If $\pi_{2ij|1}$ is joint inclusion probability of elements $i$
-%    and $j$ in $A_2$, the Phase 2 sample and 
-%    $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$ then,
-%    $$\limsup_{n \to \infty} \max_{i,j \in U: i \neq j} |\Delta_{2ij|1}| <
-%    \infty,$$
-%    and if $\pi_{1ij}$ is the joint inclusion probability of elements $i$ and
-%    $j$ in $A_1$, the Phase 1 sample and 
-%    $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ then,
-%    $$\limsup_{n \to \infty} \max_{i,j \in U: i \neq j} |\Delta_{1ij}| <
-%    \infty,$$
-%  \item[(A4)] Assume that $\Sigma_z = \lim_{N \to \infty} \sum_{i \in U} z_i
-%    z_i^T/N$ exists and is positive definite, the average fourth moment of $(y_i,
-%    x_i^T)$ is finite ($\limsup_{N \to \infty} \sum_{i \in U} ||(y_i, x_i^T)||^4
-%    / N < \infty$), and $\Gamma(\lambda) = \lim_{N \to \infty}  \sum_{i \in U}
-%    f'(\lambda^T z_i) z_i z_i^T / N$ exists in a neighborhood around $\lambda_0
-%    = (\bf 0, 1)$.
-%\end{itemize}
+\begin{theorem}[Design Consistency]
+  Under regularity conditions,
 
-%\begin{theorem}[Design Consistency]
-%  Suppose that Conditions (A1) - (A4) hold. Then the solution $\hat w$ to 
-%  Equation~\ref{eq:primalloss} subject to Equation~\ref{eq:calconst1} and 
-%  Equation~\ref{eq:calconst2} 
-%  exists and is unique with probability approaching 1. Furthermore, the
-%  estimator $\hat \theta_{DCE} = \sum_{i \in A_2} \frac{\hat w_i y_i}{\pi_{1i}}$
-%  satisfies
-%
-%  $$\hat \theta_{DCE} = \hat \theta_{DC} + o_p(n_2^{-1})$$
-%
-%  where $\hat \theta_{DC}$ is the debiased calibration estimator of 
-%  \cite{kwon2024debiased}.
-%
-%\end{theorem}
-%
-%\begin{proof}
-%  This proof follows the proof of \cite{kwon2024debiased} in a straightforward
-%  manner. There are two main modifications that need to be made. The first is
-%  that the proof needs to be written for two-phase sampling with $n_1$ and $n_2$
-%  instead of a single $n$. The second is that we need to ensure that the result
-%  still holds with an estimated population constraint that is $O_p(n_1^{-1/2})$
-%  consistent.
-%
-%  With the Conditions (A1) - (A4) written with respect the the two-phase sample,
-%  the first challenge is done. The second step largely holds from the existing
-%  proof of \cite{kwon2024debiased} with modifications such as 
-%
-%  $$\hat Q(\lambda) = n_2^{-1} \sum_{i \in A_2} F(\lambda^T z_i^*) - n_1^{-1}
-%  \lambda^T \sum_{i \in A_1} z_i^*$$ 
-%
-%  ensuring that our extension holds.
-%
-%%  \textcolor{red}{Dr. Kim, I know that I need to write up this proof in more
-%%  detail, but I am unsure about the level of granularity that is required now. I
-%%  can basically rewrite everything from the Appendix of \cite{kwon2024debiased}
-%%  with the two-phase modification to show this result, but I don't know if I
-%%  should yet.}
-%\end{proof}
+  $$\hat Y_{DCE} = \hat Y_\ell(\lambda^*, \phi^*) + o_p(N / n_2)$$
 
-%We need the following additional assumptions to prove asymptotic normality.
-%
-%\begin{itemize}
-%  \item[(B1)] 
-%  \item[(B2)] The HT estimator is asymptotically normal under the sampling
-%    design in the sense that 
-%
-%\end{itemize}
+  where
 
-\begin{theorem}[Variance Estimation]
-  Let $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1} \pi_{2j|1}$ and 
-  $\hat \eta_i = z_i \hat \phi / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
-  - z_i \hat \phi / q_i\right)$ where
+  $$\hat Y_{\ell}(\hat \lambda, \phi^*) = \hat Y_{DCE}(\hat \lambda) + 
+  \left(\sum_{i \in A_1} w_{1i} z_i q_i = \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat
+  \lambda) z_i q_i\right)\phi^*$$
 
-  $$\hat \phi = \left[\sum_{i \in A_2} w_{1i}\left(\frac{z_i z_i^T}{
-  g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in A_2} w_{1i}
-    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
+  and
 
-  An unbiased estimator of the variance of $\hat Y_{DCE}$ is
+  $$\phi^* = 
+  \left[\sum_{i \in U} \frac{\pi_{2i|1}z_i z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
+  \sum_{i \in U} \frac{\pi_{2i|1}z_i y_i}{g'(d_{2i|1})}.$$
 
-  \begin{align*}
-    \hat V 
-    &= \sum_{i \in A_1} \sum_{j \in A_1} \left(\frac{\pi_{1ij} -
-    \pi_{1i}\pi_{1j}}{\pi_{1ij}}
-    w_{1i} w_{1j} \hat \eta_i \hat \eta_j \right) \\ 
-    &+ \sum_{i \in A_2} \sum_{j \in A_2} w_{1ij} w_{2ij|1} w_{2i|1} w_{2j|1} 
-    \Delta_{2ij|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).
-  \end{align*}
 \end{theorem}
@@
 \begin{proof}
-  First, we derive the solution for $\hat Y_{DCE}$, and then derive the
-  asymptotic variance of this estimator using the linearization technique of
-  \cite{randles1982asymptotic}.
+  In this proof, we derive the solution to Equation~\ref{eq:primal} and show
+  that it is asymptotically equivalent to a regression estimator. Using the
+  method of Lagrange multipliers, to solve Equation~\ref{eq:primal} we need to
+  minimize,
 
-  We want to solve Equation~\ref{eq:primalloss} subject to the constraints in
-  Equations~\ref{eq:calconst1} and \ref{eq:calconst2}. Using the method of
-  Lagrange multipliers we have a Lagrangian of
+  $$\mathcal{L}(w_{2|1}) = \sum_{i \in A_2} w_{1i} G(w_{2i|1}) q_i + 
+  \lambda \left(\sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i - \sum_{i \in A_1} w_{1i}
+  z_i q_i\right).$$
 
-  $$\mathcal{L}(w_2) = -\sum_{i \in A_2} w_{1i} G(w_{2i})q_i + \lambda 
-  \left(\sum_{i \in A_2} w_{1i} w_{2i} z_i - \sum_{i \in A_1} w_{1i} z_i\right).$$
+  The first order conditions show that
 
-  Our first order conditions yield,
+  $$\frac{\partial \mathcal{L}}{\partial w_{2i|1}}: g(w_{2i|1}) w_{1i}q_i +
+    \lambda w_{1i} z_i q_i = 0.$$
+
+  Hence, $\hat w_{2i}(\lambda) = g^{-1}(\lambda z_i)$ and $\hat \lambda$ is
+  determined by the calibration condition. When the sample size gets large, we
+  have $\hat w_{2i|1}(\hat \lambda) \to d_{2i|1}$ which means that $\hat \lambda
+  \to \lambda^*$ where $\lambda^* = (\bf 0, 1)$. Then using the linearization
+  technique of \cite{randles1982asymptotic}, we can construct a regression
+  estimator,
+
+  $$\hat Y_\ell(\hat \lambda, \phi)  = \hat Y_{DCE}(\hat \lambda) + 
+  \left(\sum_{i \in A_1} w_{1i} z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat
+  \lambda) z_i q_i\right)\phi.$$
+
+  Notice that $\hat Y_\ell(\hat \lambda, \phi) = \hat Y_{DCE}(\hat \lambda)$ for
+  all $\phi \in \mathbb{R}$. We choose $\phi^*$ such that
+
+  $$E\left[\frac{\partial}{\partial \lambda} \hat Y_\ell(\lambda^*, \phi^*)\right]=0.$$
+
+  Using the fact that $g^{-1}(\lambda^* z_i) = g^{-1}(g(d_{2i|1})) = d_{2i|1}$
+  and $(g^{-1})'(x) = 1 / g'(g^{-1}(x))$, we have
 
   \begin{align*}
-    0 &= \frac{\partial}{\partial w_{2i}} \mathcal{L}(w_{2}) \\ 
-      &= -w_{1i} g(w_{2i}) q_i + \lambda w_{1i} z_i.
+    \phi^*
+    &= E\left[\sum_{i \in A_2} \frac{w_{1i}z_i z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
+    E\left[\sum_{i \in A_2} \frac{w_{1i}z_i y_i}{g'(d_{2i|1})}\right]\\
+    &= \left[\sum_{i \in U} \frac{\pi_{2i|1}z_i z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
+    \left[\sum_{i \in U} \frac{\pi_{2i|1} z_i y_i}{g'(d_{2i|1})}\right]\\
   \end{align*}
 
-  This means that $\hat w_{2i} = g^{-1}(\hat \lambda z_i / q_i)$ and since $\hat
-  \lambda$ is determined from the solution to Equations~\ref{eq:calconst1} and
-  \ref{eq:calconst2}. Hence, $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} 
-  \hat w_{2i}(\hat \lambda) y_i$.
+  Thus, the linearization estimator is
 
-  To get the variance of $\hat Y_{DCE}$ we use the linearization technique of
-  \cite{randles1982asymptotic}. First, notice that
+  $$\hat Y_\ell(\lambda^*, \phi^*) = \sum_{i \in A_1} w_{1i} q_i z_i \phi^* +
+  \sum_{i \in A_2} w_{1i} d_{2i|1} (y_i - q_i z_i \phi^*).$$
 
-  $$\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i =
-  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i + \left(
-  \sum_{i \in A_1} w_{1i} z_i -
-  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) z_i\right)\phi
-  := \hat Y_\ell(\hat \lambda, \phi).$$
-
-  We can find the value of $\phi$ such that $E[\frac{\partial}{\partial
-  \lambda}\hat Y_\ell(\lambda, \phi)]_{\lambda = \lambda^*}$ where $\lambda^*$
-  is the probability limit of $\hat \lambda$. Notice, that 
-  $\lambda^* = (\bf 0, 1)$ because $\hat w_{2i} \to \pi_{2i|1}^{-1}$. This means
-  that the optimal value of $\phi$ solves,
-
-  $$E\left[\frac{\partial}{\partial \lambda} \hat Y_\ell(\lambda, \phi)\right] =
-  E\left[\sum_{i \in A_2} w_{1i} \frac{\partial}{\partial \lambda} w_{2i|1}(\lambda)
-    y_i - \phi \sum_{i \in A_2} w_{1i} z_i \frac{\partial}{\partial \lambda}
-    w_{2i|1}(\lambda) / q_i \right]$$
-
-  which means that
-
-  $$\phi^* = \left[\sum_{i \in U} \pi_{2i|1} \left(\frac{z_i z_i^T}{
-    g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in U} \pi_{2i|1}
-    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
-
-  Hence,
+  By construction using a Taylor expansion yields,
 
   \begin{align*}
-    \hat Y_\ell(\lambda^*, \phi^*) 
-    &= \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1}
-    y_i + \left(\sum_{i \in A_1} w_{1i} z_i / q_i -
-      \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} z_i / q_i\right)\phi^* \\ 
-    &= \sum_{i \in A_1} w_{1i} \eta_i
+    \hat Y_{DCE}(\hat \lambda) 
+    &= \hat Y_\ell(\lambda^*, \phi^*) + 
+    E\left[\frac{\partial}{\partial \lambda}\hat Y_\ell(\lambda^*,
+    \phi^*)\right](\hat \lambda - \lambda^*) + \frac{1}{2}
+    E\left[\frac{\partial}{\partial \lambda^2} \hat Y_{DCE}(\lambda^*)\right] (\hat
+    \lambda - \lambda^*)^2\\
+    &= \hat Y_\ell(\lambda^*, \phi^*) + O(N)o_p(n_2^{-1}).
   \end{align*}
 
-  where
-
-  $$\eta_i = z_i \phi^* / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
-  - z_i \phi^* / q_i\right).$$
-
-  To analyze the variance, consider the reverse framework of
-  \cite{fay1992inferences}, \cite{kim2006replication}, and \cite{shao1999variance}.
-  In this setup, we assume that $\mathcal{R} = \{\delta_{2i|1}\}_{i \in U}$ is
-  known at the
-  finite population level. Then with this we have,
-
-  \begin{align*}
-    V(\hat Y_{\ell}(\lambda^*, \phi^*)) 
-    &= V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\right) \\
-    &= V\left(E\left[\sum_{i \in A_1} w_{1i}\eta_i(\phi^*)\mid\mathcal{R}\right]\right) 
-    +E\left[V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\mid\mathcal{R}\right)\right]\\
-    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right) + E\left[\sum_{i \in U} 
-    \sum_{j \in U} \Cov(\delta_{1i}, \delta_{1j}) 
-    w_{1i} w_{1j} \eta_i(\phi^*) \eta_j(\phi^*)\right]\\
-  \end{align*}
-
-  We can estimate this variance using
-
-  $$\hat V_{DCE} = \sum_{i \in A_1} \sum_{j \in A_1} \left(\frac{\pi_{1ij} -
-  \pi_{1i}\pi_{1j}}{\pi_{1ij}}
-    w_{1i} w_{1j} \hat \eta_i \hat \eta_j \right)
-  $$
-
-  for $\hat \eta_i = z_i \hat \phi / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
-  - z_i \hat \phi / q_i\right)$ where
-
-  $$\hat \phi = \left[\sum_{i \in A_2} w_{1i}\left(\frac{z_i z_i^T}{
-  g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in A_2} w_{1i}
-    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
-
-  The problem with this estimator is that it is biased as $E[\hat V_{DCE}]$ is
-  asymptotically equivalent to
-  $V\left(\sum_{i \in A_1} w_{1i} \eta(\phi^*)\right)$. This yields a bias of
-
-  \begin{align*}
-    \text{Bias}(\hat V_{DCE}) 
-    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right)\\ 
-    &= \sum_{i \in U} \sum_{j \in U} \Delta_{2ij|1} w_{2i|1}
-    w_{2j|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).
-  \end{align*}
-
-  where $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1} \pi_{2j|1}$. 
-  However, we can estimate this bias with
-
-  $$\widehat{\text{Bias}}(\hat V_{DCE}) = 
-  \sum_{i \in A_2} \sum_{j \in A_2} w_{1ij} w_{2ij|1} w_{2i|1} w_{2j|1} 
-  \Delta_{2ij|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).$$
+  The final equality comes from the fact that 
+  $E\left[\frac{\partial}{\partial \lambda}\hat Y_\ell(\lambda^*, \phi^*)\right]
+  = 0$, $\frac{\partial}{\partial \lambda^2} \hat w_{2i|1}(\lambda^*)$ is
+  bounded and $|\hat \lambda - \lambda^*| = o_p(n_2^{-1/2})$, which proves our
+  result.
 
 \end{proof}
+
+%\begin{proof}
+%  First, we derive the solution for $\hat Y_{DCE}$, and then derive the
+%  asymptotic variance of this estimator using the linearization technique of
+%  \cite{randles1982asymptotic}.
+%
+%  We want to solve Equation~\ref{eq:primalloss} subject to the constraints in
+%  Equations~\ref{eq:calconst1} and \ref{eq:calconst2}. Using the method of
+%  Lagrange multipliers we have a Lagrangian of 
+%
+%  $$\mathcal{L}(w_2) = -\sum_{i \in A_2} w_{1i} G(w_{2i})q_i + \lambda 
+%  \left(\sum_{i \in A_2} w_{1i} w_{2i} z_i - \sum_{i \in A_1} w_{1i} z_i\right).$$
+%
+%  Our first order conditions yield,
+%
+%  \begin{align*}
+%    0 &= \frac{\partial}{\partial w_{2i}} \mathcal{L}(w_{2}) \\ 
+%      &= -w_{1i} g(w_{2i}) q_i + \lambda w_{1i} z_i.
+%  \end{align*}
+%
+%  This means that $\hat w_{2i} = g^{-1}(\hat \lambda z_i / q_i)$ and since $\hat
+%  \lambda$ is determined from the solution to Equations~\ref{eq:calconst1} and
+%  \ref{eq:calconst2}. Hence, $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} 
+%  \hat w_{2i}(\hat \lambda) y_i$.
+%
+%  To get the variance of $\hat Y_{DCE}$ we use the linearization technique of
+%  \cite{randles1982asymptotic}. First, notice that 
+%
+%  $$\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i =
+%  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i + \left(
+%  \sum_{i \in A_1} w_{1i} z_i -
+%  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) z_i\right)\phi
+%  := \hat Y_\ell(\hat \lambda, \phi).$$
+%
+%  We can find the value of $\phi$ such that $E[\frac{\partial}{\partial
+%  \lambda}\hat Y_\ell(\lambda, \phi)]_{\lambda = \lambda^*}$ where $\lambda^*$
+%  is the probability limit of $\hat \lambda$. Notice, that 
+%  $\lambda^* = (\bf 0, 1)$ because $\hat w_{2i} \to \pi_{2i|1}^{-1}$. This means
+%  that the optimal value of $\phi$ solves,
+%
+%  $$E\left[\frac{\partial}{\partial \lambda} \hat Y_\ell(\lambda, \phi)\right] =
+%  E\left[\sum_{i \in A_2} w_{1i} \frac{\partial}{\partial \lambda} w_{2i|1}(\lambda)
+%    y_i - \phi \sum_{i \in A_2} w_{1i} z_i \frac{\partial}{\partial \lambda}
+%    w_{2i|1}(\lambda) / q_i \right]$$
+%
+%  which means that
+%
+%  $$\phi^* = \left[\sum_{i \in U} \pi_{2i|1} \left(\frac{z_i z_i^T}{
+%    g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in U} \pi_{2i|1}
+%    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
+%
+%  Hence,
+%
+%  \begin{align*}
+%    \hat Y_\ell(\lambda^*, \phi^*) 
+%    &= \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1}
+%    y_i + \left(\sum_{i \in A_1} w_{1i} z_i / q_i -
+%      \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} z_i / q_i\right)\phi^* \\ 
+%    &= \sum_{i \in A_1} w_{1i} \eta_i
+%  \end{align*}
+%
+%  where 
+%
+%  $$\eta_i = z_i \phi^* / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
+%  - z_i \phi^* / q_i\right).$$
+%
+%  To analyze the variance, consider the reverse framework of
+%  \cite{fay1992inferences}, \cite{kim2006replication}, and \cite{shao1999variance}.
+%  In this setup, we assume that $\mathcal{R} = \{\delta_{2i|1}\}_{i \in U}$ is
+%  known at the
+%  finite population level. Then with this we have,
+%
+%  \begin{align*}
+%    V(\hat Y_{\ell}(\lambda^*, \phi^*)) 
+%    &= V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\right) \\
+%    &= V\left(E\left[\sum_{i \in A_1} w_{1i}\eta_i(\phi^*)\mid\mathcal{R}\right]\right) 
+%    +E\left[V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\mid\mathcal{R}\right)\right]\\
+%    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right) + E\left[\sum_{i \in U} 
+%    \sum_{j \in U} \Cov(\delta_{1i}, \delta_{1j}) 
+%    w_{1i} w_{1j} \eta_i(\phi^*) \eta_j(\phi^*)\right]\\
+%  \end{align*}
+%
+%  We can estimate this variance using 
+%
+%  $$\hat V_{DCE} = \sum_{i \in A_1} \sum_{j \in A_1} \left(\frac{\pi_{1ij} -
+%  \pi_{1i}\pi_{1j}}{\pi_{1ij}}
+%    w_{1i} w_{1j} \hat \eta_i \hat \eta_j \right)
+%  $$
+%
+%  for $\hat \eta_i = z_i \hat \phi / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
+%  - z_i \hat \phi / q_i\right)$ where 
+%
+%  $$\hat \phi = \left[\sum_{i \in A_2} w_{1i}\left(\frac{z_i z_i^T}{
+%  g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in A_2} w_{1i}
+%    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
+%
+%  The problem with this estimator is that it is biased as $E[\hat V_{DCE}]$ is
+%  asymptotically equivalent to
+%  $V\left(\sum_{i \in A_1} w_{1i} \eta(\phi^*)\right)$. This yields a bias of 
+%
+%  \begin{align*}
+%    \text{Bias}(\hat V_{DCE}) 
+%    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right)\\ 
+%    &= \sum_{i \in U} \sum_{j \in U} \Delta_{2ij|1} w_{2i|1}
+%    w_{2j|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).
+%  \end{align*}
+%
+%  where $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1} \pi_{2j|1}$. 
+%  However, we can estimate this bias with
+%
+%  $$\widehat{\text{Bias}}(\hat V_{DCE}) = 
+%  \sum_{i \in A_2} \sum_{j \in A_2} w_{1ij} w_{2ij|1} w_{2i|1} w_{2j|1} 
+%  \Delta_{2ij|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).$$
+%
+%\end{proof}
 
@@
 
-%In contrast to the classical two-phase sampling framework, non-nested two-phase
-%sampling involves two independent surveys conducted on the same target
-%population. The key distinction is that the two samples, denoted as $A_1$ and
-%$A_2$, are drawn independently rather than sequentially.  Table \ref{table:11-1}
-%presents the data structure for non-nested two-phase sampling. 
-%
-%In the non-nested two-phase sampling, 
-%a large probability sample $A_1$ is drawn from a finite population,  collecting
-%only the $\bx$ variable, and  a  smaller sample $A_2$ is drawn from the same
-%population, providing information  on both the $y$ and $\bx$ variables.
-%It is assumed that the observed variable $x$ is comparable in the two surveys.
-%\cite{renssen1997} formally addressed this non-nested two-phase sampling problem
-%and 
-%\cite{merkouris2004} extended the idea further to develop regression estimation
-%combining information from multiple surveys. 
-%\cite{kimrao12} considered the non-nested two-phase sampling in the context of
-%mass imputation combining two independent surveys at the population and domain
-%levels.
-%
-%
-%
-%\begin{table}[htb]
-%\caption{Data Structure for non-nested two-phase sampling}
-%\label{table:1}\par
-%\vskip .2cm
-%\centerline{\tabcolsep=3truept\begin{tabular}{|c|cc|}
-%\hline
-%Sample  & $X$ & $Y$ \\
-%\hline
-%$A_1$ & \checkmark &  \\
-%$A_2$ & \checkmark & \checkmark  \\
-%\hline
-%\end{tabular}}
-%\label{table:11-1}
-%\end{table}
-%
-%To illustrate the non-nested two-phase sampling approach, let's consider the
-%data structure shown in Table \ref{table:11-1}. This setup involves two
-%independent samples, $A_1$ and $A_2$, drawn from the same target population.
-%
-%From these two samples, we can compute two unbiased estimators of the population
-%total $\mathbf{X} = \sum_{i=1}^N \bx_i$ for the auxiliary variable x:
-%$\hat{\mathbf{X}}_1 = \sum_{i \in A_1} \pi_{1i}^{-1} \bx_i$
-%and 
-%$\hat{\mathbf{X}}_2 = \sum_{i \in A_2} \pi_{2i}^{-1} \bx_i$. 
-%Here, $\pi_{1i}$ and $\pi_{2i}$ represent the inclusion probabilities for
-%samples $A_1$ and $A_2$, respectively.
-%
-%Both $\hat{\mathbf{X}}_1$ and $\hat{\mathbf{X}}_2$ are unbiased estimators of
-%the population total $\mathbf{X}$ under the respective sampling designs. 
-%The availability of these two unbiased estimators is a key feature of the
-%non-nested two-phase sampling design, as it provides opportunities for
-%developing enhanced estimation procedures combining information from different
-%sources.  
-%
-%We can  construct a combined estimator of X, denoted as
-%$\widehat{\mathbf{X}}_c$, as follows:
-%\begin{equation}
-%\widehat{\mathbf{X}}_c = W \hat{\mathbf{X}}_1 + (I - W) \hat{\mathbf{X}}_2, 
-%\label{xgls}
-%\end{equation}
-%where $W$ is a $p \times p$ symmetric matrix of constants, and $p =
-%\text{dim}(\bx)$ is the dimension of the auxiliary variable x. The optimal
-%choice of the matrix W can be determined using the Generalized Least Squares
-%(GLS) method. However, other choices of W can also be used. The key idea is to
-%leverage the information from these two independent surveys to obtain a more
-%accurate and efficient estimator of the population total X for the auxiliary
-%variable x, compared to using only one of the surveys alone. 
-%
-%
-%Using the combined estimator  $\widehat{\mathbf{X}}_{\rm c} $ in (\ref{xgls}),
-%we can construct the following projection estimator:  
-%\begin{equation}
-% \widehat{Y}_p = \widehat{\mathbf{X}}_{\rm c}' \hat{\bm \beta}_q  
-% \label{11-proj}
-% \end{equation}
-%where the regression coefficient estimator $\hat{\bm \beta}_q$ is defined as 
-%$$ 
-%\hat{\bm \beta}_q = \left( \sum_{i \in A_2} \bx_i \bx_i' / q_i \right)^{-1} 
-%\sum_{i \in A_2} \bx_i y_i / q_i. 
-%$$
-%The choice of $q_i$ in the regression coefficient estimator is somewhat
-%arbitrary. Two possible choices are:
-%\begin{enumerate}
-%\item Using the model variance under a regression superpopulation model.
-%\item Using $q_i = \pi_{2i}^{-2} - \pi_{2i}^{-1}$ to compute the design-optimal regression estimator under Poisson sampling.
-%\end{enumerate}
-%The key idea is that by using the combined estimator $\widehat{\mathbf{X}}_c$ in
-%the projection estimator $\widehat{Y}_p$, we can leverage the information from
-%both the A1 and A2 samples to obtain a more accurate prediction of the variable
-%of interest Y. The choice of $q_i$ allows for some flexibility in how the
-%regression coefficient is estimated.
-%
-%
-%To ensure the design-consistency of the projection estimator in (\ref{11-proj}),
-%we can use the following regression estimator under non-nested two-phase
-%sampling:
-%\begin{equation}
-%\widehat{Y}_{\rm tp, reg} = \widehat{Y}_2 + \left( \widehat{\mathbf{X}}_{\rm c} - \widehat{\mathbf{X}}_2 \right)' \hat{\bm \beta}_q
-%\label{eq:11-28}
-%\end{equation} 
-%By the definition of $\widehat{\mathbf{X}}_{\rm c}$, we can also express this as: 
-%\begin{equation}
-%\widehat{Y}_{\rm tp, reg} = \widehat{Y}_2 + \left( \widehat{\mathbf{X}}_{1} - \widehat{\mathbf{X}}_2 \right)' \hat{\bm \alpha}_q,
-%\label{eq:11-28}
-%\end{equation} 
-%where $\hat{\bm \alpha}_q = W \hat{\bm \beta}_q$.
-%The key points are:
-%\begin{enumerate}
-%\item The design-consistent regression estimator $\widehat{Y}_{\rm tp, reg}$ is
-%  constructed by adding a correction term to the projection estimator
-%  $\widehat{Y}_p$ from the second sample.
-%\item The regression estimator improves the efficiency of the  design unbiased
-%  estimator  $\hat{Y}_2$ by substracting the projection of $\hat{Y}_2$ onto the
-%  augmentation space \citep{tsiatis2006}, the linear space generated by 
-% the difference between the combined estimator $\widehat{\mathbf{X}}_{\rm c}$
-% and the estimator $\widehat{\mathbf{X}}_2$ from the second sample.
-%\item 
-%Alternatively, the augmentation space  can be expressed using the difference
-%between the estimators $\widehat{\mathbf{X}}_{1}$ and $\widehat{\mathbf{X}}_2$,
-%weighted by $\hat{\bm \alpha}_q$.
-%\end{enumerate} 
-%The goal is to leverage the information from both samples to obtain a
-%design-consistent regression estimator for the variable of interest Y.
-%
-%Using the standard argument, we can obtain 
-%\begin{eqnarray}
-%\widehat{Y}_{\rm tp, reg} 
-%&=& \widehat{Y}_2 +  \left( \widehat{\mathbf{X}}_{1} - \widehat{\mathbf{X}}_2 \right)' {\bm \alpha}_q^* + O_p(n^{-1} N) \label{eq:11-29}
-%\end{eqnarray}
-%where $\bm \alpha_q^*$ is the probability limit of $\hat{\bm \alpha}_q = W \hat{\bm \beta}_q$. By (\ref{eq:11-29}), 
-% we can obtain 
-%\begin{equation}
-%V \left( \widehat{Y}_{\rm tp, reg}  \right) 
-%= ({\bm \alpha}_q^* )' V \left( \widehat{\mathbf{X}}_{1}  \right) {\bm \alpha}_q^* + V\left( \hat{u}_2 \right) 
-%\label{eq:11-31}
-%\end{equation}
-%where $\hat{u}_2 = \sum_{i \in A_2} \pi_{2i}^{-1} \left( y_i -  \bx_i'\bm
-%\alpha_q^* \right) $. From the formula in (\ref{eq:11-31}), we can construct a
-%linearized variance estimator. 
-%
-%Now, we can use the calibration weighting to construct the regression estimator
-%under non-nested two-phase sampling. For  given the design weights $d_{2i} =
-%\pi_{2i}^{-1}$, we find the minimizer of 
-%    $$Q \left( {\bm \omega } \right) = \sum_{i \in A_2} \left(  {\omega_i} - d_{2i}  \right)^2 q_i $$
-%    subject to 
-%    $$ \sum_{i \in A_2} {\omega_i} \bx_i = \widehat{\mathbf{X}}_{\rm c} .$$
-%    The solution is 
-%    $$ \hat{\omega}_i = d_{2i} + \left( \widehat{\mathbf{X}}_{\rm c} - \widehat{\mathbf{X}}_{\rm 2} \right)^{-1} \left( \sum_{i \in A_2} q_i^{-1} \bx_i \bx_i' \right)^{-1} \bx_i q_i^{-1} . $$
-%     Note that 
-%    $$ \sum_{i \in A_2} \hat{\omega}_i y_i = \widehat{Y}_{\rm tp, reg} , $$
-%    where  $\widehat{Y}_{\rm tp, reg}$ is defined in (\ref{eq:11-28}). 
-%    Thus, the algebraic equivalence between the  regression estimator and the calibration weighting estimator is established under non-nested two-phase sampling.
-
 \section{Topic 3: Multi-source Two-Phase Sampling}

### 2024-04-23 final/reports/prelim_paper.tex
Ran several basic simulations. Additional algorithms did not work.

@@
 \sum_{i \in A_2} \frac{x_i y_i}{\pi_{1i} q_i}.$$
-%\footnote{
-%  \textcolor{blue}{Caleb, we do not have to use $\pi_{2i \mid 1}^{(2)}$ in computing
-%  $\hat{\beta}_2$. Please check my book.}
-%  \textcolor{red}{Thanks for the check. I
-%checked it and I must have gotten $\pi_{1i}$ and $\pi_{2i}$ confused.}}
-%\footnote{\textcolor{red}{I was initially thinking of using $\hat \beta_2$ without $q_i$ to be
-%  more consistent with how I defined the two-phase regression estimator;
-%  however, I see your point about using $q_i$ to connect this estimator with the
-%superpopulation model.}}
 
@@
 
-\begin{theorem}[Design Consistency]
-  Under regularity conditions,
+\begin{theorem}[Design Consistency]\label{thm:dc1}
+  Under some regularity conditions,
 
-  $$\hat Y_{DCE} = \hat Y_\ell(\lambda^*, \phi^*) + o_p(N / n_2)$$
+  $$\hat Y_{DCE} = \hat Y_\ell(\lambda^*, \phi^*) + O_p(N / n_2)$$
 
@@
   $$\hat Y_{\ell}(\hat \lambda, \phi^*) = \hat Y_{DCE}(\hat \lambda) + 
-  \left(\sum_{i \in A_1} w_{1i} z_i q_i = \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat
+  \left(\sum_{i \in A_1} w_{1i} z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat
   \lambda) z_i q_i\right)\phi^*$$
@@
 
+   \begin{eqnarray*}
+   \hat Y_{\ell}( \lambda^*, \phi^*) &=&   \hat{Y}_{\pi^*} + 
+  \left(\sum_{i \in A_1} w_{1i} x_i  -  \sum_{i \in A_2} w_{1i} \pi_{2i \mid
+  1}^{-1} x_i  \right)^T \phi_1^* + \left(\sum_{i \in A_1} w_{1i} g_i   -
+\sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^T \phi_2^*  
+  \end{eqnarray*} 
+  and  $$\begin{pmatrix}
+  \phi_1^* \\
+  \phi_2^* 
+  \end{pmatrix}
+  = \left[ \sum_{i \in U} \frac{\pi_{2i \mid 1} }{ g'(d_{2i|1}) q_i} 
+  \begin{pmatrix}
+  \bx_i \bx_i^T &   \bx_i g_i   \\
+  g_i  \bx_i^T   & g_i^2     \end{pmatrix} \right]^{-1}
+  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} \begin{pmatrix}
+    \bx_i \\ g_i 
+  \end{pmatrix}y_i $$
+with $g_i = g( \pi_{2i |1}^{-1}) q_i$.
 \end{theorem}
 
+\textcolor{red}{I have tried to improve the expression of the linearization
+formula. Please check if my algebra above is correct. }
+\textcolor{blue}{This is correct. I have almost the same expression in
+Theorem~\ref{thm:dc2}.}
 \begin{proof}
@@
   = 0$, $\frac{\partial}{\partial \lambda^2} \hat w_{2i|1}(\lambda^*)$ is
-  bounded and $|\hat \lambda - \lambda^*| = o_p(n_2^{-1/2})$, which proves our
+  bounded and $|\hat \lambda - \lambda^*| = O_p(n_2^{-1/2})$, which proves our
   result.
-
 \end{proof}
-
-%\begin{proof}
-%  First, we derive the solution for $\hat Y_{DCE}$, and then derive the
-%  asymptotic variance of this estimator using the linearization technique of
-%  \cite{randles1982asymptotic}.
-%
-%  We want to solve Equation~\ref{eq:primalloss} subject to the constraints in
-%  Equations~\ref{eq:calconst1} and \ref{eq:calconst2}. Using the method of
-%  Lagrange multipliers we have a Lagrangian of 
-%
-%  $$\mathcal{L}(w_2) = -\sum_{i \in A_2} w_{1i} G(w_{2i})q_i + \lambda 
-%  \left(\sum_{i \in A_2} w_{1i} w_{2i} z_i - \sum_{i \in A_1} w_{1i} z_i\right).$$
-%
-%  Our first order conditions yield,
-%
-%  \begin{align*}
-%    0 &= \frac{\partial}{\partial w_{2i}} \mathcal{L}(w_{2}) \\ 
-%      &= -w_{1i} g(w_{2i}) q_i + \lambda w_{1i} z_i.
-%  \end{align*}
-%
-%  This means that $\hat w_{2i} = g^{-1}(\hat \lambda z_i / q_i)$ and since $\hat
-%  \lambda$ is determined from the solution to Equations~\ref{eq:calconst1} and
-%  \ref{eq:calconst2}. Hence, $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} 
-%  \hat w_{2i}(\hat \lambda) y_i$.
-%
-%  To get the variance of $\hat Y_{DCE}$ we use the linearization technique of
-%  \cite{randles1982asymptotic}. First, notice that 
-%
-%  $$\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i =
-%  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i + \left(
-%  \sum_{i \in A_1} w_{1i} z_i -
-%  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) z_i\right)\phi
-%  := \hat Y_\ell(\hat \lambda, \phi).$$
-%
-%  We can find the value of $\phi$ such that $E[\frac{\partial}{\partial
-%  \lambda}\hat Y_\ell(\lambda, \phi)]_{\lambda = \lambda^*}$ where $\lambda^*$
-%  is the probability limit of $\hat \lambda$. Notice, that 
-%  $\lambda^* = (\bf 0, 1)$ because $\hat w_{2i} \to \pi_{2i|1}^{-1}$. This means
-%  that the optimal value of $\phi$ solves,
-%
-%  $$E\left[\frac{\partial}{\partial \lambda} \hat Y_\ell(\lambda, \phi)\right] =
-%  E\left[\sum_{i \in A_2} w_{1i} \frac{\partial}{\partial \lambda} w_{2i|1}(\lambda)
-%    y_i - \phi \sum_{i \in A_2} w_{1i} z_i \frac{\partial}{\partial \lambda}
-%    w_{2i|1}(\lambda) / q_i \right]$$
-%
-%  which means that
-%
-%  $$\phi^* = \left[\sum_{i \in U} \pi_{2i|1} \left(\frac{z_i z_i^T}{
-%    g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in U} \pi_{2i|1}
-%    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
-%
-%  Hence,
-%
-%  \begin{align*}
-%    \hat Y_\ell(\lambda^*, \phi^*) 
-%    &= \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1}
-%    y_i + \left(\sum_{i \in A_1} w_{1i} z_i / q_i -
-%      \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} z_i / q_i\right)\phi^* \\ 
-%    &= \sum_{i \in A_1} w_{1i} \eta_i
-%  \end{align*}
-%
-%  where 
-%
-%  $$\eta_i = z_i \phi^* / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
-%  - z_i \phi^* / q_i\right).$$
-%
-%  To analyze the variance, consider the reverse framework of
-%  \cite{fay1992inferences}, \cite{kim2006replication}, and \cite{shao1999variance}.
-%  In this setup, we assume that $\mathcal{R} = \{\delta_{2i|1}\}_{i \in U}$ is
-%  known at the
-%  finite population level. Then with this we have,
-%
-%  \begin{align*}
-%    V(\hat Y_{\ell}(\lambda^*, \phi^*)) 
-%    &= V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\right) \\
-%    &= V\left(E\left[\sum_{i \in A_1} w_{1i}\eta_i(\phi^*)\mid\mathcal{R}\right]\right) 
-%    +E\left[V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\mid\mathcal{R}\right)\right]\\
-%    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right) + E\left[\sum_{i \in U} 
-%    \sum_{j \in U} \Cov(\delta_{1i}, \delta_{1j}) 
-%    w_{1i} w_{1j} \eta_i(\phi^*) \eta_j(\phi^*)\right]\\
-%  \end{align*}
-%
-%  We can estimate this variance using 
-%
-%  $$\hat V_{DCE} = \sum_{i \in A_1} \sum_{j \in A_1} \left(\frac{\pi_{1ij} -
-%  \pi_{1i}\pi_{1j}}{\pi_{1ij}}
-%    w_{1i} w_{1j} \hat \eta_i \hat \eta_j \right)
-%  $$
-%
-%  for $\hat \eta_i = z_i \hat \phi / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
-%  - z_i \hat \phi / q_i\right)$ where 
-%
-%  $$\hat \phi = \left[\sum_{i \in A_2} w_{1i}\left(\frac{z_i z_i^T}{
-%  g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in A_2} w_{1i}
-%    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
-%
-%  The problem with this estimator is that it is biased as $E[\hat V_{DCE}]$ is
-%  asymptotically equivalent to
-%  $V\left(\sum_{i \in A_1} w_{1i} \eta(\phi^*)\right)$. This yields a bias of 
-%
-%  \begin{align*}
-%    \text{Bias}(\hat V_{DCE}) 
-%    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right)\\ 
-%    &= \sum_{i \in U} \sum_{j \in U} \Delta_{2ij|1} w_{2i|1}
-%    w_{2j|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).
-%  \end{align*}
-%
-%  where $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1} \pi_{2j|1}$. 
-%  However, we can estimate this bias with
-%
-%  $$\widehat{\text{Bias}}(\hat V_{DCE}) = 
-%  \sum_{i \in A_2} \sum_{j \in A_2} w_{1ij} w_{2ij|1} w_{2i|1} w_{2j|1} 
-%  \Delta_{2ij|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).$$
-%
-%\end{proof}
 
@@
 where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
+
+\textcolor{red}{The simulation setup is somewhat strange.
+The first order inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$.
+In this case, your second-phase sampling is essentially equivalent to the simple random sampling.
+}
 This is a two-phase extension of the setup in \cite{kwon2024debiased}. We
@@
 
-\textcolor{blue}{
-( Materials in Section 11.4 can be used here. I have copy-and-pasted the
-textbook materials below. Please modify them.  )
-}
-\textcolor{red}{I will do this shortly.}
-
 \subsection{Background}
@@
 $\hat X_c = W \hat X_1 + (1 - W)\hat X_2$ for some matrix $W$ (see
-\cite{merkouris2004combining} for the optimal choice of $W$). We can then define
+\cite{merkouris2004combining} for the optimal choice of $W$), and can then define
 a regression estimator as
@@
 
+The methodology for the non-nested two phase sample is very similar to the setup
+described as part of Topic 1. Given a strictly convex differentiable function,
+$G: \mathcal{V} \to \mathbb{R}$, the goal is to solve
+
+\begin{equation}\label{eq:nnopt}
+\hat w_2 = \argmin_w \sum_{i \in A_2} G(w_{2i}) q_i \text{ such that } 
+\sum_{i \in A_2} w_{2i} x_i = \hat X_{GLS} \text{ and } 
+\sum_{i \in A_2} w_{2i} g(d_{2i}) q_i = \sum_{i \in U} g(d_{2i}) q_i
+\end{equation}
+
+for $g(x) = G'(x)$ and a known choice of $q_i \in \mathbb{R}$. The difference
+between solving Equation~\ref{eq:nnopt} and Equation~\ref{eq:primal} is that the
+estimator $\hat X_{GLS}$ is estimated using GLS from other sample(s). Before using
+$\hat X_{GLS}$ in the debiased calibration estimator, we need to estimate it
+from the non-nested samples.
+
+We can get multiple estimates of $\hat X$,
+
+$$\hat X_1 = \sum_{i \in A_1} \frac{x_i}{\pi_{1i}} \text{ and }
+\hat X_2 = \sum_{i \in A_2} \frac{x_i}{\pi_{2i}}.$$
+
+Let $V_1$ and $V_2$ be known variance matrices for $\hat X_1$ and $\hat X_2$,
+then the optimal combined GLS estimates is
+
+$$\hat X_c = \frac{V_1^{-1} \hat X_1 + V_2^{-1} \hat X_2}{V_1^{-1} + V_2^{-1}}.$$
+
+We can constuct a non-nested two phase estimator $\hat Y_{NNE}$ for $Y_N$ where
+$\hat Y_{NNE} = \sum_{i \in A_2} \hat w_{2i} y_i$ and $\hat w_{2i}$ solves
+Equation~\ref{eq:nnopt}.
+
 \subsection{Theoretical Results}
+
+\begin{theorem}[Design Consistency]\label{thm:dc2}
+  Under some regularity conditions, $\hat Y_{NNE} = \hat Y_{\ell, NNE}(\lambda^*,
+  \phi^*) + O_p(Nn_2^{-1})$ where
+
+  $$\hat Y_{\ell, NNE}(\lambda^*, \phi^*) = \sum_{i \in A_2} \hat w_{2i}(
+  \lambda^*) + \left(T - \sum_{i \in A_2} \hat w_{2i}(\lambda^*) z_i q_i\right)
+  \phi^*$$
+
+  and 
+  $$\phi^* =
+  \left(\sum_{i \in U} \frac{\pi_{2i} q_i}{g'(d_{2i})} 
+  \begin{bmatrix}
+    x_i^2 / q_i & x_i g(d_{2i}) / q_i \\
+    x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \end{bmatrix}
+  \right)^{-1}
+  \sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})} 
+  \begin{bmatrix} x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
+
+\end{theorem}
+
+\begin{proof}
+  The proof of this result is very similar to the proof the Theorem~\ref{thm:dc1}.
+  The biggest difference is that the total for $X$ is estimated from both samples
+  using $\hat X_c$ instead of $\hat X_{HT}$ from the Phase 1 sample.
+
+  Since $\hat Y_{NNE} = \sum_{i \in A_2} \hat w_{2i}(\hat \lambda)y_i$ where
+  $\hat \lambda$ solves
+
+  \begin{equation}
+    \sum_{i \in A_2} \hat w_{2i}(\lambda) q_i
+    \underbrace{
+    \begin{bmatrix} 
+      x_i / q_i \\ g(d_i)
+  \end{bmatrix}}_{z_i} =
+  \underbrace{\begin{bmatrix} 
+      \hat X_c \\
+      \sum_{i \in U} g(d_i) q_i
+  \end{bmatrix}}_{T}
+  \end{equation}
+
+  we have
+
+  $$\hat Y_{\ell, NNE} (\hat \lambda, \phi) = \sum_{i \in A_2} \hat w_{2i}(\hat
+  \lambda) + \left(T - \sum_{i \in A_2} \hat w_{2i}(\hat \lambda) z_i q_i\right)
+  \phi.$$
+
+  If we choose $\phi^*$ such that $E\left[\frac{\partial}{\partial \lambda} 
+    \hat Y_{\ell, NNE}(\lambda^*, \phi^*)\right] = 0$, then
+
+  $$\phi^* =
+  \begin{bmatrix}
+    \phi^*_1 \\ \phi^*_2
+  \end{bmatrix} = 
+  \left(\sum_{i \in U} \frac{\pi_{2i} q_i}{g'(d_{2i})} 
+  \begin{bmatrix}
+    x_i^2 / q_i & x_i g(d_{2i}) / q_i \\
+    x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \end{bmatrix}
+  \right)^{-1}
+  \sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})} 
+  \begin{bmatrix} x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
+
+  Hence, by a Taylor expansion around $\hat \lambda$,
+
+  $$\hat Y_{NNE}(\hat \lambda) = \hat Y_{\ell, NNE}(\lambda^*, \phi^*) +
+  O_p(Nn_2^{-1}).$$
+
+\end{proof}
+
+\begin{theorem}[Variance Estimation]
+  The variance of $\hat Y_{NNE}$ is
+
+  \begin{align*}
+    \Var(\hat Y_{NNE}(\hat \lambda)
+    &= (\phi_1^*)^T \Var(\hat X_c) \phi_1^* + 
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} (y_i - z_i \phi^* q_i)(y_j - z_j \phi^* q_j)\\
+    &\qquad + (1 - W)\phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij} \frac{x_i}{\pi_{2i}}
+    \frac{(y_j - z_j \phi^*_1 q_j)}{\pi_{2j}}\\
+  \end{align*}
+
+  We can estimate the variance using
+
+  \begin{align*}
+    \hat V_{NNE} 
+    &= (\hat \phi_1)^T \Var(\hat X_c) \hat \phi_1 + 
+    \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
+    (y_i - z_i \hat \phi q_i)(y_j - z_j \hat \phi q_j) \\
+    &\qquad + (1 - W)\hat \phi_1 \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}}
+    \frac{x_i}{\pi_{2i}} \frac{(y_j - z_j \hat \phi_1 q_j)}{\pi_{2j}}
+  \end{align*}
+
+  where
+
+  $$\hat \phi =
+  \begin{bmatrix}
+    \hat \phi_1 \\ \hat \phi_2
+  \end{bmatrix} = 
+  \left(\sum_{i \in A_2} \frac{q_i}{g'(d_{2i})} 
+  \begin{bmatrix}
+    x_i^2 / q_i & x_i g(d_{2i}) / q_i \\
+    x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \end{bmatrix}
+  \right)^{-1}
+  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})} 
+  \begin{bmatrix} x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
+\end{theorem}
+
+\begin{proof}
+  From Theorem~\ref{thm:dc2}, we know that $\hat Y_{NNE}(\hat \lambda) = 
+  \hat Y_{\ell, NNE}(\lambda^*, \phi^*) + O_p(Nn_2^{-1})$. Hence, the variance
+  of $\hat Y_{NNE}(\hat \lambda)$ is
+
+  \begin{align*}
+    \Var(\hat Y_{NNE}(\hat \lambda)) 
+    &= \Var(\hat Y_{\ell, NNE}(\lambda^*, \phi^*) + O_p(Nn_2^{-1})) \\ 
+    &= \Var\left(\sum_{i \in A_2} \hat w_{2i}(\lambda^*) y_i + \left(T - \sum_{i
+    \in A_2} \hat w_{2i}(\lambda^*) z_i q_i\right)\phi^*\right)\\
+    &= (\phi_1^*)^T \Var(\hat X_c) \phi_1^* + 
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} (y_i - z_i \phi^* q_i)(y_j - z_j \phi^* q_j)\\
+    &\qquad+ 2 \Cov\left(\hat X_c \phi_1^*, \sum_{i \in A_2} \frac{(y_i - z_i \phi^*
+    q_i)}{\pi_{2i}}\right) \\
+    &= (\phi_1^*)^T \Var(\hat X_c) \phi_1^* + 
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} (y_i - z_i \phi^* q_i)(y_j - z_j \phi^* q_j)\\
+    &\qquad + (1 - W)\phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij} \frac{x_i}{\pi_{2i}}
+    \frac{(y_j - z_j \phi^*_1 q_j)}{\pi_{2j}}\\
+  \end{align*}
+
+  where the last equality comes from the fact that $\hat X_c = W\hat X_1 + (1 -
+  W) \hat X_2$. To have an unbiased estimator of the variance we can use:
+
+  \begin{align*}
+    \hat V_{NNE} 
+    &= (\hat \phi_1)^T \Var(\hat X_c) \hat \phi_1 + 
+    \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
+    (y_i - z_i \hat \phi q_i)(y_j - z_j \hat \phi q_j) \\
+    &\qquad + (1 - W)\hat \phi_1 \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}}
+    \frac{x_i}{\pi_{2i}} \frac{(y_j - z_j \hat \phi_1 q_j)}{\pi_{2j}}
+  \end{align*}
+
+  where
+
+  $$\hat \phi =
+  \begin{bmatrix}
+    \hat \phi_1 \\ \hat \phi_2
+  \end{bmatrix} = 
+  \left(\sum_{i \in A_2} \frac{q_i}{g'(d_{2i})} 
+  \begin{bmatrix}
+    x_i^2 / q_i & x_i g(d_{2i}) / q_i \\
+    x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \end{bmatrix}
+  \right)^{-1}
+  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})} 
+  \begin{bmatrix} x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
+
+\end{proof}
 

### 2024-04-23 final/reports/prelim_paper_ov.tex
Ran several basic simulations. Additional algorithms did not work.

--- 2024-04-11 final/reports/prelim_paper_ov.tex
@@
 
-\addbibresource{references.bib}
+\usepackage{amsmath, amsthm, mathtools,commath}
+\usepackage{graphics, color}
+\usepackage{latexsym}
+\usepackage{amssymb, amsfonts, bm}
+\usepackage{mathrsfs}
 
-\hypersetup{colorlinks = true, urlcolor = black}
+\usepackage{graphicx}
+\usepackage{epsfig}
+\usepackage{makeidx}
+\usepackage{fullpage}
+\usepackage{booktabs, arydshln}
+\usepackage{comment} 
+%\makeindex
 
-\headheight     15pt
-\topmargin      -1.5cm   % read Lamport p.163
-\oddsidemargin  -0.04cm  % read Lamport p.163
-\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
-\textwidth      16.59cm
-\textheight     23.94cm
-\parskip         7.2pt   % sets spacing between paragraphs
-\parindent         0pt   % sets leading space for paragraphs
-\pagestyle{empty}        % Uncomment if don't want page numbers
-\pagestyle{fancyplain}
+\hbadness=10000 \tolerance=10000 \hyphenation{en-vi-ron-ment
+in-ven-tory e-num-er-ate char-ac-ter-is-tic}
 
+\usepackage[round]{natbib}
+%\bibliographystyle{apalike2}
+% \bibliographystyle{jmr}
+
+\newcommand{\biblist}{\begin{list}{}
+{\listparindent 0.0cm \leftmargin 0.50cm \itemindent -0.50 cm
+\labelwidth 0 cm \labelsep 0.50 cm
+\usecounter{list}}\clubpenalty4000\widowpenalty4000}
+\newcommand{\ebiblist}{\end{list}}
+
+\newcounter{list}
+
+%\usepackage{setspace}
+
+%\usepackage{hangpar}
+\newcommand{\lbl}[1]{\label{#1}{\ensuremath{^{\fbox{\tiny\upshape#1}}}}}
+% remove % from next line for final copy
+\renewcommand{\lbl}[1]{\label{#1}}
+
+\newtheorem{theorem}{Theorem}
+\newtheorem{corollary}{Corollary}
+\newtheorem{definition}{Definition}
+\newtheorem{example}{Example}
+\newtheorem{remark}{Remark}
+\newtheorem{result}{Result}
+
+\newtheorem{lemma}{Lemma}
+
+\DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\argmax}{arg\,max}
 \newcommand{\MAP}{{\text{MAP}}}
-\newcommand{\argmax}{{\text{argmax}}}
-\newcommand{\argmin}{{\text{argmin}}}
 \newcommand{\Cov}{{\text{Cov}}}
@@
 
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-
 \begin{document}
 
-\title{Debiased calibration for generalized two-Phase sampling}
+\title{Debiased Calibration for Generalized Two-Phase Sampling}
 \author{Caleb Leedy}
@@
 
+\baselineskip .3in
+
 \section{Introduction}
 
-\begin{itemize}
-\item Combining information from several sources is an important practical problem. 
-\item In many cases, we do not have direct access to the other sources.
-We can only obtain summary statistics (and their standard errors) for the external data sources.
-We wish to incorporate the information from external  sources in our in-house data effectively using calibration weighting.
-\item We formulate the problem as a generalized two-phase sampling where the first phase sample can be obtained from multiple sources.
-The second-phase sample is our in-house data in which we want to construct the calibration weights.
-\item To achieve the goal, we first consider  the  classical two-phase sampling setup where the second-phase sample is a subset of the first-phase sample.
-After that, we extend the setup to more general cases such as non-nested two-phase sampling or multiple independent surveys with some common measurements.
-\item (To simplify the presentation, I wonder whether we can just use SRS in the first-phase sampling. ) 
-\end{itemize}
+Combining information from several sources is an important practical problem. (CITEME)
+We want to incorporate information from external data sources to reduce the bias
+in our estimates or improve the estimator's efficiency. For many problems, the
+additional information consists of summary statistics with standard errors. The
+goal of this project is to incorporate external information with existing data 
+to create more efficient estimators using calibration weighting.
+
+To model this scenaro, we formulate the problem as a generalized two-phase
+sample where the first phase sample consists of data from multiple sources. The
+second phase sample contains our existing data. To motivate this setup, we
+consider the following approach: first, we consider the classical two-phase
+sampling setup where the second phase sample is a subset of the first phase
+sample; then, we extend this setup to consider non-nested two-phase samples;
+and finally, we consider the more general approach of having multiple sources.
+
+%\begin{itemize}
+%\item To achieve the goal, we first consider  the  classical two-phase sampling
+%setup where the second-phase sample is a subset of the first-phase sample.
+%  After that, we extend the setup to more general cases such as non-nested
+%  two-phase sampling or multiple independent surveys with some common
+%  measurements. 
+%\item The proposed method can be called two-step calibration. In the
+%first-step, the best linear unbiased estimators of the auxiliary variable
+%totals are computed. In the second-step, the final calibration weights are
+%constructed to match (benchmark) with the best estimators computed from Step 1. 
+%\end{itemize}
+
 \section{Topic 1: Classical Two-Phase Sampling}
 
-\subsection{Background and Introduction}
+\subsection{Background}
 
 Consider a finite population of size $N$ containing elements $(X_i, Y_i)$ where
-an initial (Phase 1) sample of size $n$ is selected and $X_i$ is observed. Then
-from the Phase 1 sample of elements, a (Phase 2) sample of size $r < n$ is
+an initial (Phase 1) sample of size $n_1$ is selected and $X_i$ is observed. Then
+from the Phase 1 sample of elements, a (Phase 2) sample of size $n_2 < n_1$ is
 selected and $Y_i$ is observed. This is two-phase sampling (See 
 \cite{fuller2009sampling}, \cite{kim2024statistics} for general references.) The
-goal of two-phase sampling is to construct an estimator of $Y$ not
-only using the observed information from the Phase 2 sample but also
-incorporating the extra auxilary information of $X$ from the Phase 1 sample, and
-the challenge is doing this efficiently.
+goal of two-phase sampling is to construct an estimator of $\bar Y_N$ 
+that uses both the observed information in the Phase 2 sample and also the extra
+auxiliary information from $X$ in the Phase 1 sample.
+The challenge is doing this efficiently.
 
-An easy-to-implement unbiased estimator in the spirit of a Horvitz-Thompson
+An easy-to-implement unbiased estimator in the spirit of a Horvitz-Thompson (HT)
 estimator (\cite{horvitz1952generalization}, \cite{narain1951sampling}) is the
@@
 denoted as $A_2$,
+%\begin{align*}
+%  \pi_i &= \sum_{A_2: i \in A_2} \Pr(A_2) \\ 
+%        &= \sum_{A_1: A_2 \subseteq A_1} \sum_{A_2: i \in A_2} \Pr(A_2 \mid
+%        A_1) \Pr(A_1) \\
+%        &= \sum_{A_1: i \in A_1} \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1) \Pr(A_1).
+%\end{align*}
+%\textcolor{red}{(I am not sure whether this part is necessary.) }
+if we define $\pi_{2i | 1} = \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1)$ and
+$\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ then,
 
-\begin{align*}
-  \pi_i &= \sum_{A_2: i \in A_2} \Pr(A_2) \\ 
-        &= \sum_{A_1: A_2 \subseteq A_1} \sum_{A_2: i \in A_2} \Pr(A_2 \mid
-        A_1) \Pr(A_1) \\
-        &= \sum_{A_1: i \in A_1} \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1) \Pr(A_1).
-\end{align*}
+$$ \pi_i^{(2)}(A_1) = \pi_{2i | 1} \pi_{1i}.$$
 
-If we define $\pi_{2i | 1}^{(2)} = \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1)$ and
-$\pi_i^{(1)} = \sum_{A_1: i \in A_1} \Pr(A_1)$ then,
-
-$$ \pi_i = \pi_{2i | 1}^{(2)} \pi_i^{(1)}.$$
-\textcolor{red}{(note: there is an abuse of notation. You do not need to introduce superscript $(2)$ on $\pi_{2i \mid 1}$.) }
 This means that we can define the $\pi^*$-estimator as the following design
@@
 
-$$ \hat Y_{\pi^*} = \sum_{i \in A_2} \frac{y_i}{\pi_{2i | 1}^{(2)} \pi_i^{(1)}}.$$
+$$ \hat Y_{\pi^*} = \sum_{i \in A_2} \frac{y_i}{\pi_{2i | 1} \pi_{1i}}.$$
 
-\textcolor{red}{Dr. Kim: Should I provide the proof for design unbiasedness of
-$\hat Y_{\pi^*}$? } \textcolor{blue}{Not necessary. You may cite a book or a paper  for this claim. }
-
-While unbiased, the $\pi^*$-estimator unfortunately does not account for the
-additional information contained in the auxiliary Phase 1 variable $X$. The
-two-phase regression estimator $\hat Y_{reg, tp}$ does incorporate \textcolor{blue}{$\hat{X}_1$ obtained from the phase one sample.
-That is, we can leverage the external information $\hat{X}_1$ to improve the $\pi^*$-estimator in the second-phase sample.
-}
-The two-phase 
-regression estimator has the form,
+While unbiased (see \cite{kim2024statistics}), the $\pi^*$-estimator
+does not account for the additional information contained in the auxiliary Phase
+1 variable $X$. The two-phase regression estimator $\hat Y_{reg, tp}$ does incorporate 
+information for $X$ by using the estimate $\hat X_1$ from the Phase 1 sample.
+This is how we can leverage the external information $\hat X_1$ to improve the
+initial $\pi^*$-estimator in the second phase sample.
+The two-phase regression estimator has the form,
 
 $$ \hat Y_{reg, tp} 
-= \sum_{i \in A_1} \frac{1}{\pi_i^{(1)}} x_i \hat \beta_2+ \sum_{i \in A_2}
-\frac{1}{\pi_i^{(1)}\pi_{2i|1}^{(2)}} (y_i - x_i \hat \beta_2)$$
-where $$\hat \beta_2 = \left(\sum_{i \in A_2} 
-  \frac{x_i x_i'}{\pi_{2i|1}^{(2)}}\right)^{-1} 
-  \sum_{i \in A_2} \frac{x_i y_i}{\pi_i^{(1)}\pi_{2i|1}^{(2)}}. 
-  \footnote{Caleb, we do not have to use $\pi_{2i \mid 1}^{(2)}$ in computing $\hat{\beta}_2$. Please check my book. }
+= \sum_{i \in A_1} \frac{1}{\pi_{1i}} x_i \hat \beta_q+ \sum_{i \in A_2}
+\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_q)$$
+
+where for $q_i = q(x_i)$ and is a function of $x_i$,
 $$
+\hat \beta_q = \left(\sum_{i \in A_2} 
+  \frac{x_i x_i'}{\pi_{1i} q_i}\right)^{-1} 
+\sum_{i \in A_2} \frac{x_i y_i}{\pi_{1i} q_i}.$$ 
+%\footnote{
+%  \textcolor{blue}{Caleb, we do not have to use $\pi_{2i \mid 1}^{(2)}$ in computing
+%  $\hat{\beta}_2$. Please check my book.}
+%  \textcolor{red}{Thanks for the check. I
+%checked it and I must have gotten $\pi_{1i}$ and $\pi_{2i}$ confused.}}
+%\footnote{\textcolor{red}{I was initially thinking of using $\hat \beta_2$ without $q_i$ to be
+%  more consistent with how I defined the two-phase regression estimator;
+%  however, I see your point about using $q_i$ to connect this estimator with the
+%superpopulation model.}}
 
-\textcolor{red}{I suggest that we use 
-$$\hat \beta_2 = \left(\sum_{i \in A_2} 
-  \frac{x_i x_i'}{\pi_{i}^{(1)} q_i }\right)^{-1} 
-  \sum_{i \in A_2} \frac{x_i y_i}{\pi_i^{(1)} q_i }
-$$ 
-where $q_i=q(\bx_i)$ is a function of $\bx_i$. This will have some connection with model-optimal regression estimator (under superpopulation model).  
-}
+The regression estimator is the minimum variance design consistent linear
+estimator which is easily shown to be the case because $\hat Y_{reg, tp} =
+\sum_{i \in A_2} \hat w_{2i} y_i / \pi_{1i}$ where
 
-  The regression
-estimator is the minumum variance design consistent linear estimator which is
-easily shown to be the case because $\hat Y_{reg, tp} = \sum_{i \in A_2} \hat
-w_{2i} y_i / \pi_i^{(1)}$ where
-
-$$\hat w_i = \argmin_{w} \sum_{i \in A_2} (w_i - \pi_{2i|1}^{-1})^2 \text{ such
-that } \sum_{i \in A_2} w_i x_i / \pi_i^{(1)} = \sum_{i \in A_1} x_i /
-\pi_i^{(1)}.$$
-
-\textcolor{red}{Using $q_i$, we can construct 
 $$\hat w_{2i} = \argmin_{w} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i \text{ such
-that } \sum_{i \in A_2} w_{2i}  x_i / \pi_i^{(1)} = \sum_{i \in A_1} x_i /
-\pi_i^{(1)}$$
-as a way to implement the two-phase regression estimator indirectly using calibration weighting. 
-}
+that } \sum_{i \in A_2} w_{2i} x_i / \pi_{1i} = \sum_{i \in A_1} x_i / \pi_{1i}.$$
 
 This means that $\hat Y_{reg, tp}$ is also a calibration estimator. The idea
-that regression estimation is a form of calibration was extended by 
-\cite{deville1992calibration} to consider loss functions other than just squared
-loss. They generalized the loss function to minimize $\sum_i G(w_i, d_i)q_i$ for
-weights $w_i$ and design-weights $d_i$ where $G()$ is non-negative, strictly
-convex function with respect to $w$, defined on an interval containing $d_i$,
-with $g(w_i, d_i) = \partial G / \partial w$ continuous.
-\footnote{The \cite{deville1992calibration} paper considers regression estimators
-for a single phase setup, which we apply to our two-phase example. } This
+that regression estimation is a form of calibration was noted by
+\cite{deville1992calibration} and extended by them to consider loss functions
+other than just squared loss. Their generalized loss function minimizes
+$\sum_i G(w_i, d_i)q_i$ for weights $w_i$ and design-weights $d_i$ where
+$G(\cdot)$ is a non-negative, strictly convex function with respect to $w$,
+defined on an interval containing $d_i$, with $g(w_i, d_i) = \partial G /
+\partial w$ continuous.\footnote{The \cite{deville1992calibration} paper
+considers regression estimators for a single phase setup, which we apply to our
+two-phase example.} This
 generalization includes empirical likelihood estimation, and maximum entropy
-estimation amoung others. The variance estimation is based on a linearization
+estimation among others. The variance estimation is based on a linearization
 that shows that minimizing the generalized loss function subject to the
@@
 
-While this generalization is useful to an analyst who may want different
-properties of their estimator from maximum entropy estimation rather than the
-minimal squared loss, unless $\pi_{2i|1}^{-1} \in C(X)$, the estimator is not
-design consistent. \footnote{ \textcolor{red}{I do not understand this part.} } 
-Furthermore, at a conceptual level, the regression estimator
-has a nice feature that its two terms can be thought about as minimizing
+Furthermore, the regression estimator
+has a nice feature that its two terms can be thought about as minimizing the
 variance and bias correction,
@@
 $$ \hat Y_{reg, tp} 
-= \underbrace{\sum_{i \in A_1} \frac{x_i \hat \beta_2}{\pi_i^{(1)}}}_{Minimizing
-the variance} + \underbrace{\sum_{i \in A_2}
-\frac{1}{\pi_i^{(1)}\pi_{2i|1}^{(2)}} (y_i - x_i \hat \beta_2)}_{Bias
-correction}.$$
+= \underbrace{\sum_{i \in A_1} \frac{x_i \hat \beta_q}{\pi_{1i}}}_{
+  \text{ Minimizing the variance}} + \underbrace{\sum_{i \in A_2}
+\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_q)}_{
+\text{Bias correction}}.$$
 
 The \cite{deville1992calibration} method incorporates the design weights into
-the loss function, which is the part minimizing the variance. Instead, there is
-a desire to get design consistency from the calibration term. In 
+the loss function, which is the part minimizing the variance. We would rather
+separate have bias calibration separate from the minimizing the variance so that
+we can control each in isolation. In
 \cite{kwon2024debiased}, the authors show that for a generalized entropy
@@
 
-Unfortunately, the method of \cite{kwon2024debiased} requires known calibration
-levels (no uncertainty) for the finite population. It does not handle the
+The method of \cite{kwon2024debiased} requires known finite population 
+calibration levels. It does not handle the
 two-phase setup where we need to estimate the finite population total of $x$
-from the Phase 1 sample. Hence, we extend this method to have a valid variance 
-estimator when including estimated Phase 1 weights.
+from the Phase 1 sample. In the rest of the section, we extend this method to 
+two phase sampling so that we have a valid 
+estimator when including estimated Phase 1 weights with appropriate variance
+estimation.
 
-\subsection*{Methodology}
+\subsection{Methodology}
 
-% TODO: Add detailed method for the setup of the debiased calibration method and
-% explain how we estimate the weights in two-phase sampling.
-% DUE: Thursday April 11
+We follow the approach of \cite{kwon2024debiased} for the debiased calibration
+method. We consider maximizing the generalized entropy \cite{gneiting2007strictly},
 
-\subsection*{Theoretical Results}
+\begin{equation}\label{eq:primalloss}
+  H(w) = - \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i
+\end{equation}
 
-% TODO: Provide asymptotic results showing the variance estimation of the
-% estimated debiased calibration method
-% DUE: Thursday April 11 (I have already showed the proof but I might finish the
-% simulation study before writing this up.)
+where $G: \mathcal{V} \to \R$ is strictly convex, differentiable function
+subject to the constraints:
 
-\subsection*{Simulation Studies}
+\begin{equation}\label{eq:calconst1}
+  \sum_{i \in A_2} \frac{x_i w_{2i}q_i}{\pi_{1i}} = 
+\sum_{i \in A_1} \frac{x_iq_i}{\pi_{1i}}
+\end{equation}
 
-% TODO: Conduct a simulation study highlighting the proposed method
-% DUE: Friday April 12
+and
 
-\section*{Topic 2: Non-nested Two-Phase Sampling}
+\begin{equation}\label{eq:calconst2}
+  \sum_{i \in A_2} \frac{g(\pi_{2i|1}^{-1})w_{2i}q_i}{\pi_{1i}} = 
+  \sum_{i \in A_1} \frac{g(\pi_{2i|1}^{-1})q_i}{\pi_{1i}}.
+\end{equation}
 
-\section*{Topic 3: Multi-source Two-Phase Sampling}
+The first constraint is the existing calibration constraint and the second
+ensures that design consistency is achieved. Here, 
+$g(w) = \partial G / \partial w$. 
+The original method of \cite{kwon2024debiased} only considered having finite
+population quantities on the right hand side of \ref{eq:calconst1}.
+Let $z_i = (x_i / q_i, g(\pi_{2i|1}^{-1}))$. Letting $w_{1i} = \pi_{1i}^{-1}$,
+the the goal is to solve,
 
-\printbibliography
+\begin{equation}\label{eq:primal}
+  \argmin_{w_{2|1}} \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i 
+  \text{ such that}
+  \sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i = \sum_{i \in A_1} w_{1i} z_i q_i.
+\end{equation}
+
+Let $\hat w_{2i|1}$ be the solution to Equation~\ref{eq:primal}, then the
+estimate of $Y_N$ is $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i|1} y_i$.
+
+\subsection{Theoretical Results}
+
+\begin{theorem}[Design Consistency]
+  Under some regularity conditions,
+
+  $$\hat Y_{DCE} = \hat Y_\ell(\lambda^*, \phi^*) + O_p(N / n_2)$$
+
+  where
+\textcolor{red}{
+   \begin{eqnarray*}
+   \hat Y_{\ell}( \lambda^*, \phi^*) &=&   \hat{Y}_{\pi^*} + 
+  \left(\sum_{i \in A_1} w_{1i} x_i  -  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} x_i  \right)^T \phi_1^* + \left(\sum_{i \in A_1} w_{1i} g_i   -  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^T \phi_2^*
+  \end{eqnarray*} 
+  and  $$\begin{pmatrix}
+  \phi_1^* \\
+  \phi_2^* 
+  \end{pmatrix}
+  = \left[ \sum_{i \in U} \frac{\pi_{2i \mid 1} }{ g'(d_{2i|1}) q_i} 
+  \begin{pmatrix}
+  \bx_i \bx_i^T &   \bx_i g_i   \\
+  g_i  \bx_i^T   & g_i^2     \end{pmatrix} \right]^{-1}
+  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} \begin{pmatrix}
+    \bx_i \\ g_i 
+  \end{pmatrix}y_i $$
+with $g_i = g( \pi_{2i |1}^{-1}) q_i$. }
+\end{theorem}
+
+\textcolor{red}{I have tried to improve the expression of the linearization formula. Please check if my algebra above is correct. }
+\begin{proof}
+  In this proof, we derive the solution to Equation~\ref{eq:primal} and show
+  that it is asymptotically equivalent to a regression estimator. Using the
+  method of Lagrange multipliers, to solve Equation~\ref{eq:primal} we need to
+  minimize,
+
+  $$\mathcal{L}(w_{2|1}) = \sum_{i \in A_2} w_{1i} G(w_{2i|1}) q_i + 
+  \lambda \left(\sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i - \sum_{i \in A_1} w_{1i}
+  z_i q_i\right).$$
+
+  The first order conditions show that
+
+  $$\frac{\partial \mathcal{L}}{\partial w_{2i|1}}: g(w_{2i|1}) w_{1i}q_i +
+    \lambda w_{1i} z_i q_i = 0.$$
+
+  Hence, $\hat w_{2i}(\lambda) = g^{-1}(\lambda z_i)$ and $\hat \lambda$ is
+  determined by the calibration condition. When the sample size gets large, we
+  have $\hat w_{2i|1}(\hat \lambda) \to d_{2i|1}$ which means that $\hat \lambda
+  \to \lambda^*$ where $\lambda^* = (\bf 0, 1)$. Then using the linearization
+  technique of \cite{randles1982asymptotic}, we can construct a regression
+  estimator,
+
+  $$\hat Y_\ell(\hat \lambda, \phi)  = \hat Y_{DCE}(\hat \lambda) + 
+  \left(\sum_{i \in A_1} w_{1i} z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat
+  \lambda) z_i q_i\right)\phi.$$
+
+  Notice that $\hat Y_\ell(\hat \lambda, \phi) = \hat Y_{DCE}(\hat \lambda)$ for
+  all $\phi \in \mathbb{R}$. We choose $\phi^*$ such that
+
+  $$E\left[\frac{\partial}{\partial \lambda} \hat Y_\ell(\lambda^*, \phi^*)\right]=0.$$
+
+  Using the fact that $g^{-1}(\lambda^* z_i) = g^{-1}(g(d_{2i|1})) = d_{2i|1}$
+  and $(g^{-1})'(x) = 1 / g'(g^{-1}(x))$, we have
+
+  \begin{align*}
+    \phi^*
+    &= E\left[\sum_{i \in A_2} \frac{w_{1i}z_i z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
+    E\left[\sum_{i \in A_2} \frac{w_{1i}z_i y_i}{g'(d_{2i|1})}\right]\\
+    &= \left[\sum_{i \in U} \frac{\pi_{2i|1}z_i z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
+    \left[\sum_{i \in U} \frac{\pi_{2i|1} z_i y_i}{g'(d_{2i|1})}\right]\\
+  \end{align*}
+
+  Thus, the linearization estimator is
+
+  $$\hat Y_\ell(\lambda^*, \phi^*) = \sum_{i \in A_1} w_{1i} q_i z_i \phi^* +
+  \sum_{i \in A_2} w_{1i} d_{2i|1} (y_i - q_i z_i \phi^*).$$
+
+  By construction using a Taylor expansion yields,
+
+  \begin{align*}
+    \hat Y_{DCE}(\hat \lambda) 
+    &= \hat Y_\ell(\lambda^*, \phi^*) + 
+    E\left[\frac{\partial}{\partial \lambda}\hat Y_\ell(\lambda^*,
+    \phi^*)\right](\hat \lambda - \lambda^*) + \frac{1}{2}
+    E\left[\frac{\partial}{\partial \lambda^2} \hat Y_{DCE}(\lambda^*)\right] (\hat
+    \lambda - \lambda^*)^2\\
+    &= \hat Y_\ell(\lambda^*, \phi^*) + O(N)o_p(n_2^{-1}).
+  \end{align*}
+
+  The final equality comes from the fact that 
+  $E\left[\frac{\partial}{\partial \lambda}\hat Y_\ell(\lambda^*, \phi^*)\right]
+  = 0$, $\frac{\partial}{\partial \lambda^2} \hat w_{2i|1}(\lambda^*)$ is
+  bounded and $|\hat \lambda - \lambda^*| = o_p(n_2^{-1/2})$, which proves our
+  result.
+\footnote{$\hat{\lambda}$ should satisfy $\hat{\lambda} - \lambda^* = O_p (n_2^{-1/2})$. }
+\end{proof}
+
+%\begin{proof}
+%  First, we derive the solution for $\hat Y_{DCE}$, and then derive the
+%  asymptotic variance of this estimator using the linearization technique of
+%  \cite{randles1982asymptotic}.
+%
+%  We want to solve Equation~\ref{eq:primalloss} subject to the constraints in
+%  Equations~\ref{eq:calconst1} and \ref{eq:calconst2}. Using the method of
+%  Lagrange multipliers we have a Lagrangian of 
+%
+%  $$\mathcal{L}(w_2) = -\sum_{i \in A_2} w_{1i} G(w_{2i})q_i + \lambda 
+%  \left(\sum_{i \in A_2} w_{1i} w_{2i} z_i - \sum_{i \in A_1} w_{1i} z_i\right).$$
+%
+%  Our first order conditions yield,
+%
+%  \begin{align*}
+%    0 &= \frac{\partial}{\partial w_{2i}} \mathcal{L}(w_{2}) \\ 
+%      &= -w_{1i} g(w_{2i}) q_i + \lambda w_{1i} z_i.
+%  \end{align*}
+%
+%  This means that $\hat w_{2i} = g^{-1}(\hat \lambda z_i / q_i)$ and since $\hat
+%  \lambda$ is determined from the solution to Equations~\ref{eq:calconst1} and
+%  \ref{eq:calconst2}. Hence, $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} 
+%  \hat w_{2i}(\hat \lambda) y_i$.
+%
+%  To get the variance of $\hat Y_{DCE}$ we use the linearization technique of
+%  \cite{randles1982asymptotic}. First, notice that 
+%
+%  $$\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i =
+%  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i + \left(
+%  \sum_{i \in A_1} w_{1i} z_i -
+%  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) z_i\right)\phi
+%  := \hat Y_\ell(\hat \lambda, \phi).$$
+%
+%  We can find the value of $\phi$ such that $E[\frac{\partial}{\partial
+%  \lambda}\hat Y_\ell(\lambda, \phi)]_{\lambda = \lambda^*}$ where $\lambda^*$
+%  is the probability limit of $\hat \lambda$. Notice, that 
+%  $\lambda^* = (\bf 0, 1)$ because $\hat w_{2i} \to \pi_{2i|1}^{-1}$. This means
+%  that the optimal value of $\phi$ solves,
+%
+%  $$E\left[\frac{\partial}{\partial \lambda} \hat Y_\ell(\lambda, \phi)\right] =
+%  E\left[\sum_{i \in A_2} w_{1i} \frac{\partial}{\partial \lambda} w_{2i|1}(\lambda)
+%    y_i - \phi \sum_{i \in A_2} w_{1i} z_i \frac{\partial}{\partial \lambda}
+%    w_{2i|1}(\lambda) / q_i \right]$$
+%
+%  which means that
+%
+%  $$\phi^* = \left[\sum_{i \in U} \pi_{2i|1} \left(\frac{z_i z_i^T}{
+%    g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in U} \pi_{2i|1}
+%    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
+%
+%  Hence,
+%
+%  \begin{align*}
+%    \hat Y_\ell(\lambda^*, \phi^*) 
+%    &= \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1}
+%    y_i + \left(\sum_{i \in A_1} w_{1i} z_i / q_i -
+%      \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} z_i / q_i\right)\phi^* \\ 
+%    &= \sum_{i \in A_1} w_{1i} \eta_i
+%  \end{align*}
+%
+%  where 
+%
+%  $$\eta_i = z_i \phi^* / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
+%  - z_i \phi^* / q_i\right).$$
+%
+%  To analyze the variance, consider the reverse framework of
+%  \cite{fay1992inferences}, \cite{kim2006replication}, and \cite{shao1999variance}.
+%  In this setup, we assume that $\mathcal{R} = \{\delta_{2i|1}\}_{i \in U}$ is
+%  known at the
+%  finite population level. Then with this we have,
+%
+%  \begin{align*}
+%    V(\hat Y_{\ell}(\lambda^*, \phi^*)) 
+%    &= V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\right) \\
+%    &= V\left(E\left[\sum_{i \in A_1} w_{1i}\eta_i(\phi^*)\mid\mathcal{R}\right]\right) 
+%    +E\left[V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\mid\mathcal{R}\right)\right]\\
+%    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right) + E\left[\sum_{i \in U} 
+%    \sum_{j \in U} \Cov(\delta_{1i}, \delta_{1j}) 
+%    w_{1i} w_{1j} \eta_i(\phi^*) \eta_j(\phi^*)\right]\\
+%  \end{align*}
+%
+%  We can estimate this variance using 
+%
+%  $$\hat V_{DCE} = \sum_{i \in A_1} \sum_{j \in A_1} \left(\frac{\pi_{1ij} -
+%  \pi_{1i}\pi_{1j}}{\pi_{1ij}}
+%    w_{1i} w_{1j} \hat \eta_i \hat \eta_j \right)
+%  $$
+%
+%  for $\hat \eta_i = z_i \hat \phi / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
+%  - z_i \hat \phi / q_i\right)$ where 
+%
+%  $$\hat \phi = \left[\sum_{i \in A_2} w_{1i}\left(\frac{z_i z_i^T}{
+%  g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in A_2} w_{1i}
+%    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
+%
+%  The problem with this estimator is that it is biased as $E[\hat V_{DCE}]$ is
+%  asymptotically equivalent to
+%  $V\left(\sum_{i \in A_1} w_{1i} \eta(\phi^*)\right)$. This yields a bias of 
+%
+%  \begin{align*}
+%    \text{Bias}(\hat V_{DCE}) 
+%    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right)\\ 
+%    &= \sum_{i \in U} \sum_{j \in U} \Delta_{2ij|1} w_{2i|1}
+%    w_{2j|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).
+%  \end{align*}
+%
+%  where $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1} \pi_{2j|1}$. 
+%  However, we can estimate this bias with
+%
+%  $$\widehat{\text{Bias}}(\hat V_{DCE}) = 
+%  \sum_{i \in A_2} \sum_{j \in A_2} w_{1ij} w_{2ij|1} w_{2i|1} w_{2j|1} 
+%  \Delta_{2ij|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).$$
+%
+%\end{proof}
+
+\subsection{Simulation Studies}
+
+We run a simulation testing the proposed method. In this approach we have the
+following simulation setup:
+
+$$
+\begin{aligned}
+X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
+X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
+X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
+X_{4i} &\stackrel{ind}{\sim} Unif(0.1, 0.9) \\
+\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
+Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
+\pi_{1i} &= \Phi_3(-x_{3i} - 2) \\
+\pi_{2i|1} &= x_{4i}.
+\end{aligned}
+$$
+
+where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
+
+\textcolor{red}{The simulation setup is somewhat strange.
+The first order inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$.
+In this case, your second-phase sampling is essentially equivalent to the simple random sampling.
+}
+This is a two-phase extension of the setup in \cite{kwon2024debiased}. We
+consider a finite population of size $N = 10,000$ with both the Phase 1 and
+Phase 2 sampling occuring under Poisson (Bernoulli) sampling. This yields a
+Phase 1 sample
+size of $E[n_1] \approx 1100$ and a Phase 2 sample size of
+$E[n_2] \approx 550$. In the Phase 1 sample, we observe 
+$(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. This
+simulation does not deal with model misspecification, and we compare the
+proposed method for the parameter $\bar Y_N$ with four approaches:
+
+\begin{itemize}
+  \item[1.] $\pi^*$-estimator: $\hat Y_{\pi^*} = N^{-1} \sum_{i \in A_2}
+    \frac{y_i}{\pi_{1i} \pi_{2i|1}},$
+  \item[2.] Two Phase Regression estimator (TP-Reg): 
+    $\hat Y_{reg} = \sum_{i \in A_1} \frac{\bf x_i' \hat \beta}{\pi_{1i}} + 
+    \sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}}(y_i - \bf x_i' \hat \beta)$ 
+    where $\hat \beta = 
+    \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$
+    and $\bf x_i = (x_{1i}, x_{2i})^T$,
+    \textcolor{red}{Dr. Kim, should I modify the simulation so that the
+    regression estimator also includes $g(\pi_{2i|1}^{-1})$ as a covariate?}
+  \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This is
+    the method from \cite{kwon2024debiased} with the true population level
+    constraints, and 
+  \item[4.] Debiased Calibration with Estimated Population Constraints (DC-Est):
+    This is the proposed method with the Phase 1 sample being used to estimate
+    the population level constraints.
+\end{itemize}
+
+In addition to estimating the mean parameter $\bar Y_N$, we also construct
+variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$. For each
+approach we have the following variance estimate\footnote{These variance
+estimates use the fact that we have Poisson sampling for both phases in the
+simulation.}:
+
+\begin{figure}[ht!]
+  \centering
+  \begin{tabular}{lcc}
+    \toprule
+    Estimator & Variance & Notes \\
+    \midrule
+    $\pi^*$ & $N^{-2} \sum_{i \in A_2} \left(\pi_{2i|1}^{-2} - \pi_{2i|1}^{-1}\right)
+    y_i^2 $ & \\
+    TP-Reg  & 
+      {\scriptsize $N^{-2}\left(\sum\limits_{i \in A_1} \left(\pi_{1i}^{-2} -
+          \pi_{1i}^{-1}\right) 
+      \eta_i^2 
+      + \sum\limits_{i \in A_2} \frac{1}{\pi_{1i} \pi_{2i|1}}(\pi_{2i|1}^{-1}
+  - 1) (y_i - \bf x_i' \hat \beta)^2 \right)$}
+    & 
+      {\scriptsize$\eta_i = \bf x_i \hat \beta + \frac{\delta_{2i}}{\pi_{2i|1}}(y_i - 
+      \bf x_i \hat \beta)$} \\
+      %$\hat \beta = \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$ \\
+      DC-Pop  & $(Y - \bf z^T \hat \gamma)^T \Pi (Y - \bf z^T \hat \gamma)$ & 
+      {\scriptsize$\Pi
+      = \text{diag}(1 - (\pi_{1}\pi_{2|1})^{-1}) \cdot \frac{\hat w^2}{N^2}$} \\
+      DC-Est  & $(Y - \bf z^T \hat \gamma)^T \Pi (Y - \bf z^T \hat \gamma) +
+      \hat \gamma_{[1:3]}^T V(\bf x) \hat \gamma_{[1:3]}/N^2$ & \\
+    \bottomrule
+  \end{tabular}
+  \caption{This table gives the formulas for each variance estimator used in
+  this simulation.}
+  \label{tab:varforms}
+\end{figure}
+
+We run this simulation $1000$ times for each of these methods and compute the
+Bias ($E[\hat Y] - \bar Y_N$), the RMSE ($\sqrt{\Var(\hat Y - \bar Y_N)}$), a 95\%
+empirical confidence interval ($\sum_{b = 1}^{1000} |\hat Y^{(b)} - \bar Y_N| \leq 
+\Phi(0.975)\sqrt{\hat V(\hat Y^{(b)})^{(b)}}$), and a T-test that assesses the
+unbiasedness of each estimator. The results are in Figure~\ref{fig:tpdc-mean}.
+
+\begin{figure}[ht!]
+  \centering
+\input{tables/tpdcsim_mean.tex}
+% This table was generated from /src/explore/20240411-tpdcsim.qmd
+\caption{This table shows the results of the simulation study. It displays the
+Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
+unbiasedness of each estimator for the estimators: $\pi^*$, TP-Reg, DC-Pop, and
+DC-Est.}
+\label{fig:tpdc-mean}
+\end{figure}
+
+\section{Topic 2: Non-nested Two-Phase Sampling}
+
+\textcolor{blue}{
+( Materials in Section 11.4 can be used here. I have copy-and-pasted the
+textbook materials below. Please modify them.  )
+}
+\textcolor{red}{I will do this shortly.}
+
+\subsection{Background}
+
+Now we consider the sampling mechanism known as non-nested two phase sampling 
+(\cite{kim2024statistics}). In the last section, we considered two phase sampling
+in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
+two phase sampling the Phase 2 sample is independent of the Phase 1 sample. It
+is a separate independent sample of the same population. Like traditional two
+phase sampling, we consider the Phase 1 sample, $A_1$, to consist of
+observations of $(X_i)_{i = 1}^{n_1}$ and the Phase 2 sample, $A_2$, to consist
+of observations of $(X_i, Y_i)_{i = 1}^{n_2}$.
+
+Whereas the classical two phase estimator uses a single Horvitz-Thompson
+estimator of the Phase 1 sample to construct estimates for calibration totals,
+in the non-nested two phase sample we have two independent Horvitz-Thompson
+estimators of the total of $X$,
+
+$$\hat X_1 = \sum_{i \in A_1}^{n_1} \frac{x_i}{\pi_{1i}} \text{ and } 
+\hat X_2 = \sum_{i \in A_2}^{n_2} \frac{x_i}{\pi_{2i}}$$
+
+where $\pi_{1i}$ is the probability of $i \in A_1$ and $\pi_{2i} = \Pr(i \in
+A_2)$. \textcolor{red}{I don't know if it makes sense to change the Phase 2
+selection probability as $\pi_{2i|1}$. Should I keep it at $\pi_{2i}$? But if
+I do, how is the two phase setup useful?} We can combine these estimates to get
+$\hat X_c = W \hat X_1 + (1 - W)\hat X_2$ for some matrix $W$ (see
+\cite{merkouris2004combining} for the optimal choice of $W$). We can then define
+a regression estimator as
+
+$$
+\hat Y_{NN, reg} = \hat Y_2 + (\hat X_c - \hat X_2)^T \hat \beta_q = 
+\hat Y_2 + (\hat X_1 - \hat X_2)^T W\hat \beta_q 
+$$
+
+where
+
+$$\hat \beta_q = \left(\sum_{i \in A_2} \frac{x_i x_i^T}{q_i}\right)^{-1}
+\sum_{i \in A_2} \frac{x_i y_i}{q_i} \text{ and }\hat Y_2 = \sum_{i \in A_2}
+\frac{y_i}{\pi_{2i}}.$$
+
+Since the samples $A_1$ and $A_2$ are independent,
+
+$$V(\hat Y_{NN, reg}) = V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i -
+x_i^TW\beta^*_q)\right) + (\beta^*_q)^T W^T V(\hat X_1) W \beta_q^*$$
+
+where $\beta_q^*$ is the probability limit of $\hat \beta_q$. Like the two phase
+sample this regression estimator can be viewed as the solution to the following
+calibration equation where $d_{2i} = \pi_{2i}^{-1}$,
+
+\begin{equation}\label{eq:nncal}
+\hat w = \argmin_w Q(w) = \sum_{i \in A_2} (w_{2i} - d_{2i})^2 q_i \text{ such
+that } \sum_{i \in A_2} w_{2i} x_i = \hat X_c
+\end{equation}
+
+and $\hat Y_{NN, reg} = \sum_{i \in A_2} \hat w_{2i} y_i$ where $\hat
+w_{2i}$ is the solution to Equation~\ref{eq:nncal}.
+
+We extend the debiased calibration estimator of \cite{kwon2024debiased} to the
+non-nested two phase sampling case where we use a combined estimate $\hat X_c$
+as the calibration totals instead of using the true totals from the finite
+population.
+
+\subsection{Methodology}
+
+\subsection{Theoretical Results}
+
+\subsection{Simulation Studies}
+
+\section{Topic 3: Multi-source Two-Phase Sampling}
+
+\bibliographystyle{chicago}
+\bibliography{references}
 

### 2024-04-23 src/explore/20240411-tpdcsim.qmd
Ran several basic simulations. Additional algorithms did not work.

@@
 library(GECal)
+library(nleqslv)
 library(doParallel)
 library(doRNG)
+library(CVXR)
+
+# source("GEcalib_code.R")
 
@@
   # mean(pi_i) = 0.1
-  # ptemp <- pt(-pop_df$X3 - 2, 3)
-  # pi_i <- ifelse(ptemp > 0.7, 0.7, ptemp)
-  ptemp <- pt(-pop_df$X3, 3)
-  pi_i <- ptemp
+  ptemp <- pt(-pop_df$X3 - 2, 3)
+  pi_i <- ifelse(ptemp > 0.7, 0.7, ptemp)
   del1 <- rbinom(nrow(pop_df), 1, pi_i)
@@
     mutate(pi1 = pi_i) %>%
-    rename(pi2 = X4) %>%
     mutate(del1 = del1) %>%
@@
   # Phase 2
-  # del2 <- rbinom(nrow(p1_df), 1, p1_df$pi2)
-  del2 <- rbinom(nrow(p1_df), 1, p1_df$pi2 / 2)
+  # pi2_1 <- p1_df$X4
+  p2tmp <- pt(-p1_df$X1 + 1, 3)
+  pi2_i <- ifelse(p2tmp > 0.9, 0.9, p2tmp)
+  del2 <- rbinom(nrow(p1_df), 1, pi2_i)
+
+  p1_df <- mutate(p1_df, pi2 = pi2_i)
 
@@
 #| label: estimators
+
+solveEL_prim <- function(xmat, d, xtot) {
+  w_vec <- Variable(nrow(xmat))
+  obj <- Maximize(sum(d * log(w_vec)))
+  constr <- list(t(xmat) %*% w_vec - xtot == 0)
+  prob <- Problem(obj, constr)
+  prim <- solve(prob)
+  resp <- prim$getValue(w_vec)
+  return(drop(resp))
+}
+
+# resw = resp
+# gbar = sum(-log(resw), na.rm = TRUE)
+# toremove <- which(is.na(log(resw)))
+# zbar = colSums(xmat[-toremove, ])
+# zbar = colSums(xmat)
+# lam = drop((gbar %*% zbar) %*% MASS::ginv(t(t(zbar)) %*% t(zbar)))
+# 
+# f(lam, xmat, d, xtot)
+# f(res$x, xmat, d, xtot)
+# 
+# res$x
+# lam
+# sum(p2_df$Y * resd) / pop_obs
+# sum(p2_df$Y * resp) / pop_obs
+
+solveEL <- function(xmat, d, xtot, acc2 = FALSE) {
+
+  f <- function(lam, xmat, d, xtot, returnw = FALSE) {
+    w <- -d / drop(xmat %*% lam)
+
+    if (returnw) {
+      return(w)
+    } else {
+      return(drop(w %*% xmat) - xtot)
+    }
+  }
+
+  j <- function(lam, xmat, d, xtot) {
+    t(xmat) %*% diag(d / drop(xmat %*% lam)^2) %*% xmat
+  }
+
+  init <- c(rep(0, ncol(xmat) - 1), 1)
+  res <- 
+  nleqslv(init, f, jac = j, xmat = xmat, d = d, xtot,
+          method = "Newton", control = list(maxit = 1e5, allowSingular = TRUE))
+
+  resw <- f(res$x, xmat, d, returnw = TRUE)
+
+  if (!(res$termcd %in% c(1))) {
+
+    return(NA)
+
+    w_prim <- solveEL_prim(xmat, d, xtot)
+    gbar = sum(-log(w_prim), na.rm = TRUE)
+    if (any(is.na(log(w_prim)))) {
+      toremove <- which(is.na(log(w_prim)))
+      zbar = colSums(xmat[-toremove, ])
+    } else {
+      zbar = colSums(xmat)
+    }
+    init = drop((gbar %*% zbar) %*% MASS::ginv(t(t(zbar)) %*% t(zbar)))
+
+    res2 <- 
+    nleqslv(init, f, jac = j, xmat = xmat, d = d, xtot,
+            method = "Newton", control = list(maxit = 1e5, allowSingular = TRUE))
+
+    if (!(res2$termcd %in% c(1))) {
+      return(NA)
+    }
+
+    resw <- f(res2$x, xmat, d, returnw = TRUE)
+
+  }
+
+  return(resw)
+
+}
 
@@
   samp_X <- cbind(rep(1, nrow(p2_df)), as.matrix(select(p2_df, X1, X2)))
-  dc_w <-  
-    GEcalib(cbind(samp_X, gdi),
-            d = rep(1, nrow(p2_df)),
-            extra$pop_tot,
-            entropy = entropy,
-            DS = FALSE)
+
+  # dc_w <-  
+  #   GEcalib(cbind(samp_X, gdi),
+  #           # d = rep(1, nrow(p2_df)),
+  #           d = 1 / (p2_df$pi2),
+  #           extra$pop_tot,
+  #           entropy = entropy,
+  #           DS = FALSE)
+
+  dc_w_alt <- solveEL(cbind(samp_X, gdi), 1 / p2_df$pi2, extra$pop_tot)
+  dc_w <- dc_w_alt
 
@@
 
+  if ("acc2" %in% names(extra)) {
+    acc2 <- extra$acc2
+  } else {
+    acc2 <- FALSE
+  }
+
   d_vec <- 1 / (p2_df$pi2 * p2_df$pi1)
@@
   samp_X <- cbind(rep(1, nrow(p2_df)), as.matrix(select(p2_df, X1, X2)))
-  dc_w <-  
-    GEcalib(cbind(samp_X, gdi),
-            d = rep(1, nrow(p2_df)),
-            pop_est,
-            entropy = entropy,
-            DS = FALSE)
+
+  dc_w_alt <- solveEL(cbind(samp_X, gdi), 1 / p2_df$pi2, pop_est, acc2)
+  dc_w <- dc_w_alt
 
@@
 
+  # FIXME: Change variance estimation to linearization estimate
+
   return(list(theta = theta, Var = as.numeric(vhat)))
@@
 
-mc_res <- foreach(iter=1:B_sims, .packages = c("dplyr", "GECal")) %dorng% {
+mc_res <- foreach(iter=1:B_sims,
+                  .packages = c("dplyr", "GECal", "CVXR", "nleqslv")) %dopar% {
 
+  set.seed(iter)
   samps <- gen_twophase(pop_df, n1_obs, n2_obs)
@@
 
-  # ptemp <- pt(-pop_df$X3 - 2, 3)
-  # pop_pi1 <- ifelse(ptemp > 0.7, 0.7, ptemp)
-  ptemp <- pt(-pop_df$X3, 3)
-  pop_pi1 <- ptemp
+  ptemp <- pt(-pop_df$X3 - 2, 3)
+  pop_pi1 <- ifelse(ptemp > 0.7, 0.7, ptemp)
 
@@
   tpreg <- tp_reg(p1_df, p2_df, nrow(pop_df))
+  # FIXME: GEcalib is not working. It produces results with very high
+  # values of the function.
   dcpop <- 
@@
            Theta = c(pistar[[1]], tpreg[[1]], dcpop[[1]], dcest[[1]]),
-           Var = c(pistar[[2]], tpreg[[2]], dcpop[[2]], dcest[[2]]))
+           Var = c(pistar[[2]], tpreg[[2]], dcpop[[2]], dcest[[2]]),
+           Iter = iter)
   )
@@
 which(is.na(mc_res$Var))
+mc_res %>%
+  mutate(iter = rep(1:B_sims, each = 4)) %>%
+  filter(is.na(Theta))
 
@@
 
+mc_res %>%
+  filter(Est == "dcpop") %>%
+  arrange(desc(Var))
+
 ```

### 2024-04-23 src/explore/GEcalib_code.R

@@ -1,0 +1,92 @@
+
+library(nleqslv)
+
+f = function(lambda, d, Xs, total, entropy, del, ..., returnw = F){
+  # w = d * ginv(drop(Xs %*% lambda), entropy = entropy)
+  if(entropy == "HD" & any(Xs %*% lambda >= 0)) return(rep(Inf, length(lambda)))
+  if(entropy == "PH" & any(abs(Xs %*% lambda) >= del)) return(rep(Inf, length(lambda)))
+  
+  if(entropy == "SL"){
+    w = d * drop(Xs %*% lambda)
+  }else if(entropy == "EL"){
+    w = -d / drop(Xs %*% lambda)
+  }else if(entropy == "ET"){
+    w = d * exp(drop(Xs %*% lambda))
+  }else if(entropy == "CE"){
+    w = d / (1 - exp(drop(Xs %*% lambda)))
+  }else if(entropy == "HD"){
+    w = d / drop(Xs %*% lambda)^2
+  }else if(entropy == "PH"){
+    w = d / sqrt(1 / drop(Xs %*% lambda)^2 - 1 / del^2)
+  }
+  if(entropy != "SL" & any(w <= 0)) return(rep(Inf, length(lambda)))
+  if(entropy == "CE" & any(w <= 1)) return(rep(Inf, length(lambda)))
+  if(returnw == T){
+    return(w)
+  }else{
+    return(colSums(Xs * w) - total)
+  }
+}
+
+h = function(lambda, d, Xs, total, entropy, del){
+  # return(t(Xs) %*% (Xs * d * ginvprime(drop(Xs %*% lambda), entropy = entropy)))
+  if(entropy == "SL"){
+    return(t(Xs) %*% (Xs * d))
+  }else if(entropy == "EL"){
+    w = -d / drop(Xs %*% lambda)
+    return(t(Xs) %*% (Xs * (w^2 / d)))
+  }else if(entropy == "ET"){
+    w = d * exp(drop(Xs %*% lambda))
+    return(t(Xs) %*% (Xs * w))
+  }else if(entropy == "CE"){
+    p_Stmp = 1 / (1 - exp(drop(Xs %*% lambda)))
+    return(t(Xs) %*% (Xs * d * p_Stmp * (p_Stmp - 1)))
+  }else if(entropy == "HD"){
+    return(t(Xs) %*% (Xs * (-2 * d / drop(Xs %*% lambda)^3)))
[…]

### 2024-04-27 final/reports/prelim_paper.tex
Updated Topic 1 simulation results. We are using a SRS-Poisson setup.

@@
 \end{equation}
-
 where $G: \mathcal{V} \to \R$ is strictly convex, differentiable function
@@
   \sum_{i \in A_2} \frac{g(\pi_{2i|1}^{-1})w_{2i}q_i}{\pi_{1i}} = 
-  \sum_{i \in A_1} \frac{g(\pi_{2i|1}^{-1})q_i}{\pi_{1i}}.
+  \sum_{i \in A_1} \frac{g(\pi_{2i|1}^{-1})q_i}{\pi_{1i}}, 
 \end{equation}
+where $g(w) = \partial G / \partial w$.
 
 The first constraint is the existing calibration constraint and the second
-ensures that design consistency is achieved. Here, 
-$g(w) = \partial G / \partial w$. 
+ensures that design consistency is achieved. 
 The original method of \cite{kwon2024debiased} only considered having finite
-population quantities on the right hand side of \ref{eq:calconst1}.
-Let $z_i = (x_i / q_i, g(\pi_{2i|1}^{-1}))$. Letting $w_{1i} = \pi_{1i}^{-1}$,
-the the goal is to solve,
+population quantities on the right hand side of \eqref{eq:calconst1}.
+
+Writing $w_{1i} = \pi_{1i}^{-1}$, the the goal is to solve
 
@@
   \text{ such that}
-  \sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i = \sum_{i \in A_1} w_{1i} z_i q_i.
+  \sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i = \sum_{i \in A_1} w_{1i} z_i q_i
 \end{equation}
 
-Let $\hat w_{2i|1}$ be the solution to Equation~\ref{eq:primal}, then the
-estimate of $Y_N$ is $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i|1} y_i$.
+where  $z_i = (x_i / q_i, g(\pi_{2i|1}^{-1}))$. 
+Let $\hat w_{2i|1}$ be the solution to Equation~\eqref{eq:primal}. The resulting 
+estimator of $Y_N$ is $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i|1} y_i$.
+To solve this problem we can use the method of Lagrange multipliers. We need to
+minimize the Lagrangian function
+
+\begin{equation}\label{eq:legragedc1}
+  L(w_{2i|1}, \lambda) = \sum_{i \in A_2} w_{1i} G(w_{2i|1}) q_i + \lambda 
+  \left( \sum_{i \in A_1} w_{1i} z_i q_i -
+    \sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i\right).
+\end{equation}
+
+Differntiating with respect to $w_{2i|1}$ and setting this expression equal to
+zero, yields the fact that $\hat w_{2i|1}$ satisfies
+
+$$ \hat w_{2i|1}(\hat \lambda) = g^{-1}(\hat \lambda^T z_i) $$
+
+where $\hat \lambda$ is the solution to
+
+\begin{equation}\label{eq:lamdc1}
+  \left( \sum_{i \in A_1} w_{1i} z_i q_i -
+    \sum_{i \in A_2} w_{1i} w_{2i|1}(\hat \lambda) z_i q_i\right) = 0.
+\end{equation}
 
@@
 
+% \textcolor{red}{In Theorem 1, you used $\hat{\lambda}$ without defining it.  }
 \begin{theorem}[Design Consistency]\label{thm:dc1}
+  Let $\lambda^*$ be the probability limit of $\hat \lambda$.
   Under some regularity conditions,
@@
 
-\textcolor{red}{I have tried to improve the expression of the linearization
-formula. Please check if my algebra above is correct. }
-\textcolor{blue}{This is correct. I have almost the same expression in
-Theorem~\ref{thm:dc2}.}
 \begin{proof}
@@
   that it is asymptotically equivalent to a regression estimator. Using the
-  method of Lagrange multipliers, to solve Equation~\ref{eq:primal} we need to
-  minimize,
-
-  $$\mathcal{L}(w_{2|1}) = \sum_{i \in A_2} w_{1i} G(w_{2i|1}) q_i + 
-  \lambda \left(\sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i - \sum_{i \in A_1} w_{1i}
-  z_i q_i\right).$$
-
+  method of Lagrange multipliers, to solve Equation~\eqref{eq:primal} we need to
+  minimize the Lagrangian in Equation~\eqref{eq:legragedc1}. 
   The first order conditions show that
 
-  $$\frac{\partial \mathcal{L}}{\partial w_{2i|1}}: g(w_{2i|1}) w_{1i}q_i +
+  $$\frac{\partial \mathcal{L}}{\partial w_{2i|1}}: g(w_{2i|1}) w_{1i}q_i -
     \lambda w_{1i} z_i q_i = 0.$$
 
-  Hence, $\hat w_{2i}(\lambda) = g^{-1}(\lambda z_i)$ and $\hat \lambda$ is
-  determined by the calibration condition. When the sample size gets large, we
+  Hence, $\hat w_{2i}(\lambda) = g^{-1}(\lambda^T z_i)$ and $\hat \lambda$ is
+  determined by Equation~\eqref{eq:lamdc1}. When the sample size gets large, we
   have $\hat w_{2i|1}(\hat \lambda) \to d_{2i|1}$ which means that $\hat \lambda
@@
     \lambda - \lambda^*)^2\\
-    &= \hat Y_\ell(\lambda^*, \phi^*) + O(N)o_p(n_2^{-1}).
+    &= \hat Y_\ell(\lambda^*, \phi^*) + O(N)O_p(n_2^{-1}).
   \end{align*}
-
+%\textcolor{red}{(The remainder term shoould be $O_p(N)O_p(n_2^{-1})$.)}
   The final equality comes from the fact that 
@@
 
-\subsection{Simulation Studies}
+\subsection{Simulation Study 1}
 
@@
 X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
-X_{4i} &\stackrel{ind}{\sim} Unif(0.1, 0.9) \\
 \varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
 Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
-\pi_{1i} &= \Phi_3(-x_{3i} - 2) \\
-\pi_{2i|1} &= x_{4i}.
+%\pi_{1i} &= \min(\Phi_3(-x_{3i} - 2), 0.7) \\
+\pi_{1i} &= n_1 / N \\
+\pi_{2i|1} &= \min(\Phi_3(-x_{1i} + 1), 0.9).
 \end{aligned}
@@
 where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
-
-\textcolor{red}{The simulation setup is somewhat strange.
-The first order inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$.
-In this case, your second-phase sampling is essentially equivalent to the simple random sampling.
-}
+%\textcolor{red}{The simulation setup is somewhat strange. The first order
+%inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a
+%function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$.
+%In this case, your second-phase sampling is essentially equivalent to the simple
+%random sampling. }
 This is a two-phase extension of the setup in \cite{kwon2024debiased}. We
-consider a finite population of size $N = 10,000$ with both the Phase 1 and
-Phase 2 sampling occuring under Poisson (Bernoulli) sampling. This yields a
+consider a finite population of size $N = 10,000$ with the Phase 1 sampling
+being a simple random sample (SRS) and the 
+Phase 2 sampling occuring under Poisson sampling. This yields a
 Phase 1 sample
-size of $E[n_1] \approx 1100$ and a Phase 2 sample size of
-$E[n_2] \approx 550$. In the Phase 1 sample, we observe 
+size of $n_1 = 1000$ and a Phase 2 sample size of
+$E[n_2] \approx 267$. In the Phase 1 sample, we observe 
 $(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. This
@@
     and $\bf x_i = (x_{1i}, x_{2i})^T$,
-    \textcolor{red}{Dr. Kim, should I modify the simulation so that the
-    regression estimator also includes $g(\pi_{2i|1}^{-1})$ as a covariate?}
-  \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This is
-    the method from \cite{kwon2024debiased} with the true population level
-    constraints, and 
+    % \textcolor{red}{Dr. Kim, should I modify the simulation so that the
+    % regression estimator also includes $g(\pi_{2i|1}^{-1})$ as a covariate?}
+  \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
+    solves
+
+  \begin{equation}
+    \argmin_{w_{2|1}} \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i})
+    \text{ such that}
+    \sum_{i \in A_2} w_{1i} w_{2i|1} z_i = \sum_{i \in U} z_i.
+  \end{equation}
+
   \item[4.] Debiased Calibration with Estimated Population Constraints (DC-Est):
-    This is the proposed method with the Phase 1 sample being used to estimate
-    the population level constraints.
+    This solves Equation~\eqref{eq:primal} with $q_i = 1$.
 \end{itemize}
@@
 variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$. For each
-approach we have the following variance estimate\footnote{These variance
-estimates use the fact that we have Poisson sampling for both phases in the
-simulation.}:
+approach we give the variance estimate in Table~\ref{tab:varforms}.
 
-\begin{figure}[ht!]
+\begin{table}[ht!]
   \centering
@@
     \toprule
-    Estimator & Variance & Notes \\
+    Estimator & Estimated Variance & Notes \\
     \midrule
-    $\pi^*$ & $N^{-2} \sum_{i \in A_2} \left(\pi_{2i|1}^{-2} - \pi_{2i|1}^{-1}\right)
-    y_i^2 $ & \\
-    TP-Reg  & 
-      {\scriptsize $N^{-2}\left(\sum\limits_{i \in A_1} \left(\pi_{1i}^{-2} -
-          \pi_{1i}^{-1}\right) 
-      \eta_i^2 
-      + \sum\limits_{i \in A_2} \frac{1}{\pi_{1i} \pi_{2i|1}}(\pi_{2i|1}^{-1}
-  - 1) (y_i - \bf x_i' \hat \beta)^2 \right)$}
-    & 
-      {\scriptsize$\eta_i = \bf x_i \hat \beta + \frac{\delta_{2i}}{\pi_{2i|1}}(y_i - 
-      \bf x_i \hat \beta)$} \\
-      %$\hat \beta = \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$ \\
-      DC-Pop  & $(Y - \bf z^T \hat \gamma)^T \Pi (Y - \bf z^T \hat \gamma)$ & 
-      {\scriptsize$\Pi
-      = \text{diag}(1 - (\pi_{1}\pi_{2|1})^{-1}) \cdot \frac{\hat w^2}{N^2}$} \\
-      DC-Est  & $(Y - \bf z^T \hat \gamma)^T \Pi (Y - \bf z^T \hat \gamma) +
-      \hat \gamma_{[1:3]}^T V(\bf x) \hat \gamma_{[1:3]}/N^2$ & \\
+    $\pi^*$ 
+    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_Y + 
+      \sum\limits_{i \in A_2} \frac{w_{1i}^2}{N^2} 
+        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}y_i^2$}
+    & {\scriptsize$\hat s^2_Y = \frac{1}{n_2 - 1}\sum\limits_{i \in A_2} (y_i -
+    \bar y_i)^2$} \\
+    TP-Reg  
+    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\eta + 
+      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
+        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
+    & {\scriptsize$ \eta_i = x_i'\beta + \delta_{2i} w_{2i|1}\varepsilon_i,\;
+        \varepsilon_i = (y_i - x_i' \beta)$} \\
+    DC-Pop  
+    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\varepsilon + 
+      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
+        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
+    & {\scriptsize$\eta_i = z_i' \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
+        \varepsilon_i = (y_i - z_i'\phi)$} \\
+    DC-Est  
+    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\eta + 
+      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
+        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
+    & {\scriptsize$\eta_i = z_i' \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
+        \varepsilon_i = (y_i - z_i' \phi)$} \\
     \bottomrule
@@
   \caption{This table gives the formulas for each variance estimator used in
-  this simulation.}
+  Simulation 1. For the derivation of the estimated variance of the DC-Pop
+estimatator see Appendix A.}
   \label{tab:varforms}
-\end{figure}
+\end{table}
 
-We run this simulation $1000$ times for each of these methods and compute the
+We run the simulation $1000$ times for each of these methods and compute the
 Bias ($E[\hat Y] - \bar Y_N$), the RMSE ($\sqrt{\Var(\hat Y - \bar Y_N)}$), a 95\%
@@
 \Phi(0.975)\sqrt{\hat V(\hat Y^{(b)})^{(b)}}$), and a T-test that assesses the
-unbiasedness of each estimator. The results are in Figure~\ref{fig:tpdc-mean}.
+unbiasedness of each estimator where $\hat Y^{(b)}$ is the result from the $b$th
+simulation replicate. The results are in Table~\ref{tab:tpdc-mean}.
 
-\begin{figure}[ht!]
+\begin{table}[ht!]
   \centering
 \input{tables/tpdcsim_mean.tex}
-% This table was generated from /src/explore/20240411-tpdcsim.qmd
-\caption{This table shows the results of the simulation study. It displays the
+% This table was generated from ../../src/explore/proto-dctpmcmc.R
+\caption{This table shows the results of Simulation Study 1. It displays the
 Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
@@
 DC-Est.}
-\label{fig:tpdc-mean}
-\end{figure}
+\label{tab:tpdc-mean}
+\end{table}
+
+\textcolor{red}{Dr. Kim is this confidence interval good enough? It seems
+slightly high for me.}
+
+\textcolor{red}{I also give the mean variance estimates, the Monte Carlo mean,
+as well as a test for bias.}
+\begin{table}[ht!]
+  \centering
+{\color{red} \input{tables/tpdcsim_var.tex} }
+% This table was generated from ../../src/explore/proto-dctpmcmc.R
+\caption{\textcolor{red}{This table shows the variance estimates from Simulation
+    Study 1. It displays the Monte Carlo variance of $\hat Y^{(b)}$, the average
+    estimated variance, the variance of the variance estimator, and a
+    t-statistic assessing the unbiasedness of each variance estimator.}}
+\label{tab:tpdc-var}
+\end{table}
 
@@
 Now we consider the sampling mechanism known as non-nested two phase sampling 
-(\cite{kim2024statistics}). In the last section, we considered two phase sampling
+(\cite{hidiroglou2001double}). In the last section, we considered two phase sampling
 in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
@@
 
-$$\hat X_1 = \sum_{i \in A_1}^{n_1} \frac{x_i}{\pi_{1i}} \text{ and } 
-\hat X_2 = \sum_{i \in A_2}^{n_2} \frac{x_i}{\pi_{2i}}$$
+$$\hat X_1 = \sum_{i \in A_1}^{n_1} d_{1i} x_i \text{ and } 
+\hat X_2 = \sum_{i \in A_2}^{n_2} d_{2i} x_i $$
 
-where $\pi_{1i}$ is the probability of $i \in A_1$ and $\pi_{2i} = \Pr(i \in
-A_2)$. \textcolor{red}{I don't know if it makes sense to change the Phase 2
-selection probability as $\pi_{2i|1}$. Should I keep it at $\pi_{2i}$? But if
-I do, how is the two phase setup useful?} We can combine these estimates to get
+where $d_{1i} = \pi_{1i}^{-1}$, $d_{2i} = \pi_{2i}^{-1}$, $\pi_{1i}$ is the
+probability of $i \in A_1$ and $\pi_{2i} = \Pr(i \in A_2)$. 
+%\textcolor{red}{I don't know if it makes sense to change the Phase 2
+%selection probability as $\pi_{2i|1}$. (\textcolor{blue}{No. We do not know
+%$\pi_{2i |1}$ and it is not necessary.}) Should I keep it at $\pi_{2i}$?
+%\textcolor{blue}{(Yes, $\pi_{2i}$ is available to us.) } But if
+%I do, how is the two phase setup useful?} 
+%\textcolor{blue}{Non-nested two-phase sampling is different from the classical
+%  two-phase sampling. We are not using the two-phase setup of Section 2  here.
+%  We do not have to call it two-phase sampling. However, if we define $A_c=A_1
+%  \cup A_2$ then $A_c$ serves the role of the first-phase sample in the
+%  two-phase sampling. Then, our goal is to compute an estimator of $X$ from
+%  $A_c$. The second-phase sample would be $A_2$. 
+%}
+We can combine these estimates to get
 $\hat X_c = W \hat X_1 + (1 - W)\hat X_2$ for some matrix $W$ (see
@@
 \hat Y_{NN, reg} = \hat Y_2 + (\hat X_c - \hat X_2)^T \hat \beta_q = 
-\hat Y_2 + (\hat X_1 - \hat X_2)^T W\hat \beta_q 
+\hat Y_2 + (\hat X_1 - \hat X_2)^T W^T\hat \beta_q 
 $$
 
-where
+where % \textcolor{red}{(we may need transpose in $W$)}
 
@@
 \sum_{i \in A_2} \frac{x_i y_i}{q_i} \text{ and }\hat Y_2 = \sum_{i \in A_2}
-\frac{y_i}{\pi_{2i}}.$$
+d_{2i} y_i. $$
 
@@
 sample this regression estimator can be viewed as the solution to the following
-calibration equation where $d_{2i} = \pi_{2i}^{-1}$,
+calibration equation
 
 \begin{equation}\label{eq:nncal}
-\hat w = \argmin_w Q(w) = \sum_{i \in A_2} (w_{2i} - d_{2i})^2 q_i \text{ such
-that } \sum_{i \in A_2} w_{2i} x_i = \hat X_c
+  \hat w_2 = \argmin_{w_2} Q(w_2) = \sum_{i \in A_2} (w_{2i} - d_{2i})^2 q_i 
+  \text{ such that } \sum_{i \in A_2} w_{2i} x_i = \hat X_c
 \end{equation}
@@
 \hat w_2 = \argmin_w \sum_{i \in A_2} G(w_{2i}) q_i \text{ such that } 
-\sum_{i \in A_2} w_{2i} x_i = \hat X_{GLS} \text{ and } 
+\sum_{i \in A_2} w_{2i} x_i = \hat X_{c} \text{ and } 
 \sum_{i \in A_2} w_{2i} g(d_{2i}) q_i = \sum_{i \in U} g(d_{2i}) q_i
@@
 
-for $g(x) = G'(x)$ and a known choice of $q_i \in \mathbb{R}$. The difference
+for $g(x) = G'(x)$ and a known choice of $q_i \in \mathbb{R}$. 
+%\textcolor{red}{(Instead of using $\hat{X}_{\rm GLS}$, we used $\hat{X}_c$
+%before. Please make notation consistent. )}
+
+The difference
 between solving Equation~\ref{eq:nnopt} and Equation~\ref{eq:primal} is that the
-estimator $\hat X_{GLS}$ is estimated using GLS from other sample(s). Before using
-$\hat X_{GLS}$ in the debiased calibration estimator, we need to estimate it
+estimator $\hat X_{c}$ is estimated from the combined sample $A_c= A_1 \cup
+A_2$. Before using
+$\hat X_{c}$ in the debiased calibration estimator, we need to estimate it
 from the non-nested samples.
@@
 
-$$\hat X_1 = \sum_{i \in A_1} \frac{x_i}{\pi_{1i}} \text{ and }
-\hat X_2 = \sum_{i \in A_2} \frac{x_i}{\pi_{2i}}.$$
+$$\hat X_1 = \sum_{i \in A_1} d_{1i} x_i \text{ and }
+\hat X_2 = \sum_{i \in A_2} d_{2i} x_i.$$
 
 Let $V_1$ and $V_2$ be known variance matrices for $\hat X_1$ and $\hat X_2$,
-then the optimal combined GLS estimates is
+then the optimal combined estimate is
 
@@
 $\hat Y_{NNE} = \sum_{i \in A_2} \hat w_{2i} y_i$ and $\hat w_{2i}$ solves
-Equation~\ref{eq:nnopt}.
+Equation~\ref{eq:nnopt}. Like the classical two phase approach, to solve this
+setup we minimize the Lagrangian,
+
+\begin{equation}\label{eq:legragedc2}
+  L(w_{2i}, \lambda) = \sum_{i \in A_2} G(w_{2i}) q_i + \lambda 
+  \left( \hat T - \sum_{i \in A_2} w_{2i} z_i q_i\right).
+\end{equation}
+
+with
+
+$$\hat T = 
+\begin{bmatrix}
+  \hat X_c \\ \sum_{i \in U} g(d_{2i}) q_i
+\end{bmatrix}.
+$$
+
+\textcolor{red}{I am a little worried about the second term of $\hat{T}$. It is
+  somewhat weird to assume that $T_g= \sum_{i \in U} g(d_{2i}) q_i$ is known.
+  One simple way is to use a HT estimator  of $T_g$ from sample $A_2$. That is,
+  use $\hat{T}_{g, {\rm HT}} = \sum_{i \in A_2} d_{2i} g( d_{2i}) q_i$ or even
+  use a GREG estimator of $T_g$. 
+%I wonder we can assume that an unbiased estimator of $T_g$ is available from
+  %$A_c=A_1 \cup A_2$. That is, let's assume that we have $\hat{T}_c = W
+  %\hat{T}_1 + (1- W) \hat{T}_2$ as the estimator  for $T$. 
+}
+
+Differntiating with respect to $w_{2i}$ and setting this expression equal to
+zero, yields the fact that $\hat w_{2i}$ satisfies
+
+$$ \hat w_{2i}(\hat \lambda) = g^{-1}(\hat \lambda^T z_i) $$
+
+where $\hat \lambda$ is the solution to
+
+\begin{equation}\label{eq:lamdc2}
+  \left( \hat T - \sum_{i \in A_2} w_{2i}(\hat \lambda) z_i q_i\right) = 0.
+\end{equation}
 
@@
 \begin{theorem}[Design Consistency]\label{thm:dc2}
-  Under some regularity conditions, $\hat Y_{NNE} = \hat Y_{\ell, NNE}(\lambda^*,
+  Allowing $\lambda^*$ to be the probability limit of $\hat \lambda$,
+  under some regularity conditions, $\hat Y_{NNE} = \hat Y_{\ell, NNE}(\lambda^*,
   \phi^*) + O_p(Nn_2^{-1})$ where
@@
   $$\hat Y_{\ell, NNE}(\lambda^*, \phi^*) = \sum_{i \in A_2} \hat w_{2i}(
-  \lambda^*) + \left(T - \sum_{i \in A_2} \hat w_{2i}(\lambda^*) z_i q_i\right)
+  \lambda^*) + \left(\hat T - \sum_{i \in A_2} \hat w_{2i}(\lambda^*) z_i q_i\right)
   \phi^*$$
@@
   $$
-
+% \textcolor{red}{$T$  is not defined. }
 \end{theorem}
@@
       x_i / q_i \\ g(d_i)
-  \end{bmatrix}}_{z_i} =
-  \underbrace{\begin{bmatrix} 
-      \hat X_c \\
-      \sum_{i \in U} g(d_i) q_i
-  \end{bmatrix}}_{T}
+  \end{bmatrix}}_{z_i} = T
   \end{equation}
@@
     &= (\phi_1^*)^T \Var(\hat X_c) \phi_1^* + 
-    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} (y_i - z_i \phi^* q_i)(y_j - z_j \phi^* q_j)\\
-    &\qquad + (1 - W)\phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij} \frac{x_i}{\pi_{2i}}
-    \frac{(y_j - z_j \phi^*_1 q_j)}{\pi_{2j}}\\
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
+      (y_i - z_i \phi^* q_i)(y_j - z_j \phi^* q_j)\\
+    &\qquad + (1 - W)\phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij} d_{2i} x_i
+      d_{2j}(y_j - z_j \phi^*_1 q_j)\\
   \end{align*}
@@
 
-\subsection{Simulation Studies}
+\subsection{Simulation Study 2}
+
+We run a simulation testing the proposed method. This is very similar to
+Simulation 1. We have the following simulation setup:
+
+$$
+\begin{aligned}
+X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
+X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
+X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
+\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
+Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
+\pi_{1i} &= n_1 / N \\
+\pi_{2i} &= \min(\Phi_3(-x_{1i} + 1), 0.9).
+\end{aligned}
+$$
+
+where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
+%\textcolor{red}{The simulation setup is somewhat strange. The first order
+%inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a
+%function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$.
+%In this case, your second-phase sampling is essentially equivalent to the simple
+%random sampling. }
+We consider a finite population of size $N = 10,000$ with the Phase 1 
+sampling being a simple random sample (SRS) of size $n_1 = 1000$. The Phase 2
+sample is a Poisson sample with an expected sample size of
+
+and
+Phase 2 sampling occuring under Poisson (Bernoulli) sampling. This yields a
+Phase 1 sample
+size of $E[n_1] \approx 1100$ and a Phase 2 sample size of
+$E[n_2] \approx 300$. In the Phase 1 sample, we observe 
+$(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. This
+simulation does not deal with model misspecification, and we compare the
+proposed method for the parameter $\bar Y_N$ with four approaches:
+
+\begin{itemize}
+  \item[1.] HT-estimator: $\hat Y_{HT} = N^{-1} \sum_{i \in A_2}
+    \frac{y_i}{\pi_{2i}},$
+  \item[2.] Regression estimator (Reg): (FIXME)
+  \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
+    solves
+
+  \begin{equation*}
+  \hat w_2 = \argmin_w \sum_{i \in A_2} G(w_{2i}) q_i \text{ with } 
+  \sum_{i \in A_2} w_{2i} x_i = \sum_{i \in U} x_i \text{ and } 
+  \sum_{i \in A_2} w_{2i} g(d_{2i}) q_i = \sum_{i \in U} g(d_{2i}) q_i
+  \end{equation*}
+
+  \item[4.] Debiased Calibration with Estimated Population Constraints (DC-Est):
+    This solves Equation~\eqref{eq:nnopt} with $q_i = 1$.
+\end{itemize}
+
+In addition to estimating the mean parameter $\bar Y_N$, we also construct
+variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$. For each
+approach we give the variance estimate in Table~\ref{tab:varforms}.
 
 \section{Topic 3: Multi-source Two-Phase Sampling}
+
+\newpage
 
@@
 
+\newpage
+
+\appendix
+
+\section{Derivation of the Estimated Variance of DC-Pop in Simulation 1}
+
+We know that $\hat Y_{DC-Pop} = \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat
+\lambda) y_i$ where $\hat w_{2i|1}$ solves
+
+  \begin{equation}
+    \argmin_{w_{2|1}} \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i})
+    \text{ such that}
+    \sum_{i \in A_2} w_{1i} w_{2i|1} z_i = \sum_{i \in U} z_i.
+  \end{equation}
+
+This means that using the technique of \cite{randles1982asymptotic},
+the linearization of $\hat Y_{DC-Pop}$ is
+
+$$\hat Y_{DC-Pop, \ell} = \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat \lambda) y_i + 
+\left( \sum_{i \in U} z_i - 
+\sum_{i \in A_2} w_{1i} w_{2i|1}(\hat \lambda) z_i \right) \phi.$$
+
+Since $\phi^*$ satisfies $E\left[\frac{\partial}{\partial \lambda} \hat Y_{DC-Pop,
+\ell}(\lambda^*, \phi^*)\right] = 0$, $\phi^*$ is the same as it is for $\hat Y_{DC-Est,
+\ell}$ with $q_i = 1$,
+
+  \begin{align*}
+    \phi^*
+    &= E\left[\sum_{i \in A_2} \frac{w_{1i}z_i z_i^T }{g'(d_{2i|1})}\right]^{-1}
+    E\left[\sum_{i \in A_2} \frac{w_{1i}z_i y_i}{g'(d_{2i|1})}\right]\\
+    &= \left[\sum_{i \in U} \frac{\pi_{2i|1}z_i z_i^T}{g'(d_{2i|1})}\right]^{-1}
+    \left[\sum_{i \in U} \frac{\pi_{2i|1} z_i y_i}{g'(d_{2i|1})}\right].\\
+  \end{align*}
+
+Hence,
+
+$$
+\hat Y_{DC-Pop, \ell} = \sum_{i \in U} z_i \phi^* + \sum_{i \in A_2} w_{1i} \hat
+w_{2i|1}(\lambda^*) (y_i - z_i \phi^*).
+$$
+
+Unlike the two-phase regression estimator and $\hat Y_{DC-Est}$ this
+linearization only has uncertainty in the second term. The first term is a
+population estimate. Let $\varepsilon_i = (y_i - z_i \phi^*)$. This means that
+
+\begin{align*}
+  V(\hat Y_{DC-Pop, \ell}) 
+  &= V(E[\hat Y_{DC-Pop, \ell} \mid A_1]) + E[V(\hat Y_{DC-Pop, \ell} \mid A_1)] \\
+  &= V\left(\sum_{i \in A_1} w_{1i} \pi_{2i|1} \hat w_{2i|1}(\lambda^*)
+  \varepsilon_i\right) + 
+  E\left[\sum_{i \in A_1} \pi_{2i|1}(1 - \pi_{2i|1}) w_{1i}^2 \hat w_{2i|1}(\lambda^*)^2
+    \varepsilon_i^2 \right]\\
+  &= \sum_{i \in U} \pi_{1i}(1 - \pi_{1i}) w_{1i}^2 \pi_{2i|1}^2 
+  \hat w_{2i|1}(\lambda^*)^2 \varepsilon_i^2 +
+  \sum_{i \in U} \pi_{1i}\pi_{2i|1}(1 - \pi_{2i|1}) w_{1i}^2 \hat w_{2i|1}(\lambda^*)^2
+    \varepsilon_i^2\\
+  &= \sum_{i \in U} w_{1i} w_{2i|1}(\lambda^*)^2 \pi_{2i|1}
+  ((1 - \pi_{1i})\pi_{2i|1} + (1 - \pi_{2i|1})) \varepsilon_i^2.
+\end{align*}
+
+Since $w_{2i|1}(\lambda^*) \pi_{2i|1} = 1$, there are two potential estimators
+of the variance,
+
+\begin{align*}
+  \hat V_1 &= \sum_{i \in A_2} w_{1i}^2 w_{2i|1}(\hat \lambda)^2 
+  ((1 - \pi_{1i})\pi_{2i|1} + (1 - \pi_{2i|1})) \varepsilon_i^2\\
+  \hat V_2 &= \sum_{i \in A_2}  w_{1i}^2 d_{2i|1}^2
+  ((1 - \pi_{1i})\pi_{2i|1} + (1 - \pi_{2i|1})) \varepsilon_i^2.
+\end{align*}
+
+In the current simulation, we use $\hat V_1$, but I have tried $\hat V_2$ and
+the results are similar.
 \end{document}

### 2024-04-27 final/reports/prelim_paper_ov.tex
Updated Topic 1 simulation results. We are using a SRS-Poisson setup.

@@
 
-\begin{theorem}[Design Consistency]
+% \textcolor{red}{In Theorem 1, you used $\hat{\lambda}$ without defining it.  }
+\begin{theorem}[Design Consistency]\label{thm:dc1}
+  Let $\lambda^*$ be the probability limit of $\hat \lambda$.
   Under some regularity conditions,
@@
   where
-\textcolor{red}{
+
+  $$\hat Y_{\ell}(\hat \lambda, \phi^*) = \hat Y_{DCE}(\hat \lambda) + 
+  \left(\sum_{i \in A_1} w_{1i} z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat
+  \lambda) z_i q_i\right)\phi^*$$
+
+  and
+
+  $$\phi^* = 
+  \left[\sum_{i \in U} \frac{\pi_{2i|1}z_i z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
+  \sum_{i \in U} \frac{\pi_{2i|1}z_i y_i}{g'(d_{2i|1})}.$$
+
    \begin{eqnarray*}
    \hat Y_{\ell}( \lambda^*, \phi^*) &=&   \hat{Y}_{\pi^*} + 
-  \left(\sum_{i \in A_1} w_{1i} x_i  -  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} x_i  \right)^T \phi_1^* + \left(\sum_{i \in A_1} w_{1i} g_i   -  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^T \phi_2^*
+  \left(\sum_{i \in A_1} w_{1i} x_i  -  \sum_{i \in A_2} w_{1i} \pi_{2i \mid
+  1}^{-1} x_i  \right)^T \phi_1^* + \left(\sum_{i \in A_1} w_{1i} g_i   -
+\sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^T \phi_2^*  
   \end{eqnarray*} 
@@
   \end{pmatrix}y_i $$
-with $g_i = g( \pi_{2i |1}^{-1}) q_i$. }
+with $g_i = g( \pi_{2i |1}^{-1}) q_i$.
 \end{theorem}
 
-\textcolor{red}{I have tried to improve the expression of the linearization formula. Please check if my algebra above is correct. }
 \begin{proof}
@@
   = 0$, $\frac{\partial}{\partial \lambda^2} \hat w_{2i|1}(\lambda^*)$ is
-  bounded and $|\hat \lambda - \lambda^*| = o_p(n_2^{-1/2})$, which proves our
+  bounded and $|\hat \lambda - \lambda^*| = O_p(n_2^{-1/2})$, which proves our
   result.
-\footnote{$\hat{\lambda}$ should satisfy $\hat{\lambda} - \lambda^* = O_p (n_2^{-1/2})$. }
 \end{proof}
 
-%\begin{proof}
-%  First, we derive the solution for $\hat Y_{DCE}$, and then derive the
-%  asymptotic variance of this estimator using the linearization technique of
-%  \cite{randles1982asymptotic}.
-%
-%  We want to solve Equation~\ref{eq:primalloss} subject to the constraints in
-%  Equations~\ref{eq:calconst1} and \ref{eq:calconst2}. Using the method of
-%  Lagrange multipliers we have a Lagrangian of 
-%
-%  $$\mathcal{L}(w_2) = -\sum_{i \in A_2} w_{1i} G(w_{2i})q_i + \lambda 
-%  \left(\sum_{i \in A_2} w_{1i} w_{2i} z_i - \sum_{i \in A_1} w_{1i} z_i\right).$$
-%
-%  Our first order conditions yield,
-%
-%  \begin{align*}
-%    0 &= \frac{\partial}{\partial w_{2i}} \mathcal{L}(w_{2}) \\ 
-%      &= -w_{1i} g(w_{2i}) q_i + \lambda w_{1i} z_i.
-%  \end{align*}
-%
-%  This means that $\hat w_{2i} = g^{-1}(\hat \lambda z_i / q_i)$ and since $\hat
-%  \lambda$ is determined from the solution to Equations~\ref{eq:calconst1} and
-%  \ref{eq:calconst2}. Hence, $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} 
-%  \hat w_{2i}(\hat \lambda) y_i$.
-%
-%  To get the variance of $\hat Y_{DCE}$ we use the linearization technique of
-%  \cite{randles1982asymptotic}. First, notice that 
-%
-%  $$\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i =
-%  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i + \left(
-%  \sum_{i \in A_1} w_{1i} z_i -
-%  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) z_i\right)\phi
-%  := \hat Y_\ell(\hat \lambda, \phi).$$
-%
-%  We can find the value of $\phi$ such that $E[\frac{\partial}{\partial
-%  \lambda}\hat Y_\ell(\lambda, \phi)]_{\lambda = \lambda^*}$ where $\lambda^*$
-%  is the probability limit of $\hat \lambda$. Notice, that 
-%  $\lambda^* = (\bf 0, 1)$ because $\hat w_{2i} \to \pi_{2i|1}^{-1}$. This means
-%  that the optimal value of $\phi$ solves,
-%
-%  $$E\left[\frac{\partial}{\partial \lambda} \hat Y_\ell(\lambda, \phi)\right] =
-%  E\left[\sum_{i \in A_2} w_{1i} \frac{\partial}{\partial \lambda} w_{2i|1}(\lambda)
-%    y_i - \phi \sum_{i \in A_2} w_{1i} z_i \frac{\partial}{\partial \lambda}
-%    w_{2i|1}(\lambda) / q_i \right]$$
-%
-%  which means that
-%
-%  $$\phi^* = \left[\sum_{i \in U} \pi_{2i|1} \left(\frac{z_i z_i^T}{
-%    g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in U} \pi_{2i|1}
-%    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
-%
-%  Hence,
-%
-%  \begin{align*}
-%    \hat Y_\ell(\lambda^*, \phi^*) 
-%    &= \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1}
-%    y_i + \left(\sum_{i \in A_1} w_{1i} z_i / q_i -
-%      \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} z_i / q_i\right)\phi^* \\ 
-%    &= \sum_{i \in A_1} w_{1i} \eta_i
-%  \end{align*}
-%
-%  where 
-%
-%  $$\eta_i = z_i \phi^* / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
-%  - z_i \phi^* / q_i\right).$$
-%
-%  To analyze the variance, consider the reverse framework of
-%  \cite{fay1992inferences}, \cite{kim2006replication}, and \cite{shao1999variance}.
-%  In this setup, we assume that $\mathcal{R} = \{\delta_{2i|1}\}_{i \in U}$ is
-%  known at the
-%  finite population level. Then with this we have,
-%
-%  \begin{align*}
-%    V(\hat Y_{\ell}(\lambda^*, \phi^*)) 
-%    &= V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\right) \\
-%    &= V\left(E\left[\sum_{i \in A_1} w_{1i}\eta_i(\phi^*)\mid\mathcal{R}\right]\right) 
-%    +E\left[V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\mid\mathcal{R}\right)\right]\\
-%    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right) + E\left[\sum_{i \in U} 
-%    \sum_{j \in U} \Cov(\delta_{1i}, \delta_{1j}) 
-%    w_{1i} w_{1j} \eta_i(\phi^*) \eta_j(\phi^*)\right]\\
-%  \end{align*}
-%
-%  We can estimate this variance using 
-%
-%  $$\hat V_{DCE} = \sum_{i \in A_1} \sum_{j \in A_1} \left(\frac{\pi_{1ij} -
-%  \pi_{1i}\pi_{1j}}{\pi_{1ij}}
-%    w_{1i} w_{1j} \hat \eta_i \hat \eta_j \right)
-%  $$
-%
-%  for $\hat \eta_i = z_i \hat \phi / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
-%  - z_i \hat \phi / q_i\right)$ where 
-%
-%  $$\hat \phi = \left[\sum_{i \in A_2} w_{1i}\left(\frac{z_i z_i^T}{
-%  g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in A_2} w_{1i}
-%    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
-%
-%  The problem with this estimator is that it is biased as $E[\hat V_{DCE}]$ is
-%  asymptotically equivalent to
-%  $V\left(\sum_{i \in A_1} w_{1i} \eta(\phi^*)\right)$. This yields a bias of 
-%
-%  \begin{align*}
-%    \text{Bias}(\hat V_{DCE}) 
-%    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right)\\ 
-%    &= \sum_{i \in U} \sum_{j \in U} \Delta_{2ij|1} w_{2i|1}
-%    w_{2j|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).
-%  \end{align*}
-%
-%  where $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1} \pi_{2j|1}$. 
-%  However, we can estimate this bias with
-%
-%  $$\widehat{\text{Bias}}(\hat V_{DCE}) = 
-%  \sum_{i \in A_2} \sum_{j \in A_2} w_{1ij} w_{2ij|1} w_{2i|1} w_{2j|1} 
-%  \Delta_{2ij|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).$$
-%
-%\end{proof}
-
-\subsection{Simulation Studies}
+\subsection{Simulation Study 1}
 
@@
 X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
-X_{4i} &\stackrel{ind}{\sim} Unif(0.1, 0.9) \\
 \varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
 Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
-\pi_{1i} &= \Phi_3(-x_{3i} - 2) \\
-\pi_{2i|1} &= x_{4i}.
+\pi_{1i} &= \min(\Phi_3(-x_{3i} - 2), 0.7) \\
+\pi_{2i|1} &= \min(\Phi_3(-x_{1i} + 1), 0.9).
 \end{aligned}
@@
 where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
-
-\textcolor{red}{The simulation setup is somewhat strange.
-The first order inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$.
-In this case, your second-phase sampling is essentially equivalent to the simple random sampling.
-}
+%\textcolor{red}{The simulation setup is somewhat strange. The first order
+%inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a
+%function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$.
+%In this case, your second-phase sampling is essentially equivalent to the simple
+%random sampling. }
 This is a two-phase extension of the setup in \cite{kwon2024debiased}. We
@@
 size of $E[n_1] \approx 1100$ and a Phase 2 sample size of
-$E[n_2] \approx 550$. In the Phase 1 sample, we observe 
+$E[n_2] \approx 300$. In the Phase 1 sample, we observe 
 $(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. This
@@
     \toprule
-    Estimator & Variance & Notes \\
+    Estimator & Estimated Variance & Notes \\
     \midrule
@@
     TP-Reg  & 
-      {\scriptsize $N^{-2}\left(\sum\limits_{i \in A_1} \left(\pi_{1i}^{-2} -
-          \pi_{1i}^{-1}\right) 
-      \eta_i^2 
-      + \sum\limits_{i \in A_2} \frac{1}{\pi_{1i} \pi_{2i|1}}(\pi_{2i|1}^{-1}
-  - 1) (y_i - \bf x_i' \hat \beta)^2 \right)$}
+    {\scriptsize $N^{-2}\left(\sum\limits_{i \in A_1} 
+        \left(\frac{1 - \pi_{1i}}{\pi_{1i}^{2}}\right) \eta_i^2 
+        + \sum\limits_{i \in A_2} \frac{1}{\pi_{1i} \pi_{2i|1}}(d_{2i|1} - 1)
+        (y_i - \bf x_i' \hat \beta)^2 \right)$}
     & 
@@
       %$\hat \beta = \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$ \\
-      DC-Pop  & $(Y - \bf z^T \hat \gamma)^T \Pi (Y - \bf z^T \hat \gamma)$ & 
-      {\scriptsize$\Pi
-      = \text{diag}(1 - (\pi_{1}\pi_{2|1})^{-1}) \cdot \frac{\hat w^2}{N^2}$} \\
-      DC-Est  & $(Y - \bf z^T \hat \gamma)^T \Pi (Y - \bf z^T \hat \gamma) +
-      \hat \gamma_{[1:3]}^T V(\bf x) \hat \gamma_{[1:3]}/N^2$ & \\
+    DC-Pop  & {\scriptsize$\sum\limits_{i \in A_2} c_i (y_i - q_i z_i \hat \gamma)^2 
+              (w_{1i} \hat w_{2i|1}(\hat \lambda) / N)^2$}
+            & {\scriptsize$c_i = ((1 - \pi_{1i}) \pi_{2i|1} + (1 - \pi_{2i|1}))$} \\
+      DC-Est  & 
+      \scriptsize{$N^{-2}\left(\sum\limits_{i \in A_1} \left(\frac{1 -
+        \pi_{1i}}{\pi_{1i}^2}\right) \eta_i^2 + \sum\limits_{i \in A_2}
+          \left(\frac{1}{\pi_{1i}\pi_{2i|1}}\right) (d_{2i|1} - 1) (y_i - q_i
+          z_i \hat \phi)^2\right)$} & 
+      {\scriptsize $\eta_i = q_i z_i \hat \phi +
+          \frac{\delta_{2i|1}}{\pi_{2i|1}}(y_i - q_i z_i \hat \phi)$} \\
     \bottomrule
@@
   \caption{This table gives the formulas for each variance estimator used in
-  this simulation.}
+  Simulation 1. For the derivation of the estimated variance of the DC-Pop
+estimatator see Appendix A.}
   \label{tab:varforms}
-\end{figure}
+\end{table}
 
-We run this simulation $1000$ times for each of these methods and compute the
+\textcolor{red}{Dr. Kim, please see Appendix A for my derivation of the variance
+DC-Pop.}
+
+%    Pi_dc <- 
+%      diag((1 - p2_df$pi1) * p2_df$pi2 + (1 - p2_df$pi2)) * 
+%      (w1i * dc_w / nrow(pop_df))^2
+%    err <- p2_df$Y - t(gam_dc) %*% t(z_dc) * qi
+%    vhat <- (err) %*% Pi_dc %*% t(err)
+
+We run the simulation $1000$ times for each of these methods and compute the
 Bias ($E[\hat Y] - \bar Y_N$), the RMSE ($\sqrt{\Var(\hat Y - \bar Y_N)}$), a 95\%
@@
 \Phi(0.975)\sqrt{\hat V(\hat Y^{(b)})^{(b)}}$), and a T-test that assesses the
-unbiasedness of each estimator. The results are in Figure~\ref{fig:tpdc-mean}.
+unbiasedness of each estimator where $\hat Y^{(b)}$ is the result from the $b$th
+simulation replicate. The results are in Table~\ref{tab:tpdc-mean}.
 
-\begin{figure}[ht!]
+\begin{table}[ht!]
   \centering
@@
 % This table was generated from /src/explore/20240411-tpdcsim.qmd
-\caption{This table shows the results of the simulation study. It displays the
+\caption{This table shows the results of Simulation Study 1. It displays the
 Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
@@
 DC-Est.}
-\label{fig:tpdc-mean}
-\end{figure}
+\label{tab:tpdc-mean}
+\end{table}
 
 \section{Topic 2: Non-nested Two-Phase Sampling}
-
-\textcolor{blue}{
-( Materials in Section 11.4 can be used here. I have copy-and-pasted the
-textbook materials below. Please modify them.  )
-}
-\textcolor{red}{I will do this shortly.}
 
@@
 
-$$\hat X_1 = \sum_{i \in A_1}^{n_1} \frac{x_i}{\pi_{1i}} \text{ and } 
-\hat X_2 = \sum_{i \in A_2}^{n_2} \frac{x_i}{\pi_{2i}}$$
+$$\hat X_1 = \sum_{i \in A_1}^{n_1} d_{1i} x_i \text{ and } 
+\hat X_2 = \sum_{i \in A_2}^{n_2} d_{2i} x_i $$
 
-where $\pi_{1i}$ is the probability of $i \in A_1$ and $\pi_{2i} = \Pr(i \in
-A_2)$. \textcolor{red}{I don't know if it makes sense to change the Phase 2
-selection probability as $\pi_{2i|1}$. Should I keep it at $\pi_{2i}$? But if
-I do, how is the two phase setup useful?} We can combine these estimates to get
+where $d_{1i} = \pi_{1i}^{-1}$, $d_{2i} = \pi_{2i}^{-1}$, $\pi_{1i}$ is the
+probability of $i \in A_1$ and $\pi_{2i} = \Pr(i \in A_2)$. 
+%\textcolor{red}{I don't know if it makes sense to change the Phase 2
+%selection probability as $\pi_{2i|1}$. (\textcolor{blue}{No. We do not know
+%$\pi_{2i |1}$ and it is not necessary.}) Should I keep it at $\pi_{2i}$?
+%\textcolor{blue}{(Yes, $\pi_{2i}$ is available to us.) } But if
+%I do, how is the two phase setup useful?} 
+%\textcolor{blue}{Non-nested two-phase sampling is different from the classical
+%  two-phase sampling. We are not using the two-phase setup of Section 2  here.
+%  We do not have to call it two-phase sampling. However, if we define $A_c=A_1
+%  \cup A_2$ then $A_c$ serves the role of the first-phase sample in the
+%  two-phase sampling. Then, our goal is to compute an estimator of $X$ from
+%  $A_c$. The second-phase sample would be $A_2$. 
+%}
+We can combine these estimates to get
 $\hat X_c = W \hat X_1 + (1 - W)\hat X_2$ for some matrix $W$ (see
-\cite{merkouris2004combining} for the optimal choice of $W$). We can then define
+\cite{merkouris2004combining} for the optimal choice of $W$), and can then define
 a regression estimator as
@@
 
+The methodology for the non-nested two phase sample is very similar to the setup
+described as part of Topic 1. Given a strictly convex differentiable function,
+$G: \mathcal{V} \to \mathbb{R}$, the goal is to solve
+
+\begin{equation}\label{eq:nnopt}
+\hat w_2 = \argmin_w \sum_{i \in A_2} G(w_{2i}) q_i \text{ such that } 
+\sum_{i \in A_2} w_{2i} x_i = \hat X_{c} \text{ and } 
+\sum_{i \in A_2} w_{2i} g(d_{2i}) q_i = \sum_{i \in U} g(d_{2i}) q_i
+\end{equation}
+
+for $g(x) = G'(x)$ and a known choice of $q_i \in \mathbb{R}$. 
+%\textcolor{red}{(Instead of using $\hat{X}_{\rm GLS}$, we used $\hat{X}_c$
+%before. Please make notation consistent. )}
+
+The difference
+between solving Equation~\ref{eq:nnopt} and Equation~\ref{eq:primal} is that the
+estimator $\hat X_{c}$ is estimated from the combined sample $A_c= A_1 \cup A_2$. Before using
+$\hat X_{c}$ in the debiased calibration estimator, we need to estimate it
+from the non-nested samples.
+
+We can get multiple estimates of $\hat X$,
+
+$$\hat X_1 = \sum_{i \in A_1} d_{1i} x_i \text{ and }
+\hat X_2 = \sum_{i \in A_2} d_{2i} x_i.$$
+
+Let $V_1$ and $V_2$ be known variance matrices for $\hat X_1$ and $\hat X_2$,
+then the optimal combined estimate is
+
+$$\hat X_c = \frac{V_1^{-1} \hat X_1 + V_2^{-1} \hat X_2}{V_1^{-1} + V_2^{-1}}.$$
+
+We can constuct a non-nested two phase estimator $\hat Y_{NNE}$ for $Y_N$ where
+$\hat Y_{NNE} = \sum_{i \in A_2} \hat w_{2i} y_i$ and $\hat w_{2i}$ solves
+Equation~\ref{eq:nnopt}. Like the classical two phase approach, to solve this
+setup we minimize the Lagrangian,
+
+\begin{equation}\label{eq:legragedc2}
+  L(w_{2i}, \lambda) = \sum_{i \in A_2} G(w_{2i}) q_i + \lambda 
+  \left( \hat T - \sum_{i \in A_2} w_{2i} z_i q_i\right).
+\end{equation}
+
+with
+
+$$\hat T = 
+\begin{bmatrix}
+  \hat X_c \\ \sum_{i \in U} g(d_{2i}) q_i
+\end{bmatrix}.
+$$
+
+\textcolor{red}{I am a little worried about the second term of $\hat{T}$.
+It is somewhat weird to assume that $T_g= \sum_{i \in U} g(d_{2i}) q_i$ is known.
+One simple way is to use a HT estimator  of $T_g$ from sample $A_2$.
+That is, use $\hat{T}_{g, {\rm HT}} = \sum_{i \in A_2} d_{2i} g( d_{2i}) q_i$ or even use a GREG estimator of $T_g$.
+%I wonder we can assume that an unbiased estimator of $T_g$ is available from $A_c=A_1 \cup A_2$.
+That is, let's assume that we have $\hat{T}_c = W \hat{T}_1 + (1- W) \hat{T}_2$ as the estimator  for $T$.
+}
+
+Differntiating with respect to $w_{2i}$ and setting this expression equal to
+zero, yields the fact that $\hat w_{2i}$ satisfies
+
+$$ \hat w_{2i}(\hat \lambda) = g^{-1}(\hat \lambda^T z_i) $$
+
+where $\hat \lambda$ is the solution to
+
+\begin{equation}\label{eq:lamdc2}
+  \left( \hat T - \sum_{i \in A_2} w_{2i}(\hat \lambda) z_i q_i\right) = 0.
+\end{equation}
+
 \subsection{Theoretical Results}
 
-\subsection{Simulation Studies}
+\begin{theorem}[Design Consistency]\label{thm:dc2}
+  Allowing $\lambda^*$ to be the probability limit of $\hat \lambda$,
+  under some regularity conditions, $\hat Y_{NNE} = \hat Y_{\ell, NNE}(\lambda^*,
+  \phi^*) + O_p(Nn_2^{-1})$ where
+
+  $$\hat Y_{\ell, NNE}(\lambda^*, \phi^*) = \sum_{i \in A_2} \hat w_{2i}(
+  \lambda^*) + \left(\hat T - \sum_{i \in A_2} \hat w_{2i}(\lambda^*) z_i q_i\right)
+  \phi^*$$
+
+  and 
+  $$\phi^* =
+  \left(\sum_{i \in U} \frac{\pi_{2i} q_i}{g'(d_{2i})} 
+  \begin{bmatrix}
+    x_i^2 / q_i & x_i g(d_{2i}) / q_i \\
+    x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \end{bmatrix}
+  \right)^{-1}
+  \sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})} 
+  \begin{bmatrix} x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
+% \textcolor{red}{$T$  is not defined. }
+\end{theorem}
+
+\begin{proof}
+  The proof of this result is very similar to the proof the Theorem~\ref{thm:dc1}.
+  The biggest difference is that the total for $X$ is estimated from both samples
+  using $\hat X_c$ instead of $\hat X_{HT}$ from the Phase 1 sample.
+
+  Since $\hat Y_{NNE} = \sum_{i \in A_2} \hat w_{2i}(\hat \lambda)y_i$ where
+  $\hat \lambda$ solves
+
+  \begin{equation}
+    \sum_{i \in A_2} \hat w_{2i}(\lambda) q_i
+    \underbrace{
+    \begin{bmatrix} 
+      x_i / q_i \\ g(d_i)
+  \end{bmatrix}}_{z_i} = T
+  \end{equation}
+
+  we have
+
+  $$\hat Y_{\ell, NNE} (\hat \lambda, \phi) = \sum_{i \in A_2} \hat w_{2i}(\hat
+  \lambda) + \left(T - \sum_{i \in A_2} \hat w_{2i}(\hat \lambda) z_i q_i\right)
+  \phi.$$
+
+  If we choose $\phi^*$ such that $E\left[\frac{\partial}{\partial \lambda} 
+    \hat Y_{\ell, NNE}(\lambda^*, \phi^*)\right] = 0$, then
+
+  $$\phi^* =
+  \begin{bmatrix}
+    \phi^*_1 \\ \phi^*_2
+  \end{bmatrix} = 
+  \left(\sum_{i \in U} \frac{\pi_{2i} q_i}{g'(d_{2i})} 
+  \begin{bmatrix}
+    x_i^2 / q_i & x_i g(d_{2i}) / q_i \\
+    x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \end{bmatrix}
+  \right)^{-1}
+  \sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})} 
+  \begin{bmatrix} x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
+
+  Hence, by a Taylor expansion around $\hat \lambda$,
+
+  $$\hat Y_{NNE}(\hat \lambda) = \hat Y_{\ell, NNE}(\lambda^*, \phi^*) +
+  O_p(Nn_2^{-1}).$$
+
+\end{proof}
+
+\begin{theorem}[Variance Estimation]
+  The variance of $\hat Y_{NNE}$ is
+
+  \begin{align*}
+    \Var(\hat Y_{NNE}(\hat \lambda)
+    &= (\phi_1^*)^T \Var(\hat X_c) \phi_1^* + 
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
+      (y_i - z_i \phi^* q_i)(y_j - z_j \phi^* q_j)\\
+    &\qquad + (1 - W)\phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij} d_{2i} x_i
+      d_{2j}(y_j - z_j \phi^*_1 q_j)\\
+  \end{align*}
+
+  We can estimate the variance using
+
+  \begin{align*}
+    \hat V_{NNE} 
+    &= (\hat \phi_1)^T \Var(\hat X_c) \hat \phi_1 + 
+    \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
+    (y_i - z_i \hat \phi q_i)(y_j - z_j \hat \phi q_j) \\
+    &\qquad + (1 - W)\hat \phi_1 \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}}
+    \frac{x_i}{\pi_{2i}} \frac{(y_j - z_j \hat \phi_1 q_j)}{\pi_{2j}}
+  \end{align*}
+
+  where
+
+  $$\hat \phi =
+  \begin{bmatrix}
+    \hat \phi_1 \\ \hat \phi_2
+  \end{bmatrix} = 
+  \left(\sum_{i \in A_2} \frac{q_i}{g'(d_{2i})} 
+  \begin{bmatrix}
+    x_i^2 / q_i & x_i g(d_{2i}) / q_i \\
+    x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \end{bmatrix}
+  \right)^{-1}
+  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})} 
+  \begin{bmatrix} x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
+\end{theorem}
+
+\begin{proof}
+  From Theorem~\ref{thm:dc2}, we know that $\hat Y_{NNE}(\hat \lambda) = 
+  \hat Y_{\ell, NNE}(\lambda^*, \phi^*) + O_p(Nn_2^{-1})$. Hence, the variance
+  of $\hat Y_{NNE}(\hat \lambda)$ is
+
+  \begin{align*}
+    \Var(\hat Y_{NNE}(\hat \lambda)) 
+    &= \Var(\hat Y_{\ell, NNE}(\lambda^*, \phi^*) + O_p(Nn_2^{-1})) \\ 
+    &= \Var\left(\sum_{i \in A_2} \hat w_{2i}(\lambda^*) y_i + \left(T - \sum_{i
+    \in A_2} \hat w_{2i}(\lambda^*) z_i q_i\right)\phi^*\right)\\
+    &= (\phi_1^*)^T \Var(\hat X_c) \phi_1^* + 
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} (y_i - z_i \phi^* q_i)(y_j - z_j \phi^* q_j)\\
+    &\qquad+ 2 \Cov\left(\hat X_c \phi_1^*, \sum_{i \in A_2} \frac{(y_i - z_i \phi^*
+    q_i)}{\pi_{2i}}\right) \\
+    &= (\phi_1^*)^T \Var(\hat X_c) \phi_1^* + 
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} (y_i - z_i \phi^* q_i)(y_j - z_j \phi^* q_j)\\
+    &\qquad + (1 - W)\phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij} \frac{x_i}{\pi_{2i}}
+    \frac{(y_j - z_j \phi^*_1 q_j)}{\pi_{2j}}\\
+  \end{align*}
+
+  where the last equality comes from the fact that $\hat X_c = W\hat X_1 + (1 -
+  W) \hat X_2$. To have an unbiased estimator of the variance we can use:
+
+  \begin{align*}
+    \hat V_{NNE} 
+    &= (\hat \phi_1)^T \Var(\hat X_c) \hat \phi_1 + 
+    \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
+    (y_i - z_i \hat \phi q_i)(y_j - z_j \hat \phi q_j) \\
+    &\qquad + (1 - W)\hat \phi_1 \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}}
+    \frac{x_i}{\pi_{2i}} \frac{(y_j - z_j \hat \phi_1 q_j)}{\pi_{2j}}
+  \end{align*}
+
+  where
+
+  $$\hat \phi =
+  \begin{bmatrix}
+    \hat \phi_1 \\ \hat \phi_2
+  \end{bmatrix} = 
+  \left(\sum_{i \in A_2} \frac{q_i}{g'(d_{2i})} 
+  \begin{bmatrix}
+    x_i^2 / q_i & x_i g(d_{2i}) / q_i \\
+    x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \end{bmatrix}
+  \right)^{-1}
+  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})} 
+  \begin{bmatrix} x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
+
+\end{proof}
+
+\subsection{Simulation Study 2}
+
+We run a simulation testing the proposed method. This is very similar to
+Simulation 1. We have the following simulation setup:
+
+$$
+\begin{aligned}
+X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
+X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
+X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
+\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
+Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
+\pi_{1i} &= n_1 / N \\
+\pi_{2i} &= \min(\Phi_3(-x_{1i} + 1), 0.9).
+\end{aligned}
+$$
+
+where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
+%\textcolor{red}{The simulation setup is somewhat strange. The first order
+%inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a
+%function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$.
+%In this case, your second-phase sampling is essentially equivalent to the simple
+%random sampling. }
+We consider a finite population of size $N = 10,000$ with the Phase 1 
+sampling being a simple random sample (SRS) of size $n_1 = 1000$. The Phase 2
+sample is a Poisson sample with an expected sample size of
+
+and
+Phase 2 sampling occuring under Poisson (Bernoulli) sampling. This yields a
+Phase 1 sample
+size of $E[n_1] \approx 1100$ and a Phase 2 sample size of
+$E[n_2] \approx 300$. In the Phase 1 sample, we observe 
+$(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. This
+simulation does not deal with model misspecification, and we compare the
+proposed method for the parameter $\bar Y_N$ with four approaches:
+
+\begin{itemize}
+  \item[1.] HT-estimator: $\hat Y_{HT} = N^{-1} \sum_{i \in A_2}
+    \frac{y_i}{\pi_{2i}},$
+  \item[2.] Regression estimator (Reg): (FIXME)
+  \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
+    solves
+
+  \begin{equation*}
+  \hat w_2 = \argmin_w \sum_{i \in A_2} G(w_{2i}) q_i \text{ with } 
+  \sum_{i \in A_2} w_{2i} x_i = \sum_{i \in U} x_i \text{ and } 
+  \sum_{i \in A_2} w_{2i} g(d_{2i}) q_i = \sum_{i \in U} g(d_{2i}) q_i
+  \end{equation*}
+
+  \item[4.] Debiased Calibration with Estimated Population Constraints (DC-Est):
+    This solves Equation~\eqref{eq:nnopt} with $q_i = 1$.
+\end{itemize}
+
+In addition to estimating the mean parameter $\bar Y_N$, we also construct
+variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$. For each
+approach we give the variance estimate in Table~\ref{tab:varforms}.
 
 \section{Topic 3: Multi-source Two-Phase Sampling}
+
+\newpage
 

### 2024-04-27 src/explore/20240411-tpdcsim.qmd
Updated Topic 1 simulation results. We are using a SRS-Poisson setup.

@@
 
+  # TODO: I am not sure if d_vec is correct
   d_vec <- 1 / (p2_df$pi2 * p2_df$pi1)
@@
 
+  if (any(is.na(dc_w))) {
+    return(list(theta = NA, Var = NA))
+  }
+
   # Mean Estimation
@@
   # Variance Estimation
-  ex_var_mat <- matrix(nrow = 2, ncol = 2)
-  ex_var_mat[1, 1] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X1^2)
-  ex_var_mat[1, 2] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X1 * p1_df$X2)
-  ex_var_mat[2, 1] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X1 * p1_df$X2)
-  ex_var_mat[2, 2] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X2^2)
+  # ex_var_mat <- matrix(nrow = 2, ncol = 2)
+  # ex_var_mat[1, 1] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X1^2)
+  # ex_var_mat[1, 2] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X1 * p1_df$X2)
+  # ex_var_mat[2, 1] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X1 * p1_df$X2)
+  # ex_var_mat[2, 2] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X2^2)
 
-  Pi_dc <- diag(1 - 1 / d_vec) * (dc_w / pop_obs)^2
+  # Pi_dc <- diag(1 - 1 / d_vec) * (dc_w / pop_obs)^2
   z_dc <- model.matrix(~1 + p2_df$X1 + p2_df$X2 + gdi)
-  gam_dc <- 
-    solve(t(z_dc) %*% diag(gpinv) %*% z_dc, t(z_dc) %*% diag(gpinv) %*% p2_df$Y)
-  vhat <- 
-  (p2_df$Y - t(gam_dc) %*% t(z_dc)) %*% Pi_dc %*% t(p2_df$Y - t(gam_dc) %*% t(z_dc)) +
-  (gam_dc[2:3] %*% ex_var_mat %*% gam_dc[2:3] / pop_obs^2)
+  # gam_dc <- 
+  #   solve(t(z_dc) %*% diag(gpinv) %*% z_dc, t(z_dc) %*% diag(gpinv) %*% p2_df$Y)
+  # vhat <- 
+  # (p2_df$Y - t(gam_dc) %*% t(z_dc)) %*% Pi_dc %*% t(p2_df$Y - t(gam_dc) %*% t(z_dc)) +
+  # (gam_dc[2:3] %*% ex_var_mat %*% gam_dc[2:3] / pop_obs^2)
 
-  # FIXME: Change variance estimation to linearization estimate
+  phi_est <- 
+    base::solve(t(z_dc) %*% diag(gpinv) %*% z_dc, t(z_dc) %*% (p2_df$Y * gpinv))
+
+  p12_df <- 
+    left_join(p1_df, p2_df, by = join_by(X1, X2, X3, X4, Y, pi1, pi2, del1)) %>%
+    mutate(del2 = ifelse(is.na(del2), 0, 1))
+
+  pred12 <- 
+  model.matrix(~1 + p12_df$X1 + p12_df$X2 + I(-1/(p12_df$pi2 + p12_df$pi1))) %*%
+    phi_est
+
+  model.matrix(~1 + p12_df$X1 + p12_df$X2 + I(-1/(p12_df$pi2 + p12_df$pi1))) %>%
+    head
+  pred2 <- 
+  model.matrix(~1 + p2_df$X1 + p2_df$X2 + I(-1/(p2_df$pi2 + p2_df$pi1))) %*%
+    phi_est
+  eta <- pred12 + p12_df$del2 / p12_df$pi2 * (p12_df$Y - pred12)
+  v1 <- sum((1 - p12_df$pi1) / p12_df$pi1^2 * eta^2)
+  v2 <- sum(1 / (p2_df$pi1 * p2_df$pi2) * (1 / p2_df$pi2 - 1) * (p2_df$Y - pred2)^2)
+  vhat <- (v1 + v2) / pop_obs^2
 
@@
   tpreg <- tp_reg(p1_df, p2_df, nrow(pop_df))
-  # FIXME: GEcalib is not working. It produces results with very high
-  # values of the function.
   dcpop <- 
@@
 
+mc_res %>%
+  filter(Est == "dcpop") %>%
+  arrange(desc(Theta))
+
 ```

### 2024-04-27 src/explore/20240425-nndcsim.qmd
Updated Topic 1 simulation results. We are using a SRS-Poisson setup.

---
title: "Non-Nested Two Phase Debiased Calibration Simulation"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    code-fold: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

# Summary

This report simulates non-nested two-phase sampling with the debiased
calibration of @kwon2024debiased. We check for the following:

1. Bias of $\hat Y$ and,
2. Bias of $\hat V(\hat Y)$.

For the estimator $\hat Y$ we construct

1. The $\pi^*$-estimator,
2. The two-phase regression estimator,
3. The debiased calibration estimator (with known finite population total), and
4. The debiased calibration estimator with estimated population totals from the
   Phase 1 sample.

# Simulation Setup

```{r}
#| label: setup

B_sims <- 1000

N_obs <- 10000
n1_obs <- 1000
n2_obs <- 500

library(dplyr)
library(GECal)
library(nleqslv)
library(doParallel)
library(doRNG)
library(CVXR)
library(rlang)

# source("GEcalib_code.R")

```

We have the following superpopulation model:

$$
\begin{aligned}
X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
X_{4i} &\stackrel{ind}{\sim} Unif(0.1, 0.9) \\
\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
\pi_{1i} &= \Phi(-x_{3i} - 2)\\
\pi_{2i|1} &= x_{4i}.
\end{aligned}
$$

and the non-nested two-phase sampling scheme is to select a SRS of 
size $n = $`{r} n1_obs`
from a finite population of size $N = $`{r} N_obs` for the Phase 1 sample,
$A_1$, and a Poisson sample with expected size $r =$`{r} n2_obs` for the Phase 2
sample, $A_2$. We only observe $Y$ in the Phase 2 sample, but we observe 
$(X_1, X_2)$ in both $A_1$ and $A_2$.

```{r}
#| label: generate population and samples

gen_pop <- function(obs) {

  x1 <- rnorm(obs, 2, 1)
  x2 <- runif(obs, 0, 4)
  x3 <- rnorm(obs, 0, 1)
  x4 <- runif(obs, 0.1, 0.9)
  eps <- rnorm(obs)

  y <- 3 * x1 + 2 * x2 + eps

  return(tibble(X1 = x1, X2 = x2, X3 = x3, X4 = x4, Y = y))
}

#' This function adds sampling probabilites to a population data frame.
#'
#' @details The phase 2 formula will be included for every element. There is no
#' information about if an element is part of the Phase 1 or Phase 2 sample.
update_pop <- function(pop_df, p1_formula, p2_formula) {

  pi1 <- eval(parse_expr(p1_formula))
  pi2 <- eval(parse_expr(p2_formula))

  pop_df %>%
    mutate(pi1 = pi1, pi2 = pi2)
}

#' This function generates Phase 1 and Phase 2 samples from an updated
#' population with selection probabilites.
gen_samps <- function(upop_df, p1_type, p2_type) {

  if (p1_type == "poisson") {
    del1 <- rbinom(nrow(upop_df), 1, upop_df$pi1)
  } else if (p1_type == "srs") {
    ind <- sample(1:nrow(upop_df), size = upop_df$pi1[1] * nrow(upop_df))
    del1 <- as.numeric(1:nrow(upop_df) %in% ind)
  } else {
    stop("We have only implemented poisson and srs for Phase 1.")
  }

  p1_df <- mutate(upop_df, del1 = del1) %>% filter(del1 == 1)

  if (p2_type == "poisson") {
    del2 <- rbinom(nrow(upop_df), 1, upop_df$pi2)
  } else if (p2_type == "srs") {
    ind <- sample(1:nrow(upop_df), size = round(upop_df$pi2[1] * nrow(upop_df)))
    del2 <- as.numeric(1:nrow(upop_df) %in% ind)
  } else {
    stop("We have only implemented poisson and srs for Phase 2.")
  }

  p2_df <- mutate(upop_df, del2 = del2) %>% filter(del2 == 1)

  return(list(p1_df, p2_df))
}

```

```{r}
#| label: estimators

```

```{r}
#| label: run MCMC

set.seed(1)
pop_df <- gen_pop(N_obs)
true_theta <- mean(pop_df$Y)

upop_df <-
  update_pop(
    pop_df,
    "n1_obs / N_obs",
    "ifelse(pt(-pop_df$X1 + 1, 3) > 0.9, 0.9, pt(-pop_df$X1 + 1, 3))"
  )

gen_samps(upop_df, "srs", "poisson")

```

### 2024-04-27 src/explore/proto-dctp.R

@@ -1,0 +1,364 @@
+# Title: proto-dctp.R
+# Created by: Caleb Leedy
+# Created on: April 24, 2024
+# Purpose: This file contains some of the initial code for an R package to run
+# simulations of the debiased calibration algorithm with two-phase sampling.
+
+# *********
+# * Goals *
+# *********
+# 1. [ ] Create consistent functions to generate two-phase samples.
+# 2. [ ] Be able to run two-phase sampling debiased calibration (estimation and
+# variance) with minimal code repetitions.
+
+# *************
+# * Libraries *
+# *************
+
+library(rlang)
+library(dplyr)
+library(nleqslv)
+
+# ******************************
+# * Generate Two-Phase Samples *
+# ******************************
+# The tricky part is that when we use MCMC, we need to generate additional
+# samples from a fixed population. But, should the sampling probabilities be
+# included in the population data set?
+# Answer: Yes, with the update_pop function we can add the probabilities later.
+
+#' This function generates a population data frame with the requisite variables.
+#'
+#' @param obs - This is the number of observations in the population.
+#'
+#' @details - This function does not include any sampling probabilities because
+#' it does not take into account the sampling type or formula.
+gen_pop <- function(obs) {
+
+  x1 <- rnorm(obs, 2, 1)
+  x2 <- runif(obs, 0, 4)
+  x3 <- rnorm(obs, 0, 1)
+  x4 <- runif(obs, 0.1, 0.9)
+  eps <- rnorm(obs)
+
+  y <- 3 * x1 + 2 * x2 + eps
+
+  return(tibble(X1 = x1, X2 = x2, X3 = x3, X4 = x4, Y = y))
+}
+
+#' This function adds sampling probabilites to a population data frame.
+#'
[…]

### 2024-04-27 src/explore/proto-dctpmcmc.R

@@ -1,0 +1,151 @@
+# Title: proto-dctpmcmc.R
+# Created by: Caleb Leedy
+# Created on: April 24, 2024
+# Purpose: This script runs the code for MCMC analysis of two-phase samples.
+
+# *************
+# * Libraries *
+# *************
+
+library(dplyr)
+library(doParallel)
+library(doRNG)
+
+source("proto-dctp.R")
+
+# **************
+# * Parameters *
+# **************
+
+B_sims <- 1000
+N_obs <- 10000
+n1_obs <- 1000
+
+# ************
+# * Run MCMC *
+# ************
+
+set.seed(1)
+pop_df <- gen_pop(N_obs)
+true_theta <- mean(pop_df$Y)
+
+upop_df <-
+  update_pop(
+    pop_df,
+    "1000 / nrow(pop_df)",
+    "ifelse(pt(-pop_df$X1 + 1, 3) > 0.9, 0.9, pt(-pop_df$X1 + 1, 3))"
+  )
+
+clust <- makeCluster(min(detectCores() - 2, 100), outfile = "")
+registerDoParallel(clust)
+
+mc_res <- 
+  foreach(iter=1:B_sims,
+          .packages = c("dplyr", "nleqslv", "rlang")) %dopar% {
+
+  if (iter %% 100 == 0) {
+    print(paste0("Iter: ", iter))
+  }
+
+  set.seed(iter)
+
+  samps <- gen_samps(upop_df, "srs", "poisson")
+  p1_df <- samps[[1]]
+  p2_df <- samps[[2]]
+
+  pistar <- pi_star(p1_df, p2_df, N_obs)
+  tpreg <- tp_reg(p1_df, p2_df, N_obs)
+  dcpop <- dc_ybar(p1_df, p2_df, upop_df, qi = 1, entropy = "EL", estT1 = FALSE)
+  dcest <- dc_ybar(p1_df, p2_df, upop_df, qi = 1, entropy = "EL", estT1 = TRUE)
+
+  return(
+    tibble(Est = c("pistar", "tpreg", "dcpop", "dcest"),
+           Theta = c(pistar[[1]], tpreg[[1]], dcpop[[1]], dcest[[1]]),
+           Var = c(pistar[[2]], tpreg[[2]], dcpop[[2]], dcest[[2]]),
+           Iter = iter)
+  )
+
+} %>% bind_rows()
[…]

### 2024-04-29 final/reports/prelim_paper.tex
Non-nested simulation working for HT and regression estimators.

@@
 %}
-We can combine these estimates to get
-$\hat X_c = W \hat X_1 + (1 - W)\hat X_2$ for some matrix $W$ (see
-\cite{merkouris2004combining} for the optimal choice of $W$), and can then define
-a regression estimator as
+We can combine these estimates using the effective sample size 
+(\cite{kish1965survey}) to get
+$\hat X_c = (n_{1, eff} \hat X_1 + n_{2, eff}\hat X_2) / (n_{1, eff} + n_{2,
+eff})$ where $n_{1, eff}$ and $n_{2, eff}$ are the effective sample size for
+$A_1$ and $A_2$ respectively. Then we can define a regression estimator as
 
@@
 
-where % \textcolor{red}{(we may need transpose in $W$)}
+where, $W = n_{1, eff} / (n_{1, eff} + n_{2, eff})$, and
 
@@
 $$V(\hat Y_{NN, reg}) = V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i -
-x_i^TW\beta^*_q)\right) + (\beta^*_q)^T W^T V(\hat X_1) W \beta_q^*$$
+x_i^TW\beta^*_q)\right) + (\beta^*_q)^T W V(\hat X_1) W \beta_q^*$$
 

### 2024-04-29 src/explore/20240425-nndcsim.qmd
Non-nested simulation working for HT and regression estimators.

@@
 library(rlang)
-
-# source("GEcalib_code.R")
 
@@
 
+ht_est <- function(p1_df, p2_df, pop_obs, sampling = "srs-poisson") {
+
+  theta <- sum(p2_df$Y / p2_df$pi2) / pop_obs
+
+  if (sampling == "srs-poisson") {
+
+    v_hat <- sum((1 - p2_df$pi2) / p2_df$pi2^2 * p2_df$Y^2) / pop_obs^2
+
+  } else {
+
+    v_hat <- NA
+
+  }
+
+  return(list(theta = theta, v_hat = v_hat))
+}
+
+reg_est <- function(p1_df, p2_df, pop_obs, sampling = "srs-poisson") {
+
+  v_poisson <- sum((1 - p2_df$pi2) / p2_df$pi2^2 * p2_df$Y^2) / pop_obs^2
+  v_srs <- (1 / nrow(p2_df) - 1 / pop_obs) * var(p2_df$Y)
+  d_eff <- v_poisson / v_srs
+  n2_eff <- nrow(p2_df) / d_eff
+  n1_eff <- nrow(p1_df)
+
+  x1_hat <- 
+    c(sum(p1_df$X1 / p1_df$pi1) / pop_obs,
+      sum(p1_df$X2 / p1_df$pi1) / pop_obs)
+
+  x2_hat <- 
+    c(sum(p2_df$X1 / p2_df$pi2) / pop_obs,
+      sum(p2_df$X2 / p2_df$pi2) / pop_obs)
+
+  w <- n1_eff / (n1_eff + n2_eff)
+  xc_hat <- (n1_eff * x1_hat + n2_eff * x2_hat) / (n1_eff + n2_eff)
+
+  # Estimate beta
+  mod <- lm(Y ~ X1 + X2, data = p2_df)
+
+  # Regression Estimation
+  y_ht <- sum(p2_df$Y / p2_df$pi2) / pop_obs
+  theta <- y_ht + sum((xc_hat - x2_hat) * mod$coefficients[2:3])
+
+  # Variance Estimation
+  coefs <- mod$coefficients
+  eps <- (p2_df$Y - coefs[1] * w - coefs[2] * w * p2_df$X1 - coefs[3] * w * p2_df$X2)
+  var_x <- 
+  (1 / nrow(p1_df) - 1 / pop_obs) * var(matrix(c(p1_df$X1, p1_df$X2), ncol = 2))
+
+  v1 <- w^2 * matrix(coefs[2:3], nrow = 1) %*% var_x %*% matrix(coefs[2:3], ncol = 1)
+  v2 <- sum((1 - p2_df$pi2) / p2_df$pi2^2 * eps^2) / pop_obs^2
+
+  cov_xy <- c(sum((1 - p2_df$pi2) / p2_df$pi2^2 * coefs[2] * 
+                 (n2_eff) / (n1_eff + n2_eff) * p2_df$X1 * eps) / pop_obs^2,
+              sum((1 - p2_df$pi2) / p2_df$pi2^2 * coefs[3] * 
+                 (n2_eff) / (n1_eff + n2_eff) * p2_df$X3 * eps) / pop_obs^2)
+  v3 <- sum(cov_xy)
+
+  v_hat <- v1 + v2 + v3
+
+  return(list(theta = theta, v_hat = as.numeric(v_hat)))
+}
+
 ```
@@
 
-gen_samps(upop_df, "srs", "poisson")
+clust <- makeCluster(min(detectCores() - 2, 100), outfile = "")
+registerDoParallel(clust)
+
+mc_res <- 
+  foreach(iter=1:B_sims,
+          .packages = c("dplyr", "nleqslv", "rlang")) %dopar% {
+
+    if (iter %% 100 == 0) {
+      print(paste0("Iter: ", iter))
+    }
+
+    set.seed(iter)
+
+    samps <- gen_samps(upop_df, "srs", "poisson")
+    p1_df <- samps[[1]]
+    p2_df <- samps[[2]]
+
+    ht <- ht_est(p1_df, p2_df, nrow(upop_df))
+    reg <- reg_est(p1_df, p2_df, nrow(upop_df))
+
+    return(
+      tibble(Est = c("HT", "Reg"),
+             Theta = c(ht[[1]], reg[[1]]),
+             Var = c(ht[[2]], reg[[2]]),
+             Iter = iter)
+    )
+  } %>% bind_rows()
+
+stopCluster(clust)
 
 ```
+
+```{r}
+#| label: analyze results
+
+# Analyze Mean
+mc_res %>%
+  mutate(err = Theta - true_theta) %>%
+  mutate(CI = abs(err) < qnorm(0.975) * sqrt(Var)) %>%
+  group_by(Est) %>%
+  summarize(Bias = mean(err, na.rm = TRUE),
+            RMSE = sqrt(mean(err^2, na.rm = TRUE)),
+            EmpCI = mean(CI, na.rm = TRUE),
+            sdest = sd(Theta, na.rm = TRUE)) %>%
+  ungroup() %>%
+  mutate(Ttest = abs(Bias) / sqrt(sdest^2 / B_sims)) %>%
+  select(-sdest)
+
+%>%
+  mutate(Ind = case_when(
+    Est == "dcest" ~ 4,
+    Est == "dcpop" ~ 3,
+    Est == "Reg" ~ 2,
+    Est == "HT" ~ 1)
+  ) %>%
+  arrange(Ind) %>%
+  select(Est, Bias, RMSE, EmpCI, Ttest)
+
+%>%
+  knitr::kable("latex", booktabs = TRUE, escape = FALSE, digits = 3) %>%
+  cat(file = "tables/nndcsim_mean.tex")
+
+# Analyze Variance
+mc_res %>%
+  group_by(Est) %>%
+  summarize(MCVar = var(Theta, na.rm = TRUE),
+            EstVar = mean(Var, na.rm = TRUE),
+            VarVar = var(Var, na.rm = TRUE)) %>%
+  ungroup() %>%
+  mutate(Ttest = abs(MCVar - EstVar) / sqrt(VarVar / B_sims))
+
+%>%
+  mutate(Ind = case_when(
+    Est == "dcest" ~ 4,
+    Est == "dcpop" ~ 3,
+    Est == "Reg" ~ 2,
+    Est == "HT" ~ 1)
+  ) %>%
+  arrange(Ind) %>%
+  select(Est, MCVar, EstVar, VarVar, Ttest)
+
+%>%
+  knitr::kable("latex", booktabs = TRUE, escape = FALSE, digits = 4) %>%
+  cat(file = "tables/nndcsim_var.tex")
+
+```

### 2024-04-29 final/reports/prelim_paper.tex
Added NN simulations. They are working well.

@@
     \frac{y_i}{\pi_{2i}},$
-  \item[2.] Regression estimator (Reg): (FIXME)
+  \item[2.] Regression estimator (Reg): Let $\hat Y_{NN, reg} = \hat Y_{HT} + (\hat
+    X_c - \hat X_{2, HT}) \hat \beta_2$ where $\hat Y_{HT} = \sum_{i \in A_2} d_{2i}
+    y_i$, $\hat X_c = W \hat X_{1, HT} + (1 - W)\hat X_{2, HT}$, $W = n_{1,eff}
+    / (n_{1,eff} + n_{2, eff})$, $\hat X_{1, HT} = \sum_{i \in A_1} d_{1i} x_{i}$,
+    $\hat X_{2, HT} = \sum_{i \in A_2} d_{2i} x_{i}$ and
+
+    $$ \hat \beta_2 = \left(\sum_{i \in A_2} x_i x_i^T\right)^{-1} \sum_{i \in
+    A_2} x_i y_i.$$
+
+    Then $\hat{\bar Y}_{NN, reg} = \hat Y_{NN, reg} / N$.
   \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
@@
 In addition to estimating the mean parameter $\bar Y_N$, we also construct
-variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$. For each
-approach we give the variance estimate in Table~\ref{tab:varforms}.
+variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$.
+% For each approach we give the variance estimate in Table~\ref{tab:varforms2}.
+
+We run the simulation $1000$ times for each of these methods and compute the
+Bias ($E[\hat Y] - \bar Y_N$), the RMSE ($\sqrt{\Var(\hat Y - \bar Y_N)}$), a 95\%
+empirical confidence interval ($\sum_{b = 1}^{1000} |\hat Y^{(b)} - \bar Y_N| \leq 
+\Phi(0.975)\sqrt{\hat V(\hat Y^{(b)})^{(b)}}$), and a T-test that assesses the
+unbiasedness of each estimator where $\hat Y^{(b)}$ is the result from the $b$th
+simulation replicate. The results are in Table~\ref{tab:nndc-mean}.
+
+\begin{table}[ht!]
+  \centering
+\input{tables/nndcsim_mean.tex}
+% This table was generated from ../../src/explore/proto-dctpmcmc.R
+\caption{This table shows the results of Simulation Study 2. It displays the
+Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
+unbiasedness of each estimator for the estimators: HT, Reg, DC-Pop, and
+DC-Est.}
+\label{tab:nndc-mean}
+\end{table}
 

### 2024-04-29 src/explore/20240425-nndcsim.qmd
Added NN simulations. They are working well.

@@
 
+#' This  function does the debiased calibration
+#'
+#' @param zmat - A n x p matrix
+#' @param w1i - A vector of length n
+#' @param d2i - A vector of length n
+#' @param T1 - A vector of length p
+#' @param qi - A scaler or a vector of length n
+#' @param entropy - One of "EL", "ET"
+solveGE <- function(zmat, w1i, d2i, T1, qi, entropy) {
+
+  f <- function(lam, zmat, w1i, d2i, T1, qi, entropy, returnw = FALSE) {
+
+    if (entropy == "EL") {
+      w <- -d2i / drop(zmat %*% lam)
+    } else if (entropy == "ET") {
+      w <- d2i * exp(drop(zmat %*% lam))
+    } else {
+      stop("We only accept entropy of EL or ET.")
+    }
+
+    if (returnw) {
+      return(w)
+    } else {
+      return(drop((w * w1i * qi) %*% zmat) - T1)
+    }
+  }
+
+  j <- function(lam, zmat, w1i, d2i, T1, qi, entropy) {
+    if (entropy == "EL") {
+      res <- t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
+    } else if (entropy == "ET") {
+      res <- t(zmat) %*% diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
+    }
+
+    return(res)
+  }
+
+  init <- c(rep(0, ncol(zmat) - 1), 1)
+  res <- 
+  nleqslv(init, f, jac = j, zmat = zmat, w1i = w1i, d2i = d2i,
+          T1 = T1, qi = qi, entropy = entropy,
+          method = "Newton", control = list(maxit = 1e5, allowSingular = TRUE))
+
+  resw <- f(res$x, zmat, w1i, d2i, T1, qi, entropy, returnw = TRUE)
+
+  if (!(res$termcd %in% c(1))) {
+    return(NA)
+  }
+
+  return(resw)
+
+}
+
+dc_nn_ybar <- 
+  function(p1_df,
+           p2_df,
+           pop_df,
+           qi = 1,
+           entropy = "EL",
+           estT1 = TRUE,
+           sampling = "srs-poisson") {
+
+  d_vec <- 1 / (p2_df$pi2)
+  if (entropy == "EL") {
+    gdi <- -1 / d_vec
+    gdi1 <- -1 / (1 / p1_df$pi2)
+    gdip <- -1 / (1 / pop_df$pi2)
+    gpinv <- d_vec^2
+    gdi12 <- "-1 / d2i"
+  } else if (entropy == "ET") {
+    gdi <- log(d_vec)
+    gdi1 <- log(1 / p1_df$pi2)
+    gdip <- log(1 / pop_df$pi2)
+    gpinv <- d_vec
+    gdi12 <- "log(d2i)"
+  } else {
+    stop("This type of entropy has not yet been implemented.")
+  }
+
+  samp_X <- cbind(rep(1, nrow(p2_df)), as.matrix(select(p2_df, X1, X2)))
+  zmat <- cbind(samp_X, gdi)
+  w1i <- 1
+  d2i <- 1 / p2_df$pi2
+
+  if (estT1) {
+    v_poisson <- sum((1 - p2_df$pi2) / p2_df$pi2^2 * p2_df$Y^2) / nrow(pop_df)^2
+    v_srs <- (1 / nrow(p2_df) - 1 / nrow(pop_df)) * var(p2_df$Y)
+    d_eff <- v_poisson / v_srs
+    n2_eff <- nrow(p2_df) / d_eff
+    n1_eff <- nrow(p1_df)
+
+    x1_hat <- 
+      c(sum(p1_df$X1 * qi / p1_df$pi1),
+        sum(p1_df$X2 * qi / p1_df$pi1))
+
+    x2_hat <- 
+      c(sum(p2_df$X1 * qi / p2_df$pi2),
+        sum(p2_df$X2 * qi / p2_df$pi2))
+
+    w <- n1_eff / (n1_eff + n2_eff)
+    xc_hat <- (n1_eff * x1_hat + n2_eff * x2_hat) / (n1_eff + n2_eff)
+
+    T1 <- c(nrow(pop_df),
+            xc_hat[1],
+            xc_hat[2],
+            sum(gdip * qi))
+  } else {
+    w <- 1
+    T1 <- c(nrow(pop_df),
+            sum(pop_df$X1 * qi),
+            sum(pop_df$X2 * qi),
+            sum(gdip * qi))
+  }
+
+  dc_w <- solveGE(zmat, w1i, d2i, T1, qi, entropy)
+
+  # Mean Estimation
+  theta <- sum(p2_df$Y * dc_w * w1i) / nrow(pop_df)
+
+  z_dc <- model.matrix(~1 + p2_df$X1 + p2_df$X2 + gdi)
+  gam_dc <- 
+    solve(t(z_dc) %*% diag(d2i * gpinv * qi) %*% z_dc,
+          t(z_dc) %*% diag(d2i * gpinv) %*% p2_df$Y)
+
+  if (sampling == "srs-poisson") {
+
+    if (estT1) {
+      # Variance Estimation
+      coefs <- as.numeric(gam_dc)
+      u_mat <- 
+      matrix(c(rep(1, nrow(p2_df)), p2_df$X1 * w, p2_df$X2 * w, gdi * qi), ncol = 4)
+      eps <- (p2_df$Y - drop(u_mat %*% coefs))
+      var_x <- 
+      (1/nrow(p1_df) - 1/nrow(pop_df)) * var(matrix(c(p1_df$X1, p1_df$X2), ncol = 2))
+
+      v1 <- 
+      w^2 * matrix(coefs[2:3], nrow = 1) %*% var_x %*% matrix(coefs[2:3], ncol = 1)
+      v2 <- sum((1 - p2_df$pi2) / p2_df$pi2^2 * eps^2) / nrow(pop_df)^2
+
+      cov_xy <- c(sum((1 - p2_df$pi2) / p2_df$pi2^2 * coefs[2] * 
+                     (n2_eff) / (n1_eff + n2_eff) * p2_df$X1 * eps) / nrow(pop_df)^2,
+                  sum((1 - p2_df$pi2) / p2_df$pi2^2 * coefs[3] * 
+                     (n2_eff) / (n1_eff + n2_eff) * p2_df$X3 * eps) / nrow(pop_df)^2)
+      v3 <- sum(cov_xy)
+
+      v_hat <- v1 + v2 + v3
+
+    } else {
+      coefs <- as.numeric(gam_dc)
+      u_mat <- 
+      matrix(c(rep(1, nrow(p2_df)), p2_df$X1, p2_df$X2, gdi * qi), ncol = 4)
+      eps <- (p2_df$Y - drop(u_mat %*% coefs))
+      v2 <- sum((1 - p2_df$pi2) / p2_df$pi2^2 * eps^2) / nrow(pop_df)^2
+
+      v_hat <- v2
+
+    }
+
+  } else {
+    v_hat <- NA
+  }
+
+  return(list(theta = theta, v_hat = as.numeric(v_hat)))
+}
+
 ```
@@
     reg <- reg_est(p1_df, p2_df, nrow(upop_df))
+    dcpop <- dc_nn_ybar(p1_df, p2_df, upop_df, entropy = "EL", estT1 = FALSE)
+    dcest <- dc_nn_ybar(p1_df, p2_df, upop_df, entropy = "EL", estT1 = TRUE)
 
     return(
-      tibble(Est = c("HT", "Reg"),
-             Theta = c(ht[[1]], reg[[1]]),
-             Var = c(ht[[2]], reg[[2]]),
+      tibble(Est = c("HT", "Reg", "dcpop", "dcest"),
+             Theta = c(ht[[1]], reg[[1]], dcpop[[1]], dcest[[1]]),
+             Var = c(ht[[2]], reg[[2]], dcpop[[2]], dcest[[2]]),
              Iter = iter)
@@
 #| label: analyze results
+
+# Check for failures
+filter(mc_res, is.na(Theta))
 
@@
   mutate(Ttest = abs(Bias) / sqrt(sdest^2 / B_sims)) %>%
-  select(-sdest)
-
-%>%
+  select(-sdest) %>%
   mutate(Ind = case_when(
@@
     Est == "HT" ~ 1)
+  ) %>%
+  mutate(Est = case_when(
+    Est == "dcest" ~ "DC-Est",
+    Est == "dcpop" ~ "DC-Pop",
+    Est == "HT" ~ "HT",
+    Est == "Reg" ~ "Reg")
   ) %>%
@@
   ungroup() %>%
-  mutate(Ttest = abs(MCVar - EstVar) / sqrt(VarVar / B_sims))
-
-%>%
+  mutate(Ttest = abs(MCVar - EstVar) / sqrt(VarVar / B_sims)) %>%
   mutate(Ind = case_when(

### 2024-04-29 src/explore/proto-dctp.R

@@ -277,7 +277,7 @@ dc_ybar <-
     T1 <- c(nrow(pop_df),
             sum(p1_df$X1 * qi / p1_df$pi1),
             sum(p1_df$X2 * qi / p1_df$pi1),
-            sum(gdi1 * qi / p1_df$pi1))
+            sum(gdip * qi)) # NOTE: We do NOT want to estimate gdi
   } else {
     T1 <- c(nrow(pop_df),
             sum(pop_df$X1 * qi),

### 2024-04-30 final/reports/prelim_paper.tex
Changed DC to G(w_i) instead of d_i G(w_i / d_i). It works better.

@@
 
-Consider a finite population of size $N$ containing elements $(X_i, Y_i)$ where
-an initial (Phase 1) sample of size $n_1$ is selected and $X_i$ is observed. Then
+Consider a finite population of size $N$ containing elements $(\bf X_i, Y_i)$ where
+an initial (Phase 1) sample of size $n_1$ is selected and $\bf X_i$ is observed. Then
 from the Phase 1 sample of elements, a (Phase 2) sample of size $n_2 < n_1$ is
@@
 that uses both the observed information in the Phase 2 sample and also the extra
-auxiliary information from $X$ in the Phase 1 sample.
+auxiliary information from $\bf X$ in the Phase 1 sample.
 The challenge is doing this efficiently.
@@
 
-While unbiased (see \cite{kim2024statistics}), the $\pi^*$-estimator
-does not account for the additional information contained in the auxiliary Phase
-1 variable $X$. The two-phase regression estimator $\hat Y_{reg, tp}$ does incorporate 
-information for $X$ by using the estimate $\hat X_1$ from the Phase 1 sample.
-This is how we can leverage the external information $\hat X_1$ to improve the
-initial $\pi^*$-estimator in the second phase sample.
-The two-phase regression estimator has the form,
+While unbiased (see \cite{kim2024statistics}), the $\pi^*$-estimator does not
+account for the additional information contained in the auxiliary Phase 1
+variable $\bf X$. The two-phase regression estimator $\hat Y_{reg, tp}$ does
+incorporate information for $\bf X$ by using the estimate $\hat{\bf X_1}$ from
+the Phase 1 sample. This is how we can leverage the external information 
+$\hat{\bf{X_1}}$ to improve the initial $\pi^*$-estimator in the second phase
+sample. The two-phase regression estimator has the form,
 
 $$ \hat Y_{reg, tp} 
-= \sum_{i \in A_1} \frac{1}{\pi_{1i}} x_i \hat \beta_q+ \sum_{i \in A_2}
-\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_q)$$
+= \sum_{i \in A_1} \frac{1}{\pi_{1i}} \bf x_i \hat{\bm \beta_q} + 
+\sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i \hat{\bm \beta_q})$$
 
-where for $q_i = q(x_i)$ and is a function of $x_i$,
+where for $q_i = q(\bf x_i)$ and is a function of $\bf x_i$,
 $$
-\hat \beta_q = \left(\sum_{i \in A_2} 
-  \frac{x_i x_i'}{\pi_{1i} q_i}\right)^{-1} 
-\sum_{i \in A_2} \frac{x_i y_i}{\pi_{1i} q_i}.$$
+\hat{\bm \beta_q} = \left(\sum_{i \in A_2} 
+  \frac{\bf x_i \bf x_i'}{\pi_{1i} q_i}\right)^{-1} 
+\sum_{i \in A_2} \frac{\bf x_i y_i}{\pi_{1i} q_i}.$$
 
@@
 
-$$\hat w_{2i} = \argmin_{w} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i \text{ such
-that } \sum_{i \in A_2} w_{2i} x_i / \pi_{1i} = \sum_{i \in A_1} x_i / \pi_{1i}.$$
+$$\hat w_{2i} = \argmin_{w} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i
+\text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i / \pi_{1i} = \sum_{i \in A_1}
+\bf x_i / \pi_{1i}.$$
 
@@
 $$ \hat Y_{reg, tp} 
-= \underbrace{\sum_{i \in A_1} \frac{x_i \hat \beta_q}{\pi_{1i}}}_{
+= \underbrace{\sum_{i \in A_1} \frac{\bf x_i \hat{\bm \beta_q}}{\pi_{1i}}}_{
   \text{ Minimizing the variance}} + \underbrace{\sum_{i \in A_2}
-\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_q)}_{
+\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i \hat{\bm \beta_q})}_{
 \text{Bias correction}}.$$
@@
 calibration levels. It does not handle the
-two-phase setup where we need to estimate the finite population total of $x$
+two-phase setup where we need to estimate the finite population total of $\bf x$
 from the Phase 1 sample. In the rest of the section, we extend this method to 
@@
 \begin{equation}\label{eq:calconst1}
-  \sum_{i \in A_2} \frac{x_i w_{2i}q_i}{\pi_{1i}} = 
-\sum_{i \in A_1} \frac{x_iq_i}{\pi_{1i}}
+  \sum_{i \in A_2} \frac{\bf x_i w_{2i}q_i}{\pi_{1i}} = 
+\sum_{i \in A_1} \frac{\bf x_iq_i}{\pi_{1i}}
 \end{equation}
@@
 \begin{equation}\label{eq:primal}
-  \argmin_{w_{2|1}} \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i 
+  \argmin_{w_{2|1}} \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
   \text{ such that}
-  \sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i = \sum_{i \in A_1} w_{1i} z_i q_i
+  \sum_{i \in A_2} w_{1i} w_{2i|1} \bf z_i q_i = \sum_{i \in A_1} w_{1i} \bf z_i q_i
 \end{equation}
 
-where  $z_i = (x_i / q_i, g(\pi_{2i|1}^{-1}))$. 
+where  $\bf z_i = (\bf x_i / q_i, g(\pi_{2i|1}^{-1}))$. 
 Let $\hat w_{2i|1}$ be the solution to Equation~\eqref{eq:primal}. The resulting 
@@
 \begin{equation}\label{eq:legragedc1}
-  L(w_{2i|1}, \lambda) = \sum_{i \in A_2} w_{1i} G(w_{2i|1}) q_i + \lambda 
-  \left( \sum_{i \in A_1} w_{1i} z_i q_i -
-    \sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i\right).
+  L(w_{2i|1}, \bm \lambda) = \sum_{i \in A_2} w_{1i} G(w_{2i|1}) q_i 
+  + \bm \lambda \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
+    \sum_{i \in A_2} w_{1i} w_{2i|1} \bf z_i q_i\right).
 \end{equation}
@@
 
-$$ \hat w_{2i|1}(\hat \lambda) = g^{-1}(\hat \lambda^T z_i) $$
+$$ \hat w_{2i|1}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda}^T \bf z_i) $$
 
-where $\hat \lambda$ is the solution to
+where $\hat{\bm \lambda}$ is the solution to
 
 \begin{equation}\label{eq:lamdc1}
-  \left( \sum_{i \in A_1} w_{1i} z_i q_i -
-    \sum_{i \in A_2} w_{1i} w_{2i|1}(\hat \lambda) z_i q_i\right) = 0.
+  \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
+  \sum_{i \in A_2} w_{1i} w_{2i|1}(\hat{\bm \lambda}) \bf z_i q_i\right) = 0.
 \end{equation}
@@
 \begin{theorem}[Design Consistency]\label{thm:dc1}
-  Let $\lambda^*$ be the probability limit of $\hat \lambda$.
+  Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$.
   Under some regularity conditions,
 
-  $$\hat Y_{DCE} = \hat Y_\ell(\lambda^*, \phi^*) + O_p(N / n_2)$$
+  $$\hat Y_{DCE} = \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + O_p(N / n_2)$$
 
@@
 
-  $$\hat Y_{\ell}(\hat \lambda, \phi^*) = \hat Y_{DCE}(\hat \lambda) + 
-  \left(\sum_{i \in A_1} w_{1i} z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat
-  \lambda) z_i q_i\right)\phi^*$$
+  $$\hat Y_{\ell}(\bm \lambda, \bm \phi^*) = \hat Y_{DCE}(\hat{\bm \lambda}) + 
+  \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(
+  \bm \lambda) z_i q_i\right)\bm \phi^*$$
 
@@
 
-  $$\phi^* = 
-  \left[\sum_{i \in U} \frac{\pi_{2i|1}z_i z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
-  \sum_{i \in U} \frac{\pi_{2i|1}z_i y_i}{g'(d_{2i|1})}.$$
+  $$\bm \phi^* = 
+  \left[\sum_{i \in U} \frac{\pi_{2i|1}\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
+  \sum_{i \in U} \frac{\pi_{2i|1}\bf z_i y_i}{g'(d_{2i|1})}.$$
 
    \begin{eqnarray*}
-   \hat Y_{\ell}( \lambda^*, \phi^*) &=&   \hat{Y}_{\pi^*} + 
-  \left(\sum_{i \in A_1} w_{1i} x_i  -  \sum_{i \in A_2} w_{1i} \pi_{2i \mid
-  1}^{-1} x_i  \right)^T \phi_1^* + \left(\sum_{i \in A_1} w_{1i} g_i   -
+   \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) &=&   \hat{Y}_{\pi^*} + 
+  \left(\sum_{i \in A_1} w_{1i} \bf x_i  -  \sum_{i \in A_2} w_{1i} \pi_{2i \mid
+  1}^{-1} \bf x_i  \right)^T \bm \phi_1^* + \left(\sum_{i \in A_1} w_{1i} g_i  -
   \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^T \phi_2^*  
@@
   and  $$\begin{pmatrix}
-  \phi_1^* \\
+  \bm \phi_1^* \\
   \phi_2^* 
@@
   $$\frac{\partial \mathcal{L}}{\partial w_{2i|1}}: g(w_{2i|1}) w_{1i}q_i -
-    \lambda w_{1i} z_i q_i = 0.$$
+  w_{1i} \bm \lambda \bf z_i q_i = 0.$$
 
-  Hence, $\hat w_{2i}(\lambda) = g^{-1}(\lambda^T z_i)$ and $\hat \lambda$ is
-  determined by Equation~\eqref{eq:lamdc1}. When the sample size gets large, we
-  have $\hat w_{2i|1}(\hat \lambda) \to d_{2i|1}$ which means that $\hat \lambda
-  \to \lambda^*$ where $\lambda^* = (\bf 0, 1)$. Then using the linearization
-  technique of \cite{randles1982asymptotic}, we can construct a regression
-  estimator,
+  Hence, $\hat w_{2i}(\bm \lambda) = g^{-1}(\bm \lambda^T \bf z_i)$ and
+  $\hat{\bm \lambda}$ is determined by Equation~\eqref{eq:lamdc1}. When the
+  sample size gets large, we have $\hat w_{2i|1}(\hat{\bm \lambda}) \to
+  d_{2i|1}$ which means that $\hat{\bm \lambda} \to \bm \lambda^*$ where $\bm
+  \lambda^* = (\bf 0, 1)$. Then using the linearization technique of
+  \cite{randles1982asymptotic}, we can construct a regression estimator,
 
-  $$\hat Y_\ell(\hat \lambda, \phi)  = \hat Y_{DCE}(\hat \lambda) + 
-  \left(\sum_{i \in A_1} w_{1i} z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat
-  \lambda) z_i q_i\right)\phi.$$
+  $$\hat Y_\ell(\hat{\bm \lambda}, \bm \phi)  = \hat Y_{DCE}(\hat{\bm \lambda}) + 
+  \left(\sum_{i \in A_1} w_{1i} \bm z_i q_i - \sum_{i \in A_2} w_{1i} 
+  \hat w_{2i|1}(\hat{\bm \lambda}) \bf z_i q_i\right)\bm \phi.$$
 
-  Notice that $\hat Y_\ell(\hat \lambda, \phi) = \hat Y_{DCE}(\hat \lambda)$ for
-  all $\phi \in \mathbb{R}$. We choose $\phi^*$ such that
+  Notice that $\hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = \hat Y_{DCE}(\hat{\bm
+  \lambda})$ for all $\bm \phi$. We choose $\bm \phi^*$ such that
 
-  $$E\left[\frac{\partial}{\partial \lambda} \hat Y_\ell(\lambda^*, \phi^*)\right]=0.$$
+  $$E\left[\frac{\partial}{\partial \bm \lambda} \hat Y_\ell(\bm \lambda^*,
+  \bm \phi^*)\right]=0.$$
 
-  Using the fact that $g^{-1}(\lambda^* z_i) = g^{-1}(g(d_{2i|1})) = d_{2i|1}$
+  Using the fact that $g^{-1}(\bm \lambda^* \bf z_i) = g^{-1}(g(d_{2i|1})) = d_{2i|1}$
   and $(g^{-1})'(x) = 1 / g'(g^{-1}(x))$, we have
@@
   \begin{align*}
-    \phi^*
-    &= E\left[\sum_{i \in A_2} \frac{w_{1i}z_i z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
-    E\left[\sum_{i \in A_2} \frac{w_{1i}z_i y_i}{g'(d_{2i|1})}\right]\\
-    &= \left[\sum_{i \in U} \frac{\pi_{2i|1}z_i z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
-    \left[\sum_{i \in U} \frac{\pi_{2i|1} z_i y_i}{g'(d_{2i|1})}\right]\\
+    \bm \phi^*
+    &= E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
+    E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i y_i}{g'(d_{2i|1})}\right]\\
+    &= \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
+    \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i y_i}{g'(d_{2i|1})}\right]\\
   \end{align*}
@@
 
-  $$\hat Y_\ell(\lambda^*, \phi^*) = \sum_{i \in A_1} w_{1i} q_i z_i \phi^* +
-  \sum_{i \in A_2} w_{1i} d_{2i|1} (y_i - q_i z_i \phi^*).$$
+  $$\hat Y_\ell(\bm \lambda^*, \bm \phi^*) = 
+  \sum_{i \in A_1} w_{1i} q_i \bf z_i \bm \phi^* +
+  \sum_{i \in A_2} w_{1i} d_{2i|1} (y_i - q_i \bf z_i \bm \phi^*).$$
 
@@
   \begin{align*}
-    \hat Y_{DCE}(\hat \lambda) 
-    &= \hat Y_\ell(\lambda^*, \phi^*) + 
-    E\left[\frac{\partial}{\partial \lambda}\hat Y_\ell(\lambda^*,
-    \phi^*)\right](\hat \lambda - \lambda^*) + \frac{1}{2}
-    E\left[\frac{\partial}{\partial \lambda^2} \hat Y_{DCE}(\lambda^*)\right] (\hat
-    \lambda - \lambda^*)^2\\
-    &= \hat Y_\ell(\lambda^*, \phi^*) + O(N)O_p(n_2^{-1}).
+    \hat Y_{DCE}(\hat{\bm \lambda}) 
+    &= \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + 
+    E\left[\frac{\partial}{\partial \bm \lambda}\hat Y_\ell(\bm \lambda^*,
+    \bm \phi^*)\right](\hat{\bm \lambda} - \bm \lambda^*) + \frac{1}{2}
+    E\left[\frac{\partial}{\partial \bm \lambda^2} \hat Y_{DCE}(\bm \lambda^*)\right] 
+    (\hat{\bm \lambda} - \bm \lambda^*)^2\\
+    &= \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + O(N)O_p(n_2^{-1}).
   \end{align*}
-%\textcolor{red}{(The remainder term shoould be $O_p(N)O_p(n_2^{-1})$.)}
+
   The final equality comes from the fact that 
-  $E\left[\frac{\partial}{\partial \lambda}\hat Y_\ell(\lambda^*, \phi^*)\right]
-  = 0$, $\frac{\partial}{\partial \lambda^2} \hat w_{2i|1}(\lambda^*)$ is
-  bounded and $|\hat \lambda - \lambda^*| = O_p(n_2^{-1/2})$, which proves our
+  $E\left[\frac{\partial}{\partial \bm \lambda}\hat Y_\ell(\bm \lambda^*, \bm \phi^*)\right]
+  = 0$, $\frac{\partial}{\partial \bm \lambda^2} \hat w_{2i|1}(\bm \lambda^*)$ is
+  bounded and $||\hat{\bm \lambda} - \bm \lambda^*|| = O_p(n_2^{-1/2})$, which proves our
   result.
@@
     \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$
-    and $\bf x_i = (x_{1i}, x_{2i})^T$,
+    and $\bf x_i = (1, x_{1i}, x_{2i})^T$,
     % \textcolor{red}{Dr. Kim, should I modify the simulation so that the
     % regression estimator also includes $g(\pi_{2i|1}^{-1})$ as a covariate?}
+
+    \textcolor{red}{You should include an intercept in $\bf x_i$. }
+    \textcolor{blue}{I have been including an intercept, but I forgot to add it
+    into these notes.}
   \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
@@
   \begin{equation}
-    \argmin_{w_{2|1}} \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i})
-    \text{ such that}
-    \sum_{i \in A_2} w_{1i} w_{2i|1} z_i = \sum_{i \in U} z_i.
+    \argmin_{w_{2|1}} \sum_{i \in A_2} w_{1i} G(w_{2i}) \text{ such that}
+    \sum_{i \in A_2} w_{1i} w_{2i|1} \bf z_i = \sum_{i \in U} \bf z_i.
   \end{equation}
@@
         \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
-    & {\scriptsize$ \eta_i = x_i'\beta + \delta_{2i} w_{2i|1}\varepsilon_i,\;
-        \varepsilon_i = (y_i - x_i' \beta)$} \\
+    & {\scriptsize$ \eta_i = \bf x_i'\bm \beta + \delta_{2i} w_{2i|1}\varepsilon_i,\;
+        \varepsilon_i = (y_i - \bf x_i' \bm \beta)$} \\
     DC-Pop  
@@
         \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
-    & {\scriptsize$\eta_i = z_i' \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
-        \varepsilon_i = (y_i - z_i'\phi)$} \\
+    & {\scriptsize$\eta_i = \bf z_i' \bm \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
+        \varepsilon_i = (y_i - \bf z_i'\bm \phi)$} \\
     DC-Est  
@@
         \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
-    & {\scriptsize$\eta_i = z_i' \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
-        \varepsilon_i = (y_i - z_i' \phi)$} \\
+    & {\scriptsize$\eta_i = \bf z_i' \bm \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
+        \varepsilon_i = (y_i - \bf z_i' \bm \phi)$} \\
     \bottomrule
@@
   \label{tab:varforms}
+\end{table}
+
+\textcolor{red}{I am not sure about the variance estimation formula in Table 1. 
+They are different from what we discussed before. I wonder whether you can make
+the sampling rate $n/N$ negligible so that the second term becomes
+insignificant. How about making $N=50,000$, $n_1=500$, and $n_2=200$?}
+
+\textcolor{blue}{The result with $N = 50,000$, $n_1 = 500$ and $E[n_2] \approx
+200$}
+
+\begin{table}[ht!]
+  \centering
+  {\color{blue}\input{tables/tpdcsim50000_mean.tex}}
+% This table was generated from ../../src/explore/proto-dctpmcmc.R
+  \caption{\textcolor{blue}{This table shows the results of Simulation Study 1 with
+      a different finite population and Phase 2 selection probability.
+      It displays the Bias, RMSE, empirical 95\% confidence interval, and a
+      t-statistic assessing the unbiasedness of each estimator for the
+      estimators: $\pi^*$, TP-Reg, DC-Pop, and DC-Est.}}
 \end{table}
@@
 phase sampling, we consider the Phase 1 sample, $A_1$, to consist of
-observations of $(X_i)_{i = 1}^{n_1}$ and the Phase 2 sample, $A_2$, to consist
-of observations of $(X_i, Y_i)_{i = 1}^{n_2}$.
+observations of $(\bf X_i)_{i = 1}^{n_1}$ and the Phase 2 sample, $A_2$, to consist
+of observations of $(\bf X_i, Y_i)_{i = 1}^{n_2}$.
 
@@
 in the non-nested two phase sample we have two independent Horvitz-Thompson
-estimators of the total of $X$,
+estimators of the total of $\bf X$,
 
-$$\hat X_1 = \sum_{i \in A_1}^{n_1} d_{1i} x_i \text{ and } 
-\hat X_2 = \sum_{i \in A_2}^{n_2} d_{2i} x_i $$
+$$\hat{\bf X_1} = \sum_{i \in A_1}^{n_1} d_{1i} \bf x_i \text{ and } 
+\hat{\bf X_2} = \sum_{i \in A_2}^{n_2} d_{2i} \bf x_i $$
 
@@
 probability of $i \in A_1$ and $\pi_{2i} = \Pr(i \in A_2)$. 
-%\textcolor{red}{I don't know if it makes sense to change the Phase 2
-%selection probability as $\pi_{2i|1}$. (\textcolor{blue}{No. We do not know
-%$\pi_{2i |1}$ and it is not necessary.}) Should I keep it at $\pi_{2i}$?
-%\textcolor{blue}{(Yes, $\pi_{2i}$ is available to us.) } But if
-%I do, how is the two phase setup useful?} 
-%\textcolor{blue}{Non-nested two-phase sampling is different from the classical
-%  two-phase sampling. We are not using the two-phase setup of Section 2  here.
-%  We do not have to call it two-phase sampling. However, if we define $A_c=A_1
-%  \cup A_2$ then $A_c$ serves the role of the first-phase sample in the
-%  two-phase sampling. Then, our goal is to compute an estimator of $X$ from
-%  $A_c$. The second-phase sample would be $A_2$. 
-%}
 We can combine these estimates using the effective sample size 
 (\cite{kish1965survey}) to get
-$\hat X_c = (n_{1, eff} \hat X_1 + n_{2, eff}\hat X_2) / (n_{1, eff} + n_{2,
-eff})$ where $n_{1, eff}$ and $n_{2, eff}$ are the effective sample size for
-$A_1$ and $A_2$ respectively. Then we can define a regression estimator as
+$\hat{\bf X_c} = (n_{1, eff} \hat{\bf X_1} + n_{2, eff}\hat{\bf X_2}) / 
+(n_{1, eff} + n_{2, eff})$ where $n_{1, eff}$ and $n_{2, eff}$ are the effective
+sample size for $A_1$ and $A_2$ respectively. Then we can define a regression
+estimator as
 
 $$
-\hat Y_{NN, reg} = \hat Y_2 + (\hat X_c - \hat X_2)^T \hat \beta_q = 
-\hat Y_2 + (\hat X_1 - \hat X_2)^T W^T\hat \beta_q 
+\hat Y_{NN, reg} = \hat Y_2 + (\hat{\bf X_c} - \hat{\bf X_2})^T \hat{\bm \beta_q} = 
+\hat Y_2 + (\hat{\bf X_1} - \hat{\bf X_2})^T W\hat{\bm \beta_q}
 $$
@@
 
-$$\hat \beta_q = \left(\sum_{i \in A_2} \frac{x_i x_i^T}{q_i}\right)^{-1}
-\sum_{i \in A_2} \frac{x_i y_i}{q_i} \text{ and }\hat Y_2 = \sum_{i \in A_2}
-d_{2i} y_i. $$
+$$\hat{\bm \beta_q} = \left(\sum_{i \in A_2} 
+\frac{\bf x_i \bf x_i^T}{q_i}\right)^{-1} 
+\sum_{i \in A_2} \frac{\bf x_i y_i}{q_i} \text{ and
+}\hat Y_2 = \sum_{i \in A_2} d_{2i} y_i. $$
 
@@
 $$V(\hat Y_{NN, reg}) = V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i -
-x_i^TW\beta^*_q)\right) + (\beta^*_q)^T W V(\hat X_1) W \beta_q^*$$
+\bf x_i^TW\bm \beta^*_q)\right) + (\bm \beta^*_q)^T W^T V(\hat{\bf X_1}) W \bm \beta_q^*$$
 
-where $\beta_q^*$ is the probability limit of $\hat \beta_q$. Like the two phase
+where $\bm \beta_q^*$ is the probability limit of $\hat{\bm \beta_q}$. Like the two phase
 sample this regression estimator can be viewed as the solution to the following
@@
   \hat w_2 = \argmin_{w_2} Q(w_2) = \sum_{i \in A_2} (w_{2i} - d_{2i})^2 q_i 
-  \text{ such that } \sum_{i \in A_2} w_{2i} x_i = \hat X_c
+  \text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X_c}
 \end{equation}
@@
 We extend the debiased calibration estimator of \cite{kwon2024debiased} to the
-non-nested two phase sampling case where we use a combined estimate $\hat X_c$
-as the calibration totals instead of using the true totals from the finite
-population.
+non-nested two phase sampling case where we use a combined estimate 
+$\hat{\bm X_c}$ as the calibration totals instead of using the true totals from
+the finite population.
 
@@
 \begin{equation}\label{eq:nnopt}
-\hat w_2 = \argmin_w \sum_{i \in A_2} G(w_{2i}) q_i \text{ such that } 
-\sum_{i \in A_2} w_{2i} x_i = \hat X_{c} \text{ and } 
+\hat w_2 = \argmin_w \sum_{i \in A_2} G\left(w_{2i}\right) q_i 
+\text{ with } 
+\sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X_{c}} \text{ and } 
 \sum_{i \in A_2} w_{2i} g(d_{2i}) q_i = \sum_{i \in U} g(d_{2i}) q_i
@@
 for $g(x) = G'(x)$ and a known choice of $q_i \in \mathbb{R}$. 
-%\textcolor{red}{(Instead of using $\hat{X}_{\rm GLS}$, we used $\hat{X}_c$
-%before. Please make notation consistent. )}
-
 The difference
 between solving Equation~\ref{eq:nnopt} and Equation~\ref{eq:primal} is that the
-estimator $\hat X_{c}$ is estimated from the combined sample $A_c= A_1 \cup
-A_2$. Before using
-$\hat X_{c}$ in the debiased calibration estimator, we need to estimate it
-from the non-nested samples.
+estimator $\hat{\bf X_c}$ is estimated from the combined sample $A_c= A_1 \cup
+A_2$. Before using $\hat{\bf X_c}$ in the debiased calibration estimator, we
+need to estimate it from the non-nested samples. We can get multiple estimates
+of $\hat{\bf X}$,
 
-We can get multiple estimates of $\hat X$,
+$$\hat{\bf X_1} = \sum_{i \in A_1} d_{1i} \bf x_i \text{ and }
+\hat{\bf X_2} = \sum_{i \in A_2} d_{2i} \bf x_i.$$
 
-$$\hat X_1 = \sum_{i \in A_1} d_{1i} x_i \text{ and }
-\hat X_2 = \sum_{i \in A_2} d_{2i} x_i.$$
+Let $V_1$ and $V_2$ be known variance matrices for $\hat{\bf X_1}$ and 
+$\hat{\bf X_2}$, then the optimal combined estimate is
 
-Let $V_1$ and $V_2$ be known variance matrices for $\hat X_1$ and $\hat X_2$,
-then the optimal combined estimate is
-
-$$\hat X_c = \frac{V_1^{-1} \hat X_1 + V_2^{-1} \hat X_2}{V_1^{-1} + V_2^{-1}}.$$
+$$\hat{\bf X_c} = \frac{V_1^{-1} \hat{\bf X_1} + V_2^{-1} 
+\hat{\bf X_2}}{V_1^{-1} + V_2^{-1}}.$$
 
@@
 \begin{equation}\label{eq:legragedc2}
-  L(w_{2i}, \lambda) = \sum_{i \in A_2} G(w_{2i}) q_i + \lambda 
-  \left( \hat T - \sum_{i \in A_2} w_{2i} z_i q_i\right).
+  L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} G(w_{2i}) q_i + 
+  \bm \lambda \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i} \bf z_i q_i\right).
 \end{equation}
@@
 
-$$\hat T = 
+$$\hat{\bf T} = 
 \begin{bmatrix}
-  \hat X_c \\ \sum_{i \in U} g(d_{2i}) q_i
+  \hat{\bf X_c} \\ \sum_{i \in U} g(d_{2i}) q_i
 \end{bmatrix}.
@@
 
-$$ \hat w_{2i}(\hat \lambda) = g^{-1}(\hat \lambda^T z_i) $$
+$$ \hat w_{2i}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda^T} \bf z_i) $$
 
-where $\hat \lambda$ is the solution to
+where $\hat{\bm \lambda}$ is the solution to
 
 \begin{equation}\label{eq:lamdc2}
-  \left( \hat T - \sum_{i \in A_2} w_{2i}(\hat \lambda) z_i q_i\right) = 0.
+  \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i}(\hat{\bm \lambda}) \bf z_i
+  q_i\right) = 0.
 \end{equation}
@@
 \begin{theorem}[Design Consistency]\label{thm:dc2}
-  Allowing $\lambda^*$ to be the probability limit of $\hat \lambda$,
-  under some regularity conditions, $\hat Y_{NNE} = \hat Y_{\ell, NNE}(\lambda^*,
-  \phi^*) + O_p(Nn_2^{-1})$ where
+  Allowing $\bm \lambda^*$ to be the probability limit of $\hat{\bm \lambda}$,
+  under some regularity conditions, $\hat Y_{NNE} = \hat Y_{\ell, NNE}(\bm \lambda^*,
+  \bm \phi^*) + O_p(Nn_2^{-1})$ where
 
-  $$\hat Y_{\ell, NNE}(\lambda^*, \phi^*) = \sum_{i \in A_2} \hat w_{2i}(
-  \lambda^*) + \left(\hat T - \sum_{i \in A_2} \hat w_{2i}(\lambda^*) z_i q_i\right)
-  \phi^*$$
+  $$\hat Y_{\ell, NNE}(\bm \lambda^*, \bm \phi^*) = \sum_{i \in A_2} \hat w_{2i}(
+  \bm \lambda^*) + \left(\hat{\bf T} - \sum_{i \in A_2} \hat w_{2i}(
+  \bm \lambda^*) \bf z_i q_i\right) \bm \phi^*$$
 
   and 
-  $$\phi^* =
+  $$\bm \phi^* =
   \left(\sum_{i \in U} \frac{\pi_{2i} q_i}{g'(d_{2i})} 
   \begin{bmatrix}
-    x_i^2 / q_i & x_i g(d_{2i}) / q_i \\
-    x_i g(d_{2i}) / q_i & g(d_{2i})^2
+    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
+    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
   \end{bmatrix}
@@
   \sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})} 
-  \begin{bmatrix} x_i / q_i \\ g(d_i) \end{bmatrix}.
+  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
   $$
-% \textcolor{red}{$T$  is not defined. }
 \end{theorem}
@@
 \begin{proof}
-  The proof of this result is very similar to the proof the Theorem~\ref{thm:dc1}.
-  The biggest difference is that the total for $X$ is estimated from both samples
-  using $\hat X_c$ instead of $\hat X_{HT}$ from the Phase 1 sample.
+  The proof of this result is very similar to the proof the
+  Theorem~\ref{thm:dc1}. The biggest difference is that the total for $\bf X$ is
+  estimated from both samples using $\hat{\bf X_c}$ instead of $\hat{\bf
+  X_{HT}}$ from the Phase 1 sample.
 
-  Since $\hat Y_{NNE} = \sum_{i \in A_2} \hat w_{2i}(\hat \lambda)y_i$ where
-  $\hat \lambda$ solves
+  Since $\hat Y_{NNE} = \sum_{i \in A_2} \hat w_{2i}(\hat{\bm \lambda})y_i$ where
+  $\hat{\bm \lambda}$ solves
 
   \begin{equation}
-    \sum_{i \in A_2} \hat w_{2i}(\lambda) q_i
+    \sum_{i \in A_2} \hat w_{2i}(\bm \lambda) q_i
     \underbrace{
     \begin{bmatrix} 
-      x_i / q_i \\ g(d_i)
-  \end{bmatrix}}_{z_i} = T
+      \bf x_i / q_i \\ g(d_i)
+  \end{bmatrix}}_{\bf z_i} = \bf T
   \end{equation}
@@
 
-  $$\hat Y_{\ell, NNE} (\hat \lambda, \phi) = \sum_{i \in A_2} \hat w_{2i}(\hat
-  \lambda) + \left(T - \sum_{i \in A_2} \hat w_{2i}(\hat \lambda) z_i q_i\right)
-  \phi.$$
+  $$\hat Y_{\ell, NNE} (\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} \hat
+  w_{2i}(\hat{\bm \lambda}) + \left(\bf T - \sum_{i \in A_2} \hat
+  w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right) \bm \phi.$$
 
-  If we choose $\phi^*$ such that $E\left[\frac{\partial}{\partial \lambda} 
-    \hat Y_{\ell, NNE}(\lambda^*, \phi^*)\right] = 0$, then
+  If we choose $\bm \phi^*$ such that $E\left[\frac{\partial}{\partial \bm \lambda} 
+    \hat Y_{\ell, NNE}(\bm \lambda^*, \bm \phi^*)\right] = 0$, then
 
-  $$\phi^* =
+  $$\bm \phi^* =
   \begin{bmatrix}
-    \phi^*_1 \\ \phi^*_2
+    \bm \phi^*_1 \\ \bm \phi^*_2
   \end{bmatrix} = 
@@
   \begin{bmatrix}
-    x_i^2 / q_i & x_i g(d_{2i}) / q_i \\
-    x_i g(d_{2i}) / q_i & g(d_{2i})^2
+    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
+    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
   \end{bmatrix}
@@
   \sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})} 
-  \begin{bmatrix} x_i / q_i \\ g(d_i) \end{bmatrix}.
+  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
   $$
 
-  Hence, by a Taylor expansion around $\hat \lambda$,
+  Hence, by a Taylor expansion around $\hat{\bm \lambda}$,
 
-  $$\hat Y_{NNE}(\hat \lambda) = \hat Y_{\ell, NNE}(\lambda^*, \phi^*) +
-  O_p(Nn_2^{-1}).$$
+  $$\hat Y_{NNE}(\hat{\bm \lambda}) = \hat Y_{\ell, NNE}(\bm \lambda^*, 
+  \bm \phi^*) + O_p(Nn_2^{-1}).$$
 
@@
     \Var(\hat Y_{NNE}(\hat \lambda)
-    &= (\phi_1^*)^T \Var(\hat X_c) \phi_1^* + 
+    &= (\bm \phi_1^*)^T \Var(\hat{\bf X_c}) \bm \phi_1^* + 
     \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
-      (y_i - z_i \phi^* q_i)(y_j - z_j \phi^* q_j)\\
-    &\qquad + (1 - W)\phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij} d_{2i} x_i
-      d_{2j}(y_j - z_j \phi^*_1 q_j)\\
+      (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
+    &\qquad + (1 - W)\bm \phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij}
+    d_{2i} \bf x_i d_{2j}(y_j - \bf z_j \bm \phi^*_1 q_j)\\
   \end{align*}
@@
     \hat V_{NNE} 
-    &= (\hat \phi_1)^T \Var(\hat X_c) \hat \phi_1 + 
+    &= (\hat{\bm \phi_1})^T \Var(\hat{\bf X_c}) \hat{\bm \phi_1} + 
     \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
-    (y_i - z_i \hat \phi q_i)(y_j - z_j \hat \phi q_j) \\
-    &\qquad + (1 - W)\hat \phi_1 \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}}
-    \frac{x_i}{\pi_{2i}} \frac{(y_j - z_j \hat \phi_1 q_j)}{\pi_{2j}}
+    (y_i - \bf z_i \hat{\bm \phi} q_i)(y_j - \bf z_j \hat{\bm \phi} q_j) \\
+    &\qquad + (1 - W)\hat{\bm \phi_1} \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}}
+    \frac{\bf x_i}{\pi_{2i}} \frac{(y_j - \bf z_j \hat{\bm \phi_1} q_j)}{\pi_{2j}}
   \end{align*}
@@
 
-  $$\hat \phi =
+  $$\hat{\bm \phi} =
   \begin{bmatrix}
-    \hat \phi_1 \\ \hat \phi_2
+    \hat{\bm \phi_1} \\ \hat \phi_2
   \end{bmatrix} = 
@@
   \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})} 
-  \begin{bmatrix} x_i / q_i \\ g(d_i) \end{bmatrix}.
+  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
   $$
@@
   From Theorem~\ref{thm:dc2}, we know that $\hat Y_{NNE}(\hat \lambda) = 
-  \hat Y_{\ell, NNE}(\lambda^*, \phi^*) + O_p(Nn_2^{-1})$. Hence, the variance
-  of $\hat Y_{NNE}(\hat \lambda)$ is
+  \hat Y_{\ell, NNE}(\bm \lambda^*, \bm \phi^*) + O_p(Nn_2^{-1})$. Hence, the
+  variance of $\hat Y_{NNE}(\hat{\bm \lambda})$ is
 
   \begin{align*}
-    \Var(\hat Y_{NNE}(\hat \lambda)) 
-    &= \Var(\hat Y_{\ell, NNE}(\lambda^*, \phi^*) + O_p(Nn_2^{-1})) \\ 
-    &= \Var\left(\sum_{i \in A_2} \hat w_{2i}(\lambda^*) y_i + \left(T - \sum_{i
-    \in A_2} \hat w_{2i}(\lambda^*) z_i q_i\right)\phi^*\right)\\
-    &= (\phi_1^*)^T \Var(\hat X_c) \phi_1^* + 
-    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} (y_i - z_i \phi^* q_i)(y_j - z_j \phi^* q_j)\\
-    &\qquad+ 2 \Cov\left(\hat X_c \phi_1^*, \sum_{i \in A_2} \frac{(y_i - z_i \phi^*
-    q_i)}{\pi_{2i}}\right) \\
-    &= (\phi_1^*)^T \Var(\hat X_c) \phi_1^* + 
-    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} (y_i - z_i \phi^* q_i)(y_j - z_j \phi^* q_j)\\
-    &\qquad + (1 - W)\phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij} \frac{x_i}{\pi_{2i}}
-    \frac{(y_j - z_j \phi^*_1 q_j)}{\pi_{2j}}\\
+    \Var(\hat Y_{NNE}(\hat{\bm \lambda})) 
+    &= \Var(\hat Y_{\ell, NNE}(\bm \lambda^*, \bm \phi^*) + O_p(Nn_2^{-1})) \\ 
+    &= \Var\left(\sum_{i \in A_2} \hat w_{2i}(\bm \lambda^*) y_i + 
+      \left(\bf T - \sum_{i \in A_2} \hat w_{2i}(\bm \lambda^*) \bf z_i
+    q_i\right)\bm \phi^*\right)\\
+    &= (\bm \phi_1^*)^T \Var(\hat{\bf X_c}) \bm \phi_1^* + 
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
+    (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
+    &\qquad+ 2 \Cov\left(\hat{\bf X_c} \bm \phi_1^*, \sum_{i \in A_2} 
+      \frac{(y_i - \bf z_i \bm \phi^* q_i)}{\pi_{2i}}\right) \\
+    &= (\bm \phi_1^*)^T \Var(\hat{\bf X_c}) \bm \phi_1^* + 
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
+    (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
+    &\qquad + (1 - W)\bm \phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij} 
+    \frac{x_i}{\pi_{2i}} \frac{(y_j - \bf z_j \bm \phi^*_1 q_j)}{\pi_{2j}}\\
   \end{align*}
 
-  where the last equality comes from the fact that $\hat X_c = W\hat X_1 + (1 -
-  W) \hat X_2$. To have an unbiased estimator of the variance we can use:
+  where the last equality comes from the fact that $\hat{\bf X_c} = W\hat{\bf
+  X_1} + (1 - W) \hat{\bf X_2}$. To have an unbiased estimator of the variance
+  we can use:
 
@@
     \hat V_{NNE} 
-    &= (\hat \phi_1)^T \Var(\hat X_c) \hat \phi_1 + 
+    &= (\hat{\bm \phi_1})^T \Var(\hat{\bf X_c}) \hat{\bm \phi_1} + 
     \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
-    (y_i - z_i \hat \phi q_i)(y_j - z_j \hat \phi q_j) \\
-    &\qquad + (1 - W)\hat \phi_1 \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}}
-    \frac{x_i}{\pi_{2i}} \frac{(y_j - z_j \hat \phi_1 q_j)}{\pi_{2j}}
+    (y_i - \bf z_i \hat{\bm \phi} q_i)(y_j - \bf z_j \hat{\bm \phi} q_j) \\
+    &\qquad + (1 - W)\hat \phi_1 \sum_{i \in A_2} \sum_{j \in A_2} 
+    \frac{\Delta_{2ij}}{\pi_{2ij}} \frac{x_i}{\pi_{2i}} 
+    \frac{(y_j - \bf z_j \hat{\bm \phi_1} q_j)}{\pi_{2j}}
   \end{align*}
@@
     \frac{y_i}{\pi_{2i}},$
-  \item[2.] Regression estimator (Reg): Let $\hat Y_{NN, reg} = \hat Y_{HT} + (\hat
-    X_c - \hat X_{2, HT}) \hat \beta_2$ where $\hat Y_{HT} = \sum_{i \in A_2} d_{2i}
-    y_i$, $\hat X_c = W \hat X_{1, HT} + (1 - W)\hat X_{2, HT}$, $W = n_{1,eff}
-    / (n_{1,eff} + n_{2, eff})$, $\hat X_{1, HT} = \sum_{i \in A_1} d_{1i} x_{i}$,
-    $\hat X_{2, HT} = \sum_{i \in A_2} d_{2i} x_{i}$ and
+  \item[2.] Regression estimator (Reg): Let $\hat Y_{NN, reg} = \hat Y_{HT} + (
+    \hat{\bf X_c} - \hat{\bf X_{2, HT}}) \hat{\bm \beta_2}$ where $\hat Y_{HT} =
+    \sum_{i \in A_2} d_{2i} y_i$, $\hat{\bf X_c} = W \hat{\bf X_{1, HT}} + (1 -
+    W)\hat{\bf X_{2, HT}}$, $W = n_{1,eff} / (n_{1,eff} + n_{2, eff})$,
+    $\hat{\bf X_{1, HT}} = \sum_{i \in A_1} d_{1i} \bf x_{i}$, $\hat{\bf X_{2,
+    HT}} = \sum_{i \in A_2} d_{2i} \bf x_{i}$, $\bf x_i = (1, x_{1i}, x_{2i})^T$
+    and
 
-    $$ \hat \beta_2 = \left(\sum_{i \in A_2} x_i x_i^T\right)^{-1} \sum_{i \in
-    A_2} x_i y_i.$$
+    $$ \hat{\bm \beta_2} = \left(\sum_{i \in A_2} \bf x_i \bf x_i^T\right)^{-1}
+    \sum_{i \in A_2} \bf x_i y_i.$$
 
@@
   \hat w_2 = \argmin_w \sum_{i \in A_2} G(w_{2i}) q_i \text{ with } 
-  \sum_{i \in A_2} w_{2i} x_i = \sum_{i \in U} x_i \text{ and } 
+  \sum_{i \in A_2} w_{2i} \bf x_i = \sum_{i \in U} \bf x_i \text{ and } 
   \sum_{i \in A_2} w_{2i} g(d_{2i}) q_i = \sum_{i \in U} g(d_{2i}) q_i

### 2024-04-30 final/reports/prelim_paper_ov.tex
Changed DC to G(w_i) instead of d_i G(w_i / d_i). It works better.

@@
 
-  $$\hat Y_{\ell}(\hat \lambda, \phi^*) = \hat Y_{DCE}(\hat \lambda) + 
-  \left(\sum_{i \in A_1} w_{1i} z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat
+  $$\hat Y_{\ell}(\lambda, \phi^*) = \hat Y_{DCE}(\hat \lambda) + 
+  \left(\sum_{i \in A_1} w_{1i} z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(
   \lambda) z_i q_i\right)\phi^*$$
@@
 Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
-\pi_{1i} &= \min(\Phi_3(-x_{3i} - 2), 0.7) \\
+%\pi_{1i} &= \min(\Phi_3(-x_{3i} - 2), 0.7) \\
+\pi_{1i} &= n_1 / N \\
 \pi_{2i|1} &= \min(\Phi_3(-x_{1i} + 1), 0.9).
@@
 This is a two-phase extension of the setup in \cite{kwon2024debiased}. We
-consider a finite population of size $N = 10,000$ with both the Phase 1 and
-Phase 2 sampling occuring under Poisson (Bernoulli) sampling. This yields a
+consider a finite population of size $N = 10,000$ with the Phase 1 sampling
+being a simple random sample (SRS) and the 
+Phase 2 sampling occuring under Poisson sampling. This yields a
 Phase 1 sample
-size of $E[n_1] \approx 1100$ and a Phase 2 sample size of
-$E[n_2] \approx 300$. In the Phase 1 sample, we observe 
+size of $n_1 = 1000$ and a Phase 2 sample size of
+$E[n_2] \approx 267$. In the Phase 1 sample, we observe 
 $(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. This
@@
     % regression estimator also includes $g(\pi_{2i|1}^{-1})$ as a covariate?}
+
+    \textcolor{red}{You should include an intercept in $\bf x_i$. }
   \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
@@
     \midrule
-    $\pi^*$ & $N^{-2} \sum_{i \in A_2} \left(\pi_{2i|1}^{-2} - \pi_{2i|1}^{-1}\right)
-    y_i^2 $ & \\
-    TP-Reg  & 
-    {\scriptsize $N^{-2}\left(\sum\limits_{i \in A_1} 
-        \left(\frac{1 - \pi_{1i}}{\pi_{1i}^{2}}\right) \eta_i^2 
-        + \sum\limits_{i \in A_2} \frac{1}{\pi_{1i} \pi_{2i|1}}(d_{2i|1} - 1)
-        (y_i - \bf x_i' \hat \beta)^2 \right)$}
-    & 
-      {\scriptsize$\eta_i = \bf x_i \hat \beta + \frac{\delta_{2i}}{\pi_{2i|1}}(y_i - 
-      \bf x_i \hat \beta)$} \\
-      %$\hat \beta = \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$ \\
-    DC-Pop  & {\scriptsize$\sum\limits_{i \in A_2} c_i (y_i - q_i z_i \hat \gamma)^2 
-              (w_{1i} \hat w_{2i|1}(\hat \lambda) / N)^2$}
-            & {\scriptsize$c_i = ((1 - \pi_{1i}) \pi_{2i|1} + (1 - \pi_{2i|1}))$} \\
-      DC-Est  & 
-      \scriptsize{$N^{-2}\left(\sum\limits_{i \in A_1} \left(\frac{1 -
-        \pi_{1i}}{\pi_{1i}^2}\right) \eta_i^2 + \sum\limits_{i \in A_2}
-          \left(\frac{1}{\pi_{1i}\pi_{2i|1}}\right) (d_{2i|1} - 1) (y_i - q_i
-          z_i \hat \phi)^2\right)$} & 
-      {\scriptsize $\eta_i = q_i z_i \hat \phi +
-          \frac{\delta_{2i|1}}{\pi_{2i|1}}(y_i - q_i z_i \hat \phi)$} \\
+    $\pi^*$ 
+    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_Y + 
+      \sum\limits_{i \in A_2} \frac{w_{1i}^2}{N^2} 
+        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}y_i^2$}
+    & {\scriptsize$\hat s^2_Y = \frac{1}{n_2 - 1}\sum\limits_{i \in A_2} (y_i -
+    \bar y_i)^2$} \\
+    TP-Reg  
+    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\eta + 
+      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
+        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
+    & {\scriptsize$ \eta_i = x_i'\beta + \delta_{2i} w_{2i|1}\varepsilon_i,\;
+        \varepsilon_i = (y_i - x_i' \beta)$} \\
+    DC-Pop  
+    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\varepsilon + 
+      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
+        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
+    & {\scriptsize$\eta_i = z_i' \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
+        \varepsilon_i = (y_i - z_i'\phi)$} \\
+    DC-Est  
+    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\eta + 
+      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
+        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
+    & {\scriptsize$\eta_i = z_i' \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
+        \varepsilon_i = (y_i - z_i' \phi)$} \\
     \bottomrule
@@
 
-\textcolor{red}{Dr. Kim, please see Appendix A for my derivation of the variance
-DC-Pop.}
-
-%    Pi_dc <- 
-%      diag((1 - p2_df$pi1) * p2_df$pi2 + (1 - p2_df$pi2)) * 
-%      (w1i * dc_w / nrow(pop_df))^2
-%    err <- p2_df$Y - t(gam_dc) %*% t(z_dc) * qi
-%    vhat <- (err) %*% Pi_dc %*% t(err)
+%\textcolor{red}{I am not sure about the variance estimation formula in Table 1.
+They are different from what we discussed before.
+I wonder whether you can make the sampling rate $n/N$ negligible so that the second term becomes insignificant.
+How about making $N=50,000$, $n_1=500$, and $n_2=200$?}
 
@@
 \input{tables/tpdcsim_mean.tex}
-% This table was generated from /src/explore/20240411-tpdcsim.qmd
+% This table was generated from ../../src/explore/proto-dctpmcmc.R
 \caption{This table shows the results of Simulation Study 1. It displays the
@@
 \label{tab:tpdc-mean}
+\end{table}
+
+\textcolor{red}{Dr. Kim is this confidence interval good enough? It seems
+slightly high for me.}
+
+\textcolor{red}{I also give the mean variance estimates, the Monte Carlo mean,
+as well as a test for bias.}
+\begin{table}[ht!]
+  \centering
+{\color{red} \input{tables/tpdcsim_var.tex} }
+% This table was generated from ../../src/explore/proto-dctpmcmc.R
+\caption{\textcolor{red}{This table shows the variance estimates from Simulation
+    Study 1. It displays the Monte Carlo variance of $\hat Y^{(b)}$, the average
+    estimated variance, the variance of the variance estimator, and a
+    t-statistic assessing the unbiasedness of each variance estimator.}}
+\label{tab:tpdc-var}
 \end{table}
@@
 between solving Equation~\ref{eq:nnopt} and Equation~\ref{eq:primal} is that the
-estimator $\hat X_{c}$ is estimated from the combined sample $A_c= A_1 \cup A_2$. Before using
+estimator $\hat X_{c}$ is estimated from the combined sample $A_c= A_1 \cup
+A_2$. Before using
 $\hat X_{c}$ in the debiased calibration estimator, we need to estimate it
@@
 
-\textcolor{red}{I am a little worried about the second term of $\hat{T}$.
-It is somewhat weird to assume that $T_g= \sum_{i \in U} g(d_{2i}) q_i$ is known.
-One simple way is to use a HT estimator  of $T_g$ from sample $A_2$.
-That is, use $\hat{T}_{g, {\rm HT}} = \sum_{i \in A_2} d_{2i} g( d_{2i}) q_i$ or even use a GREG estimator of $T_g$.
-%I wonder we can assume that an unbiased estimator of $T_g$ is available from $A_c=A_1 \cup A_2$.
-That is, let's assume that we have $\hat{T}_c = W \hat{T}_1 + (1- W) \hat{T}_2$ as the estimator  for $T$.
+\textcolor{red}{I am a little worried about the second term of $\hat{T}$. It is
+  somewhat weird to assume that $T_g= \sum_{i \in U} g(d_{2i}) q_i$ is known.
+  One simple way is to use a HT estimator  of $T_g$ from sample $A_2$. That is,
+  use $\hat{T}_{g, {\rm HT}} = \sum_{i \in A_2} d_{2i} g( d_{2i}) q_i$ or even
+  use a GREG estimator of $T_g$. 
+%I wonder we can assume that an unbiased estimator of $T_g$ is available from
+  %$A_c=A_1 \cup A_2$. That is, let's assume that we have $\hat{T}_c = W
+  %\hat{T}_1 + (1- W) \hat{T}_2$ as the estimator  for $T$. 
 }

### 2024-04-30 src/explore/proto-dctp.R

@@ -197,12 +197,12 @@ tp_reg <- function(p1_df, p2_df, pop_obs, sampling = "srs-poisson") {
 #' @param entropy - One of "EL", "ET"
 solveGE <- function(zmat, w1i, d2i, T1, qi, entropy) {
 
-  f <- function(lam, zmat, w1i, d2i, T1, qi, entropy, returnw = FALSE) {
+  f <- function(lam, zmat, w1i, T1, qi, entropy, returnw = FALSE) {
 
     if (entropy == "EL") {
-      w <- -d2i / drop(zmat %*% lam)
+      w <- -1 / drop(zmat %*% lam)
     } else if (entropy == "ET") {
-      w <- d2i * exp(drop(zmat %*% lam))
+      w <- exp(drop(zmat %*% lam))
     } else {
       stop("We only accept entropy of EL or ET.")
     }
@@ -214,11 +214,11 @@ solveGE <- function(zmat, w1i, d2i, T1, qi, entropy) {
     }
   }
 
-  j <- function(lam, zmat, w1i, d2i, T1, qi, entropy) {
+  j <- function(lam, zmat, w1i, T1, qi, entropy) {
     if (entropy == "EL") {
-      res <- t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
+      res <- t(zmat) %*% diag((w1i * qi) / drop(zmat %*% lam)^2) %*% zmat
     } else if (entropy == "ET") {
-      res <- t(zmat) %*% diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
+      res <- t(zmat) %*% diag((w1i * qi) * exp(drop(zmat %*% lam))) %*% zmat
     }
 
     return(res)
@@ -226,11 +226,11 @@ solveGE <- function(zmat, w1i, d2i, T1, qi, entropy) {
 
   init <- c(rep(0, ncol(zmat) - 1), 1)
   res <- 
-  nleqslv(init, f, jac = j, zmat = zmat, w1i = w1i, d2i = d2i,
+  nleqslv(init, f, jac = j, zmat = zmat, w1i = w1i,
           T1 = T1, qi = qi, entropy = entropy,
           method = "Newton", control = list(maxit = 1e5, allowSingular = TRUE))
 
-  resw <- f(res$x, zmat, w1i, d2i, T1, qi, entropy, returnw = TRUE)
[…]

### 2024-04-30 src/explore/proto-dctpmcmc.R

@@ -19,7 +19,7 @@ source("proto-dctp.R")
 
 B_sims <- 1000
 N_obs <- 10000
-n1_obs <- 1000
+# n1_obs <- 1000
 
 # ************
 # * Run MCMC *
@@ -109,7 +109,7 @@ mc_res %>%
 
 %>%
   knitr::kable("latex", booktabs = TRUE, escape = FALSE, digits = 3) %>%
-  cat(file = "tables/tpdcsim_mean.tex")
+  cat(file = paste0("tables/tpdcsim",  "_mean.tex"))
 
 # Analyze Variance
 mc_res %>%
@@ -136,7 +136,7 @@ mc_res %>%
 
 %>%
   knitr::kable("latex", booktabs = TRUE, escape = FALSE, digits = 4) %>%
-  cat(file = "tables/tpdcsim_var.tex")
+  cat(file = paste0("tables/tpdcsim", "_var.tex"))
 
 
 mc_res %>%

### 2024-05-01 src/explore/proto-dctp.R
Updated simulation study of Topic 1. In this case we tried to get TP-Reg to match DC-Est. They look very similar, but TP-Reg is still better.

@@ -123,10 +123,15 @@ pi_star <- function(p1_df, p2_df, pop_obs, sampling = "srs-poisson") {
 }
 
 #' This function implements the two-phase regression estimator
-tp_reg <- function(p1_df, p2_df, pop_obs, sampling = "srs-poisson") {
+tp_reg <- function(p1_df, p2_df, pop_df, sampling = "srs-poisson") {
 
   # Step 1: Get beta
-  mod <- lm(Y ~ X1 + X2, data = p2_df)
+  
+  gpinv <- 1 / p2_df$pi2^2
+  p2_df <- p2_df %>%
+    mutate(gpinv = gpinv) %>%
+    mutate(weight_v = gpinv / pi1)
+  mod <- lm(Y ~ X1 + X2 + I(1 / pi2), weights = weight_v, data = p2_df)
 
   # Step 2: Predict and Get Variance
   if (sampling == "srs-srs") {
@@ -134,7 +139,7 @@ tp_reg <- function(p1_df, p2_df, pop_obs, sampling = "srs-poisson") {
     theta <- mean(predict(mod, p1_df))
 
     vhat <-
-      (1 / n1 - 1 / pop_obs) * (mod$coefficients[2]^2 * var(p1_df$X1) + 
+      (1 / n1 - 1 / nrow(pop_df)) * (mod$coefficients[2]^2 * var(p1_df$X1) + 
       mod$coefficients[3]^2 * var(p1_df$X2)) +
       (1 / n2 - 1 / n1) * var(p2_df$Y - predict(mod, p2_df))
 
@@ -142,7 +147,7 @@ tp_reg <- function(p1_df, p2_df, pop_obs, sampling = "srs-poisson") {
 
     theta <- 
       (sum(predict(mod, p1_df) / p1_df$pi1) + 
-      sum(1 / (p2_df$pi1 * p2_df$pi2) * (p2_df$Y - predict(mod, p2_df)))) / pop_obs
+      sum(1 / (p2_df$pi1 * p2_df$pi2) * (p2_df$Y - predict(mod, p2_df)))) / nrow(pop_df)
 
     p12_df <- 
       left_join(p1_df, p2_df, by = join_by(X1, X2, X3, X4, Y, pi1, pi2, del1)) %>%
[…]

### 2024-05-01 src/explore/proto-dctpmcmc.R

@@ -54,7 +54,7 @@ mc_res <-
   p2_df <- samps[[2]]
 
   pistar <- pi_star(p1_df, p2_df, N_obs)
-  tpreg <- tp_reg(p1_df, p2_df, N_obs)
+  tpreg <- tp_reg(p1_df, p2_df, upop_df)
   dcpop <- dc_ybar(p1_df, p2_df, upop_df, qi = 1, entropy = "EL", estT1 = FALSE)
   dcest <- dc_ybar(p1_df, p2_df, upop_df, qi = 1, entropy = "EL", estT1 = TRUE)
 

### 2024-05-04 final/reports/prelim_paper.tex
Updated prelim_paper to include fixed simulation study for topic 1.

@@
 
-    \textcolor{red}{You should include an intercept in $\bf x_i$. }
-    \textcolor{blue}{I have been including an intercept, but I forgot to add it
-    into these notes.}
+    %\textcolor{red}{You should include an intercept in $\bf x_i$. }
+    %\textcolor{blue}{I have been including an intercept, but I forgot to add it
+    %into these notes.}
   \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
@@
     This solves Equation~\eqref{eq:primal} with $q_i = 1$.
+  \item[5.] Linearized Debiased Calibration with Population Constraints (Eta-Pop):
+    This is the linearized estimate DC-Pop using the linearization in
+    Theorem~\ref{thm:dc1}.
+  \item[6.] Linearized Debiased Calibration with Estimated Population
+    Constraints (Eta-Est): This is the linearized estimate DC-Est using the
+    linearization in Theorem~\ref{thm:dc1}.
 \end{itemize}
@@
         \varepsilon_i = (y_i - \bf z_i' \bm \phi)$} \\
+    Eta-Pop & Same as DC-Pop & \\
+    Eta-Est & Same as DC-Est & \\
     \bottomrule
@@
 
-\textcolor{red}{I am not sure about the variance estimation formula in Table 1. 
-They are different from what we discussed before. I wonder whether you can make
-the sampling rate $n/N$ negligible so that the second term becomes
-insignificant. How about making $N=50,000$, $n_1=500$, and $n_2=200$?}
-
-\textcolor{blue}{The result with $N = 50,000$, $n_1 = 500$ and $E[n_2] \approx
-200$}
-
-\begin{table}[ht!]
-  \centering
-  {\color{blue}\input{tables/tpdcsim50000_mean.tex}}
-% This table was generated from ../../src/explore/proto-dctpmcmc.R
-  \caption{\textcolor{blue}{This table shows the results of Simulation Study 1 with
-      a different finite population and Phase 2 selection probability.
-      It displays the Bias, RMSE, empirical 95\% confidence interval, and a
-      t-statistic assessing the unbiasedness of each estimator for the
-      estimators: $\pi^*$, TP-Reg, DC-Pop, and DC-Est.}}
-\end{table}
+%\textcolor{red}{I am not sure about the variance estimation formula in Table 1. 
+%They are different from what we discussed before. I wonder whether you can make
+%the sampling rate $n/N$ negligible so that the second term becomes
+%insignificant. How about making $N=50,000$, $n_1=500$, and $n_2=200$?}
+%
+%\textcolor{blue}{The result with $N = 50,000$, $n_1 = 500$ and $E[n_2] \approx
+%200$}
+%
+%\begin{table}[ht!]
+%  \centering
+%  {\color{blue}\input{tables/tpdcsim50000_mean.tex}}
+%% This table was generated from ../../src/explore/proto-dctpmcmc.R
+%  \caption{\textcolor{blue}{This table shows the results of Simulation Study 1 with
+%      a different finite population and Phase 2 selection probability.
+%      It displays the Bias, RMSE, empirical 95\% confidence interval, and a
+%      t-statistic assessing the unbiasedness of each estimator for the
+%      estimators: $\pi^*$, TP-Reg, DC-Pop, and DC-Est.}}
+%\end{table}
 
@@
 
-\textcolor{red}{Dr. Kim is this confidence interval good enough? It seems
-slightly high for me.}
+\textcolor{red}{While DC-Est still does not outperform TP-Reg, it is equivalent
+to Eta-Est, which is a regression estimator.}
+
+\textcolor{red}{The variance estimates look good now. I needed to think of
+DC-Pop as a regression estimator instead of a two-phase regression estimator.}
 

### 2024-05-04 src/explore/proto-dctp.R

@@ -95,7 +95,7 @@ gen_samps <- function(uppop_df, p1_type, p2_type) {
 ## Alternative Estimators
 
 #' This function implements the $\pi^*$-estimator for two-phase sampling.
-pi_star <- function(p1_df, p2_df, pop_obs, sampling = "srs-poisson") {
+pi_star <- function(p1_df, p2_df, pop_obs, sampling = "poisson-poisson") {
 
   theta <- sum(p2_df$Y / (p2_df$pi1 * p2_df$pi2)) / pop_obs
 
@@ -110,10 +110,32 @@ pi_star <- function(p1_df, p2_df, pop_obs, sampling = "srs-poisson") {
 
   } else if (sampling == "srs-poisson") {
 
-    var_y <- var(p2_df$Y)
-    v1 <- (1 / nrow(p1_df) - 1 / pop_obs) * var_y
-    v2 <- sum((1 - p2_df$pi2) / (p2_df$pi1^2 * p2_df$pi2^2) * p2_df$Y^2) / pop_obs^2
-    vhat <- v1 + v2
+    # var_y <- var(p2_df$Y)
+    # v1 <- (1 / nrow(p1_df) - 1 / pop_obs) * var_y
+    # v2 <- sum((1 - p2_df$pi2) / (p2_df$pi1^2 * p2_df$pi2^2) * p2_df$Y^2) / pop_obs^2
+    # vhat <- v1 + v2
+
+    tmp <- matrix(NA, nrow = nrow(p2_df), ncol = nrow(p2_df))
+    for (i in 1:nrow(p2_df)) {
+      for (j in 1:nrow(p2_df)) {
+        if (i == j) {
+          pi2i <- p2_df$pi2[i] * p2_df$pi1[i]
+          tmp[i, j] <- ((pi2i - (pi2i)^2 / pi2i) * p2_df$Y[i]^2 / (pi2i^2)
+        } else {
+          pi2i <- p2_df$pi2[i] * p2_df$pi1[i]
+          pi2j <- p2_df$pi2[j] * p2_df$pi1[j]
+          pi2ijp1ij <- 
+            p2_df$pi2[i] * p2_df$pi2[j] * p2_df$pi1[i] * 
+            (nrow(p1_df) - 1) / (nrow(pop_df) - 1)
+          tmp[i, j] <- 
+            ((pi2ijp1ij - pi2i * pi2j) / pi2ijp1ij) * 
[…]

### 2024-05-04 src/explore/proto-dctpmcmc.R

@@ -19,7 +19,13 @@ source("proto-dctp.R")
 
 B_sims <- 1000
 N_obs <- 10000
-# n1_obs <- 1000
+
+p1_type <- "srs"
+# p1_pi <- "ifelse(pnorm(-pop_df$X2) < 0.10, 0.10, pnorm(-pop_df$X2))"
+p1_pi <- "1200 / nrow(pop_df)"
+p2_type <- "poisson"
+p2_pi <- "ifelse(pt(-pop_df$X1 + 1, 3) > 0.9, 0.9, pt(-pop_df$X1 + 1, 3))"
+comb <- paste0(p1_type, "-", p2_type)
 
 # ************
 # * Run MCMC *
@@ -29,12 +35,7 @@ set.seed(1)
 pop_df <- gen_pop(N_obs)
 true_theta <- mean(pop_df$Y)
 
-upop_df <-
-  update_pop(
-    pop_df,
-    "1000 / nrow(pop_df)",
-    "ifelse(pt(-pop_df$X1 + 1, 3) > 0.9, 0.9, pt(-pop_df$X1 + 1, 3))"
-  )
+upop_df <- update_pop(pop_df, p1_pi, p2_pi)
 
 clust <- makeCluster(min(detectCores() - 2, 100), outfile = "")
 registerDoParallel(clust)
@@ -49,20 +50,28 @@ mc_res <-
 
   set.seed(iter)
 
-  samps <- gen_samps(upop_df, "srs", "poisson")
+  samps <- gen_samps(upop_df, p1_type, p2_type)
   p1_df <- samps[[1]]
   p2_df <- samps[[2]]
 
-  pistar <- pi_star(p1_df, p2_df, N_obs)
-  tpreg <- tp_reg(p1_df, p2_df, upop_df)
-  dcpop <- dc_ybar(p1_df, p2_df, upop_df, qi = 1, entropy = "EL", estT1 = FALSE)
-  dcest <- dc_ybar(p1_df, p2_df, upop_df, qi = 1, entropy = "EL", estT1 = TRUE)
+  pistar <- pi_star(p1_df, p2_df, N_obs, sampling = comb)
+  tpreg <- tp_reg(p1_df, p2_df, upop_df, sampling = comb)
+  dcpop <- dc_ybar(p1_df, p2_df, upop_df,
+                   qi = 1, entropy = "EL", estT1 = FALSE, sampling = comb)
+  dcest <- dc_ybar(p1_df, p2_df, upop_df,
[…]

### 2024-05-04 final/reports/twophaseridge.tex
Added note connecting non-nested sampling and ridge regression.

--- 2024-04-23 final/reports/prelim_paper_ov.tex
@@
 
-\title{Debiased Calibration for Generalized Two-Phase Sampling}
+\title{Similarities between Nonnested Regression Estimation and Ridge Regression}
 \author{Caleb Leedy}
@@
 
-\section{Introduction}
+\section{Nonnested Regression}
 
-Combining information from several sources is an important practical problem. (CITEME)
-We want to incorporate information from external data sources to reduce the bias
-in our estimates or improve the estimator's efficiency. For many problems, the
-additional information consists of summary statistics with standard errors. The
-goal of this project is to incorporate external information with existing data 
-to create more efficient estimators using calibration weighting.
-
-To model this scenaro, we formulate the problem as a generalized two-phase
-sample where the first phase sample consists of data from multiple sources. The
-second phase sample contains our existing data. To motivate this setup, we
-consider the following approach: first, we consider the classical two-phase
-sampling setup where the second phase sample is a subset of the first phase
-sample; then, we extend this setup to consider non-nested two-phase samples;
-and finally, we consider the more general approach of having multiple sources.
-
-%\begin{itemize}
-%\item To achieve the goal, we first consider  the  classical two-phase sampling
-%setup where the second-phase sample is a subset of the first-phase sample.
-%  After that, we extend the setup to more general cases such as non-nested
-%  two-phase sampling or multiple independent surveys with some common
-%  measurements. 
-%\item The proposed method can be called two-step calibration. In the
-%first-step, the best linear unbiased estimators of the auxiliary variable
-%totals are computed. In the second-step, the final calibration weights are
-%constructed to match (benchmark) with the best estimators computed from Step 1. 
-%\end{itemize}
-
-\section{Topic 1: Classical Two-Phase Sampling}
-
-\subsection{Background}
-
-Consider a finite population of size $N$ containing elements $(X_i, Y_i)$ where
-an initial (Phase 1) sample of size $n_1$ is selected and $X_i$ is observed. Then
-from the Phase 1 sample of elements, a (Phase 2) sample of size $n_2 < n_1$ is
-selected and $Y_i$ is observed. This is two-phase sampling (See 
-\cite{fuller2009sampling}, \cite{kim2024statistics} for general references.) The
-goal of two-phase sampling is to construct an estimator of $\bar Y_N$ 
-that uses both the observed information in the Phase 2 sample and also the extra
-auxiliary information from $X$ in the Phase 1 sample.
-The challenge is doing this efficiently.
-
-An easy-to-implement unbiased estimator in the spirit of a Horvitz-Thompson (HT)
-estimator (\cite{horvitz1952generalization}, \cite{narain1951sampling}) is the
-$\pi^*$-estimator. Let $\pi_i^{(2)}$ be the response probability of element $i$
-being observed in the Phase 2 sample. Then, allowing the elements in the Phase 1
-sample to be represented by $A_1$ and the elements in the Phase 2 sample to be
-denoted as $A_2$,
-%\begin{align*}
-%  \pi_i &= \sum_{A_2: i \in A_2} \Pr(A_2) \\ 
-%        &= \sum_{A_1: A_2 \subseteq A_1} \sum_{A_2: i \in A_2} \Pr(A_2 \mid
-%        A_1) \Pr(A_1) \\
-%        &= \sum_{A_1: i \in A_1} \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1) \Pr(A_1).
-%\end{align*}
-%\textcolor{red}{(I am not sure whether this part is necessary.) }
-if we define $\pi_{2i | 1} = \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1)$ and
-$\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ then,
-
-$$ \pi_i^{(2)}(A_1) = \pi_{2i | 1} \pi_{1i}.$$
-
-This means that we can define the $\pi^*$-estimator as the following design
-unbiased estimator:
-
-$$ \hat Y_{\pi^*} = \sum_{i \in A_2} \frac{y_i}{\pi_{2i | 1} \pi_{1i}}.$$
-
-While unbiased (see \cite{kim2024statistics}), the $\pi^*$-estimator
-does not account for the additional information contained in the auxiliary Phase
-1 variable $X$. The two-phase regression estimator $\hat Y_{reg, tp}$ does incorporate 
-information for $X$ by using the estimate $\hat X_1$ from the Phase 1 sample.
-This is how we can leverage the external information $\hat X_1$ to improve the
-initial $\pi^*$-estimator in the second phase sample.
-The two-phase regression estimator has the form,
-
-$$ \hat Y_{reg, tp} 
-= \sum_{i \in A_1} \frac{1}{\pi_{1i}} x_i \hat \beta_q+ \sum_{i \in A_2}
-\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_q)$$
-
-where for $q_i = q(x_i)$ and is a function of $x_i$,
-$$
-\hat \beta_q = \left(\sum_{i \in A_2} 
-  \frac{x_i x_i'}{\pi_{1i} q_i}\right)^{-1} 
-\sum_{i \in A_2} \frac{x_i y_i}{\pi_{1i} q_i}.$$ 
-%\footnote{
-%  \textcolor{blue}{Caleb, we do not have to use $\pi_{2i \mid 1}^{(2)}$ in computing
-%  $\hat{\beta}_2$. Please check my book.}
-%  \textcolor{red}{Thanks for the check. I
-%checked it and I must have gotten $\pi_{1i}$ and $\pi_{2i}$ confused.}}
-%\footnote{\textcolor{red}{I was initially thinking of using $\hat \beta_2$ without $q_i$ to be
-%  more consistent with how I defined the two-phase regression estimator;
-%  however, I see your point about using $q_i$ to connect this estimator with the
-%superpopulation model.}}
-
-The regression estimator is the minimum variance design consistent linear
-estimator which is easily shown to be the case because $\hat Y_{reg, tp} =
-\sum_{i \in A_2} \hat w_{2i} y_i / \pi_{1i}$ where
-
-$$\hat w_{2i} = \argmin_{w} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i \text{ such
-that } \sum_{i \in A_2} w_{2i} x_i / \pi_{1i} = \sum_{i \in A_1} x_i / \pi_{1i}.$$
-
-This means that $\hat Y_{reg, tp}$ is also a calibration estimator. The idea
-that regression estimation is a form of calibration was noted by
-\cite{deville1992calibration} and extended by them to consider loss functions
-other than just squared loss. Their generalized loss function minimizes
-$\sum_i G(w_i, d_i)q_i$ for weights $w_i$ and design-weights $d_i$ where
-$G(\cdot)$ is a non-negative, strictly convex function with respect to $w$,
-defined on an interval containing $d_i$, with $g(w_i, d_i) = \partial G /
-\partial w$ continuous.\footnote{The \cite{deville1992calibration} paper
-considers regression estimators for a single phase setup, which we apply to our
-two-phase example.} This
-generalization includes empirical likelihood estimation, and maximum entropy
-estimation among others. The variance estimation is based on a linearization
-that shows that minimizing the generalized loss function subject to the
-calibration constraints is asymptotically equivalent to a regression estimator.
-
-Furthermore, the regression estimator
-has a nice feature that its two terms can be thought about as minimizing the
-variance and bias correction,
-
-$$ \hat Y_{reg, tp} 
-= \underbrace{\sum_{i \in A_1} \frac{x_i \hat \beta_q}{\pi_{1i}}}_{
-  \text{ Minimizing the variance}} + \underbrace{\sum_{i \in A_2}
-\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_q)}_{
-\text{Bias correction}}.$$
-
-The \cite{deville1992calibration} method incorporates the design weights into
-the loss function, which is the part minimizing the variance. We would rather
-separate have bias calibration separate from the minimizing the variance so that
-we can control each in isolation. In
-\cite{kwon2024debiased}, the authors show that for a generalized entropy
-function $G(w)$, including a term of $g(\pi_{2i|1}^{-1})$ into the calibration
-for $g = \partial G / \partial w$ not only creates a design consistent
-estimator, but it also has better efficiency than the generalized regression
-estimators of \cite{deville1992calibration}.
-
-The method of \cite{kwon2024debiased} requires known finite population 
-calibration levels. It does not handle the
-two-phase setup where we need to estimate the finite population total of $x$
-from the Phase 1 sample. In the rest of the section, we extend this method to 
-two phase sampling so that we have a valid 
-estimator when including estimated Phase 1 weights with appropriate variance
-estimation.
-
-\subsection{Methodology}
-
-We follow the approach of \cite{kwon2024debiased} for the debiased calibration
-method. We consider maximizing the generalized entropy \cite{gneiting2007strictly},
-
-\begin{equation}\label{eq:primalloss}
-  H(w) = - \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i
-\end{equation}
-
-where $G: \mathcal{V} \to \R$ is strictly convex, differentiable function
-subject to the constraints:
-
-\begin{equation}\label{eq:calconst1}
-  \sum_{i \in A_2} \frac{x_i w_{2i}q_i}{\pi_{1i}} = 
-\sum_{i \in A_1} \frac{x_iq_i}{\pi_{1i}}
-\end{equation}
-
-and
-
-\begin{equation}\label{eq:calconst2}
-  \sum_{i \in A_2} \frac{g(\pi_{2i|1}^{-1})w_{2i}q_i}{\pi_{1i}} = 
-  \sum_{i \in A_1} \frac{g(\pi_{2i|1}^{-1})q_i}{\pi_{1i}}.
-\end{equation}
-
-The first constraint is the existing calibration constraint and the second
-ensures that design consistency is achieved. Here, 
-$g(w) = \partial G / \partial w$. 
-The original method of \cite{kwon2024debiased} only considered having finite
-population quantities on the right hand side of \ref{eq:calconst1}.
-Let $z_i = (x_i / q_i, g(\pi_{2i|1}^{-1}))$. Letting $w_{1i} = \pi_{1i}^{-1}$,
-the the goal is to solve,
-
-\begin{equation}\label{eq:primal}
-  \argmin_{w_{2|1}} \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i 
-  \text{ such that}
-  \sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i = \sum_{i \in A_1} w_{1i} z_i q_i.
-\end{equation}
-
-Let $\hat w_{2i|1}$ be the solution to Equation~\ref{eq:primal}, then the
-estimate of $Y_N$ is $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i|1} y_i$.
-
-\subsection{Theoretical Results}
-
-\begin{theorem}[Design Consistency]
-  Under some regularity conditions,
-
-  $$\hat Y_{DCE} = \hat Y_\ell(\lambda^*, \phi^*) + O_p(N / n_2)$$
-
-  where
-\textcolor{red}{
-   \begin{eqnarray*}
-   \hat Y_{\ell}( \lambda^*, \phi^*) &=&   \hat{Y}_{\pi^*} + 
-  \left(\sum_{i \in A_1} w_{1i} x_i  -  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} x_i  \right)^T \phi_1^* + \left(\sum_{i \in A_1} w_{1i} g_i   -  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^T \phi_2^*
-  \end{eqnarray*} 
-  and  $$\begin{pmatrix}
-  \phi_1^* \\
-  \phi_2^* 
-  \end{pmatrix}
-  = \left[ \sum_{i \in U} \frac{\pi_{2i \mid 1} }{ g'(d_{2i|1}) q_i} 
-  \begin{pmatrix}
-  \bx_i \bx_i^T &   \bx_i g_i   \\
-  g_i  \bx_i^T   & g_i^2     \end{pmatrix} \right]^{-1}
-  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} \begin{pmatrix}
-    \bx_i \\ g_i 
-  \end{pmatrix}y_i $$
-with $g_i = g( \pi_{2i |1}^{-1}) q_i$. }
-\end{theorem}
-
-\textcolor{red}{I have tried to improve the expression of the linearization formula. Please check if my algebra above is correct. }
-\begin{proof}
-  In this proof, we derive the solution to Equation~\ref{eq:primal} and show
-  that it is asymptotically equivalent to a regression estimator. Using the
-  method of Lagrange multipliers, to solve Equation~\ref{eq:primal} we need to
-  minimize,
-
-  $$\mathcal{L}(w_{2|1}) = \sum_{i \in A_2} w_{1i} G(w_{2i|1}) q_i + 
-  \lambda \left(\sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i - \sum_{i \in A_1} w_{1i}
-  z_i q_i\right).$$
-
-  The first order conditions show that
-
-  $$\frac{\partial \mathcal{L}}{\partial w_{2i|1}}: g(w_{2i|1}) w_{1i}q_i +
-    \lambda w_{1i} z_i q_i = 0.$$
-
-  Hence, $\hat w_{2i}(\lambda) = g^{-1}(\lambda z_i)$ and $\hat \lambda$ is
-  determined by the calibration condition. When the sample size gets large, we
-  have $\hat w_{2i|1}(\hat \lambda) \to d_{2i|1}$ which means that $\hat \lambda
-  \to \lambda^*$ where $\lambda^* = (\bf 0, 1)$. Then using the linearization
-  technique of \cite{randles1982asymptotic}, we can construct a regression
-  estimator,
-
-  $$\hat Y_\ell(\hat \lambda, \phi)  = \hat Y_{DCE}(\hat \lambda) + 
-  \left(\sum_{i \in A_1} w_{1i} z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat
-  \lambda) z_i q_i\right)\phi.$$
-
-  Notice that $\hat Y_\ell(\hat \lambda, \phi) = \hat Y_{DCE}(\hat \lambda)$ for
-  all $\phi \in \mathbb{R}$. We choose $\phi^*$ such that
-
-  $$E\left[\frac{\partial}{\partial \lambda} \hat Y_\ell(\lambda^*, \phi^*)\right]=0.$$
-
-  Using the fact that $g^{-1}(\lambda^* z_i) = g^{-1}(g(d_{2i|1})) = d_{2i|1}$
-  and $(g^{-1})'(x) = 1 / g'(g^{-1}(x))$, we have
+In the case of non-nested two phase sampling, we have $A_1 = (X_i)_{i = 1}^{n_1}$
+and $A_2 = (X_1, Y_i)_{i = 1}^{n_2}$ with $A_1$ and $A_2$ being selected
+independently from the same sampling frame. The regression estimator is then
 
 \begin{align*}
-    \phi^*
-    &= E\left[\sum_{i \in A_2} \frac{w_{1i}z_i z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
-    E\left[\sum_{i \in A_2} \frac{w_{1i}z_i y_i}{g'(d_{2i|1})}\right]\\
-    &= \left[\sum_{i \in U} \frac{\pi_{2i|1}z_i z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
-    \left[\sum_{i \in U} \frac{\pi_{2i|1} z_i y_i}{g'(d_{2i|1})}\right]\\
+\hat{\bar{Y}}_{reg} 
+  &= \hat{\bar{Y}}_{HT} + (\hat{\bar{X}}_c - \hat{\bar{X}}_2)' \hat \beta_2 
+  & \text{for } \hat \beta_2 = \left(\sum_{i \in A_2} x_i x_i'\right)^{-1} 
+    \sum_{i \in A2} x_i y_i \text{ with } x_{1i} = 1 \\
+  &= \hat{\bar{Y}}_{HT} + (\hat{\bar{X}}_1 - \hat{\bar{X}}_2)'W'\hat \beta_2 
+  & \text{ if } \hat{\bar{X}}_c = W \hat{\bar{X}}_1 + (I - W) \hat{\bar{X}}_2\\
+  &= \sum_{i \in A_1} x_i' W' \hat \beta_2 + \sum_{i \in A_2} (y_i - x_i' W'
+  \hat \beta_2).
 \end{align*}
 
-  Thus, the linearization estimator is
+While the matrix $W$ controls the interaction between $A_1$ and $A_2$ it also
+plays the role of shrinkage on $\hat \beta_2$.
 
-  $$\hat Y_\ell(\lambda^*, \phi^*) = \sum_{i \in A_1} w_{1i} q_i z_i \phi^* +
-  \sum_{i \in A_2} w_{1i} d_{2i|1} (y_i - q_i z_i \phi^*).$$
+\section{Ridge Regression}
 
-  By construction using a Taylor expansion yields,
+Given a sample $A$, ridge regression solves the optimization problem,
 
-  \begin{align*}
-    \hat Y_{DCE}(\hat \lambda) 
-    &= \hat Y_\ell(\lambda^*, \phi^*) + 
-    E\left[\frac{\partial}{\partial \lambda}\hat Y_\ell(\lambda^*,
-    \phi^*)\right](\hat \lambda - \lambda^*) + \frac{1}{2}
-    E\left[\frac{\partial}{\partial \lambda^2} \hat Y_{DCE}(\lambda^*)\right] (\hat
-    \lambda - \lambda^*)^2\\
-    &= \hat Y_\ell(\lambda^*, \phi^*) + O(N)o_p(n_2^{-1}).
-  \end{align*}
+$$\hat \beta = \argmin_{\beta \in \R^p} \sum_{i \in A} (y_i - x_i'\beta)^2 + \beta^T
+(\lambda I_p) \beta.$$
 
-  The final equality comes from the fact that 
-  $E\left[\frac{\partial}{\partial \lambda}\hat Y_\ell(\lambda^*, \phi^*)\right]
-  = 0$, $\frac{\partial}{\partial \lambda^2} \hat w_{2i|1}(\lambda^*)$ is
-  bounded and $|\hat \lambda - \lambda^*| = o_p(n_2^{-1/2})$, which proves our
-  result.
-\footnote{$\hat{\lambda}$ should satisfy $\hat{\lambda} - \lambda^* = O_p (n_2^{-1/2})$. }
-\end{proof}
+Differentiating with respect to $\beta$ and setting this equal to zero yields a
+solution,
 
-%\begin{proof}
-%  First, we derive the solution for $\hat Y_{DCE}$, and then derive the
-%  asymptotic variance of this estimator using the linearization technique of
-%  \cite{randles1982asymptotic}.
-%
-%  We want to solve Equation~\ref{eq:primalloss} subject to the constraints in
-%  Equations~\ref{eq:calconst1} and \ref{eq:calconst2}. Using the method of
-%  Lagrange multipliers we have a Lagrangian of 
-%
-%  $$\mathcal{L}(w_2) = -\sum_{i \in A_2} w_{1i} G(w_{2i})q_i + \lambda 
-%  \left(\sum_{i \in A_2} w_{1i} w_{2i} z_i - \sum_{i \in A_1} w_{1i} z_i\right).$$
-%
-%  Our first order conditions yield,
-%
-%  \begin{align*}
-%    0 &= \frac{\partial}{\partial w_{2i}} \mathcal{L}(w_{2}) \\ 
-%      &= -w_{1i} g(w_{2i}) q_i + \lambda w_{1i} z_i.
-%  \end{align*}
-%
-%  This means that $\hat w_{2i} = g^{-1}(\hat \lambda z_i / q_i)$ and since $\hat
-%  \lambda$ is determined from the solution to Equations~\ref{eq:calconst1} and
-%  \ref{eq:calconst2}. Hence, $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} 
-%  \hat w_{2i}(\hat \lambda) y_i$.
-%
-%  To get the variance of $\hat Y_{DCE}$ we use the linearization technique of
-%  \cite{randles1982asymptotic}. First, notice that 
-%
-%  $$\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i =
-%  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i + \left(
-%  \sum_{i \in A_1} w_{1i} z_i -
-%  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) z_i\right)\phi
-%  := \hat Y_\ell(\hat \lambda, \phi).$$
-%
-%  We can find the value of $\phi$ such that $E[\frac{\partial}{\partial
-%  \lambda}\hat Y_\ell(\lambda, \phi)]_{\lambda = \lambda^*}$ where $\lambda^*$
-%  is the probability limit of $\hat \lambda$. Notice, that 
-%  $\lambda^* = (\bf 0, 1)$ because $\hat w_{2i} \to \pi_{2i|1}^{-1}$. This means
-%  that the optimal value of $\phi$ solves,
-%
-%  $$E\left[\frac{\partial}{\partial \lambda} \hat Y_\ell(\lambda, \phi)\right] =
-%  E\left[\sum_{i \in A_2} w_{1i} \frac{\partial}{\partial \lambda} w_{2i|1}(\lambda)
-%    y_i - \phi \sum_{i \in A_2} w_{1i} z_i \frac{\partial}{\partial \lambda}
-%    w_{2i|1}(\lambda) / q_i \right]$$
-%
-%  which means that
-%
-%  $$\phi^* = \left[\sum_{i \in U} \pi_{2i|1} \left(\frac{z_i z_i^T}{
-%    g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in U} \pi_{2i|1}
-%    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
-%
-%  Hence,
-%
-%  \begin{align*}
-%    \hat Y_\ell(\lambda^*, \phi^*) 
-%    &= \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1}
-%    y_i + \left(\sum_{i \in A_1} w_{1i} z_i / q_i -
-%      \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} z_i / q_i\right)\phi^* \\ 
-%    &= \sum_{i \in A_1} w_{1i} \eta_i
-%  \end{align*}
-%
-%  where 
-%
-%  $$\eta_i = z_i \phi^* / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
-%  - z_i \phi^* / q_i\right).$$
-%
-%  To analyze the variance, consider the reverse framework of
-%  \cite{fay1992inferences}, \cite{kim2006replication}, and \cite{shao1999variance}.
-%  In this setup, we assume that $\mathcal{R} = \{\delta_{2i|1}\}_{i \in U}$ is
-%  known at the
-%  finite population level. Then with this we have,
-%
-%  \begin{align*}
-%    V(\hat Y_{\ell}(\lambda^*, \phi^*)) 
-%    &= V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\right) \\
-%    &= V\left(E\left[\sum_{i \in A_1} w_{1i}\eta_i(\phi^*)\mid\mathcal{R}\right]\right) 
-%    +E\left[V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\mid\mathcal{R}\right)\right]\\
-%    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right) + E\left[\sum_{i \in U} 
-%    \sum_{j \in U} \Cov(\delta_{1i}, \delta_{1j}) 
-%    w_{1i} w_{1j} \eta_i(\phi^*) \eta_j(\phi^*)\right]\\
-%  \end{align*}
-%
-%  We can estimate this variance using 
-%
-%  $$\hat V_{DCE} = \sum_{i \in A_1} \sum_{j \in A_1} \left(\frac{\pi_{1ij} -
-%  \pi_{1i}\pi_{1j}}{\pi_{1ij}}
-%    w_{1i} w_{1j} \hat \eta_i \hat \eta_j \right)
-%  $$
-%
-%  for $\hat \eta_i = z_i \hat \phi / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
-%  - z_i \hat \phi / q_i\right)$ where 
-%
-%  $$\hat \phi = \left[\sum_{i \in A_2} w_{1i}\left(\frac{z_i z_i^T}{
-%  g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in A_2} w_{1i}
-%    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
-%
-%  The problem with this estimator is that it is biased as $E[\hat V_{DCE}]$ is
-%  asymptotically equivalent to
-%  $V\left(\sum_{i \in A_1} w_{1i} \eta(\phi^*)\right)$. This yields a bias of 
-%
-%  \begin{align*}
-%    \text{Bias}(\hat V_{DCE}) 
-%    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right)\\ 
-%    &= \sum_{i \in U} \sum_{j \in U} \Delta_{2ij|1} w_{2i|1}
-%    w_{2j|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).
-%  \end{align*}
-%
-%  where $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1} \pi_{2j|1}$. 
-%  However, we can estimate this bias with
-%
-%  $$\widehat{\text{Bias}}(\hat V_{DCE}) = 
-%  \sum_{i \in A_2} \sum_{j \in A_2} w_{1ij} w_{2ij|1} w_{2i|1} w_{2j|1} 
-%  \Delta_{2ij|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).$$
-%
-%\end{proof}
+$$\hat \beta = \left(\sum_{i \in A} x_i x_i' + \lambda I_p\right)^{-1} \sum_{i
+\in A} x_i' y_i = (X'X + \lambda I_p)^{-1} X'Y.$$
 
-\subsection{Simulation Studies}
+Let $\hat \beta_{OLS} = \hat \beta_2 = (X'X)^{-1} X'Y$, then using the
+Sherman-Morrison-Woodbury inverse formula,
 
-We run a simulation testing the proposed method. In this approach we have the
-following simulation setup:
+$$\hat \beta = (I_p - (X'X)^{-1}(\lambda^{-1}I_p + (X'X)^{-1}))^{-1} \hat \beta_{OLS}.$$
+
+\section{Discussion}
+
+The previous two sections suggest that if we let
 
@@
 \begin{aligned}
-X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
-X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
-X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
-X_{4i} &\stackrel{ind}{\sim} Unif(0.1, 0.9) \\
-\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
-Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
-\pi_{1i} &= \Phi_3(-x_{3i} - 2) \\
-\pi_{2i|1} &= x_{4i}.
+  W &= (I_p - (X'X)^{-1}(\lambda^{-1}I_p + (X'X)^{-1}))^{-1} \\
+  \iff \lambda W &= (X'X)(I_p - W) \\ 
+  \iff \lambda &= (X'X)(W^{-1} - I_p)
 \end{aligned}
@@
 
-where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
-
-\textcolor{red}{The simulation setup is somewhat strange.
-The first order inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$.
-In this case, your second-phase sampling is essentially equivalent to the simple random sampling.
-}
-This is a two-phase extension of the setup in \cite{kwon2024debiased}. We
-consider a finite population of size $N = 10,000$ with both the Phase 1 and
-Phase 2 sampling occuring under Poisson (Bernoulli) sampling. This yields a
-Phase 1 sample
-size of $E[n_1] \approx 1100$ and a Phase 2 sample size of
-$E[n_2] \approx 550$. In the Phase 1 sample, we observe 
-$(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. This
-simulation does not deal with model misspecification, and we compare the
-proposed method for the parameter $\bar Y_N$ with four approaches:
-
-\begin{itemize}
-  \item[1.] $\pi^*$-estimator: $\hat Y_{\pi^*} = N^{-1} \sum_{i \in A_2}
-    \frac{y_i}{\pi_{1i} \pi_{2i|1}},$
-  \item[2.] Two Phase Regression estimator (TP-Reg): 
-    $\hat Y_{reg} = \sum_{i \in A_1} \frac{\bf x_i' \hat \beta}{\pi_{1i}} + 
-    \sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}}(y_i - \bf x_i' \hat \beta)$ 
-    where $\hat \beta = 
-    \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$
-    and $\bf x_i = (x_{1i}, x_{2i})^T$,
-    \textcolor{red}{Dr. Kim, should I modify the simulation so that the
-    regression estimator also includes $g(\pi_{2i|1}^{-1})$ as a covariate?}
-  \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This is
-    the method from \cite{kwon2024debiased} with the true population level
-    constraints, and 
-  \item[4.] Debiased Calibration with Estimated Population Constraints (DC-Est):
-    This is the proposed method with the Phase 1 sample being used to estimate
-    the population level constraints.
-\end{itemize}
-
-In addition to estimating the mean parameter $\bar Y_N$, we also construct
-variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$. For each
-approach we have the following variance estimate\footnote{These variance
-estimates use the fact that we have Poisson sampling for both phases in the
-simulation.}:
-
-\begin{figure}[ht!]
-  \centering
-  \begin{tabular}{lcc}
-    \toprule
-    Estimator & Variance & Notes \\
-    \midrule
-    $\pi^*$ & $N^{-2} \sum_{i \in A_2} \left(\pi_{2i|1}^{-2} - \pi_{2i|1}^{-1}\right)
-    y_i^2 $ & \\
-    TP-Reg  & 
-      {\scriptsize $N^{-2}\left(\sum\limits_{i \in A_1} \left(\pi_{1i}^{-2} -
-          \pi_{1i}^{-1}\right) 
-      \eta_i^2 
-      + \sum\limits_{i \in A_2} \frac{1}{\pi_{1i} \pi_{2i|1}}(\pi_{2i|1}^{-1}
-  - 1) (y_i - \bf x_i' \hat \beta)^2 \right)$}
-    & 
-      {\scriptsize$\eta_i = \bf x_i \hat \beta + \frac{\delta_{2i}}{\pi_{2i|1}}(y_i - 
-      \bf x_i \hat \beta)$} \\
-      %$\hat \beta = \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$ \\
-      DC-Pop  & $(Y - \bf z^T \hat \gamma)^T \Pi (Y - \bf z^T \hat \gamma)$ & 
-      {\scriptsize$\Pi
-      = \text{diag}(1 - (\pi_{1}\pi_{2|1})^{-1}) \cdot \frac{\hat w^2}{N^2}$} \\
-      DC-Est  & $(Y - \bf z^T \hat \gamma)^T \Pi (Y - \bf z^T \hat \gamma) +
-      \hat \gamma_{[1:3]}^T V(\bf x) \hat \gamma_{[1:3]}/N^2$ & \\
-    \bottomrule
-  \end{tabular}
-  \caption{This table gives the formulas for each variance estimator used in
-  this simulation.}
-  \label{tab:varforms}
-\end{figure}
-
-We run this simulation $1000$ times for each of these methods and compute the
-Bias ($E[\hat Y] - \bar Y_N$), the RMSE ($\sqrt{\Var(\hat Y - \bar Y_N)}$), a 95\%
-empirical confidence interval ($\sum_{b = 1}^{1000} |\hat Y^{(b)} - \bar Y_N| \leq 
-\Phi(0.975)\sqrt{\hat V(\hat Y^{(b)})^{(b)}}$), and a T-test that assesses the
-unbiasedness of each estimator. The results are in Figure~\ref{fig:tpdc-mean}.
-
-\begin{figure}[ht!]
-  \centering
-\input{tables/tpdcsim_mean.tex}
-% This table was generated from /src/explore/20240411-tpdcsim.qmd
-\caption{This table shows the results of the simulation study. It displays the
-Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
-unbiasedness of each estimator for the estimators: $\pi^*$, TP-Reg, DC-Pop, and
-DC-Est.}
-\label{fig:tpdc-mean}
-\end{figure}
-
-\section{Topic 2: Non-nested Two-Phase Sampling}
-
-\textcolor{blue}{
-( Materials in Section 11.4 can be used here. I have copy-and-pasted the
-textbook materials below. Please modify them.  )
-}
-\textcolor{red}{I will do this shortly.}
-
-\subsection{Background}
-
-Now we consider the sampling mechanism known as non-nested two phase sampling 
-(\cite{kim2024statistics}). In the last section, we considered two phase sampling
-in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
-two phase sampling the Phase 2 sample is independent of the Phase 1 sample. It
-is a separate independent sample of the same population. Like traditional two
-phase sampling, we consider the Phase 1 sample, $A_1$, to consist of
-observations of $(X_i)_{i = 1}^{n_1}$ and the Phase 2 sample, $A_2$, to consist
-of observations of $(X_i, Y_i)_{i = 1}^{n_2}$.
-
-Whereas the classical two phase estimator uses a single Horvitz-Thompson
-estimator of the Phase 1 sample to construct estimates for calibration totals,
-in the non-nested two phase sample we have two independent Horvitz-Thompson
-estimators of the total of $X$,
-
-$$\hat X_1 = \sum_{i \in A_1}^{n_1} \frac{x_i}{\pi_{1i}} \text{ and } 
-\hat X_2 = \sum_{i \in A_2}^{n_2} \frac{x_i}{\pi_{2i}}$$
-
-where $\pi_{1i}$ is the probability of $i \in A_1$ and $\pi_{2i} = \Pr(i \in
-A_2)$. \textcolor{red}{I don't know if it makes sense to change the Phase 2
-selection probability as $\pi_{2i|1}$. Should I keep it at $\pi_{2i}$? But if
-I do, how is the two phase setup useful?} We can combine these estimates to get
-$\hat X_c = W \hat X_1 + (1 - W)\hat X_2$ for some matrix $W$ (see
-\cite{merkouris2004combining} for the optimal choice of $W$). We can then define
-a regression estimator as
-
-$$
-\hat Y_{NN, reg} = \hat Y_2 + (\hat X_c - \hat X_2)^T \hat \beta_q = 
-\hat Y_2 + (\hat X_1 - \hat X_2)^T W\hat \beta_q 
-$$
-
-where
-
-$$\hat \beta_q = \left(\sum_{i \in A_2} \frac{x_i x_i^T}{q_i}\right)^{-1}
-\sum_{i \in A_2} \frac{x_i y_i}{q_i} \text{ and }\hat Y_2 = \sum_{i \in A_2}
-\frac{y_i}{\pi_{2i}}.$$
-
-Since the samples $A_1$ and $A_2$ are independent,
-
-$$V(\hat Y_{NN, reg}) = V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i -
-x_i^TW\beta^*_q)\right) + (\beta^*_q)^T W^T V(\hat X_1) W \beta_q^*$$
-
-where $\beta_q^*$ is the probability limit of $\hat \beta_q$. Like the two phase
-sample this regression estimator can be viewed as the solution to the following
-calibration equation where $d_{2i} = \pi_{2i}^{-1}$,
-
-\begin{equation}\label{eq:nncal}
-\hat w = \argmin_w Q(w) = \sum_{i \in A_2} (w_{2i} - d_{2i})^2 q_i \text{ such
-that } \sum_{i \in A_2} w_{2i} x_i = \hat X_c
-\end{equation}
-
-and $\hat Y_{NN, reg} = \sum_{i \in A_2} \hat w_{2i} y_i$ where $\hat
-w_{2i}$ is the solution to Equation~\ref{eq:nncal}.
-
-We extend the debiased calibration estimator of \cite{kwon2024debiased} to the
-non-nested two phase sampling case where we use a combined estimate $\hat X_c$
-as the calibration totals instead of using the true totals from the finite
-population.
-
-\subsection{Methodology}
-
-\subsection{Theoretical Results}
-
-\subsection{Simulation Studies}
-
-\section{Topic 3: Multi-source Two-Phase Sampling}
-
-\bibliographystyle{chicago}
-\bibliography{references}
+then we would have equivalent results. In this way, non-nested two phase
+sampling is like ridge regression.
 

### 2024-05-06 final/reports/prelim_paper.tex
Added Topic 3 methods.

@@
 from the Phase 1 sample. In the rest of the section, we extend this method to 
-two phase sampling so that we have a valid 
+two-phase sampling so that we have a valid 
 estimator when including estimated Phase 1 weights with appropriate variance
@@
     \frac{y_i}{\pi_{1i} \pi_{2i|1}},$
-  \item[2.] Two Phase Regression estimator (TP-Reg): 
+  \item[2.] Two-Phase Regression estimator (TP-Reg): 
     $\hat Y_{reg} = \sum_{i \in A_1} \frac{\bf x_i' \hat \beta}{\pi_{1i}} + 
@@
 
-Now we consider the sampling mechanism known as non-nested two phase sampling 
-(\cite{hidiroglou2001double}). In the last section, we considered two phase sampling
+Now we consider the sampling mechanism known as non-nested two-phase sampling 
+(\cite{hidiroglou2001double}). In the last section, we considered two-phase sampling
 in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
-two phase sampling the Phase 2 sample is independent of the Phase 1 sample. It
+two-phase sampling the Phase 2 sample is independent of the Phase 1 sample. It
 is a separate independent sample of the same population. Like traditional two
@@
 
-Whereas the classical two phase estimator uses a single Horvitz-Thompson
+Whereas the classical two-phase estimator uses a single Horvitz-Thompson
 estimator of the Phase 1 sample to construct estimates for calibration totals,
-in the non-nested two phase sample we have two independent Horvitz-Thompson
+in the non-nested two-phase sample we have two independent Horvitz-Thompson
 estimators of the total of $\bf X$,
@@
 
-where $\bm \beta_q^*$ is the probability limit of $\hat{\bm \beta_q}$. Like the two phase
+where $\bm \beta_q^*$ is the probability limit of $\hat{\bm \beta_q}$. Like the two-phase
 sample this regression estimator can be viewed as the solution to the following
@@
 We extend the debiased calibration estimator of \cite{kwon2024debiased} to the
-non-nested two phase sampling case where we use a combined estimate 
+non-nested two-phase sampling case where we use a combined estimate 
 $\hat{\bm X_c}$ as the calibration totals instead of using the true totals from
@@
 
-The methodology for the non-nested two phase sample is very similar to the setup
+The methodology for the non-nested two-phase sample is very similar to the setup
 described as part of Topic 1. Given a strictly convex differentiable function,
@@
 
-We can constuct a non-nested two phase estimator $\hat Y_{NNE}$ for $Y_N$ where
+We can constuct a non-nested two-phase estimator $\hat Y_{NNE}$ for $Y_N$ where
 $\hat Y_{NNE} = \sum_{i \in A_2} \hat w_{2i} y_i$ and $\hat w_{2i}$ solves
-Equation~\ref{eq:nnopt}. Like the classical two phase approach, to solve this
+Equation~\ref{eq:nnopt}. Like the classical two-phase approach, to solve this
 setup we minimize the Lagrangian,
@@
 
-\section{Topic 3: Multi-source Two-Phase Sampling}
+\section{Topic 3: Multi-Source Two-Phase Sampling}
+
+\subsection{Background}
+
+When considering non-nested two-phase sampling, we focused on the case of having
+two samples. Now, we consider incorporating more than two independent samples
+together with a debiasing constraint. First, we consider the case in which we
+want to estimate the $\theta = E[Y]$ and $Y$ is only observed in one sample.
+Then we will consider the case where $Y$ may be in multiple samples.
+
+Consider the setup in which we have independent samples $A_0, A_1, \dots, A_M$
+where $Y$ is observed only in $A_0$ but $\bf X$ is observed in all of the
+samples with elements $\bf X^{(0)}$ observed in $A_0$ and $\bf X^{(m)}$ observed
+in $A_m$ for each $m = 1, \dots, M$. Like the non-nested case, we assume that
+each survey is sampling independently from the same population sampling frame.
+
+The traditional multi-source approach (\cite{kim2024statistics}) is to use
+generalized least squares (GLS) to obtain an optimal estimator of $X$ from the
+samples $A_1, \dots, A_M$. Then we can incorporate this information in the
+estimation of $\theta$ by using the following estimate
+
+$$\hat \theta = \sum_{i \in A_0} \hat w_i y_i$$
+
+where
+
+\begin{equation}\label{eq:mstps}
+\hat w = \argmin_w \sum_{i \in A_0} G(w_i)q_i \text{ such that } \sum_{i \in
+A_0} w_i \bf x_i = \hat X_{GLS}^{(0)}
+\end{equation}
+
+where $\hat X_{GLS}^{(0)}$ is the GLS estimate of $X$ for all of the
+$X$-variables measured in sample $A_0$, and $G$ is a generalized entropy
+function.
+
+% Outline
+% Instead of incorporating data from one additional source, we want to combine
+% data from multiple sources
+% * Use GLS (generalized from effective sample size)
+% * We need to add debiasing constraint 
+%   * What do we do if we have multiple surveys for which we observe Y?
+%   * How do we incorporate the selection probability?
+
+\subsection{Methodology}
+
+In order to have a design consistent estimator that incorporates information from
+samples $A_0, A_1, \dots, A_M$, we want to add the debiasing constraints,
+
+$$ \sum_{i \in A_0} w_{i} \bf x_i q_i = \sum_{i \in U} g(d_i) q_i $$
+
+where $g(x) = \partial G(x) / \partial x$ and $d_i$ are the design weights for
+sample $A_0$. This yields the optimization problem:
+
+\begin{equation}\label{eq:dc3}
+  \hat w = \argmin_w \sum_{i \in A_0} G(w_i) q_i + \lambda\left(\hat T - \sum_{i
+  \in A_0} w_i \bf z_i\right)
+\end{equation}
+
+where $\hat T = (\hat{\bf X}^{(0)}, \sum_{i \in U} g(d_i) q_i)^T$ and 
+$\bf z_i = (\bf x_i^T, g(d_i)q_i)^T$. A more general approach considers the case
+in which multiple samples can observe $Y$. In this case, assume that we have
+samples $A_1, \dots A_{M_1}$ for which we observe elements of $\bf X$ and $Y$
+and samples $B_1, \dots, B_{M_2}$ for which we only observed elements of $\bf
+X$. We denote $\bf X^{(A_m)}$ and $\bf X^{(B_m)}$ as the columns of $\bf X$
+observed in samples $A_m$ and $B_m$ respectively. Then the final
+estimator\footnote{This is sort of like stratified sampling with unknown strata.}
+is
+
+$$ \hat Y = M_1^{-1} \sum_{m = 1}^{M_1} \sum_{i \in A_m} \hat w_{im} y_i $$
+
+and $\hat w_{im}$ solves
+
+\begin{equation}\label{eq:dc32}
+  \hat w = \argmin_w \sum_{i \in A_m} G(w_i) q_i + \lambda\left(\hat T_m - \sum_{i
+  \in A_m} w_i \bf z_i\right)
+\end{equation}
+
+where $\hat T = (\hat{\bf X}^{(m)}, \sum_{i \in U} g(d_{im}) q_i)^T$ and $d_{im}$
+are the design weights for $A_m$.
+
+\subsection{Theoretical Results}
+
+\subsection{Simulation Study 3.1}
+
+\subsection{Simulation Study 3.2}
 

### 2024-05-08 final/reports/Debiased_two_phase.tex
Added simulations for debiased nonparametric regression. We still need to find a good setup to show the strength of our result.

\documentclass[12pt]{article}

\usepackage{amsmath, amssymb, amsthm, mathrsfs, fancyhdr}
\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
%\usepackage[style=authoryear]{biblatex}
\usepackage{booktabs}
\usepackage{float}

\usepackage{amsmath, amsthm, mathtools,commath}
\usepackage{graphics, color}
\usepackage{latexsym}
\usepackage{amssymb, amsfonts, bm}
\usepackage{mathrsfs}

\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{makeidx}
\usepackage{fullpage}
\usepackage{booktabs, arydshln}
\usepackage{comment} 
%\makeindex

\hbadness=10000 \tolerance=10000 \hyphenation{en-vi-ron-ment
in-ven-tory e-num-er-ate char-ac-ter-is-tic}

\usepackage[round]{natbib}
%\bibliographystyle{apalike2}
% \bibliographystyle{jmr}

\newcommand{\biblist}{\begin{list}{}
{\listparindent 0.0cm \leftmargin 0.50cm \itemindent -0.50 cm
\labelwidth 0 cm \labelsep 0.50 cm
\usecounter{list}}\clubpenalty4000\widowpenalty4000}
\newcommand{\ebiblist}{\end{list}}

\newcounter{list}

%\usepackage{setspace}

%\usepackage{hangpar}
\newcommand{\lbl}[1]{\label{#1}{\ensuremath{^{\fbox{\tiny\upshape#1}}}}}
% remove % from next line for final copy
\renewcommand{\lbl}[1]{\label{#1}}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{remark}{Remark}
\newtheorem{result}{Result}

\newtheorem{lemma}{Lemma}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}

\newcommand{\bx}{\mathbf{x}}
\newcommand{\R}{\mathbb{R}}
\renewcommand{\bf}[1]{\mathbf{#1}}

\begin{document}

\title{Debiased two-phase regression estimation}
\author{Caleb Leedy \and Jae Kwang Kim}
\maketitle

\baselineskip .3in

\section{Introduction}

\begin{itemize}
\item Introduce two-phase sampling
\item Introduce classical two-phase regression estimation under parametric regression model 
$$ \hat{Y}_{\rm reg} = \sum_{i \in A_1} w_{1i} m(\bx_i ; \hat{\beta}) + \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \left( y_i - m(\bx_i; \hat{\beta}) \right) $$
 Properties: If $\mbox{dim} (\beta)=p$ is fixed in the asymptotic sense, then $\hat{Y}_{\rm reg}$ is asymptotically equivalent to 
$$ \hat{Y}_{\rm reg, \ell} = \sum_{i \in A_1} w_{1i} m(\bx_i ; {\beta}^*) + \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \left( y_i - m(\bx_i; {\beta}^*) \right). $$
Note that $\hat{Y}_{\rm reg, \ell}$ is design unbiased and achieves the optimality if $E(Y \mid \bx_i)= m(\bx_i; \beta_0)$ for some $\beta_0$ and $\hat{\beta}$ is consistent for $\beta_0$.

\item Two-phase regression estimation under non-parametric regression 
$$ \hat{Y}_{\rm np, reg} = \sum_{i \in A_1} w_{1i} \hat{m}(\bx_i ) + \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \left( y_i - \hat{m}(\bx_i) \right) $$
where $\hat{m}(\bx)$ is a consistent estimator of $m(\bx)= E(Y \mid \bx)$, but the convergence rate is slow. 
\item Generally speaking, $\hat{Y}_{\rm rep, np}$ is not asymptotically equivalent to the oracle regression estimator given by 
\begin{equation}
\hat{Y}_{\rm oracle} = \sum_{i \in A_1} w_{1i} {m}(\bx_i ) + \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \left( y_i - {m}(\bx_i) \right). 
\label{oracle}
\end{equation}
Note that 
\begin{eqnarray*}
 \hat{Y}_{\rm np, reg} - \hat{Y}_{\rm  oracle} &=&\sum_{i \in A_1} w_{1i} \left\{ \hat{m} (\bx_i) - m(\bx_i) \right\} -\sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} \left\{ \hat{m} (\bx_i) - m(\bx_i) \right\} \\
 &=& \sum_{i \in A_1} w_{1i} \left\{ \hat{m} (\bx_i) - m(\bx_i) \right\} \left( 1- \delta_i \pi_{2i|1}^{-1} \right) \\
 &:=& D_n .\end{eqnarray*} 
\end{itemize}
In general, the estimation error term $D_n$ is not asymptotically negligible. 
\section{Proposed estimator}

One possible idea is to develop a debiased estimation under two-phase sampling. If the covariates are high dimensional or the regression employs nonparametric regression (such as spline or random forest, etc), then the resulting two-phase regression estimator can be biased. To correct for the bias, we can use the following approach.

\begin{enumerate}
\item Split the sample $A_2$ into two parts: $A_2= A_2^{(a)} \cup A_2^{(b)}$.
We can use SRS to split the sample, but other sampling designs can be used. 
\item Use the observations in $A_2^{(a)}$ only to obtain a predictor of $y_i$, $\hat{m}^{(a)} (\bx_i)$. Also, use the observations in $A_2^{(b)}$ only to obtain a predictor of $y_i$, $\hat{m}^{(b)} (\bx_i)$. 
\item Let 
$$ \hat{m}(\bx_i) = \left( \hat{m}^{(a)} (\bx_i) + \hat{m}^{(b)} (\bx_i) \right)/ 2$$
be the predictor combining two samples. 
\item The final debiased two-phase regression estimator is given by 
\begin{equation}
\hat{Y}_{\rm d, reg} = \sum_{i \in A_1} w_{1i}  \hat{m}(\bx_i) +  \sum_{i \in A_2^{(a)} } w_{1i} \pi_{2i \mid 1}^{-1}   \left\{y_i - \hat{m}^{(b)} (\bx_i) \right\} +\sum_{i \in A_2^{(b)} } w_{1i} \pi_{2i \mid 1}^{-1}   \left\{y_i - \hat{m}^{(a)} (\bx_i) \right\}
\label{debiased}
\end{equation}
\end{enumerate}

Note that $\hat{Y}_{\rm d, reg}$ can be expresssed as 
$$ \hat{Y}_{\rm d, reg}= \left( \hat{Y}_{\rm d, reg}^{(a)} + \hat{Y}_{\rm d, reg}^{(b)} \right)/ 2 $$
where 
$$\hat{Y}_{\rm d, reg}^{(a)} =\sum_{i \in A_1} w_{1i} \hat{m}^{(b)}(\bx_i ) + \sum_{i \in A_2^{(a)}} w_{1i} 2\pi_{2i \mid 1}^{-1} \left( y_i - \hat{m}^{(b)}(\bx_i) \right)$$
and $$\hat{Y}_{\rm d, reg}^{(b)} =\sum_{i \in A_1} w_{1i} \hat{m}^{(a)}(\bx_i ) + \sum_{i \in A_2^{(b)}} w_{1i} 2\pi_{2i \mid 1}^{-1} \left( y_i - \hat{m}^{(a)}(\bx_i) \right).$$

Now, we can establish the following result.

\begin{theorem}
Under some regularity conditions (to be explained later), we obtain 
\begin{equation}
N^{-1} \left( \hat{Y}_{\rm d, reg} - \hat{Y}_{\rm oracle} \right) =o_p( n^{-1/2} ),
\label{result}
\end{equation}
where $\hat{Y}_{\rm oracle}$ is defined in (\ref{oracle}). 
\end{theorem}
\begin{proof}
Define 
$$ \hat{Y}_{\rm oracle}^{(a)} = \sum_{i \in A_1} w_{1i} m(\bx_i) + \sum_{i \in A_2^{(a)} }w_{1i} 2 \pi_{2i \mid 1}^{-1} \left( y_i - m(\bx_i) \right) $$
and
$$ \hat{Y}_{\rm oracle}^{(b)} = \sum_{i \in A_1} w_{1i} m(\bx_i) + \sum_{i \in A_2^{(b)} }w_{1i} 2 \pi_{2i \mid 1}^{-1} \left( y_i - m(\bx_i) \right) .$$
Note that 
$$ \hat{Y}_{\rm oracle} = \left( \hat{Y}_{\rm oracle}^{(a)} + \hat{Y}_{\rm oracle}^{(b)} \right)/ 2$$
where $\hat{Y}_{\rm oracle}$ is defined in (\ref{oracle}).

Note that  
\begin{eqnarray*}
 \hat{Y}_{\rm d, reg}^{(a)} - \hat{Y}_{\rm  oracle}^{(a)} &=&\sum_{i \in A_1} w_{1i} \left\{ \hat{m}^{(b)} (\bx_i) - m(\bx_i) \right\} -\sum_{i \in A_2^{(a)}} w_{1i} 2\pi_{2i|1}^{-1} \left\{ \hat{m}^{(b)}  (\bx_i) - m(\bx_i) \right\} \\
 &=& \sum_{i \in A_1} w_{1i} \left\{ \hat{m}^{(b)}  (\bx_i) - m(\bx_i) \right\} \left( 1- 2 \delta_i I_i^{(a)} \pi_{2i|1}^{-1} \right) \\
 &:=& D_n^{(a)} .\end{eqnarray*} 
It can be shown that, under some regularity conditions, we obtain 
\begin{equation}
 N^{-1} D_n^{(a)} = o_p(n^{-1/2} ).
 \label{res1}
 \end{equation}
%( We may present the above result as a lemma.)

Also, writing 
$$ D_n^{(b)} = \hat{Y}_{\rm d, reg}^{(b)} - \hat{Y}_{\rm  oracle}^{(b)},  $$
we can establish 
\begin{equation}
 N^{-1} D_n^{(b)} = o_p(n^{-1/2} ).
 \label{res2}
 \end{equation}
Combining (\ref{res1}) and (\ref{res2}), we can establish that 
\begin{equation}
N^{-1} \left( \hat{Y}_{\rm d, reg} - \hat{Y}_{\rm oracle} \right) =o_p( n^{-1/2} ).
\label{result}
\end{equation}
Therefore, asymptotic unbiasedness of the $\hat{Y}_{\rm d, reg}$ in (\ref{debiased}) is established. 
 \end{proof}

Theorem 1 means that the estimation error of $\hat{m}(\bx)$ can be safely ignored in the asymptotic sense. There are several advantages of the debiased two-phase regression estimator in (\ref{debiased}). 
\begin{enumerate}
\item Unlike the classical two-phase regression estimator using nonparametric regression, we can establish asymptotic unbiasedness and  $\sqrt{n}$-consistency. 
\item Even if we use the sample split, there is no efficiency loss. That is, the asymptotic variance is equal to 
$$ V \left( \hat{Y}_{\rm d, reg} \right) = V \left( \hat{Y}_1 \right) + E\left[ V \left\{ \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \left( y_i - m(\bx_i) \right) \mid A_1 \right\} \right] 
$$
where $m(\bx_i)$ is the probability limit of $\hat{m}(\bx_i)$. 
\item Variance estimation is also straightforward. We can compute 
$$ \hat{\eta}_i = \hat{m}(\bx_i) +\delta_i  \pi_{2i \mid 1}^{-1} I_i^{(a)} \left\{ y_i - \hat{m}^{(b)}(\bx_i) \right\} +\delta_i \pi_{2i \mid 1}^{-1} I_i^{(b)} \left\{ y_i - \hat{m}^{(a)} (\bx_i) \right\}$$
and apply to the usual variance estimation formula for the first-phase sample, 
where $I_i^{(a)}$ is the indicator function for $A_2^{(a)}$ such that $I_i^{(a)} = 1$ if $i \in A_2^{(a)}$ and $I_i^{(a)}=0$ otherwise. Also, $I_i^{(b)}= 1- I_i^{(a)}$. 
\end{enumerate}

My variance estimator is 
$$ \hat{V} = \frac{1}{n_1 (n_1-1) } \sum_{i \in A_1} \left( \hat{\eta}_i - \bar{\eta}_n \right)^2 $$

I also have an idea on how to implement the above debiased regression estimator using calibration. I will give more details once we are confident in the proposed idea. 
\end{document}

### 2024-05-08 final/reports/twophaseridge.tex
Updated twophaseridge.tex
Added simulations for debiased nonparametric regression. We still need to find a good setup to show the strength of our result.

@@
 
+\textcolor{red}{If the population total of $X$ is known, then the regression
+  estimator 
+  $$ \hat{\bar{Y}}_{reg} = \hat{\bar{Y}}_{HT} + ({\bar{X}} - \hat{\bar{X}}_2)' \hat \beta_2$$
+  can be used. In this case, $\hat{\beta}_2$ does not have the shrinkage
+  interpretation. Now, if $\bar{X}$ is unknown, we need to use an estimator. If
+  we use the best estimator $\hat{\bar{X}}_c$, then the residual is orthogonal.
+  Otherwise, if we use an inefficient estimator $\hat{\bar{X}}_1$, we pay its
+  price. It is more related to attenuation effect in the measurement error model
+  problem. 
+}
+
 \section{Ridge Regression}
@@
 
+\begin{itemize}
+  \item We could connect this to LASSO?
+  \item This could be a connection to Bayesian statistics and choosing priors on
+    our estimates?
+  \item Maybe this can help us choose priors for ridge regression better?
+\end{itemize}
+
+\section{Alternative idea for non-parametric regression estimation}
+
+One possible idea is to develop a debiased estimation under two-phase sampling.
+If the covariates are high dimensional or the regression employs nonparametric
+regression (such as spline or random forest, etc), then the resulting two-phase
+regression estimator can be biased. To correct for the bias, we can use the
+following approach.
+
+\begin{enumerate}
+\item Split the sample $A_2$ into two parts: $A_2= A_2^{(a)} \cup A_2^{(b)}$.
+  We can use SRS to split the sample, but other sampling designs can be used.
+
+\item Use the observations in $A_2^{(a)}$ only to obtain a predictor of $y_i$,
+  $\hat{f}^{(a)} (\bx_i)$. Also, use the observations in $A_2^{(b)}$ only to
+  obtain a predictor of $y_i$, $\hat{f}^{(b)} (\bx_i)$.
+
+\item Let 
+$$ \hat{f}(\bx_i) = \left( \hat{f}^{(a)} (\bx_i) + \hat{f}^{(b)} (\bx_i) \right)/ 2$$
+be the predictor combining two samples.
+
+\item The final debiased two-phase regression estimator is given by 
+\begin{equation}
+\hat{Y}_{\rm d, reg} = \sum_{i \in A_1} w_{1i}  \hat{f}(\bx_i) +  
+\sum_{i \in A_2^{(a)} } w_{1i} \pi_{2i \mid 1}^{-1} \left\{y_i - \hat{f}^{(b)} (\bx_i) \right\} +
+\sum_{i \in A_2^{(b)} } w_{1i} \pi_{2i \mid 1}^{-1} \left\{y_i - \hat{f}^{(a)} (\bx_i) \right\}
+\label{debiased}
+\end{equation}
+\end{enumerate}
+
+ If I am not mistaken, there are several advantages of the debiased two-phase
+ regression estimator in (\ref{debiased}).
+
+\begin{enumerate}
+\item Unlike the classical two-phase regression estimator using nonparametric
+  regression, we can establish asymptotic unbiasedness and
+  $\sqrt{n}$-consistency. 
+\item Even if we use the sample split, there is no efficiency loss. That is, the
+  asymptotic variance is equal to
+
+$$ V \left( \hat{Y}_{\rm d, reg} \right) = V \left( \hat{Y}_1 \right) + E\left[
+V \left\{ \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \left( y_i - f(\bx_i)
+\right) \mid A_1 \right\} \right] 
+$$
+
+where $f(\bx_i)$ is the probability limit of $\hat{f}(\bx_i)$.
+
+\item Variance estimation is also straightforward. We can compute
+
+$$ \hat{\eta}_i = \hat{f}(\bx_i) +\delta_i  \pi_{2i \mid 1}^{-1} I_i^{(a)}
+\left\{ y_i - \hat{f}^{(b)}(\bx_i) \right\} +\delta_i \pi_{2i \mid 1}^{-1}
+I_i^{(b)} \left\{ y_i - \hat{f}^{(a)} (\bx_i) \right\}$$
+
+and apply to the variance estimation formula for the first-phase sample, 
+where $I_i^{(a)}$ is the indicator function for $A_2^{(a)}$ such that $I_i^{(a)}
+= 1$ if $i \in A_2^{(a)}$ and $I_i^{(a)}=0$ otherwise. Also, $I_i^{(b)}= 1-
+I_i^{(a)}$. 
+\end{enumerate}
+
+I also have an idea on how to implement the above debiased regression estimator
+using calibration. I will give more details once we are confident in the
+proposed idea.
+
+\subsection{Simulation Study}
+
+We generate a finite population of size $N = 10,000$ from a superpopulation
+model of:
+
+$$\begin{aligned}
+X_i &\stackrel{ind}{\sim} N(0, 1) \\
+\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
+Y_i = m(x) + 0.3 \varepsilon_i
+\end{aligned}
+$$
+
+for $m(x) = 0.2 x + \sin(x)$.
+The Phase 1 sample is a simple random sample (SRS) of size $n_1 = 2000$
+from the finite population and the Phase 2 sample is a Poisson sample with the
+probability of selection into the Phase 2 sample from the Phase 1 sample being
+$\max(\min(0.7, \Phi(X_i - 0.5)), 0.02)$ where $\Phi$ is the CDF function of a
+normal distribution. These sampling probabilities yield $E[n_2] \approx 700$.
+
+The goal of this simulation is to estimate $\theta = E[Y]$, which is challenging
+because $Y$ is a nonlinear function of $X$ as seen in Figure~\ref{fig:popplot}
+and Figure~\ref{fig:p2plot}.
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{tables/popplot.png}
+  \caption{Population plot of relationship between $X$ and $Y$.}
+  \label{fig:popplot}
+\end{figure}
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{tables/p2plot.png}
+  \caption{Phase 2 plot of relationship between $X$ and $Y$.}
+  \label{fig:p2plot}
+\end{figure}
+
+We estimate $\theta$ using the estimator:
+
+$$\hat Y = \sum_{i \in A_1} \frac{\hat \eta_i}{\pi_{1i}}$$
+
+where $\hat \eta_i$ differs between three methods:
+
+\begin{itemize}
+  \item[1.] Oracle: $\hat \eta_i = m(x_i) + \frac{\delta_{2i|1}}{\pi_{2i|1}}(y_i
+    - m(x_i))$ and $m(x)$ is the true mean function of $Y$.
+  \item[2.] Naive: $\hat \eta_i = \hat m(x_i) + \frac{\delta_{2i|1}}{\pi_{2i|1}}(y_i
+    - \hat m(x_i))$ where $\hat m(x)$ is estimated using the data from the Phase 2
+    sample using nonparametric regression. 
+  \item[3.] Proposed: $\hat \eta_i = \hat m(x_i) + 
+    \frac{\delta_{2i|1}\delta_i^{(a)}}{\pi_{2i|1}}(y_i - \hat m^{(b)}(x_i)) +
+    \frac{\delta_{2i|1}\delta_i^{(b)}}{\pi_{2i|1}}(y_i - \hat m^{(a)}(x_i))$
+    where $\hat m(x) = \frac 1 2 (\hat m^{(a)}(x) + \hat m^{(b)}(x))$ and $\hat
+    m^{(a)}(x)$ is the nonparametric model estimated using $A_2^{(a)}$ and $\hat
+    m^{(b)}(x)$ is the nonparametric model estimated using $A_2^{(b)}$. We also
+    have $\delta_i^{(a)}$ and $\delta_i^{(b)}$ which denote is $i \in A_2$ was
+    part of $A_2^{(a)}$ or $A_2^{(b)}$ respectively.
+\end{itemize}
+
+For the Naive and Proposed methods, we use three nonparametric regression
+estimators: loess, random forest, and a smoothing spline. These functions come
+from the \texttt{stats}, \texttt{randomForest}, and \texttt{npreg} packages in
+\texttt{R}. To estimate the variance we use the following:
+
+$$\hat V = \hat S^2_\eta / n_1$$
+
+where $\hat S^2_\eta$ is the estimated variane of $\hat \eta$ in $A_1$.
+%\begin{table}[ht!]
+%  \begin{tabular}{lrr}
+%    \toprule
+%    Method & Estimated Variance & Notes \\
+%    \midrule
+%    Oracle & $\hat V = (\hat V_1 + \hat V_2) / N^2$ & 
+%    $\hat V_1 = N^2\left(n_1^{-1} - N^{-1}\right) \hat S_{\eta}$ \\
+%           & & $\hat V_2 = \sum_{i \in A_2} 
+%           \frac{(1 - \pi_{2i|1})(y_i - m(x_i))^2}{\pi_{1i} \pi_{2i|1}^2}$\\
+%    Naive & $\hat V = (\hat V_1 + \hat V_2) / N^2$ & Same as Oracle with $\hat m(x)$
+%    instead of $m(x)$.\\
+%    Proposed & $\hat V = (\hat V_1 + \hat V_{2a} + \hat V_{2b} + \hat V_{2ab}) / N^2$ 
+%             & $\hat V_1$ is the same as $\hat V_1$ from Naive. \\
+%             & & $\hat V_{2a} = \frac{1}{4}\sum_{i \in A_2}\sum_{i \in A_2}
+%             \frac{\Delta_{2ij|1} + \pi_{2ij|1}}{\pi_{1ij}\pi_{2ij|1}} 
+%             \frac{\varepsilon_i^{(b)}}{\pi_{2i|1}}
+%             \frac{\varepsilon_j^{(b)}}{\pi_{2j|1}}$\\
+%             & & $\hat V_{2b} = \frac{1}{4}\sum_{i \in A_2}\sum_{i \in A_2}
+%             \frac{\Delta_{2ij|1} + \pi_{2ij|1}}{\pi_{1ij}\pi_{2ij|1}} 
+%             \frac{\varepsilon_i^{(a)}}{\pi_{2i|1}}
+%             \frac{\varepsilon_j^{(a)}}{\pi_{2j|1}}$\\
+%             & & {\scriptsize$\hat V_{2ab} = 
+%             \sum\limits_{i \in A_2} \sum\limits_{i \in A_2} \frac{n_2^{(a)} 
+%             \varepsilon_i^{(b)} \varepsilon_j^{(a)}}{n_2
+%             \pi_{2ij|1}\pi_{1ij}} \left(\frac{n_2 - n_2^{(a)}}{n_2 - 1}
+%             I(i \neq j) - \frac{1 - n_2^{(a)}}{n_2}\right)$}\\
+%             & & $\varepsilon_i^{(a)} = (y_i - m^{(a)}(x_i))$,
+%             $\varepsilon_i^{(b)} = (y_i - m^{(b)}(x_i))$ \\
+%    \bottomrule
+%  \end{tabular}
+%\end{table}
+
+The results of this simulation for the mean are displayed in
+Table~\ref{tbl:npmean} and the variance results are in Table~\ref{tbl:npvar}.
+
+\begin{table}[ht!]
+  \centering
+  \input{tables/npreg_mean.tex}
+  \caption{This shows the results of our simulation. The Est column displays the
+  name of the estimator with examples of NaiveLoess indicating the Naive method
+  with the loess nonparametric estimator. The Bias column computes $E_{MC}(\hat
+  \theta) - \theta_N$ where $E_{MC}$ is the Monte Carlo mean of the 1000
+  simulations and $\theta_N$ is the true value from the population. The RMSE
+  column computes $\sqrt{B^{-1}\sum_{b = 1}^B (\hat \theta^{(b)} - \theta_N)^2}$
+  where $\hat \theta^{(b)}$ is the estimate of $\theta_N$ for the $b$th
+  iteration of the Monte Carlo sample. The EmpCI column indicates the fraction
+  of iterations for which $|\hat \theta^{(b)} - \theta_N| < 1.96 * \sqrt{\hat V^{(b)}}$
+  is true where $\hat V^{(b)}$ is the estimated variance of $\hat \theta^{(b)}$.
+  The final column of Ttest gives the test statistic of a test for the bias of the
+  estimator being zero.}
+  \label{tbl:npmean}
+\end{table}
+
+\begin{table}[ht!]
+  \centering
+  \input{tables/npreg_var.tex}
+  \caption{This displays the variance results from the simulation. The MCVar
+  colum is the Monte Carlo variance of $\hat \theta^{(b)}$ while the EstVar
+  column is the average of the estimated variance $\hat V^{(b)}$. The VarVar
+  column is the Monte Carlo variance of the variance estimator and the Ttest
+  column tests for equality of MCVar and EstVar.}
+  \label{tbl:npvar}
+\end{table}
+
+Overall, these results seem very promising. From Table~\ref{tbl:npmean}, we can
+see the there is evidence for bias in the naive algorithm in both the random
+forest and the spline method. Yet, this bias is no longer there under the
+proposed method and the empirical confidence intervals are approximately
+correct. A further investigation of the variance in Table~\ref{tbl:npvar}
+indicates that the estimated variance for the proposed method is too large when
+the regression function is either the random forest or smoothing spline, but we
+might be approximately correct.
+
+For this simulation, we are still doing well. The only problem is that the naive
+method is also working well (especially for the random forest). Maybe I can
+choose a more nonlinear function to estimate?
+
+\subsubsection{Simulation Update}
+
+I updated the simulation so that we use the same approach as
+\cite{wang2023statistical}. They use a multivariate approach with $X_{1i}$,
+$X_{2i}$, $X_{3i}$, and $X_{4i}$ being simulated independently from a $U(1, 3)$
+distribution. They have three models for $Y$,
+
+\begin{table}[ht!]
+\centering
+\begin{tabular}{lr}
+\toprule
+Model & $Y$ \\
+\midrule
+A & $Y_i = 3 + 2.5 x_{1i} + 2.75 x_{2i} + 2.5 x_{3i} + 2.25 x_{4i} + \sqrt{3} \varepsilon_i$ \\
+B & $Y_i = 3 + (1/35) x_{1i}^2 x_{2i}^3 x_{3i} + 0.1 x_{4i} + \sqrt{3} \varepsilon_i$\\
+C & $Y_i = 3 + (1/180) x_{1i}^2 x_{2i}^3 x_{3i} x_{4i}^2 + \sqrt{3} \varepsilon_i$ \\ 
+\bottomrule
+\end{tabular}
+\end{table}
+
+The first phase sample is a SRS with $n_1 = 1000$ and the second phase sample is
+a Poisson sample with $\pi_{2i|1} = \logistic(\bf x_i^T \bm \beta + 2.5)$ for $\bm
+\beta = (-1.1, 0.5, -0.25, -0.1)^T$. We have modified the spline method to be a
+cubic spline with 15 knots for each coordinate and implemented by the \texttt{mgcv}
+package. Table~\ref{tbl:np2mean} shows the results
+of the simulation under Model C. The results for Model B are similar.
+
+\begin{table}[ht!]
+  \centering
+  \input{tables/npsim2C1_mean.tex}
+  \caption{This shows the results of our simulation. The Est column displays the
+  name of the estimator with examples of NaiveLoess indicating the Naive method
+  with the loess nonparametric estimator. The Bias column computes $E_{MC}(\hat
+  \theta) - \theta_N$ where $E_{MC}$ is the Monte Carlo mean of the 1000
+  simulations and $\theta_N$ is the true value from the population. The RMSE
+  column computes $\sqrt{B^{-1}\sum_{b = 1}^B (\hat \theta^{(b)} - \theta_N)^2}$
+  where $\hat \theta^{(b)}$ is the estimate of $\theta_N$ for the $b$th
+  iteration of the Monte Carlo sample. The EmpCI column indicates the fraction
+  of iterations for which $|\hat \theta^{(b)} - \theta_N| < 1.96 * \sqrt{\hat V^{(b)}}$
+  is true where $\hat V^{(b)}$ is the estimated variance of $\hat \theta^{(b)}$.
+  The final column of Ttest gives the test statistic of a test for the bias of the
+  estimator being zero.}
+  \label{tbl:np2mean}
+\end{table}
+
+Unfortunately, with this simulation (and also the simulation with Model B but
+not shown here), the naive models are not biased for the loess and spline
+methods. It does look like we can debias the result for the random forest
+though.
+
+\newpage
+
+\bibliographystyle{chicago}
+\bibliography{references}
+
 \end{document}

### 2024-05-08 final/reports/twophaseridge_ov.tex
Added simulations for debiased nonparametric regression. We still need to find a good setup to show the strength of our result.

--- 2024-05-04 final/reports/twophaseridge.tex
@@
 
+\textcolor{red}{If the population total of $X$ is known, then the regression estimator $$ \hat{\bar{Y}}_{reg} 
+  = \hat{\bar{Y}}_{HT} + ({\bar{X}} - \hat{\bar{X}}_2)' \hat \beta_2$$
+  can be used.
+In this case, $\hat{\beta}_2$ does not have the shrinkage interpretation.
+Now, if $\bar{X}$ is unknown, we need to use an estimator.
+If we use the best estimator $\hat{\bar{X}}_c$, then the residual is orthogonal.
+Otherwise, if we use an inefficient estimator $\hat{\bar{X}}_1$, we pay its price.
+It is more related to attenuation effect in the measurement error model problem.
+  }
 \section{Ridge Regression}
@@
 
+\begin{itemize}
+  \item We could connect this to LASSO?
+  \item This could be a connection to Bayesian statistics and choosing priors on
+    our estimates?
+  \item Maybe this can help us choose priors for ridge regression better?
+\end{itemize}
+
+\section{Alternative idea for non-parametric regression estimation}
+
+One possible idea is to develop a debiased estimation under two-phase sampling.
+If the covariates are high dimensional or the regression employs nonparametric regression (such as spline or random forest, etc), then the resulting two-phase regression estimator can be biased.
+To correct for the bias, we can use the following approach.
+
+\begin{enumerate}
+\item Split the sample $A_2$ into two parts: $A_2= A_2^{(a)} \cup A_2^{(b)}$.
+We can use SRS to split the sample, but other sampling designs can be used. 
+\item Use the observations in $A_2^{(a)}$ only to obtain a predictor of $y_i$, $\hat{f}^{(a)} (\bx_i)$.
+Also, use the observations in $A_2^{(b)}$ only to obtain a predictor of $y_i$, $\hat{f}^{(b)} (\bx_i)$.
+\item Let 
+$$ \hat{f}(\bx_i) = \left( \hat{f}^{(a)} (\bx_i) + \hat{f}^{(b)} (\bx_i) \right)/ 2$$
+be the predictor combining two samples. 
+\item The final debiased two-phase regression estimator is given by 
+\begin{equation}
+\hat{Y}_{\rm d, reg} = \sum_{i \in A_1} w_{1i}  \hat{f}(\bx_i) +  \sum_{i \in A_2^{(a)} } w_{1i} \pi_{2i \mid 1}^{-1}   \left\{y_i - \hat{f}^{(b)} (\bx_i) \right\} +\sum_{i \in A_2^{(b)} } w_{1i} \pi_{2i \mid 1}^{-1}   \left\{y_i - \hat{f}^{(a)} (\bx_i) \right\}
+\label{debiased}
+\end{equation}
+\end{enumerate}
+ If I am not mistaken, there are several advantages of the debiased two-phase regression estimator in (\ref{debiased}). 
+\begin{enumerate}
+\item Unlike the classical two-phase regression estimator using nonparametric regression, we can establish asymptotic unbiasedness and  $\sqrt{n}$-consistency. 
+\item Even if we use the sample split, there is no efficiency loss. That is, the asymptotic variance is equal to 
+$$ V \left( \hat{Y}_{\rm d, reg} \right) = V \left( \hat{Y}_1 \right) + E\left[ V \left\{ \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \left( y_i - f(\bx_i) \right) \mid A_1 \right\} \right] 
+$$
+where $f(\bx_i)$ is the probability limit of $\hat{f}(\bx_i)$. 
+\item Variance estimation is also straightforward. We can compute 
+$$ \hat{\eta}_i = \hat{f}(\bx_i) +\delta_i  \pi_{2i \mid 1}^{-1} I_i^{(a)} \left\{ y_i - \hat{f}^{(b)}(\bx_i) \right\} +\delta_i \pi_{2i \mid 1}^{-1} I_i^{(b)} \left\{ y_i - \hat{f}^{(a)} (\bx_i) \right\}$$
+and apply to the variance estimation formula for the first-phase sample, 
+where $I_i^{(a)}$ is the indicator function for $A_2^{(a)}$ such that $I_i^{(a)} = 1$ if $i \in A_2^{(a)}$ and $I_i^{(a)}=0$ otherwise. Also, $I_i^{(b)}= 1- I_i^{(a)}$. 
+\end{enumerate}
+
+I also have an idea on how to implement the above debiased regression estimator using calibration. I will give more details once we are confident in the proposed idea. 
 \end{document}

### 2024-05-08 src/explore/20240506-msdcsim.qmd
Added simulations for debiased nonparametric regression. We still need to find a good setup to show the strength of our result.

--- 2024-04-27 src/explore/20240425-nndcsim.qmd
@@
 ---
-title: "Non-Nested Two Phase Debiased Calibration Simulation"
+title: "Multi-Source Debiased Calibration Simulation"
 author: "Caleb Leedy"
@@
 
-This report simulates non-nested two-phase sampling with the debiased
+This report simulates multi-source sampling with the debiased
 calibration of @kwon2024debiased. We check for the following:
@@
 n1_obs <- 1000
-n2_obs <- 500
+n2_obs <- 2000
 
@@
 
-# source("GEcalib_code.R")
-
 ```
 
-We have the following superpopulation model:
+We have the following superpopulation model with $N=$`{r} N_obs` elements:
 
@@
 Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
-\pi_{1i} &= \Phi(-x_{3i} - 2)\\
-\pi_{2i|1} &= x_{4i}.
+\pi_{0i} &= \Phi(-x_{1i} - 2)\\
+\pi_{1i} &= |x_{1i}| / \sum_{i \in U} |x_{i1}|\\
+\pi_{2i} &= n_2 / N.\\
 \end{aligned}
@@
 
-and the non-nested two-phase sampling scheme is to select a SRS of 
-size $n = $`{r} n1_obs`
-from a finite population of size $N = $`{r} N_obs` for the Phase 1 sample,
-$A_1$, and a Poisson sample with expected size $r =$`{r} n2_obs` for the Phase 2
-sample, $A_2$. We only observe $Y$ in the Phase 2 sample, but we observe 
-$(X_1, X_2)$ in both $A_1$ and $A_2$.
+We observe the following columns in each sample
+
+| Sample | $X_1$ | $X_2$ | $X_3$ |  $Y$ |
+|:-------|------:|------:|------:|-----:|
+| $A_0$  | ✓     | ✓     | ✓     |  ✓   |  
+| $A_1$  | ✓     | ✓     |       |      |  
+| $A_2$  | ✓     |       | ✓     |      |
+
+For the sampling mechanism both $A_0$ and $A_1$ are selected using a Poisson
+sample with response probabilities $\pi_{0i}$ and $\pi_{1i}$ respectively. The
+sample $A_2$ is a simple random sample.
 
@@
 #' information about if an element is part of the Phase 1 or Phase 2 sample.
-update_pop <- function(pop_df, p1_formula, p2_formula) {
+update_pop <- function(pop_df, p0_formula, p1_formula, p2_formula) {
 
+  pi0 <- eval(parse_expr(p0_formula))
   pi1 <- eval(parse_expr(p1_formula))
@@
   pop_df %>%
-    mutate(pi1 = pi1, pi2 = pi2)
+    mutate(pi0 = pi0, pi1 = pi1, pi2 = pi2)
 }
@@
 #' population with selection probabilites.
-gen_samps <- function(upop_df, p1_type, p2_type) {
+gen_samps <- function(upop_df, p0_type, p1_type, p2_type) {
+
+  if (p0_type == "poisson") {
+    del0 <- rbinom(nrow(upop_df), 1, upop_df$pi0)
+  } else if (p1_type == "srs") {
+    ind <- sample(1:nrow(upop_df), size = upop_df$pi0[1] * nrow(upop_df))
+    del0 <- as.numeric(1:nrow(upop_df) %in% ind)
+  } else {
+    stop("We have only implemented poisson and srs for Sample 0.")
+  }
+
+  p0_df <- mutate(upop_df, del0 = del0) %>% filter(del0 == 1)
 
@@
   } else {
-    stop("We have only implemented poisson and srs for Phase 1.")
+    stop("We have only implemented poisson and srs for Sample 1.")
   }
@@
   } else {
-    stop("We have only implemented poisson and srs for Phase 2.")
+    stop("We have only implemented poisson and srs for Sample 2.")
   }
@@
 
-  return(list(p1_df, p2_df))
+  return(list(p0_df, p1_df, p2_df))
 }
@@
 ```
-
-```{r}
-#| label: run MCMC
-
-set.seed(1)
-pop_df <- gen_pop(N_obs)
-true_theta <- mean(pop_df$Y)
-
-upop_df <-
-  update_pop(
-    pop_df,
-    "n1_obs / N_obs",
-    "ifelse(pt(-pop_df$X1 + 1, 3) > 0.9, 0.9, pt(-pop_df$X1 + 1, 3))"
-  )
-
-gen_samps(upop_df, "srs", "poisson")
-
-```

### 2024-05-08 src/explore/20240506-npregest.qmd
Added simulations for debiased nonparametric regression. We still need to find a good setup to show the strength of our result.

---
title: "Debiased Nonparametric Regression Estimation"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    code-fold: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

# Summary

This report implements a simulation study in which we assess the bias of a
nonparametric regression estimator. We compare an oracle regression estimate,
naive two-phase nonparametric regression, and our proposed method--debiased
nonparametric regression estimation.

# Simulation Setup

```{r}
#| label: setup

B_sims <- 1000

N_obs <- 10000
n1_obs <- 2000
n2_obs <- 1000

library(dplyr)
library(doParallel)
library(doRNG)
library(npreg)
library(mgcv)
library(randomForest)

```

We generate a finite population of size $N =$`{r} N_obs` from a superpopulation
model of:

$$\begin{aligned}
X_i &\stackrel{ind}{\sim} N(0, 1) \\
\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
Y_i = 0.2 x + \sin(x) + 0.3 \varepsilon_i
\end{aligned}
$$

The Phase 1 sample is a simple random sample (SRS) of size $n_1 =$`{r} n1_obs`
from the finite population and the Phase 2 sample is a Poisson sample with the
probability of selection into the Phase 2 sample from the Phase 1 sample being
$\Phi_3(X_i)$ where $\Phi_3$ is the CDF function of a t-distribution with 3
degrees of freedom.

```{r}
#| label: generate data and samples

gen_pop <- function(N_obs, n1_obs) {
  x <- runif(N_obs, 0, 8 * pi)
  eps <- rnorm(N_obs)
  y <- 0.2 * x + sin(x) + 0.3 * eps

  pi1 <- n1_obs / N_obs
  z <- (x - mean(x)) / sd(x)
  pi2 <- ifelse(pnorm(z - 0.5) > 0.7, 0.7, pnorm(z - 0.5))
  pi2 <- ifelse(pi2 < 0.02, 0.02, pi2)

  return(tibble(X = x, Y = y, pi1 = pi1, pi2 = pi2))

}

gen_samps <- function(pop_df, p1_samp = "srs", p2_samp = "poisson") {

  if (p1_samp == "srs") {
    ind <- sample(1:nrow(pop_df), size = pop_df$pi1[1] * nrow(pop_df))
    p1_df <- filter(pop_df, row_number() %in% ind) %>%
      mutate(del1 = 1)
  } else {
    stop("p1_samp must be srs.")
  }

  if (p2_samp == "poisson") {
    del2 <- rbinom(nrow(p1_df), 1, p1_df$pi2)
    p1_df <- mutate(p1_df, del2 = del2)
    p2_df <- filter(p1_df, del2 == 1)
  } else {
    stop("p2_samp must be poisson")
  }

  return(list(p1_df, p2_df))
}

```

```{r}
#| labels: estimators

oracle_est <- function(p1_df, p2_df, N_obs) {

  # Steps:
  # 1. Get mhat.
  # 2. Predict missing values
  # 3. Estimate variance via linearization

  # 1. Get mhat.
  mx <- 0.2 * p1_df$X + sin(p1_df$X)
  mx2 <- 0.2 * p2_df$X + sin(p2_df$X)

  # 2. Predict missing values
  eta <- mx + p1_df$del2 / p1_df$pi2 * (p1_df$Y - mx)
  theta <- sum(eta / p1_df$pi1) / N_obs

  # 3. Estimate variance via linearization
  v1 <- (1 / nrow(p1_df) - 1 / N_obs) * var(eta)
  v2 <- sum(1 / (p2_df$pi1 * p2_df$pi2^2) * 
           (1 - p2_df$pi2) * (p2_df$Y - mx2)^2) / N_obs^2

  vhat <- (v1 + v2)

  return(list(theta, vhat))
}

naive_npreg_est <- function(p1_df, p2_df, N_obs, type = "loess") {

  # Steps:
  # 1. Estimate mhat.
  # 2. Predict missing values
  # 3. Estimate variance via linearization

  # 1. Estimate mhat.
  if (type == "loess") {
    mod <- loess(Y ~ X, p2_df, control = loess.control(surface="direct"))
  } else if (type == "randomforest") {
    mod <- randomForest(Y ~ X, p2_df)
  } else if (type == "spline") {
    # mod <- ss(x = p2_df$X, y = p2_df$Y)
    # mx <- predict(mod, x = p1_df$X)$y
    # mx2 <- predict(mod, x = p2_df$X)$y
    mod <- gam(Y ~ s(X, k = 15, bs = "cr"), data = p2_df)
  } else if (type == "linear") {
    mod <- lm(Y ~ X, data = p2_df)
  }

  mx <- predict(mod, p1_df)
  mx2 <- predict(mod, p2_df)
  # 2. Predict missing values
  eta <- mx + p1_df$del2 / p1_df$pi2 * (p1_df$Y - mx)
  theta <- sum(eta / p1_df$pi1) / N_obs

  # 3. Estimate variance via linearization
  v1 <- (1 / nrow(p1_df) - 1 / N_obs) * var(eta)
  v2 <- sum(1 / (p2_df$pi1 * p2_df$pi2^2) * 
           (1 - p2_df$pi2) * (p2_df$Y - mx2)^2) / N_obs^2

  vhat <- (v1 + v2)

  return(list(theta, vhat))
}

ss_npreg_est <- function(p1_df, p2_df, N_obs, type = "loess") {

  # Steps:
  # 0. Split A_2
  # 1. Estimate mhat in both splits
  # 2. Predict missing values
  # 3. Estimate variance via linearization

  # 0. Split A_2
  inda <- sample(1:nrow(p2_df), size = round(nrow(p2_df) / 2))
  p2a_df <- filter(p2_df, row_number() %in% inda)
  p2b_df <- filter(p2_df, !(row_number() %in% inda))
  p2_df <- p2_df %>%
    mutate(del2a = row_number() %in% inda) %>%
    mutate(del2b = !(row_number() %in% inda))

  p1_df <- p1_df %>%
    left_join(p2_df, by = join_by(X, Y, pi1, pi2, del1, del2)) %>%
    mutate(del2a = ifelse(is.na(del2a), 0, del2a)) %>%
    mutate(del2b = ifelse(is.na(del2b), 0, del2b))

  # 1. Estimate mhat in both splits
  if (type == "loess") {
    moda <- loess(Y ~ X, p2a_df, control = loess.control(surface="direct"))
    modb <- loess(Y ~ X, p2b_df, control = loess.control(surface="direct"))
  } else if (type == "randomforest") {
    moda <- randomForest(Y ~ X, p2a_df)
    modb <- randomForest(Y ~ X, p2b_df)
  } else if (type == "spline") {
    # moda <- ss(x = p2a_df$X, y = p2a_df$Y)
    # modb <- ss(x = p2b_df$X, y = p2b_df$Y)
    # ma_2 <- predict(moda, x = p2_df$X)$y
    # mb_2 <- predict(modb, x = p2_df$X)$y
    # ma_1 <- predict(moda, x = p1_df$X)$y
    # mb_1 <- predict(modb, x = p1_df$X)$y
    # mx <- (ma_1 + mb_1) / 2
    moda <- gam(Y ~ s(X, k = 15, bs = "cr"), data = p2a_df)
    modb <- gam(Y ~ s(X, k = 15, bs = "cr"), data = p2b_df)
  } else if (type == "linear") {
    moda <- lm(Y ~ X, data = p2a_df)
    modb <- lm(Y ~ X, data = p2b_df)
  }

  ma_2 <- predict(moda, p2_df)
  mb_2 <- predict(modb, p2_df)
  ma_1 <- predict(moda, p1_df)
  mb_1 <- predict(modb, p1_df)
  mx <- (ma_1 + mb_1) / 2

  # 2. Predict missing values
  eta <- 
    mx + 
    p1_df$del2a / p1_df$pi2 * (p1_df$Y - mb_1) + 
    p1_df$del2b / p1_df$pi2 * (p1_df$Y - ma_1)

  # plot(p1_df$Y, eta)

  theta <- sum(eta / p1_df$pi1) / N_obs

  # 3. Estimate variance via linearization
  vhat <- var(eta) / nrow(p1_df)
  # v0 <- (1 / nrow(p1_df) - 1 / N_obs) * var(eta)

  # p2a_df <- p2a_df %>%
  #   mutate(pi2a = pi2 * nrow(p2a_df) / nrow(p2_df))
  # p2b_df <- p2b_df %>%
  #   mutate(pi2b = pi2 * nrow(p2b_df) / nrow(p2_df))

  # v1_t <- matrix(NA, nrow = nrow(p2_df), ncol = nrow(p2_df))
  # v2_t <- matrix(NA, nrow = nrow(p2_df), ncol = nrow(p2_df))
  # cov_t <- matrix(NA, nrow = nrow(p2_df), ncol = nrow(p2_df))

  # # Since Phase 1 is SRS
  # p1ij <- nrow(p1_df) / N_obs * (nrow(p1_df) - 1) / (N_obs - 1)
  # eps_a <- p2_df$Y - ma_2
  # eps_b <- p2_df$Y - mb_2

  # n2a <- nrow(p2a_df)
  # n2b <- nrow(p2b_df)
  # n2 <- nrow(p2_df)

  # for (i in 1:nrow(p2_df)) {
  #   for (j in 1:nrow(p2_df)) {
  #     if (i == j) {
  #       v1_t[i, j] <- 
  #       1 / 4 * (2 - p2_df$pi2[i]) / p2_df$pi1[i] * eps_b[i]^2 / p2_df$pi2[i]^2
  #       v2_t[i, j] <- 
  #       1 / 4 * (2 - p2_df$pi2[i]) / p2_df$pi1[i] * eps_a[i]^2 / p2_df$pi2[i]^2
  #       cov_t[i, j] <- 
  #       1/(p2_df$pi2[i] * p2_df$pi1[i]) * n2a/n2 * (n2a - 1)/n2 * eps_b[i] * eps_a[i]
  #     } else {
  #       v1_t[i, j] <- 1 / 4 * (p2_df$pi2[i]) / (p2_df$pi1[i] * p2_df$pi1[j]) *
  #         eps_b[i] * eps_b[j] / (p2_df$pi2[i] * p2_df$pi2[j])
  #       v2_t[i, j] <- 1 / 4 * 1 / (p2_df$pi1[i] * p2_df$pi1[j]) *
  #         eps_a[i] * eps_a[j] / (p2_df$pi2[i] * p2_df$pi2[j])
  #       cov_t[i, j] <- 1 / (p2_df$pi2[i] * p2_df$pi1[i]) * n2a / n2 * 
  #       ((n2 - n2a) / (n2 - 1) + (n2a - 1)/n2) * eps_b[i] * eps_a[i]
  #     }
  #   }
  # }

  # v1 <- sum(as.numeric(v1_t)) / N_obs^2
  # v2 <- sum(as.numeric(v2_t)) / N_obs^2
  # v3 <- sum(as.numeric(cov_t)) / N_obs^2

  # vhat <- (v0 + v1 + v2 + v3)

  return(list(theta, vhat))
}

```

```{r}
#| label: MCMC

set.seed(1)
pop_df <- gen_pop(N_obs, n1_obs)
true_theta <- mean(pop_df$Y)

# png("tables/popplot.png")
# plot(pop_df$X, pop_df$Y)
# dev.off()

clust <- makeCluster(min(detectCores() - 2, 100), outfile = "")
registerDoParallel(clust)

mc_res <- 
  foreach(iter=1:B_sims,
          .packages = c("dplyr", "randomForest", "mgcv")) %dopar% {

  if (iter %% 100 == 0) {
    print(paste0("Iter: ", iter))
  }

  samps <- gen_samps(pop_df)
  p1_df <- samps[[1]]
  p2_df <- samps[[2]]

  # png("tables/p2plot.png")
  # plot(p2_df$X, p2_df$Y)
  # dev.off()

  oest <- oracle_est(p1_df, p2_df, N_obs)
  nest_lm <- naive_npreg_est(p1_df, p2_df, N_obs, type = "linear")
  pest_lm <- ss_npreg_est(p1_df, p2_df, N_obs, type = "linear")
  nest_lo <- naive_npreg_est(p1_df, p2_df, N_obs, type = "loess")
  pest_lo <- ss_npreg_est(p1_df, p2_df, N_obs, type = "loess")
  nest_rf <- naive_npreg_est(p1_df, p2_df, N_obs, type = "randomforest")
  pest_rf <- ss_npreg_est(p1_df, p2_df, N_obs, type = "randomforest")
  nest_sp <- naive_npreg_est(p1_df, p2_df, N_obs, type = "spline")
  pest_sp <- ss_npreg_est(p1_df, p2_df, N_obs, type = "spline")

  return(tibble(
      Est = c("Oracle",
              "NaiveLinear", "PropLinear",
              "NaiveLoess", "PropLoess",
              "NaiveRF", "PropRF",
              "NaiveSpline", "PropSpline"),
      Theta = c(oest[[1]],
                nest_lm[[1]], pest_lm[[1]],
                nest_lo[[1]], pest_lo[[1]],
                nest_rf[[1]], pest_rf[[1]],
                nest_sp[[1]], pest_sp[[1]]),
      Var = c(oest[[2]],
              nest_lm[[2]], pest_lm[[2]],
              nest_lo[[2]], pest_lo[[2]],
              nest_rf[[2]], pest_rf[[2]],
              nest_sp[[2]], pest_sp[[2]]),
      Iter = iter)
  )

} %>% bind_rows()

stopCluster(clust)

```

```{r}
#| label: Analyse MCMC

# Find any failed estimates
which(is.na(mc_res$Theta))
which(is.na(mc_res$Var))

# Analyze Mean
mc_res %>%
  mutate(err = Theta - true_theta) %>%
  mutate(CI = abs(err) < qnorm(0.975) * sqrt(Var)) %>%
  group_by(Est) %>%
  summarize(Bias = mean(err, na.rm = TRUE),
            RMSE = sqrt(mean(err^2, na.rm = TRUE)),
            EmpCI = mean(CI, na.rm = TRUE),
            sdest = sd(Theta, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Ttest = abs(Bias) / sqrt(sdest^2 / B_sims)) %>%
  select(-sdest) %>%
  mutate(Ind = case_when(
    Est == "NaiveLinear" ~ 2,
    Est == "NaiveLoess" ~ 4,
    Est == "NaiveRF" ~ 6,
    Est == "NaiveSpline" ~ 8,
    Est == "PropLinear" ~ 3,
    Est == "PropLoess" ~ 5,
    Est == "PropRF" ~ 7,
    Est == "PropSpline" ~ 9,
    Est == "Oracle" ~ 1
  )) %>%
  arrange(Ind) %>%
  select(-Ind)

%>%
  knitr::kable(digits = 3, "latex", booktabs = TRUE, linesep = "") %>%
  cat(file = "tables/npreg_mean.tex")

# Analyze Variance
mc_res %>%
  group_by(Est) %>%
  summarize(MCVar = var(Theta, na.rm = TRUE),
            EstVar = mean(Var, na.rm = TRUE),
            VarVar = var(Var, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Ttest = abs(MCVar - EstVar) / sqrt(VarVar / B_sims)) %>%
  mutate(Ind = case_when(
    Est == "NaiveLoess" ~ 2,
    Est == "NaiveRF" ~ 4,
    Est == "NaiveSpline" ~ 6,
    Est == "PropLoess" ~ 3,
    Est == "PropRF" ~ 5,
    Est == "PropSpline" ~ 7,
    Est == "Oracle" ~ 1
  )) %>%
  arrange(Ind) %>%
  select(-Ind)

%>%
  knitr::kable(digits = c(1, 4, 4, 7, 1), "latex", booktabs = TRUE, linesep = "") %>%
  cat(file = "tables/npreg_var.tex")

```

```{r}
#| label: individual plots

```

### 2024-05-08 src/explore/20240508-npsim2.qmd
Added simulations for debiased nonparametric regression. We still need to find a good setup to show the strength of our result.

---
title: "Debiased Nonparametric Regression Estimation"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    code-fold: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

# Summary

This report implements the simulation study in @wang2023statistical
in which we assess the bias of a
nonparametric regression estimator. We compare an oracle regression estimate,
naive two-phase nonparametric regression, and our proposed method--debiased
nonparametric regression estimation. This is very similar to
`20240506-npregest.qmd` except with a different setup.

# Simulation Setup

```{r}
#| label: setup

B_sims <- 1000

N_obs <- 10000
n1_obs <- 500

library(dplyr)
library(doParallel)
library(doRNG)
library(randomForest)
library(mgcv)

```

We generate a finite population of size $N =$`{r} N_obs` from a superpopulation
model of:

$$\begin{aligned}
X_{1i} &\stackrel{ind}{\sim} U(1, 3) \\
X_{2i} &\stackrel{ind}{\sim} U(1, 3) \\
X_{3i} &\stackrel{ind}{\sim} U(1, 3) \\
X_{4i} &\stackrel{ind}{\sim} U(1, 3) \\
\varepsilon_i &\stackrel{ind}{\sim} N(0, 1),
\end{aligned}
$$

The construction of $Y$ depends on the chosen model,

| Model | $Y$ |
|:------|----:|
| A     | $Y_i = 3 + 2.5 x_{1i} + 2.75 x_{2i} + 2.5 x_{3i} + 2.25 x_{4i} + \sqrt{3} \varepsilon_i$|
| B     | $Y_i = 3 + (1 / 35) x_{1i}^2 x_{2i}^3 x_{3i} + 0.1 x_{4i} + \sqrt{3} \varepsilon_i $|
| C     | $Y_i = 3 + (1 / 180) x_{1i}^2 x_{2i}^3 x_{3i} x_{4i}^2 + \sqrt{3} \varepsilon_i$ |

The Phase 1 sample is a simple random sample (SRS) of size $n_1 =$`{r} n1_obs`
from the finite population and the Phase 2 sample is a Poisson sample with the
probability of selection into the Phase 2 sample from the Phase 1 sample being
$\text{logit}(\bf x_i^T \bm \beta + 2.5)$ where $\bm \beta = (-1.1, 0.5, -0.25,
-0.1)^T$ or $\text{logit}(-0.3 + 0.7 x_{1i}^2 - 0.5 x_{2i} - 0.25 x_{3i} -0.25
x_{4i})$.

```{r}
#| label: generate data and samples

gen_pop <- function(N_obs, n1_obs, model, response) {

  x1 <- runif(N_obs, 1, 3)
  x2 <- runif(N_obs, 1, 3)
  x3 <- runif(N_obs, 1, 3)
  x4 <- runif(N_obs, 1, 3)
  eps <- rnorm(N_obs)

  if (model == "A") {
    y <- 3 + 2.5 * x1 + 2.75 * x2 + 2.5 * x3 + 2.25 * x4 + sqrt(3) * eps
  } else if (model == "B") {
    y <- 3 + (1 / 35) * x1^2 * x2^3 * x3 + 0.1 * x4 + sqrt(3) * eps
  } else if (model == "C") {
    y <- 3 + (1 / 180) * x1^2 * x2^3 * x3 * x4^2 + sqrt(3) * eps
  } else {
    stop("We must have model be A, B, or C.")
  }

  pi1 <- n1_obs / N_obs

  if (response == 1) {
    beta <- c(-1.1, 0.5, -0.25, -0.1)
    xbeta <- 2.5 + x1 * beta[1] + x2 * beta[2] + x3 * beta[3] + x4 * beta[4]
  } else if (response == 2) {
    beta <- c(0.7, -0.5, -0.25, -0.25)
    xbeta <- -0.3 + x1^2 * beta[1] + x2 * beta[2] + x3 * beta[3] + x4 * beta[4]
  } else {
    stop("response must be either 1 or 2.")
  }

  expit <- function(x) {1 / (1 + exp(-x))}
  pi2 <- expit(xbeta)

  return(tibble(X1 = x1, X2 = x2, X3 = x3, X4 = x4, Y = y, pi1 = pi1, pi2 = pi2))

}

gen_samps <- function(pop_df, p1_samp = "srs", p2_samp = "poisson") {

  if (p1_samp == "srs") {
    ind <- sample(1:nrow(pop_df), size = pop_df$pi1[1] * nrow(pop_df))
    p1_df <- filter(pop_df, row_number() %in% ind) %>%
      mutate(del1 = 1)
  } else {
    stop("p1_samp must be srs.")
  }

  if (p2_samp == "poisson") {
    del2 <- rbinom(nrow(p1_df), 1, p1_df$pi2)
    p1_df <- mutate(p1_df, del2 = del2)
    p2_df <- filter(p1_df, del2 == 1)
  } else {
    stop("p2_samp must be poisson")
  }

  return(list(p1_df, p2_df))
}

```

```{r}
#| labels: estimators

oracle_est <- function(p1_df, p2_df, N_obs, model) {

  # Steps:
  # 1. Get mhat.
  # 2. Predict missing values
  # 3. Estimate variance via linearization

  # 1. Get mhat.
  if (model == "A") {
    mx <- 3 + 2.5 * p1_df$X1 + 2.75 * p1_df$X2 + 2.5 * p1_df$X3 + 2.25 * p1_df$X4 
    mx2 <- 3 + 2.5 * p2_df$X1 + 2.75 * p2_df$X2 + 2.5 * p2_df$X3 + 2.25 * p2_df$X4 
  } else if (model == "B") {
    mx <- 3 + (1 / 35) * p1_df$X1^2 * p1_df$X2^3 * p1_df$X3 + 0.1 * p1_df$X4
    mx2 <- 3 + (1 / 35) * p2_df$X1^2 * p2_df$X2^3 * p2_df$X3 + 0.1 * p2_df$X4
  } else if (model == "C") {
    mx <- 3 + (1 / 180) * p1_df$X1^2 * p1_df$X2^3 * p1_df$X3 * p1_df$X4^2
    mx2 <- 3 + (1 / 180) * p2_df$X1^2 * p2_df$X2^3 * p2_df$X3 * p2_df$X4^2
  } else {
    stop("We must have model be A, B, or C.")
  }

  # 2. Predict missing values
  eta <- mx + p1_df$del2 / p1_df$pi2 * (p1_df$Y - mx)
  theta <- sum(eta / p1_df$pi1) / N_obs

  # 3. Estimate variance via linearization
  v1 <- (1 / nrow(p1_df) - 1 / N_obs) * var(eta)
  v2 <- sum(1 / (p2_df$pi1 * p2_df$pi2^2) * 
           (1 - p2_df$pi2) * (p2_df$Y - mx2)^2) / N_obs^2

  vhat <- (v1 + v2)

  return(list(theta, vhat))
}

naive_npreg_est <- function(p1_df, p2_df, N_obs, type = "loess") {

  # Steps:
  # 1. Estimate mhat.
  # 2. Predict missing values
  # 3. Estimate variance via linearization

  # 1. Estimate mhat.
  if (type == "loess") {
    mod <- loess(Y ~ X1 + X2 + X3 + X4, p2_df,
                 control = loess.control(surface="direct"))
  } else if (type == "randomforest") {
    mod <- randomForest(Y ~ X1 + X2 + X3 + X4, p2_df)
  } else if (type == "spline") {
    # mod <- gam(Y ~ s(X1, X2, X3, X4), data = p2_df)
    mod <- gam(Y ~ s(X1, k = 15, bs = "cr") +
                   s(X2, k = 15, bs = "cr") +
                   s(X3, k = 15, bs = "cr") +
                   s(X4, k = 15, bs = "cr"), data = p2_df)
  }

  mx <- predict(mod, p1_df)
  mx2 <- predict(mod, p2_df)

  # 2. Predict missing values
  eta <- mx + p1_df$del2 / p1_df$pi2 * (p1_df$Y - mx)
  theta <- sum(eta / p1_df$pi1) / N_obs

  # 3. Estimate variance via linearization
  v1 <- (1 / nrow(p1_df) - 1 / N_obs) * var(eta)
  v2 <- sum(1 / (p2_df$pi1 * p2_df$pi2^2) * 
           (1 - p2_df$pi2) * (p2_df$Y - mx2)^2) / N_obs^2

  vhat <- (v1 + v2)

  return(list(theta, vhat))
}

ss_npreg_est <- function(p1_df, p2_df, N_obs, type = "loess") {

  # Steps:
  # 0. Split A_2
  # 1. Estimate mhat in both splits
  # 2. Predict missing values
  # 3. Estimate variance via linearization

  # 0. Split A_2
  inda <- sample(1:nrow(p2_df), size = round(nrow(p2_df) / 2))
  p2a_df <- filter(p2_df, row_number() %in% inda)
  p2b_df <- filter(p2_df, !(row_number() %in% inda))
  p2_df <- p2_df %>%
    mutate(del2a = row_number() %in% inda) %>%
    mutate(del2b = !(row_number() %in% inda))

  p1_df <- p1_df %>%
    left_join(p2_df, by = join_by(X1, X2, X3, X4, Y, pi1, pi2, del1, del2)) %>%
    mutate(del2a = ifelse(is.na(del2a), 0, del2a)) %>%
    mutate(del2b = ifelse(is.na(del2b), 0, del2b))

  # 1. Estimate mhat in both splits
  if (type == "loess") {
    moda <- loess(Y ~ X1 + X2 + X3 + X4, p2a_df,
                  control = loess.control(surface="direct"))
    modb <- loess(Y ~ X1 + X2 + X3 + X4, p2b_df,
                  control = loess.control(surface="direct"))
  } else if (type == "randomforest") {
    moda <- randomForest(Y ~ X1 + X2 + X3 + X4, p2a_df)
    modb <- randomForest(Y ~ X1 + X2 + X3 + X4, p2b_df)
  } else if (type == "spline") {
    # moda <- gam(Y ~ s(X1, X2, X3, X4), data = p2a_df)
    # modb <- gam(Y ~ s(X1, X2, X3, X4), data = p2b_df)
    moda <- gam(Y ~ s(X1, k = 15, bs = "cr") +
                   s(X2, k = 15, bs = "cr") +
                   s(X3, k = 15, bs = "cr") +
                   s(X4, k = 15, bs = "cr"), data = p2a_df)
    modb <- gam(Y ~ s(X1, k = 15, bs = "cr") +
                   s(X2, k = 15, bs = "cr") +
                   s(X3, k = 15, bs = "cr") +
                   s(X4, k = 15, bs = "cr"), data = p2b_df)
  }

  ma_2 <- predict(moda, p2_df)
  mb_2 <- predict(modb, p2_df)
  ma_1 <- predict(moda, p1_df)
  mb_1 <- predict(modb, p1_df)
  mx <- (ma_1 + mb_1) / 2

  # 2. Predict missing values
  eta <- 
    mx + 
    p1_df$del2a / p1_df$pi2 * (p1_df$Y - mb_1) + 
    p1_df$del2b / p1_df$pi2 * (p1_df$Y - ma_1)

  # plot(p1_df$Y, eta)

  theta <- sum(eta / p1_df$pi1) / N_obs

  # 3. Estimate variance via linearization
  vhat <- var(eta) / nrow(p1_df)

  return(list(theta, vhat))
}

```

```{r}
#| label: MCMC

set.seed(1)
model <- "C"
response <- 1
pop_df <- gen_pop(N_obs, n1_obs, model = model, response = response)
true_theta <- mean(pop_df$Y)

# png("tables/popplot.png")
# plot(pop_df$X, pop_df$Y)
# dev.off()

clust <- makeCluster(min(detectCores() - 2, 100), outfile = "")
registerDoParallel(clust)

mc_res <- 
  foreach(iter=1:B_sims,
          .packages = c("dplyr", "randomForest", "mgcv")) %dopar% {

  if (iter %% 100 == 0) {
    print(paste0("Iter: ", iter))
  }

  samps <- gen_samps(pop_df)
  p1_df <- samps[[1]]
  p2_df <- samps[[2]]

  # png("tables/p2plot.png")
  # plot(p2_df$X, p2_df$Y)
  # dev.off()

  oest <- oracle_est(p1_df, p2_df, N_obs, model = model)
  nest_l <- naive_npreg_est(p1_df, p2_df, N_obs, type = "loess")
  pest_l <- ss_npreg_est(p1_df, p2_df, N_obs, type = "loess")
  nest_rf <- naive_npreg_est(p1_df, p2_df, N_obs, type = "randomforest")
  pest_rf <- ss_npreg_est(p1_df, p2_df, N_obs, type = "randomforest")
  nest_sp <- naive_npreg_est(p1_df, p2_df, N_obs, type = "spline")
  pest_sp <- ss_npreg_est(p1_df, p2_df, N_obs, type = "spline")

  return(tibble(
      Est = c("Oracle",
              "NaiveLoess", "PropLoess",
              "NaiveRF", "PropRF",
              "NaiveSpline", "PropSpline"),
      Theta = c(oest[[1]],
                nest_l[[1]], pest_l[[1]],
                nest_rf[[1]], pest_rf[[1]],
                nest_sp[[1]], pest_sp[[1]]),
      Var = c(oest[[2]],
              nest_l[[2]], pest_l[[2]],
              nest_rf[[2]], pest_rf[[2]],
              nest_sp[[2]], pest_sp[[2]]),
      Iter = iter)
  )

} %>% bind_rows()

stopCluster(clust)

```

```{r}
#| label: Analyse MCMC

# Find any failed estimates
which(is.na(mc_res$Theta))
which(is.na(mc_res$Var))

# Analyze Mean
mc_res %>%
  mutate(err = Theta - true_theta) %>%
  mutate(CI = abs(err) < qnorm(0.975) * sqrt(Var)) %>%
  group_by(Est) %>%
  summarize(Bias = mean(err, na.rm = TRUE),
            RMSE = sqrt(mean(err^2, na.rm = TRUE)),
            EmpCI = mean(CI, na.rm = TRUE),
            sdest = sd(Theta, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Ttest = abs(Bias) / sqrt(sdest^2 / B_sims)) %>%
  select(-sdest) %>%
  mutate(Ind = case_when(
    Est == "NaiveLoess" ~ 2,
    Est == "NaiveRF" ~ 4,
    Est == "NaiveSpline" ~ 6,
    Est == "PropLoess" ~ 3,
    Est == "PropRF" ~ 5,
    Est == "PropSpline" ~ 7,
    Est == "Oracle" ~ 1
  )) %>%
  arrange(Ind) %>%
  select(-Ind)

%>%
  knitr::kable(digits = 3, "latex", booktabs = TRUE, linesep = "") %>%
  cat(file = paste0("tables/npsim2", model, response, "_mean.tex"))

# Analyze Variance
mc_res %>%
  group_by(Est) %>%
  summarize(MCVar = var(Theta, na.rm = TRUE),
            EstVar = mean(Var, na.rm = TRUE),
            VarVar = var(Var, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Ttest = abs(MCVar - EstVar) / sqrt(VarVar / B_sims)) %>%
  mutate(Ind = case_when(
    Est == "NaiveLoess" ~ 2,
    Est == "NaiveRF" ~ 4,
    Est == "NaiveSpline" ~ 6,
    Est == "PropLoess" ~ 3,
    Est == "PropRF" ~ 5,
    Est == "PropSpline" ~ 7,
    Est == "Oracle" ~ 1
  )) %>%
  arrange(Ind) %>%
  select(-Ind)

%>%
  knitr::kable(digits = c(1, 4, 4, 7, 1), "latex", booktabs = TRUE, linesep = "") %>%
  cat(file = paste0("tables/npsim2", model, response, "_var.tex"))

```

### 2024-05-13 final/reports/prelim_paper.tex
Fixed multisource simulation 2. It works, but I need to change the estimator to be a GLS estimate instead of a simple average.

@@
 
-where $\hat T = (\hat{\bf X}^{(m)}, \sum_{i \in U} g(d_{im}) q_i)^T$ and $d_{im}$
+where $\hat T = (\hat{\bf X}^{(A_m)}, \sum_{i \in U} g(d_{im}) q_i)^T$ and $d_{im}$
 are the design weights for $A_m$.
@@
 
+We first run a simulation for the case in which we only have one sample in which
+we observe $Y$. We have the following superpopulation model with $N=10000$
+elements:
+
+$$
+\begin{aligned}
+X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
+X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
+X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
+Z_i &\stackrel{ind}{\sim} N(0, 1) \\
+\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
+Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
+\pi_{0i} &= \min(\max(\Phi(-x_{1i}), 0.02), 0.9)\\
+\pi_{1i} &= n_1 / N.\\
+\pi_{2i} &= \Phi(x_{2i} - 2) \\
+\end{aligned}
+$$
+
+We observe the following columns in each sample
+
+\begin{table}[ht!]
+  \centering
+  \begin{tabular}{lrrrr}
+    \toprule
+    Sample & $X_1$ & $X_2$ & $X_3$ & $Y$ \\
+    \midrule
+   $A_0$  & \checkmark     & \checkmark     & \checkmark     &  \checkmark \\    
+   $A_1$  & \checkmark     &       & \checkmark     &    \\  
+   $A_2$  & \checkmark     & \checkmark     &       &    \\  
+   \bottomrule
+  \end{tabular}
+\end{table}
+
+For the sampling mechanism both $A_0$ and $A_1$ are selected using a Poisson
+sample with response probabilities $\pi_{0i}$ and $\pi_{1i}$ respectively. The
+sample $A_2$ is a simple random sample with $n_1 = 2000$. We compare five
+different estimators for $\theta = E[Y]$.
+
+\begin{itemize}
+  \item[1.] Horvitz-Thompson estimator (HT): $\hat Y = N^{-1} \sum_{i \in A_0}
+    \frac{y_i}{\pi_{0i}}$,
+  \item[2.] Non-nested regression (NNReg): This is the non-nested regression from
+    Equation~\eqref{eq:nnopt} with only using information from Samples $A_0$ and
+    $A_1$,
+  \item[3.] Multi-Source proposed (MSEst): This is the proposed estimator from
+    Equation~\eqref{eq:dc3},
+  \item[4.] Multi-Source population (MSPop): This is the proposed estimator with
+    using the true value of $T_1$ from the population,
+  \item[5.] Multi-Source regression (MSReg): This is the linearized regression
+    estimator from FIXME.
+\end{itemize}
+
+\begin{table}[ht!]
+  \centering
+  \input{tables/msdcsim1.tex}
+% This table was generated from ../../src/explore/20240506-msdcsim.qmd
+\caption{This table shows the results of Simulation Study 2. It displays the
+Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
+unbiasedness of each estimator for the estimators: HT, NNReg, MSPop, MSEst, and
+MSReg.}
+\label{tab:msdc-mean}
+\end{table}
+
+As expected MSPop has the lowest RMSE because it also uses the population
+totals. It seems like MSEst and MSReg are almost equivalent, which is good
+because they are also supposed to be identical. The MSEst estimator outperforms
+NNReg because it also uses information from $A_2$, even though it implicitly
+uses a regression estimator with $X_3$ as a covariate, which is unnecessary.
+
 \subsection{Simulation Study 3.2}
+
+Building off of the previous simulation study, we now consider the case for
+which we observe $Y$ in separate samples.
+We have the following superpopulation model with $N = 10,000$ elements:
+
+$$
+\begin{aligned}
+X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
+X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
+X_{3i} &\stackrel{ind}{\sim} N(5, 1) \\
+\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
+Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
+\pi_{A1i} &= \min(\max(\Phi(-x_{1i}), 0.02), 0.9)\\
+\pi_{A2i} &= \min(\max(\Phi(x_{3i} - 6), 0.02), 0.95) \\
+\pi_{B1i} &= n_1 / N\\
+\pi_{B2i} &= \Phi(x_{2i} - 2) \\
+\end{aligned}
+$$
+
+We observe the following columns in each sample
+
+\begin{table}[ht!]
+  \centering
+  \begin{tabular}{lrrrr}
+    \toprule
+    Sample & $X_1$ & $X_2$ & $X_3$ & $Y$ \\
+    \midrule
+   $A_1$  & \checkmark     & \checkmark     & \checkmark     &  \checkmark \\    
+   $A_2$  & \checkmark     & & \checkmark     &  \checkmark \\    
+   $B_1$  & \checkmark     &       & \checkmark     &    \\  
+   $B_2$  & \checkmark     & \checkmark     &       &    \\  
+   \bottomrule
+  \end{tabular}
+\end{table}
+
+For the sampling mechanism $A_1$, $A_2$ and $B_2$ are selected using a Poisson
+sample with response probabilities $\pi_{A1i}$, $\pi_{A2i}$ and $\pi_{B1i}$
+respectively. The $B_2$ sample is a simple random sample (SRS). We estimate the
+following methods:
+
+\begin{itemize}
+  \item[1.] Horvitz-Thompson estimator for $A_1$ (HTA1):
+
+    $$\hat Y_{HT, A_1} = \sum_{i \in A_1} \frac{y_i}{\pi_{A1i}}.$$
+  \item[2.] Horvitz-Thompson estimator for $A_2$ (HTA1):
+    $$\hat Y_{HT, A_2} = \sum_{i \in A_2} \frac{y_i}{\pi_{A2i}}.$$
+  \item[3.] Weighted Horvitz-Thompson estimator (WHT):
+    $$\hat Y_{WHT} = \frac{\hat Y_{HT, A_1} / \hat V_{A_1} + \hat Y_{HT, A_2} /
+    \hat V_{A_2}}{1 / \hat V_{A_1} + 1 / \hat V_{A_2}}$$
+
+    where $\hat V_{A_m}$ is the estimated variance of the Horvitz-Thompson
+    estimator $\hat Y_{HT, A_m}$.
+  \item[4.] Multi-Source proposed (MSEst): This is the proposed estimator from
+    Equation~FIXME.
+  \item[5.] Multi-Source population (MSPop): This is the proposed estimator with
+    using the true value of $T_1$ from the population,
+  \item[6.] Multi-Source regression (MSReg): This is the linearized regression
+    estimator from FIXME.
+\end{itemize}
+
+\begin{table}[ht!]
+  \centering
+  \input{tables/msdcsim2.tex}
+% This table was generated from ../../src/explore/20240511-msdcsim2.qmd
+\caption{This table shows the results of Simulation Study 2. It displays the
+Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
+unbiasedness of each estimator for the estimators: HT, NNReg, MSPop, MSEst, and
+MSReg.}
+\label{tab:msdc2-mean}
+\end{table}
+
+% FIXME: Change to GLS estimate of Y.
 
@@
 
-\newpage
-
-\appendix
-
-\section{Derivation of the Estimated Variance of DC-Pop in Simulation 1}
-
-We know that $\hat Y_{DC-Pop} = \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat
-\lambda) y_i$ where $\hat w_{2i|1}$ solves
-
-  \begin{equation}
-    \argmin_{w_{2|1}} \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i})
-    \text{ such that}
-    \sum_{i \in A_2} w_{1i} w_{2i|1} z_i = \sum_{i \in U} z_i.
-  \end{equation}
-
-This means that using the technique of \cite{randles1982asymptotic},
-the linearization of $\hat Y_{DC-Pop}$ is
-
-$$\hat Y_{DC-Pop, \ell} = \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat \lambda) y_i + 
-\left( \sum_{i \in U} z_i - 
-\sum_{i \in A_2} w_{1i} w_{2i|1}(\hat \lambda) z_i \right) \phi.$$
-
-Since $\phi^*$ satisfies $E\left[\frac{\partial}{\partial \lambda} \hat Y_{DC-Pop,
-\ell}(\lambda^*, \phi^*)\right] = 0$, $\phi^*$ is the same as it is for $\hat Y_{DC-Est,
-\ell}$ with $q_i = 1$,
-
-  \begin{align*}
-    \phi^*
-    &= E\left[\sum_{i \in A_2} \frac{w_{1i}z_i z_i^T }{g'(d_{2i|1})}\right]^{-1}
-    E\left[\sum_{i \in A_2} \frac{w_{1i}z_i y_i}{g'(d_{2i|1})}\right]\\
-    &= \left[\sum_{i \in U} \frac{\pi_{2i|1}z_i z_i^T}{g'(d_{2i|1})}\right]^{-1}
-    \left[\sum_{i \in U} \frac{\pi_{2i|1} z_i y_i}{g'(d_{2i|1})}\right].\\
-  \end{align*}
-
-Hence,
-
-$$
-\hat Y_{DC-Pop, \ell} = \sum_{i \in U} z_i \phi^* + \sum_{i \in A_2} w_{1i} \hat
-w_{2i|1}(\lambda^*) (y_i - z_i \phi^*).
-$$
-
-Unlike the two-phase regression estimator and $\hat Y_{DC-Est}$ this
-linearization only has uncertainty in the second term. The first term is a
-population estimate. Let $\varepsilon_i = (y_i - z_i \phi^*)$. This means that
-
-\begin{align*}
-  V(\hat Y_{DC-Pop, \ell}) 
-  &= V(E[\hat Y_{DC-Pop, \ell} \mid A_1]) + E[V(\hat Y_{DC-Pop, \ell} \mid A_1)] \\
-  &= V\left(\sum_{i \in A_1} w_{1i} \pi_{2i|1} \hat w_{2i|1}(\lambda^*)
-  \varepsilon_i\right) + 
-  E\left[\sum_{i \in A_1} \pi_{2i|1}(1 - \pi_{2i|1}) w_{1i}^2 \hat w_{2i|1}(\lambda^*)^2
-    \varepsilon_i^2 \right]\\
-  &= \sum_{i \in U} \pi_{1i}(1 - \pi_{1i}) w_{1i}^2 \pi_{2i|1}^2 
-  \hat w_{2i|1}(\lambda^*)^2 \varepsilon_i^2 +
-  \sum_{i \in U} \pi_{1i}\pi_{2i|1}(1 - \pi_{2i|1}) w_{1i}^2 \hat w_{2i|1}(\lambda^*)^2
-    \varepsilon_i^2\\
-  &= \sum_{i \in U} w_{1i} w_{2i|1}(\lambda^*)^2 \pi_{2i|1}
-  ((1 - \pi_{1i})\pi_{2i|1} + (1 - \pi_{2i|1})) \varepsilon_i^2.
-\end{align*}
-
-Since $w_{2i|1}(\lambda^*) \pi_{2i|1} = 1$, there are two potential estimators
-of the variance,
-
-\begin{align*}
-  \hat V_1 &= \sum_{i \in A_2} w_{1i}^2 w_{2i|1}(\hat \lambda)^2 
-  ((1 - \pi_{1i})\pi_{2i|1} + (1 - \pi_{2i|1})) \varepsilon_i^2\\
-  \hat V_2 &= \sum_{i \in A_2}  w_{1i}^2 d_{2i|1}^2
-  ((1 - \pi_{1i})\pi_{2i|1} + (1 - \pi_{2i|1})) \varepsilon_i^2.
-\end{align*}
-
-In the current simulation, we use $\hat V_1$, but I have tried $\hat V_2$ and
-the results are similar.
 \end{document}

### 2024-05-13 src/explore/20240506-msdcsim.qmd
Fixed multisource simulation 2. It works, but I need to change the estimator to be a GLS estimate instead of a simple average.

@@
 
-1. The $\pi^*$-estimator,
-2. The two-phase regression estimator,
+1. The a HT-estimator,
+2. The non-nested two-phase regression estimator ($A_0$ and $A_2$),
 3. The debiased calibration estimator (with known finite population total), and
@@
 N_obs <- 10000
-n1_obs <- 1000
-n2_obs <- 2000
+n1_obs <- 2000
 
 library(dplyr)
-library(GECal)
 library(nleqslv)
@@
 library(doRNG)
-library(CVXR)
-library(rlang)
 
@@
 X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
-X_{4i} &\stackrel{ind}{\sim} Unif(0.1, 0.9) \\
+Z_i &\stackrel{ind}{\sim} N(0, 1) \\
 \varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
 Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
-\pi_{0i} &= \Phi(-x_{1i} - 2)\\
-\pi_{1i} &= |x_{1i}| / \sum_{i \in U} |x_{i1}|\\
-\pi_{2i} &= n_2 / N.\\
+\pi_{0i} &= \min(\max(\Phi(-x_{1i}), 0.02), 0.9)\\
+\pi_{1i} &= n_1 / N\\
+\pi_{2i} &= \Phi(x_{2i} - 2) \\
 \end{aligned}
@@
 | $A_0$  | ✓     | ✓     | ✓     |  ✓   |  
-| $A_1$  | ✓     | ✓     |       |      |  
-| $A_2$  | ✓     |       | ✓     |      |
+| $A_1$  | ✓     |       | ✓     |      |  
+| $A_2$  | ✓     | ✓     |       |      |
 
@@
 
-gen_pop <- function(obs) {
+gen_pop <- function(N_obs, n1_obs) {
 
-  x1 <- rnorm(obs, 2, 1)
-  x2 <- runif(obs, 0, 4)
-  x3 <- rnorm(obs, 0, 1)
-  x4 <- runif(obs, 0.1, 0.9)
-  eps <- rnorm(obs)
+  x1 <- rnorm(N_obs, 2, 1)
+  x2 <- runif(N_obs, 0, 4)
+  x3 <- rnorm(N_obs, 5, 1)
+  z <- rnorm(N_obs, 0, 1)
+
+  eps <- rnorm(N_obs)
 
@@
 
-  return(tibble(X1 = x1, X2 = x2, X3 = x3, X4 = x4, Y = y))
-}
+  pi0 <- pmin(pmax(pnorm(-x1), 0.02), 0.9)
+  # pi1 <- pnorm(-z - 1) # Informative sampling. Do we want this?
+  pi1 <- n1_obs / N_obs
+  # pi2 <- pmin(pmax(pnorm(x2 - 4), 0.01), 0.9)
+  pi2 <- pnorm(x2 - 2)
 
-#' This function adds sampling probabilites to a population data frame.
-#'
-#' @details The phase 2 formula will be included for every element. There is no
-#' information about if an element is part of the Phase 1 or Phase 2 sample.
-update_pop <- function(pop_df, p0_formula, p1_formula, p2_formula) {
-
-  pi0 <- eval(parse_expr(p0_formula))
-  pi1 <- eval(parse_expr(p1_formula))
-  pi2 <- eval(parse_expr(p2_formula))
-
-  pop_df %>%
-    mutate(pi0 = pi0, pi1 = pi1, pi2 = pi2)
+  return(tibble(X1 = x1, X2 = x2, X3 = x3, Y = y,
+                pi0 = pi0, pi1 = pi1, pi2 = pi2))
 }
@@
 #' population with selection probabilites.
-gen_samps <- function(upop_df, p0_type, p1_type, p2_type) {
+gen_samps <- function(pop_df, p0_type, p1_type, p2_type) {
 
   if (p0_type == "poisson") {
-    del0 <- rbinom(nrow(upop_df), 1, upop_df$pi0)
-  } else if (p1_type == "srs") {
-    ind <- sample(1:nrow(upop_df), size = upop_df$pi0[1] * nrow(upop_df))
-    del0 <- as.numeric(1:nrow(upop_df) %in% ind)
+    del0 <- rbinom(nrow(pop_df), 1, pop_df$pi0)
+  } else if (p0_type == "srs") {
+    ind <- sample(1:nrow(pop_df), size = pop_df$pi0[1] * nrow(pop_df))
+    del0 <- as.numeric(1:nrow(pop_df) %in% ind)
   } else {
@@
 
-  p0_df <- mutate(upop_df, del0 = del0) %>% filter(del0 == 1)
+  p0_df <- mutate(pop_df, del0 = del0) %>% filter(del0 == 1)
 
   if (p1_type == "poisson") {
-    del1 <- rbinom(nrow(upop_df), 1, upop_df$pi1)
+    del1 <- rbinom(nrow(pop_df), 1, pop_df$pi1)
   } else if (p1_type == "srs") {
-    ind <- sample(1:nrow(upop_df), size = upop_df$pi1[1] * nrow(upop_df))
-    del1 <- as.numeric(1:nrow(upop_df) %in% ind)
+    ind <- sample(1:nrow(pop_df), size = pop_df$pi1[1] * nrow(pop_df))
+    del1 <- as.numeric(1:nrow(pop_df) %in% ind)
   } else {
@@
 
-  p1_df <- mutate(upop_df, del1 = del1) %>% filter(del1 == 1)
+  p1_df <- mutate(pop_df, del1 = del1) %>% filter(del1 == 1)
 
   if (p2_type == "poisson") {
-    del2 <- rbinom(nrow(upop_df), 1, upop_df$pi2)
+    del2 <- rbinom(nrow(pop_df), 1, pop_df$pi2)
   } else if (p2_type == "srs") {
-    ind <- sample(1:nrow(upop_df), size = round(upop_df$pi2[1] * nrow(upop_df)))
-    del2 <- as.numeric(1:nrow(upop_df) %in% ind)
+    ind <- sample(1:nrow(pop_df), size = round(pop_df$pi2[1] * nrow(pop_df)))
+    del2 <- as.numeric(1:nrow(pop_df) %in% ind)
   } else {
@@
 
-  p2_df <- mutate(upop_df, del2 = del2) %>% filter(del2 == 1)
+  p2_df <- mutate(pop_df, del2 = del2) %>% filter(del2 == 1)
 
@@
 
+ht_est <- function(p0_df, N_obs, p0_type = "poisson") {
+
+  theta <- sum(p0_df$Y / p0_df$pi0) / N_obs
+
+  if (p0_type == "poisson") {
+
+    v_hat <- sum((1 - p0_df$pi0) / p0_df$pi0^2 * p0_df$Y^2) / N_obs^2
+
+  } else {
+
+    v_hat <- NA
+
+  }
+
+  return(list(theta = theta, v_hat = v_hat))
+}
+
+reg_est <- function(p0_df, p1_df, N_obs, p0_samp = "poisson", p1_samp = "srs") {
+
+  v_poisson <- sum((1 - p0_df$pi0) / p0_df$pi0^2 * p0_df$Y^2) / N_obs^2
+  v_srs <- (1 / nrow(p0_df) - 1 / N_obs) * var(p0_df$Y)
+  d_eff <- v_poisson / v_srs
+  n0_eff <- nrow(p0_df) / d_eff
+  n1_eff <- nrow(p1_df)
+
+  x1_hat <- 
+    c(sum(p1_df$X1 / p1_df$pi1) / N_obs,
+      sum(p1_df$X2 / p1_df$pi1) / N_obs)
+
+  x0_hat <- 
+    c(sum(p0_df$X1 / p0_df$pi0) / N_obs,
+      sum(p0_df$X2 / p0_df$pi0) / N_obs)
+
+  w <- n1_eff / (n1_eff + n0_eff)
+  xc_hat <- (n1_eff * x1_hat + n0_eff * x0_hat) / (n1_eff + n0_eff)
+
+  # Estimate beta
+  mod <- lm(Y ~ X1 + X2, data = p0_df)
+
+  # Regression Estimation
+  y_ht <- sum(p0_df$Y / p0_df$pi0) / N_obs
+  theta <- y_ht + sum((xc_hat - x0_hat) * mod$coefficients[2:3])
+
+  # Variance Estimation
+  coefs <- mod$coefficients
+  eps <- (p0_df$Y - coefs[1] * w - coefs[2] * w * p0_df$X1 - coefs[3] * w * p0_df$X2)
+  var_x <- 
+  (1 / nrow(p1_df) - 1 / N_obs) * var(matrix(c(p1_df$X1, p1_df$X2), ncol = 2))
+
+  v1 <- w^2 * matrix(coefs[2:3], nrow = 1) %*% var_x %*% matrix(coefs[2:3], ncol = 1)
+  v2 <- sum((1 - p0_df$pi0) / p0_df$pi0^2 * eps^2) / N_obs^2
+
+  cov_xy <- c(sum((1 - p0_df$pi0) / p0_df$pi0^2 * coefs[2] * 
+                 (n0_eff) / (n1_eff + n0_eff) * p0_df$X1 * eps) / N_obs^2,
+              sum((1 - p0_df$pi0) / p0_df$pi0^2 * coefs[3] * 
+                 (n0_eff) / (n1_eff + n0_eff) * p0_df$X3 * eps) / N_obs^2)
+  v3 <- sum(cov_xy)
+
+  v_hat <- v1 + v2 + v3
+
+  return(list(theta = theta, v_hat = as.numeric(v_hat)))
+}
+
+#' This  function does the debiased calibration
+#'
+#' @param zmat - A n x p matrix
+#' @param w1i - A vector of length n
+#' @param d2i - A vector of length n
+#' @param T1 - A vector of length p
+#' @param qi - A scaler or a vector of length n
+#' @param entropy - One of "EL", "ET"
+solveGE <- function(zmat, w1i, d2i, T1, qi, entropy) {
+
+  f <- function(lam, zmat, w1i, d2i, T1, qi, entropy, returnw = FALSE) {
+
+    if (entropy == "EL") {
+      w <- -d2i / drop(zmat %*% lam)
+    } else if (entropy == "ET") {
+      w <- d2i * exp(drop(zmat %*% lam))
+    } else {
+      stop("We only accept entropy of EL or ET.")
+    }
+
+    if (returnw) {
+      return(w)
+    } else {
+      return(drop((w * w1i * qi) %*% zmat) - T1)
+    }
+  }
+
+  j <- function(lam, zmat, w1i, d2i, T1, qi, entropy) {
+    if (entropy == "EL") {
+      res <- t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
+    } else if (entropy == "ET") {
+      res <- t(zmat) %*% diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
+    }
+
+    return(res)
+  }
+
+  init <- c(rep(0, ncol(zmat) - 1), 1)
+  res <- 
+  nleqslv(init, f, jac = j, zmat = zmat, w1i = w1i, d2i = d2i,
+          T1 = T1, qi = qi, entropy = entropy,
+          method = "Newton", control = list(maxit = 1e5, allowSingular = TRUE))
+
+  resw <- f(res$x, zmat, w1i, d2i, T1, qi, entropy, returnw = TRUE)
+
+  if (!(res$termcd %in% c(1))) {
+    return(NA)
+  }
+
+  return(resw)
+
+}
+
+dc_ms_ybar <- 
+  function(p0_df, p1_df, p2_df, pop_df, estT1, qi = 1, entropy = "EL", reg = FALSE) {
+
+  d_vec <- 1 / (p0_df$pi0)
+  if (entropy == "EL") {
+    gdi <- -1 / d_vec
+    gdip <- -1 / (1 / pop_df$pi0)
+    gpinv <- d_vec^2
+  } else if (entropy == "ET") {
+    gdi <- log(d_vec)
+    gdip <- log(1 / pop_df$pi0)
+    gpinv <- d_vec
+  } else {
+    stop("This type of entropy has not yet been implemented.")
+  }
+
+  samp_X <- cbind(rep(1, nrow(p0_df)), as.matrix(select(p0_df, X1, X2, X3)))
+  zmat <- cbind(samp_X, gdi)
+  w1i <- 1
+  d0i <- 1 / p0_df$pi0
+
+  if (estT1) {
+    x1_0 <- sum(p0_df$X1 / p0_df$pi0) / N_obs
+    x2_0 <- sum(p0_df$X2 / p0_df$pi0) / N_obs
+    x3_0 <- sum(p0_df$X3 / p0_df$pi0) / N_obs
+
+    x1_1 <- sum(p1_df$X1 / p1_df$pi1) / N_obs
+    x3_1 <- sum(p1_df$X3 / p1_df$pi1) / N_obs
+
+    x1_2 <- sum(p2_df$X1 / p2_df$pi2) / N_obs
+    x2_2 <- sum(p2_df$X2 / p2_df$pi2) / N_obs
+
+    xhat <- c(x1_0, x2_0, x3_0, x1_1, x3_1, x1_2, x2_2)
+    x_mat <- 
+    matrix(c(1, 0, 0,
+             0, 1, 0,
+             0, 0, 1,
+             1, 0, 0,
+             0, 0, 1,
+             1, 0, 0,
+             0, 1, 0), ncol = 3, byrow = TRUE)
+    vmat <- matrix(0, nrow = 7, ncol = 7)
+    vmat[1, 1] <- sum((1 / p0_df$pi0^2 - 1 / p0_df$pi0) * p0_df$X1^2) / nrow(pop_df)^2
+    vmat[1, 2] <- 
+      sum((1 / p0_df$pi0^2 - 1 / p0_df$pi0) * p0_df$X1 * p0_df$X2) / nrow(pop_df)^2
+    vmat[1, 3] <-
+      sum((1 / p0_df$pi0^2 - 1 / p0_df$pi0) * p0_df$X1 * p0_df$X3) / nrow(pop_df)^2
+    vmat[2, 1] <- vmat[1, 2]
+    vmat[3, 1] <- vmat[1, 3]
+    vmat[2, 2] <- sum((1 / p0_df$pi0^2 - 1 / p0_df$pi0) * p0_df$X2^2) / nrow(pop_df)^2
+    vmat[2, 3] <-
+      sum((1 / p0_df$pi0^2 - 1 / p0_df$pi0) * p0_df$X2 * p0_df$X3) / nrow(pop_df)^2
+    vmat[3, 2] <- vmat[2, 3]
+    vmat[3, 3] <- sum((1 / p0_df$pi0^2 - 1 / p0_df$pi0) * p0_df$X3^2) / nrow(pop_df)^2
+
+    vmat[4, 4] <- (1 / nrow(p1_df) - 1 / nrow(pop_df)) * var(p1_df$X1)
+    vmat[5, 5] <- (1 / nrow(p1_df) - 1 / nrow(pop_df)) * var(p1_df$X3)
+    vmat[4, 5] <- (1 / nrow(p1_df) - 1 / nrow(pop_df)) * cov(p1_df$X1, p1_df$X3)
+    vmat[5, 4] <- vmat[4, 5]
+
+    vmat[6, 6] <- sum((1 / p2_df$pi2^2 - 1 / p2_df$pi2) * p2_df$X1^2) / nrow(pop_df)^2
+    vmat[7, 7] <- sum((1 / p2_df$pi2^2 - 1 / p2_df$pi2) * p2_df$X2^2) / nrow(pop_df)^2
+    vmat[6, 7] <- 
+      sum((1 / p2_df$pi2^2 - 1 / p2_df$pi2) * p2_df$X1 * p2_df$X2) / nrow(pop_df)^2
+    vmat[7, 6] <- vmat[6, 7]
+
+    xgls <- drop(MASS::ginv(t(x_mat) %*% solve(vmat) %*% x_mat) %*% 
+                (t(x_mat) %*% solve(vmat) %*% xhat))
+
+    T1 <- c(nrow(pop_df),
+            xgls[1] * nrow(pop_df),
+            xgls[2] * nrow(pop_df),
+            xgls[3] * nrow(pop_df),
+            sum(gdip * qi))
+  } else {
+    T1 <- c(nrow(pop_df),
+            sum(pop_df$X1),
+            sum(pop_df$X2),
+            sum(pop_df$X3),
+            sum(gdip * qi))
+  }
+
+  dc_w <- solveGE(zmat, w1i, d0i, T1, qi, entropy)
+
+  # Mean Estimation
+  theta <- sum(p0_df$Y * dc_w * w1i) / nrow(pop_df)
+  z_dc <- model.matrix(~1 + p0_df$X1 + p0_df$X2 + p0_df$X3 + gdi)
+  gam_dc <- 
+    solve(t(z_dc) %*% diag(d0i * gpinv * qi) %*% z_dc,
+          t(z_dc) %*% diag(d0i * gpinv) %*% p0_df$Y)
+  coefs <- as.numeric(gam_dc)
+  u_mat <- 
+  matrix(c(rep(1, nrow(p0_df)), p0_df$X1, p0_df$X2, p0_df$X3, gdi * qi), ncol = 5)
+  eps <- (p0_df$Y - drop(u_mat %*% coefs))
+  if (estT1) {
+    # Variance Estimation
+    # This is the variance of bar x gls not x gls total.
+    var_x <- MASS::ginv(t(x_mat) %*% MASS::ginv(vmat) %*% x_mat)
+
+    v1 <- matrix(coefs[2:4], nrow = 1) %*% var_x %*% matrix(coefs[2:4], ncol = 1)
+    v2 <- sum((1 - p0_df$pi0) / p0_df$pi0^2 * eps^2) / nrow(pop_df)^2
+    v_hat <- v1 + v2 
+  } else {
+    v2 <- sum((1 - p0_df$pi0) / p0_df$pi0^2 * eps^2) / nrow(pop_df)^2
+    v_hat <- v2
+  }
+  if (reg) {
+    theta <- 
+    (drop(T1 %*% gam_dc) + 
+      sum(1 / p0_df$pi0 * (p0_df$Y - drop(z_dc %*% gam_dc)))) / nrow(pop_df) 
+  }
+
+  return(list(theta = theta, v_hat = as.numeric(v_hat)))
+}
+
 ```
+
+```{r}
+#| label: run MCMC
+
+set.seed(1)
+pop_df <- gen_pop(N_obs, n1_obs)
+true_theta <- mean(pop_df$Y)
+
+clust <- makeCluster(min(detectCores() - 2, 100), outfile = "")
+registerDoParallel(clust)
+
+mc_res <- 
+  foreach(iter=1:B_sims,
+          .packages = c("dplyr", "nleqslv"),
+          .options.RNG = 1) %dorng% {
+
+    if (iter %% 100 == 0) {
+      print(paste0("Iter: ", iter))
+    }
+
+    samps <- gen_samps(pop_df, "poisson", "srs", "poisson")
+    p0_df <- samps[[1]]
+    p1_df <- samps[[2]]
+    p2_df <- samps[[3]]
+
+    ht <- ht_est(p0_df, N_obs)
+    reg <- reg_est(p0_df, p1_df, N_obs)
+    ms_pop <- dc_ms_ybar(p0_df, p1_df, p2_df, pop_df, estT1 = FALSE)
+    ms_est <- dc_ms_ybar(p0_df, p1_df, p2_df, pop_df, estT1 = TRUE)
+    ms_reg <- dc_ms_ybar(p0_df, p1_df, p2_df, pop_df, estT1 = TRUE, reg = TRUE)
+
+    return(tibble(Est = c("HT", "NNReg", "MSPop", "MSEst", "MSReg"),
+                  Theta = c(ht[[1]], reg[[1]], ms_pop[[1]], ms_est[[1]], ms_reg[[1]]),
+                  Var = c(ht[[2]], reg[[2]], ms_pop[[2]], ms_est[[2]], ms_reg[[2]]),
+                  Iter = iter)
+    )
+
+  } %>% bind_rows()
+
+stopCluster(clust)
+
+```
+
+```{r}
+#| label: analyze results
+
+# Check for failures
+filter(mc_res, is.na(Theta))
+
+# Analyze Mean
+mc_res %>%
+  mutate(err = Theta - true_theta) %>%
+  mutate(CI = abs(err) < qnorm(0.975) * sqrt(Var)) %>%
+  group_by(Est) %>%
+  summarize(Bias = mean(err, na.rm = TRUE),
+            RMSE = sqrt(mean(err^2, na.rm = TRUE)),
+            EmpCI = mean(CI, na.rm = TRUE),
+            sdest = sd(Theta, na.rm = TRUE)) %>%
+  ungroup() %>%
+  mutate(Ttest = abs(Bias) / sqrt(sdest^2 / B_sims)) %>%
+  select(-sdest) %>%
+  mutate(Ind = case_when(
+    Est == "MSReg" ~ 5,
+    Est == "MSEst" ~ 4,
+    Est == "MSPop" ~ 3,
+    Est == "NNReg" ~ 2,
+    Est == "HT" ~ 1)
+  ) %>%
+  arrange(Ind) %>%
+  select(Est, Bias, RMSE, EmpCI, Ttest) %>%
+  knitr::kable("latex", booktabs = TRUE, digits = 4) %>%
+  cat(file = "tables/msdcsim1.tex")
+
+```

### 2024-05-13 src/explore/20240506-npregest.qmd
Fixed multisource simulation 2. It works, but I need to change the estimator to be a GLS estimate instead of a simple average.

@@
 library(doRNG)
-library(npreg)
 library(mgcv)
@@
 gen_pop <- function(N_obs, n1_obs) {
-  x <- runif(N_obs, 0, 8 * pi)
+  x <- runif(N_obs, 1, 2)
+  z <- rnorm(N_obs)
   eps <- rnorm(N_obs)
-  y <- 0.2 * x + sin(x) + 0.3 * eps
+  # y <- 0.05 * x^2 + 2 * eps + exp(x - max(x) + 4) + 2 * sin(x)
+  y <- x + z + eps
 
   pi1 <- n1_obs / N_obs
-  z <- (x - mean(x)) / sd(x)
-  pi2 <- ifelse(pnorm(z - 0.5) > 0.7, 0.7, pnorm(z - 0.5))
+  pi2 <- ifelse(pnorm(z - 1) > 0.7, 0.7, pnorm(z - 1))
   pi2 <- ifelse(pi2 < 0.02, 0.02, pi2)
 
-  return(tibble(X = x, Y = y, pi1 = pi1, pi2 = pi2))
+  w1 <- 1 / pi1
+  w2 <- 1 / pi2
+
+  return(tibble(X = x, Y = y, pi1 = pi1, pi2 = pi2, w1 = w1, w2 = w2))
 
@@
   # 1. Get mhat.
-  mx <- 0.2 * p1_df$X + sin(p1_df$X)
-  mx2 <- 0.2 * p2_df$X + sin(p2_df$X)
+  mx <- p1_df$X
+  mx2 <- p2_df$X
 
@@
   if (type == "loess") {
-    mod <- loess(Y ~ X, p2_df, control = loess.control(surface="direct"))
+    mod <- loess(Y ~ X, data = p2_df, weights = w2,
+                 control = loess.control(surface="direct"))
   } else if (type == "randomforest") {
-    mod <- randomForest(Y ~ X, p2_df)
+    mod <- randomForest(Y ~ X, data = p2_df)
   } else if (type == "spline") {
-    # mod <- ss(x = p2_df$X, y = p2_df$Y)
-    # mx <- predict(mod, x = p1_df$X)$y
-    # mx2 <- predict(mod, x = p2_df$X)$y
-    mod <- gam(Y ~ s(X, k = 15, bs = "cr"), data = p2_df)
+    mod <- gam(Y ~ s(X, k = 15, bs = "cr"), data = p2_df, weights = w2)
   } else if (type == "linear") {
-    mod <- lm(Y ~ X, data = p2_df)
+    mod <- lm(Y ~ X, data = p2_df, weights = w2)
   }
@@
   p1_df <- p1_df %>%
-    left_join(p2_df, by = join_by(X, Y, pi1, pi2, del1, del2)) %>%
+    left_join(p2_df, by = join_by(X, Y, pi1, pi2, w1, w2, del1, del2)) %>%
     mutate(del2a = ifelse(is.na(del2a), 0, del2a)) %>%
@@
   if (type == "loess") {
-    moda <- loess(Y ~ X, p2a_df, control = loess.control(surface="direct"))
-    modb <- loess(Y ~ X, p2b_df, control = loess.control(surface="direct"))
+    moda <- loess(Y ~ X, data = p2a_df, weights = w2,
+                  control = loess.control(surface="direct"))
+    modb <- loess(Y ~ X, data = p2b_df, weights = w2,
+                  control = loess.control(surface="direct"))
   } else if (type == "randomforest") {
-    moda <- randomForest(Y ~ X, p2a_df)
-    modb <- randomForest(Y ~ X, p2b_df)
+    moda <- randomForest(Y ~ X, data = p2a_df)
+    modb <- randomForest(Y ~ X, data = p2b_df)
   } else if (type == "spline") {
-    # moda <- ss(x = p2a_df$X, y = p2a_df$Y)
-    # modb <- ss(x = p2b_df$X, y = p2b_df$Y)
-    # ma_2 <- predict(moda, x = p2_df$X)$y
-    # mb_2 <- predict(modb, x = p2_df$X)$y
-    # ma_1 <- predict(moda, x = p1_df$X)$y
-    # mb_1 <- predict(modb, x = p1_df$X)$y
-    # mx <- (ma_1 + mb_1) / 2
-    moda <- gam(Y ~ s(X, k = 15, bs = "cr"), data = p2a_df)
-    modb <- gam(Y ~ s(X, k = 15, bs = "cr"), data = p2b_df)
+    moda <- gam(Y ~ s(X, k = 15, bs = "cr"), data = p2a_df, weights = w2)
+    modb <- gam(Y ~ s(X, k = 15, bs = "cr"), data = p2b_df, weights = w2)
   } else if (type == "linear") {
-    moda <- lm(Y ~ X, data = p2a_df)
-    modb <- lm(Y ~ X, data = p2b_df)
+    moda <- lm(Y ~ X, data = p2a_df, weights = w2)
+    modb <- lm(Y ~ X, data = p2b_df, weights = w2)
   }
@@
 
-  # p2a_df <- p2a_df %>%
-  #   mutate(pi2a = pi2 * nrow(p2a_df) / nrow(p2_df))
-  # p2b_df <- p2b_df %>%
-  #   mutate(pi2b = pi2 * nrow(p2b_df) / nrow(p2_df))
-
-  # v1_t <- matrix(NA, nrow = nrow(p2_df), ncol = nrow(p2_df))
-  # v2_t <- matrix(NA, nrow = nrow(p2_df), ncol = nrow(p2_df))
-  # cov_t <- matrix(NA, nrow = nrow(p2_df), ncol = nrow(p2_df))
-
-  # # Since Phase 1 is SRS
-  # p1ij <- nrow(p1_df) / N_obs * (nrow(p1_df) - 1) / (N_obs - 1)
-  # eps_a <- p2_df$Y - ma_2
-  # eps_b <- p2_df$Y - mb_2
-
-  # n2a <- nrow(p2a_df)
-  # n2b <- nrow(p2b_df)
-  # n2 <- nrow(p2_df)
-
-  # for (i in 1:nrow(p2_df)) {
-  #   for (j in 1:nrow(p2_df)) {
-  #     if (i == j) {
-  #       v1_t[i, j] <- 
-  #       1 / 4 * (2 - p2_df$pi2[i]) / p2_df$pi1[i] * eps_b[i]^2 / p2_df$pi2[i]^2
-  #       v2_t[i, j] <- 
-  #       1 / 4 * (2 - p2_df$pi2[i]) / p2_df$pi1[i] * eps_a[i]^2 / p2_df$pi2[i]^2
-  #       cov_t[i, j] <- 
-  #       1/(p2_df$pi2[i] * p2_df$pi1[i]) * n2a/n2 * (n2a - 1)/n2 * eps_b[i] * eps_a[i]
-  #     } else {
-  #       v1_t[i, j] <- 1 / 4 * (p2_df$pi2[i]) / (p2_df$pi1[i] * p2_df$pi1[j]) *
-  #         eps_b[i] * eps_b[j] / (p2_df$pi2[i] * p2_df$pi2[j])
-  #       v2_t[i, j] <- 1 / 4 * 1 / (p2_df$pi1[i] * p2_df$pi1[j]) *
-  #         eps_a[i] * eps_a[j] / (p2_df$pi2[i] * p2_df$pi2[j])
-  #       cov_t[i, j] <- 1 / (p2_df$pi2[i] * p2_df$pi1[i]) * n2a / n2 * 
-  #       ((n2 - n2a) / (n2 - 1) + (n2a - 1)/n2) * eps_b[i] * eps_a[i]
-  #     }
-  #   }
-  # }
-
-  # v1 <- sum(as.numeric(v1_t)) / N_obs^2
-  # v2 <- sum(as.numeric(v2_t)) / N_obs^2
-  # v3 <- sum(as.numeric(cov_t)) / N_obs^2
-
-  # vhat <- (v0 + v1 + v2 + v3)
-
   return(list(theta, vhat))
@@
   foreach(iter=1:B_sims,
-          .packages = c("dplyr", "randomForest", "mgcv")) %dopar% {
+          .packages = c("dplyr", "randomForest", "mgcv"),
+          .options.RNG  = 2) %dorng% {
 
@@
 
+library(ggplot2)
+
+set.seed(1)
+pop_df <- gen_pop(N_obs, n1_obs)
+true_theta <- mean(pop_df$Y)
+
+  samps <- gen_samps(pop_df)
+  p1_df <- samps[[1]]
+  p2_df <- samps[[2]]
+
+ggplot(pop_df, aes(x = X, y = Y)) + 
+geom_point() +
+geom_smooth(method = "lm", formula = y ~ x)
+
+ggplot(p2_df, aes(x = X, y = Y)) + 
+geom_point() +
+geom_smooth(method = "lm", formula = y ~ x)
+
 ```

### 2024-05-13 src/explore/20240511-msdcsim2.qmd
Fixed multisource simulation 2. It works, but I need to change the estimator to be a GLS estimate instead of a simple average.

---
title: "Multi-Source Debiased Calibration Simulation 2: Multiple $A_m$"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    code-fold: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

# Summary

This report simulates multi-source sampling with the debiased
calibration of @kwon2024debiased. We check for the following:

1. Bias of $\hat Y$ and,
2. Bias of $\hat V(\hat Y)$.

For the estimator $\hat Y$ we construct

1. Two individual HT-estimators,
2. A weighted HT-estimator,
3. The debiased calibration estimator (with known finite population total), and
4. The debiased calibration estimator with estimated population totals from the
   Phase 1 sample.

# Simulation Setup

```{r}
#| label: setup

B_sims <- 1000

N_obs <- 10000
B1_obs <- 2000
A2_obs <- 200

library(dplyr)
library(nleqslv)
library(doParallel)
library(doRNG)

```

We have the following superpopulation model with $N=$`{r} N_obs` elements:

$$
\begin{aligned}
X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
X_{3i} &\stackrel{ind}{\sim} N(5, 1) \\
\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
\pi_{A1i} &= \min(\max(\Phi(-x_{1i}), 0.02), 0.9)\\
\pi_{A2i} &= \min(\max(\Phi(x_{3i} - 6), 0.02), 0.95) \\
\pi_{B1i} &= n_1 / N\\
\pi_{B2i} &= \Phi(x_{2i} - 2) \\
\end{aligned}
$$

We observe the following columns in each sample

| Sample | $X_1$ | $X_2$ | $X_3$ |  $Y$ |
|:-------|------:|------:|------:|-----:|
| $A_1$  | ✓     | ✓     | ✓     |  ✓   |  
| $A_2$  | ✓     |       | ✓     |  ✓   |  
| $B_1$  | ✓     |       | ✓     |      |  
| $B_2$  | ✓     | ✓     |       |      |

For the sampling mechanism $A_1$, $A_2$ and $B_2$ are selected using a Poisson
sample with response probabilities $\pi_{A1i}$, $\pi_{A2i}$ and $\pi_{B1i}$
respectively. The $B_1$ sample is a simple random sample (SRS).

```{r}
#| label: generate population and samples

gen_pop <- function(N_obs, B1_obs) {

  x1 <- rnorm(N_obs, 2, 1)
  x2 <- runif(N_obs, 0, 4)
  x3 <- rnorm(N_obs, 5, 1)
  z <- rnorm(N_obs, 0, 1)

  eps <- rnorm(N_obs)

  y <- 3 * x1 + 2 * x2 + eps

  piA1 <- pmin(pmax(pnorm(-x1), 0.02), 0.9)
  # piA2 <- A2_obs / N_obs
  piA2 <- pmin(pmax(pnorm(x3 - 6), 0.02), 0.95)
  piB1 <- B1_obs / N_obs
  piB2 <- pnorm(x2 - 2)

  return(tibble(X1 = x1, X2 = x2, X3 = x3, Y = y,
                piA1 = piA1, piA2 = piA2, piB1 = piB1, piB2 = piB2))
}

#' This function generates Phase 1 and Phase 2 samples from an updated
#' population with selection probabilites.
gen_samps <- function(pop_df, A1_type, A2_type, B1_type, B2_type) {

  if (A1_type == "poisson") {
    delA1 <- rbinom(nrow(pop_df), 1, pop_df$piA1)
  } else if (A1_type == "srs") {
    ind <- sample(1:nrow(pop_df), size = pop_df$piA1[1] * nrow(pop_df))
    delA1 <- as.numeric(1:nrow(pop_df) %in% ind)
  } else {
    stop("We have only implemented poisson and srs for Sample A1.")
  }

  A1_df <- mutate(pop_df, delA1 = delA1) %>% filter(delA1 == 1)

  if (A2_type == "poisson") {
    delA2 <- rbinom(nrow(pop_df), 1, pop_df$piA2)
  } else if (A2_type == "srs") {
    ind <- sample(1:nrow(pop_df), size = pop_df$piA2[1] * nrow(pop_df))
    delA2 <- as.numeric(1:nrow(pop_df) %in% ind)
  } else {
    stop("We have only implemented poisson and srs for Sample A2.")
  }

  A2_df <- mutate(pop_df, delA2 = delA2) %>% filter(delA2 == 1)

  if (B1_type == "poisson") {
    delB1 <- rbinom(nrow(pop_df), 1, pop_df$piB1)
  } else if (B1_type == "srs") {
    ind <- sample(1:nrow(pop_df), size = pop_df$piB1[1] * nrow(pop_df))
    delB1 <- as.numeric(1:nrow(pop_df) %in% ind)
  } else {
    stop("We have only implemented poisson and srs for Sample B1.")
  }

  B1_df <- mutate(pop_df, delB1 = delB1) %>% filter(delB1 == 1)

  if (B2_type == "poisson") {
    delB2 <- rbinom(nrow(pop_df), 1, pop_df$piB2)
  } else if (B2_type == "srs") {
    ind <- sample(1:nrow(pop_df), size = round(pop_df$piB2[1] * nrow(pop_df)))
    delB2 <- as.numeric(1:nrow(pop_df) %in% ind)
  } else {
    stop("We have only implemented poisson and srs for Sample B2.")
  }

  B2_df <- mutate(pop_df, delB2 = delB2) %>% filter(delB2 == 1)

  return(list(A1_df, A2_df, B1_df, B2_df))
}

```

```{r}
#| label: estimators

ht_est <- function(p_df, piAn, N_obs, type) {

  theta <- sum(p_df$Y / p_df[[piAn]]) / N_obs

  if (type == "poisson") {

    v_hat <- sum((1 - p_df[[piAn]]) / p_df[[piAn]]^2 * p_df$Y^2) / N_obs^2

  } else if (type == "srs") {

    v_hat <- (1 / nrow(p_df) - 1 / N_obs) * var(p_df$Y)

  } else {

    v_hat <- NA

  }

  return(list(theta = theta, v_hat = v_hat))
}

wht_est <- function(A1_df, A2_df, N_obs) {
  # Inverse-variance weighted average
  theta1 <- sum(A1_df$Y / A1_df$piA1) / N_obs
  theta2 <- sum(A2_df$Y / A2_df$piA2) / N_obs
  vhat1 <- sum((1 - A1_df$piA1) / A1_df$piA1^2 * A1_df$Y^2) / N_obs^2
  vhat2 <- sum((1 - A2_df$piA2) / A2_df$piA2^2 * A2_df$Y^2) / N_obs^2
  # vhat2 <- (1 / nrow(A2_df) - 1 / N_obs) * var(A2_df$Y)

  theta <- (theta1 / vhat1 + theta2 / vhat2) / (1 / vhat1 + 1 / vhat2)
  v_hat <- 1 / (1 / vhat1 + 1 / vhat2)

  return(list(theta = theta, v_hat = v_hat))
}

#' This  function does the debiased calibration
#'
#' @param zmat - A n x p matrix
#' @param w1i - A vector of length n
#' @param d2i - A vector of length n
#' @param T1 - A vector of length p
#' @param qi - A scaler or a vector of length n
#' @param entropy - One of "EL", "ET"
solveGE <- function(zmat, w1i, d2i, T1, qi, entropy) {

  f <- function(lam, zmat, w1i, d2i, T1, qi, entropy, returnw = FALSE) {

    if (entropy == "EL") {
      w <- -d2i / drop(zmat %*% lam)
    } else if (entropy == "ET") {
      w <- d2i * exp(drop(zmat %*% lam))
    } else {
      stop("We only accept entropy of EL or ET.")
    }

    if (returnw) {
      return(w)
    } else {
      return(drop((w * w1i * qi) %*% zmat) - T1)
    }
  }

  j <- function(lam, zmat, w1i, d2i, T1, qi, entropy) {
    if (entropy == "EL") {
      res <- t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
    } else if (entropy == "ET") {
      res <- t(zmat) %*% diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
    }

    return(res)
  }

  init <- c(rep(0, ncol(zmat) - 1), 1)
  res <- 
  nleqslv(init, f, jac = j, zmat = zmat, w1i = w1i, d2i = d2i,
          T1 = T1, qi = qi, entropy = entropy,
          method = "Newton", control = list(maxit = 1e5, allowSingular = TRUE))

  resw <- f(res$x, zmat, w1i, d2i, T1, qi, entropy, returnw = TRUE)

  if (!(res$termcd %in% c(1))) {
    return(NA)
  }

  return(resw)

}

dc_ms_ybar <- 
  function(A1_df, A2_df, B1_df, B2_df, pop_df, estT1, qi = 1, entropy = "EL", reg = FALSE) {

  dA1_vec <- 1 / (A1_df$piA1)
  dA2_vec <- 1 / (A2_df$piA2)
  if (entropy == "EL") {
    gdA1 <- -1 / dA1_vec
    gdA1p <- -1 / (1 / pop_df$piA1)
    gA1inv <- dA1_vec^2
    gdA2 <- -1 / dA2_vec
    gdA2p <- -1 / (1 / pop_df$piA2)
    gA2inv <- dA2_vec^2
  } else if (entropy == "ET") {
    gdA1 <- log(dA1_vec)
    gdA1p <- log(1 / pop_df$piA1)
    gA1inv <- dA1_vec
    gdA2 <- log(dA2_vec)
    gdA2p <- log(1 / pop_df$piA2)
    gA2inv <- dA2_vec
  } else {
    stop("This type of entropy has not yet been implemented.")
  }

  sampA1 <- cbind(rep(1, nrow(A1_df)), as.matrix(select(A1_df, X1, X2, X3)))
  sampA2 <- cbind(rep(1, nrow(A2_df)), as.matrix(select(A2_df, X1, X3)))
  zmatA1 <- cbind(sampA1, gdA1)
  zmatA2 <- cbind(sampA2, gdA2)
  w1i <- 1
  dA1i <- 1 / A1_df$piA1
  dA2i <- 1 / A2_df$piA2

  if (estT1) {
    x1_A1 <- sum(A1_df$X1 / A1_df$piA1) / N_obs
    x2_A1 <- sum(A1_df$X2 / A1_df$piA1) / N_obs
    x3_A1 <- sum(A1_df$X3 / A1_df$piA1) / N_obs

    x1_A2 <- sum(A2_df$X1 / A2_df$piA2) / N_obs
    x3_A2 <- sum(A2_df$X3 / A2_df$piA2) / N_obs

    x1_B1 <- sum(B1_df$X1 / B1_df$piB1) / N_obs
    x3_B1 <- sum(B1_df$X3 / B1_df$piB1) / N_obs

    x1_B2 <- sum(B2_df$X1 / B2_df$piB2) / N_obs
    x2_B2 <- sum(B2_df$X2 / B2_df$piB2) / N_obs

    xhat <- c(x1_A1, x2_A1, x3_A1, x1_A2, x3_A2, x1_B1, x3_B1, x1_B2, x2_B2)
    x_mat <- 
    matrix(c(1, 0, 0,
             0, 1, 0,
             0, 0, 1,
             1, 0, 0,
             0, 0, 1,
             1, 0, 0,
             0, 0, 1,
             1, 0, 0,
             0, 1, 0), ncol = 3, byrow = TRUE)
    vmat <- matrix(0, nrow = 9, ncol = 9)
    vmat[1, 1] <- sum((1 / A1_df$piA1^2 - 1 / A1_df$piA1) * A1_df$X1^2) / nrow(pop_df)^2
    vmat[1, 2] <- 
      sum((1 / A1_df$piA1^2 - 1 / A1_df$piA1) * A1_df$X1 * A1_df$X2) / nrow(pop_df)^2
    vmat[1, 3] <-
      sum((1 / A1_df$piA1^2 - 1 / A1_df$piA1) * A1_df$X1 * A1_df$X3) / nrow(pop_df)^2
    vmat[2, 1] <- vmat[1, 2]
    vmat[3, 1] <- vmat[1, 3]
    vmat[2, 2] <- sum((1 / A1_df$piA1^2 - 1 / A1_df$piA1) * A1_df$X2^2) / nrow(pop_df)^2
    vmat[2, 3] <-
      sum((1 / A1_df$piA1^2 - 1 / A1_df$piA1) * A1_df$X2 * A1_df$X3) / nrow(pop_df)^2
    vmat[3, 2] <- vmat[2, 3]
    vmat[3, 3] <- sum((1 / A1_df$piA1^2 - 1 / A1_df$piA1) * A1_df$X3^2) / nrow(pop_df)^2

    vmat[4, 4] <- sum((1 / A2_df$piA2^2 - 1 / A2_df$piA2) * A2_df$X1^2) / nrow(pop_df)^2
    vmat[5, 5] <- sum((1 / A2_df$piA2^2 - 1 / A2_df$piA2) * A2_df$X3^2) / nrow(pop_df)^2
    vmat[4, 5] <- sum((1 / A2_df$piA2^2 - 1 / A2_df$piA2) * 
      A2_df$X1 * A2_df$X3) / nrow(pop_df)^2
    vmat[5, 4] <- vmat[4, 5]

    vmat[6, 6] <- (1 / nrow(B1_df) - 1 / nrow(pop_df)) * var(B1_df$X1)
    vmat[7, 7] <- (1 / nrow(B1_df) - 1 / nrow(pop_df)) * var(B1_df$X3)
    vmat[6, 7] <- (1 / nrow(B1_df) - 1 / nrow(pop_df)) * cov(B1_df$X1, B1_df$X3)
    vmat[7, 6] <- vmat[6, 7]

    vmat[8, 8] <- sum((1 / B2_df$piB2^2 - 1 / B2_df$piB2) * B2_df$X1^2) / nrow(pop_df)^2
    vmat[9, 9] <- sum((1 / B2_df$piB2^2 - 1 / B2_df$piB2) * B2_df$X2^2) / nrow(pop_df)^2
    vmat[8, 9] <- 
      sum((1 / B2_df$piB2^2 - 1 / B2_df$piB2) * B2_df$X1 * B2_df$X2) / nrow(pop_df)^2
    vmat[9, 8] <- vmat[8, 9]

    xgls <- drop(MASS::ginv(t(x_mat) %*% solve(vmat) %*% x_mat) %*% 
                (t(x_mat) %*% solve(vmat) %*% xhat))

    TA1 <- c(nrow(pop_df),
            xgls[1] * nrow(pop_df),
            xgls[2] * nrow(pop_df),
            xgls[3] * nrow(pop_df),
            sum(gdA1p * qi))
    TA2 <- c(nrow(pop_df),
            xgls[1] * nrow(pop_df),
            xgls[3] * nrow(pop_df),
            sum(gdA2p * qi))
  } else {
    TA1 <- c(nrow(pop_df),
            sum(pop_df$X1),
            sum(pop_df$X2),
            sum(pop_df$X3),
            sum(gdA1p * qi))
    # We have some misspecification in A2
    TA2 <- c(nrow(pop_df),
            sum(pop_df$X1),
            sum(pop_df$X3),
            sum(gdA2p * qi))
  }

  dc_wA1 <- solveGE(zmatA1, w1i, dA1i, TA1, qi, entropy)
  dc_wA2 <- solveGE(zmatA2, w1i, dA2i, TA2, qi, entropy)

  # Mean Estimation
  thetaA1 <- sum(A1_df$Y * dc_wA1 * w1i) / nrow(pop_df)
  thetaA2 <- sum(A2_df$Y * dc_wA2 * w1i) / nrow(pop_df)
  theta <- (thetaA1 + thetaA2) / 2

  zA1_dc <- model.matrix(~1 + A1_df$X1 + A1_df$X2 + A1_df$X3 + gdA1)
  zA2_dc <- model.matrix(~1 + A2_df$X1 + A2_df$X3 + gdA2)

  gamA1_dc <- 
    solve(t(zA1_dc) %*% diag(dA1i * gA1inv * qi) %*% zA1_dc,
          t(zA1_dc) %*% diag(dA1i * gA1inv) %*% A1_df$Y)
  gamA2_dc <- 
    solve(t(zA2_dc) %*% diag(dA2i * gA2inv * qi) %*% zA2_dc,
          t(zA2_dc) %*% diag(dA2i * gA2inv) %*% A2_df$Y)

  coefsA1 <- as.numeric(gamA1_dc)
  coefsA2 <- as.numeric(gamA2_dc)
  uA1_mat <- 
  matrix(c(rep(1, nrow(A1_df)), A1_df$X1, A1_df$X2, A1_df$X3, gdA1 * qi), ncol = 5)
  uA2_mat <- 
  matrix(c(rep(1, nrow(A2_df)), A2_df$X1, A2_df$X3, gdA2 * qi), ncol = 4)
  epsA1 <- (A1_df$Y - drop(uA1_mat %*% coefsA1))
  epsA2 <- (A2_df$Y - drop(uA2_mat %*% coefsA2))

  if (estT1) {
    # Variance Estimation
    # This is the variance of bar x gls not x gls total.
    var_x <- MASS::ginv(t(x_mat) %*% MASS::ginv(vmat) %*% x_mat)
    coefs <- c(coefsA1[2] + coefsA2[2], coefsA1[3], coefsA1[4] + coefsA2[3])

    v1 <- matrix(coefs, nrow = 1) %*% var_x %*% matrix(coefs, ncol = 1)
    v2A1 <- sum((1 - A1_df$piA1) / A1_df$piA1^2 * epsA1^2) / nrow(pop_df)^2
    v2A2 <- sum((1 - A2_df$piA2) / A2_df$piA2^2 * epsA2^2) / nrow(pop_df)^2
    # v2A2 <- (1 / nrow(A2_df) - 1 / nrow(pop_df)) * var(epsA2)
    v_hat <- (v1 + v2A1 + v2A2) / 4
  } else {
    v2A1 <- sum((1 - A1_df$piA1) / A1_df$piA1^2 * epsA1^2) / nrow(pop_df)^2
    v2A2 <- sum((1 - A2_df$piA2) / A2_df$piA2^2 * epsA2^2) / nrow(pop_df)^2
    # v2A2 <- (1 / nrow(A2_df) - 1 / nrow(pop_df)) * var(epsA2)
    v_hat <- (v2A1 + v2A2) / 4
  }
  if (reg) {
    thetaA1 <- 
    (drop(TA1 %*% gamA1_dc) + 
      sum(1 / A1_df$piA1 * (A1_df$Y - drop(zA1_dc %*% gamA1_dc)))) / nrow(pop_df) 
    thetaA2 <- 
    (drop(TA2 %*% gamA2_dc) + 
      sum(1 / A2_df$piA2 * (A2_df$Y - drop(zA2_dc %*% gamA2_dc)))) / nrow(pop_df) 
    theta <- (thetaA1 + thetaA2) / 2
  }

  return(list(theta = theta, v_hat = as.numeric(v_hat)))
}

```

```{r}
#| label: run MCMC

set.seed(1)
pop_df <- gen_pop(N_obs, B1_obs)
true_theta <- mean(pop_df$Y)

clust <- makeCluster(min(detectCores() - 2, 100), outfile = "")
registerDoParallel(clust)

mc_res <- 
  foreach(iter=1:B_sims,
          .packages = c("dplyr", "nleqslv"),
          .options.RNG = 1) %dorng% {

    if (iter %% 100 == 0) {
      print(paste0("Iter: ", iter))
    }

    samps <- gen_samps(pop_df, "poisson", "poisson", "srs", "poisson")
    A1_df <- samps[[1]]
    A2_df <- samps[[2]]
    B1_df <- samps[[3]]
    B2_df <- samps[[4]]

    htA1 <- ht_est(A1_df, "piA1", N_obs, "poisson")
    htA2 <- ht_est(A2_df, "piA2", N_obs, "poisson")
    wht <- wht_est(A1_df, A2_df, N_obs)
    ms_pop <- dc_ms_ybar(A1_df, A2_df, B1_df, B2_df, pop_df, estT1 = FALSE)
    ms_est <- dc_ms_ybar(A1_df, A2_df, B1_df, B2_df, pop_df, estT1 = TRUE)
    ms_reg <- dc_ms_ybar(A1_df, A2_df, B1_df, B2_df, pop_df, estT1 = TRUE, reg = TRUE)

    return(
    tibble(
      Est = c("HTA1", "HTA2", "WHT", "MSPop", "MSEst", "MSReg"),
      Theta = c(htA1[[1]], htA2[[1]], wht[[1]], ms_pop[[1]], ms_est[[1]], ms_reg[[1]]),
      Var = c(htA1[[2]], htA2[[2]], wht[[2]], ms_pop[[2]], ms_est[[2]], ms_reg[[2]]),
      Iter = iter)
    )

  } %>% bind_rows()

stopCluster(clust)

```

```{r}
#| label: analyze results

# Check for failures
filter(mc_res, is.na(Theta))

# Analyze Mean
mc_res %>%
  mutate(err = Theta - true_theta) %>%
  mutate(CI = abs(err) < qnorm(0.975) * sqrt(Var)) %>%
  group_by(Est) %>%
  summarize(Bias = mean(err, na.rm = TRUE),
            RMSE = sqrt(mean(err^2, na.rm = TRUE)),
            EmpCI = mean(CI, na.rm = TRUE),
            sdest = sd(Theta, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Ttest = abs(Bias) / sqrt(sdest^2 / B_sims)) %>%
  select(-sdest) %>%
  mutate(Ind = case_when(
    Est == "MSReg" ~ 6,
    Est == "MSEst" ~ 5,
    Est == "MSPop" ~ 4,
    Est == "WHT" ~ 3,
    Est == "HTA2" ~ 2,
    Est == "HTA1" ~ 1)
  ) %>%
  arrange(Ind) %>%
  select(Est, Bias, RMSE, EmpCI, Ttest)

%>%
  knitr::kable("latex", booktabs = TRUE, digits = 4) %>%
  cat(file = "tables/msdcsim2.tex")

# Analyze Variance
mc_res %>%
  group_by(Est) %>%
  summarize(MCVar = var(Theta, na.rm = TRUE),
            EstVar = mean(Var, na.rm = TRUE),
            VarVar = var(Var, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Ttest = abs(MCVar - EstVar) / sqrt(VarVar / B_sims))

%>%
  knitr::kable("latex", booktabs = TRUE, escape = FALSE, digits = 4) %>%
  cat(file = "tables/msdcsim2_var.tex")

```

### 2024-07-29 final/reports/prelim_paper.tex
Updated July 29, 2024

@@
 
-\begin{theorem}[Variance Estimation]
+\begin{theorem}[Variance Estimation]\label{thm:nnve}
   The variance of $\hat Y_{NNE}$ is
@@
 sampling being a simple random sample (SRS) of size $n_1 = 1000$. The Phase 2
-sample is a Poisson sample with an expected sample size of
+sample is a Poisson sample with an expected sample size of about 300.
 
@@
 where $\hat T = (\hat{\bf X}^{(0)}, \sum_{i \in U} g(d_i) q_i)^T$ and 
-$\bf z_i = (\bf x_i^T, g(d_i)q_i)^T$. A more general approach considers the case
-in which multiple samples can observe $Y$. In this case, assume that we have
-samples $A_1, \dots A_{M_1}$ for which we observe elements of $\bf X$ and $Y$
-and samples $B_1, \dots, B_{M_2}$ for which we only observed elements of $\bf
-X$. We denote $\bf X^{(A_m)}$ and $\bf X^{(B_m)}$ as the columns of $\bf X$
+$\bf z_i = ((\bf x_i^{(0)})^T, g(d_i)q_i)^T$. Then the estimator is $\hat Y_{MS}
+= \sum_{i \in A_0} w_{i} y_i$.
+
+A more general approach considers
+the case in which multiple samples can observe $Y$. In this case, assume that we
+have samples $A_1, \dots A_{M_1}$ for which we observe elements of $\bf X$ and
+$Y$ and samples $B_1, \dots, B_{M_2}$ for which we only observed elements of
+$\bf X$. We denote $\bf X^{(A_m)}$ and $\bf X^{(B_m)}$ as the columns of $\bf X$
 observed in samples $A_m$ and $B_m$ respectively. Then the final
-estimator\footnote{This is sort of like stratified sampling with unknown strata.}
-is
-
-$$ \hat Y = M_1^{-1} \sum_{m = 1}^{M_1} \sum_{i \in A_m} \hat w_{im} y_i $$
+estimator\footnote{This is sort of like stratified sampling with unknown
+strata.} is
+
+$$ \hat Y_{MSE} = M_1^{-1} \sum_{m = 1}^{M_1} \sum_{i \in A_m} \hat w_{im} y_i $$
 
@@
 
-where $\hat T = (\hat{\bf X}^{(A_m)}, \sum_{i \in U} g(d_{im}) q_i)^T$ and $d_{im}$
+where $\hat T = (\hat{\bf X}^{(A_m)}, \sum_{i \in U} g(d_{im}) q_i)^T$, $\bf
+z_i^{T} = ((\bf x_i^{(A_m)})^T, g(d_{im})q_i)^T$ and $d_{im}$
 are the design weights for $A_m$.
@@
 \subsection{Theoretical Results}
+
+We prove the theory for the simpler case in which we have samples $A_0, A_1,
+\dots, A_{M}$.
+
+\begin{theorem}[Design Consistency]
+  Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$.
+  Under regularity conditions,
+
+  $$\hat Y_{MS} = \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) + O_p(N / n_0)$$
+
+  where
+
+  $$\hat Y_\ell(\bm \lambda^*, \bm \phi^*) = \hat Y_{DCE} + \left(
+    \hat T - \sum_{i \in A_0} \hat w_{0i}(\bm \lambda^*) \bf z_i q_i
+  \right)\bm \phi^*$$
+
+  and
+
+  $$\bm \phi^* = \left[\sum_{i \in U} \frac{\pi_{0i} \bf z_i \bf z_i^T
+  q_i}{g'(d_{0i})}\right]^{-1} \sum_{i \in U} \frac{\pi_{0i} \bf z_i
+  y_i}{g'(d_{0i})}.$$
+\end{theorem}
+
+\begin{proof}
+  This proof follows from the proof of Theorem~\ref{thm:dc2}. The only
+  difference is that we have $\hat X_{GLS}$ instead of $\hat X_c$, but the final
+  result still holds.
+\end{proof}
+
+Since this result holds for each simple case, and our general case is the
+average of each of the simple cases that we observe, the result holds for
+Equation~\eqref{eq:dc32} as we assume each $A_m$ is independent of each other.
+
+\begin{theorem}[Variance Estimation]
+  Under regularity conditions,
+
+  $$
+  \begin{aligned}
+    V(\hat Y_{MS}) 
+    &= M_1^{-2} \left(\sum_{m = 1}^{M_1} 
+    \Big\{(\bm \phi_1^{(m)*})^T\Var(\hat{\bf X}_{GLS}^{(m)})
+      (\bm \phi_1^{(m)*})^T \right.\\
+    &+ \left.
+      \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{mij}}{\pi_{mi}\pi_{mj}}
+      (y_i - \bf z_i^{(m)}\bm \phi^{(m)*}q_i)
+      (y_j - \bf z_j^{(m)}\bm \phi^{(m)*}q_j) \Big\} \right).
+  \end{aligned}
+  $$
+
+  We can estimate the variance with
+
+  $$
+  \begin{aligned}
+    \hat V(\hat Y_{MS}) 
+    &= M_1^{-2} \left(\sum_{m = 1}^{M_1} 
+      \Big\{(\hat{\bm \phi_1}^{(m)})^T\hat \Var(\hat{\bf X}_{GLS}^{(m)})
+        (\hat{\bm \phi_1}^{(m)})^T \right.\\
+    &+ \left.
+      \sum_{i \in A_m} \sum_{j \in A_m} \frac{\Delta_{mij}}{\pi_{mij}\pi_{mi}\pi_{mj}}
+      (y_i - \bf z_i^{(m)}\hat{\bm \phi}^{(m)}q_i)
+      (y_j - \bf z_j^{(m)}\hat{\bm \phi}^{(m)}q_j) \Big\} \right).
+  \end{aligned}
+  $$
+
+\end{theorem}
+
+\begin{proof}
+  This result follows the same argument as Theorem~\eqref{thm:nnve}. The result
+  holds because each sample $A_m$ is independent from the other $A_{m'}$. Since,
+  the score equation for the regression estimator asserts that $\sum_{i \in A_m}
+  \bf z_i (y_i - \bf z_i \hat{\bm \phi}^{(m)}) = \bf 0$, the covariance terms
+  should be approximately zero.
+\end{proof}
 

### 2024-07-29 final/reports/prelim_paper_ov.tex
Updated July 29, 2024

@@
 
-While unbiased (see \cite{kim2024statistics}), the $\pi^*$-estimator
-does not account for the additional information contained in the auxiliary Phase
-1 variable $X$. The two-phase regression estimator $\hat Y_{reg, tp}$ does incorporate 
-information for $X$ by using the estimate $\hat X_1$ from the Phase 1 sample.
-This is how we can leverage the external information $\hat X_1$ to improve the
-initial $\pi^*$-estimator in the second phase sample.
-The two-phase regression estimator has the form,
+While unbiased (see \cite{kim2024statistics}), the $\pi^*$-estimator does not
+account for the additional information contained in the auxiliary Phase 1
+variable $\bf X$. The two-phase regression estimator $\hat Y_{reg, tp}$ does
+incorporate information for $\bf X$ by using the estimate $\hat{\bf X_1}$ from
+the Phase 1 sample. This is how we can leverage the external information 
+$\hat{\bf{X_1}}$ to improve the initial $\pi^*$-estimator in the second phase
+sample. The two-phase regression estimator has the form,
 
 $$ \hat Y_{reg, tp} 
-= \sum_{i \in A_1} \frac{1}{\pi_{1i}} x_i \hat \beta_q+ \sum_{i \in A_2}
-\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_q)$$
-
-where for $q_i = q(x_i)$ and is a function of $x_i$,
-$$
-\hat \beta_q = \left(\sum_{i \in A_2} 
-  \frac{x_i x_i'}{\pi_{1i} q_i}\right)^{-1} 
-\sum_{i \in A_2} \frac{x_i y_i}{\pi_{1i} q_i}.$$
+= \sum_{i \in A_1} \frac{1}{\pi_{1i}} \bf x_i \hat{\bm \beta_q} + 
+\sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i \hat{\bm \beta_q})$$
+
+where for $q_i = q(\bf x_i)$ and is a function of $\bf x_i$,
+$$
+\hat{\bm \beta_q} = \left(\sum_{i \in A_2} 
+  \frac{\bf x_i \bf x_i'}{\pi_{1i} q_i}\right)^{-1} 
+\sum_{i \in A_2} \frac{\bf x_i y_i}{\pi_{1i} q_i}.$$
 
@@
 calibration levels. It does not handle the
-two-phase setup where we need to estimate the finite population total of $x$
+two-phase setup where we need to estimate the finite population total of $\bf x$
 from the Phase 1 sample. In the rest of the section, we extend this method to 
-two phase sampling so that we have a valid 
+two-phase sampling so that we have a valid 
 estimator when including estimated Phase 1 weights with appropriate variance
@@
 \begin{equation}\label{eq:primal}
-  \argmin_{w_{2|1}} \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i 
+  \argmin_{w_{2|1}} \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
   \text{ such that}
-  \sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i = \sum_{i \in A_1} w_{1i} z_i q_i
-\end{equation}
-
-where  $z_i = (x_i / q_i, g(\pi_{2i|1}^{-1}))$. 
+  \sum_{i \in A_2} w_{1i} w_{2i|1} \bf z_i q_i = \sum_{i \in A_1} w_{1i} \bf z_i q_i
+\end{equation}
+
+where  $\bf z_i = (\bf x_i / q_i, g(\pi_{2i|1}^{-1}))$. 
 Let $\hat w_{2i|1}$ be the solution to Equation~\eqref{eq:primal}. The resulting 
@@
 
-$$ \hat w_{2i|1}(\hat \lambda) = g^{-1}(\hat \lambda^T z_i) $$
-
-where $\hat \lambda$ is the solution to
+$$ \hat w_{2i|1}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda}^T \bf z_i) $$
+
+where $\hat{\bm \lambda}$ is the solution to
 
 \begin{equation}\label{eq:lamdc1}
-  \left( \sum_{i \in A_1} w_{1i} z_i q_i -
-    \sum_{i \in A_2} w_{1i} w_{2i|1}(\hat \lambda) z_i q_i\right) = 0.
+  \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
+  \sum_{i \in A_2} w_{1i} w_{2i|1}(\hat{\bm \lambda}) \bf z_i q_i\right) = 0.
 \end{equation}
@@
 
-  $$\hat Y_{\ell}(\lambda, \phi^*) = \hat Y_{DCE}(\hat \lambda) + 
-  \left(\sum_{i \in A_1} w_{1i} z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(
-  \lambda) z_i q_i\right)\phi^*$$
+  $$\hat Y_{\ell}(\bm \lambda, \bm \phi^*) = \hat Y_{DCE}(\hat{\bm \lambda}) + 
+  \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(
+  \bm \lambda) z_i q_i\right)\bm \phi^*$$
 
@@
   $$\frac{\partial \mathcal{L}}{\partial w_{2i|1}}: g(w_{2i|1}) w_{1i}q_i -
-    \lambda w_{1i} z_i q_i = 0.$$
-
-  Hence, $\hat w_{2i}(\lambda) = g^{-1}(\lambda^T z_i)$ and $\hat \lambda$ is
-  determined by Equation~\eqref{eq:lamdc1}. When the sample size gets large, we
-  have $\hat w_{2i|1}(\hat \lambda) \to d_{2i|1}$ which means that $\hat \lambda
-  \to \lambda^*$ where $\lambda^* = (\bf 0, 1)$. Then using the linearization
-  technique of \cite{randles1982asymptotic}, we can construct a regression
-  estimator,
-
-  $$\hat Y_\ell(\hat \lambda, \phi)  = \hat Y_{DCE}(\hat \lambda) + 
-  \left(\sum_{i \in A_1} w_{1i} z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat
-  \lambda) z_i q_i\right)\phi.$$
-
-  Notice that $\hat Y_\ell(\hat \lambda, \phi) = \hat Y_{DCE}(\hat \lambda)$ for
-  all $\phi \in \mathbb{R}$. We choose $\phi^*$ such that
-
-  $$E\left[\frac{\partial}{\partial \lambda} \hat Y_\ell(\lambda^*, \phi^*)\right]=0.$$
-
-  Using the fact that $g^{-1}(\lambda^* z_i) = g^{-1}(g(d_{2i|1})) = d_{2i|1}$
+  w_{1i} \bm \lambda \bf z_i q_i = 0.$$
+
+  Hence, $\hat w_{2i}(\bm \lambda) = g^{-1}(\bm \lambda^T \bf z_i)$ and
+  $\hat{\bm \lambda}$ is determined by Equation~\eqref{eq:lamdc1}. When the
+  sample size gets large, we have $\hat w_{2i|1}(\hat{\bm \lambda}) \to
+  d_{2i|1}$ which means that $\hat{\bm \lambda} \to \bm \lambda^*$ where $\bm
+  \lambda^* = (\bf 0, 1)$. Then using the linearization technique of
+  \cite{randles1982asymptotic}, we can construct a regression estimator,
+
+  $$\hat Y_\ell(\hat{\bm \lambda}, \bm \phi)  = \hat Y_{DCE}(\hat{\bm \lambda}) + 
+  \left(\sum_{i \in A_1} w_{1i} \bm z_i q_i - \sum_{i \in A_2} w_{1i} 
+  \hat w_{2i|1}(\hat{\bm \lambda}) \bf z_i q_i\right)\bm \phi.$$
+
+  Notice that $\hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = \hat Y_{DCE}(\hat{\bm
+  \lambda})$ for all $\bm \phi$. We choose $\bm \phi^*$ such that
+
+  $$E\left[\frac{\partial}{\partial \bm \lambda} \hat Y_\ell(\bm \lambda^*,
+  \bm \phi^*)\right]=0.$$
+
+  Using the fact that $g^{-1}(\bm \lambda^* \bf z_i) = g^{-1}(g(d_{2i|1})) = d_{2i|1}$
   and $(g^{-1})'(x) = 1 / g'(g^{-1}(x))$, we have
@@
     \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$
-    and $\bf x_i = (x_{1i}, x_{2i})^T$,
+    and $\bf x_i = (1, x_{1i}, x_{2i})^T$,
     % \textcolor{red}{Dr. Kim, should I modify the simulation so that the
@@
 
-    \textcolor{red}{You should include an intercept in $\bf x_i$. }
+    %\textcolor{red}{You should include an intercept in $\bf x_i$. }
+    %\textcolor{blue}{I have been including an intercept, but I forgot to add it
+    %into these notes.}
   \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
@@
         \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
-    & {\scriptsize$\eta_i = z_i' \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
-        \varepsilon_i = (y_i - z_i' \phi)$} \\
+    & {\scriptsize$\eta_i = \bf z_i' \bm \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
+        \varepsilon_i = (y_i - \bf z_i' \bm \phi)$} \\
+    Eta-Pop & Same as DC-Pop & \\
+    Eta-Est & Same as DC-Est & \\
     \bottomrule
@@
 %\textcolor{red}{I am not sure about the variance estimation formula in Table 1. 
-They are different from what we discussed before.
-I wonder whether you can make the sampling rate $n/N$ negligible so that the second term becomes insignificant.
-How about making $N=50,000$, $n_1=500$, and $n_2=200$?}
+%They are different from what we discussed before. I wonder whether you can make
+%the sampling rate $n/N$ negligible so that the second term becomes
+%insignificant. How about making $N=50,000$, $n_1=500$, and $n_2=200$?}
+%
+%\textcolor{blue}{The result with $N = 50,000$, $n_1 = 500$ and $E[n_2] \approx
+%200$}
+%
+%\begin{table}[ht!]
+%  \centering
+%  {\color{blue}\input{tables/tpdcsim50000_mean.tex}}
+%% This table was generated from ../../src/explore/proto-dctpmcmc.R
+%  \caption{\textcolor{blue}{This table shows the results of Simulation Study 1 with
+%      a different finite population and Phase 2 selection probability.
+%      It displays the Bias, RMSE, empirical 95\% confidence interval, and a
+%      t-statistic assessing the unbiasedness of each estimator for the
+%      estimators: $\pi^*$, TP-Reg, DC-Pop, and DC-Est.}}
+%\end{table}
 
@@
 
-Now we consider the sampling mechanism known as non-nested two phase sampling 
-(\cite{hidiroglou2001double}). In the last section, we considered two phase sampling
+Now we consider the sampling mechanism known as non-nested two-phase sampling 
+(\cite{hidiroglou2001double}). In the last section, we considered two-phase sampling
 in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
-two phase sampling the Phase 2 sample is independent of the Phase 1 sample. It
+two-phase sampling the Phase 2 sample is independent of the Phase 1 sample. It
 is a separate independent sample of the same population. Like traditional two
 phase sampling, we consider the Phase 1 sample, $A_1$, to consist of
-observations of $(X_i)_{i = 1}^{n_1}$ and the Phase 2 sample, $A_2$, to consist
-of observations of $(X_i, Y_i)_{i = 1}^{n_2}$.
-
-Whereas the classical two phase estimator uses a single Horvitz-Thompson
+observations of $(\bf X_i)_{i = 1}^{n_1}$ and the Phase 2 sample, $A_2$, to consist
+of observations of $(\bf X_i, Y_i)_{i = 1}^{n_2}$.
+
+Whereas the classical two-phase estimator uses a single Horvitz-Thompson
 estimator of the Phase 1 sample to construct estimates for calibration totals,
-in the non-nested two phase sample we have two independent Horvitz-Thompson
-estimators of the total of $X$,
-
-$$\hat X_1 = \sum_{i \in A_1}^{n_1} d_{1i} x_i \text{ and } 
-\hat X_2 = \sum_{i \in A_2}^{n_2} d_{2i} x_i $$
+in the non-nested two-phase sample we have two independent Horvitz-Thompson
+estimators of the total of $\bf X$,
+
+$$\hat{\bf X_1} = \sum_{i \in A_1}^{n_1} d_{1i} \bf x_i \text{ and } 
+\hat{\bf X_2} = \sum_{i \in A_2}^{n_2} d_{2i} \bf x_i $$
 
@@
 probability of $i \in A_1$ and $\pi_{2i} = \Pr(i \in A_2)$. 
-%\textcolor{red}{I don't know if it makes sense to change the Phase 2
-%selection probability as $\pi_{2i|1}$. (\textcolor{blue}{No. We do not know
-%$\pi_{2i |1}$ and it is not necessary.}) Should I keep it at $\pi_{2i}$?
-%\textcolor{blue}{(Yes, $\pi_{2i}$ is available to us.) } But if
-%I do, how is the two phase setup useful?} 
-%\textcolor{blue}{Non-nested two-phase sampling is different from the classical
-%  two-phase sampling. We are not using the two-phase setup of Section 2  here.
-%  We do not have to call it two-phase sampling. However, if we define $A_c=A_1
-%  \cup A_2$ then $A_c$ serves the role of the first-phase sample in the
-%  two-phase sampling. Then, our goal is to compute an estimator of $X$ from
-%  $A_c$. The second-phase sample would be $A_2$. 
-%}
-We can combine these estimates to get
-$\hat X_c = W \hat X_1 + (1 - W)\hat X_2$ for some matrix $W$ (see
-\cite{merkouris2004combining} for the optimal choice of $W$), and can then define
-a regression estimator as
-
-$$
-\hat Y_{NN, reg} = \hat Y_2 + (\hat X_c - \hat X_2)^T \hat \beta_q = 
-\hat Y_2 + (\hat X_1 - \hat X_2)^T W^T\hat \beta_q 
-$$
-
-where % \textcolor{red}{(we may need transpose in $W$)}
-
-$$\hat \beta_q = \left(\sum_{i \in A_2} \frac{x_i x_i^T}{q_i}\right)^{-1}
-\sum_{i \in A_2} \frac{x_i y_i}{q_i} \text{ and }\hat Y_2 = \sum_{i \in A_2}
-d_{2i} y_i. $$
+We can combine these estimates using the effective sample size 
+(\cite{kish1965survey}) to get
+$\hat{\bf X_c} = (n_{1, eff} \hat{\bf X_1} + n_{2, eff}\hat{\bf X_2}) / 
+(n_{1, eff} + n_{2, eff})$ where $n_{1, eff}$ and $n_{2, eff}$ are the effective
+sample size for $A_1$ and $A_2$ respectively. Then we can define a regression
+estimator as
+
+$$
+\hat Y_{NN, reg} = \hat Y_2 + (\hat{\bf X_c} - \hat{\bf X_2})^T \hat{\bm \beta_q} = 
+\hat Y_2 + (\hat{\bf X_1} - \hat{\bf X_2})^T W\hat{\bm \beta_q}
+$$
+
+where, $W = n_{1, eff} / (n_{1, eff} + n_{2, eff})$, and
+
+$$\hat{\bm \beta_q} = \left(\sum_{i \in A_2} 
+\frac{\bf x_i \bf x_i^T}{q_i}\right)^{-1} 
+\sum_{i \in A_2} \frac{\bf x_i y_i}{q_i} \text{ and
+}\hat Y_2 = \sum_{i \in A_2} d_{2i} y_i. $$
 
@@
 $$V(\hat Y_{NN, reg}) = V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i -
-x_i^TW\beta^*_q)\right) + (\beta^*_q)^T W^T V(\hat X_1) W \beta_q^*$$
-
-where $\beta_q^*$ is the probability limit of $\hat \beta_q$. Like the two phase
+\bf x_i^TW\bm \beta^*_q)\right) + (\bm \beta^*_q)^T W^T V(\hat{\bf X_1}) W \bm \beta_q^*$$
+
+where $\bm \beta_q^*$ is the probability limit of $\hat{\bm \beta_q}$. Like the two-phase
 sample this regression estimator can be viewed as the solution to the following
@@
 We extend the debiased calibration estimator of \cite{kwon2024debiased} to the
-non-nested two phase sampling case where we use a combined estimate $\hat X_c$
-as the calibration totals instead of using the true totals from the finite
-population.
+non-nested two-phase sampling case where we use a combined estimate 
+$\hat{\bm X_c}$ as the calibration totals instead of using the true totals from
+the finite population.
 
@@
 for $g(x) = G'(x)$ and a known choice of $q_i \in \mathbb{R}$. 
-%\textcolor{red}{(Instead of using $\hat{X}_{\rm GLS}$, we used $\hat{X}_c$
-%before. Please make notation consistent. )}
-
 The difference
 between solving Equation~\ref{eq:nnopt} and Equation~\ref{eq:primal} is that the
-estimator $\hat X_{c}$ is estimated from the combined sample $A_c= A_1 \cup
-A_2$. Before using
-$\hat X_{c}$ in the debiased calibration estimator, we need to estimate it
-from the non-nested samples.
-
-We can get multiple estimates of $\hat X$,
-
-$$\hat X_1 = \sum_{i \in A_1} d_{1i} x_i \text{ and }
-\hat X_2 = \sum_{i \in A_2} d_{2i} x_i.$$
-
-Let $V_1$ and $V_2$ be known variance matrices for $\hat X_1$ and $\hat X_2$,
-then the optimal combined estimate is
-
-$$\hat X_c = \frac{V_1^{-1} \hat X_1 + V_2^{-1} \hat X_2}{V_1^{-1} + V_2^{-1}}.$$
-
-We can constuct a non-nested two phase estimator $\hat Y_{NNE}$ for $Y_N$ where
+estimator $\hat{\bf X_c}$ is estimated from the combined sample $A_c= A_1 \cup
+A_2$. Before using $\hat{\bf X_c}$ in the debiased calibration estimator, we
+need to estimate it from the non-nested samples. We can get multiple estimates
+of $\hat{\bf X}$,
+
+$$\hat{\bf X_1} = \sum_{i \in A_1} d_{1i} \bf x_i \text{ and }
+\hat{\bf X_2} = \sum_{i \in A_2} d_{2i} \bf x_i.$$
+
+Let $V_1$ and $V_2$ be known variance matrices for $\hat{\bf X_1}$ and 
+$\hat{\bf X_2}$, then the optimal combined estimate is
+
+$$\hat{\bf X_c} = \frac{V_1^{-1} \hat{\bf X_1} + V_2^{-1} 
+\hat{\bf X_2}}{V_1^{-1} + V_2^{-1}}.$$
+
+We can constuct a non-nested two-phase estimator $\hat Y_{NNE}$ for $Y_N$ where
 $\hat Y_{NNE} = \sum_{i \in A_2} \hat w_{2i} y_i$ and $\hat w_{2i}$ solves
-Equation~\ref{eq:nnopt}. Like the classical two phase approach, to solve this
+Equation~\ref{eq:nnopt}. Like the classical two-phase approach, to solve this
 setup we minimize the Lagrangian,
@@
 
-$$ \hat w_{2i}(\hat \lambda) = g^{-1}(\hat \lambda^T z_i) $$
-
-where $\hat \lambda$ is the solution to
+$$ \hat w_{2i}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda^T} \bf z_i) $$
+
+where $\hat{\bm \lambda}$ is the solution to
 
 \begin{equation}\label{eq:lamdc2}
-  \left( \hat T - \sum_{i \in A_2} w_{2i}(\hat \lambda) z_i q_i\right) = 0.
+  \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i}(\hat{\bm \lambda}) \bf z_i
+  q_i\right) = 0.
 \end{equation}
@@
 \begin{theorem}[Design Consistency]\label{thm:dc2}
-  Allowing $\lambda^*$ to be the probability limit of $\hat \lambda$,
-  under some regularity conditions, $\hat Y_{NNE} = \hat Y_{\ell, NNE}(\lambda^*,
-  \phi^*) + O_p(Nn_2^{-1})$ where
-
-  $$\hat Y_{\ell, NNE}(\lambda^*, \phi^*) = \sum_{i \in A_2} \hat w_{2i}(
-  \lambda^*) + \left(\hat T - \sum_{i \in A_2} \hat w_{2i}(\lambda^*) z_i q_i\right)
-  \phi^*$$
+  Allowing $\bm \lambda^*$ to be the probability limit of $\hat{\bm \lambda}$,
+  under some regularity conditions, $\hat Y_{NNE} = \hat Y_{\ell, NNE}(\bm \lambda^*,
+  \bm \phi^*) + O_p(Nn_2^{-1})$ where
+
+  $$\hat Y_{\ell, NNE}(\bm \lambda^*, \bm \phi^*) = \sum_{i \in A_2} \hat w_{2i}(
+  \bm \lambda^*) + \left(\hat{\bf T} - \sum_{i \in A_2} \hat w_{2i}(
+  \bm \lambda^*) \bf z_i q_i\right) \bm \phi^*$$
 
   and 
-  $$\phi^* =
+  $$\bm \phi^* =
   \left(\sum_{i \in U} \frac{\pi_{2i} q_i}{g'(d_{2i})} 
   \begin{bmatrix}
-    x_i^2 / q_i & x_i g(d_{2i}) / q_i \\
-    x_i g(d_{2i}) / q_i & g(d_{2i})^2
+    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
+    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
   \end{bmatrix}
@@
   \sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})} 
-  \begin{bmatrix} x_i / q_i \\ g(d_i) \end{bmatrix}.
-  $$
-% \textcolor{red}{$T$  is not defined. }
+  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
 \end{theorem}
@@
 \begin{proof}
-  The proof of this result is very similar to the proof the Theorem~\ref{thm:dc1}.
-  The biggest difference is that the total for $X$ is estimated from both samples
-  using $\hat X_c$ instead of $\hat X_{HT}$ from the Phase 1 sample.
-
-  Since $\hat Y_{NNE} = \sum_{i \in A_2} \hat w_{2i}(\hat \lambda)y_i$ where
-  $\hat \lambda$ solves
+  The proof of this result is very similar to the proof the
+  Theorem~\ref{thm:dc1}. The biggest difference is that the total for $\bf X$ is
+  estimated from both samples using $\hat{\bf X_c}$ instead of $\hat{\bf
+  X_{HT}}$ from the Phase 1 sample.
+
+  Since $\hat Y_{NNE} = \sum_{i \in A_2} \hat w_{2i}(\hat{\bm \lambda})y_i$ where
+  $\hat{\bm \lambda}$ solves
 
   \begin{equation}
-    \sum_{i \in A_2} \hat w_{2i}(\lambda) q_i
+    \sum_{i \in A_2} \hat w_{2i}(\bm \lambda) q_i
     \underbrace{
     \begin{bmatrix} 
-      x_i / q_i \\ g(d_i)
-  \end{bmatrix}}_{z_i} = T
+      \bf x_i / q_i \\ g(d_i)
+  \end{bmatrix}}_{\bf z_i} = \bf T
   \end{equation}
@@
 
-  $$\hat Y_{\ell, NNE} (\hat \lambda, \phi) = \sum_{i \in A_2} \hat w_{2i}(\hat
-  \lambda) + \left(T - \sum_{i \in A_2} \hat w_{2i}(\hat \lambda) z_i q_i\right)
-  \phi.$$
-
-  If we choose $\phi^*$ such that $E\left[\frac{\partial}{\partial \lambda} 
-    \hat Y_{\ell, NNE}(\lambda^*, \phi^*)\right] = 0$, then
-
-  $$\phi^* =
+  $$\hat Y_{\ell, NNE} (\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} \hat
+  w_{2i}(\hat{\bm \lambda}) + \left(\bf T - \sum_{i \in A_2} \hat
+  w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right) \bm \phi.$$
+
+  If we choose $\bm \phi^*$ such that $E\left[\frac{\partial}{\partial \bm \lambda} 
+    \hat Y_{\ell, NNE}(\bm \lambda^*, \bm \phi^*)\right] = 0$, then
+
+  $$\bm \phi^* =
   \begin{bmatrix}
-    \phi^*_1 \\ \phi^*_2
+    \bm \phi^*_1 \\ \bm \phi^*_2
   \end{bmatrix} = 
@@
   \sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})} 
-  \begin{bmatrix} x_i / q_i \\ g(d_i) \end{bmatrix}.
-  $$
-
-  Hence, by a Taylor expansion around $\hat \lambda$,
-
-  $$\hat Y_{NNE}(\hat \lambda) = \hat Y_{\ell, NNE}(\lambda^*, \phi^*) +
-  O_p(Nn_2^{-1}).$$
+  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
+
+  Hence, by a Taylor expansion around $\hat{\bm \lambda}$,
+
+  $$\hat Y_{NNE}(\hat{\bm \lambda}) = \hat Y_{\ell, NNE}(\bm \lambda^*, 
+  \bm \phi^*) + O_p(Nn_2^{-1}).$$
 
@@
     \frac{y_i}{\pi_{2i}},$
-  \item[2.] Regression estimator (Reg): (FIXME)
+  \item[2.] Regression estimator (Reg): Let $\hat Y_{NN, reg} = \hat Y_{HT} + (
+    \hat{\bf X_c} - \hat{\bf X_{2, HT}}) \hat{\bm \beta_2}$ where $\hat Y_{HT} =
+    \sum_{i \in A_2} d_{2i} y_i$, $\hat{\bf X_c} = W \hat{\bf X_{1, HT}} + (1 -
+    W)\hat{\bf X_{2, HT}}$, $W = n_{1,eff} / (n_{1,eff} + n_{2, eff})$,
+    $\hat{\bf X_{1, HT}} = \sum_{i \in A_1} d_{1i} \bf x_{i}$, $\hat{\bf X_{2,
+    HT}} = \sum_{i \in A_2} d_{2i} \bf x_{i}$, $\bf x_i = (1, x_{1i}, x_{2i})^T$
+    and
+
+    $$ \hat{\bm \beta_2} = \left(\sum_{i \in A_2} \bf x_i \bf x_i^T\right)^{-1}
+    \sum_{i \in A_2} \bf x_i y_i.$$
+
+    Then $\hat{\bar Y}_{NN, reg} = \hat Y_{NN, reg} / N$.
   \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
@@
 In addition to estimating the mean parameter $\bar Y_N$, we also construct
-variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$. For each
-approach we give the variance estimate in Table~\ref{tab:varforms}.
-
-\section{Topic 3: Multi-source Two-Phase Sampling}
+variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$.
+% For each approach we give the variance estimate in Table~\ref{tab:varforms2}.
+
+We run the simulation $1000$ times for each of these methods and compute the
+Bias ($E[\hat Y] - \bar Y_N$), the RMSE ($\sqrt{\Var(\hat Y - \bar Y_N)}$), a 95\%
+empirical confidence interval ($\sum_{b = 1}^{1000} |\hat Y^{(b)} - \bar Y_N| \leq 
+\Phi(0.975)\sqrt{\hat V(\hat Y^{(b)})^{(b)}}$), and a T-test that assesses the
+unbiasedness of each estimator where $\hat Y^{(b)}$ is the result from the $b$th
+simulation replicate. The results are in Table~\ref{tab:nndc-mean}.
+
+\begin{table}[ht!]
+  \centering
+\input{tables/nndcsim_mean.tex}
+% This table was generated from ../../src/explore/proto-dctpmcmc.R
+\caption{This table shows the results of Simulation Study 2. It displays the
+Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
+unbiasedness of each estimator for the estimators: HT, Reg, DC-Pop, and
+DC-Est.}
+\label{tab:nndc-mean}
+\end{table}
+
+\section{Topic 3: Multi-Source Two-Phase Sampling}
+
+\subsection{Background}
+
+When considering non-nested two-phase sampling, we focused on the case of having
+two samples. Now, we consider incorporating more than two independent samples
+together with a debiasing constraint. First, we consider the case in which we
+want to estimate the $\theta = E[Y]$ and $Y$ is only observed in one sample.
+Then we will consider the case where $Y$ may be in multiple samples.
+
+Consider the setup in which we have independent samples $A_0, A_1, \dots, A_M$
+where $Y$ is observed only in $A_0$ but $\bf X$ is observed in all of the
+samples with elements $\bf X^{(0)}$ observed in $A_0$ and $\bf X^{(m)}$ observed
+in $A_m$ for each $m = 1, \dots, M$. Like the non-nested case, we assume that
+each survey is sampling independently from the same population sampling frame.
+
+The traditional multi-source approach (\cite{kim2024statistics}) is to use
+generalized least squares (GLS) to obtain an optimal estimator of $X$ from the
+samples $A_1, \dots, A_M$. Then we can incorporate this information in the
+estimation of $\theta$ by using the following estimate
+
+$$\hat \theta = \sum_{i \in A_0} \hat w_i y_i$$
+
+where
+
+\begin{equation}\label{eq:mstps}
+\hat w = \argmin_w \sum_{i \in A_0} G(w_i)q_i \text{ such that } \sum_{i \in
+A_0} w_i \bf x_i = \hat X_{GLS}^{(0)}
+\end{equation}
+
+where $\hat X_{GLS}^{(0)}$ is the GLS estimate of $X$ for all of the
+$X$-variables measured in sample $A_0$, and $G$ is a generalized entropy
+function.
+
+% Outline
+% Instead of incorporating data from one additional source, we want to combine
+% data from multiple sources
+% * Use GLS (generalized from effective sample size)
+% * We need to add debiasing constraint 
+%   * What do we do if we have multiple surveys for which we observe Y?
+%   * How do we incorporate the selection probability?
+
+\subsection{Methodology}
+
+In order to have a design consistent estimator that incorporates information from
+samples $A_0, A_1, \dots, A_M$, we want to add the debiasing constraints,
+
+$$ \sum_{i \in A_0} w_{i} \bf x_i q_i = \sum_{i \in U} g(d_i) q_i $$
+
+where $g(x) = \partial G(x) / \partial x$ and $d_i$ are the design weights for
+sample $A_0$. This yields the optimization problem:
+
+\begin{equation}\label{eq:dc3}
+  \hat w = \argmin_w \sum_{i \in A_0} G(w_i) q_i + \lambda\left(\hat T - \sum_{i
+  \in A_0} w_i \bf z_i\right)
+\end{equation}
+
+where $\hat T = (\hat{\bf X}^{(0)}, \sum_{i \in U} g(d_i) q_i)^T$ and 
+$\bf z_i = (\bf x_i^T, g(d_i)q_i)^T$. A more general approach considers the case
+in which multiple samples can observe $Y$. In this case, assume that we have
+samples $A_1, \dots A_{M_1}$ for which we observe elements of $\bf X$ and $Y$
+and samples $B_1, \dots, B_{M_2}$ for which we only observed elements of $\bf
+X$. We denote $\bf X^{(A_m)}$ and $\bf X^{(B_m)}$ as the columns of $\bf X$
+observed in samples $A_m$ and $B_m$ respectively. Then the final
+estimator\footnote{This is sort of like stratified sampling with unknown strata.}
+is
+
+$$ \hat Y = M_1^{-1} \sum_{m = 1}^{M_1} \sum_{i \in A_m} \hat w_{im} y_i $$
+
+and $\hat w_{im}$ solves
+
+\begin{equation}\label{eq:dc32}
+  \hat w = \argmin_w \sum_{i \in A_m} G(w_i) q_i + \lambda\left(\hat T_m - \sum_{i
+  \in A_m} w_i \bf z_i\right)
+\end{equation}
+
+where $\hat T = (\hat{\bf X}^{(A_m)}, \sum_{i \in U} g(d_{im}) q_i)^T$ and $d_{im}$
+are the design weights for $A_m$.
+
+\subsection{Theoretical Results}
+
+\subsection{Simulation Study 3.1}
+
+We first run a simulation for the case in which we only have one sample in which
+we observe $Y$. We have the following superpopulation model with $N=10000$
+elements:
+
+$$
+\begin{aligned}
+X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
+X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
+X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
+Z_i &\stackrel{ind}{\sim} N(0, 1) \\
+\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
+Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
+\pi_{0i} &= \min(\max(\Phi(-x_{1i}), 0.02), 0.9)\\
+\pi_{1i} &= n_1 / N.\\
+\pi_{2i} &= \Phi(x_{2i} - 2) \\
+\end{aligned}
+$$
+
+We observe the following columns in each sample
+
+\begin{table}[ht!]
+  \centering
+  \begin{tabular}{lrrrr}
+    \toprule
+    Sample & $X_1$ & $X_2$ & $X_3$ & $Y$ \\
+    \midrule
+   $A_0$  & \checkmark     & \checkmark     & \checkmark     &  \checkmark \\    
+   $A_1$  & \checkmark     &       & \checkmark     &    \\  
+   $A_2$  & \checkmark     & \checkmark     &       &    \\  
+   \bottomrule
+  \end{tabular}
+\end{table}
+
+For the sampling mechanism both $A_0$ and $A_1$ are selected using a Poisson
+sample with response probabilities $\pi_{0i}$ and $\pi_{1i}$ respectively. The
+sample $A_2$ is a simple random sample with $n_1 = 2000$. We compare five
+different estimators for $\theta = E[Y]$.
+
+\begin{itemize}
+  \item[1.] Horvitz-Thompson estimator (HT): $\hat Y = N^{-1} \sum_{i \in A_0}
+    \frac{y_i}{\pi_{0i}}$,
+  \item[2.] Non-nested regression (NNReg): This is the non-nested regression from
+    Equation~\eqref{eq:nnopt} with only using information from Samples $A_0$ and
+    $A_1$,
+  \item[3.] Multi-Source proposed (MSEst): This is the proposed estimator from
+    Equation~\eqref{eq:dc3},
+  \item[4.] Multi-Source population (MSPop): This is the proposed estimator with
+    using the true value of $T_1$ from the population,
+  \item[5.] Multi-Source regression (MSReg): This is the linearized regression
+    estimator from FIXME.
+\end{itemize}
+
+\begin{table}[ht!]
+  \centering
+  \input{tables/msdcsim1.tex}
+% This table was generated from ../../src/explore/20240506-msdcsim.qmd
+\caption{This table shows the results of Simulation Study 2. It displays the
+Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
+unbiasedness of each estimator for the estimators: HT, NNReg, MSPop, MSEst, and
+MSReg.}
+\label{tab:msdc-mean}
+\end{table}
+
+As expected MSPop has the lowest RMSE because it also uses the population
+totals. It seems like MSEst and MSReg are almost equivalent, which is good
+because they are also supposed to be identical. The MSEst estimator outperforms
+NNReg because it also uses information from $A_2$, even though it implicitly
+uses a regression estimator with $X_3$ as a covariate, which is unnecessary.
+
+\subsection{Simulation Study 3.2}
+
+Building off of the previous simulation study, we now consider the case for
+which we observe $Y$ in separate samples.
+We have the following superpopulation model with $N = 10,000$ elements:
+
+$$
+\begin{aligned}
+X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
+X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
+X_{3i} &\stackrel{ind}{\sim} N(5, 1) \\
+\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
+Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
+\pi_{A1i} &= \min(\max(\Phi(-x_{1i}), 0.02), 0.9)\\
+\pi_{A2i} &= \min(\max(\Phi(x_{3i} - 6), 0.02), 0.95) \\
+\pi_{B1i} &= n_1 / N\\
+\pi_{B2i} &= \Phi(x_{2i} - 2) \\
+\end{aligned}
+$$
+
+We observe the following columns in each sample
+
+\begin{table}[ht!]
+  \centering
+  \begin{tabular}{lrrrr}
+    \toprule
+    Sample & $X_1$ & $X_2$ & $X_3$ & $Y$ \\
+    \midrule
+   $A_1$  & \checkmark     & \checkmark     & \checkmark     &  \checkmark \\    
+   $A_2$  & \checkmark     & & \checkmark     &  \checkmark \\    
+   $B_1$  & \checkmark     &       & \checkmark     &    \\  
+   $B_2$  & \checkmark     & \checkmark     &       &    \\  
+   \bottomrule
+  \end{tabular}
+\end{table}
+
+For the sampling mechanism $A_1$, $A_2$ and $B_2$ are selected using a Poisson
+sample with response probabilities $\pi_{A1i}$, $\pi_{A2i}$ and $\pi_{B1i}$
+respectively. The $B_2$ sample is a simple random sample (SRS). We estimate the
+following methods:
+
+\begin{itemize}
+  \item[1.] Horvitz-Thompson estimator for $A_1$ (HTA1):
+
+    $$\hat Y_{HT, A_1} = \sum_{i \in A_1} \frac{y_i}{\pi_{A1i}}.$$
+  \item[2.] Horvitz-Thompson estimator for $A_2$ (HTA1):
+    $$\hat Y_{HT, A_2} = \sum_{i \in A_2} \frac{y_i}{\pi_{A2i}}.$$
+  \item[3.] Weighted Horvitz-Thompson estimator (WHT):
+    $$\hat Y_{WHT} = \frac{\hat Y_{HT, A_1} / \hat V_{A_1} + \hat Y_{HT, A_2} /
+    \hat V_{A_2}}{1 / \hat V_{A_1} + 1 / \hat V_{A_2}}$$
+
+    where $\hat V_{A_m}$ is the estimated variance of the Horvitz-Thompson
+    estimator $\hat Y_{HT, A_m}$.
+  \item[4.] Multi-Source proposed (MSEst): This is the proposed estimator from
+    Equation~FIXME.
+  \item[5.] Multi-Source population (MSPop): This is the proposed estimator with
+    using the true value of $T_1$ from the population,
+  \item[6.] Multi-Source regression (MSReg): This is the linearized regression
+    estimator from FIXME.
+\end{itemize}
+
+\begin{table}[ht!]
+  \centering
+  \input{tables/msdcsim2.tex}
+% This table was generated from ../../src/explore/20240511-msdcsim2.qmd
+\caption{This table shows the results of Simulation Study 2. It displays the
+Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
+unbiasedness of each estimator for the estimators: HT, NNReg, MSPop, MSEst, and
+MSReg.}
+\label{tab:msdc2-mean}
+\end{table}
+
+% FIXME: Change to GLS estimate of Y.
 

### 2024-07-29 src/explore/20240425-nndcsim.qmd
Added simulations for debiased nonparametric regression. We still need to find a good setup to show the strength of our result.
Updated July 29, 2024

@@
 library(dplyr)
-library(GECal)
+# library(GECal)
 library(nleqslv)
@@
 Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
-\pi_{1i} &= \Phi(-x_{3i} - 2)\\
-\pi_{2i|1} &= x_{4i}.
+\pi_{1i} &= FIXME\\
+\pi_{2i|1} &= FIXME.
 \end{aligned}
@@
   x3 <- rnorm(obs, 0, 1)
+  z <- rnorm(obs, 0, 1)
   x4 <- runif(obs, 0.1, 0.9)
@@
 
-  y <- 3 * x1 + 2 * x2 + eps
+  y <- 3 * x1 + 2 * x2 + z + eps
 
-  return(tibble(X1 = x1, X2 = x2, X3 = x3, X4 = x4, Y = y))
+  return(tibble(X1 = x1, X2 = x2, X3 = x3, X4 = x4, Z = z, Y = y))
 }
@@
     "n1_obs / N_obs",
-    "ifelse(pt(-pop_df$X1 + 1, 3) > 0.9, 0.9, pt(-pop_df$X1 + 1, 3))"
+    "pmin(pmax(pt(pop_df$Z - 2.5, 3), 0.01), 0.9)"
   )

### 2024-07-29 src/explore/20240506-msdcsim.qmd
Updated July 29, 2024

@@
 \varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
-Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
-\pi_{0i} &= \min(\max(\Phi(-x_{1i}), 0.02), 0.9)\\
+Y_{i} &= 3 X_{1i} + 2 X_{2i} + Z_i + \varepsilon_i \\
+\pi_{0i} &= \min(\max(\Phi(z_{i} - 2), 0.02), 0.9)\\
 \pi_{1i} &= n_1 / N\\
@@
 
-  y <- 3 * x1 + 2 * x2 + eps
+  y <- 3 * x1 + 2 * x2 + z + eps
 
-  pi0 <- pmin(pmax(pnorm(-x1), 0.02), 0.9)
+  pi0 <- pmin(pmax(pnorm(z - 2), 0.02), 0.9)
   # pi1 <- pnorm(-z - 1) # Informative sampling. Do we want this?
@@
   arrange(Ind) %>%
-  select(Est, Bias, RMSE, EmpCI, Ttest) %>%
+  select(Est, Bias, RMSE, EmpCI, Ttest)
+
+%>%
   knitr::kable("latex", booktabs = TRUE, digits = 4) %>%

### 2024-07-29 src/explore/proto-dctp.R

@@ -39,11 +39,12 @@ gen_pop <- function(obs) {
   x2 <- runif(obs, 0, 4)
   x3 <- rnorm(obs, 0, 1)
   x4 <- runif(obs, 0.1, 0.9)
+  z <- rnorm(obs, 0, 1)
   eps <- rnorm(obs)
 
-  y <- 3 * x1 + 2 * x2 + eps
+  y <- 3 * x1 + 2 * x2 + 0.5 * z + eps
 
-  return(tibble(X1 = x1, X2 = x2, X3 = x3, X4 = x4, Y = y))
+  return(tibble(X1 = x1, X2 = x2, X3 = x3, X4 = x4, Z = z, Y = y))
 }
 
 #' This function adds sampling probabilites to a population data frame.
@@ -120,7 +121,7 @@ pi_star <- function(p1_df, p2_df, pop_obs, sampling = "poisson-poisson") {
       for (j in 1:nrow(p2_df)) {
         if (i == j) {
           pi2i <- p2_df$pi2[i] * p2_df$pi1[i]
-          tmp[i, j] <- ((pi2i - (pi2i)^2 / pi2i) * p2_df$Y[i]^2 / (pi2i^2)
+          tmp[i, j] <- ((pi2i - (pi2i)^2) / pi2i) * p2_df$Y[i]^2 / (pi2i^2)
         } else {
           pi2i <- p2_df$pi2[i] * p2_df$pi1[i]
           pi2j <- p2_df$pi2[j] * p2_df$pi1[j]
@@ -381,32 +382,32 @@ dc_ybar <-
 
       # # This works but why 1 / nrow(p2_df) in v1 instead of 1 / nrow(p1_df)?
       # Because, we essentially do not use the first stage.
-      # var_eps <- var(eps2)
-      # v1 <- (1 / nrow(p2_df) - 1 / nrow(pop_df)) * var_eps
-      # v2 <- 
-      #   sum((1 - p2_df$pi2) / (p2_df$pi1^2 * (1 / dc_w)^2) * eps2^2) / nrow(pop_df)^2
-      # vhat <- v1 + v2
+       var_eps <- var(eps2)
+       v1 <- (1 / nrow(p2_df) - 1 / nrow(pop_df)) * var_eps
+       v2 <- 
+         sum((1 - p2_df$pi2) / (p2_df$pi1^2 * (1 / dc_w)^2) * eps2^2) / nrow(pop_df)^2
+       vhat <- v1 + v2
 
[…]

### 2024-07-29 src/explore/proto-dctpmcmc.R

@@ -24,7 +24,8 @@ p1_type <- "srs"
 # p1_pi <- "ifelse(pnorm(-pop_df$X2) < 0.10, 0.10, pnorm(-pop_df$X2))"
 p1_pi <- "1200 / nrow(pop_df)"
 p2_type <- "poisson"
-p2_pi <- "ifelse(pt(-pop_df$X1 + 1, 3) > 0.9, 0.9, pt(-pop_df$X1 + 1, 3))"
+# p2_pi <- "ifelse(pt(-pop_df$X1 + 1, 3) > 0.9, 0.9, pt(-pop_df$X1 + 1, 3))"
+p2_pi <- "pmin(pmax(pt(pop_df$Z - 1, 3), 0.02), 0.7)"
 comb <- paste0(p1_type, "-", p2_type)
 
 # ************
@@ -118,9 +119,7 @@ mc_res %>%
     Est == "pistar" ~ "$\\pi^*$")
   ) %>%
   arrange(Ind) %>%
-  select(Est, Bias, RMSE, EmpCI, Ttest) 
-
-%>%
+  select(Est, Bias, RMSE, EmpCI, Ttest) %>%
   knitr::kable("latex", booktabs = TRUE, escape = FALSE, digits = 3) %>%
   cat(file = paste0("tables/tpdcsim",  "_mean.tex"))
 

### 2024-07-30 final/jsm_poster.tex
Added JSM poster. Rough draft complete.

% Unofficial ISU Poster template.
% A fork of https://github.com/anishathalye/gemini
% also refer to https://github.com/yhkwon/gemini-isu

% .pptx or illustrator version is available on https://www.print.iastate.edu/PrintingResources/DesignTemplates/

\documentclass[final]{beamer}

% ====================
% Packages
% ====================

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[size=custom,width=120,height=72,scale=1.0]{beamerposter}
\usetheme{gemini}
\usecolortheme{gemini}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{amsmath}
\usepackage{bm}
\pgfplotsset{compat=1.14}
\usepackage{anyfontsize}
\usepackage[backend=biber, style=bwl-FU]{biblatex}
\addbibresource{poster.bib}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\suchthat}{such\; that}
% ====================
% Lengths
% ====================

% If you have N columns, choose \sepwidth and \colwidth such that
% (N+1)*\sepwidth + N*\colwidth = \paperwidth
\newlength{\sepwidth}
\newlength{\colwidth}
\setlength{\sepwidth}{0.025\paperwidth}
\setlength{\colwidth}{0.3\paperwidth}

\newcommand{\separatorcolumn}{\begin{column}{\sepwidth}\end{column}}

% ====================
% Title
% ====================

\title{\hspace{25cm}Debiased Calibration for Generalized Two-Phase Sampling}

%\author{Your Name}
\author{Caleb Leedy \inst{1} \and Jae-Kwang Kim \inst{1}}

%\institute[shortinst]{Some Institute}
\institute[shortinst]{\inst{1} Iowa State University}

% ====================
% Footer (optional)
% ====================

\footercontent{
  \href{https://github.com}{Github: https://github.com/calebleedy} \hfill
JSM Poster Session \hfill
  \href{mailto:cleedy@iastate.edu}{cleedy@iastate.edu}}
% (can be left out to remove footer)

% ====================
% Logo (optional)
% ====================

% use this to include logos on the left and/or right side of the header:
% \logoright{\includegraphics[height=7cm]{logo1.pdf}}
% \logoleft{\includegraphics[height=7cm]{logo2.pdf}}

% ====================
% Body
% ====================

\begin{document}
\addtobeamertemplate{headline}{}
{
    \begin{tikzpicture}[remember picture,overlay]
        \node [anchor=north west, inner sep=3cm] at ([xshift=0.0cm,yshift=1.0cm]current page.north west){\includegraphics[height=2.5cm]{logos/watermark.png}}; % Iowa State University watermark
        % ISU watermark can be found on https://www.brandmarketing.iastate.edu/wordmark/primary/
        \node [anchor=north west, inner sep=3cm] at ([xshift=0.0cm,yshift=-2.5cm]current page.north west){\huge \textbf{Statistics Department}}; % Department name
        %\node [anchor=north east, inner sep=3cm] at ([xshift=0.0cm,yshift=1.0cm]current page.north east){\includegraphics[height=5cm]{logos/logo.png}}; % Use this if you want to add a logo on your right top corner.
    \end{tikzpicture}
}

\begin{frame}[t]
\begin{columns}[t]
\separatorcolumn

\begin{column}{\colwidth}

  \begin{block}{Introduction}

    In classical two-phase sampling, one first selects a sample $A_1$ from a
    finite population $U$ of size $N$, and observes $({\bf X}_i)_{i = 1}^{n_1}$.
    Then one selects a sample $A_2$ from $A_1$ and observes 
    $({\bf X}_i, Y_i)_{i = 1}^{n_2}$. We take the
    framework of two-phase sampling and view data integration as specific case
    of observing a variable of interest $Y$ within a single data set while we
    also observe common covariates $({\bf X}_i)$ between both the data set with
    $Y$ as well as an outside auxilary sample. This is an important practical
    problem \cite{yang2020statistical}, 
    \cite{dagdoug2023model}. We focus on the case:

    \begin{itemize}
      \item Where all of the surveys are probability samples, 
      \item Where we only have summary level information instead of individual
        observation values,
      \item When we want the estimate the finite population total of $Y$,
        $$Y_T = \sum_{i \in U} y_i.$$
    \end{itemize}

    \heading{Notation}

    \begin{itemize}
      \item Let $\pi_{1i}$ be the probability that element $i$ is selected into
        the Phase 1 sample, $A_1$.
      \item Let $\pi_{2i|1}$ be the probability that element $i$ is selected into
        the Phase 1 sample, $A_2$ conditional on the fact that $i \in A_1$.
      \item Let $d_{1i} = 1 / \pi_{1i}$ and $d_{2i|1} = 1 / \pi_{2i|1}$.
      \item We use $\delta_{1i}$ and $\delta_{2i}$ to indicate if an observation
        in contained within $A_1$ and $A_2$ respectively.
    \end{itemize}

  \end{block}

  \begin{block}{Goal}

    We want a two-phase sampling framework to

    \begin{itemize}
      \item[1.] Combine information from multiple data sources,
      \item[2.] In a way that is efficient, and
      \item[3.] Approximately design unbiased.
    \end{itemize}

  \end{block}

  % NOTE: Use the "alertblock" to highlight a particular section in green.
  \begin{block}{Comparable Methods}

    \begin{itemize}
      \item The \textbf{double expansion estimator} of \cite{kott1997can}
        is a Horvitz-Thompson like estimator that is design unbiased, but
        inefficient:

        $$
        \hat Y_{\rm{DEE}} = \sum_{i \in A_2} \frac{y_i}{\pi_{1i}\pi_{2i|1}}.
        $$
      \item The \textbf{two-phase regression estimator} is approximately design
        unbiased but not as efficient as our proposed method:
        $$
        \hat Y_{\rm{Reg}} = 
        \sum_{i \in A_1} \frac{{\bf{x}}_i \hat{\bm{\beta}}_q}{\pi_{1i}} + 
        \sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}}(y_i - 
        {\bf{x}}_i \hat{\bm{\beta}}_q)
        $$

        where $q_i = q(\bf x_i)$ and

        $$
        \hat{\bm \beta}_q = \left(\sum_{i \in A_2} 
        \frac{{\bf{x}}_i {\bf{x}}_i'}{\pi_{1i} q_i}\right)^{-1} 
        \sum_{i \in A_2} \frac{{\bf{x}}_i y_i}{\pi_{1i} q_i}.
        $$

    \end{itemize}

  \end{block}

\end{column}

\separatorcolumn

\begin{column}{\colwidth}

  \begin{block}{Methodology}

    \heading{Generalized Calibration}

    \begin{itemize}
      \item In a seminal paper, \cite{deville1992calibration} generalized the
        regression estimator to other loss functions besides squared-error loss
        for a sample $A$ with auxilary information about $X$. Their generalized
        loss function minimizes,

        $$
        \sum_{i \in A} G(w_{i}, d_{i}) \suchthat \sum_{i \in
        A} d_{i} w_{i} {\bf x_i} = \sum_{i \in U} {\bf x_i}.
        $$

        for a non-negative, strictly convex function with respect to $w$
        function $G$, with a minimum at $g(w_i, d_i) = \frac{\partial G}{\partial
        w}$ defined on an interval containing $d_{i}$ with $g(w_i, d_i)$
        continuous.
    \end{itemize}

    \heading{Calibration with Generalized Entropy}

    \begin{itemize}
      \item Recently, \cite{kwon2024debiased} proposed a calibration estimator that
        uses a generalized entropy function $G(w)$ (\cite{gneiting2007strictly})
        instead of the generalized loss function $G(w, d)$ of
        \cite{deville1992calibration}. They separate the bias calibration from
        the minimization term and solve the following equation for estimated
        sample weights:

        $$
        \hat w_{i} = \argmin_{w} \sum_{i \in A} G(w_i)
        \suchthat \sum_{i \in A} w_{i} {\bf{x}}_i = 
        \sum_{i \in U} {\bf{z}}_i,
        \sum_{i \in A} w_i g(d_i) = \sum_{i \in U} g(d_i) .
        $$

      \item Our proposed method extends their result to two-phase sampling.
    \end{itemize}

  \end{block}

  \begin{alertblock}{Proposal}

    Let ${\bf{z}}_i = ({\bf x}_i / q_i, g(d_{2i|1}))^T$, the proposed debiased
    calibration estimator is 
    \vspace{-1.0cm}

    \begin{equation}
    \hat Y_{\rm{DCE}} = \sum_{i \in A_2} d_{1i} \hat w_{2i|1} y_i
    \end{equation}

    \vspace{-1.0cm}

    where 
    \vspace{-0.5cm}

    \begin{equation}\label{eq:wmin}
    \hat w_{2i|1} = \argmin_{w_{2|1}} \sum_{i \in A_2} w_{1i} G(w_{2i|1}) q_i
    \suchthat \sum_{i \in A_2} d_{1i} w_{2i|1} {\bf{z}}_i q_i = 
    \sum_{i \in A_1} d_{1i} {\bf{z}}_i q_i.
    \end{equation}

    \vspace{-0.5cm}
  \end{alertblock}

  \begin{exampleblock}{Theoretical Results: Asymptotic Design Consistency}

  %\begin{theorem}[Design Consistency]\label{thm:dc1}
    Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$,
    where $\hat{\bm \lambda}$ are the corresponding Lagrange multipliers from
    Equation~\ref{eq:wmin}.
    Under some regularity conditions, $\bm \lambda^* = ({\bf 0}^T, 1)^T$ and

    $$\hat Y_{\rm{DCE}} = \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + O_p(N / n_2)$$

    where
    \vspace{-1.3cm}

    $$\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) = \hat Y_{\rm{DEE}}+ 
    \left(\sum_{i \in A_1} d_{1i} {\bf z}_i q_i - \sum_{i \in A_2} d_{1i} 
    \pi_{2i \mid 1}^{-1}  {\bf z}_i q_i\right)\bm \phi^*$$

    \vspace{-1.3cm}
    and
    \vspace{-0.3cm}

    $$
    \bm \phi^* = 
    \left[\sum_{i \in U} \frac{\pi_{2i|1}{\bf z}_i {\bf z}_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
    \sum_{i \in U} \frac{\pi_{2i|1}{\bf z}_i y_i}{g'(d_{2i|1})}.
    $$

    \vspace{-0.5cm}
  %\end{theorem}

  \end{exampleblock}

\end{column}

\separatorcolumn

\begin{column}{\colwidth}

  \begin{block}{Simulation Study}

  For a finite population of size $N = 10,000$, and $n_1 = 1000$,

  \begin{itemize}
    \item $X_{1i} \stackrel{ind}{\sim} N(2, 1)$,
      $X_{2i} \stackrel{ind}{\sim} \rm{Unif}(0, 4)$,
      $Z_{i} \stackrel{ind}{\sim} N(0, 1)$,
      $\varepsilon_i \stackrel{ind}{\sim} N(0, 1)$
    \item $Y_{i} = 3 X_{1i} + 2 X_{2i} + 0.5 Z_i + \varepsilon_i$
    \item $\pi_{1i} = n_1 / N$,
      $\pi_{2i|1} = \max(\min(\Phi_3(z_{i} - 1), 0.7), 0.02)$.
  \end{itemize}

  where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom. We
  compare the following algorithms:

  \begin{itemize}
    \item[1.] Double Expansion Estimator (DEE)
    \item[2.] Two-Phase Regression estimator (TP-Reg)
    \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
      solves

    \begin{equation*}
      \argmin_{w_{2|1}} \sum_{i \in A_2} d_{1i} G(w_{2i}) \suchthat
      \sum_{i \in A_2} d_{1i} w_{2i|1} {\bf z}_i = \sum_{i \in U} {\bf z}_i.
    \end{equation*}

    \item[4.] Debiased Calibration with Estimated Population Constraints (DC-Est):
      This solves Equation~\eqref{eq:wmin} with $q_i = 1$.
  \end{itemize}

  \begin{table}[ht!]
    \centering
  \input{tpdcsim_mean.tex}
  % This table was generated from ../../src/explore/proto-dctpmcmc.R
  \caption{This table shows the results of the simulation study. It displays the
  Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
  unbiasedness of each estimator for the estimators: DEE, TP-Reg, DC-Pop, and
  DC-Est.}
  \label{tab:tpdc-mean}
  \end{table}
  \end{block}

  \begin{block}{Extensions}

    We also consider the following extensions:

    \begin{itemize}
      \item Non-nested two-phase sampling: when $A_1$ and $A_2$ are independent.
      \item Multi-source sampling: when $Y$ is contained in a sample $A_0$ that
        shares common covariates $\bf X$, with samples $A_1$, $A_2$, \dots, $A_M$.
    \end{itemize}
  \end{block}

  \vspace{-1.0cm}

  \begin{block}{References}

    %\footnotesize{\bibliographystyle{plainnat}\bibliography{poster}}
    \printbibliography

  \end{block}

\end{column}

\separatorcolumn
\end{columns}
\end{frame}

\end{document}

### 2024-07-30 final/jsm_poster_ov.tex
Updated JSM poster. I think that it looks good.

--- 2024-07-30 final/jsm_poster.tex
@@
         \suchthat \sum_{i \in A} w_{i} {\bf{x}}_i = 
-        \sum_{i \in U} {\bf{z}}_i,
+        \sum_{i \in U} {\bf{x}}_i,
         \sum_{i \in A} w_i g(d_i) = \sum_{i \in U} g(d_i) .
@@
 
-    Let ${\bf{z}}_i = ({\bf x}_i / q_i, g(d_{2i|1}))^T$, the proposed debiased
+    Let ${\bf{z}}_i = ({\bf x}_i, g(d_{2i|1}))^T$, the proposed debiased
     calibration estimator is 
@@
 
-    where 
+    where $d_{1i} = \pi_{1i}^{-1}, d_{2i \mid 1}= \pi_{2i \mid 1}^{-1}$ and 
     \vspace{-0.5cm}
@@
     \begin{equation}\label{eq:wmin}
-    \hat w_{2i|1} = \argmin_{w_{2|1}} \sum_{i \in A_2} w_{1i} G(w_{2i|1}) q_i
-    \suchthat \sum_{i \in A_2} d_{1i} w_{2i|1} {\bf{z}}_i q_i = 
-    \sum_{i \in A_1} d_{1i} {\bf{z}}_i q_i.
+    \hat w_{2i|1} = \argmin_{w_{2|1}} \sum_{i \in A_2} d_{1i} G(w_{2i|1})
+    \suchthat \sum_{i \in A_2} d_{1i} w_{2i|1} {\bf{z}}_i = 
+    \sum_{i \in A_1} d_{1i} {\bf{z}}_i.
     \end{equation}
@@
     $$\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) = \hat Y_{\rm{DEE}}+ 
-    \left(\sum_{i \in A_1} d_{1i} {\bf z}_i q_i - \sum_{i \in A_2} d_{1i} 
-    \pi_{2i \mid 1}^{-1}  {\bf z}_i q_i\right)\bm \phi^*$$
+    \left(\sum_{i \in A_1} d_{1i} {\bf z}_i - \sum_{i \in A_2} d_{1i} 
+    \pi_{2i \mid 1}^{-1}  {\bf z}_i \right)\bm \phi^*$$
 
@@
     \bm \phi^* = 
-    \left[\sum_{i \in U} \frac{\pi_{2i|1}{\bf z}_i {\bf z}_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
+    \left[\sum_{i \in U} \frac{\pi_{2i|1}{\bf z}_i {\bf z}_i^T}{g'(d_{2i|1})}\right]^{-1}
     \sum_{i \in U} \frac{\pi_{2i|1}{\bf z}_i y_i}{g'(d_{2i|1})}.
@@
     \item[4.] Debiased Calibration with Estimated Population Constraints (DC-Est):
-      This solves Equation~\eqref{eq:wmin} with $q_i = 1$.
+      This solves Equation~\eqref{eq:wmin}.
   \end{itemize}

### 2024-08-02 final/jsm_poster.tex
Updated JSM poster size. Submitted final to print.

@@
 \usepackage{lmodern}
-\usepackage[size=custom,width=120,height=72,scale=1.0]{beamerposter}
+\usepackage[size=custom,width=96,height=72,scale=1.0]{beamerposter}
 \usetheme{gemini}
@@
 \usepackage{amsmath}
-\usepackage{bm}
+\usepackage{bm, ulem}
 \pgfplotsset{compat=1.14}
@@
 
-\title{\hspace{25cm}Debiased Calibration for Generalized Two-Phase Sampling}
+\title{\hspace{32cm}Debiased Calibration for Generalized Two-Phase Sampling}
 
@@
     \begin{tikzpicture}[remember picture,overlay]
-        \node [anchor=north west, inner sep=3cm] at ([xshift=0.0cm,yshift=1.0cm]current page.north west){\includegraphics[height=2.5cm]{logos/watermark.png}}; % Iowa State University watermark
+        \node [anchor=north west, inner sep=3cm] at ([xshift=0.0cm,yshift=1.0cm]current page.north west){\includegraphics[height=2.0cm]{logos/watermark.png}}; % Iowa State University watermark
         % ISU watermark can be found on https://www.brandmarketing.iastate.edu/wordmark/primary/
@@
     also observe common covariates $({\bf X}_i)$ between both the data set with
-    $Y$ as well as an outside auxilary sample. This is an important practical
-    problem \cite{yang2020statistical}, 
-    \cite{dagdoug2023model}. We focus on the case:
+    $Y$ as well as an outside auxiliary sample. 
+    %\sout{This is an important practical problem \cite{yang2020statistical}, 
+    %\cite{dagdoug2023model}.} 
+    We focus on the case:
 
@@
 
+    \vspace{-0.3cm}
     \heading{Notation}
@@
 
+  \vspace{-0.7cm}
   \begin{block}{Goal}
@@
   % NOTE: Use the "alertblock" to highlight a particular section in green.
+  \vspace{-0.7cm}
   \begin{block}{Comparable Methods}
@@
         where $q_i = q(\bf x_i)$ and 
+        \vspace{-0.5cm}
 
@@
 
+    \vspace{-0.3cm}
     \begin{itemize}
@@
         regression estimator to other loss functions besides squared-error loss
-        for a sample $A$ with auxilary information about $X$. Their generalized
+        for a sample $A$ with auxiliary information about $X$. Their generalized
         loss function minimizes,
@@
   %\begin{theorem}[Design Consistency]\label{thm:dc1}
-    Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$,
-    where $\hat{\bm \lambda}$ are the corresponding Lagrange multipliers from
-    Equation~\ref{eq:wmin}.
-    Under some regularity conditions, $\bm \lambda^* = ({\bf 0}^T, 1)^T$ and
+    %\sout{Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$,
+    %where $\hat{\bm \lambda}$ are the corresponding Lagrange multipliers from
+    %Equation~\ref{eq:wmin}.}
+    Under some regularity conditions, %\sout{$\bm \lambda^* = ({\bf 0}^T, 1)^T$} and
+    \vspace{-0.7cm}
 
-    $$\hat Y_{\rm{DCE}} = \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + O_p(N / n_2)$$
+    $$\hat Y_{\rm{DCE}} = \hat Y_\ell + O_p(N / n_2), $$
 
+    \vspace{-1.0cm}
     where 
@@
 
-    $$\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) = \hat Y_{\rm{DEE}}+ 
-    \left(\sum_{i \in A_1} d_{1i} {\bf z}_i q_i - \sum_{i \in A_2} d_{1i} 
-    \pi_{2i \mid 1}^{-1}  {\bf z}_i q_i\right)\bm \phi^*$$
+    $$\hat Y_{\ell} = \hat Y_{\rm{DEE}}+ 
+    \left(\sum_{i \in A_1} d_{1i} {\bf z}_i - \sum_{i \in A_2} d_{1i} 
+    \pi_{2i \mid 1}^{-1}  {\bf z}_i \right)\bm \phi^*$$
 
@@
     and
-    \vspace{-0.3cm}
+    \vspace{-0.7cm}
 
@@
 
-  \vspace{-1.0cm}
+  %\vspace{-1.0cm}
 

### 2024-08-02 final/jsm_poster_ov.tex
Updated JSM poster size. Submitted final to print.

@@
     also observe common covariates $({\bf X}_i)$ between both the data set with
-    $Y$ as well as an outside auxilary sample. This is an important practical
+    $Y$ as well as an outside auxiliary sample. \sout{This is an important practical
     problem \cite{yang2020statistical}, 
-    \cite{dagdoug2023model}. We focus on the case:
+    \cite{dagdoug2023model}.} We focus on the case:
 
@@
   %\begin{theorem}[Design Consistency]\label{thm:dc1}
-    Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$,
+   \sout{Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$,
     where $\hat{\bm \lambda}$ are the corresponding Lagrange multipliers from
-    Equation~\ref{eq:wmin}.
-    Under some regularity conditions, $\bm \lambda^* = ({\bf 0}^T, 1)^T$ and
+    Equation~\ref{eq:wmin}.}
+    Under some regularity conditions, \sout{$\bm \lambda^* = ({\bf 0}^T, 1)^T$} and
 
-    $$\hat Y_{\rm{DCE}} = \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + O_p(N / n_2)$$
+    $$\hat Y_{\rm{DCE}} = \hat Y_\ell + O_p(N / n_2), $$
 
@@
 
-    $$\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) = \hat Y_{\rm{DEE}}+ 
+    $$\hat Y_{\ell} = \hat Y_{\rm{DEE}}+ 
     \left(\sum_{i \in A_1} d_{1i} {\bf z}_i - \sum_{i \in A_2} d_{1i} 
@@
 
+%\textcolor{red}{The linearization expression does not involve $\blambda^*$. }
 \end{column}

### 2024-11-04 src/explore/20241025-mcbssim.qmd
Added real data simulation. Computation still needs work. It takes too long to transfer data.

---
title: "Real Data Analysis: MCBS Data"
author: "Caleb Leedy"
date: "October 25, 2024"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    code-fold: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

# Summary

This report

1. Understands the MCBS data,
2. Understands the ACS data,
3. Applies our data integration method to these surveys.

# Data

```{r}
#| label: setup

library(dplyr)
library(readr)
library(haven)
library(stringr)
library(purrr)
library(nleqslv)

```

```{r}
#| label: load data

# # mcbs_df <- read_delim("data/MCBS/MCBSPUF_2021_1_fall.txt") # Did not work
# mcbs_fall_df <- read_csv("data/MCBS/puf2021_1_fall.csv")
# mcbs_winter_df <- read_csv("data/MCBS/puf2021_2_winter.csv")
# mcbs_summer_df <- read_csv("data/MCBS/puf2021_3_summer.csv")

```

The summer supplement contains information about whether an individual took a
COVID vaccine. The variable is `CVS_ONEDOSE`. Demographic information is found
in the fall supplement. Individuals between data sets can be matched using the
`PUF_ID`. The corresponding `MCBSPUF_2021_<NUM>_<SEASON>.txt` file contains a
description of data responses.

I also downloaded CPS data to get information about people below the poverty
line: Table A-5 from 
[here](https://www.census.gov/data/tables/2022/demo/income-poverty/p60-277.html).
This provides a mean for the percentage of older Americans are below the poverty
line in 2021. (Note: The ACS also has the variable `POVPIP` which is the
income-to-poverty percentage.)

Futhermore, we have data from the American Community Survey to get additional
demographic information.

```{r}
#| label: acs data

# acs_a <- read_csv("data/ACS/psam_pusa.csv")
# acs_b <- read_csv("data/ACS/psam_pusb.csv")
# acs_df <- bind_rows(acs_a, mutate(acs_b, ST = as.character(ST)))

```

We can see the variable codes in the `data/ACS/PUMS_Data_Dictionary_2021.pdf`.
We have the following common variables between the MCBS and ACS:

| Variable | MCBS Code | ACS Code |
|----------|-----------|----------|
| Sex      | DEM_SEX   | SEX      |
| Education | DEM_EDU  | SCHL     |
| Marital Status | DEM_MARSTA | MAR (MSP?) |
| Race / Ethnicity | DEM_RACE | RAC1P |
| On Medicare | ADM_H_MEDSTA  | HINS3 | 
| On Medicaid | ADM_DUAL_FLAG_YR | FHINS4C |

Since ACS contains information about age, we can reweight the data to only
include people 65 years or older. We also have data from the 
~~CPS containing information about sex, education, and marital status.~~
CMS containing information about the total number of people enrolled in Medicare
in 2021.

## Cleaning the Data

We need to ensure that the data from MCBS and ACS are encoded in the same
way.

```{r}
#| label: understanding acs data

# acs_df %>%
#   group_by(SERIALNO) %>%
#   summarize(Count = n()) %>%
#   arrange(desc(Count))
# 
# # What is the unique identifier for the ACS data?
# # RT - record type
# # SERIALNO - HU or GQ serial number
# # DIVISION - division based on Census. Ex. New England, Middle Atlantic
# # SPORDER - person number (1 - 20)
# # PUMA - public use microdata area code
# #
# # I believe SERIALNO + SPORDER will be unique
# 
# acs_df %>%
#   mutate(tmp_id = str_c(SERIALNO, SPORDER)) %>%
#   group_by(tmp_id) %>%
#   summarize(Count = n()) %>%
#   arrange(desc(Count))
# 
# # It is unique!

```

```{r}
#| label: cleaning sex category

# unique(acs_df$SEX)            # Male: 1, Female: 2
# unique(mcbs_fall_df$DEM_SEX)  # Male: 1, Female: 2

```

Cleaning the school category

MCBS: `DEM_EDU` contains the following categories:

| Category | Response |
|----------|----------|
| Less than high school | 1 |
| High school or vocational | 2 |
| More than high school | 3 |
| Refused | R |
| Don't know | D |

The ACS needs to be transformed into this scale:

| Category | Response |
|----------|----------|
| Less than high school | bb and 01 - 15 | 
| High school or vocational | 16 - 19 | 
| More than high school | 20 - 24 |

Cleaning Marital Status

MCBS: `DEM_MARSTA` contains the following categories:

| Category | Response |
|----------|----------|
| Married | 1 |
| Widowed | 2 |
| Divorced / Separated | 3 |
| Never married | 4 |
| Refused | R |

The ACS needs to be transformed into this scale:

| Category | ACS Response |
|----------|----------|
| Married | 1 |
| Widowed | 2 |
| Divorced / Separated | 3, 4 |
| Never married | 5 |

The actual cleaning of the data is found in `../build/clean_acs_data.R` with the
associated saved data set.

# Getting Uncertainty Estimates

According to the PUMS Handbook (see `references/` folder) the way to get
uncertainty estimates in the ACS is to compute the estimate with the weight
`PWGTP`. Call the $X$. Then compute the estimate with the weights `PWGTP1`
through `PWGTP80`. Call these $X_1, \dots, X_{80}$. Then an estimate for the
variance is

$$\hat V = \frac{4}{80} \sum_{i = 1}^{80} (X_i - X)^2.$$

We need to compute the uncertainty for the following variables: `DEM_EDU`,
`DEM_MARSTA`, and `DEM_RACE`.

```{r}
#| label: acs uncertainty
acs_cl <- read_csv("../build/cleaned_acs_data.csv")
acs_med <- filter(acs_cl, HINS3 == 1)

# Transform categorical to logical vectors
acs_med <- acs_med %>%
  mutate(edulths = (DEM_EDU == "1"),
         edueqhs = (DEM_EDU == "2"),
         edumths = (DEM_EDU == "3")) %>%
  mutate(marmar = (DEM_MARSTA == "1"),
         marwid = (DEM_MARSTA == "2"),
         mardiv = (DEM_MARSTA == "3"),
         marsin = (DEM_MARSTA == "4")) %>%
  mutate(sexm = (DEM_SEX == "1"),
         sexf = (DEM_SEX == "2")) %>%
  mutate(racew = (DEM_RACE == "1"),
         raceb = (DEM_RACE == "2"),
         raceh = (DEM_RACE == "3"))

get_est_uncertainty <- function(acs_med, varname, def_w = "PWGTP") {

  # Estimate Mean
  # est <- sum(acs_med[[varname]] * acs_med[[def_w]]) / sum(acs_med[[def_w]])
  est <- sum(acs_med[[varname]] * acs_med[[def_w]])

  # Estimate Variance
  alt_est <- rep(NA, 80)
  for (i in 1:80) {
    w_str <- str_c("PWGTP", i)
    # alt_est[i] <- sum(acs_med[[varname]] * acs_med[[w_str]]) / sum(acs_med[[w_str]])
    alt_est[i] <- sum(acs_med[[varname]] * acs_med[[w_str]])
  }

  v_est <- 4 / 80 * sum((est - alt_est)^2)

  return(list(est = est, var = v_est))

}

param_df <- 
  tibble(varname = c("HINS3", "edulths", "edueqhs", "edumths",
                     "marmar", "marwid", "mardiv", "marsin",
                     "sexm", "sexf",
                     "racew", "raceb", "raceh"),
         type = c("TOT", "EDU", "EDU", "EDU", "MARSTA", "MARSTA", "MARSTA", "MARSTA",
                  "SEX", "SEX", "RACE", "RACE", "RACE")
  )

```

## Clean MCBS Data

```{r}
#| label: clean MCBS

mcbs_fall_df <- read_csv("data/MCBS/puf2021_1_fall.csv")
mcbs_summer_df <- read_csv("data/MCBS/puf2021_3_summer.csv")

# The fall segment contains demographic information while the summer segment
# contains the covid questions

mcbs_all <- 
  left_join(select(mcbs_summer_df, PUF_ID, starts_with("CVS_"), starts_with("PUFS")),
            select(mcbs_fall_df, PUF_ID, starts_with("DEM_")),
            by = join_by(PUF_ID))

mcbs_all <- mcbs_all %>%
  mutate(edulths = (DEM_EDU == "1"),
         edueqhs = (DEM_EDU == "2"),
         edumths = (DEM_EDU == "3")) %>%
  mutate(marmar = (DEM_MARSTA == "1"),
         marwid = (DEM_MARSTA == "2"),
         mardiv = (DEM_MARSTA == "3"),
         marsin = (DEM_MARSTA == "4")) %>%
  mutate(sexm = (DEM_SEX == "1"),
         sexf = (DEM_SEX == "2")) %>%
  mutate(racew = (DEM_RACE == "1"),
         raceb = (DEM_RACE == "2"),
         raceh = (DEM_RACE == "3"))

```

# Data Integration

The CMS data is considered a population level data. There is no uncertainty in
this result.

```{r}
#| label: add CMS data
cov_ests <- 
  bind_rows(
  purrr::map_dfr(param_df$varname, function(vn) {
    out <- get_est_uncertainty(acs_med, vn)
    return(tibble(Data = "ACS", Name = vn, Est = out$est, Var = out$var))
  }),
  purrr::map_dfr(param_df$varname[-1], function(vn) {
    out <- get_est_uncertainty(mcbs_all, vn, def_w = "PUFSWGT")
    return(tibble(Data = "MCBS", Name = vn, Est = out$est, Var = out$var))
  }))

mcbs_tot_est <- 
  get_est_uncertainty(mutate(mcbs_all, ones = 1), "ones", def_w = "PUFSWGT")

cov_ests <- cov_ests %>% 
  add_row(Data = "CMS", Name = "tot", Est = 63892626, Var = 1) %>%
  add_row(Data = "CMS", Name = "sexm", Est = 29159084, Var = 1) %>%
  add_row(Data = "CMS", Name = "sexf", Est = 34733542, Var = 1) %>%
  mutate(Name = ifelse(Name %in% c("HINS3", "ones"), "tot", Name)) %>%
  add_row(Data = "MCBS", Name = "tot", Est = mcbs_tot_est$est, Var = mcbs_tot_est$var)

```

```{r}
#| label: data integration

#' This  function does the debiased calibration
#'
#' @param zmat - A n x p matrix
#' @param w1i - A vector of length n
#' @param d2i - A vector of length n
#' @param T1 - A vector of length p
#' @param qi - A scaler or a vector of length n
#' @param entropy - One of "EL", "ET"
solveGE <- function(zmat, w1i, d2i, T1, qi, entropy) {

  f <- function(lam, zmat, w1i, d2i, T1, qi, entropy, returnw = FALSE) {

    if (entropy == "EL") {
      w <- -d2i / drop(zmat %*% lam)
    } else if (entropy == "ET") {
      w <- d2i * exp(drop(zmat %*% lam))
    } else {
      stop("We only accept entropy of EL or ET.")
    }

    if (returnw) {
      return(w)
    } else {
      return(drop((w * w1i * qi) %*% zmat) - T1)
    }
  }

  j <- function(lam, zmat, w1i, d2i, T1, qi, entropy) {
    if (entropy == "EL") {
      res <- t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
    } else if (entropy == "ET") {
      res <- t(zmat) %*% diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
    }

    return(res)
  }

  init <- c(rep(0, ncol(zmat) - 1), 1)
  res <- 
  nleqslv(init, f, jac = j, zmat = zmat, w1i = w1i, d2i = d2i,
          T1 = T1, qi = qi, entropy = entropy,
          method = "Newton", control = list(maxit = 1e5, allowSingular = TRUE))

  resw <- f(res$x, zmat, w1i, d2i, T1, qi, entropy, returnw = TRUE)

  if (!(res$termcd %in% c(1))) {
    return(NA)
  }

  return(resw)

}

data_int_mcbsacs <- function(mcbs_all, cov_ests) {

  qi = 1
  entropy = "EL"

  # Default: p0_df -> mcbs_all
  # d_vec <- 1 / p0_df$pi0
  d_vec <- mcbs_all$PUFSWGT
  if (entropy == "EL") {
    gdi <- -1 / d_vec
    # FIXME: We need an estimate of gdip
    gdip <- -1 / (1 / pop_df$pi0)
    # END FIXME:
    gpinv <- d_vec^2
  } else if (entropy == "ET") {
    gdi <- log(d_vec)
    gdip <- log(1 / pop_df$pi0)
    gpinv <- d_vec
  } else {
    stop("This type of entropy has not yet been implemented.")
  }

  samp_X <- model.matrix(~edulths + edueqhs + 
                         marmar + marwid + mardiv + 
                         sexm + racew + raceb + raceh, data = mcbs_all)
  zmat <- cbind(samp_X, gdi)
  w1i <- 1

  c_ests <- filter(cov_ests, !(Name %in% c("edumths", "sexf", "marsin")))
  xhat <- c_ests$Est
  # Columns: tot, edulths, edueqhs, marmar, marwid, mardiv, sexm,
  #          racew, raceb, raceh
  # Rows: same dimension as length(xhat)
  x_mat <- matrix(c(as.numeric(c_ests$Name == "tot"),
                    as.numeric(c_ests$Name == "edulths"),
                    as.numeric(c_ests$Name == "edueqhs"),
                    as.numeric(c_ests$Name == "marmar"),
                    as.numeric(c_ests$Name == "marwid"),
                    as.numeric(c_ests$Name == "mardiv"),
                    as.numeric(c_ests$Name == "sexm"),
                    as.numeric(c_ests$Name == "racew"),
                    as.numeric(c_ests$Name == "raceb"),
                    as.numeric(c_ests$Name == "raceh")),
                  ncol = 10)

  vmat <- diag(c_ests$Var)
  vinv <- diag(1 / c_ests$Var)
  xgls <- drop(solve(t(x_mat) %*% vinv %*% x_mat) %*% 
              (t(x_mat) %*% vinv %*% xhat))
  T1 <- c(xgls, sum(gdip * qi))
  dc_w <- solveGE(zmat, w1i, d_vec, T1, qi, entropy)

  # Mean Estimation
  N_pop <- filter(cov_ests, Data == "CMS", Name = "tot") %>% pull(Est)
  Y_vec <- as.numeric(mcbs_all$CVS_ONEDOSE == "1")
  theta <- sum(Y_vec * dc_w * w1i) / N_pop

  z_dc <- sampX
  gam_dc <- 
    solve(t(z_dc) %*% diag(d_vec * gpinv * qi) %*% z_dc,
          t(z_dc) %*% diag(d_vec * gpinv) %*% Y_vec)
  coefs <- as.numeric(gam_dc)
  u_mat <- sampX
  eps <- (Y_vec - drop(u_mat %*% coefs))

  # Variance Estimation
  var_x <- solve(t(x_mat) %*% vinv %*% x_mat)

  v1 <- matrix(coefs, nrow = 1) %*% var_x %*% matrix(coefs, ncol = 1)
  v2 <- sum((1 - (1 / d_vec)) / (1 / d_vec)^2 * eps^2) / (N_pop)^2
  v_hat <- v1 + v2

  return(list(theta = theta, v_hat = as.numeric(v_hat)))
}

```

Next steps:
* Use data integration model.

### 2024-11-04 src/explore/20241101-msdcsim_estalpha.qmd
Added real data simulation. Computation still needs work. It takes too long to transfer data.

--- 2024-07-29 src/explore/20240506-msdcsim.qmd
@@
 ---
-title: "Multi-Source Debiased Calibration Simulation"
+title: "Multi-Source Debiased Calibration Simulation with Estimated Alpha"
 author: "Caleb Leedy"
@@
 This report simulates multi-source sampling with the debiased
-calibration of @kwon2024debiased. We check for the following:
-
-1. Bias of $\hat Y$ and,
-2. Bias of $\hat V(\hat Y)$.
+calibration of @kwon2024debiased with an estimated $\alpha$ value. 
+With this modification we do not need the design weights for the population.
 
@@
     if (entropy == "EL") {
-      w <- -d2i / drop(zmat %*% lam)
+      w <- (-d2i) / drop(zmat %*% lam)
     } else if (entropy == "ET") {
@@
     } else {
-      return(drop((w * w1i * qi) %*% zmat) - T1)
+      return(T1 - drop((w * w1i * qi) %*% zmat))
     }
@@
   j <- function(lam, zmat, w1i, d2i, T1, qi, entropy) {
+    end <- ncol(zmat)
     if (entropy == "EL") {
-      res <- t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
+      res <- (-1) * t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
     } else if (entropy == "ET") {
-      res <- t(zmat) %*% diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
+      res <- (-1) * t(zmat) %*% 
+        diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
     }
@@
 
+  # zmat still contains g(di) as the last column
   init <- c(rep(0, ncol(zmat) - 1), 1)
@@
 
+# The solveGE_alpha function estimates alpha and the weights.
+# This is useful for when we do not have design weights for the entire
+# population.
+solveGE_alpha <- function(alpha, zmat, w1i, d2i, T1, qi, entropy, retw = FALSE) {
+
+  N <- T1[1]
+  T1_adj <- T1
+  T1_adj[length(T1_adj)] <- alpha * N
+  dc_w <- solveGE(zmat, w1i, d2i, T1_adj, qi, entropy)
+
+  if (retw) {
+    return(dc_w)
+  }
+
+  Ka_type <- "alpha"
+  if (Ka_type == "alpha") {
+    # HACK: G(x) = -log(x) for EL only
+    G <- function(x) {-log(x)}
+    return(sum(G(dc_w)) - N * alpha)
+  }
+}
+
 dc_ms_ybar <- 
-  function(p0_df, p1_df, p2_df, pop_df, estT1, qi = 1, entropy = "EL", reg = FALSE) {
+  function(p0_df, p1_df, p2_df, pop_df, qi = 1, entropy = "EL", reg = FALSE) {
 
+  # Since we have a SRS, it is fine to use this info.
+  N_obs <- nrow(pop_df)
   d_vec <- 1 / (p0_df$pi0)
@@
     gdi <- -1 / d_vec
-    gdip <- -1 / (1 / pop_df$pi0)
     gpinv <- d_vec^2
@@
 
-  samp_X <- cbind(rep(1, nrow(p0_df)), as.matrix(select(p0_df, X1, X2, X3)))
+  samp_X <- model.matrix(~X1 + X2 + X3, data = p0_df)
   zmat <- cbind(samp_X, gdi)
   w1i <- 1
-  d0i <- 1 / p0_df$pi0
 
-  if (estT1) {
   x1_0 <- sum(p0_df$X1 / p0_df$pi0) / N_obs
@@
 
-    T1 <- c(nrow(pop_df),
-            xgls[1] * nrow(pop_df),
-            xgls[2] * nrow(pop_df),
-            xgls[3] * nrow(pop_df),
-            sum(gdip * qi))
-  } else {
-    T1 <- c(nrow(pop_df),
-            sum(pop_df$X1),
-            sum(pop_df$X2),
-            sum(pop_df$X3),
-            sum(gdip * qi))
+  T1 <- c(N_obs,
+          xgls[1] * N_obs,
+          xgls[2] * N_obs,
+          xgls[3] * N_obs,
+          NA)
+
+  # Estimating Alpha
+  alpha_init <- (-1) * nrow(p0_df) / N_obs
+  a_res <- 
+    nlm(f = solveGE_alpha, p = alpha_init,
+        zmat = zmat, w1i = w1i, d2i = d_vec, T1 = T1, qi = qi, entropy = entropy)
+
+  if (a_res$code != 1) {
+    warning(paste0("a_res has code", a_res$code))
   }
 
-  dc_w <- solveGE(zmat, w1i, d0i, T1, qi, entropy)
+  dc_w <- 
+    solveGE_alpha(a_res$estimate, zmat, w1i, d_vec, T1, qi, entropy, retw = TRUE)
 
   # Mean Estimation
-  theta <- sum(p0_df$Y * dc_w * w1i) / nrow(pop_df)
-  z_dc <- model.matrix(~1 + p0_df$X1 + p0_df$X2 + p0_df$X3 + gdi)
+  # Do we want N_obs? Or an estimated N_hat? 
+  # N_obs is probably fine since we have SRS.
+  theta <- sum(p0_df$Y * dc_w * w1i) / N_obs
+  z_dc <- samp_X
   gam_dc <- 
-    solve(t(z_dc) %*% diag(d0i * gpinv * qi) %*% z_dc,
-          t(z_dc) %*% diag(d0i * gpinv) %*% p0_df$Y)
+    solve(t(z_dc) %*% diag(d_vec * gpinv * qi) %*% z_dc,
+          t(z_dc) %*% diag(d_vec * gpinv) %*% p0_df$Y)
   coefs <- as.numeric(gam_dc)
   u_mat <- 
-  matrix(c(rep(1, nrow(p0_df)), p0_df$X1, p0_df$X2, p0_df$X3, gdi * qi), ncol = 5)
+  matrix(c(rep(1, nrow(p0_df)), p0_df$X1, p0_df$X2, p0_df$X3), ncol = 4)
   eps <- (p0_df$Y - drop(u_mat %*% coefs))
-  if (estT1) {
   # Variance Estimation
@@
   v_hat <- v1 + v2 
-  } else {
-    v2 <- sum((1 - p0_df$pi0) / p0_df$pi0^2 * eps^2) / nrow(pop_df)^2
-    v_hat <- v2
-  }
   if (reg) {
@@
     reg <- reg_est(p0_df, p1_df, N_obs)
-    ms_pop <- dc_ms_ybar(p0_df, p1_df, p2_df, pop_df, estT1 = FALSE)
-    ms_est <- dc_ms_ybar(p0_df, p1_df, p2_df, pop_df, estT1 = TRUE)
-    ms_reg <- dc_ms_ybar(p0_df, p1_df, p2_df, pop_df, estT1 = TRUE, reg = TRUE)
+    ms_est <- dc_ms_ybar(p0_df, p1_df, p2_df, pop_df)
 
-    return(tibble(Est = c("HT", "NNReg", "MSPop", "MSEst", "MSReg"),
-                  Theta = c(ht[[1]], reg[[1]], ms_pop[[1]], ms_est[[1]], ms_reg[[1]]),
-                  Var = c(ht[[2]], reg[[2]], ms_pop[[2]], ms_est[[2]], ms_reg[[2]]),
+    return(tibble(Est = c("HT", "NNReg", "MSEst"),
+                  Theta = c(ht[[1]], reg[[1]], ms_est[[1]]), 
+                  Var = c(ht[[2]], reg[[2]], ms_est[[2]]), 
                   Iter = iter)
@@
 
-# Check for failures
-filter(mc_res, is.na(Theta))
-
 # Analyze Mean
@@
   summarize(Bias = mean(err, na.rm = TRUE),
+            SE = sd(err, na.rm = TRUE),
             RMSE = sqrt(mean(err^2, na.rm = TRUE)),
             EmpCI = mean(CI, na.rm = TRUE),
-            sdest = sd(Theta, na.rm = TRUE)) %>%
+            MCVar = var(Theta, na.rm = TRUE),
+            EstVar = mean(Var, na.rm = TRUE),
+            sdest = sd(Theta, na.rm = TRUE),
+            RelBias = (EstVar - MCVar) / MCVar) %>%
   ungroup() %>%
@@
     Est == "HT" ~ 1)
-  ) %>%
-  arrange(Ind) %>%
-  select(Est, Bias, RMSE, EmpCI, Ttest)
+  )
 
@@
   knitr::kable("latex", booktabs = TRUE, digits = 4) %>%
-  cat(file = "tables/msdcsim1.tex")
+  cat(file = "tables/msdc_estalphasim1.tex")
 

### 2024-11-04 src/R/debiased_calibration.R

@@ -1,0 +1,120 @@
+# File: debiased_calibration.R
+# Author: Caleb Leedy
+# Created on: November 2, 2024
+# Purpose:
+# This file contains the updated debiased calibration code to be used in the
+# scripts for my simulation study.
+#
+# The functions in this script have been developed and ran in the previous
+# exploration documents:
+# * ../explore/20240411-tpdcsim.qmd
+# * ../explore/20240425-nndcsim.qmd
+# * ../explore/20240506-msdcsim.qmd
+# * ../explore/20240511-msdcsim2.qmd
+# * ../explore/20241025-mcbssim.qmd
+# * ../explore/20241101-msdcsim_estalpha.qmd
+# * ../explore/proto-dctp.R
+
+# *************
+# * Libraries *
+# *************
+
+library(nleqslv)
+
+# *************
+# * Functions *
+# *************
+
+#' This  function does the debiased calibration
+#'
+#' @param zmat - A n x p matrix
+#' @param w1i - A vector of length n
+#' @param d2i - A vector of length n
+#' @param T1 - A vector of length p
+#' @param qi - A scaler or a vector of length n
+#' @param entropy - One of "EL", "ET"
+solveGE <- function(zmat, w1i, d2i, T1, qi, entropy) {
+
+  f <- function(lam, zmat, w1i, d2i, T1, qi, entropy, returnw = FALSE) {
+
+    if (entropy == "EL") {
+      w <- (-d2i) / drop(zmat %*% lam)
+    } else if (entropy == "ET") {
+      w <- d2i * exp(drop(zmat %*% lam))
+    } else {
+      stop("We only accept entropy of EL or ET.")
+    }
+
+    if (returnw) {
+      return(w)
+    } else {
+      return(T1 - drop((w * w1i * qi) %*% zmat))
+    }
+  }
+
+  j <- function(lam, zmat, w1i, d2i, T1, qi, entropy) {
+    end <- ncol(zmat)
+    if (entropy == "EL") {
+      res <- (-1) * t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
+    } else if (entropy == "ET") {
+      res <- (-1) * t(zmat) %*% 
+        diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
+    }
+
+    return(res)
+  }
+
[…]

### 2024-11-04 src/build/clean_acs_data.R

@@ -1,0 +1,104 @@
+# File: clean_acs_data.R
+# Created by: Caleb Leedy
+# Created on: October 28, 2024
+# Purpose:
+# This document cleans the ACS data so that it can be more easily used. The data
+# set is quite large and even simply tidyverse operations cause R to close.
+
+# *************
+# * Libraries *
+# *************
+
+library(dplyr)
+library(readr)
+library(stringr)
+
+# ********
+# * Data *
+# ********
+
+acs_a <- read_csv("data/ACS/psam_pusa.csv")
+acs_b <- read_csv("data/ACS/psam_pusb.csv")
+
+# *********************
+# * Cleaning the Data *
+# *********************
+
+clean_acs_data <- function(df) {
+
+  df <- df %>%
+    mutate(DEM_SEX = as.character(SEX))
+
+  # Clean Education
+  df <- df %>%
+    mutate(DEM_EDU = case_when(
+      SCHL == "bb" ~ "1",
+      SCHL == "01" ~ "1", 
+      SCHL == "02" ~ "1", 
+      SCHL == "03" ~ "1", 
+      SCHL == "04" ~ "1", 
+      SCHL == "05" ~ "1", 
+      SCHL == "06" ~ "1", 
+      SCHL == "07" ~ "1", 
+      SCHL == "08" ~ "1", 
+      SCHL == "09" ~ "1", 
+      SCHL == "10" ~ "1", 
+      SCHL == "11" ~ "1", 
+      SCHL == "12" ~ "1", 
+      SCHL == "13" ~ "1", 
+      SCHL == "14" ~ "1", 
+      SCHL == "15" ~ "1", 
+      SCHL == "16" ~ "2", 
+      SCHL == "17" ~ "2", 
+      SCHL == "18" ~ "2", 
+      SCHL == "19" ~ "2", 
+      SCHL == "20" ~ "3", 
+      SCHL == "21" ~ "3", 
+      SCHL == "22" ~ "3", 
+      SCHL == "23" ~ "3", 
+      SCHL == "24" ~ "3", 
+      TRUE ~ "M" # There are some NAs in the data
+    ))
+
+  # Clean marital statu
+  df <- df %>%
+    mutate(DEM_MARSTA = case_when(
+      MAR == 1 ~ "1",
+      MAR == 2 ~ "2",
+      MAR == 3 ~ "3",
+      MAR == 4 ~ "3",
+      MAR == 5 ~ "4",
+    ))
+
+  # Clean race / ethnicity
[…]

### 2024-11-12 src/explore/20241025-mcbssim.qmd
Finish all of the simulations. Added simulation script.

@@
 library(nleqslv)
+library(doParallel)
 
@@
 #| label: acs uncertainty
-acs_cl <- read_csv("../build/cleaned_acs_data.csv")
+
+acs_cl <- read_csv("data/cleaned_acs_data.csv")
 acs_med <- filter(acs_cl, HINS3 == 1)
@@
 
+mcbs_all <- mcbs_all %>%
+  mutate(CVS_ONEDOSE = case_when(
+    CVS_ONEDOSE == "1" ~ 1,
+    CVS_ONEDOSE == "2" ~ 2,
+    is.na(CVS_ONEDOSE) & (CVS_TWODOSE %in% c("1", "2")) ~ as.numeric(CVS_TWODOSE),
+    CVS_ONEDOSE %in% c("R", "D") & !is.na(CVS_VCNUMS) ~ 1,
+    TRUE ~ 2
+  ))
+
 ```
@@
 #| label: add CMS data
+
 cov_ests <- 
-  bind_rows(
+  # bind_rows(
   purrr::map_dfr(param_df$varname, function(vn) {
@@
     return(tibble(Data = "ACS", Name = vn, Est = out$est, Var = out$var))
-  }),
-  purrr::map_dfr(param_df$varname[-1], function(vn) {
-    out <- get_est_uncertainty(mcbs_all, vn, def_w = "PUFSWGT")
-    return(tibble(Data = "MCBS", Name = vn, Est = out$est, Var = out$var))
-  }))
+  })#,
+  # purrr::map_dfr(param_df$varname[-1], function(vn) {
+  #   out <- get_est_uncertainty(mcbs_all, vn, def_w = "PUFSWGT")
+  #   return(tibble(Data = "MCBS", Name = vn, Est = out$est, Var = out$var))
+  # }))
 
@@
   add_row(Data = "CMS", Name = "sexf", Est = 34733542, Var = 1) %>%
-  mutate(Name = ifelse(Name %in% c("HINS3", "ones"), "tot", Name)) %>%
-  add_row(Data = "MCBS", Name = "tot", Est = mcbs_tot_est$est, Var = mcbs_tot_est$var)
+  mutate(Name = ifelse(Name %in% c("HINS3", "ones"), "tot", Name))
+
+# %>%
+#  add_row(Data = "MCBS", Name = "tot", Est = mcbs_tot_est$est, Var = mcbs_tot_est$var)
 
@@
     if (entropy == "EL") {
-      res <- t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
+      res <- (-1) * t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
     } else if (entropy == "ET") {
-      res <- t(zmat) %*% diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
+      res <- (-1) * t(zmat) %*% 
+        diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
     }
@@
 
-  if (!(res$termcd %in% c(1))) {
+  if (!(res$termcd %in% c(1, 2))) {
     return(NA)
@@
 
-data_int_mcbsacs <- function(mcbs_all, cov_ests) {
+# The solveGE_alpha function estimates alpha and the weights.
+# This is useful for when we do not have design weights for the entire
+# population.
+solveGE_alpha <- function(alpha, zmat, w1i, d2i, T1, qi, entropy, retw = FALSE) {
+
+  N <- T1[1]
+  T1_adj <- T1
+  T1_adj[length(T1_adj)] <- alpha * N
+  dc_w <- solveGE(zmat, w1i, d2i, T1_adj, qi, entropy)
+
+  if (retw) {
+    return(dc_w)
+  }
+
+  Ka_type <- "alpha"
+  if (Ka_type == "alpha") {
+    # HACK: G(x) = -log(x) for EL only
+    G <- function(x) {-log(x)}
+    return(sum(G(dc_w)) - N * alpha)
+  }
+}
+
+data_int_mcbsacs <- function(mcbs_all, cov_ests, weight_v = "PUFSWGT") {
 
@@
   entropy = "EL"
+  N_pop <- filter(cov_ests, Data == "CMS", Name == "tot") %>% pull(Est)
 
@@
   # d_vec <- 1 / p0_df$pi0
-  d_vec <- mcbs_all$PUFSWGT
+  d_vec <- mcbs_all[[weight_v]]
   if (entropy == "EL") {
     gdi <- -1 / d_vec
-    # FIXME: We need an estimate of gdip
-    gdip <- -1 / (1 / pop_df$pi0)
-    # END FIXME:
     gpinv <- d_vec^2
@@
     gdi <- log(d_vec)
-    gdip <- log(1 / pop_df$pi0)
     gpinv <- d_vec
@@
 
-  samp_X <- model.matrix(~edulths + edueqhs + 
-                         marmar + marwid + mardiv + 
-                         sexm + racew + raceb + raceh, data = mcbs_all)
+  samp_X <- model.matrix(~edulths + edueqhs + marmar + marwid + mardiv + sexm,
+                        data = mcbs_all)
   zmat <- cbind(samp_X, gdi)
@@
 
-  c_ests <- filter(cov_ests, !(Name %in% c("edumths", "sexf", "marsin")))
-  xhat <- c_ests$Est
+  c_ests <- 
+    filter(cov_ests, !(Name %in% c("edumths", "sexf", "marsin"))) %>%
+    filter(!(Name %in% c("racew", "raceb", "raceh")))
+
+  # We want the mean for X not the total
+  xhat <- c_ests$Est / N_pop
   # Columns: tot, edulths, edueqhs, marmar, marwid, mardiv, sexm,
@@
                     as.numeric(c_ests$Name == "mardiv"),
-                    as.numeric(c_ests$Name == "sexm"),
-                    as.numeric(c_ests$Name == "racew"),
-                    as.numeric(c_ests$Name == "raceb"),
-                    as.numeric(c_ests$Name == "raceh")),
-                  ncol = 10)
+                    as.numeric(c_ests$Name == "sexm")),
+                  ncol = length(unique(c_ests$Name)))
 
-  vmat <- diag(c_ests$Var)
-  vinv <- diag(1 / c_ests$Var)
+  # We want the variance of the mean not the variance of the total
+  vmat <- diag(c_ests$Var / N_pop^2)
+  vinv <- diag(N_pop^2 / c_ests$Var)
   xgls <- drop(solve(t(x_mat) %*% vinv %*% x_mat) %*% 
               (t(x_mat) %*% vinv %*% xhat))
-  T1 <- c(xgls, sum(gdip * qi))
-  dc_w <- solveGE(zmat, w1i, d_vec, T1, qi, entropy)
+
+  # We want T1 to be the totals
+  T1 <- c(xgls * N_pop, NA)
+
+  # Estimating Alpha
+  alpha_init <- (-1) * nrow(mcbs_all) / N_pop
+  a_res <- 
+    optim(par = alpha_init, fn = solveGE_alpha, 
+        zmat = zmat, w1i = w1i, d2i = d_vec, T1 = T1, qi = qi, entropy = entropy)
+
+  if (a_res$convergence != 0) {
+    warning(paste0("a_res has convergence", a_res$convergence))
+  }
+
+  dc_w <- 
+    solveGE_alpha(a_res$par, zmat, w1i, d_vec, T1, qi, entropy, retw = TRUE)
 
   # Mean Estimation
-  N_pop <- filter(cov_ests, Data == "CMS", Name = "tot") %>% pull(Est)
-  Y_vec <- as.numeric(mcbs_all$CVS_ONEDOSE == "1")
+  Y_vec <- as.numeric(mcbs_all$CVS_ONEDOSE == 1)
   theta <- sum(Y_vec * dc_w * w1i) / N_pop
 
-  z_dc <- sampX
+  # FIXME: Do we want samp_X or zmat?
   gam_dc <- 
-    solve(t(z_dc) %*% diag(d_vec * gpinv * qi) %*% z_dc,
-          t(z_dc) %*% diag(d_vec * gpinv) %*% Y_vec)
+    solve(t(samp_X) %*% diag(d_vec * gpinv * qi) %*% samp_X,
+          t(samp_X) %*% diag(d_vec * gpinv) %*% Y_vec)
   coefs <- as.numeric(gam_dc)
-  u_mat <- sampX
-  eps <- (Y_vec - drop(u_mat %*% coefs))
+  eps <- (Y_vec - drop(samp_X %*% coefs))
 
@@
 
+reg_est <- function(mcbs_all, cov_ests, weight_v = "PUFSWGT") {
+
+  N_pop <- filter(cov_ests, Data == "CMS", Name == "tot") %>% pull(Est)
+  Y_vec <- as.numeric(mcbs_all$CVS_ONEDOSE == 1)
+  mod_df <- mcbs_all %>%
+    mutate(Y = (CVS_ONEDOSE == 1))
+
+  mod <- lm(Y ~ sexm, data = mod_df, weights = mod_df[[weight_v]])
+  yhat <- predict(mod, mod_df)
+
+  # TEST:
+  # mod_t <- lm(Y ~ edulths + edueqhs + marmar + marwid + mardiv + sexm,
+  #   data = mod_df, weights = mod_df[[weight_v]])
+  # yhat <- predict(mod_t, mod_df) 
+  # END TEST:
+
+  # Regression Estimation
+  cms_tot <- filter(cov_ests, Data == "CMS", Name == "tot") %>% pull(Est)
+  cms_sexm <- filter(cov_ests, Data == "CMS", Name == "sexm") %>% pull(Est)
+  cms_mean <- cms_tot / N_pop
+  cms_meansexm <- cms_sexm / N_pop
+
+  err_term <- sum((mod_df$Y - yhat) * mod_df[[weight_v]]) / N_pop
+
+  theta <- 
+    cms_mean * mod$coefficients[1] + cms_meansexm * mod$coefficients[2] + err_term
+
+  # Variance Estimation
+  pihat <- 1 / mod_df[[weight_v]]
+  v_hat <- sum((1 - pihat) * (mod_df$Y - yhat)^2 * (1 / pihat)^2) / N_pop^2
+
+  return(list(theta = theta, v_hat = v_hat))
+}
+
+ht_est <- function(mcbs_all, cov_ests, weight_v = "PUFSWGT") {
+
+  N_pop <- filter(cov_ests, Data == "CMS", Name == "tot") %>% pull(Est)
+  Y_vec <- as.numeric(mcbs_all$CVS_ONEDOSE == 1)
+  theta <- sum(Y_vec * mcbs_all[[weight_v]]) / N_pop
+  pihat <- 1 / mod_df[[weight_v]]
+  v_hat <- sum((1 - pihat) * (Y_vec  / pihat)^2) / N_pop^2
+
+  return(list(theta = theta, v_hat = v_hat))
+}
+
+```
+
+# Assessing the Variance
+
+```{r}
+#| label: variance analysis of mcbs
+
+res <- data_int_mcbsacs(mcbs_all, cov_ests, weight_v = "PUFSWGT") 
+ms <- reg_est(mcbs_all, cov_ests)
+ht <- ht_est(mcbs_all, cov_ests)
+
+cl <- makeCluster(min(detectCores() - 1, 100), outfile = "")
+registerDoParallel(cl)
+
+packs <- c("dplyr", "stringr", "purrr", "nleqslv")
+
+mc_res <- 
+  foreach(b = 1:100, .packages = packs,
+          .errorhandling = "remove", .inorder = FALSE) %dopar% {
+
+  if ((b %% 10) == 0) {
+    print(paste0("Iteration: ", b))
+  }
+
+  weight_v <- paste0("PUFS", str_sub(paste0("00", b), -3))
+  ms <- data_int_mcbsacs(mcbs_all, cov_ests, weight_v = weight_v) 
+  # ms <- reg_est(mcbs_all, cov_ests, weight_v = weight_v)
+
+  return(
+      tibble(Est = c("MSEst"),
+             Theta = c(ms$theta),
+             Var = c(ms$v_hat), 
+             Iter = b)
+  )
+
+  } %>% bind_rows()
+
+stopCluster(cl)
+
+# Weighted variance estimate
+var(mc_res$Theta)
+
+# Estimated variance estimate
+res$v_hat
+
+# Relative Bias
+(res$v_hat - var(mc_res$Theta)) / var(mc_res$Theta)
+
 ```
@@
 * Use data integration model.
+* Check variance assessment using other weight models (PUFS001, etc.).

### 2024-11-12 src/explore/20241101-msdcsim_estalpha.qmd
Finish all of the simulations. Added simulation script.

@@
   j <- function(lam, zmat, w1i, d2i, T1, qi, entropy) {
-    end <- ncol(zmat)
     if (entropy == "EL") {
@@
   a_res <- 
-    nlm(f = solveGE_alpha, p = alpha_init,
+    optim(par = alpha_init, fn = solveGE_alpha, 
         zmat = zmat, w1i = w1i, d2i = d_vec, T1 = T1, qi = qi, entropy = entropy)
 
-  if (a_res$code != 1) {
-    warning(paste0("a_res has code", a_res$code))
+  if (a_res$convergence != 0) {
+    warning(paste0("a_res has convergence", a_res$convergence))
   }
@@
   dc_w <- 
-    solveGE_alpha(a_res$estimate, zmat, w1i, d_vec, T1, qi, entropy, retw = TRUE)
+    solveGE_alpha(a_res$par, zmat, w1i, d_vec, T1, qi, entropy, retw = TRUE)
 
@@
 mc_res <- 
-  foreach(iter=1:B_sims,
-          .packages = c("dplyr", "nleqslv"),
-          .options.RNG = 1) %dorng% {
+  foreach(iter=1:B_sims, .packages = c("dplyr", "nleqslv"), .options.RNG = 1) %dopar% {
 
-    if (iter %% 100 == 0) {
+    if (iter %% 50 == 0) {
       print(paste0("Iter: ", iter))
+      # set.seed(iter)
     }
@@
     reg <- reg_est(p0_df, p1_df, N_obs)
-    ms_est <- dc_ms_ybar(p0_df, p1_df, p2_df, pop_df)
+    ms_est <- suppressWarnings(dc_ms_ybar(p0_df, p1_df, p2_df, pop_df))
 
@@
   mutate(Ttest = abs(Bias) / sqrt(sdest^2 / B_sims)) %>%
-  select(-sdest) %>%
-  mutate(Ind = case_when(
-    Est == "MSReg" ~ 5,
-    Est == "MSEst" ~ 4,
-    Est == "MSPop" ~ 3,
-    Est == "NNReg" ~ 2,
-    Est == "HT" ~ 1)
-  )
+  select(-sdest)
 

### 2024-11-12 src/R/alternative_algorithms.R

@@ -1,0 +1,380 @@
+# File: alternative_algorithms.R
+# Created by: Caleb Leedy
+# Created on: November 4, 2024
+# Purpose:
+# This file contains all of the alternative algorithms that are used in the
+# paper simulations procedure found in ../final_sims.R.
+
+# *********************************
+# * Two-Phase Sampling Algorithms *
+# *********************************
+# Exploration file: ../explore/proto-dctp.R
+
+#' This function implements the $\pi^*$-estimator for two-phase sampling.
+tp_pistar <- function(vars) {
+
+  p1_df <- vars[[1]]
+  p2_df <- vars[[2]]
+  pop_df <- vars[[3]]
+  sampling = "srs-poisson"
+
+  pop_obs <- nrow(pop_df)
+  theta <- sum(p2_df$Y / (p2_df$pi1 * p2_df$pi2)) / pop_obs
+
+  if (sampling == "srs-srs") {
+    
+    vhat <- (1 / nrow(p2_df) - 1 / pop_obs) * var(p2_df$Y)
+
+  } else if (sampling == "poisson-poisson") {
+
+    pis <- p2_df$pi1 * p2_df$pi2
+    vhat <- sum((1 / pis^2 - 1 / pis) * p2_df$Y^2) / pop_obs^2
+
+  } else if (sampling == "srs-poisson") {
+
+    # var_y <- var(p2_df$Y)
+    # v1 <- (1 / nrow(p1_df) - 1 / pop_obs) * var_y
+    # v2 <- sum((1 - p2_df$pi2) / (p2_df$pi1^2 * p2_df$pi2^2) * p2_df$Y^2) / pop_obs^2
+    # vhat <- v1 + v2
+
+    tmp <- matrix(NA, nrow = nrow(p2_df), ncol = nrow(p2_df))
+    for (i in 1:nrow(p2_df)) {
+      for (j in 1:nrow(p2_df)) {
+        if (i == j) {
+          pi2i <- p2_df$pi2[i] * p2_df$pi1[i]
+          tmp[i, j] <- ((pi2i - (pi2i)^2) / pi2i) * p2_df$Y[i]^2 / (pi2i^2)
+        } else {
[…]

### 2024-11-12 src/R/debiased_calibration.R

@@ -87,29 +87,38 @@ solveGE <- function(zmat, w1i, d2i, T1, qi, entropy) {
 #' This is useful for when we do not have design weights for the entire
 #' population.
 #'
-#' @param alpha
-#' @param zmat
-#' @param w1i
-#' @param d2i
-#' @param T1
-#' @param qi
-#' @param entropy
-#' @param retw
+#' @param alpha A vector of length c.
+#' @param zmat A matrix of dimension n x c.
+#' @param w1i A vector of length n or a scaler value.
+#' @param d2i A vector of length n.
+#' @param T1 A vector of length c.
+#' @param qi A numeric value. Currently, qi must be 1.
+#' @param entropy A string indicating the entropy used. Currently on "EL" is
+#' accepted.
+#' @param retw A boolean value indicating if the function should return the
+#' estimated weights (TRUE) or if it should return the minimum value of the
+#' objective function (FALSE).
 #'
 #' @return If retw is TRUE, then the function will return the estimated weights.
 #' Otherwise, the function will return the value of $\sum_{i \in A} G(w_i) - N
 #' * alpha$. The second case if used for the optimization procedure.
 solveGE_alpha <- function(alpha, zmat, w1i, d2i, T1, qi, entropy, retw = FALSE) {
 
+  # The first element of T1 should be the population size N, or an estimate of
+  # it.
   N <- T1[1]
   T1_adj <- T1
   T1_adj[length(T1_adj)] <- alpha * N
+
+  # Estimate the weights given that T1 is correct.
   dc_w <- solveGE(zmat, w1i, d2i, T1_adj, qi, entropy)
 
   if (retw) {
     return(dc_w)
   }
 
+  # Return the minimum value of sum(G(w)) - K(alpha)N for use in an optimization
+  # routine.
   Ka_type <- "alpha"
   if (Ka_type == "alpha") {
     # HACK: G(x) = -log(x) for EL only

### 2024-11-12 src/R/integration_algorithms.R

@@ -1,0 +1,539 @@
+# File: integration_algorithms.R
+# Created by: Caleb Leedy
+# Created on: November 4, 2024
+# Purpose:
+# This file contains all of the data integration algorithms that are used in the
+# paper simulations procedure found in ../final_sims.R.
+
+# *********************************
+# * Two-Phase Sampling Algorithms *
+# *********************************
+# Exploration file: ../explore/proto-dctp.R
+
+#' This function estimates the mean of Y_N from a two-phase sample using
+#' debiased calibration.
+tp_est <- function(vars) {
+  p1_df <- vars[[1]]
+  p2_df <- vars[[2]]
+  pop_df <- vars[[3]]
+  estT1 <- vars[[4]]
+
+  qi = 1
+  entropy = "EL"
+  sampling = "srs-poisson"
+  linearized = FALSE
+
+  d_vec <- 1 / (p2_df$pi2)
+  if (entropy == "EL") {
+    gdi <- (-1) / d_vec
+    gdi1 <- (-1) / (1 / p1_df$pi2)
+    gdip <- (-1) / (1 / pop_df$pi2)
+    gpinv <- d_vec^2
+    gdi12 <- "-1 / d2i"
+  } else if (entropy == "ET") {
+    gdi <- log(d_vec)
+    gdi1 <- log(1 / p1_df$pi2)
+    gdip <- log(1 / pop_df$pi2)
+    gpinv <- d_vec
+    gdi12 <- "log(d2i)"
+  } else {
+    stop("This type of entropy has not yet been implemented.")
+  }
+
+  samp_X <- cbind(rep(1, nrow(p2_df)), as.matrix(select(p2_df, X1, X2)))
+  zmat <- cbind(samp_X, gdi)
+  w1i <- 1 / p2_df$pi1
+  d2i <- 1 / p2_df$pi2
+
+  if (estT1) {
+    T1 <- c(nrow(pop_df),
+            sum(p1_df$X1 * qi / p1_df$pi1),
+            sum(p1_df$X2 * qi / p1_df$pi1),
+            sum(gdip * qi)) # We do NOT want to estimate gdi
+  } else {
[…]

### 2024-11-12 src/R/simulation_algorithms.R

@@ -1,0 +1,270 @@
+# File: simulation_algorithms.R
+# Created by: Caleb Leedy
+# Created on: November 4, 2024
+# Purpose:
+# This file contains all of the simulation algorithms that are used in the
+# paper simulations procedure found in ../final_sims.R.
+
+# *********************************
+# * Two-Phase Sampling Algorithms *
+# *********************************
+# Exploration file: ../explore/proto-dctp.R
+
+#' This function generates a population data frame with the requisite variables.
+#'
+#' @param obs - This is the number of observations in the population.
+#'
+#' @details - This function does not include any sampling probabilities because
+#' it does not take into account the sampling type or formula.
+tp_pop <- function(obs, vars) {
+  delta <- vars$delta
+
+  x1 <- rnorm(obs, 2, 1)
+  x2 <- runif(obs, 0, 4)
+  x3 <- rnorm(obs, 0, 1)
+  x4 <- runif(obs, 0.1, 0.9)
+  z <- rnorm(obs, 0, 1)
+  eps <- rnorm(obs)
+
+  y <- 3 * x1 + 2 * x2 + delta * 0.5 * z + eps
+  n1 <- 1200
+  p1_pi <- n1 / obs
+  p2_pi <- pmin(pmax(pt(z - 1, 3), 0.02), 0.7)
+
+  return(tibble(X1 = x1, X2 = x2, X3 = x3, X4 = x4, Z = z, Y = y,
+                pi1 = p1_pi, pi2 = p2_pi))
+}
+
+#' This function generates Phase 1 and Phase 2 samples from an updated
+#' population with selection probabilites.
+tp_samps <- function(uppop_df, vars = list("srs", "poisson")) {
+  p1_type <- vars[[1]]
+  p2_type <- vars[[2]]
+
+  if (p1_type == "poisson") {
+    del1 <- rbinom(nrow(uppop_df), 1, uppop_df$pi1)
+  } else if (p1_type == "srs") {
+    ind <- sample(1:nrow(uppop_df), size = uppop_df$pi1[1] * nrow(uppop_df))
+    del1 <- as.numeric(1:nrow(uppop_df) %in% ind)
+  } else {
[…]

### 2024-11-12 src/final_sims.R

@@ -1,0 +1,229 @@
+# File: final_sims.R
+# Created by: Caleb Leedy
+# Created on: November 4, 2024
+# Purpose:
+# This file contains the code to run all of the simulations for the Data
+# Integration paper. It also saves the output of all of the tables that are used
+# in the final paper.
+#
+# The functions in this script have been developed and ran in the previous
+# exploration documents:
+# * ./explore/20240411-tpdcsim.qmd
+# * ./explore/20240425-nndcsim.qmd
+# * ./explore/20240506-msdcsim.qmd
+# * ./explore/20240511-msdcsim2.qmd
+# * ./explore/20241025-mcbssim.qmd
+# * ./explore/20241101-msdcsim_estalpha.qmd
+# * ./explore/proto-dctp.R
+
+# **************
+# * Parameters *
+# **************
+
+c_opts <- commandArgs(trailingOnly = TRUE)
+if (length(c_opts) > 0) {
+  sim_setup <- c_opts[1]
+  pvars <- list(delta = as.numeric(c_opts[2]))
+} else {
+  # Options: "twophasesamp", "nonnestedtp", "multisource", "multisourcealpha"
+  sim_setup <- "multisource"
+  pvars <- list(delta = 1)
+}
+
+
+# Number of Monte Carlo replicates
+B_sims <- 1000
+
+# Finite population size
+N_obs <- 10000
+
+# ********************************
+# * Custom Simulation Parameters *
+# ********************************
+
+if (sim_setup == "twophasesamp") {
+  # delta controls model misspecification.
+  # If delta = 1, the model is misspecified.
+  name_add <- ifelse(pvars$delta == 1, "d1", "d0")
+} else if (sim_setup == "nonnestedtp") {
+  name_add <- ifelse(pvars$delta == 1, "d1", "d0")
+} else if (sim_setup == "multisource") {
+  pvars$n1_obs <- 2000
+  name_add <- ifelse(pvars$delta == 1, "d1", "d0")
+} else if (sim_setup == "multisourcealpha") {
+  pvars$n1_obs <- 2000
+  name_add <- ""
+} else {
+
+}
+
+# *************
+# * Libraries *
+# *************
+
+library(dplyr)
+library(doParallel)
+library(doRNG)
+library(nleqslv)
+
[…]

### 2024-11-12 src/sim_script.sh

@@ -1,0 +1,16 @@
+#!/bin/bash
+
+# File: sim_script.sh
+# Created by: Caleb Leedy
+# Created on: November 12, 2024
+# Purpose: This script runs all of the code for the simulations for the project:
+# Debiased Calibration for Generalized Two-Phase Sampling.
+
+Rscript final_sims.R "twophasesamp" 0
+Rscript final_sims.R "twophasesamp" 1
+Rscript final_sims.R "nonnestedtp" 0
+Rscript final_sims.R "nonnestedtp" 1
+Rscript final_sims.R "multisource" 0
+Rscript final_sims.R "multisource" 1
+Rscript final_sims.R "multisourcealpha" 0
+Rscript final_sims.R "multisourcealpha" 1

### 2024-11-12 final/paper/paper.tex
Finished rough draft of final paper.

--- 2024-05-13 final/reports/prelim_paper.tex
@@
 
-\usepackage{amsmath, amsthm, mathtools,commath}
+\usepackage{amsmath, amsthm, mathtools}
 \usepackage{graphics, color}
@@
 \title{Debiased Calibration for Generalized Two-Phase Sampling}
-\author{Caleb Leedy}
+\author{Caleb Leedy, Jae-Kwang Kim}
 \maketitle
@@
 
+\begin{abstract}
+Generalized entropy calibration \cite{gneiting2007strictly} is a nonparametric
+method to estimate survey weights without using the design weights in the
+specified loss function. While \cite{kwon2024debiased} explore generalized
+entropy calibration to estimate survey weights with known population totals, by
+using a two-phase sampling framework, we extend their method to allow for
+estimated population totals. We then extend this two-phase sampling framework to
+two additional cases: first, when when have two non-nested samples, and
+second, when we want to combine three or more samples. 
+We apply our method to the Medicare Current Beneficiary Survey for which we
+get more efficient estimates by combining ths data set with population level
+controls obtained from the Center of Medicare and Medicaid Services as well as
+survey data from the U.S. Census' American Community Survey.
+\end{abstract}
+
 \section{Introduction}
 
-Combining information from several sources is an important practical problem. (CITEME)
+Combining information from multiple samples is an important practical problem
+(\cite{yang2020statistical}, \cite{yang2023elastic}, \cite{dagdoug2023model}).
 We want to incorporate information from external data sources to reduce the bias
-in our estimates or improve the estimator's efficiency. For many problems, the
+in our estimates and/or improve the estimator's efficiency. For many problems, the
 additional information consists of summary statistics with standard errors. The
@@
 
-To model this scenaro, we formulate the problem as a generalized two-phase
-sample where the first phase sample consists of data from multiple sources. The
-second phase sample contains our existing data. To motivate this setup, we
-consider the following approach: first, we consider the classical two-phase
+To model this scenario, we formulate the problem as a generalized two-phase
+sample. The first phase sample consists of data from multiple sources, and the
+second phase sample contains the existing data. To motivate this setup, we
+take the following approach: first, we consider the classical two-phase
 sampling setup where the second phase sample is a subset of the first phase
 sample; then, we extend this setup to consider non-nested two-phase samples;
-and finally, we consider the more general approach of having multiple sources.
-
-%\begin{itemize}
-%\item To achieve the goal, we first consider  the  classical two-phase sampling
-%setup where the second-phase sample is a subset of the first-phase sample.
-%  After that, we extend the setup to more general cases such as non-nested
-%  two-phase sampling or multiple independent surveys with some common
-%  measurements. 
-%\item The proposed method can be called two-step calibration. In the
-%first-step, the best linear unbiased estimators of the auxiliary variable
-%totals are computed. In the second-step, the final calibration weights are
-%constructed to match (benchmark) with the best estimators computed from Step 1. 
-%\end{itemize}
+and finally, we consider the more general approach of having multiple data sets.
 
@@
 Consider a finite population of size $N$ containing elements $(\bf X_i, Y_i)$ where
-an initial (Phase 1) sample of size $n_1$ is selected and $\bf X_i$ is observed. Then
-from the Phase 1 sample of elements, a (Phase 2) sample of size $n_2 < n_1$ is
-selected and $Y_i$ is observed. This is two-phase sampling (See 
+an initial (Phase 1) sample of size $n_1$ is selected and $\bf X_i$ is observed.
+From the Phase 1 sample of elements, a (Phase 2) sample of size $n_2 < n_1$ is
+selected and $Y_i$ is observed. This is two-phase sampling. (See 
 \cite{fuller2009sampling}, \cite{kim2024statistics} for general references.) The
@@
 denoted as $A_2$,
-%\begin{align*}
-%  \pi_i &= \sum_{A_2: i \in A_2} \Pr(A_2) \\ 
-%        &= \sum_{A_1: A_2 \subseteq A_1} \sum_{A_2: i \in A_2} \Pr(A_2 \mid
-%        A_1) \Pr(A_1) \\
-%        &= \sum_{A_1: i \in A_1} \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1) \Pr(A_1).
-%\end{align*}
-%\textcolor{red}{(I am not sure whether this part is necessary.) }
-if we define $\pi_{2i | 1} = \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1)$ and
+we define $\pi_{2i | 1} = \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1)$ and
 $\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ then,
@@
 incorporate information for $\bf X$ by using the estimate $\hat{\bf X_1}$ from
-the Phase 1 sample. This is how we can leverage the external information 
+the Phase 1 sample. This is one way we can leverage the external information of 
 $\hat{\bf{X_1}}$ to improve the initial $\pi^*$-estimator in the second phase
@@
 
-$$ \hat Y_{reg, tp} 
+$$ 
+\hat Y_{reg, tp} 
 = \sum_{i \in A_1} \frac{1}{\pi_{1i}} \bf x_i \hat{\bm \beta_q} + 
-\sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i \hat{\bm \beta_q})$$
+\sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i \hat{\bm \beta_q})
+$$
 
 where for $q_i = q(\bf x_i)$ and is a function of $\bf x_i$,
+
 $$
@@
   \frac{\bf x_i \bf x_i'}{\pi_{1i} q_i}\right)^{-1} 
-\sum_{i \in A_2} \frac{\bf x_i y_i}{\pi_{1i} q_i}.$$
+\sum_{i \in A_2} \frac{\bf x_i y_i}{\pi_{1i} q_i}.
+$$
 
@@
 
-$$\hat w_{2i} = \argmin_{w} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i
+$$
+\hat w_{2i} = \argmin_{w} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i
 \text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i / \pi_{1i} = \sum_{i \in A_1}
-\bf x_i / \pi_{1i}.$$
+\bf x_i / \pi_{1i}.
+$$
 
@@
 
-$$ \hat Y_{reg, tp} 
+$$ 
+\hat Y_{reg, tp} 
 = \underbrace{\sum_{i \in A_1} \frac{\bf x_i \hat{\bm \beta_q}}{\pi_{1i}}}_{
@@
 \frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i \hat{\bm \beta_q})}_{
-\text{Bias correction}}.$$
+\text{Bias correction}}.
+$$
 
@@
 
-The method of \cite{kwon2024debiased} requires known finite population 
+However, the method of \cite{kwon2024debiased} requires known finite population 
 calibration levels. It does not handle the
@@
 from the Phase 1 sample. In the rest of the section, we extend this method to 
-two-phase sampling so that we have a valid 
-estimator when including estimated Phase 1 weights with appropriate variance
-estimation.
+two-phase sampling so that we construct asymptotically valid confidence
+intervals when using estimate finite population totals derived from the Phase 1
+sample.
 
@@
 We follow the approach of \cite{kwon2024debiased} for the debiased calibration
-method. We consider maximizing the generalized entropy \cite{gneiting2007strictly},
+method. We consider maximizing a generalized entropy function
+\cite{gneiting2007strictly},
 
@@
 \end{equation}
+
 where $G: \mathcal{V} \to \R$ is strictly convex, differentiable function
@@
 ensures that design consistency is achieved. 
-The original method of \cite{kwon2024debiased} only considered having finite
+The original method of \cite{kwon2024debiased} only considers known finite
 population quantities on the right hand side of \eqref{eq:calconst1}.
 
-Writing $w_{1i} = \pi_{1i}^{-1}$, the the goal is to solve
+By writing $w_{1i} = \pi_{1i}^{-1}$, the the goal is to solve
 
@@
 estimator of $Y_N$ is $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i|1} y_i$.
-To solve this problem we can use the method of Lagrange multipliers. We need to
+We can get an estimator using the method of Lagrange multipliers. We need to
 minimize the Lagrangian function
@@
   + \bm \lambda \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
-    \sum_{i \in A_2} w_{1i} w_{2i|1} \bf z_i q_i\right).
-\end{equation}
-
-Differntiating with respect to $w_{2i|1}$ and setting this expression equal to
+    \sum_{i \in A_2} w_{1i} w_{2i|1} \bf z_i q_i\right)
+\end{equation}
+
+where $\bm \lambda$ is a vector of Lagrange multipliers.
+Differentiating with respect to $w_{2i|1}$ and setting this expression equal to
 zero, yields the fact that $\hat w_{2i|1}$ satisfies
@@
 
-% \textcolor{red}{In Theorem 1, you used $\hat{\lambda}$ without defining it.  }
+The first result that we show is that the construction of the weights in
+Equation~\ref{eq:primal} leads to a design consistent estimator $\hat Y$.
+
 \begin{theorem}[Design Consistency]\label{thm:dc1}
@@
 
-  $$\hat Y_{\ell}(\bm \lambda, \bm \phi^*) = \hat Y_{DCE}(\hat{\bm \lambda}) + 
+  $$
+  \hat Y_{\ell}(\bm \lambda, \bm \phi^*) = \hat Y_{DCE}(\hat{\bm \lambda}) + 
   \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(
-  \bm \lambda) z_i q_i\right)\bm \phi^*$$
+  \bm \lambda) z_i q_i\right)\bm \phi^*
+  $$
 
@@
 
-  $$\bm \phi^* = 
+  $$
+  \bm \phi^* = 
   \left[\sum_{i \in U} \frac{\pi_{2i|1}\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
-  \sum_{i \in U} \frac{\pi_{2i|1}\bf z_i y_i}{g'(d_{2i|1})}.$$
+  \sum_{i \in U} \frac{\pi_{2i|1}\bf z_i y_i}{g'(d_{2i|1})}.
+  $$
 
@@
   \end{eqnarray*}
-  and  $$\begin{pmatrix}
+
+  and
+
+  $$
+  \begin{pmatrix}
   \bm \phi_1^* \\
@@
   \bx_i \bx_i^T &   \bx_i g_i   \\
-  g_i  \bx_i^T   & g_i^2     \end{pmatrix} \right]^{-1}
-  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} \begin{pmatrix}
+  g_i  \bx_i^T   & g_i^2     
+  \end{pmatrix} \right]^{-1}
+  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} 
+  \begin{pmatrix}
     \bx_i \\ g_i 
-  \end{pmatrix}y_i $$
+  \end{pmatrix}y_i 
+  $$
+
 with $g_i = g( \pi_{2i |1}^{-1}) q_i$.
@@
 
-  $$\frac{\partial \mathcal{L}}{\partial w_{2i|1}}: g(w_{2i|1}) w_{1i}q_i -
-  w_{1i} \bm \lambda \bf z_i q_i = 0.$$
+  $$
+  \frac{\partial \mathcal{L}}{\partial w_{2i|1}}: g(w_{2i|1}) w_{1i}q_i -
+  w_{1i} \bm \lambda \bf z_i q_i = 0.
+  $$
 
@@
 
-  $$\hat Y_\ell(\hat{\bm \lambda}, \bm \phi)  = \hat Y_{DCE}(\hat{\bm \lambda}) + 
+  $$
+  \hat Y_\ell(\hat{\bm \lambda}, \bm \phi)  = \hat Y_{DCE}(\hat{\bm \lambda}) + 
   \left(\sum_{i \in A_1} w_{1i} \bm z_i q_i - \sum_{i \in A_2} w_{1i} 
-  \hat w_{2i|1}(\hat{\bm \lambda}) \bf z_i q_i\right)\bm \phi.$$
+  \hat w_{2i|1}(\hat{\bm \lambda}) \bf z_i q_i\right)\bm \phi.
+  $$
 
@@
 
-  $$E\left[\frac{\partial}{\partial \bm \lambda} \hat Y_\ell(\bm \lambda^*,
-  \bm \phi^*)\right]=0.$$
+  $$
+  E\left[\frac{\partial}{\partial \bm \lambda} \hat Y_\ell(\bm \lambda^*,
+  \bm \phi^*)\right]=0.
+  $$
 
@@
     \bm \phi^*
-    &= E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
+    &= E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}
+    \right]^{-1}
     E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i y_i}{g'(d_{2i|1})}\right]\\
-    &= \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
+    &= \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}
+    \right]^{-1}
     \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i y_i}{g'(d_{2i|1})}\right]\\
@@
   The final equality comes from the fact that 
-  $E\left[\frac{\partial}{\partial \bm \lambda}\hat Y_\ell(\bm \lambda^*, \bm \phi^*)\right]
-  = 0$, $\frac{\partial}{\partial \bm \lambda^2} \hat w_{2i|1}(\bm \lambda^*)$ is
-  bounded and $||\hat{\bm \lambda} - \bm \lambda^*|| = O_p(n_2^{-1/2})$, which proves our
-  result.
+  $E\left[\frac{\partial}{\partial \bm \lambda}
+  \hat Y_\ell(\bm \lambda^*, \bm \phi^*)\right] = 0$,
+  $\frac{\partial}{\partial \bm \lambda^2} \hat w_{2i|1}(\bm \lambda^*)$ is
+  bounded and $||\hat{\bm \lambda} - \bm \lambda^*|| = O_p(n_2^{-1/2})$, which
+  proves our result.
 \end{proof}
@@
 \varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
-Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
-%\pi_{1i} &= \min(\Phi_3(-x_{3i} - 2), 0.7) \\
+Y_{i} &= 3 X_{1i} + 2 X_{2i} + \delta 0.5 X_{3i} + \varepsilon_i \\
 \pi_{1i} &= n_1 / N \\
-\pi_{2i|1} &= \min(\Phi_3(-x_{1i} + 1), 0.9).
+\pi_{2i|1} &= \max(\min(\Phi_3(x_{3i} - 1), 0.7), 0.02).
 \end{aligned}
@@
 where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
-%\textcolor{red}{The simulation setup is somewhat strange. The first order
-%inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a
-%function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$.
-%In this case, your second-phase sampling is essentially equivalent to the simple
-%random sampling. }
 This is a two-phase extension of the setup in \cite{kwon2024debiased}. We
@@
 being a simple random sample (SRS) and the 
-Phase 2 sampling occuring under Poisson sampling. This yields a
-Phase 1 sample
-size of $n_1 = 1000$ and a Phase 2 sample size of
-$E[n_2] \approx 267$. In the Phase 1 sample, we observe 
-$(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. This
-simulation does not deal with model misspecification, and we compare the
+Phase 2 sampling occuring under Poisson sampling. The Phase 1 sample has
+$n_1 = 1200$ and the Phase 2 sample has an expected size of
+$E[n_2] \approx 2588$. In the Phase 1 sample, we observe 
+$(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. 
+The parameter $\delta \in \{0, 1\}$ controls the effect of model misspecification
+in the simulation. Let $\bf x_i = (1, x_{1i}, x_{2i})^T$. We compare the
 proposed method for the parameter $\bar Y_N$ with four approaches:
@@
 \begin{itemize}
-  \item[1.] $\pi^*$-estimator: $\hat Y_{\pi^*} = N^{-1} \sum_{i \in A_2}
+  \item[1.] $\pi^*$-estimator (PiStar): $\hat Y_{\pi^*} = N^{-1} \sum_{i \in A_2}
     \frac{y_i}{\pi_{1i} \pi_{2i|1}},$
-  \item[2.] Two-Phase Regression estimator (TP-Reg): 
-    $\hat Y_{reg} = \sum_{i \in A_1} \frac{\bf x_i' \hat \beta}{\pi_{1i}} + 
-    \sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}}(y_i - \bf x_i' \hat \beta)$ 
-    where $\hat \beta = 
-    \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$
-    and $\bf x_i = (1, x_{1i}, x_{2i})^T$,
-    % \textcolor{red}{Dr. Kim, should I modify the simulation so that the
-    % regression estimator also includes $g(\pi_{2i|1}^{-1})$ as a covariate?}
-
-    %\textcolor{red}{You should include an intercept in $\bf x_i$. }
-    %\textcolor{blue}{I have been including an intercept, but I forgot to add it
-    %into these notes.}
-  \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
+  \item[2.] Two-Phase Regression estimator (Reg): 
+    $\hat Y_{reg} = \sum_{i \in A_1} \frac{\bf x_i' \widehat{\bm \beta}}{\pi_{1i}} + 
+    \sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}}(y_i - \bf x_i' \widehat{\bm \beta})$ 
+    where $\widehat{\bm \beta} = 
+    \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$.
+  \item[3.] Debiased Calibration with Population Constraints (EstPop): This 
     solves
@@
 
-  \item[4.] Debiased Calibration with Estimated Population Constraints (DC-Est):
+  \item[4.] Debiased Calibration with Estimated Population Constraints (Est):
     This solves Equation~\eqref{eq:primal} with $q_i = 1$.
-  \item[5.] Linearized Debiased Calibration with Population Constraints (Eta-Pop):
-    This is the linearized estimate DC-Pop using the linearization in
-    Theorem~\ref{thm:dc1}.
-  \item[6.] Linearized Debiased Calibration with Estimated Population
-    Constraints (Eta-Est): This is the linearized estimate DC-Est using the
-    linearization in Theorem~\ref{thm:dc1}.
 \end{itemize}
 
-In addition to estimating the mean parameter $\bar Y_N$, we also construct
-variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$. For each
-approach we give the variance estimate in Table~\ref{tab:varforms}.
-
-\begin{table}[ht!]
-  \centering
-  \begin{tabular}{lcc}
-    \toprule
-    Estimator & Estimated Variance & Notes \\
-    \midrule
-    $\pi^*$ 
-    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_Y + 
-      \sum\limits_{i \in A_2} \frac{w_{1i}^2}{N^2} 
-        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}y_i^2$}
-    & {\scriptsize$\hat s^2_Y = \frac{1}{n_2 - 1}\sum\limits_{i \in A_2} (y_i -
-    \bar y_i)^2$} \\
-    TP-Reg  
-    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\eta + 
-      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
-        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
-    & {\scriptsize$ \eta_i = \bf x_i'\bm \beta + \delta_{2i} w_{2i|1}\varepsilon_i,\;
-        \varepsilon_i = (y_i - \bf x_i' \bm \beta)$} \\
-    DC-Pop  
-    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\varepsilon + 
-      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
-        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
-    & {\scriptsize$\eta_i = \bf z_i' \bm \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
-        \varepsilon_i = (y_i - \bf z_i'\bm \phi)$} \\
-    DC-Est  
-    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\eta + 
-      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
-        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
-    & {\scriptsize$\eta_i = \bf z_i' \bm \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
-        \varepsilon_i = (y_i - \bf z_i' \bm \phi)$} \\
-    Eta-Pop & Same as DC-Pop & \\
-    Eta-Est & Same as DC-Est & \\
-    \bottomrule
-  \end{tabular}
-  \caption{This table gives the formulas for each variance estimator used in
-  Simulation 1. For the derivation of the estimated variance of the DC-Pop
-estimatator see Appendix A.}
-  \label{tab:varforms}
-\end{table}
-
-%\textcolor{red}{I am not sure about the variance estimation formula in Table 1. 
-%They are different from what we discussed before. I wonder whether you can make
-%the sampling rate $n/N$ negligible so that the second term becomes
-%insignificant. How about making $N=50,000$, $n_1=500$, and $n_2=200$?}
-%
-%\textcolor{blue}{The result with $N = 50,000$, $n_1 = 500$ and $E[n_2] \approx
-%200$}
+%In addition to estimating the mean parameter $\bar Y_N$, we also construct
+%variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$. For each
+%approach we give the variance estimate in Table~\ref{tab:varforms}.
+%% Move variance calculations into an appendix?
 %
@@
 %  \centering
-%  {\color{blue}\input{tables/tpdcsim50000_mean.tex}}
-%% This table was generated from ../../src/explore/proto-dctpmcmc.R
-%  \caption{\textcolor{blue}{This table shows the results of Simulation Study 1 with
-%      a different finite population and Phase 2 selection probability.
-%      It displays the Bias, RMSE, empirical 95\% confidence interval, and a
-%      t-statistic assessing the unbiasedness of each estimator for the
-%      estimators: $\pi^*$, TP-Reg, DC-Pop, and DC-Est.}}
+%  \begin{tabular}{lcc}
+%    \toprule
+%    Estimator & Estimated Variance & Notes \\
+%    \midrule
+%    $\pi^*$ 
+%    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_Y + 
+%      \sum\limits_{i \in A_2} \frac{w_{1i}^2}{N^2} 
+%        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}y_i^2$}
+%    & {\scriptsize$\hat s^2_Y = \frac{1}{n_2 - 1}\sum\limits_{i \in A_2} (y_i -
+%    \bar y_i)^2$} \\
+%    TP-Reg  
+%    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\eta + 
+%      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
+%        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
+%    & {\scriptsize$ \eta_i = \bf x_i'\bm \beta + \delta_{2i} w_{2i|1}\varepsilon_i,\;
+%        \varepsilon_i = (y_i - \bf x_i' \bm \beta)$} \\
+%    DC-Pop  
+%    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\varepsilon + 
+%      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
+%        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
+%    & {\scriptsize$\eta_i = \bf z_i' \bm \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
+%        \varepsilon_i = (y_i - \bf z_i'\bm \phi)$} \\
+%    DC-Est  
+%    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\eta + 
+%      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
+%        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
+%    & {\scriptsize$\eta_i = \bf z_i' \bm \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
+%        \varepsilon_i = (y_i - \bf z_i' \bm \phi)$} \\
+%    \bottomrule
+%  \end{tabular}
+%  \caption{This table gives the formulas for each variance estimator used in
+%  Simulation 1. For the derivation of the estimated variance of the DC-Pop
+%estimatator see Appendix A.}
+%  \label{tab:varforms}
 %\end{table}
 
-We run the simulation $1000$ times for each of these methods and compute the
+We run the simulation $B = 1000$ times for each of these methods and compute the
 Bias ($E[\hat Y] - \bar Y_N$), the RMSE ($\sqrt{\Var(\hat Y - \bar Y_N)}$), a 95\%
@@
 unbiasedness of each estimator where $\hat Y^{(b)}$ is the result from the $b$th
-simulation replicate. The results are in Table~\ref{tab:tpdc-mean}.
+simulation replicate. We also include the Monte Carlo variance of the estimated
+value, ($V_{MC} = (B - 1)^{-1} \sum_{b = 1}^B (\hat Y^{(b)} - \bar Y_N)^2$), the
+mean of the estimated variance, 
+($\bar{\hat V} = B^{-1} \sum_{b = 1}^B \hat V^{(b)}$), and the relative bias of
+the estimated variance, ($(V_{MC} - \bar{\hat V}) / V_{MC}$). The results are in
+Table~\ref{tab:tpdc0-mean} and Table~\ref{tab:tpdc1-mean}.
 
@@
   \centering
-\input{tables/tpdcsim_mean.tex}
-% This table was generated from ../../src/explore/proto-dctpmcmc.R
-\caption{This table shows the results of Simulation Study 1. It displays the
-Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
-unbiasedness of each estimator for the estimators: $\pi^*$, TP-Reg, DC-Pop, and
-DC-Est.}
-\label{tab:tpdc-mean}
+  \input{tables/twophasesampd0.tex}
+\caption{This table shows the results of Simulation Study 1 with $\delta = 0$.
+  It displays the Bias, RMSE, empirical 95\% confidence interval, a t-statistic
+  assessing the unbiasedness, the Monte Carlo variance, mean estimated variance
+  and relative bias of the variance estimator for the estimators: PiStar, Reg,
+  EstPop, and Est.}
+\label{tab:tpdc0-mean}
 \end{table}
 
-\textcolor{red}{While DC-Est still does not outperform TP-Reg, it is equivalent
-to Eta-Est, which is a regression estimator.}
-
-\textcolor{red}{The variance estimates look good now. I needed to think of
-DC-Pop as a regression estimator instead of a two-phase regression estimator.}
-
-\textcolor{red}{I also give the mean variance estimates, the Monte Carlo mean,
-as well as a test for bias.}
 \begin{table}[ht!]
   \centering
-{\color{red} \input{tables/tpdcsim_var.tex} }
-% This table was generated from ../../src/explore/proto-dctpmcmc.R
-\caption{\textcolor{red}{This table shows the variance estimates from Simulation
-    Study 1. It displays the Monte Carlo variance of $\hat Y^{(b)}$, the average
-    estimated variance, the variance of the variance estimator, and a
-    t-statistic assessing the unbiasedness of each variance estimator.}}
-\label{tab:tpdc-var}
+  \input{tables/twophasesampd1.tex}
+\caption{This table shows the results of Simulation Study 1 with $\delta = 1$.
+  It displays the Bias, RMSE, empirical 95\% confidence interval, a t-statistic
+  assessing the unbiasedness, the Monte Carlo variance, mean estimated variance
+  and relative bias of the variance estimator for the estimators: PiStar, Reg,
+  EstPop, and Est.}
+\label{tab:tpdc1-mean}
 \end{table}
@@
 in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
-two-phase sampling the Phase 2 sample is independent of the Phase 1 sample. It
-is a separate independent sample of the same population. Like traditional two
-phase sampling, we consider the Phase 1 sample, $A_1$, to consist of
-observations of $(\bf X_i)_{i = 1}^{n_1}$ and the Phase 2 sample, $A_2$, to consist
-of observations of $(\bf X_i, Y_i)_{i = 1}^{n_2}$.
+two-phase sampling the Phase 2 sample is independent of the Phase 1 sample. 
+Like traditional two-phase sampling, we consider the Phase 1 sample, $A_1$, to
+consist of observations of $(\bf X_i)_{i = 1}^{n_1}$ and the Phase 2 sample,
+$A_2$, to consist of observations of $(\bf X_i, Y_i)_{i = 1}^{n_2}$.
 
@@
 (\cite{kish1965survey}) to get
-$\hat{\bf X_c} = (n_{1, eff} \hat{\bf X_1} + n_{2, eff}\hat{\bf X_2}) / 
-(n_{1, eff} + n_{2, eff})$ where $n_{1, eff}$ and $n_{2, eff}$ are the effective
-sample size for $A_1$ and $A_2$ respectively. Then we can define a regression
-estimator as
+$\hat{\bf X_c} = (n_{1, \rm{eff}} \hat{\bf X_1} + n_{2, \rm{eff}}\hat{\bf X_2}) / 
+(n_{1, \rm{eff}} + n_{2, \rm{eff}})$ where $n_{1, \rm{eff}}$ and 
+$n_{2, \rm{eff}}$ are the effective sample size for $A_1$ and $A_2$ respectively.
+Then we can define a regression estimator as
 
@@
 
-where, $W = n_{1, eff} / (n_{1, eff} + n_{2, eff})$, and
-
-$$\hat{\bm \beta_q} = \left(\sum_{i \in A_2} 
-\frac{\bf x_i \bf x_i^T}{q_i}\right)^{-1} 
-\sum_{i \in A_2} \frac{\bf x_i y_i}{q_i} \text{ and
-}\hat Y_2 = \sum_{i \in A_2} d_{2i} y_i. $$
+where, $W = n_{1, \rm{eff}} / (n_{1, \rm{eff}} + n_{2, \rm{eff}})$, and
+
+$$
+\hat{\bm \beta_q} = 
+\left(\sum_{i \in A_2} \frac{\bf x_i \bf x_i^T}{q_i}\right)^{-1} 
+\sum_{i \in A_2} \frac{\bf x_i y_i}{q_i} \text{ and }
+\hat Y_2 = \sum_{i \in A_2} d_{2i} y_i.
+$$
+
+From an optimality perspective the choice of using the effective sample size to
+weight the estimates from $A_1$ and $A_2$ is reasonable because often the
+effective sample sizes are proportional to the variance of the estimates 
+of $\bf X$. While we know that the inverse variance weighted estimate is optimal
+to combine independent samples for a linear estimate, using the effective sample
+size approximates this procedure without requiring the actual variance of an
+estimator to be known.
 
@@
 
-$$V(\hat Y_{NN, reg}) = V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i -
-\bf x_i^TW\bm \beta^*_q)\right) + (\bm \beta^*_q)^T W^T V(\hat{\bf X_1}) W \bm \beta_q^*$$
-
-where $\bm \beta_q^*$ is the probability limit of $\hat{\bm \beta_q}$. Like the two-phase
-sample this regression estimator can be viewed as the solution to the following
-calibration equation
+$$
+V(\hat Y_{NN, reg}) = 
+V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i - \bf x_i^TW\bm \beta^*_q)\right)
++ (\bm \beta^*_q)^T W^T V(\hat{\bf X_1}) W \bm \beta_q^*
+$$
+
+where $\bm \beta_q^*$ is the probability limit of $\hat{\bm \beta_q}$. Like the
+two-phase sample this regression estimator can be viewed as the solution to the
+following calibration equation
 
@@
 between solving Equation~\ref{eq:nnopt} and Equation~\ref{eq:primal} is that the
-estimator $\hat{\bf X_c}$ is estimated from the combined sample $A_c= A_1 \cup
+estimator $\hat{\bf X_c}$ is estimated from the combined sample $A_c = A_1 \cup
 A_2$. Before using $\hat{\bf X_c}$ in the debiased calibration estimator, we
@@
 
-$$\hat{\bf X_1} = \sum_{i \in A_1} d_{1i} \bf x_i \text{ and }
-\hat{\bf X_2} = \sum_{i \in A_2} d_{2i} \bf x_i.$$
-
-Let $V_1$ and $V_2$ be known variance matrices for $\hat{\bf X_1}$ and 
-$\hat{\bf X_2}$, then the optimal combined estimate is
-
-$$\hat{\bf X_c} = \frac{V_1^{-1} \hat{\bf X_1} + V_2^{-1} 
-\hat{\bf X_2}}{V_1^{-1} + V_2^{-1}}.$$
+$$
+\hat{\bf X_1} = \sum_{i \in A_1} d_{1i} \bf x_i \text{ and }
+\hat{\bf X_2} = \sum_{i \in A_2} d_{2i} \bf x_i.
+$$
+
+Let $n_{1, \rm{eff}}$ and $n_{2, \rm{eff}}$ be the effective samples sizes for
+$A_1$ and $A_2$ respectively. Then the optimal combined estimate is
+
+$$
+\hat{\bf X_c} = (n_{1, \rm{eff}} \hat{\bf X_1} + n_{2, \rm{eff}}\hat{\bf X_2}) / 
+(n_{1, \rm{eff}} + n_{2, \rm{eff}})
+$$
 
@@
 
-with
+where $\bm \lambda$ are the Lagrange multipliers, with
 
@@
 
-\textcolor{red}{I am a little worried about the second term of $\hat{T}$. It is
-  somewhat weird to assume that $T_g= \sum_{i \in U} g(d_{2i}) q_i$ is known.
-  One simple way is to use a HT estimator  of $T_g$ from sample $A_2$. That is,
-  use $\hat{T}_{g, {\rm HT}} = \sum_{i \in A_2} d_{2i} g( d_{2i}) q_i$ or even
-  use a GREG estimator of $T_g$. 
-%I wonder we can assume that an unbiased estimator of $T_g$ is available from
-  %$A_c=A_1 \cup A_2$. That is, let's assume that we have $\hat{T}_c = W
-  %\hat{T}_1 + (1- W) \hat{T}_2$ as the estimator  for $T$. 
-}
-
 Differntiating with respect to $w_{2i}$ and setting this expression equal to
@@
 
-  $$\hat Y_{\ell, NNE}(\bm \lambda^*, \bm \phi^*) = \sum_{i \in A_2} \hat w_{2i}(
+  $$
+  \hat Y_{\ell, NNE}(\bm \lambda^*, \bm \phi^*) = \sum_{i \in A_2} \hat w_{2i}(
   \bm \lambda^*) + \left(\hat{\bf T} - \sum_{i \in A_2} \hat w_{2i}(
-  \bm \lambda^*) \bf z_i q_i\right) \bm \phi^*$$
+  \bm \lambda^*) \bf z_i q_i\right) \bm \phi^*
+  $$
 
   and
-  $$\bm \phi^* =
+
+  $$
+  \bm \phi^* =
   \left(\sum_{i \in U} \frac{\pi_{2i} q_i}{g'(d_{2i})} 
@@
 \varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
-Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
+Y_{i} &= 3 X_{1i} + 2 X_{2i} + \delta X_{3i} \varepsilon_i \\
 \pi_{1i} &= n_1 / N \\
-\pi_{2i} &= \min(\Phi_3(-x_{1i} + 1), 0.9).
+\pi_{2i} &= \max(\min(\Phi_3(X_{3i} - 2.5), 0.9), 0.01).
 \end{aligned}
@@
 where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
-%\textcolor{red}{The simulation setup is somewhat strange. The first order
-%inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a
-%function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$.
-%In this case, your second-phase sampling is essentially equivalent to the simple
-%random sampling. }
 We consider a finite population of size $N = 10,000$ with the Phase 1 
 sampling being a simple random sample (SRS) of size $n_1 = 1000$. The Phase 2
-sample is a Poisson sample with an expected sample size of
-
-and
-Phase 2 sampling occuring under Poisson (Bernoulli) sampling. This yields a
-Phase 1 sample
-size of $E[n_1] \approx 1100$ and a Phase 2 sample size of
-$E[n_2] \approx 300$. In the Phase 1 sample, we observe 
-$(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. This
-simulation does not deal with model misspecification, and we compare the
-proposed method for the parameter $\bar Y_N$ with four approaches:
+sample is a Poisson sample with an expected sample size of about 300.
+In the Phase 1 sample, we observe 
+$(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. 
+If $\delta = 0$ then there is no model misspecification. However, if $\delta =
+1$, then there is some model misspecification. We estimate the
+the parameter $\bar Y_N$ with four approaches:
 
 \begin{itemize}
-  \item[1.] HT-estimator: $\hat Y_{HT} = N^{-1} \sum_{i \in A_2}
+  \item[1.] HT-estimator (HT): $\hat Y_{HT} = N^{-1} \sum_{i \in A_2}
     \frac{y_i}{\pi_{2i}},$
@@
     \sum_{i \in A_2} d_{2i} y_i$, $\hat{\bf X_c} = W \hat{\bf X_{1, HT}} + (1 -
-    W)\hat{\bf X_{2, HT}}$, $W = n_{1,eff} / (n_{1,eff} + n_{2, eff})$,
+    W)\hat{\bf X_{2, HT}}$, $W = n_{1,\rm{eff}} / (n_{1,\rm{eff}} + n_{2, \rm{eff}})$,
     $\hat{\bf X_{1, HT}} = \sum_{i \in A_1} d_{1i} \bf x_{i}$, $\hat{\bf X_{2,
@@
     Then $\hat{\bar Y}_{NN, reg} = \hat Y_{NN, reg} / N$.
-  \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
+  \item[3.] Debiased Calibration with Population Constraints (EstPop): This 
     solves
@@
 
-  \item[4.] Debiased Calibration with Estimated Population Constraints (DC-Est):
+  \item[4.] Debiased Calibration with Estimated Population Constraints (Est):
     This solves Equation~\eqref{eq:nnopt} with $q_i = 1$.
@@
 variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$.
-% For each approach we give the variance estimate in Table~\ref{tab:varforms2}.
-
-We run the simulation $1000$ times for each of these methods and compute the
+
+We run the simulation $B = 1000$ times for each of these methods and compute the
 Bias ($E[\hat Y] - \bar Y_N$), the RMSE ($\sqrt{\Var(\hat Y - \bar Y_N)}$), a 95\%
@@
 unbiasedness of each estimator where $\hat Y^{(b)}$ is the result from the $b$th
-simulation replicate. The results are in Table~\ref{tab:nndc-mean}.
+simulation replicate. We also include the Monte Carlo variance of the estimated
+value, ($V_{MC} = (B - 1)^{-1} \sum_{b = 1}^B (\hat Y^{(b)} - \bar Y_N)^2$), the
+mean of the estimated variance, 
+($\bar{\hat V} = B^{-1} \sum_{b = 1}^B \hat V^{(b)}$), and the relative bias of
+the estimated variance, ($(V_{MC} - \bar{\hat V}) / V_{MC}$). The results are in
+Table~\ref{tab:nndc0-mean} and Table~\ref{tab:nndc1-mean}.
 
@@
   \centering
-\input{tables/nndcsim_mean.tex}
-% This table was generated from ../../src/explore/proto-dctpmcmc.R
-\caption{This table shows the results of Simulation Study 2. It displays the
-Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
-unbiasedness of each estimator for the estimators: HT, Reg, DC-Pop, and
-DC-Est.}
-\label{tab:nndc-mean}
+  \input{tables/nonnestedtpd0.tex}
+\caption{This table shows the results of Simulation Study 2 with $\delta = 0$.
+  It displays the Bias, RMSE, empirical 95\% confidence interval, a t-statistic
+  assessing the unbiasedness, the Monte Carlo variance, mean estimated variance
+  and relative bias of the variance estimator for the estimators: HT, Reg,
+  EstPop, and Est.}
+\label{tab:nndc0-mean}
+\end{table}
+
+\begin{table}[ht!]
+  \centering
+  \input{tables/nonnestedtpd1.tex}
+\caption{This table shows the results of Simulation Study 2 with $\delta = 1$.
+  It displays the Bias, RMSE, empirical 95\% confidence interval, a t-statistic
+  assessing the unbiasedness, the Monte Carlo variance, mean estimated variance
+  and relative bias of the variance estimator for the estimators: HT, Reg,
+  EstPop, and Est.}
+\label{tab:nndc1-mean}
 \end{table}
@@
 two samples. Now, we consider incorporating more than two independent samples
-together with a debiasing constraint. First, we consider the case in which we
+together with a debiasing constraint. We consider the case in which we
 want to estimate the $\theta = E[Y]$ and $Y$ is only observed in one sample.
-Then we will consider the case where $Y$ may be in multiple samples.
 
@@
 \begin{equation}\label{eq:mstps}
-\hat w = \argmin_w \sum_{i \in A_0} G(w_i)q_i \text{ such that } \sum_{i \in
-A_0} w_i \bf x_i = \hat X_{GLS}^{(0)}
-\end{equation}
-
-where $\hat X_{GLS}^{(0)}$ is the GLS estimate of $X$ for all of the
+\hat w = \argmin_w \sum_{i \in A_0} G(w_i)q_i \text{ such that } 
+\sum_{i \in A_0} w_i \bf x_i = \hat X_{\rm{GLS}}^{(0)}
+\end{equation}
+
+where $\hat{\bf X}_{\rm{GLS}}^{(0)}$ is the GLS estimate of $X$ for all of the
 $X$-variables measured in sample $A_0$, and $G$ is a generalized entropy
@@
 
-% Outline
-% Instead of incorporating data from one additional source, we want to combine
-% data from multiple sources
-% * Use GLS (generalized from effective sample size)
-% * We need to add debiasing constraint 
-%   * What do we do if we have multiple surveys for which we observe Y?
-%   * How do we incorporate the selection probability?
-
 \subsection{Methodology}
@@
 
-$$ \sum_{i \in A_0} w_{i} \bf x_i q_i = \sum_{i \in U} g(d_i) q_i $$
+\begin{equation}\label{eq:msdebconstr}
+\sum_{i \in A_0} w_{i} \bf x_i q_i = \sum_{i \in U} g(d_i) q_i
+\end{equation}
 
@@
 \begin{equation}\label{eq:dc3}
-  \hat w = \argmin_w \sum_{i \in A_0} G(w_i) q_i + \lambda\left(\hat T - \sum_{i
-  \in A_0} w_i \bf z_i\right)
-\end{equation}
-
-where $\hat T = (\hat{\bf X}^{(0)}, \sum_{i \in U} g(d_i) q_i)^T$ and 
-$\bf z_i = (\bf x_i^T, g(d_i)q_i)^T$. A more general approach considers the case
-in which multiple samples can observe $Y$. In this case, assume that we have
-samples $A_1, \dots A_{M_1}$ for which we observe elements of $\bf X$ and $Y$
-and samples $B_1, \dots, B_{M_2}$ for which we only observed elements of $\bf
-X$. We denote $\bf X^{(A_m)}$ and $\bf X^{(B_m)}$ as the columns of $\bf X$
-observed in samples $A_m$ and $B_m$ respectively. Then the final
-estimator\footnote{This is sort of like stratified sampling with unknown strata.}
-is
-
-$$ \hat Y = M_1^{-1} \sum_{m = 1}^{M_1} \sum_{i \in A_m} \hat w_{im} y_i $$
-
-and $\hat w_{im}$ solves
-
-\begin{equation}\label{eq:dc32}
-  \hat w = \argmin_w \sum_{i \in A_m} G(w_i) q_i + \lambda\left(\hat T_m - \sum_{i
-  \in A_m} w_i \bf z_i\right)
-\end{equation}
-
-where $\hat T = (\hat{\bf X}^{(A_m)}, \sum_{i \in U} g(d_{im}) q_i)^T$ and $d_{im}$
-are the design weights for $A_m$.
+  \hat w = \argmin_w \sum_{i \in A_0} G(w_i) q_i \text{ such that }
+  \widehat{\bf{T}} = \sum_{i \in A_0} w_i \bf z_i.
+\end{equation}
+
+where $\widehat{\bf{T}} = (\widehat{\bf X}_{\rm{GLS}}^{(0)},
+\sum_{i \in U} g(d_i) q_i)^T$,
+and $\bf z_i = ((\bf x_i^{(0)})^T, g(d_i)q_i)^T$. Then
+the estimator is $\hat Y_{MS} = \sum_{i \in A_0} w_{i} y_i$ and we can construct
+this estimate minimizing the Lagrangian:
+
+$$
+L(w_{0}, \bm \lambda) = \sum_{i \in A_0} G(w_{0i}) q_i + {\bm \lambda}^T
+\left(\widehat{\bf{T}} - \sum_{i \in A_0} w_i \bf z_i\right)
+$$
+
+where $\bm \lambda$ are the Lagrange multipliers.
 
@@
 
-\subsection{Simulation Study 3.1}
-
-We first run a simulation for the case in which we only have one sample in which
-we observe $Y$. We have the following superpopulation model with $N=10000$
-elements:
+\begin{theorem}[Design Consistency]
+  Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$.
+  Under regularity conditions,
+
+  $$\hat Y_{MS} = \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) + O_p(N / n_0)$$
+
+  where
+
+  $$
+  \hat Y_\ell(\bm \lambda^*, \bm \phi^*) = \hat Y_{DCE} + \left(
+    \widehat{\bf T} - \sum_{i \in A_0} \hat w_{0i}(\bm \lambda^*) \bf z_i q_i
+  \right)\bm \phi^*
+  $$
+
+  and
+
+  $$
+  \bm \phi^* = 
+  \left[\sum_{i \in U} \frac{\pi_{0i} \bf z_i \bf z_i^T q_i}{g'(d_{0i})}\right]^{-1}
+  \sum_{i \in U} \frac{\pi_{0i} \bf z_i y_i}{g'(d_{0i})}.
+  $$
+\end{theorem}
+
+\begin{proof}
+  This proof follows from the proof of Theorem~\ref{thm:dc2}. The only
+  difference is that we have $\hat X_{\rm GLS}$ instead of $\hat X_c$.
+\end{proof}
+
+\begin{theorem}[Variance Estimation]
+  Under regularity conditions,
+
+  $$
+  \begin{aligned}
+    V(\hat Y_{MS}) 
+    &= (\bm \phi^{*})^T\Var(\hat{\bf X}_{\rm GLS}^{(0)}) (\bm \phi^{*}) + 
+      \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{0ij}}{\pi_{0i}\pi_{0j}}
+      (y_i - \bf z_i\bm \phi^{*}q_i)
+      (y_j - \bf z_j\bm \phi^{*}q_j). 
+  \end{aligned}
+  $$
+
+  We can estimate the variance with
+
+  $$
+  \begin{aligned}
+    \hat V(\hat Y_{MS}) 
+    &= (\hat{\bm \phi})^T \hat \Var(\hat{\bf X}_{\rm GLS}^{(0)}) (\hat{\bm \phi})
+    + \sum_{i \in A_0} \sum_{j \in A_0} \frac{\Delta_{0ij}}{\pi_{0ij}\pi_{0i}\pi_{0j}}
+      (y_i - \bf z_i \hat{\bm \phi}q_i)
+      (y_j - \bf z_j \hat{\bm \phi}q_j). 
+  \end{aligned}
+  $$
+
+\end{theorem}
+
+\begin{proof}
+  This result follows the same argument as Theorem~\eqref{thm:nnve}. The result
+  holds because each sample $A_m$ is independent from the other $A_{m'}$.
+\end{proof}
+
+\subsection{Simulation Study 3}
+
+We have the following superpopulation model with $N = 10000$ elements:
 
@@
 X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
-X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
+X_{3i} &\stackrel{ind}{\sim} N(5, 1) \\
 Z_i &\stackrel{ind}{\sim} N(0, 1) \\
 \varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
-Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
-\pi_{0i} &= \min(\max(\Phi(-x_{1i}), 0.02), 0.9)\\
-\pi_{1i} &= n_1 / N.\\
-\pi_{2i} &= \Phi(x_{2i} - 2) \\
+Y_{i} &= 3 X_{1i} + 2 X_{2i} + \delta Z_{i} + \varepsilon_i \\
+\pi_{0i} &= \min(\max(\Phi(Z_i - 2), 0.02), 0.9)\\
+\pi_{1i} &= n_1 / N\\
+\pi_{2i} &= \Phi(X_{2i} - 2) \\
 \end{aligned}
@@
 
-We observe the following columns in each sample
+Like the previous simulation studies, when $\delta = 1$, there is model
+misspecification for the outcome model because we observe the following columns
+in each sample
 
@@
 
-For the sampling mechanism both $A_0$ and $A_1$ are selected using a Poisson
+For the sampling mechanism both $A_0$ and $A_2$ are selected using a Poisson
 sample with response probabilities $\pi_{0i}$ and $\pi_{1i}$ respectively. The
-sample $A_2$ is a simple random sample with $n_1 = 2000$. We compare five
+sample $A_1$ is a simple random sample with $n_1 = 2000$. We compare four
 different estimators for $\theta = E[Y]$.
@@
     $A_1$,
-  \item[3.] Multi-Source proposed (MSEst): This is the proposed estimator from
-    Equation~\eqref{eq:dc3},
-  \item[4.] Multi-Source population (MSPop): This is the proposed estimator with
-    using the true value of $T_1$ from the population,
-  \item[5.] Multi-Source regression (MSReg): This is the linearized regression
-    estimator from FIXME.
+  \item[3.] Multi-Source proposed (Est): This is the proposed estimator from
+    Equation~\eqref{eq:dc3}, and
+  \item[4.] Multi-Source population (EstPop): This is the proposed estimator with
+    using the true value of $T_1$ from the population.
 \end{itemize}
+
+The simulation results are displayed in Table~\ref{tab:msdc0-mean} and
+Table~\ref{tab:msdc1-mean}.
 
@@
   \centering
-  \input{tables/msdcsim1.tex}
-% This table was generated from ../../src/explore/20240506-msdcsim.qmd
-\caption{This table shows the results of Simulation Study 2. It displays the
-Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
-unbiasedness of each estimator for the estimators: HT, NNReg, MSPop, MSEst, and
-MSReg.}
-\label{tab:msdc-mean}
+  \input{tables/multisourced0.tex}
+\caption{This table shows the results of Simulation Study 3 with $\delta = 0$.
+  It displays the Bias, RMSE, empirical 95\% confidence interval, a t-statistic
+  assessing the unbiasedness, the Monte Carlo variance, mean estimated variance
+  and relative bias of the variance estimator for the estimators: HT, NNReg,
+  EstPop, and Est.}
+\label{tab:msdc0-mean}
+\end{table}
+
+\begin{table}[ht!]
+  \centering
+  \input{tables/multisourced1.tex}
+\caption{This table shows the results of Simulation Study 3 with $\delta = 1$.
+  It displays the Bias, RMSE, empirical 95\% confidence interval, a t-statistic
+  assessing the unbiasedness, the Monte Carlo variance, mean estimated variance
+  and relative bias of the variance estimator for the estimators: HT, NNReg,
+  EstPop, and Est.}
+\label{tab:msdc1-mean}
 \end{table}
@@
 
-\subsection{Simulation Study 3.2}
-
-Building off of the previous simulation study, we now consider the case for
-which we observe $Y$ in separate samples.
-We have the following superpopulation model with $N = 10,000$ elements:
-
-$$
-\begin{aligned}
-X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
-X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
-X_{3i} &\stackrel{ind}{\sim} N(5, 1) \\
-\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
-Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
-\pi_{A1i} &= \min(\max(\Phi(-x_{1i}), 0.02), 0.9)\\
-\pi_{A2i} &= \min(\max(\Phi(x_{3i} - 6), 0.02), 0.95) \\
-\pi_{B1i} &= n_1 / N\\
-\pi_{B2i} &= \Phi(x_{2i} - 2) \\
-\end{aligned}
-$$
-
-We observe the following columns in each sample
+\section{Real-Data Analysis}
+
+We apply our method to Medicare Current Beneficiary Survey (MCBS). The MCBS
+is an ongoing representative national survey of the population of people on
+Medicare in the United States, and in 2020, NORC started issuing an MCBS
+supplement to identify the impact of COVID-19 on the Medicare population
+\cite{norc2020mcbs}. This analysis looks at whether
+respondents had obtained at least one dose of a COVID-19 vaccine.
+
+Assessing the vaccine rate of Medicare patients is important from a public
+health perspective. Older people\footnote{Medicare is available to people ages
+65 years and older and select designated populations between the ages of 18 and
+65. However, the vast majority of Medicare patients are 65 years old or older.}
+are known to have a higher all-cause mortality from
+COVID-19 (\cite{bonanad2020effect}). Thus, understanding how older people get
+vaccinated can help save lives and reduce medical costs.
+
+The MCBS is a rotating panel survey in which participants are interviewed up to
+three times within a four-year period (\cite{cms2021mcbs}). The participants
+are selected from Medicare enrollment data and enter a cohort in the fall round
+of the survey. They then respond to the winter, summer,
+and fall rounds of the survey for three consecutive years before exiting the
+survey in the winter round of year four (\cite{cms2021mcbs}). Each year a new
+cohort is added. Participants are selected based on a three-stage cluster sample
+design. The first stage of the sample consists of major metropolitan areas and
+groups of rural counties (\cite{cms2021mcbs}). The second stage are census
+tracts within the primary sampling units. Finally, Medicare beneficiaries are
+selected using a stratified systematic sample with random starts 
+(\cite{cms2021mcbs}). Medicare beneficiaries are selected from strata based on
+age and whether the beneficiary is Hispanic. Strata with Hispanic beneficiaries
+and strata with beneficiaries over the age of 85 and under the age of 65 are
+oversampled.
+The COVID supplement was added to the traditional MCBS survey in March of
+2020 to assess the impact of COVID on the Medicare
+population (\cite{cms2021covid}). These questions about vaccines were added to
+participants who were enrolled in the Summer 2021 cohort.
+
+We combine the 2021 Summer MCBS survey with data from the 2021 American
+Community Survey (ACS)
+issued by the U.S. Census Bureau along with administrative data about people on
+Medicare from the Center of Medicare and Medicaid Services (CMS). Like the MCBS
+survey, the ACS data contains information about the overall percentage of
+people's sex, race and ethnicity, education, and marital status for people on
+Medicare. We combine this information with official totals of people on Medicare
+by sex from the CMS. A complete display of the data available can be seen in
+Table~\ref{tab:mcbscols}.
+
+\begin{table}[ht!]
+  \label{tab:mcbscols}
+  \centering
+  \begin{tabular}{lrrrrr}
+  \toprule
+  Data & Sex & Race and Ethnicity & Education & Marital Status & COVID Vaccine Dose \\
+  \midrule
+  MCBS & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
+  ACS & \checkmark & \checkmark & \checkmark & \checkmark &  \\
+  CMS & \checkmark &  &  &  & \\
+  \bottomrule
+  \end{tabular}
+  \caption{This table shows the data available to us in each data set. A
+  checkmark in a particular column indicates that the data set listed in the
+  left column contains information about the participant characteristic in the
+  given column.}
+\end{table}
+
+To make the results between the ACS and MCBS compatible, we recode the 
+ACS variables with the map found in Table~\ref{tab:acsvars}.
 
@@
   \centering
-  \begin{tabular}{lrrrr}
+  \label{tab:acsvars}
+  \begin{tabular}{lr}
   \toprule
-    Sample & $X_1$ & $X_2$ & $X_3$ & $Y$ \\
+  Variable & Consolidated ACS Variables \\
   \midrule
-   $A_1$  & \checkmark     & \checkmark     & \checkmark     &  \checkmark \\    
-   $A_2$  & \checkmark     & & \checkmark     &  \checkmark \\    
-   $B_1$  & \checkmark     &       & \checkmark     &    \\  
-   $B_2$  & \checkmark     & \checkmark     &       &    \\  
+  Education & SCHL:  01-15, 16-19, 20-24 \\
+  Marital Status & MAR: 1, 2, 3-4, 5 \\
   \bottomrule
   \end{tabular}
+  \caption{This table shows how we consolidated ACS variables. The right column
+  displays the ACS category code and the values used in a particular category to
+  match a MCBS category. For example, SCHL: 01-15 means that the values of 01 -
+  15 of the SCHL ACS column were combined to match an individual MCBS value.}
 \end{table}
 
-For the sampling mechanism $A_1$, $A_2$ and $B_2$ are selected using a Poisson
-sample with response probabilities $\pi_{A1i}$, $\pi_{A2i}$ and $\pi_{B1i}$
-respectively. The $B_2$ sample is a simple random sample (SRS). We estimate the
-following methods:
-
-\begin{itemize}
-  \item[1.] Horvitz-Thompson estimator for $A_1$ (HTA1):
-
-    $$\hat Y_{HT, A_1} = \sum_{i \in A_1} \frac{y_i}{\pi_{A1i}}.$$
-  \item[2.] Horvitz-Thompson estimator for $A_2$ (HTA1):
-    $$\hat Y_{HT, A_2} = \sum_{i \in A_2} \frac{y_i}{\pi_{A2i}}.$$
-  \item[3.] Weighted Horvitz-Thompson estimator (WHT):
-    $$\hat Y_{WHT} = \frac{\hat Y_{HT, A_1} / \hat V_{A_1} + \hat Y_{HT, A_2} /
-    \hat V_{A_2}}{1 / \hat V_{A_1} + 1 / \hat V_{A_2}}$$
-
-    where $\hat V_{A_m}$ is the estimated variance of the Horvitz-Thompson
-    estimator $\hat Y_{HT, A_m}$.
-  \item[4.] Multi-Source proposed (MSEst): This is the proposed estimator from
-    Equation~FIXME.
-  \item[5.] Multi-Source population (MSPop): This is the proposed estimator with
-    using the true value of $T_1$ from the population,
-  \item[6.] Multi-Source regression (MSReg): This is the linearized regression
-    estimator from FIXME.
-\end{itemize}
+While the MCBS data is largely free of missing values, there are a small number
+that need to be imputed for the $Y$ variable about whether someone received at
+least one COVID vaccination dose. This MCBS data was updated to include
+information about if a participant had received a second vaccine dose and then
+later to include how many total vaccine doses patients had recieved. If the
+participant had received at least one more dose and their answer was missing (because
+they refused to answer, did not know, or was just missing) the missing value was
+imputed to say that they yes, they had received a COVID vaccine dose. Otherwise,
+the missing value was imputed to say no, the participant had not received a
+COVID vaccine dose.
+
+We analyze the data with three approaches: a Horvitz-Thompson method, a
+regression algorithm, and our preposed multi-source two-phase sampling technique. 
+The Horvitz-Thompson estimator uses the design weights, the MCBS observation if
+an individual received a COVID-19 vaccine dose, and the population total from
+the CMS about the total number of people on Medicare in 2021. The regression
+estimator uses all of this and a population level total for the number of males.
+Our multi-source algorithm also includes information from the ACS about
+education and marital status. Due to the fact that we do not observe the
+probability that an individual will be selected into the MCBS survey within the
+population of all Medicare patients, we have to estimate the term 
+$\sum_{i \in U} g(d_i)$ from Equation~\ref{eq:msdebconstr}. This technique is
+adopted from \cite{kwon2024debiased} and described in more details in the
+appendix.
+The results of the real data analysis are found in Table~\ref{tab:mcbsres}.
 
@@
   \centering
-  \input{tables/msdcsim2.tex}
-% This table was generated from ../../src/explore/20240511-msdcsim2.qmd
-\caption{This table shows the results of Simulation Study 2. It displays the
-Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
-unbiasedness of each estimator for the estimators: HT, NNReg, MSPop, MSEst, and
-MSReg.}
-\label{tab:msdc2-mean}
+  \label{tab:mcbsres}
+  \begin{tabular}{lrr}
+    \toprule
+    Method & Point Estimate & Standard Error \\
+    \midrule
+    HT  & 0.817 & 0.01223 \\
+    Reg & 0.877 & 0.00439 \\
+    Est & 0.863 & 0.00436 \\ 
+    \bottomrule
+  \end{tabular}
+  \caption{This table shows the point estimate and standard error of the
+  percentage of people on Medicare who had at least one COVID-19 vaccine in
+  2021 as estimated by a Horvitz-Thompson estimator (HT), a regression estimator
+  (Reg) and the proposed multi-source estimator (Est).}
 \end{table}
 
-% FIXME: Change to GLS estimate of Y.
-
-\newpage
-
-\bibliographystyle{chicago}
+\section{Conclusion}
+
+Overall, using debiased calibration for generalized two-phase sampling seems to
+be a promising approach for combining multiple samples efficiently. Not only can
+this method be used to combine surveys focusing on the same population frame,
+but because it only does not require individual responses levels from the
+supplemental surveys, this method can be used to construct domain level
+estimates as exemplified in the real data analysis of Medicare patients.
+
+\bibliographystyle{apa}
 \bibliography{references}
 
+\section{Appendix A: Estimating the Population Weights}
+
+In Equation~\ref{eq:primal}, Equation~\ref{eq:nnopt}, and
+Equation~\ref{eq:msdebconstr}, we see the need for knowing $\sum_{i \in U}
+g(d_i)$. (In the case of Equation~\ref{eq:primal}, we only need $\sum_{i \in
+A_1} g(d_{2i|1})$.) However, this quantity is not always known. When it is
+unknown, we need to estimate it.
+
+We follow the approach of \cite{kwon2024debiased}. Define
+
+$$N \alpha = \sum_{i \in U} g(d_i).$$
+
+Assuming $N$ is known (or that we can estimate it), we need to estimate
+$\alpha$. Since our estimating equation is convex for any fixed $\alpha$, we
+adopt a two-step procedure for which we estimate $\alpha$ and $\bf w$. In this
+approach we have the following loss function for the case of non-nested sampling:
+
+\begin{equation}\label{eq:nnalpha}
+  (\hat \alpha, \hat w_2) = 
+  \argmin_{\alpha} \argmin_{w_2} \sum_{i \in A_2} \{G(w_{2i})\} - NK(\alpha),
+  \text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X_c},
+  \text{ and } \sum_{i \in A_2} w_{2i} g(d_{2i}) = N\alpha.
+\end{equation}
+
+The choice of $K(\alpha)$ is up to the analyst. We consider $K(\alpha) =
+\alpha$.
+
 \end{document}

### 2024-12-12 final/paper/paper_ov.tex
Updated final paper.

--- 2024-11-12 final/paper/paper.tex
@@
 estimated population totals. We then extend this two-phase sampling framework to
-two additional cases: first, when when have two non-nested samples, and
+two additional cases: first, when we have two non-nested samples, and
 second, when we want to combine three or more samples. 
 We apply our method to the Medicare Current Beneficiary Survey for which we
-get more efficient estimates by combining ths data set with population level
+get more efficient estimates by combining this data set with population level
 controls obtained from the Center of Medicare and Medicaid Services as well as
@@
 and finally, we consider the more general approach of having multiple data sets.
+
+\textcolor{red}{The overall structure needs to be changed.
+You can change the current section 2 to introduce the notation and review the basic properties of the regression estimator in two phase sampling.
+Section 3 is the main proposal of the debiasing calibration for two-phase sampling.
+Section 4 is an extension of the proposal to non-nested two-phase sampling and multi-source data integration.
+The simulation studies in the subsections of Sections 2-4 should be moved to Section 5.
+The real data analysis in Section 5 should be changed to Section 6.
+Concluding remarks are made in Section 7.
+In summary, here is my suggestion for the structure.
+\begin{enumerate}
+\item Introduction
+\item Basic setup: introduce the notation, two-phase regression estimator and its calibration form
+\item Main Proposal: introduce debiased calibration for two-phase sampling and discuss its statistical properties 
+\item Extensions: \begin{enumerate}
+\item Extension to non-nested two-phase sampling 
+\item Extension to multi-source data integration
+\end{enumerate}
+\item Simulation study 
+\item Real data analysis 
+\item Concluding Remarks 
+\end{enumerate}
+}
 
@@
 where  $\bf z_i = (\bf x_i / q_i, g(\pi_{2i|1}^{-1}))$. 
+\textcolor{red}{Did you use $w_{2i \mid 1}$ to denote $w_{2i}$? Why do you introduce new notation?  } 
 Let $\hat w_{2i|1}$ be the solution to Equation~\eqref{eq:primal}. The resulting 
@@
 
+  %% FIXME: This needs to be defined in terms of \hat Y_{\pi^*}
   $$
@@
   \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(
-  \bm \lambda) z_i q_i\right)\bm \phi^*
+  \bm \lambda) \bf z_i q_i\right)\bm \phi^*
   $$
@@
 
+\textcolor{red}{This is not enough. We need to discuss variance estimation. }
 \subsection{Simulation Study 1}
+
+\textcolor{red}{This subsection should be moved to Section 5. }
 
@@
 \bibliographystyle{apa}
-\bibliography{references}
+\bibliography{references.bib}
 

### 2025-01-07 final/paper/paper.tex
Updated final paper. Fixed Dr. Kim's suggestions.

@@
 \title{Debiased Calibration for Generalized Two-Phase Sampling}
-\author{Caleb Leedy, Jae-Kwang Kim}
+\author{Caleb Leedy \and Jae-Kwang Kim}
 \maketitle
@@
 \begin{abstract}
-Generalized entropy calibration \cite{gneiting2007strictly} is a nonparametric
+Generalized entropy calibration \citep{gneiting2007strictly} is a nonparametric
 method to estimate survey weights without using the design weights in the
-specified loss function. While \cite{kwon2024debiased} explore generalized
-entropy calibration to estimate survey weights with known population totals, by
-using a two-phase sampling framework, we extend their method to allow for
-estimated population totals. We then extend this two-phase sampling framework to
-two additional cases: first, when when have two non-nested samples, and
+specified loss function. \cite{kwon2024debiased} explore generalized
+entropy calibration to estimate survey weights with known population totals.
+By using a two-phase sampling framework, we extend this method to allow for
+estimated population totals. We then generalize the two-phase sampling framework to
+two additional cases: first, when we have two non-nested samples, and
 second, when we want to combine three or more samples. 
 We apply our method to the Medicare Current Beneficiary Survey for which we
-get more efficient estimates by combining ths data set with population level
+get more efficient estimates by combining this data set with population level
 controls obtained from the Center of Medicare and Medicaid Services as well as
@@
 
+\newpage
+
 \section{Introduction}
 
-Combining information from multiple samples is an important practical problem
-(\cite{yang2020statistical}, \cite{yang2023elastic}, \cite{dagdoug2023model}).
-We want to incorporate information from external data sources to reduce the bias
-in our estimates and/or improve the estimator's efficiency. For many problems, the
-additional information consists of summary statistics with standard errors. The
-goal of this project is to incorporate external information with existing data 
-to create more efficient estimators using calibration weighting.
-
-To model this scenario, we formulate the problem as a generalized two-phase
-sample. The first phase sample consists of data from multiple sources, and the
-second phase sample contains the existing data. To motivate this setup, we
-take the following approach: first, we consider the classical two-phase
-sampling setup where the second phase sample is a subset of the first phase
-sample; then, we extend this setup to consider non-nested two-phase samples;
-and finally, we consider the more general approach of having multiple data sets.
-
-\section{Topic 1: Classical Two-Phase Sampling}
-
-\subsection{Background}
-
-Consider a finite population of size $N$ containing elements $(\bf X_i, Y_i)$ where
-an initial (Phase 1) sample of size $n_1$ is selected and $\bf X_i$ is observed.
-From the Phase 1 sample of elements, a (Phase 2) sample of size $n_2 < n_1$ is
-selected and $Y_i$ is observed. This is two-phase sampling. (See 
-\cite{fuller2009sampling}, \cite{kim2024statistics} for general references.) The
-goal of two-phase sampling is to construct an estimator of $\bar Y_N$ 
+Government surveys in the United States provide the public with diverse
+information about a range of population characteristics. These include health
+indicators from the National Health and Nutrition Examination Survey, population
+and housing data from the American Community Survey, and labor force statistics
+from the Current Population Survey. All of these surveys are probability
+surveys. Probability sampling is considered the
+gold-standard for survey sampling. However, obtaining a large probability
+sample is getting more difficult because people are more likely to choose not to
+respond. One solution to this problem is data integration. Instead of analyzing
+a probability sample by itself, we can incorporate information from other
+probability samples in order to have an analysis with higher statistical power.
+
+The goal of statistical data integration is to combine information from one
+survey with the variable of interest to another survey that only contains some
+shared covariates. 
+This integration usually involves one of two contexts: incorporating item
+level responses or using summary statistics. These mirror the micro and macro
+approaches described in \cite{yang2020statistical} respectively. The main
+technique to incorporate item level responses is mass imputation. With access to
+individual item level responses, the traditional approach is to build a
+predictive model of the variable of interest from the covariates shared between
+the two surveys and predict the missing outcome variable in the survey without
+the variable of interest. An early use of mass imputation is found in
+\cite{breidt1996two} where they used mass imputation to improve estimation in
+the NRI survey. \cite{kim2012combining} used a parametric model to develop a
+novel resampling technique to construct confidence intervals of population level
+characteristics. \cite{chen2022nonparametric} use a nonparametric model to
+estimate population characteristics. In the macro approach, one uses summary
+statistics from other studies to adjust the current estimation. This is commonly
+done using generalized least squares (GLS). 
+\cite{hidiroglou2001double} applied the GLS technique to non-nested two-phase
+samples. \cite{merkouris2004combining} used GLS with multiple surveys and showed
+an optimality result. This was later extended to area level domains in 
+\cite{merkouris2010combining}.
+
+In this paper, we incorporate summary statistics of other surveys by
+using a generalized two-phase sample similar to the GLS method. The first phase
+sample consists of data
+from multiple sources, and the second phase sample contains the existing
+data. Unlike previous data integration methods, however, we also include a
+pre-chosen function of the second phase design weights like
+\cite{kwon2024debiased}. This allows us to develop theory showing that our
+estimates are asymptotically design unbiased and that we have valid confidence
+intervals. Moreover, we can show that this technique outperforms traditional
+regression methods in finite samples.
+
+This paper is organized as follows. Section 2 describes the basic setup of data
+integration using summary statistics. Section 3 details the generalized
+two-phase sampling framework and provides theory for our estimator being design
+unbiased and the variance estimation to be valid. In Section 4, we extend this
+framework first to non-nested two-phase sampling and second to the multiple
+samples. Section 5 provides several simulation studies which show that our
+method works, and Section 6 applies the method the Medicare Current Beneficiary
+Survey.
+
+\newpage
+
+\section{Basic Setup}
+
+Consider a finite population, $U$, of size $N$ 
+with measurement  $(\bx_i, y_i)$, where $\bx_i$ is the realization of the
+auxiliary variables and $y_i$ is the realization of the study variable of
+interest for element  $i$ in the finite population. The auxiliary variables are
+cheap to measure but the measurement for the study variable is expensive.
+Two-phase sampling is a popular method for cost effective sampling.
+In the two-phase sampling, the first phase sample $A_1$  is selected from the
+target population to obtain the measurement of $\bx_i$ in the sample. The
+sampling design for the first-phase sample can be denoted by $P(A_1)$. After
+that, based on the observed auxiliary information in the first-phase  sample,
+the second-phase sample $A_2 \subseteq A_1$ is selected from the first-phase
+sample to observe $y_i$. The sampling design for the second-phase sample is
+denoted as $P( A_2 \mid A_1)$.  
+The
+goal of two-phase sampling is to construct an estimator of $Y_N =  \sum_{i=1}^N y_i$ 
 that uses both the observed information in the Phase 2 sample and also the extra
-auxiliary information from $\bf X$ in the Phase 1 sample.
-The challenge is doing this efficiently.
-
-An easy-to-implement unbiased estimator in the spirit of a Horvitz-Thompson (HT)
-estimator (\cite{horvitz1952generalization}, \cite{narain1951sampling}) is the
-$\pi^*$-estimator. Let $\pi_i^{(2)}$ be the response probability of element $i$
-being observed in the Phase 2 sample. Then, allowing the elements in the Phase 1
-sample to be represented by $A_1$ and the elements in the Phase 2 sample to be
-denoted as $A_2$,
-we define $\pi_{2i | 1} = \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1)$ and
-$\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ then,
-
-$$ \pi_i^{(2)}(A_1) = \pi_{2i | 1} \pi_{1i}.$$
-
-This means that we can define the $\pi^*$-estimator as the following design
-unbiased estimator:
-
-$$ \hat Y_{\pi^*} = \sum_{i \in A_2} \frac{y_i}{\pi_{2i | 1} \pi_{1i}}.$$
-
-While unbiased (see \cite{kim2024statistics}), the $\pi^*$-estimator does not
-account for the additional information contained in the auxiliary Phase 1
-variable $\bf X$. The two-phase regression estimator $\hat Y_{reg, tp}$ does
-incorporate information for $\bf X$ by using the estimate $\hat{\bf X_1}$ from
-the Phase 1 sample. This is one way we can leverage the external information of 
-$\hat{\bf{X_1}}$ to improve the initial $\pi^*$-estimator in the second phase
-sample. The two-phase regression estimator has the form,
-
-$$ 
-\hat Y_{reg, tp} 
-= \sum_{i \in A_1} \frac{1}{\pi_{1i}} \bf x_i \hat{\bm \beta_q} + 
-\sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i \hat{\bm \beta_q})
-$$
-
-where for $q_i = q(\bf x_i)$ and is a function of $\bf x_i$,
-
+auxiliary information from $\bf X$ in the Phase 1 sample. How to combine
+information efficiently is an important problem in two-phase sampling.
+
+Let $\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ be the first-order inclusion
+probability for the first-phase sample and let $\pi_{2i | 1} = \sum_{A_2: i \in
+A_2} \Pr(A_2 \mid A_1)$ be the first-order inclusion probability for the
+second-phase sample. From the first-phase sample, we can compute
+$\widehat{\mathbf{X}}_1 = \sum_{i \in A_1} \pi_{1i}^{-1} \bx_i$ as an unbiased
+estimator of $\mathbf{X}_N = \sum_{i=1}^N  \bx_i$. Also, from the second-phase
+sample, we can compute $\left( \widehat{\mathbf{X}}_2 , \hat{Y}_2 \right) =
+\sum_{i \in A_2} \pi_{1i}^{-1} \pi_{2i \mid 1}^{-1} \left( \bx_i, y_i\right)$ as
+an unbiased estimator of $(\mathbf{X}_N, Y_N)$.
+We therefore have three estimators for
+two parameters. The two-phase regression estimator can be  used to combine this
+information efficiently and it can be expressed as  
+\begin{equation} \label{eq:1}
+\hat Y_{\rm reg, tp} 
+= \sum_{i \in A_1} \frac{1}{\pi_{1i}} \bf x_i^\top  \hat{\bm \beta_q} + 
+\sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}} \left(y_i - \bf x_i^\top  \hat{\bm \beta_q}\right)
+\end{equation} 
+where $q_i = q(\bf x_i)$  is a function of $\bf x_i$ and  
 $$
 \hat{\bm \beta_q} = \left(\sum_{i \in A_2} 
-  \frac{\bf x_i \bf x_i'}{\pi_{1i} q_i}\right)^{-1} 
-\sum_{i \in A_2} \frac{\bf x_i y_i}{\pi_{1i} q_i}.
-$$
-
-The regression estimator is the minimum variance design consistent linear
-estimator which is easily shown to be the case because $\hat Y_{reg, tp} =
+ \frac{1}{\pi_{1i} q_i } \bf x_i \bf x_i^\top \right)^{-1} 
+\sum_{i \in A_2} \frac{1}{\pi_{1i} q_i} \bf x_i y_i .
+$$ 
+Regarding the choice of $q_i$, one can consider a linear regression model 
+\begin{equation}
+y_i = \bx_i^\top \bm \beta + e_i 
+\end{equation}
+where $e_i \sim \left(0, \sigma^2 c_i \right)$, where $c_i=c( \bx_i)$ is a known
+function. In this case,   $q_i=c_i$ is a reasonable choice.
+
+One way to understand the regression estimator is to use the following decomposition:
+
+$$ 
+\hat Y_{\rm reg, tp} 
+= \underbrace{\sum_{i \in A_1} \frac{\bf x_i \hat{\bm \beta_q}}{\pi_{1i}}}_{
+  \text{ Prediction  }} + \underbrace{\sum_{i \in A_2}
+\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i \hat{\bm \beta_q})}_{
+\text{Bias correction}}.
+$$
+The prediction term is obtained by finding the best linear unbiased predictor
+under the working superpopulation model. The bias correction term 
+makes the two-phase regression estimator asymptotically design unbiased
+regardless of the underlying the superpopulation model. 
+Note that the two-phase regression estimator in (\ref{eq:1})  can be
+expressed as  $\hat Y_{\rm reg, tp} =
 \sum_{i \in A_2} \hat w_{2i} y_i / \pi_{1i}$ where
@@
 $$
-\hat w_{2i} = \argmin_{w} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i
+\hat w_{2} = \argmin_{w_2} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i
 \text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i / \pi_{1i} = \sum_{i \in A_1}
@@
 
-Furthermore, the regression estimator
-has a nice feature that its two terms can be thought about as minimizing the
-variance and bias correction,
-
-$$ 
-\hat Y_{reg, tp} 
-= \underbrace{\sum_{i \in A_1} \frac{\bf x_i \hat{\bm \beta_q}}{\pi_{1i}}}_{
-  \text{ Minimizing the variance}} + \underbrace{\sum_{i \in A_2}
-\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i \hat{\bm \beta_q})}_{
-\text{Bias correction}}.
-$$
-
 The \cite{deville1992calibration} method incorporates the design weights into
 the loss function, which is the part minimizing the variance. We would rather
-separate have bias calibration separate from the minimizing the variance so that
-we can control each in isolation. In
-\cite{kwon2024debiased}, the authors show that for a generalized entropy
+have bias calibration separate from the minimizing the variance so that
+we can control each in isolation. 
+\cite{kwon2024debiased} shows that for a generalized entropy
 function $G(w)$, including a term of $g(\pi_{2i|1}^{-1})$ into the calibration
@@
 
-However, the method of \cite{kwon2024debiased} requires known finite population 
+However, the method of \cite{kwon2024debiased} assumes known finite population 
 calibration levels. It does not handle the
 two-phase setup where we need to estimate the finite population total of $\bf x$
-from the Phase 1 sample. In the rest of the section, we extend this method to 
+from the Phase 1 sample. In the rest of this paper, we extend their method to 
 two-phase sampling so that we construct asymptotically valid confidence
-intervals when using estimate finite population totals derived from the Phase 1
+intervals when using estimated finite population totals derived from the Phase 1
 sample.
 
-\subsection{Methodology}
+To construct asymptotically unbiased estimate of the variance we assume that the
+sampling designs for $A_1$ and $A_2$ are measurable and we define the following
+joint inclusion probabilities:
+
+$\Pr(i \in A_1, j \in A_1) = \pi_{1ij}, \text{ and }
+\Pr(i \in A_2, j \in A_2 \mid i \in A_1, j \in A_1) = \pi_{2ij|1}.$
+
+\section{Main Proposal}
 
 We follow the approach of \cite{kwon2024debiased} for the debiased calibration
-method. We consider maximizing a generalized entropy function
-\cite{gneiting2007strictly},
+method.  Let $G: \mathcal{V} \to \R$ be the generalized entropy function that is
+strictly convex and differentiable. We consider maximizing
 
@@
 
-where $G: \mathcal{V} \to \R$ is strictly convex, differentiable function
-subject to the constraints:
+subject to the following constraints:
 
 \begin{equation}\label{eq:calconst1}
-  \sum_{i \in A_2} \frac{\bf x_i w_{2i}q_i}{\pi_{1i}} = 
-\sum_{i \in A_1} \frac{\bf x_iq_i}{\pi_{1i}}
+  \sum_{i \in A_2} \frac{\bf x_i w_{2i}}{\pi_{1i}} = 
+\sum_{i \in A_1} \frac{\bf x_i}{\pi_{1i}}
 \end{equation}
@@
 \begin{equation}\label{eq:primal}
-  \argmin_{w_{2|1}} \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
+  \hat w_2  = \argmin_{w_{2}} \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
   \text{ such that}
-  \sum_{i \in A_2} w_{1i} w_{2i|1} \bf z_i q_i = \sum_{i \in A_1} w_{1i} \bf z_i q_i
-\end{equation}
-
-where  $\bf z_i = (\bf x_i / q_i, g(\pi_{2i|1}^{-1}))$. 
-Let $\hat w_{2i|1}$ be the solution to Equation~\eqref{eq:primal}. The resulting 
-estimator of $Y_N$ is $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i|1} y_i$.
-We can get an estimator using the method of Lagrange multipliers. We need to
+  \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i = \sum_{i \in A_1} w_{1i} \bf z_i q_i
+\end{equation}
+
+where  $\bf z_i = (\bf x_i / q_i, g(\pi_{2i|1}^{-1}))$. The resulting 
+estimator of $Y_N$ is
+
+\begin{equation}\label{eq:tpest}
+\hat Y_{\rm DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i} y_i.
+\end{equation}
+
+\subsection{Theoretical Results}
+
+We can understand the estimated weights, $\hat w_{2i}$ from
+Equation~\ref{eq:primal} using the method of Lagrange multipliers. We need to
 minimize the Lagrangian function
@@
 \begin{equation}\label{eq:legragedc1}
-  L(w_{2i|1}, \bm \lambda) = \sum_{i \in A_2} w_{1i} G(w_{2i|1}) q_i 
+  L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
   + \bm \lambda \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
-    \sum_{i \in A_2} w_{1i} w_{2i|1} \bf z_i q_i\right)
+    \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i\right)
 \end{equation}
@@
 where $\bm \lambda$ is a vector of Lagrange multipliers.
-Differentiating with respect to $w_{2i|1}$ and setting this expression equal to
-zero, yields the fact that $\hat w_{2i|1}$ satisfies
-
-$$ \hat w_{2i|1}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda}^T \bf z_i) $$
+Differentiating with respect to $w_{2i}$ and setting this expression equal to
+zero, yields the fact that $\hat w_{2i}$ satisfies
+
+$$ \hat w_{2i} = g^{-1}(\hat{\bm \lambda}^T \bf z_i) $$
 
@@
   \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
-  \sum_{i \in A_2} w_{1i} w_{2i|1}(\hat{\bm \lambda}) \bf z_i q_i\right) = 0.
-\end{equation}
-
-\subsection{Theoretical Results}
-
-The first result that we show is that the construction of the weights in
-Equation~\ref{eq:primal} leads to a design consistent estimator $\hat Y$.
+  \sum_{i \in A_2} w_{1i} w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right) = 0.
+\end{equation}
+
+The first result we show is that the construction of the weights in
+Equation~\ref{eq:primal} leads to a design consistent estimator 
+$\hat Y_{\rm DCE}$ in Equation~\ref{eq:tpest}.
 
@@
 
-  $$\hat Y_{DCE} = \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + O_p(N / n_2)$$
+  $$
+  \hat Y_{\rm DCE} = \hat Y_\ell(\hat{\bm{\lambda}}, \bm \phi^*) + O_p(N / n_2)
+  $$
 
@@
   $$
-  \hat Y_{\ell}(\bm \lambda, \bm \phi^*) = \hat Y_{DCE}(\hat{\bm \lambda}) + 
-  \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(
-  \bm \lambda) z_i q_i\right)\bm \phi^*
+  \hat Y_{\ell}(\hat{\bm \lambda}, \bm \phi^*) = 
+  \sum_{i \in A_2} w_{1i} \hat w_{2i} y_i
+  + \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - 
+    \sum_{i \in A_2} w_{1i} \hat w_{2i} \bf{z}_i q_i\right)\bm \phi^*,
   $$
@@
 
-   \begin{eqnarray*}
-   \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) &=&   \hat{Y}_{\pi^*} + 
-  \left(\sum_{i \in A_1} w_{1i} \bf x_i  -  \sum_{i \in A_2} w_{1i} \pi_{2i \mid
-  1}^{-1} \bf x_i  \right)^T \bm \phi_1^* + \left(\sum_{i \in A_1} w_{1i} g_i  -
-  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^T \phi_2^*  
-  \end{eqnarray*}
+%   \begin{align*}
+%   \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) 
+%   &= \hat{Y}_{\rm DCE} + \left(\sum_{i \in A_1} w_{1i} \bf x_i  -
+%     \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \bf x_i  \right)^T \bm \phi_1^* \\
+%   &\quad+ \left(\sum_{i \in A_1} w_{1i} g_i  -
+%  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^T \phi_2^*  
+%  \end{align*} 
+%\textcolor{red}{(You introduced $\hat{Y}_{\ell}$ twice!)}
+%  and  
+%
+%  $$
+%  \begin{pmatrix}
+%  \bm \phi_1^* \\
+%  \phi_2^* 
+%  \end{pmatrix}
+%  = \left[ \sum_{i \in U} \frac{\pi_{2i \mid 1} }{ g'(d_{2i|1}) q_i} 
+%  \begin{pmatrix}
+%  \bx_i \bx_i^T &   \bx_i g_i   \\
+%  g_i  \bx_i^T   & g_i^2     
+%  \end{pmatrix} \right]^{-1}
+%  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} 
+%  \begin{pmatrix}
+%    \bx_i \\ g_i 
+%  \end{pmatrix}y_i 
+%  $$
+
+% with $g_i = g( \pi_{2i |1}^{-1}) q_i$.
+\end{theorem}
+
+See the appendix for the proof of Theorem 1. Notice that this theorem implies
+that $\hat Y_{\ell}(\hat{\bm \lambda}, \bm \phi^*)$ and $\hat Y_{\rm DCE}$ have
+the same asymptotic expectation and variance, which means that we can estimate
+the variance of $\hat Y_{\ell}(\hat{\bm \lambda}, \bm \phi^*)$ to get an
+estimate of the variance of $\hat Y_{\rm DCE}$. To estimate $\bm \phi^*$, we can
+use
+
+$$
+\hat{\bm \phi} = 
+\left[\sum_{i \in A_2} 
+w_{1i}\frac{\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
+\sum_{i \in A_2} w_{1i}\frac{\bf z_i y_i}{g'(d_{2i|1})}.
+$$
+
+The next theorem discusses how to compute an estimate for the variance of 
+$\hat Y_{\rm DCE}$. For the proof see the appendix.
+
+\begin{theorem}[Variance Estimation]\label{thm:mainvar}
+
+  Define $\hat \eta_i = \bf z_i q_i \hat{\bm \phi} + 
+  \frac{\delta_i}{\pi_{2i|1}}(y_i - \bf z_i q_i \hat{\bm \phi})$.
+  We can construct an unbiased estimate the variance of $\hat Y_{\rm DCE}$ with
+
+  $$
+  \hat V_{\rm DCE} 
+  = \sum_{i \in A_1} \sum_{j \in A_1} \frac{\Delta_{1ij}}{\pi_{1ij}} 
+  \hat \eta_i \hat \eta_j
+  + \sum_{i \in A_2} \sum_{j \in A_2} w_{1i} \frac{\Delta_{2ij|1}}{\pi_{2ij|1}} 
+  \frac{(y_i - \bf z_i q_i \hat{\bm \phi})}{\pi_{2i|1}}
+  \frac{(y_j - \bf z_j q_j \hat{\bm \phi})}{\pi_{2j|1}}
+  $$
+
+  where $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ and 
+  $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$.
+
+\end{theorem}
+
+\section{Extensions}
+
+\subsection{Non-nested Two-Phase Sampling}
+
+Now we consider the sampling mechanism known as non-nested two-phase sampling 
+(\cite{hidiroglou2001double}). In the last section, we considered two-phase sampling
+in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
+two-phase sampling the Phase 2 sample is independent of the Phase 1 sample. 
+Like traditional two-phase sampling, we consider the Phase 1 sample, $A_1$, to
+consist of observations of $(\bf X_i)_{i = 1}^{n_1}$ and the Phase 2 sample,
+$A_2$, to consist of observations of $(\bf X_i, Y_i)_{i = 1}^{n_2}$.
+
+Whereas the classical two-phase estimator uses a single Horvitz-Thompson
+estimator of the Phase 1 sample to construct estimates for calibration totals,
+in the non-nested two-phase sample we have two independent Horvitz-Thompson
+estimators of the total of $\bf X_N$,
+
+$$\hat{\bf X}_1 = \sum_{i \in A_1}^{n_1} d_{1i} \bf x_i \text{ and } 
+\hat{\bf X}_2 = \sum_{i \in A_2}^{n_2} d_{2i} \bf x_i $$
+
+where $d_{1i} = \pi_{1i}^{-1}$, $d_{2i} = \pi_{2i}^{-1}$, 
+$\pi_{1i} = \Pr(i \in A_1)$ and $\pi_{2i} = \Pr(i \in A_2)$. 
+We can combine these estimates using the effective sample size 
+(\cite{kish1965survey}) to get
+$\hat{\bf X}_c = (n_{1, \rm{eff}} \hat{\bf X}_1 + n_{2, \rm{eff}}\hat{\bf X}_2) / 
+(n_{1, \rm{eff}} + n_{2, \rm{eff}})$ where $n_{1, \rm{eff}}$ and 
+$n_{2, \rm{eff}}$ are the effective sample size for $A_1$ and $A_2$ respectively.
+Then we can define a regression estimator as
+
+$$
+\hat Y_{\rm NN, reg} = \hat Y_2 + (\hat{\bf X}_c - \hat{\bf X}_2)^T \hat{\bm \beta_q} = 
+\hat Y_2 + (\hat{\bf X}_1 - \hat{\bf X}_2)^T W\hat{\bm \beta_q}
+$$
+
+where, $W = n_{1, \rm{eff}} / (n_{1, \rm{eff}} + n_{2, \rm{eff}})$, and
+
+$$
+\hat{\bm \beta_q} = 
+\left(\sum_{i \in A_2} \frac{\bf x_i \bf x_i^T}{q_i}\right)^{-1} 
+\sum_{i \in A_2} \frac{\bf x_i y_i}{q_i} \text{ and }
+\hat Y_2 = \sum_{i \in A_2} d_{2i} y_i.
+$$
+
+From an optimality perspective the choice of using the effective sample size to
+weight the estimates from $A_1$ and $A_2$ is reasonable because the
+effective sample sizes are often proportional to the variance of the estimates 
+of $\bf X_N$. While the inverse variance weighted estimate is optimal
+to combine independent samples for a linear estimate, using the effective sample
+size approximates this procedure without requiring the actual variance of an
+estimator to be known. Since the samples $A_1$ and $A_2$ are independent,
+
+$$
+V(\hat Y_{\rm NN, reg}) = 
+V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i - \bf x_i^TW\bm \beta^*_q)\right)
++ (\bm \beta^*_q)^T W^T V(\hat{\bf X}_1) W \bm \beta_q^*
+$$
+
+where $\bm \beta_q^*$ is the probability limit of $\hat{\bm \beta_q}$. Like the
+two-phase sample this regression estimator can be viewed as the solution to the
+following calibration equation
+
+\begin{equation}\label{eq:nncal}
+  \hat w_2 = \argmin_{w_2} \sum_{i \in A_2} (w_{2i} - d_{2i})^2 q_i 
+  \text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X}_c
+\end{equation}
+
+and $\hat Y_{\rm NN, reg} = \sum_{i \in A_2} \hat w_{2i} y_i$ where $\hat
+w_{2i}$ is the solution to Equation~\ref{eq:nncal}.
+
+We can extend the debiased calibration estimator of \cite{kwon2024debiased} to
+the non-nested two-phase sampling case where we use a combined estimate
+$\hat{\bm X_c}$ as the calibration totals instead of using the true totals from
+the finite population.
+
+%\subsubsection{Theoretical Results}
+
+The methodology for the non-nested two-phase sample is very similar to the setup
+described as part of Topic 1. Given a strictly convex differentiable function,
+$G: \mathcal{V} \to \mathbb{R}$, the goal is to solve
+
+\begin{equation}\label{eq:nnopt}
+\hat w_2 = \argmin_{w_2} \sum_{i \in A_2} G\left(w_{2i}\right) q_i 
+\text{ such that } 
+\sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X}_{c} \text{ and } 
+\sum_{i \in A_2} w_{2i} g(d_{2i}) q_i = \sum_{i \in U} g(d_{2i}) q_i
+\end{equation}
+
+for $g(x) = G'(x)$ and a known choice of $q_i \in \mathbb{R}$. 
+The difference
+between solving Equation~\ref{eq:nnopt} and Equation~\ref{eq:primal} is that the
+estimator $\hat{\bf X}_c$ is estimated from the combined sample $A_c = A_1 \cup
+A_2$. Before using $\hat{\bf X}_c$ in the debiased calibration estimator, we
+need to estimate it from the non-nested samples. We can get multiple estimates
+of $\bf X_N$,
+
+$$
+\hat{\bf X}_1 = \sum_{i \in A_1} d_{1i} \bf x_i \text{ and }
+\hat{\bf X}_2 = \sum_{i \in A_2} d_{2i} \bf x_i.
+$$
+
+Let $n_{1, \rm{eff}}$ and $n_{2, \rm{eff}}$ be the effective samples sizes for
+$A_1$ and $A_2$ respectively. Then the optimal combined estimate is
+
+$$
+\hat{\bf X}_c = (n_{1, \rm{eff}} \hat{\bf X}_1 + n_{2, \rm{eff}}\hat{\bf X}_2) / 
+(n_{1, \rm{eff}} + n_{2, \rm{eff}})
+$$
+
+We can construct a non-nested two-phase estimator $\hat Y_{\rm NNE}$ for $Y_N$ where
+$\hat Y_{\rm NNE} = \sum_{i \in A_2} \hat w_{2i} y_i$ and $\hat w_{2i}$ solves
+Equation~\ref{eq:nnopt}. Like the classical two-phase approach, to solve this
+setup we minimize the Lagrangian,
+
+\begin{equation}\label{eq:legragedc2}
+  L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} G(w_{2i}) q_i + 
+  \bm \lambda \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i} \bf z_i q_i\right).
+\end{equation}
+
+where $\bm \lambda$ are the Lagrange multipliers, with
+
+$$\hat{\bf T} = 
+\begin{bmatrix}
+\hat{\bf X}_c \\ \sum_{i \in U} g(d_{2i}) q_i
+\end{bmatrix}.
+$$
+
+Differentiating with respect to $w_{2i}$ and setting this expression equal to
+zero, yields the fact that $\hat w_{2i}$ satisfies
+
+$$ \hat w_{2i}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda^T} \bf z_i) $$
+
+where $\hat{\bm \lambda}$ is the solution to
+
+\begin{equation}\label{eq:lamdc2}
+  \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i}(\hat{\bm \lambda}) \bf z_i
+  q_i\right) = 0.
+\end{equation}
+
+Like the previous section we first prove that the estimator is design consistent.
+Then we provide a result showing how to estimate the variance.
+
+\begin{theorem}[Design Consistency]\label{thm:dc2}
+  Allowing $\bm \lambda^*$ to be the probability limit of $\hat{\bm \lambda}$,
+  under some regularity conditions, 
+  $\hat Y_{\rm NNE} = \hat Y_{\ell, \rm NNE}(\bm \lambda^*,
+  \bm \phi^*) + O_p(Nn_2^{-1})$ where
+
+  $$
+  \hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*) = \sum_{i \in A_2} \hat w_{2i}(
+  \bm \lambda^*) y_i + \left(\hat{\bf T} - \sum_{i \in A_2} \hat w_{2i}(
+  \bm \lambda^*) \bf z_i q_i\right) \bm \phi^*
+  $$
 
@@
   $$
-  \begin{pmatrix}
-  \bm \phi_1^* \\
-  \phi_2^* 
-  \end{pmatrix}
-  = \left[ \sum_{i \in U} \frac{\pi_{2i \mid 1} }{ g'(d_{2i|1}) q_i} 
-  \begin{pmatrix}
-  \bx_i \bx_i^T &   \bx_i g_i   \\
-  g_i  \bx_i^T   & g_i^2     
-  \end{pmatrix} \right]^{-1}
-  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} 
-  \begin{pmatrix}
-    \bx_i \\ g_i 
-  \end{pmatrix}y_i 
-  $$
-
-with $g_i = g( \pi_{2i |1}^{-1}) q_i$.
+  \bm \phi^* =
+  \left(\sum_{i \in U} \frac{\pi_{2i} q_i}{g'(d_{2i})} 
+  \begin{bmatrix}
+    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
+    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \end{bmatrix}
+  \right)^{-1}
+  \sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})} 
+  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
 \end{theorem}
 
-\begin{proof}
-  In this proof, we derive the solution to Equation~\ref{eq:primal} and show
-  that it is asymptotically equivalent to a regression estimator. Using the
-  method of Lagrange multipliers, to solve Equation~\eqref{eq:primal} we need to
-  minimize the Lagrangian in Equation~\eqref{eq:legragedc1}. 
-  The first order conditions show that
-
-  $$
-  \frac{\partial \mathcal{L}}{\partial w_{2i|1}}: g(w_{2i|1}) w_{1i}q_i -
-  w_{1i} \bm \lambda \bf z_i q_i = 0.
-  $$
-
-  Hence, $\hat w_{2i}(\bm \lambda) = g^{-1}(\bm \lambda^T \bf z_i)$ and
-  $\hat{\bm \lambda}$ is determined by Equation~\eqref{eq:lamdc1}. When the
-  sample size gets large, we have $\hat w_{2i|1}(\hat{\bm \lambda}) \to
-  d_{2i|1}$ which means that $\hat{\bm \lambda} \to \bm \lambda^*$ where $\bm
-  \lambda^* = (\bf 0, 1)$. Then using the linearization technique of
-  \cite{randles1982asymptotic}, we can construct a regression estimator,
-
-  $$
-  \hat Y_\ell(\hat{\bm \lambda}, \bm \phi)  = \hat Y_{DCE}(\hat{\bm \lambda}) + 
-  \left(\sum_{i \in A_1} w_{1i} \bm z_i q_i - \sum_{i \in A_2} w_{1i} 
-  \hat w_{2i|1}(\hat{\bm \lambda}) \bf z_i q_i\right)\bm \phi.
-  $$
-
-  Notice that $\hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = \hat Y_{DCE}(\hat{\bm
-  \lambda})$ for all $\bm \phi$. We choose $\bm \phi^*$ such that
-
-  $$
-  E\left[\frac{\partial}{\partial \bm \lambda} \hat Y_\ell(\bm \lambda^*,
-  \bm \phi^*)\right]=0.
-  $$
-
-  Using the fact that $g^{-1}(\bm \lambda^* \bf z_i) = g^{-1}(g(d_{2i|1})) = d_{2i|1}$
-  and $(g^{-1})'(x) = 1 / g'(g^{-1}(x))$, we have
+The proof of this result is very similar to the proof the
+Theorem~\ref{thm:dc1}. The biggest difference is that the total for $\bf X_N$ is
+estimated from both samples using $\hat{\bf X}_c$ instead of 
+$\hat{\bf X}_{HT}$ from the Phase 1 sample. 
+Since $\hat Y_{\rm NNE} = \sum_{i \in A_2} \hat w_{2i}(\hat{\bm \lambda})y_i$
+where $\hat{\bm \lambda}$ solves
+
+\begin{equation}
+  \sum_{i \in A_2} \hat w_{2i}(\bm \lambda) q_i
+  \underbrace{
+  \begin{bmatrix} 
+    \bf x_i / q_i \\ g(d_i)
+\end{bmatrix}}_{\bf z_i} = \bf T
+\end{equation}
+
+we have
+
+$$\hat Y_{\ell, \rm NNE} (\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} \hat
+w_{2i}(\hat{\bm \lambda})y_i + \left(\bf T - \sum_{i \in A_2} \hat
+w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right) \bm \phi.$$
+
+If we choose $\bm \phi^*$ such that $E\left[\frac{\partial}{\partial \bm \lambda} 
+  \hat Y_{\ell, NNE}(\bm \lambda^*, \bm \phi^*)\right] = 0$, then
+
+$$\bm \phi^* =
+\begin{bmatrix}
+  \bm \phi^*_1 \\ \bm \phi^*_2
+\end{bmatrix} = 
+\left(\sum_{i \in U} \frac{\pi_{2i} q_i}{g'(d_{2i})} 
+\begin{bmatrix}
+  \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
+  \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
+\end{bmatrix}
+\right)^{-1}
+\sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})} 
+\begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
+$$
+
+Hence, by a Taylor expansion around $\hat{\bm \lambda}$,
+
+$$\hat Y_{\rm NNE}(\hat{\bm \lambda}) = \hat Y_{\ell, \rm NNE}(\bm \lambda^*, 
+\bm \phi^*) + O_p(Nn_2^{-1}).$$
+
+The next theorem shows how to estimate the variance of $\hat Y_{\rm NNE}$.
+The proof of this result is in the appendix.
+
+\begin{theorem}[Variance Estimation]\label{thm:nnve}
+  The variance of $\hat Y_{\rm NNE}$ is
 
   \begin{align*}
-    \bm \phi^*
-    &= E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}
-    \right]^{-1}
-    E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i y_i}{g'(d_{2i|1})}\right]\\
-    &= \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}
-    \right]^{-1}
-    \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i y_i}{g'(d_{2i|1})}\right]\\
+    \Var(\hat Y_{\rm NNE}(\hat \lambda))
+    &= (\bm \phi_1^*)^T \Var(\hat{\bf X}_c) \bm \phi_1^* + 
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
+      (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
+    &\qquad + (1 - W)\bm \phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij}
+    d_{2i} \bf x_i d_{2j}(y_j - \bf z_j \bm \phi^*_1 q_j)\\
   \end{align*}
 
-  Thus, the linearization estimator is
-
-  $$\hat Y_\ell(\bm \lambda^*, \bm \phi^*) = 
-  \sum_{i \in A_1} w_{1i} q_i \bf z_i \bm \phi^* +
-  \sum_{i \in A_2} w_{1i} d_{2i|1} (y_i - q_i \bf z_i \bm \phi^*).$$
-
-  By construction using a Taylor expansion yields,
+  We can estimate the variance using
 
   \begin{align*}
-    \hat Y_{DCE}(\hat{\bm \lambda}) 
-    &= \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + 
-    E\left[\frac{\partial}{\partial \bm \lambda}\hat Y_\ell(\bm \lambda^*,
-    \bm \phi^*)\right](\hat{\bm \lambda} - \bm \lambda^*) + \frac{1}{2}
-    E\left[\frac{\partial}{\partial \bm \lambda^2} \hat Y_{DCE}(\bm \lambda^*)\right] 
-    (\hat{\bm \lambda} - \bm \lambda^*)^2\\
-    &= \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + O(N)O_p(n_2^{-1}).
+    \hat V_{\rm NNE} 
+    &= (\hat{\bm \phi_1})^T \Var(\hat{\bf X}_c) \hat{\bm \phi_1} + 
+    \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
+    (y_i - \bf z_i \hat{\bm \phi} q_i)(y_j - \bf z_j \hat{\bm \phi} q_j) \\
+    &\qquad + (1 - W)\hat{\bm \phi_1} \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}}
+    \frac{\bf x_i}{\pi_{2i}} \frac{(y_j - \bf z_j \hat{\bm \phi_1} q_j)}{\pi_{2j}}
   \end{align*}
 
-  The final equality comes from the fact that 
-  $E\left[\frac{\partial}{\partial \bm \lambda}
-  \hat Y_\ell(\bm \lambda^*, \bm \phi^*)\right] = 0$,
-  $\frac{\partial}{\partial \bm \lambda^2} \hat w_{2i|1}(\bm \lambda^*)$ is
-  bounded and $||\hat{\bm \lambda} - \bm \lambda^*|| = O_p(n_2^{-1/2})$, which
-  proves our result.
-\end{proof}
+  where
+
+  $$\hat{\bm \phi} =
+  \begin{bmatrix}
+    \hat{\bm \phi_1} \\ \hat \phi_2
+  \end{bmatrix} = 
+  \left(\sum_{i \in A_2} \frac{q_i}{g'(d_{2i})} 
+  \begin{bmatrix}
+    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
+    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \end{bmatrix}
+  \right)^{-1}
+  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})} 
+  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
+\end{theorem}
+
+\subsection{Multi-Source Two-Phase Sampling}
+
+When considering non-nested two-phase sampling, we focused on the case of having
+two samples. Now, we incorporate more than two independent samples
+together with a debiasing constraint. We examine the case in which we
+want to estimate $\theta = E[Y]$ where $Y$ is only observed in one sample.
+
+Consider the setup in which we have independent samples $A_0, A_1, \dots, A_M$
+where $Y$ is observed only in $A_0$ but $\bf X$ is observed in all of the
+samples with variables $\bf X^{(0)}$ observed in $A_0$ and variables $\bf X^{(m)}$
+observed in $A_m$ for each $m = 1, \dots, M$. Like the non-nested case, we
+assume that each survey is sampled independently from the same population
+sampling frame.
+
+The traditional multi-source approach (\cite{kim2024statistics}) is to use
+GLS to obtain an optimal estimator of $\bf X_N$ from the
+samples $A_1, \dots, A_M$. Then we can incorporate this information in the
+estimation of $\theta$ by using the following estimate
+
+$$\hat \theta = \sum_{i \in A_0} \hat w_i y_i$$
+
+where
+
+\begin{equation}\label{eq:mstps}
+\hat w = \argmin_w \sum_{i \in A_0} G(w_i)q_i \text{ such that } 
+\sum_{i \in A_0} w_i \bf x_i = \hat{\bf X}_{\rm{GLS}}^{(0)}
+\end{equation}
+
+where $\hat{\bf X}_{\rm{GLS}}^{(0)}$ is the GLS estimate of $\bf X$ for all of
+the $X$-variables measured in sample $A_0$, and $G$ is a generalized entropy
+function. 
+In order to have a design consistent estimator that incorporates information from
+samples $A_0, A_1, \dots, A_M$, we want to add the debiasing constraint,
+
+\begin{equation}\label{eq:msdebconstr}
+\sum_{i \in A_0} w_{i} g(d_i) q_i = \sum_{i \in U} g(d_i) q_i
+\end{equation}
+
+where $g(x) = \partial G(x) / \partial x$ and $d_i$ are the design weights for
+sample $A_0$. This yields the optimization problem:
+
+\begin{equation}\label{eq:dc3}
+  \hat w_0 = \argmin_{w_0} \sum_{i \in A_0} G(w_{0i}) q_i \text{ such that }
+  \widehat{\bf{T}} = \sum_{i \in A_0} w_{0i} \bf z_i.
+\end{equation}
+
+where $\widehat{\bf{T}} = (\widehat{\bf X}_{\rm{GLS}}^{(0)},
+\sum_{i \in U} g(d_i) q_i)^T$,
+and $\bf z_i = ((\bf x_i^{(0)})^T, g(d_i)q_i)^T$. Then
+the estimator is $\hat Y_{\rm MS} = \sum_{i \in A_0} \hat w_{0i} y_i$ and we can
+construct this estimate minimizing the Lagrangian:
+
+$$
+L(w_{0}, \bm \lambda) = \sum_{i \in A_0} G(w_{0i}) q_i + {\bm \lambda}^T
+\left(\widehat{\bf{T}} - \sum_{i \in A_0} w_{0i} \bf z_i\right)
+$$
+
+where $\bm \lambda$ are the Lagrange multipliers from Equation~\ref{eq:dc3}.
+We now show that $\hat Y_{\rm MS}$ is design consistent.
+
+\begin{theorem}[Design Consistency]
+  Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$.
+  Under regularity conditions,
+
+  $$
+  \hat Y_{\rm MS} = \hat Y_{\ell, \rm MS}(\bm \lambda^*, \bm \phi^*) + O_p(N / n_0)
+  $$
+
+  where
+
+  $$
+  \hat Y_{\ell, \rm MS}(\bm \lambda^*, \bm \phi^*) = \sum_{i \in A_0} \hat w_{0i} y_i + \left(
+    \widehat{\bf T} - \sum_{i \in A_0} \hat w_{0i}(\bm \lambda^*) \bf z_i q_i
+  \right)\bm \phi^*
+  $$
+
+  and
+
+  $$
+  \bm \phi^* = 
+  \left[\sum_{i \in U} \frac{\pi_{0i} \bf z_i \bf z_i^T q_i}{g'(d_{0i})}\right]^{-1}
+  \sum_{i \in U} \frac{\pi_{0i} \bf z_i y_i}{g'(d_{0i})}.
+  $$
+\end{theorem}
+
+This proof follows from the proof of Theorem~\ref{thm:dc2}. The only
+difference is that we have $\hat{\bf X}_{\rm GLS}$ instead of $\hat{\bf X}_c$.
+The next theorem gives a method for how to estimate the variance of 
+$\hat Y_{\rm MS}$.
+
+\begin{theorem}[Variance Estimation]
+  Under regularity conditions,
+
+  $$
+  \begin{aligned}
+    V(\hat Y_{\rm MS}) 
+    &= (\bm \phi^{*})^T\Var(\hat{\bf X}_{\rm GLS}^{(0)}) (\bm \phi^{*}) + 
+      \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{0ij}}{\pi_{0i}\pi_{0j}}
+      (y_i - \bf z_i\bm \phi^{*}q_i)
+      (y_j - \bf z_j\bm \phi^{*}q_j). 
+  \end{aligned}
+  $$
+
+  We can estimate the variance with
+
+  $$
+  \begin{aligned}
+    \hat V(\hat Y_{\rm MS}) 
+    &= (\hat{\bm \phi})^T \hat \Var(\hat{\bf X}_{\rm GLS}^{(0)}) (\hat{\bm \phi})
+    + \sum_{i \in A_0} \sum_{j \in A_0} \frac{\Delta_{0ij}}{\pi_{0ij}\pi_{0i}\pi_{0j}}
+      (y_i - \bf z_i \hat{\bm \phi}q_i)
+      (y_j - \bf z_j \hat{\bm \phi}q_j). 
+  \end{aligned}
+  $$
+
+\end{theorem}
+
+This result follows the same argument as Theorem~\eqref{thm:nnve}. The result
+holds because each sample $A_m$ is independent from the other $A_{m'}$.
+
+\section{Simulation Studies}
 
@@
 
-We run a simulation testing the proposed method. In this approach we have the
+We run a simulation testing the main proposed method. In this approach we have the
 following simulation setup:
@@
 X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
-X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
+X_{2i} &\stackrel{ind}{\sim} \rm{Unif}(0, 4) \\
 X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
@@
 being a simple random sample (SRS) and the 
-Phase 2 sampling occuring under Poisson sampling. The Phase 1 sample has
+Phase 2 sampling occurring under Poisson sampling. The Phase 1 sample has
 $n_1 = 1200$ and the Phase 2 sample has an expected size of
@@
   \item[2.] Two-Phase Regression estimator (Reg): 
-    $\hat Y_{reg} = \sum_{i \in A_1} \frac{\bf x_i' \widehat{\bm \beta}}{\pi_{1i}} + 
+    $\hat Y_{\rm reg} = 
+    \sum_{i \in A_1} \frac{\bf x_i' \widehat{\bm \beta}}{\pi_{1i}} + 
     \sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}}(y_i - \bf x_i' \widehat{\bm \beta})$ 
@@
     \argmin_{w_{2|1}} \sum_{i \in A_2} w_{1i} G(w_{2i}) \text{ such that}
-    \sum_{i \in A_2} w_{1i} w_{2i|1} \bf z_i = \sum_{i \in U} \bf z_i.
+    \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i = \sum_{i \in U} \bf z_i.
   \end{equation}
@@
 %        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
-%    & {\scriptsize$ \eta_i = \bf x_i'\bm \beta + \delta_{2i} w_{2i|1}\varepsilon_i,\;
+%    & {\scriptsize$ \eta_i = \bf x_i'\bm \beta + \delta_{2i} w_{2i}\varepsilon_i,\;
 %        \varepsilon_i = (y_i - \bf x_i' \bm \beta)$} \\
@@
 %        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
-%    & {\scriptsize$\eta_i = \bf z_i' \bm \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
+%    & {\scriptsize$\eta_i = \bf z_i' \bm \phi + \delta_{2i}w_{2i}\varepsilon_i,\;
 %        \varepsilon_i = (y_i - \bf z_i'\bm \phi)$} \\
@@
 %        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
-%    & {\scriptsize$\eta_i = \bf z_i' \bm \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
+%    & {\scriptsize$\eta_i = \bf z_i' \bm \phi + \delta_{2i}w_{2i}\varepsilon_i,\;
 %        \varepsilon_i = (y_i - \bf z_i' \bm \phi)$} \\
@@
 
-\section{Topic 2: Non-nested Two-Phase Sampling}
-
-\subsection{Background}
-
-Now we consider the sampling mechanism known as non-nested two-phase sampling 
-(\cite{hidiroglou2001double}). In the last section, we considered two-phase sampling
-in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
-two-phase sampling the Phase 2 sample is independent of the Phase 1 sample. 
-Like traditional two-phase sampling, we consider the Phase 1 sample, $A_1$, to
-consist of observations of $(\bf X_i)_{i = 1}^{n_1}$ and the Phase 2 sample,
-$A_2$, to consist of observations of $(\bf X_i, Y_i)_{i = 1}^{n_2}$.
-
-Whereas the classical two-phase estimator uses a single Horvitz-Thompson
-estimator of the Phase 1 sample to construct estimates for calibration totals,
-in the non-nested two-phase sample we have two independent Horvitz-Thompson
-estimators of the total of $\bf X$,
-
-$$\hat{\bf X_1} = \sum_{i \in A_1}^{n_1} d_{1i} \bf x_i \text{ and } 
-\hat{\bf X_2} = \sum_{i \in A_2}^{n_2} d_{2i} \bf x_i $$
-
-where $d_{1i} = \pi_{1i}^{-1}$, $d_{2i} = \pi_{2i}^{-1}$, $\pi_{1i}$ is the
-probability of $i \in A_1$ and $\pi_{2i} = \Pr(i \in A_2)$. 
-We can combine these estimates using the effective sample size 
-(\cite{kish1965survey}) to get
-$\hat{\bf X_c} = (n_{1, \rm{eff}} \hat{\bf X_1} + n_{2, \rm{eff}}\hat{\bf X_2}) / 
-(n_{1, \rm{eff}} + n_{2, \rm{eff}})$ where $n_{1, \rm{eff}}$ and 
-$n_{2, \rm{eff}}$ are the effective sample size for $A_1$ and $A_2$ respectively.
-Then we can define a regression estimator as
-
-$$
-\hat Y_{NN, reg} = \hat Y_2 + (\hat{\bf X_c} - \hat{\bf X_2})^T \hat{\bm \beta_q} = 
-\hat Y_2 + (\hat{\bf X_1} - \hat{\bf X_2})^T W\hat{\bm \beta_q}
-$$
-
-where, $W = n_{1, \rm{eff}} / (n_{1, \rm{eff}} + n_{2, \rm{eff}})$, and
-
-$$
-\hat{\bm \beta_q} = 
-\left(\sum_{i \in A_2} \frac{\bf x_i \bf x_i^T}{q_i}\right)^{-1} 
-\sum_{i \in A_2} \frac{\bf x_i y_i}{q_i} \text{ and }
-\hat Y_2 = \sum_{i \in A_2} d_{2i} y_i.
-$$
-
-From an optimality perspective the choice of using the effective sample size to
-weight the estimates from $A_1$ and $A_2$ is reasonable because often the
-effective sample sizes are proportional to the variance of the estimates 
-of $\bf X$. While we know that the inverse variance weighted estimate is optimal
-to combine independent samples for a linear estimate, using the effective sample
-size approximates this procedure without requiring the actual variance of an
-estimator to be known.
-
-Since the samples $A_1$ and $A_2$ are independent,
-
-$$
-V(\hat Y_{NN, reg}) = 
-V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i - \bf x_i^TW\bm \beta^*_q)\right)
-+ (\bm \beta^*_q)^T W^T V(\hat{\bf X_1}) W \bm \beta_q^*
-$$
-
-where $\bm \beta_q^*$ is the probability limit of $\hat{\bm \beta_q}$. Like the
-two-phase sample this regression estimator can be viewed as the solution to the
-following calibration equation
-
-\begin{equation}\label{eq:nncal}
-  \hat w_2 = \argmin_{w_2} Q(w_2) = \sum_{i \in A_2} (w_{2i} - d_{2i})^2 q_i 
-  \text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X_c}
-\end{equation}
-
-and $\hat Y_{NN, reg} = \sum_{i \in A_2} \hat w_{2i} y_i$ where $\hat
-w_{2i}$ is the solution to Equation~\ref{eq:nncal}.
-
-We extend the debiased calibration estimator of \cite{kwon2024debiased} to the
-non-nested two-phase sampling case where we use a combined estimate 
-$\hat{\bm X_c}$ as the calibration totals instead of using the true totals from
-the finite population.
-
-\subsection{Methodology}
-
-The methodology for the non-nested two-phase sample is very similar to the setup
-described as part of Topic 1. Given a strictly convex differentiable function,
-$G: \mathcal{V} \to \mathbb{R}$, the goal is to solve
-
-\begin{equation}\label{eq:nnopt}
-\hat w_2 = \argmin_w \sum_{i \in A_2} G\left(w_{2i}\right) q_i 
-\text{ with } 
-\sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X_{c}} \text{ and } 
-\sum_{i \in A_2} w_{2i} g(d_{2i}) q_i = \sum_{i \in U} g(d_{2i}) q_i
-\end{equation}
-
-for $g(x) = G'(x)$ and a known choice of $q_i \in \mathbb{R}$. 
-The difference
-between solving Equation~\ref{eq:nnopt} and Equation~\ref{eq:primal} is that the
-estimator $\hat{\bf X_c}$ is estimated from the combined sample $A_c = A_1 \cup
-A_2$. Before using $\hat{\bf X_c}$ in the debiased calibration estimator, we
-need to estimate it from the non-nested samples. We can get multiple estimates
-of $\hat{\bf X}$,
-
-$$
-\hat{\bf X_1} = \sum_{i \in A_1} d_{1i} \bf x_i \text{ and }
-\hat{\bf X_2} = \sum_{i \in A_2} d_{2i} \bf x_i.
-$$
-
-Let $n_{1, \rm{eff}}$ and $n_{2, \rm{eff}}$ be the effective samples sizes for
-$A_1$ and $A_2$ respectively. Then the optimal combined estimate is
-
-$$
-\hat{\bf X_c} = (n_{1, \rm{eff}} \hat{\bf X_1} + n_{2, \rm{eff}}\hat{\bf X_2}) / 
-(n_{1, \rm{eff}} + n_{2, \rm{eff}})
-$$
-
-We can constuct a non-nested two-phase estimator $\hat Y_{NNE}$ for $Y_N$ where
-$\hat Y_{NNE} = \sum_{i \in A_2} \hat w_{2i} y_i$ and $\hat w_{2i}$ solves
-Equation~\ref{eq:nnopt}. Like the classical two-phase approach, to solve this
-setup we minimize the Lagrangian,
-
-\begin{equation}\label{eq:legragedc2}
-  L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} G(w_{2i}) q_i + 
-  \bm \lambda \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i} \bf z_i q_i\right).
-\end{equation}
-
-where $\bm \lambda$ are the Lagrange multipliers, with
-
-$$\hat{\bf T} = 
-\begin{bmatrix}
-  \hat{\bf X_c} \\ \sum_{i \in U} g(d_{2i}) q_i
-\end{bmatrix}.
-$$
-
-Differntiating with respect to $w_{2i}$ and setting this expression equal to
-zero, yields the fact that $\hat w_{2i}$ satisfies
-
-$$ \hat w_{2i}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda^T} \bf z_i) $$
-
-where $\hat{\bm \lambda}$ is the solution to
-
-\begin{equation}\label{eq:lamdc2}
-  \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i}(\hat{\bm \lambda}) \bf z_i
-  q_i\right) = 0.
-\end{equation}
-
-\subsection{Theoretical Results}
-
-\begin{theorem}[Design Consistency]\label{thm:dc2}
-  Allowing $\bm \lambda^*$ to be the probability limit of $\hat{\bm \lambda}$,
-  under some regularity conditions, $\hat Y_{NNE} = \hat Y_{\ell, NNE}(\bm \lambda^*,
-  \bm \phi^*) + O_p(Nn_2^{-1})$ where
-
-  $$
-  \hat Y_{\ell, NNE}(\bm \lambda^*, \bm \phi^*) = \sum_{i \in A_2} \hat w_{2i}(
-  \bm \lambda^*) + \left(\hat{\bf T} - \sum_{i \in A_2} \hat w_{2i}(
-  \bm \lambda^*) \bf z_i q_i\right) \bm \phi^*
-  $$
-
-  and
-
-  $$
-  \bm \phi^* =
-  \left(\sum_{i \in U} \frac{\pi_{2i} q_i}{g'(d_{2i})} 
-  \begin{bmatrix}
-    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
-    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
-  \end{bmatrix}
-  \right)^{-1}
-  \sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})} 
-  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
-  $$
-\end{theorem}
-
-\begin{proof}
-  The proof of this result is very similar to the proof the
-  Theorem~\ref{thm:dc1}. The biggest difference is that the total for $\bf X$ is
-  estimated from both samples using $\hat{\bf X_c}$ instead of $\hat{\bf
-  X_{HT}}$ from the Phase 1 sample.
-
-  Since $\hat Y_{NNE} = \sum_{i \in A_2} \hat w_{2i}(\hat{\bm \lambda})y_i$ where
-  $\hat{\bm \lambda}$ solves
-
-  \begin{equation}
-    \sum_{i \in A_2} \hat w_{2i}(\bm \lambda) q_i
-    \underbrace{
-    \begin{bmatrix} 
-      \bf x_i / q_i \\ g(d_i)
-  \end{bmatrix}}_{\bf z_i} = \bf T
-  \end{equation}
-
-  we have
-
-  $$\hat Y_{\ell, NNE} (\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} \hat
-  w_{2i}(\hat{\bm \lambda}) + \left(\bf T - \sum_{i \in A_2} \hat
-  w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right) \bm \phi.$$
-
-  If we choose $\bm \phi^*$ such that $E\left[\frac{\partial}{\partial \bm \lambda} 
-    \hat Y_{\ell, NNE}(\bm \lambda^*, \bm \phi^*)\right] = 0$, then
-
-  $$\bm \phi^* =
-  \begin{bmatrix}
-    \bm \phi^*_1 \\ \bm \phi^*_2
-  \end{bmatrix} = 
-  \left(\sum_{i \in U} \frac{\pi_{2i} q_i}{g'(d_{2i})} 
-  \begin{bmatrix}
-    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
-    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
-  \end{bmatrix}
-  \right)^{-1}
-  \sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})} 
-  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
-  $$
-
-  Hence, by a Taylor expansion around $\hat{\bm \lambda}$,
-
-  $$\hat Y_{NNE}(\hat{\bm \lambda}) = \hat Y_{\ell, NNE}(\bm \lambda^*, 
-  \bm \phi^*) + O_p(Nn_2^{-1}).$$
-
-\end{proof}
-
-\begin{theorem}[Variance Estimation]\label{thm:nnve}
-  The variance of $\hat Y_{NNE}$ is
-
-  \begin{align*}
-    \Var(\hat Y_{NNE}(\hat \lambda)
-    &= (\bm \phi_1^*)^T \Var(\hat{\bf X_c}) \bm \phi_1^* + 
-    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
-      (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
-    &\qquad + (1 - W)\bm \phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij}
-    d_{2i} \bf x_i d_{2j}(y_j - \bf z_j \bm \phi^*_1 q_j)\\
-  \end{align*}
-
-  We can estimate the variance using
-
-  \begin{align*}
-    \hat V_{NNE} 
-    &= (\hat{\bm \phi_1})^T \Var(\hat{\bf X_c}) \hat{\bm \phi_1} + 
-    \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
-    (y_i - \bf z_i \hat{\bm \phi} q_i)(y_j - \bf z_j \hat{\bm \phi} q_j) \\
-    &\qquad + (1 - W)\hat{\bm \phi_1} \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}}
-    \frac{\bf x_i}{\pi_{2i}} \frac{(y_j - \bf z_j \hat{\bm \phi_1} q_j)}{\pi_{2j}}
-  \end{align*}
-
-  where
-
-  $$\hat{\bm \phi} =
-  \begin{bmatrix}
-    \hat{\bm \phi_1} \\ \hat \phi_2
-  \end{bmatrix} = 
-  \left(\sum_{i \in A_2} \frac{q_i}{g'(d_{2i})} 
-  \begin{bmatrix}
-    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
-    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
-  \end{bmatrix}
-  \right)^{-1}
-  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})} 
-  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
-  $$
-\end{theorem}
-
-\begin{proof}
-  From Theorem~\ref{thm:dc2}, we know that $\hat Y_{NNE}(\hat \lambda) = 
-  \hat Y_{\ell, NNE}(\bm \lambda^*, \bm \phi^*) + O_p(Nn_2^{-1})$. Hence, the
-  variance of $\hat Y_{NNE}(\hat{\bm \lambda})$ is
-
-  \begin{align*}
-    \Var(\hat Y_{NNE}(\hat{\bm \lambda})) 
-    &= \Var(\hat Y_{\ell, NNE}(\bm \lambda^*, \bm \phi^*) + O_p(Nn_2^{-1})) \\ 
-    &= \Var\left(\sum_{i \in A_2} \hat w_{2i}(\bm \lambda^*) y_i + 
-      \left(\bf T - \sum_{i \in A_2} \hat w_{2i}(\bm \lambda^*) \bf z_i
-    q_i\right)\bm \phi^*\right)\\
-    &= (\bm \phi_1^*)^T \Var(\hat{\bf X_c}) \bm \phi_1^* + 
-    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
-    (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
-    &\qquad+ 2 \Cov\left(\hat{\bf X_c} \bm \phi_1^*, \sum_{i \in A_2} 
-      \frac{(y_i - \bf z_i \bm \phi^* q_i)}{\pi_{2i}}\right) \\
-    &= (\bm \phi_1^*)^T \Var(\hat{\bf X_c}) \bm \phi_1^* + 
-    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
-    (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
-    &\qquad + (1 - W)\bm \phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij} 
-    \frac{x_i}{\pi_{2i}} \frac{(y_j - \bf z_j \bm \phi^*_1 q_j)}{\pi_{2j}}\\
-  \end{align*}
-
-  where the last equality comes from the fact that $\hat{\bf X_c} = W\hat{\bf
-  X_1} + (1 - W) \hat{\bf X_2}$. To have an unbiased estimator of the variance
-  we can use:
-
-  \begin{align*}
-    \hat V_{NNE} 
-    &= (\hat{\bm \phi_1})^T \Var(\hat{\bf X_c}) \hat{\bm \phi_1} + 
-    \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
-    (y_i - \bf z_i \hat{\bm \phi} q_i)(y_j - \bf z_j \hat{\bm \phi} q_j) \\
-    &\qquad + (1 - W)\hat \phi_1 \sum_{i \in A_2} \sum_{j \in A_2} 
-    \frac{\Delta_{2ij}}{\pi_{2ij}} \frac{x_i}{\pi_{2i}} 
-    \frac{(y_j - \bf z_j \hat{\bm \phi_1} q_j)}{\pi_{2j}}
-  \end{align*}
-
-  where
-
-  $$\hat{\bm \phi} =
-  \begin{bmatrix}
-    \hat{\bm \phi_1} \\ \hat \phi_2
-  \end{bmatrix} = 
-  \left(\sum_{i \in A_2} \frac{q_i}{g'(d_{2i})} 
-  \begin{bmatrix}
-    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
-    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
-  \end{bmatrix}
-  \right)^{-1}
-  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})} 
-  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
-  $$
-
-\end{proof}
-
 \subsection{Simulation Study 2}
 
-We run a simulation testing the proposed method. This is very similar to
+We run a simulation testing the non-nested extension. This is very similar to
 Simulation 1. We have the following simulation setup:
@@
 If $\delta = 0$ then there is no model misspecification. However, if $\delta =
-1$, then there is some model misspecification. We estimate the
+1$, then there is some model misspecification. We estimate
 the parameter $\bar Y_N$ with four approaches:
@@
 \begin{itemize}
-  \item[1.] HT-estimator (HT): $\hat Y_{HT} = N^{-1} \sum_{i \in A_2}
+  \item[1.] HT-estimator (HT): $\hat Y_{\rm HT} = N^{-1} \sum_{i \in A_2}
     \frac{y_i}{\pi_{2i}},$
-  \item[2.] Regression estimator (Reg): Let $\hat Y_{NN, reg} = \hat Y_{HT} + (
-    \hat{\bf X_c} - \hat{\bf X_{2, HT}}) \hat{\bm \beta_2}$ where $\hat Y_{HT} =
-    \sum_{i \in A_2} d_{2i} y_i$, $\hat{\bf X_c} = W \hat{\bf X_{1, HT}} + (1 -
-    W)\hat{\bf X_{2, HT}}$, $W = n_{1,\rm{eff}} / (n_{1,\rm{eff}} + n_{2, \rm{eff}})$,
-    $\hat{\bf X_{1, HT}} = \sum_{i \in A_1} d_{1i} \bf x_{i}$, $\hat{\bf X_{2,
-    HT}} = \sum_{i \in A_2} d_{2i} \bf x_{i}$, $\bf x_i = (1, x_{1i}, x_{2i})^T$
+  \item[2.] Regression estimator (Reg): Let $\hat Y_{\rm NN, reg} = \hat Y_{\rm HT} + (
+  \hat{\bf X}_c - \hat{\bf X}_{2, \rm HT}) \hat{\bm \beta_2}$ where $\hat Y_{HT} =
+\sum_{i \in A_2} d_{2i} y_i$, $\hat{\bf X}_c = W \hat{\bf X}_{1, \rm HT} + (1 -
+W)\hat{\bf X}_{2, \rm HT}$, $W = n_{1,\rm{eff}} / (n_{1,\rm{eff}} + n_{2, \rm{eff}})$,
+$\hat{\bf X}_{1, \rm HT} = \sum_{i \in A_1} d_{1i} \bf x_{i}$, $\hat{\bf X}_{2,
+    \rm HT} = \sum_{i \in A_2} d_{2i} \bf x_{i}$, $\bf x_i = (1, x_{1i}, x_{2i})^T$
     and
@@
 
-    Then $\hat{\bar Y}_{NN, reg} = \hat Y_{NN, reg} / N$.
+    Then $\hat{\bar Y}_{\rm NN, reg} = \hat Y_{\rm NN, reg} / N$.
   \item[3.] Debiased Calibration with Population Constraints (EstPop): This 
@@
 
-  \begin{equation*}
-  \hat w_2 = \argmin_w \sum_{i \in A_2} G(w_{2i}) q_i \text{ with } 
+  \begin{align*}
+    \hat w_2 &= \argmin_{w_2} \sum_{i \in A_2} G(w_{2i}) \\
+    &\quad\text{ such that } 
   \sum_{i \in A_2} w_{2i} \bf x_i = \sum_{i \in U} \bf x_i \text{ and } 
-  \sum_{i \in A_2} w_{2i} g(d_{2i}) q_i = \sum_{i \in U} g(d_{2i}) q_i
-  \end{equation*}
+  \sum_{i \in A_2} w_{2i} g(d_{2i}) = \sum_{i \in U} g(d_{2i})
+  \end{align*}
 
@@
 variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$.
-
 We run the simulation $B = 1000$ times for each of these methods and compute the
@@
 
-\section{Topic 3: Multi-Source Two-Phase Sampling}
-
-\subsection{Background}
-
-When considering non-nested two-phase sampling, we focused on the case of having
-two samples. Now, we consider incorporating more than two independent samples
-together with a debiasing constraint. We consider the case in which we
-want to estimate the $\theta = E[Y]$ and $Y$ is only observed in one sample.
-
-Consider the setup in which we have independent samples $A_0, A_1, \dots, A_M$
-where $Y$ is observed only in $A_0$ but $\bf X$ is observed in all of the
-samples with elements $\bf X^{(0)}$ observed in $A_0$ and $\bf X^{(m)}$ observed
-in $A_m$ for each $m = 1, \dots, M$. Like the non-nested case, we assume that
-each survey is sampling independently from the same population sampling frame.
-
-The traditional multi-source approach (\cite{kim2024statistics}) is to use
-generalized least squares (GLS) to obtain an optimal estimator of $X$ from the
-samples $A_1, \dots, A_M$. Then we can incorporate this information in the
-estimation of $\theta$ by using the following estimate
-
-$$\hat \theta = \sum_{i \in A_0} \hat w_i y_i$$
-
-where
-
-\begin{equation}\label{eq:mstps}
-\hat w = \argmin_w \sum_{i \in A_0} G(w_i)q_i \text{ such that } 
-\sum_{i \in A_0} w_i \bf x_i = \hat X_{\rm{GLS}}^{(0)}
-\end{equation}
-
-where $\hat{\bf X}_{\rm{GLS}}^{(0)}$ is the GLS estimate of $X$ for all of the
-$X$-variables measured in sample $A_0$, and $G$ is a generalized entropy
-function.
-
-\subsection{Methodology}
-
-In order to have a design consistent estimator that incorporates information from
-samples $A_0, A_1, \dots, A_M$, we want to add the debiasing constraints,
-
-\begin{equation}\label{eq:msdebconstr}
-\sum_{i \in A_0} w_{i} \bf x_i q_i = \sum_{i \in U} g(d_i) q_i
-\end{equation}
-
-where $g(x) = \partial G(x) / \partial x$ and $d_i$ are the design weights for
-sample $A_0$. This yields the optimization problem:
-
-\begin{equation}\label{eq:dc3}
-  \hat w = \argmin_w \sum_{i \in A_0} G(w_i) q_i \text{ such that }
-  \widehat{\bf{T}} = \sum_{i \in A_0} w_i \bf z_i.
-\end{equation}
-
-where $\widehat{\bf{T}} = (\widehat{\bf X}_{\rm{GLS}}^{(0)},
-\sum_{i \in U} g(d_i) q_i)^T$,
-and $\bf z_i = ((\bf x_i^{(0)})^T, g(d_i)q_i)^T$. Then
-the estimator is $\hat Y_{MS} = \sum_{i \in A_0} w_{i} y_i$ and we can construct
-this estimate minimizing the Lagrangian:
-
-$$
-L(w_{0}, \bm \lambda) = \sum_{i \in A_0} G(w_{0i}) q_i + {\bm \lambda}^T
-\left(\widehat{\bf{T}} - \sum_{i \in A_0} w_i \bf z_i\right)
-$$
-
-where $\bm \lambda$ are the Lagrange multipliers.
-
-\subsection{Theoretical Results}
-
-\begin{theorem}[Design Consistency]
-  Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$.
-  Under regularity conditions,
-
-  $$\hat Y_{MS} = \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) + O_p(N / n_0)$$
-
-  where
-
-  $$
-  \hat Y_\ell(\bm \lambda^*, \bm \phi^*) = \hat Y_{DCE} + \left(
-    \widehat{\bf T} - \sum_{i \in A_0} \hat w_{0i}(\bm \lambda^*) \bf z_i q_i
-  \right)\bm \phi^*
-  $$
-
-  and
-
-  $$
-  \bm \phi^* = 
-  \left[\sum_{i \in U} \frac{\pi_{0i} \bf z_i \bf z_i^T q_i}{g'(d_{0i})}\right]^{-1}
-  \sum_{i \in U} \frac{\pi_{0i} \bf z_i y_i}{g'(d_{0i})}.
-  $$
-\end{theorem}
-
-\begin{proof}
-  This proof follows from the proof of Theorem~\ref{thm:dc2}. The only
-  difference is that we have $\hat X_{\rm GLS}$ instead of $\hat X_c$.
-\end{proof}
-
-\begin{theorem}[Variance Estimation]
-  Under regularity conditions,
-
-  $$
-  \begin{aligned}
-    V(\hat Y_{MS}) 
-    &= (\bm \phi^{*})^T\Var(\hat{\bf X}_{\rm GLS}^{(0)}) (\bm \phi^{*}) + 
-      \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{0ij}}{\pi_{0i}\pi_{0j}}
-      (y_i - \bf z_i\bm \phi^{*}q_i)
-      (y_j - \bf z_j\bm \phi^{*}q_j). 
-  \end{aligned}
-  $$
-
-  We can estimate the variance with
-
-  $$
-  \begin{aligned}
-    \hat V(\hat Y_{MS}) 
-    &= (\hat{\bm \phi})^T \hat \Var(\hat{\bf X}_{\rm GLS}^{(0)}) (\hat{\bm \phi})
-    + \sum_{i \in A_0} \sum_{j \in A_0} \frac{\Delta_{0ij}}{\pi_{0ij}\pi_{0i}\pi_{0j}}
-      (y_i - \bf z_i \hat{\bm \phi}q_i)
-      (y_j - \bf z_j \hat{\bm \phi}q_j). 
-  \end{aligned}
-  $$
-
-\end{theorem}
-
-\begin{proof}
-  This result follows the same argument as Theorem~\eqref{thm:nnve}. The result
-  holds because each sample $A_m$ is independent from the other $A_{m'}$.
-\end{proof}
-
 \subsection{Simulation Study 3}
 
+This simulation study tests the multi-source extension.
 We have the following superpopulation model with $N = 10000$ elements:
@@
 X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
-X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
+X_{2i} &\stackrel{ind}{\sim} \rm{Unif}(0, 4) \\
 X_{3i} &\stackrel{ind}{\sim} N(5, 1) \\
@@
 
-As expected MSPop has the lowest RMSE because it also uses the population
-totals. It seems like MSEst and MSReg are almost equivalent, which is good
-because they are also supposed to be identical. The MSEst estimator outperforms
+As expected EstPop has the lowest RMSE because it also uses the population
+totals. The Est estimator outperforms
 NNReg because it also uses information from $A_2$, even though it implicitly
@@
 
-\section{Real-Data Analysis}
+\section{Real Data Analysis}
 
@@
 65 years and older and select designated populations between the ages of 18 and
-65. However, the vast majority of Medicare patients are 65 years old or older.}
+65.}
 are known to have a higher all-cause mortality from
@@
   \caption{This table shows the data available to us in each data set. A
-  checkmark in a particular column indicates that the data set listed in the
+  check mark in a particular column indicates that the data set listed in the
   left column contains information about the participant characteristic in the
@@
 information about if a participant had received a second vaccine dose and then
-later to include how many total vaccine doses patients had recieved. If the
+later to include how many total vaccine doses patients had received. If the
 participant had received at least one more dose and their answer was missing (because
@@
 \bibliographystyle{apa}
-\bibliography{references}
-
-\section{Appendix A: Estimating the Population Weights}
+\bibliography{references.bib}
+
+\newpage
+
+\section*{Appendix A: Proof of Theorem 1}
+
+In this proof, we derive the solution to Equation~\ref{eq:primal} and show
+that it is asymptotically equivalent to a regression estimator. Using the
+method of Lagrange multipliers, to solve Equation~\eqref{eq:primal} we need to
+minimize the Lagrangian in Equation~\eqref{eq:legragedc1}. 
+The first order conditions show that at the solution,
+
+$$
+\frac{\partial \mathcal{L}}{\partial w_{2i}}: g(w_{2i}) w_{1i}q_i -
+w_{1i} \hat{\bm \lambda}^T \bf z_i q_i = 0.
+$$
+
+Hence, $\hat w_{2i}(\hat{\bm \lambda}) = 
+g^{-1}(\hat{\bm \lambda}^T \bf z_i)$ where
+$\hat{\bm \lambda}$ is determined by Equation~\eqref{eq:lamdc1}. When the
+sample size gets large, we have $\hat w_{2i}(\hat{\bm \lambda}) \to
+d_{2i|1}$ in probability because both terms in Equation~\eqref{eq:lamdc1} must
+become equal by the Central Limit Theorem. Hence,
+$\hat{\bm \lambda} \to \bm \lambda^*$ in probability and 
+$\bm \lambda^* = (\bf 0^T, 1)^T$. 
+Using the linearization technique of
+\cite{randles1982asymptotic}, we can construct a regression estimator,
+
+$$
+\hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = 
+\sum_{i \in A_2} w_{1i} \hat w_{2i} y_i + 
+\left(\sum_{i \in A_1} w_{1i} \bm z_i q_i - \sum_{i \in A_2} w_{1i} 
+\hat w_{2i} \bf z_i q_i\right)\bm \phi.
+$$
+
+Notice that $\hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} w_{1i}
+\hat w_{2i} y_i$ for all $\bm \phi$. We choose $\bm \phi^*$ such that
+
+$$
+E\left[\frac{\partial}{\partial \bm \lambda} \hat Y_\ell(\bm \lambda^*,
+\bm \phi^*)\right]=0.
+$$
+
+Using the fact that $g^{-1}(\bm \lambda^* \bf z_i) = g^{-1}(g(d_{2i|1})) = d_{2i|1}$
+and $(g^{-1})'(x) = 1 / g'(g^{-1}(x))$, we have
+
+\begin{align*}
+  \bm \phi^*
+  &= E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}
+  \right]^{-1}
+  E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i y_i}{g'(d_{2i|1})}\right]\\
+  &= \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}
+  \right]^{-1}
+  \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i y_i}{g'(d_{2i|1})}\right]\\
+\end{align*}
+
+Thus, the linearization estimator is
+
+$$\hat Y_\ell(\bm \lambda^*, \bm \phi^*) = 
+\sum_{i \in A_1} w_{1i} q_i \bf z_i \bm \phi^* +
+\sum_{i \in A_2} w_{1i} d_{2i|1} (y_i - q_i \bf z_i \bm \phi^*).$$
+
+By construction using a Taylor expansion yields,
+
+\begin{align*}
+  \hat Y_{DCE}(\hat{\bm \lambda}) 
+  &= \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + 
+  E\left[\frac{\partial}{\partial \bm \lambda}\hat Y_\ell(\bm \lambda^*,
+  \bm \phi^*)\right](\hat{\bm \lambda} - \bm \lambda^*) + \frac{1}{2}
+  E\left[\frac{\partial}{\partial \bm \lambda^2} \hat Y_{DCE}(\bm \lambda^*)\right] 
+  (\hat{\bm \lambda} - \bm \lambda^*)^2\\
+  &= \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + O(N)O_p(n_2^{-1}).
+\end{align*}
+
+The final equality comes from the fact that 
+$E\left[\frac{\partial}{\partial \bm \lambda}
+\hat Y_\ell(\bm \lambda^*, \bm \phi^*)\right] = 0$,
+$\frac{\partial}{\partial \bm \lambda^2} \hat w_{2i}(\bm \lambda^*)$ is
+bounded and $||\hat{\bm \lambda} - \bm \lambda^*|| = O_p(n_2^{-1/2})$, which
+proves our result.
+
+\section*{Appendix B: Proof of Theorem 2}
+
+\begin{proof}
+  By Theorem~\ref{thm:dc1}, $\hat Y_{\rm DCE}$ is asymptotically equivalent to
+  $\hat Y_{\ell}$. Notice that we can define
+
+  $$
+  \eta_i^* = 
+  \bf z_i q_i \bm \phi^* + \frac{\delta_i}{\pi_{2i|1}}(y_i - \bf z_i q_i \bm \phi^*)
+  $$
+
+  where $\delta_i$ is one if $i \in A_2$ and zero otherwise. Then,
+
+  $$
+  \hat Y_{\ell} = \sum_{i \in A_1} w_{1i} \eta_i^*.
+  $$
+
+  Then using the reverse framework of \cite{fay1992inferences} and 
+  \cite{shao1999variance}, and noting that 
+  $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ and 
+  $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$,
+
+  \begin{align*}
+    \Var(\hat Y_{\ell, \rm{DCE}}) 
+    &= \Var\left(\sum_{i \in A_1} w_{1i} \eta_i^*\right) \\
+    &= \Var\left(E\left[\sum_{i \in A_1} w_{1i} \eta_i^* \mid A_2 \right]\right) 
+    + E\left[\Var\left(\sum_{i \in A_1} w_{1i} \eta_i^* \mid A_2 \right)\right] \\
+    &= \Var\left(\sum_{i \in U} \eta_i^*\right) 
+    + E\left[\sum_{i \in U} \sum_{j \in U} \Delta_{1ij} w_{1i} \eta_i^* w_{1j}
+      \eta_j^*\right]\\
+  \end{align*}
+
+  This means that we can estimate the variance of $\hat Y_{\rm DCE}$ with
+
+  $$
+  \hat V_{\rm DCE} 
+  = \sum_{i \in A_1} \sum_{j \in A_1} \frac{\Delta_{1ij}}{\pi_{1ij}} 
+  \hat \eta_i \hat \eta_j
+  + \sum_{i \in A_2} \sum_{j \in A_2} w_{1i} \frac{\Delta_{2ij|1}}{\pi_{2ij|1}} 
+  \frac{(y_i - \bf z_i q_i \hat{\bm \phi})}{\pi_{2i|1}}
+  \frac{(y_j - \bf z_j q_j \hat{\bm \phi})}{\pi_{2j|1}}.
+  $$
+
+\end{proof}
+
+\section*{Appendix C: Proof of Theorem 4}
+
+\begin{proof}
+  From Theorem~\ref{thm:dc2}, we know that $\hat Y_{\rm NNE}(\hat \lambda) = 
+  \hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*) + O_p(Nn_2^{-1})$. Hence, the
+  variance of $\hat Y_{\rm NNE}(\hat{\bm \lambda})$ is
+
+  \begin{align*}
+    \Var(\hat Y_{\rm NNE}(\hat{\bm \lambda})) 
+    &\doteq \Var(\hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*)) \\ 
+    &= \Var\left(\sum_{i \in A_2} \hat w_{2i}(\bm \lambda^*) y_i + 
+      \left(\bf T - \sum_{i \in A_2} \hat w_{2i}(\bm \lambda^*) \bf z_i
+    q_i\right)\bm \phi^*\right)\\
+    &= (\bm \phi_1^*)^T \Var(\hat{\bf X}_c) \bm \phi_1^* + 
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
+    (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
+    &\qquad+ 2 \Cov\left(\hat{\bf X}_c \bm \phi_1^*, \sum_{i \in A_2} 
+      \frac{(y_i - \bf z_i \bm \phi^* q_i)}{\pi_{2i}}\right) \\
+    &= (\bm \phi_1^*)^T \Var(\hat{\bf X}_c) \bm \phi_1^* + 
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
+    (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
+    &\qquad + (1 - W)\bm \phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij} 
+    \frac{\bf{x}_i}{\pi_{2i}} \frac{(y_j - \bf z_j \bm \phi^*_1 q_j)}{\pi_{2j}}\\
+  \end{align*}
+
+where the last equality comes from the fact that $\hat{\bf X}_c = W\hat{\bf
+X}_1 + (1 - W) \hat{\bf X}_2$. To have an unbiased estimator of the variance
+  we can use:
+
+  \begin{align*}
+    \hat V_{\rm NNE} 
+    &= (\hat{\bm \phi_1})^T \Var(\hat{\bf X_c}) \hat{\bm \phi_1} + 
+    \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
+    (y_i - \bf z_i \hat{\bm \phi} q_i)(y_j - \bf z_j \hat{\bm \phi} q_j) \\
+    &\qquad + (1 - W)\hat{\bm \phi}_1 \sum_{i \in A_2} \sum_{j \in A_2} 
+    \frac{\Delta_{2ij}}{\pi_{2ij}} \frac{\bf{x}_i}{\pi_{2i}} 
+    \frac{(y_j - \bf z_j \hat{\bm \phi_1} q_j)}{\pi_{2j}}
+  \end{align*}
+
+  where
+
+  $$\hat{\bm \phi} =
+  \begin{bmatrix}
+    \hat{\bm \phi_1} \\ \hat \phi_2
+  \end{bmatrix} = 
+  \left(\sum_{i \in A_2} \frac{q_i}{g'(d_{2i})} 
+  \begin{bmatrix}
+    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
+    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \end{bmatrix}
+  \right)^{-1}
+  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})} 
+  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
+
+\end{proof}
+
+\section*{Appendix D: Estimating the Population Weights}
 
@@
 
-\begin{equation}\label{eq:nnalpha}
-  (\hat \alpha, \hat w_2) = 
-  \argmin_{\alpha} \argmin_{w_2} \sum_{i \in A_2} \{G(w_{2i})\} - NK(\alpha),
-  \text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X_c},
+\begin{align*}
+  (\hat \alpha, \hat w_2) 
+  &= 
+  \argmin_{\alpha} \argmin_{w_2} \sum_{i \in A_2} \{G(w_{2i})\} - NK(\alpha), \\
+  &\text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X}_c,
   \text{ and } \sum_{i \in A_2} w_{2i} g(d_{2i}) = N\alpha.
-\end{equation}
+\end{align*}
 

### 2025-01-07 final/paper/paper_ov.tex
Updated final paper. Fixed Dr. Kim's suggestions.

--- 2024-11-12 final/paper/paper.tex
@@
 \title{Debiased Calibration for Generalized Two-Phase Sampling}
-\author{Caleb Leedy, Jae-Kwang Kim}
+\author{Caleb Leedy \and  Jae-Kwang Kim}
 \maketitle
@@
 \end{abstract}
+\newpage
 
@@
 We want to incorporate information from external data sources to reduce the bias
-in our estimates and/or improve the estimator's efficiency. For many problems, the
+in our estimates and improve the estimator's efficiency. For many problems, the
 additional information consists of summary statistics with standard errors. The
@@
 sample. The first phase sample consists of data from multiple sources, and the
-second phase sample contains the existing data. To motivate this setup, we
-take the following approach: first, we consider the classical two-phase
-sampling setup where the second phase sample is a subset of the first phase
-sample; then, we extend this setup to consider non-nested two-phase samples;
-and finally, we consider the more general approach of having multiple data sets.
-
-\section{Topic 1: Classical Two-Phase Sampling}
-
-\subsection{Background}
-
-Consider a finite population of size $N$ containing elements $(\bf X_i, Y_i)$ where
-an initial (Phase 1) sample of size $n_1$ is selected and $\bf X_i$ is observed.
-From the Phase 1 sample of elements, a (Phase 2) sample of size $n_2 < n_1$ is
-selected and $Y_i$ is observed. This is two-phase sampling. (See 
-\cite{fuller2009sampling}, \cite{kim2024statistics} for general references.) The
-goal of two-phase sampling is to construct an estimator of $\bar Y_N$ 
+second phase sample contains the existing data. We first consider the classical
+two-pahse sampling setup where the second phase sample is a subset of the first
+phase sample. Then we extend this setup to consider non-nested samples and
+multiple non-nested samples.
+
+\textcolor{red}{ Introduction is too short. We need to make 4-5 paragraphs. } 
+\begin{enumerate}
+\item Narrow down (like a funnel) to the subject that you want to address. 
+\item Explain the context of the research (and cite a few key papers).
+\item Explain what is proposed in the paper.  
+\item Significance of the research (why this paper should be published)
+\item Outline of the paper 
+\end{enumerate}
+
+\newpage
+
+\section{Basic Setup}
+
+Consider a finite population of size $N$ 
+with measurement  $(\bx_i, y_i)$, where $\bx_i$ is the realization of the auxiliary variables and $y_i$ is the realization of the study variable of interest for element  $i$ in finite population.
+The auxiliary variables are cheap to measure but the measurement for the study variable is expensive.
+Thus,  two-phase sampling is popular as a tool for cost effective sampling technique.
+In the two-phase sampling, the first phase sample $A_1$  is selected from the target population to obtain the measurement of $\bx_i$ in the sample.
+The sampling design for the first-phase sample can be denoted by $P(A_1)$.
+After that, based on the observed auxiliary information in the first-phase  sample, the second-phase sample $A_1$ is selected from the first-phase sample to observe $y_i$.
+The sampling design for the second-phase sample is denoted as $P( A_2 \mid A_1)$.
+The
+goal of two-phase sampling is to construct an estimator of $Y =  \sum_{i=1}^N y_i$ 
 that uses both the observed information in the Phase 2 sample and also the extra
-auxiliary information from $\bf X$ in the Phase 1 sample.
-The challenge is doing this efficiently.
-
-An easy-to-implement unbiased estimator in the spirit of a Horvitz-Thompson (HT)
-estimator (\cite{horvitz1952generalization}, \cite{narain1951sampling}) is the
-$\pi^*$-estimator. Let $\pi_i^{(2)}$ be the response probability of element $i$
-being observed in the Phase 2 sample. Then, allowing the elements in the Phase 1
-sample to be represented by $A_1$ and the elements in the Phase 2 sample to be
-denoted as $A_2$,
-we define $\pi_{2i | 1} = \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1)$ and
-$\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ then,
-
-$$ \pi_i^{(2)}(A_1) = \pi_{2i | 1} \pi_{1i}.$$
-
-This means that we can define the $\pi^*$-estimator as the following design
-unbiased estimator:
-
-$$ \hat Y_{\pi^*} = \sum_{i \in A_2} \frac{y_i}{\pi_{2i | 1} \pi_{1i}}.$$
-
-While unbiased (see \cite{kim2024statistics}), the $\pi^*$-estimator does not
-account for the additional information contained in the auxiliary Phase 1
-variable $\bf X$. The two-phase regression estimator $\hat Y_{reg, tp}$ does
-incorporate information for $\bf X$ by using the estimate $\hat{\bf X_1}$ from
-the Phase 1 sample. This is one way we can leverage the external information of 
-$\hat{\bf{X_1}}$ to improve the initial $\pi^*$-estimator in the second phase
-sample. The two-phase regression estimator has the form,
-
-$$ 
-\hat Y_{reg, tp} 
-= \sum_{i \in A_1} \frac{1}{\pi_{1i}} \bf x_i \hat{\bm \beta_q} + 
-\sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i \hat{\bm \beta_q})
-$$
-
-where for $q_i = q(\bf x_i)$ and is a function of $\bf x_i$,
-
+auxiliary information from $\bf X$ in the Phase 1 sample. How to combine  information efficiently is an important problem in two-phase sampling.
+
+Let $\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ be the first-order inclusion probability for the first-phase sample and let $\pi_{2i | 1} = \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1)$ be the first-order inclusion probability for the second-phase sample.
+From the first-phase sample, we can compute $\widehat{\mathbf{X}}_1 = \sum_{i \in A_1} \pi_{1i}^{-1} \bx_i$ as an unbiased estimator of $\mathbf{X}= \sum_{i=1}^N  \bx_i$.
+Also, from the second-phase sample, we can compute $\left( \widehat{\mathbf{X}}_2 , \widehat{Y}_2 \right) = \sum_{i \in A_2} \pi_{1i}^{-1} \pi_{2i \mid 1}^{-1} \left( \bx_i, y_i\right)$ as an unbiased estimator of $( \mathbf{X}, Y )$.
+Thus, we have three estimators for two parameters.
+The two-phase regression estimator can be  used to combine the information efficiently.
+The two-phase regression estimator can be expressed as
+\begin{equation}
+\hat Y_{\rm reg, tp} 
+= \sum_{i \in A_1} \frac{1}{\pi_{1i}} \bf x_i^\top  \hat{\bm \beta_q} + 
+\sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}} \left(y_i - \bf x_i^\top  \hat{\bm \beta_q}\right)
+\label{eq:1}
+\end{equation} 
+where $q_i = q(\bf x_i)$  is a function of $\bf x_i$ and  
 $$
 \hat{\bm \beta_q} = \left(\sum_{i \in A_2} 
-  \frac{\bf x_i \bf x_i'}{\pi_{1i} q_i}\right)^{-1} 
-\sum_{i \in A_2} \frac{\bf x_i y_i}{\pi_{1i} q_i}.
-$$
-
-The regression estimator is the minimum variance design consistent linear
-estimator which is easily shown to be the case because $\hat Y_{reg, tp} =
+ \frac{1}{\pi_{1i} q_i } \bf x_i \bf x_i^\top \right)^{-1} 
+\sum_{i \in A_2} \frac{1}{\pi_{1i} q_i} \bf x_i y_i .
+$$ 
+Regarding the choice of $q_i$, one can consider a linear regression model 
+\begin{equation}
+y_i = \bx_i^\top \bm \beta + e_i 
+\end{equation}
+where $e_i \sim \left(0, \sigma^2 c_i \right)$, where $c_i=c( \bx_i)$ is a known function. In this case,   $q_i=c_i$ is a reasonable choice.
+
+One way to understand the regression estimator is to use the following decomposition: 
+$$ 
+\hat Y_{\rm reg, tp} 
+= \underbrace{\sum_{i \in A_1} \frac{\bf x_i \hat{\bm \beta_q}}{\pi_{1i}}}_{
+  \text{ Prediction  }} + \underbrace{\sum_{i \in A_2}
+\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i \hat{\bm \beta_q})}_{
+\text{Bias correction}}.
+$$
+The prediction term is obtained by finding the best linear unbiased predictor under the working superpopulation model.
+The bias correction term is used to make  the two-phase regression estimator be asymptotically design unbiased regardless of the underlying the superpopulation model.
+
+%\textcolor{red}{I would move the calibration expression to the next subsection. }
+
+Now, note that the two-phase regression estimator in (\ref{eq:1})  can be expressed as  $\hat Y_{\rm reg, tp} =
 \sum_{i \in A_2} \hat w_{2i} y_i / \pi_{1i}$ where
@@
 $$
-\hat w_{2i} = \argmin_{w} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i
+\hat w_{2i} = \argmin_{w_2} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i
 \text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i / \pi_{1i} = \sum_{i \in A_1}
@@
 
-Furthermore, the regression estimator
-has a nice feature that its two terms can be thought about as minimizing the
-variance and bias correction,
-
-$$ 
-\hat Y_{reg, tp} 
-= \underbrace{\sum_{i \in A_1} \frac{\bf x_i \hat{\bm \beta_q}}{\pi_{1i}}}_{
-  \text{ Minimizing the variance}} + \underbrace{\sum_{i \in A_2}
-\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i \hat{\bm \beta_q})}_{
-\text{Bias correction}}.
-$$
-
 The \cite{deville1992calibration} method incorporates the design weights into
 the loss function, which is the part minimizing the variance. We would rather
-separate have bias calibration separate from the minimizing the variance so that
+have bias calibration separate from the minimizing the variance so that
 we can control each in isolation. In
@@
 
-However, the method of \cite{kwon2024debiased} requires known finite population 
+However, the method of \cite{kwon2024debiased} assumes known finite population 
 calibration levels. It does not handle the
 two-phase setup where we need to estimate the finite population total of $\bf x$
-from the Phase 1 sample. In the rest of the section, we extend this method to 
+from the Phase 1 sample. In the rest of this paper, we extend their method to 
 two-phase sampling so that we construct asymptotically valid confidence
-intervals when using estimate finite population totals derived from the Phase 1
+intervals when using estimated finite population totals derived from the Phase 1
 sample.
 
-\subsection{Methodology}
+\section{Main Proposal}
 
 We follow the approach of \cite{kwon2024debiased} for the debiased calibration
-method. We consider maximizing a generalized entropy function
-\cite{gneiting2007strictly},
-
+method. Let $G: \mathcal{V} \to \R$ be the generalized entropy function that is strictly convex, differentiable.  
+We consider maximizing  
 \begin{equation}\label{eq:primalloss}
@@
 \end{equation} 
-
-where $G: \mathcal{V} \to \R$ is strictly convex, differentiable function
-subject to the constraints:
+subject to the following constraints:
 
@@
 \begin{equation}\label{eq:primal}
-  \argmin_{w_{2|1}} \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
+  \hat w_2  = \argmin_{w_{2}} \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
   \text{ such that}
-  \sum_{i \in A_2} w_{1i} w_{2i|1} \bf z_i q_i = \sum_{i \in A_1} w_{1i} \bf z_i q_i
-\end{equation}
-
-where  $\bf z_i = (\bf x_i / q_i, g(\pi_{2i|1}^{-1}))$. 
-Let $\hat w_{2i|1}$ be the solution to Equation~\eqref{eq:primal}. The resulting 
-estimator of $Y_N$ is $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i|1} y_i$.
-We can get an estimator using the method of Lagrange multipliers. We need to
-minimize the Lagrangian function
+  \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i = \sum_{i \in A_1} w_{1i} \bf z_i q_i
+\end{equation}
+where  $\bf z_i = (\bf x_i / q_i, g(\pi_{2i|1}^{-1}))$. The resulting 
+estimator of $Y_N$ is
+
+\begin{equation}\label{eq:tpest}
+\hat Y_{\rm DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i} y_i.
+\end{equation}
+
+%\subsection{Theoretical Results}
+
+We can understand the estimated weights, $\hat w_{2i}$ from 
+Equation~\ref{eq:primal}
+using the method of Legrange multipliers.
+We need to minimize the Lagrangian function
 
 \begin{equation}\label{eq:legragedc1}
-  L(w_{2i|1}, \bm \lambda) = \sum_{i \in A_2} w_{1i} G(w_{2i|1}) q_i 
+  L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
   + \bm \lambda \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
-    \sum_{i \in A_2} w_{1i} w_{2i|1} \bf z_i q_i\right)
-\end{equation}
-
+    \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i\right)
+\end{equation}
 where $\bm \lambda$ is a vector of Lagrange multipliers.
-Differentiating with respect to $w_{2i|1}$ and setting this expression equal to
-zero, yields the fact that $\hat w_{2i|1}$ satisfies
-
-$$ \hat w_{2i|1}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda}^T \bf z_i) $$
+Differentiating with respect to $w_{2i}$ and setting this expression equal to
+zero, yields the fact that $\hat w_{2i}$ satisfies
+
+$$ \hat w_{2i}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda}^T \bf z_i) $$
 
@@
   \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
-  \sum_{i \in A_2} w_{1i} w_{2i|1}(\hat{\bm \lambda}) \bf z_i q_i\right) = 0.
-\end{equation}
-
-\subsection{Theoretical Results}
+  \sum_{i \in A_2} w_{1i} w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right) = 0.
+\end{equation}
 
 The first result that we show is that the construction of the weights in
-Equation~\ref{eq:primal} leads to a design consistent estimator $\hat Y$.
+Equation~\ref{eq:primal} leads to a design consistent estimator 
+$\hat Y_{\rm DCE}$ in Equation~\ref{eq:tpest}.
 
@@
 
-  $$\hat Y_{DCE} = \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + O_p(N / n_2)$$
+  $$\hat Y_{\rm DCE} = \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + O_p(N / n_2)$$
 
@@
   $$
-  \hat Y_{\ell}(\bm \lambda, \bm \phi^*) = \hat Y_{DCE}(\hat{\bm \lambda}) + 
-  \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(
-  \bm \lambda) z_i q_i\right)\bm \phi^*
-  $$
-
+  \hat Y_{\ell}(\bm \lambda, \bm \phi^*) = \hat Y_{\rm DCE}(\hat{\bm \lambda}) + 
+  \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i}(
+  \hat{\bm \lambda}) \bf z_i q_i\right)\bm \phi^*,
+  $$
+\textcolor{red}{(The above expression is not good.
+For example, the left-hand side is a function of $\bm \lambda$ and the right-hand side is a function of $\hat{\bm \lambda}$.
+Also, since we have $\bm \lambda^*=( \mathbf{0}, 1)$, the above expression can be simplified further.
+That is, there is no need to use $\bm \lambda^*$ in the above expression.
+) }
   and
@@
 
-   \begin{eqnarray*}
-   \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) &=&   \hat{Y}_{\pi^*} + 
-  \left(\sum_{i \in A_1} w_{1i} \bf x_i  -  \sum_{i \in A_2} w_{1i} \pi_{2i \mid
-  1}^{-1} \bf x_i  \right)^T \bm \phi_1^* + \left(\sum_{i \in A_1} w_{1i} g_i  -
+   \begin{align*}
+   \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) 
+   &= \hat{Y}_{\rm DCE} + \left(\sum_{i \in A_1} w_{1i} \bf x_i  -
+     \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \bf x_i  \right)^T \bm \phi_1^* \\
+   &\quad+ \left(\sum_{i \in A_1} w_{1i} g_i  -
   \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^T \phi_2^*  
-  \end{eqnarray*}
-
+  \end{align*} 
+\textcolor{red}{(You introduced $\hat{Y}_{\ell}$ twice!)}
   and
@@
 
+\textcolor{red}{The presentation should be improved further. Since we
+
+}
+
 \begin{proof}
@@
   $$
-  \frac{\partial \mathcal{L}}{\partial w_{2i|1}}: g(w_{2i|1}) w_{1i}q_i -
+  \frac{\partial \mathcal{L}}{\partial w_{2i}}: g(w_{2i}) w_{1i}q_i -
   w_{1i} \bm \lambda \bf z_i q_i = 0.
@@
   $\hat{\bm \lambda}$ is determined by Equation~\eqref{eq:lamdc1}. When the
-  sample size gets large, we have $\hat w_{2i|1}(\hat{\bm \lambda}) \to
+  sample size gets large, we have $\hat w_{2i}(\hat{\bm \lambda}) \to
   d_{2i|1}$ which means that $\hat{\bm \lambda} \to \bm \lambda^*$ where $\bm
-  \lambda^* = (\bf 0, 1)$. Then using the linearization technique of
+  \lambda^* = (\bf 0, 1)$. \textcolor{red}{(You should prove it rather than just saying it. ) }  
+  Then using the linearization technique of
   \cite{randles1982asymptotic}, we can construct a regression estimator,
@@
   $$
-  \hat Y_\ell(\hat{\bm \lambda}, \bm \phi)  = \hat Y_{DCE}(\hat{\bm \lambda}) + 
+  \hat Y_\ell(\hat{\bm \lambda}, \bm \phi)  = \hat Y_{\rm DCE}(\hat{\bm \lambda}) + 
   \left(\sum_{i \in A_1} w_{1i} \bm z_i q_i - \sum_{i \in A_2} w_{1i} 
-  \hat w_{2i|1}(\hat{\bm \lambda}) \bf z_i q_i\right)\bm \phi.
-  $$
-
-  Notice that $\hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = \hat Y_{DCE}(\hat{\bm
+  \hat w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right)\bm \phi.
+  $$
+
+  Notice that $\hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = \hat Y_{\rm DCE}(\hat{\bm
   \lambda})$ for all $\bm \phi$. We choose $\bm \phi^*$ such that
@@
   \hat Y_\ell(\bm \lambda^*, \bm \phi^*)\right] = 0$,
-  $\frac{\partial}{\partial \bm \lambda^2} \hat w_{2i|1}(\bm \lambda^*)$ is
+  $\frac{\partial}{\partial \bm \lambda^2} \hat w_{2i}(\bm \lambda^*)$ is
   bounded and $||\hat{\bm \lambda} - \bm \lambda^*|| = O_p(n_2^{-1/2})$, which
@@
 
+\textcolor{red}{(Please move the proof to Appendix. Also, we need to discuss the implication of the theorem. Specifically, what is the expectation and variance of the linearization term? ) }
+
+To construct asymptotically unbiased estimate of the variance we assume that the
+sampling designs for $A_1$ and $A_2$ are measurable and we define the following
+joint inclusion probabilities:
+
+$\Pr(i \in A_1, j \in A_1) = \pi_{1ij}, \text{ and }
+\Pr(i \in A_2, j \in A_2 \mid i \in A_1, j \in A_1) = \pi_{2ij|1}.$
+
+\begin{theorem}[Variance Estimation]\label{thm:mainvar}
+
+  Define $\hat \eta_i = \bf z_i q_i \hat{\bm \phi} + 
+  \frac{\delta_i}{\pi_{2i|1}}(y_i - \bf z_i q_i \hat{\bm \phi})$.
+  We can construct an unbiased estimate the variance of $\hat Y_{\rm DCE}$ with
+
+  $$
+  \hat V_{\rm DCE} 
+  = \sum_{i \in A_1} \sum_{j \in A_1} \frac{\Delta_{1ij}}{\pi_{1ij}} 
+  \hat \eta_i \hat \eta_j
+  + \sum_{i \in A_2} \sum_{j \in A_2} w_{1i} \frac{\Delta_{2ij|1}}{\pi_{2ij|1}} 
+  \frac{(y_i - \bf z_i q_i \bm \phi^*)}{\pi_{2i|1}}
+  \frac{(y_j - \bf z_j q_j \bm \phi^*)}{\pi_{2j|1}}
+  $$
+
+  where $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ and 
+  $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$.
+
+\end{theorem}
+
+\begin{proof}
+  By Theorem~\ref{thm:dc1}, $\hat Y_{\rm DCE}$ is asymptotically equivalent to
+  $\hat Y_{\ell, \rm{DCE}}$. Notice that we can define
+
+  $$
+  \eta_i^* = 
+  \bf z_i q_i \bm \phi^* + \frac{\delta_i}{\pi_{2i|1}}(y_i - \bf z_i q_i \bm \phi^*)
+  $$
+
+  where $\delta_i$ is one if $i \in A_2$ and zero otherwise. Then,
+
+  $$
+  \hat Y_{\ell, \rm{DCE}} = \sum_{i \in A_1} w_{1i} \eta_i^*.
+  $$
+
+  Then using the reverse framework of \cite{fay1992inferences} and 
+  \cite{shao1999variance}, and noting that 
+  $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ and 
+  $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$,
+
+  \begin{align*}
+    \Var(\hat Y_{\ell, \rm{DCE}}) 
+    &= \Var\left(\sum_{i \in A_1} w_{1i} \eta_i^*\right) \\
+    &= \Var\left(E\left[\sum_{i \in A_1} w_{1i} \eta_i^* \mid A_2 \right]\right) 
+    + E\left[\Var\left(\sum_{i \in A_1} w_{1i} \eta_i^* \mid A_2 \right)\right] \\
+    &= \Var\left(\sum_{i \in U} \eta_i^*\right) 
+    + E\left[\sum_{i \in U} \sum_{j \in U} \Delta_{1ij} w_{1i} \eta_i^* w_{1j}
+      \eta_j^*\right]\\
+  \end{align*}
+
+  This means that we can estimate the variance of $\hat Y_{\rm DCE}$ with
+
+  $$
+  \hat V_{\rm DCE} 
+  = \sum_{i \in A_1} \sum_{j \in A_1} \frac{\Delta_{1ij}}{\pi_{1ij}} 
+  \hat \eta_i \hat \eta_j
+  + \sum_{i \in A_2} \sum_{j \in A_2} w_{1i} \frac{\Delta_{2ij|1}}{\pi_{2ij|1}} 
+  \frac{(y_i - \bf z_i q_i \bm \phi^*)}{\pi_{2i|1}}
+  \frac{(y_j - \bf z_j q_j \bm \phi^*)}{\pi_{2j|1}}
+  $$
+
+  where $\hat \eta_i = \bf z_i q_i \hat{\bm \phi} + 
+  \frac{\delta_i}{\pi_{2i|1}}(y_i - \bf z_i q_i \hat{\bm \phi})$.
+
+\end{proof}
+
+\section{Extensions}
+\subsection{Non-nested Two-Phase Sampling}
+
+Now we consider the sampling mechanism known as non-nested two-phase sampling 
+(\cite{hidiroglou2001double}). In the last section, we considered two-phase sampling
+in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
+two-phase sampling the Phase 2 sample is independent of the Phase 1 sample. 
+Like traditional two-phase sampling, we consider the Phase 1 sample, $A_1$, to
+consist of observations of $(\bf X_i)_{i = 1}^{n_1}$ and the Phase 2 sample,
+$A_2$, to consist of observations of $(\bf X_i, Y_i)_{i = 1}^{n_2}$.
+
+Whereas the classical two-phase estimator uses a single Horvitz-Thompson
+estimator of the Phase 1 sample to construct estimates for calibration totals,
+in the non-nested two-phase sample we have two independent Horvitz-Thompson
+estimators of the total of $\bf X$,
+
+$$\hat{\bf X}_1 = \sum_{i \in A_1}^{n_1} d_{1i} \bf x_i \text{ and } 
+\hat{\bf X}_2 = \sum_{i \in A_2}^{n_2} d_{2i} \bf x_i $$
+
+where $d_{1i} = \pi_{1i}^{-1}$, $d_{2i} = \pi_{2i}^{-1}$, $\pi_{1i}$ is the
+probability of $i \in A_1$ and $\pi_{2i} = \Pr(i \in A_2)$. 
+We can combine these estimates using the effective sample size 
+(\cite{kish1965survey}) to get
+$\hat{\bf X}_c = (n_{1, \rm{eff}} \hat{\bf X}_1 + n_{2, \rm{eff}}\hat{\bf X}_2) / 
+(n_{1, \rm{eff}} + n_{2, \rm{eff}})$ where $n_{1, \rm{eff}}$ and 
+$n_{2, \rm{eff}}$ are the effective sample size for $A_1$ and $A_2$ respectively.
+Then we can define a regression estimator as
+
+$$
+\hat Y_{\rm NN, reg} = \hat Y_2 + (\hat{\bf X}_c - \hat{\bf X}_2)^T \hat{\bm \beta_q} = 
+\hat Y_2 + (\hat{\bf X}_1 - \hat{\bf X}_2)^T W\hat{\bm \beta_q}
+$$
+
+where, $W = n_{1, \rm{eff}} / (n_{1, \rm{eff}} + n_{2, \rm{eff}})$, and
+
+$$
+\hat{\bm \beta_q} = 
+\left(\sum_{i \in A_2} \frac{\bf x_i \bf x_i^T}{q_i}\right)^{-1} 
+\sum_{i \in A_2} \frac{\bf x_i y_i}{q_i} \text{ and }
+\hat Y_2 = \sum_{i \in A_2} d_{2i} y_i.
+$$
+
+From an optimality perspective the choice of using the effective sample size to
+weight the estimates from $A_1$ and $A_2$ is reasonable because often the
+effective sample sizes are proportional to the variance of the estimates 
+of $\bf X$. While we know that the inverse variance weighted estimate is optimal
+to combine independent samples for a linear estimate, using the effective sample
+size approximates this procedure without requiring the actual variance of an
+estimator to be known.
+
+Since the samples $A_1$ and $A_2$ are independent,
+
+$$
+V(\hat Y_{\rm NN, reg}) = 
+V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i - \bf x_i^TW\bm \beta^*_q)\right)
++ (\bm \beta^*_q)^T W^T V(\hat{\bf X}_1) W \bm \beta_q^*
+$$
+
+where $\bm \beta_q^*$ is the probability limit of $\hat{\bm \beta_q}$. Like the
+two-phase sample this regression estimator can be viewed as the solution to the
+following calibration equation
+
+\begin{equation}\label{eq:nncal}
+  \hat w_2 = \argmin_{w_2} Q(w_2) = \sum_{i \in A_2} (w_{2i} - d_{2i})^2 q_i 
+  \text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X}_c
+\end{equation}
+
+and $\hat Y_{\rm NN, reg} = \sum_{i \in A_2} \hat w_{2i} y_i$ where $\hat
+w_{2i}$ is the solution to Equation~\ref{eq:nncal}.
+
+We can extend the debiased calibration estimator of \cite{kwon2024debiased} to
+the non-nested two-phase sampling case where we use a combined estimate
+$\hat{\bm X_c}$ as the calibration totals instead of using the true totals from
+the finite population.
+
+%\subsubsection{Theoretical Results}
+
+The methodology for the non-nested two-phase sample is very similar to the setup
+described as part of Topic 1. Given a strictly convex differentiable function,
+$G: \mathcal{V} \to \mathbb{R}$, the goal is to solve
+
+\begin{equation}\label{eq:nnopt}
+\hat w_2 = \argmin_w \sum_{i \in A_2} G\left(w_{2i}\right) q_i 
+\text{ with } 
+\sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X}_{c} \text{ and } 
+\sum_{i \in A_2} w_{2i} g(d_{2i}) q_i = \sum_{i \in U} g(d_{2i}) q_i
+\end{equation}
+
+for $g(x) = G'(x)$ and a known choice of $q_i \in \mathbb{R}$. 
+The difference
+between solving Equation~\ref{eq:nnopt} and Equation~\ref{eq:primal} is that the
+estimator $\hat{\bf X}_c$ is estimated from the combined sample $A_c = A_1 \cup
+A_2$. Before using $\hat{\bf X}_c$ in the debiased calibration estimator, we
+need to estimate it from the non-nested samples. We can get multiple estimates
+of $\hat{\bf X}$,
+
+$$
+\hat{\bf X}_1 = \sum_{i \in A_1} d_{1i} \bf x_i \text{ and }
+\hat{\bf X}_2 = \sum_{i \in A_2} d_{2i} \bf x_i.
+$$
+
+Let $n_{1, \rm{eff}}$ and $n_{2, \rm{eff}}$ be the effective samples sizes for
+$A_1$ and $A_2$ respectively. Then the optimal combined estimate is
+
+$$
+\hat{\bf X}_c = (n_{1, \rm{eff}} \hat{\bf X}_1 + n_{2, \rm{eff}}\hat{\bf X}_2) / 
+(n_{1, \rm{eff}} + n_{2, \rm{eff}})
+$$
+
+We can constuct a non-nested two-phase estimator $\hat Y_{\rm NNE}$ for $Y_N$ where
+$\hat Y_{\rm NNE} = \sum_{i \in A_2} \hat w_{2i} y_i$ and $\hat w_{2i}$ solves
+Equation~\ref{eq:nnopt}. Like the classical two-phase approach, to solve this
+setup we minimize the Lagrangian,
+
+\begin{equation}\label{eq:legragedc2}
+  L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} G(w_{2i}) q_i + 
+  \bm \lambda \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i} \bf z_i q_i\right).
+\end{equation}
+
+where $\bm \lambda$ are the Lagrange multipliers, with
+
+$$\hat{\bf T} = 
+\begin{bmatrix}
+\hat{\bf X}_c \\ \sum_{i \in U} g(d_{2i}) q_i
+\end{bmatrix}.
+$$
+
+Differentiating with respect to $w_{2i}$ and setting this expression equal to
+zero, yields the fact that $\hat w_{2i}$ satisfies
+
+$$ \hat w_{2i}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda^T} \bf z_i) $$
+
+where $\hat{\bm \lambda}$ is the solution to
+
+\begin{equation}\label{eq:lamdc2}
+  \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i}(\hat{\bm \lambda}) \bf z_i
+  q_i\right) = 0.
+\end{equation}
+
+\begin{theorem}[Design Consistency]\label{thm:dc2}
+  Allowing $\bm \lambda^*$ to be the probability limit of $\hat{\bm \lambda}$,
+  under some regularity conditions, 
+  $\hat Y_{\rm NNE} = \hat Y_{\ell, \rm NNE}(\bm \lambda^*,
+  \bm \phi^*) + O_p(Nn_2^{-1})$ where
+
+  $$
+  \hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*) = \sum_{i \in A_2} \hat w_{2i}(
+  \bm \lambda^*) y_i + \left(\hat{\bf T} - \sum_{i \in A_2} \hat w_{2i}(
+  \bm \lambda^*) \bf z_i q_i\right) \bm \phi^*
+  $$
+
+  and
+
+  $$
+  \bm \phi^* =
+  \left(\sum_{i \in U} \frac{\pi_{2i} q_i}{g'(d_{2i})} 
+  \begin{bmatrix}
+    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
+    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \end{bmatrix}
+  \right)^{-1}
+  \sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})} 
+  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
+\end{theorem}
+
+\begin{proof}
+  The proof of this result is very similar to the proof the
+  Theorem~\ref{thm:dc1}. The biggest difference is that the total for $\bf X$ is
+  estimated from both samples using $\hat{\bf X}_c$ instead of 
+  $\hat{\bf X}_{HT}$ from the Phase 1 sample.
+
+  Since $\hat Y_{\rm NNE} = \sum_{i \in A_2} \hat w_{2i}(\hat{\bm \lambda})y_i$ where
+  $\hat{\bm \lambda}$ solves
+
+  \begin{equation}
+    \sum_{i \in A_2} \hat w_{2i}(\bm \lambda) q_i
+    \underbrace{
+    \begin{bmatrix} 
+      \bf x_i / q_i \\ g(d_i)
+  \end{bmatrix}}_{\bf z_i} = \bf T
+  \end{equation}
+
+  we have
+
+  $$\hat Y_{\ell, \rm NNE} (\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} \hat
+  w_{2i}(\hat{\bm \lambda})y_i + \left(\bf T - \sum_{i \in A_2} \hat
+  w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right) \bm \phi.$$
+
+  If we choose $\bm \phi^*$ such that $E\left[\frac{\partial}{\partial \bm \lambda} 
+    \hat Y_{\ell, NNE}(\bm \lambda^*, \bm \phi^*)\right] = 0$, then
+
+  $$\bm \phi^* =
+  \begin{bmatrix}
+    \bm \phi^*_1 \\ \bm \phi^*_2
+  \end{bmatrix} = 
+  \left(\sum_{i \in U} \frac{\pi_{2i} q_i}{g'(d_{2i})} 
+  \begin{bmatrix}
+    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
+    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \end{bmatrix}
+  \right)^{-1}
+  \sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})} 
+  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
+
+  Hence, by a Taylor expansion around $\hat{\bm \lambda}$,
+
+  $$\hat Y_{\rm NNE}(\hat{\bm \lambda}) = \hat Y_{\ell, \rm NNE}(\bm \lambda^*, 
+  \bm \phi^*) + O_p(Nn_2^{-1}).$$
+
+\end{proof}
+
+\begin{theorem}[Variance Estimation]\label{thm:nnve}
+  The variance of $\hat Y_{\rm NNE}$ is
+
+  \begin{align*}
+    \Var(\hat Y_{\rm NNE}(\hat \lambda))
+    &= (\bm \phi_1^*)^T \Var(\hat{\bf X}_c) \bm \phi_1^* + 
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
+      (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
+    &\qquad + (1 - W)\bm \phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij}
+    d_{2i} \bf x_i d_{2j}(y_j - \bf z_j \bm \phi^*_1 q_j)\\
+  \end{align*}
+
+  We can estimate the variance using
+
+  \begin{align*}
+    \hat V_{\rm NNE} 
+    &= (\hat{\bm \phi_1})^T \Var(\hat{\bf X}_c) \hat{\bm \phi_1} + 
+    \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
+    (y_i - \bf z_i \hat{\bm \phi} q_i)(y_j - \bf z_j \hat{\bm \phi} q_j) \\
+    &\qquad + (1 - W)\hat{\bm \phi_1} \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}}
+    \frac{\bf x_i}{\pi_{2i}} \frac{(y_j - \bf z_j \hat{\bm \phi_1} q_j)}{\pi_{2j}}
+  \end{align*}
+
+  where
+
+  $$\hat{\bm \phi} =
+  \begin{bmatrix}
+    \hat{\bm \phi_1} \\ \hat \phi_2
+  \end{bmatrix} = 
+  \left(\sum_{i \in A_2} \frac{q_i}{g'(d_{2i})} 
+  \begin{bmatrix}
+    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
+    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \end{bmatrix}
+  \right)^{-1}
+  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})} 
+  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
+\end{theorem}
+
+\begin{proof}
+  From Theorem~\ref{thm:dc2}, we know that $\hat Y_{\rm NNE}(\hat \lambda) = 
+  \hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*) + O_p(Nn_2^{-1})$. Hence, the
+  variance of $\hat Y_{\rm NNE}(\hat{\bm \lambda})$ is
+
+  \begin{align*}
+    \Var(\hat Y_{\rm NNE}(\hat{\bm \lambda})) 
+    &\doteq \Var(\hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*)) \\ 
+    &= \Var\left(\sum_{i \in A_2} \hat w_{2i}(\bm \lambda^*) y_i + 
+      \left(\bf T - \sum_{i \in A_2} \hat w_{2i}(\bm \lambda^*) \bf z_i
+    q_i\right)\bm \phi^*\right)\\
+    &= (\bm \phi_1^*)^T \Var(\hat{\bf X}_c) \bm \phi_1^* + 
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
+    (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
+    &\qquad+ 2 \Cov\left(\hat{\bf X}_c \bm \phi_1^*, \sum_{i \in A_2} 
+      \frac{(y_i - \bf z_i \bm \phi^* q_i)}{\pi_{2i}}\right) \\
+    &= (\bm \phi_1^*)^T \Var(\hat{\bf X}_c) \bm \phi_1^* + 
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
+    (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
+    &\qquad + (1 - W)\bm \phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij} 
+    \frac{\bf{x}_i}{\pi_{2i}} \frac{(y_j - \bf z_j \bm \phi^*_1 q_j)}{\pi_{2j}}\\
+  \end{align*}
+
+where the last equality comes from the fact that $\hat{\bf X}_c = W\hat{\bf
+X}_1 + (1 - W) \hat{\bf X}_2$. To have an unbiased estimator of the variance
+  we can use:
+
+  \begin{align*}
+    \hat V_{\rm NNE} 
+    &= (\hat{\bm \phi_1})^T \Var(\hat{\bf X_c}) \hat{\bm \phi_1} + 
+    \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
+    (y_i - \bf z_i \hat{\bm \phi} q_i)(y_j - \bf z_j \hat{\bm \phi} q_j) \\
+    &\qquad + (1 - W)\hat{\bm \phi}_1 \sum_{i \in A_2} \sum_{j \in A_2} 
+    \frac{\Delta_{2ij}}{\pi_{2ij}} \frac{\bf{x}_i}{\pi_{2i}} 
+    \frac{(y_j - \bf z_j \hat{\bm \phi_1} q_j)}{\pi_{2j}}
+  \end{align*}
+
+  where
+
+  $$\hat{\bm \phi} =
+  \begin{bmatrix}
+    \hat{\bm \phi_1} \\ \hat \phi_2
+  \end{bmatrix} = 
+  \left(\sum_{i \in A_2} \frac{q_i}{g'(d_{2i})} 
+  \begin{bmatrix}
+    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
+    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \end{bmatrix}
+  \right)^{-1}
+  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})} 
+  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
+
+\end{proof}
+
+\subsection{Multi-Source Two-Phase Sampling}
+
+When considering non-nested two-phase sampling, we focused on the case of having
+two samples. Now, we consider incorporating more than two independent samples
+together with a debiasing constraint. We consider the case in which we
+want to estimate the $\theta = E[Y]$ and $Y$ is only observed in one sample.
+
+Consider the setup in which we have independent samples $A_0, A_1, \dots, A_M$
+where $Y$ is observed only in $A_0$ but $\bf X$ is observed in all of the
+samples with elements $\bf X^{(0)}$ observed in $A_0$ and $\bf X^{(m)}$ observed
+in $A_m$ for each $m = 1, \dots, M$. Like the non-nested case, we assume that
+each survey is sampling independently from the same population sampling frame.
+
+The traditional multi-source approach (\cite{kim2024statistics}) is to use
+generalized least squares (GLS) to obtain an optimal estimator of $X$ from the
+samples $A_1, \dots, A_M$. Then we can incorporate this information in the
+estimation of $\theta$ by using the following estimate
+
+$$\hat \theta = \sum_{i \in A_0} \hat w_i y_i$$
+
+where
+
+\begin{equation}\label{eq:mstps}
+\hat w = \argmin_w \sum_{i \in A_0} G(w_i)q_i \text{ such that } 
+\sum_{i \in A_0} w_i \bf x_i = \hat X_{\rm{GLS}}^{(0)}
+\end{equation}
+
+where $\hat{\bf X}_{\rm{GLS}}^{(0)}$ is the GLS estimate of $X$ for all of the
+$X$-variables measured in sample $A_0$, and $G$ is a generalized entropy
+function.
+
+In order to have a design consistent estimator that incorporates information from
+samples $A_0, A_1, \dots, A_M$, we want to add the debiasing constraints,
+
+\begin{equation}\label{eq:msdebconstr}
+\sum_{i \in A_0} w_{i} \bf x_i q_i = \sum_{i \in U} g(d_i) q_i
+\end{equation}
+
+where $g(x) = \partial G(x) / \partial x$ and $d_i$ are the design weights for
+sample $A_0$. This yields the optimization problem:
+
+\begin{equation}\label{eq:dc3}
+  \hat w_0 = \argmin_w \sum_{i \in A_0} G(w_{0i}) q_i \text{ such that }
+  \widehat{\bf{T}} = \sum_{i \in A_0} w_{0i} \bf z_i.
+\end{equation}
+
+where $\widehat{\bf{T}} = (\widehat{\bf X}_{\rm{GLS}}^{(0)},
+\sum_{i \in U} g(d_i) q_i)^T$,
+and $\bf z_i = ((\bf x_i^{(0)})^T, g(d_i)q_i)^T$. Then
+the estimator is $\hat Y_{\rm MS} = \sum_{i \in A_0} \hat w_{0i} y_i$ and we can construct
+this estimate minimizing the Lagrangian:
+
+$$
+L(w_{0}, \bm \lambda) = \sum_{i \in A_0} G(w_{0i}) q_i + {\bm \lambda}^T
+\left(\widehat{\bf{T}} - \sum_{i \in A_0} w_{0i} \bf z_i\right)
+$$
+
+where $\bm \lambda$ are the Lagrange multipliers.
+
+%\subsection{Theoretical Results}
+
+\begin{theorem}[Design Consistency]
+  Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$.
+  Under regularity conditions,
+
+  $$\hat Y_{\rm MS} = \hat Y_{\ell, \rm MS}(\bm \lambda^*, \bm \phi^*) + O_p(N / n_0)$$
+
+  where
+
+  $$
+  \hat Y_{\ell, \rm MS}(\bm \lambda^*, \bm \phi^*) = \sum_{i \in A_0} \hat w_{0i} y_i + \left(
+    \widehat{\bf T} - \sum_{i \in A_0} \hat w_{0i}(\bm \lambda^*) \bf z_i q_i
+  \right)\bm \phi^*
+  $$
+
+  and
+
+  $$
+  \bm \phi^* = 
+  \left[\sum_{i \in U} \frac{\pi_{0i} \bf z_i \bf z_i^T q_i}{g'(d_{0i})}\right]^{-1}
+  \sum_{i \in U} \frac{\pi_{0i} \bf z_i y_i}{g'(d_{0i})}.
+  $$
+\end{theorem}
+
+\begin{proof}
+  This proof follows from the proof of Theorem~\ref{thm:dc2}. The only
+  difference is that we have $\hat X_{\rm GLS}$ instead of $\hat X_c$.
+\end{proof}
+
+\begin{theorem}[Variance Estimation]
+  Under regularity conditions,
+
+  $$
+  \begin{aligned}
+    V(\hat Y_{\rm MS}) 
+    &= (\bm \phi^{*})^T\Var(\hat{\bf X}_{\rm GLS}^{(0)}) (\bm \phi^{*}) + 
+      \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{0ij}}{\pi_{0i}\pi_{0j}}
+      (y_i - \bf z_i\bm \phi^{*}q_i)
+      (y_j - \bf z_j\bm \phi^{*}q_j). 
+  \end{aligned}
+  $$
+
+  We can estimate the variance with
+
+  $$
+  \begin{aligned}
+    \hat V(\hat Y_{\rm MS}) 
+    &= (\hat{\bm \phi})^T \hat \Var(\hat{\bf X}_{\rm GLS}^{(0)}) (\hat{\bm \phi})
+    + \sum_{i \in A_0} \sum_{j \in A_0} \frac{\Delta_{0ij}}{\pi_{0ij}\pi_{0i}\pi_{0j}}
+      (y_i - \bf z_i \hat{\bm \phi}q_i)
+      (y_j - \bf z_j \hat{\bm \phi}q_j). 
+  \end{aligned}
+  $$
+
+\end{theorem}
+
+\begin{proof}
+  This result follows the same argument as Theorem~\eqref{thm:nnve}. The result
+  holds because each sample $A_m$ is independent from the other $A_{m'}$.
+\end{proof}
+
+\section{Simulation Studies}
+
 \subsection{Simulation Study 1}
 
-We run a simulation testing the proposed method. In this approach we have the
+We run a simulation testing the main proposed method. In this approach we have the
 following simulation setup:
@@
 uses a regression estimator with $X_3$ as a covariate, which is unnecessary.
-
 \section{Real-Data Analysis}

### 2025-01-20 final/paper/paper.tex
Updated paper with Dr. Kim corrections.

@@
 
+Updates:
+\begin{itemize}
+  \item Modified the end of paragraph two of the introduction and the beginning
+    of paragraph three of the introduction to mirror more of the content of the
+    rest of the paper.
+  \item Fixed spacing issues with the beginning of paragraphs.
+  \item Added Table~\ref{tab:msexample}
+\end{itemize}
 \newpage
@@
 The goal of statistical data integration is to combine information from one
-survey with the variable of interest to another survey that only contains some
+survey with the variable of interest to another survey with
 shared covariates. 
@@
 the variable of interest. An early use of mass imputation is found in
-\cite{breidt1996two} where they used mass imputation to improve estimation in
-the NRI survey. \cite{kim2012combining} used a parametric model to develop a
-novel resampling technique to construct confidence intervals of population level
-characteristics. \cite{chen2022nonparametric} use a nonparametric model to
-estimate population characteristics. In the macro approach, one uses summary
-statistics from other studies to adjust the current estimation. This is commonly
-done using generalized least squares (GLS). 
-\cite{hidiroglou2001double} applied the GLS technique to non-nested two-phase
-samples. \cite{merkouris2004combining} used GLS with multiple surveys and showed
+\cite{breidt1996two} where they use mass imputation to improve estimation in
+the National Resources Inventory survey. Later, \cite{kim2012combining} use a
+parametric model to develop a novel resampling technique to construct confidence
+intervals of population level characteristics. \cite{chen2022nonparametric} use
+a nonparametric model to estimate population characteristics. In the macro
+approach, one uses summary statistics from other studies to adjust the current
+estimation. This is commonly done using generalized least squares (GLS).
+\cite{hidiroglou2001double} apply the GLS technique to non-nested two-phase
+samples. \cite{merkouris2004combining} use GLS with multiple surveys and showed
 an optimality result. This was later extended to area level domains in
-\cite{merkouris2010combining}.
-
-In this paper, we incorporate summary statistics of other surveys by
+\cite{merkouris2010combining}. The main idea of the GLS approach is to construct
+a regression estimator for two-phase samples. In Section 2 of the paper, we
+explain the basic approach of regression estimation and show how it can be
+generalized to a calibration estimator.
+
+This paper, however, generalizes the two-phase regression estimator using the
+debiased calibration technique of \cite{kwon2024debiased}.
+We incorporate summary statistics of other surveys by
 using a generalized two-phase sample similar to the GLS method. The first phase
@@
 data. Unlike previous data integration methods, however, we also include a
-pre-chosen function of the second phase design weights like
-\cite{kwon2024debiased}. This allows us to develop theory showing that our
+pre-chosen function of the second phase design weights, which allows us to
+develop theory showing that our
 estimates are asymptotically design unbiased and that we have valid confidence
@@
 
-This paper is organized as follows. Section 2 describes the basic setup of data
-integration using summary statistics. Section 3 details the generalized
-two-phase sampling framework and provides theory for our estimator being design
+\textcolor{red}{(I am not sure whether the current introduction provides a good
+  guide for the main contents of the paper. The introduction does not match
+  well with the main contents of the paper.)
+}
+
+\textcolor{red}{CL: I have modified the last two paragraphs to help tie into the
+main work of the paper.}
+
+This paper is organized as follows. 
+Section 2 describes the existing methods to combine probability samples using
+regression estimators or calibration.
+Section 3 details the generalized
+two-phase sampling framework and provides theory for our estimator to be design
 unbiased and the variance estimation to be valid. In Section 4, we extend this
-framework first to non-nested two-phase sampling and second to the multiple
+framework first to non-nested two-phase sampling and second to multiple
 samples. Section 5 provides several simulation studies which show that our
-method works, and Section 6 applies the method the Medicare Current Beneficiary
-Survey.
+method works, and Section 6 applies the method to the Medicare Current
+Beneficiary Survey.
 
@@
 
-Consider a finite population, $U$, of size $N$ 
+Consider a finite population, $U = \{1, \dots, N\}$, of size $N$ 
 with measurement  $(\bx_i, y_i)$, where $\bx_i$ is the realization of the
@@
 Two-phase sampling is a popular method for cost effective sampling.
-In the two-phase sampling, the first phase sample $A_1$  is selected from the
+In two-phase sampling, the first phase sample $A_1$  is selected from the
 target population to obtain the measurement of $\bx_i$ in the sample. The
@@
 goal of two-phase sampling is to construct an estimator of $Y_N =  \sum_{i=1}^N y_i$ 
-that uses both the observed information in the Phase 2 sample and also the extra
-auxiliary information from $\bf X$ in the Phase 1 sample. How to combine
-information efficiently is an important problem in two-phase sampling.
+that uses both the observed information in the Phase 2 sample and also the 
+auxiliary information from $\bf X \in \R^p$ in the Phase 1 sample. How to
+combine information efficiently is an important problem in two-phase sampling.
 
@@
 where $q_i = q(\bf x_i)$  is a function of $\bf x_i$ and
+
 $$
@@
 $$
+
 Regarding the choice of $q_i$, one can consider a linear regression model 
@@
 y_i = \bx_i^\top \bm \beta + e_i 
+\label{regression}
 \end{equation}
 where $e_i \sim \left(0, \sigma^2 c_i \right)$, where $c_i=c( \bx_i)$ is a known
-function. In this case,   $q_i=c_i$ is a reasonable choice.
-
-One way to understand the regression estimator is to use the following decomposition:
+function. In this case,  $q_i = c_i$ is a reasonable choice. 
+One way to understand the two-phase regression estimator is to use the following
+decomposition:
 
@@
 \hat Y_{\rm reg, tp} 
-= \underbrace{\sum_{i \in A_1} \frac{\bf x_i \hat{\bm \beta_q}}{\pi_{1i}}}_{
+= \underbrace{\sum_{i \in A_1} \frac{\bf x_i^\top \hat{\bm \beta_q}}{\pi_{1i}}}_{
   \text{ Prediction  }} + \underbrace{\sum_{i \in A_2}
-\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i \hat{\bm \beta_q})}_{
+\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i^\top \hat{\bm \beta_q})}_{
 \text{Bias correction}}.
@@
 regardless of the underlying the superpopulation model. 
-Note that the two-phase regression estimator in (\ref{eq:1})  can be
+Note that the two-phase regression estimator in Equation~\ref{eq:1} can be
 expressed as  $\hat Y_{\rm reg, tp} =
-\sum_{i \in A_2} \hat w_{2i} y_i / \pi_{1i}$ where
+\sum_{i \in A_2} \pi_{1i}^{-1} \hat w_{2i} y_i$ where
 
@@
 
-This means that $\hat Y_{reg, tp}$ is also a calibration estimator. The idea
+This means that $\hat Y_{\rm reg, tp}$ is also a calibration estimator. The idea
 that regression estimation is a form of calibration was noted by
 \cite{deville1992calibration} and extended by them to consider loss functions
-other than just squared loss. Their generalized loss function minimizes
+other than squared loss. Their generalized loss function minimizes
 $\sum_i G(w_i, d_i)q_i$ for weights $w_i$ and design-weights $d_i$ where
@@
 defined on an interval containing $d_i$, with $g(w_i, d_i) = \partial G /
-\partial w$ continuous.\footnote{The \cite{deville1992calibration} paper
+\partial w$ being continuous.\footnote{The \cite{deville1992calibration} paper
 considers regression estimators for a single phase setup, which we apply to our
@@
 generalization includes empirical likelihood estimation, and maximum entropy
-estimation among others. The variance estimation is based on a linearization
-that shows that minimizing the generalized loss function subject to the
-calibration constraints is asymptotically equivalent to a regression estimator.
+estimation among others \citep{schennach2007point}.
 
@@
 
-To construct asymptotically unbiased estimate of the variance we assume that the
-sampling designs for $A_1$ and $A_2$ are measurable and we define the following
-joint inclusion probabilities:
-
-$\Pr(i \in A_1, j \in A_1) = \pi_{1ij}, \text{ and }
-\Pr(i \in A_2, j \in A_2 \mid i \in A_1, j \in A_1) = \pi_{2ij|1}.$
-
 \section{Main Proposal}
@@
 strictly convex and differentiable. We consider maximizing 
-
 \begin{equation}\label{eq:primalloss}
-  H(w) = - \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i
-\end{equation}
-
+  H(w_2) = - \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i
+\end{equation}
 subject to the following constraints:
-
 \begin{equation}\label{eq:calconst1}
@@
 \end{equation}
-
 and 
-
 \begin{equation}\label{eq:calconst2}
@@
 
-The first constraint is the existing calibration constraint and the second
+The first constraint is the usual calibration constraint motivated from the
+regression model in (\ref{regression}) and the second
 ensures that design consistency is achieved. 
@@
 population quantities on the right hand side of \eqref{eq:calconst1}.
-
 By writing $w_{1i} = \pi_{1i}^{-1}$, the the goal is to solve
-
 \begin{equation}\label{eq:primal}
@@
 \end{equation}
-
 where  $\bf z_i = (\bf x_i / q_i, g(\pi_{2i|1}^{-1}))$. The resulting 
 estimator of $Y_N$ is 
-
 \begin{equation}\label{eq:tpest}
@@
 minimize the Lagrangian function
-
 \begin{equation}\label{eq:legragedc1}
   L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
-  + \bm \lambda \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
+  + \bm \lambda^\top \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
     \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i\right)
 \end{equation}
-
 where $\bm \lambda$ is a vector of Lagrange multipliers.
 Differentiating with respect to $w_{2i}$ and setting this expression equal to
-zero, yields the fact that $\hat w_{2i}$ satisfies
-
-$$ \hat w_{2i} = g^{-1}(\hat{\bm \lambda}^T \bf z_i) $$
-
-where $\hat{\bm \lambda}$ is the solution to
-
+zero, we can obtain 
+$$ \hat w_{2i} = g^{-1}(\hat{\bm \lambda}^\top \bf z_i) $$
+where $g(w) = \partial G / \partial w$ and 
+$\hat{\bm \lambda}$ is the solution to
 \begin{equation}\label{eq:lamdc1}
@@
 \end{equation}
-
-The first result we show is that the construction of the weights in
+The following theorem shows that the construction of the weights in
 Equation~\ref{eq:primal} leads to a design consistent estimator 
@@
   Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$. 
-  Under some regularity conditions,
-
-  $$
-  \hat Y_{\rm DCE} = \hat Y_\ell(\hat{\bm{\lambda}}, \bm \phi^*) + O_p(N / n_2)
-  $$
-
+  Under the regularity conditions stated in the appendix
+  $$
+  \hat Y_{\rm DCE} = \hat Y_\ell(\bm{\lambda}^*, \bm \phi^*) + O_p(N / n_2)
+  $$
   where
-
-  $$
-  \hat Y_{\ell}(\hat{\bm \lambda}, \bm \phi^*) = 
-  \sum_{i \in A_2} w_{1i} \hat w_{2i} y_i
+  $$
+  \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) = 
+  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} y_i
   + \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - 
-    \sum_{i \in A_2} w_{1i} \hat w_{2i} \bf{z}_i q_i\right)\bm \phi^*,
-  $$
-
+  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} \bf{z}_i q_i\right)\bm \phi^*,
+  $$
   and
-
   $$
   \bm \phi^* = 
-  \left[\sum_{i \in U} \frac{\pi_{2i|1}\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
+  \left[\sum_{i \in U} \frac{\pi_{2i|1}\bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}\right]^{-1}
   \sum_{i \in U} \frac{\pi_{2i|1}\bf z_i y_i}{g'(d_{2i|1})}.
@@
 %   &= \hat{Y}_{\rm DCE} + \left(\sum_{i \in A_1} w_{1i} \bf x_i  -
-%     \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \bf x_i  \right)^T \bm \phi_1^* \\
+%     \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \bf x_i  \right)^\top \bm \phi_1^* \\
 %   &\quad+ \left(\sum_{i \in A_1} w_{1i} g_i  -
-%  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^T \phi_2^*  
+%  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^\top \phi_2^*  
 %  \end{align*} 
@@
 %  \begin{pmatrix}
-%  \bx_i \bx_i^T &   \bx_i g_i   \\
-%  g_i  \bx_i^T   & g_i^2     
+%  \bx_i \bx_i^\top &   \bx_i g_i   \\
+%  g_i  \bx_i^\top   & g_i^2     
 %  \end{pmatrix} \right]^{-1}
@@
 
-See the appendix for the proof of Theorem 1. Notice that this theorem implies
-that $\hat Y_{\ell}(\hat{\bm \lambda}, \bm \phi^*)$ and $\hat Y_{\rm DCE}$ have
-the same asymptotic expectation and variance, which means that we can estimate
-the variance of $\hat Y_{\ell}(\hat{\bm \lambda}, \bm \phi^*)$ to get an
-estimate of the variance of $\hat Y_{\rm DCE}$. To estimate $\bm \phi^*$, we can
-use
-
-$$
+See the appendix for the proof of Theorem~\ref{thm:dc1}. Notice that this
+theorem implies
+that $\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ and $\hat Y_{\rm DCE}$ have
+the same asymptotic expectation and variance, 
+which means that we can estimate
+the variance of $\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ to get an
+estimate of the variance of $\hat Y_{\rm DCE}$. The basic idea of 
+Theorem~\ref{thm:dc1} is that asymptotically the two-phase debiased calibration
+method is equivalent to a regression estimator. By constraints 
+(\ref{eq:calconst1}) and (\ref{eq:calconst2}), 
+$\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ is design consistent, and the
+variance is asymptotically equivalent to the two-phase regression estimator. 
+We can estimate $\bm \phi^*$ with
+\begin{equation}
 \hat{\bm \phi} = 
 \left[\sum_{i \in A_2} 
-w_{1i}\frac{\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
+w_{1i}\frac{\bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}\right]^{-1}
 \sum_{i \in A_2} w_{1i}\frac{\bf z_i y_i}{g'(d_{2i|1})}.
-$$
-
+\label{eq:8}
+\end{equation}
+Therefore, the proposed calibration estimator in (\ref{eq:primal}) is
+asymptotically equivalent to the two-phase regression estimator given by 
+  $$
+  \hat Y_{\rm tp, reg}  = 
+  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} y_i
+  + \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - 
+  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} \bf{z}_i q_i\right)\hat{\bm \phi},
+  $$
+where $\hat{\bm \phi}$ is defined in (\ref{eq:8}).
+
+For specific sample designs, we can construct the optimal entropy function
+$G(\cdot)$. From Theorem~\ref{thm:dc1}, the choice of entropy function affects
+the variance of the debiased calibration estimator through the form of $g'(d_{2i|1})$
+in $\bm \phi^*$. Consider a class of generalized regression estimators,
+$$
+\mathbf{B} = 
+\left\{\sum_{i \in A_1} w_{1i} \bf{z}_i^\top \bm \gamma + 
+\sum_{i \in A_2} \frac{w_{1i}}{\pi_{2i|1}}(y_i - \bf{z}_i^\top \bm \gamma):
+\bm \gamma \in \R^{p+1}\right\}
+$$
+Then the design optimal estimator is 
+$$
+\hat \theta_{\ell, \rm{opt}} = 
+\sum_{i \in A_1} w_{1i} \bf{z}_i^\top \bm \phi_{\rm opt} + 
+\frac{w_{1i}}{\pi_{2i|1}}\sum_{i \in A_2} (y_i - \bf{z}_i^\top \bm \phi_{\rm opt}):
+$$
+where 
+$$
+\bm \phi_{\rm opt} = 
+\Var\left(\sum_{i \in A_2} d_{2i|1} \bf{z}_i\right)^{-1} 
+\Cov\left(\sum_{i \in A_2} d_{2i|1} \bf{z}_i, \sum_{i \in A_2}, d_{2i|1} y_i \right).
+$$
+We assume that both $A_1$ and $A_2$ are Poisson samples. Then,
+$$
+\bm \phi_{\rm opt} = 
+\left\{
+\sum_{i \in U} \left(d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1)\right) \bf{z_i}\bf{z_i}^\top
+\right\}^{-1}
+\sum_{i \in U} \left(d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1)\right) \bf{z_i} y_i.
+$$
+To find the optimal entropy function, we solve the differential equation
+$$
+\frac{1}{d_{2i|1} g'(d_{2i|1})} = d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1).
+$$
+Using the fact that $(y(xy-1))^{-1} = -y^{-1} + x(xy-1)^{-1}$, the optimal
+entropy function is 
+$$
+G(w_{2i|1}) = - \log(w_{2i|1}) + \log(w_{2i|1}w_{1i} - 1).
+$$
+To construct an asymptotically unbiased estimate of the variance we assume that
+the sampling designs for $A_1$ and $A_2$ are measurable and we define the
+following joint inclusion probabilities:
+$$
+\Pr(i \in A_1, j \in A_1) = \pi_{1ij}, \text{ and }
+\Pr(i \in A_2, j \in A_2 \mid i \in A_1, j \in A_1) = \pi_{2ij|1}.
+$$
 The next theorem discusses how to compute an estimate for the variance of 
@@
   We can construct an unbiased estimate the variance of $\hat Y_{\rm DCE}$ with
-
   $$
@@
   $$
-
   where $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ and 
   $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$.
-
 \end{theorem}
@@
 Now we consider the sampling mechanism known as non-nested two-phase sampling 
-(\cite{hidiroglou2001double}). In the last section, we considered two-phase sampling
+\citep{hidiroglou2001double}. In the last section, we considered two-phase sampling
 in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
@@
 estimators of the total of $\bf X_N$,
-
 $$\hat{\bf X}_1 = \sum_{i \in A_1}^{n_1} d_{1i} \bf x_i \text{ and } 
 \hat{\bf X}_2 = \sum_{i \in A_2}^{n_2} d_{2i} \bf x_i $$
-
 where $d_{1i} = \pi_{1i}^{-1}$, $d_{2i} = \pi_{2i}^{-1}$, 
@@
 We can combine these estimates using the effective sample size 
-(\cite{kish1965survey}) to get
+\citep{kish1965survey} to get
 $\hat{\bf X}_c = (n_{1, \rm{eff}} \hat{\bf X}_1 + n_{2, \rm{eff}}\hat{\bf X}_2) / 
@@
 Then we can define a regression estimator as
-
-$$
-\hat Y_{\rm NN, reg} = \hat Y_2 + (\hat{\bf X}_c - \hat{\bf X}_2)^T \hat{\bm \beta_q} = 
-\hat Y_2 + (\hat{\bf X}_1 - \hat{\bf X}_2)^T W\hat{\bm \beta_q}
-$$
-
+$$
+\hat Y_{\rm NN, reg} = \hat Y_2 + (\hat{\bf X}_c - \hat{\bf X}_2)^\top \hat{\bm \beta_q} = 
+\hat Y_2 + (\hat{\bf X}_1 - \hat{\bf X}_2)^\top W\hat{\bm \beta_q}
+$$
 where, $W = n_{1, \rm{eff}} / (n_{1, \rm{eff}} + n_{2, \rm{eff}})$, and
-
 $$
 \hat{\bm \beta_q} = 
-\left(\sum_{i \in A_2} \frac{\bf x_i \bf x_i^T}{q_i}\right)^{-1} 
+\left(\sum_{i \in A_2} \frac{\bf x_i \bf x_i^\top}{q_i}\right)^{-1} 
 \sum_{i \in A_2} \frac{\bf x_i y_i}{q_i} \text{ and }
@@
 estimator to be known. Since the samples $A_1$ and $A_2$ are independent, 
-
 $$
 V(\hat Y_{\rm NN, reg}) = 
-V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i - \bf x_i^TW\bm \beta^*_q)\right)
-+ (\bm \beta^*_q)^T W^T V(\hat{\bf X}_1) W \bm \beta_q^*
-$$
-
+V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i - \bf x_i^\top W\bm \beta^*_q)\right)
++ (\bm \beta^*_q)^\top W^\top V(\hat{\bf X}_1) W \bm \beta_q^*
+$$
 where $\bm \beta_q^*$ is the probability limit of $\hat{\bm \beta_q}$. Like the
-two-phase sample this regression estimator can be viewed as the solution to the
+two-phase sample, this regression estimator can be viewed as the solution to the
 following calibration equation 
-
 \begin{equation}\label{eq:nncal}
@@
 \end{equation}
-
 and $\hat Y_{\rm NN, reg} = \sum_{i \in A_2} \hat w_{2i} y_i$ where $\hat
 w_{2i}$ is the solution to Equation~\ref{eq:nncal}.
-
 We can extend the debiased calibration estimator of \cite{kwon2024debiased} to
@@
 $G: \mathcal{V} \to \mathbb{R}$, the goal is to solve
-
 \begin{equation}\label{eq:nnopt}
@@
 \end{equation}
-
 for $g(x) = G'(x)$ and a known choice of $q_i \in \mathbb{R}$. 
@@
 of $\bf X_N$, 
-
 $$
@@
 $A_1$ and $A_2$ respectively. Then the optimal combined estimate is 
-
 $$
@@
 $$
-
 We can construct a non-nested two-phase estimator $\hat Y_{\rm NNE}$ for $Y_N$ where
@@
 setup we minimize the Lagrangian,
-
 \begin{equation}\label{eq:legragedc2}
   L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} G(w_{2i}) q_i + 
-  \bm \lambda \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i} \bf z_i q_i\right).
-\end{equation}
-
+  \bm \lambda^\top \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i} \bf z_i q_i\right).
+\end{equation}
 where $\bm \lambda$ are the Lagrange multipliers, with 
-
 $$\hat{\bf T} = 
@@
 $$
-
 Differentiating with respect to $w_{2i}$ and setting this expression equal to
-zero, yields the fact that $\hat w_{2i}$ satisfies
-
-$$ \hat w_{2i}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda^T} \bf z_i) $$
-
+zero yields the fact that $\hat w_{2i}$ satisfies 
+$$ \hat w_{2i}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda}^\top \bf z_i) $$
 where $\hat{\bm \lambda}$ is the solution to
-
 \begin{equation}\label{eq:lamdc2}
@@
 \end{equation}
-
 Like the previous section we first prove that the estimator is design consistent.
@@
   \bm \phi^*) + O_p(Nn_2^{-1})$ where
-
-  $$
-  \hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*) = \sum_{i \in A_2} \hat w_{2i}(
-  \bm \lambda^*) y_i + \left(\hat{\bf T} - \sum_{i \in A_2} \hat w_{2i}(
-  \bm \lambda^*) \bf z_i q_i\right) \bm \phi^*
-  $$
-
+  $$
+  \hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*) = 
+  \sum_{i \in A_2} \pi_{2i}^{-1} y_i + 
+  \left(\hat{\bf T} - \sum_{i \in A_2} \pi_{2i}^{-1} \bf z_i q_i\right)^\top \bm \phi^*
+  $$
   and 
-
   $$
@@
 where $\hat{\bm \lambda}$ solves
-
 \begin{equation}
@@
 \end{equation}
-
 we have 
-
-$$\hat Y_{\ell, \rm NNE} (\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} \hat
+$$\hat Y_{\rm NNE} (\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} \hat
 w_{2i}(\hat{\bm \lambda})y_i + \left(\bf T - \sum_{i \in A_2} \hat
-w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right) \bm \phi.$$
-
+w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right)^\top \bm \phi.$$
 If we choose $\bm \phi^*$ such that $E\left[\frac{\partial}{\partial \bm \lambda} 
-  \hat Y_{\ell, NNE}(\bm \lambda^*, \bm \phi^*)\right] = 0$, then
-
+  \hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*)\right] = 0$, then
 $$\bm \phi^* =
@@
 $$
-
 Hence, by a Taylor expansion around $\hat{\bm \lambda}$,
-
 $$\hat Y_{\rm NNE}(\hat{\bm \lambda}) = \hat Y_{\ell, \rm NNE}(\bm \lambda^*, 
 \bm \phi^*) + O_p(Nn_2^{-1}).$$
-
 The next theorem shows how to estimate the variance of $\hat Y_{\rm NNE}$.
@@
   The variance of $\hat Y_{\rm NNE}$ is 
-
   \begin{align*}
-    \Var(\hat Y_{\rm NNE}(\hat \lambda))
-    &= (\bm \phi_1^*)^T \Var(\hat{\bf X}_c) \bm \phi_1^* + 
+    \Var(\hat Y_{\rm NNE}(\hat{\bm \lambda}))
+    &= (\bm \phi_1^*)^\top \Var(\hat{\bf X}_c) \bm \phi_1^* + 
     \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
@@
   \end{align*}
-
   We can estimate the variance using
-
   \begin{align*}
     \hat V_{\rm NNE} 
-    &= (\hat{\bm \phi_1})^T \Var(\hat{\bf X}_c) \hat{\bm \phi_1} + 
+    &= (\hat{\bm \phi_1})^\top \Var(\hat{\bf X}_c) \hat{\bm \phi_1} + 
     \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
@@
   \end{align*}
-
   where 
-
   $$\hat{\bm \phi} =
@@
 assume that each survey is sampled independently from the same population
-sampling frame.
-
-The traditional multi-source approach (\cite{kim2024statistics}) is to use
+sampling frame. Some basic examples of data structures that would benefit from
+the multi-source two-phase sampling can be found in Table~\ref{tab:msexample}.
+
+\begin{table}[ht!]
+  \centering
+  \begin{tabular}{lrrrrrr}
+    \toprule 
+           & \multicolumn{3}{c}{Example 1} & \multicolumn{3}{c}{Example 2} \\
+           \cmidrule(r){2-4}  
+           \cmidrule(r){5-7}
+    Sample & $X_1$ & $X_2$ & $Y$ & $X_1$ & $X_2$ & $Y$ \\
+    \midrule
+    $A_0$ & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark \\
+    $A_1$ & \checkmark & &  & \checkmark & \checkmark & \\
+    $A_2$ & & \checkmark &  & & \checkmark & \\
+    \bottomrule
+  \end{tabular}
+  \caption{This tables shows two examples configuations of multi-source sampling.
+  In each example, $A_0$ contains the variable of interest $Y$. For Example 1,
+  the covariates in $A_1$ and $A_2$ are both contained in $A_0$. In Example 2,
+  sample $A_2$ improves the estimation of $Y$ via improving the estimation of
+  $X_1$ in $A_1$.}
+  \label{tab:msexample}
+\end{table}
+
+The traditional multi-source approach \citep{kim2024statistics} is to use
 GLS to obtain an optimal estimator of $\bf X_N$ from the
@@
 estimation of $\theta$ by using the following estimate 
-
 $$\hat \theta = \sum_{i \in A_0} \hat w_i y_i$$
-
 where 
-
 \begin{equation}\label{eq:mstps}
@@
 \end{equation}
-
 where $\hat{\bf X}_{\rm{GLS}}^{(0)}$ is the GLS estimate of $\bf X$ for all of
-the $X$-variables measured in sample $A_0$, and $G$ is a generalized entropy
+the $X$-variables measured in sample $A_0$, and $G(\cdot)$ is a generalized entropy
 function. 
@@
 samples $A_0, A_1, \dots, A_M$, we want to add the debiasing constraint, 
-
 \begin{equation}\label{eq:msdebconstr}
@@
 \end{equation}
-
 where $g(x) = \partial G(x) / \partial x$ and $d_i$ are the design weights for
 sample $A_0$. This yields the optimization problem:
-
 \begin{equation}\label{eq:dc3}
@@
 \end{equation}
-
 where $\widehat{\bf{T}} = (\widehat{\bf X}_{\rm{GLS}}^{(0)},
-\sum_{i \in U} g(d_i) q_i)^T$,
-and $\bf z_i = ((\bf x_i^{(0)})^T, g(d_i)q_i)^T$. Then
+\sum_{i \in U} g(d_i) q_i)^\top$,
+and $\bf z_i = ((\bf x_i^{(0)})^\top, g(d_i)q_i)^\top$. Then
 the estimator is $\hat Y_{\rm MS} = \sum_{i \in A_0} \hat w_{0i} y_i$ and we can
 construct this estimate minimizing the Lagrangian:
-
-$$
-L(w_{0}, \bm \lambda) = \sum_{i \in A_0} G(w_{0i}) q_i + {\bm \lambda}^T
+$$
+L(w_{0}, \bm \lambda) = \sum_{i \in A_0} G(w_{0i}) q_i + {\bm \lambda}^\top
 \left(\widehat{\bf{T}} - \sum_{i \in A_0} w_{0i} \bf z_i\right)
 $$
-
 where $\bm \lambda$ are the Lagrange multipliers from Equation~\ref{eq:dc3}.
@@
   Under regularity conditions,
-
   $$
@@
   $$
-
   where
-
-  $$
-  \hat Y_{\ell, \rm MS}(\bm \lambda^*, \bm \phi^*) = \sum_{i \in A_0} \hat w_{0i} y_i + \left(
-    \widehat{\bf T} - \sum_{i \in A_0} \hat w_{0i}(\bm \lambda^*) \bf z_i q_i
-  \right)\bm \phi^*
-  $$
-
+  $$
+  \hat Y_{\ell, \rm MS}(\bm \lambda^*, \bm \phi^*) = 
+  \sum_{i \in A_0} d_{0i} y_i + \left(
+    \widehat{\bf T} - \sum_{i \in A_0} d_{0i} \bf z_i q_i
+  \right)^\top \bm \phi^*
+  $$
   and 
-
   $$
   \bm \phi^* = 
-  \left[\sum_{i \in U} \frac{\pi_{0i} \bf z_i \bf z_i^T q_i}{g'(d_{0i})}\right]^{-1}
+  \left[\sum_{i \in U} \frac{\pi_{0i} \bf z_i \bf z_i^\top q_i}{g'(d_{0i})}\right]^{-1}
   \sum_{i \in U} \frac{\pi_{0i} \bf z_i y_i}{g'(d_{0i})}.
@@
     V(\hat Y_{\rm MS}) 
-    &= (\bm \phi^{*})^T\Var(\hat{\bf X}_{\rm GLS}^{(0)}) (\bm \phi^{*}) + 
+    &= (\bm \phi^{*})^\top \Var(\hat{\bf X}_{\rm GLS}^{(0)}) (\bm \phi^{*}) + 
       \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{0ij}}{\pi_{0i}\pi_{0j}}
@@
   $$
-
   We can estimate the variance with
-
   $$
@@
     \hat V(\hat Y_{\rm MS}) 
-    &= (\hat{\bm \phi})^T \hat \Var(\hat{\bf X}_{\rm GLS}^{(0)}) (\hat{\bm \phi})
+    &= (\hat{\bm \phi})^\top \hat \Var(\hat{\bf X}_{\rm GLS}^{(0)}) (\hat{\bm \phi})
     + \sum_{i \in A_0} \sum_{j \in A_0} \frac{\Delta_{0ij}}{\pi_{0ij}\pi_{0i}\pi_{0j}}
@@
 following simulation setup:
-
 $$
@@
 $$
-
 where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
@@
 The parameter $\delta \in \{0, 1\}$ controls the effect of model misspecification
-in the simulation. Let $\bf x_i = (1, x_{1i}, x_{2i})^T$. We compare the
+in the simulation. Let $\bf x_i = (1, x_{1i}, x_{2i})^\top$. We compare the
 proposed method for the parameter $\bar Y_N$ with four approaches:
@@
 Simulation 1. We have the following simulation setup:
-
 $$
@@
 $(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. 
-If $\delta = 0$ then there is no model misspecification. However, if $\delta =
+If $\delta = 0$, then there is no model misspecification. However, if $\delta =
 1$, then there is some model misspecification. We estimate
@@
 $\hat{\bf X}_{1, \rm HT} = \sum_{i \in A_1} d_{1i} \bf x_{i}$, $\hat{\bf X}_{2,
-    \rm HT} = \sum_{i \in A_2} d_{2i} \bf x_{i}$, $\bf x_i = (1, x_{1i}, x_{2i})^T$
+    \rm HT} = \sum_{i \in A_2} d_{2i} \bf x_{i}$, $\bf x_i = (1, x_{1i}, x_{2i})^\top$
     and
 
-    $$ \hat{\bm \beta_2} = \left(\sum_{i \in A_2} \bf x_i \bf x_i^T\right)^{-1}
+    $$ \hat{\bm \beta_2} = \left(\sum_{i \in A_2} \bf x_i \bf x_i^\top\right)^{-1}
     \sum_{i \in A_2} \bf x_i y_i.$$
@@
 This simulation study tests the multi-source extension.
-We have the following superpopulation model with $N = 10000$ elements:
-
+We have the following superpopulation model with $N = 10,000$ elements:
 $$
@@
 $$
-
 Like the previous simulation studies, when $\delta = 1$, there is model
 misspecification for the outcome model because we observe the following columns
-in each sample
-
+in each sample:
 \begin{table}[ht!]
@@
 different estimators for $\theta = E[Y]$.
-
 \begin{itemize}
@@
 \end{itemize}
-
 The simulation results are displayed in Table~\ref{tab:msdc0-mean} and
@@
 
-As expected EstPop has the lowest RMSE because it also uses the population
+As expected, EstPop has the lowest RMSE because it also uses the population
 totals. The Est estimator outperforms
@@
 supplement to identify the impact of COVID-19 on the Medicare population
-\cite{norc2020mcbs}. This analysis looks at whether
+\citep{norc2020mcbs}. This analysis looks at whether
 respondents had obtained at least one dose of a COVID-19 vaccine.
@@
 are known to have a higher all-cause mortality from
-COVID-19 (\cite{bonanad2020effect}). Thus, understanding how older people get
-vaccinated can help save lives and reduce medical costs.
+COVID-19 \citep{bonanad2020effect}. Thus, understanding what causes older people
+to get vaccinated can help save lives and reduce medical costs.
 
 The MCBS is a rotating panel survey in which participants are interviewed up to
-three times within a four-year period (\cite{cms2021mcbs}). The participants
+three times within a four-year period \citep{cms2021mcbs}. The participants
 are selected from Medicare enrollment data and enter a cohort in the fall round
@@
 and fall rounds of the survey for three consecutive years before exiting the
-survey in the winter round of year four (\cite{cms2021mcbs}). Each year a new
+survey in the winter round of year four \citep{cms2021mcbs}. Each year a new
 cohort is added. Participants are selected based on a three-stage cluster sample
 design. The first stage of the sample consists of major metropolitan areas and
-groups of rural counties (\cite{cms2021mcbs}). The second stage are census
+groups of rural counties \citep{cms2021mcbs}. The second stage are census
 tracts within the primary sampling units. Finally, Medicare beneficiaries are
 selected using a stratified systematic sample with random starts 
-(\cite{cms2021mcbs}). Medicare beneficiaries are selected from strata based on
+\citep{cms2021mcbs}. Medicare beneficiaries are selected from strata based on
 age and whether the beneficiary is Hispanic. Strata with Hispanic beneficiaries
@@
 2020 to assess the impact of COVID on the Medicare
-population (\cite{cms2021covid}). These questions about vaccines were added to
+population \citep{cms2021covid}. These questions about vaccines were added to
 participants who were enrolled in the Summer 2021 cohort.
@@
 Medicare from the Center of Medicare and Medicaid Services (CMS). Like the MCBS
-survey, the ACS data contains information about the overall percentage of
-people's sex, race and ethnicity, education, and marital status for people on
+survey, the ACS data contains information about the overall distribution of
+sex, race and ethnicity, education, and marital status for people on
 Medicare. We combine this information with official totals of people on Medicare
@@
 least one COVID vaccination dose. This MCBS data was updated to include
-information about if a participant had received a second vaccine dose and then
-later to include how many total vaccine doses patients had received. If the
-participant had received at least one more dose and their answer was missing (because
-they refused to answer, did not know, or was just missing) the missing value was
-imputed to say that they yes, they had received a COVID vaccine dose. Otherwise,
-the missing value was imputed to say no, the participant had not received a
-COVID vaccine dose.
+information about whether a participant had received a second vaccine dose. 
+Later, the survey included the total vaccine doses patients had received. If the
+participant had received at least one more dose and their answer was missing
+(because they refused to answer, did not know, or was missing) the missing value
+was imputed to say that they yes, they had received a COVID vaccine dose.
+Otherwise, the missing value was imputed to say no, the participant had not
+received a COVID vaccine dose.
 
@@
 regression algorithm, and our preposed multi-source two-phase sampling technique. 
-The Horvitz-Thompson estimator uses the design weights, the MCBS observation if
-an individual received a COVID-19 vaccine dose, and the population total from
-the CMS about the total number of people on Medicare in 2021. The regression
-estimator uses all of this and a population level total for the number of males.
-Our multi-source algorithm also includes information from the ACS about
-education and marital status. Due to the fact that we do not observe the
-probability that an individual will be selected into the MCBS survey within the
-population of all Medicare patients, we have to estimate the term 
-$\sum_{i \in U} g(d_i)$ from Equation~\ref{eq:msdebconstr}. This technique is
-adopted from \cite{kwon2024debiased} and described in more details in the
-appendix.
+The Horvitz-Thompson estimator uses the design weights, the MCBS observation
+whether an individual received a COVID-19 vaccine dose, and the population total
+from the CMS data about the total number of people on Medicare in 2021. The
+regression estimator uses all of this and a population level total for the
+number of males. Our multi-source algorithm also includes information from the
+ACS about education and marital status. Due to the fact that we do not observe
+the probability that an individual will be selected into the MCBS survey within
+the population of all Medicare patients, we have to estimate the term $\sum_{i
+\in U} g(d_i)$ from Equation~\ref{eq:msdebconstr}. This technique is adopted
+from \cite{kwon2024debiased} and described in more details in the appendix.
 The results of the real data analysis are found in Table~\ref{tab:mcbsres}.
@@
 this method be used to combine surveys focusing on the same population frame,
-but because it only does not require individual responses levels from the
-supplemental surveys, this method can be used to construct domain level
+but because it does not require individual responses levels from the
+supplemental surveys, this method can also be used to construct domain level
 estimates as exemplified in the real data analysis of Medicare patients.
@@
 \newpage
-
-\section*{Appendix A: Proof of Theorem 1}
+\appendix
+
+\section{Appendix: Regularity Conditions}
+
+We need the following regularity conditions to prove Theorem 1.
+
+\begin{itemize}
+  \item[A1]: $\mathcal{V} \in \R$ is open,
+  \item[A2]: $G(w)$ is strictly convex and twice continuously differentiable on
+    $\mathcal{V}$ with a bounded second derivative,
+  \item[A3]: There exist positives constants $c_1, c_2 \in \mathcal{V}$ such
+    that $c_1 < \pi_{2i|1} < c_2$ for all $\pi_i$,
+  \item[A4]: For $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$,
+    $$
+    \limsup_{N \to \infty} N \max_{i, j \in U, i \neq j} |\Delta_{2ij|1}| <
+    \infty,
+    $$
+  \item[A5]: The matrix $\bm \Sigma_{\bf z} := 
+    \lim_{N \to \infty} N^{-1} \sum_{i \in U} {\bf z}_i {\bf z}_i^\top$ exists
+    and is positive definite, 
+  \item[A6]: The average fourth moment of $({\bf x}_i^\top, y_i)^\top$ is finite
+    such that 
+    $$
+    \limsup_{N \to \infty} N^{-1} \sum_{i = 1}^N ({\bf x}_i^\top, y_i)^\top < \infty,
+    $$
+  \item[A7]: $||\hat{\bm \lambda} - \bm \lambda^*|| = O_p(n_2^{-1/2})$
+\end{itemize}
+
+Assumption A1 is needed for the differentiability of $G(w)$ in A2.
+Assumption A2 guarentees a unique solution for $\hat{\bm \lambda}$ and ensures
+that $g(w)$ is differentiable. The boundedness part of the assumption ensures
+that $\frac{\partial}{\partial \bm \lambda} g^{-1}(\bm \lambda^\top {\bf z})$ is
+also bounded. Conditions A3 and A4 are standard conditions which control the
+first-order and second-order inclusion probability for the second phase sample.
+These conditions are met by many traditional survey designs. Assumptions A5 and
+A6 are technical conditions and common in the survey literature. Assumption A7
+is added to simplify the proof. We could remove this condition and follow the
+approach of Theorem 1 in \cite{kwon2024debiased}.
+
+\subsection{Proof of Theorem 1}
 
@@
 The first order conditions show that at the solution,
-
 $$
 \frac{\partial \mathcal{L}}{\partial w_{2i}}: g(w_{2i}) w_{1i}q_i -
-w_{1i} \hat{\bm \lambda}^T \bf z_i q_i = 0.
-$$
-
+w_{1i} \hat{\bm \lambda}^\top \bf z_i q_i = 0.
+$$
 Hence, $\hat w_{2i}(\hat{\bm \lambda}) = 
-g^{-1}(\hat{\bm \lambda}^T \bf z_i)$ where
+g^{-1}(\hat{\bm \lambda}^\top \bf z_i)$ where
 $\hat{\bm \lambda}$ is determined by Equation~\eqref{eq:lamdc1}. When the
@@
 $\hat{\bm \lambda} \to \bm \lambda^*$ in probability and 
-$\bm \lambda^* = (\bf 0^T, 1)^T$. 
+$\bm \lambda^* = (\bf 0^\top, 1)^\top$. 
 Using the linearization technique of
 \cite{randles1982asymptotic}, we can construct a regression estimator, 
-
 $$
@@
 $$
-
 Notice that $\hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} w_{1i}
 \hat w_{2i} y_i$ for all $\bm \phi$. We choose $\bm \phi^*$ such that
-
 $$
@@
 $$
-
 Using the fact that $g^{-1}(\bm \lambda^* \bf z_i) = g^{-1}(g(d_{2i|1})) = d_{2i|1}$
 and $(g^{-1})'(x) = 1 / g'(g^{-1}(x))$, we have
-
 \begin{align*}
   \bm \phi^*
-  &= E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}
+  &= E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}
   \right]^{-1}
   E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i y_i}{g'(d_{2i|1})}\right]\\
-  &= \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}
+  &= \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}
   \right]^{-1}
@@
 \end{align*}
-
 Thus, the linearization estimator is
-
 $$\hat Y_\ell(\bm \lambda^*, \bm \phi^*) = 
@@
 \sum_{i \in A_2} w_{1i} d_{2i|1} (y_i - q_i \bf z_i \bm \phi^*).$$
-
 By construction using a Taylor expansion yields,
-
 \begin{align*}
@@
 \end{align*}
-
 The final equality comes from the fact that 
@@
 
-\section*{Appendix B: Proof of Theorem 2}
+\subsection{Proof of Theorem 2}
 
@@
   $\hat Y_{\ell}$. Notice that we can define
-
   $$
@@
   $$
-
   where $\delta_i$ is one if $i \in A_2$ and zero otherwise. Then,
-
   $$
@@
   $$
-
-  Then using the reverse framework of \cite{fay1992inferences} and 
+  Using the reverse framework of \cite{fay1992inferences} and 
   \cite{shao1999variance}, and noting that 
@@
   $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$,
-
   \begin{align*}
@@
   \end{align*}
-
   This means that we can estimate the variance of $\hat Y_{\rm DCE}$ with
-
   $$
@@
   $$
-
 \end{proof}
 
-\section*{Appendix C: Proof of Theorem 4}
+\subsection{Proof of Theorem 4}
 
@@
   variance of $\hat Y_{\rm NNE}(\hat{\bm \lambda})$ is 
-
   \begin{align*}
@@
     q_i\right)\bm \phi^*\right)\\
-    &= (\bm \phi_1^*)^T \Var(\hat{\bf X}_c) \bm \phi_1^* + 
+    &= (\bm \phi_1^*)^\top \Var(\hat{\bf X}_c) \bm \phi_1^* + 
     \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
@@
       \frac{(y_i - \bf z_i \bm \phi^* q_i)}{\pi_{2i}}\right) \\
-    &= (\bm \phi_1^*)^T \Var(\hat{\bf X}_c) \bm \phi_1^* + 
+    &= (\bm \phi_1^*)^\top \Var(\hat{\bf X}_c) \bm \phi_1^* + 
     \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
@@
   \end{align*}
-
 where the last equality comes from the fact that $\hat{\bf X}_c = W\hat{\bf
@@
   we can use:
-
   \begin{align*}
     \hat V_{\rm NNE} 
-    &= (\hat{\bm \phi_1})^T \Var(\hat{\bf X_c}) \hat{\bm \phi_1} + 
+    &= (\hat{\bm \phi_1})^\top \Var(\hat{\bf X_c}) \hat{\bm \phi_1} + 
     \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
@@
 
-\section*{Appendix D: Estimating the Population Weights}
+\section{Estimating the Population Weights}
 
@@
 unknown, we need to estimate it.
-
 We follow the approach of \cite{kwon2024debiased}. Define 
-
 $$N \alpha = \sum_{i \in U} g(d_i).$$
-
 Assuming $N$ is known (or that we can estimate it), we need to estimate
@@
 approach we have the following loss function for the case of non-nested sampling:
-
 \begin{align*}
@@
 \end{align*}
-
 The choice of $K(\alpha)$ is up to the analyst. We consider $K(\alpha) =

### 2025-01-20 final/paper/paper_ov.tex
Updated paper with Dr. Kim corrections.

@@
 \begin{abstract}
-Generalized entropy calibration \cite{gneiting2007strictly} is a nonparametric
+Generalized entropy calibration \citep{gneiting2007strictly} is a nonparametric
 method to estimate survey weights without using the design weights in the
-specified loss function. While \cite{kwon2024debiased} explore generalized
-entropy calibration to estimate survey weights with known population totals, by
-using a two-phase sampling framework, we extend their method to allow for
-estimated population totals. We then extend this two-phase sampling framework to
+specified loss function. \cite{kwon2024debiased} explore generalized
+entropy calibration to estimate survey weights with known population totals.
+By using a two-phase sampling framework, we extend this method to allow for
+estimated population totals. We then generalize the two-phase sampling framework to
 two additional cases: first, when we have two non-nested samples, and
@@
 \end{abstract}
+
 \newpage
@@
 
-Combining information from multiple samples is an important practical problem
-(\cite{yang2020statistical}, \cite{yang2023elastic}, \cite{dagdoug2023model}).
-We want to incorporate information from external data sources to reduce the bias
-in our estimates and improve the estimator's efficiency. For many problems, the
-additional information consists of summary statistics with standard errors. The
-goal of this project is to incorporate external information with existing data 
-to create more efficient estimators using calibration weighting.
-
-To model this scenario, we formulate the problem as a generalized two-phase
-sample. The first phase sample consists of data from multiple sources, and the
-second phase sample contains the existing data. We first consider the classical
-two-pahse sampling setup where the second phase sample is a subset of the first
-phase sample. Then we extend this setup to consider non-nested samples and
-multiple non-nested samples.
-
-\textcolor{red}{ Introduction is too short. We need to make 4-5 paragraphs. } 
-\begin{enumerate}
-\item Narrow down (like a funnel) to the subject that you want to address. 
-\item Explain the context of the research (and cite a few key papers).
-\item Explain what is proposed in the paper.  
-\item Significance of the research (why this paper should be published)
-\item Outline of the paper 
-\end{enumerate}
+Government surveys in the United States provide the public with diverse
+information about a range of population characteristics. These include health
+indicators from the National Health and Nutrition Examination Survey, population
+and housing data from the American Community Survey, and labor force statistics
+from the Current Population Survey. All of these surveys are probability
+surveys. Probability sampling is considered the
+gold-standard for survey sampling. However, obtaining a large probability
+sample is getting more difficult because people are more likely to choose not to
+respond. One solution to this problem is data integration. Instead of analyzing
+a probability sample by itself, we can incorporate information from other
+probability samples in order to have an analysis with higher statistical power.
+
+The goal of statistical data integration is to combine information from one
+survey with the variable of interest to another survey with
+shared covariates. 
+This integration usually involves one of two contexts: incorporating item
+level responses or using summary statistics. These mirror the micro and macro
+approaches described in \cite{yang2020statistical} respectively. The main
+technique to incorporate item level responses is mass imputation. With access to
+individual item level responses, the traditional approach is to build a
+predictive model of the variable of interest from the covariates shared between
+the two surveys and predict the missing outcome variable in the survey without
+the variable of interest. An early use of mass imputation is found in
+\cite{breidt1996two} where they use mass imputation to improve estimation in
+the National Resources Inventory survey. Later, \cite{kim2012combining} use a
+parametric model to develop a novel resampling technique to construct confidence
+intervals of population level characteristics. \cite{chen2022nonparametric} use
+a nonparametric model to estimate population characteristics. In the macro
+approach, one uses summary statistics from other studies to adjust the current
+estimation. This is commonly done using generalized least squares (GLS).
+\cite{hidiroglou2001double} apply the GLS technique to non-nested two-phase
+samples. \cite{merkouris2004combining} use GLS with multiple surveys and showed
+an optimality result. This was later extended to area level domains in
+\cite{merkouris2010combining}.
+
+In this paper, we incorporate summary statistics of other surveys by
+using a generalized two-phase sample similar to the GLS method. The first phase
+sample consists of data
+from multiple sources, and the second phase sample contains the existing
+data. Unlike previous data integration methods, however, we also include a
+pre-chosen function of the second phase design weights like
+\cite{kwon2024debiased}. This allows us to develop theory showing that our
+estimates are asymptotically design unbiased and that we have valid confidence
+intervals. Moreover, we can show that this technique outperforms traditional
+regression methods in finite samples.
+
+\textcolor{red}{(I am not sure whether the current introduction provides a good guide for the main contents of the paper. The introduction does not match  well with the main contents of the paper.)
+}
+
+This paper is organized as follows. Section 2 describes the basic setup of data
+integration using summary statistics. \textcolor{red}{(This is not true.)} Section 3 details the generalized
+two-phase sampling framework and provides theory for our estimator to be design
+unbiased and the variance estimation to be valid. In Section 4, we extend this
+framework first to non-nested two-phase sampling and second to multiple
+samples. Section 5 provides several simulation studies which show that our
+method works, and Section 6 applies the method to the Medicare Current
+Beneficiary Survey.
 
@@
 
-Consider a finite population of size $N$ 
-with measurement  $(\bx_i, y_i)$, where $\bx_i$ is the realization of the auxiliary variables and $y_i$ is the realization of the study variable of interest for element  $i$ in finite population.
-The auxiliary variables are cheap to measure but the measurement for the study variable is expensive.
-Thus,  two-phase sampling is popular as a tool for cost effective sampling technique.
-In the two-phase sampling, the first phase sample $A_1$  is selected from the target population to obtain the measurement of $\bx_i$ in the sample.
-The sampling design for the first-phase sample can be denoted by $P(A_1)$.
-After that, based on the observed auxiliary information in the first-phase  sample, the second-phase sample $A_1$ is selected from the first-phase sample to observe $y_i$.
-The sampling design for the second-phase sample is denoted as $P( A_2 \mid A_1)$.
+Consider a finite population, $U = \{1, \dots, N\}$, of size $N$ 
+with measurement  $(\bx_i, y_i)$, where $\bx_i$ is the realization of the
+auxiliary variables and $y_i$ is the realization of the study variable of
+interest for element  $i$ in the finite population. The auxiliary variables are
+cheap to measure but the measurement for the study variable is expensive.
+Two-phase sampling is a popular method for cost effective sampling.
+In two-phase sampling, the first phase sample $A_1$  is selected from the
+target population to obtain the measurement of $\bx_i$ in the sample. The
+sampling design for the first-phase sample can be denoted by $P(A_1)$. After
+that, based on the observed auxiliary information in the first-phase  sample,
+the second-phase sample $A_2 \subseteq A_1$ is selected from the first-phase
+sample to observe $y_i$. The sampling design for the second-phase sample is
+denoted as $P( A_2 \mid A_1)$.  
 The
-goal of two-phase sampling is to construct an estimator of $Y =  \sum_{i=1}^N y_i$ 
-that uses both the observed information in the Phase 2 sample and also the extra
-auxiliary information from $\bf X$ in the Phase 1 sample. How to combine  information efficiently is an important problem in two-phase sampling.
-
-Let $\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ be the first-order inclusion probability for the first-phase sample and let $\pi_{2i | 1} = \sum_{A_2: i \in A_2} \Pr(A_2 \mid A_1)$ be the first-order inclusion probability for the second-phase sample.
-From the first-phase sample, we can compute $\widehat{\mathbf{X}}_1 = \sum_{i \in A_1} \pi_{1i}^{-1} \bx_i$ as an unbiased estimator of $\mathbf{X}= \sum_{i=1}^N  \bx_i$.
-Also, from the second-phase sample, we can compute $\left( \widehat{\mathbf{X}}_2 , \widehat{Y}_2 \right) = \sum_{i \in A_2} \pi_{1i}^{-1} \pi_{2i \mid 1}^{-1} \left( \bx_i, y_i\right)$ as an unbiased estimator of $( \mathbf{X}, Y )$.
-Thus, we have three estimators for two parameters.
-The two-phase regression estimator can be  used to combine the information efficiently.
-The two-phase regression estimator can be expressed as
-\begin{equation}
+goal of two-phase sampling is to construct an estimator of $Y_N =  \sum_{i=1}^N y_i$ 
+that uses both the observed information in the Phase 2 sample and also the 
+auxiliary information from $\bf X \in \R^p$ in the Phase 1 sample. How to
+combine information efficiently is an important problem in two-phase sampling.
+
+Let $\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ be the first-order inclusion
+probability for the first-phase sample and let $\pi_{2i | 1} = \sum_{A_2: i \in
+A_2} \Pr(A_2 \mid A_1)$ be the first-order inclusion probability for the
+second-phase sample. From the first-phase sample, we can compute
+$\widehat{\mathbf{X}}_1 = \sum_{i \in A_1} \pi_{1i}^{-1} \bx_i$ as an unbiased
+estimator of $\mathbf{X}_N = \sum_{i=1}^N  \bx_i$. Also, from the second-phase
+sample, we can compute $\left( \widehat{\mathbf{X}}_2 , \hat{Y}_2 \right) =
+\sum_{i \in A_2} \pi_{1i}^{-1} \pi_{2i \mid 1}^{-1} \left( \bx_i, y_i\right)$ as
+an unbiased estimator of $(\mathbf{X}_N, Y_N)$.
+We therefore have three estimators for
+two parameters. The two-phase regression estimator can be  used to combine this
+information efficiently and it can be expressed as  
+\begin{equation} \label{eq:1}
 \hat Y_{\rm reg, tp} 
@@
 \sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}} \left(y_i - \bf x_i^\top  \hat{\bm \beta_q}\right)
-\label{eq:1}
 \end{equation}  
 where $q_i = q(\bf x_i)$  is a function of $\bf x_i$ and
+
 $$
@@
 $$
+
 Regarding the choice of $q_i$, one can consider a linear regression model
+
 \begin{equation}
 y_i = \bx_i^\top \bm \beta + e_i 
-\end{equation}
-where $e_i \sim \left(0, \sigma^2 c_i \right)$, where $c_i=c( \bx_i)$ is a known function. In this case,   $q_i=c_i$ is a reasonable choice.
-
-One way to understand the regression estimator is to use the following decomposition: 
+\label{regression}
+\end{equation}
+
+\textcolor{red}{(Why do you have a space here?) }
+where $e_i \sim \left(0, \sigma^2 c_i \right)$, where $c_i=c( \bx_i)$ is a known
+function. In this case,  $q_i = c_i$ is a reasonable choice. 
+One way to understand the two-phase regression estimator is to use the following
+decomposition:
+
 $$ 
 \hat Y_{\rm reg, tp} 
-= \underbrace{\sum_{i \in A_1} \frac{\bf x_i \hat{\bm \beta_q}}{\pi_{1i}}}_{
+= \underbrace{\sum_{i \in A_1} \frac{\bf x_i^\top \hat{\bm \beta_q}}{\pi_{1i}}}_{
   \text{ Prediction  }} + \underbrace{\sum_{i \in A_2}
-\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i \hat{\bm \beta_q})}_{
+\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i^\top \hat{\bm \beta_q})}_{
 \text{Bias correction}}.
 $$
-The prediction term is obtained by finding the best linear unbiased predictor under the working superpopulation model.
-The bias correction term is used to make  the two-phase regression estimator be asymptotically design unbiased regardless of the underlying the superpopulation model.
-
-%\textcolor{red}{I would move the calibration expression to the next subsection. }
-
-Now, note that the two-phase regression estimator in (\ref{eq:1})  can be expressed as  $\hat Y_{\rm reg, tp} =
-\sum_{i \in A_2} \hat w_{2i} y_i / \pi_{1i}$ where
-
-$$
-\hat w_{2i} = \argmin_{w_2} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i
+The prediction term is obtained by finding the best linear unbiased predictor
+under the working superpopulation model. The bias correction term 
+makes the two-phase regression estimator asymptotically design unbiased
+regardless of the underlying the superpopulation model. 
+Note that the two-phase regression estimator in Equation~\ref{eq:1} can be
+expressed as  $\hat Y_{\rm reg, tp} =
+\sum_{i \in A_2} \pi_{1i}^{-1} \hat w_{2i} y_i$ where
+
+$$
+\hat w_{2} = \argmin_{w_2} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i
 \text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i / \pi_{1i} = \sum_{i \in A_1}
@@
 have bias calibration separate from the minimizing the variance so that
-we can control each in isolation. In
-\cite{kwon2024debiased}, the authors show that for a generalized entropy
+we can control each in isolation. 
+\cite{kwon2024debiased} shows that for a generalized entropy
 function $G(w)$, including a term of $g(\pi_{2i|1}^{-1})$ into the calibration
@@
 We follow the approach of \cite{kwon2024debiased} for the debiased calibration
-method. Let $G: \mathcal{V} \to \R$ be the generalized entropy function that is strictly convex, differentiable.  
-We consider maximizing  
+method.  Let $G: \mathcal{V} \to \R$ be the generalized entropy function that is
+strictly convex and differentiable. We consider maximizing
+
 \begin{equation}\label{eq:primalloss}
-  H(w) = - \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i
-\end{equation} 
+  H(w_2) = - \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i
+\end{equation}
+
 subject to the following constraints:
+\textcolor{red}{(Please do not put a space between a displayed equation and the following sentence, unless you want to make it as a separate paragraph. ) }
 
 \begin{equation}\label{eq:calconst1}
-  \sum_{i \in A_2} \frac{\bf x_i w_{2i}q_i}{\pi_{1i}} = 
-\sum_{i \in A_1} \frac{\bf x_iq_i}{\pi_{1i}}
+  \sum_{i \in A_2} \frac{\bf x_i w_{2i}}{\pi_{1i}} = 
+\sum_{i \in A_1} \frac{\bf x_i}{\pi_{1i}}
 \end{equation}
@@
 \end{equation}
+
 where  $\bf z_i = (\bf x_i / q_i, g(\pi_{2i|1}^{-1}))$. The resulting 
@@
 
-%\subsection{Theoretical Results}
+\subsection{Theoretical Results}
 
 We can understand the estimated weights, $\hat w_{2i}$ from
-Equation~\ref{eq:primal}
-using the method of Legrange multipliers.
-We need to minimize the Lagrangian function
+Equation~\ref{eq:primal} using the method of Lagrange multipliers. We need to
+minimize the Lagrangian function
 
@@
   L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
-  + \bm \lambda \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
+  + \bm \lambda^\top \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
     \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i\right)
 \end{equation}
+
 where $\bm \lambda$ is a vector of Lagrange multipliers.
 Differentiating with respect to $w_{2i}$ and setting this expression equal to
-zero, yields the fact that $\hat w_{2i}$ satisfies
-
-$$ \hat w_{2i}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda}^T \bf z_i) $$
-
-where $\hat{\bm \lambda}$ is the solution to
+zero, we can obtain
+
+$$ \hat w_{2i} = g^{-1}(\hat{\bm \lambda}^\top \bf z_i) $$
+
+where $g(w) = \partial G / \partial w$ and 
+$\hat{\bm \lambda}$ is the solution to
 
@@
 
-The first result that we show is that the construction of the weights in
+The following theorem shows that the construction of the weights in
 Equation~\ref{eq:primal} leads to a design consistent estimator 
@@
   Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$. 
-  Under some regularity conditions,
-
-  $$\hat Y_{\rm DCE} = \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + O_p(N / n_2)$$
+  Under the regularity conditions stated in the appendix
+
+  $$
+  \hat Y_{\rm DCE} = \hat Y_\ell(\bm{\lambda}^*, \bm \phi^*) + O_p(N / n_2)
+  $$
 
@@
   $$
-  \hat Y_{\ell}(\bm \lambda, \bm \phi^*) = \hat Y_{\rm DCE}(\hat{\bm \lambda}) + 
-  \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i}(
-  \hat{\bm \lambda}) \bf z_i q_i\right)\bm \phi^*,
-  $$
-\textcolor{red}{(The above expression is not good.
-For example, the left-hand side is a function of $\bm \lambda$ and the right-hand side is a function of $\hat{\bm \lambda}$.
-Also, since we have $\bm \lambda^*=( \mathbf{0}, 1)$, the above expression can be simplified further.
-That is, there is no need to use $\bm \lambda^*$ in the above expression.
-) }
+  \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) = 
+  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} y_i
+  + \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - 
+  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} \bf{z}_i q_i\right)\bm \phi^*,
+  $$
+
   and
@@
   \bm \phi^* = 
-  \left[\sum_{i \in U} \frac{\pi_{2i|1}\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
+  \left[\sum_{i \in U} \frac{\pi_{2i|1}\bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}\right]^{-1}
   \sum_{i \in U} \frac{\pi_{2i|1}\bf z_i y_i}{g'(d_{2i|1})}.
@@
 
-   \begin{align*}
-   \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) 
-   &= \hat{Y}_{\rm DCE} + \left(\sum_{i \in A_1} w_{1i} \bf x_i  -
-     \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \bf x_i  \right)^T \bm \phi_1^* \\
-   &\quad+ \left(\sum_{i \in A_1} w_{1i} g_i  -
-  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^T \phi_2^*  
-  \end{align*} 
-\textcolor{red}{(You introduced $\hat{Y}_{\ell}$ twice!)}
-  and
-
-  $$
-  \begin{pmatrix}
-  \bm \phi_1^* \\
-  \phi_2^* 
-  \end{pmatrix}
-  = \left[ \sum_{i \in U} \frac{\pi_{2i \mid 1} }{ g'(d_{2i|1}) q_i} 
-  \begin{pmatrix}
-  \bx_i \bx_i^T &   \bx_i g_i   \\
-  g_i  \bx_i^T   & g_i^2     
-  \end{pmatrix} \right]^{-1}
-  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} 
-  \begin{pmatrix}
-    \bx_i \\ g_i 
-  \end{pmatrix}y_i 
-  $$
-
-with $g_i = g( \pi_{2i |1}^{-1}) q_i$.
+%   \begin{align*}
+%   \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) 
+%   &= \hat{Y}_{\rm DCE} + \left(\sum_{i \in A_1} w_{1i} \bf x_i  -
+%     \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \bf x_i  \right)^\top \bm \phi_1^* \\
+%   &\quad+ \left(\sum_{i \in A_1} w_{1i} g_i  -
+%  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^\top \phi_2^*  
+%  \end{align*} 
+%\textcolor{red}{(You introduced $\hat{Y}_{\ell}$ twice!)}
+%  and  
+%
+%  $$
+%  \begin{pmatrix}
+%  \bm \phi_1^* \\
+%  \phi_2^* 
+%  \end{pmatrix}
+%  = \left[ \sum_{i \in U} \frac{\pi_{2i \mid 1} }{ g'(d_{2i|1}) q_i} 
+%  \begin{pmatrix}
+%  \bx_i \bx_i^\top &   \bx_i g_i   \\
+%  g_i  \bx_i^\top   & g_i^2     
+%  \end{pmatrix} \right]^{-1}
+%  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} 
+%  \begin{pmatrix}
+%    \bx_i \\ g_i 
+%  \end{pmatrix}y_i 
+%  $$
+
+% with $g_i = g( \pi_{2i |1}^{-1}) q_i$.
 \end{theorem}
 
-\textcolor{red}{The presentation should be improved further. Since we
-
-}
-
-\begin{proof}
-  In this proof, we derive the solution to Equation~\ref{eq:primal} and show
-  that it is asymptotically equivalent to a regression estimator. Using the
-  method of Lagrange multipliers, to solve Equation~\eqref{eq:primal} we need to
-  minimize the Lagrangian in Equation~\eqref{eq:legragedc1}. 
-  The first order conditions show that
-
-  $$
-  \frac{\partial \mathcal{L}}{\partial w_{2i}}: g(w_{2i}) w_{1i}q_i -
-  w_{1i} \bm \lambda \bf z_i q_i = 0.
-  $$
-
-  Hence, $\hat w_{2i}(\bm \lambda) = g^{-1}(\bm \lambda^T \bf z_i)$ and
-  $\hat{\bm \lambda}$ is determined by Equation~\eqref{eq:lamdc1}. When the
-  sample size gets large, we have $\hat w_{2i}(\hat{\bm \lambda}) \to
-  d_{2i|1}$ which means that $\hat{\bm \lambda} \to \bm \lambda^*$ where $\bm
-  \lambda^* = (\bf 0, 1)$. \textcolor{red}{(You should prove it rather than just saying it. ) }  
-  Then using the linearization technique of
-  \cite{randles1982asymptotic}, we can construct a regression estimator,
-
-  $$
-  \hat Y_\ell(\hat{\bm \lambda}, \bm \phi)  = \hat Y_{\rm DCE}(\hat{\bm \lambda}) + 
-  \left(\sum_{i \in A_1} w_{1i} \bm z_i q_i - \sum_{i \in A_2} w_{1i} 
-  \hat w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right)\bm \phi.
-  $$
-
-  Notice that $\hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = \hat Y_{\rm DCE}(\hat{\bm
-  \lambda})$ for all $\bm \phi$. We choose $\bm \phi^*$ such that
-
-  $$
-  E\left[\frac{\partial}{\partial \bm \lambda} \hat Y_\ell(\bm \lambda^*,
-  \bm \phi^*)\right]=0.
-  $$
-
-  Using the fact that $g^{-1}(\bm \lambda^* \bf z_i) = g^{-1}(g(d_{2i|1})) = d_{2i|1}$
-  and $(g^{-1})'(x) = 1 / g'(g^{-1}(x))$, we have
-
-  \begin{align*}
-    \bm \phi^*
-    &= E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}
-    \right]^{-1}
-    E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i y_i}{g'(d_{2i|1})}\right]\\
-    &= \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}
-    \right]^{-1}
-    \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i y_i}{g'(d_{2i|1})}\right]\\
-  \end{align*}
-
-  Thus, the linearization estimator is
-
-  $$\hat Y_\ell(\bm \lambda^*, \bm \phi^*) = 
-  \sum_{i \in A_1} w_{1i} q_i \bf z_i \bm \phi^* +
-  \sum_{i \in A_2} w_{1i} d_{2i|1} (y_i - q_i \bf z_i \bm \phi^*).$$
-
-  By construction using a Taylor expansion yields,
-
-  \begin{align*}
-    \hat Y_{DCE}(\hat{\bm \lambda}) 
-    &= \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + 
-    E\left[\frac{\partial}{\partial \bm \lambda}\hat Y_\ell(\bm \lambda^*,
-    \bm \phi^*)\right](\hat{\bm \lambda} - \bm \lambda^*) + \frac{1}{2}
-    E\left[\frac{\partial}{\partial \bm \lambda^2} \hat Y_{DCE}(\bm \lambda^*)\right] 
-    (\hat{\bm \lambda} - \bm \lambda^*)^2\\
-    &= \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + O(N)O_p(n_2^{-1}).
-  \end{align*}
-
-  The final equality comes from the fact that 
-  $E\left[\frac{\partial}{\partial \bm \lambda}
-  \hat Y_\ell(\bm \lambda^*, \bm \phi^*)\right] = 0$,
-  $\frac{\partial}{\partial \bm \lambda^2} \hat w_{2i}(\bm \lambda^*)$ is
-  bounded and $||\hat{\bm \lambda} - \bm \lambda^*|| = O_p(n_2^{-1/2})$, which
-  proves our result.
-\end{proof}
-
-\textcolor{red}{(Please move the proof to Appendix. Also, we need to discuss the implication of the theorem. Specifically, what is the expectation and variance of the linearization term? ) }
-
-To construct asymptotically unbiased estimate of the variance we assume that the
-sampling designs for $A_1$ and $A_2$ are measurable and we define the following
-joint inclusion probabilities:
-
-$\Pr(i \in A_1, j \in A_1) = \pi_{1ij}, \text{ and }
-\Pr(i \in A_2, j \in A_2 \mid i \in A_1, j \in A_1) = \pi_{2ij|1}.$
+See the appendix for the proof of Theorem~\ref{thm:dc1}. Notice that this
+theorem implies
+that $\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ and $\hat Y_{\rm DCE}$ have
+the same asymptotic expectation and variance, 
+which means that we can estimate
+the variance of $\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ to get an
+estimate of the variance of $\hat Y_{\rm DCE}$. The basic idea of 
+Theorem~\ref{thm:dc1} is that asymptotically the two-phase debiased calibration
+method is equivalent to a regression estimator. By constraints 
+(\ref{eq:calconst1}) and (\ref{eq:calconst2}), 
+$\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ is design consistent, and the
+variance is asymptotically equivalent to the two-phase regression estimator. 
+We can estimate $\bm \phi^*$ with
+\begin{equation}
+\hat{\bm \phi} = 
+\left[\sum_{i \in A_2} 
+w_{1i}\frac{\bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}\right]^{-1}
+\sum_{i \in A_2} w_{1i}\frac{\bf z_i y_i}{g'(d_{2i|1})}.
+\label{eq:8}
+\end{equation}
+Therefore, the proposed calibration estimator in (\ref{eq:primal}) is asymptotically equivalent to the two-phase regression estimator given by 
+  $$
+  \hat Y_{\rm tp, reg}  = 
+  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} y_i
+  + \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - 
+  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} \bf{z}_i q_i\right)\hat{\bm \phi},
+  $$
+  where $\hat{\bm \phi}$ is defined in (\ref{eq:8}).
+
+For specific sample designs, we can construct the optimal entropy function
+$G(\cdot)$. From Theorem~\ref{thm:dc1}, the choice of entropy function affects
+the variance of the debiased calibration estimator through the form of $g'(d_{2i|1})$
+in $\bm \phi^*$. Consider a class of generalized regression estimators,
+
+$$
+\mathbf{B} = 
+\left\{\sum_{i \in A_1} w_{1i} \bf{z}_i^\top \bm \gamma + 
+\sum_{i \in A_2} \frac{w_{1i}}{\pi_{2i|1}}(y_i - \bf{z}_i^\top \bm \gamma):
+\bm \gamma \in \R^{p+1}\right\}
+$$
+
+Then the design optimal estimator is
+
+$$
+\hat \theta_{\ell, \rm{opt}} = 
+\sum_{i \in A_1} w_{1i} \bf{z}_i^\top \bm \phi_{\rm opt} + 
+\frac{w_{1i}}{\pi_{2i|1}}\sum_{i \in A_2} (y_i - \bf{z}_i^\top \bm \phi_{\rm opt}):
+$$
+
+where
+
+$$
+\bm \phi_{\rm opt} = 
+\Var\left(\sum_{i \in A_2} d_{2i|1} \bf{z}_i\right)^{-1} 
+\Cov\left(\sum_{i \in A_2} d_{2i|1} \bf{z}_i, \sum_{i \in A_2}, d_{2i|1} y_i \right).
+$$
+
+We assume that both $A_1$ and $A_2$ are Poisson samples. Then,
+
+$$
+\bm \phi_{\rm opt} = 
+\left\{
+\sum_{i \in U} \left(d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1)\right) \bf{z_i}\bf{z_i}^\top
+\right\}^{-1}
+\sum_{i \in U} \left(d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1)\right) \bf{z_i} y_i.
+$$
+
+To find the optimal entropy function, we solve the differential equation
+
+$$
+\frac{1}{d_{2i|1} g'(d_{2i|1})} = d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1).
+$$
+
+Using the fact that $(y(xy-1))^{-1} = -y^{-1} + x(xy-1)^{-1}$, the optimal
+entropy function is
+
+$$
+G(w_{2i|1}) = - \log(w_{2i|1}) + \log(w_{2i|1}w_{1i} - 1).
+$$
+
+To construct an asymptotically unbiased estimate of the variance we assume that
+the sampling designs for $A_1$ and $A_2$ are measurable and we define the
+following joint inclusion probabilities:
+
+$$
+\Pr(i \in A_1, j \in A_1) = \pi_{1ij}, \text{ and }
+\Pr(i \in A_2, j \in A_2 \mid i \in A_1, j \in A_1) = \pi_{2ij|1}.
+$$
+
+The next theorem discusses how to compute an estimate for the variance of 
+$\hat Y_{\rm DCE}$. For the proof see the appendix.
 
@@
   + \sum_{i \in A_2} \sum_{j \in A_2} w_{1i} \frac{\Delta_{2ij|1}}{\pi_{2ij|1}} 
-  \frac{(y_i - \bf z_i q_i \bm \phi^*)}{\pi_{2i|1}}
-  \frac{(y_j - \bf z_j q_j \bm \phi^*)}{\pi_{2j|1}}
+  \frac{(y_i - \bf z_i q_i \hat{\bm \phi})}{\pi_{2i|1}}
+  \frac{(y_j - \bf z_j q_j \hat{\bm \phi})}{\pi_{2j|1}}
   $$
@@
 
-\begin{proof}
-  By Theorem~\ref{thm:dc1}, $\hat Y_{\rm DCE}$ is asymptotically equivalent to
-  $\hat Y_{\ell, \rm{DCE}}$. Notice that we can define
-
-  $$
-  \eta_i^* = 
-  \bf z_i q_i \bm \phi^* + \frac{\delta_i}{\pi_{2i|1}}(y_i - \bf z_i q_i \bm \phi^*)
-  $$
-
-  where $\delta_i$ is one if $i \in A_2$ and zero otherwise. Then,
-
-  $$
-  \hat Y_{\ell, \rm{DCE}} = \sum_{i \in A_1} w_{1i} \eta_i^*.
-  $$
-
-  Then using the reverse framework of \cite{fay1992inferences} and 
-  \cite{shao1999variance}, and noting that 
-  $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ and 
-  $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$,
-
-  \begin{align*}
-    \Var(\hat Y_{\ell, \rm{DCE}}) 
-    &= \Var\left(\sum_{i \in A_1} w_{1i} \eta_i^*\right) \\
-    &= \Var\left(E\left[\sum_{i \in A_1} w_{1i} \eta_i^* \mid A_2 \right]\right) 
-    + E\left[\Var\left(\sum_{i \in A_1} w_{1i} \eta_i^* \mid A_2 \right)\right] \\
-    &= \Var\left(\sum_{i \in U} \eta_i^*\right) 
-    + E\left[\sum_{i \in U} \sum_{j \in U} \Delta_{1ij} w_{1i} \eta_i^* w_{1j}
-      \eta_j^*\right]\\
-  \end{align*}
-
-  This means that we can estimate the variance of $\hat Y_{\rm DCE}$ with
-
-  $$
-  \hat V_{\rm DCE} 
-  = \sum_{i \in A_1} \sum_{j \in A_1} \frac{\Delta_{1ij}}{\pi_{1ij}} 
-  \hat \eta_i \hat \eta_j
-  + \sum_{i \in A_2} \sum_{j \in A_2} w_{1i} \frac{\Delta_{2ij|1}}{\pi_{2ij|1}} 
-  \frac{(y_i - \bf z_i q_i \bm \phi^*)}{\pi_{2i|1}}
-  \frac{(y_j - \bf z_j q_j \bm \phi^*)}{\pi_{2j|1}}
-  $$
-
-  where $\hat \eta_i = \bf z_i q_i \hat{\bm \phi} + 
-  \frac{\delta_i}{\pi_{2i|1}}(y_i - \bf z_i q_i \hat{\bm \phi})$.
-
-\end{proof}
-
 \section{Extensions}
+
 \subsection{Non-nested Two-Phase Sampling}
@@
 in the non-nested two-phase sample we have two independent Horvitz-Thompson
-estimators of the total of $\bf X$,
+estimators of the total of $\bf X_N$,
 
@@
 
-where $d_{1i} = \pi_{1i}^{-1}$, $d_{2i} = \pi_{2i}^{-1}$, $\pi_{1i}$ is the
-probability of $i \in A_1$ and $\pi_{2i} = \Pr(i \in A_2)$. 
+where $d_{1i} = \pi_{1i}^{-1}$, $d_{2i} = \pi_{2i}^{-1}$, 
+$\pi_{1i} = \Pr(i \in A_1)$ and $\pi_{2i} = \Pr(i \in A_2)$. 
 We can combine these estimates using the effective sample size 
-(\cite{kish1965survey}) to get
+\citep{kish1965survey} to get
 $\hat{\bf X}_c = (n_{1, \rm{eff}} \hat{\bf X}_1 + n_{2, \rm{eff}}\hat{\bf X}_2) / 
@@
 $$
-\hat Y_{\rm NN, reg} = \hat Y_2 + (\hat{\bf X}_c - \hat{\bf X}_2)^T \hat{\bm \beta_q} = 
-\hat Y_2 + (\hat{\bf X}_1 - \hat{\bf X}_2)^T W\hat{\bm \beta_q}
+\hat Y_{\rm NN, reg} = \hat Y_2 + (\hat{\bf X}_c - \hat{\bf X}_2)^\top \hat{\bm \beta_q} = 
+\hat Y_2 + (\hat{\bf X}_1 - \hat{\bf X}_2)^\top W\hat{\bm \beta_q}
 $$
@@
 \hat{\bm \beta_q} = 
-\left(\sum_{i \in A_2} \frac{\bf x_i \bf x_i^T}{q_i}\right)^{-1} 
+\left(\sum_{i \in A_2} \frac{\bf x_i \bf x_i^\top}{q_i}\right)^{-1} 
 \sum_{i \in A_2} \frac{\bf x_i y_i}{q_i} \text{ and }
@@
 From an optimality perspective the choice of using the effective sample size to
-weight the estimates from $A_1$ and $A_2$ is reasonable because often the
-effective sample sizes are proportional to the variance of the estimates 
-of $\bf X$. While we know that the inverse variance weighted estimate is optimal
+weight the estimates from $A_1$ and $A_2$ is reasonable because the
+effective sample sizes are often proportional to the variance of the estimates 
+of $\bf X_N$. While the inverse variance weighted estimate is optimal
 to combine independent samples for a linear estimate, using the effective sample
 size approximates this procedure without requiring the actual variance of an
-estimator to be known.
-
-Since the samples $A_1$ and $A_2$ are independent,
+estimator to be known. Since the samples $A_1$ and $A_2$ are independent,
 
@@
 V(\hat Y_{\rm NN, reg}) = 
-V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i - \bf x_i^TW\bm \beta^*_q)\right)
-+ (\bm \beta^*_q)^T W^T V(\hat{\bf X}_1) W \bm \beta_q^*
+V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i - \bf x_i^\top W\bm \beta^*_q)\right)
++ (\bm \beta^*_q)^\top W^\top V(\hat{\bf X}_1) W \bm \beta_q^*
 $$
@@
 where $\bm \beta_q^*$ is the probability limit of $\hat{\bm \beta_q}$. Like the
-two-phase sample this regression estimator can be viewed as the solution to the
+two-phase sample, this regression estimator can be viewed as the solution to the
 following calibration equation
@@
 \begin{equation}\label{eq:nncal}
-  \hat w_2 = \argmin_{w_2} Q(w_2) = \sum_{i \in A_2} (w_{2i} - d_{2i})^2 q_i 
+  \hat w_2 = \argmin_{w_2} \sum_{i \in A_2} (w_{2i} - d_{2i})^2 q_i 
   \text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X}_c
@@
 \begin{equation}\label{eq:nnopt}
-\hat w_2 = \argmin_w \sum_{i \in A_2} G\left(w_{2i}\right) q_i 
-\text{ with } 
+\hat w_2 = \argmin_{w_2} \sum_{i \in A_2} G\left(w_{2i}\right) q_i 
+\text{ such that } 
 \sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X}_{c} \text{ and } 
@@
 need to estimate it from the non-nested samples. We can get multiple estimates
-of $\hat{\bf X}$,
+of $\bf X_N$,
 
@@
 
-We can constuct a non-nested two-phase estimator $\hat Y_{\rm NNE}$ for $Y_N$ where
+We can construct a non-nested two-phase estimator $\hat Y_{\rm NNE}$ for $Y_N$ where
 $\hat Y_{\rm NNE} = \sum_{i \in A_2} \hat w_{2i} y_i$ and $\hat w_{2i}$ solves
@@
   L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} G(w_{2i}) q_i + 
-  \bm \lambda \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i} \bf z_i q_i\right).
+  \bm \lambda^\top \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i} \bf z_i q_i\right).
 \end{equation}
@@
 Differentiating with respect to $w_{2i}$ and setting this expression equal to
-zero, yields the fact that $\hat w_{2i}$ satisfies
-
-$$ \hat w_{2i}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda^T} \bf z_i) $$
+zero yields the fact that $\hat w_{2i}$ satisfies
+
+$$ \hat w_{2i}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda}^\top \bf z_i) $$
 
@@
 \end{equation}
+
+Like the previous section we first prove that the estimator is design consistent.
+Then we provide a result showing how to estimate the variance.
 
@@
   $$
-  \hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*) = \sum_{i \in A_2} \hat w_{2i}(
-  \bm \lambda^*) y_i + \left(\hat{\bf T} - \sum_{i \in A_2} \hat w_{2i}(
-  \bm \lambda^*) \bf z_i q_i\right) \bm \phi^*
+  \hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*) = 
+  \sum_{i \in A_2} \pi_{2i}^{-1} y_i + 
+  \left(\hat{\bf T} - \sum_{i \in A_2} \pi_{2i}^{-1} \bf z_i q_i\right)^\top \bm \phi^*
   $$
@@
 
-\begin{proof}
 The proof of this result is very similar to the proof the
-  Theorem~\ref{thm:dc1}. The biggest difference is that the total for $\bf X$ is
+Theorem~\ref{thm:dc1}. The biggest difference is that the total for $\bf X_N$ is
 estimated from both samples using $\hat{\bf X}_c$ instead of 
 $\hat{\bf X}_{HT}$ from the Phase 1 sample. 
-
-  Since $\hat Y_{\rm NNE} = \sum_{i \in A_2} \hat w_{2i}(\hat{\bm \lambda})y_i$ where
-  $\hat{\bm \lambda}$ solves
+Since $\hat Y_{\rm NNE} = \sum_{i \in A_2} \hat w_{2i}(\hat{\bm \lambda})y_i$
+where $\hat{\bm \lambda}$ solves
 
@@
 
-  $$\hat Y_{\ell, \rm NNE} (\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} \hat
+$$\hat Y_{\rm NNE} (\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} \hat
 w_{2i}(\hat{\bm \lambda})y_i + \left(\bf T - \sum_{i \in A_2} \hat
-  w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right) \bm \phi.$$
+w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right)^\top \bm \phi.$$
 
 If we choose $\bm \phi^*$ such that $E\left[\frac{\partial}{\partial \bm \lambda} 
-    \hat Y_{\ell, NNE}(\bm \lambda^*, \bm \phi^*)\right] = 0$, then
+  \hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*)\right] = 0$, then
 
@@
 
-\end{proof}
+The next theorem shows how to estimate the variance of $\hat Y_{\rm NNE}$.
+The proof of this result is in the appendix.
 
@@
   \begin{align*}
-    \Var(\hat Y_{\rm NNE}(\hat \lambda))
-    &= (\bm \phi_1^*)^T \Var(\hat{\bf X}_c) \bm \phi_1^* + 
+    \Var(\hat Y_{\rm NNE}(\hat{\bm \lambda}))
+    &= (\bm \phi_1^*)^\top \Var(\hat{\bf X}_c) \bm \phi_1^* + 
     \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
@@
     \hat V_{\rm NNE} 
-    &= (\hat{\bm \phi_1})^T \Var(\hat{\bf X}_c) \hat{\bm \phi_1} + 
+    &= (\hat{\bm \phi_1})^\top \Var(\hat{\bf X}_c) \hat{\bm \phi_1} + 
     \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
@@
 
-\begin{proof}
-  From Theorem~\ref{thm:dc2}, we know that $\hat Y_{\rm NNE}(\hat \lambda) = 
-  \hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*) + O_p(Nn_2^{-1})$. Hence, the
-  variance of $\hat Y_{\rm NNE}(\hat{\bm \lambda})$ is
-
-  \begin{align*}
-    \Var(\hat Y_{\rm NNE}(\hat{\bm \lambda})) 
-    &\doteq \Var(\hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*)) \\ 
-    &= \Var\left(\sum_{i \in A_2} \hat w_{2i}(\bm \lambda^*) y_i + 
-      \left(\bf T - \sum_{i \in A_2} \hat w_{2i}(\bm \lambda^*) \bf z_i
-    q_i\right)\bm \phi^*\right)\\
-    &= (\bm \phi_1^*)^T \Var(\hat{\bf X}_c) \bm \phi_1^* + 
-    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
-    (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
-    &\qquad+ 2 \Cov\left(\hat{\bf X}_c \bm \phi_1^*, \sum_{i \in A_2} 
-      \frac{(y_i - \bf z_i \bm \phi^* q_i)}{\pi_{2i}}\right) \\
-    &= (\bm \phi_1^*)^T \Var(\hat{\bf X}_c) \bm \phi_1^* + 
-    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
-    (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
-    &\qquad + (1 - W)\bm \phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij} 
-    \frac{\bf{x}_i}{\pi_{2i}} \frac{(y_j - \bf z_j \bm \phi^*_1 q_j)}{\pi_{2j}}\\
-  \end{align*}
-
-where the last equality comes from the fact that $\hat{\bf X}_c = W\hat{\bf
-X}_1 + (1 - W) \hat{\bf X}_2$. To have an unbiased estimator of the variance
-  we can use:
-
-  \begin{align*}
-    \hat V_{\rm NNE} 
-    &= (\hat{\bm \phi_1})^T \Var(\hat{\bf X_c}) \hat{\bm \phi_1} + 
-    \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
-    (y_i - \bf z_i \hat{\bm \phi} q_i)(y_j - \bf z_j \hat{\bm \phi} q_j) \\
-    &\qquad + (1 - W)\hat{\bm \phi}_1 \sum_{i \in A_2} \sum_{j \in A_2} 
-    \frac{\Delta_{2ij}}{\pi_{2ij}} \frac{\bf{x}_i}{\pi_{2i}} 
-    \frac{(y_j - \bf z_j \hat{\bm \phi_1} q_j)}{\pi_{2j}}
-  \end{align*}
-
-  where
-
-  $$\hat{\bm \phi} =
-  \begin{bmatrix}
-    \hat{\bm \phi_1} \\ \hat \phi_2
-  \end{bmatrix} = 
-  \left(\sum_{i \in A_2} \frac{q_i}{g'(d_{2i})} 
-  \begin{bmatrix}
-    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
-    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
-  \end{bmatrix}
-  \right)^{-1}
-  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})} 
-  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
-  $$
-
-\end{proof}
-
 \subsection{Multi-Source Two-Phase Sampling}
@@
 When considering non-nested two-phase sampling, we focused on the case of having
-two samples. Now, we consider incorporating more than two independent samples
-together with a debiasing constraint. We consider the case in which we
-want to estimate the $\theta = E[Y]$ and $Y$ is only observed in one sample.
+two samples. Now, we incorporate more than two independent samples
+together with a debiasing constraint. We examine the case in which we
+want to estimate $\theta = E[Y]$ where $Y$ is only observed in one sample.
 
@@
 where $Y$ is observed only in $A_0$ but $\bf X$ is observed in all of the
-samples with elements $\bf X^{(0)}$ observed in $A_0$ and $\bf X^{(m)}$ observed
-in $A_m$ for each $m = 1, \dots, M$. Like the non-nested case, we assume that
-each survey is sampling independently from the same population sampling frame.
-
-The traditional multi-source approach (\cite{kim2024statistics}) is to use
-generalized least squares (GLS) to obtain an optimal estimator of $X$ from the
+samples with variables $\bf X^{(0)}$ observed in $A_0$ and variables $\bf X^{(m)}$
+observed in $A_m$ for each $m = 1, \dots, M$. Like the non-nested case, we
+assume that each survey is sampled independently from the same population
+sampling frame.
+
+The traditional multi-source approach \citep{kim2024statistics} is to use
+GLS to obtain an optimal estimator of $\bf X_N$ from the
 samples $A_1, \dots, A_M$. Then we can incorporate this information in the
@@
 \hat w = \argmin_w \sum_{i \in A_0} G(w_i)q_i \text{ such that } 
-\sum_{i \in A_0} w_i \bf x_i = \hat X_{\rm{GLS}}^{(0)}
-\end{equation}
-
-where $\hat{\bf X}_{\rm{GLS}}^{(0)}$ is the GLS estimate of $X$ for all of the
-$X$-variables measured in sample $A_0$, and $G$ is a generalized entropy
+\sum_{i \in A_0} w_i \bf x_i = \hat{\bf X}_{\rm{GLS}}^{(0)}
+\end{equation}
+
+where $\hat{\bf X}_{\rm{GLS}}^{(0)}$ is the GLS estimate of $\bf X$ for all of
+the $X$-variables measured in sample $A_0$, and $G(\cdot)$ is a generalized entropy
 function. 
-
 In order to have a design consistent estimator that incorporates information from
-samples $A_0, A_1, \dots, A_M$, we want to add the debiasing constraints,
+samples $A_0, A_1, \dots, A_M$, we want to add the debiasing constraint,
 
 \begin{equation}\label{eq:msdebconstr}
-\sum_{i \in A_0} w_{i} \bf x_i q_i = \sum_{i \in U} g(d_i) q_i
+\sum_{i \in A_0} w_{i} g(d_i) q_i = \sum_{i \in U} g(d_i) q_i
 \end{equation}
@@
 \begin{equation}\label{eq:dc3}
-  \hat w_0 = \argmin_w \sum_{i \in A_0} G(w_{0i}) q_i \text{ such that }
+  \hat w_0 = \argmin_{w_0} \sum_{i \in A_0} G(w_{0i}) q_i \text{ such that }
   \widehat{\bf{T}} = \sum_{i \in A_0} w_{0i} \bf z_i.
@@
 where $\widehat{\bf{T}} = (\widehat{\bf X}_{\rm{GLS}}^{(0)},
-\sum_{i \in U} g(d_i) q_i)^T$,
-and $\bf z_i = ((\bf x_i^{(0)})^T, g(d_i)q_i)^T$. Then
-the estimator is $\hat Y_{\rm MS} = \sum_{i \in A_0} \hat w_{0i} y_i$ and we can construct
-this estimate minimizing the Lagrangian:
-
-$$
-L(w_{0}, \bm \lambda) = \sum_{i \in A_0} G(w_{0i}) q_i + {\bm \lambda}^T
+\sum_{i \in U} g(d_i) q_i)^\top$,
+and $\bf z_i = ((\bf x_i^{(0)})^\top, g(d_i)q_i)^\top$. Then
+the estimator is $\hat Y_{\rm MS} = \sum_{i \in A_0} \hat w_{0i} y_i$ and we can
+construct this estimate minimizing the Lagrangian:
+
+$$
+L(w_{0}, \bm \lambda) = \sum_{i \in A_0} G(w_{0i}) q_i + {\bm \lambda}^\top
 \left(\widehat{\bf{T}} - \sum_{i \in A_0} w_{0i} \bf z_i\right)
@@
 
-where $\bm \lambda$ are the Lagrange multipliers.
-
-%\subsection{Theoretical Results}
+where $\bm \lambda$ are the Lagrange multipliers from Equation~\ref{eq:dc3}.
+We now show that $\hat Y_{\rm MS}$ is design consistent.
 
@@
 
-  $$\hat Y_{\rm MS} = \hat Y_{\ell, \rm MS}(\bm \lambda^*, \bm \phi^*) + O_p(N / n_0)$$
+  $$
+  \hat Y_{\rm MS} = \hat Y_{\ell, \rm MS}(\bm \lambda^*, \bm \phi^*) + O_p(N / n_0)
+  $$
 
@@
   $$
-  \hat Y_{\ell, \rm MS}(\bm \lambda^*, \bm \phi^*) = \sum_{i \in A_0} \hat w_{0i} y_i + \left(
-    \widehat{\bf T} - \sum_{i \in A_0} \hat w_{0i}(\bm \lambda^*) \bf z_i q_i
-  \right)\bm \phi^*
+  \hat Y_{\ell, \rm MS}(\bm \lambda^*, \bm \phi^*) = 
+  \sum_{i \in A_0} d_{0i} y_i + \left(
+    \widehat{\bf T} - \sum_{i \in A_0} d_{0i} \bf z_i q_i
+  \right)^\top \bm \phi^*
   $$
@@
   \bm \phi^* = 
-  \left[\sum_{i \in U} \frac{\pi_{0i} \bf z_i \bf z_i^T q_i}{g'(d_{0i})}\right]^{-1}
+  \left[\sum_{i \in U} \frac{\pi_{0i} \bf z_i \bf z_i^\top q_i}{g'(d_{0i})}\right]^{-1}
   \sum_{i \in U} \frac{\pi_{0i} \bf z_i y_i}{g'(d_{0i})}.
@@
 
-\begin{proof}
 This proof follows from the proof of Theorem~\ref{thm:dc2}. The only
-  difference is that we have $\hat X_{\rm GLS}$ instead of $\hat X_c$.
-\end{proof}
+difference is that we have $\hat{\bf X}_{\rm GLS}$ instead of $\hat{\bf X}_c$.
+The next theorem gives a method for how to estimate the variance of 
+$\hat Y_{\rm MS}$.
 
@@
 
-\begin{proof}
 This result follows the same argument as Theorem~\eqref{thm:nnve}. The result
 holds because each sample $A_m$ is independent from the other $A_{m'}$.
-\end{proof}
+
+\textcolor{red}{(I think we need more explanations about the setup of Section 4.2. A toy example might be helpful. Please check Section 11.5 of my sampling book.) }
 
@@
 $(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. 
-If $\delta = 0$ then there is no model misspecification. However, if $\delta =
-1$, then there is some model misspecification. We estimate the
+If $\delta = 0$, then there is no model misspecification. However, if $\delta =
+1$, then there is some model misspecification. We estimate
 the parameter $\bar Y_N$ with four approaches:
@@
 \begin{itemize}
-  \item[1.] HT-estimator (HT): $\hat Y_{HT} = N^{-1} \sum_{i \in A_2}
+  \item[1.] HT-estimator (HT): $\hat Y_{\rm HT} = N^{-1} \sum_{i \in A_2}
     \frac{y_i}{\pi_{2i}},$
-  \item[2.] Regression estimator (Reg): Let $\hat Y_{NN, reg} = \hat Y_{HT} + (
-    \hat{\bf X_c} - \hat{\bf X_{2, HT}}) \hat{\bm \beta_2}$ where $\hat Y_{HT} =
-    \sum_{i \in A_2} d_{2i} y_i$, $\hat{\bf X_c} = W \hat{\bf X_{1, HT}} + (1 -
-    W)\hat{\bf X_{2, HT}}$, $W = n_{1,\rm{eff}} / (n_{1,\rm{eff}} + n_{2, \rm{eff}})$,
-    $\hat{\bf X_{1, HT}} = \sum_{i \in A_1} d_{1i} \bf x_{i}$, $\hat{\bf X_{2,
-    HT}} = \sum_{i \in A_2} d_{2i} \bf x_{i}$, $\bf x_i = (1, x_{1i}, x_{2i})^T$
+  \item[2.] Regression estimator (Reg): Let $\hat Y_{\rm NN, reg} = \hat Y_{\rm HT} + (
+  \hat{\bf X}_c - \hat{\bf X}_{2, \rm HT}) \hat{\bm \beta_2}$ where $\hat Y_{HT} =
+\sum_{i \in A_2} d_{2i} y_i$, $\hat{\bf X}_c = W \hat{\bf X}_{1, \rm HT} + (1 -
+W)\hat{\bf X}_{2, \rm HT}$, $W = n_{1,\rm{eff}} / (n_{1,\rm{eff}} + n_{2, \rm{eff}})$,
+$\hat{\bf X}_{1, \rm HT} = \sum_{i \in A_1} d_{1i} \bf x_{i}$, $\hat{\bf X}_{2,
+    \rm HT} = \sum_{i \in A_2} d_{2i} \bf x_{i}$, $\bf x_i = (1, x_{1i}, x_{2i})^\top$
     and
 
-    $$ \hat{\bm \beta_2} = \left(\sum_{i \in A_2} \bf x_i \bf x_i^T\right)^{-1}
+    $$ \hat{\bm \beta_2} = \left(\sum_{i \in A_2} \bf x_i \bf x_i^\top\right)^{-1}
     \sum_{i \in A_2} \bf x_i y_i.$$
 
-    Then $\hat{\bar Y}_{NN, reg} = \hat Y_{NN, reg} / N$.
+    Then $\hat{\bar Y}_{\rm NN, reg} = \hat Y_{\rm NN, reg} / N$.
   \item[3.] Debiased Calibration with Population Constraints (EstPop): This 
@@
 This simulation study tests the multi-source extension.
-We have the following superpopulation model with $N = 10000$ elements:
+We have the following superpopulation model with $N = 10,000$ elements:
 
@@
 
-As expected MSPop has the lowest RMSE because it also uses the population
-totals. It seems like MSEst and MSReg are almost equivalent, which is good
-because they are also supposed to be identical. The MSEst estimator outperforms
+As expected, EstPop has the lowest RMSE because it also uses the population
+totals. The Est estimator outperforms
 NNReg because it also uses information from $A_2$, even though it implicitly
 uses a regression estimator with $X_3$ as a covariate, which is unnecessary.
-\section{Real-Data Analysis}
+
+\section{Real Data Analysis}
 
@@
 65 years and older and select designated populations between the ages of 18 and
-65. However, the vast majority of Medicare patients are 65 years old or older.}
+65.}
 are known to have a higher all-cause mortality from
-COVID-19 (\cite{bonanad2020effect}). Thus, understanding how older people get
-vaccinated can help save lives and reduce medical costs.
+COVID-19 \citep{bonanad2020effect}. Thus, understanding what causes older people
+to get vaccinated can help save lives and reduce medical costs.
 
 The MCBS is a rotating panel survey in which participants are interviewed up to
-three times within a four-year period (\cite{cms2021mcbs}). The participants
+three times within a four-year period \citep{cms2021mcbs}. The participants
 are selected from Medicare enrollment data and enter a cohort in the fall round
@@
 least one COVID vaccination dose. This MCBS data was updated to include
-information about if a participant had received a second vaccine dose and then
-later to include how many total vaccine doses patients had recieved. If the
-participant had received at least one more dose and their answer was missing (because
-they refused to answer, did not know, or was just missing) the missing value was
-imputed to say that they yes, they had received a COVID vaccine dose. Otherwise,
-the missing value was imputed to say no, the participant had not received a
-COVID vaccine dose.
+information about whether a participant had received a second vaccine dose. 
+Later, the survey included the total vaccine doses patients had received. If the
+participant had received at least one more dose and their answer was missing
+(because they refused to answer, did not know, or was missing) the missing value
+was imputed to say that they yes, they had received a COVID vaccine dose.
+Otherwise, the missing value was imputed to say no, the participant had not
+received a COVID vaccine dose.
 
@@
 
-\section{Appendix A: Estimating the Population Weights}
+\newpage
+\appendix
+
+\section{Appendix: Regularity Conditions}
+
+We need the following regularity conditions to prove Theorem 1.
+
+\begin{itemize}
+  \item[A1]: $\mathcal{V} \in \R$ is open,
+  \item[A2]: $G(w)$ is strictly convex and twice continuously differentiable on
+    $\mathcal{V}$ with a bounded second derivative,
+  \item[A3]: There exist positives constants $c_1, c_2 \in \mathcal{V}$ such
+    that $c_1 < \pi_{2i|1} < c_2$ for all $\pi_i$,
+  \item[A4]: For $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$,
+    $$
+    \limsup_{N \to \infty} N \max_{i, j \in U, i \neq j} |\Delta_{2ij|1}| <
+    \infty,
+    $$
+  \item[A5]: The matrix $\bm \Sigma_{\bf z} := 
+    \lim_{N \to \infty} N^{-1} \sum_{i \in U} {\bf z}_i {\bf z}_i^\top$ exists
+    and is positive definite, 
+  \item[A6]: The average fourth moment of $({\bf x}_i^\top, y_i)^\top$ is finite
+    such that 
+    $$
+    \limsup_{N \to \infty} N^{-1} \sum_{i = 1}^N ({\bf x}_i^\top, y_i)^\top < \infty,
+    $$
+  \item[A7]: $||\hat{\bm \lambda} - \bm \lambda^*|| = O_p(n_2^{-1/2})$
+\end{itemize}
+
+Assumption A1 is needed for the differentiability of $G(w)$ in A2.
+Assumption A2 guarentees a unique solution for $\hat{\bm \lambda}$ and ensures
+that $g(w)$ is differentiable. The boundedness part of the assumption ensures
+that $\frac{\partial}{\partial \bm \lambda} g^{-1}(\bm \lambda^\top {\bf z})$ is
+also bounded. Conditions A3 and A4 are standard conditions which control the
+first-order and second-order inclusion probability for the second phase sample.
+These conditions are met by many traditional survey designs. Assumptions A5 and
+A6 are technical conditions and common in the survey literature. Assumption A7
+is added to simplify the proof. We could remove this condition and follow the
+approach of Theorem 1 in \cite{kwon2024debiased}.
+
+\subsection{Proof of Theorem 1}
+
+In this proof, we derive the solution to Equation~\ref{eq:primal} and show
+that it is asymptotically equivalent to a regression estimator. Using the
+method of Lagrange multipliers, to solve Equation~\eqref{eq:primal} we need to
+minimize the Lagrangian in Equation~\eqref{eq:legragedc1}. 
+The first order conditions show that at the solution,
+
+$$
+\frac{\partial \mathcal{L}}{\partial w_{2i}}: g(w_{2i}) w_{1i}q_i -
+w_{1i} \hat{\bm \lambda}^\top \bf z_i q_i = 0.
+$$
+
+Hence, $\hat w_{2i}(\hat{\bm \lambda}) = 
+g^{-1}(\hat{\bm \lambda}^\top \bf z_i)$ where
+$\hat{\bm \lambda}$ is determined by Equation~\eqref{eq:lamdc1}. When the
+sample size gets large, we have $\hat w_{2i}(\hat{\bm \lambda}) \to
+d_{2i|1}$ in probability because both terms in Equation~\eqref{eq:lamdc1} must
+become equal by the Central Limit Theorem. Hence,
+$\hat{\bm \lambda} \to \bm \lambda^*$ in probability and 
+$\bm \lambda^* = (\bf 0^\top, 1)^\top$. 
+Using the linearization technique of
+\cite{randles1982asymptotic}, we can construct a regression estimator,
+
+$$
+\hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = 
+\sum_{i \in A_2} w_{1i} \hat w_{2i} y_i + 
+\left(\sum_{i \in A_1} w_{1i} \bm z_i q_i - \sum_{i \in A_2} w_{1i} 
+\hat w_{2i} \bf z_i q_i\right)\bm \phi.
+$$
+
+Notice that $\hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} w_{1i}
+\hat w_{2i} y_i$ for all $\bm \phi$. We choose $\bm \phi^*$ such that
+
+$$
+E\left[\frac{\partial}{\partial \bm \lambda} \hat Y_\ell(\bm \lambda^*,
+\bm \phi^*)\right]=0.
+$$
+
+Using the fact that $g^{-1}(\bm \lambda^* \bf z_i) = g^{-1}(g(d_{2i|1})) = d_{2i|1}$
+and $(g^{-1})'(x) = 1 / g'(g^{-1}(x))$, we have
+
+\begin{align*}
+  \bm \phi^*
+  &= E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}
+  \right]^{-1}
+  E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i y_i}{g'(d_{2i|1})}\right]\\
+  &= \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}
+  \right]^{-1}
+  \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i y_i}{g'(d_{2i|1})}\right]\\
+\end{align*}
+
+Thus, the linearization estimator is
+
+$$\hat Y_\ell(\bm \lambda^*, \bm \phi^*) = 
+\sum_{i \in A_1} w_{1i} q_i \bf z_i \bm \phi^* +
+\sum_{i \in A_2} w_{1i} d_{2i|1} (y_i - q_i \bf z_i \bm \phi^*).$$
+
+By construction using a Taylor expansion yields,
+
+\begin{align*}
+  \hat Y_{DCE}(\hat{\bm \lambda}) 
+  &= \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + 
+  E\left[\frac{\partial}{\partial \bm \lambda}\hat Y_\ell(\bm \lambda^*,
+  \bm \phi^*)\right](\hat{\bm \lambda} - \bm \lambda^*) + \frac{1}{2}
+  E\left[\frac{\partial}{\partial \bm \lambda^2} \hat Y_{DCE}(\bm \lambda^*)\right] 
+  (\hat{\bm \lambda} - \bm \lambda^*)^2\\
+  &= \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + O(N)O_p(n_2^{-1}).
+\end{align*}
+
+The final equality comes from the fact that 
+$E\left[\frac{\partial}{\partial \bm \lambda}
+\hat Y_\ell(\bm \lambda^*, \bm \phi^*)\right] = 0$,
+$\frac{\partial}{\partial \bm \lambda^2} \hat w_{2i}(\bm \lambda^*)$ is
+bounded and $||\hat{\bm \lambda} - \bm \lambda^*|| = O_p(n_2^{-1/2})$, which
+proves our result.
+
+\subsection{Proof of Theorem 2}
+
+\begin{proof}
+  By Theorem~\ref{thm:dc1}, $\hat Y_{\rm DCE}$ is asymptotically equivalent to
+  $\hat Y_{\ell}$. Notice that we can define
+
+  $$
+  \eta_i^* = 
+  \bf z_i q_i \bm \phi^* + \frac{\delta_i}{\pi_{2i|1}}(y_i - \bf z_i q_i \bm \phi^*)
+  $$
+
+  where $\delta_i$ is one if $i \in A_2$ and zero otherwise. Then,
+
+  $$
+  \hat Y_{\ell} = \sum_{i \in A_1} w_{1i} \eta_i^*.
+  $$
+
+  Then using the reverse framework of \cite{fay1992inferences} and 
+  \cite{shao1999variance}, and noting that 
+  $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ and 
+  $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$,
+
+  \begin{align*}
+    \Var(\hat Y_{\ell, \rm{DCE}}) 
+    &= \Var\left(\sum_{i \in A_1} w_{1i} \eta_i^*\right) \\
+    &= \Var\left(E\left[\sum_{i \in A_1} w_{1i} \eta_i^* \mid A_2 \right]\right) 
+    + E\left[\Var\left(\sum_{i \in A_1} w_{1i} \eta_i^* \mid A_2 \right)\right] \\
+    &= \Var\left(\sum_{i \in U} \eta_i^*\right) 
+    + E\left[\sum_{i \in U} \sum_{j \in U} \Delta_{1ij} w_{1i} \eta_i^* w_{1j}
+      \eta_j^*\right]\\
+  \end{align*}
+
+  This means that we can estimate the variance of $\hat Y_{\rm DCE}$ with
+
+  $$
+  \hat V_{\rm DCE} 
+  = \sum_{i \in A_1} \sum_{j \in A_1} \frac{\Delta_{1ij}}{\pi_{1ij}} 
+  \hat \eta_i \hat \eta_j
+  + \sum_{i \in A_2} \sum_{j \in A_2} w_{1i} \frac{\Delta_{2ij|1}}{\pi_{2ij|1}} 
+  \frac{(y_i - \bf z_i q_i \hat{\bm \phi})}{\pi_{2i|1}}
+  \frac{(y_j - \bf z_j q_j \hat{\bm \phi})}{\pi_{2j|1}}.
+  $$
+
+\end{proof}
+
+\subsection{Proof of Theorem 4}
+
+\begin{proof}
+  From Theorem~\ref{thm:dc2}, we know that $\hat Y_{\rm NNE}(\hat \lambda) = 
+  \hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*) + O_p(Nn_2^{-1})$. Hence, the
+  variance of $\hat Y_{\rm NNE}(\hat{\bm \lambda})$ is
+
+  \begin{align*}
+    \Var(\hat Y_{\rm NNE}(\hat{\bm \lambda})) 
+    &\doteq \Var(\hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*)) \\ 
+    &= \Var\left(\sum_{i \in A_2} \hat w_{2i}(\bm \lambda^*) y_i + 
+      \left(\bf T - \sum_{i \in A_2} \hat w_{2i}(\bm \lambda^*) \bf z_i
+    q_i\right)\bm \phi^*\right)\\
+    &= (\bm \phi_1^*)^\top \Var(\hat{\bf X}_c) \bm \phi_1^* + 
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
+    (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
+    &\qquad+ 2 \Cov\left(\hat{\bf X}_c \bm \phi_1^*, \sum_{i \in A_2} 
+      \frac{(y_i - \bf z_i \bm \phi^* q_i)}{\pi_{2i}}\right) \\
+    &= (\bm \phi_1^*)^\top \Var(\hat{\bf X}_c) \bm \phi_1^* + 
+    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
+    (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
+    &\qquad + (1 - W)\bm \phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij} 
+    \frac{\bf{x}_i}{\pi_{2i}} \frac{(y_j - \bf z_j \bm \phi^*_1 q_j)}{\pi_{2j}}\\
+  \end{align*}
+
+where the last equality comes from the fact that $\hat{\bf X}_c = W\hat{\bf
+X}_1 + (1 - W) \hat{\bf X}_2$. To have an unbiased estimator of the variance
+  we can use:
+
+  \begin{align*}
+    \hat V_{\rm NNE} 
+    &= (\hat{\bm \phi_1})^\top \Var(\hat{\bf X_c}) \hat{\bm \phi_1} + 
+    \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
+    (y_i - \bf z_i \hat{\bm \phi} q_i)(y_j - \bf z_j \hat{\bm \phi} q_j) \\
+    &\qquad + (1 - W)\hat{\bm \phi}_1 \sum_{i \in A_2} \sum_{j \in A_2} 
+    \frac{\Delta_{2ij}}{\pi_{2ij}} \frac{\bf{x}_i}{\pi_{2i}} 
+    \frac{(y_j - \bf z_j \hat{\bm \phi_1} q_j)}{\pi_{2j}}
+  \end{align*}
+
+  where
+
+  $$\hat{\bm \phi} =
+  \begin{bmatrix}
+    \hat{\bm \phi_1} \\ \hat \phi_2
+  \end{bmatrix} = 
+  \left(\sum_{i \in A_2} \frac{q_i}{g'(d_{2i})} 
+  \begin{bmatrix}
+    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
+    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \end{bmatrix}
+  \right)^{-1}
+  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})} 
+  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
+  $$
+
+\end{proof}
+
+\section{Estimating the Population Weights}
 

### 2025-01-23 final/paper/paper.tex
Updated paper with Dr. Kim's suggestions.

@@
 
-Updates:
-\begin{itemize}
-  \item Modified the end of paragraph two of the introduction and the beginning
-    of paragraph three of the introduction to mirror more of the content of the
-    rest of the paper.
-  \item Fixed spacing issues with the beginning of paragraphs.
-  \item Added Table~\ref{tab:msexample}
-\end{itemize}
 \newpage
@@
 A_2} \Pr(A_2 \mid A_1)$ be the first-order inclusion probability for the
-second-phase sample. From the first-phase sample, we can compute
-$\widehat{\mathbf{X}}_1 = \sum_{i \in A_1} \pi_{1i}^{-1} \bx_i$ as an unbiased
+second-phase sample. We write $d_{1i} = \pi_{1i}^{-1}$ and $d_{2i|1} =
+\pi_{2i|1}^{-1}$ to denote the design weights.
+From the first-phase sample, we can compute
+$\widehat{\mathbf{X}}_1 = \sum_{i \in A_1} d_{1i}\bx_i$ as an unbiased
 estimator of $\mathbf{X}_N = \sum_{i=1}^N  \bx_i$. Also, from the second-phase
 sample, we can compute $\left( \widehat{\mathbf{X}}_2 , \hat{Y}_2 \right) =
-\sum_{i \in A_2} \pi_{1i}^{-1} \pi_{2i \mid 1}^{-1} \left( \bx_i, y_i\right)$ as
+\sum_{i \in A_2} d_{1i} d_{2i \mid 1} \left( \bx_i, y_i\right)$ as
 an unbiased estimator of $(\mathbf{X}_N, Y_N)$.
@@
 \hat Y_{\rm reg, tp} 
-= \sum_{i \in A_1} \frac{1}{\pi_{1i}} \bf x_i^\top  \hat{\bm \beta_q} + 
-\sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}} \left(y_i - \bf x_i^\top  \hat{\bm \beta_q}\right)
+= \sum_{i \in A_1} d_{1i} \bf x_i^\top  \hat{\bm \beta_q} + 
+\sum_{i \in A_2} d_{1i} d_{2i|1} \left(y_i - \bf x_i^\top  \hat{\bm \beta_q}\right)
 \end{equation} 
 where $q_i = q(\bf x_i)$  is a function of $\bf x_i$ and  
-
 $$
 \hat{\bm \beta_q} = \left(\sum_{i \in A_2} 
- \frac{1}{\pi_{1i} q_i } \bf x_i \bf x_i^\top \right)^{-1} 
-\sum_{i \in A_2} \frac{1}{\pi_{1i} q_i} \bf x_i y_i .
+d_{1i} \bf x_i \bf x_i^\top q_i \right)^{-1} 
+ \sum_{i \in A_2} d_{1i}\bf{x}_i y_i q_i.
 $$
@@
 where $e_i \sim \left(0, \sigma^2 c_i \right)$, where $c_i=c( \bx_i)$ is a known
-function. In this case,  $q_i = c_i$ is a reasonable choice. 
-One way to understand the two-phase regression estimator is to use the following
-decomposition:
-
+function. In this case,  $q_i = c_i^{-1}$ is a reasonable choice. 
+One way to understand the two-phase regression estimator in \eqref{eq:1}
+is to use the following decomposition: 
 $$ 
 \hat Y_{\rm reg, tp} 
-= \underbrace{\sum_{i \in A_1} \frac{\bf x_i^\top \hat{\bm \beta_q}}{\pi_{1i}}}_{
+= \underbrace{\sum_{i \in A_1} d_{1i}\bf x_i^\top \hat{\bm \beta_q}}_{
   \text{ Prediction  }} + \underbrace{\sum_{i \in A_2}
-\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i^\top \hat{\bm \beta_q})}_{
+d_{1i} d_{2i|1}(y_i - \bf x_i^\top \hat{\bm \beta_q})}_{
 \text{Bias correction}}.
@@
 makes the two-phase regression estimator asymptotically design unbiased
-regardless of the underlying the superpopulation model. 
-Note that the two-phase regression estimator in Equation~\ref{eq:1} can be
-expressed as  $\hat Y_{\rm reg, tp} =
-\sum_{i \in A_2} \pi_{1i}^{-1} \hat w_{2i} y_i$ where
-
-$$
-\hat w_{2} = \argmin_{w_2} \sum_{i \in A_2} (w_{2i} - \pi_{2i|1}^{-1})^2 q_i
-\text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i / \pi_{1i} = \sum_{i \in A_1}
-\bf x_i / \pi_{1i}.
+regardless if the underlying the superpopulation model is correct or not.
+We can find an equivalent estimator in the form $\hat Y_{\rm reg, tp} =
+\sum_{i \in A_2} d_{1i} \hat w_{2i} y_i$. This is useful because it shows that
+the estimator is linear and it is interesting because the estimated weights
+$\hat w_{2i}$ are the weights closest to the design weights that satisfy the
+constraints from \eqref{eq:constr}.
+\begin{equation}\label{eq:constr}
+\sum_{i \in A_2} d_{1i} w_{2i} \bf{x}_i 
+= \sum_{i \in A_1} d_{1i} \bf{x}_i.
+\end{equation}
+Closeness is in an $L^2$-sense. Overall, the weights $\hat w_{2i}$ are the
+solution to the following optimization problem:
+$$
+\hat w_{2} = \argmin_{w_2} \sum_{i \in A_2} (w_{2i} - d_{2i|1})^2 q_i
+\text{ such that } \sum_{i \in A_2} d_{1i} w_{2i} \bf{x}_i 
+= \sum_{i \in A_1} d_{1i} \bf{x}_i.
 $$
@@
 defined on an interval containing $d_i$, with $g(w_i, d_i) = \partial G /
-\partial w$ being continuous.\footnote{The \cite{deville1992calibration} paper
+\partial w$ being continuous with respect 
+to $w_i$.\footnote{The \cite{deville1992calibration} paper
 considers regression estimators for a single phase setup, which we apply to our
@@
 \cite{kwon2024debiased} shows that for a generalized entropy
-function $G(w)$, including a term of $g(\pi_{2i|1}^{-1})$ into the calibration
+function $G(w)$, including a term of $g(d_{2i|1})$ into the calibration
 for $g = \partial G / \partial w$ not only creates a design consistent
@@
 \begin{equation}\label{eq:primalloss}
-  H(w_2) = - \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i
+  H(w_2) = - \sum_{i \in A_2} d_{1i} G(w_{2i}) q_i
 \end{equation}
@@
 \begin{equation}\label{eq:calconst1}
-  \sum_{i \in A_2} \frac{\bf x_i w_{2i}}{\pi_{1i}} = 
-\sum_{i \in A_1} \frac{\bf x_i}{\pi_{1i}}
+  \sum_{i \in A_2} d_{1i} w_{2i} \bf{x}_i = 
+  \sum_{i \in A_1} d_{1i}\bf{x}_i
 \end{equation}
@@
 \begin{equation}\label{eq:calconst2}
-  \sum_{i \in A_2} \frac{g(\pi_{2i|1}^{-1})w_{2i}q_i}{\pi_{1i}} = 
-  \sum_{i \in A_1} \frac{g(\pi_{2i|1}^{-1})q_i}{\pi_{1i}}, 
-\end{equation}
-where $g(w) = \partial G / \partial w$.
-
-The first constraint is the usual calibration constraint motivated from the
-regression model in (\ref{regression}) and the second
+  \sum_{i \in A_2} d_{1i} w_{2i} g(d_{2i|1})q_i = 
+  \sum_{i \in A_1} d_{1i} g(d_{2i|1})q_i, 
+\end{equation}
+where $g(w) = \frac{\partial G}{\partial w}$.
+
+Also known as covariate balancing \citep{imai2014covariate},
+constraint~\eqref{eq:calconst1} ensures that the weighted mean of the
+covariates, $\bf{x}_i$, are the same.
+This is also the usual calibration constraint motivated from
+the regression model in \eqref{regression}.
+Constraint~\eqref{eq:calconst2} is how we incorporate the design weights. Using
+$g(w) = \frac{\partial G}{\partial w}$, ultimately ensures that $\hat w_{2i}$
+converges to $d_{2i}$ in probability. This 
 ensures that design consistency is achieved. 
 The original method of \cite{kwon2024debiased} only considers known finite
-population quantities on the right hand side of \eqref{eq:calconst1}.
-By writing $w_{1i} = \pi_{1i}^{-1}$, the the goal is to solve
+population quantities on the right hand side of \eqref{eq:calconst1}. We extend
+their method within our two-phase framework to allow for an estimated quantity
+on the right hand side of \eqref{eq:calconst1}.
+To estimate the weights, we solve
 \begin{equation}\label{eq:primal}
-  \hat w_2  = \argmin_{w_{2}} \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
+  \hat w_2  = \argmin_{w_{2}} \sum_{i \in A_2} d_{1i} G(w_{2i}) q_i 
   \text{ such that}
-  \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i = \sum_{i \in A_1} w_{1i} \bf z_i q_i
-\end{equation}
-where  $\bf z_i = (\bf x_i / q_i, g(\pi_{2i|1}^{-1}))$. The resulting 
-estimator of $Y_N$ is 
+  \sum_{i \in A_2} d_{1i} w_{2i} \bf z_i = \sum_{i \in A_1} d_{1i} \bf z_i 
+\end{equation}
+where  $\bf z_i = (\bf x_i , g(d_{2i|1})q_i)$. The resulting 
+estimator of $\bar Y_N$ is 
 \begin{equation}\label{eq:tpest}
-\hat Y_{\rm DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i} y_i.
+  \hat Y_{\rm DCE} = N^{-1} \sum_{i \in A_2} d_{1i} \hat w_{2i} y_i.
 \end{equation}
@@
 
-We can understand the estimated weights, $\hat w_{2i}$ from
-Equation~\ref{eq:primal} using the method of Lagrange multipliers. We need to
+We compute the estimated weights, $\hat w_{2i}$ from
+\eqref{eq:primal} using the method of Lagrange multipliers. We 
 minimize the Lagrangian function 
 \begin{equation}\label{eq:legragedc1}
-  L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
-  + \bm \lambda^\top \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
-    \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i\right)
-\end{equation}
-where $\bm \lambda$ is a vector of Lagrange multipliers.
+  L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} d_{1i} G(w_{2i}) q_i 
+  + \bm \lambda^\top \left( \sum_{i \in A_1} d_{1i} \bf z_i -
+    \sum_{i \in A_2} d_{1i} w_{2i} \bf z_i\right)
+\end{equation}
+where $\bm \lambda \in \R^{p+1}$ is a vector of Lagrange multipliers and
+$\bf{z}_i = (\bf{x}_i^\top, g(d_{2i|1})q_i)^\top$.
 Differentiating with respect to $w_{2i}$ and setting this expression equal to
@@
 $$ \hat w_{2i} = g^{-1}(\hat{\bm \lambda}^\top \bf z_i) $$
-where $g(w) = \partial G / \partial w$ and 
+where $g(w) = \frac{\partial G}{\partial w}$ and 
 $\hat{\bm \lambda}$ is the solution to
 \begin{equation}\label{eq:lamdc1}
-  \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
-  \sum_{i \in A_2} w_{1i} w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right) = 0.
-\end{equation}
-The following theorem shows that the construction of the weights in
-Equation~\ref{eq:primal} leads to a design consistent estimator 
-$\hat Y_{\rm DCE}$ in Equation~\ref{eq:tpest}.
+  \sum_{i \in A_1} d_{1i} \bf{z}_i  -
+  \sum_{i \in A_2} d_{1i} w_{2i}(\hat{\bm \lambda}) \bf{z}_i = 0.
+\end{equation}
+Under regularity conditions presented in the appendix, this solution is unique.
+The following theorem presents a linearization of
+\eqref{eq:primal} which leads to a design consistent estimator 
+$\hat Y_{\rm DCE}$ in \eqref{eq:tpest}.
 
@@
   \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) = 
-  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} y_i
-  + \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - 
-  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} \bf{z}_i q_i\right)\bm \phi^*,
+  \sum_{i \in A_2} d_{1i} d_{2i|1} y_i
+  + \left(\sum_{i \in A_1} d_{1i} \bf z_i - 
+  \sum_{i \in A_2} d_{1i} d_{2i|1} \bf{z}_i \right)\bm \phi^*,
   $$
@@
   \bm \phi^* = 
-  \left[\sum_{i \in U} \frac{\pi_{2i|1}\bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}\right]^{-1}
-  \sum_{i \in U} \frac{\pi_{2i|1}\bf z_i y_i}{g'(d_{2i|1})}.
+  \left[\sum_{i \in U} \frac{\pi_{2i|1}\bf z_i \bf z_i^\top}{g'(d_{2i|1})q_i}\right]^{-1}
+  \sum_{i \in U} \frac{\pi_{2i|1}\bf z_i y_i}{g'(d_{2i|1})q_i}.
   $$
@@
 %   \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) 
-%   &= \hat{Y}_{\rm DCE} + \left(\sum_{i \in A_1} w_{1i} \bf x_i  -
-%     \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \bf x_i  \right)^\top \bm \phi_1^* \\
-%   &\quad+ \left(\sum_{i \in A_1} w_{1i} g_i  -
-%  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^\top \phi_2^*  
+%   &= \hat{Y}_{\rm DCE} + \left(\sum_{i \in A_1} d_{1i} \bf x_i  -
+%     \sum_{i \in A_2} d_{1i} \pi_{2i \mid 1}^{-1} \bf x_i  \right)^\top \bm \phi_1^* \\
+%   &\quad+ \left(\sum_{i \in A_1} d_{1i} g_i  -
+%  \sum_{i \in A_2} d_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^\top \phi_2^*  
 %  \end{align*} 
@@
 Theorem~\ref{thm:dc1} is that asymptotically the two-phase debiased calibration
-method is equivalent to a regression estimator. By constraints 
-(\ref{eq:calconst1}) and (\ref{eq:calconst2}), 
-$\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ is design consistent, and the
-variance is asymptotically equivalent to the two-phase regression estimator. 
-We can estimate $\bm \phi^*$ with
+method is equivalent to a two-phase regression estimator. By Theorem~\ref{thm:dc1}, 
+$\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ is design consistent,
+and the proposed calibration estimator in \eqref{eq:primal} is
+asymptotically equivalent to the two-phase regression estimator given by 
+  $$
+  \hat Y_{\rm tp, reg}  = 
+  \sum_{i \in A_2} d_{1i} d_{2i|1} y_i
+  + \left(\sum_{i \in A_1} d_{1i} \bf z_i  - 
+  \sum_{i \in A_2} d_{1i} d_{2i|1} \bf{z}_i \right)\hat{\bm \phi},
+  $$
+where $\hat{\bm \phi}$ is defined in \eqref{eq:8}
 \begin{equation}
@@
 \left[\sum_{i \in A_2} 
-w_{1i}\frac{\bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}\right]^{-1}
-\sum_{i \in A_2} w_{1i}\frac{\bf z_i y_i}{g'(d_{2i|1})}.
+d_{1i}\frac{\bf z_i \bf z_i^\top}{g'(d_{2i|1})q_i}\right]^{-1}
+\sum_{i \in A_2} d_{1i}\frac{\bf z_i y_i}{g'(d_{2i|1})q_i}.
 \label{eq:8}
 \end{equation}
-Therefore, the proposed calibration estimator in (\ref{eq:primal}) is
-asymptotically equivalent to the two-phase regression estimator given by 
-  $$
-  \hat Y_{\rm tp, reg}  = 
-  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} y_i
-  + \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - 
-  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} \bf{z}_i q_i\right)\hat{\bm \phi},
-  $$
-where $\hat{\bm \phi}$ is defined in (\ref{eq:8}).
 
@@
 \mathbf{B} = 
-\left\{\sum_{i \in A_1} w_{1i} \bf{z}_i^\top \bm \gamma + 
-\sum_{i \in A_2} \frac{w_{1i}}{\pi_{2i|1}}(y_i - \bf{z}_i^\top \bm \gamma):
+\left\{\sum_{i \in A_1} d_{1i} \bf{z}_i^\top \bm \gamma + 
+\sum_{i \in A_2} d_{1i}d_{2i|1}(y_i - \bf{z}_i^\top \bm \gamma):
 \bm \gamma \in \R^{p+1}\right\}
 $$
-Then the design optimal estimator is 
+The design optimal estimator is 
 $$
 \hat \theta_{\ell, \rm{opt}} = 
-\sum_{i \in A_1} w_{1i} \bf{z}_i^\top \bm \phi_{\rm opt} + 
-\frac{w_{1i}}{\pi_{2i|1}}\sum_{i \in A_2} (y_i - \bf{z}_i^\top \bm \phi_{\rm opt}):
+\sum_{i \in A_1} d_{1i} \bf{z}_i^\top \bm \phi_{\rm opt} + 
+\sum_{i \in A_2} d_{1i}d_{2i|1}(y_i - \bf{z}_i^\top \bm \phi_{\rm opt}):
 $$
@@
 $$
-G(w_{2i|1}) = - \log(w_{2i|1}) + \log(w_{2i|1}w_{1i} - 1).
+G(w_{2i|1}) = - \log(w_{2i|1}) + \log(w_{2i|1}d_{1i} - 1).
 $$
@@
 The next theorem discusses how to compute an estimate for the variance of 
-$\hat Y_{\rm DCE}$. For the proof see the appendix.
+$\hat Y_{\rm DCE}$. For the proof see the appendix. The idea behind the proof is
+to use the two-phase sampling approach to estimate the variance. If the sampling
+fraction $n_2 / n_1 \to 0$ as $N$ increases, then the reverse framework of
+\cite{fay1992inferences} allows us to make a good approximation.
 
@@
 
-  Define $\hat \eta_i = \bf z_i q_i \hat{\bm \phi} + 
-  \frac{\delta_i}{\pi_{2i|1}}(y_i - \bf z_i q_i \hat{\bm \phi})$.
+  Define $\hat \eta_i = \bf z_i \hat{\bm \phi} + 
+  \delta_i d_{2i|1}(y_i - \bf z_i \hat{\bm \phi})$.
   We can construct an unbiased estimate the variance of $\hat Y_{\rm DCE}$ with
@@
   \hat V_{\rm DCE} 
-  = \sum_{i \in A_1} \sum_{j \in A_1} \frac{\Delta_{1ij}}{\pi_{1ij}} 
+  = \sum_{i \in A_1} \sum_{j \in A_1} d_{1ij} \Delta_{1ij}
   \hat \eta_i \hat \eta_j
-  + \sum_{i \in A_2} \sum_{j \in A_2} w_{1i} \frac{\Delta_{2ij|1}}{\pi_{2ij|1}} 
-  \frac{(y_i - \bf z_i q_i \hat{\bm \phi})}{\pi_{2i|1}}
-  \frac{(y_j - \bf z_j q_j \hat{\bm \phi})}{\pi_{2j|1}}
-  $$
-  where $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ and 
-  $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$.
+  + \sum_{i \in A_2} \sum_{j \in A_2} d_{1i} d_{2ij|1}\Delta_{2ij|1}
+  d_{2i|1}(y_i - \bf z_i \hat{\bm \phi})
+  d_{2j|1}(y_j - \bf z_j \hat{\bm \phi})
+  $$
+  where $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$,
+  $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$,
+  $d_{1ij} = \pi_{1ij}^{-1}$, $d_{2ij|1} = \pi_{2ij|1}^{-1}$,
+  $\pi_{1ij} = \Pr(i \in A_1, j \in A_1)$, and 
+  $\pi_{1ij} = \Pr(i \in A_2, j \in A_2 | i \in A_1, j \in A_1)$.
 \end{theorem}
@@
 
-Now we consider the sampling mechanism known as non-nested two-phase sampling 
+We now consider the sampling mechanism known as non-nested two-phase sampling 
 \citep{hidiroglou2001double}. In the last section, we considered two-phase sampling
@@
 Like traditional two-phase sampling, we consider the Phase 1 sample, $A_1$, to
-consist of observations of $(\bf X_i)_{i = 1}^{n_1}$ and the Phase 2 sample,
-$A_2$, to consist of observations of $(\bf X_i, Y_i)_{i = 1}^{n_2}$.
+consist of observations of $(\bf X_i)$ and the Phase 2 sample,
+$A_2$, to consist of observations of $(\bf X_i, Y_i)$. We denote the sample size
+of $A_1$ as $n_1$ and the sample size of $A_2$ as $n_2$.
 
@@
 Then we can define a regression estimator as
-$$
-\hat Y_{\rm NN, reg} = \hat Y_2 + (\hat{\bf X}_c - \hat{\bf X}_2)^\top \hat{\bm \beta_q} = 
+\begin{equation}\label{eq:nntpreg}
+\hat Y_{\rm NN, reg} = 
+\hat Y_2 + (\hat{\bf X}_c - \hat{\bf X}_2)^\top \hat{\bm \beta_q} = 
 \hat Y_2 + (\hat{\bf X}_1 - \hat{\bf X}_2)^\top W\hat{\bm \beta_q}
-$$
+\end{equation}
 where, $W = n_{1, \rm{eff}} / (n_{1, \rm{eff}} + n_{2, \rm{eff}})$, and
@@
 \hat{\bm \beta_q} = 
-\left(\sum_{i \in A_2} \frac{\bf x_i \bf x_i^\top}{q_i}\right)^{-1} 
-\sum_{i \in A_2} \frac{\bf x_i y_i}{q_i} \text{ and }
+\left(\sum_{i \in A_2} \bf x_i \bf x_i^\top q_i\right)^{-1} 
+\sum_{i \in A_2} \bf x_i y_i q_i \text{ and }
 \hat Y_2 = \sum_{i \in A_2} d_{2i} y_i.
@@
 V(\hat Y_{\rm NN, reg}) = 
-V\left(\sum_{i \in A_2} \frac{1}{\pi_{2i}}(y_i - \bf x_i^\top W\bm \beta^*_q)\right)
+V\left(\sum_{i \in A_2} d_{2i}(y_i - \bf x_i^\top W\bm \beta^*_q)\right)
 + (\bm \beta^*_q)^\top W^\top V(\hat{\bf X}_1) W \bm \beta_q^*
@@
 where $\bm \beta_q^*$ is the probability limit of $\hat{\bm \beta_q}$. Like the
-two-phase sample, this regression estimator can be viewed as the solution to the
-following calibration equation 
+two-phase sample, the regression estimator in \eqref{eq:nntpreg} can be viewed
+as a linear estimator given a particular choice of weights $\hat w_{2i}$. These
+weights are the solution to the calibration equation in \eqref{eq:nncal}.
 \begin{equation}\label{eq:nncal}
@@
 \end{equation}
-and $\hat Y_{\rm NN, reg} = \sum_{i \in A_2} \hat w_{2i} y_i$ where $\hat
-w_{2i}$ is the solution to Equation~\ref{eq:nncal}.
+These estimated weights yield an estimator of $\theta = E[Y]$ of 
+$\hat Y_{\rm NN, reg} = N^{-1} \sum_{i \in A_2} \hat w_{2i} y_i$. 
 We can extend the debiased calibration estimator of \cite{kwon2024debiased} to
@@
 described as part of Topic 1. Given a strictly convex differentiable function,
-$G: \mathcal{V} \to \mathbb{R}$, the goal is to solve
+$G: \mathcal{V} \to \mathbb{R}$, the goal is to minimize the variance of the
+estimator while ensuring that the covariate balancing condition is met. This
+results in $\hat w_2$ being the solution to the optimization problem in 
+\eqref{eq:nnopt}
 \begin{equation}\label{eq:nnopt}
 \hat w_2 = \argmin_{w_2} \sum_{i \in A_2} G\left(w_{2i}\right) q_i 
-\text{ such that } 
+\text{ where } 
 \sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X}_{c} \text{ and } 
@@
 The difference
-between solving Equation~\ref{eq:nnopt} and Equation~\ref{eq:primal} is that the
+between solving \eqref{eq:nnopt} and \eqref{eq:primal} is that the
 estimator $\hat{\bf X}_c$ is estimated from the combined sample $A_c = A_1 \cup
@@
 $$
-
-Let $n_{1, \rm{eff}}$ and $n_{2, \rm{eff}}$ be the effective samples sizes for
-$A_1$ and $A_2$ respectively. Then the optimal combined estimate is 
+If $n_{1, \rm{eff}}$ and $n_{2, \rm{eff}}$ are the effective samples sizes for
+$A_1$ and $A_2$ respectively, then the optimal combined estimate is 
 $$
@@
 $$
-We can construct a non-nested two-phase estimator $\hat Y_{\rm NNE}$ for $Y_N$ where
-$\hat Y_{\rm NNE} = \sum_{i \in A_2} \hat w_{2i} y_i$ and $\hat w_{2i}$ solves
-Equation~\ref{eq:nnopt}. Like the classical two-phase approach, to solve this
+We can construct a non-nested two-phase estimator $\hat Y_{\rm NNE}$ for 
+$\bar Y_N$ where
+$\hat Y_{\rm NNE} = N^{-1} \sum_{i \in A_2} \hat w_{2i} y_i$ and $\hat w_{2i}$
+solves \eqref{eq:nnopt}. Like the classical two-phase approach, to solve this
 setup we minimize the Lagrangian,
@@
   L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} G(w_{2i}) q_i + 
-  \bm \lambda^\top \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i} \bf z_i q_i\right).
-\end{equation}
-where $\bm \lambda$ are the Lagrange multipliers, with 
+  \bm \lambda^\top \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i} \bf z_i \right).
+\end{equation}
+where $\bm \lambda$ are the Lagrange multipliers, $\bf{z}_i = (\bf{x}_i^\top, g(d_{2i})q_i)^\top$,
+and
 $$\hat{\bf T} = 
@@
 \begin{equation}\label{eq:lamdc2}
-  \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i}(\hat{\bm \lambda}) \bf z_i
-  q_i\right) = 0.
-\end{equation}
-Like the previous section we first prove that the estimator is design consistent.
-Then we provide a result showing how to estimate the variance.
+  \hat{\bf T} - \sum_{i \in A_2} w_{2i}(\hat{\bm \lambda}) \bf z_i = 0.
+\end{equation}
+Like the previous section we first prove that the estimator is 
+asymptotically equivalent to a regression estimator. This ensures that it is
+design consistent. The asymptotic equivalence is useful because our estimator
+now has the same variance as a regression estimator. Using this fact,
+we provide a result showing how to estimate the variance.
 
 \begin{theorem}[Design Consistency]\label{thm:dc2}
-  Allowing $\bm \lambda^*$ to be the probability limit of $\hat{\bm \lambda}$,
+  Let $\bm \lambda^*$ to be the probability limit of $\hat{\bm \lambda}$,
   under some regularity conditions,
@@
   \hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*) = 
-  \sum_{i \in A_2} \pi_{2i}^{-1} y_i + 
-  \left(\hat{\bf T} - \sum_{i \in A_2} \pi_{2i}^{-1} \bf z_i q_i\right)^\top \bm \phi^*
+  \sum_{i \in A_2} d_{2i} y_i + 
+  \left(\hat{\bf T} - \sum_{i \in A_2} d_{2i} \bf z_i \right)^\top \bm \phi^*
   $$
@@
   \bm \phi^* =
-  \left(\sum_{i \in U} \frac{\pi_{2i} q_i}{g'(d_{2i})} 
+  \left(\sum_{i \in U} \frac{\pi_{2i}}{g'(d_{2i})q_i} 
   \begin{bmatrix}
-    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
-    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
+    \bf x_i^2 & \bf x_i g(d_{2i}) q_i\\
+    \bf x_i g(d_{2i}) q_i & g(d_{2i})^2 q_i^2
   \end{bmatrix}
   \right)^{-1}
-  \sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})} 
-  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
+  \sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})q_i} 
+  \begin{bmatrix} \bf x_i \\ g(d_i) q_i \end{bmatrix}.
   $$
@@
 estimated from both samples using $\hat{\bf X}_c$ instead of 
-$\hat{\bf X}_{HT}$ from the Phase 1 sample. 
-Since $\hat Y_{\rm NNE} = \sum_{i \in A_2} \hat w_{2i}(\hat{\bm \lambda})y_i$
-where $\hat{\bm \lambda}$ solves
-\begin{equation}
-  \sum_{i \in A_2} \hat w_{2i}(\bm \lambda) q_i
-  \underbrace{
-  \begin{bmatrix} 
-    \bf x_i / q_i \\ g(d_i)
-\end{bmatrix}}_{\bf z_i} = \bf T
-\end{equation}
-we have 
+$\hat{\bf X}_{\rm HT}$ from the Phase 1 sample. 
+By the dual formulation of \eqref{eq:legragedc2}, for any $\bm \phi$
 $$\hat Y_{\rm NNE} (\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} \hat
 w_{2i}(\hat{\bm \lambda})y_i + \left(\bf T - \sum_{i \in A_2} \hat
-w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right)^\top \bm \phi.$$
+w_{2i}(\hat{\bm \lambda}) \bf z_i \right)^\top \bm \phi.$$
 If we choose $\bm \phi^*$ such that $E\left[\frac{\partial}{\partial \bm \lambda} 
@@
 \end{bmatrix} = 
-\left(\sum_{i \in U} \frac{\pi_{2i} q_i}{g'(d_{2i})} 
+\left(\sum_{i \in U} \frac{\pi_{2i}}{g'(d_{2i})q_i} 
 \begin{bmatrix}
-  \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
-  \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
+  \bf x_i^2  & \bf x_i g(d_{2i}) q_i \\
+  \bf x_i g(d_{2i}) q_i & g(d_{2i})^2 q_i^2
 \end{bmatrix}
 \right)^{-1}
-\sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})} 
-\begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
+\sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})q_i} 
+\begin{bmatrix} \bf x_i \\ g(d_i) q_i\end{bmatrix}.
 $$
@@
 The next theorem shows how to estimate the variance of $\hat Y_{\rm NNE}$.
+Since we have already proves that the estimator is asymptotically equivalent to
+a type of regression estimator, we only need to compute the variance of the
+regression estimator to get the variance of $\hat Y_{\rm NNE}$.
 The proof of this result is in the appendix.
@@
     &= (\bm \phi_1^*)^\top \Var(\hat{\bf X}_c) \bm \phi_1^* + 
-    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
-      (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
+    \sum_{i \in U} \sum_{j \in U} \Delta_{2ij}d_{2i}
+    (y_i - \bf z_i \bm \phi^* q_i)d_{2j}(y_j - \bf z_j \bm \phi^* q_j)\\
     &\qquad + (1 - W)\bm \phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij}
@@
     &= (\hat{\bm \phi_1})^\top \Var(\hat{\bf X}_c) \hat{\bm \phi_1} + 
-    \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
-    (y_i - \bf z_i \hat{\bm \phi} q_i)(y_j - \bf z_j \hat{\bm \phi} q_j) \\
-    &\qquad + (1 - W)\hat{\bm \phi_1} \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}}
-    \frac{\bf x_i}{\pi_{2i}} \frac{(y_j - \bf z_j \hat{\bm \phi_1} q_j)}{\pi_{2j}}
+    \sum_{i \in A_2} \sum_{j \in A_2} d_{2ij} \Delta_{2ij} d_{2i}
+    (y_i - \bf z_i \hat{\bm \phi} q_i) d_{2j} (y_j - \bf z_j \hat{\bm \phi} q_j) \\
+    &\qquad + (1 - W)\hat{\bm \phi_1} \sum_{i \in A_2} \sum_{j \in A_2} d_{2ij} \Delta_{2ij}
+    d_{2i}\bf x_i d_{2j}(y_j - \bf z_j \hat{\bm \phi_1} q_j)
   \end{align*}
-  where 
+  where $\Delta_{2ij} = \pi_{2ij} - \pi_{2i}\pi_{2j}$, $d_{2ij} = \pi_{2ij}^{-1}$, 
+  $\pi_{2ij} = \Pr(i \in A_2, j \in A_2)$, and 
   $$\hat{\bm \phi} =
@@
   \end{bmatrix} = 
-  \left(\sum_{i \in A_2} \frac{q_i}{g'(d_{2i})} 
+  \left(\sum_{i \in A_2} \frac{1}{g'(d_{2i})q_i} 
   \begin{bmatrix}
-    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
-    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
+    \bf x_i^2 & \bf x_i g(d_{2i}) q_i \\
+    \bf x_i g(d_{2i}) q_i & g(d_{2i})^2 q_i^2
   \end{bmatrix}
   \right)^{-1}
-  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})} 
-  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
+  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i}) q_i} 
+  \begin{bmatrix} \bf x_i \\ g(d_i) q_i \end{bmatrix}.
   $$
@@
 the $X$-variables measured in sample $A_0$, and $G(\cdot)$ is a generalized entropy
-function. 
+function. In this setup, all of the data integration occurs when estimating
+$\hat{\bf{X}}_{\rm GLS}$, which is used as a calibration constraint.
 In order to have a design consistent estimator that incorporates information from
-samples $A_0, A_1, \dots, A_M$, we want to add the debiasing constraint, 
+samples $A_0, A_1, \dots, A_M$, we want to add the debiasing constraint from 
+\eqref{eq:msdebconstr} because it ensures design consistency of the estimated
+weights
 \begin{equation}\label{eq:msdebconstr}
@@
 \end{equation}
-where $\widehat{\bf{T}} = (\widehat{\bf X}_{\rm{GLS}}^{(0)},
+where $\widehat{\bf{T}} = ((\hat{\bf X}_{\rm{GLS}}^{(0)})^\top,
 \sum_{i \in U} g(d_i) q_i)^\top$,
@@
 $$
-where $\bm \lambda$ are the Lagrange multipliers from Equation~\ref{eq:dc3}.
-We now show that $\hat Y_{\rm MS}$ is design consistent.
-
+where $\bm \lambda$ are the Lagrange multipliers from \eqref{eq:dc3}.
+We now show that $\hat Y_{\rm MS}$ is asymptotically equivalent to a regression
+estimator. This linearization technique shows design consistency.
 \begin{theorem}[Design Consistency]
@@
   \sum_{i \in A_0} d_{0i} y_i + \left(
-    \widehat{\bf T} - \sum_{i \in A_0} d_{0i} \bf z_i q_i
+    \widehat{\bf T} - \sum_{i \in A_0} d_{0i} \bf z_i
   \right)^\top \bm \phi^*
@@
 difference is that we have $\hat{\bf X}_{\rm GLS}$ instead of $\hat{\bf X}_c$.
+Since we have shown equivalence between our estimator and a regression estimator
+we can estimate the variance by giving the variance of the regression estimator.
 The next theorem gives a method for how to estimate the variance of 
 $\hat Y_{\rm MS}$.
-
 \begin{theorem}[Variance Estimation]
@@
     &= (\bm \phi^{*})^\top \Var(\hat{\bf X}_{\rm GLS}^{(0)}) (\bm \phi^{*}) + 
-      \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{0ij}}{\pi_{0i}\pi_{0j}}
-      (y_i - \bf z_i\bm \phi^{*}q_i)
-      (y_j - \bf z_j\bm \phi^{*}q_j). 
+      \sum_{i \in U} \sum_{j \in U} \Delta_{0ij}
+      d_{0i}(y_i - \bf z_i\bm \phi^{*}q_i)
+      d_{0j}(y_j - \bf z_j\bm \phi^{*}q_j). 
   \end{aligned}
@@
     &= (\hat{\bm \phi})^\top \hat \Var(\hat{\bf X}_{\rm GLS}^{(0)}) (\hat{\bm \phi})
-    + \sum_{i \in A_0} \sum_{j \in A_0} \frac{\Delta_{0ij}}{\pi_{0ij}\pi_{0i}\pi_{0j}}
-      (y_i - \bf z_i \hat{\bm \phi}q_i)
-      (y_j - \bf z_j \hat{\bm \phi}q_j). 
+  + \sum_{i \in A_0} \sum_{j \in A_0} d_{0ij}\Delta_{0ij}
+    d_{0i}(y_i - \bf z_i \hat{\bm \phi}q_i)
+    d_{0j}(y_j - \bf z_j \hat{\bm \phi}q_j)
   \end{aligned}
   $$
-
+  where $\Delta_{0ij} = \pi_{0ij} - \pi_{0i}\pi_{0j}$, $d_{0ij} = \pi_{0ij}^{-1}$,
+  and $\pi_{0ij} = \Pr(i \in A_0, j\in A_0)$.
 \end{theorem}
-
-This result follows the same argument as Theorem~\eqref{thm:nnve}. The result
+The result follows the same argument as Theorem~\ref{thm:nnve}. The result
 holds because each sample $A_m$ is independent from the other $A_{m'}$.
@@
   \begin{equation}
-    \argmin_{w_{2|1}} \sum_{i \in A_2} w_{1i} G(w_{2i}) \text{ such that}
-    \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i = \sum_{i \in U} \bf z_i.
+    \argmin_{w_{2|1}} \sum_{i \in A_2} d_{1i} G(w_{2i}) \text{ such that}
+    \sum_{i \in A_2} d_{1i} w_{2i} \bf z_i = \sum_{i \in U} \bf z_i.
   \end{equation}
@@
   \item[4.] Debiased Calibration with Estimated Population Constraints (Est):
-    This solves Equation~\eqref{eq:primal} with $q_i = 1$.
+    This solves \eqref{eq:primal} with $q_i = 1$.
 \end{itemize}
@@
 %    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_Y + 
-%      \sum\limits_{i \in A_2} \frac{w_{1i}^2}{N^2} 
+%      \sum\limits_{i \in A_2} \frac{d_{1i}^2}{N^2} 
 %        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}y_i^2$}
@@
 %    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\eta + 
-%      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
+%      \sum\limits_{i \in A_2} \frac{d_{1i}}{N^2} 
 %        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
@@
 %    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\varepsilon + 
-%      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
+%      \sum\limits_{i \in A_2} \frac{d_{1i}}{N^2} 
 %        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
@@
   \item[4.] Debiased Calibration with Estimated Population Constraints (Est):
-    This solves Equation~\eqref{eq:nnopt} with $q_i = 1$.
+    This solves \eqref{eq:nnopt} with $q_i = 1$.
 \end{itemize}
@@
   \item[2.] Non-nested regression (NNReg): This is the non-nested regression from
-    Equation~\eqref{eq:nnopt} with only using information from Samples $A_0$ and
+    \eqref{eq:nnopt} with only using information from Samples $A_0$ and
     $A_1$,
   \item[3.] Multi-Source proposed (Est): This is the proposed estimator from
-    Equation~\eqref{eq:dc3}, and
+    \eqref{eq:dc3}, and
   \item[4.] Multi-Source population (EstPop): This is the proposed estimator with
@@
 the population of all Medicare patients, we have to estimate the term $\sum_{i
-\in U} g(d_i)$ from Equation~\ref{eq:msdebconstr}. This technique is adopted
+\in U} g(d_i)$ from \eqref{eq:msdebconstr}. This technique is adopted
 from \cite{kwon2024debiased} and described in more details in the appendix.
@@
 
-In this proof, we derive the solution to Equation~\ref{eq:primal} and show
+In this proof, we derive the solution to \eqref{eq:primal} and show
 that it is asymptotically equivalent to a regression estimator. Using the
-method of Lagrange multipliers, to solve Equation~\eqref{eq:primal} we need to
-minimize the Lagrangian in Equation~\eqref{eq:legragedc1}. 
+method of Lagrange multipliers, to solve \eqref{eq:primal} we need to
+minimize the Lagrangian in \eqref{eq:legragedc1}. 
 The first order conditions show that at the solution,
 $$
-\frac{\partial \mathcal{L}}{\partial w_{2i}}: g(w_{2i}) w_{1i}q_i -
-w_{1i} \hat{\bm \lambda}^\top \bf z_i q_i = 0.
+\frac{\partial \mathcal{L}}{\partial w_{2i}}: g(w_{2i}) d_{1i}q_i -
+d_{1i} \hat{\bm \lambda}^\top \bf z_i = 0.
 $$
@@
 g^{-1}(\hat{\bm \lambda}^\top \bf z_i)$ where
-$\hat{\bm \lambda}$ is determined by Equation~\eqref{eq:lamdc1}. When the
+$\hat{\bm \lambda}$ is determined by \eqref{eq:lamdc1}. When the
 sample size gets large, we have $\hat w_{2i}(\hat{\bm \lambda}) \to
-d_{2i|1}$ in probability because both terms in Equation~\eqref{eq:lamdc1} must
+d_{2i|1}$ in probability because both terms in \eqref{eq:lamdc1} must
 become equal by the Central Limit Theorem. Hence,
@@
 \hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = 
-\sum_{i \in A_2} w_{1i} \hat w_{2i} y_i + 
-\left(\sum_{i \in A_1} w_{1i} \bm z_i q_i - \sum_{i \in A_2} w_{1i} 
-\hat w_{2i} \bf z_i q_i\right)\bm \phi.
-$$
-Notice that $\hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} w_{1i}
+\sum_{i \in A_2} d_{1i} \hat w_{2i} y_i + 
+\left(\sum_{i \in A_1} d_{1i} \bm z_i  - \sum_{i \in A_2} d_{1i} 
+\hat w_{2i} \bf z_i \right)^\top \bm \phi.
+$$
+Notice that $\hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} d_{1i}
 \hat w_{2i} y_i$ for all $\bm \phi$. We choose $\bm \phi^*$ such that
@@
 E\left[\frac{\partial}{\partial \bm \lambda} \hat Y_\ell(\bm \lambda^*,
-\bm \phi^*)\right]=0.
-$$
-Using the fact that $g^{-1}(\bm \lambda^* \bf z_i) = g^{-1}(g(d_{2i|1})) = d_{2i|1}$
+\bm \phi^*)\right]=\bf{0}.
+$$
+Using the fact that $g^{-1}((\bm \lambda^*)^\top \bf z_i) = g^{-1}(g(d_{2i|1})) = d_{2i|1}$
 and $(g^{-1})'(x) = 1 / g'(g^{-1}(x))$, we have
@@
   \bm \phi^*
-  &= E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}
+  &= E\left[\sum_{i \in A_2} \frac{d_{1i}\bf z_i \bf z_i^\top }{g'(d_{2i|1})q_i}
   \right]^{-1}
-  E\left[\sum_{i \in A_2} \frac{w_{1i}\bf z_i y_i}{g'(d_{2i|1})}\right]\\
-  &= \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}
+  E\left[\sum_{i \in A_2} \frac{d_{1i}\bf z_i y_i}{g'(d_{2i|1})q_i}\right]\\
+  &= \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i \bf z_i^\top }{g'(d_{2i|1})q_i}
   \right]^{-1}
-  \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i y_i}{g'(d_{2i|1})}\right]\\
+  \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i y_i}{g'(d_{2i|1})q_i}\right]\\
 \end{align*}
@@
 $$\hat Y_\ell(\bm \lambda^*, \bm \phi^*) = 
-\sum_{i \in A_1} w_{1i} q_i \bf z_i \bm \phi^* +
-\sum_{i \in A_2} w_{1i} d_{2i|1} (y_i - q_i \bf z_i \bm \phi^*).$$
+\sum_{i \in A_1} d_{1i} \bf{z}_i^\top \bm \phi^* +
+\sum_{i \in A_2} d_{1i} d_{2i|1} (y_i - \bf{z}_i^\top \bm \phi^*).$$
 By construction using a Taylor expansion yields,
@@
   \eta_i^* = 
-  \bf z_i q_i \bm \phi^* + \frac{\delta_i}{\pi_{2i|1}}(y_i - \bf z_i q_i \bm \phi^*)
+  \bf z_i \bm \phi^* + \delta_i d_{2i|1}(y_i - \bf{z}_i^\top \bm \phi^*)
   $$
@@
   $$
-  \hat Y_{\ell} = \sum_{i \in A_1} w_{1i} \eta_i^*.
+  \hat Y_{\ell} = \sum_{i \in A_1} d_{1i} \eta_i^*.
   $$
@@
     \Var(\hat Y_{\ell, \rm{DCE}}) 
-    &= \Var\left(\sum_{i \in A_1} w_{1i} \eta_i^*\right) \\
-    &= \Var\left(E\left[\sum_{i \in A_1} w_{1i} \eta_i^* \mid A_2 \right]\right) 
-    + E\left[\Var\left(\sum_{i \in A_1} w_{1i} \eta_i^* \mid A_2 \right)\right] \\
+    &= \Var\left(\sum_{i \in A_1} d_{1i} \eta_i^*\right) \\
+    &= \Var\left(E\left[\sum_{i \in A_1} d_{1i} \eta_i^* \mid A_2 \right]\right) 
+    + E\left[\Var\left(\sum_{i \in A_1} d_{1i} \eta_i^* \mid A_2 \right)\right] \\
     &= \Var\left(\sum_{i \in U} \eta_i^*\right) 
-    + E\left[\sum_{i \in U} \sum_{j \in U} \Delta_{1ij} w_{1i} \eta_i^* w_{1j}
+    + E\left[\sum_{i \in U} \sum_{j \in U} \Delta_{1ij} d_{1i} \eta_i^* w_{1j}
       \eta_j^*\right]\\
@@
   \hat V_{\rm DCE} 
-  = \sum_{i \in A_1} \sum_{j \in A_1} \frac{\Delta_{1ij}}{\pi_{1ij}} 
+= \sum_{i \in A_1} \sum_{j \in A_1} d_{1ij} \Delta_{1ij}
   \hat \eta_i \hat \eta_j
-  + \sum_{i \in A_2} \sum_{j \in A_2} w_{1i} \frac{\Delta_{2ij|1}}{\pi_{2ij|1}} 
-  \frac{(y_i - \bf z_i q_i \hat{\bm \phi})}{\pi_{2i|1}}
-  \frac{(y_j - \bf z_j q_j \hat{\bm \phi})}{\pi_{2j|1}}.
++ \sum_{i \in A_2} \sum_{j \in A_2} d_{1i} d_{2ij|1}\Delta_{2ij|1}
+  d_{2i|1}(y_i - \bf z_i \hat{\bm \phi})
+  d_{2j|1}(y_j - \bf z_j \hat{\bm \phi}).
   $$
@@
       \left(\bf T - \sum_{i \in A_2} \hat w_{2i}(\bm \lambda^*) \bf z_i
-    q_i\right)\bm \phi^*\right)\\
+    \right)^\top \bm \phi^*\right)\\
     &= (\bm \phi_1^*)^\top \Var(\hat{\bf X}_c) \bm \phi_1^* + 
-    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
-    (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
-    &\qquad+ 2 \Cov\left(\hat{\bf X}_c \bm \phi_1^*, \sum_{i \in A_2} 
-      \frac{(y_i - \bf z_i \bm \phi^* q_i)}{\pi_{2i}}\right) \\
+    \sum_{i \in U} \sum_{j \in U} \Delta_{2ij}d_{2i} 
+    (y_i - \bf{z}_i^\top \bm \phi^*)d_{2j}(y_j - \bf{z}_j^\top \bm \phi^*)\\
+    &\qquad+ 2 \Cov\left(\hat{\bf X}_c^\top \bm \phi_1^*, \sum_{i \in A_2} 
+    d_{2i}(y_i - \bf{z}_i^\top \bm \phi^*)\right) \\
     &= (\bm \phi_1^*)^\top \Var(\hat{\bf X}_c) \bm \phi_1^* + 
-    \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{2ij}}{\pi_{2i}\pi_{2j}} 
-    (y_i - \bf z_i \bm \phi^* q_i)(y_j - \bf z_j \bm \phi^* q_j)\\
-    &\qquad + (1 - W)\bm \phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij} 
-    \frac{\bf{x}_i}{\pi_{2i}} \frac{(y_j - \bf z_j \bm \phi^*_1 q_j)}{\pi_{2j}}\\
+    \sum_{i \in U} \sum_{j \in U} \Delta_{2ij}d_{2i}
+    (y_i - \bf{z}_i^\top \bm \phi^*)d_{2j}(y_j - \bf{z}_j^\top \bm \phi^*)\\
+    &\qquad + (1 - W)\left(\sum_{i \in U} \sum_{j \in U} \Delta_{2ij} 
+    d_{2i}\bf{x}_i d_{2j}(y_j - \bf{z}_j^\top \bm \phi^*_1)\right)^\top \bm \phi_1^* \\
   \end{align*}
@@
     &= (\hat{\bm \phi_1})^\top \Var(\hat{\bf X_c}) \hat{\bm \phi_1} + 
-    \sum_{i \in A_2} \sum_{j \in A_2} \frac{\Delta_{2ij}}{\pi_{2ij}\pi_{2i}\pi_{2j}} 
-    (y_i - \bf z_i \hat{\bm \phi} q_i)(y_j - \bf z_j \hat{\bm \phi} q_j) \\
-    &\qquad + (1 - W)\hat{\bm \phi}_1 \sum_{i \in A_2} \sum_{j \in A_2} 
-    \frac{\Delta_{2ij}}{\pi_{2ij}} \frac{\bf{x}_i}{\pi_{2i}} 
-    \frac{(y_j - \bf z_j \hat{\bm \phi_1} q_j)}{\pi_{2j}}
+    \sum_{i \in A_2} \sum_{j \in A_2} d_{2ij}\Delta_{2ij}d_{2i}
+    (y_i - \bf{z}^\top_i \hat{\bm \phi})d_{2j}(y_j - \bf{z}^\top_j \hat{\bm \phi}) \\
+    &\qquad + (1 - W)\left(\sum_{i \in A_2} \sum_{j \in A_2} 
+      d_{2ij} \Delta_{2ij} d_{2i}\bf{x}_i
+    d_{2j}(y_j - \bf{z}^\top_j \hat{\bm \phi_1})\right)^\top \hat{\bm \phi}_1 
   \end{align*}
@@
   \end{bmatrix} = 
-  \left(\sum_{i \in A_2} \frac{q_i}{g'(d_{2i})} 
+  \left(\sum_{i \in A_2} \frac{1}{g'(d_{2i})q_i} 
   \begin{bmatrix}
-    \bf x_i^2 / q_i & \bf x_i g(d_{2i}) / q_i \\
-    \bf x_i g(d_{2i}) / q_i & g(d_{2i})^2
+    \bf x_i^2 & \bf x_i g(d_{2i}) q_i \\
+    \bf x_i g(d_{2i}) q_i & g(d_{2i})^2 q_i^2
   \end{bmatrix}
   \right)^{-1}
-  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})} 
-  \begin{bmatrix} \bf x_i / q_i \\ g(d_i) \end{bmatrix}.
+  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})q_i} 
+  \begin{bmatrix} \bf x_i \\ g(d_i) q_i \end{bmatrix}.
   $$
@@
 
-In Equation~\ref{eq:primal}, Equation~\ref{eq:nnopt}, and
-Equation~\ref{eq:msdebconstr}, we see the need for knowing $\sum_{i \in U}
-g(d_i)$. (In the case of Equation~\ref{eq:primal}, we only need $\sum_{i \in
+In \eqref{eq:primal}, \eqref{eq:nnopt}, and
+\eqref{eq:msdebconstr}, we see the need for knowing $\sum_{i \in U}
+g(d_i)$. (In the case of \eqref{eq:primal}, we only need $\sum_{i \in
 A_1} g(d_{2i|1})$.) However, this quantity is not always known. When it is

### 2025-05-18 final/paper/paper.tex
Updated on May 17, 2025

@@
 regression methods in finite samples.
-
-\textcolor{red}{(I am not sure whether the current introduction provides a good
-  guide for the main contents of the paper. The introduction does not match
-  well with the main contents of the paper.)
-}
-
-\textcolor{red}{CL: I have modified the last two paragraphs to help tie into the
-main work of the paper.}
 
